### 2022-06-25 sandp23/appendix.tex
update paper.

\section{Appendix}

\subsection{Differences with the Coq and Redex Models}\label{app:model-diffs}

% \begin{figure}[h]
% {\small
% {\captionsetup[lstlisting]{margin = 8 mm}
%   \begin{lstlisting}[xleftmargin=8 mm]
% Inductive expression : Type :=
%   | ELit : Z -> type -> expression
%   | EVar : var -> expression
%   | EStrlen : var -> expression
%   | ECall : funid -> list expression -> expression
%   | ERet : var -> Z* type -> expression -> expression
%   | EDynCast : type -> expression -> expression
%   | ELet : var -> expression -> expression -> expression
%   | EMalloc : type -> expression
%   | ECast : type -> expression -> expression
%   | EPlus : expression -> expression -> expression
%   | EFieldAddr : expression -> field -> expression
%   | EDeref : expression -> expression (*  * e *)
%   | EAssign : expression -> expression -> expression (* *e = e *)
%   | EIfDef : var -> expression -> expression -> expression (* if *x then e1 else e2. *)
%   | EIf : expression -> expression -> expression -> expression (* if e1 then e2 else e3. *)
%   | EUnchecked : expression -> expression.
%   \end{lstlisting}
% }
% }
% \caption{Expression Syntax in Coq}
% \label{fig:coq-syntax}
% \end{figure}

The Coq and Redex models of \lang may be found at \url{https://github.com/plum-umd/checkedc}.
The Coq model's syntax is slightly different from that in
Fig.~\ref{fig:checkc-syn}. In particular, the arguments in a function call
are restricted to variables and constants, according to
a separate well-formedness condition. A function call \code{f(e)} can
always be written in \code{let x = e in f(x)} to cope.
% In the Coq model, we do not restrict the function argument syntax. Instead, we achieve the restriction
% by restricting the type rule for each function argument element to only deal with constants and variables. 
% To avoid recursive inductive relations and simplify proofs, we define different inductive relations for typing function argument lists and elements. In the inductive relation of function argument elements, we only define rules for typing constants and variables, and the rules are copied from the main inductive type rules for constants and variables. 
%MWH: The above is nonsensical. The answer is: We have an expresion
%well-formedness condition that restricts this syntax, among other things.
In addition,
conditionals have two syntactic forms: \code{EIf} is a normal
conditional, and \code{EIfDef} is one whose boolean guard is of the
form \estar{x}. By syntactically distinguishing these two
cases, the Coq model does not need the \emph{[prefer]} rule for
\code{if (*x)}$...$ forms as in
Fig.~\ref{fig:c-context}. The Redex model \emph{does} prioritize such
forms but not the same way as in the figure. It uses a variation of
the \textsc{S-Var} rule: The 
modified rule is equipped with a precondition that is false whenever
\textsc{S-IfNTT} is applicable.
% \mwh{Do we need
%   to talk about other semantic judgments that are not in the paper,
%   e.g., well-formedness? Maybe not, but should make sure the code is
%   commented to clarify that these things are not strictly needed --
%   well-formedness is subsumed by well typing (if that's true)?} \liyi{I will add comments to make sure it is clear in Coq.}
% \yiyun{No, well-formedness is not subsumed by the typing relation in
%   the Coq model. I
%   checked the latest coq code; they must be explicitly stated in the theorems.}
% \mwh{I mean: We might have it in the coq model, but I suspect it is
%   not necessary. If it is, we should be saying something in the main
%   paper.

The Coq model uses a runtime stack $\varphi$ as described at the start
of Sec.~\ref{sec:semantics}.
The Redex model introduces let bindings during evaluation to
simulate a runtime stack. For example, consider the expression 
$e \equiv \elet{x}{(\evalue{5}{\tint})}{\ebinop{x}{x}}$. Expression $e$ first steps to
$\elet{x}{(\evalue{5}{\tint})}{\ebinop{(\evalue{5}{\tint})}{x}}$, which in
turns steps to
$\elet{x}{(\evalue{5}{\tint})}{\ebinop{(\evalue{5}{\tint})}{(\evalue{5}{\tint})}}$. Since
the rhs of $x$ is a value, the let binding in $e$ effectively
functions as a stack that maps from $x$ to $\evalue{5}{\tint}$. The
let form remains in the expression and lazily replaces the variables
in its body. The let form can be removed from the expression only if its
body is evaluated to a value, e.g., 
$\elet{x}{(\evalue{5}{\tint})}{(\evalue{10}{\tint})}$
steps to 
$\evalue{10}{\tint}$. The rule for popping let bindings in this manner
corresponds to the \textsc{S-Ret} rule in Fig.~\ref{fig:semantics}.
Leveraging let bindings adds complexity to the semantics but
simplifies typing/consistency and term generation during randomized testing.
%  \mwh{I don't understand the last sentence. Nowhere in the
%   paper do we talk about ANF. Also: Isn't it the case that there are
%   rules for let bindings do not use substitution, but rather ``look
%   up'' the binding on the definition chain? E.g.,
%   \code{let x=5 in x+x} does not step to \code{5+5} but instead steps
%   to \code{let x=5 in 5+5} and then to \code{let x=5 in 10} ? }
% \liyi{Yiyun: please see here. If you do not know, please explain to me and then we can make a correct statement. }
% \yiyun{It is supposed to refer to the ANF transformation of the erased language, not the source
%   language. I moved it to the discussion of ANF form below in \ref{appx:comp1}. }

\subsection{Typing Rules for Literal Pointers}\label{sec:literal-pointer-typing}

The typing of integer literals, which can also be pointers to the
heap, was presented in Sec.~\ref{sec:theorem} in Fig.~\ref{fig:const-type}. Here
we describe these rules further.

The variable type rule
($\textsc{T-Var}$) simply checks if a given variable has the defined
type in $\Gamma$; the constant rule ($\textsc{T-Const}$) is slightly
more involved.  First, it ensures that the type annotation $\tau$ does
not contain any free variables. More importantly, it ensures that the
literal itself is well typed using an auxilliary typing relation
$\heap;\sigma \vdash n : \tau$.

If the literal's type is an integer, an unchecked pointer, or a null
pointer, it is well typed, as shown by the top three rules in
Fig.~\ref{fig:const-type}. However, if it is a checked pointer
$\tptr{\omega}{\cmode}$, we need to ensure that what it points to in
the heap is of the appropriate pointed-to type ($\omega$), and also
recursively ensure that any literal pointers reachable this way are
also well-typed. This is captured by the bottom rule in the figure,
which states that for every location $n+i$ in the pointers' range
%
$[n, n+\size(\omega))$, where $\size$ yields the size of its argument,
  then the value at the location $\heap(n+i)$ is also well-typed.
  However, as heap snapshots can contain cyclic structures (which
  would lead to infinite typing deriviations), we use a scope $\sigma$
  to assume that the original pointer is well-typed when checking the
  types of what it points to. The middle rule then accesses the scope
  to tie the knot and keep the derivation finite, just like in
  \citet{ruef18checkedc-incr}.

\subsection{Other Semantic Rules}\label{sec:rem-semantics}

Fig.~\ref{fig:rem-semantics} shows the remaining semantic rules for
$\lang$. We explain a selected few rules in this subsection.
% other few low-level semantic rules for variable and dereference and $\emalloctext$ operations in \checkedc. Other operations are defined in the same manner.

\begin{figure*}[t]
{\small
\begin{mathpar}

\inferrule [S-Var]{} {(\varphi,\heap,x)\longrightarrow (\varphi,\heap,\varphi(x))}

    \inferrule[S-DefArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}

    \inferrule[S-DefArrayBound]{0 \not\in [n_l,n_h)}
     { (\varphi,\heap,\estar{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}

    \inferrule[S-DefNTArrayBound]{0 \notin [n_l,n_h]}
    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\ebounds)}

    \inferrule[S-Assign]{\heap(n)=\evalue{n_a}{\tau_a} }
      {(\varphi,\heap,\eassign{\evalue{n}{\tptr{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}

    \inferrule[S-AssignNull]{}
      {(\varphi,\heap,\eassign{\evalue{0}{\tptr{\omega}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\enull)}

    \inferrule[S-AssignArrBound]{0 \not\in [n_l,n_h) }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\ebounds)}

  \inferrule[S-Malloc]{\varphi(\omega)=\omega_a \\ \mathtt{alloc}(\heap,\omega_a)=(n,\heap')}
   { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\evalue{n}{\tptr{\omega_a}{\cmode}})}

  \inferrule[S-MallocBound]{\varphi(\omega)=\tallarray{n_l}{n_h}{\tau}\\ (n_l \neq 0 \vee n_h \le 0)}
    { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\ebounds)}

    \inferrule[S-IfT]{n \neq 0 }
    {(\varphi,\heap,\eif{\evalue{n}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}

    \inferrule[S-IfF]{}
    {(\varphi,\heap,\eif{\evalue{0}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}

    \inferrule[S-Unchecked]{}
    {(\varphi,\heap,\eunchecked{\evalue{n}{\tau}} \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}

    \inferrule[S-Str]{
        0 \in [n_l,n_h]
\\ n_a \le n_h
\\ \heap(n+n_a) = 0 
\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tint})}

    \inferrule[S-StrBounds]{
        0 \notin [n_l,n_h]
}
    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}

    \inferrule[S-StrNull]{}
    {(\varphi,\heap,\estrlen{\evalue{0}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\enull)}

    \inferrule[S-Add]{n = n_1 + n_2}
    {(\varphi,\heap,\evalue{n_1}{\tint} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, n)}

    \inferrule[S-AddArr]{n = n_1 + n_2\\ n_l' = n_l - n_2 \\ n_h' = n_h - n_2}
    {(\varphi,\heap,\evalue{n_1}{\tallarrayptr{n_l}{n_h}{\tau}{m}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau}{m}})}

n    \inferrule[S-AddArrNull]{}
    {(\varphi,\heap,\evalue{0}{\tallarrayptr{n_l}{n_h}{\tau}{c}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \enull)}

\end{mathpar}

}
\caption{Remaining \lang Semantics Rules (extends Fig.~\ref{fig:semantics})}
\label{fig:rem-semantics}
\end{figure*}

Rule \textsc{S-Var} loads the value for $x$ in stack $\varphi$.
Rule \textsc{S-DefArray} dereferences an array pointer, which is similar to the Rule \textsc{S-DefNTArray} in Fig.~\ref{fig:semantics} (dealing with null-terminated array pointers).
The only difference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
Rules \textsc{DefArrayBound} and \textsc{DefNTArrayBound} describe an error case for a dereference operation.
If we are dereferencing an array/NT-array pointer and the mode is $\cmode$, $0$ must be in the range from $n_l$ to $n_h$ (meaning that the dereference is in-bound); if not, the system results in a $\ebounds$ error. Obviously, the dereference of an array/NT-array pointer also experiences a $\enull$ state transition if $n\le 0$.

Rules \textsc{S-Malloc} and \textsc{S-MallocBound} describe the $\emalloctext$ semantics. Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$, and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation. The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\cmode}}$ and new updated heap. It is possible for $\emalloctext$ to transition to a $\ebounds$ error if the $\omega_a$ is an array/NT-array type $\tallarray{n_l}{n_h}{\tau}$, and either $n_l \neq 0$ or $n_h \le 0$. This can happen when the bound variable is evaluated to a bound constant that is not desired.

\subsection{Subtyping for dependent types}
\label{app:le}

The subtyping relation given in Fig.~\ref{fig:checkc-subtype} involves
dependent bounds, i.e., bounds that may refer to variables. To decide
premises $b \leq b'$, we need a decision procedure that accounts for
the possible values of these variables. This process considers
$\Theta$, tracked by the typing judgment, and $\varphi$, the current
stack snapshot (when performing subtyping as part of the type
preservation proof).

\begin{defi}[Inequality]

\begin{itemize}

\item $n \le m$ if $n$ is less than or equal to $m$.
\item $x+n \le x + m$ if $n$ is less than or equal to $m$.
\item All other cases result in $\efalse$.

\end{itemize}
\end{defi}

To capture bound variables in dependent types, the \checkedc subtyping
relation ($\sqsubseteq$) is parameterized by a restricted stack
snapshot $\varphi|_{\rho}$ and the predicate map $\Theta$, where
$\varphi$ is a stack and $\rho$ is a set of
variables. $\varphi|_{\rho}$ means to restrict the domain of $\varphi$
to the variable set $\rho$. Clearly, we have the relation:
$\varphi|_{\rho} \subseteq \varphi$. $\sqsubseteq$
being parameterized by $\varphi|_{\rho}$ refers to that when we
compare two bounds $b \le b'$, we actually do
$\varphi|_{\rho}(b) \le \varphi|_{\rho}(b')$ by interpreting the
variables in $b$ and $b'$ with possible values in $\varphi|_{\rho}$.
Let's define a subset relation $\preceq$ for two restricted stack
snapshot $\varphi|_{\rho}$ and $\varphi'|_{\rho}$:

\begin{defi}[Subset of Stack Snapshots]
  Given two $\varphi|_{\rho}$ and $\varphi'|_{\rho}$,
  $\varphi|_{\rho} \preceq \varphi'|_{\rho}$, iff for $x\in\rho$ and
  $y$,
  $(x,y) \in \varphi|_{\rho} \Rightarrow (x,y) \in \varphi'|_{\rho}$.
\end{defi}

For every two restricted stack snapshots $\varphi|_{\rho}$ and
$\varphi'|_{\rho}$, such that
$\varphi|_{\rho} \preceq \varphi'|_{\rho}$, we have the following
theorem in \checkedc (proved in Coq):

\begin{thm}[Stack Snapshot Theorem]
  Given two types $\tau$ and $\tau'$, two restricted stack snapshots
  $\varphi|_{\rho}$ and $\varphi'|_{\rho}$, if
  $\varphi|_{\rho}\preceq \varphi'|_{\rho}$, and
  $\tau \sqsubseteq \tau'$ under the parameterization of
  $\varphi|_{\rho}$, then $\tau \sqsubseteq \tau'$ under the
  parameterization of $\varphi'|_{\rho}$.
\end{thm}

Clearly, for every $\varphi|_{\rho}$, we have
$\emptyset \preceq \varphi|_{\rho}$. The type checking stage is a
compile-time process, so $\varphi|_{\rho}$
is $\emptyset$ at the type checking stage. Stack snapshots are needed
for proving type preserving, as variables in bounds expressions are
evaluated away.

\begin{figure}[h!]
{\small
  \begin{mathpar}
    \inferrule[T-Def]
              {\Gamma;\Theta \vdash_m e : \tptr{\tau}{m'} \\
                m \leq m'}
              {\Gamma;\Theta \vdash_m \estar{e} : \tau}

    \inferrule[T-Mac]
              {}
              {\Gamma; \Theta \vdash_m \emalloc{\omega} : \tptr{\omega}{\cmode}}

    \inferrule[T-Add]
              {\Gamma; \Theta \vdash_m e_1 : \tint \\
                \Gamma; \Theta \vdash_m e_2 : \tint}
              {\Gamma; \Theta \vdash_m (e_1 \plus e_2) : \tint }

    \inferrule[T-Ind] 
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'} \\
                \Gamma; \Theta \vdash_m e_2 : \tint \\
                m \leq m'}              
              {\Gamma; \Theta \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}

    \inferrule[T-Assign]
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tau}{m'} \\
                \Gamma; \Theta \vdash_m e_2 : \tau' \\
                \tau'\sqsubseteq \tau \\
                m \leq m'}
              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}

   \inferrule[T-IndAssign]
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
                \Gamma; \Theta \vdash_m e_2 : \tint \\
                \Gamma; \Theta \vdash_m e_3 : \tau' \\
                \tau'\sqsubseteq \tau \\
                m \leq m'}
              {\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}

  \end{mathpar}
}
\caption{Remaining \lang Type Rules (extends Fig.~\ref{fig:type-system-1})}
\label{fig:rem-type-system}
\end{figure}

As mentioned in the main text, $\sqsubseteq$ is also parameterized by
$\Theta$, which provides the range of allowed values for a bound
variable; thus, more $\sqsubseteq$ relation is provable. For example,
in Fig.~\ref{fig:strcat-ex}, the \code{strlen} operation in line 4
turns the type of \code{dst} to be $\tntarrayptr{0}{x}{\tint}{\cmode}$
and extends the upper bound to \code{x}. In the \code{strlen} type
rule, it also inserts a predicate \code{x}$\ge 0$ in $\Theta$; thus,
the cast operation in line 16 is valid because
$\tntarrayptr{0}{x}{\tint}{\cmode} \sqsubseteq
\tntarrayptr{0}{0}{\tint}{\cmode}$ is provable when we know
\code{x}$\ge 0$.

Note that if $\varphi$ and $\Theta$ are $\emptyset$, we do only the
syntactic $\le$ comparison; otherwise, we apply $\varphi$ to both
sides of $\sqsubseteq$, and then determine the $\le$ comparasion based
on a Boolean predicate decision procedure on top of $\Theta$. This
process allows us to type check both an input expression and the
intermediate expression after evaluating an expression.

\subsection{Other Type Rules}\label{rem-type}

Here we show the type rules for other \checkedc operations in Fig.~\ref{fig:rem-type-system}.
Rule \textsc{T-Def} is for dereferencing a non-array pointer. 
The statement $m \leq m'$ ensures that no unchecked pointers are used in checked regions.
Rule \textsc{T-Mac} deals with
$\emalloctext$ operations. There is a well-formedness check to require
that the possible bound variables in $\omega$ must be in the domain of
$\Gamma$ (see Fig.~\ref{fig:wftypesandbounds}). This is similar to the well-formedness assumption of the type environment (Definition~\ref{type-wellformed}) Rule \textsc{T-Add} deals with binary operations whose sub-terms are integer expressions, while rule \textsc{T-Ind} serves the case for pointer arithmetic. For simplicity, in the \checkedc formalization, we do not allow arbitrary pointer arithmetic. The only pointer arithmetic operations allowed are the forms shown in rules \textsc{T-Ind} and \textsc{T-IndAssign} in Fig.~\ref{fig:rem-type-system}. Rule \textsc{T-Assign} assigns a value to a non-array pointer location. The predicate $\tau'\sqsubseteq \tau$ requires that the value being assigned is a subtype of the pointer type.
The \textsc{T-IndAssign} rule is an extended assignment operation for handling assignments for array/NT-array pointers with pointer arithmetic. Rule \textsc{T-Unchecked} type checks \code{unchecked} blocks.

\subsection{Struct Pointers}\label{appx:struct}

\checkedc has \kw{struct} types and \kw{struct} pointers. Fig.~\ref{fig:checkc-struct} contains the syntax of \kw{struct} types as well as new subtyping relations built on the \kw{struct} values.
For a \kw{struct} typed value, \checkedc has a special operation for it, which is $\eamper{e}{f}$. This operation indexes the $f$-th position \kw{struct}~$T$ item, if the expression $e$ is evaluated to a \kw{struct} pointer $\tptr{\tstruct{T}}{m}$. Rule \textsc{T-Struct} in Fig.~\ref{fig:checkc-struct} describes its typing behavior.
Rules \textsc{S-StructChecked} and \textsc{S-StructUnChecked} describe the semantic behaviors of $\eamper{e}{f}$ on a given \kw{struct} \code{checked}/\code{unchecked} pointers, while rule \textsc{S-StructNull} describes a \code{checked} \kw{struct} null-pointer case.
In our Coq/Redex formalization, we include the \kw{struct} values and the operation $\eamper{e}{f}$. We omit it in the main text due to the paper length limitation.

\begin{figure}
{\small
$\begin{array}{l}
\text{  Struct Syntax: }\\[0.5em]
  \begin{array}{ll}
 \mathtt{Type} & \tstruct{T}
\\[0.2em]
     \text{Structdefs} & D \; \in \; T \rightharpoonup fs \\[0.2em]
      \text{Fields} & fs \; ::= \; \tau~\mathtt{f} \mid \tau~\mathtt{f}; fs 
    \end{array}\\[2em]
\text{  Struct Subtype: }\\[0.5em]
\begin{array}{l}
    D(T) = fs \wedge fs(0) = \tnat \Rightarrow  \tptr{\tstruct{T}}{m} \sqsubseteq \tptr{\tnat}{m}\\[0.5em]
    D(T) = fs \wedge fs(0) = \tnat \wedge 0 \le b_l \wedge b_h \le 1 \\[0.2em]
 \qquad\qquad \Rightarrow 
       \tptr{\tstruct{T}}{m} \sqsubseteq \tarrayptr{b_l}{b_h}{\tnat}{m}
    \end{array}
\\[3em]
\text{  Struct Type Rule: }\\[0.5em]
    \end{array}
$

{
\begin{mathpar}
  \inferrule [T-Struct]
  {\Gamma; \Theta \vdash_m e :  {\tptr{\tstruct{T}}{m}} \\
    D(T) = fs \\ fs(f)=\tau_f}
  {\Gamma; \Theta \vdash_m \eamper{e}{f} : \tptr{\tau_f}{m}}

\end{mathpar}
}
{
$\begin{array}{l}
\text{Struct Semantics: }
\end{array}
$
}
{
\begin{mathpar}
  \inferrule [S-StructChecked]
  {n > 0 \\ D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\cmode}})}

  \inferrule [S-StructNull]
  {n = 0}
  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\enull)}

  \inferrule [S-StructUnChecked]
  {D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\umode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\umode}})}

\end{mathpar}
}
}
  \caption{\lang Struct Definitions}
  \label{fig:checkc-struct}
\end{figure}

\begin{figure}[t]
{\small
  \begin{mathpar}

    \inferrule[]
    {}
    {\Gamma \vdash n}

    \inferrule[]
    {x:\tint \in \Gamma}
    {\Gamma \vdash x + n}

    \inferrule[]
    {\Gamma \vdash b_l\\
    \Gamma \vdash b_h}
  {\Gamma \vdash (b_l,b_h)}

  \inferrule[]
  {}
  {\Gamma \vdash \tint}

  \inferrule[]
  {\Gamma \vdash \bvar \\
  \Gamma \vdash \tau}
  {\Gamma \vdash \tptr{\tallarrayb{\bvar}{\tau}}{m}}

  \inferrule[]
  {\Gamma \vdash \tau}
  {\Gamma \vdash \tptr{\tau}{m}}

  \inferrule[]
  {T \in D}
  {\Gamma \vdash \tptr{\tstruct{T}}{m}}
  \end{mathpar}
}
 \caption{Well-formedness for Types and Bounds}
\label{fig:wftypesandbounds}
\end{figure}

% \inferrule[T-Fun]
%     {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\\\
%         \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
%              \overline{\tau'} \sqsubseteq 
%                \overline{\tau}[\overline{e} / \overline{x}]}
%     {\Gamma; \Theta \vdash_m f(\overline{e}) : \tau[\overline{e} / \overline{x}]}

\begin{figure}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    {\Gamma \vdash \overline{x}:\overline{\tau} \\
      \Gamma[\overline{x} \mapsto \overline{\tau}] \vdash \tau \\
    \Gamma[\overline{x} \mapsto \overline{\tau}]; \Theta  \vdash_{\cmode} e : \tau}
    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e}

    \inferrule[]
    {}
    {\Gamma \vdash \cdot}

    \inferrule[]
    {\Gamma \vdash \tau \\
    \Gamma[x \mapsto \tau] \vdash \overline{x}:\overline{\tau}}
    {\Gamma \vdash x:\tau, \overline{x}:\overline{\tau}}
  \end{mathpar}
}
 \caption{Well-formedness for functions}
\label{fig:wffunctions}
\end{figure}

\begin{figure}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    {\Gamma \vdash \tau}
    {\Gamma \vdash \tau~\mathtt{f} }

    \inferrule[]
    {\Gamma \vdash \tau\\
     \Gamma \vdash fs}
    {\Gamma \vdash \tau~\mathtt{f}; fs }

  \end{mathpar}
}
 \caption{Well-formedness for structs}
\label{fig:wfstructs}

{\small
  \begin{mathpar}

    \inferrule[]
    {\Gamma[\overline{x} \mapsto \overline{\tau}]; \emptyset  \vdash
      e \gg \dot e : \tau}
    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e \gg
      (\overline{x})\;\dot e}
  \end{mathpar}
}
 \caption{Compilation Rules for Functions}
\label{fig:compilefunctions}
\end{figure}

\subsection{The Compilation Rules}\label{appx:comp1}
Fig.~\ref{fig:syntaxerased} and Fig.~\ref{fig:semanticserased} shows
the syntax for $\elang$, the target language for compilation. We syntactically
restrict the expressions to be in A-normal form to simplify the
presentation of the compilation rules. In the Redex model, we
occasionally break this constraint to speed up the performance of
random testing by removing unnecessary let bindings.
To allow explicit runtime
checks, we include $\ebounds$ and $\enull$ as part of $\elang$
expressions which, once evaluated, result in an corresponding error state. $\eassignstack{x}{\dot a}$ is a new syntactic form that
modifies the stack variable $x$ with the result of $\dot a$. It is
essential for bounds widening. $\leq$ and $-$ are introduced
to operate on bounds and decide whether we need to halt with a bounds error
or widen a null-terminated string.

\begin{figure}[h]
{\small
  \[\begin{array}{llcl}
      \text{Atoms}       & \dot a & ::= & n \mid x \\

      \text{C-Expressions} & \dot c & ::= & \dot a \mid \estrlen{\dot a}  \mid \emalloc {\dot  a}  \mid \ecall{f} {\overline{\dot a}}  \\
                         & & \mid &  \mid  \dot a \circ \dot a \mid \estar{\dot a} \\
                         & & \mid & \eassign{\dot a}{\dot a} \mid \eassignstack{x}{\dot a}  \mid \eif{\dot a}{\dot e}{\dot e} \\
                         & & \mid & \ebounds \mid \enull \\
      \text{Expressions}  & \dot e & ::= & \dot c \mid \elet{x}{\dot c}{\dot e}\\
      \text{Binops} & \circ & ::= & + \mid - \mid \leq  \\
      \text{Closure} & \dot C & ::= & \hole \mid \elet{x}{\dot a}{\dot C} \\
       & & \mid & \eif{\dot a}{\dot e}{\dot C} \mid \eif{\dot a}{\dot    C}{\dot e} \\
      \text{Bounds Map} & \rho & \in & \texttt{Var} \rightharpoonup \texttt{Var} \times \texttt{Var}
  \end{array}
  \]
}
  \caption{\elang{} Syntax}
  \label{fig:syntaxerased}
{\small
  \[\begin{array}{lll}
    \dot \mu & ::= & n \mid \bot \\
    \dot c & ::= & \ldots  \mid \eret{x}{\dot \mu}{\dot e} \\
    \dot H & \in & \mathbb{Z} \rightharpoonup \mathbb{Z} \\
    \dot r & ::= & \dot e \mid \enull \mid \ebounds \\
    \dot E & ::= & \hole \mid  \elet{x}{\dot E}{\dot e}  \mid \eret{x}{i}{\dot E} \\
             & \mid & \eif{\dot E}{\dot e}{\dot e} \mid \estrlen{\dot E} \\
             & \mid & \emalloc{\dot E} \mid \ecall{f}{\overline{\dot
                      E}} \mid \dot E \circ \dot a \mid n \circ \dot E \\
             & \mid & \estar{\dot E} \mid \eassign{\dot E}{\dot a}
                      \mid \eassign{n}{\dot E} \mid
                      \eassignstack{x}{\dot E} \\
      \overline{\dot E } & ::= & \dot E \mid n,\overline{\dot E} \mid
                                 \overline{\dot E}, \dot a

  \end{array} \]
}
  \caption{\elang Semantic Defs}
  \label{fig:semanticserased}
\end{figure}

$\elang$ does not include any annotations. We remove structs from
$\elang$ because we can always statically convert expressions of the form
$\eamper{n:\tau}{f}$ into $n + n_f$, where $n_f$ is the statically
determined offset of $f$ within the struct. We ellide the semantics of
$\elang$ because it is self-evident and mirrors the semantics
$\lang$. The difference is that in $\elang$, only $\ebounds$ and
$\enull$ can step into an error state. All failed dereferences and
assignments would result in a stuck state and therefore we rely on the
compiler to explicitly insert checks for checked pointers.

Fig.~\ref{fig:compilation} and Fig.~\ref{fig:compilation2}
 shows the rules for the compilation judgment for expressions,
\[\Gamma;\rho \vdash e \gg \dot C, \dot a\]
The judgment is presented differently from the one in
Sec.~\ref{sec:compilation}, which was simplified for presentation purposes. First,
we remove $\Theta$ and $m$ because these
parameters are only used for checking and have no
impact on compilation. Second, the judgment includes two
outputs, a closure $\dot C$ and an atom expression $\dot a$, instead of
a single $\elang$ expression $\dot e$. $\dot C$ can be intuitively understood as a
partially constructed program or context. Whereas $\dot E$ is used for
evaluation, $\dot C$
is used purely as a device for compilation. As an example,
when compiling $(1 : \tint) \plus (2 : \tint)$, 
we would first create a fresh variable $x$, and then produce two outputs:
\[    \dot C = \elet{x}{1 \plus 2}{\hole}\]
\[    \dot a = x\]
To obtain the compiled expression $\dot e$, we plug $\dot a$ into
$\dot C$ using the usual notation $\dot C[\dot a]$. We can also use $\dot C$ to represent runtime checks, which
usually take the form $\elet{x}{\dot c}{\hole}$, where $\dot c$
contains the check whose evaluation must not trigger $\ebounds$ or $\enull$ for the program to
continue (see Fig.~\ref{fig:compilationhelpers} for the metafunctions that create
those checks).

This unconventional output format enables us to separate the
evaluation of the term and the computation that relies on the term's
evaluated result. Since effects and reduction (except for variables) happen only
within closures, we can precisely control the order in which effects
and evaluation happen by composing the contexts in a specific order.
Given two closures $\dot C_1$
and $\dot C_2$, we write $\dot C_1[\dot C_2]$ to denote the meta
operation of plugging $\dot C_2$ into $\dot C_1$. We also use
$\dot C_{a;b;c}$ as a shorthand for $\dot C_a[\dot C_b[\dot C_c]]$. In
the \textsc{C-Ind} rule,
we first evaluate the expressions that correspond to $e_1$ and $e_2$
through $\dot C_1$ and $\dot C_2$, and then perform a null check and
an addition through $\dot C_n$ and $\dot C_3$. Finally, we dereference
the result through $\dot C_4$ before returning the pair $\dot C_4,\dot
x_4$, propagating the flexibility to the compilation rule that recursively
calls \textsc{C-Ind}.

Fig.~\ref{fig:compilationhelpers} shows the metafunctions that create
closures representing dynamic checks. These functions first examine
whether the pointer is a checked. If the pointer is unchecked, an empty closure $\hole$
will be returned, because there is no need to perform a check. For bounds
checking, there is a special case for NT-array pointers, where the
bounds are retrived from the \textcolor{purple}{shadow} variables (found by
looking up $\rho$) on the stack  rather than
using the bounds specified in the type annotation. This is how we
achieve the same precise runtime behavior as $\lang$ in our compiled expressions.

Fig.~\ref{fig:compilationhelpers2} shows the metafunctions related to
bounds widening. $\vdash_{extend}$ takes $\rho$,
a checked NT-array pointer variable $x$, and its bounds $(b_l,b_h)$ as inputs, and returns an
extended $\rho'$ that maps $x$ to two fresh variables $x_l$, $x_h$,
together with a closure $\dot C$ that initializes $x_l$ and $x_h$ to
$b_l$ and $b_h$ respectively. This function is used in the
\textsc{C-Let} rule to extend $\rho$ before compiling the body of the
$\elettext$ binding. The updated $\rho'$ can be used for generating
precise bounds checks, and for inserting expressions that can potentially
widen the upper bounds, as seen in the $\vdash_{widenstr}$
metafunction used in the \textsc{C-Str} compilation rule.

\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    { x_l, x_h = \fresh \\
      \rho' = \rho[x \mapsto (x_l, x_h)]\\
    \dot C = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}}
  {\cextend{\rho}{x}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot C, \rho'}\\}

  \inferrule[]
  {x_l,x_h = \rho(x) \\
    x_w = \fresh \\
  \dot C = \elet{x_w}{\eif{x_h}{0}{\eassignstack{x_h}{1}}}{\hole}}
{\ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C}}

\inferrule[]
{e \notin dom(\rho)}
{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{m}}{\hole}}

\inferrule[]
{x_l,x_h = \rho(e) \\
  x_a = \fresh \\
  \dot C = \elet{x_a}{\eif{\dot a \leq x_h}{0}{\eassignstack{x_h}{\dot a}}}{\hole}}
{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{\cmode}}{\dot C}}

  \end{mathpar}

}
 \caption{Metafunctions for Widening}
\label{fig:compilationhelpers2}
\end{figure*}

\begin{figure*}[h]
{\small
  \begin{mathpar}
    \inferrule[]
    {x = \fresh \\
      \dot C = \elet{x}{\eif{\dot a}{0}{\enull}}{\hole}}
    {\echecknull{\dot a}{c}{\dot C}}

    \inferrule[]
    {}
    {\echecknull{\dot a}{u}{\hole}} \\

    \inferrule[]
    {}
    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}

    \inferrule[]
    { x_l,x_h = \rho(e) \\
      x_{cl},x_{ch} = \fresh\\
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

    \inferrule[]
    {}
    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}

    \inferrule[]
    { x_l,x_h = \rho(e) \\
      x_{cl},x_{ch} = \fresh\\
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
    }
    {\echeckboundsw{\rho}{e}{\tallarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

\inferrule[]
{e \notin dom(\rho) \\
  x_l,x_l',x_h,x_h' = \fresh \\
  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
  \dot C_2 = \elet{x_l'}{b_l'}{\elet{x_h'}{b_h'}{\hole}}\\
  \dot C_3 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
  \dot C_4 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
}
{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3;4}}}

\inferrule[]
{x_l',x_h' =\rho(e) \\
  x_l,x_h = \fresh \\
  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
  \dot C_2 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
  \dot C_3 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
}
{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3}}}
  \end{mathpar}
}
\caption{Metafunctions for Dynamic Checks}
\label{fig:compilationhelpers}
\end{figure*}

% the closure $\dot
% C_{1;2;n;3;b;4}$ encodes the idea that $e_1$ and $e_2$ are evaluated to values
% before a null check (included in $\dot C_n$) is performed on
% $\dot a_1$ , the result of evaluating $e_1$. $\dot C_3$ adds $\dot
% a_1$ and $\dot a_2$ before a bounds check (included in $\dot C_b$) is performed on $\dot a_1$
% with $\dot a_2$ as the index. Finally, we use $\dot C_4$ 
% to store the result of the dereference in $x_4$ and return the pair
% $\dot C_4,x_4$ as the result of the compilation. 
% Further checks and
% widening can be performed using $x_4$ and we don't have to worry about
% the effects being triggered multiple times because the evaluation only
% happens once in $\dot C_4$.

% insert runtime checks and bounds widening
% code in a straightforward way. Take the \textbf{C-Deref} rule as an example,
% if we were to output $\dot e$ directly, we would have to rebind the
% expression to a variable before we can perform the null check and
% bounds check since $\dot e$ is not always a pure expression. Adding
% $\elettext$ bindings in such an ad-hoc fashion for avoiding side
% effects can clutter the compilation rules. Our approach of
% outputting a context and an atom threads the ANF transformation into
% each individual rule. Fig.~\ref{fig:compilationhelpers} shows helper
% functions that take atom expressions and return a context that
% contains the checks. The advantage of using explicit contexts becomes
% evident when we start to compose them. For the simulation property to
% hold, we want to ensure that the effects happen in the exact same
% order.  In
% the C-Deref rule, we first evaluate the expression that corresponds to
% $e$ through $\dot C_1$, and then perform a bounds check followed by a
% null check through $\dot C_b$ and $\dot C_n$. Finally, we dereference
% the expression through $\dot C_2$, potentially widen the upper bound
% by checking the dereferenced
% result.

% One big difference is that
% our compilation strategy is not parameterized by the heap. Given a
% literal annotated with a type, the compilation judgment simply
% discards the type without validating whether the literal's type is
% valid. This kind of relaxation not only makes the compilation rules
% simpler (we no longer need to keep track of $\defscope$ and the mode),
% but also allows us to formulate a stronger simulation theorem that
% applies even after the program state is corrupted. More concretely,
% the compilation judgment satisfies preservation regardless of whether
% an unchecked step is being taken. On the other hand, if we were to
% embed the compilation into the full-fledged typechecking rules, we
% would not always be able to compile the expression after a step has
% been taken.

% When our source
% program is stuck because of the execution of unchecked code, we can
% conclude that the compiled program will also get stuck with the
% same erased heap state. It is possible to embed our compilation
% strategy directly into the typing judgment, but our simulation would
% be limited to states that are not yet corrupted to be untypable.

% Since the erased language does not have type annotations, we need
% to explicitly represent bounds of null-terminated arrays as stack
% variables to support bounds widening at runtime. Every time a checked
% null-terminated array pointer is bound to a variable, we generate two
% fresh new variables that are assigned the lower and upper bounds of
% the array. The upper bound can be mutated into
% a larger bound during execution. We use $\rho$ to keep track of the
% bounds variables that belong to null-terminated array variables.

% The compilation judgment does not produce expressions
% directly. Instead, it outputs a tuple of a context $\dot C$ and an atom
% expression $\dot a$ such that $\dot C[\dot a]$ corresponds to the
% compiled target expression. In the case where the atom is a variable, $\dot
% C$ will contain a let binding that defines the variable. This unconventional
% output format enables us to insert runtime checks and bounds widening
% code in a straightforward way. Take the C-Deref rule as an example,
% if we were to output $\dot e$ directly, we would have to bind the
% expression to a variable before we can perform the null check and
% bounds check since $\dot e$ is not always a pure expression. Adding
% $\kw{let}$ bindings in such an ad-hoc fashion for avoiding side
% effects can clutter the compilation rules. Our approach of
% outputting a context and an atom threads the ANF transformation into
% each individual rule. Fig.~\ref{fig:compilationhelpers} shows helper
% functions that take atom expressions and return a context that
% contains the checks. The advantage of using explicit contexts becomes
% evident when we start to compose them. For the simulation property to
% hold, we want to ensure that the effects happen in the exact same
% order. Since effects and reduction (except for variables?) happen only
% within contexts, we can precisely control the order in which effects
% and evaluation happen by composing the contexts in a specific order. In
% the C-Deref rule, we first evaluate the expression that corresponds to
% $e$ through $\dot C_1$, and then perform a bounds check followed by a
% null check through $\dot C_b$ and $\dot C_n$. Finally, we dereference
% the expression through $\dot C_2$, potentially widen the upper bound
% by checking the dereferenced
% result.

\begin{figure*}[t]
{\small
  \begin{mathpar}

    \inferrule[C-Const]
      {}
      {\Gamma;\rho\vdash \evalue{n}{\tau} \gg \hole, n : \tau}

    \inferrule[C-Var]
      {x : \tau \in \Gamma}
      {\Gamma;\rho \vdash x \gg \hole, x : \tau}

    \inferrule[C-Cast]
              {\Gamma;\rho \vdash e \gg  \dot C , \dot a : \tau'}
              {\Gamma;\rho \vdash \ecast{\tau}{e} \gg \dot C, \dot a  : \tau}

      \inferrule[C-DynCast]
      {\Gamma;\rho \vdash e \gg \dot C_1, \dot a : \tptr{\tallarrayb{\bvar'}{\tau}}{m} \\
      \echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\tptr{\tallarrayb{\bvar'}{\tau}}{m}}{\dot C_{b}}}
      {\Gamma;\rho \vdash \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e}
            \gg \dot C_{1;b}, \dot a : \tptr{\tallarrayb{\bvar}{\tau}}{m}}

      \inferrule[C-Str]
      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} \\
        \echecknull{\dot a_1}{m}{\dot C_{n}} \\
        \echeckbounds{\rho}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{0}{\dot C_{b}} \\
        x_2 = \fresh \\
          \dot C_2 = \elet{x_2}{\estrlen{\dot a_1}}{\hole} \\
      \ewidenstrlen{\rho}{e}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{\dot C_{w}} } 
                {\Gamma;\rho \vdash \estrlen{e} \gg \dot C_{1;n;b;2;w}, x_2 : \tint}

      \inferrule[C-LetStr]
      {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\
        \Gamma; \rho \vdash \estrlen{y} \gg \dot C_1, \dot a_1 : \tint\\
        \dot C_2 = \elet{x}{\dot a_1}{\hole} \\
        \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}]];\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau}
      %   x_3 = \fresh \\
      %   \dot C_3 = \elet{x_3}{\dot a_2}{\hole}}
                {\Gamma;\rho \vdash \elet{x}{\estrlen{y}}{e} \gg \dot C_{1;2;3},\dot a_3 : \tau}

      \inferrule[C-If ]
      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
        \Gamma; \rho \vdash e_1 \gg \dot C_2, \dot a_2 : \tau_2 \\
        \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
        x_4 = \fresh \\
        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_2[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}
      }
        {\Gamma; \rho \vdash \eif{e_1}{e_2}{e_3} \gg \dot C_{1;4}, x_4 : \tau_2 \sqcup \tau_3}

        \inferrule[C-IfNT]
        {\Gamma; \rho \vdash x : \tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c} \\
          b_h = 0 \Rightarrow \Gamma' = \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}] \\
          b_h \neq 0 \Rightarrow \Gamma' = \Gamma \\
          \Gamma; \rho \vdash \estar{x} \gg \dot C_1, \dot a_1 : \tau_1\\
          \Gamma'; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau_2\\
          \Gamma;\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
        \ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C_{w}} \\
        x_4 = \fresh \\
        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_{2;w}[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}}
        {\Gamma;\rho \vdash \eif{\estar{x}}{e_1}{e_2} \gg \dot C_{1;4}, x_4 : \tau_1 \sqcup \tau_2}

   \inferrule[C-Let]
    { (x\in \fv(\tau') \Rightarrow e_1 \in Bound) \\
        \Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tau_1 \\ 
        \cextend{\rho}{x}{\tau_1}{\dot C_2,\rho'}\\
        \dot C_3 = \elet{x}{\dot a_1}{\hole}\\
        \Gamma[x\mapsto \tau];\rho' \vdash e_4 \gg \dot C_4, \dot a_4  : \tau_4}
    {\Gamma;\rho' \vdash \elet{x}{e_1}{e_4} \gg \dot C_{1;2;3;4},\dot a_4 : \tau_4[\tau_1 = \tint \Rightarrow x \mapsto e_1]}

  \inferrule[C-Ret]
    {\Gamma(x)\neq \bot \\
      \Gamma;\rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
      x_2 = \fresh \\
      \mu \gg \dot \mu \\
    \dot C_2 = \elet{x_2}{\eret{x}{\dot \mu}{\dot C_1[\dot a_1]}}{\hole}}
  {\Gamma;\rho \vdash \eret{x}{\mu}{e} \gg \dot C_2, x_2 : \tau}

    \inferrule[C-Fun]
    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\
      % \dot \Xi(f) = (\overline{x})\; \dot C_f; \dot a_f \\
      (\forall e_i\in\overline{e}\;\;\tau_i\in\overline{\tau}\;.\; \Gamma; \rho \vdash {e_i} \gg \dot C_i, \dot a_i  : \\ \tau_i' \wedge
      \tau_i' \sqsubseteq 
      \tau_i[\overline{e} / \overline{x}]%  \\
      ) \\
      x_f = \fresh \\
      \dot C_f = \elet{x_f}{f(\overline{a})}{\hole}}
    {\Gamma; \rho \vdash f(\overline{e}) \gg \overline{\dot C}[\dot C_f] , x_f  : \tau[\overline{e} / \overline{x}]}

  \inferrule[C-Def]
  {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m} \\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    x_2 = \fresh \\
    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}
   }
   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;2}, x_2 : \tau}

   \inferrule[C-DefArr]
   {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tallarrayptr{b_l}{b_h}{\tau}{m} \\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    \echeckbounds{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
    x_2 = \fresh \\
    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}}
   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;b;2},x_2 : \tau}

    \inferrule[C-Mac]
    {\dot C_{1}, \dot a_1 = \esizeof{\omega} \\
      x_2 = \fresh \\
    \dot C_2 = \elet{x_2}{\emalloc{\dot a_1}}{\hole}}
     {\Gamma; \rho \vdash \emalloc{\omega} \gg \dot C_{1;2}, x_2 : \tptr{\omega}{\cmode}}

  \end{mathpar}
}
\caption{Compilation}
\label{fig:compilation}
\end{figure*}

\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[C-Add]
              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tint \\
                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint\\
                x_3 = \fresh \\
              \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}}
              {\Gamma; \rho \vdash \dot C_3, x_3 : \tint }

   \inferrule[C-Ind] 
              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
                \echecknull{\dot a_1}{m}{\dot C_{n}}\\
                \echeckbounds{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
                x_3, x_4 = \fresh\\
                \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}\\
                \dot C_4 = \elet{x_4}{\estar{x_3}}{\hole}
                }
                {\Gamma; \rho \vdash \estar{(\ebinop{e_1}{e_2})} \gg \dot C_{1;2;n;3;b;4} , x_4: \tau}

  \inferrule[C-Assign]
  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m'} \\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
    \tau'\sqsubseteq \tau \\
    x_3 = \fresh \\
    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
  }
  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;3}, x_3 : \tau}

  \inferrule[C-AssignArr]
  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    \echeckboundsw{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
    x_3 = \fresh \\
    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
    \tau'\sqsubseteq \tau}
  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;b;3},x_3 : \tau}

    \inferrule[C-IndAssign]
    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
      \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
      \echeckboundsw{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
      \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau' \\
      x_4, x_5 = \fresh \\
      \dot C_4 = \elet{x_4}{\dot a_1 \plus \dot a_2}{\hole}\\
      \dot C_5 = \elet{x_5}{\eassign{x_4}{x_3}{\hole}}
      \tau'\sqsubseteq \tau}
    {\Gamma; \rho \vdash \eassign{(e_1 \plus e_2)}{e_3} \gg \dot C_{1;2;n;3;4;b;5} : \tau}

    \inferrule [C-Struct]
    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1  :  {\tptr{\tstruct{T}}{m}} \\
      D(T) = \tau_0~f_0\ldots; \tau_j~f; ...  \\
      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
      x_2 = \fresh\\
    \dot C_2 = \elet{x_2}{\dot a_1 \plus \dot j}{\hole}}
    {\Gamma; \rho \vdash \eamper{e_1}{f} \gg \dot C_2, x_2  : \tptr{\tau_f}{m}}

    \inferrule[C-Unchecked]
              {\Gamma;\rho \vdash e \gg \dot C, \dot a : \tau}
              {\Gamma;\rho \vdash \eunchecked{e} \gg \dot C, \dot a : \tau}

  \end{mathpar}
    % \begin{prooftree}
    %   \hypo{
    %     \Gamma; \rho \vdash_u e \gg \dot C, a : \tau
    %   }
    %   \infer1[C-Unchecked]{\Gamma; \rho \vdash \eunchecked{e} \gg \dot C, a : \tau}
    % \end{prooftree}\\\\

    % \begin{prooftree}
    %   \hypo{
    %     \begin{matrix}
    %         \Gamma; \rho \vdash e \gg \dot C_1, a_1 :  {\tptr[m]{\tstruct{T}}} \\
    %         % \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
    %         D(T) = \tau_0\ldots; \tau_j~f; ...\\
    %         x_2 = \fresh \\
    %         \dot C_2 = \elet{x_2}{a_1 + j}{\hole} \\
    %         \echecknull{a_1}{m}{\dot C_{null}}
    %     \end{matrix}
    %   }
    %   \infer1[C-Amper]{\Gamma; \rho \vdash \eamper{e}{f} \gg \dot C_{1;null;2}, x_2  : \tptr[m]{\tau_f}}
    % \end{prooftree} \\ \\

    % \begin{prooftree}
    %   \hypo{
    %     \begin{matrix}
    %       \Gamma; \rho \vdash e_1 \gg \dot C_1, a_1 : \tptr[m']{\omega} \\
    %       \Gamma; \rho \vdash e_2 \gg \dot C_2, a_2 : \tau \\
    %       m' = u \Rightarrow m = u \\
    %       \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}\\
    %       \echecknull{a_1}{m}{\dot C_{n}} \\
    %       % \echeckbounds{\rho}{a_1}{\tptr[m]{(\tntarray{le}{he}{\tau})}}{\dot C_{b}} \\
    %       x_3 = \fresh\\
    %       \dot C_3 = \elet{x_3}{\eassign{a_1}{a_2}}{\hole} \\
    %     \end{matrix}
    %   }
    %   \infer1[C-Assign]{\Gamma; \rho \vdash \dot C_{1;b;n;2;3}, x_3 : \tau}
    % \end{prooftree}\\\\

    % \begin{prooftree}
    %   \hypo{
    %     \begin{matrix}
    %       \Gamma; \rho \vdash e_1 : \tptr[m']{\omega} \\
    %       \Gamma; \rho \vdash e_2 : \tint \\
    %       \Gamma; \rho \vdash e_3 : \tau \\
    %       m' = u \Rightarrow m = u \\
    %       \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
    %     \end{matrix}
    %   }
    %   \infer1[C-IndAssign]{\Gamma; \rho \vdash \eassign{(e_1 + e_2)}{e_3} : \tau}
    % \end{prooftree}
}
\caption{Compilation (Continued)}
\label{fig:compilation2}
\end{figure*}

### 2022-06-25 sandp23/checkedc-formal/compilation.tex
update paper.

\section{Compilation}\label{sec:compilation}

% \review{
% The first reviewer said:
% \begin{itemize}
% \item I had a really hard time understanding precisely the "formalized" compilation
%   from CoreChkC to CoreC. Specifically: is CoreC intended to model LLVM-IR, or
%   is it a subset of C? Is CheckedC compiling to C or is it a new frontend like
%   `clang`? See remark about undefined behavior (UB) below which got me even more
%   confused. \mwh{CoreC is meant to model CoreChkC but with
%   annotations, and checks removed; it is not meant to model C or LLVM-IR.}
% \item Is the compilation scheme from CoreChkC to CoreC is faithful to the actual
%   compilation scheme of the checkedc-clang compiler to ...? I feel like the
%   paper is missing an actual description of what happens in the compiler to
%   allow us to connect the dots and understand how the formalization illuminates
%   the implementation. \mwh{We did not model compilation on the real
%   implementation; our purpose was to show that annotations in CoreChkC
%   do not necessitate fat pointers in an implementation; that said,
%   our formalization does show how a real implementation can be
%   carried out}
% \item The paper doesn't seem upfront about what is *shown* (theorem in Coq) and what
%   is *tested* (via PLT-Redex), and thus remains a
%   conjecture. \mwh{Updated the intro and the individual sections}
% \item  Missing discussion of why Coq vs. PLT-Redex, effort involved, any plans to
%   formally prove compilation from CoreChkC to CoreC, any hopes of integrating
%   that in the official implementation, etc. \mwh{Added note at the
%   start of section 3.}
% \end{itemize} }

% \review{IV: ghost variables in other contexts (e.g. Why3, Dafny) are used for things
%   that do not exist at run-time, but this doesn't seem to be the case here.}
% \yiyun{Agreed: We changed the name to ``shadow variables'' to avoid confusion}
% \review{Do your bug report and github links break anonymity?}

% \review{From reviewer C:
% \begin{itemize}
% \item I'd like to have seen a bit more motivation for using PLT redex:
%   what aspects made the use of this tool preferable to formulating the
%   compilation in Coq and using Quickchick to do random testing of the
%   simulation result.
%   \mwh{Added some text to the end of III.A and IV.C}
% \yiyun{Some reasons I can think of:\begin{itemize}
%     \item Redex is highly optimized for specifying judgments that are algorithmic. By writing down a typing relation, we can immediately obtain a typechecker
%     that is executable. Same applies for the small-step evaluation relation. Translating the relations into functions in Coq is definitely doable but time-consuming, especially since compilation is embedded as part of the typing rules. It is also hard to see whether the function we define really corresponds to the relation unless we formally prove it. This issue is particularly relevant at the early stage of the development when the compilation rules were buggy and the simulation property was violated often as we added new generator cases.
%     In Redex, we don't have any formal guarantee either, but at least we can more easily see the correspondence because Redex is able to convert the relation into an executable version so we can specify the relation literally. This feature of Redex helped us speed up our development significantly when our compilation rules were constantly changing.
% \end{itemize} }
% \item there's a funny change of line spacing in column 2 of page 6, about 2/3 down.
% \end{itemize}
% }

The semantics of \lang uses annotations on pointer literals in order
to keep track of array bounds information, which is used in premises
of rules like \textsc{S-DefArray} and
  \textsc{S-AssignArr} to prevent spatial safety violations. However, in the real
implementation of \checkedc, which extends Clang/LLVM, these annotations
are not present---pointers are represented as a single
machine word with no extra metadata, and bounds
  checks are not handled by the machine, but inserted by the
  compiler.

This section shows how \lang
annotations can be safely erased: using static information a compiler
can insert code to manage and check bounds metadata, with no loss of
expressiveness. We present a compilation algorithm that converts from
\lang to \elang, an untyped language without metadata
annotations. The syntax and semantics \elang
  closely mirrors that of \lang; it differs only in that literals lack
  type annotations and its operational rules perform no
  bounds and null checks, which are instead inserted during
  compilation. Our compilation algorithm is evidence that \lang's
  semantics, despite its apparent use of fat pointers, faithfully
  represents Checked C's intended behavior. The algorithm also sheds
  light on how compilation can be implemented in the real Checked C
  compiler, while eschewing many important details (\elang has many 
  differences with LLVM IR).

Compilation is defined by extending \lang's
typing judgment thusly:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each \code{nt_array_ptr} variable \code{p} to a pair of \emph{shadow
  variables} that keep \code{p}'s up-to-date upper and lower bounds; these
may differ from the bounds in \code{p}'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound shadow variable is unnecessary; we include it for uniformity.}
% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
% complete judgment, implicitly assuming that $e$ is a well-typed and closed
% term.

We formalize rules for this judgment in PLT Redex~\cite{pltredex},
following and extending our Coq development for \lang. To give
confidence that compilation is correct, we use Redex's property-based
random testing support to show that compiled-to $\dot e $ simulates
$e$, for all $e$.

% We developed a \checkedc compiler to compile a \checkedc program to a C program.
% \mwh{We formalized compilation from CoreChkC to a version of CoreChkC
%   but with the metadata removed, right? This is not a Checked C
%   compiler. You go on to see stuff about CompCert, CLight,
%   etc. This is confusing. We should be talking about how this relates
%   to what was just presented. Pick definitive names for things. }
% Given a \checkedc program $e$, we build a compilation process ($\gg$), such that $e \gg \dot e$, where $\dot e$ is the corresponding C program for $e$ in A-normal form (ANF). 
% The compilation process ($\gg$) relies on the type checking step $\Gamma;\Theta\vdash_m e:\tau$. 
% Especially, it relies on $\Gamma$ to provide the type information for variables in $e$. 
% We utilize CompCert/CLight syntax and semantics \cite{Leroy:2009:FVC:1666192.1666216,Blazy2009} as our translation target language.
%  Besides, we defined two data structures in the CompCert format for representing $\enull$ and $\ebounds$ states.
% We write $\xrightarrow{c}$ for the semantics of CLight. \mwh{Of our
%   target language, which I presume does not have all the features that
%   CLight has. Should mention up front that we did all of this in PLT Redex.}

\subsection{Approach}

Due to space constraints, we explain the rules for compilation by
example, using a C-like syntax; the complete rules are given in
\iftr
Appendix~\ref{appx:comp1}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks; and (c) insertion of
bounds widening expressions.
%
A-normal form conversion is straightforward: compound expressions are
handled by storing results of subexpressions into temporary variables,
as in the following example.

{\vspace*{-0.5em}
{\small
\begin{center}
$
\begin{array}{l}
$\code{let y=(x+1)+(6+1)}$
\;
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\;
\begin{array}{l}
$\code{let a=x+1;}$\\
$\code{let b=6+1;}$\\
$\code{let y=a+b}$\\
\end{array}
\end{array}
$
\end{center}
}
}

This simplifies the management of effects from subexpressions. The
next two steps of compilation are more interesting.

\begin{figure}[t!]
  \begin{small}
\begin{lstlisting}[mathescape,xleftmargin=4 mm]
/* p : $\color{purple!40!black}\tntarrayptr{0}{0}{\tint}{\cmode}$  */
/* $\color{purple!40!black}\rho$(p) = p_lo,p_hi */
{
  let x = strlen(p);
  if (x > 1) putchar(*(p+1));
}
\end{lstlisting}
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\begin{lstlisting}[xleftmargin=4 mm]
{
  assert(p_lo <= 0 && 0 <= p_hi); // bounds check
  assert(p != 0); // null check
  let x = strlen(p);
  let p_hi_new = x;
  p_hi = max(p_hi, p_hi_new);
  if (x > 1) {
    assert(p != 0); // null check for p + 1
    let p_1 = p + 1;
    assert(p_lo <= 1 && // bounds check for p + 1
     1 <= p_hi);    
    putchar(*p_1);
  }
}
\end{lstlisting}
\end{small}
\caption{Compilation Example for Check Insertions
% \review{- Fig 8, line 1: did you mean `bounds` instead of `count`? I'm confused
% - Fig 8, line 8: why null-check `p` again?}
% \yiyun{Yes, fixed. The second null check is due to the pointer
%   arithmetic on p; our transformation does not attempt to optimize
%   away provably redundant checks.}
}
\label{fig:compilationexample}
\end{figure}

\begin{figure}[t!]
  \begin{small}
\begin{lstlisting}[mathescape,xleftmargin=4 mm]
int deref_array(n : int,
     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$) {
  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi */
  if (* p)
    (* (p + 1))
    else 0
}
...
/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
deref_array(5, p0);
    \end{lstlisting}
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\begin{lstlisting}[xleftmargin=4 mm]
deref_array(n, p) {
  let p_lo = 0;
  let p_hi = n;
  /* runtime checks */
  assert(p_lo <= 0 && 0 <= p_hi);
  assert(p != 0);
  let p_derefed = *p;
  if (p_derefed != 0) {
    /* widening */
    if (p_hi == 0) {
      p_hi = p_hi + 1;
    }
    /* null check before pointer arithmetic */
    assert(p != 0);
    let p0 = p + 1;
    assert(p_lo <= 1 && 1 <= p_hi);
    (* p0)
  }
  else {
    0
  }
}
...
deref_array(5, p0);
    \end{lstlisting}
\end{small}
    \caption{Compilation Example for Dependent Functions}
\label{fig:compilationexample1}
\end{figure}

% \review{Fig 9: if this is actual C code, then your null-check at line 6 will be
%   eliminated by the compiler. At line 3, you performed a pointer addition, which
%   is only defined when `p` is non-null. So, either `p` is non-null, and the
%   NULL-check can be eliminated; or, `p` is NULL, but line 3 was undefined
%   behavior, meaning the compiler is allowed to do anything, notably eliminate
%   the NULL-check. This is where I am super confused, and either:
%   - CoreC is not really the C language, and has different semantics...? but is
%     this well-defined in the context of LLVM?
%   - there is a problem that was not caught by the PLT-Redex-based testing.
% \yiyun{We have clarified at the start of IV that CoreC is an untyped
%   variant of CoreChkC, and does not aim to represent C per se, or LLVM
%   IR. We aimed to avoid confusion by rewriting the examples in a way that is more closely
%     related to the syntax presented in Fig 3. Pointer arithmetic between 0 and a non-zero
%     index is always valid because CoreC there is technically only
%     integer arithmetic.}}

During compilation, $\Gamma$ tracks the lower and upper bound
associated with every pointer variable according to its type. At each
declaration of a \code{nt_array_ptr} variable \code{p}, the
compiler allocates two \emph{shadow variables}, 
stored in $\rho(p)$; these are initialized to \code{p}'s declared bounds
and will be updated during bounds widening.\footnote{Shadow variables
  are not used for \code{array_ptr} types (the bounds expressions are)
  since they are not subject to bounds widening.} 
Fig.~\ref{fig:compilationexample} shows how an invocation of
\code{strlen} on a null-terminated string is compiled into C
code. Each dereference of a checked pointer requires a null check
(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
compiler makes explicit: Line~$3$ of the generated code has the null
check on pointer \code{p} due to the \code{strlen},
  and a similar check happens
  at line~$8$ due to the pointer arithmetic on \code{p}.
Dereferences also require bounds checks: line~$2$ checks \code{p} is
in bounds before computing \code{strlen(p)}, while line~$10$ does
likewise before computing \code{*(p+1)}.

For \code{strlen(p)} and conditionals \code{if(*p)}, the \lang
semantics allows the upper bound of \code{p} to be extended.
The compiler explicitly inserts statements to do so on \code{p}'s
shadow bound variables. For example,
Fig.~\ref{fig:compilationexample}~line~$6$ widens \code{p}'s upper
bound if \code{strlen}'s result is larger than the existing
bound. 
Lines 7--12 of the generated code in
Fig.~\ref{fig:compilationexample1}
show how bounds are 
widened when compiling expression \code{if(*p)}. If we find that the
current \code{p}'s relative upper bound is equal to $0$ (line 10),
and \code{p}'s content is not null (line 8), we then increase the
upper bound by $1$ (line 11).

Fig.~\ref{fig:compilationexample1} also shows a dependent function call.
Notice that the bounds for the array pointer \code{p} are not passed as
arguments. Instead, they are initialized according to \code{p}'s
type---see line~3 of the original \lang program at the top of the figure.
Line~$2$ of the generated code
sets the lower bound  to \code{0} and line~$3$ sets the
upper bound to \code{n}.

\subsection{Comparison with Checked C Specification}
\label{sec:disc}

\begin{figure}[t]
{\small
{\captionsetup[lstlisting]{margin = 8 mm}
  \begin{lstlisting}[xleftmargin=8 mm]
nt_array_ptr<char> safe_strcat_c 
   (nt_array_ptr<char> dst : count(n),
    nt_array_ptr<char> src : count(0), int n) {
  nt_array_ptr<char> tmp : count(n) = dst;
  int x = strlen(tmp);
  /* tmp now has x as its upper bound */
  /* dst still has n as its upper bound */
  int y = strlen(src);

  if (x+y < n) {
    for (int i = 0; i < y; ++i)
      *(dst+x+i) = *(src+i);
    *(dst+x+y) = '\0';
    return dst;
  }
  return null;
}
\end{lstlisting}
}
}
\caption{Safe \code{strcat} in Checked C that avoids a run-time error
  exhibited by \code{safe_strcat} (Fig.~\ref{fig:strcat-ex}) when
  compiled with the current Checked C compiler}
\label{fig:strcatc-ex}
\end{figure}

The use of shadow variables for bounds widening is a key novelty of our
compilation approach, and adds more precision to bounds checking at runtime
compared to the official specification and current implementation of
\checkedc~\cite[5.1.2, pg 85]{checkedc}.  For example, the
\code{safe_strcat} example of Fig.~\ref{fig:strcat-ex} compiles with the
current Clang \checkedc compiler but will fail with a runtime
error. The statement \code{int x = strlen(dst)} at line 4 changes the
statically determined upper bound of \code{dst} to \code{x}, which can be smaller than
\code{n}, the full capacity of \code{dst}. The attempt to recover
the full capacity of \code{dst} through a dynamic cast
at line 7 will always fail if the capacity \code{n} is checked
against the statically determined new upper bound \code{x}. % This problem can be worked around by
% effectively inlining the definition of \code{strlen} as 
% in \code{safe_strcat_c} in Fig.~\ref{fig:strcatc-ex} (lines
% 6-7).
This problem can be worked around by invoking \code{strlen} on a temporary
variable \code{tmp} instead of \code{dst}
as in \code{safe_strcat_c} in Fig.~\ref{fig:strcatc-ex} (lines 4-5).
Likewise, if we were to add line \code{putchar(*(p+1));} 
after line 6 in the original code at the top of
Fig.~\ref{fig:compilationexample}, the code will always fail: the Clang \checkedc
compiler (with the transliterated C code as its input) would 
check \code{p} against its \emph{original} bounds \code{(0,0)} since the
updated upper bound \code{x} is now out of
the scope. Shadow variables address these problems because
they retain widened bounds beyond the scope of variables that store
them (i.e., \code{x} in both examples).

To make it match the specification, our compilation definition could
easily eschew
shadow variables and rely only on the type-based
bounds expressions available in $\Gamma$ for checking. However, doing so would 
force us to weaken the simulation theorem, reduce expressiveness,
and/or force the semantics to be more awkward. We plan to work with
the \checkedc team to implement our approach in a future revision.

% \textcolor{purple}{Despite its name, \elang
%   does not model C. As mentioned at the beginning of this section,
%   \elang is an untyped version of \lang with the annotations on
%   literals removed. The lack of types make it impossible to capture
%   the undefinedness of certain operations such as pointer
%   arithmetic. This restriction is not an issue for establishing
%   the metatheory covered in the next section, but it does indicate
%   that we have to be careful if we want to embed \elang programs into C
%   programs.  As discussed in \cite[2.10, pg 24]{checkedc}, the C
%   standard can be overly
%   conservative when it comes to pointer arithmetic. For example,
%   it should be perfectly safe to increment a pointer out of its allocated
%   region, as long as the runtime guarantees that
%   the pointer is never dereferenced. However, a standard C compiler
%   would treat this type of incrementing operation as undefined, regardless of
%   whether the resulting pointer is dereferenced or not. Embedding
%   \elang naively into C would give us this type of undefined programs.
%   This can be mitigated through casting or rearranging the code during
%   the embedding process.
% }
% \mwh{Don't think all this was necessary. Added clarifying text at the
%   start at exactly what CoreC is. No need to go on at length about
%   what it is not.}

\subsection{Metatheory}
\label{sec:meta}

% \review{ 
% My interpretation of section IV.C is that the authors have a pen-and-paper
%   proof of Theorem 4, and that random testing using the Redex models has been
%   used to gain confidence in such a proof. Is this correct?}
% \liyi{Yes, this is correct.}
% \mwh{No, not correct: We have no pen-and-paper proof, just the
%   model. Clarify here}
% showing that
% extensive random testing fails to falsify the bisimilarity
% property. \mwh{Better way to state the previous?}

We formalize both the compilation procedure and the simulation
theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
and then attempt to falsify it via Redex's support for random
testing. Redex allows us
  to specify compilation as logical rules (essentially, an extension
  of typing), but then execute it algorithmically to
  automatically test whether simulation holds. This process revealed
  several bugs in compilation and the theorem statement.
%
  % us gain confidence that our original pen and paper proof of
  % simulation remains true with the addition of variable bounds. }
We ultimately plan to prove simulation in the Coq model.

%Turning to the simulation theorem: We first introduce notation
%used to specify the theorem.
We use the notation $\gg$ to
indicate the \emph{erasure} of stack and heap---the rhs is the same as
the lhs but with type annotations removed:
\begin{equation*}
  \begin{split}
    \heap  \gg & \dot \heap \\
    \varphi \gg & \dot \varphi
  \end{split}
\end{equation*}
In addition, when $\Gamma;\emptyset\vdash
\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg (\dot \varphi, \dot \heap,
\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
and $\Gamma;\Theta;\emptyset \vdash e \gg \dot e : \tau$ for some $\tau$ respectively. $\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
reduction relation of $\elang$. Unlike the $\lang$, the semantics of
$\elang$ does not distinguish checked and unchecked regions.

Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
that is similar to the one by~\citet{merigoux2021catala}.
An ordinary simulation property would
replace the middle and bottom parts of the figure with the
following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
Instead, we relate two erased configurations using the relation $\sim$,
which only requires that the two configurations will eventually reduce
to the same state. We formulate our simulation theorem differently
because the standard simulation theorem imposes a very strong
syntactic restriction to the compilation strategy. Very often, $(\dot
\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
but we are unable to syntactically equate the two configurations due
to the extra binders generated for dynamic checks and ANF
transformation. In earlier versions of the Redex model, we attempted
to change the compilation rules so the configurations could match
syntactically. However, the approach scaled poorly as we added
additional rules. 
This slight relaxation on the equivalence relation
between target configurations allows us to specify compilation more
naturally without having to worry about syntactic constraints.

% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg(\dot \varphi_0,\dot \heap_0, \dot e_0)$,
and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
\rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, then the following facts hold:

\begin{itemize}

\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
\varphi,\dot \heap,\dot e)$ and $(\dot
\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
\dot \heap,\dot e)$.

\item if $r_1 = \ebounds$ or $\enull$, then we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
\varphi_1$, $\heap_1 \gg \dot \heap_1$.

\end{itemize}
\end{thm}

% when $r_1 = e_1$ for
% some $e_1$ and
% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
% \varphi,\dot \heap,\dot e)$ and $(\dot
% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
% \varphi_1$, $\heap_1 \gg \dot \heap_1$.

% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
% some $e_1$ and
% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
% , then
% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
% \varphi,\dot \heap,\dot e)$ and $(\dot
% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
% \end{thm}

Our random generator (discussed in the next section) never produces
$\euncheckedtext$ expressions (whose behavior could be undefined), so we
can only test a the simulation theorem 
as it applies to checked code. This limitation makes it
unnecessary to state the other direction of the simulation theorem
where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
that $e_0$ will never enter a stuck state if it is well-typed in
checked mode.

The current version of the Redex model has been tested against $20000$
expressions with depth less than $10$. Each expression can
reduce multiple steps, and we test simulation between every two
adjacent steps to cover a wider range of programs, particularly the
ones that have a non-empty heap.

\begin{figure}[t]
{\small
\[
\begin{array}{c}
\begin{tikzpicture}[
            > = stealth, % arrow head style
            shorten > = 1pt, % don't touch arrow head to node
            auto,
            node distance = 3cm
        ]

\begin{scope}[every node/.style={draw}]
    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
\end{scope}
\begin{scope}[every edge/.style={draw=black}]

    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
    \path [<->] (A) edge node {$\gg$} (C);
    \path [<->] (B) edge node {$\gg$} (D);
    \path [dashed,<->] (C) edge node {$\sim$} (D);
    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
\end{scope}

\end{tikzpicture}
\end{array}
\]
}
\caption{Simulation between \lang and \elang }
\label{fig:checkedc-simulation-ref}
\end{figure}

% \ignore
% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
% % The diagram also indicates the simulation relation ($\sim$) among our formalization of the \checkedc Type System ($\vdash$), Semantics ($\longrightarrow$), Compilation ($\gg$), as well as the compiled C semantics ($\xrightarrow{c}$).
% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist. The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.

%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C. The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).

% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation. We generate tens of thousands cases and cover all possible corner cases of the compilation. 
% }
% \ignore{
% \begin{figure*}[t]
%   \begin{subfigure}[b]{1\textwidth}
%     \begin{lstlisting}
% int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
%   return *p;
% }

% ...
% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
% deref_array(5, p0);
%     \end{lstlisting}
%     \label{fig:chkcexamplederef}
%     \caption{\texttt{deref\_array} in \checkedc}
%   \end{subfigure}

%   \begin{subfigure}[b]{1\textwidth}
% \begin{lstlisting}
% int deref_array(size_t n, int *p) {
%   /* statically determine the definitions p_lo and p_hi */
%   int *p_lo = p;
%   int *p_hi = p + n;
%   /* runtime checks */
%   assert(p_lo <= p && p <= p_hi);
%   assert(p != NULL);
%   /* widening */
%   int p_derefed = *p;
%   if (p_derefed != '\0') {
%     if (p_hi == p) {
%       ++p_hi;
%     }
%   }
%   return p_derefed;
% }
% ...
% /* int *p0 */
% /* rho(p0) = p_lo, p_hi */
% deref_array(5, p0);
% \end{lstlisting}
% \caption{\texttt{deref\_array} in C}
%     \label{fig:cexamplederef}
%   \end{subfigure}

%   \begin{subfigure}[b]{\textwidth}
%     \begin{lstlisting}
% /* nt_array_ptr<int> p : bounds(e0, e1) */
% /* rho(p) = p_lo, p_hi */
% size_t x = strlen(p);
% if(x >= 1) {
%   *(p+1);
% }
%     \end{lstlisting}
%     \caption{\texttt{strlen} in \checkedc}
%     \label{fig:chkcexamplestrlen}
%   \end{subfigure}

%   \begin{subfigure}[b]{\textwidth}
%     \begin{lstlisting}
% /* nt_array_ptr<int> p : bounds(e0, e1) */
% /* rho(p) = p_lo, p_hi */
% /* runtime checks omitted */
% ...
% size_t x = strlen(p);
% int *p_hi_new = p + x;
% p_hi = max(p_hi, p_hi_new);
% if (x >= 1) {
%   /* null check for pointer arithmetic */
%   assert(p != NULL);
%   int *p_1 = p + 1;
%   /* null check for dereference */
%   assert(p_1 != NULL);
%   /* bounds check for dereference */
%   /* note how we are using p_hi rather than p + x */
%   assert(p_lo <= p_1 && p_1 <= p_hi);
%   ...
% }
%     \end{lstlisting}
%     \caption{\texttt{strlen} in C}
%     \label{fig:cexamplestrlen}
%   \end{subfigure}
% \caption{Compilation of null-terminated string dereference}
% \label{fig:compilationexample}
% \end{figure*}
% }

% \ignore{
% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
% dereference of null-terminated strings by inserting explicit checks and
% bounds widening code. For readability, we present the example in real
% \checkedc and C syntax. At
% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
% bounds are defined in terms of the arguments \code{p} and \code{n}
% according to the bounds annotations obtained during typechecking. This
% avoids the need to pass in bounds as extra arguments, thus maintaining
% compatibility with C code. When we call \code{deref_array} on
% \code{p0}, a string with size 5 (excluding the null character), there
% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
% the stack. It is possible that \code{p_hi} has been widened
% before we reach line 21. The programmer will have to perform a
% \code{dynamic_bounds_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
% attempts to widen the bound when the dereferenced result is not
% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
% \code{deref_array}, in contrast to the T-IfNT rule, which only
% remembers the widened bounds within the scope of the then branch of
% the if statement.

% \todo[inline]{Widening can happen at every
%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
% shows how an invocation of \code{strlen} on null-termianted strings
% is compiled into C code. We perform the same runtime checks that
% happen during dereferencing. The widening code at line
% 8 updates \code{p}'s upper bound only if the result of
% \code{strlen} is larger than the value of the upper bound stored on
% the stack. This is another scenario where the runtime can be more precise
% than the statically determined bounds information.

% }

### 2022-06-25 sandp23/checkedc-formal/evaluation.tex
update paper.

\section{Random Testing via the Implementation}\label{sec:evaluation}

% \review{While I found the idea behind section V very interesting, the current version
%   of this section lacks some details that would help in better understanding (1) 
%   how the approach works, and (2) the overall scope of the approach. 
%   $\\$
%   For instance, the authors state that, following [19], they try to "exercise
%   interesting patterns" by adding "admissible but redundant typing rules" like
%   G-ASTR. There are a few points that are unclear here: (1) are these rules
%   discovered manually or automatically (starting from the Redex semantics)?, (2)
%   are there any guiding principles for coming up with rules that lead to
%   interesting cases?
%   $\\$
%   Later, the authors refer to "generation rules modified to be slightly more
%   permissive" to generate "a little" ill-typed terms. Again, are these rules
%   obtained automatically or defined manually? If the latter, did you follow any
%   methodology to derive such rules? Are these rules the same as the "admissible
%   but redundant typing rules" from above?}
% \liyi{Deena? Leo? }

In addition to using the \lang Redex model to establish simulation of
compilation (Section~\ref{sec:meta}), we also used it to gain confidence
that our model matches the Clang \checkedc implementation.
Disagreement on outcomes signals a bug in either the model or the
compiler itself. Doing so allowed us to quickly iterate on the
design of the model while adding new features, and revealed
several bugs in the Clang \checkedc implementation.

\myparagraph{Generating Well Typed Terms} For this random generation, we follow
the approach of~\citet{PalkaAST11} to generate well-typed \checkedc
terms by viewing the typing rules as generation rules.
%
%Suppose we have a context $\textcolor{red}{\Gamma}$, a mode
%$\textcolor{red}{m}$ and a type $\textcolor{red}{\tau}$, and we are
Suppose we have a context $\Gamma$, a mode
$m$ and a type $\tau$, and we are
trying to generate a well-typed expression. We can do that by
reversing the process of type checking, selecting a typing rule and
building up an expression in a way that satisfies the rule's premises.

Recall the typing rule for dereferencing an
array pointer, which we depict below as
\textsc{G-DefArr}\footnote{Generator rules G-* correspond one to one
  with the type rules T-* in Sec.~\ref{sec:type-system}.}, color-coded to represent \textcolor{red}{inputs} and \textcolor{blue}{outputs} of the
generation process:%
\footnote{This input-output marking is commonly called a mode in
  the literature, but we eschew this term to avoid
  confusion with our pointer mode annotation.}
%
%\begin{figure}[t]
\begin{mathpar}
    \inferrule[G-DefArr]
              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} : \tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m'}} \\
\textcolor{red}{m} \leq \textcolor{cyan}{m'} 
              }
              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\estar{e}} : \textcolor{red}{\tau}}
%  
%     \inferrule[T-Dfa] {
% %      \textcolor{cyan}{m'} = \texttt{choose}(C,U) \text{ if } \textcolor{red}{m} = U
% %      \text{, otherwise } \textcolor{cyan}{m'} = C
% %      \\
%       \textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} : \tptr{\textcolor{red}{\tau}}{\textcolor{cyan}{m'}} \\      
%       \textcolor{red}{m} \leq \textcolor{cyan}{m'} \\
% }
%     {\textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\estar{e}} : \textcolor{red}{\tau}}
  \end{mathpar}
%   \caption{T-DFA as a generation rule}
%   \label{fig:tdfagen}
% \end{figure}
%
If we selected \textsc{G-DefArr} for generating an expression, the
generated expression has to have the form $\textcolor{blue}{*e}$,
for some $\textcolor{blue}{e}$, to be generated according
to the rule's premises.
%
To satisfy the premise
$\textcolor{red}{\Gamma};\textcolor{red}{\Theta}
\vdash_{\textcolor{red}{m}} \textcolor{blue}{e} :
\tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m'}}$,
we essentially need to make a recursive call to the generator, with
appropriately adjusted inputs.
%
However, the type in this judgment is not fixed yet---it contains
three unknown variables: $\textcolor{cyan}{m'}$,
$\textcolor{cyan}{\bvar}$, and $\textcolor{cyan}{\kappa}$---that need
to be generated before making the call.
%
Looking at the second premise informs that generation:
if the input mode $\textcolor{red}{m}$ is $\umode$, then $\textcolor{cyan}{m'}$
needs to be $\umode$ as well; if not, it is unconstrained,
just like $\textcolor{cyan}{\bvar}$ and $\textcolor{cyan}{\kappa}$, and therefore all three are free to be generated at random.
%
Thus, the recursive call to generate $\textcolor{blue}{e}$ can now
be made, and the \textsc{G-DefArr} rule returns $\textcolor{blue}{*e}$
as its output.

% If we selected \textbf{T-Dfa} for generating an expression, the
% expression we generate has to have the form $\textcolor{blue}{*e}$,
% for some expression $\textcolor{blue}{e}$, to be generated according
% to the rule's premises.
% %
% To satisfy the premise $\textcolor{red}{\Gamma}
% \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} :
% \tptr{\textcolor{red}{\tau}}{\textcolor{cyan}{m'}}$, we essentially
% need to make a recursive call to the generator, with appropriately
% adjusted inputs.
% %
% However, the type in this judgement is not fixed yet---it contains an
% unknown variable $\textcolor{cyan}{m'}$ that needs to be generated
% before making the call.
% %
% Looking at the second premise informs that generation:
% if the input mode $\textcolor{red}{m}$ is $\umode$, then $\textcolor{cyan}{m'}$
% needs to be $\umode$ as well; if not, it is unconstrained and we're free
% to generate it at random to be $\cmode$ or $\umode$.
% %
% Thus, the recursive call to generate $\textcolor{blue}{e}$ can now
% be made, and the \textbf{T-Dfa} rule returns $\textcolor{blue}{*e}$
% as its output.

Using such generator rules, we can create a generator for random
well-typed terms of a given type in a straightforward manner: find all
rules whose conclusion matches the given type and then randomly choose
a candidate rule to perform the generation. To ensure that this
process terminates, we follow the standard practice of using ``fuel'' to
bound the depth of the generated terms; once the fuel is exhausted,
only rules without recursive premises are
selected~\cite{Pierce:SF4}. Similar methods were used for
generating top level functions and \kw{struct} definitions.

%^ \leo{The following is an insanely large, undigestible paragraph, but now moved to the correct place.}
%^ The generator follows the typing rules in the redex model to generate
%^ well typed terms. Since the redex model only supports top level
%^ functions the functions were generated separately, as were struct
%^ definitions. Given a list of struct definitions, a list of functions,
%^ and a type the generator randomly chooses the structure of a term that
%^ can have that type. This is based on the typing rules. Then depending
%^ on the holes needed to be filled in the term, new expressions are
%^ generated. For example when attempting to generate an expression of
%^ type (ptr c int) with fuel 2 the following steps might be taken. First
%^ a conditional expression is randomly chosen. In order for this to have
%^ the correct type a random expression of type int with fuel 1 is
%^ generated as the guard. Then both branched must have type (ptr c
%^ int). They will recursively generate but with reduced fuel. If the
%^ fuel is less than or equal to zero then either a variable with the
%^ correct type will be chosen from the enviornment or a malloc
%^ expression will be generated. The generator takes fuel so that
%^ generating expressions will always terminate.

While using just the typing-turned-generation rules is in theory
enough to generate all well-typed terms, it's more effective in
practice to try and exercise interesting patterns.  As
in~\citet{PalkaAST11} this can be viewed as a way of adding admissible
but redundant typing rules, with the sole purpose of
using them for generation. For example, below is one such rule,
\textsc{G-ASTR}, which creates an initialized null-terminated string that
is statically cast into an array with bounds $(0,0)$.
%
%\begin{figure}[t]
  \begin{mathpar}
    \inferrule[G-ASTR]
    {
      \textcolor{cyan}{i} \in \mathbb{N}^* \quad\quad
      \textcolor{cyan}{n_0,\ldots,n_{i-1} } \in \mathbb{Z} \quad\quad
      \texttt{fresh}(\textcolor{cyan}{x})\\
      \textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}}
      \textcolor{blue}{e'} :
      \tptr{\textcolor{red}{\tntarray{0}{\textcolor{cyan}{i}}{\tint}}}{c}
      \\      
% \eassign{(\ebinop{x}{0})}{n_0};\ldots;\eassign{(\ebinop{x}{i-1})}
      \textcolor{blue}{e} = \elet{\textcolor{cyan}{x}}{\textcolor{blue}{e'}}{\texttt{(init $\textcolor{cyan}{x}$ with }\textcolor{cyan}{n_0,\ldots
        n_{i-1}}\texttt{)};\textcolor{cyan}{x}}
      % \textcolor{cyan}{m'} = \texttt{choose}(C,U) \text{ if } \textcolor{red}{m} = U
      % \text{, otherwise } \textcolor{cyan}{m'} = C
      % \\
}
    {\textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\ecast{\tptr{\tntarray{0}{0}{\tint}}{c}}{e}} : \textcolor{red}{\tptr{\tntarray{0}{0}{\tint}}{c}}}
  \end{mathpar}
%  \caption{Specialized rule for generating initialized strings}
%  \label{fig:gastr}
%\end{figure}
%
Given some positive number $\textcolor{cyan}{i}$, numbers
$\textcolor{cyan}{n_0, \ldots, n_{i-1}}$, and a fresh variable
$\textcolor{cyan}{x}$ (which are arbitrarily generated), we can
recursively generate a pointer $\textcolor{blue}{e'}$ with bounds
$(0,\textcolor{cyan}{i})$, and initialize it with the generated $n_j$
using $\textcolor{cyan}{x}$ to temporarily store the pointer.

This rule is particularly useful when combined with \textsc{G-IfNT}
since there is a much higher chance of obtaining a non-zero value when
evaluating $\estar{p}$ in the guard of $\eiftext$, skewing the
distribution towards programs that enter the $\ethentext$ branch.
Relying solely on the type-based rules, entering the $\ethentext$
branch requires that \textsc{G-AssignArr} was chosen before \textsc{G-IfNT},
and that assignment would have to appear before $\eiftext$, which
means additional \textsc{G-Let} rules would need to be chosen: this
combination would therefore be essentially impossible to generate in
isolation.

Adding admissible generation rules like \textsc{G-Astr} in this
manner, as described in~\citet{PalkaAST11}, is a manual process. It is
guided by gathering statistics on the generated data and focusing on
language constructs that appear underrepresented in the posterior
distribution. For example, we arrived at the \textsc{G-Astr} rule by
recognizing that the pure type-based generation was not generating
non-trivial null-terminated strings, and then analyzing the sequence
of random choices that could lead to their generation.

% Originally only the
% typing rules were used to generate expressions. Eventualy new options
% were added to create more interesting programs. Interesting programs
% include allocating memory for an array and then initializing it with
% values, and casting arrays to a smaller type so that bounds widening
% can be triggered. There were also some issues with generating strlen
% expressions. After a variable is assigned to the strlen of a pointer
% that bounds for that pointer contain the variable. This caused issues
% with scope as an array with bounds dependent on that variable could be
% returned. In order to prevent these issues expressions that use strlen
% are only generated in the form \code{let \_ = let x = (strlen y)} in
% (expression generated with updated bounds for y) in (some literal
% expression that does not contain y). Once expressions from the redex
% model are generated they are compiled into CheckedC and compiled with
% the clang compiler. One problem is that type annotations are needed in
% Checked-C but do not exist in the Redex model. To properly translate
% an enviornment is passed around. As various parts are translated types
% are added to the enviornment so when declarations are needed their
% types can be found.  Several small changes had to be made to
% expressions since the redex model which used immutable values, had
% expressions that were not allowed in the clang version (e.g. let
% expressions, conditional expressions with complex bound types in the
% branches, using calloc to initialize nt-arrays, and casting motivating
% expressions). Let, if and cast expressions got lifted to functions. We
% defined a calloc for nt-arrays using a dynamic bounds cast. The int
% type was changed to char. This is because while in the specification
% any nt-array can be passed to strlen the clang compiler only allows
% arrays of type const char *. The expression was evaluated and then
% printed to stdout.
%  
%  
%  
% \leo{TODO: adding interesting patterns (mentioned later), just like Palka
%   adds seq}
% \leo{I don't understand why the conversion tool is important here. I
%   moved the following para up, as it has to do with generation.}
% % \yiyun{I'm not sure if patterns can be written as a clean rule. For
% %   example, we want to add rules that creates an ntarray of size n,
% %   initializes the array, and then }
% \yiyun{I'm not sure if T-ASTR is exactly implemented like that in
%   Deena's random generator, but we do have something that performs
%   this type of initialization.}
% The conversion tool can be used separately from the random generator
% to speed up the manual testing process, where we construct programs
% that we believe are likely to trigger rules that have subtle
% definitions and are likely to cause failure in the experimental Clang
% implementation. For example, we were able to find that % \yiyun{is this
% % accurate? not that familiar with itype} the \code{itype} 
% \checkedc wrapper
% of \code{strlen} does not perform null check at the
% interface, which would lead to undefined behavior within the checked
% region. We integrate
% these manually found patterns into the random generator to complement the
% type-based generation approach, allowing us to cover more interseting
% cases during random testing.

\myparagraph{Generating Ill-typed Terms}
%
We can use generated well-typed terms to test our simulation theorem
(Section~\ref{sec:compilation}) and test that \lang and \checkedc
Clang agree on what is type-correct. But it is also useful to generate
ill-typed terms to test that \lang and \checkedc Clang also agree on
what is not.  However, while it is easy to generate arbitrary
ill-typed terms, they would be very unlikely to trigger any
inconsistencies; those are far more likely to exist on the boundary
between well- and ill-typedness. Therefore, we also manually added
variations of existing generation rules modified to be slightly more
permissive, e.g., by relaxing a single premise, thus allowing terms
that are ``a little'' ill-typed to be generated. Unlike coming up with
admissible generation rules like \textsc{G-Astr} (which is quite
challenging to automate), systematically and automatically relaxing
premises of existing rules seems feasible, and worthwhile future work.

% \myparagraph{Testing Model Properties}
% The generator is used to establish confidence in properties that are
% expected to hold in our model, such as progress and preservation of
% \lang, or more notably the simulation property between \lang and
% \elang (Theorem~\ref{simulation-thm}).
% %
% As discussed in
% Section~\ref{sec:compilation}, we use the simulation property to
% ensure that our source language can be compiled into a language that
% does not rely on fat pointers. For simulation to hold, 
% the semantics must avoid accessing information that would
% necessitate the use of fat pointers during compilation. This is unsurprisingly
% a tricky process, and we found mistakes in early versions of our
% desing where we wrote widened bounds onto the heap, or passed
% widened bounds to functions, making them accessible outside of the
% current stack frame (thus implying the existence of fat
% pointers). Random testing also helped us discover the limitation of
% the standard simulation theorem and motivated us to rephrase it
% in such a way that is compatible with our definition of compilation.
% % relax our simulation theorem to require only the confluence of the
% % two compiled programs by providing examples where the compilation
% % process can generate redundant stack variables that would break the more
% % standard simulation property
% . \yiyun{does this deserve a figure to
%   explain? or maybe I should mention why our simulation is slightly
%   off from what a ``standard'' simulation should look like in
%   compilation? }\leo{This should have already been discussed in the
%   statement of simulation - point there?}

\myparagraph{Random Testing for Language Design}
%
We used our Redex model and random generator to successfully guide the
design of our formal model, and indeed the Clang Checked C
implementation itself, which is being actively developed. To that end, we implemented a
conversion tool that converts \lang into a subset of the \checkedc
language and ensured that model and implementation exhibit the same
behavior (accept and reject the same programs and yield the same return
value).

This approach constitutes an interesting twist to traditional
model-based checking approaches.  Usually, one checks that the
implementation and model agree on all inputs \emph{of the
  implementation}, with the goal of covering as many behaviors as
possible. This is the case, for example, in \citet{lambdajs}, where
they use real test suites to demonstrate the faithfullness of their
core calculus to Javascript. Our approach and goal in this work is
essentially the opposite: as the Clang Checked C implementation does
not fully implement the Checked C spec, there is little hope of
covering all terms that are generated by Clang Checked C. Instead,
we're looking for \emph{inconsistencies}, which could be caused by
bugs either in the Clang Checked C compiler or our own model.

One inconsistency we found comes from the following:

{\small 
\begin{lstlisting}[xleftmargin=4 mm]
 array_ptr<char> fun(void) : count(3) {
   array_ptr<char> x : count(3);
   x = calloc(3, sizeof(char));
   return x+3;
 }
 int main(void) {
   *(fun()) = 0;
   return 0;
 }
\end{lstlisting}
}% 
%
\noindent
In this code, the function \code{fun} is supposed to return a
checked array pointer of size 3. Internally, it allocates such an
array, but instead of returning the pointer \code{x} to that array, it
increments that pointer by 3. Then, the \code{main} function just
calls \code{fun}, and tries to assign 0 to its result. Our model
correctly rules out this program, while the Clang Checked C
implementation happily accepted this out-of-bounds
assignment. Interestingly, it correctly rejected programs where the
array had size 1 or 2. This inconsistency has been fixed in the latest
version of the compiler.

We also found the opposite kind of inconsistency---programs that
the Clang Checked C implementation rejects contrary to the spec.
For instance:%
\footnote{After minimization, this turned out to be a known issue: 
  \url{https://github.com/microsoft/checkedc-clang/issues/1008}}

{\small 
\begin{lstlisting}[xleftmargin=4 mm]
array_ptr<int> f(void) : count(5) {
  array_ptr<int> x : count(5) =
    calloc<int>(5, sizeof(int)); 
  return x;
}
array_ptr<int> g(void ) : count(5) {
  array_ptr<int> x : count(5) =
    calloc<int>(5, sizeof(int)); 
  return x+3;
}
int main(void) {
  return *(0 ? g() : f() + 3);
}    
\end{lstlisting}
}% 
\noindent
In this piece of code both \code{f} and \code{g} functions compute a
pointer to the same index in an array of size 5 (as \code{f} calls
\code{g}). The \code{main} function then creates a ternary expression
whose branches call \code{f} and \code{g}, but the Clang Checked C
implementation rejects this program, as its static analysis is not
sophisticated enough to detect that both branches have the same type.
% 
%\leo{The following is extremely
%  weak. ``Most of them'', were there any that weren't? Which ones? For
%  the ones that were, mention github issues.}  The random generator,
%equipped with the conversion tool, successfully found a few minor
%errors in the clang compiler, most of them were already issues in the
%git bug reports. For example, we discovered that while the ternary
%operator is implemented in the compiler it cannot handle complex
%bounds types in the branches. The static analysis is not sophisticated
%enough to properly detect that both branches have the same type. While
%not precisely a bug, the clang compiler does not permit memory for
%null terminated arrays to be allocated with calloc. Although calloc
%fills all spaces in memory with null, the compiler does not recognize
%this and claims that it is an unsafe cast.

% Recall that our
% formal model makes liberal use of bounds annotations in literals and
% the heap.

% In order to get a better understanding of the formalism we wrote it in
% redex. This allowed us to make sure that expressions were well-typed
% and evaluated to what we expected. It also was helpful for use in
% prototyping; new features could first be added to the redex model to
% see how they interacted with the existing language. This model was
% slightly larger than the Coq model and there are some differences in
% the type systems. We included top level functions and conditional
% expressions. All of these extra expressions are still expressible in
% the coq model, for example functions can be represented as nested let
% expressions. In the Coq model variables are stored on a stack while in
% the Redex model the variables are simply looked up in the context. In
% general the Redex model is easier to modify and slightly closer to the
% actual Checked-C specifications. Instead of using the model for a
% static proof, we used it to increase our certainty of the accuracy of
% the model.

% \item Describe the random testing generator setup and the properties
%   to test.
% \yiyun{Deena's description of the implementation details. I tried
%   integrating the ones that I find relevant/interesting to the text above. Maybe we can
%   add more if we have some space to fill in.}
% In order for our guarantee of safety to hold, we need to know that our
% model acurately reflects the CheckedC clang compiler. Safety is proved
% for the Coq model, but it is significantly smaller than the actual
% language. The Redex model is a combination of both. It is written in
% the same style as the formalism but has slightly more of Checked-C's
% extra features. If expressions from the Redex model display the same
% behavior as equivalent programs in Checked-C then we have greater
% certainty that our model is useful. We built a random testing
% generator to increase this certainty.

  % \item Describe the bug findings from the random testing against the Checked-C compiler.
%   \leo{This is now integrated above}
% The generator was helpful in finding bugs in the redex model. Several things failed to typecheck that should have been well typed, and the generator was able to catch them. The generated code also found a few minor errors in the clang compiler, most of them were already issues in the git bug reports. For example we discovered that while the ternary operator is implemented in the compiler it cannot handle complex bounds types in the branches. The static analysis is not sophisticated enough to properly detect that both branches have the same type. While not precisely a bug, the clang compiler does not permit memory for null terminated arrays to be allocated with calloc. Although calloc fills all spaces in memory with null, the compiler does not recognize this and claims that it is an unsafe cast. In the Redex model there is no issue with this. A few other minor things were brought to light in the implementation of the generator. The main use was to increase certainty that the behavior in the formal model accurately matched the clang compiler.
%  
% % \end{itemize}

% \begin{itemize}
%  
% \item Show that why the formal semantics/type-system defined for Checked-C is useful. 
% Since we have certainty that our model reflects the clang compiler the model is very useful. Proofs are easier on the smaller model, so we can show  that certain things are true for it. Since the Redex model is between the formalism and the clang version we can have certainty that properties we expect are actually true for the clang version.
%  
% \begin{itemize}
% \item Show some bug findings. 
% \item Show the properties that we can guarantee for Checked-C based on the type-system and blame theorem.
% \item Maybe other useful tools that can be extracted from the Redex model.
%  
% \end{itemize}
%  
% \end{itemize}

### 2022-06-25 sandp23/formal.tex
update paper.

\section{Formalization}\label{sec:formal}

% \begin{itemize}
% \item Describe the types for checked-C as a graph. \dvh{I don't know what this means.}

% \item Describe the subtyping for checked-C types.
% \begin{itemize}
% \item State that the subtypes in Checked-C are transitive.
% \end{itemize}

% \item Describe the syntax of Checked-C

% \item Describe the semantics of Checked-C

% \item Describe the type system of Checked-C

% \item Describe the progress and preservation theorems, and outline the proofs.

% \item Describe the blame theorem and proof.

% \end{itemize}

% \ignore{
% \begin{figure}
%   \begin{center}
%     \includegraphics[height=6in]{syntax.pdf}
%   \end{center}
%   \caption{\lang: Syntax}
% \end{figure}

% \begin{figure}
%   \begin{center}
%     \includegraphics[height=6in]{types.pdf}
%   \end{center}
%   \caption{\lang: Typing}
% \end{figure}
% }
% \liyi{main text begins here. }

% \review{
% While reading the semantics, I found the fact that S-Def and S-DefNull are
%   applicable non-deterministically if n is 0 a bit confusing. Only when
%   reading the meta-theory section I realized that this is not a concrete issue
%   because well-formed heaps are such that $\mathcal{H}(0)$ is never defined. It
%   might be worth pointing this out early on. }
% \mwh{Done.}

\begin{figure}
  \small \centering
  $\begin{array}{l}
\begin{array}{lll}
\text{Function names:}~f&
       \text{Variables:}~ x
& \text{Integers:}~n::=\mathbb{Z} 
\end{array}
\\[0.5em]

\begin{array}{llcllcl}

\text{Mode:} & m & ::= & \cmode \mid \umode \\[0.5em]

\text{Bound:} & b & ::= & n \mid x \plus n \\
              & \bvar & ::= & (b,b) \\[0.5em]

     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{m}
\\[0.5em]

\text{Type Flag:}&\kappa &::=& nt \mid \cdot
\\[0.5em]

\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau}
\\[0.5em]

\text{Expression:}& e & ::= & 
\evalue{n}{\tau} \mid x \mid \emalloc{\omega} \mid\elet{x}{e}{e} \\[0.2em]
&&\mid&
\ecast{\tau}{e} \mid \edyncast{\tau}{e}\mid \ecall{f}{\overline{e}} \mid \estrlen{x} \\[0.2em]
&&\mid&
\ebinop{e}{e} \mid\estar{e}\mid\eassign{e}{e}\mid\eunchecked{e}
\\[0.2em]
&&\mid&\eif{e}{e}{e}
\end{array}
    \end{array}
  $
  \caption{\lang Syntax}
  \label{fig:checkc-syn}
\end{figure}

%% \dvh{I don't understand the variable grammar.  What is $T$?  What is $\eta$?  I think $\cmode$ and $\umode$ should be in tt font.}
%% \liyi{T and $\eta$ can be moved to the appendix, they are useful only for struct types.}

% \review{
% - Furthermore, inspecting the code also suggests that the expression type (line 155) does not contain constructors for function calls (and I don't see a way to define functions either), conditionals, or strlen, and doesn't distinguish between the two forms of casting. All this contradicts figure 2, and should be clarified
% }
% \liyi{It is in the CheckedC.v file, 392. }
% \mwh{How is this answer helping the reviewer since you've added
%   nothing to the text? Maybe we should add something to an appendix
%   that matches the formalism shown in the paper to definitions in the
%   Coq file?}
% \liyi{The detailed explanation is in the appendix.}

This section describes our formal model of \checkedc, called
\lang, making precise its syntax, semantics, and type system. It also
develops \lang's meta-theory, including type soundness and the blame
theorem.

\subsection{Syntax}\label{sec:syntax}

The syntax of \lang is given by the expression-based
language presented in Fig.~\ref{fig:checkc-syn}.

There are two notions of type in \lang.  Types $\tau$ classify
word-sized values including integers and pointers, while types
$\omega$ classify multi-word values such as arrays, null-terminated
arrays, and single-word-size values.
%
Pointer types ($\tptr{\omega}{m}$) include a mode annotation ($m$)
which is either checked (\cmode) or unchecked (\umode) and a type
($\omega$) denoting valid values that can be pointed to. Array types include both the type of
elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
limited to integer literals $n$ and expressions $x + n$.
Whether an array pointer is null terminated or not is determined by annotation
$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
otherwise (we elide the $\cdot$ when writing the type). Here is the
corresponding \checkedc syntax for these types:
\[
\begin{array}{rcl}
$\code{array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\cmode}
\\[0.2em]
$\code{nt_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
\end{array}
\]
As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.

\lang expressions include literals ($n\!:\!\tau$), variables ($x$), memory
allocation ($\emalloc{\omega}$), let binding ($\elet{x}{e_1}{e_2}$),
static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$)
(assumed at compile-time and verified at run-time, see Sec.~\ref{sec:example}),
function calls ($\ecall{f}{\overline{e}}$), addition
($\ebinop{e_1}{e_2}$), pointer dereference and assignment ($\estar{e}$)
and ($\eassign{e_1}{e_2}$), resp.), unchecked blocks ($\eunchecked{e}$),
the \texttt{strlen} operation ($\estrlen{x}$), and conditionals ($\eif{e}{e_1}{e_2}$).
% \review{III.A.: the "dynamic cast" terminology may be briefly confused for C++'s
%   RTTI-based dynamic cast feature}
% \mwh{Added reference back to Section 2-B}

Integer literals $n$ are annotated with a type $\tau$ which can be either
$\tint$, or $\tptr{\omega}{m}$ in the case $n$ is being used as
a heap address (this is useful for the semantics);
$\evalue{0}{\tptr{\omega}{m}}$ (for any $m$ and $\omega$) represents
the $\enull$ pointer, as usual. 
The
$\texttt{strlen}$ expression operates on variables $x$
rather than arbitrary expressions to simplify managing
bounds information in the type system; the more general case can be
encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
casts; e.g., the following %

\code{dyn_bounds_cast<array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}

\noindent
becomes $\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$.

\lang aims to be simple enough to work with, but powerful enough to
encode realistic \checkedc idioms. For example, mutable local
variables can be encoded as immutable locals that point to the heap;
the use of \code{&} can be simulated with \code{malloc};
and loops can be encoded as recursive function calls. \code{struct}s are
not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
actually in our model, and developed in
\iftr
Appendix~\ref{appx:struct}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
C-style \code{union}s have no safe typing
in \checkedc, so we omit them. By default, functions are assumed to
be within checked regions; placing the body in an \code{unchecked}
expression relaxes this, and within that, checked regions can be
nested via function calls. Bounds are restricted slightly:
rather than allowing arbitrary sub-expressions, bounds must be either
integer literals or variables plus an integer offset, which accounts
for most uses of \code{bounds} in \checkedc programs. \lang bounds are
defined as relative offsets, not absolute ones, as in the second part
of Fig.~\ref{fig:checkedc-example}. We see no technical problem to
modeling absolute bounds, but it would be a pervasive change so we
have not done so.

  We have mechanized two models of \lang, one in Coq
  and one in PLT Redex~\cite{pltredex}, which is a semantic
  engineering framework implemented in Racket. Redex provides direct
  support for specifying the operational semantics and typing with logical
  rules, but then automatically makes them executable and subject to
  randomized testing, which is very useful during development.
  The model we present in the paper faithfully
  represents both mechanizations, but there are some differences for
  presentation purposes. For example, the paper and the Coq model use an explicit
  stack, whereas the Redex model uses \code{let} bindings to simulate
  one (simplifying term generation for randomized
  testing).
\iftr
  Appendix~\ref{app:model-diffs} 
\else
  The supplemental report~\cite{checkedc-tech-report}
\fi
  outlines the differences
  between the two models and the paper formalism.

% \mwh{NOTE: COULD WORK THE FOLLOWING INTO THE ABOVE DESCRIPTION OF
%   POINTERS: Array and NT-array types have two relative bounds, whose structures
% can be either an integer or a variable plus an integer. For example,
% if we have the expression
% $x\texttt{=}\emalloc{\tarrayb{(0,10)}{\tint}}$, $x$ then has the type
% $\tptr{\tarrayb{10}{\tint}}{\cmode}$, which represents an array
% pointer of size $10$. The bounds ($(b_l,b_h)$ in
% $\evalue{p}{\tptr{\tarrayb{(b_l,b_h)}{\tau}}{m}}$) indicate relative
% offsets from this pointer of the accessible memory, i.e., $p+b_l$ and
% $p+b_h$. As such, a pointer can only be directly dereferenced if 0 is
% included within the annotated range.  If following the $\emalloctext$
% operation, we execute $*(x-1)$ and $*(x\plus 10)$. The two expressions
% are not valid, because the type of the two expressions $(x-1)$ and
% $(x\plus 10)$ are $\tarrayptr{1}{11}{\tint}{\cmode}$ and
% $\tarrayptr{-10}{0}{\tint}{\cmode}$ and $0$ in these two cases are not
% in the ranges: $[1,11)$ and $[-10,0)$. Thus, in order for a $\cmode$
% pointer with type $\tarrayptr{b_l}{b_h}{\tint}{\cmode}$ to be
% accessible in \checkedc, $0$ must be in the range of $[b_l,b_h)$.}

\begin{figure}
{\small
$    \begin{array}{l}
    \begin{array}{lll}
\mu & ::= & \evalue{n}{\tau} \mid \bot\\
e & ::= & \ldots \mid \ret{x}{\mu}{e}\\
r & ::= & e \mid \enull \mid \ebounds\\
E &::=& \Box \mid \elet{x}{E}{e} \mid \ecall{f}{\overline{E}}\mid\ecast{\tau}{E} \mid \edyncast{\tau}{E} \\[0.2em]
&&\mid \ret{x}{\evalue{n}{\tau}}{E}\mid\ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid\estar{E}\mid\eassign{E}{e} \\[0.2em]
&&\mid\eassign{\evalue{n}{\tau}}{E}\mid\eunchecked{E}\mid\eif{E}{e}{e}\\[0.2em]
\overline{E} &::=& E \mid  \evalue{n}{\tau},\overline{E} \mid \overline{E}, e
\end{array}
\\ \\
    \end{array} 
$
  \begin{mathpar}
    \inferrule{ m=\mode(E) \\
      e=E[e'] \\
      (\varphi,\heap,e') \longrightarrow (\varphi',\heap',e'')}
    {(\varphi,\heap,e)\longrightarrow_{m} (\varphi',\heap',E[e''])}

    \inferrule{ m=\mode(E) \\
      e=E[\eif{\estar{x}}{e_1}{e_2}] \\
      (\varphi,\heap,\eif{\estar{x}}{e_1}{e_2})
              \longrightarrow (\varphi',\heap',e')}
    {(\varphi,\heap,e)\longrightarrow_{m} (\varphi',\heap',E[e'])}
    \;\;\;{[\emph{prefer}]}

  \end{mathpar}
}
  \caption{\lang Semantics: Evaluation}
  \label{fig:c-context}
\end{figure}

\subsection{Semantics}\label{sec:semantics}

% The semantics
% gives an independent account of spatial safety in \lang by
% checking pointer bounds based on the annotations carried on types at
% run-time.  While this account makes clear that bounds checking occurs
% as expected, it suggests an implementation that uses fat pointers to
% carry bounds.  We resolve this tension in the subsequent section on
% compilation and show that an implementation faithful to the semantics
% can be obtained without fat pointers.  
% \review{repeat that the stack is immutable at this point?}
% \liyi{Is it? Is the stack immutable? What does the immutable mean? 
%   In a stack, the variable values can be changed? Right?
%   The pointer address itself cannot be changed once it is created, but the stack variable content can be updated?  }
% \mwh{It certainly seems to be immutable: Your create stack frames
%   using let binding, and the let-bound variables will always be bound
%   to the same things. I.e., stack cells are immutable.}

% \review{this raises a fair amount of questions regarding the treatment of the
%   NULL pointer at this stage of the paper... is it modeled as 0, as returned by
%   `malloc`? are dynamic checks inserted by CheckedC to guarantee that no NULL
%   pointer is dereferenced?}
% \mwh{Yes, it is modeled as 0, and the semantics checks for
%   dereferences of 0. }

The operational semantics for \lang is defined as a small-step
transition relation with the judgment $ (\varphi,\heap,e)
\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
$\heap$ is a \emph{heap} mapping addresses (integer literals) to
values $\evalue{n}{\tau}$; for both we ensure
$\fv(\tau)=\emptyset$.
While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated; we can model mutable stack variables as pointers into the
mutable heap.
As mentioned, value $\evalue{0}{\tau}$
represents a $\enull$ pointer when $\tau$ is a pointer type;
correspondingly, $\heap(0)$ should always be undefined.
The relation steps to a \emph{result} $r$,
which is either an expression or a $\enull$ or $\ebounds$ failure,
representing a null-pointer dereference or out-of-bounds access,
respectively. Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterize undefined behavior.
%
The mode $m$ indicates whether the
stepped redex within $e$ was in a checked ($\cmode$) or
unchecked ($\umode$) region.

The rules for the main operational semantics
judgment---\emph{evaluation}---are given at the bottom of
Fig.~\ref{fig:c-context}.
The first rule takes an expression $e$, decomposes
it into an \emph{evaluation context} $E$ and a sub-expression $e'$
(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
$e$), and then evaluates $e'$ according to the \emph{computation}
  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
whose rules are given in Fig.~\ref{fig:semantics}, discussed
shortly. 
The second rule handles conditionals $\eif{\estar{x}}{e_2}{e_3}$ in
redex position specially, delegating directly to the \textsc{S-IfNTT} computation
rule, which supports bounds widening; we discuss this rule
shortly. When the second and first rules could both apply, we always
prefer the second.\footnote{This approach is that of the PLT Redex model of \lang; the Coq
development uses a slightly simpler syntax to achieve the same
effect.}
% \review{the special case raises questions, e.g. why is this syntax-driven and
%   not type-driven? }
% \liyi{This describes the semantic transition rules. We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3. like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.  }
% \mwh{Don't follow the above. I don't see this ``context transition
%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
The $\mode$ function
determines the mode when evaluating $e'$ based on the context $E$:
if the $\Box$ occurs within $(\eunchecked{E'})$ inside $E$, then the mode is
$\umode$; otherwise, it is $\cmode$. Evaluation contexts $E$ define a
standard left-to-right evaluation order. (We explain the
$\ret{x}{\mu}{e}$ syntax shortly.)
% For a term $e$, $C$ is a context if any
% only if it contains a $\Box$ hole term in one of the subterm position
% in $e$. $\Box$ indicates the break point where we split $e$ into $C$
% and a subterm $e'$ such that $C[e'] = e$.  For a function call
% $\ecall{f}{\overline{e}}$, a valid context ($\overline{C}$) is defined
% as any one of the elements in the list $\overline{e}$ being a
% context. We show a rule of the upper level \checkedc semantics. The
% rule splits a term $e$ into a context $C$ and redex $e'$.  If subterm
% $e'$ is transitioned to $e_a$, $C[e']$ is transitioned to
% $C[e_a]$.

\begin{DIFnomarkup}
\begin{figure*}[t]
{\small
  \begin{mathpar}
        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} }
    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}

    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}

        \inferrule[S-DefNTArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h]}
    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}

    \inferrule[S-AssignArr]{\heap(n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}

    \inferrule[S-Cast]
              {}
              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}

  \inferrule[S-DynCast]{
\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{m})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{m} \\ n_l' \le n_l\ \\ n_h \le n_h'}
    { (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{m}}}) \longrightarrow (\varphi',\heap',\evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau_b}{m}})}

 \inferrule[S-DynCastBound]{\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{\cmode})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{\cmode} \\ n_l' > n_l \vee n_h > n_h'}{ (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{\cmode}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{\cmode}}}) 
 \longrightarrow (\varphi',\heap',\ebounds)}

        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}

    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}

    \inferrule[S-Fun]{ \Xi(f) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
        {(\varphi,\heap,\ecall{f}{{\evalue{\overline{n}}{\overline{\tau_a}}}}) \longrightarrow
   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}

     % \inferrule[S-IfNTF]{\varphi(x)=\evalue{n}{\tau} \\ \heap(n)=0}
     %      {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}
      \inferrule[S-StrWiden]{
         \varphi(x) = \evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{\cmode}}\\ 0 \in [n_l,n_h] \\ n_a > n_h
\\ \heap(n+n_a) = 0 
\\\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
    {(\varphi,\heap,\estrlen{x}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tarrayptr{n_l}{n_a}{\tau}{\cmode}}],\heap,\evalue{n_a}{\tint})}

     \inferrule[S-IfNTT]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(n)\neq 0}
    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}

 % \inferrule[S-IfNTNot]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}} \\ \heap(n)\neq 0\\ 0 < n_h}
 %            {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}

\end{mathpar}
}
% {\footnotesize
% \begin{center}
% $
% \begin{array}{l}
% \tau[\overline{n} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall n_i\in\overline{n}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \Rightarrow \tau[n_i / x_i]\\[0.2em]
% \mathtt{let}\;\overline{x}=\overline{e}\;\mathtt{in}...\triangleq \mathtt{let}\;x_0=e_0\;\mathtt{in}\;\mathtt{let}\;x_1=e_1\;\mathtt{in}...
% \end{array}
% $
% \end{center}
% }
\caption{\lang Semantics: Computation (Selected Rules)}
\label{fig:semantics}
\end{figure*}
\end{DIFnomarkup}

% \review{Fig4: it was hard to tell which cases were stuck states, or could reduce
%   owing to a rule that was not shown}
% \mwh{Stuck states are those where the expression is a non-value and
%   not $\enull$ or $\ebounds$. Updated III-B and III-D.}
% \review{Fig4: can the rules be presented in the same order they were introduced in
%   the paper?}
% \liyi{ Reordered }

% \review{Fig4: S-FUN: $\vec\tau_a$ seems unused; why?}
% \liyi{the list of $\tau_a$ is a list of input argument types. These
%   are used during type checking, but not during evaluation (as is
%   typical).}

% LEO: This has an overfull line...?
\begin{figure}[t]
{\small
{\captionsetup[lstlisting]{margin = 8 mm}
  \begin{lstlisting}[xleftmargin=8 mm]
nt_array_ptr<char> safe_strcat
   (nt_array_ptr<char> dst : count(n),
    nt_array_ptr<char> src : count(0), int n) {
  int x = strlen(dst);
  int y = strlen(src);
  nt_array_ptr<char> c : count(n) =
    dyn_bounds_cast
           <nt_array_ptr<char>>(dst,count(n));
    // sets c == dst with bound n (not x)
  if (x+y < n) {
    for (int i = 0; i < y; ++i)
      *(c+x+i) = *(src+i);
    *(c+x+y) = '\0';
    return dst;
  }
  return null;
}
  \end{lstlisting}
}
}
\caption{Implementation of safe \code{strcat}}
\label{fig:strcat-ex}
\end{figure}

Fig.~\ref{fig:semantics} shows selected rules for the computation
relation; we explain them with the help of the example in
Fig.~\ref{fig:strcat-ex},
  which defines a 
  safe version of \code{strcat} (using actual Checked C syntax).  The
  function takes a target 
  pointer \code{dst} of capacity \code{n}, where the first null
  character (determined by \code{strlen}) is at index \code{x} where
  $0 \leq $\code{x}$ \leq n$. It concatenates the \code{src} buffer to
  the end of \code{dst} as long as \code{dst} has sufficient space.

% Below, we introduce low-level transition semantics for some case operations. The design of the low-level individual operation semantics is carefully engineered to perform match our compiler's behavior, such as correctly characterizing the bound widening behaviors for NT-array pointers, even though it is written in terms of fat-pointer formalization.

\myparagraph{Pointer accesses}
%
The rules for dereference and assignment operations---\textsc{S-Def},
\textsc{S-DefNull}, \textsc{S-DefNTArray}, and
\textsc{S-AssignArr}---illustrate how the semantics checks bounds.
Rule \textsc{S-DefNull} transitions attempted null-pointer
dereferences to $\enull$, whereas \textsc{S-Def} dereferences a
non-null (single) pointer.
  When $\enull$ is returned by the
computation relation, the evaluation relation halts the entire
evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
does likewise when $\ebounds$ is returned (see below).

% \review{- on page 6, section IIIC, paragraph "Pointer Access" mentions that checked pointers cannot be dereferenced in unchecked blocks - this looks funny, shouldn't it be the other way around? The Coq code contains the hypothesis m'=Unchecked -> m Unchecked in various rules of definition well-typed (BoundCheckedC, line 669; rule TyDeref in the code seems closest to the figure's T-DefArr and T-Def in the appendix, although it's a bit concerning that there's no 1-to-1 correspondence of the rules in the code and the paper).
% }
% \liyi{Yeah. It is another typo. It should be the other way around.  }

\textsc{S-AssignArr} assigns to an array as long as 0 (the point of
dereference) is within the bounds designated by the pointer's annotation
and strictly less than the upper bound. For the assignment rule,
arrays are treated uniformly whether they are null-terminated or not
($\kappa$ can be $\cdot$ or $nt$)---the semantics does not search past
the current position for a null terminator. The program
can widen the bounds as needed, if they currently precede the null
terminator: \textsc{S-DefNTArray}, which dereferences an NT array
pointer, allows an upper bound of $0$, since the program may read, but
not write, the null terminator. A separate rule (not shown) handles
normal arrays.

\myparagraph{Casts}
%
Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
by \textsc{S-Cast}. In a type-correct program, such casts are
confirmed safe by the type system. To evaluate a cast, the rule
updates the type annotation on $n$. Before doing so, it must
``evaluate'' any variables that occur in $\tau$ according to their
bindings in $\varphi$. For example, if $\tau$ was
$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.

Dynamic casts are accounted for by \textsc{S-DynCast} and
\textsc{S-DynCastBound}. In a type-correct program, such casts are
assumed correct by the type system, and later confirmed by the
semantics. As such, a dynamic cast will cause a $\ebounds$ failure if
the cast-to type is incompatible with the type of the target pointer,
as per the $n_l' > n_l \vee n_h > n_h'$ condition in
\textsc{S-DynCastBound}. An example use of dynamic casts is given on
line~7 in Fig.~\ref{fig:strcat-ex}. The values of \code{x} and
\code{n} might not be known statically, so the type system cannot
confirm that \code{x <= n}; the dynamic cast assumes this inequality
holds, but then checks it at run-time.

\myparagraph{Binding and Function Calls}
%
The semantics handles variable scopes using the special $\erettext$
form. \textsc{S-Let} evaluates to a configuration whose stack
is $\varphi$ extended with a binding for $x$, and whose expression is
$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
$\varphi(x)$; if it had no previous binding, $\varphi(x) =
\bot$. Evaluation proceeds on $e$ until it becomes a literal
$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
binding (or $\bot$) in the new stack, and evaluates to
$n\!:\!\tau$.

Function calls are handled by \textsc{S-Fun}. Recall that array
bounds in types may refer to in-scope variables; e.g., parameter
\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
2-3 in Fig.~\ref{fig:strcat-ex}. A call to function $f$ causes $f$'s
definition to be retrieved from $\Xi$,
% LEO: This is the first occurence of \Xi. It is explained,
% but reads a bit awkward as it's unclear this is its "introduction"
which maps function names to
forms $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e$, where
$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
is the parameter list of variables and their types, and $e$ is the
function body. The call is expanded into a \texttt{let} which binds
parameter variables $\overline{x}$ to the actual arguments
$\overline{n}$, but annotated with the parameter types
$\overline{\tau}$ (this will be safe for type-correct programs). The
function body $e$ is wrapped in a static cast
$(\tau[\overline{n} / \overline{x}])$ which is the function's return
type but with any parameter variables $\overline{x}$ appearing in that
type substituted with the call's actual arguments $\overline{n}$. To
see why this is needed, suppose that \code{safe_strcat} in
Fig.~\ref{fig:strcat-ex} is defined to return a
\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
perform a \code{safe_strcat} function call as
\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
function returns a value with type \code{nt_array_ptr<int>:count(10)}
because we substitute bound variable \code{n} in the 
defined return type with \code{10} from the function call's
argument list.
Note that the \textsc{S-Fun} rule replaces the
  annotations $\overline{\tau_a}$ with
  $\overline{\tau}$ (after instantiation) from the function's
  signature. Using $\overline{\tau_a}$ when executing the body of
the function has no impact on the soundness of \lang, but will violate
Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.

\myparagraph{Bounds Widening}
%
Bounds widening occurs when branching on a dereference of a NT array
pointer, or when performing $\estrlentext$. The latter is most useful
when assigned to a local variable so that subsequent code can use the
result, e.g., $e$ in $\elet{x}{\estrlen{y}}{e}$.  Lines 4 and 5 in
Fig.~\ref{fig:strcat-ex} are examples. 
The widened upper bound
precipitated by $\estrlen{y}$ is extended beyond the lifetime of $x$, as long as $y$ is live.
% Even if $x$ returns, if $y$'s upper bound keeps the widened bound that is caused by $\estrlen{y}$.
For example, \code{x}'s scope in line 4 at runtime is
the whole function body in \code{safe_strcat} because the lifetime of the pointer \code{dst} is in the function body.
% Allowing widening to persist
% outside the scope is sound, % but would precipitate the need for
% % fat pointers during compilation, which we want to avoid.
% and we take this approach to give more precise run-time behavior. We
% show in Sec.~\ref{sec:compilation}
This is different from the \checkedc specification, 
which only allows bound widening to happen within the scope of \code{x}, and restoring old bound values once \code{x} dies.
We allow widening to persist outside the scope at run-time as long as
we are within the stack frame, and we show
this does not necessarily require the use of fat pointers in
Sec.~\ref{sec:compilation}.

Rule \textsc{S-StrWiden} implements $\estrlentext$ widening. The
predicate
$\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i)
= \evalue{n_i}{\tau_i} \wedge n_i \neq 0))$ aims to find a position
$n+n_a$ in the NT array that stores a null character, where no
character as indexes between $n$ and $n+n_a$ contains one. (This rule
handles the case when $n_a > n_h$, the $n_a \le n_h$ case is handled
by a normal $\estrlentext$ rule; see 
\iftr
Appendix~\ref{sec:rem-semantics}.)
\else
the supplemental report~\cite{checkedc-tech-report}.)
\fi

Rule \textsc{S-IfNTT} performs bounds widening on $x$ when the
dereference $\estar{x}$ is not at the null terminator, but the
pointer's upper bound is 0 (i.e., it's at the end of its known
range). $x$'s upper bound is incremented to 1, and this count persists
as long as $x$ is live.  For example, \code{s}'s increment (lines
5--8) is live until the return of the function in
Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
\code{s}'s upper bound is properly extended. %  For this rule and
% \textsc{S-StrWiden}, this widening persists in the current stack
% frame. When $x$ goes out of scope, .

% \textsc{S-IfNTF} does not widen when seeing null; rule
% \textsc{S-IfNTNot} sees a non-null character, but the pointer is not
% at its upper bound, so the bounds cannot be widened.

% \ignore{
% Fig.~\ref{fig:semantics} provides the low-level semantic rules for operations involving NT-array pointers, mainly, the $\estrlentext$ and $\eiftext$ operations. The semantics has concurred the ambiguity in the \checkedc specification, e.g., we define the exact behavior of the $\estrlentext$ operation to return the length between the current pointer position and the first null-character.
% We also utilize new technique in our compiler so that the scope of the bound widening behavior in our formalization is a little longer. More details are in Sec.~\ref{sec:compilation}.

% The first rule defines the evaluation behavior of a $\estrlentext$ operation. Given a pointer $x$ with its type $\tntarrayptr{0}{n_h}{\tau}{m}$, the application of such operation takes the address of the pointer $x$, and search incrementally the heap positions next to the address $x$ until we find a $0$ value (representing a null character). We return the value $n_a$ as the length, and update the bound information in the stack for $x$. In the compilation, we use a ghost variable to record such bound changes without using fat-pointer implementations.

% The last three rules in Fig.~\ref{fig:semantics} describe the semantic behaviors of an $\eiftext$ branching operation when the Boolean guard is a dereference of an NT-array pointer. The first one states that if the type upper bound of the pointer $x$ is $0$, and the pointer data value $n_a$ is not $0$, we can conclude that the upper bound is not the last position of the NT-array pointer, so we can then update $1$ in the upper bound while jump to the $\etrue$ branch. The second rule describes that we do not extend the upper bound if the upper-bound of the type of $x$ is not zero because we know that we are not in the NT-array's last position. The third rule describes the behavior of jumping to the $\efalse$ branch when the pointer content is $0$. In this case, we also do not need to increase the upper-bound of the type of $x$.}

\subsection{Typing}\label{sec:type-system}

%% DVH: covered above.
%% Any \checkedc expression is a word type object ($\tau$ in Fig.~\ref{fig:checkc-syn}), which is either an \tint{} or a pointer. A pointer can be either a word type $\tau$, an array $\tarray{b}{b}{\tau}$, or an NT-array $\tntarray{b}{b}{\tau}$. Each pointer type is associated with a mode $m$ indicating whether the pointer is checked (\cmode) or unchecked ($\umode$).

We now turn to the \lang type system.
% The design of the type
% system has been carefully constructed to ensure the expected
% properties of progress and preservation, but also so that the
% type-based compilation strategy detailed later is correct and
% economical in its representation of pointers.
%
The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
which states that in a type environment $\Gamma$ (mapping variables to
their types) and a predicate environment $\Theta$ (mapping integer-typed
variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
in mode $m$. Key rules for this judgment are given in
Fig.~\ref{fig:type-system-1}. In the rules, $m \le m'$ uses
the two-point lattice with $\umode < \cmode$.
% \review{text says $u < c$ but this seems to contradict the paragraph right
%   after, because if $m = u$ then $u <= c$ and it seems like T-DefArr *does*
%   allow me to dereference a checked pointer in unchecked mode...? Or am I
%   missing something?}
% \liyi{ Why it will be a problem in dereferencing a checked pointer in unchecked mode?
%    unchecked regions contain history C code that is not translated to Checked-C syntax, but it could also mean that the translation is under construction, so that we are free to put checked-C code there, and we gaurantee that the checked-pointers in unchecked-mode are also spatially safe. The opposite is not true, where you cannot use (dereference/assign) a unchecked-pointer in a checked mode. }
All remaining rules
are given in
\iftr
Appendix~\ref{sec:literal-pointer-typing}~and~\ref{rem-type}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi

\begin{DIFnomarkup}
\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[T-DefArr]
              {m \leq m' \\\\ \Gamma;\Theta \vdash_{m} e : \tptr{\tallarrayb{\bvar}{\tau}}{m'}}
              {\Gamma;\Theta \vdash_m \estar{e} : \tau}

    \inferrule[T-AssignArr]
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\\\
                \Gamma; \Theta \vdash_m e_2 : \tau' \\
                \tau'\sqsubseteq \tau\\
                m \leq m'}
              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}

    \inferrule[T-Unchecked]
              {\Gamma;\Theta \vdash_u e : \tau}
              {\Gamma;\Theta \vdash_m \eunchecked{e} : \tau}

    \inferrule[T-Cast]
              {m = \cmode \Rightarrow \tau \neq\tptr{\tau''}{\cmode} \text{ for any $\tau''$} \\\\
                \Gamma;\Theta \vdash_m e : \tau'}
              {\Gamma;\Theta \vdash_m \ecast{\tau}{e} : \tau}

     \inferrule[T-CastCheckedPtr]
               {\Gamma;\Theta \vdash_m e : \tau' \\
                 \tau' \sqsubseteq \tptr{\tau}{\cmode}}
               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\cmode}}{e} : \tptr{\tau}{\cmode}}

      \inferrule[T-DynCast]
                {\Gamma;\Theta \vdash_m e : \tptr{\tallarrayb{\bvar'}{\tau}}{m}}
                {\Gamma;\Theta \vdash_m \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e} : \tptr{\tallarrayb{\bvar}{\tau}}{m}}

      \inferrule[T-If ]
        {\Gamma; \Theta \vdash_m e : \tau \\\\ \Gamma; \Theta \vdash_m e_1 : \tau_1 \\\\ \Gamma; \Theta \vdash_m e_2 : \tau_2}
        {\Gamma; \Theta \vdash_m \eif{e}{e_1}{e_2} : \tau_1 \sqcup \tau_2}

      \inferrule [T-IfNT]
                { \Gamma; \Theta \vdash_m x : \tptr{\tntarrayb{(b_l,0)}{\tau}}{c} \\\\
                   \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}]; \Theta \vdash_m e_1 : \tau_1\\\\
                  \Gamma;\Theta \vdash_m e_2 : \tau_2 }
                {\Gamma;\Theta \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_1 \sqcup \tau_2}

      \inferrule[T-Str]
                {\Gamma; \Theta \vdash_m e :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} 
                }
                {\Gamma;\Theta \vdash_m \estrlen{e} : \tint}

      \inferrule[T-LetStr]
                {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\\\
                 \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}];\Theta[x\mapsto \texttt{ge}\_0] \vdash_m e : \tau}
                {\Gamma;\Theta \vdash_m \elet{x}{\estrlen{y}}{e} : \tau}

   \inferrule[T-Let]
    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
        \Gamma;\Theta \vdash_m e_1 : \tau \\
           \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
             }
    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}

\inferrule[T-Fun]
    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\
        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
             \overline{\tau'} \sqsubseteq 
               \overline{\tau}[\overline{e} / \overline{x}]}
    {\Gamma; \Theta \vdash_m f(\overline{e}) : \tau[\overline{e} / \overline{x}]}

                \inferrule[T-Ret]
    {\Gamma(x)\neq \bot \\
          \Gamma;\Theta \vdash_m e : \tau}
    {\Gamma;\Theta \vdash_m \eret{x}{\mu}{e} : \tau}

  \end{mathpar}
}
% {\footnotesize
% \begin{center}
% $
% \begin{array}{l}
% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
% \\[0.2em]
% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
% \end{array}
% $
% \end{center}
% }
\caption{Selected type rules}
\label{fig:type-system-1}
\end{figure*}
\end{DIFnomarkup}

\myparagraph{Pointer Access}
%
Rules \textsc{T-DefArr} and \textsc{T-AssignArr} type-check array
dereference and assignment operations resp., returning the type of
pointed-to objects; rules for pointers to single objects are
similar.
The condition $m\le m'$ ensures that unchecked pointers can only
be dereferenced in unchecked blocks;
the type rule for
$\eunchecked{e}$ sets $m = \umode$ when checking $e$.
The rules do not attempt to reason whether the access is in bounds;
this check is deferred to the semantics.

% Subtyping and casting operations are briefly introduced in
% Sec.~\ref{sec:intros}~and~\ref{sec:overview}.  Subtyping is useful in
% static casting operations that allow users to view a pointer in one
% type as another, such as casting an NT-array pointer to an array
% one. \checkedc provides a set of safe static casting operations that
% have no cost in execution.  Moreover, subtyping acts as oracles for
% bound widening and dynamic casting operations; thus, \checkedc is
% different from a complete static array pointer bound system.  For
% example, if $e$ has type $\tau'$ and $\varphi$ is the current stack
% snapshot, the semantics of $\edyncast{\tau}{e}$ does not transition to
% an error state when $\varphi(\tau')\sqsubseteq\varphi(\tau)$.  In a
% function call, for every argument, \lang permits users to input a
% subtype entity and we prove that this does not affect the correctness
% of the program.

\myparagraph{Casting and Subtyping}
%
Rule \textsc{T-Cast} rule forbids casting to checked pointers when in checked
regions (when $m = \cmode$), but $\tau$ is unrestricted when
$m = \umode$. The \textsc{T-CastCheckedPtr} rule
permits casting from an expression of type $\tau'$ to a checked pointer when
$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype}; the many
rules ensure the relation is transitive. Most of the rules handle
casting between array pointer types. The second rule 
$0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
\tarrayptr{b_l}{b_h}{\tau}{m}$ permits treating a singleton
pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.

Since bounds expressions may
contain variables, determining assumptions like $b_l \leq b_l'$
requires reasoning about those variables' possible values. The type
system uses $\Theta$ to make such reasoning more precise.\footnote{Technically, the subtyping relation $\sqsubseteq$ and the bounds
  ordering relation $\leq$ are parameterized by $\Theta$; this fact is
  implicit to avoid clutter.} $\Theta$ is a map from variables $x$ to
predicates $P$, which have the form $P ::= \top \;|\; \texttt{ge}\_0$.
If $\Theta$ maps $x$ to $\top$, that means that the variable can
possibly be any value; $\texttt{ge}\_0$ means that $x \ge 0$. We will
see how $\Theta$ gets populated and give a detailed example of
subtyping below.\footnote{As it turns out, the subtyping relation is
  also parameterized by $\varphi$, which is needed when type checking
  intermediate results to prove type preservation; source programs
  would always have $\varphi = \emptyset$. Details are in
\iftr
  Appendix~\ref{app:le}.
\else
  the supplemental report~\cite{checkedc-tech-report}.
\fi
}

Rule \textsc{T-DynCast} typechecks dynamic casting operations, which
apply to array pointer types only. The cast is accepted by the type
system, as its legality will be checked by the semantics.

\begin{DIFnomarkup}
\begin{figure}
{\small
\begin{center}
  $\begin{array}{r@{~}c@{~}l@{~}c@{~}l}
    &&\multicolumn{1}{r}{\tau} &\sqsubseteq& \tau\\[0.6em]
    0\le b_l \wedge b_h \le 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.6em]
    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
    b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
    b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
    \end{array}
  $
\end{center}
}
  \caption{Subtyping}
  \label{fig:checkc-subtype}
\end{figure}
\end{DIFnomarkup}

\myparagraph{Bounds Widening}
%
The bounds of NT array pointers may be widened at conditionals, and calls to $\estrlentext$.  Rule \textsc{T-If} handles normal
branching operations; rule \textsc{T-IfNT} is specialized to the case
of branching on $\estar{x}$ when $x$ is a NT array pointer whose upper
bound is 0. In this case, true-branch $e_1$ is checked with $x$'s type
updated so that its upper bound is incremented by 1; the else-branch
$e_2$ is type-checked under the existing assumptions. For both rules,
the resulting type is the join of the types of the two branches
(according to subtyping). This is important for the situation when $x$
itself is part of the result, since $x$ will have
different types in the two branches.

% \review{"Bounds Widening": can the join be always computed? this may be a hard
%   problem...?}
%   \mwh{If bounds expressions on two versions of an array pointer
%   are very different, a precise join may not be possible, but an
%   imprecise one is always possible by dropping the bounds
%   expression entirely. Right Liyi?}

Rule \textsc{T-Str} handles the case for when $\estrlentext(y)$ does
not appear in a let binding. Rule \textsc{T-LetStr} handles the case
when it does, and performs bounds widening. The result of the call is
stored in variable $x$, and the type of $y$ is updated in $\Gamma$ when
checking the let-body $e$ to indicate that $x$ is $y$'s upper
bound. Notice that the lower bound $b_l$ is unaffected by the call to
$\estrlentext(y)$; this is sound because we know that $\estrlentext$
will always return a result $n$ such that $n \ge b_h$, the current
view of $x$'s upper bound.
The type rule tracks $\estrlentext$'s widened bounds
within the scope of $x$, while the bound-widening effect in the semantics applies to the lifetime of $y$.
Our type preservation theorem in Sec.~\ref{sec:theorem}
shows that our type system is a sound model of the \CoreChkC semantics,
and we discuss how we guarantee that the behavior of our compiler formalization and the semantics matches in Sec.~\ref{sec:compilation}.

This rule also extends $\Theta$ when checking $e$, adding a predicate
indicating that $x\ge 0$. To see how this information is used,
consider this example.  The \code{return} on line 14 of
Fig.~\ref{fig:strcat-ex} has an implicit static cast from the
returned expression to the declared function type (see rule
\textsc{T-Fun}, described below). In type checking the $\estrlentext$
on line 4, we insert a predicate in $\Theta$ showing
\code{x}$\;\ge 0$.  The static cast on line 14 is valid according to
the last line in Fig.~\ref{fig:checkc-subtype}:

{\small
$$
\tallarrayptr{0}{x}{\tau}{c} \sqsubseteq \tallarrayptr{0}{0}{\tau}{c}
$$
}
because $0 \leq 0$ and $0 \leq x$, where the latter holds since
$\Theta$ proves $x \ge 0$.  Without $\Theta$, we would need a
dynamic cast.

In our formal presentation, $\Theta$ is quite simple and is just meant
to illustrate how static information can be used to avoid dynamic
checks; it is easy to imagine richer environments of facts that can be
leveraged by, say, an SMT solver as part of the subtyping
check \cite{LiquidHaskell,Refinementlh}.

% There is a difference between the \checkedc and \lang type systems in
% terms of recognizing subtyping due to $\Theta$. \code{strncat_c} in
% Fig.~\ref{fig:strcat-ex} represents what the \checkedc specification
% can do, while \code{strncat} represents what \lang can do. In the
% usage of a static casting operation in line 16 of
% Fig.~\ref{fig:strcat-ex}, \lang formalization is smart enough to
% infer that \code{x} must be a non-negative number because the creation
% of \code{x} is from a \code{strlen} call.  \checkedc does not perform
% predicate decision procedure to reason about the bounds statically;
% thus, a dynamic casting is required in the case, such as line 30 in
% \code{strncat_c} in Fig.~\ref{fig:strcat-ex}.  More differences are
% in Sec.~\ref{sec:disc}.

% Here, we introduce
% two most common NT-array operations: $\estrlentext$ and branching
% operations.  Line 4 in Fig.~\ref{fig:strcat-ex} shows the bound
% widening behavior of the NT-array pointer usage in a $\estrlentext$
% expression.  The upper-bound of pointer \code{a} is updated to
% \code{x} through out the whole function body, which is the scope of
% variable \code{x}.  Line 5-8 in Fig.~\ref{fig:checkedc-example} shows
% the bound widening behavior of using NT-array pointers in a branching
% operation. In \checkedc, if a Boolean guard has the form $\estar{x}$,
% and $x$'s upper bound is $0$, then the \code{true} branch recognizes
% the upper bound of $x$ incremented by $1$, because the end position in
% the NT-array pointed to by $x$ is not a null-character in the
% \code{true} branch; thus, there is at least one more valid position in
% the NT-array after the current end position. More details discussion
% are in Sec.~\ref{sec:disc}.

\myparagraph{Dependent Functions and Let Bindings}
%
Rule \textsc{T-Fun} is the standard dependent function call rule. It
looks up the definition of the function in the function environment
$\Xi$, type-checks the actual arguments $\overline{e}$ which have
types $\overline{\tau'}$, and then confirms that each of these types
is a subtype of the declared type of $f$'s corresponding parameter. Because
functions have dependent types, we substitute each parameter $e_i$ for
its corresponding parameter $x_i$ in both the parameter types and the
return type. Consider the \code{safe_strcat} function in
Fig.~\ref{fig:strcat-ex}; its parameter type for \code{dst} 
depends on \code{n}. The \textsc{T-Fun} rule will substitute 
\code{n} with the argument at a call-site.

Rule \textsc{T-Let} types a $\elettext$ expression, which also admits
type dependency. In particular, the result of evaluating a $\elettext$
may have a type that refers to one of its bound variables (e.g., if
the result is a checked pointer with a variable-defined bound); if so,
we must substitute away this variable once it goes out of scope. Note
that we restrict the expression $e_1$ to syntactically match the
structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).

Rule \textsc{T-Ret} types a $\erettext$ expression, which does not
appear in source programs but is introduced by the semantics when
evaluating a let binding (rule \textsc{S-Let} in
Fig.~\ref{fig:semantics}); this rule is needed for the preservation
proof. After the evaluation of a let binding a variable $x$ concludes,
we need to restore any prior binding of $x$, which is either
$\bot$ (meaning that there is no $x$ originally) or some value
$\evalue{n}{\tau}$.

\subsection{Type Soundness and Blame}\label{sec:theorem}

% Before we present our main theorems, we need to first
% discuss the meaning what a pointer being well-typed in a given heap
% snapshot $\heap$ means, which is captured by rules in
% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
% simply checks if a given variable has the defined type in $\Gamma$;
% the constant rule ($\textsc{T-Const}$) is slightly more involved.
% First, it ensures that the type annotation $\tau$ does not contain any
% free variables. More importantly, it ensures that the pointer points
% to a location that makes sense in a given heap.
%  
%  
%  The $\size$ function in Fig.~\ref{fig:const-type}
% refers to the \code{sizeof} function in C computing the number of
% bytes for a type.
%  
%  
%  Second, we
% require that any constant ($\evalue{n}{\tau}$) should make sense in
% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
% constant set containing the constants that have been verified by the
% relation. For every constant $\evalue{n}{\tau}$, it is either an
% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
% zero-valued number ($n=0$), checked in $\sigma$
% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
% location $n+i$ in the range of the pointer (if $\omega$ is a word
% type, range is $[0,1)$; if $\omega$ is an array type
%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
%       \vdash n_a : \tau_a$.
%  
%  
% \begin{figure}[t]
% {\small
% \text{Type Rules for Constants and Variables:}
% \begin{mathpar}
%   \inferrule[T-Var]
%       {x : \tau \in \Gamma}
%       {\Gamma;\Theta \vdash_m x : \tau}
%  
%   \inferrule[T-Const]
%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
% \end{mathpar}
%     
% \text{Rules for Checking Constant Pointers In Heap:}
% \begin{mathpar}
%   \inferrule
%       {}
%       {\sigma \vdash n : \tint}
%  
%   \inferrule
%       {}
%       {\sigma \vdash n : \tptr{\omega}{\umode}}
%  
%   \inferrule
%       {}
%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
%  
%   \inferrule
%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
%  
%   \inferrule
%       {\forall i \in [0,\size(\omega)) .
%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
% \end{mathpar}
% }
% \caption{Type Rules for Checking Constants/Variables}
% \label{fig:const-type}
% \end{figure}

% \review{
%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
%   something, I didn't see such a definition in the paper.}
% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
%   assumption of well typing; we have added more details about that.}

In this subsection, we focus on our main meta-theoretic results about
\lang: type soundness (progress and preservation) and
blame.
  These proofs have been carried out in our
  Coq model, found at \url{https://github.com/plum-umd/checkedc}.

The type soundness theorems rely on several notions of
\emph{well-formedness}:

\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
A type environment $\Gamma$ is well-formed iff every variable mentioned as type bounds in $\Gamma$ are bounded by $\tnat$ typed variables in $\Gamma$.
\end{defi}

\begin{defi}[Heap Well-formedness]
A heap $\heap$ is well-formed iff (i) $\heap(0)$ is undefined, and
(ii) for all $\evalue{n}{\tau}$ in the range of $\heap$, type $\tau$
contains no free variables. 
\end{defi}

\begin{defi}[Stack Well-formedness]
A stack snapshot $\varphi$ is well-formed iff
for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
contains no free variables. 
\end{defi}

We also need to introduce a notion of
\emph{consistency}, relating heap environments before and after a
reduction step, and type environments, predicate sets, and stack
snapshots together.

\begin{defi}[Stack Consistency]
A type environment $\Gamma$, variable predicate set $\Theta$, and
stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
\varphi$---iff for every variable $x$, $\Theta(x)$ is defined implies
$\Gamma(x) = \tau$ for some $\tau$ and 
$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq \tau$. 
\end{defi}

\begin{defi}[Stack-Heap Consistency]
A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---iff
for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ implies $\heap;\emptyset \vdash n:\tau$.
\end{defi}

\begin{defi}[Heap-Heap Consistency]
A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---iff
for every constant $n$, $\heap;\emptyset \vdash n:\tau$ implies $\heap';\emptyset \vdash n:\tau$.
\end{defi}

Moreover, as a program evaluates, its expression may contain literals
$\evalue{n}{\tau}$ where $\tau$ is a pointer type, i.e., $n$ is an
index in $\heap$ (perhaps because $n$ was chosen by
$\mathtt{malloc}$). The normal type-checking judgment for $e$ is
implicitly parameterized by $\heap$, and the rules for type-checking
literals confirm that pointed-to heap cells are compatible with
(subtypes of) the pointer's type annotation; in turn this check may
precipitate checking the type consistency of the heap itself. We
follow the same approach as \citet{ruef18checkedc-incr}, and show the
rules in
  Fig.~\ref{fig:const-type}; the judgment
  $\heap;\sigma \vdash n : \tau$ is used to confirm literal
  well-typing, where $\sigma$ is a set of pointer literals already
  checked in $\heap$ (to allow pointer cycles). See
  \iftr
  Appendix~\ref{sec:literal-pointer-typing}
  \else
  the supplemental report~\cite{checkedc-tech-report}
  \fi
   for further discussion.

Progress now states that terms that don't reduce are either values or their
mode is unchecked:

\begin{thm}[Progress]\label{thm:progress}

For any \checkedc program $e$, heap $\heap$, stack
$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
that are all are well-formed, consistent
($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
one of the following holds:

\begin{itemize}

\item $e$ is a value ($\evalue{n}{\tau}$).

\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.

\item $m = \umode$, or there exists $E$ and $e'$, such that $e = E[e']$ and $\mode(E) = \umode$.

\end{itemize}
\end{thm}
{\em Proof:} By induction on the typing derivation.

\begin{DIFnomarkup}
 \begin{figure}[t]
 {\small
 \text{Type rules for constants and variables:}
 \begin{mathpar}
   \inferrule[T-Var]
       {x : \tau \in \Gamma}
       {\Gamma;\Theta \vdash_m x : \tau}

   \inferrule[T-Const]
       {\fv(\tau) = \emptyset \\ \heap;\emptyset \vdash n : \tau}
       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
 \end{mathpar}

 \text{Rules for checking constant pointers:}
 \begin{mathpar}
   \inferrule
       {}
       {\heap;\sigma \vdash n : \tint}

   \inferrule
       {}
       {\heap;\sigma \vdash n : \tptr{\omega}{\umode}}

   \inferrule
       {}
       {\heap;\sigma \vdash 0 : \tptr{\omega}{\cmode}}

   \inferrule
       {(\evalue{n}{\tptr{\omega}{\cmode}})\in \sigma}
       {\heap;\sigma \vdash n : \tptr{\omega}{\cmode}}

   \inferrule
       {\forall i \in [0,\size(\omega)) .
            \heap;(\sigma \cup \{(n:\tptr{\omega}{\cmode})) \}\vdash \heap(n+i)}
       {\heap;\sigma \vdash n : \tptr{\omega}{\cmode}}
 \end{mathpar}
 }
 \caption{Type Rules for Constants/Variables}
 \label{fig:const-type}
 \end{figure}
\end{DIFnomarkup}

\noindent
Preservation states that a reduction step preserves both the
type and consistency of the program being reduced.

\begin{thm}[Preservation]
For any \checkedc program $e$, heap $\heap$, stack
$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
that are all are well-formed, consistent
($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
$\heap'$ and $e'$, such that $(\varphi,\heap,e)
\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
$\Gamma'$, $\Theta'$ and $\tau'$ that are well formed, consistent
($\Gamma';\Theta'\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
well typed ($\Gamma';\Theta' \vdash_{\cmode} e: \tau'$), where
$\tau'\sqsubseteq \tau$.
\end{thm}
{\em Proof:} By induction on the typing derivation.
\smallskip

Using these two theorems we can prove our main result, {\em blame},
which states that
  if a well-typed program is \emph{stuck}---expression
$e$ is a non-value that cannot take a step\footnote{Note that
  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
  safety.}---the cause must be the
(past or imminent) execution of code in an unchecked region.

% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r. Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
% \liyi{It is a typo. We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region. This is a bit due to the space limitation. The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement. }

\begin{thm}[The Blame Theorem]\label{thm:blame} For any \checkedc
  program $e$, heap $\heap$, stack
$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
that are well-formed and consistent
($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
\tau$ for some $\tau$) and there exists
$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$ and $r$ is \emph{stuck}, then
there exists $j \in [1,k]$, such that $m_j = \umode$, or there exists $E$ and $e'$, such that $r = E[e']$ and $\mode(E) = \umode$.
\end{thm}
{\em Proof:} By induction on the number of steps of the \checkedc
evaluation ($\longrightarrow_m^*$), using progress and preservation to
maintain the invariance of the assumptions.

% \review{
%   add a paragraph that discusses what are the main changes from [21] in terms
%   of the technical development (if there are any), e.g., are there any new
%   challenges that needed to be solved while proving the blame theorem for this
%   paper's semantics?
% }
  Compared to \citet{ruef18checkedc-incr}, proofs for
  \lang were made challenging by the addition of dependently typed
  functions and dynamic arrays, and the need to handle bounds widening for NT
  array pointers. These features required changes in the runtime
  semantics (adding a stack, and dynamically changing bounds) and in
  compile-time knowledge of them (to soundly typing widened bounds).

% \ignore{
% \begin{figure}[t!]
%   \begin{prooftree}
%     \hypo{\evalue{n}{\tau} \in \defscope}
%     \infer1[T-VConst]{\Gamma;\defscope \vdash_m \evalue{n}{\tau}  : \tau}
%   \end{prooftree}
%   \qquad

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma;\defscope \vdash_m e : \tau \\
%         \Gamma;\defscope \vdash_m e_1 : \tau_1 
%       \end{matrix}
%     }
%     \hypo{
%       \begin{matrix}
% \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2 \\
%             \Gamma;\defscope \vdash_m e_2 : \tau_2
%           \end{matrix}
%         }
%     \infer2[T-If]{\Gamma;\defscope \vdash_m \eif{e}{e_1}{e_2} : \tau_3}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma;\defscope \vdash_m x : \tptr[c]{(\tntarray{l}{0}{\tau})}, l \leq 0 \\
%         \Gamma, x : \tptr[c]{(\tntarray{l}{1}{\tau})};\defscope \vdash_m e_1 : \tau_1 \\
%         \Gamma;\defscope \vdash_m e_2 : \tau_2 \\
%         \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2
%       \end{matrix}
%     }
%     \infer1[T-IfNT]{\Gamma;\defscope \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_3}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         F(f) = \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e  \\
%         \Gamma; \defscope \vdash_m \tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}] ~~~ 0 \leq i \leq j\\
%         \Gamma; \defscope \vdash_m e_i : \tau_i' ~~~ 0 \leq i \leq j\\
%         \Gamma; \defscope \vdash_m \subtype{\tau_i'}{\tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}]}  ~~~ 0 \leq i < j
%       \end{matrix}
%     }
%     \infer1[T-VCall]{\Gamma; \defscope \vdash_m f(\overline{e}) : \tau_j[x_0,\ldots,x_{j-1} \mapsto e_0,\ldots,e_{j-1}]}
%   \end{prooftree} \\ \\
% % \inferrule*[lab=T-PtrC]
% % {
% %   \tau = \tptr[c]{\omega} \\
% %   \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\\\
% %   \wt[\Gamma][\defscope, n^\tau]{H(n+k)}{\tau_k} ~~~ 0 \leq k < j
% % %  \Gamma,n^\tau \proves H(n+k) : \tau_k~~~0 \leq k < j
% % }

%   \begin{prooftree}
%     \hypo{
%           \Gamma; \defscope \vdash_m e_1 : \tau_1
%       }
%       \hypo{        \Gamma, x = e_1 : \tau_1; \defscope \vdash_m e_2 : \tau_2
% }
%     \infer2[T-Let]{\Gamma; \defscope \vdash_m \elet{x}{e_1}{e_2} : \tau[x \mapsto e_1]}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m \estrlen{y} : \tau_1\\
%         \Gamma; \defscope \vdash_m y : \tptr[c]{(\tntarray{le}{\_}{\tau_3})}\\
%         \Gamma, x = \estrlen{y} : \tau_1, \\ y : \tptr[c]{(\tntarray{le}{x}{\tau_3})} ;  \defscope  \vdash_m e_2 : \tau_2
%         \end{matrix}
%         }
%         \infer1[T-LetStr]{\Gamma; \defscope \vdash_m \elet{x}{\estrlen{y}}{e_2} : \tau[x \mapsto e_1]}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%         \Gamma; \defscope \vdash_m e : \tptr[c]{(\tntarray{le}{he}{\tau})}
%       }
%         \infer1[T-Str]{\Gamma; \defscope \vdash_m \estrlen{y} : \tint}
%       \end{prooftree} \\ \\

%       \begin{prooftree}
%         \hypo{
%           \begin{matrix}
%             \tau = \tint \vee \tau = \tptr[u]{\omega}~\vee  n=0~ \vee \\ \tau = \tptr[c]{(\tarray{0}{0}{\tau'})} \vee \\
%             \tau = \tptr[c]{(\tntarray{0}{0}{\tau'})}
%         \end{matrix}
%       }
%       \infer1[T-Base]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
%     \end{prooftree} \\ \\

%     \begin{prooftree}
%       \hypo{
%         \begin{matrix}
%         \tau = \tptr[c]{\omega} \\
%         \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\
%         \Gamma;\defscope, \evalue{n}{\tau}  \vdash_m {H(n+k)} : {\tau_k} ~~~ 0 \leq k < j
%       \end{matrix}
%       }
%       \infer1[T-PtrC]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
%     \end{prooftree} \\ \\

%     \begin{prooftree}
%       \hypo{
%         \begin{matrix}
%             \Gamma; \defscope \vdash_m e :  {\tptr[m]{\tstruct{T}}} \\
%             % \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
%             D(T) = ...; \tau_f~f; ...
%         \end{matrix}
%       }
%       \infer1[T-Amper]{\Gamma; \defscope \vdash_m \eamper{e}{f} : \tptr[m]{\tau_f}}
%     \end{prooftree} \\ \\

%     \todo[inline]{YL: how to express le - n as a metafunction?}
%     \begin{prooftree}
%       \hypo{
%         \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}\\
%         \Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tint \\
%         le' = le - n, he' = he - n
%       \end{matrix}
%     }
%       \infer1[T-BinopInd]{\Gamma; \defscope \vdash_m (e_1 \plus \evalue{n}{\tau} ) : \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})}}
%     \end{prooftree}
% \caption{Typing}
% \label{fig:typing}
% \end{figure}

% \begin{figure}[t!]
%   \todo[inline]{no sizeof check because we don't know statically whether the allocation would be null}
%   \begin{prooftree}
%     \hypo{
%       \omega = \tgarray{\alpha}{le}{he}{\tau} \Rightarrow le = 0
%     }
%     \infer1[T-Malloc]{\Gamma; \defscope \vdash_m \emalloc{\omega} : \tptr[c]{\omega}}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{
%       \Gamma; \defscope \vdash_u e : \tau
%     }
%     \infer1[T-Unchecked]{\Gamma; \defscope \vdash_m \eunchecked{e} : \tau}
%   \end{prooftree}\\\\

%   \todo[inline]{any constraints on m'?}
%   \begin{prooftree}
%     \hypo{\Gamma; \defscope \vdash_m e :  \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})} }
%     \infer1[T-DynCast]{\Gamma; \defscope \vdash_m \edyncast{\tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}}{e} : \tau}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{\Gamma; \defscope \vdash_m e : \tau'}
%     \infer1[T-Cast]{\Gamma; \defscope \vdash_m \ecast{\tau}{e} : \tau}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\tgarray{\alpha}{le}{he}{\tau}} \\
%         \Gamma; \defscope \vdash_m e_2 : \tint \\
%         m' = u \Rightarrow m = u \\
%       \end{matrix}
%     }
%     \infer1[T-Index]{\Gamma; \defscope \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
%         \Gamma; \defscope \vdash_m e_2 : \tau \\
%         m' = u \Rightarrow m = u \\
%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
%       \end{matrix}
%     }
%     \infer1[T-Assign]{\Gamma; \defscope \vdash_m \eassign{e_1}{e_2} : \tau}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
%         \Gamma; \defscope \vdash_m e_2 : \tint \\
%         \Gamma; \defscope \vdash_m e_3 : \tau \\
%         m' = u \Rightarrow m = u \\
%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
%       \end{matrix}
%     }
%     \infer1[T-IndAssign]{\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
%   \end{prooftree}

% \caption{Typing Cont.}
% \label{fig:typing2}
% \end{figure}
% \ignore{

% \begin{figure}[t!]
%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j} \\
%         \Gamma, x_0 = \kw{none} : \tau_0, \ldots, x_{j-1} = \kw{none} : \tau_{j-1}; \defscope \vdash e : \tau_{r}
%       \end{matrix}
%     }
%     \infer1[WF-Fun]{\Gamma;\defscope \vdash \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \infer0[WF-Nil]{\Gamma \vdash \circ}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma \vdash \tau_0 \\
%         \Gamma, x_0 = \kw{none} :\tau_0 \vdash x_1:\tau_{1}, \ldots, x_{j}:\tau_{j}
%       \end{matrix}
%     }
%     \infer1[WF-Cons]{\Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \infer0[WF-Int]{\Gamma \vdash \tint}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{   \Gamma \vdash le }
%      \hypo{   \Gamma \vdash he }
%       \hypo{  \Gamma \vdash \tau}

%     \infer3[WF-Array]{\Gamma \vdash \tptr[m]{(\tgarray{\alpha}{le}{he}{\tau})}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{T \in dom(D)}
%     \infer1[WF-Struct]{\Gamma \vdash \tptr[m]{\tstruct{T}}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{\Gamma \vdash \tau}
%     \infer1[WF-Ptr]{\Gamma \vdash \tptr[m]{\tau}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \infer0[WFB-Int]{\Gamma \vdash i}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{x = e? : \tint \in \Gamma}
%     \infer1[WFB-Var]{\Gamma \vdash x \plus i}
%   \end{prooftree} \\\\

%   \caption{Well-Formedness}
%   \label{fig:wf}
% \end{figure}

% \begin{figure}[t!]
%   \begin{prooftree}
%     \infer0[Sub-Nt]{\subtype{\tptr[c]{(\tntarray{le}{he}{\tau})}}{\tptr[c]{(\tarray{le}{he}{\tau})}}}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \infer0[Sub-Refl]{\subtype{\tau}{\tau}}
%   \end{prooftree} \\ \\

%   \todo[inline]{what about ntarrays?}
%   \begin{prooftree}
%     \infer0[Sub-Ptr]{\subtype{\tptr[c]{\tau}}{\tptr{\tarray{0}{1}{\tau}}}}
%   \end{prooftree} \\ \\

%   \todo[inline]{the subtyping relation is not anti-symmetric..}
%   \begin{prooftree}
%     \infer0[Sub-Arr]{\subtype{\tptr[c]{\tarray{0}{1}{\tau}}}{\tptr{\tau}}}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{\rboundle{le_0}{le_1}}
%     \hypo{\rboundle{he_1}{he_0}}
%     \infer2[Sub-Subsume]{\subtype{\tptr{(\tgarray{\alpha}{le}{he}{\tau})}{\cmode}}{ \tptr{(\tgarray{\alpha}{ le_1} {he_1} {\tau})}}}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     % check t-amper below
%     \hypo{D(T) = \tau_f~f; ...}
%     \hypo{\subtype{\tau_f}{\tau}}
%     \infer2{\subtype{\tptr[c]{\tstruct{T}}}{\tau}}
%   \end{prooftree}
%   \caption{Subtyping}
%   \label{fig:sub}
% \end{figure}
% }
% }

% \ignore{

% \begin{figure*}[t]
%   \begin{lstlisting}
% int foo(nt_array_ptr<char> p : count(0)) {
%   if (* p) {
%     dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
%   }
%   dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
%   return 0;
% }
%   \end{lstlisting}
% \caption{The example where clang Checked C fails at run-time}
% \label{fig:clangbad1}
% \end{figure*}

% \begin{figure*}[t]
%   \begin{lstlisting}
% /* nt_array_ptr<char> p : bounds(p,p) */
% size_t cnt = 0;

% while(*(p+cnt)) {
%   ++cnt;
% }
% dyn_bounds_cast<nt_array_ptr<char>>(p, count(cnt));
%   \end{lstlisting}
% \caption{A useful program that the Checked C spec doesn't allow at run-time}
% \label{fig:clangbad2}
% \end{figure*}
% }

% \ignore{
% The Clang CheckedC implementation uses statically determined
% bounds to insert run-time checks. In Fig.~\ref{fig:clangbad1}, the
% \code{dyn_bounds_cast} at line 3 will always succeed, because the
% compiler knows that within the scope of then branch, the pointer
% \code{p} must have at least one element. The same cast at line 5,
% however, will always fail, since there is no way to tell statically
% whether the program has entered the then branch before. The compiler
% will check whether the \code{count(1)} bounds specification is
% contained within the earlier \code{count(0)} specification, resulting
% in a run-time failure even when we pass in a non-empty string.

% Our formalization diverges from this run-time behavior and instead keeps
% track of the bounds on the stack. After entering the then branch, we
% increment the upper bound for \code{p}, effectively making the
% updated bounds information available even after we exit the if
% statement. The cast at line 5 will be checking the new bounds against
% the incremented bounds for non-empty strings.

% Fig.~\ref{fig:clangbad2} gives a more practical example of why keeping
% track of the bounds on the stack is useful. The program snippet
% implements the functionality of the \code{strlen} function using a
% while loop and a \code{cnt} variable. Even though the type system is
% unable to reason about the while loop, as long as the runtime system
% updates the bounds in-place, the user can apply a
% \code{dyn_bounds_cast} to soundly recover the more precise bounds
% information.
% }

% \inferrule*[lab=T-Amper]
% {
%   \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
%   D(T) = ...; \tau_f~f; ...
% }
% {
%   \Gamma \proves \eamper{e}{f} : \tptr[m]{\tau_f}
% }

### 2022-06-25 sandp23/old.tex
update paper.

\documentclass[conference]{IEEEtran}

\makeatletter
\def\endthebibliography{\def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}\endlist
}
\makeatother

\IEEEpeerreviewmaketitle

\usepackage{filecontents}
\usepackage[noadjust]{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{ebproof}
\usepackage{todonotes}
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{textcomp}
\usepackage{mathpartir}
\usepackage{color}
\usepackage{xcolor}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\usepackage{subcaption} \usepackage{epsfig}
\usepackage{listings}
\usepackage[override]{cmtt}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{url}
\usepackage{xparse}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{chains,fit,shapes,positioning,calc,arrows.meta,shapes.arrows}

\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}

\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  numbers=left,
  xleftmargin=\parindent,
  language=C,
  columns=flexible,      
  showstringspaces=false,
  basicstyle=\small\sffamily,
  otherkeywords={uint},
  literate={{<-}{{$\leftarrow\,$}}2
            {->}{{$\rightarrow\,$}}2
            {<=}{{$\leq\,$}}2},
  numberstyle=\tiny\rmfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue!80!black},
}
\lstset{language=C,style=customc}

\newenvironment{DIFnomarkup}{}{}
\newcommand{\myparagraph}[1]{\textbf{#1}.\xspace}

\newcommand{\code}[1]{\lstinline|#1|}
%\newcommand{\ignore}[1]{{}}
\newcommand{\kwchecked}{\lstinline|\_Checked|}
\newcommand{\kwunchecked}{\lstinline|\_Unchecked|}
\newcommand{\kwdynamiccheck}{\lstinline|\_Dynamic\_check|}

\newcommand{\var}[1]{#1}
\newcommand{\Arrayptr}[1]{\lstinline|_Array_ptr<|{#1}\lstinline|>|}
\newcommand{\ArrayptrT}{\Arrayptr{$\tau$}}
\newcommand{\Ptr}[1]{\lstinline|_Ptr<|{#1}\lstinline|>|}
\newcommand{\PtrName}{\lstinline|_Ptr|}
\newcommand{\PtrT}{\Ptr{$\tau$}}
\newcommand{\Ntarrayptr}[1]{\lstinline|_Nt_array_ptr<|{#1}\lstinline|>|}
\newcommand{\NtarrayptrT}{\Ntarrayptr{$\tau$}}
\newcommand{\Checkedarr}[2]{{#1}~\lstinline|_Checked[|{#2}\lstinline|]|}
\newcommand{\CheckedarrT}[1]{\Checkedarr{$\tau$}{#1}}

\newif\ifsubmit\submitfalse

\ifsubmit
\newcommand{\mwh}[1]{}
\newcommand{\ashe}[1]{}
\newcommand{\dtarditi}[1]{}
\newcommand{\dvh}[1]{}
\newcommand{\leo}[1]{}
\newcommand{\liyi}[1]{}
\newcommand{\yiyun}[1]{}
\else
\newcommand{\mwh}[1]{\textcolor{red}{Mike: #1}}
\newcommand{\dtarditi}[1]{\textcolor{purple}{David: #1}}
\newcommand{\leo}[1]{\textcolor{green!80!blue}{Leo: #1}}
\newcommand{\liyi}[1]{\textbf{\textcolor{orange}{[ Liyi: #1 ]}}}
\newcommand{\yiyun}[1]{\textcolor{cyan}{Yiyun: #1}}
\newcommand{\dvh}[1]{\textcolor{magenta}{DVH: #1}}
\fi

\newcommand{\lang}{\textsc{CoreChkC}\xspace}
\newcommand{\elang}{\textsc{CoreC}\xspace}

\newtheorem{defi}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{prop}{Proposition}

\newcommand{\CoreChkC}{\textsc{CoreChkC}\xspace}
\newcommand{\checkedc}{\text{Checked C}\xspace}

 \newcommand{\rulelab}[1]{{\small \textsc{#1}}}
\newcommand{\steps}{\ensuremath{\longrightarrow}}
\newcommand{\tbool}{\texttt{bool}}
\newcommand{\tsizeof}{\texttt{sizeof}}
\newcommand{\vbool}{\texttt{Bool}}
\newcommand{\verror}{\texttt{Error}}
\newcommand{\vval}{\mathpzc{val}}
\newcommand{\tfixed}{\texttt{fixedp}}
\newcommand{\tnat}{\texttt{nat}}
\newcommand{\tarr}[2]{\texttt{array}~{#1}~{#2}}
\newcommand{\econst}[2]{({#1}){#2}}
\newcommand{\eindex}[2]{{#1}\texttt{[}{#2}\texttt{]}}
\newcommand{\sassign}[4]{{#1} \leftarrow {#3}~{#2}~{#4}}
\newcommand{\ssassign}[3]{{#1} \xleftarrow{#2} {#3}}
\newcommand{\sif}[3]{\texttt{if}~{#1}~{#2}~{#3}}
\newcommand{\sfor}[3]{\texttt{for}~{#1}~{#2}~{#3}}
\newcommand{\scall}[3]{{#1}\leftarrow {#2}~{#3}}
\newcommand{\sseq}[2]{{#1}\,\texttt{;}\,{#2}}
\newcommand{\sinv}[1]{\texttt{inv}~{#1}}
\newcommand{\inst}[3][ ]{\texttt{#2}^{#1}~{#3}}
\newcommand{\insttwo}[4][ ]{\texttt{#2}^{#1}~{#3}~{#4}}
\newcommand{\instthree}[5][ ]{\texttt{#2}^{#1}~{#3}~{#4}~{#5}}
\newcommand{\iskip}[1]{\inst{ID}{#1}}
\newcommand{\inot}[1]{\inst{X}{#1}}
\newcommand{\ictrl}[2]{\insttwo{CU}{#1}{#2}}
\newcommand{\irz}[3][ ]{\insttwo[#1]{RZ}{#2}{#3}}
\newcommand{\isr}[3][ ]{\insttwo[#1]{SR}{#2}{#3}}
\newcommand{\icnot}[2]{\insttwo{CNOT}{#1}{#2}}
\newcommand{\ilshift}[1]{\inst{Lshift}{#1}}
\newcommand{\irshift}[1]{\inst{Rshift}{#1}}
\newcommand{\irev}[1]{\inst{Rev}{#1}}
\newcommand{\iqft}[2][ ]{\inst[#1]{QFT}{#2}}
\newcommand{\ihad}[1]{\inst{H}{#1}}

\newcommand{\iseq}[2]{{#1}\,\texttt{;}\,{#2}}
\newcommand{\inval}[2]{\insttwo{Nval}{#1}{#2}}
\newcommand{\ihval}[3]{\instthree{Hval}{#1}{#2}{#3}}
\newcommand{\iqval}[2]{\insttwo{Qval}{#1}{#2}}

\newcommand{\instr}{\iota}

\newcommand{\kw}[1]{\ensuremath{\mathtt{#1}}}
\newcommand{\subtype}[2]{\ensuremath{\vdash{#1}\prec{#2}}}
\newcommand{\estrlen}[1]{\ensuremath{\kw{strlen}({#1})}}
\newcommand{\estrlentext}{\ensuremath{\kw{strlen}}}

\newcommand{\tarray}[3]{\tarrayb{({#1},{#2})}{#3}}
\newcommand{\tarrayb}[2]{\ensuremath{[{#1}~{#2}]}}
\newcommand{\tntarray}[3]{\tntarrayb{({#1},{#2})}{#3}}
\newcommand{\tntarrayb}[2]{\tarrayb{#1}{#2}_{nt}}
\newcommand{\tallarrayb}[2]{\ensuremath{[{#1}~{#2}]_{\kappa}}}
\newcommand{\tallarraybc}[2]{\ensuremath{[{#1}~{#2}]_{\textcolor{cyan}{\kappa}}}}

\newcommand{\tallarray}[3]{\tallarrayb{({#1},{#2})}{#3}}

\newcommand{\tptr}[2]{\ensuremath{\mathtt{ptr}^{#2}~{#1}}}
\newcommand{\tntptr}[4]{\ensuremath{[({#1},{#2})~{#3}]^{#4}_{nt}*}}

\newcommand{\tarrayptr}[4]{{\tptr{\tarray{#1}{#2}{#3}}{#4}}}
\newcommand{\tntarrayptr}[4]{{\tptr{\tntarray{#1}{#2}{#3}}{#4}}}
\newcommand{\tallarrayptr}[4]{{\tptr{\tallarray{#1}{#2}{#3}}{#4}}}

\newcommand{\tgarray}[4]{\ensuremath{\mathtt{#1}~{#2}~{#3}~{#4}}}
\newcommand{\tstruct}[1]{\ensuremath{\kw{struct}~{#1}}}

\newcommand{\evalue}[2]{\ensuremath{{#1}\!:\!{#2}}}

\newcommand{\emalloc}[1]{\ensuremath{\kw{malloc}({#1})}}
\newcommand{\emalloctext}{\ensuremath{\kw{malloc}}}
\newcommand{\ecall}[2]{\ensuremath{{#1}({#2})}}
\newcommand{\ret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
\newcommand{\rettext}{\ensuremath{\kw{ret}}}
\newcommand{\ecast}[2]{\ensuremath{\kw{(}{#1}\kw{)}{#2}}}
\newcommand{\edyncast}[2]{\ensuremath{\langle{#1}\rangle{#2}}}
\newcommand{\edcast}[2]{\ensuremath{\kw{(D},{#1}\kw{)}{#2}}}
\newcommand{\elet}[3]{\ensuremath{\kw{let}~#1\, \texttt{=}\, #2~\kw{in}\;{#3}}}
\newcommand{\elettext}{\ensuremath{\kw{let}}}
\newcommand{\ebinop}[2]{\ensuremath{#1 \plus #2}}
\newcommand{\eassign}[2]{\ensuremath{\texttt{*}{#1}\,\texttt{=}\, {#2}}}
\newcommand{\eassignstack}[2]{\ensuremath{{#1}\,\texttt{=}\, {#2}}}
\newcommand{\efield}[2]{\ensuremath{{#1}\kw{\rightarrow}{#2}}}
\newcommand{\estar}[1]{\ensuremath{\texttt{*}{#1}}}
\newcommand{\eamper}[2]{\ensuremath{\kw{\&}{#1}\kw{\rightarrow}{#2}}}
\newcommand{\eunchecked}[1]{\ensuremath{\kw{unchecked}\;{#1}}}
\newcommand{\euncheckedtext}{\ensuremath{\kw{unchecked}}}
\newcommand{\eif}[3]{\ensuremath{\kw{if\;}(#1)\;{#2}\;\kw{else}\;{#3}}}
\newcommand{\eifa}[3]{\ensuremath{\kw{IF\;}(#1)\;{#2}\;\kw{ELSE}\;{#3}}}
\newcommand{\eifatext}{\ensuremath{\kw{IF}}}

\newcommand{\eiftext}{\ensuremath{\kw{if}}}
\newcommand{\ethentext}{\ensuremath{\kw{then}}}
\newcommand{\eelse}{\ensuremath{\kw{else}}}
\newcommand{\ebreak}{\ensuremath{\kw{break}}}
\newcommand{\ebounds}{\ensuremath{\kw{bounds}}}
\newcommand{\enull}{\ensuremath{\kw{null}}}
\newcommand{\hole}{\ensuremath{\Box}}
\newcommand{\defscope}{\sigma}
\newcommand{\wf}[1]{wf \, #1}
\newcommand{\tint}{\ensuremath{\mathtt{int}}}
\newcommand{\efor}{\ensuremath{\mathtt{for}}}
\newcommand{\ewhile}{\ensuremath{\mathtt{while}}}
\newcommand{\etrue}{\ensuremath{\mathtt{true}}}
\newcommand{\efalse}{\ensuremath{\mathtt{false}}}
\newcommand{\edplus}{\ensuremath{\texttt{++}}}
\newcommand{\edminus}{\ensuremath{\texttt{--}}}

\newcommand{\ememcpy}{\ensuremath{\mathtt{memcpy}}}
\newcommand{\ereturn}{\ensuremath{\mathtt{return}}}
\newcommand{\emain}{\ensuremath{\mathtt{main}}}
\newcommand{\estrcat}{\ensuremath{\mathtt{strcat}}}
\newcommand{\estrcatbad}{\ensuremath{\mathtt{strcat\_b}}}
\newcommand{\heap}{\ensuremath{\mathpzc{H}}}
\newcommand{\eret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
\newcommand{\erettext}{\ensuremath{\kw{ret}}}

\newcommand{\plus}{\mathbin{\texttt{+}}}

\newcommand{\fv}{\mathit{FV}}
\newcommand{\dom}{\mathit{dom}}

\newcommand{\fm}{\mathit{FM}}
\newcommand{\size}{\mathit{size}}

\newcommand{\cmode}{\texttt{c}}
\newcommand{\umode}{\texttt{u}}
\newcommand{\bvar}{\ensuremath{\beta}}
\newcommand{\mode}{\textit{mode}}

\newcommand{\cextend}[4]{\ensuremath{#4 = ~ \vdash_{extend} #1, #2, #3}}
\newcommand{\fresh}{\ensuremath{\kw{fresh}}}
\newcommand{\echecknull}[3]{\ensuremath{#3 = ~\vdash_{null}#1, #2}}
\newcommand{\echeckboundsdyn}[5]{\ensuremath{#5 = ~ \vdash_{boundsD}#1, #2, #3, #4}}
\newcommand{\echeckbounds}[5]{\ensuremath{#5 = ~ \vdash_{boundsR}#1, #2, #3, #4}}
\newcommand{\echeckboundsw}[5]{\ensuremath{#5 = ~ \vdash_{boundsW}#1, #2, #3, #4}}
\newcommand{\esizeof}[1]{\ensuremath{\kw{sizeof}(#1)}}
\newcommand{\ewidenstrlen}[5]{\ensuremath{#5 = ~ \vdash_{widenstr}#1, #2, #3, #4}}
\newcommand{\ewidenderef}[4]{\ensuremath{#4 = ~ \vdash_{widenderef}#1, #2, #3}}
\newcommand{\eleq}[2]{\ensuremath{#1 \leq #2}}

\newcommand{\rboundle}[2]{\ensuremath{\vdash_{bounds}#1 \leq #2}}

\def\codesize{\normalsize}
\definecolor{programs}{gray}{0.1}

\lstset{
  basicstyle=\ttfamily,  
}

\makeatletter
\lst@CCPutMacro
    \lst@ProcessOther {"22}{\lst@ifupquote \textquotedbl
                                     \else \char34\relax \fi}
    \@empty\z@\@empty
\makeatother

\lstloadlanguages{C}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{A Formal Model of Checked C}

% \ignore{
% \author{\IEEEauthorblockN{1\textsuperscript{st} Leonidas Lampropoulos}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Maryland}\\
% llampro@cs.umd.edu}
% \and
% \IEEEauthorblockN{1\textsuperscript{st} Michael Hicks}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Maryland}\\
% llampro@cs.umd.edu}
% }
% \author{Leonidas Lampropoulos \and
% Michael Hicks}
% }
\maketitle

\begin{abstract}

  We present a formal model of Checked C, a dialect of C
  that aims to enforce spatial memory safety. Our model pays particular
  attention to the semantics of dynamically sized, potentially null-terminated
  arrays.
We formalize this model in Coq, and prove that any spatial memory
  safety errors can be \emph{blamed} on portions of the program
  labeled \emph{unchecked}; this is a Checked C feature that supports
  incremental porting and backward compatibility.
Our model develops an operational semantics that uses fat pointers to
  guarantee spatial safety. However, we formalize a compilation scheme
  that can yield thin pointers, with bounds information managed using
  inserted code. We show that the generated code faithfully simulates
  the original.
Finally, we build an executable version of our model in PLT Redex, 
  and use a custom random generator
  for well-typed and almost-well-typed terms to find
  inconsistencies between our model and the Clang Checked C
  implementation. We find these steps to be a useful way to co-develop a language
  (Checked C is still in development) and a core model of it.
\end{abstract}

\section{Introduction}\label{sec:intros}

The C programming language remains extremely popular despite the
emergence of new, modern languages. Unfortunately, C programs lack
spatial memory safety, which makes them susceptible to a host
of devastating vulnerabilities, including buffer overflows and
out-of-bounds reads/writes. Despite their long history, buffer
overflows and other spatial safety violations are among the most
prevalent and dangerous vulnerabilities on the Internet today \cite{Zeng:2013:SRF:2534766.2534798}.

Several industrial and research efforts---including CCured~\cite{Necula2005},
Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
explored means to compile C programs 
to automatically enforce spatial safety. These
approaches all impose performance overheads that are deemed too high for
use in deployment. Recently, Microsoft introduced \checkedc, an
open-source extension to C with new types and
annotations whose use can ensure a program’s spatial
safety~\cite{Elliott2018}. Importantly, \checkedc supports development that is
incremental and compositional. Code regions (e.g.,
functions or whole files) designated as \emph{checked} are sure to enforce
spatial safety, a property which is preserved via composition with
other checked regions. But not all regions must be checked: Checked
C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
may coexist in the same code, which permits a deliberate (and
semi-automated) refactoring process. Parts of the FreeBSD kernel have
been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
overhead seems low enough for practical deployment.

While \checkedc promises to enforce spatial safety, we might wonder
whether its design and implementation deliver on this promise, or even
what ``spatial safety'' means when a program contains both checked and
unchecked code. In prior work, \citet{ruef18checkedc-incr} developed a core
formalization of \checkedc and with it proved that \emph{checked code cannot 
  be blamed}: any spatial safety violation can only be attributed to
code that is not in a checked region. While their work is a good start,
it fails to model important aspects of \checkedc's functionality,
particularly those involving pointers to arrays. In this paper, we
cover this gap, making three main contributions.

\myparagraph{Dynamically bounded and null-terminated arrays}
Our first contribution is a core formalism called \lang, which extends
\citet{ruef18checkedc-incr} with several new features, most notably 
\emph{dynamically bounded arrays} (Section~\ref{sec:formal}). 
Dynamically bounded arrays are those whose size is
known only at run time, as designated by in-scope variables using
dependent types. A pointer's accessible memory is bounded both above
and below, to admit arbitrary pointer arithmetic.

\lang also models
\emph{null-terminated} arrays, a kind of dynamically bounded array whose upper bound 
defines the array's \emph{minimum} length---additional space is available 
up to a null terminator. For example, the \checkedc type
\lstinline{nt_array_ptr<char> p:count(n)} says that
\lstinline{p} has length \emph{at least} \lstinline{n} (excluding the
null terminator), but further capacity is present if \lstinline|p[n]| is not null. \checkedc (and \lang) supports flow-sensitive
\emph{bounds widening}: statements of the form \lstinline|if (*p) | $s$,
where \lstinline{p}'s type is \lstinline{nt_array_ptr<T> count(0)},
typecheck statement $s$ under the assumption that \lstinline{p} has
type \lstinline{nt_array_ptr<T> count(1)}, 
 i.e., one more than it was,
since the character at the current-known length is
non-null. Similarly, the call \lstinline|n = strlen(p)| will widen
\lstinline|p|'s bounds to \lstinline|n|. Subtyping permits treating
null-terminated arrays as normal arrays of the same size (which
does not include, and thereby protects, the null terminator).

We prove, in Coq, a blame theorem for \lang.  As far as we are aware,
ours is the first formalized type system and proof of soundness for
pointers to null-terminated arrays with expandable bounds.

\myparagraph{Sound compilation of checked pointers} Our second
contribution is a formalization of bounds-check insertion for array
accesses (Section~\ref{sec:compilation}). Our operational semantics
annotates each pointer with metadata that describes its bounds, and
the assignment and dereference rules have premises to confirm the
access is in bounds. An obvious compilation scheme (taken by
Cyclone~\cite{Jim2002,GrossmanMJHWC02}, CCured~\cite{Necula2005}, and
earlier works) would be to translate annotated pointers to multi-word
objects: one word for the pointer, and 1-2 words to describe its lower
and upper bounds. Inserted checks reference these bounds. While
convenient, such ``fat'' pointers are expensive, and break backward
binary compatibility with legacy pointers. We formalize \checkedc's
compilation approach, which uses a single machine word for the
pointer, and adds checks involving the declared bounds (e.g., in a
dependent type) or additional stack-allocated \emph{ghost variables}
to accommodate bounds widening. We show that the compiled program
\emph{simulates} the original by mechanizing \lang and the compilation
judgment in PLT Redex~\cite{pltredex}, and use its random testing
feature to give confidence that simulation holds.

As far as we are aware, \lang is the first formalism to cleanly
separate bounds-checking compilation from the core semantics; prior
work merged the two, conflating \emph{meaning} with
\emph{mechanism}~\cite{Feng2006,Condit2007}. In carrying out the 
formalization, we discovered that our compilation approach is 
more expressive than that proposed in the Checked C
specification~\cite{checkedc} (Section~\ref{sec:disc}); we 
would not have discovered this improvement had we not separated checks from 
semantics.

\myparagraph{Model-based randomized testing} Our third
and final contribution is a strategy and implementation of model-based
randomized testing (Section~\ref{sec:evaluation}).  To check the
correctness of our formal model, we compare the behavior between the
existing Clang \checkedc implementation and our own model. This is
done by a conversion tool that converts expressions from \lang into
actual \checkedc code that can be compiled by the Clang \checkedc
compiler. We build a random generator of programs largely based on the
typing rules of \lang and make sure that, both statically and
dynamically, \lang and Clang Checked C are consistent after
conversion.  This helped rapidly prototype the model and uncovered
several issues in the Checked C compiler.

We begin with a review of \checkedc (Section~\ref{sec:overview}),
present our main contributions (Sections~\ref{sec:formal}--\ref{sec:evaluation}),
and conclude with a discussion of
related and future work (Sections~\ref{sec:related}, \ref{sec:conclude}).

\section{\checkedc Overview}\label{sec:overview}

This section describes Checked C, which extends C with new pointer
types and annotations that ensure spatial safety. More
details can be found in a prior overview~\cite{Elliott2018} or the
full specification~\cite{checkedc}.
Checked C is implemented as a fork of Clang/LLVM and is freely
available.\footnote{\url{https://github.com/Microsoft/checkedc-clang}}

\begin{figure}[t]
{\small
  \begin{lstlisting}[xleftmargin=4 mm]
nt_array_ptr<const char>
parse_utf16_hex(nt_array_ptr<const char> s,
                  ptr<uint> result) {
 int x1, x2, x3, x4;
 if (s[0] != 0) { x1 = hex_char_to_int(s[0]);
 if (s[1] != 0) { x2 = hex_char_to_int(s[1]);
 if (s[2] != 0) { x3 = hex_char_to_int(s[2]);
 if (s[3] != 0) { x4 = hex_char_to_int(s[3]);
 if (x1 != -1 && x2 != -1 && x3 != -1 && x4 != -1){
   *result = (uint)((x1<<12)|(x2<<8)|(x3<<4)|x4);
   return s+4;
  ...// several } braces
 } 
 return 0;
}
void parse(nt_array_ptr<const char> s,
            array_ptr<uint> p : count(n), 
            int n) {
 array_ptr<uint> q : bounds(p,p+n) = p;
 while (s && q < p+n) {
   array_ptr<uint> r : count(1) =
     dyn_bounds_cast<array_ptr<uint>>(q,count(1));
   s = parse_utf16_hex(s,r);
   q++;
 }
}
  \end{lstlisting}

}

\caption{Parsing a string of UTF16 hex characters in \checkedc}
\label{fig:checkedc-example}
\end{figure}

\subsection{Checked Pointer Types}
\checkedc introduces three varieties of \emph{checked pointer}:
\begin{itemize}
\item \code{ptr<}$T$\code{>} types a pointer that is either null or
  points to a single object of type $T$.
\item \code{array_ptr<}$T$\code{>} types a pointer that is either null
  or points to an array of $T$ objects. The array width is defined
  by a \emph{bounds} expression, discussed below.
\item \code{nt_array_ptr<}$T$\code{>} is like
  \code{array_ptr<}$T$\code{>} except that the bounds expression
  defines the \emph{minimum} array width---additional objects may
  be available past the upper bound, up to a null terminator.
\end{itemize}
A bounds expression used with the latter two pointer types has three
forms:
\begin{itemize}
\item \code{count(}$e$\code{)} where $e$ defines the array's
  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
  expression $e$ must be side-effect free and may only refer to
  variables whose addresses are not taken, or adjacent \code{struct}
  fields.
\item \code{byte_count(}$e$\code{)} is like \code{count}, but
  expresses arithmetic using bytes, no objects; i.e.,
  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
  pointers that bound the accessible region $[e_l,e_h)$ (the
  expressions are similarly restricted). Bounds
  \code{count(}$e$\code{)} is shorthand for
  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
  expression is useful for supporting pointer arithmetic.
\end{itemize}
The \checkedc compiler will instrument loads and stores of checked
pointers to confirm the pointer is non-null, and the access is within
the specified bounds. For pointers $p$ of type
\code{nt_array_ptr<}$T$\code{>}, such a check could spuriously fail if
the index is past $p$'s specified upper bound, but before the null
terminator. To address this problem, Checked C supports \emph{bounds
  widening}. If $p$ has bounds expression
\code{bounds(}$e_l$,$e_h$\code{)} a program may read from (but not
write to) $e_h$; when the compiler notices that a non-null character
is read at the upper bound, it will extend that bound to $e_h+1$.

\subsection{Example}

Fig.~\ref{fig:checkedc-example} gives an example \checkedc
program.\footnote{Ported from the Parson JSON
parser, \url{https://github.com/kgabis/parson}} 
The function \code{parse_utf16_hex} on lines 1-17 takes as its argument
null-terminated pointer \code{s} from which it attempts to read four
characters. These are interpreted as hex digits and converted to an
\code{uint} returned via parameter \code{result}. At first,
\code{s} has no specific bounds annotation, which we can interpret as
\code{count(0)}; this means that \code{s[0]} may be read on line
5. The true branch of the conditional (which extends all the way to
the brace on line 15) is thus typechecked with \code{s} given a
\emph{widened} bound of \code{count(1)}. Likewise, the conditionals on
lines 6-8 each widen it one further; the widened pointer
(\code{s+4}) is returned on success.

The \code{parse} function on lines 18-26 repeatedly invokes
\code{parse_utf16_hex} with its parameter \code{s}, and fills out
array \code{p} whose declared length is the parameter \code{n}. Writes
happen via pointer \code{q}, which is updated using pointer
arithmetic. We specify its bounds as
\code{bounds(p,p+n)} to support this: even as \code{q} changes, 
variables \code{p} and \code{n} do not. Converting from an
\code{array_ptr<uint>} to a \code{ptr<uint>}, done for the call on
line 25, requires proving the array has size at least 1. While this is true
because of the loop condition \code{q < p+n}, which is \code{q}'s
upper bound, the compiler is not smart enough to figure this
out. To convince it, we manually insert a bounds check using
\code{dyn_bounds_cast}.

While bounds checks are \emph{conceptually} inserted on every array
load and store, many of these are eliminated by LLVM\@. For example,
all of the pointer accesses to \code{s} on lines 5-8 are proved safe
at compile-time, so no bounds checks are inserted for
them. \citet{Elliott2018}
reported average run-time overheads of 8.6\% on a pointer-intensive
benchmark suite (49.3\% in one case); \citet{duanrefactoring} measured
no overhead at all on a port of FreeBSD's UDP and IP stacks to \checkedc.

\subsection{Other features}

\checkedc has other features not modeled in this paper. Two in regular
use are \emph{interop types}, which ascribe checked pointer types to
unported legacy code, notably in libraries; and \emph{generic types}
on both functions and \code{struct}s, for type-safe polymorphism. More
details about these can be found in the language specification.

\subsection{Spatial Safety and Backward Compatibility}
\checkedc is backward compatible with legacy C in the sense that all
legacy code will typecheck and compile. However, only code that
appears in \emph{checked regions}, which we call \emph{checked code},
is spatially safe. Checked regions can be designated at the level of
files, functions, or individual code blocks, the first with a
\code{#pragma} and the latter two using the \code{checked}
keyword.\footnote{You can also designate \emph{unchecked} regions
  within checked ones.}  Within checked regions, both
legacy pointers and certain unsafe idioms (e.g., variadic function
calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
satisfies these conditions, and will typecheck in a checked region.

How should we think about code that contains both checked and legacy
components? \citet{ruef18checkedc-incr} proved, for a simple
formalization of \checkedc, that \emph{checked code cannot be blamed}:
Any spatial safety violation owes to the execution of unchecked
code. In this paper we extend that result to a richer formalization of
\checkedc.

% \ignore{
% \liyi{Good information about pointer types here:
% https://github.com/microsoft/checkedc/wiki/New-pointer-and-array-types}

% This code creates an array pointer of length $10$, and reads the pointer $(x+y)$. 
% If the integer value $y$ is greater than $10$, executing the program in C results in undefined behavior. 
% In \checkedc, however, we links every pointer in a program with their bound information statically, and inserts checks to dynamically verify if the usage of the pointer violates the spatial safety. For example, In compiling the above fragment, we inserts the bound's check for $x$ in the read operation $(x+y)$, and if $y$ results in a value $11$, the execution of the read operation $(x+y)$ results in an dynamic error.

% There are many existing language development trying to guarantee the C spatial safety property through different approaches. For example, Cyclone \cite{Jim2002} and Cerberus \cite{cerberus} are the fat pointer approaches for guaranteeing the spatial safety and eliminating undefined behaviors in C. Rust \cite{Rust2016}, on the other hand, tries to restrict the possible way of writing programs to guarantee such property statically.
% The problem of these two approaches are effectiveness of executing programs and user practicality of writing programs. The fat-pointer approach completely relies on dynamic checks to guarantee the safety properties but it might pay a huge price on execution efficiency. In our experiment (Sec.\ref{sec:evaluation}), we found that the execution of fat pointer approaches cost the speed a 50\% overhead. 
% On the other hand, Rust uses a static approach by restricting the usage of pointers to guarantee the properties, which results in a user UN-friendly system.  
% The development of \checkedc is an investigation on keeping the balance between execution efficiency and user-friendly systems. We develop a system to utilize static type systems to allow the \checkedc compiler to insert dynamic checks to guarantee the safety property. The efficiency of \checkedc is in the same level of Rust while we maintain a user-friendly system and inherited most of the operation features from C.

% Another feature of \checkedc is the division of the checked and unchecked blocks. When users try to rewrite their C programs into \checkedc programs, they can develop the code incrementally by rewriting sub-parts of their programs to \checkedc code fragments and placing them as the checked blocks, while keeping the rest C programs as unchecked C programs. \checkedc guarantees that if there is any safety violation, this comes from the unchecked blocks (the blame theory).

% Here is the contributions of the paper. First, we develop the code formalization of \checkedc and show that the type system in \checkedc is type-sound and satisfies the blame theory with respect to the \checkedc semantics. Specifically, we formalize null-terminated array pointers. To our best knowledge, this is the first work defining such feature in C-related languages. 
% Second, in formalizing the \checkedc language, we investigate the balance between program execution efficiency and user practicality. We re-investigate the \checkedc type system with subtyping relations to make it more efficient and user-friendly.
% Second, we also develop random testing tools in Redex \cite{pltredex} to test the \checkedc compiler and make sure it is properly developed. We are able to find and reproduce many bugs/faults in the \checkedc compiler \liyi{numbers?}. The random testing tool is able to generate tens of thousands of programs to properly validate the \checkedc compiler. Third, we also spent a great amount of work in developing experiments to compare \checkedc and other existing C-like languages that guarantee the spatial safety and eliminating undefined behaviors. We found that \checkedc keeps the best balance between execution efficiency and user practicality.
% }

% \ignore{
% \section{Null-Terminated Array Pointers}

% Key ideas, formalized:
% \begin{itemize}
% \item Can read one past the end: x[N] can be read, but only written if
%   the value being written is 0.
% \item NT arrays are a refinement of normal arrays: Can convert
%   x:(ntarray N t) to (array N t).
% \item NT arrays can expand their length flow sensitively: Can convert
%   x:(ntarray N t) to (ntarray N+1 t) assuming x[N] != 0.
% \item In general unsafe to convert x:(array N t) to (ntarray N-1 t)
%   even if x[N-1] == 0 because aliases of x could eliminate the NULL
%   terminator. Would need some sort of linearity/alias tracking to
%   support this idea.
% \item Question? What is the syntax of doing a malloc on a NT-array?
%   Should we just do malloc(NT-array l h type) as the malloc of a
%   normal array or should we also include something like int [] nt =
%   “abcd”, where the bound size is not specified?
% \end{itemize}

% See also the Checked C specification
% \ignore{
% https://github.com/microsoft/checkedc/tree/master/spec/bounds\_safety}
% }

\section{Formalization}\label{sec:formal}

% \ignore{
% \begin{figure}
%   \begin{center}
%     \includegraphics[height=6in]{syntax.pdf}
%   \end{center}
%   \caption{\lang: Syntax}
% \end{figure}

% \begin{figure}
%   \begin{center}
%     \includegraphics[height=6in]{types.pdf}
%   \end{center}
%   \caption{\lang: Typing}
% \end{figure}
% }

\begin{figure}
  \small \centering
  $\begin{array}{l}
\begin{array}{lll}
\text{Function names:}~f&
       \text{Variables:}~ x
& \text{Integers:}~n::=\mathbb{Z} 
\end{array}
\\[0.5em]

\begin{array}{llcllcl}

\text{Mode:} & m & ::= & \cmode \mid \umode \\[0.5em]

\text{Bound:} & b & ::= & n \mid x \plus n \\
              & \bvar & ::= & (b,b) \\[0.5em]

     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{m}
\\[0.5em]

\text{Type Flag:}&\kappa &::=& nt \mid \cdot
\\[0.5em]

\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau}
\\[0.5em]

\text{Expression:}& e & ::= & 
\evalue{n}{\tau} \mid x \mid \emalloc{\omega} \mid\elet{x}{e}{e} \\[0.2em]
&&\mid&
\ecast{\tau}{e} \mid \edyncast{\tau}{e}\mid \ecall{f}{\overline{e}} \mid \estrlen{x} \\[0.2em]
&&\mid&
\ebinop{e}{e} \mid\estar{e}\mid\eassign{e}{e}\mid\eunchecked{e}
\\[0.2em]
&&\mid&\eif{e}{e}{e}
\end{array}
    \end{array}
  $
  \caption{\lang Syntax}
  \label{fig:checkc-syn}
\end{figure}

This section describes our formal model of \checkedc, called
\lang, making precise its syntax, semantics, and type system. It also
develops \lang's metatheory, including type soundness and the blame
theorem.

\subsection{Syntax}\label{sec:syntax}

The syntax of \lang is given by the expression-based
language presented in Fig.~\ref{fig:checkc-syn}.

There are two notions of type in \lang.  Types $\tau$ classify
word-sized values including the integers and pointers, while types
$\omega$ classify multi-word values such as arrays, null-terminated
arrays, and single-word-size values.
Pointer types ($\tptr{\omega}{m}$) include a mode annotation ($m$)
which is either checked (\cmode) or unchecked (\umode) and a type
($\omega$) denoting the type of value to which is pointed. Array types include both the type of
elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
limited to integer literals $n$ and expressions $x + n$.
Whether an array pointer is null terminated or not is determined by annotation
$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
otherwise (we elide the $\cdot$ when writing the type). Here is the
corresponding \checkedc syntax for these types:
\[
\begin{array}{rcl}
$\code{array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\cmode}
\\[0.2em]
$\code{nt_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
\end{array}
\]
As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.

\lang expressions include literals ($n\!:\!\tau$), variables ($x$), memory
allocation ($\emalloc{\omega}$), let binding ($\elet{x}{e_1}{e_2}$),
static and dynamic casts ($\ecast{\tau}{e}$ and $\edyncast{\tau}{e}$,
resp.), function calls ($\ecall{f}{\overline{e}}$), addition
($\ebinop{e_1}{e_2}$), pointer dereference and assignment ($\estar{e}$
and $\eassign{e_1}{e_2}$, resp.), unchecked blocks ($\eunchecked{e}$),
the \texttt{strlen} operation ($\estrlen{x}$), and conditionals $\eif{e}{e_1}{e_2}$.

Integer literals $n$ are annotated with a type $\tau$ which can be either
$\tint$, or $\tptr{\omega}{m}$ in the case $n$ is being used as
a heap address (this is useful for the semantics). The
$\texttt{strlen}$ expression operates on variables $x$
rather than arbitrary expressions to simplify managing
bounds information in the type system; the more general case can be
encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
casts; e.g., the following

\code{dyn_bounds_cast<array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}

\noindent
becomes $\edyncast{\tptr{\tntarrayb{n}{\tau}}{\cmode}}{e}$.

\lang aims to be simple enough to work with, but powerful enough to
encode realistic \checkedc idioms. For example: mutable local
variables can be encoded as immutable locals that point to the heap;
the use of \code{&} can be simulated with \code{malloc};
and loops can be encoded as recursive function calls. \code{struct}s are
not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
actually in our model, and developed in
Appendix~\ref{appx:struct}. C-style \code{union}s have no safe typing
in \checkedc, so we elide them. By default, functions are assumed to
be within checked regions; placing the body in an \code{unchecked}
expression relaxes this, and within that, checked regions can be
nested within via function calls. Bounds are restricted slightly:
rather than allowing arbitrary subexpressions, bounds must be either
integer literals or variables plus an integer offset, which accounts
for most uses of \code{bounds} in \checkedc programs. \lang bounds are
defined as relative offsets, not absolute ones, as in the second part
of Fig.~\ref{fig:checkedc-example}. We see no technical problem to
modeling absolute bounds, but it would be a pervasive change so we
have not done so.

\begin{figure}
{\small
$    \begin{array}{l}
    \begin{array}{lll}
\mu & ::= & \evalue{n}{\tau} \mid \bot\\
e & ::= & \ldots \mid \ret{x}{\mu}{e}\\
r & ::= & e \mid \enull \mid \ebounds\\
E &::=& \Box \mid \elet{x}{E}{e} \mid \ecall{f}{\overline{E}}\mid\ecast{\tau}{E} \mid \edyncast{\tau}{E} \\[0.2em]
&&\mid \ret{x}{\evalue{n}{\tau}}{E}\mid\ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid\estar{E}\mid\eassign{E}{e} \\[0.2em]
&&\mid\eassign{\evalue{n}{\tau}}{E}\mid\eunchecked{E}\mid\eif{E}{e}{e}\\[0.2em]
\overline{E} &::=& E \mid  \evalue{n}{\tau},\overline{E} \mid \overline{E}, e
\end{array}
\\ \\
    \end{array} 
$
  \begin{mathpar}
    \inferrule{ m=\mode(E) \\
      e=E[e'] \\
      (\varphi,\heap,e') \longrightarrow (\varphi',\heap',e'')}
    {(\varphi,\heap,e)\longrightarrow_{m} (\varphi',\heap',E[e''])}

  \end{mathpar}
}
  \caption{\lang semantics: Evaluation}
  \label{fig:c-context}
\end{figure}

\subsection{Semantics}\label{sec:semantics}

The operational semantics for \lang is defined as a small-step
transition relation with the judgment $ (\varphi,\heap,e)
\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
$\heap$ is a \emph{heap} mapping addresses (integer literals) to
values $\evalue{n}{\tau}$; for both we ensure
$\fv(\tau)=\emptyset$. The relation steps to a \emph{result} $r$,
which is either an expression or a $\enull$ or $\ebounds$ failure,
representing a null-pointer dereference or out-of-bounds access,
respectively. Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterize undefined behavior.
The mode $m$ indicates whether the
stepped redex within $e$ was in a checked ($\cmode$) or
unchecked ($\umode$) region.

The rules for the main operational semantics judgment---\emph{evaluation}---are given at the bottom of
Fig.~\ref{fig:c-context}. The rule takes an expression $e$, decomposes
it into an \emph{evaluation context} $E$ and a subexpression $e'$
(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
$e$), and then evaluates $e'$ according to the \emph{computation} relation $
(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$, discussed
shortly. The computation relation can transition to any $r$; the rule
in the figure just considers $r = e$. For the cases $r = \enull$ and
$r = \ebounds$, two other rules (not shown) will cause the whole
evaluation to terminate with $r$.  The semantics has a special case
for a conditional $\eif{e_0}{e_1}{e_2}$.  If $e_0=\estar{x}$, the
conditional itself is considered as the redex with special handling
for bound widening.  Otherwise, the standard conditional behavior is
used.
The $\mode$ function
determines the mode of the evaluating $e'$ based on the context $E$:
if the $\Box$ in $E$ occurs in context $\eunchecked{E'}$, the mode is
$\umode$, otherwise, it is $\cmode$. Evaluation contexts $E$ define a
standard left-to-right evaluation order. (We explain the
$\ret{x}{\mu}{e}$ syntax shortly.)

\begin{DIFnomarkup}
\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[S-Cast]
              {}
              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}

    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}

        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}

    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}

        \inferrule[S-DefNTArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h]}
    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}

    \inferrule[S-Fun]{ \Xi(f) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
        {(\varphi,\heap,\ecall{f}{{\evalue{\overline{n}}{\overline{\tau_a}}}}) \longrightarrow
   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}

  \inferrule[S-DynCast]{
\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{m})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{m} \\ n_l' \le n_l\ \\ n_h \le n_h'}
    { (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{m}}}) \longrightarrow (\varphi',\heap',\evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau_b}{m}})}

 \inferrule[S-DynCastBound]{\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{\cmode})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{\cmode} \\ n_l' > n_l \vee n_h > n_h'}{ (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{\cmode}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{\cmode}}}) 
 \longrightarrow (\varphi',\heap',\ebounds)}

      \inferrule[S-StrWiden]{
         \varphi(x) = \evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{\cmode}}\\ 0 \in [n_l,n_h] \\ n_a > n_h
\\ \heap(n+n_a) = 0 
\\\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
    {(\varphi,\heap,\estrlen{x}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tarrayptr{n_l}{n_a}{\tau}{\cmode}}],\heap,\evalue{n_a}{\tint})}

     \inferrule[S-IfNTT]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(n)\neq 0}
    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}

    \inferrule[S-AssignArr]{\heap(n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}

        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} }
    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
\end{mathpar}
}
\caption{\lang semantics: Computation (selected rules)}
\label{fig:semantics}
\end{figure*}
\end{DIFnomarkup}

\begin{figure}[t]
{\small
{\captionsetup[lstlisting]{margin = 8 mm}
  \begin{lstlisting}[xleftmargin=8 mm]
nt_array_ptr<int> strncat : count(0)
  (nt_array_ptr<int> a : count(n),
   nt_array_ptr<int> b : count(0), int n) {
  int x = strlen(a);
  int y = strlen(b);
  if (x <= n)
    nt_array_ptr<int> c : count(n) =
      dynamic_bounds_cast
        <nt_array_ptr<int>>(a,count(n));
    else return null;

  if (x+y <= n) 
      for (int i = 0; i < y; ++i)
         * (c+x+i) = * (b+i);
    else return null;
  return a;
}

nt_array_ptr<int> strncat_c : count(0)
  (nt_array_ptr<int> a : count(n),
   nt_array_ptr<int> b : count(0), int n) {
   int x = 0;
   int y = strlen(b);
   while (*x != '/0')
    a++; x++;

   for (int i = 0; i < y; ++i)
     if (i + x < n)
         * (a+i) = * (b+i);
       else return null;
   return a;
}
  \end{lstlisting}
}
}
\caption{Implementations for strncat}
\label{fig:strncat-ex}
\end{figure}

Fig.~\ref{fig:semantics} shows rules for the computation
relation. We explain the rules in turn using the
example of Fig.~\ref{fig:strncat-ex}.

\myparagraph{Pointer accesses}
The rules for dereference and assignment operations---\textsc{S-Def},
\textsc{S-DefNull}, \textsc{S-DefNTArray}, and
\textsc{S-AssignArr}---illustrate how the semantics checks bounds.
Rule \textsc{S-DefNull} transitions attempted null-pointer
dereferences to $\enull$, whereas \textsc{S-Def} dereferences a
non-null (single) pointer.

\textsc{S-AssignArr} assigns to an array as long as 0 (the point of
dereference) is within the bounds designated by the pointer's annotation
and strictly less than the upper bound. For the assignment rule,
arrays are treated uniformly whether they are null-terminated or not
($\kappa$ can be $\cdot$ or $nt$)---the semantics does not search past
the current position for a null terminator. The program
can widen the bounds as needed, if they currently precede the null
terminator: \textsc{S-DefNTArray}, which dereferences an NT array
pointer, allows an upper bound of $0$, since the program may read, but
not write, the null terminator. A separate rule (not shown) handles
normal arrays.

\myparagraph{Casts}
Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
by \textsc{S-Cast}. In a type-correct program, such casts are
confirmed safe by the type system. To evaluate a cast, the rule
updates the type annotation on $n$. Before doing so, it must
``evaluate'' any variables that occur in $\tau$ according to their
bindings in $\varphi$. For example, if $\tau$ was
$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.

Dynamic casts are accounted for by \textsc{S-DynCast} and
\textsc{S-DynCastBound}. In a type-correct program, such casts are
assumed correct by the type system, and later confirmed by the
semantics. As such, a dynamic cast will cause a $\ebounds$ failure if
the cast-to type is incompatible with the type of the target pointer,
as per the $n_l' > n_l \vee n_h > n_h'$ condition in
\textsc{S-DynCastBound}. An example use of dynamic casts is given on
line~7 in Fig.~\ref{fig:strncat-ex}. The values of \code{x} and
\code{n} might not be known statically, so the type system cannot
confirm that \code{x <= n}; the dynamic cast assumes this inequality
holds, but then checks it at run-time.

\myparagraph{Binding and Function Calls}
The semantics handles variable scopes using the special $\erettext$
form. \textsc{S-Let} evaluates to a configuration whose stack
is $\varphi$ extended with a binding for $x$, and whose expression is
$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
$\varphi(x)$; if it had no previous binding, $\varphi(x) =
\bot$. Evaluation proceeds on $e$ until it becomes a literal
$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
binding (or $\bot$) in the new stack, and evaluates to
$n\!:\!\tau$.

Function calls are handled by \textsc{S-Fun}. Recall that array
bounds in types may refer to in-scope variables; e.g., parameter
\code{a}'s bound \code{count(n)} refers to parameter \code{n} on lines
2-3 in Fig.~\ref{fig:strncat-ex}. A call to function $f$ causes $f$'s
definition to be retrieved from $\Xi$, which maps function names to
forms $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e$, where
$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
is the parameter list of variables and their types, and $e$ is the
function body. The call is expanded into a \texttt{let} which binds
parameter variables $\overline{x}$ to the actual arguments
$\overline{n}$, but annotated with the parameter types
$\overline{\tau}$ (this will be safe for type-correct programs). The
function body $e$ is wrapped in a static cast
$(\tau[\overline{n} / \overline{x}])$, which is the function's return
type but with any parameter variables $\overline{x}$ appearing in that
type substituted with the call's actual arguments $\overline{n}$. To
see why this is needed, suppose that \code{strncat} in
Fig.~\ref{fig:strncat-ex} is defined to return a
\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
perform a \code{strncat} function call as
\code{x=strncat(a,b,10)}. After the evaluation of \code{strncat}, the
function returns a value with type \code{nt_array_ptr<int>:count(10)}
because we substitute bound variable \code{n} in the 
defined return type with \code{10} from the function call's
argument list.

\myparagraph{Bounds Widening}
Bounds widening occurs when branching on a dereference of a NT array
pointer, or when performing $\estrlentext$. The latter is most useful
when assigned to a local variable so that subsequent code can use the
result, e.g., $e$ in $\elet{x}{\estrlen{y}}{e}$.  Lines 4 and 5 in
Fig.~\ref{fig:strncat-ex} are examples. 
The widened upper bound
precipitated by $\estrlen{y}$ is extended beyond the lifetime of $x$, as long as $y$ is live.
For example, \code{x}'s scope in line 4 is
the whole function body in \code{strncat} because the lifetime of the pointer $y$ is in the function body.
This is different from the \checkedc specification, 
which only allows bound widening to happen within the scope of $x$, and restoring old bound values once $x$ dies.
We allow widening to persist outside the scope at run-time as long as
we are within the stack frame, and we show
this does not necessarily require the use of fat pointers in
Sec.~\ref{sec:compilation}.

Rule \textsc{S-StrWiden} implements $\estrlentext$ widening. The
predicate
$\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i)
= \evalue{n_i}{\tau_i} \wedge n_i \neq 0))$ aims to find a position
$n+n_a$ in the NT array that stores a null character, where no
character as indexes between $n$ and $n+n_a$ contains one. (This rule
handles the case when $n_a > n_h$, the $n_a \le n_h$ case is handled
by a normal $\estrlentext$ rule; see Appx.~\ref{fig:rem-semantics}.)

Rule \textsc{S-IfNTT} performs bounds widening on $x$ when the
dereference $\estar{x}$ is not at the null terminator, but the
pointer's upper bound is 0 (i.e., it's at the end of its known
range). $x$'s upper bound is incremented to 1, and this count persists
as long as $x$ is live.  For example, \code{s}'s increment (lines
5--8) is live until the return of the function in
Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
\code{s}'s upper bound is properly extended.

\subsection{Typing}\label{sec:type-system}

We now turn to the \lang type system.
The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
which states that in type environment $\Gamma$ (mapping variables to
their types) and predicate environment $\Theta$ (mapping integer-typed
variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
in mode $m$. Key rules for this judgment are given in
Fig.~\ref{fig:type-system-1}. In the rules, $m \le m'$ uses
the two-point lattice with $\umode < \cmode$. All remaining rules
  are given in Appx.~\ref{sec:literal-pointer-typing}~and~\ref{rem-type}.

\begin{DIFnomarkup}
\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[T-Cast]
              {m = \cmode \Rightarrow \tau \neq\tptr{\tau''}{\cmode} \text{ for any $\tau''$} \\\\
                \Gamma;\Theta \vdash_m e : \tau'}
              {\Gamma;\Theta \vdash_m \ecast{\tau}{e} : \tau}

     \inferrule[T-CastCheckedPtr]
               {\Gamma;\Theta \vdash_m e : \tau' \\
                 \tau' \sqsubseteq \tptr{\tau}{\cmode}}
               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\cmode}}{e} : \tptr{\tau}{\cmode}}

      \inferrule[T-DynCast]
                {\Gamma;\Theta \vdash_m e : \tptr{\tallarrayb{\bvar'}{\tau}}{m}}
                {\Gamma;\Theta \vdash_m \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e} : \tptr{\tallarrayb{\bvar}{\tau}}{m}}

      \inferrule[T-Str]
                {\Gamma; \Theta \vdash_m e :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} 
                }
                {\Gamma;\Theta \vdash_m \estrlen{e} : \tint}

      \inferrule[T-LetStr]
                {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\\\
                 \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}];\Theta[x\mapsto \texttt{ge}\_0] \vdash_m e : \tau}
                {\Gamma;\Theta \vdash_m \elet{x}{\estrlen{y}}{e} : \tau}

                \inferrule[T-Ret]
    {\Gamma(x)\neq \bot \\
          \Gamma;\Theta \vdash_m e : \tau}
    {\Gamma;\Theta \vdash_m \eret{x}{\mu}{e} : \tau}

      \inferrule[T-If ]
        {\Gamma; \Theta \vdash_m e : \tau \\\\ \Gamma; \Theta \vdash_m e_1 : \tau_1 \\\\ \Gamma; \Theta \vdash_m e_2 : \tau_2}
        {\Gamma; \Theta \vdash_m \eif{e}{e_1}{e_2} : \tau_1 \sqcup \tau_2}

      \inferrule [T-IfNT]
                { \Gamma; \Theta \vdash_m x : \tptr{\tntarrayb{(b_l,0)}{\tau}}{c} \\\\
                   \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}]; \Theta \vdash_m e_1 : \tau_1\\\\
                  \Gamma;\Theta \vdash_m e_2 : \tau_2 }
                {\Gamma;\Theta \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_1 \sqcup \tau_2}

   \inferrule[T-Let]
    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
           \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
             }
    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}

    \inferrule[T-DefArr]
              {m \leq m' \\\\ \Gamma;\Theta \vdash_{m} e : \tptr{\tallarrayb{\bvar}{\tau}}{m'}}
              {\Gamma;\Theta \vdash_m \estar{e} : \tau}

    \inferrule[T-AssignArr]
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\\\
                \Gamma; \Theta \vdash_m e_2 : \tau' \\
                \tau'\sqsubseteq \tau\\
                m \leq m'}
              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}

\inferrule[T-Fun]
    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\\\
        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
             \overline{\tau'} \sqsubseteq 
               \overline{\tau}[\overline{e} / \overline{x}]}
    {\Gamma; \Theta \vdash_m f(\overline{e}) : \tau[\overline{e} / \overline{x}]}
  \end{mathpar}
}
\caption{Selected Type Rules}
\label{fig:type-system-1}
\end{figure*}
\end{DIFnomarkup}

\myparagraph{Pointer Access}
Rules \textsc{T-DefArr} and \textsc{T-AssignArr} typecheck array
dereference and assignment operations resp. returning the type of
pointed-to objects; rules for pointers to single objects are
similar. The condition $m\le m'$ ensures that checked pointers cannot
be dereferenced in $\eunchecked{e}$ blocks; the type rule for
$\eunchecked{e}$  (not shown) sets $m = \umode$ when checking $e$.
The rules do not attempt to reason whether the access is in bounds;
this check is deferred to the semantics.

\myparagraph{Casting and Subtyping}
Rule \textsc{T-Cast} rule forbids casting to checked pointers when in checked
regions (when $m = \cmode$), but $\tau$ is unrestricted when
$m = \umode$. The \textsc{T-CastCheckedPtr} rule
permits casting from an expression of type $\tau'$ to a checked pointer when
$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype}; the many
rules ensure the relation is transitive. Most of the rules handle
casting between array pointer types; the second rule 
$0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
\tarrayptr{b_l}{b_h}{\tau}{m}$ permits treating a singleton
pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.

Since bounds expressions may
contain variables, determining assumptions like $b_l \leq b_l'$
requires reasoning about those variables' possible values. The type
system uses $\Theta$ to make such reasoning more precise.\footnote{So,
  technically, the subtyping relation $\sqsubseteq$ and the bounds
  ordering relation $\leq$ are parameterized by $\Theta$; this fact is
  implicit to avoid clutter.} $\Theta$ is a map from variables $x$ to
predicates $P$, which have the form $P ::= \top \;|\; \texttt{ge}\_0$.
If $\Theta$ maps $x$ to $\top$, that means that the variable can
possibly be any value; $\texttt{ge}\_0$ means that $x \ge 0$. We will
see how $\Theta$ gets populated and give a detailed example of
subtyping below.\footnote{As it turns out, the subtyping relation is
  also parameterized by $\varphi$, which is needed when type checking
  intermediate results to prove type preservation; source programs
  would always have $\varphi = \emptyset$. Details are in
Appendix~\ref{app:le}.}

Rule \textsc{T-DynCast} typechecks dynamic casting operations, which
apply to array pointer types only. The cast is accepted by the type
system, as its legality will be checked by the semantics.

\begin{DIFnomarkup}
\begin{figure}
{\small
\begin{center}
  $\begin{array}{r@{~}c@{~}l@{~}c@{~}l}
    &&\multicolumn{1}{r}{\tau} &\sqsubseteq& \tau\\[0.6em]
    0\le b_l \wedge b_h \le 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.6em]
    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
    b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
    \end{array}
  $
\end{center}
}
  \caption{Subtyping Relation}
  \label{fig:checkc-subtype}
\end{figure}
\end{DIFnomarkup}

\myparagraph{Bounds Widening}
The bounds of NT array pointers may be widened at conditionals, and
due to calls to $\estrlentext$.  Rule \textsc{T-If} handles normal
branching operations; rule \textsc{T-IfNT} is specialized to the case
of branching on $\estar{x}$ when $x$ is a NT array pointer whose upper
bound is 0. In this case, true-branch $e_1$ is checked with $x$'s type
updated so that its upper bound is incremented by 1; the else-branch
$e_2$ is typechecked under the existing assumptions. For both rules,
the resulting type is the join of the types of the two branches
(according to subtyping). This is important for the situation when $x$
itself is part of the result, since $x$ will have
different types in the two branches.

Rule \textsc{T-Str} handles the case for when $\estrlentext(y)$ does
not appear in a let binding. Rule \textsc{T-LetStr} handles the case
when it does, and performs bounds widening. The result of the call is
stored in variable $x$, and the type of $y$ is updated in $\Gamma$ when
checking the let-body $e$ to indicate that $x$ is $y$'s upper
bound. Notice that the lower bound $b_l$ is unaffected by the call to
$\estrlentext(y)$; this is sound because we know that $\estrlentext$
will always return a result $n$ such that $n \ge b_h$, the current
view of $x$'s upper bound.
The type rule tracks $\estrlentext$'s widened bounds
within the scope of $x$, while the bound-widening effect in the semantics applies to the lifetime of $y$.
Our type preservation theorem in Sec.~\ref{sec:theorem}
shows that our type system is a sound model of the \CoreChkC semantics,
and we discuss how we guarantee that the behavior of our compiler formalization and the semantics matches in Sec.~\ref{sec:compilation}.

This rule also extends $\Theta$ when checking $e$, adding a predicate
indicating that $x\ge 0$. To see how this information is used,
consider this example.  The \code{return} on line 16 of
Fig.~\ref{fig:strncat-ex} has an implicit static cast from the
returned expression to the declared function type (see rule
\textsc{T-Fun}, described below). In type checking the $\estrlentext$
on line 4, we insert a predicate in $\Theta$ showing
\code{n}$\;\ge 0$.  The static cast on line 16 is valid according ot
the last line in Fig.~\ref{fig:checkc-subtype}:
{\small
$$
\tallarrayptr{0}{n}{\tau}{c} \sqsubseteq \tallarrayptr{0}{0}{\tau}{c}
$$
}
because $0 \leq 0$ and $0 \leq n$, where the latter holds since
$\Theta$ proves $n \ge 0$.  Without $\Theta$, we would need a
dynamic cast.

In our formal presentation, $\Theta$ is quite simple and is just meant
to illustrate how static information can be used to avoid dynamic
checks; it is easy to imagine richer environments of facts that can be
leveraged by, say, an SMT solver as part of the subtyping
check \cite{LiquidHaskell,Refinementlh}

\myparagraph{Dependent Functions and Let Bindings}
Rule \textsc{T-Fun} is the standard dependent function call rule. It
looks up the definition of the function in the function environment
$\Xi$, typechecks the actual arguments $\overline{e}$ which have
types $\overline{\tau'}$, and then confirms that each of these types
is a subtype of the declared type of $f$'s corresponding parameter. Because
functions have dependent types, we substitute each parameter $e_i$ for
its corresponding parameter $x_i$ in both the parameter types and the
return type. Consider the \code{strncat} function in
Fig.~\ref{fig:strncat-ex}; its parameter type for \code{a} 
depends on \code{n}. The \textsc{T-Fun} rule will substitute 
\code{n} with the argument at a call-site.

Rule \textsc{T-Let} types a $\elettext$ expression, which also admits
type dependency. In particular, the result of evaluating a $\elettext$
may have a type that refers to one of its bound variables (e.g., if
the result is a checked pointer with a variable-defined bound); if so,
we must substitute away this variable once it goes out of scope. Note
that we restrict the expression $e_1$ to syntactically match the
structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).

Rule \textsc{T-Ret} types a $\erettext$ expression, which does not
appear in source programs but is introduced by the semantics when
evaluating a let binding (rule \textsc{S-Let} in
Fig.~\ref{fig:semantics}); this rule is needed for the preservation
proof. After the evaluation of a let binding a variable $x$ concludes,
we need to restore any prior binding of $x$, which is either
$\bot$ (meaning that there is no $x$ originally) or some value
$\evalue{n}{\tau}$.

\subsection{Type Soundness and Blame}\label{sec:theorem}

In this subsection, we focus on our main metatheoretic results about
\lang: type soundness (progress and preservation) and blame. The type
soundness theorems rely on a notion of heap and stack
\emph{well-formedness}:

\smallskip
\begin{defi}[Heap Well-formedness]
A heap $\heap$ is well-formed, iff (i) the null position
($0$) is not defined in $\heap$, and (ii) every type annotation in it
contains no free variables.
\end{defi}
\smallskip
\begin{defi}[Stack Well-formedness]
A stack snapshot $\varphi$ is well-formed, iff every type annotation in
it contains no free variables.
\end{defi}
\smallskip

Moreover, as a program evaluates its expression may contain literals
$n\!:\!\tau$ where $\tau$ is a pointer type, i.e., $n$ is an index in
$\heap$ (perhaps because $n$ was chosen by $\mathtt{malloc}$). The
normal typechecking judgment for $e$ is implicitly parameterized by
$\heap$, and the rules for typechecking literals confirm that
pointed-to heap cells are compatible with (subtypes of) the pointer's
type annotation; in turn this check may precipitate checking the type
consistency of the heap itself. We follow the same approach as
\citet{ruef18checkedc-incr}, and show the rules in
Appendix~\ref{sec:literal-pointer-typing}.

Progress now states that terms that don't reduce are either values or their
mode is unchecked:

\smallskip
\begin{thm}[Type Progress Theorem]\label{thm:progress}
For any \checkedc program $e$ and heap $\heap$, if $e$ and $\heap$ are
well-formed, and $\emptyset;\emptyset\vdash_m e : \tau$, then $e$ is
a value ($\evalue{n}{\tau}$), \code{unchecked} ($m=\umode$), or
there exists $\varphi'$ $\heap'$ $e'$, such that $(\emptyset,\heap,e)
\longrightarrow_m (\varphi',\heap',e')$.
\end{thm}
{\em Proof:} By induction on the typing derivation.

\smallskip

For preservation, we also need to introduce a notion of
\emph{consistency}, relating heap environments before and after a
reduction step, and type environments, predicate sets, and stack
snapshots together.

\smallskip
\begin{defi}[Type-Stack Consistency]
A type environment $\Gamma$, variable predicate set $\Theta$, and stack snapshot $\varphi$ are consistent, iff every variable defined in $\Theta$ is defined in $\Gamma$, and for every variable $x$, $\Gamma(x) = \tau$ implies that $\varphi(x)$ is defined and there exists $n$ and $\tau'$, such that $\varphi(x)=\evalue{n}{\tau'}$ and $\tau' \sqsubseteq \tau$.
\end{defi}
\smallskip
\begin{defi}[Heap Consistency]
A heap $\heap'$ is consistent with $\heap$ iff every address defined
in $\heap$ is defined in $\heap'$.
\end{defi}
\smallskip

\noindent
Armed with the definitions of consistency, we can now prove
preservation, which states that a reduction step preserves both the
type of the expression being reduced, as well as well-formedness and
consistency of environments:

\smallskip
\begin{thm}[Type Preservation Theorem]
For any \checkedc program $e$, heap $\heap$, stack
$\varphi$, type environment $\Gamma$, variable predicate set $\Theta$,
and a type $\tau$, that are all are well-formed, if $\Gamma, \Theta$,
and $\varphi$ are consistent, $e$ is well typed
$\Gamma;\Theta\vdash_{\cmode} e : \tau$, and if there exists $\varphi'$,
$\heap'$ and $e'$, such that $(\varphi,\heap,e)
\longrightarrow_{\cmode} (\varphi',\heap',e')$, then there exists
$\Gamma'$, $\Theta'$ and $\tau'$, such that $\Gamma'$, $\Theta'$,
$\varphi'$, $\heap'$ and $e'$ are well-formed, $\Gamma'$, $\Theta'$
and $\varphi'$ are consistent, $\heap'$ is consistent with $\heap$,
$\Gamma';\Theta' \vdash_{\cmode} e: \tau'$, and $\tau'\sqsubseteq
\tau$.
\end{thm}
{\em Proof:} By induction on the typing derivation.
\smallskip

Using type soundness we can prove our main result, {\em blame}, which
states that if there is any spatial memory safety violation is
triggered, it must necessarily come from the unchecked region.

\smallskip
\begin{thm}\label{thm:blame}
[The Blame Theorem] For any \checkedc program $e$, heap
$\heap$, type $\tau$, if $\heap$ and $e$ are well-formed,
$\emptyset;\emptyset\vdash_{\cmode} e : \tau$, and if there exists
$\varphi'$, $\heap'$, a failure result $r$, and $m$, such that
$(\varphi,\heap,e) \longrightarrow_m^* (\varphi',\heap',r)$, then
there exist $E$ and $e_a$, such that $e' = E[e_a]$, and $\mode(E) = \umode$.
\end{thm}
{\em Proof:} By induction on the number of steps of the \checkedc
evaluation ($\longrightarrow_m^*$), using progress and preservation to
maintain the invariance of the assumptions.

These proofs have been carried out in a Coq development.

% \ignore{
% \begin{figure}[t!]
%   \begin{prooftree}
%     \hypo{\evalue{n}{\tau} \in \defscope}
%     \infer1[T-VConst]{\Gamma;\defscope \vdash_m \evalue{n}{\tau}  : \tau}
%   \end{prooftree}
%   \qquad

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma;\defscope \vdash_m e : \tau \\
%         \Gamma;\defscope \vdash_m e_1 : \tau_1 
%       \end{matrix}
%     }
%     \hypo{
%       \begin{matrix}
% \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2 \\
%             \Gamma;\defscope \vdash_m e_2 : \tau_2
%           \end{matrix}
%         }
%     \infer2[T-If]{\Gamma;\defscope \vdash_m \eif{e}{e_1}{e_2} : \tau_3}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma;\defscope \vdash_m x : \tptr[c]{(\tntarray{l}{0}{\tau})}, l \leq 0 \\
%         \Gamma, x : \tptr[c]{(\tntarray{l}{1}{\tau})};\defscope \vdash_m e_1 : \tau_1 \\
%         \Gamma;\defscope \vdash_m e_2 : \tau_2 \\
%         \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2
%       \end{matrix}
%     }
%     \infer1[T-IfNT]{\Gamma;\defscope \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_3}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         F(f) = \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e  \\
%         \Gamma; \defscope \vdash_m \tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}] ~~~ 0 \leq i \leq j\\
%         \Gamma; \defscope \vdash_m e_i : \tau_i' ~~~ 0 \leq i \leq j\\
%         \Gamma; \defscope \vdash_m \subtype{\tau_i'}{\tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}]}  ~~~ 0 \leq i < j
%       \end{matrix}
%     }
%     \infer1[T-VCall]{\Gamma; \defscope \vdash_m f(\overline{e}) : \tau_j[x_0,\ldots,x_{j-1} \mapsto e_0,\ldots,e_{j-1}]}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%           \Gamma; \defscope \vdash_m e_1 : \tau_1
%       }
%       \hypo{        \Gamma, x = e_1 : \tau_1; \defscope \vdash_m e_2 : \tau_2
% }
%     \infer2[T-Let]{\Gamma; \defscope \vdash_m \elet{x}{e_1}{e_2} : \tau[x \mapsto e_1]}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m \estrlen{y} : \tau_1\\
%         \Gamma; \defscope \vdash_m y : \tptr[c]{(\tntarray{le}{\_}{\tau_3})}\\
%         \Gamma, x = \estrlen{y} : \tau_1, \\ y : \tptr[c]{(\tntarray{le}{x}{\tau_3})} ;  \defscope  \vdash_m e_2 : \tau_2
%         \end{matrix}
%         }
%         \infer1[T-LetStr]{\Gamma; \defscope \vdash_m \elet{x}{\estrlen{y}}{e_2} : \tau[x \mapsto e_1]}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{
%         \Gamma; \defscope \vdash_m e : \tptr[c]{(\tntarray{le}{he}{\tau})}
%       }
%         \infer1[T-Str]{\Gamma; \defscope \vdash_m \estrlen{y} : \tint}
%       \end{prooftree} \\ \\

%       \begin{prooftree}
%         \hypo{
%           \begin{matrix}
%             \tau = \tint \vee \tau = \tptr[u]{\omega}~\vee  n=0~ \vee \\ \tau = \tptr[c]{(\tarray{0}{0}{\tau'})} \vee \\
%             \tau = \tptr[c]{(\tntarray{0}{0}{\tau'})}
%         \end{matrix}
%       }
%       \infer1[T-Base]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
%     \end{prooftree} \\ \\

%     \begin{prooftree}
%       \hypo{
%         \begin{matrix}
%         \tau = \tptr[c]{\omega} \\
%         \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\
%         \Gamma;\defscope, \evalue{n}{\tau}  \vdash_m {H(n+k)} : {\tau_k} ~~~ 0 \leq k < j
%       \end{matrix}
%       }
%       \infer1[T-PtrC]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
%     \end{prooftree} \\ \\

%     \begin{prooftree}
%       \hypo{
%         \begin{matrix}
%             \Gamma; \defscope \vdash_m e :  {\tptr[m]{\tstruct{T}}} \\
% D(T) = ...; \tau_f~f; ...
%         \end{matrix}
%       }
%       \infer1[T-Amper]{\Gamma; \defscope \vdash_m \eamper{e}{f} : \tptr[m]{\tau_f}}
%     \end{prooftree} \\ \\

%     \todo[inline]{YL: how to express le - n as a metafunction?}
%     \begin{prooftree}
%       \hypo{
%         \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}\\
%         \Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tint \\
%         le' = le - n, he' = he - n
%       \end{matrix}
%     }
%       \infer1[T-BinopInd]{\Gamma; \defscope \vdash_m (e_1 \plus \evalue{n}{\tau} ) : \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})}}
%     \end{prooftree}
% \caption{Typing}
% \label{fig:typing}
% \end{figure}

% \begin{figure}[t!]
%   \todo[inline]{no sizeof check because we don't know statically whether the allocation would be null}
%   \begin{prooftree}
%     \hypo{
%       \omega = \tgarray{\alpha}{le}{he}{\tau} \Rightarrow le = 0
%     }
%     \infer1[T-Malloc]{\Gamma; \defscope \vdash_m \emalloc{\omega} : \tptr[c]{\omega}}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{
%       \Gamma; \defscope \vdash_u e : \tau
%     }
%     \infer1[T-Unchecked]{\Gamma; \defscope \vdash_m \eunchecked{e} : \tau}
%   \end{prooftree}\\\\

%   \todo[inline]{any constraints on m'?}
%   \begin{prooftree}
%     \hypo{\Gamma; \defscope \vdash_m e :  \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})} }
%     \infer1[T-DynCast]{\Gamma; \defscope \vdash_m \edyncast{\tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}}{e} : \tau}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{\Gamma; \defscope \vdash_m e : \tau'}
%     \infer1[T-Cast]{\Gamma; \defscope \vdash_m \ecast{\tau}{e} : \tau}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\tgarray{\alpha}{le}{he}{\tau}} \\
%         \Gamma; \defscope \vdash_m e_2 : \tint \\
%         m' = u \Rightarrow m = u \\
%       \end{matrix}
%     }
%     \infer1[T-Index]{\Gamma; \defscope \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
%         \Gamma; \defscope \vdash_m e_2 : \tau \\
%         m' = u \Rightarrow m = u \\
%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
%       \end{matrix}
%     }
%     \infer1[T-Assign]{\Gamma; \defscope \vdash_m \eassign{e_1}{e_2} : \tau}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
%         \Gamma; \defscope \vdash_m e_2 : \tint \\
%         \Gamma; \defscope \vdash_m e_3 : \tau \\
%         m' = u \Rightarrow m = u \\
%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
%       \end{matrix}
%     }
%     \infer1[T-IndAssign]{\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
%   \end{prooftree}

% \caption{Typing Cont.}
% \label{fig:typing2}
% \end{figure}
% \ignore{

% \begin{figure}[t!]
%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j} \\
%         \Gamma, x_0 = \kw{none} : \tau_0, \ldots, x_{j-1} = \kw{none} : \tau_{j-1}; \defscope \vdash e : \tau_{r}
%       \end{matrix}
%     }
%     \infer1[WF-Fun]{\Gamma;\defscope \vdash \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \infer0[WF-Nil]{\Gamma \vdash \circ}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{
%       \begin{matrix}
%         \Gamma \vdash \tau_0 \\
%         \Gamma, x_0 = \kw{none} :\tau_0 \vdash x_1:\tau_{1}, \ldots, x_{j}:\tau_{j}
%       \end{matrix}
%     }
%     \infer1[WF-Cons]{\Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \infer0[WF-Int]{\Gamma \vdash \tint}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{   \Gamma \vdash le }
%      \hypo{   \Gamma \vdash he }
%       \hypo{  \Gamma \vdash \tau}

%     \infer3[WF-Array]{\Gamma \vdash \tptr[m]{(\tgarray{\alpha}{le}{he}{\tau})}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{T \in dom(D)}
%     \infer1[WF-Struct]{\Gamma \vdash \tptr[m]{\tstruct{T}}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \hypo{\Gamma \vdash \tau}
%     \infer1[WF-Ptr]{\Gamma \vdash \tptr[m]{\tau}}
%   \end{prooftree} \\\\

%   \begin{prooftree}
%     \infer0[WFB-Int]{\Gamma \vdash i}
%   \end{prooftree}\\\\

%   \begin{prooftree}
%     \hypo{x = e? : \tint \in \Gamma}
%     \infer1[WFB-Var]{\Gamma \vdash x \plus i}
%   \end{prooftree} \\\\

%   \caption{Well-Formedness}
%   \label{fig:wf}
% \end{figure}

% \begin{figure}[t!]
%   \begin{prooftree}
%     \infer0[Sub-Nt]{\subtype{\tptr[c]{(\tntarray{le}{he}{\tau})}}{\tptr[c]{(\tarray{le}{he}{\tau})}}}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \infer0[Sub-Refl]{\subtype{\tau}{\tau}}
%   \end{prooftree} \\ \\

%   \todo[inline]{what about ntarrays?}
%   \begin{prooftree}
%     \infer0[Sub-Ptr]{\subtype{\tptr[c]{\tau}}{\tptr{\tarray{0}{1}{\tau}}}}
%   \end{prooftree} \\ \\

%   \todo[inline]{the subtyping relation is not anti-symmetric..}
%   \begin{prooftree}
%     \infer0[Sub-Arr]{\subtype{\tptr[c]{\tarray{0}{1}{\tau}}}{\tptr{\tau}}}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
%     \hypo{\rboundle{le_0}{le_1}}
%     \hypo{\rboundle{he_1}{he_0}}
%     \infer2[Sub-Subsume]{\subtype{\tptr{(\tgarray{\alpha}{le}{he}{\tau})}{\cmode}}{ \tptr{(\tgarray{\alpha}{ le_1} {he_1} {\tau})}}}
%   \end{prooftree} \\ \\

%   \begin{prooftree}
% \hypo{D(T) = \tau_f~f; ...}
%     \hypo{\subtype{\tau_f}{\tau}}
%     \infer2{\subtype{\tptr[c]{\tstruct{T}}}{\tau}}
%   \end{prooftree}
%   \caption{Subtyping}
%   \label{fig:sub}
% \end{figure}
% }
% }

% \ignore{

% \begin{figure*}[t]
%   \begin{lstlisting}
% int foo(nt_array_ptr<char> p : count(0)) {
%   if (* p) {
%     dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
%   }
%   dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
%   return 0;
% }
%   \end{lstlisting}
% \caption{The example where clang Checked C fails at run-time}
% \label{fig:clangbad1}
% \end{figure*}

% \begin{figure*}[t]
%   \begin{lstlisting}
% /* nt_array_ptr<char> p : bounds(p,p) */
% size_t cnt = 0;

% while(*(p+cnt)) {
%   ++cnt;
% }
% dyn_bounds_cast<nt_array_ptr<char>>(p, count(cnt));
%   \end{lstlisting}
% \caption{A useful program that the Checked C spec doesn't allow at run-time}
% \label{fig:clangbad2}
% \end{figure*}
% }

% \ignore{
% The Clang CheckedC implementation uses statically determined
% bounds to insert run-time checks. In Fig.~\ref{fig:clangbad1}, the
% \code{dyn_bounds_cast} at line 3 will always succeed, because the
% compiler knows that within the scope of then branch, the pointer
% \code{p} must have at least one element. The same cast at line 5,
% however, will always fail, since there is no way to tell statically
% whether the program has entered the then branch before. The compiler
% will check whether the \code{count(1)} bounds specification is
% contained within the earlier \code{count(0)} specification, resulting
% in a run-time failure even when we pass in a non-empty string.

% Our formalization diverges from this run-time behavior and instead keeps
% track of the bounds on the stack. After entering the then branch, we
% increment the upper bound for \code{p}, effectively making the
% updated bounds information available even after we exit the if
% statement. The cast at line 5 will be checking the new bounds against
% the incremented bounds for non-empty strings.

% Fig.~\ref{fig:clangbad2} gives a more practical example of why keeping
% track of the bounds on the stack is useful. The program snippet
% implements the functionality of the \code{strlen} function using a
% while loop and a \code{cnt} variable. Even though the type system is
% unable to reason about the while loop, as long as the runtime system
% updates the bounds in-place, the user can apply a
% \code{dyn_bounds_cast} to soundly recover the more precise bounds
% information.
% }

 \section{Compilation}\label{sec:compilation}

The semantics of \lang uses annotations on pointer literals in
order to keep track of array bounds information, which is checked at
dereferences and changed during widening. However, in the real
implementation of \checkedc, these annotations
are not present---pointers are represented as a single
machine word with no extra metadata. We show how the
annotations can be safely erased: using static information a compiler
can insert code to manage and check bounds metadata, with no loss of
expressiveness.

This section sketches our compilation algorithm that converts from
\lang to \elang, an untyped language without metadata annotations.
Compilation is defined by extending \lang's
typing judgment thusly:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each \code{nt_array_ptr} variable $p$ to a pair of \emph{ghost
  variables} that keep $p$'s up-to-date upper and lower bounds; these
may differ from the bounds in $p$'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound ghost variable is unnecessary; we include it for uniformity.}
When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
complete judgment, implicitly assuming that $e$ is a well-typed and closed
term.

We formalize rules for this judgment in PLT Redex~\cite{pltredex},
following and extending our Coq development for \lang. To give
confidence that compilation is correct, we use Redex's property-based
random testing support to show that compiled-to $\dot e $ simulates
$e$, for all $e$.

\subsection{Approach}

Due to space constraints, we explain the rules for compilation by
example; the complete rules are given in Appendix~\ref{appx:comp1}. 
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks; and (c) insertion of
bounds widening expressions.
A-normal form conversion is straightforward: compound expressions are
handled by storing results of subexpressions into temporary variables,
as in the following example.

{\vspace*{-0.5em}
{\small
\begin{center}
$
\begin{array}{l}
$\code{y=(x+1)+(6+1);}$
\;
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\;
\begin{array}{l}
$\code{a=x+1;}$\\
$\code{b=6+1;}$\\
$\code{y=a+b;}$\\
\end{array}
\end{array}
$
\end{center}
}
}

This simplifies the management of effects from subexpressions. The
next two steps of compilation are more interesting.

\begin{figure}[t!]
  \begin{small}
\begin{lstlisting}[mathescape,xleftmargin=4 mm]
/* nt_array_ptr<int> p : count(p,p) */
/* $\rho$(p) = p_lo,p_hi */
{
  int x = strlen(p);
  if (x > 1) putchar(*(p+1));
}
\end{lstlisting}
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\begin{lstlisting}[xleftmargin=4 mm]
{
  assert(p_lo <= p && p <= p_hi);  //bound check
  assert(p != NULL);               //null check
  int x = strlen(p);
  int *p_hi_new = p + x;
  p_hi = max(p_hi, p_hi_new);
  if (x > 1) {
    assert(p != NULL);             //null check p
    int *p_1 = p + 1;
    assert(p_lo <= p_1 && p_1 <= p_hi); // p+1
    assert(p_1 != NULL);          //null check p+1
    putchar(*p_1);
  }
}
\end{lstlisting}
\end{small}
\caption{Compilation example for check insertions}
\label{fig:compilationexample}
\end{figure}

\begin{figure}[t!]
  \begin{small}
\begin{lstlisting}[mathescape,xleftmargin=4 mm]
int deref_array (int n, 
      nt_array_ptr<int> p : bounds(p, p+n)) {
  /* $\rho$(p) = p_lo,p_hi */
  if (*p) return *(p+1);
  return 0;
}
...
// nt_array_ptr<int> p0 : bounds(p0, p0+5)
deref_array(5, p0);
    \end{lstlisting}
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\begin{lstlisting}[xleftmargin=4 mm]
int deref_array(int n, int *p) {
  int *p_lo = p;
  int *p_hi = p + n;
  /* runtime checks */
  assert(p_lo <= p && p <= p_hi);
  assert(p != NULL);
  int p_derefed = *p;
  if (p_derefed != '\0') {
    /* widening */
    if (p_hi == p) {
      ++p_hi;
    }
    int *p0 = p + 1;
    assert(p_lo <= p0 && p0 <= p_hi);
    assert(p0 != NULL);
    return *p0;
  }
  return 0;
...
//int *p0, set_bounds(p0) = p_lo, p_hi
deref_array(5, p0);
    \end{lstlisting}
\end{small}
    \caption{Compilation example for dependent functions}
\label{fig:compilationexample1}
\end{figure}

During compilation, $\Gamma$ tracks the lower and upper bound
associated with every pointer variable according to its type. At each
declaration of a \code{nt_array_ptr} variable $p$, the
compiler allocates two \emph{ghost variables}, 
stored in $\rho(p)$; these are initialized to $p$'s declared bounds
and will be updated during bounds widening.\footnote{Ghost variables
  are not used for \code{array_ptr} types (the bounds expressions are)
  since they are not subject to bounds widening.} 
Fig.~\ref{fig:compilationexample} shows how an invocation of
\code{strlen} on a null-terminated string is compiled into C
code. Each dereference of a checked pointer requires a null check
(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
compiler makes explicit: Line~$3$ of the generated code has the null
check on pointer $p$, and similar checks happen at line~$8$ and
line~$11$.
Dereferences also require bounds checks: line~$2$ checks \code{p} is
in bounds before computing \code{strlen(p)}, while line~$10$ does
likewise before computing \code{*(p+1)}.

For \code{strlen(p)} and conditionals \code{if(*p)}, the \lang
semantics allows the upper bound of \code{p} to be extended.
The compiler explicitly inserts statements to do so on \code{p}'s
ghost bound variables. For example,
Fig.~\ref{fig:compilationexample}~line~$6$ widens \code{p}'s upper
bound if \code{strlen}'s result is larger than the existing
bound. 
Lines 7--12 of the generated code in
Fig.~\ref{fig:compilationexample1} show how bounds are 
widened when compiling expression \code{if(*p)}. If we find that the
current \code{p} address is equal to the upper bound (line 10),
and \code{p}'s content is not null (line 8), we then increase the
upper bound by $1$ (line 11).
Fig.~\ref{fig:compilationexample1} also shows a dependent function call.
Notice that the bounds for the array pointer \code{p} are not passed as
arguments. Instead, they are initialized according to \code{p}'s
type---see line~3 of the original \lang program at the top of the figure.
Line~$2$ of the generated code
sets the lower bound  to \code{p} and line~$3$ sets the
upper bound to \code{p+n}.

\subsection{Comparison with Checked C Specification}
\label{sec:disc}

The use of ghost variables for bounds widening is a key novelty of our
compilation approach, and adds more precision to bounds checking at runtime
compared to the official specification and current implementation of
\checkedc~\cite[5.1.2, pg 85]{checkedc}.  For example, the
\code{strncat} example of Fig.~\ref{fig:strncat-ex} compiles with the
current Clang \checkedc compiler but will fail with a runtime
error. The statement \code{int x = strlen(a)} at line 4 changes the
upper bound of \code{a} to \code{x}, which can be smaller than
\code{n}, the capacity of the array pointer \code{a}. The assignment
at line 14 will always fail if the index \code{x + i} is checked
against the statically determined upper bound
\code{x}. This forces us to inline the definition of \code{strlen} as
in \code{strncat_c} to avoid runtime errors when running code compiled with the
Clang \checkedc compiler.  Likewise, if we were to add another dereference to
$p$ after line 6 in the original code at the top of
Fig.~\ref{fig:compilationexample}, the Clang \checkedc
compiler would 
check \code{p} against its original bounds \code{(p,p)} since the
updated upper bound \code{p+x} cannot be retained with \code{x} out of
the scope. In the presence of ghost variables,
these bounds have been widened by the assignment in line $5$ (assuming
the null-terminator was not the first element of the string) and
remain available in the entire stack frame, and
therefore the check will succeed. In contrast, in the actual
implementation of Checked C, the scope of the widening is limited to
the scope of the conditional at both runtime and compile time, which means that the inserted dynamic
check would fail.
To make it match the specification, our compilation definition could
rely only on the type-based bounds expressions available in $\Gamma$
for checking, and eschew ghost variables. However, doing so would
force us to weaken the simulation theorem, reduce expressiveness,
and/or force the semantics to be more awkward. We plan to work with
the \checkedc team to implement our approach in a future revision.

\subsection{Metatheory}
\label{sec:meta}

While designing our Coq model of \lang, we also designed a model in
PLT Redex.\footnote{The two models, in Redex and Coq, are
  equivalent, with the only difference being in the representation of
  stacks: as we saw, the Coq model uses an explicit map for
  representing stacks to ease the effort of theorem proving; on the
  other hand, the Redex model uses \texttt{let} bindings to simulate a
  stack, which removes the need to account for the stack during random
  generation of terms.}  Redex~\cite{pltredex} is a semantic
engineering framework implemented in Racket, which allows for
concisely specifying semantics and typing rules. We
formalize the simulation theorem in this model, and then attempt to
falsify it via Redex's support for random testing. We ultimately plan
to prove simulation in the Coq model.

Turning to the simulation theorem: We first introduce notation
used to specify the theorem. We use the notation $\gg$ to
indicate the \emph{erasure} of stack and heap---the rhs is the same as
the lhs but with type annotations removed:
\begin{equation*}
  \begin{split}
    \heap  \gg & \dot \heap \\
    \varphi \gg & \dot \varphi
  \end{split}
\end{equation*}
In addition, we write $(\varphi,\heap,e) \gg (\dot \varphi, \dot \heap,
\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
and $e \gg \dot e$ respectively.

We use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
reduction relation of $\elang$. Unlike the $\lang$, the semantics of
$\elang$ does not distinguish checked and unchecked regions.

Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
that is similar to the one by
~\citet{merigoux2021catala}. An ordinary simulation property would
replace the middle and bottom parts of the figure with the
following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
Instead, we relate two erased configurations using the relation $\sim$,
which only requires that the two configurations will eventually reduce
to the same state. We formulate our simulation theorem differently
because the standard simulation theorem imposes a very strong
syntactic restriction to the compilation strategy. Very often, $(\dot
\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
but we are unable to syntactically equate the two configurations due
to the extra binders generated for dynamic checks and ANF
transformation. In earlier versions of the Redex model, we attempted
to change the compilation rules so the configurations could match
syntactically. However, the approach scaled poorly as we added
additional rules. 
This slight relaxation on the equivalence relation
between target configurations allows us to specify compilation more
naturally without having to worry about syntactic constraints.

\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
\rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
some $e_1$ and
$\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
, then
there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
\varphi,\dot \heap,\dot e)$ and $(\dot
\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
\dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
\varphi_1$, $\heap_1 \gg \dot \heap_1$.
\end{thm}

Our random generator (discussed in the next section) never produces
$\euncheckedtext$ expressions (whose behavior could be undefined), so we
can only test a the simulation theorem 
as it applies to checked code. This limitation makes it
unnecessary to state the other direction of the simulation theorem
where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
that $e_0$ will never enter a stuck state if it is well-typed in
checked mode.

The current version of the Redex model has been tested against $20000$
expressions with depth less than $10$. Each expression can
reduce multiple steps, and we test simulation between every two
adjacent steps to cover a wider range of programs, particularly the
ones that have a non-empty heap.

\begin{figure}[t]
{\small
\[
\begin{array}{c}
\begin{tikzpicture}[
            > = stealth, shorten > = 1pt, auto,
            node distance = 3cm
        ]

\begin{scope}[every node/.style={draw}]
    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
\end{scope}
\begin{scope}[every edge/.style={draw=black}]

    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
    \path [<->] (A) edge node {$\gg$} (C);
    \path [<->] (B) edge node {$\gg$} (D);
    \path [dashed,<->] (C) edge node {$\sim$} (D);
    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
\end{scope}

\end{tikzpicture}
\end{array}
\]
}
\caption{Simulation between \lang and \elang }
\label{fig:checkedc-simulation-ref}
\end{figure}

% \ignore
% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist. The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.

%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C. The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).

% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation. We generate tens of thousands cases and cover all possible corner cases of the compilation. 
% }
% \ignore{
% \begin{figure*}[t]
%   \begin{subfigure}[b]{1\textwidth}
%     \begin{lstlisting}
% % int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
%   return *p;
% }

% ...
% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
% deref_array(5, p0);
%     \end{lstlisting}
%     \label{fig:chkcexamplederef}
%     \caption{\texttt{deref\_array} in \checkedc}
%   \end{subfigure}

%   \begin{subfigure}[b]{1\textwidth}
% \begin{lstlisting}
% % int deref_array(size_t n, int *p) {
%   /* statically determine the definitions p_lo and p_hi */
%   int *p_lo = p;
%   int *p_hi = p + n;
%   /* runtime checks */
%   assert(p_lo <= p && p <= p_hi);
%   assert(p != NULL);
%   /* widening */
%   int p_derefed = *p;
%   if (p_derefed != '\0') {
%     if (p_hi == p) {
%       ++p_hi;
%     }
%   }
%   return p_derefed;
% }
% ...
% /* int *p0 */
% /* rho(p0) = p_lo, p_hi */
% deref_array(5, p0);
% \end{lstlisting}
% \caption{\texttt{deref\_array} in C}
%     \label{fig:cexamplederef}
%   \end{subfigure}

%   \begin{subfigure}[b]{\textwidth}
%     \begin{lstlisting}
% % /* nt_array_ptr<int> p : bounds(e0, e1) */
% /* rho(p) = p_lo, p_hi */
% size_t x = strlen(p);
% if(x >= 1) {
%   *(p+1);
% }
%     \end{lstlisting}
%     \caption{\texttt{strlen} in \checkedc}
%     \label{fig:chkcexamplestrlen}
%   \end{subfigure}

%   \begin{subfigure}[b]{\textwidth}
%     \begin{lstlisting}
% % /* nt_array_ptr<int> p : bounds(e0, e1) */
% /* rho(p) = p_lo, p_hi */
% /* runtime checks omitted */
% ...
% size_t x = strlen(p);
% int *p_hi_new = p + x;
% p_hi = max(p_hi, p_hi_new);
% if (x >= 1) {
%   /* null check for pointer arithmetic */
%   assert(p != NULL);
%   int *p_1 = p + 1;
%   /* null check for dereference */
%   assert(p_1 != NULL);
%   /* bounds check for dereference */
%   /* note how we are using p_hi rather than p + x */
%   assert(p_lo <= p_1 && p_1 <= p_hi);
%   ...
% }
%     \end{lstlisting}
%     \caption{\texttt{strlen} in C}
%     \label{fig:cexamplestrlen}
%   \end{subfigure}
% \caption{Compilation of null-terminated string dereference}
% \label{fig:compilationexample}
% \end{figure*}
% }

% \ignore{
% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
% dereference of null-terminated strings by inserting explicit checks and
% bounds widening code. For readability, we present the example in real
% \checkedc and C syntax. At
% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
% bounds are defined in terms of the arguments \code{p} and \code{n}
% according to the bounds annotations obtained during typechecking. This
% avoids the need to pass in bounds as extra arguments, thus maintaining
% compatibility with C code. When we call \code{deref_array} on
% \code{p0}, a string with size 5 (excluding the null character), there
% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
% the stack. It is possible that \code{p_hi} has been widened
% before we reach line 21. The programmer will have to perform a
% \code{dynamic_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
% attempts to widen the bound when the dereferenced result is not
% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
% \code{deref_array}, in contrast to the T-IfNT rule, which only
% remembers the widened bounds within the scope of the then branch of
% the if statement.

% \todo[inline]{Widening can happen at every
%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
% shows how an invocation of \code{strlen} on null-termianted strings
% is compiled into C code. We perform the same runtime checks that
% happen during dereferencing. The widening code at line
% 8 updates \code{p}'s upper bound only if the result of
% \code{strlen} is larger than the value of the upper bound stored on
% the stack. This is another scenario where the runtime can be more precise
% than the statically determined bounds information.

% }
 \section{Random Testing via the Implementation}\label{sec:evaluation}

In addition to using the \lang Redex model to establish simulation of
compilation (Section~\ref{sec:meta}), we also used it to gain confidence
that our model matches the Clang \checkedc implementation;
disagreement on outcomes signals a bug in either the model or the
compiler itself. Doing so allowed us to quickly iterate on the
design of the model while adding new features, and revealed
several bugs in the Clang \checkedc implementation.

\myparagraph{Generating Well Typed Terms} For this random generation, we follow
the approach of~\citet{PalkaAST11} to generate well-typed \checkedc
terms by viewing the typing rules as generation rules.
Suppose we have a context $\textcolor{red}{\Gamma}$, a mode
$\textcolor{red}{m}$ and a type $\textcolor{red}{\tau}$, and we are
trying to generate a well-typed expression. We can do that by
reversing the process of type checking, selecting a typing rule and
building up an expression in a way that satisfies the rule's premises.

Recall the typing rule for dereferencing an
array pointer, which we depict below as
\textsc{G-DefArr}\footnote{Generator rules G-* correspond one to one
  with the type rules T-* in Sec.~\ref{sec:type-system}.}, color-coded to represent \textcolor{red}{inputs} and \textcolor{blue}{outputs} of the
generation process:\footnote{This input-output marking is commonly called a mode in
  the literature, but we eschew this term to avoid
  confusion with our pointer mode annotation.}
\begin{mathpar}
    \inferrule[G-DefArr]
              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} : \tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m_a}} \\
\textcolor{red}{m} \leq \textcolor{cyan}{m_a} 
              }
              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\estar{e}} : \textcolor{red}{\tau}}
\end{mathpar}
If we selected \textsc{G-DefArr} for generating an expression, the
generated expression has to have the form $\textcolor{blue}{*e}$,
for some $\textcolor{blue}{e}$, to be generated according
to the rule's premises.
To satisfy the premise
$\textcolor{red}{\Gamma};\textcolor{red}{\Theta}
\vdash_{\textcolor{red}{m}} \textcolor{blue}{e} :
\tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m_a}}$,
we essentially need to make a recursive call to the generator, with
appropriately adjusted inputs.
However, the type in this judgment is not fixed yet---it contains
three unknown variables: $\textcolor{cyan}{m_a}$,
$\textcolor{cyan}{\bvar}$, and $\textcolor{cyan}{\kappa}$---that need
to be generated before making the call.
Looking at the second premise informs that generation:
if the input mode $\textcolor{red}{m}$ is $\umode$, then $\textcolor{cyan}{m_a}$
needs to be $\umode$ as well; if not, it is unconstrained,
just like $\textcolor{cyan}{\bvar}$ and $\textcolor{cyan}{\kappa}$, and therefore all three are free to be generated at random.
Thus, the recursive call to generate $\textcolor{blue}{e}$ can now
be made, and the \textsc{G-DefArr} rule returns $\textcolor{blue}{*e}$
as its output.

Using such generator rules, we can create a generator for random
well-typed terms of a given type in a straightforward manner: find all
rules whose conclusion matches the given type and then randomly choose
a candidate rule to perform the generation. To ensure that this
process terminates, we follow the standard practice of using ``fuel'' to
bound the depth of the generated terms; once the fuel is exhausted,
only rules without recursive premises are
selected~\cite{Pierce:SF4}. Similar methods were used for
generating top level functions and \kw{struct} definitions.

While using just the typing-turned-generation rules is in theory
enough to generate all well-typed terms, it's more effective in
practice to try and exercise interesting patterns.  As
in~\citet{PalkaAST11} this can be viewed as a way of adding admissible
but redundant typing rules, with the sole purpose of
using them for generation. For example, below is one such rule,
\textsc{G-ASTR}, which creates an initialized null-terminated string that
is statically cast into an array with bounds $(0,0)$.
\begin{mathpar}
    \inferrule[G-ASTR]
    {
      \textcolor{cyan}{i} \in \mathbb{N}^* \quad\quad
      \textcolor{cyan}{n_0,\ldots,n_{i-1} } \in \mathbb{Z} \quad\quad
      \texttt{fresh}(\textcolor{cyan}{x})\\
      \textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}}
      \textcolor{blue}{e'} :
      \tptr{\textcolor{red}{\tntarray{0}{\textcolor{cyan}{i}}{\tint}}}{c}
      \\      
\textcolor{blue}{e} = \elet{\textcolor{cyan}{x}}{\textcolor{blue}{e'}}{\texttt{(init $\textcolor{cyan}{x}$ with }\textcolor{cyan}{n_0,\ldots
        n_{i-1}}\texttt{)};\textcolor{cyan}{x}}
}
    {\textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\ecast{\tptr{\tntarray{0}{0}{\tint}}{c}}{e}} : \textcolor{red}{\tptr{\tntarray{0}{0}{\tint}}{c}}}
  \end{mathpar}
Given some positive number $\textcolor{cyan}{i}$, numbers
$\textcolor{cyan}{n_0, \ldots, n_{i-1}}$, and a fresh variable
$\textcolor{cyan}{x}$ (which are arbitrarily generated), we can
recursively generate a pointer $\textcolor{blue}{e'}$ with bounds
$(0,\textcolor{cyan}{i})$, and initialize it with the generated $n_j$
using $\textcolor{cyan}{x}$ to temporarily store the pointer.

This rule is particularly useful when combined with \textsc{G-IfNT}
since there is a much higher chance of obtaining a non-zero value when
evaluating $\estar{p}$ in the guard of $\eiftext$, skewing the
distribution towards programs that enter the $\ethentext$ branch.
Relying solely on the type-based rules, entering the $\ethentext$
branch requires \textsc{G-AssignArr} was chosen before \textsc{G-IfNT},
and that assignment would have to appear before $\eiftext$, which
means additional \textsc{G-Let} rules would need to be chosen: this
combination would therefore be essentially impossible to generate in
isolation.

\myparagraph{Generating Ill-typed Terms}
We can use generated well-typed terms to test our simulation theorem
(Section~\ref{sec:compilation}) and test that \lang and \checkedc Clang agree
on what is type-correct. But it is also useful to generate ill-typed
terms to test that \lang and \checkedc Clang agree on those.  However,
while it is easy to generate arbitrary ill-typed terms, they would be
very unlikely to trigger any inconsistencies; those are far more
likely to exist on the boundary between well- and
ill-typedness. Therefore, we also include generation rules modified to
be slightly more permissive, which results in sometimes generating
terms that are ``a little'' ill-typed.

\myparagraph{Random Testing for Language Design}
We used our Redex model and random generator to successfully guide the
design of our formal model, and indeed the Clang Checked C
implementation itself, which is being actively developed. To that end, we implemented a
conversion tool that converts \lang into a subset of the \checkedc
language and ensured that model and implementation exhibit the same
behavior (accept and reject the same programs and yield the same return
value).

This approach constitutes an interesting twist to traditional
model-based checking approaches.  Usually, one checks that the
implementation and model agree on all inputs \emph{of the
  implementation}, with the goal of covering as many behaviors as
possible. This is the case, for example, in \citet{lambdajs}, where
they use real test suites to demonstrate the faithfullness of their
core calculus to Javascript. Our approach and goal in this work is
essentially the opposite: as the Clang Checked C implementation does
not fully implement the Checked C spec, there is little hope of
covering all terms that are generated by Clang Checked C. Instead,
we're looking for \emph{inconsistencies}, which could be caused by
bugs either in the Clang Checked C compiler or our own model.

One inconsistency we found comes from the following:

{\small 
\begin{lstlisting}[xleftmargin=4 mm]
array_ptr<char> fun(void) : count(3) {
   array_ptr<char> x : count(3);
   x = calloc(3, sizeof(char));
   return x+3;
 }
 int main(void) {
   *(fun()) = 0;
   return 0;
 }
\end{lstlisting}
}\noindent
In this code, the function \code{fun} is supposed to return a
checked array pointer of size 3. Internally, it allocates such an
array, but instead of returning the pointer \code{x} to that array, it
increments that pointer by 3. Then, the \code{main} function just
calls \code{fun}, and tries to assign 0 to its result. Our model
correctly rules out this program, while the Clang Checked C
implementation happily accepted this out-of-bounds
assignment. Interestingly, it correctly rejected programs where the
array had size 1 or 2. This inconsistency has been fixed in the latest
version of the compiler.

We also found the opposite kind of inconsistency---programs that
the Clang Checked C implementation rejects contrary to the spec.
For instance:\footnote{After minimization, this turned out to be a known issue: 
  \url{https://github.com/microsoft/checkedc-clang/issues/1008}}

{\small 
\begin{lstlisting}[xleftmargin=4 mm]
array_ptr<int> f(void) : count(5) {
  array_ptr<int> x : count(5) =
    calloc<int>(5, sizeof(int)); 
  return x;
}
array_ptr<int> g(void ) : count(5) {
  array_ptr<int> x : count(5) =
    calloc<int>(5, sizeof(int)); 
  return x+3;
}
int main(void) {
  return *(0 ? g() : f() + 3);
}    
\end{lstlisting}
}\noindent
In this piece of code both \code{f} and \code{g} functions compute a
pointer to the same index in an array of size 5 (as \code{f} calls
\code{g}). The \code{main} function then creates a ternary expression
whose branches call \code{f} and \code{g}, but the Clang Checked C
implementation rejects this program, as its static analysis is not
sophisticated enough to detect that both branches have the same type.

 \section{Related Work}
\label{sec:related}

Our work is most closely related to prior formalizations of C(-like)
languages that aim to enforce memory safety, but it also touches on
C-language formalization in general.

\myparagraph{Formalizing C and Low-level code}
A number of prior works have looked at formalizing the semantics of C,
including CompCert~\cite{Blazy2009,leroy:hal-00703441},
\citet{ellison-rosu-2012-popl}, \citet{Kang:2015:FCM:2813885.2738005},
and \citet{10.1145/2980983.2908081, Memarian:2019:ECS:3302515.3290380}. These works also model
pointers as logically coupled with either the bounds of the blocks
they point to, or provenance information from which bounds can be
derived. None of these is directly concerned with enforcing
spatial safety, and that is reflected in the design. For example,
memory itself is not be represented as a flat address space, as in our
model or real machines, so memory corruption due to spatial safety
violations, which Checked C's type system aims to prevent, may not be
expressible. That said, these formalizations consider much more of the
C language than does \lang, since they are interested in the entire
language's behavior.

\myparagraph{Spatially Safe C Formalizations}
Several prior works formalize C-language transformations or C-language
dialects aiming to ensure spatial safety. 
\citet{10.1145/2813885.2737979} extends the formalization
of \citet{ellison-rosu-2012-popl} to produce a semantics that detects
violations of spatial safety (and other forms of undefinedness). It
uses a CompCert-style memory model, but ``fattens'' logical pointer
representations to facilitate adding side conditions similar to \lang's.
Its concern is bug finding, not compiling programs to
use this semantics.

CCured~\cite{Necula2005} and Softbound~\cite{softbound} implement
spatially safe semantics for normal C via program transformation. Like
\lang, both systems' operational semantics annotate pointers with
their bounds. CCured's equivalent of array pointers are compiled to be
``fat,'' while SoftBound compiles bounds metadata to a separate
hashtable, thus retaining binary compatibility at higher checking
cost. Checked C uses static type information to enable bounds checks
without need of pointer-attached metadata, as we show in
Section~\ref{sec:compilation}. Neither CCured nor Softbound models
null-terminated array pointers, whereas our semantics ensures that
such pointers respect the zero-termination invariant, leveraging
bounds widening to enhance expressiveness.

Cyclone \cite{Jim2002,GrossmanMJHWC02} is a C dialect that aims to
ensure memory safety; its pointer types are similar to
CCured. Cyclone's formalization~\cite{GrossmanMJHWC02} focuses on the
use of \emph{regions} to ensure temporal safety; it does not formalize
arrays or threats to spatial safety. Deputy~\cite{Feng2006,Condit2007}
is another safe-C dialect that aims to avoid fat pointers; it was an
initial inspiration for Checked C's design~\cite{Elliott2018}, though
it provides no specific modeling for null-terminated array
pointers. Deputy's formalization~\cite{Condit2007} defines its
semantics directly in terms of compilation, similar in style to what
we present in Section~\ref{sec:compilation}. Doing so tightly couples
typing, compilation, and semantics, which are treated independently in
\lang. Separating semantics from compilation isolates meaning 
from mechanism, easing understandability. Indeed, it was this
separation that led us to notice the 
limitation with Checked C's handling of bounds widening.

The most closely related work is the formalization of \checkedc done
by \citet{ruef18checkedc-incr}. They were the first to formalize and
prove \emph{blame} for a core model of \checkedc, which shows that any
spatial safety violation owes to invariants violated by unchecked
code. Our Coq-based development (Section~\ref{sec:formal})
substantially extends
theirs,\footnote{\url{https://github.com/plum-umd/checkedc/tree/master/coq}}
re-proving the blame theorem after adding dynamically bounded array
pointers with dependent types, null-terminated pointers, and
dependently typed functions. They postulate that pointer metadata can
be erased, but do not show it; indeed, we found it nontrivial once
null-terminated pointers were considered.

% \ignore{

% Other work that might be relavent but for guaranteeing temporal safety, I have recorded some of them in the related work if they are relavent to spatial safety, but I might missed some.

% Purify operates on binaries, but only ensures the safety of heap-allocated objects.

% Safe C ensures complete temporal safety by using fat-pointers.

% MSCC uses a fat-pointer approach for handling the spatial and temporal safeties regarding stack/heap pointers. According to the experiment in MSCC, it is supposed to be faster than CCured, but it seems that they are in the same level or CCured might be faster in some cases according to.

% Fail-Safe C is an updated version of Safe C, and it maintains complete compatibility with ANSI C but incurs significant runtime overhead.

% Cyclone is a new dialect designed to minimise the changes required to port over C programs. Cyclone’s region-based memory management system provides a foundation for solving the temporal safety issues for stack pointers in Checked-C. The only difference between the Cyclone’s region-based system and the Checked-C will be time when dangling pointers cause errors. Dangling pointers should be banned at the time when they are used to access memory. If a program contains a function call with a stack pointer that is created and returned to the upper level function call and it is never used to access memory, the program should be considered as valid.

% Here are several ideas of using metadata instead of fat-pointers, I (Liyi) think they are basically another form of fat-pointer approaches:

% Patil and Fischer address the spatial and temporal safeties by maintaining disjoint metadata and performing checks in a separate “shadow process,” but this requires an additional CPU.

% MemSafe provides a framework for checking spatial and temporal memory safety at runtime. 
% Their work is based on transferring an original C program to an updated CFG-based program with extra metadata. These metadata has similar functionality as fat pointer and there are extra
% checks allowing them to place extra checks during the program execution. For each pointer, the updated CFG stores information about all possible values of the pointer through possible aliasing. The efficiency of MemSafe can be as effective as CCured.

% Write integrity testing -- similar to regional memory management. The so-called colored memory space is essentially giving regional flags to different pointers, and then check if these pointers have potential security threats.

% The following approaches require hardware support:

% Clause et al. describe an efficient technique for detecting memory errors, but it requires custom hardware.

% Dhurjati and Adve describe a technique based on the Electric Fence [11] malloc debugger: Their system assigns a unique virtual page to every dynamically allocated object and relies on hardware page protection to detect dangling pointer dereferences.

% There are also few static approaches to the temporal safety in C. However, they are not complete and a lot of them are not even sound. There are four kinds of static approach methods: model checking, abstract interpretation, pattern matching, and pointer analysis.

% Model Checking. CBMC is a bounded model checker that reasons about all the program paths for C/C++ programs as constraints that can be solved by an SMT solver. When used in finding UAF bugs, CBMC is sound (in a bounded manner) and highly precise but scales only to small programs whose “sizes are restricted” (according to its user manual).

% Thomas Ball and Sriram K. Rajamani developed a model checking framework (C2bp) for checking Temporal Safety Properties in C. The framework allows pointer analysis by using a pointer aliasing algorithm. The problem is that it only deals with small programs, and the pointer aliasing algorithm only detects a small part of aliasing pointers.

% MOPS (MOdel checking Program for Security Properties) makes use of the model checking technique to check for violation of security rules, that are defined as temporal safety properties. As for program representation, MOPS models the program in the form of Push Down Automaton (PDA), that contains all the possible execution paths. Push Down Automata are used as tools to analyze procedural sequential programs, and more specifically those having recursive procedures. As for automata, they are according to Schneider used in the objective of specifying security policies that can be enforced by mechanisms. MOPS makes use of this approach to model security properties in the form of Finite State Automata (FSA), that dictate the order of security-relevant operations sequence. The modularity of security properties was also proposed by MOPS; this approach allows the decomposition of complex security properties into simpler and reusable basic security properties that are easy to model and to extend (such as role based access). MOPS verifies that the security properties are properly respected in all the execution paths of the analyzed program, making use of the model checking on the PDA, and checks if risky states are reachable within the PDA. MOPS does not know that a pointer is an alias of another, so the property is not sound for this program.

% Abstract Interpretation. Clang Static Analyzer is an abstract interpreter for analyzing C/C++ programs. It adopts a highly unsound model by analyzing only a small subset of the program paths in order to achieve scalability and precision. To scale for large codebases with few false alarms, Clang limits its UAF-bug-finding ability by performing an intraprocedural analysis (with inlining). In general, such tools refrain from reporting too many false alarms, but at the expense of missing many UAF bugs.

% Pattern Matching. Coccinelle is a pattern-based tool for analyzing and certifying C programs. Coccinelle can find UAF bugs based on some patterns given. Due to the lack of the points-to information, Coccinelle can be both fairly unsound and imprecise but is highly scalable (due to its pattern-matching nature).

% Pointer Analysis. Supa is a state-of-the-art demand-driven pointer analysis that is field-, flow- and context-sensitive but path- insensitive for C programs. When used in finding UAF bugs, Supa can be regarded as reasoning about all the program paths with an extremely coarse abstraction, in order to achieve soundness and scalability.

% Yan et al. develops a Spatio-Temporal context reduction technique for  doing pointer-analysis for detecting use-after-free vulnerabilities, which are essentially temporal safety. They mainly focus on the heap pointer allocation-free detection to  check if a pointer is used after the free function call for the memory field the pointer points to. The technique is based on static analysis. The main idea is still based on intraprocedural control flow analysis. It relies on context sensitive reduction technique to reduce the search function calling path spaces for analyzing pointer aliasing.  The context sensitivity and their pointer analysis algorithm is applied in a state-of-the-art manner. The information for the algorithm of search pointer aliasing reduces the calling context path, while the context path information enhances the efficiency of the algorithm.  However, the technique is not complete. It will ban a large number of valid C programs due to false use-after-free vulnerabilities.

% It seems that they might not detect the use-after-free error in this case. In addition, the work does not handle stack pointer use-after-free vulnerabilities.

% ASAP: As Static As Possible memory management -- use data flow analysis, garbage collection and region memory management tools to ensure the security safety of C programs by static analysis. In order to guarantee the use-after-free, it requires the program written in a special way: in a linearly typed program, each value must be used exactly once. It does not include examples of dealing with stack pointers. It shows that in a pointer management world, a C program temporal safety cannot be fully analyzed by only static analysis.

% Jia-Ju Bai et al. develops a static method to detect use-after-free bugs in Linux device drivers. It is a UAF bug in the concurrent system. The basic algorithm is simple. It first goes through each single-threaded code to collect a read/write set storing point-to information. The information collection is through traditional intraprocedural CFG analysis. Then, it uses a trace analysis tool to discover UAF bugs. This is achievable only in dealing with linux device drivers, since usually, these drivers are simple and the concurrency is very formulated.  I believe if we apply the same process in dealing with C, the method will not be sound.

% Industrial tools:

% There are also some industrial level tools for static checking bugs in C. However, they are either not able to check UAF bugs or they only use heuristics to check very small sets of UAF bugs. They include CppCheck, Clang Core, Clang Alpha, CodeSonar, Facebook Infer,  Splint standard, Splint weak. For example, CodeSonar state-based checkers do not check temporal safety. They require insertion of extra code in order for the tool to check if a supposed deadcode will be reached or not. Coverity is a dataflow analysis tool that relies on inter-procedural analysis techniques. The analysis is neither sound nor complete, that is, there may be both defects which are not reported and there may be false alarms. Coverity Prevent implements an incremental analysis which means that the system automatically infers what parts of the source code that have to be re-analyzed after the code has been modified. This typically reduces the analysis time substantially, but may of course imply a complete re-analysis in the worst case. Klocwork does similar things to Coverity and Coverity Prevent.
% }

\section{Conclusion and Future Work}
\label{sec:conclude}

This paper presented \CoreChkC, a formalization of an extended core of
the Checked C language which aims to provide spatial memory
safety. \lang models dynamically sized and
null-terminated arrays with dependently typed bounds that can
additionally be widened at runtime. We prove, in Coq, the key safety
property of Checked C for our formalization, {\em blame}: if a mix of
checked and unchecked code gives rise to a spatial memory safety
violation, then this violation originated in an unchecked part of the
code. We also show how programs written in \CoreChkC (whose
semantics leverage fat pointers) can be compiled to \elang (which does
not) while preserving their behavior. Finally, we developed a random
testing framework to guide the design of our formal model by comparing
it against the \checkedc compiler, finding multiple inconsistencies
in the process.

As future work, we are interested in designing an a way to
automatically port legacy C code to \checkedc.  We also want to
further extend our \CoreChkC model to include more C behaviors,
such as function pointers, with our testing framework guiding the
design process.

%  \ignore{
% \begin{acks}                            This material is based upon work supported by the
%   \grantsponsor{GS100000001}{National Science
%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%   conclusions or recommendations expressed in this material are those
%   of the author and do not necessarily reflect the views of the
%   National Science Foundation.
% \end{acks}
% }

\begin{small}
\bibliography{IEEEabrv,paper,sources}
\end{small}
\newpage
\appendix
\section{Appendix}

\subsection{Typing Rules for Literal Pointers}\label{sec:literal-pointer-typing}

One thing we elided from the main presentation is the typing of
integer literals (which can also be pointers to the heap). These rules
are shown in Fig.~\ref{fig:const-type}. The variable type rule
($\textsc{T-Var}$) simply checks if a given variable has the defined
type in $\Gamma$; the constant rule ($\textsc{T-Const}$) is slightly
more involved.  First, it ensures that the type annotation $\tau$ does
not contain any free variables. More importantly, it ensures that the
literal itself is well typed using an auxilliary typing relation
$\sigma \vdash n : \tau$, which is implicitly indexed by a given heap
$\heap$.

If the literal's type is an integer, an unchecked pointer, or a null
pointer, it is well typed, as shown by the top three rules in
Fig.~\ref{fig:const-type}. However, if it is a checked pointer
$\tptr{\omega}{\cmode}$, we need to ensure that what it points to in
the heap is of the appropriate pointed-to type ($\omega$), and also
recursively ensure that any literal pointers reachable this way are
also well-typed. This is captured by the bottom rule in the figure,
which states that for every location $n+i$ in the pointers' range
$[n, n+\size(\omega))$, where $\size$ yields the size of its argument,
  then the value at the location $\heap(n+i)$ is also well-typed.
  However, as heap snapshots can contain cyclic structures (which
  would lead to infinite typing deriviations), we use a scope $\sigma$
  to assume that the original pointer is well-typed when checking the
  types of what it points to. The middle rule then accesses the scope
  to tie the knot and keep the derivation finite, just like in
  \citet{ruef18checkedc-incr}.

 \begin{figure}[h]
 {\small
 \text{Type Rules for Constants and Variables:}
 \begin{mathpar}
   \inferrule[T-Var]
       {x : \tau \in \Gamma}
       {\Gamma;\Theta \vdash_m x : \tau}

   \inferrule[T-Const]
       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
 \end{mathpar}

 \text{Rules for Checking Constant Pointers In Heap:}
 \begin{mathpar}
   \inferrule
       {}
       {\sigma \vdash n : \tint}

   \inferrule
       {}
       {\sigma \vdash n : \tptr{\omega}{\umode}}

   \inferrule
       {}
       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}

   \inferrule
       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
       {\sigma \vdash n : \tptr{\omega}{\cmode}}

   \inferrule
       {\forall i \in [0,\size(\omega)) .
            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
       {\sigma \vdash n : \tptr{\omega}{\cmode}}
 \end{mathpar}
 }
 \caption{Type Rules for Checking Constants/Variables}
 \label{fig:const-type}
 \end{figure}

\subsection{Other Semantic Rules}\label{sec:rem-semantics}

Fig.~\ref{fig:rem-semantics} shows the remaining semantic rules for
$\lang$. We explain a selected few rules in this subsection.

\begin{figure*}[t]
{\small
\begin{mathpar}

\inferrule [S-Var]{} {(\varphi,\heap,x)\longrightarrow (\varphi,\heap,\varphi(x))}

    \inferrule[S-DefArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}

    \inferrule[S-DefArrayBound]{0 \not\in [n_l,n_h)}
     { (\varphi,\heap,\estar{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}

    \inferrule[S-DefNTArrayBound]{0 \notin [n_l,n_h]}
    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\ebounds)}

    \inferrule[S-Assign]{\heap(n)=\evalue{n_a}{\tau_a} }
      {(\varphi,\heap,\eassign{\evalue{n}{\tptr{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}

    \inferrule[S-AssignNull]{}
      {(\varphi,\heap,\eassign{\evalue{0}{\tptr{\omega}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\enull)}

    \inferrule[S-AssignArrBound]{0 \not\in [n_l,n_h) }
      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\ebounds)}

  \inferrule[S-Malloc]{\varphi(\omega)=\omega_a \\ \mathtt{alloc}(\heap,\omega_a)=(n,\heap')}
   { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\evalue{n}{\tptr{\omega_a}{\cmode}})}

  \inferrule[S-MallocBound]{\varphi(\omega)=\tallarray{n_l}{n_h}{\tau}\\ (n_l \neq 0 \vee n_h \le 0)}
    { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\ebounds)}

    \inferrule[S-IfT]{n \neq 0 }
    {(\varphi,\heap,\eif{\evalue{n}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}

    \inferrule[S-IfF]{}
    {(\varphi,\heap,\eif{\evalue{0}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}

    \inferrule[S-Unchecked]{}
    {(\varphi,\heap,\eunchecked{\evalue{n}{\tau}} \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}

    \inferrule[S-Str]{
        0 \in [n_l,n_h]
\\ \heap(n+n_a) = 0 
\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tint})}

    \inferrule[S-StrBounds]{
        0 \notin [n_l,n_h]
}
    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}

    \inferrule[S-StrNull]{}
    {(\varphi,\heap,\estrlen{\evalue{0}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\enull)}

    \inferrule[S-Add]{n = n_1 + n_2}
    {(\varphi,\heap,\evalue{n_1}{\tint} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, n)}

    \inferrule[S-AddArr]{n = n_1 + n_2\\ n_l' = n_l - n_2 \\ n_h' = n_h - n_2}
    {(\varphi,\heap,\evalue{n_1}{\tallarrayptr{n_l}{n_h}{\tau}{m}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau}{m}})}

n    \inferrule[S-AddArrNull]{}
    {(\varphi,\heap,\evalue{0}{\tallarrayptr{n_l}{n_h}{\tau}{c}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \enull)}

\end{mathpar}

}
\caption{Remaining \lang Semantics Rules (extends Fig.~\ref{fig:semantics})}
\label{fig:rem-semantics}
\end{figure*}

Rule \textsc{S-Var} loads the value for $x$ in stack $\varphi$.
Rule \textsc{S-DefArray} dereferences an array pointer, which is similar to the Rule \textsc{S-DefArray} in Fig.~\ref{fig:semantics}.
The only difference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
Rules \textsc{DefArrayBound} and \textsc{DefNTArrayBound} describe an error case for a dereference operation.
If we are dereferencing an array/NT-array pointer and the mode is $\cmode$, $0$ must be in the range from $n_l$ to $n_h$; if not, the system results in a $\ebounds$ error. Obviously, the dereference of an array/NT-array pointer also experiences a $\enull$ state transition if $n\le 0$.

Rules \textsc{S-Malloc} and \textsc{S-MallocBound} describe the $\emalloctext$ semantics. Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$, and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation. The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\cmode}}$ and new updated heap. It is possible for $\emalloctext$ to transition to a $\ebounds$ error if the $\omega_a$ is an array/NT-array type $\tallarray{n_l}{n_h}{\tau}$, and either $n_l \neq 0$ or $n_h \le 0$.

\subsection{Subtyping for dependent types}
\label{app:le}

The subtyping relation given in Fig.~\ref{fig:checkc-subtype} involves
dependent bounds, i.e., bounds that may refer to variables. To decide
premises $b \leq b'$, we need a decision procedure that accounts for
the possible values of these variables. This process considers
$\Theta$, tracked by the typing judgment, and $\varphi$, the current
stack snapshot (when performing subtyping as part of the type
preservation proof). We

\begin{defi}[Inequality]

\begin{itemize}

\item $n \le m$ if $n$ is less than or equal to $m$.
\item $x+n \le x + m$ if $n$ is less than or equal to $m$.
\item All other cases result in $\efalse$.

\end{itemize}
\end{defi}

To capture bound variables in dependent types, the \checkedc subtyping
relation ($\sqsubseteq$) is parameterized by a restricted stack
snapshot $\varphi|_{\rho}$ and the predicate map $\Theta$, where
$\varphi$ is a stack and $\rho$ is a set of
variables. $\varphi|_{\rho}$ means to restrict the domain of $\varphi$
to the variable set $\rho$. Clearly, we have the relation:
$\varphi|_{\rho} \subseteq \varphi$. The meaning of $\sqsubseteq$
being parameterized by $\varphi|_{\rho}$ refers to that when we
compare two bounds $b \le b'$, we actually do
$\varphi|_{\rho}(b) \le \varphi|_{\rho}(b')$ by interpreting the
variables in $b$ and $b'$ with possible values in $\varphi|_{\rho}$.
Let's define a subset relation $\preceq$ for two restricted stack
snapshot $\varphi|_{\rho}$ and $\varphi'|_{\rho}$:

\begin{defi}[Subset of Stack Snapshots]
  Given two $\varphi|_{\rho}$ and $\varphi'|_{\rho}$,
  $\varphi|_{\rho} \preceq \varphi'|_{\rho}$, iff for $x\in\rho$ and
  $y$,
  $(x,y) \in \varphi|_{\rho} \Rightarrow (x,y) \in \varphi'|_{\rho}$.
\end{defi}

For every two restricted stack snapshots $\varphi|_{\rho}$ and
$\varphi'|_{\rho}$, such that
$\varphi|_{\rho} \preceq \varphi'|_{\rho}$, we have the following
theorem in \checkedc (proved in Coq):

\begin{thm}[Stack Snapshot Theorem]
  Given two types $\tau$ and $\tau'$, two restricted stack snapshots
  $\varphi|_{\rho}$ and $\varphi'|_{\rho}$, if
  $\varphi|_{\rho}\preceq \varphi'|_{\rho}$, and
  $\tau \sqsubseteq \tau'$ under the parameterization of
  $\varphi|_{\rho}$, then $\tau \sqsubseteq \tau'$ under the
  parameterization of $\varphi'|_{\rho}$.
\end{thm}

Clearly, for every $\varphi|_{\rho}$, we have
$\emptyset \preceq \varphi|_{\rho}$. The type checking stage is a
compile-time process, so $\varphi|_{\rho}$
is $\emptyset$ at the type checking stage. Stack snapshots are needed
for proving type preserving, as variables in bounds expressions are
evaluated away.

As mentioned in the main text, $\sqsubseteq$ is also parameterized by
$\Theta$, which provides the range of allowed values for a bound
variable; thus, more $\sqsubseteq$ relation is provable. For example,
in Fig.~\ref{fig:strncat-ex}, the \code{strlen} operation in line 4
turns the type of \code{a} to be $\tntarrayptr{0}{x}{\tint}{\cmode}$
and extends the upper bound to \code{x}. In the \code{strlen} type
rule, it also inserts a predicate \code{x}$\ge 0$ in $\Theta$; thus,
the cast operation in line 16 is valid because
$\tntarrayptr{0}{x}{\tint}{\cmode} \sqsubseteq
\tntarrayptr{0}{0}{\tint}{\cmode}$ is provable when we know
\code{n}$\ge 0$.

Note that if $\varphi$ and $\Theta$ are $\emptyset$, we do only the
syntactic $\le$ comparison; otherwise, we apply $\varphi$ to both
sides of $\sqsubseteq$, and then determine the $\le$ comparasion based
on a Boolean predicate decision procedure on top of $\Theta$. This
process allows us to type check both an input expression and the
intermediate expression after evaluating an expression.

\subsection{Other Type Rules}\label{rem-type}

Here we show the type rules for other \checkedc operations in Fig.~\ref{fig:rem-type-system}.

\begin{figure}[t]
{\small
  \begin{mathpar}
    \inferrule[T-Def]
              {\Gamma;\Theta \vdash_m e : \tptr{\tau}{m'} \\
                m \leq m'}
              {\Gamma;\Theta \vdash_m \estar{e} : \tau}

    \inferrule[T-Mac]
              {}
              {\Gamma; \Theta \vdash_m \emalloc{\omega} : \tptr{\omega}{\cmode}}

    \inferrule[T-Add]
              {\Gamma; \Theta \vdash_m e_1 : \tint \\
                \Gamma; \Theta \vdash_m e_2 : \tint}
              {\Gamma; \Theta \vdash_m (e_1 \plus e_2) : \tint }

    \inferrule[T-Ind] 
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'} \\
                \Gamma; \Theta \vdash_m e_2 : \tint \\
                m \leq m'}              
              {\Gamma; \Theta \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}

    \inferrule[T-Assign]
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tau}{m'} \\
                \Gamma; \Theta \vdash_m e_2 : \tau' \\
                \tau'\sqsubseteq \tau \\
                m \leq m'}
              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}

   \inferrule[T-IndAssign]
              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
                \Gamma; \Theta \vdash_m e_2 : \tint \\
                \Gamma; \Theta \vdash_m e_3 : \tau' \\
                \tau'\sqsubseteq \tau \\
                m \leq m'}
              {\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}

    \inferrule[T-Unchecked]
              {\Gamma;\Theta \vdash_u e : \tau}
              {\Gamma;\Theta \vdash_m \eunchecked{e} : \tau}
  \end{mathpar}
}
\caption{Remaining \lang Type Rules (extends Fig.~\ref{fig:type-system-1})}
\label{fig:rem-type-system}
\end{figure}

Rule \textsc{T-Def} is for dereferencing a non-array pointer. 
The statement $m \leq m'$ relates the unchecked region for a term with
its sub-terms. We require that if the sub-term has an unchecked
region, so does the whole term. Rule \textsc{T-Mac} deals with
$\emalloctext$ operations. There is a well-formedness check to require
that the possible bound variables in $\omega$ must be in the domain of
$\Gamma$ (see Fig.~\ref{fig:wftypesandbounds}). Rule \textsc{T-Add} deals with binary operations whose sub-terms are integer expressions, while rule \textsc{T-Ind} serves the case for pointer arithmetic. For simplicity, in the \checkedc formalization, we do not allow arbitrary pointer arithmetic. The only pointer arithmetic operations allowed are the forms shown in rules \textsc{T-Ind} and \textsc{T-IndAssign} in Fig.~\ref{fig:rem-type-system}. Rule \textsc{T-Assign} is for assigning a value to a non-array pointer location. The predicate $\tau'\sqsubseteq \tau$ requires that the value being assigned is a subtype of the pointer type.
The \textsc{T-IndAssign} rule is an extended assignment operation for handling assignments for array/NT-array pointers with pointer arithmetic. Rule \textsc{T-Unchecked} type checks \code{unchecked} blocks.

\subsection{Struct Pointers}\label{appx:struct}

\checkedc has \kw{struct} types and \kw{struct} pointers. Fig.~\ref{fig:checkc-struct} contains the syntax of \kw{struct} types as well as new subtyping relations built on the \kw{struct} values.
For a \kw{struct} typed value, \checkedc has a special operation for it, which is $\eamper{e}{f}$. This operation indexes the $f$-th position \kw{struct}~$T$ item, if the expression $e$ is evaluated to a \kw{struct} pointer $\tptr{\tstruct{T}}{m}$. Rule \textsc{T-Struct} in Fig.~\ref{fig:checkc-struct} describes its typing behavior.
Rules \textsc{S-StructChecked} and \textsc{S-StructUnChecked} describe the semantic behaviors of $\eamper{e}{f}$ on a given \kw{struct} \code{checked}/\code{unchecked} pointers, while rule \textsc{S-StructNull} describes a \code{checked} \kw{struct} null-pointer case.
In our Coq/Redex formalization, we include the \kw{struct} values and the operation $\eamper{e}{f}$. We omit it in the main text due to the paper length limitation.

\begin{figure}
{\small
$\begin{array}{l}
\text{  Struct Syntax: }\\[0.5em]
  \begin{array}{ll}
 \mathtt{Type} & \tstruct{T}
\\[0.2em]
     \text{Structdefs} & D \; \in \; T \rightharpoonup fs \\[0.2em]
      \text{Fields} & fs \; ::= \; \tau~\mathtt{f} \mid \tau~\mathtt{f}; fs 
    \end{array}\\[2em]
\text{  Struct Subtype: }\\[0.5em]
\begin{array}{l}
    D(T) = fs \wedge fs(0) = \tnat \Rightarrow  \tptr{\tstruct{T}}{m} \sqsubseteq \tptr{\tnat}{m}\\[0.5em]
    D(T) = fs \wedge fs(0) = \tnat \wedge 0 \le b_l \wedge b_h \le 1 \\[0.2em]
 \qquad\qquad \Rightarrow 
       \tptr{\tstruct{T}}{m} \sqsubseteq \tarrayptr{b_l}{b_h}{\tnat}{m}
    \end{array}
\\[3em]
\text{  Struct Type Rule: }\\[0.5em]
    \end{array}
$

{
\begin{mathpar}
  \inferrule [T-Struct]
  {\Gamma; \Theta \vdash_m e :  {\tptr{\tstruct{T}}{m}} \\
    D(T) = fs \\ fs(f)=\tau_f}
  {\Gamma; \Theta \vdash_m \eamper{e}{f} : \tptr{\tau_f}{m}}

\end{mathpar}
}
{
$\begin{array}{l}
\text{Struct Semantics: }
\end{array}
$
}
{
\begin{mathpar}
  \inferrule [S-StructChecked]
  {n > 0 \\ D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\cmode}})}

  \inferrule [S-StructNull]
  {n = 0}
  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\enull)}

  \inferrule [S-StructUnChecked]
  {D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\umode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\umode}})}

\end{mathpar}
}
}
  \caption{\lang Struct Definitions}
  \label{fig:checkc-struct}
\end{figure}

\begin{figure}[t]
{\small
  \begin{mathpar}

    \inferrule[]
    {}
    {\Gamma \vdash n}

    \inferrule[]
    {x:\tint \in \Gamma}
    {\Gamma \vdash x + n}

    \inferrule[]
    {\Gamma \vdash b_l\\
    \Gamma \vdash b_h}
  {\Gamma \vdash (b_l,b_h)}

  \inferrule[]
  {}
  {\Gamma \vdash \tint}

  \inferrule[]
  {\Gamma \vdash \bvar \\
  \Gamma \vdash \tau}
  {\Gamma \vdash \tptr{\tallarrayb{\bvar}{\tau}}{m}}

  \inferrule[]
  {\Gamma \vdash \tau}
  {\Gamma \vdash \tptr{\tau}{m}}

  \inferrule[]
  {T \in D}
  {\Gamma \vdash \tptr{\tstruct{T}}{m}}
  \end{mathpar}
}
 \caption{Well-formedness for types and bounds}
\label{fig:wftypesandbounds}
\end{figure}

\begin{figure}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    {\Gamma \vdash \overline{x}:\overline{\tau} \\
      \Gamma[\overline{x} \mapsto \overline{\tau}] \vdash \tau \\
    \Gamma[\overline{x} \mapsto \overline{\tau}]; \Theta  \vdash_{\cmode} e : \tau}
    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e}

    \inferrule[]
    {}
    {\Gamma \vdash \cdot}

    \inferrule[]
    {\Gamma \vdash \tau \\
    \Gamma[x \mapsto \tau] \vdash \overline{x}:\overline{\tau}}
    {\Gamma \vdash x:\tau, \overline{x}:\overline{\tau}}
  \end{mathpar}
}
 \caption{Well-formedness for functions}
\label{fig:wffunctions}
\end{figure}

\begin{figure}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    {\Gamma \vdash \tau}
    {\Gamma \vdash \tau~\mathtt{f} }

    \inferrule[]
    {\Gamma \vdash \tau\\
     \Gamma \vdash fs}
    {\Gamma \vdash \tau~\mathtt{f}; fs }

  \end{mathpar}
}
 \caption{Well-formedness for structs}
\label{fig:wfstructs}

{\small
  \begin{mathpar}

    \inferrule[]
    {\Gamma[\overline{x} \mapsto \overline{\tau}]; \emptyset  \vdash
      e \gg \dot e : \tau}
    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e \gg
      (\overline{x})\;\dot e}
  \end{mathpar}
}
 \caption{Compilation rules for functions}
\label{fig:compilefunctions}
\end{figure}

\subsection{The Compilation Rules}\label{appx:comp1}
Fig.~\ref{fig:syntaxerased} and Fig.~\ref{fig:semanticserased} shows
the syntax for $\elang$, the target language for compilation. We syntactically
restrict the expressions to be in A-normal form because that is the
type of expression our compiler produces. To allow explicit runtime
checks, we include $\ebounds$ and $\enull$ as part of $\elang$
expressions which, once evaluated, result in an corresponding error state. $\eassignstack{x}{\dot a}$ is a new syntactic form that
modifies the stack variable $x$ with the result of $\dot a$. It is
essential for bounds widening. $\leq$ and $-$ are introduced
to operate on bounds and decide whether we need to halt with a bounds error
or widen a null-terminated string.

\begin{figure}[h]
{\small
  \[\begin{array}{llcl}
      \text{Atoms}       & \dot a & ::= & n \mid x \\

      \text{C-Expressions} & \dot c & ::= & \dot a \mid \estrlen{\dot a}  \mid \emalloc {\dot  a}  \mid \ecall{f} {\overline{\dot a}}  \\
                         & & \mid &  \mid  \dot a \circ \dot a \mid \estar{\dot a} \\
                         & & \mid & \eassign{\dot a}{\dot a} \mid \eassignstack{x}{\dot a}  \mid \eif{\dot a}{\dot e}{\dot e} \\
                         & & \mid & \ebounds \mid \enull \\
      \text{Expressions}  & \dot e & ::= & \dot c \mid \elet{x}{\dot c}{\dot e}\\
      \text{Binops} & \circ & ::= & + \mid - \mid \leq  \\
      \text{Closure} & \dot C & ::= & \hole \mid \elet{x}{\dot a}{\dot C} \\
       & & \mid & \eif{\dot a}{\dot e}{\dot C} \mid \eif{\dot a}{\dot    C}{\dot e} \\
      \text{Bounds Map} & \rho & \in & \texttt{Var} \rightharpoonup \texttt{Var} \times \texttt{Var}
  \end{array}
  \]
}
  \caption{\elang{} Syntax}
  \label{fig:syntaxerased}
{\small
  \[\begin{array}{lll}
    \dot \mu & ::= & n \mid \bot \\
    \dot c & ::= & \ldots  \mid \eret{x}{\dot \mu}{\dot e} \\
    \dot H & \in & \mathbb{Z} \rightharpoonup \mathbb{Z} \\
    \dot r & ::= & \dot e \mid \enull \mid \ebounds \\
    \dot E & ::= & \hole \mid  \elet{x}{\dot E}{\dot e}  \mid \eret{x}{i}{\dot E} \\
             & \mid & \eif{\dot E}{\dot e}{\dot e} \mid \estrlen{\dot E} \\
             & \mid & \emalloc{\dot E} \mid \ecall{f}{\overline{\dot
                      E}} \mid \dot E \circ \dot a \mid n \circ \dot E \\
             & \mid & \estar{\dot E} \mid \eassign{\dot E}{\dot a}
                      \mid \eassign{n}{\dot E} \mid
                      \eassignstack{x}{\dot E} \\
      \overline{\dot E } & ::= & \dot E \mid n,\overline{\dot E} \mid
                                 \overline{\dot E}, \dot a

  \end{array} \]
}
  \caption{\elang Semantic Defs}
  \label{fig:semanticserased}
\end{figure}

$\elang$ does not include any annotations. We remove structs from
$\elang$ because we can always statically convert expressions of the form
$\eamper{n:\tau}{f}$ into $n + n_f$, where $n_f$ is the statically
determined offset of $f$ within the struct. We ellide the semantics of
$\elang$ because it is self-evident and mirrors the semantics
$\lang$. The difference is that in $\elang$, only $\ebounds$ and
$\enull$ can step into an error state. All failed dereferences and
assignments would result in a stuck state and therefore we rely on the
compiler to explicitly insert checks for checked pointers.

Fig.~\ref{fig:compilation} and Fig.~\ref{fig:compilation2}
 shows the rules for the compilation judgment for expressions,
\[\Gamma;\rho \vdash e \gg \dot C, \dot a\]
The judgment is presented differently from the one in
Sec.~\ref{sec:compilation}, which was simplified for presentation purposes. First,
we remove $\Theta$ and $m$ because these
parameters are only used for checking and have no
impact on compilation. Second, the judgment includes two
outputs, a closure $\dot C$ and an atom expression $\dot a$, instead of
a single $\elang$ expression $\dot e$. $\dot C$ can be intuitively understood as a
partially constructed program or context. Whereas $\dot E$ is used for
evaluation, $\dot C$
is used purely as a device for compilation. As an example,
when compiling $(1 : \tint) \plus (2 : \tint)$, 
we would first create a fresh variable $x$, and then produce two outputs:
\[    \dot C = \elet{x}{1 \plus 2}{\hole}\]
\[    \dot a = x\]
To obtain the compiled expression $\dot e$, we plug $\dot a$ into
$\dot C$ using the usual notation $\dot C[\dot a]$. We can also use $\dot C$ to represent runtime checks, which
usually take the form $\elet{x}{\dot c}{\hole}$, where $\dot c$
contains the check whose evaluation must not trigger $\ebounds$ or $\enull$ for the program to
continue (see Fig.~\ref{fig:compilationhelpers} for the metafunctions that create
those checks).

This unconventional output format enables us to separate the
evaluation of the term and the computation that relies on the term's
evaluated result. Since effects and reduction (except for variables) happen only
within closures, we can precisely control the order in which effects
and evaluation happen by composing the contexts in a specific order.
Given two closures $\dot C_1$
and $\dot C_2$, we write $\dot C_1[\dot C_2]$ to denote the meta
operation of plugging $\dot C_2$ into $\dot C_1$. We also use
$\dot C_{a;b;c}$ as a shorthand for $\dot C_a[\dot C_b[\dot C_c]]$. In
the \textsc{C-Ind} rule,
we first evaluate the expressions that correspond to $e_1$ and $e_2$
through $\dot C_1$ and $\dot C_2$, and then perform a null check and
an addition through $\dot C_n$ and $\dot C_3$. Finally, we dereference
the result through $\dot C_4$ before returning the pair $\dot C_4,\dot
x_4$, propagating the flexibility to the compilation rule that recursively
calls \textsc{C-Ind}.

Fig.~\ref{fig:compilationhelpers} shows the metafunctions that create
closures representing dynamic checks. These functions first examine
whether the pointer is a checked. If the pointer is unchecked, an empty closure $\hole$
will be returned, because there is no need to perform a check. For bounds
checking, there is a special case for NT-array pointers, where the
bounds are retrived from the ghost variables (found by
looking up $\rho$) on the stack  rather than
using the bounds specified in the type annotation. This is how we
achieve the same precise runtime behavior as $\lang$ in our compiled expressions.

Fig.~\ref{fig:compilationhelpers2} shows the metafunctions related to
bounds widening. $\vdash_{extend}$ takes $\rho$,
a checked NT-array pointer variable $x$, and its bounds $(b_l,b_h)$ as inputs, and returns an
extended $\rho'$ that maps $x$ to two fresh variables $x_l$, $x_h$,
together with a closure $\dot C$ that initializes $x_l$ and $x_h$ to
$b_l$ and $b_h$ respectively. This function is used in the
\textsc{C-Let} rule to extend $\rho$ before compiling the body of the
$\elettext$ binding. The updated $\rho'$ can be used for generating
precise bounds checks, and for inserting expressions that can potentially
widen the upper bounds, as seen in the $\vdash_{widenstr}$
metafunction used in the \textsc{C-Str} compilation rule.

\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    { x_l, x_h = \fresh \\
      \rho' = \rho[x \mapsto (x_l, x_h)]\\
    \dot C = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}}
  {\cextend{\rho}{x}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot C, \rho'}\\}

  \inferrule[]
  {x_l,x_h = \rho(x) \\
    x_w = \fresh \\
  \dot C = \elet{x_w}{\eif{x_h}{0}{\eassignstack{x_h}{1}}}{\hole}}
{\ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C}}

\inferrule[]
{e \notin dom(\rho)}
{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{m}}{\hole}}

\inferrule[]
{x_l,x_h = \rho(e) \\
  x_a = \fresh \\
  \dot C = \elet{x_a}{\eif{\dot a \leq x_h}{0}{\eassignstack{x_h}{\dot a}}}{\hole}}
{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{\cmode}}{\dot C}}

  \end{mathpar}

}
 \caption{Metafunctions for widening}
\label{fig:compilationhelpers2}
\end{figure*}

\begin{figure*}[h]
{\small
  \begin{mathpar}
    \inferrule[]
    {x = \fresh \\
      \dot C = \elet{x}{\eif{\dot a}{0}{\enull}}{\hole}}
    {\echecknull{\dot a}{c}{\dot C}}

    \inferrule[]
    {}
    {\echecknull{\dot a}{u}{\hole}} \\

    \inferrule[]
    {}
    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}

    \inferrule[]
    { x_l,x_h = \rho(e) \\
      x_{cl},x_{ch} = \fresh\\
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

    \inferrule[]
    {}
    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}

    \inferrule[]
    { x_l,x_h = \rho(e) \\
      x_{cl},x_{ch} = \fresh\\
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
    }
    {\echeckboundsw{\rho}{e}{\tallarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

\inferrule[]
{e \notin dom(\rho) \\
  x_l,x_l',x_h,x_h' = \fresh \\
  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
  \dot C_2 = \elet{x_l'}{b_l'}{\elet{x_h'}{b_h'}{\hole}}\\
  \dot C_3 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
  \dot C_4 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
}
{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3;4}}}

\inferrule[]
{x_l',x_h' =\rho(e) \\
  x_l,x_h = \fresh \\
  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
  \dot C_2 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
  \dot C_3 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
}
{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3}}}
  \end{mathpar}
}
\caption{Metafunctions for dynamic checks}
\label{fig:compilationhelpers}
\end{figure*}

\begin{figure*}[t]
{\small
  \begin{mathpar}

    \inferrule[C-Const]
      {}
      {\Gamma;\rho\vdash \evalue{n}{\tau} \gg \hole, n : \tau}

    \inferrule[C-Var]
      {x : \tau \in \Gamma}
      {\Gamma;\rho \vdash x \gg \hole, x : \tau}

    \inferrule[C-Cast]
              {\Gamma;\rho \vdash e \gg  \dot C , \dot a : \tau'}
              {\Gamma;\rho \vdash \ecast{\tau}{e} \gg \dot C, \dot a  : \tau}

      \inferrule[C-DynCast]
      {\Gamma;\rho \vdash e \gg \dot C_1, \dot a : \tptr{\tallarrayb{\bvar'}{\tau}}{m} \\
      \echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\tptr{\tallarrayb{\bvar'}{\tau}}{m}}{\dot C_{b}}}
      {\Gamma;\rho \vdash \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e}
            \gg \dot C_{1;b}, \dot a : \tptr{\tallarrayb{\bvar}{\tau}}{m}}

      \inferrule[C-Str]
      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} \\
        \echecknull{\dot a_1}{m}{\dot C_{n}} \\
        \echeckbounds{\rho}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{0}{\dot C_{b}} \\
        x_2 = \fresh \\
          \dot C_2 = \elet{x_2}{\estrlen{\dot a_1}}{\hole} \\
      \ewidenstrlen{\rho}{e}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{\dot C_{w}} } 
                {\Gamma;\rho \vdash \estrlen{e} \gg \dot C_{1;n;b;2;w}, x_2 : \tint}

      \inferrule[C-LetStr]
      {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\
        \Gamma; \rho \vdash \estrlen{y} \gg \dot C_1, \dot a_1 : \tint\\
        \dot C_2 = \elet{x}{\dot a_1}{\hole} \\
        \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}]];\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau}
{\Gamma;\rho \vdash \elet{x}{\estrlen{y}}{e} \gg \dot C_{1;2;3},\dot a_3 : \tau}

      \inferrule[C-If ]
      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
        \Gamma; \rho \vdash e_1 \gg \dot C_2, \dot a_2 : \tau_2 \\
        \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
        x_4 = \fresh \\
        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_2[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}
      }
        {\Gamma; \rho \vdash \eif{e_1}{e_2}{e_3} \gg \dot C_4, x_4 : \tau_2 \sqcup \tau_3}

        \inferrule[C-IfNT]
        {\Gamma; \rho \vdash x : \tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c} \\
          b_h = 0 \Rightarrow \Gamma' = \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}] \\
          b_h \neq 0 \Rightarrow \Gamma' = \Gamma \\
          \Gamma; \rho \vdash \estar{x} \gg \dot C_1, \dot a_1 : \tau_1\\
          \Gamma'; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau_2\\
          \Gamma;\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
        \ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C_{w}} \\
        x_4 = \fresh \\
        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_{2;w}[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}}
        {\Gamma;\rho \vdash \eif{\estar{x}}{e_1}{e_2} \gg x_4, \dot C_4 : \tau_1 \sqcup \tau_2}

   \inferrule[C-Let]
    { (x\in \fv(\tau') \Rightarrow e_1 \in Bound) \\
        \Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tau_1 \\ 
        \cextend{\rho}{x}{\tau_1}{\dot C_2,\rho'}\\
        \dot C_3 = \elet{x}{\dot a_1}{\hole}\\
        \Gamma[x\mapsto \tau];\rho' \vdash e_4 \gg \dot C_4, \dot a_4  : \tau_4}
    {\Gamma;\rho' \vdash \elet{x}{e_1}{e_4} \gg \dot C_{1;2;3;4},\dot a_4 : \tau_4[\tau_1 = \tint \Rightarrow x \mapsto e_1]}

  \inferrule[C-Ret]
    {\Gamma(x)\neq \bot \\
      \Gamma;\rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
      x_2 = \fresh \\
      \mu \gg \dot \mu \\
    \dot C_2 = \elet{x_2}{\eret{x}{\dot \mu}{\dot C_1[\dot a_1]}}{\hole}}
  {\Gamma;\rho \vdash \eret{x}{\mu}{e} \gg \dot C_2, x_2 : \tau}

    \inferrule[C-Fun]
    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\
(\forall e_i\in\overline{e}\;\;\tau_i\in\overline{\tau}\;.\; \Gamma; \rho \vdash {e_i} \gg \dot C_i, \dot a_i  : \\ \tau_i' \wedge
      \tau_i' \sqsubseteq 
      \tau_i[\overline{e} / \overline{x}]) \\
      x_f = \fresh \\
      \dot C_f = \elet{x_f}{f(\overline{a})}{\hole}}
    {\Gamma; \rho \vdash f(\overline{e}) \gg \overline{\dot C}[\dot C_f] , x_f  : \tau[\overline{e} / \overline{x}]}

  \inferrule[C-Def]
  {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m} \\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    x_2 = \fresh \\
    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}
   }
   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;2}, x_2 : \tau}

   \inferrule[C-DefArr]
   {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tallarrayptr{b_l}{b_h}{\tau}{m} \\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    \echeckbounds{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
    x_2 = \fresh \\
    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}}
   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;b;2},x_2 : \tau}

    \inferrule[C-Mac]
    {\dot C_{1}, \dot a_1 = \esizeof{\omega} \\
      x_2 = \fresh \\
    \dot C_2 = \elet{x_2}{\emalloc{\dot a_1}}{\hole}}
     {\Gamma; \rho \vdash \emalloc{\omega} \gg \dot C_{1;2}, x_2 : \tptr{\omega}{\cmode}}

  \end{mathpar}
}
\caption{Compilation}
\label{fig:compilation}
\end{figure*}

\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[C-Add]
              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tint \\
                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint\\
                x_3 = \fresh \\
              \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}}
              {\Gamma; \rho \vdash \dot C_3, x_3 : \tint }

   \inferrule[C-Ind] 
              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
                \echecknull{\dot a_1}{m}{\dot C_{n}}\\
                \echeckbounds{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
                x_3, x_4 = \fresh\\
                \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}\\
                \dot C_4 = \elet{x_4}{\estar{x_3}}{\hole}
                }
                {\Gamma; \rho \vdash \estar{(\ebinop{e_1}{e_2})} \gg \dot C_{1;2;n;3;b;4} , x_4: \tau}

  \inferrule[C-Assign]
  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m'} \\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
    \tau'\sqsubseteq \tau \\
    x_3 = \fresh \\
    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
  }
  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;3}, x_3 : \tau}

  \inferrule[C-AssignArr]
  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
    \echeckboundsw{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
    x_3 = \fresh \\
    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
    \tau'\sqsubseteq \tau}
  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;b;3},x_3 : \tau}

    \inferrule[C-IndAssign]
    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
      \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
      \echeckboundsw{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
      \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau' \\
      x_4, x_5 = \fresh \\
      \dot C_4 = \elet{x_4}{\dot a_1 \plus \dot a_2}{\hole}\\
      \dot C_5 = \elet{x_5}{\eassign{x_4}{x_3}{\hole}}
      \tau'\sqsubseteq \tau}
    {\Gamma; \rho \vdash \eassign{(e_1 \plus e_2)}{e_3} \gg \dot C_{1;2;n;3;4;b;5} : \tau}

    \inferrule [C-Struct]
    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1  :  {\tptr{\tstruct{T}}{m}} \\
      D(T) = \tau_0~f_0\ldots; \tau_j~f; ...  \\
      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
      x_2 = \fresh\\
    \dot C_2 = \elet{x_2}{\dot a_1 \plus \dot j}{\hole}}
    {\Gamma; \rho \vdash \eamper{e_1}{f} \gg \dot C_2, x_2  : \tptr{\tau_f}{m}}

    \inferrule[C-Unchecked]
              {\Gamma;\rho \vdash e \gg \dot C, \dot a : \tau}
              {\Gamma;\rho \vdash \eunchecked{e} \gg \dot C, \dot a : \tau}

  \end{mathpar}

}
\caption{Compilation (continued)}
\label{fig:compilation2}
\end{figure*}

\end{document}

### 2022-06-25 sandp23/background.tex
update.

--- 2022-06-25 sandp23/old.tex
@@
-\documentclass[conference]{IEEEtran}
+\section{Overview}\label{sec:overview}
 
-\makeatletter
-\def\endthebibliography{\def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}\endlist
-}
-\makeatother
+Give an overview of Checked-C and RLBox.
 
-\IEEEpeerreviewmaketitle
+\begin{enumerate}
 
-\usepackage{filecontents}
-\usepackage[noadjust]{cite}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{ebproof}
-\usepackage{todonotes}
-\usepackage{amsfonts}
-\usepackage{algorithmic}
-\usepackage{textcomp}
-\usepackage{mathpartir}
-\usepackage{color}
-\usepackage{xcolor}
-\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
+\item discuss Checked-C and its usage.
 
-\usepackage{subcaption} \usepackage{epsfig}
-\usepackage{listings}
-\usepackage[override]{cmtt}
-\usepackage{graphicx}
-\usepackage{balance}
-\usepackage{xspace}
-\usepackage{booktabs}
-\usepackage{url}
-\usepackage{xparse}
-\usepackage{enumitem}
-\usepackage{tikz}
-\usetikzlibrary{chains,fit,shapes,positioning,calc,arrows.meta,shapes.arrows}
+\item discuss RLBox.
 
-\usepackage[numbers]{natbib}
-\bibliographystyle{plainnat}
+\item A running example (web-browser with third party libraries).
 
-\lstdefinestyle{customc}{
-  belowcaptionskip=1\baselineskip,
-  breaklines=true,
-  numbers=left,
-  xleftmargin=\parindent,
-  language=C,
-  columns=flexible,      
-  showstringspaces=false,
-  basicstyle=\small\sffamily,
-  otherkeywords={uint},
-  literate={{<-}{{$\leftarrow\,$}}2
-            {->}{{$\rightarrow\,$}}2
-            {<=}{{$\leq\,$}}2},
-  numberstyle=\tiny\rmfamily,
-  keywordstyle=\bfseries\color{green!40!black},
-  commentstyle=\itshape\color{purple!40!black},
-  identifierstyle=\color{blue!80!black},
-}
-\lstset{language=C,style=customc}
+\end{enumerate}
 
-\newenvironment{DIFnomarkup}{}{}
-\newcommand{\myparagraph}[1]{\textbf{#1}.\xspace}
+\section{Key Updates}\label{sec:key-updates}
 
-\newcommand{\code}[1]{\lstinline|#1|}
-%\newcommand{\ignore}[1]{{}}
-\newcommand{\kwchecked}{\lstinline|\_Checked|}
-\newcommand{\kwunchecked}{\lstinline|\_Unchecked|}
-\newcommand{\kwdynamiccheck}{\lstinline|\_Dynamic\_check|}
+This section discusses the key merge features between Checked-C and RLBox and the problems we are trying to solve.
 
-\newcommand{\var}[1]{#1}
-\newcommand{\Arrayptr}[1]{\lstinline|_Array_ptr<|{#1}\lstinline|>|}
-\newcommand{\ArrayptrT}{\Arrayptr{$\tau$}}
-\newcommand{\Ptr}[1]{\lstinline|_Ptr<|{#1}\lstinline|>|}
-\newcommand{\PtrName}{\lstinline|_Ptr|}
-\newcommand{\PtrT}{\Ptr{$\tau$}}
-\newcommand{\Ntarrayptr}[1]{\lstinline|_Nt_array_ptr<|{#1}\lstinline|>|}
-\newcommand{\NtarrayptrT}{\Ntarrayptr{$\tau$}}
-\newcommand{\Checkedarr}[2]{{#1}~\lstinline|_Checked[|{#2}\lstinline|]|}
-\newcommand{\CheckedarrT}[1]{\Checkedarr{$\tau$}{#1}}
+\begin{enumerate}
 
-\newif\ifsubmit\submitfalse
+\item Pointing out that the current Checked-C and C formalism is not enough to ensure the security of linking web-browser with third party libraries.
+Pointing out that even though Checked-C previously has blame theorem, but it still allows program crashing, and our new work can guarantee that every crashes can be detected and handled (non-crashing theorem).
 
-\ifsubmit
-\newcommand{\mwh}[1]{}
-\newcommand{\ashe}[1]{}
-\newcommand{\dtarditi}[1]{}
-\newcommand{\dvh}[1]{}
-\newcommand{\leo}[1]{}
-\newcommand{\liyi}[1]{}
-\newcommand{\yiyun}[1]{}
-\else
-\newcommand{\mwh}[1]{\textcolor{red}{Mike: #1}}
-\newcommand{\dtarditi}[1]{\textcolor{purple}{David: #1}}
-\newcommand{\leo}[1]{\textcolor{green!80!blue}{Leo: #1}}
-\newcommand{\liyi}[1]{\textbf{\textcolor{orange}{[ Liyi: #1 ]}}}
-\newcommand{\yiyun}[1]{\textcolor{cyan}{Yiyun: #1}}
-\newcommand{\dvh}[1]{\textcolor{magenta}{DVH: #1}}
-\fi
+\item Using an example to highlight the overhead of RLBox, and points out the advantage of building RLBox mechanism in a compiler.
 
-\newcommand{\lang}{\textsc{CoreChkC}\xspace}
-\newcommand{\elang}{\textsc{CoreC}\xspace}
+\item Discuss function pointer formalism and points out the potential security issues without a proper handling of function pointer set up.
+Mentioning what happens in RLBox and other tools.
+Pointing out that this is the first work formalizing function pointers.
 
-\newtheorem{defi}{Definition}
-\newtheorem{thm}{Theorem}
-\newtheorem{prop}{Proposition}
+\end{enumerate}
 
-\newcommand{\CoreChkC}{\textsc{CoreChkC}\xspace}
-\newcommand{\checkedc}{\text{Checked C}\xspace}
+% \paragraph*{\textbf{NT-array Pointer Formalization}} NT-array pointers point to NT-arrays whose endings are determined by a null (\code{'\0'}).
+The length of a NT-array is usually not fixed.
+The two uses of NT-array pointers in \checkedc are branching and \code{strlen} operations.
+\code{if (*x) }$e_1$\code{ else }$e_2$ branches to $e_1$/$e_2$ depending one the \code{x}'s data, while \code{strlen(x)} computes the length of the NT-array pointed to by \code{x}.
+% \liyi{move to formal and later section.}
 
- \newcommand{\rulelab}[1]{{\small \textsc{#1}}}
-\newcommand{\steps}{\ensuremath{\longrightarrow}}
-\newcommand{\tbool}{\texttt{bool}}
-\newcommand{\tsizeof}{\texttt{sizeof}}
-\newcommand{\vbool}{\texttt{Bool}}
-\newcommand{\verror}{\texttt{Error}}
-\newcommand{\vval}{\mathpzc{val}}
-\newcommand{\tfixed}{\texttt{fixedp}}
-\newcommand{\tnat}{\texttt{nat}}
-\newcommand{\tarr}[2]{\texttt{array}~{#1}~{#2}}
-\newcommand{\econst}[2]{({#1}){#2}}
-\newcommand{\eindex}[2]{{#1}\texttt{[}{#2}\texttt{]}}
-\newcommand{\sassign}[4]{{#1} \leftarrow {#3}~{#2}~{#4}}
-\newcommand{\ssassign}[3]{{#1} \xleftarrow{#2} {#3}}
-\newcommand{\sif}[3]{\texttt{if}~{#1}~{#2}~{#3}}
-\newcommand{\sfor}[3]{\texttt{for}~{#1}~{#2}~{#3}}
-\newcommand{\scall}[3]{{#1}\leftarrow {#2}~{#3}}
-\newcommand{\sseq}[2]{{#1}\,\texttt{;}\,{#2}}
-\newcommand{\sinv}[1]{\texttt{inv}~{#1}}
-\newcommand{\inst}[3][ ]{\texttt{#2}^{#1}~{#3}}
-\newcommand{\insttwo}[4][ ]{\texttt{#2}^{#1}~{#3}~{#4}}
-\newcommand{\instthree}[5][ ]{\texttt{#2}^{#1}~{#3}~{#4}~{#5}}
-\newcommand{\iskip}[1]{\inst{ID}{#1}}
-\newcommand{\inot}[1]{\inst{X}{#1}}
-\newcommand{\ictrl}[2]{\insttwo{CU}{#1}{#2}}
-\newcommand{\irz}[3][ ]{\insttwo[#1]{RZ}{#2}{#3}}
-\newcommand{\isr}[3][ ]{\insttwo[#1]{SR}{#2}{#3}}
-\newcommand{\icnot}[2]{\insttwo{CNOT}{#1}{#2}}
-\newcommand{\ilshift}[1]{\inst{Lshift}{#1}}
-\newcommand{\irshift}[1]{\inst{Rshift}{#1}}
-\newcommand{\irev}[1]{\inst{Rev}{#1}}
-\newcommand{\iqft}[2][ ]{\inst[#1]{QFT}{#2}}
-\newcommand{\ihad}[1]{\inst{H}{#1}}
-
-\newcommand{\iseq}[2]{{#1}\,\texttt{;}\,{#2}}
-\newcommand{\inval}[2]{\insttwo{Nval}{#1}{#2}}
-\newcommand{\ihval}[3]{\instthree{Hval}{#1}{#2}{#3}}
-\newcommand{\iqval}[2]{\insttwo{Qval}{#1}{#2}}
-
-\newcommand{\instr}{\iota}
-
-\newcommand{\kw}[1]{\ensuremath{\mathtt{#1}}}
-\newcommand{\subtype}[2]{\ensuremath{\vdash{#1}\prec{#2}}}
-\newcommand{\estrlen}[1]{\ensuremath{\kw{strlen}({#1})}}
-\newcommand{\estrlentext}{\ensuremath{\kw{strlen}}}
-
-\newcommand{\tarray}[3]{\tarrayb{({#1},{#2})}{#3}}
-\newcommand{\tarrayb}[2]{\ensuremath{[{#1}~{#2}]}}
-\newcommand{\tntarray}[3]{\tntarrayb{({#1},{#2})}{#3}}
-\newcommand{\tntarrayb}[2]{\tarrayb{#1}{#2}_{nt}}
-\newcommand{\tallarrayb}[2]{\ensuremath{[{#1}~{#2}]_{\kappa}}}
-\newcommand{\tallarraybc}[2]{\ensuremath{[{#1}~{#2}]_{\textcolor{cyan}{\kappa}}}}
-
-\newcommand{\tallarray}[3]{\tallarrayb{({#1},{#2})}{#3}}
-
-\newcommand{\tptr}[2]{\ensuremath{\mathtt{ptr}^{#2}~{#1}}}
-\newcommand{\tntptr}[4]{\ensuremath{[({#1},{#2})~{#3}]^{#4}_{nt}*}}
-
-\newcommand{\tarrayptr}[4]{{\tptr{\tarray{#1}{#2}{#3}}{#4}}}
-\newcommand{\tntarrayptr}[4]{{\tptr{\tntarray{#1}{#2}{#3}}{#4}}}
-\newcommand{\tallarrayptr}[4]{{\tptr{\tallarray{#1}{#2}{#3}}{#4}}}
-
-\newcommand{\tgarray}[4]{\ensuremath{\mathtt{#1}~{#2}~{#3}~{#4}}}
-\newcommand{\tstruct}[1]{\ensuremath{\kw{struct}~{#1}}}
-
-\newcommand{\evalue}[2]{\ensuremath{{#1}\!:\!{#2}}}
-
-\newcommand{\emalloc}[1]{\ensuremath{\kw{malloc}({#1})}}
-\newcommand{\emalloctext}{\ensuremath{\kw{malloc}}}
-\newcommand{\ecall}[2]{\ensuremath{{#1}({#2})}}
-\newcommand{\ret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
-\newcommand{\rettext}{\ensuremath{\kw{ret}}}
-\newcommand{\ecast}[2]{\ensuremath{\kw{(}{#1}\kw{)}{#2}}}
-\newcommand{\edyncast}[2]{\ensuremath{\langle{#1}\rangle{#2}}}
-\newcommand{\edcast}[2]{\ensuremath{\kw{(D},{#1}\kw{)}{#2}}}
-\newcommand{\elet}[3]{\ensuremath{\kw{let}~#1\, \texttt{=}\, #2~\kw{in}\;{#3}}}
-\newcommand{\elettext}{\ensuremath{\kw{let}}}
-\newcommand{\ebinop}[2]{\ensuremath{#1 \plus #2}}
-\newcommand{\eassign}[2]{\ensuremath{\texttt{*}{#1}\,\texttt{=}\, {#2}}}
-\newcommand{\eassignstack}[2]{\ensuremath{{#1}\,\texttt{=}\, {#2}}}
-\newcommand{\efield}[2]{\ensuremath{{#1}\kw{\rightarrow}{#2}}}
-\newcommand{\estar}[1]{\ensuremath{\texttt{*}{#1}}}
-\newcommand{\eamper}[2]{\ensuremath{\kw{\&}{#1}\kw{\rightarrow}{#2}}}
-\newcommand{\eunchecked}[1]{\ensuremath{\kw{unchecked}\;{#1}}}
-\newcommand{\euncheckedtext}{\ensuremath{\kw{unchecked}}}
-\newcommand{\eif}[3]{\ensuremath{\kw{if\;}(#1)\;{#2}\;\kw{else}\;{#3}}}
-\newcommand{\eifa}[3]{\ensuremath{\kw{IF\;}(#1)\;{#2}\;\kw{ELSE}\;{#3}}}
-\newcommand{\eifatext}{\ensuremath{\kw{IF}}}
-
-\newcommand{\eiftext}{\ensuremath{\kw{if}}}
-\newcommand{\ethentext}{\ensuremath{\kw{then}}}
-\newcommand{\eelse}{\ensuremath{\kw{else}}}
-\newcommand{\ebreak}{\ensuremath{\kw{break}}}
-\newcommand{\ebounds}{\ensuremath{\kw{bounds}}}
-\newcommand{\enull}{\ensuremath{\kw{null}}}
-\newcommand{\hole}{\ensuremath{\Box}}
-\newcommand{\defscope}{\sigma}
-\newcommand{\wf}[1]{wf \, #1}
-\newcommand{\tint}{\ensuremath{\mathtt{int}}}
-\newcommand{\efor}{\ensuremath{\mathtt{for}}}
-\newcommand{\ewhile}{\ensuremath{\mathtt{while}}}
-\newcommand{\etrue}{\ensuremath{\mathtt{true}}}
-\newcommand{\efalse}{\ensuremath{\mathtt{false}}}
-\newcommand{\edplus}{\ensuremath{\texttt{++}}}
-\newcommand{\edminus}{\ensuremath{\texttt{--}}}
-
-\newcommand{\ememcpy}{\ensuremath{\mathtt{memcpy}}}
-\newcommand{\ereturn}{\ensuremath{\mathtt{return}}}
-\newcommand{\emain}{\ensuremath{\mathtt{main}}}
-\newcommand{\estrcat}{\ensuremath{\mathtt{strcat}}}
-\newcommand{\estrcatbad}{\ensuremath{\mathtt{strcat\_b}}}
-\newcommand{\heap}{\ensuremath{\mathpzc{H}}}
-\newcommand{\eret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
-\newcommand{\erettext}{\ensuremath{\kw{ret}}}
-
-\newcommand{\plus}{\mathbin{\texttt{+}}}
-
-\newcommand{\fv}{\mathit{FV}}
-\newcommand{\dom}{\mathit{dom}}
-
-\newcommand{\fm}{\mathit{FM}}
-\newcommand{\size}{\mathit{size}}
-
-\newcommand{\cmode}{\texttt{c}}
-\newcommand{\umode}{\texttt{u}}
-\newcommand{\bvar}{\ensuremath{\beta}}
-\newcommand{\mode}{\textit{mode}}
-
-\newcommand{\cextend}[4]{\ensuremath{#4 = ~ \vdash_{extend} #1, #2, #3}}
-\newcommand{\fresh}{\ensuremath{\kw{fresh}}}
-\newcommand{\echecknull}[3]{\ensuremath{#3 = ~\vdash_{null}#1, #2}}
-\newcommand{\echeckboundsdyn}[5]{\ensuremath{#5 = ~ \vdash_{boundsD}#1, #2, #3, #4}}
-\newcommand{\echeckbounds}[5]{\ensuremath{#5 = ~ \vdash_{boundsR}#1, #2, #3, #4}}
-\newcommand{\echeckboundsw}[5]{\ensuremath{#5 = ~ \vdash_{boundsW}#1, #2, #3, #4}}
-\newcommand{\esizeof}[1]{\ensuremath{\kw{sizeof}(#1)}}
-\newcommand{\ewidenstrlen}[5]{\ensuremath{#5 = ~ \vdash_{widenstr}#1, #2, #3, #4}}
-\newcommand{\ewidenderef}[4]{\ensuremath{#4 = ~ \vdash_{widenderef}#1, #2, #3}}
-\newcommand{\eleq}[2]{\ensuremath{#1 \leq #2}}
-
-\newcommand{\rboundle}[2]{\ensuremath{\vdash_{bounds}#1 \leq #2}}
-
-\def\codesize{\normalsize}
-\definecolor{programs}{gray}{0.1}
-
-\lstset{
-  basicstyle=\ttfamily,  
-}
-
-\makeatletter
-\lst@CCPutMacro
-    \lst@ProcessOther {"22}{\lst@ifupquote \textquotedbl
-                                     \else \char34\relax \fi}
-    \@empty\z@\@empty
-\makeatother
-
-\lstloadlanguages{C}
-
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
-
-\title{A Formal Model of Checked C}
-
-% \ignore{
-% \author{\IEEEauthorblockN{1\textsuperscript{st} Leonidas Lampropoulos}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{University of Maryland}\\
-% llampro@cs.umd.edu}
-% \and
-% \IEEEauthorblockN{1\textsuperscript{st} Michael Hicks}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{University of Maryland}\\
-% llampro@cs.umd.edu}
-% }
-% \author{Leonidas Lampropoulos \and
-% Michael Hicks}
-% }
-\maketitle
-
-\begin{abstract}
-
-  We present a formal model of Checked C, a dialect of C
-  that aims to enforce spatial memory safety. Our model pays particular
-  attention to the semantics of dynamically sized, potentially null-terminated
-  arrays.
-We formalize this model in Coq, and prove that any spatial memory
-  safety errors can be \emph{blamed} on portions of the program
-  labeled \emph{unchecked}; this is a Checked C feature that supports
-  incremental porting and backward compatibility.
-Our model develops an operational semantics that uses fat pointers to
-  guarantee spatial safety. However, we formalize a compilation scheme
-  that can yield thin pointers, with bounds information managed using
-  inserted code. We show that the generated code faithfully simulates
-  the original.
-Finally, we build an executable version of our model in PLT Redex, 
-  and use a custom random generator
-  for well-typed and almost-well-typed terms to find
-  inconsistencies between our model and the Clang Checked C
-  implementation. We find these steps to be a useful way to co-develop a language
-  (Checked C is still in development) and a core model of it.
-\end{abstract}
-
-\section{Introduction}\label{sec:intros}
-
-The C programming language remains extremely popular despite the
-emergence of new, modern languages. Unfortunately, C programs lack
-spatial memory safety, which makes them susceptible to a host
-of devastating vulnerabilities, including buffer overflows and
-out-of-bounds reads/writes. Despite their long history, buffer
-overflows and other spatial safety violations are among the most
-prevalent and dangerous vulnerabilities on the Internet today \cite{Zeng:2013:SRF:2534766.2534798}.
-
-Several industrial and research efforts---including CCured~\cite{Necula2005},
-Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
-explored means to compile C programs 
-to automatically enforce spatial safety. These
-approaches all impose performance overheads that are deemed too high for
-use in deployment. Recently, Microsoft introduced \checkedc, an
-open-source extension to C with new types and
-annotations whose use can ensure a program’s spatial
-safety~\cite{Elliott2018}. Importantly, \checkedc supports development that is
-incremental and compositional. Code regions (e.g.,
-functions or whole files) designated as \emph{checked} are sure to enforce
-spatial safety, a property which is preserved via composition with
-other checked regions. But not all regions must be checked: Checked
-C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
-may coexist in the same code, which permits a deliberate (and
-semi-automated) refactoring process. Parts of the FreeBSD kernel have
-been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
-overhead seems low enough for practical deployment.
-
-While \checkedc promises to enforce spatial safety, we might wonder
-whether its design and implementation deliver on this promise, or even
-what ``spatial safety'' means when a program contains both checked and
-unchecked code. In prior work, \citet{ruef18checkedc-incr} developed a core
-formalization of \checkedc and with it proved that \emph{checked code cannot 
-  be blamed}: any spatial safety violation can only be attributed to
-code that is not in a checked region. While their work is a good start,
-it fails to model important aspects of \checkedc's functionality,
-particularly those involving pointers to arrays. In this paper, we
-cover this gap, making three main contributions.
-
-\myparagraph{Dynamically bounded and null-terminated arrays}
-Our first contribution is a core formalism called \lang, which extends
-\citet{ruef18checkedc-incr} with several new features, most notably 
-\emph{dynamically bounded arrays} (Section~\ref{sec:formal}). 
-Dynamically bounded arrays are those whose size is
-known only at run time, as designated by in-scope variables using
-dependent types. A pointer's accessible memory is bounded both above
-and below, to admit arbitrary pointer arithmetic.
-
-\lang also models
-\emph{null-terminated} arrays, a kind of dynamically bounded array whose upper bound 
-defines the array's \emph{minimum} length---additional space is available 
-up to a null terminator. For example, the \checkedc type
-\lstinline{nt_array_ptr<char> p:count(n)} says that
-\lstinline{p} has length \emph{at least} \lstinline{n} (excluding the
-null terminator), but further capacity is present if \lstinline|p[n]| is not null. \checkedc (and \lang) supports flow-sensitive
-\emph{bounds widening}: statements of the form \lstinline|if (*p) | $s$,
-where \lstinline{p}'s type is \lstinline{nt_array_ptr<T> count(0)},
-typecheck statement $s$ under the assumption that \lstinline{p} has
-type \lstinline{nt_array_ptr<T> count(1)}, 
- i.e., one more than it was,
-since the character at the current-known length is
-non-null. Similarly, the call \lstinline|n = strlen(p)| will widen
-\lstinline|p|'s bounds to \lstinline|n|. Subtyping permits treating
-null-terminated arrays as normal arrays of the same size (which
-does not include, and thereby protects, the null terminator).
-
-We prove, in Coq, a blame theorem for \lang.  As far as we are aware,
-ours is the first formalized type system and proof of soundness for
-pointers to null-terminated arrays with expandable bounds.
-
-\myparagraph{Sound compilation of checked pointers} Our second
-contribution is a formalization of bounds-check insertion for array
-accesses (Section~\ref{sec:compilation}). Our operational semantics
-annotates each pointer with metadata that describes its bounds, and
-the assignment and dereference rules have premises to confirm the
-access is in bounds. An obvious compilation scheme (taken by
-Cyclone~\cite{Jim2002,GrossmanMJHWC02}, CCured~\cite{Necula2005}, and
-earlier works) would be to translate annotated pointers to multi-word
-objects: one word for the pointer, and 1-2 words to describe its lower
-and upper bounds. Inserted checks reference these bounds. While
-convenient, such ``fat'' pointers are expensive, and break backward
-binary compatibility with legacy pointers. We formalize \checkedc's
-compilation approach, which uses a single machine word for the
-pointer, and adds checks involving the declared bounds (e.g., in a
-dependent type) or additional stack-allocated \emph{ghost variables}
-to accommodate bounds widening. We show that the compiled program
-\emph{simulates} the original by mechanizing \lang and the compilation
-judgment in PLT Redex~\cite{pltredex}, and use its random testing
-feature to give confidence that simulation holds.
-
-As far as we are aware, \lang is the first formalism to cleanly
-separate bounds-checking compilation from the core semantics; prior
-work merged the two, conflating \emph{meaning} with
-\emph{mechanism}~\cite{Feng2006,Condit2007}. In carrying out the 
-formalization, we discovered that our compilation approach is 
-more expressive than that proposed in the Checked C
-specification~\cite{checkedc} (Section~\ref{sec:disc}); we 
-would not have discovered this improvement had we not separated checks from 
-semantics.
-
-\myparagraph{Model-based randomized testing} Our third
-and final contribution is a strategy and implementation of model-based
-randomized testing (Section~\ref{sec:evaluation}).  To check the
-correctness of our formal model, we compare the behavior between the
-existing Clang \checkedc implementation and our own model. This is
-done by a conversion tool that converts expressions from \lang into
-actual \checkedc code that can be compiled by the Clang \checkedc
-compiler. We build a random generator of programs largely based on the
-typing rules of \lang and make sure that, both statically and
-dynamically, \lang and Clang Checked C are consistent after
-conversion.  This helped rapidly prototype the model and uncovered
-several issues in the Checked C compiler.
-
-We begin with a review of \checkedc (Section~\ref{sec:overview}),
-present our main contributions (Sections~\ref{sec:formal}--\ref{sec:evaluation}),
-and conclude with a discussion of
-related and future work (Sections~\ref{sec:related}, \ref{sec:conclude}).
-
-\section{\checkedc Overview}\label{sec:overview}
-
-This section describes Checked C, which extends C with new pointer
-types and annotations that ensure spatial safety. More
-details can be found in a prior overview~\cite{Elliott2018} or the
-full specification~\cite{checkedc}.
-Checked C is implemented as a fork of Clang/LLVM and is freely
-available.\footnote{\url{https://github.com/Microsoft/checkedc-clang}}
-
-\begin{figure}[t]
-{\small
-  \begin{lstlisting}[xleftmargin=4 mm]
-nt_array_ptr<const char>
-parse_utf16_hex(nt_array_ptr<const char> s,
-                  ptr<uint> result) {
- int x1, x2, x3, x4;
- if (s[0] != 0) { x1 = hex_char_to_int(s[0]);
- if (s[1] != 0) { x2 = hex_char_to_int(s[1]);
- if (s[2] != 0) { x3 = hex_char_to_int(s[2]);
- if (s[3] != 0) { x4 = hex_char_to_int(s[3]);
- if (x1 != -1 && x2 != -1 && x3 != -1 && x4 != -1){
-   *result = (uint)((x1<<12)|(x2<<8)|(x3<<4)|x4);
-   return s+4;
-  ...// several } braces
- } 
- return 0;
-}
-void parse(nt_array_ptr<const char> s,
-            array_ptr<uint> p : count(n), 
-            int n) {
- array_ptr<uint> q : bounds(p,p+n) = p;
- while (s && q < p+n) {
-   array_ptr<uint> r : count(1) =
-     dyn_bounds_cast<array_ptr<uint>>(q,count(1));
-   s = parse_utf16_hex(s,r);
-   q++;
- }
-}
-  \end{lstlisting}
-
-}
-
-\caption{Parsing a string of UTF16 hex characters in \checkedc}
-\label{fig:checkedc-example}
-\end{figure}
-
-\subsection{Checked Pointer Types}
-\checkedc introduces three varieties of \emph{checked pointer}:
-\begin{itemize}
-\item \code{ptr<}$T$\code{>} types a pointer that is either null or
-  points to a single object of type $T$.
-\item \code{array_ptr<}$T$\code{>} types a pointer that is either null
-  or points to an array of $T$ objects. The array width is defined
-  by a \emph{bounds} expression, discussed below.
-\item \code{nt_array_ptr<}$T$\code{>} is like
-  \code{array_ptr<}$T$\code{>} except that the bounds expression
-  defines the \emph{minimum} array width---additional objects may
-  be available past the upper bound, up to a null terminator.
-\end{itemize}
-A bounds expression used with the latter two pointer types has three
-forms:
-\begin{itemize}
-\item \code{count(}$e$\code{)} where $e$ defines the array's
-  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
-  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
-  expression $e$ must be side-effect free and may only refer to
-  variables whose addresses are not taken, or adjacent \code{struct}
-  fields.
-\item \code{byte_count(}$e$\code{)} is like \code{count}, but
-  expresses arithmetic using bytes, no objects; i.e.,
-  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
-  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
-\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
-  pointers that bound the accessible region $[e_l,e_h)$ (the
-  expressions are similarly restricted). Bounds
-  \code{count(}$e$\code{)} is shorthand for
-  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
-  expression is useful for supporting pointer arithmetic.
-\end{itemize}
-The \checkedc compiler will instrument loads and stores of checked
-pointers to confirm the pointer is non-null, and the access is within
-the specified bounds. For pointers $p$ of type
-\code{nt_array_ptr<}$T$\code{>}, such a check could spuriously fail if
-the index is past $p$'s specified upper bound, but before the null
-terminator. To address this problem, Checked C supports \emph{bounds
-  widening}. If $p$ has bounds expression
-\code{bounds(}$e_l$,$e_h$\code{)} a program may read from (but not
-write to) $e_h$; when the compiler notices that a non-null character
-is read at the upper bound, it will extend that bound to $e_h+1$.
-
-\subsection{Example}
-
-Fig.~\ref{fig:checkedc-example} gives an example \checkedc
-program.\footnote{Ported from the Parson JSON
-parser, \url{https://github.com/kgabis/parson}} 
-The function \code{parse_utf16_hex} on lines 1-17 takes as its argument
-null-terminated pointer \code{s} from which it attempts to read four
-characters. These are interpreted as hex digits and converted to an
-\code{uint} returned via parameter \code{result}. At first,
-\code{s} has no specific bounds annotation, which we can interpret as
-\code{count(0)}; this means that \code{s[0]} may be read on line
-5. The true branch of the conditional (which extends all the way to
-the brace on line 15) is thus typechecked with \code{s} given a
-\emph{widened} bound of \code{count(1)}. Likewise, the conditionals on
-lines 6-8 each widen it one further; the widened pointer
-(\code{s+4}) is returned on success.
-
-The \code{parse} function on lines 18-26 repeatedly invokes
-\code{parse_utf16_hex} with its parameter \code{s}, and fills out
-array \code{p} whose declared length is the parameter \code{n}. Writes
-happen via pointer \code{q}, which is updated using pointer
-arithmetic. We specify its bounds as
-\code{bounds(p,p+n)} to support this: even as \code{q} changes, 
-variables \code{p} and \code{n} do not. Converting from an
-\code{array_ptr<uint>} to a \code{ptr<uint>}, done for the call on
-line 25, requires proving the array has size at least 1. While this is true
-because of the loop condition \code{q < p+n}, which is \code{q}'s
-upper bound, the compiler is not smart enough to figure this
-out. To convince it, we manually insert a bounds check using
-\code{dyn_bounds_cast}.
-
-While bounds checks are \emph{conceptually} inserted on every array
-load and store, many of these are eliminated by LLVM\@. For example,
-all of the pointer accesses to \code{s} on lines 5-8 are proved safe
-at compile-time, so no bounds checks are inserted for
-them. \citet{Elliott2018}
-reported average run-time overheads of 8.6\% on a pointer-intensive
-benchmark suite (49.3\% in one case); \citet{duanrefactoring} measured
-no overhead at all on a port of FreeBSD's UDP and IP stacks to \checkedc.
-
-\subsection{Other features}
-
-\checkedc has other features not modeled in this paper. Two in regular
-use are \emph{interop types}, which ascribe checked pointer types to
-unported legacy code, notably in libraries; and \emph{generic types}
-on both functions and \code{struct}s, for type-safe polymorphism. More
-details about these can be found in the language specification.
-
-\subsection{Spatial Safety and Backward Compatibility}
-\checkedc is backward compatible with legacy C in the sense that all
-legacy code will typecheck and compile. However, only code that
-appears in \emph{checked regions}, which we call \emph{checked code},
-is spatially safe. Checked regions can be designated at the level of
-files, functions, or individual code blocks, the first with a
-\code{#pragma} and the latter two using the \code{checked}
-keyword.\footnote{You can also designate \emph{unchecked} regions
-  within checked ones.}  Within checked regions, both
-legacy pointers and certain unsafe idioms (e.g., variadic function
-calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
-satisfies these conditions, and will typecheck in a checked region.
-
-How should we think about code that contains both checked and legacy
-components? \citet{ruef18checkedc-incr} proved, for a simple
-formalization of \checkedc, that \emph{checked code cannot be blamed}:
-Any spatial safety violation owes to the execution of unchecked
-code. In this paper we extend that result to a richer formalization of
-\checkedc.
+% \paragraph*{\textbf{Dependent Functions}} \checkedc allows users to declare dependent functions.
+Fig.~\ref{fig:checkedc-example} shows an example usage of dependent functions.
+The integer argument \code{x} is used to specify the bounds of the two NT-array pointers.
+It means that when we call the \code{memcpy} function, the upper-bounds of the two NT-array pointers must be no less than \code{x}.
+If the programmer makes the mistake of writing \lstinline{i <= x} in the for loop, the runtime checks will ensure that the indexing of\lstinline|x| in either \code{a} or \code{b} results in an error.
+% Because of the dependent function features, we are able to utilize the fact that the two remaining length of the NT-array pointers are greater than \code{x}, so that the \code{for} is guaranteed to execute properly.
 
@@
 The fat-pointer approach completely relies on dynamic checks to guarantee the safety properties but it might pay a huge price on execution efficiency.
-In our experiment (Sec.\ref{sec:evaluation}), we found that the execution of fat pointer approaches cost the speed a 50\% overhead.
+In our experiment (Sec.~\ref{sec:evaluation}), we found that the execution of fat pointer approaches cost the speed a 50\% overhead.
 % On the other hand, Rust uses a static approach by restricting the usage of pointers to guarantee the properties, which results in a user UN-friendly system.  
@@
 % }
-
-% \ignore{
-% \section{Null-Terminated Array Pointers}
-
-% Key ideas, formalized:
-% \begin{itemize}
-% \item Can read one past the end: x[N] can be read, but only written if
-%   the value being written is 0.
-% \item NT arrays are a refinement of normal arrays: Can convert
-%   x:(ntarray N t) to (array N t).
-% \item NT arrays can expand their length flow sensitively: Can convert
-%   x:(ntarray N t) to (ntarray N+1 t) assuming x[N] != 0.
-% \item In general unsafe to convert x:(array N t) to (ntarray N-1 t)
-%   even if x[N-1] == 0 because aliases of x could eliminate the NULL
-%   terminator. Would need some sort of linearity/alias tracking to
-%   support this idea.
-% \item Question? What is the syntax of doing a malloc on a NT-array?
-%   Should we just do malloc(NT-array l h type) as the malloc of a
-%   normal array or should we also include something like int [] nt =
-%   “abcd”, where the bound size is not specified?
-% \end{itemize}
-
-% See also the Checked C specification
-% \ignore{
-% https://github.com/microsoft/checkedc/tree/master/spec/bounds\_safety}
-% }
-
-\section{Formalization}\label{sec:formal}
-
-% \ignore{
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[height=6in]{syntax.pdf}
-%   \end{center}
-%   \caption{\lang: Syntax}
-% \end{figure}
-
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[height=6in]{types.pdf}
-%   \end{center}
-%   \caption{\lang: Typing}
-% \end{figure}
-% }
-
-\begin{figure}
-  \small \centering
-  $\begin{array}{l}
-\begin{array}{lll}
-\text{Function names:}~f&
-       \text{Variables:}~ x
-& \text{Integers:}~n::=\mathbb{Z} 
-\end{array}
-\\[0.5em]
-
-\begin{array}{llcllcl}
-
-\text{Mode:} & m & ::= & \cmode \mid \umode \\[0.5em]
-
-\text{Bound:} & b & ::= & n \mid x \plus n \\
-              & \bvar & ::= & (b,b) \\[0.5em]
-
-     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{m}
-\\[0.5em]
-
-\text{Type Flag:}&\kappa &::=& nt \mid \cdot
-\\[0.5em]
-
-\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau}
-\\[0.5em]
-
-\text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \emalloc{\omega} \mid\elet{x}{e}{e} \\[0.2em]
-&&\mid&
-\ecast{\tau}{e} \mid \edyncast{\tau}{e}\mid \ecall{f}{\overline{e}} \mid \estrlen{x} \\[0.2em]
-&&\mid&
-\ebinop{e}{e} \mid\estar{e}\mid\eassign{e}{e}\mid\eunchecked{e}
-\\[0.2em]
-&&\mid&\eif{e}{e}{e}
-\end{array}
-    \end{array}
-  $
-  \caption{\lang Syntax}
-  \label{fig:checkc-syn}
-\end{figure}
-
-This section describes our formal model of \checkedc, called
-\lang, making precise its syntax, semantics, and type system. It also
-develops \lang's metatheory, including type soundness and the blame
-theorem.
-
-\subsection{Syntax}\label{sec:syntax}
-
-The syntax of \lang is given by the expression-based
-language presented in Fig.~\ref{fig:checkc-syn}.
-
-There are two notions of type in \lang.  Types $\tau$ classify
-word-sized values including the integers and pointers, while types
-$\omega$ classify multi-word values such as arrays, null-terminated
-arrays, and single-word-size values.
-Pointer types ($\tptr{\omega}{m}$) include a mode annotation ($m$)
-which is either checked (\cmode) or unchecked (\umode) and a type
-($\omega$) denoting the type of value to which is pointed. Array types include both the type of
-elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
-lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
-limited to integer literals $n$ and expressions $x + n$.
-Whether an array pointer is null terminated or not is determined by annotation
-$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide the $\cdot$ when writing the type). Here is the
-corresponding \checkedc syntax for these types:
-\[
-\begin{array}{rcl}
-$\code{array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\cmode}
-\\[0.2em]
-$\code{nt_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
-\end{array}
-\]
-As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
-$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
-be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
-$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-
-\lang expressions include literals ($n\!:\!\tau$), variables ($x$), memory
-allocation ($\emalloc{\omega}$), let binding ($\elet{x}{e_1}{e_2}$),
-static and dynamic casts ($\ecast{\tau}{e}$ and $\edyncast{\tau}{e}$,
-resp.), function calls ($\ecall{f}{\overline{e}}$), addition
-($\ebinop{e_1}{e_2}$), pointer dereference and assignment ($\estar{e}$
-and $\eassign{e_1}{e_2}$, resp.), unchecked blocks ($\eunchecked{e}$),
-the \texttt{strlen} operation ($\estrlen{x}$), and conditionals $\eif{e}{e_1}{e_2}$.
-
-Integer literals $n$ are annotated with a type $\tau$ which can be either
-$\tint$, or $\tptr{\omega}{m}$ in the case $n$ is being used as
-a heap address (this is useful for the semantics). The
-$\texttt{strlen}$ expression operates on variables $x$
-rather than arbitrary expressions to simplify managing
-bounds information in the type system; the more general case can be
-encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
-casts; e.g., the following
-
-\code{dyn_bounds_cast<array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}
-
-\noindent
-becomes $\edyncast{\tptr{\tntarrayb{n}{\tau}}{\cmode}}{e}$.
-
-\lang aims to be simple enough to work with, but powerful enough to
-encode realistic \checkedc idioms. For example: mutable local
-variables can be encoded as immutable locals that point to the heap;
-the use of \code{&} can be simulated with \code{malloc};
-and loops can be encoded as recursive function calls. \code{struct}s are
-not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
-actually in our model, and developed in
-Appendix~\ref{appx:struct}. C-style \code{union}s have no safe typing
-in \checkedc, so we elide them. By default, functions are assumed to
-be within checked regions; placing the body in an \code{unchecked}
-expression relaxes this, and within that, checked regions can be
-nested within via function calls. Bounds are restricted slightly:
-rather than allowing arbitrary subexpressions, bounds must be either
-integer literals or variables plus an integer offset, which accounts
-for most uses of \code{bounds} in \checkedc programs. \lang bounds are
-defined as relative offsets, not absolute ones, as in the second part
-of Fig.~\ref{fig:checkedc-example}. We see no technical problem to
-modeling absolute bounds, but it would be a pervasive change so we
-have not done so.
-
-\begin{figure}
-{\small
-$    \begin{array}{l}
-    \begin{array}{lll}
-\mu & ::= & \evalue{n}{\tau} \mid \bot\\
-e & ::= & \ldots \mid \ret{x}{\mu}{e}\\
-r & ::= & e \mid \enull \mid \ebounds\\
-E &::=& \Box \mid \elet{x}{E}{e} \mid \ecall{f}{\overline{E}}\mid\ecast{\tau}{E} \mid \edyncast{\tau}{E} \\[0.2em]
-&&\mid \ret{x}{\evalue{n}{\tau}}{E}\mid\ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid\estar{E}\mid\eassign{E}{e} \\[0.2em]
-&&\mid\eassign{\evalue{n}{\tau}}{E}\mid\eunchecked{E}\mid\eif{E}{e}{e}\\[0.2em]
-\overline{E} &::=& E \mid  \evalue{n}{\tau},\overline{E} \mid \overline{E}, e
-\end{array}
-\\ \\
-    \end{array} 
-$
-  \begin{mathpar}
-    \inferrule{ m=\mode(E) \\
-      e=E[e'] \\
-      (\varphi,\heap,e') \longrightarrow (\varphi',\heap',e'')}
-    {(\varphi,\heap,e)\longrightarrow_{m} (\varphi',\heap',E[e''])}
-
-  \end{mathpar}
-}
-  \caption{\lang semantics: Evaluation}
-  \label{fig:c-context}
-\end{figure}
-
-\subsection{Semantics}\label{sec:semantics}
-
-The operational semantics for \lang is defined as a small-step
-transition relation with the judgment $ (\varphi,\heap,e)
-\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
-\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
-$\heap$ is a \emph{heap} mapping addresses (integer literals) to
-values $\evalue{n}{\tau}$; for both we ensure
-$\fv(\tau)=\emptyset$. The relation steps to a \emph{result} $r$,
-which is either an expression or a $\enull$ or $\ebounds$ failure,
-representing a null-pointer dereference or out-of-bounds access,
-respectively. Such failures are a \emph{good} outcome; stuck states
-(non-value expressions that cannot transition to a result $r$)
-characterize undefined behavior.
-The mode $m$ indicates whether the
-stepped redex within $e$ was in a checked ($\cmode$) or
-unchecked ($\umode$) region.
-
-The rules for the main operational semantics judgment---\emph{evaluation}---are given at the bottom of
-Fig.~\ref{fig:c-context}. The rule takes an expression $e$, decomposes
-it into an \emph{evaluation context} $E$ and a subexpression $e'$
-(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
-$e$), and then evaluates $e'$ according to the \emph{computation} relation $
-(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$, discussed
-shortly. The computation relation can transition to any $r$; the rule
-in the figure just considers $r = e$. For the cases $r = \enull$ and
-$r = \ebounds$, two other rules (not shown) will cause the whole
-evaluation to terminate with $r$.  The semantics has a special case
-for a conditional $\eif{e_0}{e_1}{e_2}$.  If $e_0=\estar{x}$, the
-conditional itself is considered as the redex with special handling
-for bound widening.  Otherwise, the standard conditional behavior is
-used.
-The $\mode$ function
-determines the mode of the evaluating $e'$ based on the context $E$:
-if the $\Box$ in $E$ occurs in context $\eunchecked{E'}$, the mode is
-$\umode$, otherwise, it is $\cmode$. Evaluation contexts $E$ define a
-standard left-to-right evaluation order. (We explain the
-$\ret{x}{\mu}{e}$ syntax shortly.)
-
-\begin{DIFnomarkup}
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[S-Cast]
-              {}
-              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
-
-    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}
-
-        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
-
-    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-
-        \inferrule[S-DefNTArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h]}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-Fun]{ \Xi(f) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
-        {(\varphi,\heap,\ecall{f}{{\evalue{\overline{n}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-
-  \inferrule[S-DynCast]{
-\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{m})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{m} \\ n_l' \le n_l\ \\ n_h \le n_h'}
-    { (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{m}}}) \longrightarrow (\varphi',\heap',\evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau_b}{m}})}
-
- \inferrule[S-DynCastBound]{\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{\cmode})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{\cmode} \\ n_l' > n_l \vee n_h > n_h'}{ (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{\cmode}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{\cmode}}})
- \longrightarrow (\varphi',\heap',\ebounds)}
-
-      \inferrule[S-StrWiden]{
-         \varphi(x) = \evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{\cmode}}\\ 0 \in [n_l,n_h] \\ n_a > n_h
-\\ \heap(n+n_a) = 0 
-\\\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
-    {(\varphi,\heap,\estrlen{x}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tarrayptr{n_l}{n_a}{\tau}{\cmode}}],\heap,\evalue{n_a}{\tint})}
-
-     \inferrule[S-IfNTT]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(n)\neq 0}
-    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}
-
-    \inferrule[S-AssignArr]{\heap(n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}
-
-        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\end{mathpar}
-}
-\caption{\lang semantics: Computation (selected rules)}
-\label{fig:semantics}
-\end{figure*}
-\end{DIFnomarkup}
-
-\begin{figure}[t]
-{\small
-{\captionsetup[lstlisting]{margin = 8 mm}
-  \begin{lstlisting}[xleftmargin=8 mm]
-nt_array_ptr<int> strncat : count(0)
-  (nt_array_ptr<int> a : count(n),
-   nt_array_ptr<int> b : count(0), int n) {
-  int x = strlen(a);
-  int y = strlen(b);
-  if (x <= n)
-    nt_array_ptr<int> c : count(n) =
-      dynamic_bounds_cast
-        <nt_array_ptr<int>>(a,count(n));
-    else return null;
-
-  if (x+y <= n) 
-      for (int i = 0; i < y; ++i)
-         * (c+x+i) = * (b+i);
-    else return null;
-  return a;
-}
-
-nt_array_ptr<int> strncat_c : count(0)
-  (nt_array_ptr<int> a : count(n),
-   nt_array_ptr<int> b : count(0), int n) {
-   int x = 0;
-   int y = strlen(b);
-   while (*x != '/0')
-    a++; x++;
-
-   for (int i = 0; i < y; ++i)
-     if (i + x < n)
-         * (a+i) = * (b+i);
-       else return null;
-   return a;
-}
-  \end{lstlisting}
-}
-}
-\caption{Implementations for strncat}
-\label{fig:strncat-ex}
-\end{figure}
-
-Fig.~\ref{fig:semantics} shows rules for the computation
-relation. We explain the rules in turn using the
-example of Fig.~\ref{fig:strncat-ex}.
-
-\myparagraph{Pointer accesses}
-The rules for dereference and assignment operations---\textsc{S-Def},
-\textsc{S-DefNull}, \textsc{S-DefNTArray}, and
-\textsc{S-AssignArr}---illustrate how the semantics checks bounds.
-Rule \textsc{S-DefNull} transitions attempted null-pointer
-dereferences to $\enull$, whereas \textsc{S-Def} dereferences a
-non-null (single) pointer.
-
-\textsc{S-AssignArr} assigns to an array as long as 0 (the point of
-dereference) is within the bounds designated by the pointer's annotation
-and strictly less than the upper bound. For the assignment rule,
-arrays are treated uniformly whether they are null-terminated or not
-($\kappa$ can be $\cdot$ or $nt$)---the semantics does not search past
-the current position for a null terminator. The program
-can widen the bounds as needed, if they currently precede the null
-terminator: \textsc{S-DefNTArray}, which dereferences an NT array
-pointer, allows an upper bound of $0$, since the program may read, but
-not write, the null terminator. A separate rule (not shown) handles
-normal arrays.
-
-\myparagraph{Casts}
-Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
-by \textsc{S-Cast}. In a type-correct program, such casts are
-confirmed safe by the type system. To evaluate a cast, the rule
-updates the type annotation on $n$. Before doing so, it must
-``evaluate'' any variables that occur in $\tau$ according to their
-bindings in $\varphi$. For example, if $\tau$ was
-$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
-produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
-
-Dynamic casts are accounted for by \textsc{S-DynCast} and
-\textsc{S-DynCastBound}. In a type-correct program, such casts are
-assumed correct by the type system, and later confirmed by the
-semantics. As such, a dynamic cast will cause a $\ebounds$ failure if
-the cast-to type is incompatible with the type of the target pointer,
-as per the $n_l' > n_l \vee n_h > n_h'$ condition in
-\textsc{S-DynCastBound}. An example use of dynamic casts is given on
-line~7 in Fig.~\ref{fig:strncat-ex}. The values of \code{x} and
-\code{n} might not be known statically, so the type system cannot
-confirm that \code{x <= n}; the dynamic cast assumes this inequality
-holds, but then checks it at run-time.
-
-\myparagraph{Binding and Function Calls}
-The semantics handles variable scopes using the special $\erettext$
-form. \textsc{S-Let} evaluates to a configuration whose stack
-is $\varphi$ extended with a binding for $x$, and whose expression is
-$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
-$\varphi(x)$; if it had no previous binding, $\varphi(x) =
-\bot$. Evaluation proceeds on $e$ until it becomes a literal
-$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
-binding (or $\bot$) in the new stack, and evaluates to
-$n\!:\!\tau$.
-
-Function calls are handled by \textsc{S-Fun}. Recall that array
-bounds in types may refer to in-scope variables; e.g., parameter
-\code{a}'s bound \code{count(n)} refers to parameter \code{n} on lines
-2-3 in Fig.~\ref{fig:strncat-ex}. A call to function $f$ causes $f$'s
-definition to be retrieved from $\Xi$, which maps function names to
-forms $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e$, where
-$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
-is the parameter list of variables and their types, and $e$ is the
-function body. The call is expanded into a \texttt{let} which binds
-parameter variables $\overline{x}$ to the actual arguments
-$\overline{n}$, but annotated with the parameter types
-$\overline{\tau}$ (this will be safe for type-correct programs). The
-function body $e$ is wrapped in a static cast
-$(\tau[\overline{n} / \overline{x}])$, which is the function's return
-type but with any parameter variables $\overline{x}$ appearing in that
-type substituted with the call's actual arguments $\overline{n}$. To
-see why this is needed, suppose that \code{strncat} in
-Fig.~\ref{fig:strncat-ex} is defined to return a
-\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
-perform a \code{strncat} function call as
-\code{x=strncat(a,b,10)}. After the evaluation of \code{strncat}, the
-function returns a value with type \code{nt_array_ptr<int>:count(10)}
-because we substitute bound variable \code{n} in the 
-defined return type with \code{10} from the function call's
-argument list.
-
-\myparagraph{Bounds Widening}
-Bounds widening occurs when branching on a dereference of a NT array
-pointer, or when performing $\estrlentext$. The latter is most useful
-when assigned to a local variable so that subsequent code can use the
-result, e.g., $e$ in $\elet{x}{\estrlen{y}}{e}$.  Lines 4 and 5 in
-Fig.~\ref{fig:strncat-ex} are examples. 
-The widened upper bound
-precipitated by $\estrlen{y}$ is extended beyond the lifetime of $x$, as long as $y$ is live.
-For example, \code{x}'s scope in line 4 is
-the whole function body in \code{strncat} because the lifetime of the pointer $y$ is in the function body.
-This is different from the \checkedc specification, 
-which only allows bound widening to happen within the scope of $x$, and restoring old bound values once $x$ dies.
-We allow widening to persist outside the scope at run-time as long as
-we are within the stack frame, and we show
-this does not necessarily require the use of fat pointers in
-Sec.~\ref{sec:compilation}.
-
-Rule \textsc{S-StrWiden} implements $\estrlentext$ widening. The
-predicate
-$\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i)
-= \evalue{n_i}{\tau_i} \wedge n_i \neq 0))$ aims to find a position
-$n+n_a$ in the NT array that stores a null character, where no
-character as indexes between $n$ and $n+n_a$ contains one. (This rule
-handles the case when $n_a > n_h$, the $n_a \le n_h$ case is handled
-by a normal $\estrlentext$ rule; see Appx.~\ref{fig:rem-semantics}.)
-
-Rule \textsc{S-IfNTT} performs bounds widening on $x$ when the
-dereference $\estar{x}$ is not at the null terminator, but the
-pointer's upper bound is 0 (i.e., it's at the end of its known
-range). $x$'s upper bound is incremented to 1, and this count persists
-as long as $x$ is live.  For example, \code{s}'s increment (lines
-5--8) is live until the return of the function in
-Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
-\code{s}'s upper bound is properly extended.
-
-\subsection{Typing}\label{sec:type-system}
-
-We now turn to the \lang type system.
-The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
-which states that in type environment $\Gamma$ (mapping variables to
-their types) and predicate environment $\Theta$ (mapping integer-typed
-variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-in mode $m$. Key rules for this judgment are given in
-Fig.~\ref{fig:type-system-1}. In the rules, $m \le m'$ uses
-the two-point lattice with $\umode < \cmode$. All remaining rules
-  are given in Appx.~\ref{sec:literal-pointer-typing}~and~\ref{rem-type}.
-
-\begin{DIFnomarkup}
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[T-Cast]
-              {m = \cmode \Rightarrow \tau \neq\tptr{\tau''}{\cmode} \text{ for any $\tau''$} \\\\
-                \Gamma;\Theta \vdash_m e : \tau'}
-              {\Gamma;\Theta \vdash_m \ecast{\tau}{e} : \tau}
-
-     \inferrule[T-CastCheckedPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq \tptr{\tau}{\cmode}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\cmode}}{e} : \tptr{\tau}{\cmode}}
-
-      \inferrule[T-DynCast]
-                {\Gamma;\Theta \vdash_m e : \tptr{\tallarrayb{\bvar'}{\tau}}{m}}
-                {\Gamma;\Theta \vdash_m \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e} : \tptr{\tallarrayb{\bvar}{\tau}}{m}}
-
-      \inferrule[T-Str]
-                {\Gamma; \Theta \vdash_m e :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} 
-                }
-                {\Gamma;\Theta \vdash_m \estrlen{e} : \tint}
-
-      \inferrule[T-LetStr]
-                {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\\\
-                 \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}];\Theta[x\mapsto \texttt{ge}\_0] \vdash_m e : \tau}
-                {\Gamma;\Theta \vdash_m \elet{x}{\estrlen{y}}{e} : \tau}
-
-                \inferrule[T-Ret]
-    {\Gamma(x)\neq \bot \\
-          \Gamma;\Theta \vdash_m e : \tau}
-    {\Gamma;\Theta \vdash_m \eret{x}{\mu}{e} : \tau}
-
-      \inferrule[T-If ]
-        {\Gamma; \Theta \vdash_m e : \tau \\\\ \Gamma; \Theta \vdash_m e_1 : \tau_1 \\\\ \Gamma; \Theta \vdash_m e_2 : \tau_2}
-        {\Gamma; \Theta \vdash_m \eif{e}{e_1}{e_2} : \tau_1 \sqcup \tau_2}
-
-      \inferrule [T-IfNT]
-                { \Gamma; \Theta \vdash_m x : \tptr{\tntarrayb{(b_l,0)}{\tau}}{c} \\\\
-                   \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}]; \Theta \vdash_m e_1 : \tau_1\\\\
-                  \Gamma;\Theta \vdash_m e_2 : \tau_2 }
-                {\Gamma;\Theta \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_1 \sqcup \tau_2}
-
-   \inferrule[T-Let]
-    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
-        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
-           \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
-             }
-    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
-
-    \inferrule[T-DefArr]
-              {m \leq m' \\\\ \Gamma;\Theta \vdash_{m} e : \tptr{\tallarrayb{\bvar}{\tau}}{m'}}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-
-    \inferrule[T-AssignArr]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\\\
-                \Gamma; \Theta \vdash_m e_2 : \tau' \\
-                \tau'\sqsubseteq \tau\\
-                m \leq m'}
-              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
-
-\inferrule[T-Fun]
-    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\\\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
-             \overline{\tau'} \sqsubseteq 
-               \overline{\tau}[\overline{e} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m f(\overline{e}) : \tau[\overline{e} / \overline{x}]}
-  \end{mathpar}
-}
-\caption{Selected Type Rules}
-\label{fig:type-system-1}
-\end{figure*}
-\end{DIFnomarkup}
-
-\myparagraph{Pointer Access}
-Rules \textsc{T-DefArr} and \textsc{T-AssignArr} typecheck array
-dereference and assignment operations resp. returning the type of
-pointed-to objects; rules for pointers to single objects are
-similar. The condition $m\le m'$ ensures that checked pointers cannot
-be dereferenced in $\eunchecked{e}$ blocks; the type rule for
-$\eunchecked{e}$  (not shown) sets $m = \umode$ when checking $e$.
-The rules do not attempt to reason whether the access is in bounds;
-this check is deferred to the semantics.
-
-\myparagraph{Casting and Subtyping}
-Rule \textsc{T-Cast} rule forbids casting to checked pointers when in checked
-regions (when $m = \cmode$), but $\tau$ is unrestricted when
-$m = \umode$. The \textsc{T-CastCheckedPtr} rule
-permits casting from an expression of type $\tau'$ to a checked pointer when
-$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype}; the many
-rules ensure the relation is transitive. Most of the rules handle
-casting between array pointer types; the second rule 
-$0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
-\tarrayptr{b_l}{b_h}{\tau}{m}$ permits treating a singleton
-pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
-
-Since bounds expressions may
-contain variables, determining assumptions like $b_l \leq b_l'$
-requires reasoning about those variables' possible values. The type
-system uses $\Theta$ to make such reasoning more precise.\footnote{So,
-  technically, the subtyping relation $\sqsubseteq$ and the bounds
-  ordering relation $\leq$ are parameterized by $\Theta$; this fact is
-  implicit to avoid clutter.} $\Theta$ is a map from variables $x$ to
-predicates $P$, which have the form $P ::= \top \;|\; \texttt{ge}\_0$.
-If $\Theta$ maps $x$ to $\top$, that means that the variable can
-possibly be any value; $\texttt{ge}\_0$ means that $x \ge 0$. We will
-see how $\Theta$ gets populated and give a detailed example of
-subtyping below.\footnote{As it turns out, the subtyping relation is
-  also parameterized by $\varphi$, which is needed when type checking
-  intermediate results to prove type preservation; source programs
-  would always have $\varphi = \emptyset$. Details are in
-Appendix~\ref{app:le}.}
-
-Rule \textsc{T-DynCast} typechecks dynamic casting operations, which
-apply to array pointer types only. The cast is accepted by the type
-system, as its legality will be checked by the semantics.
-
-\begin{DIFnomarkup}
-\begin{figure}
-{\small
-\begin{center}
-  $\begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-    &&\multicolumn{1}{r}{\tau} &\sqsubseteq& \tau\\[0.6em]
-    0\le b_l \wedge b_h \le 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.6em]
-    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
-    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
-b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
-    \end{array}
-  $
-\end{center}
-}
-  \caption{Subtyping Relation}
-  \label{fig:checkc-subtype}
-\end{figure}
-\end{DIFnomarkup}
-
-\myparagraph{Bounds Widening}
-The bounds of NT array pointers may be widened at conditionals, and
-due to calls to $\estrlentext$.  Rule \textsc{T-If} handles normal
-branching operations; rule \textsc{T-IfNT} is specialized to the case
-of branching on $\estar{x}$ when $x$ is a NT array pointer whose upper
-bound is 0. In this case, true-branch $e_1$ is checked with $x$'s type
-updated so that its upper bound is incremented by 1; the else-branch
-$e_2$ is typechecked under the existing assumptions. For both rules,
-the resulting type is the join of the types of the two branches
-(according to subtyping). This is important for the situation when $x$
-itself is part of the result, since $x$ will have
-different types in the two branches.
-
-Rule \textsc{T-Str} handles the case for when $\estrlentext(y)$ does
-not appear in a let binding. Rule \textsc{T-LetStr} handles the case
-when it does, and performs bounds widening. The result of the call is
-stored in variable $x$, and the type of $y$ is updated in $\Gamma$ when
-checking the let-body $e$ to indicate that $x$ is $y$'s upper
-bound. Notice that the lower bound $b_l$ is unaffected by the call to
-$\estrlentext(y)$; this is sound because we know that $\estrlentext$
-will always return a result $n$ such that $n \ge b_h$, the current
-view of $x$'s upper bound.
-The type rule tracks $\estrlentext$'s widened bounds
-within the scope of $x$, while the bound-widening effect in the semantics applies to the lifetime of $y$.
-Our type preservation theorem in Sec.~\ref{sec:theorem}
-shows that our type system is a sound model of the \CoreChkC semantics,
-and we discuss how we guarantee that the behavior of our compiler formalization and the semantics matches in Sec.~\ref{sec:compilation}.
-
-This rule also extends $\Theta$ when checking $e$, adding a predicate
-indicating that $x\ge 0$. To see how this information is used,
-consider this example.  The \code{return} on line 16 of
-Fig.~\ref{fig:strncat-ex} has an implicit static cast from the
-returned expression to the declared function type (see rule
-\textsc{T-Fun}, described below). In type checking the $\estrlentext$
-on line 4, we insert a predicate in $\Theta$ showing
-\code{n}$\;\ge 0$.  The static cast on line 16 is valid according ot
-the last line in Fig.~\ref{fig:checkc-subtype}:
-{\small
-$$
-\tallarrayptr{0}{n}{\tau}{c} \sqsubseteq \tallarrayptr{0}{0}{\tau}{c}
-$$
-}
-because $0 \leq 0$ and $0 \leq n$, where the latter holds since
-$\Theta$ proves $n \ge 0$.  Without $\Theta$, we would need a
-dynamic cast.
-
-In our formal presentation, $\Theta$ is quite simple and is just meant
-to illustrate how static information can be used to avoid dynamic
-checks; it is easy to imagine richer environments of facts that can be
-leveraged by, say, an SMT solver as part of the subtyping
-check \cite{LiquidHaskell,Refinementlh}
-
-\myparagraph{Dependent Functions and Let Bindings}
-Rule \textsc{T-Fun} is the standard dependent function call rule. It
-looks up the definition of the function in the function environment
-$\Xi$, typechecks the actual arguments $\overline{e}$ which have
-types $\overline{\tau'}$, and then confirms that each of these types
-is a subtype of the declared type of $f$'s corresponding parameter. Because
-functions have dependent types, we substitute each parameter $e_i$ for
-its corresponding parameter $x_i$ in both the parameter types and the
-return type. Consider the \code{strncat} function in
-Fig.~\ref{fig:strncat-ex}; its parameter type for \code{a} 
-depends on \code{n}. The \textsc{T-Fun} rule will substitute 
-\code{n} with the argument at a call-site.
-
-Rule \textsc{T-Let} types a $\elettext$ expression, which also admits
-type dependency. In particular, the result of evaluating a $\elettext$
-may have a type that refers to one of its bound variables (e.g., if
-the result is a checked pointer with a variable-defined bound); if so,
-we must substitute away this variable once it goes out of scope. Note
-that we restrict the expression $e_1$ to syntactically match the
-structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
-
-Rule \textsc{T-Ret} types a $\erettext$ expression, which does not
-appear in source programs but is introduced by the semantics when
-evaluating a let binding (rule \textsc{S-Let} in
-Fig.~\ref{fig:semantics}); this rule is needed for the preservation
-proof. After the evaluation of a let binding a variable $x$ concludes,
-we need to restore any prior binding of $x$, which is either
-$\bot$ (meaning that there is no $x$ originally) or some value
-$\evalue{n}{\tau}$.
-
-\subsection{Type Soundness and Blame}\label{sec:theorem}
-
-In this subsection, we focus on our main metatheoretic results about
-\lang: type soundness (progress and preservation) and blame. The type
-soundness theorems rely on a notion of heap and stack
-\emph{well-formedness}:
-
-\smallskip
-\begin{defi}[Heap Well-formedness]
-A heap $\heap$ is well-formed, iff (i) the null position
-($0$) is not defined in $\heap$, and (ii) every type annotation in it
-contains no free variables.
-\end{defi}
-\smallskip
-\begin{defi}[Stack Well-formedness]
-A stack snapshot $\varphi$ is well-formed, iff every type annotation in
-it contains no free variables.
-\end{defi}
-\smallskip
-
-Moreover, as a program evaluates its expression may contain literals
-$n\!:\!\tau$ where $\tau$ is a pointer type, i.e., $n$ is an index in
-$\heap$ (perhaps because $n$ was chosen by $\mathtt{malloc}$). The
-normal typechecking judgment for $e$ is implicitly parameterized by
-$\heap$, and the rules for typechecking literals confirm that
-pointed-to heap cells are compatible with (subtypes of) the pointer's
-type annotation; in turn this check may precipitate checking the type
-consistency of the heap itself. We follow the same approach as
-\citet{ruef18checkedc-incr}, and show the rules in
-Appendix~\ref{sec:literal-pointer-typing}.
-
-Progress now states that terms that don't reduce are either values or their
-mode is unchecked:
-
-\smallskip
-\begin{thm}[Type Progress Theorem]\label{thm:progress}
-For any \checkedc program $e$ and heap $\heap$, if $e$ and $\heap$ are
-well-formed, and $\emptyset;\emptyset\vdash_m e : \tau$, then $e$ is
-a value ($\evalue{n}{\tau}$), \code{unchecked} ($m=\umode$), or
-there exists $\varphi'$ $\heap'$ $e'$, such that $(\emptyset,\heap,e)
-\longrightarrow_m (\varphi',\heap',e')$.
-\end{thm}
-{\em Proof:} By induction on the typing derivation.
-
-\smallskip
-
-For preservation, we also need to introduce a notion of
-\emph{consistency}, relating heap environments before and after a
-reduction step, and type environments, predicate sets, and stack
-snapshots together.
-
-\smallskip
-\begin{defi}[Type-Stack Consistency]
-A type environment $\Gamma$, variable predicate set $\Theta$, and stack snapshot $\varphi$ are consistent, iff every variable defined in $\Theta$ is defined in $\Gamma$, and for every variable $x$, $\Gamma(x) = \tau$ implies that $\varphi(x)$ is defined and there exists $n$ and $\tau'$, such that $\varphi(x)=\evalue{n}{\tau'}$ and $\tau' \sqsubseteq \tau$.
-\end{defi}
-\smallskip
-\begin{defi}[Heap Consistency]
-A heap $\heap'$ is consistent with $\heap$ iff every address defined
-in $\heap$ is defined in $\heap'$.
-\end{defi}
-\smallskip
-
-\noindent
-Armed with the definitions of consistency, we can now prove
-preservation, which states that a reduction step preserves both the
-type of the expression being reduced, as well as well-formedness and
-consistency of environments:
-
-\smallskip
-\begin{thm}[Type Preservation Theorem]
-For any \checkedc program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, variable predicate set $\Theta$,
-and a type $\tau$, that are all are well-formed, if $\Gamma, \Theta$,
-and $\varphi$ are consistent, $e$ is well typed
-$\Gamma;\Theta\vdash_{\cmode} e : \tau$, and if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\cmode} (\varphi',\heap',e')$, then there exists
-$\Gamma'$, $\Theta'$ and $\tau'$, such that $\Gamma'$, $\Theta'$,
-$\varphi'$, $\heap'$ and $e'$ are well-formed, $\Gamma'$, $\Theta'$
-and $\varphi'$ are consistent, $\heap'$ is consistent with $\heap$,
-$\Gamma';\Theta' \vdash_{\cmode} e: \tau'$, and $\tau'\sqsubseteq
-\tau$.
-\end{thm}
-{\em Proof:} By induction on the typing derivation.
-\smallskip
-
-Using type soundness we can prove our main result, {\em blame}, which
-states that if there is any spatial memory safety violation is
-triggered, it must necessarily come from the unchecked region.
-
-\smallskip
-\begin{thm}\label{thm:blame}
-[The Blame Theorem] For any \checkedc program $e$, heap
-$\heap$, type $\tau$, if $\heap$ and $e$ are well-formed,
-$\emptyset;\emptyset\vdash_{\cmode} e : \tau$, and if there exists
-$\varphi'$, $\heap'$, a failure result $r$, and $m$, such that
-$(\varphi,\heap,e) \longrightarrow_m^* (\varphi',\heap',r)$, then
-there exist $E$ and $e_a$, such that $e' = E[e_a]$, and $\mode(E) = \umode$.
-\end{thm}
-{\em Proof:} By induction on the number of steps of the \checkedc
-evaluation ($\longrightarrow_m^*$), using progress and preservation to
-maintain the invariance of the assumptions.
-
-These proofs have been carried out in a Coq development.
-
-% \ignore{
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \hypo{\evalue{n}{\tau} \in \defscope}
-%     \infer1[T-VConst]{\Gamma;\defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%   \end{prooftree}
-%   \qquad
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma;\defscope \vdash_m e : \tau \\
-%         \Gamma;\defscope \vdash_m e_1 : \tau_1 
-%       \end{matrix}
-%     }
-%     \hypo{
-%       \begin{matrix}
-% \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2 \\
-%             \Gamma;\defscope \vdash_m e_2 : \tau_2
-%           \end{matrix}
-%         }
-%     \infer2[T-If]{\Gamma;\defscope \vdash_m \eif{e}{e_1}{e_2} : \tau_3}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma;\defscope \vdash_m x : \tptr[c]{(\tntarray{l}{0}{\tau})}, l \leq 0 \\
-%         \Gamma, x : \tptr[c]{(\tntarray{l}{1}{\tau})};\defscope \vdash_m e_1 : \tau_1 \\
-%         \Gamma;\defscope \vdash_m e_2 : \tau_2 \\
-%         \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2
-%       \end{matrix}
-%     }
-%     \infer1[T-IfNT]{\Gamma;\defscope \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_3}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         F(f) = \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e  \\
-%         \Gamma; \defscope \vdash_m \tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}] ~~~ 0 \leq i \leq j\\
-%         \Gamma; \defscope \vdash_m e_i : \tau_i' ~~~ 0 \leq i \leq j\\
-%         \Gamma; \defscope \vdash_m \subtype{\tau_i'}{\tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}]}  ~~~ 0 \leq i < j
-%       \end{matrix}
-%     }
-%     \infer1[T-VCall]{\Gamma; \defscope \vdash_m f(\overline{e}) : \tau_j[x_0,\ldots,x_{j-1} \mapsto e_0,\ldots,e_{j-1}]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%           \Gamma; \defscope \vdash_m e_1 : \tau_1
-%       }
-%       \hypo{        \Gamma, x = e_1 : \tau_1; \defscope \vdash_m e_2 : \tau_2
-% }
-%     \infer2[T-Let]{\Gamma; \defscope \vdash_m \elet{x}{e_1}{e_2} : \tau[x \mapsto e_1]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m \estrlen{y} : \tau_1\\
-%         \Gamma; \defscope \vdash_m y : \tptr[c]{(\tntarray{le}{\_}{\tau_3})}\\
-%         \Gamma, x = \estrlen{y} : \tau_1, \\ y : \tptr[c]{(\tntarray{le}{x}{\tau_3})} ;  \defscope  \vdash_m e_2 : \tau_2
-%         \end{matrix}
-%         }
-%         \infer1[T-LetStr]{\Gamma; \defscope \vdash_m \elet{x}{\estrlen{y}}{e_2} : \tau[x \mapsto e_1]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%         \Gamma; \defscope \vdash_m e : \tptr[c]{(\tntarray{le}{he}{\tau})}
-%       }
-%         \infer1[T-Str]{\Gamma; \defscope \vdash_m \estrlen{y} : \tint}
-%       \end{prooftree} \\ \\
-
-%       \begin{prooftree}
-%         \hypo{
-%           \begin{matrix}
-%             \tau = \tint \vee \tau = \tptr[u]{\omega}~\vee  n=0~ \vee \\ \tau = \tptr[c]{(\tarray{0}{0}{\tau'})} \vee \\
-%             \tau = \tptr[c]{(\tntarray{0}{0}{\tau'})}
-%         \end{matrix}
-%       }
-%       \infer1[T-Base]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%     \end{prooftree} \\ \\
-
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%         \tau = \tptr[c]{\omega} \\
-%         \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\
-%         \Gamma;\defscope, \evalue{n}{\tau}  \vdash_m {H(n+k)} : {\tau_k} ~~~ 0 \leq k < j
-%       \end{matrix}
-%       }
-%       \infer1[T-PtrC]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%     \end{prooftree} \\ \\
-
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%             \Gamma; \defscope \vdash_m e :  {\tptr[m]{\tstruct{T}}} \\
-% D(T) = ...; \tau_f~f; ...
-%         \end{matrix}
-%       }
-%       \infer1[T-Amper]{\Gamma; \defscope \vdash_m \eamper{e}{f} : \tptr[m]{\tau_f}}
-%     \end{prooftree} \\ \\
-
-%     \todo[inline]{YL: how to express le - n as a metafunction?}
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}\\
-%         \Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tint \\
-%         le' = le - n, he' = he - n
-%       \end{matrix}
-%     }
-%       \infer1[T-BinopInd]{\Gamma; \defscope \vdash_m (e_1 \plus \evalue{n}{\tau} ) : \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})}}
-%     \end{prooftree}
-% \caption{Typing}
-% \label{fig:typing}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \todo[inline]{no sizeof check because we don't know statically whether the allocation would be null}
-%   \begin{prooftree}
-%     \hypo{
-%       \omega = \tgarray{\alpha}{le}{he}{\tau} \Rightarrow le = 0
-%     }
-%     \infer1[T-Malloc]{\Gamma; \defscope \vdash_m \emalloc{\omega} : \tptr[c]{\omega}}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \Gamma; \defscope \vdash_u e : \tau
-%     }
-%     \infer1[T-Unchecked]{\Gamma; \defscope \vdash_m \eunchecked{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \todo[inline]{any constraints on m'?}
-%   \begin{prooftree}
-%     \hypo{\Gamma; \defscope \vdash_m e :  \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})} }
-%     \infer1[T-DynCast]{\Gamma; \defscope \vdash_m \edyncast{\tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}}{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{\Gamma; \defscope \vdash_m e : \tau'}
-%     \infer1[T-Cast]{\Gamma; \defscope \vdash_m \ecast{\tau}{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\tgarray{\alpha}{le}{he}{\tau}} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tint \\
-%         m' = u \Rightarrow m = u \\
-%       \end{matrix}
-%     }
-%     \infer1[T-Index]{\Gamma; \defscope \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tau \\
-%         m' = u \Rightarrow m = u \\
-%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-%       \end{matrix}
-%     }
-%     \infer1[T-Assign]{\Gamma; \defscope \vdash_m \eassign{e_1}{e_2} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tint \\
-%         \Gamma; \defscope \vdash_m e_3 : \tau \\
-%         m' = u \Rightarrow m = u \\
-%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-%       \end{matrix}
-%     }
-%     \infer1[T-IndAssign]{\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
-%   \end{prooftree}
-
-% \caption{Typing Cont.}
-% \label{fig:typing2}
-% \end{figure}
-% \ignore{
-
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j} \\
-%         \Gamma, x_0 = \kw{none} : \tau_0, \ldots, x_{j-1} = \kw{none} : \tau_{j-1}; \defscope \vdash e : \tau_{r}
-%       \end{matrix}
-%     }
-%     \infer1[WF-Fun]{\Gamma;\defscope \vdash \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \infer0[WF-Nil]{\Gamma \vdash \circ}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma \vdash \tau_0 \\
-%         \Gamma, x_0 = \kw{none} :\tau_0 \vdash x_1:\tau_{1}, \ldots, x_{j}:\tau_{j}
-%       \end{matrix}
-%     }
-%     \infer1[WF-Cons]{\Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \infer0[WF-Int]{\Gamma \vdash \tint}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{   \Gamma \vdash le }
-%      \hypo{   \Gamma \vdash he }
-%       \hypo{  \Gamma \vdash \tau}
-
-%     \infer3[WF-Array]{\Gamma \vdash \tptr[m]{(\tgarray{\alpha}{le}{he}{\tau})}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{T \in dom(D)}
-%     \infer1[WF-Struct]{\Gamma \vdash \tptr[m]{\tstruct{T}}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{\Gamma \vdash \tau}
-%     \infer1[WF-Ptr]{\Gamma \vdash \tptr[m]{\tau}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \infer0[WFB-Int]{\Gamma \vdash i}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{x = e? : \tint \in \Gamma}
-%     \infer1[WFB-Var]{\Gamma \vdash x \plus i}
-%   \end{prooftree} \\\\
-
-%   \caption{Well-Formedness}
-%   \label{fig:wf}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \infer0[Sub-Nt]{\subtype{\tptr[c]{(\tntarray{le}{he}{\tau})}}{\tptr[c]{(\tarray{le}{he}{\tau})}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \infer0[Sub-Refl]{\subtype{\tau}{\tau}}
-%   \end{prooftree} \\ \\
-
-%   \todo[inline]{what about ntarrays?}
-%   \begin{prooftree}
-%     \infer0[Sub-Ptr]{\subtype{\tptr[c]{\tau}}{\tptr{\tarray{0}{1}{\tau}}}}
-%   \end{prooftree} \\ \\
-
-%   \todo[inline]{the subtyping relation is not anti-symmetric..}
-%   \begin{prooftree}
-%     \infer0[Sub-Arr]{\subtype{\tptr[c]{\tarray{0}{1}{\tau}}}{\tptr{\tau}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{\rboundle{le_0}{le_1}}
-%     \hypo{\rboundle{he_1}{he_0}}
-%     \infer2[Sub-Subsume]{\subtype{\tptr{(\tgarray{\alpha}{le}{he}{\tau})}{\cmode}}{ \tptr{(\tgarray{\alpha}{ le_1} {he_1} {\tau})}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-% \hypo{D(T) = \tau_f~f; ...}
-%     \hypo{\subtype{\tau_f}{\tau}}
-%     \infer2{\subtype{\tptr[c]{\tstruct{T}}}{\tau}}
-%   \end{prooftree}
-%   \caption{Subtyping}
-%   \label{fig:sub}
-% \end{figure}
-% }
-% }
-
-% \ignore{
-
-% \begin{figure*}[t]
-%   \begin{lstlisting}
-% int foo(nt_array_ptr<char> p : count(0)) {
-%   if (* p) {
-%     dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
-%   }
-%   dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
-%   return 0;
-% }
-%   \end{lstlisting}
-% \caption{The example where clang Checked C fails at run-time}
-% \label{fig:clangbad1}
-% \end{figure*}
-
-% \begin{figure*}[t]
-%   \begin{lstlisting}
-% /* nt_array_ptr<char> p : bounds(p,p) */
-% size_t cnt = 0;
-
-% while(*(p+cnt)) {
-%   ++cnt;
-% }
-% dyn_bounds_cast<nt_array_ptr<char>>(p, count(cnt));
-%   \end{lstlisting}
-% \caption{A useful program that the Checked C spec doesn't allow at run-time}
-% \label{fig:clangbad2}
-% \end{figure*}
-% }
-
-% \ignore{
-% The Clang CheckedC implementation uses statically determined
-% bounds to insert run-time checks. In Fig.~\ref{fig:clangbad1}, the
-% \code{dyn_bounds_cast} at line 3 will always succeed, because the
-% compiler knows that within the scope of then branch, the pointer
-% \code{p} must have at least one element. The same cast at line 5,
-% however, will always fail, since there is no way to tell statically
-% whether the program has entered the then branch before. The compiler
-% will check whether the \code{count(1)} bounds specification is
-% contained within the earlier \code{count(0)} specification, resulting
-% in a run-time failure even when we pass in a non-empty string.
-
-% Our formalization diverges from this run-time behavior and instead keeps
-% track of the bounds on the stack. After entering the then branch, we
-% increment the upper bound for \code{p}, effectively making the
-% updated bounds information available even after we exit the if
-% statement. The cast at line 5 will be checking the new bounds against
-% the incremented bounds for non-empty strings.
-
-% Fig.~\ref{fig:clangbad2} gives a more practical example of why keeping
-% track of the bounds on the stack is useful. The program snippet
-% implements the functionality of the \code{strlen} function using a
-% while loop and a \code{cnt} variable. Even though the type system is
-% unable to reason about the while loop, as long as the runtime system
-% updates the bounds in-place, the user can apply a
-% \code{dyn_bounds_cast} to soundly recover the more precise bounds
-% information.
-% }
-
- \section{Compilation}\label{sec:compilation}
-
-The semantics of \lang uses annotations on pointer literals in
-order to keep track of array bounds information, which is checked at
-dereferences and changed during widening. However, in the real
-implementation of \checkedc, these annotations
-are not present---pointers are represented as a single
-machine word with no extra metadata. We show how the
-annotations can be safely erased: using static information a compiler
-can insert code to manage and check bounds metadata, with no loss of
-expressiveness.
-
-This section sketches our compilation algorithm that converts from
-\lang to \elang, an untyped language without metadata annotations.
-Compilation is defined by extending \lang's
-typing judgment thusly:
-\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
-There is now a \elang output $\dot e$ and an input $\rho$, which maps
-each \code{nt_array_ptr} variable $p$ to a pair of \emph{ghost
-  variables} that keep $p$'s up-to-date upper and lower bounds; these
-may differ from the bounds in $p$'s type due to bounds
-widening.\footnote{Since lower bounds are never widened, the
-  lower-bound ghost variable is unnecessary; we include it for uniformity.}
-When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
-complete judgment, implicitly assuming that $e$ is a well-typed and closed
-term.
-
-We formalize rules for this judgment in PLT Redex~\cite{pltredex},
-following and extending our Coq development for \lang. To give
-confidence that compilation is correct, we use Redex's property-based
-random testing support to show that compiled-to $\dot e $ simulates
-$e$, for all $e$.
-
-\subsection{Approach}
-
-Due to space constraints, we explain the rules for compilation by
-example; the complete rules are given in Appendix~\ref{appx:comp1}. 
-Each rule performs up to three tasks: (a) conversion of $e$ to
-A-normal form; (b) insertion of dynamic checks; and (c) insertion of
-bounds widening expressions.
-A-normal form conversion is straightforward: compound expressions are
-handled by storing results of subexpressions into temporary variables,
-as in the following example.
-
-{\vspace*{-0.5em}
-{\small
-\begin{center}
-$
-\begin{array}{l}
-$\code{y=(x+1)+(6+1);}$
-\;
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\;
-\begin{array}{l}
-$\code{a=x+1;}$\\
-$\code{b=6+1;}$\\
-$\code{y=a+b;}$\\
-\end{array}
-\end{array}
-$
-\end{center}
-}
-}
-
-This simplifies the management of effects from subexpressions. The
-next two steps of compilation are more interesting.
-
-\begin{figure}[t!]
-  \begin{small}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-/* nt_array_ptr<int> p : count(p,p) */
-/* $\rho$(p) = p_lo,p_hi */
-{
-  int x = strlen(p);
-  if (x > 1) putchar(*(p+1));
-}
-\end{lstlisting}
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\begin{lstlisting}[xleftmargin=4 mm]
-{
-  assert(p_lo <= p && p <= p_hi);  //bound check
-  assert(p != NULL);               //null check
-  int x = strlen(p);
-  int *p_hi_new = p + x;
-  p_hi = max(p_hi, p_hi_new);
-  if (x > 1) {
-    assert(p != NULL);             //null check p
-    int *p_1 = p + 1;
-    assert(p_lo <= p_1 && p_1 <= p_hi); // p+1
-    assert(p_1 != NULL);          //null check p+1
-    putchar(*p_1);
-  }
-}
-\end{lstlisting}
-\end{small}
-\caption{Compilation example for check insertions}
-\label{fig:compilationexample}
-\end{figure}
-
-\begin{figure}[t!]
-  \begin{small}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-int deref_array (int n, 
-      nt_array_ptr<int> p : bounds(p, p+n)) {
-  /* $\rho$(p) = p_lo,p_hi */
-  if (*p) return *(p+1);
-  return 0;
-}
-...
-// nt_array_ptr<int> p0 : bounds(p0, p0+5)
-deref_array(5, p0);
-    \end{lstlisting}
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\begin{lstlisting}[xleftmargin=4 mm]
-int deref_array(int n, int *p) {
-  int *p_lo = p;
-  int *p_hi = p + n;
-  /* runtime checks */
-  assert(p_lo <= p && p <= p_hi);
-  assert(p != NULL);
-  int p_derefed = *p;
-  if (p_derefed != '\0') {
-    /* widening */
-    if (p_hi == p) {
-      ++p_hi;
-    }
-    int *p0 = p + 1;
-    assert(p_lo <= p0 && p0 <= p_hi);
-    assert(p0 != NULL);
-    return *p0;
-  }
-  return 0;
-...
-//int *p0, set_bounds(p0) = p_lo, p_hi
-deref_array(5, p0);
-    \end{lstlisting}
-\end{small}
-    \caption{Compilation example for dependent functions}
-\label{fig:compilationexample1}
-\end{figure}
-
-During compilation, $\Gamma$ tracks the lower and upper bound
-associated with every pointer variable according to its type. At each
-declaration of a \code{nt_array_ptr} variable $p$, the
-compiler allocates two \emph{ghost variables}, 
-stored in $\rho(p)$; these are initialized to $p$'s declared bounds
-and will be updated during bounds widening.\footnote{Ghost variables
-  are not used for \code{array_ptr} types (the bounds expressions are)
-  since they are not subject to bounds widening.} 
-Fig.~\ref{fig:compilationexample} shows how an invocation of
-\code{strlen} on a null-terminated string is compiled into C
-code. Each dereference of a checked pointer requires a null check
-(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
-compiler makes explicit: Line~$3$ of the generated code has the null
-check on pointer $p$, and similar checks happen at line~$8$ and
-line~$11$.
-Dereferences also require bounds checks: line~$2$ checks \code{p} is
-in bounds before computing \code{strlen(p)}, while line~$10$ does
-likewise before computing \code{*(p+1)}.
-
-For \code{strlen(p)} and conditionals \code{if(*p)}, the \lang
-semantics allows the upper bound of \code{p} to be extended.
-The compiler explicitly inserts statements to do so on \code{p}'s
-ghost bound variables. For example,
-Fig.~\ref{fig:compilationexample}~line~$6$ widens \code{p}'s upper
-bound if \code{strlen}'s result is larger than the existing
-bound. 
-Lines 7--12 of the generated code in
-Fig.~\ref{fig:compilationexample1} show how bounds are 
-widened when compiling expression \code{if(*p)}. If we find that the
-current \code{p} address is equal to the upper bound (line 10),
-and \code{p}'s content is not null (line 8), we then increase the
-upper bound by $1$ (line 11).
-Fig.~\ref{fig:compilationexample1} also shows a dependent function call.
-Notice that the bounds for the array pointer \code{p} are not passed as
-arguments. Instead, they are initialized according to \code{p}'s
-type---see line~3 of the original \lang program at the top of the figure.
-Line~$2$ of the generated code
-sets the lower bound  to \code{p} and line~$3$ sets the
-upper bound to \code{p+n}.
-
-\subsection{Comparison with Checked C Specification}
-\label{sec:disc}
-
-The use of ghost variables for bounds widening is a key novelty of our
-compilation approach, and adds more precision to bounds checking at runtime
-compared to the official specification and current implementation of
-\checkedc~\cite[5.1.2, pg 85]{checkedc}.  For example, the
-\code{strncat} example of Fig.~\ref{fig:strncat-ex} compiles with the
-current Clang \checkedc compiler but will fail with a runtime
-error. The statement \code{int x = strlen(a)} at line 4 changes the
-upper bound of \code{a} to \code{x}, which can be smaller than
-\code{n}, the capacity of the array pointer \code{a}. The assignment
-at line 14 will always fail if the index \code{x + i} is checked
-against the statically determined upper bound
-\code{x}. This forces us to inline the definition of \code{strlen} as
-in \code{strncat_c} to avoid runtime errors when running code compiled with the
-Clang \checkedc compiler.  Likewise, if we were to add another dereference to
-$p$ after line 6 in the original code at the top of
-Fig.~\ref{fig:compilationexample}, the Clang \checkedc
-compiler would 
-check \code{p} against its original bounds \code{(p,p)} since the
-updated upper bound \code{p+x} cannot be retained with \code{x} out of
-the scope. In the presence of ghost variables,
-these bounds have been widened by the assignment in line $5$ (assuming
-the null-terminator was not the first element of the string) and
-remain available in the entire stack frame, and
-therefore the check will succeed. In contrast, in the actual
-implementation of Checked C, the scope of the widening is limited to
-the scope of the conditional at both runtime and compile time, which means that the inserted dynamic
-check would fail.
-To make it match the specification, our compilation definition could
-rely only on the type-based bounds expressions available in $\Gamma$
-for checking, and eschew ghost variables. However, doing so would
-force us to weaken the simulation theorem, reduce expressiveness,
-and/or force the semantics to be more awkward. We plan to work with
-the \checkedc team to implement our approach in a future revision.
-
-\subsection{Metatheory}
-\label{sec:meta}
-
-While designing our Coq model of \lang, we also designed a model in
-PLT Redex.\footnote{The two models, in Redex and Coq, are
-  equivalent, with the only difference being in the representation of
-  stacks: as we saw, the Coq model uses an explicit map for
-  representing stacks to ease the effort of theorem proving; on the
-  other hand, the Redex model uses \texttt{let} bindings to simulate a
-  stack, which removes the need to account for the stack during random
-  generation of terms.}  Redex~\cite{pltredex} is a semantic
-engineering framework implemented in Racket, which allows for
-concisely specifying semantics and typing rules. We
-formalize the simulation theorem in this model, and then attempt to
-falsify it via Redex's support for random testing. We ultimately plan
-to prove simulation in the Coq model.
-
-Turning to the simulation theorem: We first introduce notation
-used to specify the theorem. We use the notation $\gg$ to
-indicate the \emph{erasure} of stack and heap---the rhs is the same as
-the lhs but with type annotations removed:
-\begin{equation*}
-  \begin{split}
-    \heap  \gg & \dot \heap \\
-    \varphi \gg & \dot \varphi
-  \end{split}
-\end{equation*}
-In addition, we write $(\varphi,\heap,e) \gg (\dot \varphi, \dot \heap,
-\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
-and $e \gg \dot e$ respectively.
-
-We use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
-reduction relation of $\elang$. Unlike the $\lang$, the semantics of
-$\elang$ does not distinguish checked and unchecked regions.
-
-Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
-the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
-that is similar to the one by
-~\citet{merigoux2021catala}. An ordinary simulation property would
-replace the middle and bottom parts of the figure with the
-following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
-  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
-Instead, we relate two erased configurations using the relation $\sim$,
-which only requires that the two configurations will eventually reduce
-to the same state. We formulate our simulation theorem differently
-because the standard simulation theorem imposes a very strong
-syntactic restriction to the compilation strategy. Very often, $(\dot
-\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
-semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
-but we are unable to syntactically equate the two configurations due
-to the extra binders generated for dynamic checks and ANF
-transformation. In earlier versions of the Redex model, we attempted
-to change the compilation rules so the configurations could match
-syntactically. However, the approach scaled poorly as we added
-additional rules. 
-This slight relaxation on the equivalence relation
-between target configurations allows us to specify compilation more
-naturally without having to worry about syntactic constraints.
-
-\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
-and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-\rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
-some $e_1$ and
-$\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
-, then
-there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\varphi,\dot \heap,\dot e)$ and $(\dot
-\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-\dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-\varphi_1$, $\heap_1 \gg \dot \heap_1$.
-\end{thm}
-
-Our random generator (discussed in the next section) never produces
-$\euncheckedtext$ expressions (whose behavior could be undefined), so we
-can only test a the simulation theorem 
-as it applies to checked code. This limitation makes it
-unnecessary to state the other direction of the simulation theorem
-where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
-that $e_0$ will never enter a stuck state if it is well-typed in
-checked mode.
-
-The current version of the Redex model has been tested against $20000$
-expressions with depth less than $10$. Each expression can
-reduce multiple steps, and we test simulation between every two
-adjacent steps to cover a wider range of programs, particularly the
-ones that have a non-empty heap.
-
-\begin{figure}[t]
-{\small
-\[
-\begin{array}{c}
-\begin{tikzpicture}[
-            > = stealth, shorten > = 1pt, auto,
-            node distance = 3cm
-        ]
-
-\begin{scope}[every node/.style={draw}]
-    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
-    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
-    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
-    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
-    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
-\end{scope}
-\begin{scope}[every edge/.style={draw=black}]
-
-    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
-    \path [<->] (A) edge node {$\gg$} (C);
-    \path [<->] (B) edge node {$\gg$} (D);
-    \path [dashed,<->] (C) edge node {$\sim$} (D);
-    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
-    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
-\end{scope}
-
-\end{tikzpicture}
-\end{array}
-\]
-}
-\caption{Simulation between \lang and \elang }
-\label{fig:checkedc-simulation-ref}
-\end{figure}
-
-% \ignore
-% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
-% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
-% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist.
-The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.
-
-%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C.
-The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
-% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).
-
-% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
-% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
-% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation.
-We generate tens of thousands cases and cover all possible corner cases of the compilation.
-% }
-% \ignore{
-% \begin{figure*}[t]
-%   \begin{subfigure}[b]{1\textwidth}
-%     \begin{lstlisting}
-% % int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
-%   return *p;
-% }
-
-% ...
-% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
-% deref_array(5, p0);
-%     \end{lstlisting}
-%     \label{fig:chkcexamplederef}
-%     \caption{\texttt{deref\_array} in \checkedc}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{1\textwidth}
-% \begin{lstlisting}
-% % int deref_array(size_t n, int *p) {
-%   /* statically determine the definitions p_lo and p_hi */
-%   int *p_lo = p;
-%   int *p_hi = p + n;
-%   /* runtime checks */
-%   assert(p_lo <= p && p <= p_hi);
-%   assert(p != NULL);
-%   /* widening */
-%   int p_derefed = *p;
-%   if (p_derefed != '\0') {
-%     if (p_hi == p) {
-%       ++p_hi;
-%     }
-%   }
-%   return p_derefed;
-% }
-% ...
-% /* int *p0 */
-% /* rho(p0) = p_lo, p_hi */
-% deref_array(5, p0);
-% \end{lstlisting}
-% \caption{\texttt{deref\_array} in C}
-%     \label{fig:cexamplederef}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{\textwidth}
-%     \begin{lstlisting}
-% % /* nt_array_ptr<int> p : bounds(e0, e1) */
-% /* rho(p) = p_lo, p_hi */
-% size_t x = strlen(p);
-% if(x >= 1) {
-%   *(p+1);
-% }
-%     \end{lstlisting}
-%     \caption{\texttt{strlen} in \checkedc}
-%     \label{fig:chkcexamplestrlen}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{\textwidth}
-%     \begin{lstlisting}
-% % /* nt_array_ptr<int> p : bounds(e0, e1) */
-% /* rho(p) = p_lo, p_hi */
-% /* runtime checks omitted */
-% ...
-% size_t x = strlen(p);
-% int *p_hi_new = p + x;
-% p_hi = max(p_hi, p_hi_new);
-% if (x >= 1) {
-%   /* null check for pointer arithmetic */
-%   assert(p != NULL);
-%   int *p_1 = p + 1;
-%   /* null check for dereference */
-%   assert(p_1 != NULL);
-%   /* bounds check for dereference */
-%   /* note how we are using p_hi rather than p + x */
-%   assert(p_lo <= p_1 && p_1 <= p_hi);
-%   ...
-% }
-%     \end{lstlisting}
-%     \caption{\texttt{strlen} in C}
-%     \label{fig:cexamplestrlen}
-%   \end{subfigure}
-% \caption{Compilation of null-terminated string dereference}
-% \label{fig:compilationexample}
-% \end{figure*}
-% }
-
-% \ignore{
-% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
-% dereference of null-terminated strings by inserting explicit checks and
-% bounds widening code. For readability, we present the example in real
-% \checkedc and C syntax. At
-% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
-% bounds are defined in terms of the arguments \code{p} and \code{n}
-% according to the bounds annotations obtained during typechecking. This
-% avoids the need to pass in bounds as extra arguments, thus maintaining
-% compatibility with C code. When we call \code{deref_array} on
-% \code{p0}, a string with size 5 (excluding the null character), there
-% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
-% the stack. It is possible that \code{p_hi} has been widened
-% before we reach line 21. The programmer will have to perform a
-% \code{dynamic_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
-% attempts to widen the bound when the dereferenced result is not
-% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
-% \code{deref_array}, in contrast to the T-IfNT rule, which only
-% remembers the widened bounds within the scope of the then branch of
-% the if statement.
-
-% \todo[inline]{Widening can happen at every
-%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
-% shows how an invocation of \code{strlen} on null-termianted strings
-% is compiled into C code. We perform the same runtime checks that
-% happen during dereferencing. The widening code at line
-% 8 updates \code{p}'s upper bound only if the result of
-% \code{strlen} is larger than the value of the upper bound stored on
-% the stack. This is another scenario where the runtime can be more precise
-% than the statically determined bounds information.
-
-% }
- \section{Random Testing via the Implementation}\label{sec:evaluation}
-
-In addition to using the \lang Redex model to establish simulation of
-compilation (Section~\ref{sec:meta}), we also used it to gain confidence
-that our model matches the Clang \checkedc implementation;
-disagreement on outcomes signals a bug in either the model or the
-compiler itself. Doing so allowed us to quickly iterate on the
-design of the model while adding new features, and revealed
-several bugs in the Clang \checkedc implementation.
-
-\myparagraph{Generating Well Typed Terms} For this random generation, we follow
-the approach of~\citet{PalkaAST11} to generate well-typed \checkedc
-terms by viewing the typing rules as generation rules.
-Suppose we have a context $\textcolor{red}{\Gamma}$, a mode
-$\textcolor{red}{m}$ and a type $\textcolor{red}{\tau}$, and we are
-trying to generate a well-typed expression. We can do that by
-reversing the process of type checking, selecting a typing rule and
-building up an expression in a way that satisfies the rule's premises.
-
-Recall the typing rule for dereferencing an
-array pointer, which we depict below as
-\textsc{G-DefArr}\footnote{Generator rules G-* correspond one to one
-  with the type rules T-* in Sec.~\ref{sec:type-system}.}, color-coded to represent \textcolor{red}{inputs} and \textcolor{blue}{outputs} of the
-generation process:\footnote{This input-output marking is commonly called a mode in
-  the literature, but we eschew this term to avoid
-  confusion with our pointer mode annotation.}
-\begin{mathpar}
-    \inferrule[G-DefArr]
-              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} : \tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m_a}} \\
-\textcolor{red}{m} \leq \textcolor{cyan}{m_a} 
-              }
-              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\estar{e}} : \textcolor{red}{\tau}}
-\end{mathpar}
-If we selected \textsc{G-DefArr} for generating an expression, the
-generated expression has to have the form $\textcolor{blue}{*e}$,
-for some $\textcolor{blue}{e}$, to be generated according
-to the rule's premises.
-To satisfy the premise
-$\textcolor{red}{\Gamma};\textcolor{red}{\Theta}
-\vdash_{\textcolor{red}{m}} \textcolor{blue}{e} :
-\tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m_a}}$,
-we essentially need to make a recursive call to the generator, with
-appropriately adjusted inputs.
-However, the type in this judgment is not fixed yet---it contains
-three unknown variables: $\textcolor{cyan}{m_a}$,
-$\textcolor{cyan}{\bvar}$, and $\textcolor{cyan}{\kappa}$---that need
-to be generated before making the call.
-Looking at the second premise informs that generation:
-if the input mode $\textcolor{red}{m}$ is $\umode$, then $\textcolor{cyan}{m_a}$
-needs to be $\umode$ as well; if not, it is unconstrained,
-just like $\textcolor{cyan}{\bvar}$ and $\textcolor{cyan}{\kappa}$, and therefore all three are free to be generated at random.
-Thus, the recursive call to generate $\textcolor{blue}{e}$ can now
-be made, and the \textsc{G-DefArr} rule returns $\textcolor{blue}{*e}$
-as its output.
-
-Using such generator rules, we can create a generator for random
-well-typed terms of a given type in a straightforward manner: find all
-rules whose conclusion matches the given type and then randomly choose
-a candidate rule to perform the generation. To ensure that this
-process terminates, we follow the standard practice of using ``fuel'' to
-bound the depth of the generated terms; once the fuel is exhausted,
-only rules without recursive premises are
-selected~\cite{Pierce:SF4}. Similar methods were used for
-generating top level functions and \kw{struct} definitions.
-
-While using just the typing-turned-generation rules is in theory
-enough to generate all well-typed terms, it's more effective in
-practice to try and exercise interesting patterns.  As
-in~\citet{PalkaAST11} this can be viewed as a way of adding admissible
-but redundant typing rules, with the sole purpose of
-using them for generation. For example, below is one such rule,
-\textsc{G-ASTR}, which creates an initialized null-terminated string that
-is statically cast into an array with bounds $(0,0)$.
-\begin{mathpar}
-    \inferrule[G-ASTR]
-    {
-      \textcolor{cyan}{i} \in \mathbb{N}^* \quad\quad
-      \textcolor{cyan}{n_0,\ldots,n_{i-1} } \in \mathbb{Z} \quad\quad
-      \texttt{fresh}(\textcolor{cyan}{x})\\
-      \textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}}
-      \textcolor{blue}{e'} :
-      \tptr{\textcolor{red}{\tntarray{0}{\textcolor{cyan}{i}}{\tint}}}{c}
-      \\      
-\textcolor{blue}{e} = \elet{\textcolor{cyan}{x}}{\textcolor{blue}{e'}}{\texttt{(init $\textcolor{cyan}{x}$ with }\textcolor{cyan}{n_0,\ldots
-        n_{i-1}}\texttt{)};\textcolor{cyan}{x}}
-}
-    {\textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\ecast{\tptr{\tntarray{0}{0}{\tint}}{c}}{e}} : \textcolor{red}{\tptr{\tntarray{0}{0}{\tint}}{c}}}
-  \end{mathpar}
-Given some positive number $\textcolor{cyan}{i}$, numbers
-$\textcolor{cyan}{n_0, \ldots, n_{i-1}}$, and a fresh variable
-$\textcolor{cyan}{x}$ (which are arbitrarily generated), we can
-recursively generate a pointer $\textcolor{blue}{e'}$ with bounds
-$(0,\textcolor{cyan}{i})$, and initialize it with the generated $n_j$
-using $\textcolor{cyan}{x}$ to temporarily store the pointer.
-
-This rule is particularly useful when combined with \textsc{G-IfNT}
-since there is a much higher chance of obtaining a non-zero value when
-evaluating $\estar{p}$ in the guard of $\eiftext$, skewing the
-distribution towards programs that enter the $\ethentext$ branch.
-Relying solely on the type-based rules, entering the $\ethentext$
-branch requires \textsc{G-AssignArr} was chosen before \textsc{G-IfNT},
-and that assignment would have to appear before $\eiftext$, which
-means additional \textsc{G-Let} rules would need to be chosen: this
-combination would therefore be essentially impossible to generate in
-isolation.
-
-\myparagraph{Generating Ill-typed Terms}
-We can use generated well-typed terms to test our simulation theorem
-(Section~\ref{sec:compilation}) and test that \lang and \checkedc Clang agree
-on what is type-correct. But it is also useful to generate ill-typed
-terms to test that \lang and \checkedc Clang agree on those.  However,
-while it is easy to generate arbitrary ill-typed terms, they would be
-very unlikely to trigger any inconsistencies; those are far more
-likely to exist on the boundary between well- and
-ill-typedness. Therefore, we also include generation rules modified to
-be slightly more permissive, which results in sometimes generating
-terms that are ``a little'' ill-typed.
-
-\myparagraph{Random Testing for Language Design}
-We used our Redex model and random generator to successfully guide the
-design of our formal model, and indeed the Clang Checked C
-implementation itself, which is being actively developed. To that end, we implemented a
-conversion tool that converts \lang into a subset of the \checkedc
-language and ensured that model and implementation exhibit the same
-behavior (accept and reject the same programs and yield the same return
-value).
-
-This approach constitutes an interesting twist to traditional
-model-based checking approaches.  Usually, one checks that the
-implementation and model agree on all inputs \emph{of the
-  implementation}, with the goal of covering as many behaviors as
-possible. This is the case, for example, in \citet{lambdajs}, where
-they use real test suites to demonstrate the faithfullness of their
-core calculus to Javascript. Our approach and goal in this work is
-essentially the opposite: as the Clang Checked C implementation does
-not fully implement the Checked C spec, there is little hope of
-covering all terms that are generated by Clang Checked C. Instead,
-we're looking for \emph{inconsistencies}, which could be caused by
-bugs either in the Clang Checked C compiler or our own model.
-
-One inconsistency we found comes from the following:
-
-{\small 
-\begin{lstlisting}[xleftmargin=4 mm]
-array_ptr<char> fun(void) : count(3) {
-   array_ptr<char> x : count(3);
-   x = calloc(3, sizeof(char));
-   return x+3;
- }
- int main(void) {
-   *(fun()) = 0;
-   return 0;
- }
-\end{lstlisting}
-}\noindent
-In this code, the function \code{fun} is supposed to return a
-checked array pointer of size 3. Internally, it allocates such an
-array, but instead of returning the pointer \code{x} to that array, it
-increments that pointer by 3. Then, the \code{main} function just
-calls \code{fun}, and tries to assign 0 to its result. Our model
-correctly rules out this program, while the Clang Checked C
-implementation happily accepted this out-of-bounds
-assignment. Interestingly, it correctly rejected programs where the
-array had size 1 or 2. This inconsistency has been fixed in the latest
-version of the compiler.
-
-We also found the opposite kind of inconsistency---programs that
-the Clang Checked C implementation rejects contrary to the spec.
-For instance:\footnote{After minimization, this turned out to be a known issue: 
-  \url{https://github.com/microsoft/checkedc-clang/issues/1008}}
-
-{\small 
-\begin{lstlisting}[xleftmargin=4 mm]
-array_ptr<int> f(void) : count(5) {
-  array_ptr<int> x : count(5) =
-    calloc<int>(5, sizeof(int)); 
-  return x;
-}
-array_ptr<int> g(void ) : count(5) {
-  array_ptr<int> x : count(5) =
-    calloc<int>(5, sizeof(int)); 
-  return x+3;
-}
-int main(void) {
-  return *(0 ? g() : f() + 3);
-}    
-\end{lstlisting}
-}\noindent
-In this piece of code both \code{f} and \code{g} functions compute a
-pointer to the same index in an array of size 5 (as \code{f} calls
-\code{g}). The \code{main} function then creates a ternary expression
-whose branches call \code{f} and \code{g}, but the Clang Checked C
-implementation rejects this program, as its static analysis is not
-sophisticated enough to detect that both branches have the same type.
-
- \section{Related Work}
-\label{sec:related}
-
-Our work is most closely related to prior formalizations of C(-like)
-languages that aim to enforce memory safety, but it also touches on
-C-language formalization in general.
-
-\myparagraph{Formalizing C and Low-level code}
-A number of prior works have looked at formalizing the semantics of C,
-including CompCert~\cite{Blazy2009,leroy:hal-00703441},
-\citet{ellison-rosu-2012-popl}, \citet{Kang:2015:FCM:2813885.2738005},
-and \citet{10.1145/2980983.2908081, Memarian:2019:ECS:3302515.3290380}. These works also model
-pointers as logically coupled with either the bounds of the blocks
-they point to, or provenance information from which bounds can be
-derived. None of these is directly concerned with enforcing
-spatial safety, and that is reflected in the design. For example,
-memory itself is not be represented as a flat address space, as in our
-model or real machines, so memory corruption due to spatial safety
-violations, which Checked C's type system aims to prevent, may not be
-expressible. That said, these formalizations consider much more of the
-C language than does \lang, since they are interested in the entire
-language's behavior.
-
-\myparagraph{Spatially Safe C Formalizations}
-Several prior works formalize C-language transformations or C-language
-dialects aiming to ensure spatial safety. 
-\citet{10.1145/2813885.2737979} extends the formalization
-of \citet{ellison-rosu-2012-popl} to produce a semantics that detects
-violations of spatial safety (and other forms of undefinedness). It
-uses a CompCert-style memory model, but ``fattens'' logical pointer
-representations to facilitate adding side conditions similar to \lang's.
-Its concern is bug finding, not compiling programs to
-use this semantics.
-
-CCured~\cite{Necula2005} and Softbound~\cite{softbound} implement
-spatially safe semantics for normal C via program transformation. Like
-\lang, both systems' operational semantics annotate pointers with
-their bounds. CCured's equivalent of array pointers are compiled to be
-``fat,'' while SoftBound compiles bounds metadata to a separate
-hashtable, thus retaining binary compatibility at higher checking
-cost. Checked C uses static type information to enable bounds checks
-without need of pointer-attached metadata, as we show in
-Section~\ref{sec:compilation}. Neither CCured nor Softbound models
-null-terminated array pointers, whereas our semantics ensures that
-such pointers respect the zero-termination invariant, leveraging
-bounds widening to enhance expressiveness.
-
-Cyclone \cite{Jim2002,GrossmanMJHWC02} is a C dialect that aims to
-ensure memory safety; its pointer types are similar to
-CCured. Cyclone's formalization~\cite{GrossmanMJHWC02} focuses on the
-use of \emph{regions} to ensure temporal safety; it does not formalize
-arrays or threats to spatial safety. Deputy~\cite{Feng2006,Condit2007}
-is another safe-C dialect that aims to avoid fat pointers; it was an
-initial inspiration for Checked C's design~\cite{Elliott2018}, though
-it provides no specific modeling for null-terminated array
-pointers. Deputy's formalization~\cite{Condit2007} defines its
-semantics directly in terms of compilation, similar in style to what
-we present in Section~\ref{sec:compilation}. Doing so tightly couples
-typing, compilation, and semantics, which are treated independently in
-\lang. Separating semantics from compilation isolates meaning 
-from mechanism, easing understandability. Indeed, it was this
-separation that led us to notice the 
-limitation with Checked C's handling of bounds widening.
-
-The most closely related work is the formalization of \checkedc done
-by \citet{ruef18checkedc-incr}. They were the first to formalize and
-prove \emph{blame} for a core model of \checkedc, which shows that any
-spatial safety violation owes to invariants violated by unchecked
-code. Our Coq-based development (Section~\ref{sec:formal})
-substantially extends
-theirs,\footnote{\url{https://github.com/plum-umd/checkedc/tree/master/coq}}
-re-proving the blame theorem after adding dynamically bounded array
-pointers with dependent types, null-terminated pointers, and
-dependently typed functions. They postulate that pointer metadata can
-be erased, but do not show it; indeed, we found it nontrivial once
-null-terminated pointers were considered.
-
-% \ignore{
-
-% Other work that might be relavent but for guaranteeing temporal safety, I have recorded some of them in the related work if they are relavent to spatial safety, but I might missed some.
-
-% Purify operates on binaries, but only ensures the safety of heap-allocated objects.
-
-% Safe C ensures complete temporal safety by using fat-pointers.
-
-% MSCC uses a fat-pointer approach for handling the spatial and temporal safeties regarding stack/heap pointers.
-According to the experiment in MSCC, it is supposed to be faster than CCured, but it seems that they are in the same level or CCured might be faster in some cases according to.
-
-% Fail-Safe C is an updated version of Safe C, and it maintains complete compatibility with ANSI C but incurs significant runtime overhead.
-
-% Cyclone is a new dialect designed to minimise the changes required to port over C programs.
-Cyclone’s region-based memory management system provides a foundation for solving the temporal safety issues for stack pointers in Checked-C.
-The only difference between the Cyclone’s region-based system and the Checked-C will be time when dangling pointers cause errors.
-Dangling pointers should be banned at the time when they are used to access memory.
-If a program contains a function call with a stack pointer that is created and returned to the upper level function call and it is never used to access memory, the program should be considered as valid.
-
-% Here are several ideas of using metadata instead of fat-pointers, I (Liyi) think they are basically another form of fat-pointer approaches:
-
-% Patil and Fischer address the spatial and temporal safeties by maintaining disjoint metadata and performing checks in a separate “shadow process,” but this requires an additional CPU.
-
-% MemSafe provides a framework for checking spatial and temporal memory safety at runtime. 
-% Their work is based on transferring an original C program to an updated CFG-based program with extra metadata. These metadata has similar functionality as fat pointer and there are extra
-% checks allowing them to place extra checks during the program execution.
-For each pointer, the updated CFG stores information about all possible values of the pointer through possible aliasing.
-The efficiency of MemSafe can be as effective as CCured.
-
-% Write integrity testing -- similar to regional memory management.
-The so-called colored memory space is essentially giving regional flags to different pointers, and then check if these pointers have potential security threats.
-
-% The following approaches require hardware support:
-
-% Clause et al. describe an efficient technique for detecting memory errors, but it requires custom hardware.
-
-% Dhurjati and Adve describe a technique based on the Electric Fence [11] malloc debugger: Their system assigns a unique virtual page to every dynamically allocated object and relies on hardware page protection to detect dangling pointer dereferences.
-
-% There are also few static approaches to the temporal safety in C.
-However, they are not complete and a lot of them are not even sound.
-There are four kinds of static approach methods: model checking, abstract interpretation, pattern matching, and pointer analysis.
-
-% Model Checking.
-CBMC is a bounded model checker that reasons about all the program paths for C/C++ programs as constraints that can be solved by an SMT solver.
-When used in finding UAF bugs, CBMC is sound (in a bounded manner) and highly precise but scales only to small programs whose “sizes are restricted” (according to its user manual).
-
-% Thomas Ball and Sriram K.
-Rajamani developed a model checking framework (C2bp) for checking Temporal Safety Properties in C.
-The framework allows pointer analysis by using a pointer aliasing algorithm.
-The problem is that it only deals with small programs, and the pointer aliasing algorithm only detects a small part of aliasing pointers.
-
-% MOPS (MOdel checking Program for Security Properties) makes use of the model checking technique to check for violation of security rules, that are defined as temporal safety properties.
-As for program representation, MOPS models the program in the form of Push Down Automaton (PDA), that contains all the possible execution paths.
-Push Down Automata are used as tools to analyze procedural sequential programs, and more specifically those having recursive procedures.
-As for automata, they are according to Schneider used in the objective of specifying security policies that can be enforced by mechanisms.
-MOPS makes use of this approach to model security properties in the form of Finite State Automata (FSA), that dictate the order of security-relevant operations sequence.
-The modularity of security properties was also proposed by MOPS; this approach allows the decomposition of complex security properties into simpler and reusable basic security properties that are easy to model and to extend (such as role based access).
-MOPS verifies that the security properties are properly respected in all the execution paths of the analyzed program, making use of the model checking on the PDA, and checks if risky states are reachable within the PDA.
-MOPS does not know that a pointer is an alias of another, so the property is not sound for this program.
-
-% Abstract Interpretation.
-Clang Static Analyzer is an abstract interpreter for analyzing C/C++ programs.
-It adopts a highly unsound model by analyzing only a small subset of the program paths in order to achieve scalability and precision.
-To scale for large codebases with few false alarms, Clang limits its UAF-bug-finding ability by performing an intraprocedural analysis (with inlining).
-In general, such tools refrain from reporting too many false alarms, but at the expense of missing many UAF bugs.
-
-% Pattern Matching.
-Coccinelle is a pattern-based tool for analyzing and certifying C programs.
-Coccinelle can find UAF bugs based on some patterns given.
-Due to the lack of the points-to information, Coccinelle can be both fairly unsound and imprecise but is highly scalable (due to its pattern-matching nature).
-
-% Pointer Analysis.
-Supa is a state-of-the-art demand-driven pointer analysis that is field-, flow- and context-sensitive but path- insensitive for C programs.
-When used in finding UAF bugs, Supa can be regarded as reasoning about all the program paths with an extremely coarse abstraction, in order to achieve soundness and scalability.
-
-% Yan et al.
-develops a Spatio-Temporal context reduction technique for  doing pointer-analysis for detecting use-after-free vulnerabilities, which are essentially temporal safety.
-They mainly focus on the heap pointer allocation-free detection to  check if a pointer is used after the free function call for the memory field the pointer points to.
-The technique is based on static analysis.
-The main idea is still based on intraprocedural control flow analysis.
-It relies on context sensitive reduction technique to reduce the search function calling path spaces for analyzing pointer aliasing.
-The context sensitivity and their pointer analysis algorithm is applied in a state-of-the-art manner.
-The information for the algorithm of search pointer aliasing reduces the calling context path, while the context path information enhances the efficiency of the algorithm.
-However, the technique is not complete.
-It will ban a large number of valid C programs due to false use-after-free vulnerabilities.
-
-% It seems that they might not detect the use-after-free error in this case. In addition, the work does not handle stack pointer use-after-free vulnerabilities.
-
-% ASAP: As Static As Possible memory management -- use data flow analysis, garbage collection and region memory management tools to ensure the security safety of C programs by static analysis.
-In order to guarantee the use-after-free, it requires the program written in a special way: in a linearly typed program, each value must be used exactly once.
-It does not include examples of dealing with stack pointers.
-It shows that in a pointer management world, a C program temporal safety cannot be fully analyzed by only static analysis.
-
-% Jia-Ju Bai et al.
-develops a static method to detect use-after-free bugs in Linux device drivers.
-It is a UAF bug in the concurrent system.
-The basic algorithm is simple.
-It first goes through each single-threaded code to collect a read/write set storing point-to information.
-The information collection is through traditional intraprocedural CFG analysis.
-Then, it uses a trace analysis tool to discover UAF bugs.
-This is achievable only in dealing with linux device drivers, since usually, these drivers are simple and the concurrency is very formulated.
-I believe if we apply the same process in dealing with C, the method will not be sound.
-
-% Industrial tools:
-
-% There are also some industrial level tools for static checking bugs in C.
-However, they are either not able to check UAF bugs or they only use heuristics to check very small sets of UAF bugs.
-They include CppCheck, Clang Core, Clang Alpha, CodeSonar, Facebook Infer,  Splint standard, Splint weak.
-For example, CodeSonar state-based checkers do not check temporal safety.
-They require insertion of extra code in order for the tool to check if a supposed deadcode will be reached or not.
-Coverity is a dataflow analysis tool that relies on inter-procedural analysis techniques.
-The analysis is neither sound nor complete, that is, there may be both defects which are not reported and there may be false alarms.
-Coverity Prevent implements an incremental analysis which means that the system automatically infers what parts of the source code that have to be re-analyzed after the code has been modified.
-This typically reduces the analysis time substantially, but may of course imply a complete re-analysis in the worst case.
-Klocwork does similar things to Coverity and Coverity Prevent.
-% }
-
-\section{Conclusion and Future Work}
-\label{sec:conclude}
-
-This paper presented \CoreChkC, a formalization of an extended core of
-the Checked C language which aims to provide spatial memory
-safety. \lang models dynamically sized and
-null-terminated arrays with dependently typed bounds that can
-additionally be widened at runtime. We prove, in Coq, the key safety
-property of Checked C for our formalization, {\em blame}: if a mix of
-checked and unchecked code gives rise to a spatial memory safety
-violation, then this violation originated in an unchecked part of the
-code. We also show how programs written in \CoreChkC (whose
-semantics leverage fat pointers) can be compiled to \elang (which does
-not) while preserving their behavior. Finally, we developed a random
-testing framework to guide the design of our formal model by comparing
-it against the \checkedc compiler, finding multiple inconsistencies
-in the process.
-
-As future work, we are interested in designing an a way to
-automatically port legacy C code to \checkedc.  We also want to
-further extend our \CoreChkC model to include more C behaviors,
-such as function pointers, with our testing framework guiding the
-design process.
-
-%  \ignore{
-% \begin{acks}                            This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-% }
-
-\begin{small}
-\bibliography{IEEEabrv,paper,sources}
-\end{small}
-\newpage
-\appendix
-\section{Appendix}
-
-\subsection{Typing Rules for Literal Pointers}\label{sec:literal-pointer-typing}
-
-One thing we elided from the main presentation is the typing of
-integer literals (which can also be pointers to the heap). These rules
-are shown in Fig.~\ref{fig:const-type}. The variable type rule
-($\textsc{T-Var}$) simply checks if a given variable has the defined
-type in $\Gamma$; the constant rule ($\textsc{T-Const}$) is slightly
-more involved.  First, it ensures that the type annotation $\tau$ does
-not contain any free variables. More importantly, it ensures that the
-literal itself is well typed using an auxilliary typing relation
-$\sigma \vdash n : \tau$, which is implicitly indexed by a given heap
-$\heap$.
-
-If the literal's type is an integer, an unchecked pointer, or a null
-pointer, it is well typed, as shown by the top three rules in
-Fig.~\ref{fig:const-type}. However, if it is a checked pointer
-$\tptr{\omega}{\cmode}$, we need to ensure that what it points to in
-the heap is of the appropriate pointed-to type ($\omega$), and also
-recursively ensure that any literal pointers reachable this way are
-also well-typed. This is captured by the bottom rule in the figure,
-which states that for every location $n+i$ in the pointers' range
-$[n, n+\size(\omega))$, where $\size$ yields the size of its argument,
-  then the value at the location $\heap(n+i)$ is also well-typed.
-  However, as heap snapshots can contain cyclic structures (which
-  would lead to infinite typing deriviations), we use a scope $\sigma$
-  to assume that the original pointer is well-typed when checking the
-  types of what it points to. The middle rule then accesses the scope
-  to tie the knot and keep the derivation finite, just like in
-  \citet{ruef18checkedc-incr}.
-
- \begin{figure}[h]
- {\small
- \text{Type Rules for Constants and Variables:}
- \begin{mathpar}
-   \inferrule[T-Var]
-       {x : \tau \in \Gamma}
-       {\Gamma;\Theta \vdash_m x : \tau}
-
-   \inferrule[T-Const]
-       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
-       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
- \end{mathpar}
-
- \text{Rules for Checking Constant Pointers In Heap:}
- \begin{mathpar}
-   \inferrule
-       {}
-       {\sigma \vdash n : \tint}
-
-   \inferrule
-       {}
-       {\sigma \vdash n : \tptr{\omega}{\umode}}
-
-   \inferrule
-       {}
-       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
-
-   \inferrule
-       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
-       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-
-   \inferrule
-       {\forall i \in [0,\size(\omega)) .
-            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
-       {\sigma \vdash n : \tptr{\omega}{\cmode}}
- \end{mathpar}
- }
- \caption{Type Rules for Checking Constants/Variables}
- \label{fig:const-type}
- \end{figure}
-
-\subsection{Other Semantic Rules}\label{sec:rem-semantics}
-
-Fig.~\ref{fig:rem-semantics} shows the remaining semantic rules for
-$\lang$. We explain a selected few rules in this subsection.
-
-\begin{figure*}[t]
-{\small
-\begin{mathpar}
-
-\inferrule [S-Var]{} {(\varphi,\heap,x)\longrightarrow (\varphi,\heap,\varphi(x))}
-
-    \inferrule[S-DefArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-DefArrayBound]{0 \not\in [n_l,n_h)}
-     { (\varphi,\heap,\estar{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-    \inferrule[S-DefNTArrayBound]{0 \notin [n_l,n_h]}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-    \inferrule[S-Assign]{\heap(n)=\evalue{n_a}{\tau_a} }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tptr{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}
-
-    \inferrule[S-AssignNull]{}
-      {(\varphi,\heap,\eassign{\evalue{0}{\tptr{\omega}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\enull)}
-
-    \inferrule[S-AssignArrBound]{0 \not\in [n_l,n_h) }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-  \inferrule[S-Malloc]{\varphi(\omega)=\omega_a \\ \mathtt{alloc}(\heap,\omega_a)=(n,\heap')}
-   { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\evalue{n}{\tptr{\omega_a}{\cmode}})}
-
-  \inferrule[S-MallocBound]{\varphi(\omega)=\tallarray{n_l}{n_h}{\tau}\\ (n_l \neq 0 \vee n_h \le 0)}
-    { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\ebounds)}
-
-    \inferrule[S-IfT]{n \neq 0 }
-    {(\varphi,\heap,\eif{\evalue{n}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
-
-    \inferrule[S-IfF]{}
-    {(\varphi,\heap,\eif{\evalue{0}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}
-
-    \inferrule[S-Unchecked]{}
-    {(\varphi,\heap,\eunchecked{\evalue{n}{\tau}} \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
-    \inferrule[S-Str]{
-        0 \in [n_l,n_h]
-\\ \heap(n+n_a) = 0 
-\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
-    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tint})}
-
-    \inferrule[S-StrBounds]{
-        0 \notin [n_l,n_h]
-}
-    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-    \inferrule[S-StrNull]{}
-    {(\varphi,\heap,\estrlen{\evalue{0}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\enull)}
-
-    \inferrule[S-Add]{n = n_1 + n_2}
-    {(\varphi,\heap,\evalue{n_1}{\tint} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, n)}
-
-    \inferrule[S-AddArr]{n = n_1 + n_2\\ n_l' = n_l - n_2 \\ n_h' = n_h - n_2}
-    {(\varphi,\heap,\evalue{n_1}{\tallarrayptr{n_l}{n_h}{\tau}{m}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau}{m}})}
-
-n    \inferrule[S-AddArrNull]{}
-    {(\varphi,\heap,\evalue{0}{\tallarrayptr{n_l}{n_h}{\tau}{c}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \enull)}
-
-\end{mathpar}
-
-}
-\caption{Remaining \lang Semantics Rules (extends Fig.~\ref{fig:semantics})}
-\label{fig:rem-semantics}
-\end{figure*}
-
-Rule \textsc{S-Var} loads the value for $x$ in stack $\varphi$.
-Rule \textsc{S-DefArray} dereferences an array pointer, which is similar to the Rule \textsc{S-DefArray} in Fig.~\ref{fig:semantics}.
-The only difference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
-Rules \textsc{DefArrayBound} and \textsc{DefNTArrayBound} describe an error case for a dereference operation.
-If we are dereferencing an array/NT-array pointer and the mode is $\cmode$, $0$ must be in the range from $n_l$ to $n_h$; if not, the system results in a $\ebounds$ error.
-Obviously, the dereference of an array/NT-array pointer also experiences a $\enull$ state transition if $n\le 0$.
-
-Rules \textsc{S-Malloc} and \textsc{S-MallocBound} describe the $\emalloctext$ semantics.
-Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$, and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation.
-The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\cmode}}$ and new updated heap.
-It is possible for $\emalloctext$ to transition to a $\ebounds$ error if the $\omega_a$ is an array/NT-array type $\tallarray{n_l}{n_h}{\tau}$, and either $n_l \neq 0$ or $n_h \le 0$.
-
-\subsection{Subtyping for dependent types}
-\label{app:le}
-
-The subtyping relation given in Fig.~\ref{fig:checkc-subtype} involves
-dependent bounds, i.e., bounds that may refer to variables. To decide
-premises $b \leq b'$, we need a decision procedure that accounts for
-the possible values of these variables. This process considers
-$\Theta$, tracked by the typing judgment, and $\varphi$, the current
-stack snapshot (when performing subtyping as part of the type
-preservation proof). We
-
-\begin{defi}[Inequality]
-
-\begin{itemize}
-
-\item $n \le m$ if $n$ is less than or equal to $m$.
-\item $x+n \le x + m$ if $n$ is less than or equal to $m$.
-\item All other cases result in $\efalse$.
-
-\end{itemize}
-\end{defi}
-
-To capture bound variables in dependent types, the \checkedc subtyping
-relation ($\sqsubseteq$) is parameterized by a restricted stack
-snapshot $\varphi|_{\rho}$ and the predicate map $\Theta$, where
-$\varphi$ is a stack and $\rho$ is a set of
-variables. $\varphi|_{\rho}$ means to restrict the domain of $\varphi$
-to the variable set $\rho$. Clearly, we have the relation:
-$\varphi|_{\rho} \subseteq \varphi$. The meaning of $\sqsubseteq$
-being parameterized by $\varphi|_{\rho}$ refers to that when we
-compare two bounds $b \le b'$, we actually do
-$\varphi|_{\rho}(b) \le \varphi|_{\rho}(b')$ by interpreting the
-variables in $b$ and $b'$ with possible values in $\varphi|_{\rho}$.
-Let's define a subset relation $\preceq$ for two restricted stack
-snapshot $\varphi|_{\rho}$ and $\varphi'|_{\rho}$:
-
-\begin{defi}[Subset of Stack Snapshots]
-  Given two $\varphi|_{\rho}$ and $\varphi'|_{\rho}$,
-  $\varphi|_{\rho} \preceq \varphi'|_{\rho}$, iff for $x\in\rho$ and
-  $y$,
-  $(x,y) \in \varphi|_{\rho} \Rightarrow (x,y) \in \varphi'|_{\rho}$.
-\end{defi}
-
-For every two restricted stack snapshots $\varphi|_{\rho}$ and
-$\varphi'|_{\rho}$, such that
-$\varphi|_{\rho} \preceq \varphi'|_{\rho}$, we have the following
-theorem in \checkedc (proved in Coq):
-
-\begin{thm}[Stack Snapshot Theorem]
-  Given two types $\tau$ and $\tau'$, two restricted stack snapshots
-  $\varphi|_{\rho}$ and $\varphi'|_{\rho}$, if
-  $\varphi|_{\rho}\preceq \varphi'|_{\rho}$, and
-  $\tau \sqsubseteq \tau'$ under the parameterization of
-  $\varphi|_{\rho}$, then $\tau \sqsubseteq \tau'$ under the
-  parameterization of $\varphi'|_{\rho}$.
-\end{thm}
-
-Clearly, for every $\varphi|_{\rho}$, we have
-$\emptyset \preceq \varphi|_{\rho}$. The type checking stage is a
-compile-time process, so $\varphi|_{\rho}$
-is $\emptyset$ at the type checking stage. Stack snapshots are needed
-for proving type preserving, as variables in bounds expressions are
-evaluated away.
-
-As mentioned in the main text, $\sqsubseteq$ is also parameterized by
-$\Theta$, which provides the range of allowed values for a bound
-variable; thus, more $\sqsubseteq$ relation is provable. For example,
-in Fig.~\ref{fig:strncat-ex}, the \code{strlen} operation in line 4
-turns the type of \code{a} to be $\tntarrayptr{0}{x}{\tint}{\cmode}$
-and extends the upper bound to \code{x}. In the \code{strlen} type
-rule, it also inserts a predicate \code{x}$\ge 0$ in $\Theta$; thus,
-the cast operation in line 16 is valid because
-$\tntarrayptr{0}{x}{\tint}{\cmode} \sqsubseteq
-\tntarrayptr{0}{0}{\tint}{\cmode}$ is provable when we know
-\code{n}$\ge 0$.
-
-Note that if $\varphi$ and $\Theta$ are $\emptyset$, we do only the
-syntactic $\le$ comparison; otherwise, we apply $\varphi$ to both
-sides of $\sqsubseteq$, and then determine the $\le$ comparasion based
-on a Boolean predicate decision procedure on top of $\Theta$. This
-process allows us to type check both an input expression and the
-intermediate expression after evaluating an expression.
-
-\subsection{Other Type Rules}\label{rem-type}
-
-Here we show the type rules for other \checkedc operations in Fig.~\ref{fig:rem-type-system}.
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[T-Def]
-              {\Gamma;\Theta \vdash_m e : \tptr{\tau}{m'} \\
-                m \leq m'}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-
-    \inferrule[T-Mac]
-              {}
-              {\Gamma; \Theta \vdash_m \emalloc{\omega} : \tptr{\omega}{\cmode}}
-
-    \inferrule[T-Add]
-              {\Gamma; \Theta \vdash_m e_1 : \tint \\
-                \Gamma; \Theta \vdash_m e_2 : \tint}
-              {\Gamma; \Theta \vdash_m (e_1 \plus e_2) : \tint }
-
-    \inferrule[T-Ind] 
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'} \\
-                \Gamma; \Theta \vdash_m e_2 : \tint \\
-                m \leq m'}              
-              {\Gamma; \Theta \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
-
-    \inferrule[T-Assign]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tau}{m'} \\
-                \Gamma; \Theta \vdash_m e_2 : \tau' \\
-                \tau'\sqsubseteq \tau \\
-                m \leq m'}
-              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
-
-   \inferrule[T-IndAssign]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
-                \Gamma; \Theta \vdash_m e_2 : \tint \\
-                \Gamma; \Theta \vdash_m e_3 : \tau' \\
-                \tau'\sqsubseteq \tau \\
-                m \leq m'}
-              {\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
-
-    \inferrule[T-Unchecked]
-              {\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{e} : \tau}
-  \end{mathpar}
-}
-\caption{Remaining \lang Type Rules (extends Fig.~\ref{fig:type-system-1})}
-\label{fig:rem-type-system}
-\end{figure}
-
-Rule \textsc{T-Def} is for dereferencing a non-array pointer. 
-The statement $m \leq m'$ relates the unchecked region for a term with
-its sub-terms. We require that if the sub-term has an unchecked
-region, so does the whole term. Rule \textsc{T-Mac} deals with
-$\emalloctext$ operations. There is a well-formedness check to require
-that the possible bound variables in $\omega$ must be in the domain of
-$\Gamma$ (see Fig.~\ref{fig:wftypesandbounds}).
-Rule \textsc{T-Add} deals with binary operations whose sub-terms are integer expressions, while rule \textsc{T-Ind} serves the case for pointer arithmetic.
-For simplicity, in the \checkedc formalization, we do not allow arbitrary pointer arithmetic.
-The only pointer arithmetic operations allowed are the forms shown in rules \textsc{T-Ind} and \textsc{T-IndAssign} in Fig.~\ref{fig:rem-type-system}.
-Rule \textsc{T-Assign} is for assigning a value to a non-array pointer location.
-The predicate $\tau'\sqsubseteq \tau$ requires that the value being assigned is a subtype of the pointer type.
-The \textsc{T-IndAssign} rule is an extended assignment operation for handling assignments for array/NT-array pointers with pointer arithmetic.
-Rule \textsc{T-Unchecked} type checks \code{unchecked} blocks.
-
-\subsection{Struct Pointers}\label{appx:struct}
-
-\checkedc has \kw{struct} types and \kw{struct} pointers. Fig.~\ref{fig:checkc-struct} contains the syntax of \kw{struct} types as well as new subtyping relations built on the \kw{struct} values.
-For a \kw{struct} typed value, \checkedc has a special operation for it, which is $\eamper{e}{f}$.
-This operation indexes the $f$-th position \kw{struct}~$T$ item, if the expression $e$ is evaluated to a \kw{struct} pointer $\tptr{\tstruct{T}}{m}$.
-Rule \textsc{T-Struct} in Fig.~\ref{fig:checkc-struct} describes its typing behavior.
-Rules \textsc{S-StructChecked} and \textsc{S-StructUnChecked} describe the semantic behaviors of $\eamper{e}{f}$ on a given \kw{struct} \code{checked}/\code{unchecked} pointers, while rule \textsc{S-StructNull} describes a \code{checked} \kw{struct} null-pointer case.
-In our Coq/Redex formalization, we include the \kw{struct} values and the operation $\eamper{e}{f}$. We omit it in the main text due to the paper length limitation.
-
-\begin{figure}
-{\small
-$\begin{array}{l}
-\text{  Struct Syntax: }\\[0.5em]
-  \begin{array}{ll}
- \mathtt{Type} & \tstruct{T}
-\\[0.2em]
-     \text{Structdefs} & D \; \in \; T \rightharpoonup fs \\[0.2em]
-      \text{Fields} & fs \; ::= \; \tau~\mathtt{f} \mid \tau~\mathtt{f}; fs 
-    \end{array}\\[2em]
-\text{  Struct Subtype: }\\[0.5em]
-\begin{array}{l}
-    D(T) = fs \wedge fs(0) = \tnat \Rightarrow  \tptr{\tstruct{T}}{m} \sqsubseteq \tptr{\tnat}{m}\\[0.5em]
-    D(T) = fs \wedge fs(0) = \tnat \wedge 0 \le b_l \wedge b_h \le 1 \\[0.2em]
- \qquad\qquad \Rightarrow 
-       \tptr{\tstruct{T}}{m} \sqsubseteq \tarrayptr{b_l}{b_h}{\tnat}{m}
-    \end{array}
-\\[3em]
-\text{  Struct Type Rule: }\\[0.5em]
-    \end{array}
-$
-
-{
-\begin{mathpar}
-  \inferrule [T-Struct]
-  {\Gamma; \Theta \vdash_m e :  {\tptr{\tstruct{T}}{m}} \\
-    D(T) = fs \\ fs(f)=\tau_f}
-  {\Gamma; \Theta \vdash_m \eamper{e}{f} : \tptr{\tau_f}{m}}
-
-\end{mathpar}
-}
-{
-$\begin{array}{l}
-\text{Struct Semantics: }
-\end{array}
-$
-}
-{
-\begin{mathpar}
-  \inferrule [S-StructChecked]
-  {n > 0 \\ D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
-  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\cmode}})}
-
-  \inferrule [S-StructNull]
-  {n = 0}
-  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\enull)}
-
-  \inferrule [S-StructUnChecked]
-  {D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
-  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\umode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\umode}})}
-
-\end{mathpar}
-}
-}
-  \caption{\lang Struct Definitions}
-  \label{fig:checkc-struct}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-
-    \inferrule[]
-    {}
-    {\Gamma \vdash n}
-
-    \inferrule[]
-    {x:\tint \in \Gamma}
-    {\Gamma \vdash x + n}
-
-    \inferrule[]
-    {\Gamma \vdash b_l\\
-    \Gamma \vdash b_h}
-  {\Gamma \vdash (b_l,b_h)}
-
-  \inferrule[]
-  {}
-  {\Gamma \vdash \tint}
-
-  \inferrule[]
-  {\Gamma \vdash \bvar \\
-  \Gamma \vdash \tau}
-  {\Gamma \vdash \tptr{\tallarrayb{\bvar}{\tau}}{m}}
-
-  \inferrule[]
-  {\Gamma \vdash \tau}
-  {\Gamma \vdash \tptr{\tau}{m}}
-
-  \inferrule[]
-  {T \in D}
-  {\Gamma \vdash \tptr{\tstruct{T}}{m}}
-  \end{mathpar}
-}
- \caption{Well-formedness for types and bounds}
-\label{fig:wftypesandbounds}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    {\Gamma \vdash \overline{x}:\overline{\tau} \\
-      \Gamma[\overline{x} \mapsto \overline{\tau}] \vdash \tau \\
-    \Gamma[\overline{x} \mapsto \overline{\tau}]; \Theta  \vdash_{\cmode} e : \tau}
-    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e}
-
-    \inferrule[]
-    {}
-    {\Gamma \vdash \cdot}
-
-    \inferrule[]
-    {\Gamma \vdash \tau \\
-    \Gamma[x \mapsto \tau] \vdash \overline{x}:\overline{\tau}}
-    {\Gamma \vdash x:\tau, \overline{x}:\overline{\tau}}
-  \end{mathpar}
-}
- \caption{Well-formedness for functions}
-\label{fig:wffunctions}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    {\Gamma \vdash \tau}
-    {\Gamma \vdash \tau~\mathtt{f} }
-
-    \inferrule[]
-    {\Gamma \vdash \tau\\
-     \Gamma \vdash fs}
-    {\Gamma \vdash \tau~\mathtt{f}; fs }
-
-  \end{mathpar}
-}
- \caption{Well-formedness for structs}
-\label{fig:wfstructs}
-
-{\small
-  \begin{mathpar}
-
-    \inferrule[]
-    {\Gamma[\overline{x} \mapsto \overline{\tau}]; \emptyset  \vdash
-      e \gg \dot e : \tau}
-    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e \gg
-      (\overline{x})\;\dot e}
-  \end{mathpar}
-}
- \caption{Compilation rules for functions}
-\label{fig:compilefunctions}
-\end{figure}
-
-\subsection{The Compilation Rules}\label{appx:comp1}
-Fig.~\ref{fig:syntaxerased} and Fig.~\ref{fig:semanticserased} shows
-the syntax for $\elang$, the target language for compilation. We syntactically
-restrict the expressions to be in A-normal form because that is the
-type of expression our compiler produces. To allow explicit runtime
-checks, we include $\ebounds$ and $\enull$ as part of $\elang$
-expressions which, once evaluated, result in an corresponding error state. $\eassignstack{x}{\dot a}$ is a new syntactic form that
-modifies the stack variable $x$ with the result of $\dot a$. It is
-essential for bounds widening. $\leq$ and $-$ are introduced
-to operate on bounds and decide whether we need to halt with a bounds error
-or widen a null-terminated string.
-
-\begin{figure}[h]
-{\small
-  \[\begin{array}{llcl}
-      \text{Atoms}       & \dot a & ::= & n \mid x \\
-
-      \text{C-Expressions} & \dot c & ::= & \dot a \mid \estrlen{\dot a}  \mid \emalloc {\dot  a}  \mid \ecall{f} {\overline{\dot a}}  \\
-                         & & \mid &  \mid  \dot a \circ \dot a \mid \estar{\dot a} \\
-                         & & \mid & \eassign{\dot a}{\dot a} \mid \eassignstack{x}{\dot a}  \mid \eif{\dot a}{\dot e}{\dot e} \\
-                         & & \mid & \ebounds \mid \enull \\
-      \text{Expressions}  & \dot e & ::= & \dot c \mid \elet{x}{\dot c}{\dot e}\\
-      \text{Binops} & \circ & ::= & + \mid - \mid \leq  \\
-      \text{Closure} & \dot C & ::= & \hole \mid \elet{x}{\dot a}{\dot C} \\
-       & & \mid & \eif{\dot a}{\dot e}{\dot C} \mid \eif{\dot a}{\dot    C}{\dot e} \\
-      \text{Bounds Map} & \rho & \in & \texttt{Var} \rightharpoonup \texttt{Var} \times \texttt{Var}
-  \end{array}
-  \]
-}
-  \caption{\elang{} Syntax}
-  \label{fig:syntaxerased}
-{\small
-  \[\begin{array}{lll}
-    \dot \mu & ::= & n \mid \bot \\
-    \dot c & ::= & \ldots  \mid \eret{x}{\dot \mu}{\dot e} \\
-    \dot H & \in & \mathbb{Z} \rightharpoonup \mathbb{Z} \\
-    \dot r & ::= & \dot e \mid \enull \mid \ebounds \\
-    \dot E & ::= & \hole \mid  \elet{x}{\dot E}{\dot e}  \mid \eret{x}{i}{\dot E} \\
-             & \mid & \eif{\dot E}{\dot e}{\dot e} \mid \estrlen{\dot E} \\
-             & \mid & \emalloc{\dot E} \mid \ecall{f}{\overline{\dot
-                      E}} \mid \dot E \circ \dot a \mid n \circ \dot E \\
-             & \mid & \estar{\dot E} \mid \eassign{\dot E}{\dot a}
-                      \mid \eassign{n}{\dot E} \mid
-                      \eassignstack{x}{\dot E} \\
-      \overline{\dot E } & ::= & \dot E \mid n,\overline{\dot E} \mid
-                                 \overline{\dot E}, \dot a
-
-  \end{array} \]
-}
-  \caption{\elang Semantic Defs}
-  \label{fig:semanticserased}
-\end{figure}
-
-$\elang$ does not include any annotations. We remove structs from
-$\elang$ because we can always statically convert expressions of the form
-$\eamper{n:\tau}{f}$ into $n + n_f$, where $n_f$ is the statically
-determined offset of $f$ within the struct. We ellide the semantics of
-$\elang$ because it is self-evident and mirrors the semantics
-$\lang$. The difference is that in $\elang$, only $\ebounds$ and
-$\enull$ can step into an error state. All failed dereferences and
-assignments would result in a stuck state and therefore we rely on the
-compiler to explicitly insert checks for checked pointers.
-
-Fig.~\ref{fig:compilation} and Fig.~\ref{fig:compilation2}
- shows the rules for the compilation judgment for expressions,
-\[\Gamma;\rho \vdash e \gg \dot C, \dot a\]
-The judgment is presented differently from the one in
-Sec.~\ref{sec:compilation}, which was simplified for presentation purposes. First,
-we remove $\Theta$ and $m$ because these
-parameters are only used for checking and have no
-impact on compilation. Second, the judgment includes two
-outputs, a closure $\dot C$ and an atom expression $\dot a$, instead of
-a single $\elang$ expression $\dot e$. $\dot C$ can be intuitively understood as a
-partially constructed program or context. Whereas $\dot E$ is used for
-evaluation, $\dot C$
-is used purely as a device for compilation. As an example,
-when compiling $(1 : \tint) \plus (2 : \tint)$, 
-we would first create a fresh variable $x$, and then produce two outputs:
-\[    \dot C = \elet{x}{1 \plus 2}{\hole}\]
-\[    \dot a = x\]
-To obtain the compiled expression $\dot e$, we plug $\dot a$ into
-$\dot C$ using the usual notation $\dot C[\dot a]$. We can also use $\dot C$ to represent runtime checks, which
-usually take the form $\elet{x}{\dot c}{\hole}$, where $\dot c$
-contains the check whose evaluation must not trigger $\ebounds$ or $\enull$ for the program to
-continue (see Fig.~\ref{fig:compilationhelpers} for the metafunctions that create
-those checks).
-
-This unconventional output format enables us to separate the
-evaluation of the term and the computation that relies on the term's
-evaluated result. Since effects and reduction (except for variables) happen only
-within closures, we can precisely control the order in which effects
-and evaluation happen by composing the contexts in a specific order.
-Given two closures $\dot C_1$
-and $\dot C_2$, we write $\dot C_1[\dot C_2]$ to denote the meta
-operation of plugging $\dot C_2$ into $\dot C_1$. We also use
-$\dot C_{a;b;c}$ as a shorthand for $\dot C_a[\dot C_b[\dot C_c]]$. In
-the \textsc{C-Ind} rule,
-we first evaluate the expressions that correspond to $e_1$ and $e_2$
-through $\dot C_1$ and $\dot C_2$, and then perform a null check and
-an addition through $\dot C_n$ and $\dot C_3$. Finally, we dereference
-the result through $\dot C_4$ before returning the pair $\dot C_4,\dot
-x_4$, propagating the flexibility to the compilation rule that recursively
-calls \textsc{C-Ind}.
-
-Fig.~\ref{fig:compilationhelpers} shows the metafunctions that create
-closures representing dynamic checks. These functions first examine
-whether the pointer is a checked. If the pointer is unchecked, an empty closure $\hole$
-will be returned, because there is no need to perform a check. For bounds
-checking, there is a special case for NT-array pointers, where the
-bounds are retrived from the ghost variables (found by
-looking up $\rho$) on the stack  rather than
-using the bounds specified in the type annotation. This is how we
-achieve the same precise runtime behavior as $\lang$ in our compiled expressions.
-
-Fig.~\ref{fig:compilationhelpers2} shows the metafunctions related to
-bounds widening. $\vdash_{extend}$ takes $\rho$,
-a checked NT-array pointer variable $x$, and its bounds $(b_l,b_h)$ as inputs, and returns an
-extended $\rho'$ that maps $x$ to two fresh variables $x_l$, $x_h$,
-together with a closure $\dot C$ that initializes $x_l$ and $x_h$ to
-$b_l$ and $b_h$ respectively. This function is used in the
-\textsc{C-Let} rule to extend $\rho$ before compiling the body of the
-$\elettext$ binding. The updated $\rho'$ can be used for generating
-precise bounds checks, and for inserting expressions that can potentially
-widen the upper bounds, as seen in the $\vdash_{widenstr}$
-metafunction used in the \textsc{C-Str} compilation rule.
-
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    { x_l, x_h = \fresh \\
-      \rho' = \rho[x \mapsto (x_l, x_h)]\\
-    \dot C = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}}
-  {\cextend{\rho}{x}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot C, \rho'}\\}
-
-  \inferrule[]
-  {x_l,x_h = \rho(x) \\
-    x_w = \fresh \\
-  \dot C = \elet{x_w}{\eif{x_h}{0}{\eassignstack{x_h}{1}}}{\hole}}
-{\ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C}}
-
-\inferrule[]
-{e \notin dom(\rho)}
-{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{m}}{\hole}}
-
-\inferrule[]
-{x_l,x_h = \rho(e) \\
-  x_a = \fresh \\
-  \dot C = \elet{x_a}{\eif{\dot a \leq x_h}{0}{\eassignstack{x_h}{\dot a}}}{\hole}}
-{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{\cmode}}{\dot C}}
-
-  \end{mathpar}
-
-}
- \caption{Metafunctions for widening}
-\label{fig:compilationhelpers2}
-\end{figure*}
-
-\begin{figure*}[h]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    {x = \fresh \\
-      \dot C = \elet{x}{\eif{\dot a}{0}{\enull}}{\hole}}
-    {\echecknull{\dot a}{c}{\dot C}}
-
-    \inferrule[]
-    {}
-    {\echecknull{\dot a}{u}{\hole}} \\
-
-    \inferrule[]
-    {}
-    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}
-
-    \inferrule[]
-    { x_l,x_h = \rho(e) \\
-      x_{cl},x_{ch} = \fresh\\
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
-    }
-    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}
-
-    \inferrule[]
-    { e \notin dom(\rho) \\
-      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
-      \dot C_l = \elet{x_l}{b_l}{\hole}\\
-      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
-    }
-    {\echeckbounds{\rho}{e}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}
-
-    \inferrule[]
-    { e \notin dom(\rho) \\
-      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
-      \dot C_l = \elet{x_l}{b_l}{\hole}\\
-      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
-    }
-    {\echeckbounds{\rho}{e}{\tarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}
-
-    \inferrule[]
-    {}
-    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}
-
-    \inferrule[]
-    { x_l,x_h = \rho(e) \\
-      x_{cl},x_{ch} = \fresh\\
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
-    }
-    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}
-
-    \inferrule[]
-    { e \notin dom(\rho) \\
-      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
-      \dot C_l = \elet{x_l}{b_l}{\hole}\\
-      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
-    }
-    {\echeckboundsw{\rho}{e}{\tallarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}
-
-\inferrule[]
-{e \notin dom(\rho) \\
-  x_l,x_l',x_h,x_h' = \fresh \\
-  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
-  \dot C_2 = \elet{x_l'}{b_l'}{\elet{x_h'}{b_h'}{\hole}}\\
-  \dot C_3 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
-  \dot C_4 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
-}
-{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3;4}}}
-
-\inferrule[]
-{x_l',x_h' =\rho(e) \\
-  x_l,x_h = \fresh \\
-  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
-  \dot C_2 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
-  \dot C_3 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
-}
-{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3}}}
-  \end{mathpar}
-}
-\caption{Metafunctions for dynamic checks}
-\label{fig:compilationhelpers}
-\end{figure*}
-
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-
-    \inferrule[C-Const]
-      {}
-      {\Gamma;\rho\vdash \evalue{n}{\tau} \gg \hole, n : \tau}
-
-    \inferrule[C-Var]
-      {x : \tau \in \Gamma}
-      {\Gamma;\rho \vdash x \gg \hole, x : \tau}
-
-    \inferrule[C-Cast]
-              {\Gamma;\rho \vdash e \gg  \dot C , \dot a : \tau'}
-              {\Gamma;\rho \vdash \ecast{\tau}{e} \gg \dot C, \dot a  : \tau}
-
-      \inferrule[C-DynCast]
-      {\Gamma;\rho \vdash e \gg \dot C_1, \dot a : \tptr{\tallarrayb{\bvar'}{\tau}}{m} \\
-      \echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\tptr{\tallarrayb{\bvar'}{\tau}}{m}}{\dot C_{b}}}
-      {\Gamma;\rho \vdash \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e}
-            \gg \dot C_{1;b}, \dot a : \tptr{\tallarrayb{\bvar}{\tau}}{m}}
-
-      \inferrule[C-Str]
-      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} \\
-        \echecknull{\dot a_1}{m}{\dot C_{n}} \\
-        \echeckbounds{\rho}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{0}{\dot C_{b}} \\
-        x_2 = \fresh \\
-          \dot C_2 = \elet{x_2}{\estrlen{\dot a_1}}{\hole} \\
-      \ewidenstrlen{\rho}{e}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{\dot C_{w}} } 
-                {\Gamma;\rho \vdash \estrlen{e} \gg \dot C_{1;n;b;2;w}, x_2 : \tint}
-
-      \inferrule[C-LetStr]
-      {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\
-        \Gamma; \rho \vdash \estrlen{y} \gg \dot C_1, \dot a_1 : \tint\\
-        \dot C_2 = \elet{x}{\dot a_1}{\hole} \\
-        \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}]];\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau}
-{\Gamma;\rho \vdash \elet{x}{\estrlen{y}}{e} \gg \dot C_{1;2;3},\dot a_3 : \tau}
-
-      \inferrule[C-If ]
-      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
-        \Gamma; \rho \vdash e_1 \gg \dot C_2, \dot a_2 : \tau_2 \\
-        \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
-        x_4 = \fresh \\
-        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_2[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}
-      }
-        {\Gamma; \rho \vdash \eif{e_1}{e_2}{e_3} \gg \dot C_4, x_4 : \tau_2 \sqcup \tau_3}
-
-        \inferrule[C-IfNT]
-        {\Gamma; \rho \vdash x : \tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c} \\
-          b_h = 0 \Rightarrow \Gamma' = \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}] \\
-          b_h \neq 0 \Rightarrow \Gamma' = \Gamma \\
-          \Gamma; \rho \vdash \estar{x} \gg \dot C_1, \dot a_1 : \tau_1\\
-          \Gamma'; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau_2\\
-          \Gamma;\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
-        \ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C_{w}} \\
-        x_4 = \fresh \\
-        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_{2;w}[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}}
-        {\Gamma;\rho \vdash \eif{\estar{x}}{e_1}{e_2} \gg x_4, \dot C_4 : \tau_1 \sqcup \tau_2}
-
-   \inferrule[C-Let]
-    { (x\in \fv(\tau') \Rightarrow e_1 \in Bound) \\
-        \Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tau_1 \\ 
-        \cextend{\rho}{x}{\tau_1}{\dot C_2,\rho'}\\
-        \dot C_3 = \elet{x}{\dot a_1}{\hole}\\
-        \Gamma[x\mapsto \tau];\rho' \vdash e_4 \gg \dot C_4, \dot a_4  : \tau_4}
-    {\Gamma;\rho' \vdash \elet{x}{e_1}{e_4} \gg \dot C_{1;2;3;4},\dot a_4 : \tau_4[\tau_1 = \tint \Rightarrow x \mapsto e_1]}
-
-  \inferrule[C-Ret]
-    {\Gamma(x)\neq \bot \\
-      \Gamma;\rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
-      x_2 = \fresh \\
-      \mu \gg \dot \mu \\
-    \dot C_2 = \elet{x_2}{\eret{x}{\dot \mu}{\dot C_1[\dot a_1]}}{\hole}}
-  {\Gamma;\rho \vdash \eret{x}{\mu}{e} \gg \dot C_2, x_2 : \tau}
-
-    \inferrule[C-Fun]
-    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\
-(\forall e_i\in\overline{e}\;\;\tau_i\in\overline{\tau}\;.\; \Gamma; \rho \vdash {e_i} \gg \dot C_i, \dot a_i  : \\ \tau_i' \wedge
-      \tau_i' \sqsubseteq 
-      \tau_i[\overline{e} / \overline{x}]) \\
-      x_f = \fresh \\
-      \dot C_f = \elet{x_f}{f(\overline{a})}{\hole}}
-    {\Gamma; \rho \vdash f(\overline{e}) \gg \overline{\dot C}[\dot C_f] , x_f  : \tau[\overline{e} / \overline{x}]}
-
-  \inferrule[C-Def]
-  {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m} \\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    x_2 = \fresh \\
-    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}
-   }
-   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;2}, x_2 : \tau}
-
-   \inferrule[C-DefArr]
-   {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tallarrayptr{b_l}{b_h}{\tau}{m} \\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    \echeckbounds{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
-    x_2 = \fresh \\
-    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}}
-   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;b;2},x_2 : \tau}
-
-    \inferrule[C-Mac]
-    {\dot C_{1}, \dot a_1 = \esizeof{\omega} \\
-      x_2 = \fresh \\
-    \dot C_2 = \elet{x_2}{\emalloc{\dot a_1}}{\hole}}
-     {\Gamma; \rho \vdash \emalloc{\omega} \gg \dot C_{1;2}, x_2 : \tptr{\omega}{\cmode}}
-
-  \end{mathpar}
-}
-\caption{Compilation}
-\label{fig:compilation}
-\end{figure*}
-
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[C-Add]
-              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tint \\
-                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint\\
-                x_3 = \fresh \\
-              \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}}
-              {\Gamma; \rho \vdash \dot C_3, x_3 : \tint }
-
-   \inferrule[C-Ind] 
-              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
-                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
-                \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-                \echeckbounds{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
-                x_3, x_4 = \fresh\\
-                \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}\\
-                \dot C_4 = \elet{x_4}{\estar{x_3}}{\hole}
-                }
-                {\Gamma; \rho \vdash \estar{(\ebinop{e_1}{e_2})} \gg \dot C_{1;2;n;3;b;4} , x_4: \tau}
-
-  \inferrule[C-Assign]
-  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m'} \\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
-    \tau'\sqsubseteq \tau \\
-    x_3 = \fresh \\
-    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
-  }
-  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;3}, x_3 : \tau}
-
-  \inferrule[C-AssignArr]
-  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    \echeckboundsw{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
-    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
-    x_3 = \fresh \\
-    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
-    \tau'\sqsubseteq \tau}
-  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;b;3},x_3 : \tau}
-
-    \inferrule[C-IndAssign]
-    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
-      \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
-      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-      \echeckboundsw{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
-      \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau' \\
-      x_4, x_5 = \fresh \\
-      \dot C_4 = \elet{x_4}{\dot a_1 \plus \dot a_2}{\hole}\\
-      \dot C_5 = \elet{x_5}{\eassign{x_4}{x_3}{\hole}}
-      \tau'\sqsubseteq \tau}
-    {\Gamma; \rho \vdash \eassign{(e_1 \plus e_2)}{e_3} \gg \dot C_{1;2;n;3;4;b;5} : \tau}
-
-    \inferrule [C-Struct]
-    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1  :  {\tptr{\tstruct{T}}{m}} \\
-      D(T) = \tau_0~f_0\ldots; \tau_j~f; ...  \\
-      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-      x_2 = \fresh\\
-    \dot C_2 = \elet{x_2}{\dot a_1 \plus \dot j}{\hole}}
-    {\Gamma; \rho \vdash \eamper{e_1}{f} \gg \dot C_2, x_2  : \tptr{\tau_f}{m}}
-
-    \inferrule[C-Unchecked]
-              {\Gamma;\rho \vdash e \gg \dot C, \dot a : \tau}
-              {\Gamma;\rho \vdash \eunchecked{e} \gg \dot C, \dot a : \tau}
-
-  \end{mathpar}
-
-}
-\caption{Compilation (continued)}
-\label{fig:compilation2}
-\end{figure*}
-
-\end{document}

### 2022-07-05 rlbox_paper/sections/introduction.tex
Adding paper

\section{Introduction}
Vulnerabilities due to memory corruption, especially spatial memory corruptions, are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe} despite a large body of work that tries
to prevent them~\cite{song2019sanitizing}.
Vulnerabilities due to memory corruption, especially spatial memory corruptions, are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe} despite a large body of work that tries
to prevent them~\cite{song2019sanitizing}.
One way to handle this is to convert these programs into safe languages such as Java. However, it requires considerable development effort. Furthermore, the performance overhead prohibits them from being used in performance-critical settings such as Operating Systems.

With the increased support for high-performant safe system languages such as Rust and Go, we can have tools that convert C to a memory-safe language, e.g., Rust, a promising memory-safe systems
language~\cite{rustlang}. Unfortunately, Rust is too different from C to constitute a practical target~\cite{zeng2019identifying}, at least for now: the (best of breed) \texttt{c2rust} tool~\cite{c2rust,c2rusttalk} transliterates C to \emph{unsafe, non-idiomatic} Rust, and recent
work~\cite{emre2021c2rust} is able to take only small steps to close the safety gap and still requires considerable work from developers.

There are other retrofitting techniques such as CCured~\cite{necula2005ccured},
Softbound~\cite{nagarakatte2009softbound}, Low Fat
pointers~\cite{duck2016heap}, and Address Sanitizer
(ASAN)~\cite{serebryany2012addresssanitizer} aim to enforce spatial safety
automatically~\emph{without any developer effort}, by analyzing a C program and compiling it to include
run-time safety checks. Unfortunately, the resulting run-time overhead is too high
for deployment (between 60\%-200\%), especially in low-powered IoT devices where the usage of C-based system software (e.g., FreeRTOS) is still rampant.
~\fig{fig:memprotections} shows the summary of these existing approaches.

\input{figures/existingmem}

\emph{Checked C}~\cite{tarditi2018checked} as a
practical target to which to convert legacy/active C code. 
Checked C
extends C with \emph{checked pointer
  types} which are restricted by the compiler to spatially safe uses (temporal safety is underway~\cite{checkedc:temporal}).
Such pointers have one of three possible types, \ptrT{|$T$|}, \arrT{|$T$|},
or~\ntarrT{|$T$|} (\ptr, \arr, and \ntarr{} for short), representing a
pointer to a single element, array of elements, or null-terminated
array of elements of type $T$, respectively. The latter two have an
associated \emph{bounds annotation}; e.g., a declaration \arrT{int}
\code{p} : \code{count(n)} says that \code{p} is a pointer to an \code{int} array
whose size is \code{n}.
%
Checked C's Clang/LLVM-based compiler represents checked
pointers as system-level memory words, i.e., without
``fattening'' metadata, ensuring backward compatibility.
The compiler uses these bounds annotations to add dynamic checks prior to
checked pointer accesses, to prevent spatial safety violations.
These run-time checks can often be proved redundant and removed by LLVM, yielding good performance.
%, and leverages LLVM to optimize away run-time
%checks, yielding good performance.
%
Tarditi et al.~\cite{tarditi2018checked} reported average
run-time overheads of 8.6\% on a small benchmark suite, and
\cite{duanrefactoring} found essentially no overhead when running
Checked C-converted portions of the FreeBSD kernel.

However, converting existing C programs to Checked C requires considerable effort~\cite{duanrefactoring}. 
Recently, Machiry et al. developed 3C~\cite{machiry2022c} that tries to automatically convert a program to Checked C by adding appropriate pointer annotations.
However, as described in 3C, complete automated conversion is infeasible and consequently requires the developer to convert some code regions manually.
Thanks to the backward compatibility of Checked C, the partially annotated program still enjoys the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
%
Although the blameless proof exists~\cite{ruef2019achieving}, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
For instance, a checked function expects a pointer to a buffer of five elements, but an unchecked region can call this function with a pointer to a buffer of three elements. This results in a spatial safety violation in the Checked region, but of course, the blame or the root cause is in the unchecked region.
Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions can take down the complete program despite having checked regions.
We need an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions.
Such an isolation mechanism should have low/no overhead and be backward compatible (i.e., pointers should have no additional metadata), preserving the advantages of Checked C.

In this work, we propose~\systemname, an isolation mechanism that enables unchecked regions or, in general, any arbitrary set of functions to be executed in an isolated address space (\ucregion).
So that checked regions can still enjoy true spatial memory safety even in the presence of unchecked or unsafe regions.
\systemname extends Checked c using~\textbf{tainted} (\taintt) types, representing the code and data that belong to the~\ucregion.

\noindent
\textbf{Automated Partitioning.} A pointer (checked or unchecked) can be marked as tainted; similarly, a function can be marked as tainted.
Along with enforcing spatial safety constraints of checked tainted pointers, the~\systemname compiler will bundle all the tainted types (functions and pointers) so that they remain in an isolated sandbox (\ucregion), and non-tainted types will remain in the process address space (\cregion).

\noindent
\textbf{Memory Access Restrictions.} The functions in the~\ucregion can only access tainted pointers, i.e., memory within the corresponding sandbox region.
However, the functions in~\cregion have no-such restrictions and can access both c- and uc-region's memory.
Similarly, the functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.

\noindent
\textbf{Marshaling free communication.} As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
We avoid marshaling as all tainted buffers (i.e., regions pointed by tainted pointers) are allocated in~\ucregion and thus can be accessed in both c- and uc-region.

\noindent
\textbf{Additional checks.} Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
However, our compiler avoids these attacks by ensuring, either statically or through dynamic checks, that tainted pointers in~\cregion can only point to~\ucregion address space, and our type system enforces that tainted pointers cannot be assigned to non-tainted pointers.

\iffalse
Recently, Narayan et al. proposed RLBox, which enables sandboxing libraries by making data- and control-flow at the sandbox interface explicit, using types.
Specifically, RLBox defines a tainted type (\code{tainted<...>}), which is a C++ template class and can be used to have any type tainted (e.g.,~\code{tainted<int>},~\code{tainted<float*>}, etc).
The tainted types represent values or pointers to buffers that can be accessed from a sandbox.
The program or unsandboxed code can use only tainted type variables to communicate (i.e., parameters and return values) with sandboxed functions.
Although RLBox provides a practical technique for sandboxing libraries, it has the following drawbacks:
\begin{itemize}
\item\textbf{Marshalling Cost.} Any non-scalar value (e.g., a pointer to a struct) that need to be passed to a sandboxed function should be first copied into a tainted buffer (\emph{marshalling}). Then the corresponding tainted value (i.e., tainted pointer) should be passed to the sandboxed function.
This can lead to considerable overhead for large objects.
\item\textbf{Developer Written Sanitization.} The unsandboxed code can access tainted type variables only through certain special (i.e., sanitization) routines.
This forces developers to sanitize every value that comes from sandboxed library functions.
However, the correctness of the sanitization is entrusted to the developer.
Consequently, any mistake in sanitization methods can still lead to vulnerabilities.
\item\textbf{Library Level Sandboxing.} The unit of sandboxing in RLBox is the whole library. However, we want to sandbox a set of functions in a program.
\end{itemize}
\fi

% How to highlighgt code in minted
% highlightlines={1-3}, highlightcolor=red
\begin{listing}[t!]
  \begin{tabular}{c c}
    \begin{minipage}[b]{.22\textwidth}
\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/orig1.c}
    \end{minipage} &
    \begin{minipage}[b]{.22\textwidth}
\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/orig2.c}
    \end{minipage} \\
   (a) Original C code & (b) After initial conversion.\\
  \end{tabular}
\caption{(Contrived) Example demonstrating various phases of.}
\label{lst:comb}
\end{listing}

\begin{listing}[t!]
\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/firstrun.c}
\caption{(Contrived) Example demonstrating various phases of.}
\label{lst:initialconv}
\end{listing}

\begin{listing}[t!]
\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/humanannotations.c}
\caption{(Contrived) Example demonstrating various phases of.}
\label{lst:humantaint}
\end{listing}

\begin{listing}[t!]
\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/humanadjustments.c}
\caption{(Contrived) Example demonstrating various phases of.}
\label{lst:humanadjust}
\end{listing}

\input{figures/taintedsplit}

%\begin{listing*}[t!]
%\inputminted[linenos, mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/originalprogram.c}
%\caption{Simple server example.}
%\label{lst:comb}
%\end{listing*}

### 2022-07-05 rlbox_paper/sections/relatedwork.tex
Adding paper

\section{Related Work}
% Safe C dialects and Checked C
Existing instrumentation based memory safety retrofitting techniques~\cite{duck2016heap,serebryany2012addresssanitizer,
  nagarakatte2009softbound,kendall1983bcc,steffen1992adding} add significant overhead~\cite{duck2016heap,serebryany2012addresssanitizer,
  nagarakatte2009softbound,kendall1983bcc,steffen1992adding}.
CCured~\cite{necula2005ccured}  and its successor Deputy~\cite{condit2007dependent}
aims to reduce overhead by adding checks only where
needed.
Nonetheless, unconverted pointers in CCured require extra metadata for pointers causing backward compatibility issues.
On the other hand, Checked C has no overhead and also is backward compatible.
% RLBox and related works
\systemname provides a Software Fault Isolation (SFI)~\cite{tan2017principles} mechanism for arbitrary functions in a program.
It uses a sandbox technique to enforce isolation.
There has been a considerable amount of work done in the area, especially for browsers~\cite{minsfi,barth-et-al:chromium:08, site-isolation-usenix}, which constantly runs code from potentially untrusted entities.
However, all these techniques use process/program-level isolation~\cite{site-isolation-usenix} and are not suited for seamlessly isolating parts of the same program.

Similar to RLBox, there are other works that provide sandboxing APIs.
These sandboxing mechanisms have different performance trade-offs \cite{wahbe-et-al:sfi:sosp93, nacl-amd64, minsfi, wasm,erim, tan-sfi-survey, shu-isolation-survey}. 
However, unlike RLBox, which enforces constraints 
through~\code{tainted} types, the other works have no such restrictions and expect the developer to take care of attacks from sandboxed regions explicitly.
% what does CheckCBox does on top of RLBox? What it provides? We should add that.
\systemname borrows the~\code{tainted} types from RLBox and merges them with Checked C types.
This enables us to have a more efficient sandboxing mechanism..what else?
\aravind{Fill this part.}

Our workflow of interactive type annotations has been explored before and is shown as an effective technique to assist developers in code conversion~\cite{machiry2022c} and vulnerability finding tasks~\cite{naik2021sporq}.

### 2022-07-05 rlbox_paper/examples/firstrun.c

@@ -1,0 +1,13 @@
+|\entrypoint| // Entry point
+void server_loop(int sock_fd) {
+ ...
+ _Array_ptr<struct queue> p : count(b_z) = NULL;
+ ...
+}
+
+int process_req2(_Array_ptr<char> msg: count(m_l), 
+                 size_t m_l) {
+ ...
+ char anum _Checked[MAX_MSG_SIZE] = {0};
+ ...
+}

### 2022-07-05 rlbox_paper/examples/humanadjustments.c

@@ -1,0 +1,9 @@
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_\_tainted} \usermods|;
+ ...
+}
+
+int process_req2(|\usermods| |\textcolor{taintcolor}{\_t\_Array\_ptr}|<char> msg: count(m_l), 
+                 size_t m_l) {
+ ...
+}

### 2022-07-05 rlbox_paper/examples/orig1.c

@@ -1,0 +1,34 @@
+|\entrypoint| // Entry point
+void server_loop(int sock_fd) {
+ unsigned b_z;
+ struct queue *p;
+ ...
+ while(1) {
+  ...
+  p = malloc(sizeof(struct queue) * b_z);
+  ...
+  if (handle_request(sock_fd)) {
+    ...  
+  }
+ }
+}
+
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE];
+ int rc = -1;
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, 
+                  MAX_MSG_SIZE);
+ if (r_len > 0) {
+  switch(buff[0]) {
+   case REQ1: 
+    rc = process_req1(buff); 
+    break;
+   case REQ2: 
+    rc = process_req2(buff); 
+    break;
+   ...
+  }
+ }
+ return rc;
+}

### 2022-07-05 rlbox_paper/examples/orig2.c

@@ -1,0 +1,34 @@
+size_t read_msg(int sock_fd, char *msg, 
+                size_t sz) {
+ size_t nr;
+ nr = read(sock_fd, |\rootcause|(void*)msg, sz);
+ ...
+ |\complexcode| // complex code
+ return nr;
+}
+
+int process_req1(char *msg, size_t m_l) {
+ int rc = -1, i;
+ if (m_l > MIN_SIZE) {
+   |\realbug| sscanf(msg, "%d", &i);
+   msg[i] = ...
+ }
+ |\complexcode| // complex code
+ return rc;
+}
+
+int process_req2(char *msg, size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}

### 2022-07-11 sandp23/background.tex
update.

@@
-\section{Overview}\label{sec:overview}
-
-Give an overview of Checked-C and RLBox.
-
-\begin{enumerate}
-
-\item discuss Checked-C and its usage.
-
-\item discuss RLBox.
-
-\item A running example (web-browser with third party libraries).
-
-\end{enumerate}
-
-\section{Key Updates}\label{sec:key-updates}
+\section{Overview and Transcendence}\label{sec:overview}
 
@@
 
-\begin{enumerate}
+\begin{figure}[t]
+{\small
+  \begin{lstlisting}[xleftmargin=4 mm]
+//in checked region
 
-\item Pointing out that the current Checked-C and C formalism is not enough to ensure the security of linking web-browser with third party libraries.
-Pointing out that even though Checked-C previously has blame theorem, but it still allows program crashing, and our new work can guarantee that every crashes can be detected and handled (non-crashing theorem).
+int compare_1(nt_array_ptr<char> x: count (0),
+  nt_array_ptr<char> y : count (0)) { 
+  int len_x = strlen(x);
+  int len_y = strlen(y);
+  return sum(x,len_x) < sum(y,len_y);
+  }
+...
 
-\item Using an example to highlight the overhead of RLBox, and points out the advantage of building RLBox mechanism in a compiler.
+int stringsort(
+ nt_array_ptr<nt_array_ptr<char>> s : count (n),
+ ptr<(int)(nt_array_ptr<char>,
+  nt_array_ptr<char>)> cmp, int n) {
+  int i, j, gap;
+  int didswap;
 
-\item Discuss function pointer formalism and points out the potential security issues without a proper handling of function pointer set up.
-Mentioning what happens in RLBox and other tools.
-Pointing out that this is the first work formalizing function pointers.
+  for(gap = n / 2; gap > 0; gap /= 2) {
+    {
+    do  {
+        didswap = 0;
+        for(i = 0; i < n - gap; i++)
+           {
+             j = i + gap;
+             if((*cmp)(s[i], s[j]) > 0)
+             {
+                int len = strlen(s[i]);
+                nt_array_ptr<char> 
+                  tmp : count (len) = s[i];
+                s[i] = s[j];
+                s[j] = tmp;
+                didswap = 1;
+             }
+           }
+        } while(didswap);
+    }
+  }
 
-\end{enumerate}
+  return 0;
+}
+  \end{lstlisting}
+}
+\caption{Checked stringsort Code}
+\label{fig:checkedc-example-1}
+\end{figure}
+
+\begin{figure}[t]
+{\small
+  \begin{lstlisting}[xleftmargin=4 mm]
+//in checked region, tainted version
+int tainted_compare_1(
+  nt_array_tptr<char> x : count (0),
+  nt_array_tptr<char> y : count (0)) {
+  int len_x = strlen(x);
+  int len_y = strlen(y);
+  nt_array_ptr<char> tx : count (len_x)
+   = malloc(nt_array<char>, len_x);
+  nt_array_ptr<char> ty : count (len_y)
+   = malloc(nt_array<char>, len_y);
+  safe_memcpy(tx,x,len_x);
+  safe_memcpy(ty,y,len_y);
+  return compare_1(tx,ty);
+}
+...
+
+//calling the function turns
+//an unchecked region to a checked region.
+int tainted_stringsort(nt_array_tptr
+     <nt_array_tptr<char>> s : count (n),
+ tptr<(int)(nt_array_tptr<char>,
+  nt_array_tptr<char>)> cmp, int n) {
+  int i;
+  nt_array_ptr<nt_array_ptr<char>> p : count (n)
+    = malloc(nt_array<nt_array_ptr<char>>, n);
+  for(i = 0; i < n; i++) {
+    int len = strlen s[i];
+    nt_array_ptr<char> tmp : count (len)
+      = new malloc(nt_array<char>, len);
+    safe_memcpy(tmp,s[i],len);
+    p[i] = tmp;
+  }
+  ptr<(int)(nt_array_ptr<char> : count (0),
+  nt_array_ptr<char> : count (0))>
+    cfun = find_check(cmp);
+
+  return stringsort(p,cfun);
+}
+  \end{lstlisting}
+}
+\caption{Tainted stringsort Code}
+\label{fig:checkedc-example-2}
+\end{figure}
+
+\begin{figure}[t]
+{\small
+  \begin{lstlisting}[xleftmargin=4 mm]
+//in unchecked region
+int f(char ** s, int (*cmp)(char *,char *),
+  int (*sort)(char **, int (*)(char *,char *),
+               int), int n) {
+...
+
+  int i = sort(s,cmp,n);
+  free(s[10]);
+...
+}
+
+int g(int (*cmp)(char *,char *)) {
+...
+  int real_addr = derandomize(cmp);
+...
+}
+
+int main(int n) {
+  nt_array_ptr<nt_array_ptr<char>> p : count(n)
+    = malloc(nt_array<nt_array_ptr<char>>, n);
+
+  nt_array_tptr<nt_array_tptr<char>>
+     tp : count(n) =
+       tmalloc(nt_array<nt_array_tptr<char>>, n);
+...
+  safe_memcpy(tp, p, n);
+
+  unchecked {
+    if (BAD) // a flag to call different funs.
+      //input checked pointers
+      f(p, compare_1, stringsort); 
+    else 
+      //input tainted pointers
+      f(tp, tainted_compare_1, tainted_stringsort);
+  }
+
+  if (!BAD) safe_memcpy(p, tp, n);
+  p[10] = "crash?";
+
+  unchecked {
+    if (BAD) g(compare_1)
+      else g(tainted_compare_1);
+  }
+
+  return 0;
+}
+  \end{lstlisting}
+}
+\caption{Tainted Pointer Usage in Calling Unchecked Fun}
+\label{fig:checkedc-example-3}
+\end{figure}
+
+\myparagraph{Maintaining Non-crashing}
+Previously, the main guarantee of \checkedc \cite{li22checkedc} was the blame theorem,
+i.e., if there is a crash, the source is from some unchecked regions.
+For example, at \Cref{fig:checkedc-example-3} line 31,
+we call a unchecked function \code{f} with a checked null-terminated array (NT-array) pointer argument.
+At line 8, depending on the NT-array size, \code{free(s[10])} might crash.
+Even if it does not crash, line 38 is doomed because of the \code{free} call.
+
+The philosophy is that unchecked regions are written in C and unstable,
+and programmers will eventually remove them by converting them to \checkedc code.
+The reality is that users might not want to convert everything.
+For some insignificant or handy history C code, such as library functions,
+they might want to keep it as long as no crashing.
+
+The sources of crashing in \checkedc are (1) unchecked regions crash themselves;
+  and (2) the misuse of checked pointers in unchecked regions.
+Enlightened by RLBox \cite{rlbox-paper},
+we sandbox the unchecked regions and then we utilize the \checkedc type system
+to disallow checked pointers to be used in an unchecked region. 
+To achieve the communication between checked and unchecked regions,
+we create tainted pointers that can be shared by different regions,
+whose data are stored in the sandboxed heap.
+Users are required to copy data to tainted pointers that are shared in unchecked regions.
+For example,  we copy the checked pointer data to the tainted pointer \code{tp}
+at \Cref{fig:checkedc-example-3} line 26,
+and input the tainted pointer to the unchecked function at line 34.
+At line 8, even if statement might crash, since tainted pointers are stored
+in the sandboxed heap, it can be recovered.
+At line 37, the use of a tainted pointer in a checked region
+requires a verification on it.
+This is handled by inserting additional checks
+and creating exception handling before the use by the \systemname compiler.
+Thus, the checked pointer \code{p} is safely used at line 38.
+
+In short, banning checked pointer uses in unchecked regions and providing tainted shared pointers
+are the solution for crashing. 
+In \systemname, we proved the non-crashing theorem based on the \systemname type system,
+.i.e, any program execution is stopped in a predictable manner
+where either the execution terminates or is stopped at a place where an exception handler exists.
+In fact, our work is \textit{the first formalism of RLBox},
+and formally shows the security guarantee that RLBox provides 
+is actually the non-crashing theorem.
+
+\myparagraph{Formalism Function Pointers}
+In C, manipulating function pointers is a major way of implementing high order functions.
+Previously, we assumed all function calls are through calling by name to a global map.
+In \systemname, we formalize function pointers and maintain the \systemname type soundness.
+To the best of our knowledge,
+this is the first work of formalizing C function pointers with security guarantee.
+
+\Cref{fig:checkedc-example-1} defines a string sorting algorithm
+depending on the input function pointer \code{cmp} defining the generic order for strings,
+and \code{compare_1} is an example \code{cmp} function
+that adds the ASCII numbers of characters in the two strings and compare the results.
+In addition, function pointers enable the callback mechanism,
+i.e., a server sends a function pointer to a client in an unchecked region,
+and allows the client to access some server resources by calling back the pointer.
+This is a common usage between a web-browser and untrusted third party libraries.
+The function call in \Cref{fig:checkedc-example-3} line 34 is one such usage.
+
+We also utilize \systemname subtyping relation to permit function pointer static auto-casting.
+Function pointer type information might contain array pointer bound information,
+for which it is inconvenient to coincide the defined types for a function implementation and the function pointer type. 
+For example, the \code{cmp} argument in \code{stringsort} (\Cref{fig:checkedc-example-1})
+has type \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>},
+meaning that the function takes two NT-array pointers with arbitrary size and outputs an integer.
+The function \code{compare_1}'s pointer has type \code{ptr<(int)(nt_array_ptr<char> : count (0), nt_array_ptr<char> : count (0) )>}
+To use \code{compare_1} in \code{stringsort}, the type is auto-cast to the \code{cmp}'s type.
+In general, if function pointer $x$ has type $* (tl \to t)$, and  $y$ has $* (tl' \to t')$,
+in order to use $x$ as $y$, $tl'$ should be a subtype of $tl$ and $t$ subtypes to $t'$.
+
+\myparagraph{Not Exposing Checked Pointer Addresses}
+The design for the non-crashing property bans the checked pointer uses in unchecked regions.
+Thus, there is no reason to permit checked pointer variable assignments in unchecked regions;
+especially, this might expose a checked pointer address to untrusted parties.
+For example, the call to function $g$ at \Cref{fig:checkedc-example-3} line 41 lives in an unchecked region,
+and $g$ might use some mechanism, such as derandomizing ASLR \cite{shacham-aslr},
+to achieve the checked pointer address.
+Thus, it enables a third party to access any checked heap and function data by simple pointer arithmetic.
+
+To prevent the checked pointer address leak,
+we prevent any unchecked regions acknowledge checked pointer variables.
+In addition, to facilitate checked function callbacks, 
+the \systemname compiler compiles every checked function
+with an additional tainted shell function.
+Users are required to serve unchecked regions with the tainted shell pointer instead of the original checked function pointer.
+For example, \code{tainted_compare_1} and \code{tainted_stringsort} in \Cref{fig:checkedc-example-2} 
+are the tainted shells of the checked functions \code{compare_1} and \code{stringsort}.
+In the tainted shells, the arguments are tainted versions of the corresponding arguments in the checked functions.
+Inside the shell body, create checked pointer copies of the tainted arguments, and call the checked functions.
+In addition, once a checked function returns,
+if the output is a checked pointer, we also its the data to a new tainted pointer and exit the shell.
+\Cref{fig:checkedc-example-3} line 42 is an example of serving the function call living in an unchecked region
+with a tainted shell pointer argument \code{tainted_compare_1}.
+Even if $g$ derndomizes its address (line 14), the shell address is in the sandbox and has no harm,
+and calling the shell never exposes any checked pointer information outside of the shell.
+
+Conceptually, the shell is run in a checked region.
+One can imagine that a tainted shell is a safe closure that contains a checked block.
+Once the closure is called, the system is turned to a checked region,
+so that the checked function call at \Cref{fig:checkedc-example-2} line 13 is safe, 
+even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.
+In the \systemname formalism, we formalize a \textit{checked} block on top of the existing unchecked regions,
+and the transition of a tainted shell call creates a checked block containing the shell body.
+We also make sure that no arguments in these tainted shell contains any checked pointers,
+as well as no output is of a checked type.
 

### 2022-07-11 sandp23/introduction.tex
update.

--- 2022-07-05 rlbox_paper/sections/introduction.tex
@@
 
-\input{figures/existingmem}
+%\input{figures/existingmem}
 
@@
   types} which are restricted by the compiler to spatially safe uses (temporal safety is underway~\cite{checkedc:temporal}).
-Such pointers have one of three possible types, \ptrT{|$T$|}, \arrT{|$T$|},
-or~\ntarrT{|$T$|} (\ptr, \arr, and \ntarr{} for short), representing a
+Such pointers have one of three possible types, \ptrT{T}, \arrT{T},
+or~\ntarrT{T} (\ptr, \arr, and \ntarr{} for short), representing a
 pointer to a single element, array of elements, or null-terminated
@@
 % highlightlines={1-3}, highlightcolor=red
+\ignore{
 \begin{listing}[t!]
@@
 \end{listing}
+}
 
-\input{figures/taintedsplit}
+%\input{figures/taintedsplit}
 

### 2022-07-15 sandp23/formal.tex
update.

@@
   \small \centering
-  $\begin{array}{l}
-\begin{array}{lll}
-\text{Function names:}~f&
+  \[ \hspace*{-1em}
+\begin{array}{l}
+\begin{array}{ll}
        \text{Variables:}~ x
@@
 
-\text{Mode:} & m & ::= & \cmode \mid \umode \\[0.5em]
+\text{Context Mode:} & m & ::= & \cmode \mid \umode \\[0.5em]
+
+\text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\[0.5em]
 
@@
 
-     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{m}
+     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{ \xi}
 \\[0.5em]
@@
 
-\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau}
+\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau} \mid \tfun{\overline{x}}{\overline{\tau}}{\tau}
 \\[0.5em]
@@
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \emalloc{\omega} \mid\elet{x}{e}{e} \\[0.2em]
-&&\mid&
-\ecast{\tau}{e} \mid \edyncast{\tau}{e}\mid \ecall{f}{\overline{e}} \mid \estrlen{x} \\[0.2em]
-&&\mid&
-\ebinop{e}{e} \mid\estar{e}\mid\eassign{e}{e}\mid\eunchecked{e}
+\evalue{n}{\tau} \mid x \mid \ecast{\tau}{e} \mid \edyncast{\tau}{e} \mid \ebinop{e}{e} \\[0.2em]
+&&\mid& \emalloc{\omega} \mid \estrlen{x} \mid\estar{e}\mid\eassign{e}{e}  \\[0.2em]
+&&\mid& \elet{x}{e}{e} \mid \eif{e}{e}{e}
 \\[0.2em]
-&&\mid&\eif{e}{e}{e}
+&&\mid&  \ecall{e}{\overline{e}} \mid
+\echecked{\overline{x}}{\tau}{e}
+\\[0.2em]
+&&\mid&\eunchecked{\overline{x}}{e}
+
 \end{array}
     \end{array}
-  $
+  \]
   \caption{\lang Syntax}
@@
 
-This section describes our formal model of \checkedc, called
+This section describes the formal model of \systemname, called
 \lang, making precise its syntax, semantics, and type system. It also
-develops \lang's meta-theory, including type soundness and the blame
+develops \lang's meta-theory, including type soundness, the non-exposure, and the non-crashing
 theorem.
@@
 $\omega$ classify multi-word values such as arrays, null-terminated
-arrays, and single-word-size values.
+arrays, functions, and single-word-size values.
 %
@@
 $\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide the $\cdot$ when writing the type). Here is the
+otherwise (we elide the $\cdot$ when writing the type).
+\systemname function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
+reflect its dependent function declrations,
+where $\overline{x}$ represents
+a list of \tint type variables in a dependent function header,
+which bind all bounds appearing in $\overline{\tau}$ and $\tau$.
+We have a well-formed requirement for a function type,
+such that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+Here is the
 corresponding \checkedc syntax for these types:
-\[
+\[\hspace*{-0.5em}
+\begin{array}{l}
 \begin{array}{rcl}
-$\code{array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\cmode}
+$\code{array_tptr<}$\tau$\code{> : count(}$n$\code{)}$
+&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
 \\[0.2em]
@@
 &\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
+\end{array}
+\\[0.2em]
+$\code{tptr<(int)(nt_array_tptr<}$\tau$\code{> : count(}$n$\code{),}$\\
+\qquad\qquad$\code{nt_array_tptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
+\\[0.2em]
+\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} * \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
 \end{array}
@@
 E &::=& \Box \mid \elet{x}{E}{e} \mid \ecall{f}{\overline{E}}\mid\ecast{\tau}{E} \mid \edyncast{\tau}{E} \\[0.2em]
-&&\mid \ret{x}{\evalue{n}{\tau}}{E}\mid\ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid\estar{E}\mid\eassign{E}{e} \\[0.2em]
+&& \mid \ebinop{\evalue{n}{\tau}}{E}\mid\estar{E}\mid\eassign{E}{e} \\[0.2em]
 &&\mid\eassign{\evalue{n}{\tau}}{E}\mid\eunchecked{E}\mid\eif{E}{e}{e}\\[0.2em]
@@
 
-    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}
+    \inferrule[S-RetCon]{\varphi(x)=\mu\\ (\varphi[x\mapsto \mu'],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\mu'}{e}) \longrightarrow (\varphi'[x\mapsto \mu],\heap',\ret{x}{\varphi'(x)}{e'})}
+
+    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
 
@@
       \inferrule[T-LetStr]
-                {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\\\
+                {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\\\
                  \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}];\Theta[x\mapsto \texttt{ge}\_0] \vdash_m e : \tau}
-                {\Gamma;\Theta \vdash_m \elet{x}{\estrlen{y}}{e} : \tau}
+                {\Gamma;\Theta \vdash_m \elet{x}{\estrlen{y}}{e} : \tau[b_h / x]}
 

### 2022-07-15 sandp23/formal.tex
update.

@@
   \small \centering
-  \[ \hspace*{-1em}
+  \[ \hspace*{-1.2em}
 \begin{array}{l}
@@
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ecast{\tau}{e} \mid \edyncast{\tau}{e} \mid \ebinop{e}{e} \\[0.2em]
-&&\mid& \emalloc{\omega} \mid \estrlen{x} \mid\estar{e}\mid\eassign{e}{e}  \\[0.2em]
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}  \\[0.2em]
+&&\mid& \estrlen{x} \mid\estar{e}\mid\eassign{e}{e}  \\[0.2em]
 &&\mid& \elet{x}{e}{e} \mid \eif{e}{e}{e}
 \\[0.2em]
-&&\mid&  \ecall{e}{\overline{e}} \mid
-\echecked{\overline{x}}{\tau}{e}
+&&\mid&  \emalloc{\xi}{\omega} \mid \ecall{e}{\overline{e}}
 \\[0.2em]
 &&\mid&\eunchecked{\overline{x}}{e}
+\mid \echecked{\overline{x}}{e}
 
@@
   \label{fig:checkc-syn}
+\end{figure}
+
+\begin{figure}[t]
+{\small
+  \begin{mathpar}
+
+  \inferrule[]
+  {}
+  {m \vdash \tint}
+
+  \inferrule[]
+  {\xi \wedge m\vdash \tau \\ \xi \le m}
+  {m \vdash \tptr{\tallarrayb{\bvar}{\tau}}{\xi}}
+
+  \inferrule[]
+  {\xi \wedge m \vdash \tau\\ \xi \le m}
+  {m \vdash \tptr{\tau}{\xi}}
+
+  \inferrule[]
+  {\xi \wedge m \vdash \tau\\ \xi \le m \\\\ \fv(\overline{\tau})\cup\fv(\tau)\subseteq \overline{x}}
+  {m \vdash \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi})}
+  \end{mathpar}
+}
+{\footnotesize
+\[
+\begin{array}{l} 
+\tmode \wedge \cmode = \umode \qquad \xi \wedge \umode = \umode
+\qquad \cmode \wedge m = m 
+\qquad  m_1 \wedge m_2 = m_2 \wedge m_1
+\\[0.2em]
+\xi \le \xi \qquad \tmode \le \xi
+\end{array}
+\\
+\]
+}
+ \caption{Well-formedness for Types}
+\label{fig:wftypes}
 \end{figure}
@@
 \\[0.2em]
-\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} * \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
+\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
 \end{array}
@@
 
-\lang expressions include literals ($n\!:\!\tau$), variables ($x$), memory
-allocation ($\emalloc{\omega}$), let binding ($\elet{x}{e_1}{e_2}$),
-static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$)
-(assumed at compile-time and verified at run-time, see Sec.~\ref{sec:example}),
-function calls ($\ecall{f}{\overline{e}}$), addition
-($\ebinop{e_1}{e_2}$), pointer dereference and assignment ($\estar{e}$)
-and ($\eassign{e_1}{e_2}$), resp.), unchecked blocks ($\eunchecked{e}$),
-the \texttt{strlen} operation ($\estrlen{x}$), and conditionals ($\eif{e}{e_1}{e_2}$).
+\lang expressions include literals ($n\!:\!\tau$), variables ($x$),
+ addition ($\ebinop{e_1}{e_2}$), static casts ($\ecast{\tau}{e}$), 
+dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}},
+the \texttt{strlen} operation ($\estrlen{x}$),
+pointer dereference and assignment ($\estar{e}$)
+and ($\eassign{e_1}{e_2}$), resp.),
+let binding ($\elet{x}{e_1}{e_2}$),
+conditionals ($\eif{e}{e_1}{e_2}$),
+memory allocation ($\emalloc{\xi}{\omega}$), 
+function calls ($\ecall{e}{\overline{e}}$),
+unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
 % \review{III.A.: the "dynamic cast" terminology may be briefly confused for C++'s
@@
 a heap address (this is useful for the semantics);
-$\evalue{0}{\tptr{\omega}{m}}$ (for any $m$ and $\omega$) represents
-the $\enull$ pointer, as usual. 
-The
-$\texttt{strlen}$ expression operates on variables $x$
+$\evalue{0}{\tptr{\omega}{m}}$ (for any $m$ and $\omega$) represents the $\enull$ pointer, as usual. 
+The $\texttt{strlen}$ expression operates on variables $x$
 rather than arbitrary expressions to simplify managing
@@
 becomes $\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$.
+
+Compared to the former \checkedc model \cite{li22checkedc},
+there are four differences.
+First, the \systemname type annotations have well-formed restrictions in \Cref{fig:wftypes}, 
+for maintaining non-exposure.
+Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi'} ...)}{\xi}$, if the outside mode $\xi$ is 
+$\tmode$ or $\umode$, the inside mode $\xi'$ cannot be $\cmode$.
+It is worth noting that pointer modes are a three point partial order ($\le$),
+where $\tmode$ is the infimum, and
+$\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
+such that a pointer mode $\tmode$ is projected as $\umode$.
+Second, $\emalloc{\xi}{\omega}$ includes a mode flag
+$\xi$ for allocating different pointers in different heap mode regions.
+For simplicity, we disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$),
+Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
+Fourth, we update $\eunchecked{\overline{x}}{e}$ 
+to specify $\overline{x}$ restricting all free variables appearing in $e$.
+Additionally, we create $\echeckedtext$ blocks indicating the context mode change from $\umode$ to $\cmode$.
+Both the free variables in the $\euncheckedtext$ and $\echeckedtext$ blocks
+and the return types cannot be $\cmode$ mode for maintaining non-exposure,
+i.e., no checked pointer can be communicated between $\euncheckedtext$ and $\echeckedtext$ blocks.
 
@@
 C-style \code{union}s have no safe typing
-in \checkedc, so we omit them. By default, functions are assumed to
+in \checkedc, so we omit them.
+Additional syntactic descrition is listed in the \checkedc formalism paper \cite{li22checkedc}.
+
+\ignore{
+ By default, functions are assumed to
 be within checked regions; placing the body in an \code{unchecked}
@@
   between the two models and the paper formalism.
-
+}
 % \mwh{NOTE: COULD WORK THE FOLLOWING INTO THE ABOVE DESCRIPTION OF
@@
 {\small
-$    \begin{array}{l}
+$\hspace*{-1.2em}
+    \begin{array}{l}
     \begin{array}{lll}
-\mu & ::= & \evalue{n}{\tau} \mid \bot\\
+\mu & ::= & \evalue{n}{\tau}\\
 e & ::= & \ldots \mid \ret{x}{\mu}{e}\\
 r & ::= & e \mid \enull \mid \ebounds\\
-E &::=& \Box \mid \elet{x}{E}{e} \mid \ecall{f}{\overline{E}}\mid\ecast{\tau}{E} \mid \edyncast{\tau}{E} \\[0.2em]
-&& \mid \ebinop{\evalue{n}{\tau}}{E}\mid\estar{E}\mid\eassign{E}{e} \\[0.2em]
-&&\mid\eassign{\evalue{n}{\tau}}{E}\mid\eunchecked{E}\mid\eif{E}{e}{e}\\[0.2em]
-\overline{E} &::=& E \mid  \evalue{n}{\tau},\overline{E} \mid \overline{E}, e
+E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
+&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid \eif{E}{e}{e}\\[0.2em]
+&&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 
+\eunchecked{\overline{x}}{E}
+\mid \echecked{\overline{x}}{E}
+
 \end{array}
@@
 
+    \inferrule{ \umode=\mode(E) \\
+      e=E[e'] \\
+      \tau=\type(e')}
+    {(\varphi,\heap,e)\longrightarrow_{\umode} (\varphi,\heap,E[\evalue{0}{\tau}])}
+
   \end{mathpar}
+}
+{\footnotesize
+\[
+\begin{array}{l} 
+\mode(E)=\mode'(E,\cmode)
+\\[0.2em]
+\mode'(\Box,m)=m
+\\[0.2em]
+\mode'(\eunchecked{\overline{x}}{E},m)=\mode'(E,\umode)
+\\[0.2em]
+\mode'(\echecked{\overline{x}}{E},m)=\mode'(E,\cmode)
+\\[0.2em]
+\mode'(\alpha({E}),m)=\mode'(E,m)\;\;{[\emph{owise}]}
+\end{array}
+\\
+\]
 }
@@
 \emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
-$\heap$ is a \emph{heap} mapping addresses (integer literals) to
-values $\evalue{n}{\tau}$; for both we ensure
-$\fv(\tau)=\emptyset$.
+$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ regions), each of which
+maps addresses (integer literals) to values $\evalue{n}{\tau}$.
+If a pointer has mode $\cmode$, it lives in the $\cmode$ region; otherwise, it lives in the $\umode$ region.
+We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ region,
+and $\heap(m)[n\mapsto \mu]$ to update location $n$ with the value $\mu$.
+It is worth noting that \systemname is not a fat-pointer system;
+thus, in every heap update, the value type annotation remains the same through program executions.
+Additionally, for both stack and heap, 
+we ensure $\fv(\tau)=\emptyset$ for all the value type annotations $\tau$.
+
 While heap bindings can change, stack bindings are immutable---once
@@
 %
-The mode $m$ indicates whether the
+The context mode $m$ indicates whether the
 stepped redex within $e$ was in a checked ($\cmode$) or
@@
 rule, which supports bounds widening; we discuss this rule
-shortly. When the second and first rules could both apply, we always
-prefer the second.\footnote{This approach is that of the PLT Redex model of \lang; the Coq
-development uses a slightly simpler syntax to achieve the same
-effect.}
+shortly. When the second and first rules could both apply, we always prefer the second.
+The third rule describes the exception handling 
+for possible crashing behaviors in unchecked region.
+A program in $\umode$ mode can non-deterministically crash
+and the \systemname sandbox mechanism recorvers
+the program to a safe point ($\evalue{0}{\tau}$)
+and continues with the existing program state.
+
+\begin{DIFnomarkup}
+ \begin{figure}[t]
+ {\small
+
+ \begin{mathpar}
+   \inferrule
+       {}
+       {m;\Theta;\heap;\sigma \vdash n : \tint}
+
+   \inferrule
+       {}
+       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\umode}}
+
+   \inferrule
+       {}
+       {m;\Theta;\heap;\sigma \vdash 0 : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
+       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
+            \\ m;\Theta;\heap;\sigma \vdash n : \tptr{\omega'}{\xi'}}
+       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       { \xi \le m 
+     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
+       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
+       {m;\Theta;\heap;\sigma \vdash n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
+
+   \inferrule
+       {\neg\funptr(\omega)\\ \xi \le m\\
+        \forall i \in [0,\size(\omega)) \;.\;
+            m;\Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash \heap(m,n+i)}
+       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\xi}}
+ \end{mathpar}
+ }
+{\footnotesize
+\[
+\begin{array}{l} 
+\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
+\qquad
+\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
+\end{array}
+\\
+\]
+}
+ \caption{Verification/Type Rules for Constants}
+ \label{fig:const-type}
+ \end{figure}
+\end{DIFnomarkup}
+
+%\footnote{This approach is that of the PLT Redex model of \lang; the Coq
+%development uses a slightly simpler syntax to achieve the same
+%effect.}
 % \review{the special case raises questions, e.g. why is this syntax-driven and
@@
 The $\mode$ function
-determines the mode when evaluating $e'$ based on the context $E$:
-if the $\Box$ occurs within $(\eunchecked{E'})$ inside $E$, then the mode is
-$\umode$; otherwise, it is $\cmode$. Evaluation contexts $E$ define a
-standard left-to-right evaluation order. (We explain the
+determines the context mode when evaluating $e'$ based on the context $E$.
+For any program execution, the top context mode always starts with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
+The result context mode is depeneded on where the $\Box$ locates.
+If it occurs within $(\eunchecked{\overline{x}}{E'})$ inside $E$ without
+a surrounding $\echeckedtext$ block, then the mode is
+$\umode$; if the exact opposite happens (within $(\echecked{\overline{x}}{E'})$ without surrounding $\euncheckedtext$ blocks),
+the mode is $\cmode$; the other cases ($\mode'(\alpha({E}),m)=\mode'(E,m)$) are recursively defined by the above base cases ($\alpha$ represents any syntactic categories in \Cref{fig:c-context} other than $\Box$, $\euncheckedtext$, and $\echeckedtext$).
+Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
 $\ret{x}{\mu}{e}$ syntax shortly.)
@@
   \begin{mathpar}
-        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-AssignArrC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\cmode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}
+
+        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
+         \\  \umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_a}{\tau} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-AssignArrT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h)
+               \\  \umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_1}{\tau} }
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\umode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}
 
@@
 
-        \inferrule[S-DefNTArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h]}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-AssignArr]{\heap(n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}
+     \inferrule[S-IfNTTC]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(\cmode,n)\neq 0}
+    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}
 
@@
 
-  \inferrule[S-DynCast]{
-\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{m})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{m} \\ n_l' \le n_l\ \\ n_h \le n_h'}
-    { (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{m}}}) \longrightarrow (\varphi',\heap',\evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau_b}{m}})}
-
- \inferrule[S-DynCastBound]{\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{\cmode})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{\cmode} \\ n_l' > n_l \vee n_h > n_h'}{ (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{\cmode}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{\cmode}}})
- \longrightarrow (\varphi',\heap',\ebounds)}
-
         \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
@@
 
-    \inferrule[S-Fun]{ \Xi(f) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
-        {(\varphi,\heap,\ecall{f}{{\evalue{\overline{n}}{\overline{\tau_a}}}}) \longrightarrow
+    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
+        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
    (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
 
-     % \inferrule[S-IfNTF]{\varphi(x)=\evalue{n}{\tau} \\ \heap(n)=0}
-     %      {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}
-      \inferrule[S-StrWiden]{
-         \varphi(x) = \evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{\cmode}}\\ 0 \in [n_l,n_h] \\ n_a > n_h
-\\ \heap(n+n_a) = 0 
-\\\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
-    {(\varphi,\heap,\estrlen{x}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tarrayptr{n_l}{n_a}{\tau}{\cmode}}],\heap,\evalue{n_a}{\tint})}
-
-     \inferrule[S-IfNTT]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(n)\neq 0}
-    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}
+    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
+                  \\ \umode;\emptyset;\heap ; \emptyset \vdash\evalue{n}{\tptr{\tau}{\tmode}}}
+        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
 
@@
 Fig.~\ref{fig:semantics} shows selected rules for the computation
-relation; we explain them with the help of the example in
+relation; we explain them with the help of the example in \Cref{fig:checkedc-example-1} and \Cref{fig:checkedc-example-2}.
+
+\ignore{
+ in
 Fig.~\ref{fig:strcat-ex},
@@
   $0 \leq $\code{x}$ \leq n$. It concatenates the \code{src} buffer to
-  the end of \code{dst} as long as \code{dst} has sufficient space.
+  the end of \code{dst} as long as \code{dst} has sufficient space.}
 
@@
 
-\myparagraph{Pointer accesses}
-%
-The rules for dereference and assignment operations---\textsc{S-Def},
-\textsc{S-DefNull}, \textsc{S-DefNTArray}, and
-\textsc{S-AssignArr}---illustrate how the semantics checks bounds.
+\myparagraph{Checked and Tainted Pointer Operations}
+%
+The rules for pointer related operations---\textsc{S-DefC},
+\textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
+\textsc{S-DefNull}, \textsc{S-IfNTTC} and \textsc{S-Cast}.
+The first five are deference and assignment operations---illustrate how the semantics checks bounds.
 Rule \textsc{S-DefNull} transitions attempted null-pointer
-dereferences to $\enull$, whereas \textsc{S-Def} dereferences a
+dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a \cmode-mode
 non-null (single) pointer.
@@
 evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
-does likewise when $\ebounds$ is returned (see below).
+does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
 
@@
 
-\textsc{S-AssignArr} assigns to an array as long as 0 (the point of
+\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
 dereference) is within the bounds designated by the pointer's annotation
@@
 
-\begin{DIFnomarkup}
- \begin{figure}[t]
- {\small
- \text{Type rules for constants and variables:}
- \begin{mathpar}
-   \inferrule[T-Var]
-       {x : \tau \in \Gamma}
-       {\Gamma;\Theta \vdash_m x : \tau}
-
-   \inferrule[T-Const]
-       {\fv(\tau) = \emptyset \\ \heap;\emptyset \vdash n : \tau}
-       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
- \end{mathpar}
-
- \text{Rules for checking constant pointers:}
- \begin{mathpar}
-   \inferrule
-       {}
-       {\heap;\sigma \vdash n : \tint}
-
-   \inferrule
-       {}
-       {\heap;\sigma \vdash n : \tptr{\omega}{\umode}}
-
-   \inferrule
-       {}
-       {\heap;\sigma \vdash 0 : \tptr{\omega}{\cmode}}
-
-   \inferrule
-       {(\evalue{n}{\tptr{\omega}{\cmode}})\in \sigma}
-       {\heap;\sigma \vdash n : \tptr{\omega}{\cmode}}
-
-   \inferrule
-       {\forall i \in [0,\size(\omega)) .
-            \heap;(\sigma \cup \{(n:\tptr{\omega}{\cmode})) \}\vdash \heap(n+i)}
-       {\heap;\sigma \vdash n : \tptr{\omega}{\cmode}}
- \end{mathpar}
- }
- \caption{Type Rules for Constants/Variables}
- \label{fig:const-type}
- \end{figure}
-\end{DIFnomarkup}
-
 \noindent

### 2022-07-15 sandp23/formal.tex
update.

@@
 
+\begin{figure}[t]
+{\small
+  \begin{lstlisting}[xleftmargin=4 mm]
+nt_array_ptr<const char>
+parse_utf16_hex(nt_array_ptr<const char> s,
+                  ptr<uint> result) {
+ int x1, x2, x3, x4;
+ if (s[0] != 0) { x1 = hex_char_to_int(s[0]);
+ if (s[1] != 0) { x2 = hex_char_to_int(s[1]);
+ if (s[2] != 0) { x3 = hex_char_to_int(s[2]);
+ if (s[3] != 0) { x4 = hex_char_to_int(s[3]);
+ if (x1 != -1 && x2 != -1 && x3 != -1 && x4 != -1){
+   *result = (uint)((x1<<12)|(x2<<8)|(x3<<4)|x4);
+   return s+4;
+  ...// several } braces
+ } 
+ return 0;
+}
+void parse(nt_array_ptr<const char> s,
+            array_ptr<uint> p : count(n), 
+            int n) {
+ array_ptr<uint> q : bounds(p,p+n) = p;
+ while (s && q < p+n) {
+   array_ptr<uint> r : count(1) =
+     dyn_bounds_cast<array_ptr<uint>>(q,count(1));
+   s = parse_utf16_hex(s,r);
+   q++;
+ }
+}
+  \end{lstlisting}
+
+% {\captionsetup[lstlisting]{margin = 8 mm}
+%   \begin{lstlisting}[xleftmargin=8 mm]
+% array_ptr<int> memcpy(int x,
+%      array_ptr<int> a : count(x),
+%      array_ptr<int> b : count(x)) : count(x) {
+%  for(int i = 0; i < x; ++i) {
+%     (*a = *b); a++; b++;
+%  }
+%  return a;
+}
+
+% static int idx = 0;
+% int buf (int * reg,
+%           nt_array_ptr<int> x,
+%           nt_array_ptr<int> y) {
+%   if (idx < n)
+%     unchecked { * x = &(reg + idx); idx++; }
+%   else bug();
+%   memcpy(n,x,y);
+%   return 0;
+% }
+% }
+\caption{Parsing a String of UTF16 Hex Characters in \checkedc}
+\label{fig:checkedc-example}
+\end{figure}
+
 % \review{Fig4: it was hard to tell which cases were stuck states, or could reduce
@@
 % LEO: This has an overfull line...?
-\begin{figure}[t]
-{\small
-{\captionsetup[lstlisting]{margin = 8 mm}
-  \begin{lstlisting}[xleftmargin=8 mm]
-nt_array_ptr<char> safe_strcat
-   (nt_array_ptr<char> dst : count(n),
-    nt_array_ptr<char> src : count(0), int n) {
-  int x = strlen(dst);
-  int y = strlen(src);
-  nt_array_ptr<char> c : count(n) =
-    dyn_bounds_cast
-           <nt_array_ptr<char>>(dst,count(n));
-    // sets c == dst with bound n (not x)
-  if (x+y < n) {
-    for (int i = 0; i < y; ++i)
-      *(c+x+i) = *(src+i);
-    *(c+x+y) = '\0';
-    return dst;
-  }
-  return null;
-}
-  \end{lstlisting}
-}
-}
-\caption{Implementation of safe \code{strcat}}
-\label{fig:strcat-ex}
-\end{figure}
 
@@
 does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
-
+\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
+dereference) is within the bounds designated by the pointer's annotation
+and strictly less than the upper bound.
+
+\textsc{S-DefC} and \textsc{S-AssignArrT} are the deference and assignment rules for tainted pointers.
+Any dynamic heap use of a tainted pointer requires to perform a verification.
+Performing such a verification equates to performing a type check for a pointer constant in \Cref{fig:const-type}.
+We explain this shortly in \Cref{sec:type-system}.
+For now, the verification step, e.g. $\umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_a}{\tau}$ in \textsc{S-DefC},
+means we verify that the value $n_a$ has type $\tau$ in the $\umode$ region heap.
+
+Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
+by \textsc{S-Cast}. In a type-correct program, such casts are
+confirmed safe by the type system no matter
+if the target is a $\tmode$ or $\cmode$ pointer. To evaluate a cast, the rule
+updates the type annotation on $n$. Before doing so, it must
+``evaluate'' any variables that occur in $\tau$ according to their
+bindings in $\varphi$. For example, if $\tau$ was
+$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
+produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
+
+Bounds widening occurs when branching on a dereference of a NT array
+pointer, or when performing $\estrlentext$, which has been explained in \cite{li22checkedc}. 
+Here, we explain Rule \textsc{S-IfNTTC} that 
+performs bounds widening on a $\cmode$-mode pointer $x$ when the
+dereference $\estar{x}$ is not at the null terminator, but the
+pointer's upper bound is 0 (i.e., it's at the end of its known
+range). $x$'s upper bound is incremented to 1, and this count persists
+as long as $x$ is live.  For example, \code{s}'s increment (lines
+5--8) is live until the return of the function in
+Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
+\code{s}'s upper bound is properly extended.
+
+\ignore{
+The latter is most useful
+when assigned to a local variable so that subsequent code can use the
+result, e.g., $e$ in $\elet{x}{\estrlen{y}}{e}$.  Lines 4 and 5 in
+Fig.~\ref{fig:strcat-ex} are examples. 
+The widened upper bound
+precipitated by $\estrlen{y}$ is extended beyond the lifetime of $x$, as long as $y$ is live.
+% Even if $x$ returns, if $y$'s upper bound keeps the widened bound that is caused by $\estrlen{y}$.
+For example, \code{x}'s scope in line 4 at runtime is
+the whole function body in \code{safe_strcat} because the lifetime of the pointer \code{dst} is in the function body.
+% Allowing widening to persist
+% outside the scope is sound, % but would precipitate the need for
+% % fat pointers during compilation, which we want to avoid.
+% and we take this approach to give more precise run-time behavior. We
+% show in Sec.~\ref{sec:compilation}
+This is different from the \checkedc specification, 
+which only allows bound widening to happen within the scope of \code{x}, and restoring old bound values once \code{x} dies.
+We allow widening to persist outside the scope at run-time as long as
+we are within the stack frame, and we show
+this does not necessarily require the use of fat pointers in
+Sec.~\ref{sec:compilation}.
+
+Rule \textsc{S-StrWiden} implements $\estrlentext$ widening. The
+predicate
+$\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i)
+= \evalue{n_i}{\tau_i} \wedge n_i \neq 0))$ aims to find a position
+$n+n_a$ in the NT array that stores a null character, where no
+character as indexes between $n$ and $n+n_a$ contains one. (This rule
+handles the case when $n_a > n_h$, the $n_a \le n_h$ case is handled
+by a normal $\estrlentext$ rule; see 
+\iftr
+Appendix~\ref{sec:rem-semantics}.)
+\else
+the supplemental report~\cite{checkedc-tech-report}.)
+\fi
+
+Rule \textsc{S-IfNTT} performs bounds widening on $x$ when the
+dereference $\estar{x}$ is not at the null terminator, but the
+pointer's upper bound is 0 (i.e., it's at the end of its known
+range). $x$'s upper bound is incremented to 1, and this count persists
+as long as $x$ is live.  For example, \code{s}'s increment (lines
+5--8) is live until the return of the function in
+Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
+\code{s}'s upper bound is properly extended.
+}
 % \review{- on page 6, section IIIC, paragraph "Pointer Access" mentions that checked pointers cannot be dereferenced in unchecked blocks - this looks funny, shouldn't it be the other way around?
@@
 
+\ignore{
 \textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
@@
 not write, the null terminator. A separate rule (not shown) handles
-normal arrays.
-
+normal arrays.}
+
+\ignore{
 \myparagraph{Casts}
 %
-Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
-by \textsc{S-Cast}. In a type-correct program, such casts are
-confirmed safe by the type system. To evaluate a cast, the rule
-updates the type annotation on $n$. Before doing so, it must
-``evaluate'' any variables that occur in $\tau$ according to their
-bindings in $\varphi$. For example, if $\tau$ was
-$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
-produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
 
@@
 holds, but then checks it at run-time.
+}
 
@@
 
+\ignore{
 \myparagraph{Bounds Widening}
@@
 Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
-\code{s}'s upper bound is properly extended. %  For this rule and
+\code{s}'s upper bound is properly extended.
+} %  For this rule and
 % \textsc{S-StrWiden}, this widening persists in the current stack

### 2022-07-15 sandp23/formal.tex
update.

@@
 
-        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e})}
 
@@
 is $\varphi$ extended with a binding for $x$, and whose expression is
-$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
-$\varphi(x)$; if it had no previous binding, $\varphi(x) =
-\bot$. Evaluation proceeds on $e$ until it becomes a literal
-$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
-binding (or $\bot$) in the new stack, and evaluates to
+$\ret{x}{\evalue{n}{\tau}}{e})$ which keeps $\varphi$ unchanged, but
+remember $x$ with the new value $\evalue{n}{\tau}$ in the scope $e$.
+Every time when Evaluation proceeds on $e$ (Rule \textsc{S-RetCon}),
+we install the current stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$,
+after one-step evaluation is completed, we restore the result $\varphi'(x)$ 
+in the result $\ret{x}{\varphi'(x)}{e'})$.
+This procedure continues until it becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-RetEnd} remove the $\kw{ret}$ frame and evaluates to
 $n\!:\!\tau$.

### 2022-07-16 sandp23/formal.tex
update.

--- 2022-07-15 sandp23/formal.tex
@@
 Here is the
-corresponding \checkedc syntax for these types:
+corresponding \systemname syntax for these types:
 \[\hspace*{-0.5em}
@@
 \lang aims to be simple enough to work with, but powerful enough to
-encode realistic \checkedc idioms. For example, mutable local
+encode realistic \systemname idioms. For example, mutable local
 variables can be encoded as immutable locals that point to the heap;
@@
 
-\ignore{
- By default, functions are assumed to
-be within checked regions; placing the body in an \code{unchecked}
-expression relaxes this, and within that, checked regions can be
-nested via function calls. Bounds are restricted slightly:
-rather than allowing arbitrary sub-expressions, bounds must be either
-integer literals or variables plus an integer offset, which accounts
-for most uses of \code{bounds} in \checkedc programs. \lang bounds are
-defined as relative offsets, not absolute ones, as in the second part
-of Fig.~\ref{fig:checkedc-example}. We see no technical problem to
-modeling absolute bounds, but it would be a pervasive change so we
-have not done so.
-
-  We have mechanized two models of \lang, one in Coq
-  and one in PLT Redex~\cite{pltredex}, which is a semantic
-  engineering framework implemented in Racket. Redex provides direct
-  support for specifying the operational semantics and typing with logical
-  rules, but then automatically makes them executable and subject to
-  randomized testing, which is very useful during development.
-  The model we present in the paper faithfully
-  represents both mechanizations, but there are some differences for
-  presentation purposes. For example, the paper and the Coq model use an explicit
-  stack, whereas the Redex model uses \code{let} bindings to simulate
-  one (simplifying term generation for randomized
-  testing).
-\iftr
-  Appendix~\ref{app:model-diffs} 
-\else
-  The supplemental report~\cite{checkedc-tech-report}
-\fi
-  outlines the differences
-  between the two models and the paper formalism.
-}
 % \mwh{NOTE: COULD WORK THE FOLLOWING INTO THE ABOVE DESCRIPTION OF
@@
 
-    \inferrule{ m=\mode(E) \\
-      e=E[\eif{\estar{x}}{e_1}{e_2}] \\
-      (\varphi,\heap,\eif{\estar{x}}{e_1}{e_2})
-              \longrightarrow (\varphi',\heap',e')}
-    {(\varphi,\heap,e)\longrightarrow_{m} (\varphi',\heap',E[e'])}
-    \;\;\;{[\emph{prefer}]}
-
     \inferrule{ \umode=\mode(E) \\
@@
 represents a $\enull$ pointer when $\tau$ is a pointer type;
-correspondingly, $\heap(0)$ should always be undefined.
+correspondingly, $\heap(m,0)$ should always be undefined.
 The relation steps to a \emph{result} $r$,
@@
 shortly. 
-The second rule handles conditionals $\eif{\estar{x}}{e_2}{e_3}$ in
-redex position specially, delegating directly to the \textsc{S-IfNTT} computation
-rule, which supports bounds widening; we discuss this rule
-shortly. When the second and first rules could both apply, we always prefer the second.
-The third rule describes the exception handling 
+The second rule describes the exception handling 
 for possible crashing behaviors in unchecked region.
@@
 and continues with the existing program state.
-
-\begin{DIFnomarkup}
- \begin{figure}[t]
- {\small
-
- \begin{mathpar}
-   \inferrule
-       {}
-       {m;\Theta;\heap;\sigma \vdash n : \tint}
-
-   \inferrule
-       {}
-       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\umode}}
-
-   \inferrule
-       {}
-       {m;\Theta;\heap;\sigma \vdash 0 : \tptr{\omega}{\xi}}
-
-   \inferrule
-       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
-       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\xi}}
-
-   \inferrule
-       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
-            \\ m;\Theta;\heap;\sigma \vdash n : \tptr{\omega'}{\xi'}}
-       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\xi}}
-
-   \inferrule
-       { \xi \le m 
-     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
-       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
-       {m;\Theta;\heap;\sigma \vdash n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
-
-   \inferrule
-       {\neg\funptr(\omega)\\ \xi \le m\\
-        \forall i \in [0,\size(\omega)) \;.\;
-            m;\Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash \heap(m,n+i)}
-       {m;\Theta;\heap;\sigma \vdash n : \tptr{\omega}{\xi}}
- \end{mathpar}
- }
-{\footnotesize
-\[
-\begin{array}{l} 
-\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
-\qquad
-\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
-\end{array}
-\\
-\]
-}
- \caption{Verification/Type Rules for Constants}
- \label{fig:const-type}
- \end{figure}
-\end{DIFnomarkup}
+There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
 
@@
 For any program execution, the top context mode always starts with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
-The result context mode is depeneded on where the $\Box$ locates.
+The result context mode is depeneded on where $\Box$ locates.
 If it occurs within $(\eunchecked{\overline{x}}{E'})$ inside $E$ without
@@
         \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
-         \\  \umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_a}{\tau} }
+         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
     {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
@@
     \inferrule[S-AssignArrT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h)
-               \\  \umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_1}{\tau} }
+               \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_1}{\tau} }
       {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\umode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}
@@
     \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-
-     \inferrule[S-IfNTTC]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(\cmode,n)\neq 0}
-    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}
 
@@
 
-        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
+    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\mu}{\mu'}) \longrightarrow (\varphi,\heap,\mu')}
+
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e})}
 
@@
 
-    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\mu}) \longrightarrow (\varphi,\heap,\mu)}
 
@@
 
-    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\mu}) \longrightarrow (\varphi,\heap,\mu)}
 
     \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
-                  \\ \umode;\emptyset;\heap ; \emptyset \vdash\evalue{n}{\tptr{\tau}{\tmode}}}
+                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
         {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
@@
 
-% \review{Fig4: it was hard to tell which cases were stuck states, or could reduce
-%   owing to a rule that was not shown}
-% \mwh{Stuck states are those where the expression is a non-value and
-%   not $\enull$ or $\ebounds$. Updated III-B and III-D.}
-% \review{Fig4: can the rules be presented in the same order they were introduced in
-%   the paper?}
-% \liyi{ Reordered }
-
-% \review{Fig4: S-FUN: $\vec\tau_a$ seems unused; why?}
-% \liyi{the list of $\tau_a$ is a list of input argument types. These
-%   are used during type checking, but not during evaluation (as is
-%   typical).}
-
-% LEO: This has an overfull line...?
 \begin{figure}[t]
@@
 
+% \review{Fig4: it was hard to tell which cases were stuck states, or could reduce
+%   owing to a rule that was not shown}
+% \mwh{Stuck states are those where the expression is a non-value and
+%   not $\enull$ or $\ebounds$. Updated III-B and III-D.}
+% \review{Fig4: can the rules be presented in the same order they were introduced in
+%   the paper?}
+% \liyi{ Reordered }
+
+% \review{Fig4: S-FUN: $\vec\tau_a$ seems unused; why?}
+% \liyi{the list of $\tau_a$ is a list of input argument types. These
+%   are used during type checking, but not during evaluation (as is
+%   typical).}
+
+% LEO: This has an overfull line...?
+
 Fig.~\ref{fig:semantics} shows selected rules for the computation
@@
 \textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
-\textsc{S-DefNull}, \textsc{S-IfNTTC} and \textsc{S-Cast}.
+\textsc{S-DefNull}, and \textsc{S-Cast}.
 The first five are deference and assignment operations---illustrate how the semantics checks bounds.
@@
 does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
-
-% \review{- on page 6, section IIIC, paragraph "Pointer Access" mentions that checked pointers cannot be dereferenced in unchecked blocks - this looks funny, shouldn't it be the other way around?
-The Coq code contains the hypothesis m'=Unchecked -> m Unchecked in various rules of definition well-typed (BoundCheckedC, line 669; rule TyDeref in the code seems closest to the figure's T-DefArr and T-Def in the appendix, although it's a bit concerning that there's no 1-to-1 correspondence of the rules in the code and the paper).
-% }
-% \liyi{Yeah. It is another typo. It should be the other way around.  }
-
 \textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
 dereference) is within the bounds designated by the pointer's annotation
-and strictly less than the upper bound. For the assignment rule,
-arrays are treated uniformly whether they are null-terminated or not
-($\kappa$ can be $\cdot$ or $nt$)---the semantics does not search past
-the current position for a null terminator. The program
-can widen the bounds as needed, if they currently precede the null
-terminator: \textsc{S-DefNTArray}, which dereferences an NT array
-pointer, allows an upper bound of $0$, since the program may read, but
-not write, the null terminator. A separate rule (not shown) handles
-normal arrays.
-
-\myparagraph{Casts}
-%
+and strictly less than the upper bound.
+
+\textsc{S-DefT} and \textsc{S-AssignArrT} are similar rules to \textsc{S-DefC} and \textsc{S-AssignArrC} for tainted pointers.
+Any dynamic heap use of a tainted pointer requires a verification.
+Performing such a verification equates to performing a type check for a pointer constant in \Cref{fig:const-type}.
+We explain this shortly in \Cref{sec:type-system}.
+For now, the verification step, e.g. $\umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_a}{\tau}$ in \textsc{S-DefC},
+means we verify that the value $n_a$ has type $\tau$ in the $\umode$ region heap.
+
 Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
 by \textsc{S-Cast}. In a type-correct program, such casts are
-confirmed safe by the type system. To evaluate a cast, the rule
+confirmed safe by the type system no matter
+if the target is a $\tmode$ or $\cmode$ pointer. To evaluate a cast, the rule
 updates the type annotation on $n$. Before doing so, it must
@@
 produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
-
-Dynamic casts are accounted for by \textsc{S-DynCast} and
-\textsc{S-DynCastBound}. In a type-correct program, such casts are
-assumed correct by the type system, and later confirmed by the
-semantics. As such, a dynamic cast will cause a $\ebounds$ failure if
-the cast-to type is incompatible with the type of the target pointer,
-as per the $n_l' > n_l \vee n_h > n_h'$ condition in
-\textsc{S-DynCastBound}. An example use of dynamic casts is given on
-line~7 in Fig.~\ref{fig:strcat-ex}. The values of \code{x} and
-\code{n} might not be known statically, so the type system cannot
-confirm that \code{x <= n}; the dynamic cast assumes this inequality
-holds, but then checks it at run-time.
+The full formalism, including \kw{struct}
+and null-terminated bound widening pointer operations, is given in \Cref{app:main}.
+
+% \review{- on page 6, section IIIC, paragraph "Pointer Access" mentions that checked pointers cannot be dereferenced in unchecked blocks - this looks funny, shouldn't it be the other way around?
+The Coq code contains the hypothesis m'=Unchecked -> m Unchecked in various rules of definition well-typed (BoundCheckedC, line 669; rule TyDeref in the code seems closest to the figure's T-DefArr and T-Def in the appendix, although it's a bit concerning that there's no 1-to-1 correspondence of the rules in the code and the paper).
+% }
+% \liyi{Yeah. It is another typo. It should be the other way around.  }
+
+\myparagraph{Unchecked and Checked Blocks}
+%
+Semantically, $\euncheckedtext$ and $\echeckedtext$ blocks 
+(\textsc{S-Unchecked} and \textsc{S-Checked}) act as a classical C block structures.
+The interesting part is their type rules in \Cref{sec:type-system}.
 
@@
 is $\varphi$ extended with a binding for $x$, and whose expression is
-$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
-$\varphi(x)$; if it had no previous binding, $\varphi(x) =
-\bot$. Evaluation proceeds on $e$ until it becomes a literal
-$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
-binding (or $\bot$) in the new stack, and evaluates to
+$\ret{x}{\evalue{n}{\tau}}{e})$ which keeps $\varphi$ unchanged, but
+remember $x$ with the new value $\evalue{n}{\tau}$ in the scope $e$.
+Every time when Evaluation proceeds on $e$ (Rule \textsc{S-RetCon}),
+we install the current stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$,
+after one-step evaluation is completed, we restore the result $\varphi'(x)$ 
+in the result $\ret{x}{\varphi'(x)}{e'})$.
+This procedure continues until it becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-RetEnd} remove the $\kw{ret}$ frame and evaluates to
 $n\!:\!\tau$.
 
-Function calls are handled by \textsc{S-Fun}. Recall that array
-bounds in types may refer to in-scope variables; e.g., parameter
-\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
-2-3 in Fig.~\ref{fig:strcat-ex}. A call to function $f$ causes $f$'s
+Function calls are handled by \textsc{S-FunC} and \textsc{S-FunT},
+for \cmode and \tmode mode function pointers, respectively. 
+A call to function $f$ causes $f$'s
 definition to be retrieved from $\Xi$,
@@
 which maps function names to
-forms $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e$, where
+function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
 $\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
-is the parameter list of variables and their types, and $e$ is the
-function body. The call is expanded into a \texttt{let} which binds
+is the parameter list of variables and their types, 
+$\xi$ determines the mode of the function, and $e$ is the
+function body. 
+Similar to \heap, the global function store $\Xi$ is also partitioned into
+two parts ($\cmode$ and $\umode$ regions), each of which
+maps addresses (integer literals) to the function data described above.
+
+The \systemname functions are dependent functions.
+Recall that array
+bounds in types may refer to in-scope variables; e.g., parameter
+\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
+2-3 in Fig.~\ref{fig:strcat-ex}. 
+%All dependent bound variables appearing in function headers 
+%are bounded by a variable list in the function pointer type, which is checked in the type rule.
+Semantically,
+the call is expanded into a \texttt{let} which binds
 parameter variables $\overline{x}$ to the actual arguments
@@
 argument list.
-Note that the \textsc{S-Fun} rule replaces the
+Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
   annotations $\overline{\tau_a}$ with
@@
 
-\myparagraph{Bounds Widening}
-%
-Bounds widening occurs when branching on a dereference of a NT array
-pointer, or when performing $\estrlentext$. The latter is most useful
-when assigned to a local variable so that subsequent code can use the
-result, e.g., $e$ in $\elet{x}{\estrlen{y}}{e}$.  Lines 4 and 5 in
-Fig.~\ref{fig:strcat-ex} are examples. 
-The widened upper bound
-precipitated by $\estrlen{y}$ is extended beyond the lifetime of $x$, as long as $y$ is live.
-% Even if $x$ returns, if $y$'s upper bound keeps the widened bound that is caused by $\estrlen{y}$.
-For example, \code{x}'s scope in line 4 at runtime is
-the whole function body in \code{safe_strcat} because the lifetime of the pointer \code{dst} is in the function body.
-% Allowing widening to persist
-% outside the scope is sound, % but would precipitate the need for
-% % fat pointers during compilation, which we want to avoid.
-% and we take this approach to give more precise run-time behavior. We
-% show in Sec.~\ref{sec:compilation}
-This is different from the \checkedc specification, 
-which only allows bound widening to happen within the scope of \code{x}, and restoring old bound values once \code{x} dies.
-We allow widening to persist outside the scope at run-time as long as
-we are within the stack frame, and we show
-this does not necessarily require the use of fat pointers in
-Sec.~\ref{sec:compilation}.
-
-Rule \textsc{S-StrWiden} implements $\estrlentext$ widening. The
-predicate
-$\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i)
-= \evalue{n_i}{\tau_i} \wedge n_i \neq 0))$ aims to find a position
-$n+n_a$ in the NT array that stores a null character, where no
-character as indexes between $n$ and $n+n_a$ contains one. (This rule
-handles the case when $n_a > n_h$, the $n_a \le n_h$ case is handled
-by a normal $\estrlentext$ rule; see 
-\iftr
-Appendix~\ref{sec:rem-semantics}.)
-\else
-the supplemental report~\cite{checkedc-tech-report}.)
-\fi
-
-Rule \textsc{S-IfNTT} performs bounds widening on $x$ when the
-dereference $\estar{x}$ is not at the null terminator, but the
-pointer's upper bound is 0 (i.e., it's at the end of its known
-range). $x$'s upper bound is incremented to 1, and this count persists
-as long as $x$ is live.  For example, \code{s}'s increment (lines
-5--8) is live until the return of the function in
-Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
-\code{s}'s upper bound is properly extended. %  For this rule and
+ %  For this rule and
 % \textsc{S-StrWiden}, this widening persists in the current stack
@@
 in mode $m$. Key rules for this judgment are given in
-Fig.~\ref{fig:type-system-1}. In the rules, $m \le m'$ uses
-the two-point lattice with $\umode < \cmode$.
+Fig.~\ref{fig:type-system-1}.
 % \review{text says $u < c$ but this seems to contradict the paragraph right
@@
   \begin{mathpar}
-    \inferrule[T-DefArr]
-              {m \leq m' \\\\ \Gamma;\Theta \vdash_{m} e : \tptr{\tallarrayb{\bvar}{\tau}}{m'}}
+   \inferrule[T-ConstU]
+       { }
+       {\Gamma;\Theta\vdash_c \evalue{u}{\tau} : \tau}
+
+   \inferrule[T-ConstC]
+       {\Theta;\heap;\emptyset \vdash_c n : \tau}
+       {\Gamma;\Theta\vdash_c \evalue{n}{\tau} : \tau}
+
+    \inferrule[T-Def]
+              {m' \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{m'}}
               {\Gamma;\Theta \vdash_m \estar{e} : \tau}
@@
               {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\\\
-                \Gamma; \Theta \vdash_m e_2 : \tau' \\
-                \tau'\sqsubseteq \tau\\
-                m \leq m'}
+                \Gamma; \Theta \vdash_m e_2 : \tau' \\\\
+                \tau'\sqsubseteq_{\Theta} \tau\\
+                m' \leq m}
               {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
 
+     \inferrule[T-CastPtr]
+               {\Gamma;\Theta \vdash_m e : \tau' \\
+                 \tau' \sqsubseteq \tptr{\tau}{\xi}}
+               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
+
+   \inferrule[T-Let]
+    { x\not\in \fv(\tau') \\
+        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
+          \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'}
+
+   \inferrule[T-LetInt]
+    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
+        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
+           \Gamma[x\mapsto \tau];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
+
+   \inferrule[T-Ret]
+    {m;\emptyset;\Gamma;\emptyset\vdash \mu : \tau' \\\\
+          \Gamma[x\mapsto \tau'];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
+    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tau'}}{e} : \tau}
+
+    \inferrule[T-Checked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
+              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
+
     \inferrule[T-Unchecked]
-              {\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{e} : \tau}
-
-    \inferrule[T-Cast]
-              {m = \cmode \Rightarrow \tau \neq\tptr{\tau''}{\cmode} \text{ for any $\tau''$} \\\\
-                \Gamma;\Theta \vdash_m e : \tau'}
-              {\Gamma;\Theta \vdash_m \ecast{\tau}{e} : \tau}
-
-     \inferrule[T-CastCheckedPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq \tptr{\tau}{\cmode}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\cmode}}{e} : \tptr{\tau}{\cmode}}
-
-      \inferrule[T-DynCast]
-                {\Gamma;\Theta \vdash_m e : \tptr{\tallarrayb{\bvar'}{\tau}}{m}}
-                {\Gamma;\Theta \vdash_m \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e} : \tptr{\tallarrayb{\bvar}{\tau}}{m}}
-
-      \inferrule[T-If ]
-        {\Gamma; \Theta \vdash_m e : \tau \\\\ \Gamma; \Theta \vdash_m e_1 : \tau_1 \\\\ \Gamma; \Theta \vdash_m e_2 : \tau_2}
-        {\Gamma; \Theta \vdash_m \eif{e}{e_1}{e_2} : \tau_1 \sqcup \tau_2}
-
-      \inferrule [T-IfNT]
-                { \Gamma; \Theta \vdash_m x : \tptr{\tntarrayb{(b_l,0)}{\tau}}{c} \\\\
-                   \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}]; \Theta \vdash_m e_1 : \tau_1\\\\
-                  \Gamma;\Theta \vdash_m e_2 : \tau_2 }
-                {\Gamma;\Theta \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_1 \sqcup \tau_2}
-
-      \inferrule[T-Str]
-                {\Gamma; \Theta \vdash_m e :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} 
-                }
-                {\Gamma;\Theta \vdash_m \estrlen{e} : \tint}
-
-      \inferrule[T-LetStr]
-                {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\\\
-                 \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}];\Theta[x\mapsto \texttt{ge}\_0] \vdash_m e : \tau}
-                {\Gamma;\Theta \vdash_m \elet{x}{\estrlen{y}}{e} : \tau[b_h / x]}
-
-   \inferrule[T-Let]
-    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
-        \Gamma;\Theta \vdash_m e_1 : \tau \\
-           \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
-             }
-    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\\
+                  \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
+              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
 
 \inferrule[T-Fun]
-    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
-             \overline{\tau'} \sqsubseteq 
-               \overline{\tau}[\overline{e} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m f(\overline{e}) : \tau[\overline{e} / \overline{x}]}
-
-                \inferrule[T-Ret]
-    {\Gamma(x)\neq \bot \\
-          \Gamma;\Theta \vdash_m e : \tau}
-    {\Gamma;\Theta \vdash_m \eret{x}{\mu}{e} : \tau}
+    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
+        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\\\
+         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
+         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
+             \overline{\tau'} \sqsubseteq_{\Theta}
+               \overline{\tau}[\overline{e'} / \overline{x}]}
+    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
 
@@
 % }
+{\footnotesize
+\[
+\begin{array}{l} 
+\cmode(\tint)=\texttt{false}
+\qquad
+\cmode(\tptr{\omega}{\cmode})=\texttt{true}
+\qquad
+\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
+\end{array}
+\\
+\]
+}
 \caption{Selected type rules}
@@
 similar.
-The condition $m\le m'$ ensures that unchecked pointers can only
-be dereferenced in unchecked blocks;
+The condition $m\le m'$ ensures that checked pointers and unchecked pointers 
+can only be deferenced in checked and unchecked regions, respectively;
 the type rule for
-$\eunchecked{e}$ sets $m = \umode$ when checking $e$.
 The rules do not attempt to reason whether the access is in bounds;
@@
 
-\myparagraph{Casting and Subtyping}
-%
-Rule \textsc{T-Cast} rule forbids casting to checked pointers when in checked
-regions (when $m = \cmode$), but $\tau$ is unrestricted when
-$m = \umode$. The \textsc{T-CastCheckedPtr} rule
+\myparagraph{Type Equality and Subtyping and Casting}
+%
+In \systemname, we define the type equality $\tau=_{\Theta}\tau'$
+is an equivalent relation over the bound equality ($=_{\Theta}$) of the bounds in (NT-)array types
+and alpha equivalence of two function types in \Cref{fig:checkc-subtype}.
+Two (NT-)array pointers $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$; two function types 
+$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
+are equivalent, if we can find a same length variable set $\overline{z}$ that is substituted for  
+$\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, respectively,
+and the substitution results are equal.
+
+ The \textsc{T-CastPtr} rule
 permits casting from an expression of type $\tau'$ to a checked pointer when
 $\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype}; the many
+$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} built on the type equality
+($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$); the many
 rules ensure the relation is transitive. Most of the rules handle
@@
 pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
+Two function pointers are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are oppositely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
 
@@
 requires reasoning about those variables' possible values. The type
-system uses $\Theta$ to make such reasoning more precise.\footnote{Technically, the subtyping relation $\sqsubseteq$ and the bounds
-  ordering relation $\leq$ are parameterized by $\Theta$; this fact is
-  implicit to avoid clutter.} $\Theta$ is a map from variables $x$ to
-predicates $P$, which have the form $P ::= \top \;|\; \texttt{ge}\_0$.
-If $\Theta$ maps $x$ to $\top$, that means that the variable can
-possibly be any value; $\texttt{ge}\_0$ means that $x \ge 0$. We will
-see how $\Theta$ gets populated and give a detailed example of
-subtyping below.\footnote{As it turns out, the subtyping relation is
-  also parameterized by $\varphi$, which is needed when type checking
-  intermediate results to prove type preservation; source programs
-  would always have $\varphi = \emptyset$. Details are in
+system uses $\Theta$ to make such reasoning more precise.
+$\Theta$ is a map from variables $x$ to
+equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
+It maps variables to equations that are recorded along the type checking procedure.
+If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
+$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
+such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-Ret}.
 \iftr
@@
   the supplemental report~\cite{checkedc-tech-report}.
-\fi
-}
-
-Rule \textsc{T-DynCast} typechecks dynamic casting operations, which
-apply to array pointer types only. The cast is accepted by the type
-system, as its legality will be checked by the semantics.
+\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
 
@@
 {\small
-\begin{center}
-  $\begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-    &&\multicolumn{1}{r}{\tau} &\sqsubseteq& \tau\\[0.6em]
-    0\le b_l \wedge b_h \le 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.6em]
-    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
-    b_l \le 0 \wedge 1 \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tptr{\tau}{m}\\[0.6em]
+\[\hspace*{-1.2em}
+\begin{array}{l}
+\text{Bound Inequality and Equality:}\\
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
+     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
+     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
+     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
+     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
+     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
+    \end{array}
+  \\[0.5em]
+\text{Type Equility:}\\
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     && \tint & =_{\Theta} & \tint\\
+     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
+     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
+             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
+
+    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
+
+ &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
+                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
+    \end{array}
+  \\[0.5em]
+\text{Subtype:}\\
+
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
+    &&\tptr{\omega}{\tmode} &\sqsubseteq_{\Theta}& \tptr{\omega'}{\umode}\\[0.2em]
+
+    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
     %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
+\\[0.2em]
+\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
+
     \end{array}
-  $
-\end{center}
-}
-  \caption{Subtyping}
+\end{array}
+  \]
+}
+{\footnotesize
+\[
+\begin{array}{l}
+n'+n = add(n',n)
+\qquad
+(x+n')+n = x+add(n',n)\\
+\textit{cond}(\overline{x},\tau,\overline{y},\tau')
+=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
+  \wedge \overline{y}\cupdot\overline{z}
+  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
+\\\qquad\qquad\qquad\qquad\qquad
+  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
+\end{array}
+\]
+}
+  \caption{Type Equality and Subtyping}
   \label{fig:checkc-subtype}
@@
 
-\myparagraph{Bounds Widening}
-%
-The bounds of NT array pointers may be widened at conditionals, and calls to $\estrlentext$.  Rule \textsc{T-If} handles normal
-branching operations; rule \textsc{T-IfNT} is specialized to the case
-of branching on $\estar{x}$ when $x$ is a NT array pointer whose upper
-bound is 0. In this case, true-branch $e_1$ is checked with $x$'s type
-updated so that its upper bound is incremented by 1; the else-branch
-$e_2$ is type-checked under the existing assumptions. For both rules,
-the resulting type is the join of the types of the two branches
-(according to subtyping). This is important for the situation when $x$
-itself is part of the result, since $x$ will have
-different types in the two branches.
-
-% \review{"Bounds Widening": can the join be always computed? this may be a hard
-%   problem...?}
-%   \mwh{If bounds expressions on two versions of an array pointer
-%   are very different, a precise join may not be possible, but an
-%   imprecise one is always possible by dropping the bounds
-%   expression entirely. Right Liyi?}
-
-Rule \textsc{T-Str} handles the case for when $\estrlentext(y)$ does
-not appear in a let binding. Rule \textsc{T-LetStr} handles the case
-when it does, and performs bounds widening. The result of the call is
-stored in variable $x$, and the type of $y$ is updated in $\Gamma$ when
-checking the let-body $e$ to indicate that $x$ is $y$'s upper
-bound. Notice that the lower bound $b_l$ is unaffected by the call to
-$\estrlentext(y)$; this is sound because we know that $\estrlentext$
-will always return a result $n$ such that $n \ge b_h$, the current
-view of $x$'s upper bound.
-The type rule tracks $\estrlentext$'s widened bounds
-within the scope of $x$, while the bound-widening effect in the semantics applies to the lifetime of $y$.
-Our type preservation theorem in Sec.~\ref{sec:theorem}
-shows that our type system is a sound model of the \CoreChkC semantics,
-and we discuss how we guarantee that the behavior of our compiler formalization and the semantics matches in Sec.~\ref{sec:compilation}.
-
-This rule also extends $\Theta$ when checking $e$, adding a predicate
-indicating that $x\ge 0$. To see how this information is used,
-consider this example.  The \code{return} on line 14 of
-Fig.~\ref{fig:strcat-ex} has an implicit static cast from the
-returned expression to the declared function type (see rule
-\textsc{T-Fun}, described below). In type checking the $\estrlentext$
-on line 4, we insert a predicate in $\Theta$ showing
-\code{x}$\;\ge 0$.  The static cast on line 14 is valid according to
-the last line in Fig.~\ref{fig:checkc-subtype}:
-
+\begin{DIFnomarkup}
+ \begin{figure}[t]
  {\small
-$$
-\tallarrayptr{0}{x}{\tau}{c} \sqsubseteq \tallarrayptr{0}{0}{\tau}{c}
-$$
-}
-because $0 \leq 0$ and $0 \leq x$, where the latter holds since
-$\Theta$ proves $x \ge 0$.  Without $\Theta$, we would need a
-dynamic cast.
-
-In our formal presentation, $\Theta$ is quite simple and is just meant
-to illustrate how static information can be used to avoid dynamic
-checks; it is easy to imagine richer environments of facts that can be
-leveraged by, say, an SMT solver as part of the subtyping
-check \cite{LiquidHaskell,Refinementlh}.
-
-% There is a difference between the \checkedc and \lang type systems in
-% terms of recognizing subtyping due to $\Theta$. \code{strncat_c} in
-% Fig.~\ref{fig:strcat-ex} represents what the \checkedc specification
-% can do, while \code{strncat} represents what \lang can do. In the
-% usage of a static casting operation in line 16 of
-% Fig.~\ref{fig:strcat-ex}, \lang formalization is smart enough to
-% infer that \code{x} must be a non-negative number because the creation
-% of \code{x} is from a \code{strlen} call.  \checkedc does not perform
-% predicate decision procedure to reason about the bounds statically;
-% thus, a dynamic casting is required in the case, such as line 30 in
-% \code{strncat_c} in Fig.~\ref{fig:strcat-ex}.  More differences are
-% in Sec.~\ref{sec:disc}.
-
-% Here, we introduce
-% two most common NT-array operations: $\estrlentext$ and branching
-% operations.  Line 4 in Fig.~\ref{fig:strcat-ex} shows the bound
-% widening behavior of the NT-array pointer usage in a $\estrlentext$
-% expression.  The upper-bound of pointer \code{a} is updated to
-% \code{x} through out the whole function body, which is the scope of
-% variable \code{x}.  Line 5-8 in Fig.~\ref{fig:checkedc-example} shows
-% the bound widening behavior of using NT-array pointers in a branching
-% operation. In \checkedc, if a Boolean guard has the form $\estar{x}$,
-% and $x$'s upper bound is $0$, then the \code{true} branch recognizes
-% the upper bound of $x$ incremented by $1$, because the end position in
-% the NT-array pointed to by $x$ is not a null-character in the
-% \code{true} branch; thus, there is at least one more valid position in
-% the NT-array after the current end position. More details discussion
-% are in Sec.~\ref{sec:disc}.
-
-\myparagraph{Dependent Functions and Let Bindings}
-%
+
+ \begin{mathpar}
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m n : \tint}
+
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\umode}}
+
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
+            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       { \xi \le m 
+     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
+       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
+
+   \inferrule
+       {\neg\funptr(\omega)\\ \xi \le m\\
+        \forall i \in [0,\size(\omega)) \;.\;
+            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+ \end{mathpar}
+ }
+{\footnotesize
+\[
+\begin{array}{l} 
+\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
+\qquad
+\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
+\end{array}
+\\
+\]
+}
+ \caption{Verification/Type Rules for Constants}
+ \label{fig:const-type}
+ \end{figure}
+\end{DIFnomarkup}
+
+\myparagraph{Constant Validity}
+Given a constant $\evalue{n}{\tau}$, 
+the type judgment $\Theta;\heap;\sigma \vdash_m n : \tau$ checks in \Cref{fig:const-type}
+if the constant is valid, 
+where $\heap$ is the initial heap that the constant resides on and
+$\sigma$ is a set of constant assumed to be checked.
+A global function store $\Xi$ is also required in checking the validity of a function pointer.
+A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
+and the address stores a function with the right type.
+
+The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
+where every element in the pointer range ($[0,\size(\omega))$) should be well typed.
+A checked pointer checks validity in type step as rule \textsc{T-ConstC},
+while a tainted/unchecked pointer does not check for such during the type checking.
+Tainted pointers are verified through the validity check in dynamic execution as we mentioned above.
+
+\myparagraph{Unchecked and Checked Blocks}
+%
+During the type checking,
+Both $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$
+check all free variables are within $\overline{x}$,
+and the types for $\overline{x}$ have no checked pointers.
+A $\echeckedtext$ and $\euncheckedtext$ block represents 
+the transition stage from a checked to an checked region, or vice versa.
+We need to make sure no checked pointers are information exposed to unsafe code regions.
+
+\myparagraph{Dependent Function Pointers and Let Bindings}
+%
+Rule \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
+type dependency. In particular, the result of evaluating a $\elettext$
+may have a type that refers to one of its bound variables (e.g., if
+the result is a checked pointer with a variable-defined bound); if so,
+we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). Note
+that we restrict the expression $e_1$ to syntactically match the
+structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
+
+Rule \textsc{T-Ret} types a $\erettext$ expression, which does not
+appear in source programs but is introduced by the semantics when
+evaluating a let binding (rule \textsc{S-Let} in
+Fig.~\ref{fig:semantics}); this rule is needed for the preservation
+proof.
+% After the evaluation of a let binding a variable $x$ concludes,
+%we need to restore any prior binding of $x$, which is either
+%$\bot$ (meaning that there is no $x$ originally) or some value
+%$\evalue{n}{\tau}$.
+
 Rule \textsc{T-Fun} is the standard dependent function call rule. It
-looks up the definition of the function in the function environment
-$\Xi$, type-checks the actual arguments $\overline{e}$ which have
+learns the function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$) of $e$ by a recursive type check,
+type-checks the actual arguments $\overline{e}$ which have
 types $\overline{\tau'}$, and then confirms that each of these types
-is a subtype of the declared type of $f$'s corresponding parameter. Because
-functions have dependent types, we substitute each parameter $e_i$ for
+is a subtype of the declared type of $e$'s corresponding parameter ($\overline{\tau}$). 
+Because functions have dependent types bounded by $\overline{x}$, 
+we substitute each parameter $e_i$ for
 its corresponding parameter $x_i$ in both the parameter types and the
@@
 
-Rule \textsc{T-Let} types a $\elettext$ expression, which also admits
-type dependency. In particular, the result of evaluating a $\elettext$
-may have a type that refers to one of its bound variables (e.g., if
-the result is a checked pointer with a variable-defined bound); if so,
-we must substitute away this variable once it goes out of scope. Note
-that we restrict the expression $e_1$ to syntactically match the
-structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
-
-Rule \textsc{T-Ret} types a $\erettext$ expression, which does not
-appear in source programs but is introduced by the semantics when
-evaluating a let binding (rule \textsc{S-Let} in
-Fig.~\ref{fig:semantics}); this rule is needed for the preservation
-proof. After the evaluation of a let binding a variable $x$ concludes,
-we need to restore any prior binding of $x$, which is either
-$\bot$ (meaning that there is no $x$ originally) or some value
-$\evalue{n}{\tau}$.
-
-\subsection{Type Soundness and Blame}\label{sec:theorem}
+\subsection{Type Soundness, Non-exposure, Non-crashing}\label{sec:theorem}
 
@@
 In this subsection, we focus on our main meta-theoretic results about
-\lang: type soundness (progress and preservation) and
-blame.
-  These proofs have been carried out in our
-  Coq model, found at \url{https://github.com/plum-umd/checkedc}.
-
-The type soundness theorems rely on several notions of
-\emph{well-formedness}:
+\lang: type soundness (progress and preservation),
+non-exposure, and non-crashing.
+These proofs have been carried out in our Coq model.
+
+Type soundness relies on several \emph{well-formedness}:
 
@@
 \begin{defi}[Heap Well-formedness]
-A heap $\heap$ is well-formed iff (i) $\heap(0)$ is undefined, and
-(ii) for all $\evalue{n}{\tau}$ in the range of $\heap$, type $\tau$
+For every $m$, A heap $\heap$ is well-formed iff (i) $\heap(m,0)$ is undefined, and
+(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
 contains no free variables. 
@@
 $\Gamma(x) = \tau$ for some $\tau$ and 
-$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq \tau$. 
+$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
 \end{defi}
 
-\begin{defi}[Stack-Heap Consistency]
+\begin{defi}[Checked Stack-Heap Consistency]
 A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---iff
-for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ implies $\heap;\emptyset \vdash n:\tau$.
+for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
 \end{defi}
 
-\begin{defi}[Heap-Heap Consistency]
+\begin{defi}[Checked Heap-Heap Consistency]
 A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---iff
-for every constant $n$, $\heap;\emptyset \vdash n:\tau$ implies $\heap';\emptyset \vdash n:\tau$.
+for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
 \end{defi}
@@
   Fig.~\ref{fig:const-type}; the judgment
-  $\heap;\sigma \vdash n : \tau$ is used to confirm literal
+  $\Theta;\heap;\sigma \vdash_m n : \tau$ is used to confirm literal
   well-typing, where $\sigma$ is a set of pointer literals already
@@
 
-Progress now states that terms that don't reduce are either values or their
-mode is unchecked:
+Progress states that a \systemname program can always make a move:
 
@@
 
-For any \checkedc program $e$, heap $\heap$, stack
+For any \systemname program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 
-\item $m = \umode$, or there exists $E$ and $e'$, such that $e = E[e']$ and $\mode(E) = \umode$.
-
 \end{itemize}
 \end{thm}
-{\em Proof:} By induction on the typing derivation.
+%{\em Proof:} By induction on the typing derivation.
 
 \noindent
-Preservation states that a reduction step preserves both the
+There are two forms of preservation regarding the checked and unchecked regions.
+Checked Preservation states that a reduction step preserves both the
 type and consistency of the program being reduced.
 
-\begin{thm}[Preservation]
-For any \checkedc program $e$, heap $\heap$, stack
+\begin{thm}[Checked Preservation]
+For any \systemname program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 \longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
-consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$, $\Theta'$ and $\tau'$ that are well formed, consistent
+checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
+$\Gamma'$, $\Theta'$ and $\tau'$ that are well formed, checked region consistent
 ($\Gamma';\Theta'\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
@@
 \end{thm}
-{\em Proof:} By induction on the typing derivation.
-\smallskip
-
-Using these two theorems we can prove our main result, {\em blame},
-which states that
-  if a well-typed program is \emph{stuck}---expression
+%{\em Proof:} By induction on the typing derivation.
+%\smallskip
+\begin{thm}[Unchecked Preservation]
+For any \systemname program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
+\end{thm}
+
+Using the above theorem, we first show the non-exposure theorem,
+where code in unchecked region cannot observe know a valid checked pointer address.
+
+\begin{thm}[Non-Exposure]
+For any \systemname program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
+where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$, if $x$ is a checked pointer, it is a null one.
+\end{thm}
+
+Using these two theorems we can prove our main result, {\em non-crashing},
+which states that a well-typed program can never be \emph{stuck} (expression
 $e$ is a non-value that cannot take a step\footnote{Note that
@@
   but $\heap$ is undefined at $n$ will be stuck, and violates spatial
-  safety.}---the cause must be the
-(past or imminent) execution of code in an unchecked region.
+  safety.}).
 
@@
 
-\begin{thm}[The Blame Theorem]\label{thm:blame} For any \checkedc
+\begin{thm}[Non-Crashing]\label{thm:blame} For any \systemname
   program $e$, heap $\heap$, stack
@@
 $\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
-$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$ and $r$ is \emph{stuck}, then
-there exists $j \in [1,k]$, such that $m_j = \umode$, or there exists $E$ and $e'$, such that $r = E[e']$ and $\mode(E) = \umode$.
+$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
 \end{thm}
-{\em Proof:} By induction on the number of steps of the \checkedc
-evaluation ($\longrightarrow_m^*$), using progress and preservation to
-maintain the invariance of the assumptions.
+%{\em Proof:} By induction on the number of steps of the \checkedc
+%evaluation ($\longrightarrow_m^*$), using progress and preservation to
+%maintain the invariance of the assumptions.
 
@@
 % }
-  Compared to \citet{ruef18checkedc-incr}, proofs for
-  \lang were made challenging by the addition of dependently typed
-  functions and dynamic arrays, and the need to handle bounds widening for NT
-  array pointers. These features required changes in the runtime
-  semantics (adding a stack, and dynamically changing bounds) and in
-  compile-time knowledge of them (to soundly typing widened bounds).
+%  Compared to \citet{ruef18checkedc-incr}, proofs for
+%  \lang were made challenging by the addition of dependently typed
+%  functions and dynamic arrays, and the need to handle bounds widening for NT
+%  array pointers. These features required changes in the runtime
+%%  semantics (adding a stack, and dynamically changing bounds) and in
+ % compile-time knowledge of them (to soundly typing widened bounds).
 

### 2022-07-16 sandp23/evaluation.tex
update.

--- 2022-06-25 sandp23/checkedc-formal/evaluation.tex
@@
-\section{Random Testing via the Implementation}\label{sec:evaluation}
+\section{Evaluation}\label{sec:evaluation}
 
@@
 
-In addition to using the \lang Redex model to establish simulation of
-compilation (Section~\ref{sec:meta}), we also used it to gain confidence
-that our model matches the Clang \checkedc implementation.
-Disagreement on outcomes signals a bug in either the model or the
-compiler itself. Doing so allowed us to quickly iterate on the
-design of the model while adding new features, and revealed
-several bugs in the Clang \checkedc implementation.
+\begin{itemize}
 
-\myparagraph{Generating Well Typed Terms} For this random generation, we follow
-the approach of~\citet{PalkaAST11} to generate well-typed \checkedc
-terms by viewing the typing rules as generation rules.
-%
-%Suppose we have a context $\textcolor{red}{\Gamma}$, a mode
-%$\textcolor{red}{m}$ and a type $\textcolor{red}{\tau}$, and we are
-Suppose we have a context $\Gamma$, a mode
-$m$ and a type $\tau$, and we are
-trying to generate a well-typed expression. We can do that by
-reversing the process of type checking, selecting a typing rule and
-building up an expression in a way that satisfies the rule's premises.
+\item provide evidence that the \systemname compiler is efficient. Compare the compiler with respect to other work, like RLBox, also the previous \checkedc compiler.
 
-Recall the typing rule for dereferencing an
-array pointer, which we depict below as
-\textsc{G-DefArr}\footnote{Generator rules G-* correspond one to one
-  with the type rules T-* in Sec.~\ref{sec:type-system}.}, color-coded to represent \textcolor{red}{inputs} and \textcolor{blue}{outputs} of the
-generation process:%
-\footnote{This input-output marking is commonly called a mode in
-  the literature, but we eschew this term to avoid
-  confusion with our pointer mode annotation.}
-%
-%\begin{figure}[t]
-\begin{mathpar}
-    \inferrule[G-DefArr]
-              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} : \tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m'}} \\
-\textcolor{red}{m} \leq \textcolor{cyan}{m'} 
-              }
-              {\textcolor{red}{\Gamma};\textcolor{red}{\Theta} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\estar{e}} : \textcolor{red}{\tau}}
-%  
-%     \inferrule[T-Dfa] {
-% %      \textcolor{cyan}{m'} = \texttt{choose}(C,U) \text{ if } \textcolor{red}{m} = U
-% %      \text{, otherwise } \textcolor{cyan}{m'} = C
-% %      \\
-%       \textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} : \tptr{\textcolor{red}{\tau}}{\textcolor{cyan}{m'}} \\      
-%       \textcolor{red}{m} \leq \textcolor{cyan}{m'} \\
-% }
-%     {\textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\estar{e}} : \textcolor{red}{\tau}}
-  \end{mathpar}
-%   \caption{T-DFA as a generation rule}
-%   \label{fig:tdfagen}
-% \end{figure}
-%
-If we selected \textsc{G-DefArr} for generating an expression, the
-generated expression has to have the form $\textcolor{blue}{*e}$,
-for some $\textcolor{blue}{e}$, to be generated according
-to the rule's premises.
-%
-To satisfy the premise
-$\textcolor{red}{\Gamma};\textcolor{red}{\Theta}
-\vdash_{\textcolor{red}{m}} \textcolor{blue}{e} :
-\tptr{\tallarraybc{\textcolor{cyan}{\bvar}}{\textcolor{red}{\tau}}}{\textcolor{cyan}{m'}}$,
-we essentially need to make a recursive call to the generator, with
-appropriately adjusted inputs.
-%
-However, the type in this judgment is not fixed yet---it contains
-three unknown variables: $\textcolor{cyan}{m'}$,
-$\textcolor{cyan}{\bvar}$, and $\textcolor{cyan}{\kappa}$---that need
-to be generated before making the call.
-%
-Looking at the second premise informs that generation:
-if the input mode $\textcolor{red}{m}$ is $\umode$, then $\textcolor{cyan}{m'}$
-needs to be $\umode$ as well; if not, it is unconstrained,
-just like $\textcolor{cyan}{\bvar}$ and $\textcolor{cyan}{\kappa}$, and therefore all three are free to be generated at random.
-%
-Thus, the recursive call to generate $\textcolor{blue}{e}$ can now
-be made, and the \textsc{G-DefArr} rule returns $\textcolor{blue}{*e}$
-as its output.
+\item provide user experience of \systemname.
+We restrict the use of checked pointers compared to previous checked-c compiler.
+Is the restriction arrangable.
+We can say that the tainted shells are auto-matically generated, so we have a mechanism for auto-generating tainted pointers if necessary.
 
-% If we selected \textbf{T-Dfa} for generating an expression, the
-% expression we generate has to have the form $\textcolor{blue}{*e}$,
-% for some expression $\textcolor{blue}{e}$, to be generated according
-% to the rule's premises.
-% %
-% To satisfy the premise $\textcolor{red}{\Gamma}
-% \vdash_{\textcolor{red}{m}} \textcolor{blue}{e} :
-% \tptr{\textcolor{red}{\tau}}{\textcolor{cyan}{m'}}$, we essentially
-% need to make a recursive call to the generator, with appropriately
-% adjusted inputs.
-% %
-% However, the type in this judgement is not fixed yet---it contains an
-% unknown variable $\textcolor{cyan}{m'}$ that needs to be generated
-% before making the call.
-% %
-% Looking at the second premise informs that generation:
-% if the input mode $\textcolor{red}{m}$ is $\umode$, then $\textcolor{cyan}{m'}$
-% needs to be $\umode$ as well; if not, it is unconstrained and we're free
-% to generate it at random to be $\cmode$ or $\umode$.
-% %
-% Thus, the recursive call to generate $\textcolor{blue}{e}$ can now
-% be made, and the \textbf{T-Dfa} rule returns $\textcolor{blue}{*e}$
-% as its output.
+\item if we have space, we can re-introduce random testing a little, saying that how it helps us to develop the compiler.
 
-Using such generator rules, we can create a generator for random
-well-typed terms of a given type in a straightforward manner: find all
-rules whose conclusion matches the given type and then randomly choose
-a candidate rule to perform the generation. To ensure that this
-process terminates, we follow the standard practice of using ``fuel'' to
-bound the depth of the generated terms; once the fuel is exhausted,
-only rules without recursive premises are
-selected~\cite{Pierce:SF4}. Similar methods were used for
-generating top level functions and \kw{struct} definitions.
+\item we can then talk about the possible bugs we find in the \checkedc compiler for function pointer or the RLBox bugs.
 
-%^ \leo{The following is an insanely large, undigestible paragraph, but now moved to the correct place.}
-%^ The generator follows the typing rules in the redex model to generate
-%^ well typed terms. Since the redex model only supports top level
-%^ functions the functions were generated separately, as were struct
-%^ definitions. Given a list of struct definitions, a list of functions,
-%^ and a type the generator randomly chooses the structure of a term that
-%^ can have that type. This is based on the typing rules. Then depending
-%^ on the holes needed to be filled in the term, new expressions are
-%^ generated. For example when attempting to generate an expression of
-%^ type (ptr c int) with fuel 2 the following steps might be taken. First
-%^ a conditional expression is randomly chosen. In order for this to have
-%^ the correct type a random expression of type int with fuel 1 is
-%^ generated as the guard. Then both branched must have type (ptr c
-%^ int). They will recursively generate but with reduced fuel. If the
-%^ fuel is less than or equal to zero then either a variable with the
-%^ correct type will be chosen from the enviornment or a malloc
-%^ expression will be generated. The generator takes fuel so that
-%^ generating expressions will always terminate.
+\end{itemize}
 
-While using just the typing-turned-generation rules is in theory
-enough to generate all well-typed terms, it's more effective in
-practice to try and exercise interesting patterns.  As
-in~\citet{PalkaAST11} this can be viewed as a way of adding admissible
-but redundant typing rules, with the sole purpose of
-using them for generation. For example, below is one such rule,
-\textsc{G-ASTR}, which creates an initialized null-terminated string that
-is statically cast into an array with bounds $(0,0)$.
-%
-%\begin{figure}[t]
-  \begin{mathpar}
-    \inferrule[G-ASTR]
-    {
-      \textcolor{cyan}{i} \in \mathbb{N}^* \quad\quad
-      \textcolor{cyan}{n_0,\ldots,n_{i-1} } \in \mathbb{Z} \quad\quad
-      \texttt{fresh}(\textcolor{cyan}{x})\\
-      \textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}}
-      \textcolor{blue}{e'} :
-      \tptr{\textcolor{red}{\tntarray{0}{\textcolor{cyan}{i}}{\tint}}}{c}
-      \\      
-% \eassign{(\ebinop{x}{0})}{n_0};\ldots;\eassign{(\ebinop{x}{i-1})}
-      \textcolor{blue}{e} = \elet{\textcolor{cyan}{x}}{\textcolor{blue}{e'}}{\texttt{(init $\textcolor{cyan}{x}$ with }\textcolor{cyan}{n_0,\ldots
-        n_{i-1}}\texttt{)};\textcolor{cyan}{x}}
-      % \textcolor{cyan}{m'} = \texttt{choose}(C,U) \text{ if } \textcolor{red}{m} = U
-      % \text{, otherwise } \textcolor{cyan}{m'} = C
-      % \\
-}
-    {\textcolor{red}{\Gamma} \vdash_{\textcolor{red}{m}} \textcolor{blue}{\ecast{\tptr{\tntarray{0}{0}{\tint}}{c}}{e}} : \textcolor{red}{\tptr{\tntarray{0}{0}{\tint}}{c}}}
-  \end{mathpar}
-%  \caption{Specialized rule for generating initialized strings}
-%  \label{fig:gastr}
-%\end{figure}
-%
-Given some positive number $\textcolor{cyan}{i}$, numbers
-$\textcolor{cyan}{n_0, \ldots, n_{i-1}}$, and a fresh variable
-$\textcolor{cyan}{x}$ (which are arbitrarily generated), we can
-recursively generate a pointer $\textcolor{blue}{e'}$ with bounds
-$(0,\textcolor{cyan}{i})$, and initialize it with the generated $n_j$
-using $\textcolor{cyan}{x}$ to temporarily store the pointer.
-
-This rule is particularly useful when combined with \textsc{G-IfNT}
-since there is a much higher chance of obtaining a non-zero value when
-evaluating $\estar{p}$ in the guard of $\eiftext$, skewing the
-distribution towards programs that enter the $\ethentext$ branch.
-Relying solely on the type-based rules, entering the $\ethentext$
-branch requires that \textsc{G-AssignArr} was chosen before \textsc{G-IfNT},
-and that assignment would have to appear before $\eiftext$, which
-means additional \textsc{G-Let} rules would need to be chosen: this
-combination would therefore be essentially impossible to generate in
-isolation.
-
-Adding admissible generation rules like \textsc{G-Astr} in this
-manner, as described in~\citet{PalkaAST11}, is a manual process. It is
-guided by gathering statistics on the generated data and focusing on
-language constructs that appear underrepresented in the posterior
-distribution. For example, we arrived at the \textsc{G-Astr} rule by
-recognizing that the pure type-based generation was not generating
-non-trivial null-terminated strings, and then analyzing the sequence
-of random choices that could lead to their generation.
-
-% Originally only the
-% typing rules were used to generate expressions. Eventualy new options
-% were added to create more interesting programs. Interesting programs
-% include allocating memory for an array and then initializing it with
-% values, and casting arrays to a smaller type so that bounds widening
-% can be triggered. There were also some issues with generating strlen
-% expressions. After a variable is assigned to the strlen of a pointer
-% that bounds for that pointer contain the variable. This caused issues
-% with scope as an array with bounds dependent on that variable could be
-% returned. In order to prevent these issues expressions that use strlen
-% are only generated in the form \code{let \_ = let x = (strlen y)} in
-% (expression generated with updated bounds for y) in (some literal
-% expression that does not contain y). Once expressions from the redex
-% model are generated they are compiled into CheckedC and compiled with
-% the clang compiler. One problem is that type annotations are needed in
-% Checked-C but do not exist in the Redex model. To properly translate
-% an enviornment is passed around. As various parts are translated types
-% are added to the enviornment so when declarations are needed their
-% types can be found.  Several small changes had to be made to
-% expressions since the redex model which used immutable values, had
-% expressions that were not allowed in the clang version (e.g. let
-% expressions, conditional expressions with complex bound types in the
-% branches, using calloc to initialize nt-arrays, and casting motivating
-% expressions). Let, if and cast expressions got lifted to functions. We
-% defined a calloc for nt-arrays using a dynamic bounds cast. The int
-% type was changed to char. This is because while in the specification
-% any nt-array can be passed to strlen the clang compiler only allows
-% arrays of type const char *. The expression was evaluated and then
-% printed to stdout.
-%  
-%  
-%  
-% \leo{TODO: adding interesting patterns (mentioned later), just like Palka
-%   adds seq}
-% \leo{I don't understand why the conversion tool is important here. I
-%   moved the following para up, as it has to do with generation.}
-% % \yiyun{I'm not sure if patterns can be written as a clean rule. For
-% %   example, we want to add rules that creates an ntarray of size n,
-% %   initializes the array, and then }
-% \yiyun{I'm not sure if T-ASTR is exactly implemented like that in
-%   Deena's random generator, but we do have something that performs
-%   this type of initialization.}
-% The conversion tool can be used separately from the random generator
-% to speed up the manual testing process, where we construct programs
-% that we believe are likely to trigger rules that have subtle
-% definitions and are likely to cause failure in the experimental Clang
-% implementation. For example, we were able to find that % \yiyun{is this
-% % accurate? not that familiar with itype} the \code{itype} 
-% \checkedc wrapper
-% of \code{strlen} does not perform null check at the
-% interface, which would lead to undefined behavior within the checked
-% region. We integrate
-% these manually found patterns into the random generator to complement the
-% type-based generation approach, allowing us to cover more interseting
-% cases during random testing.
-
-\myparagraph{Generating Ill-typed Terms}
-%
-We can use generated well-typed terms to test our simulation theorem
-(Section~\ref{sec:compilation}) and test that \lang and \checkedc
-Clang agree on what is type-correct. But it is also useful to generate
-ill-typed terms to test that \lang and \checkedc Clang also agree on
-what is not.  However, while it is easy to generate arbitrary
-ill-typed terms, they would be very unlikely to trigger any
-inconsistencies; those are far more likely to exist on the boundary
-between well- and ill-typedness. Therefore, we also manually added
-variations of existing generation rules modified to be slightly more
-permissive, e.g., by relaxing a single premise, thus allowing terms
-that are ``a little'' ill-typed to be generated. Unlike coming up with
-admissible generation rules like \textsc{G-Astr} (which is quite
-challenging to automate), systematically and automatically relaxing
-premises of existing rules seems feasible, and worthwhile future work.
-
-% \myparagraph{Testing Model Properties}
-% The generator is used to establish confidence in properties that are
-% expected to hold in our model, such as progress and preservation of
-% \lang, or more notably the simulation property between \lang and
-% \elang (Theorem~\ref{simulation-thm}).
-% %
-% As discussed in
-% Section~\ref{sec:compilation}, we use the simulation property to
-% ensure that our source language can be compiled into a language that
-% does not rely on fat pointers. For simulation to hold, 
-% the semantics must avoid accessing information that would
-% necessitate the use of fat pointers during compilation. This is unsurprisingly
-% a tricky process, and we found mistakes in early versions of our
-% desing where we wrote widened bounds onto the heap, or passed
-% widened bounds to functions, making them accessible outside of the
-% current stack frame (thus implying the existence of fat
-% pointers). Random testing also helped us discover the limitation of
-% the standard simulation theorem and motivated us to rephrase it
-% in such a way that is compatible with our definition of compilation.
-% % relax our simulation theorem to require only the confluence of the
-% % two compiled programs by providing examples where the compilation
-% % process can generate redundant stack variables that would break the more
-% % standard simulation property
-% . \yiyun{does this deserve a figure to
-%   explain? or maybe I should mention why our simulation is slightly
-%   off from what a ``standard'' simulation should look like in
-%   compilation? }\leo{This should have already been discussed in the
-%   statement of simulation - point there?}
-
-\myparagraph{Random Testing for Language Design}
-%
-We used our Redex model and random generator to successfully guide the
-design of our formal model, and indeed the Clang Checked C
-implementation itself, which is being actively developed. To that end, we implemented a
-conversion tool that converts \lang into a subset of the \checkedc
-language and ensured that model and implementation exhibit the same
-behavior (accept and reject the same programs and yield the same return
-value).
-
-This approach constitutes an interesting twist to traditional
-model-based checking approaches.  Usually, one checks that the
-implementation and model agree on all inputs \emph{of the
-  implementation}, with the goal of covering as many behaviors as
-possible. This is the case, for example, in \citet{lambdajs}, where
-they use real test suites to demonstrate the faithfullness of their
-core calculus to Javascript. Our approach and goal in this work is
-essentially the opposite: as the Clang Checked C implementation does
-not fully implement the Checked C spec, there is little hope of
-covering all terms that are generated by Clang Checked C. Instead,
-we're looking for \emph{inconsistencies}, which could be caused by
-bugs either in the Clang Checked C compiler or our own model.
-
-One inconsistency we found comes from the following:
-
-{\small 
-\begin{lstlisting}[xleftmargin=4 mm]
- array_ptr<char> fun(void) : count(3) {
-   array_ptr<char> x : count(3);
-   x = calloc(3, sizeof(char));
-   return x+3;
- }
- int main(void) {
-   *(fun()) = 0;
-   return 0;
- }
-\end{lstlisting}
-}% 
-%
-\noindent
-In this code, the function \code{fun} is supposed to return a
-checked array pointer of size 3. Internally, it allocates such an
-array, but instead of returning the pointer \code{x} to that array, it
-increments that pointer by 3. Then, the \code{main} function just
-calls \code{fun}, and tries to assign 0 to its result. Our model
-correctly rules out this program, while the Clang Checked C
-implementation happily accepted this out-of-bounds
-assignment. Interestingly, it correctly rejected programs where the
-array had size 1 or 2. This inconsistency has been fixed in the latest
-version of the compiler.
-
-We also found the opposite kind of inconsistency---programs that
-the Clang Checked C implementation rejects contrary to the spec.
-For instance:%
-\footnote{After minimization, this turned out to be a known issue: 
-  \url{https://github.com/microsoft/checkedc-clang/issues/1008}}
-
-{\small 
-\begin{lstlisting}[xleftmargin=4 mm]
-array_ptr<int> f(void) : count(5) {
-  array_ptr<int> x : count(5) =
-    calloc<int>(5, sizeof(int)); 
-  return x;
-}
-array_ptr<int> g(void ) : count(5) {
-  array_ptr<int> x : count(5) =
-    calloc<int>(5, sizeof(int)); 
-  return x+3;
-}
-int main(void) {
-  return *(0 ? g() : f() + 3);
-}    
-\end{lstlisting}
-}% 
-\noindent
-In this piece of code both \code{f} and \code{g} functions compute a
-pointer to the same index in an array of size 5 (as \code{f} calls
-\code{g}). The \code{main} function then creates a ternary expression
-whose branches call \code{f} and \code{g}, but the Clang Checked C
-implementation rejects this program, as its static analysis is not
-sophisticated enough to detect that both branches have the same type.
 % 

### 2022-07-17 sandp23/compilation.tex
update compilation.

--- 2022-06-25 sandp23/checkedc-formal/compilation.tex
@@
 
-The semantics of \lang uses annotations on pointer literals in order
-to keep track of array bounds information, which is used in premises
+The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
+that track array bound information, which is used in premises
 of rules like \textsc{S-DefArray} and
-  \textsc{S-AssignArr} to prevent spatial safety violations. However, in the real
-implementation of \checkedc, which extends Clang/LLVM, these annotations
-are not present---pointers are represented as a single
-machine word with no extra metadata, and bounds
-  checks are not handled by the machine, but inserted by the
-  compiler.
+  \textsc{S-AssignArr} to prevent spatial safety violations.
+The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
+for verifying pointers being not null and the bounds are within their limits.
+The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.
 
-This section shows how \lang
-annotations can be safely erased: using static information a compiler
-can insert code to manage and check bounds metadata, with no loss of
-expressiveness. We present a compilation algorithm that converts from
-\lang to \elang, an untyped language without metadata
-annotations. The syntax and semantics \elang
-  closely mirrors that of \lang; it differs only in that literals lack
-  type annotations and its operational rules perform no
-  bounds and null checks, which are instead inserted during
-  compilation. Our compilation algorithm is evidence that \lang's
-  semantics, despite its apparent use of fat pointers, faithfully
-  represents Checked C's intended behavior. The algorithm also sheds
+\begin{figure}[t!]
+{\small
+\begin{tabular}{|c|c|c|c|}
+\hline
+& \cmode & \tmode & \umode \\
+\hline
+& \text{Impl} / \text{Foml} & \text{Impl} / \text{Foml} & \text{Impl} / \text{Foml} \\
+\hline
+\cmode & $\estar{x}$ / $\getstar{\cmode}{x}$ 
+ & $\texttt{sand\_get}(x)$ / $\getstar{\umode}{x}$ &  $\times$ \\
+\hline
+\umode & $\times$
+ & $\estar{x}$ / $\getstar{\umode}{x}$ &  $\estar{x}$ / $\getstar{\umode}{x}$ \\
+\hline
+\end{tabular}
+
+}
+\caption{Compiled Targets for Dereference}
+\label{fig:flagtable}
+\end{figure}
+
+In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
+i.e., \cmode pointer points to checked regions, while \tmode and \umode pointers point to unchecked regions.
+Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
+In the compiled LLVM code, pointer access operations have different syntax when the modes are different.
+\Cref{fig:flagtable} lists the different compiled syntax for a deference operation ($\estar{x}$) for the compiler implementation (\text{Impl}) and formalism (\text{Foml}).
+The columns represent different pointer modes and the rows represent context modes.
+For example, when we have a \tmode pointer in a \cmode-mode region, we use the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data.
+In formalism, we create a new deference data-structure in LLVM on top of the existing $\estar{x}$ operation: $\getstar{m}{x}$.
+if the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
+
+In the compilation, we use a function \code{mode} to get the pointer mode of a function learning from the pointer's type.
+Thus, the compiler can determine the correct compiled syntax for a pointer access operation with 
+the knowledge of the current context mode.
+We also build a context switching mechanism to change 
+the context mode when we meet $\echeckedtext$ and $\euncheckedtext$ blocks.
+We describe the mechanism shortly.
+
+This section shows how \systemname deals with pointer modes, mode switching and function pointer compilations, 
+with no loss of expressiveness;
+as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
+For the compiler formalism, 
+we present a compilation algorithm that converts from
+\systemname to \elang, an untyped language without metadata
+annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
+In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
+$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$.
+The algorithm sheds
   light on how compilation can be implemented in the real Checked C
@@
   differences with LLVM IR).
+
+%This section shows how \systemname deals with 
+%annotations can be safely erased: using static information a compiler
+%can insert code to manage and check bounds metadata, with no loss of
+%expressiveness. We present a compilation algorithm that converts from
+%\lang to \elang, an untyped language without metadata
+%annotations. The syntax and semantics \elang
+  %closely mirrors that of \lang; it differs only in that literals lack
+  %type annotations and its operational rules perform no
+  %bounds and null checks, which are instead inserted during
+  %compilation. Our compilation algorithm is evidence that \lang's
+  %semantics, despite its apparent use of fat pointers, faithfully
+  %represents Checked C's intended behavior. The algorithm also sheds
+  %light on how compilation can be implemented in the real Checked C
+  %compiler, while eschewing many important details (\elang has many 
+  %differences with LLVM IR).
 
@@
 There is now a \elang output $\dot e$ and an input $\rho$, which maps
-each \code{nt_array_ptr} variable \code{p} to a pair of \emph{shadow
-  variables} that keep \code{p}'s up-to-date upper and lower bounds; these
-may differ from the bounds in \code{p}'s type due to bounds
+each (NT-)array pointer variable to its mode and
+each variable \code{p} to a pair of \emph{shadow
+  variables} that keep \code{p}'s up-to-date upper and lower bounds; 
+these may differ from the bounds in \code{p}'s type due to bounds
 widening.\footnote{Since lower bounds are never widened, the
@@
 Each rule performs up to three tasks: (a) conversion of $e$ to
-A-normal form; (b) insertion of dynamic checks; and (c) insertion of
-bounds widening expressions.
+A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
+and (c) generate right pointer accessing expressions based on modes.
 %
@@
 next two steps of compilation are more interesting.
+We state them based on different \systemname operations.
 
@@
 \begin{lstlisting}[mathescape,xleftmargin=4 mm]
-/* p : $\color{purple!40!black}\tntarrayptr{0}{0}{\tint}{\cmode}$  */
-/* $\color{purple!40!black}\rho$(p) = p_lo,p_hi */
-{
-  let x = strlen(p);
-  if (x > 1) putchar(*(p+1));
+int deref_array(n : int,
+     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
+     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
+  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
+  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
+    * p;
+    * q = 1;
 }
+...
+/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
+/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
+deref_array(5, p0, q0);
     \end{lstlisting}
@@
 \end{frame}
-\begin{lstlisting}[xleftmargin=4 mm]
-{
-  assert(p_lo <= 0 && 0 <= p_hi); // bounds check
-  assert(p != 0); // null check
-  let x = strlen(p);
-  let p_hi_new = x;
-  p_hi = max(p_hi, p_hi_new);
-  if (x > 1) {
-    assert(p != 0); // null check for p + 1
-    let p_1 = p + 1;
-    assert(p_lo <= 1 && // bounds check for p + 1
-     1 <= p_hi);    
-    putchar(*p_1);
-  }
-}
-\end{lstlisting}
-\end{small}
-\caption{Compilation Example for Check Insertions
-% \review{- Fig 8, line 1: did you mean `bounds` instead of `count`? I'm confused
-% - Fig 8, line 8: why null-check `p` again?}
-% \yiyun{Yes, fixed. The second null check is due to the pointer
-%   arithmetic on p; our transformation does not attempt to optimize
-%   away provably redundant checks.}
-}
-\label{fig:compilationexample}
-\end{figure}
-
-\begin{figure}[t!]
-  \begin{small}
 \begin{lstlisting}[mathescape,xleftmargin=4 mm]
-int deref_array(n : int,
-     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$) {
-  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi */
-  if (* p)
-    (* (p + 1))
-    else 0
-}
-...
-/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
-deref_array(5, p0);
-    \end{lstlisting}
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\begin{lstlisting}[xleftmargin=4 mm]
-deref_array(n, p) {
-  let p_lo = 0;
-  let p_hi = n;
+deref_array(int n, int* p, int * q) {
+  //m is the current context mode
+  let p_lo = 0; let p_hi = n; 
+  let q_lo = 0; let q_hi = n; 
   /* runtime checks */
@@
   assert(p != 0);
-  let p_derefed = *p;
-  if (p_derefed != 0) {
-    /* widening */
-    if (p_hi == 0) {
-      p_hi = p_hi + 1;
-    }
-    /* null check before pointer arithmetic */
-    assert(p != 0);
-    let p0 = p + 1;
-    assert(p_lo <= 1 && 1 <= p_hi);
-    (* p0)
-  }
-  else {
-    0
-  }
+  *(mode(p) $\wedge$ m,p);
+  verify(q, not_null && q_lo <= 0 && 0 <= q_hi);
+  *(mode(q) $\wedge$ m,q)=1;
 }
 ...
-deref_array(5, p0);
+deref_array(5, p0, q0);
     \end{lstlisting}
@@
 
-During compilation, $\Gamma$ tracks the lower and upper bound
-associated with every pointer variable according to its type. At each
-declaration of a \code{nt_array_ptr} variable \code{p}, the
-compiler allocates two \emph{shadow variables}, 
-stored in $\rho(p)$; these are initialized to \code{p}'s declared bounds
-and will be updated during bounds widening.\footnote{Shadow variables
-  are not used for \code{array_ptr} types (the bounds expressions are)
-  since they are not subject to bounds widening.} 
-Fig.~\ref{fig:compilationexample} shows how an invocation of
-\code{strlen} on a null-terminated string is compiled into C
-code. Each dereference of a checked pointer requires a null check
-(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
-compiler makes explicit: Line~$3$ of the generated code has the null
-check on pointer \code{p} due to the \code{strlen},
-  and a similar check happens
-  at line~$8$ due to the pointer arithmetic on \code{p}.
-Dereferences also require bounds checks: line~$2$ checks \code{p} is
-in bounds before computing \code{strlen(p)}, while line~$10$ does
-likewise before computing \code{*(p+1)}.
+\myparagraph{Pointer Accesses and Modes}
+%
+In every declaration (or the beginning of a function body) of a pointer,
+if the poniter is an (NT-)array one,
+we first allocates two \emph{shadow variables}
+to track the lower and upper bounds potentially changed for pointer arithmetic and NT-array bound widening.
+We also associated with every \cmode-mode pointer variable according to its type;
+we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
+In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9), 
+there is an inserted verification step similar to \Cref{fig:const-type},
+which checks a pointer is well defined in the heap for every element in its defined range.
+In implementation, we actually optimize the verification away and substitute it
+with the bounds and null-pointer checks.
+Because tainted pointer is checked every time before it is used,
+so that we only need to check the top pointer is 
+well defined without recursively looking at the sub-terms in a nested pointer case.
+The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
+ and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
+are computed based on the meet 
+operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).
 
-For \code{strlen(p)} and conditionals \code{if(*p)}, the \lang
-semantics allows the upper bound of \code{p} to be extended.
-The compiler explicitly inserts statements to do so on \code{p}'s
-shadow bound variables. For example,
-Fig.~\ref{fig:compilationexample}~line~$6$ widens \code{p}'s upper
-bound if \code{strlen}'s result is larger than the existing
-bound. 
-Lines 7--12 of the generated code in
-Fig.~\ref{fig:compilationexample1}
-show how bounds are 
-widened when compiling expression \code{if(*p)}. If we find that the
-current \code{p}'s relative upper bound is equal to $0$ (line 10),
-and \code{p}'s content is not null (line 8), we then increase the
-upper bound by $1$ (line 11).
+\myparagraph{Checked and Unchecked Blocks}
+%
+In the implementation,
+$\euncheckedtext$ and $\echeckedtext$ blocks 
+are compiled as context switching functions provided by the sandbox mechanism.
+$\eunchecked{\overline{x}}{e}$ is compiled to 
+$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
+to execute expression $e$ with the arguments $\overline{x}$.
+$\echecked{\overline{x}}{e}$ is compiled to 
+$\texttt{callback}(\overline{x},e)$, where we perform 
+a \texttt{callback} to a checked block code $e$ inside a sandbox.
+In \systemname, We adopt an aggressive execution scheme that
+directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
+In the formalism, we rely on the type system to 
+guarantee the context switching without creating the extra function calls for simplicity.
 
-Fig.~\ref{fig:compilationexample1} also shows a dependent function call.
+%Fig.~\ref{fig:compilationexample} shows how an invocation of
+%\code{strlen} on a null-terminated string is compiled into C
+%code. Each dereference of a checked pointer requires a null check
+%(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
+%compiler makes explicit: Line~$3$ of the generated code has the null
+%check on pointer \code{p} due to the \code{strlen},
+%  and a similar check happens
+%  at line~$8$ due to the pointer arithmetic on \code{p}.
+%Dereferences also require bounds checks: line~$2$ checks \code{p} is
+%in bounds before computing \code{strlen(p)}, while line~$10$ does
+%likewise before computing \code{*(p+1)}.
+
+\myparagraph{Function Pointers and Calls}
+%
+Function pointers are dealt with similarly to normal pointers,
+but we insert checks to check if the pointer address is not null in 
+the function store and if the type is correctly represented, 
+for both $\cmode$ and $\tmode$ mode pointers 
+\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
+Fpr example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
+we place a check \code{verify_fun(cmp, not_null && type_match)},
+and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
+The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
+is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
+
+For compiling dependent function calls,
+\Cref{fig:compilationexample1} also shows an example.
 Notice that the bounds for the array pointer \code{p} are not passed as
 arguments. Instead, they are initialized according to \code{p}'s
-type---see line~3 of the original \lang program at the top of the figure.
-Line~$2$ of the generated code
-sets the lower bound  to \code{0} and line~$3$ sets the
+type---see line~4 of the original \lang program at the top of the figure.
+Line~$3$ of the generated code
+sets the lower bound  to \code{0} and the
 upper bound to \code{n}.
-
-\subsection{Comparison with Checked C Specification}
-\label{sec:disc}
-
-\begin{figure}[t]
-{\small
-{\captionsetup[lstlisting]{margin = 8 mm}
-  \begin{lstlisting}[xleftmargin=8 mm]
-nt_array_ptr<char> safe_strcat_c 
-   (nt_array_ptr<char> dst : count(n),
-    nt_array_ptr<char> src : count(0), int n) {
-  nt_array_ptr<char> tmp : count(n) = dst;
-  int x = strlen(tmp);
-  /* tmp now has x as its upper bound */
-  /* dst still has n as its upper bound */
-  int y = strlen(src);
-
-  if (x+y < n) {
-    for (int i = 0; i < y; ++i)
-      *(dst+x+i) = *(src+i);
-    *(dst+x+y) = '\0';
-    return dst;
-  }
-  return null;
-}
-\end{lstlisting}
-}
-}
-\caption{Safe \code{strcat} in Checked C that avoids a run-time error
-  exhibited by \code{safe_strcat} (Fig.~\ref{fig:strcat-ex}) when
-  compiled with the current Checked C compiler}
-\label{fig:strcatc-ex}
-\end{figure}
-
-The use of shadow variables for bounds widening is a key novelty of our
-compilation approach, and adds more precision to bounds checking at runtime
-compared to the official specification and current implementation of
-\checkedc~\cite[5.1.2, pg 85]{checkedc}.  For example, the
-\code{safe_strcat} example of Fig.~\ref{fig:strcat-ex} compiles with the
-current Clang \checkedc compiler but will fail with a runtime
-error. The statement \code{int x = strlen(dst)} at line 4 changes the
-statically determined upper bound of \code{dst} to \code{x}, which can be smaller than
-\code{n}, the full capacity of \code{dst}. The attempt to recover
-the full capacity of \code{dst} through a dynamic cast
-at line 7 will always fail if the capacity \code{n} is checked
-against the statically determined new upper bound \code{x}. % This problem can be worked around by
-% effectively inlining the definition of \code{strlen} as 
-% in \code{safe_strcat_c} in Fig.~\ref{fig:strcatc-ex} (lines
-% 6-7).
-This problem can be worked around by invoking \code{strlen} on a temporary
-variable \code{tmp} instead of \code{dst}
-as in \code{safe_strcat_c} in Fig.~\ref{fig:strcatc-ex} (lines 4-5).
-Likewise, if we were to add line \code{putchar(*(p+1));} 
-after line 6 in the original code at the top of
-Fig.~\ref{fig:compilationexample}, the code will always fail: the Clang \checkedc
-compiler (with the transliterated C code as its input) would 
-check \code{p} against its \emph{original} bounds \code{(0,0)} since the
-updated upper bound \code{x} is now out of
-the scope. Shadow variables address these problems because
-they retain widened bounds beyond the scope of variables that store
-them (i.e., \code{x} in both examples).
-
-To make it match the specification, our compilation definition could
-easily eschew
-shadow variables and rely only on the type-based
-bounds expressions available in $\Gamma$ for checking. However, doing so would 
-force us to weaken the simulation theorem, reduce expressiveness,
-and/or force the semantics to be more awkward. We plan to work with
-the \checkedc team to implement our approach in a future revision.
-
-% \textcolor{purple}{Despite its name, \elang
-%   does not model C. As mentioned at the beginning of this section,
-%   \elang is an untyped version of \lang with the annotations on
-%   literals removed. The lack of types make it impossible to capture
-%   the undefinedness of certain operations such as pointer
-%   arithmetic. This restriction is not an issue for establishing
-%   the metatheory covered in the next section, but it does indicate
-%   that we have to be careful if we want to embed \elang programs into C
-%   programs.  As discussed in \cite[2.10, pg 24]{checkedc}, the C
-%   standard can be overly
-%   conservative when it comes to pointer arithmetic. For example,
-%   it should be perfectly safe to increment a pointer out of its allocated
-%   region, as long as the runtime guarantees that
-%   the pointer is never dereferenced. However, a standard C compiler
-%   would treat this type of incrementing operation as undefined, regardless of
-%   whether the resulting pointer is dereferenced or not. Embedding
-%   \elang naively into C would give us this type of undefined programs.
-%   This can be mitigated through casting or rearranging the code during
-%   the embedding process.
-% }
-% \mwh{Don't think all this was necessary. Added clarifying text at the
-%   start at exactly what CoreC is. No need to go on at length about
-%   what it is not.}
 
@@
 In addition, when $\Gamma;\emptyset\vdash
-\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg (\dot \varphi, \dot \heap,
+\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
 \dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
-and $\Gamma;\Theta;\emptyset \vdash e \gg \dot e : \tau$ for some $\tau$ respectively.
+and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively.
 $\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
@@
 that is similar to the one by~\citet{merigoux2021catala}.
+
 An ordinary simulation property would
@@
 which only requires that the two configurations will eventually reduce
-to the same state. We formulate our simulation theorem differently
+to the same state.
+
+\ignore{
+ We formulate our simulation theorem differently
 because the standard simulation theorem imposes a very strong
@@
 naturally without having to worry about syntactic constraints.
+}
 
@@
 For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg(\dot \varphi_0,\dot \heap_0, \dot e_0)$,
+if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
 and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-\rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, then the following facts hold:
+\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
 

### 2022-07-17 sandp23/background.tex

@@
   nt_array_tptr<char> y : count (0)) {
+ checked (x,y) {
   int len_x = strlen(x);
@@
  }
+}
 ...
@@
   nt_array_tptr<char>)> cmp, int n) {
+ checked (s,cmp,n) {
   int i;
@@
   return stringsort(p,cfun);
+ }
 }
@@
 Once the closure is called, the system is turned to a checked region,
-so that the checked function call at \Cref{fig:checkedc-example-2} line 13 is safe, 
+so that the checked function call at \Cref{fig:checkedc-example-2} line 14 is safe, 
 even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.

### 2022-07-19 sandp23/introduction.tex
Updates from Overleaf

@@
 to prevent them~\cite{song2019sanitizing}.
-Vulnerabilities due to memory corruption, especially spatial memory corruptions, are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe} despite a large body of work that tries
-to prevent them~\cite{song2019sanitizing}.
-One way to handle this is to convert these programs into safe languages such as Java.
-However, it requires considerable development effort.
-Furthermore, the performance overhead prohibits them from being used in performance-critical settings such as Operating Systems.
-
-With the increased support for high-performant safe system languages such as Rust and Go, we can have tools that convert C to a memory-safe language, e.g., Rust, a promising memory-safe systems
-language~\cite{rustlang}.
-Unfortunately, Rust is too different from C to constitute a practical target~\cite{zeng2019identifying}, at least for now: the (best of breed) \texttt{c2rust} tool~\cite{c2rust,c2rusttalk} transliterates C to \emph{unsafe, non-idiomatic} Rust, and recent
-work~\cite{emre2021c2rust} is able to take only small steps to close the safety gap and still requires considerable work from developers.
-
+Converting these existing programs into safe languages, such as Java, requires considerable effort and is not (yet) feasible for performance-critical environments such as Embedded Operating Systems.
+Although high-performant safe system languages such as Rust~\cite{rustlang} and Go exist, these a too different to constitute a practical target~\cite{zeng2019identifying}, and rewriting in these languages requires considerable effort.
+Existing automated conversion tools~\cite{c2rust,c2rusttalk, emre2021c2rust} are able to take only small steps (i.e., to unsafe, non-idiomatic Rust) and still~\emph{require considerable work from developers}.
 There are other retrofitting techniques such as CCured~\cite{necula2005ccured},
@@
 automatically~\emph{without any developer effort}, by analyzing a C program and compiling it to include
-run-time safety checks. Unfortunately, the resulting run-time overhead is too high
-for deployment (between 60\%-200\%), especially in low-powered IoT devices where the usage of C-based system software (e.g., FreeRTOS) is still rampant.
-~\fig{fig:memprotections} shows the summary of these existing approaches.
+run-time safety checks. Unfortunately, the resulting~\emph{run-time overhead is too high
+for deployment (between 60\%-200\%)}, especially in low-powered IoT devices where the usage of C-based system software (e.g., FreeRTOS) is still rampant.
 
-%\input{figures/existingmem}
-
-\emph{Checked C}~\cite{tarditi2018checked} as a
-practical target to which to convert legacy/active C code. 
-Checked C
-extends C with \emph{checked pointer
+\noindent
+\textbf{Checked C.} Tarditi et al. developed~\emph{Checked C}, a safe dialect of C that prevents spatial memory issues (temporal safety is underway~\cite{checkedc:temporal}) with no overhead.
+Checked C extends C with \emph{checked pointer
   types} which are restricted by the compiler to spatially safe uses (temporal safety is underway~\cite{checkedc:temporal}).
@@
 associated \emph{bounds annotation}; e.g., a declaration \arrT{int}
-\code{p} : \code{count(n)} says that \code{p} is a pointer to an \code{int} array
-whose size is \code{n}.
-%
-Checked C's Clang/LLVM-based compiler represents checked
-pointers as system-level memory words, i.e., without
+\code{p} : \code{count(n)} says that \code{p} is a pointer to an \code{int} array whose size is \code{n}.
+The compiler uses these bounds annotations to add dynamic checks prior to checked pointer accesses, to prevent spatial safety violations.
+These run-time checks can often be proved redundant and removed by LLVM, yielding good performance.
+Specifically, Tarditi et al.~\cite{tarditi2018checked} reported average
+run-time overheads of 8.6\% on a small benchmark suite, and
+Duan et al.~\cite{duanrefactoring} found essentially~\emph{no overhead} when running
+Checked C-converted portions of the FreeBSD kernel.
+Furthermore, Checked C is backward compatible as the compiler  represents checked pointers as system-level memory words, i.e., without
 ``fattening'' metadata, ensuring backward compatibility. 
-The compiler uses these bounds annotations to add dynamic checks prior to
-checked pointer accesses, to prevent spatial safety violations.
-These run-time checks can often be proved redundant and removed by LLVM, yielding good performance.
-%, and leverages LLVM to optimize away run-time
-%checks, yielding good performance.
-%
-Tarditi et al.~\cite{tarditi2018checked} reported average
-run-time overheads of 8.6\% on a small benchmark suite, and
-\cite{duanrefactoring} found essentially no overhead when running
-Checked C-converted portions of the FreeBSD kernel.
+The backward compatibility also enables incremental or partial conversion where Checked annotations can be provided to a few pointers.
+Such partially annotated programs still enjoy the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
+~\fig{fig:memprotections} shows the summary of these existing approaches along with Checked C.
+We argue that~\emph{converting to Checked C provides a reasonable and practical approach to adding safety to existing system software written in C}.
 
+\input{figures/existingmem}
+
+\noindent
+\textbf{Converting C to Checked C.} The safety guarantees of Checked C come with certain restrictions.
+For instance, as shown below, Checked C programs cannot use address-taken variables in a bounds expression as the bounds relations may not hold because of possible modifications through pointers.
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize}{c}
+...
+array_ptr<int> p : count (n) = NULL;
+|\textcolor{red}{\faTimes}|..,&n,.
+\end{minted}
+Consequently, converting existing C programs to Checked C might require refactoring, e.g., modifying the above program to not use~\inlinecode{&n} expression, which might require considerable effort~\cite{duanrefactoring} depending on the program's complexity.
+Recently, Machiry et al. developed~\threec~\cite{machiry2022c} that tries to automatically convert a program to Checked C by adding appropriate pointer annotations.
+However, as described in \threec, complete automated conversion is infeasible and requires the developer to convert some code regions manually.
+Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
+But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+Although the blameless proof exists~\cite{ruef2019achieving}, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
+Consider the following example:
+\begin{minted}[mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+// Checked code
+int func(array_ptr<char> p : count(5)) {
+|\textcolor{red}{\faChainBroken}|..p[4]..
+}
+
+// unchecked code
+...
+str = "he";
+...
+|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5); 
+...
+char ptr[16];
+...
+len <- derived from user input
+...
+|\textcolor{red}{\faBug}| memcpy(ptr, buff, len);
+\end{minted}
+Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function with a buffer of 2 elements.
+This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
+Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (line ...) can take down the complete program despite having checked regions.
+We need an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions.
 However, converting existing C programs to Checked C requires considerable effort~\cite{duanrefactoring}. 

### 2022-07-20 sandp23/introduction.tex
Updates from Overleaf
Minor intro?

@@
 For instance, as shown below, Checked C programs cannot use address-taken variables in a bounds expression as the bounds relations may not hold because of possible modifications through pointers.
-\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize}{c}
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
 ...
@@
 Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
-But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+
+\noindent
+\textbf{No safety against unchecked code.} But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
 Although the blameless proof exists~\cite{ruef2019achieving}, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
 Consider the following example:
-\begin{minted}[mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
 // Checked code
@@
 }
-
 // unchecked code
@@
 ...
-|\textcolor{red}{\faBug}| memcpy(ptr, buff, len);
+|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
 \end{minted}
@@
 This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (line ...) can take down the complete program despite having checked regions.
-We need an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions.
+Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
+We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+
 However, converting existing C programs to Checked C requires considerable effort~\cite{duanrefactoring}. 

### 2022-07-20 sandp23/introduction.tex
minor

@@
 
-However, converting existing C programs to Checked C requires considerable effort~\cite{duanrefactoring}. 
-Recently, Machiry et al. developed 3C~\cite{machiry2022c} that tries to automatically convert a program to Checked C by adding appropriate pointer annotations.
-However, as described in 3C, complete automated conversion is infeasible and consequently requires the developer to convert some code regions manually.
-Thanks to the backward compatibility of Checked C, the partially annotated program still enjoys the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
-But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
-%
-Although the blameless proof exists~\cite{ruef2019achieving}, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
-For instance, a checked function expects a pointer to a buffer of five elements, but an unchecked region can call this function with a pointer to a buffer of three elements.
-This results in a spatial safety violation in the Checked region, but of course, the blame or the root cause is in the unchecked region.
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions can take down the complete program despite having checked regions.
-We need an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions.
-Such an isolation mechanism should have low/no overhead and be backward compatible (i.e., pointers should have no additional metadata), preserving the advantages of Checked C.
+This mechanism is called program partitioning~\cite{rul2009towards}, and there has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
+Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
+E.g., Given sensitive data in a program, the goal is to partition functions into two parts or partitions based on whether a function can access the sensitive data.
+The performance overhead of these approaches is dominated by marshaling costs and depends on the usage of sensitive data.
+The overhead of state-of-the-art approaches~\cite{lind2017glamdring, liu2017ptrsplit} is prohibitive and varies from 37\%-163\%.
+But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
+We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
 

### 2022-07-20 sandp23/introduction.tex
Adding

@@
 
-In this work, we propose~\systemname, an isolation mechanism that enables unchecked regions or, in general, any arbitrary set of functions to be executed in an isolated address space (\ucregion).
-So that checked regions can still enjoy true spatial memory safety even in the presence of unchecked or unsafe regions.
-\systemname extends Checked c using~\textbf{tainted} (\taintt) types, representing the code and data that belong to the~\ucregion.
+In this work, we propose a type-directed code-centric program partitioning approach.
+Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
+The tainted types long with Checked C allow annotating pointer along two dimensions, i.e., (i) taintedness: a pointer can be either tainted or not (untainted), and (ii) checkedness: it can be either checked or not.
+The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.
+The developer starts by marking desired (unchecked or risky) functions and pointers used in these functions as tainted.
+Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
+\begin{itemize}
+\item~\ucregion (low privilege tainted region): This partition contains tainted types (i.e., functions and pointers) and can only access tainted pointers.
+\item~\cregion (high privilege untainted region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
+The functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
+\end{itemize}
+Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (e.g., WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
 
-\noindent
-\textbf{Automated Partitioning.} A pointer (checked or unchecked) can be marked as tainted; similarly, a function can be marked as tainted.
-Along with enforcing spatial safety constraints of checked tainted pointers, the~\systemname compiler will bundle all the tainted types (functions and pointers) so that they remain in an isolated sandbox (\ucregion), and non-tainted types will remain in the process address space (\cregion).
+The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
+As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
+We avoid marshaling by allocating all tainted pointers (i.e., tainted buffers) in~\ucregion and thus can be accessed in both partitions.
+Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
+Our compiler avoids these attacks by ensuring, either statically or through dynamic checks, that tainted pointers can only point to~\ucregion address space.
 
-\noindent
-\textbf{Memory Access Restrictions.} The functions in the~\ucregion can only access tainted pointers, i.e., memory within the corresponding sandbox region.
-However, the functions in~\cregion have no-such restrictions and can access both c- and uc-region's memory.
-Similarly, the functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
-
-\noindent
-\textbf{Marshaling free communication.} As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
-We avoid marshaling as all tainted buffers (i.e., regions pointed by tainted pointers) are allocated in~\ucregion and thus can be accessed in both c- and uc-region.
-
-\noindent
-\textbf{Additional checks.} Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
-However, our compiler avoids these attacks by ensuring, either statically or through dynamic checks, that tainted pointers in~\cregion can only point to~\ucregion address space, and our type system enforces that tainted pointers cannot be assigned to non-tainted pointers.
-
-\iffalse
-Recently, Narayan et al. proposed RLBox, which enables sandboxing libraries by making data- and control-flow at the sandbox interface explicit, using types.
-Specifically, RLBox defines a tainted type (\code{tainted<...>}), which is a C++ template class and can be used to have any type tainted (e.g.,~\code{tainted<int>},~\code{tainted<float*>}, etc).
-The tainted types represent values or pointers to buffers that can be accessed from a sandbox.
-The program or unsandboxed code can use only tainted type variables to communicate (i.e., parameters and return values) with sandboxed functions.
-Although RLBox provides a practical technique for sandboxing libraries, it has the following drawbacks:
-\begin{itemize}
-\item\textbf{Marshalling Cost.} Any non-scalar value (e.g., a pointer to a struct) that need to be passed to a sandboxed function should be first copied into a tainted buffer (\emph{marshalling}).
+%In this work, we propose~\systemname, an isolation mechanism that enables unchecked regions or, in general, any arbitrary set of functions to be executed in an isolated address space (\ucregion).
+%So that checked regions can still enjoy true spatial memory safety even in the presence of unchecked or unsafe regions.
+%\systemname extends Checked c using~\textbf{tainted} (\taintt) types, representing the code and data that belong to the~\ucregion.
+%
+%\noindent
+%\textbf{Automated Partitioning.} A pointer (checked or unchecked) can be marked as tainted; similarly, a function can be marked as tainted.
+%Along with enforcing spatial safety constraints of checked tainted pointers, the~\systemname compiler will bundle all the tainted types (functions and pointers) so that they remain in an isolated sandbox (\ucregion), and non-tainted types will remain in the process address space (\cregion).
+%
+%\noindent
+%\textbf{Memory Access Restrictions.} The functions in the~\ucregion can only access tainted pointers, i.e., memory within the corresponding sandbox region.
+%However, the functions in~\cregion have no-such restrictions and can access both c- and uc-region's memory.
+%Similarly, the functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
+%
+%\noindent
+%\textbf{Marshaling free communication.} As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
+%We avoid marshaling as all tainted buffers (i.e., regions pointed by tainted pointers) are allocated in~\ucregion and thus can be accessed in both c- and uc-region.
+%
+%\noindent
+%\textbf{Additional checks.} Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
+%However, our compiler avoids these attacks by ensuring, either statically or through dynamic checks, that tainted pointers in~\cregion can only point to~\ucregion address space, and our type system enforces that tainted pointers cannot be assigned to non-tainted pointers.
+%
+%%\iffalse
+%Recently, Narayan et al. proposed RLBox, which enables sandboxing libraries by making data- and control-flow at the sandbox interface explicit, using types.
+%Specifically, RLBox defines a tainted type (\code{tainted<...>}), which is a C++ template class and can be used to have any type tainted (e.g.,~\code{tainted<int>},~\code{tainted<float*>}, etc).
+%The tainted types represent values or pointers to buffers that can be accessed from a sandbox.
+%The program or unsandboxed code can use only tainted type variables to communicate (i.e., parameters and return values) with sandboxed functions.
+%Although RLBox provides a practical technique for sandboxing libraries, it has the following drawbacks:
+%\begin{itemize}
+%\item\textbf{Marshalling Cost.} Any non-scalar value (e.g., a pointer to a struct) that need to be passed to a sandboxed function should be first copied into a tainted buffer (\emph{marshalling}).
 Then the corresponding tainted value (i.e., tainted pointer) should be passed to the sandboxed function.
-This can lead to considerable overhead for large objects.
-\item\textbf{Developer Written Sanitization.} The unsandboxed code can access tainted type variables only through certain special (i.e., sanitization) routines.
-This forces developers to sanitize every value that comes from sandboxed library functions.
-However, the correctness of the sanitization is entrusted to the developer.
-Consequently, any mistake in sanitization methods can still lead to vulnerabilities.
-\item\textbf{Library Level Sandboxing.} The unit of sandboxing in RLBox is the whole library. However, we want to sandbox a set of functions in a program.
-\end{itemize}
-\fi
+%This can lead to considerable overhead for large objects.
+%\item\textbf{Developer Written Sanitization.} The unsandboxed code can access tainted type variables only through certain special (i.e., sanitization) routines.
+%This forces developers to sanitize every value that comes from sandboxed library functions.
+%However, the correctness of the sanitization is entrusted to the developer.
+%Consequently, any mistake in sanitization methods can still lead to vulnerabilities.
+%\item\textbf{Library Level Sandboxing.} The unit of sandboxing in RLBox is the whole library. However, we want to sandbox a set of functions in a program.
+%\end{itemize}
+%\fi
 

### 2022-08-01 sandp23/formal.tex
update.

@@
 
-This section describes the formal model of \systemname, called
+This section describes the formal model of \systemname, named
 \lang, making precise its syntax, semantics, and type system. It also
-develops \lang's meta-theory, including type soundness, the non-exposure, and the non-crashing
-theorem.
+develops \lang's meta-theory, including the type soundness, non-exposure, and non-crashing
+theorems.
 
@@
 
-There are two notions of type in \lang.  Types $\tau$ classify
+There are two type notions in \lang.  Types $\tau$ classify
 word-sized values including integers and pointers, while types
@@
 %
-Pointer types ($\tptr{\omega}{m}$) include a mode annotation ($m$)
-which is either checked (\cmode) or unchecked (\umode) and a type
-($\omega$) denoting valid values that can be pointed to. Array types include both the type of
+Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
+($\xi$, the difference between context and pointer modes is introduced shortly below)
+that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
+and a type ($\omega$) denoting valid values that can be pointed to.
+Array types include both the type of
 elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
@@
 $\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide the $\cdot$ when writing the type).
-\systemname function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
-reflect its dependent function declrations,
+otherwise (we elide $\cdot$ when writing types).
+\lang function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
+reflect its dependent function declarations,
 where $\overline{x}$ represents
-a list of \tint type variables in a dependent function header,
-which bind all bounds appearing in $\overline{\tau}$ and $\tau$.
+a list of \tint type variables in a dependent function header
+that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
 We have a well-formed requirement for a function type,
@@
 Integer literals $n$ are annotated with a type $\tau$ which can be either
-$\tint$, or $\tptr{\omega}{m}$ in the case $n$ is being used as 
+$\tint$, or $\tptr{\omega}{\xi}$ in the case $n$ is being used as 
 a heap address (this is useful for the semantics);
-$\evalue{0}{\tptr{\omega}{m}}$ (for any $m$ and $\omega$) represents the $\enull$ pointer, as usual. 
+$\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer, as usual. 
 The $\texttt{strlen}$ expression operates on variables $x$
@@
 for maintaining non-exposure.
-Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi'} ...)}{\xi}$, if the outside mode $\xi$ is 
-$\tmode$ or $\umode$, the inside mode $\xi'$ cannot be $\cmode$.
+Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi'} ...)}{\xi}$, $\xi'\le \xi$.
 It is worth noting that pointer modes are a three point partial order ($\le$),
@@
 $\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
-such that a pointer mode $\tmode$ is projected as $\umode$.
+such that $\tmode$ is projected as $\umode$.
 Second, $\emalloc{\xi}{\omega}$ includes a mode flag
 $\xi$ for allocating different pointers in different heap mode regions.
-For simplicity, we disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$),
+We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
 Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
-Fourth, we update $\eunchecked{\overline{x}}{e}$ 
-to specify $\overline{x}$ restricting all free variables appearing in $e$.
-Additionally, we create $\echeckedtext$ blocks indicating the context mode change from $\umode$ to $\cmode$.
-Both the free variables in the $\euncheckedtext$ and $\echeckedtext$ blocks
-and the return types cannot be $\cmode$ mode for maintaining non-exposure,
-i.e., no checked pointer can be communicated between $\euncheckedtext$ and $\echeckedtext$ blocks.
+Fourth, $\echeckedtext$ blocks are added to the system, 
+which permits the nested switching between $\echeckedtext$ (represented by context mode $\cmode$)
+and $\euncheckedtext$ (represented by context mode $\umode$) code regions.
+One example usage of the nested switching is the checked function callbacks inside 
+an unsafe region in \Cref{fig:checkedc-example-2} and \ref{fig:checkedc-example-3}.
+To guarantee the non-exposure safety,
+ we extend the $\echeckedtext$ and $\euncheckedtext$ block syntax to be 
+$\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$:
+$\overline{x}$ restricts all free variables appearing in $e$, and they cannot be checked pointers.
 
@@
 in \checkedc, so we omit them.
-Additional syntactic descrition is listed in the \checkedc formalism paper \cite{li22checkedc}.
 
@@
     \begin{array}{lll}
-\mu & ::= & \evalue{n}{\tau}\\
-e & ::= & \ldots \mid \ret{x}{\mu}{e}\\
+e & ::= & \ldots \mid \ret{x}{\evalue{n}{\tau}}{e}\\
 r & ::= & e \mid \enull \mid \ebounds\\
@@
 maps addresses (integer literals) to values $\evalue{n}{\tau}$.
-If a pointer has mode $\cmode$, it lives in the $\cmode$ region; otherwise, it lives in the $\umode$ region.
+A $\cmode$pointer is mapped to a heap location in the $\cmode$ region, 
+while a $\tmode$ and $\umode$ pointer represents a $\umode$ region location.
 We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ region,
-and $\heap(m)[n\mapsto \mu]$ to update location $n$ with the value $\mu$.
+and $\heapup{m}{n}{\evalue{n'}{\tau}}$ 
+to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ region.
 It is worth noting that \systemname is not a fat-pointer system;
@@
 %
-The context mode $m$ indicates whether the
+The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the
 stepped redex within $e$ was in a checked ($\cmode$) or
@@
 The rules for the main operational semantics
-judgment---\emph{evaluation}---are given at the bottom of
+judgment---\emph{evaluation}---are given at the middle of
 Fig.~\ref{fig:c-context}.
@@
 for possible crashing behaviors in unchecked region.
-A program in $\umode$ mode can non-deterministically crash
-and the \systemname sandbox mechanism recorvers
+A $\umode$ mode operation can non-deterministically crash
+and the \systemname sandbox mechanism recovers
 the program to a safe point ($\evalue{0}{\tau}$)
 and continues with the existing program state.
+Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
+$\ret{x}{\mu}{e}$ syntax shortly.)
 There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
@@
 % \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
-The $\mode$ function
-determines the context mode when evaluating $e'$ based on the context $E$.
-For any program execution, the top context mode always starts with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
-The result context mode is depeneded on where $\Box$ locates.
-If it occurs within $(\eunchecked{\overline{x}}{E'})$ inside $E$ without
-a surrounding $\echeckedtext$ block, then the mode is
-$\umode$; if the exact opposite happens (within $(\echecked{\overline{x}}{E'})$ without surrounding $\euncheckedtext$ blocks),
-the mode is $\cmode$; the other cases ($\mode'(\alpha({E}),m)=\mode'(E,m)$) are recursively defined by the above base cases ($\alpha$ represents any syntactic categories in \Cref{fig:c-context} other than $\Box$, $\euncheckedtext$, and $\echeckedtext$).
-Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
-$\ret{x}{\mu}{e}$ syntax shortly.)
+The $\mode$ function  at the bottom of Fig.~\ref{fig:c-context}
+describes the context mode determination in each evaluation step based on the context $E$.
+For any program execution, the function starts the mode computation with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
+The result context mode depends on where $\Box$ locates.
+If it occurs within $E$ in $(\eunchecked{\overline{x}}{E})$
+that has no surrounding $\echeckedtext$ block,
+the mode is $\umode$; otherwise, the mode is $\cmode$.
+$\mode'(\alpha({E}),m)=\mode'(E,m)$ represent other construct cases 
+that are not $\echeckedtext$ and $\euncheckedtext$; in such case, 
+the function recursively traverses the sub-context to find the context mode.
+
 % For a term $e$, $C$ is a context if any
@@
     \inferrule[S-AssignArrC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\cmode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\cmode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}
 
@@
                \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_1}{\tau} }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,(\heap,\umode)[n \mapsto \evalue{n_1}{\tau_a}],\evalue{n_1}{\tau})}
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\umode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}
 
@@
 
-    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\mu}{\mu'}) \longrightarrow (\varphi,\heap,\mu')}
+    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{\evalue{n'}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n'}{\tau'})}
 
@@
 
-    \inferrule[S-RetCon]{\varphi(x)=\mu\\ (\varphi[x\mapsto \mu'],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\mu'}{e}) \longrightarrow (\varphi'[x\mapsto \mu],\heap',\ret{x}{\varphi'(x)}{e'})}
-
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\mu}) \longrightarrow (\varphi,\heap,\mu)}
+    \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}
+
+    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
 
@@
 
-    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\mu}) \longrightarrow (\varphi,\heap,\mu)}
+    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
 
@@
 
-Fig.~\ref{fig:semantics} shows selected rules for the computation
-relation; we explain them with the help of the example in \Cref{fig:checkedc-example-1} and \Cref{fig:checkedc-example-2}.
+Fig.~\ref{fig:semantics} shows selected rules for the computation relation.
 
@@
 \textsc{S-DefNull}, and \textsc{S-Cast}.
-The first five are deference and assignment operations---illustrate how the semantics checks bounds.
+The first five define deference and assignment operations---illustrate how the semantics checks bounds.
 Rule \textsc{S-DefNull} transitions attempted null-pointer
-dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a \cmode-mode
+dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a $\cmode$-mode
 non-null (single) pointer.
@@
 We explain this shortly in \Cref{sec:type-system}.
-For now, the verification step, e.g. $\umode;\emptyset;\heap ; \emptyset \vdash\evalue{n_a}{\tau}$ in \textsc{S-DefC},
-means we verify that the value $n_a$ has type $\tau$ in the $\umode$ region heap.
+For now, the verification step, e.g. $\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$ in \textsc{S-DefC},
+means we verify that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.
 
@@
 %
-Semantically, $\euncheckedtext$ and $\echeckedtext$ blocks 
-(\textsc{S-Unchecked} and \textsc{S-Checked}) act as a classical C block structures.
-The interesting part is their type rules in \Cref{sec:type-system}.
+Semantically, $\euncheckedtext$ and $\echeckedtext$ blocks act as classical C blocks
+as rules \textsc{S-Unchecked} and \textsc{S-Checked} in \Cref{fig:semantics}.
 
@@
 The semantics handles variable scopes using the special $\erettext$
-form. \textsc{S-Let} evaluates to a configuration whose stack
-is $\varphi$ extended with a binding for $x$, and whose expression is
-$\ret{x}{\evalue{n}{\tau}}{e})$ which keeps $\varphi$ unchanged, but
-remember $x$ with the new value $\evalue{n}{\tau}$ in the scope $e$.
-Every time when Evaluation proceeds on $e$ (Rule \textsc{S-RetCon}),
-we install the current stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$,
-after one-step evaluation is completed, we restore the result $\varphi'(x)$ 
-in the result $\ret{x}{\varphi'(x)}{e'})$.
-This procedure continues until it becomes a literal
-$n\!:\!\tau$, in which case \textsc{S-RetEnd} remove the $\kw{ret}$ frame and evaluates to
-$n\!:\!\tau$.
+form. \textsc{S-Let} evaluates to a configuration whose expression is
+$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
+and remember $x$ and its the new value $\evalue{n}{\tau}$
+in $e$'s scope that is defined by the $\erettext$ operation.
+Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
+we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope;
+after one-step evaluation is completed, 
+we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
+and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
+This procedure continues until $e'$ becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
+the literal.
 
 Function calls are handled by \textsc{S-FunC} and \textsc{S-FunT},
-for \cmode and \tmode mode function pointers, respectively. 
-A call to function $f$ causes $f$'s
-definition to be retrieved from $\Xi$,
+for $\cmode$ and $\tmode$ mode function pointers, respectively. 
+A call to a function pointer $n$ retrieves
+ the function definition in $n$'s location in the global function store $\Xi$,
 % LEO: This is the first occurence of \Xi. It is explained,
 % but reads a bit awkward as it's unclear this is its "introduction"
-which maps function names to
+which maps function pointers to
 function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
@@
 \code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
-2-3 in Fig.~\ref{fig:strcat-ex}. 
+2-3 in \Cref{fig:strcat-ex}. 
 %All dependent bound variables appearing in function headers 
@@
 $(\tau[\overline{n} / \overline{x}])$ which is the function's return
-type but with any parameter variables $\overline{x}$ appearing in that
-type substituted with the call's actual arguments $\overline{n}$. To
-see why this is needed, suppose that \code{safe_strcat} in
+type but with any integer parameter variables $\overline{x}$ appearing in that
+type, as type bound variables,
+substituted with the call's actual arguments $\overline{n}$.
+To see why this is needed, suppose that \code{safe_strcat} in
 Fig.~\ref{fig:strcat-ex} is defined to return a
@@
 Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
+Rule \textsc{S-FunT} defines the tainted version of function call semantics.
+In such case, the verification process 
+$\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}$
+makes sure that the function in the global store is well-defined and has the right type.
 
@@
 variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-in mode $m$. Key rules for this judgment are given in
+in context mode $m$. Key rules for this judgment are given in
 Fig.~\ref{fig:type-system-1}.
@@
    \inferrule[T-ConstU]
-       { }
-       {\Gamma;\Theta\vdash_c \evalue{u}{\tau} : \tau}
+       { \neg \cmode(\tau)}
+       {\Gamma;\Theta\vdash_u \evalue{n}{\tau} : \tau}
 
@@
     \inferrule[T-Def]
-              {m' \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{m'}}
+              {\xi \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
               {\Gamma;\Theta \vdash_m \estar{e} : \tau}
@@
     \inferrule[T-AssignArr]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\\\
+              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{\xi}\\\\
                 \Gamma; \Theta \vdash_m e_2 : \tau' \\\\
                 \tau'\sqsubseteq_{\Theta} \tau\\
-                m' \leq m}
+                \xi \leq m}
               {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
@@
                {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq \tptr{\tau}{\xi}}
+                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
                {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
@@
     { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
-        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
-           \Gamma[x\mapsto \tau];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
+        \Gamma;\Theta \vdash_m e_1 : \tint \\\\
+           \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
              }
@@
 
-   \inferrule[T-Ret]
-    {m;\emptyset;\Gamma;\emptyset\vdash \mu : \tau' \\\\
-          \Gamma[x\mapsto \tau'];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
-    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tau'}}{e} : \tau}
+   \inferrule[T-RetInt]
+    { \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
+    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tint}}{e} : \tau}
 
     \inferrule[T-Checked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
               {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
@@
     \inferrule[T-Unchecked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\\
-                  \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
               {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
@@
 %
-Rules \textsc{T-DefArr} and \textsc{T-AssignArr} type-check array
+Rules \textsc{T-Def} and \textsc{T-AssignArr} check array
 dereference and assignment operations resp., returning the type of
-pointed-to objects; rules for pointers to single objects are
+pointed-to objects; rules for pointers for other object types are
 similar.
-The condition $m\le m'$ ensures that checked pointers and unchecked pointers 
-can only be deferenced in checked and unchecked regions, respectively;
-the type rule for
-The rules do not attempt to reason whether the access is in bounds;
-this check is deferred to the semantics.
+The condition $m\le m'$ ensures that checked and unchecked pointers 
+can only be deferenced in checked and unchecked regions, resp.;
+The type rules do not attempt to reason whether the access is in bounds;
+such check is deferred to the semantics.
 
@@
 %
-In \systemname, we define the type equality $\tau=_{\Theta}\tau'$
-is an equivalent relation over the bound equality ($=_{\Theta}$) of the bounds in (NT-)array types
-and alpha equivalence of two function types in \Cref{fig:checkc-subtype}.
-Two (NT-)array pointers $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
-$\bvar =_{\Theta} \bvar'$; two function types 
+In \lang, type equality $\tau=_{\Theta}\tau'$
+is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
+and the alpha equivalence of two function types in \Cref{fig:checkc-subtype}.
+Two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
 $\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
-are equivalent, if we can find a same length variable set $\overline{z}$ that is substituted for  
-$\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, respectively,
+are equivalent, if we can find a same length, as $\overline{x}$ and $\overline{y}$, variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
 and the substitution results are equal.
@@
 $\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} built on the type equality
+$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and it is built on the type equality
 ($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$); the many
@@
 pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
-Two function pointers are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
-if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are oppositely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
+Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
+There is another casting rule in \Cref{app:main} states that
+ users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
 
@@
 $\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
-such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-Ret}.
+such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
 \iftr
@@
 \begin{array}{l}
-\text{Bound Inequality and Equality:}\\
+\text{Bound Inequality and Equality:}\\[0.3em]
   \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
@@
     \end{array}
-  \\[0.5em]
-\text{Type Equility:}\\
+  \\[0.7em]
+\text{Type Equility:}\\[0.3em]
   \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
@@
     \end{array}
-  \\[0.5em]
-\text{Subtype:}\\
+  \\[0.7em]
+\text{Subtype:}\\[0.3em]
 
@@
     \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
-    &&\tptr{\omega}{\tmode} &\sqsubseteq_{\Theta}& \tptr{\omega'}{\umode}\\[0.2em]
 
@@
        {}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\umode}}
+       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
 
    \inferrule
-       {}
-       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
+       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
+       {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
 
@@
 \myparagraph{Constant Validity}
+Rules \textsc{T-ConstU} and \textsc{T-ConstC}
+describes type assumptions for constants appearing in a program.
+$\cmode(\tau)$ judges a constant pointer 
+in an unchecked region cannot be of a checked type,
+which represents an assumption that programmers 
+cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+In rule \textsc{T-ConstC}, we requires a static 
+verification procedure for validating a constant pointer, 
+which is similar to the dynamic verification process in \Cref{sec:semantics}.
+
 Given a constant $\evalue{n}{\tau}$, 
-the type judgment $\Theta;\heap;\sigma \vdash_m n : \tau$ checks in \Cref{fig:const-type}
+the verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
 if the constant is valid, 
-where $\heap$ is the initial heap that the constant resides on and
+where $\heap(m)$ is the initial heap that the constant resides on and
 $\sigma$ is a set of constant assumed to be checked.
-A global function store $\Xi$ is also required in checking the validity of a function pointer.
+A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
 A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
 and the address stores a function with the right type.
-
 The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
 where every element in the pointer range ($[0,\size(\omega))$) should be well typed.
+
 A checked pointer checks validity in type step as rule \textsc{T-ConstC},
@@
 Both $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$
-check all free variables are within $\overline{x}$,
-and the types for $\overline{x}$ have no checked pointers.
-A $\echeckedtext$ and $\euncheckedtext$ block represents 
-the transition stage from a checked to an checked region, or vice versa.
+check all free variables in $e$ are within $\overline{x}$;
+the types for $\overline{x}$ and the final return type $\tau$ have no checked pointers.
+A $\echeckedtext$ or $\euncheckedtext$ block represents 
+the context switching from a checked to an checked region, or vice versa.
 We need to make sure no checked pointers are information exposed to unsafe code regions.
 
-\myparagraph{Dependent Function Pointers and Let Bindings}
-%
-Rule \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
+\myparagraph{Let Bindings and Dependent Function Pointers}
+%
+Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
 type dependency. In particular, the result of evaluating a $\elettext$
@@
 the result is a checked pointer with a variable-defined bound); if so,
-we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). Note
-that we restrict the expression $e_1$ to syntactically match the
+we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
+Note that we restrict the expression $e_1$ to syntactically match the
 structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
 
-Rule \textsc{T-Ret} types a $\erettext$ expression, which does not
-appear in source programs but is introduced by the semantics when
+Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
+$\erettext$ does not appear in source programs but is introduced by the semantics when
 evaluating a let binding (rule \textsc{S-Let} in
@@
 
-Rule \textsc{T-Fun} is the standard dependent function call rule. It
-learns the function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$) of $e$ by a recursive type check,
-type-checks the actual arguments $\overline{e}$ which have
-types $\overline{\tau'}$, and then confirms that each of these types
-is a subtype of the declared type of $e$'s corresponding parameter ($\overline{\tau}$). 
-Because functions have dependent types bounded by $\overline{x}$, 
-we substitute each parameter $e_i$ for
-its corresponding parameter $x_i$ in both the parameter types and the
-return type. Consider the \code{safe_strcat} function in
+Rule \textsc{T-Fun} is the dependent function call rule. 
+Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
+from a type-check for $e$ and the types $\overline{\tau'}$ from the argument type checks for $\overline{e}$,
+we confirm that each of $\overline{\tau'}$ is
+a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$,
+which replaces possible integer bound variables $\overline{x}$ with bound expressions $\overline{e'}$.
+The final result type is the defined target type $\tau$ appearing in the function pointer type
+also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
+Consider the \code{safe_strcat} function in
 Fig.~\ref{fig:strcat-ex}; its parameter type for \code{dst} 
@@
 \begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
-A type environment $\Gamma$ is well-formed iff every variable mentioned as type bounds in $\Gamma$ are bounded by $\tnat$ typed variables in $\Gamma$.
+A type environment $\Gamma$ is well-formed iff every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
 \end{defi}
@@
 
-Moreover, as a program evaluates, its expression may contain literals
-$\evalue{n}{\tau}$ where $\tau$ is a pointer type, i.e., $n$ is an
-index in $\heap$ (perhaps because $n$ was chosen by
-$\mathtt{malloc}$). The normal type-checking judgment for $e$ is
-implicitly parameterized by $\heap$, and the rules for type-checking
-literals confirm that pointed-to heap cells are compatible with
-(subtypes of) the pointer's type annotation; in turn this check may
-precipitate checking the type consistency of the heap itself. We
-follow the same approach as \citet{ruef18checkedc-incr}, and show the
-rules in
-  Fig.~\ref{fig:const-type}; the judgment
-  $\Theta;\heap;\sigma \vdash_m n : \tau$ is used to confirm literal
-  well-typing, where $\sigma$ is a set of pointer literals already
-  checked in $\heap$ (to allow pointer cycles). See
-  \iftr
-  Appendix~\ref{sec:literal-pointer-typing}
-  \else
-  the supplemental report~\cite{checkedc-tech-report}
-  \fi
-   for further discussion.
-
-Progress states that a \systemname program can always make a move:
+%Moreover, as a program evaluates, its expression may contain literals
+%$\evalue{n}{\tau}$ where $\tau$ is a pointer type, i.e., $n$ is an
+%index in $\heap$ (perhaps because $n$ was chosen by
+%$\mathtt{malloc}$). The normal type-checking judgment for $e$ is
+%implicitly parameterized by $\heap$, and the rules for type-checking
+%literals confirm that pointed-to heap cells are compatible with
+%(subtypes of) the pointer's type annotation; in turn this check may
+%precipitate checking the type consistency of the heap itself. We
+%follow the same approach as \citet{ruef18checkedc-incr}, and show the
+%rules in
+%  Fig.~\ref{fig:const-type}; the judgment
+%  $\Theta;\heap;\sigma \vdash_m n : \tau$ is used to confirm literal
+%  well-typing, where $\sigma$ is a set of pointer literals already
+%  checked in $\heap$ (to allow pointer cycles). See
+%  \iftr
+%  Appendix~\ref{sec:literal-pointer-typing}
+%  \else
+%  the supplemental report~\cite{checkedc-tech-report}
+%  \fi
+%   for further discussion.
+
+Progress states that a \lang program can always make a move:
 
@@
 
-For any \systemname program $e$, heap $\heap$, stack
+For any \lang program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 type and consistency of the program being reduced.
+Unchecked Preservation states that any evaluation happens at unchecked region does not affect the checked heap.
 
 \begin{thm}[Checked Preservation]
-For any \systemname program $e$, heap $\heap$, stack
+For any \lang program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$, $\Theta'$ and $\tau'$ that are well formed, checked region consistent
-($\Gamma';\Theta'\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
-well typed ($\Gamma';\Theta' \vdash_{\cmode} e: \tau'$), where
-$\tau'\sqsubseteq \tau$.
+$\Gamma'$ and $\tau'$ that are well formed, checked region consistent
+($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
+well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
+$\tau'\sqsubseteq_{\Theta} \tau$.
 \end{thm}
@@
 \begin{thm}[Unchecked Preservation]
-For any \systemname program $e$, heap $\heap$, stack
+For any \lang program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 Using the above theorem, we first show the non-exposure theorem,
-where code in unchecked region cannot observe know a valid checked pointer address.
+where code in unchecked region cannot observe a valid checked pointer address.
 
 \begin{thm}[Non-Exposure]
-For any \systemname program $e$, heap $\heap$, stack
+For any \lang program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 \longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
-where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$, if $x$ is a checked pointer, it is a null one.
+where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
+thus, it is not a checked pointer.
 \end{thm}
 
-Using these two theorems we can prove our main result, {\em non-crashing},
-which states that a well-typed program can never be \emph{stuck} (expression
+We now state our main result, {\em non-crashing},
+which suggests that a well-typed program can never be \emph{stuck} (expression
 $e$ is a non-value that cannot take a step\footnote{Note that
@@
 
-\begin{thm}[Non-Crashing]\label{thm:blame} For any \systemname
+\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
   program $e$, heap $\heap$, stack

### 2022-08-01 sandp23/compilation.tex
update compilation.

@@
 {\small
+\hspace*{-0.5em}
 \begin{tabular}{|c|c|c|c|}
@@
 \hline
-& \text{Impl} / \text{Foml} & \text{Impl} / \text{Foml} & \text{Impl} / \text{Foml} \\
+& \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} \\
 \hline
@@
 In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
-i.e., \cmode pointer points to checked regions, while \tmode and \umode pointers point to unchecked regions.
+i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
 Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
 In the compiled LLVM code, pointer access operations have different syntax when the modes are different.
-\Cref{fig:flagtable} lists the different compiled syntax for a deference operation ($\estar{x}$) for the compiler implementation (\text{Impl}) and formalism (\text{Foml}).
+\Cref{fig:flagtable} lists the different compiled syntax for a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
 The columns represent different pointer modes and the rows represent context modes.
-For example, when we have a \tmode pointer in a \cmode-mode region, we use the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data.
-In formalism, we create a new deference data-structure in LLVM on top of the existing $\estar{x}$ operation: $\getstar{m}{x}$.
+For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation.
+In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
 if the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
 
-In the compilation, we use a function \code{mode} to get the pointer mode of a function learning from the pointer's type.
-Thus, the compiler can determine the correct compiled syntax for a pointer access operation with 
-the knowledge of the current context mode.
-We also build a context switching mechanism to change 
-the context mode when we meet $\echeckedtext$ and $\euncheckedtext$ blocks.
-We describe the mechanism shortly.
-
-This section shows how \systemname deals with pointer modes, mode switching and function pointer compilations, 
+This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
 with no loss of expressiveness;
@@
 we present a compilation algorithm that converts from
-\systemname to \elang, an untyped language without metadata
+\lang to \elang, an untyped language without metadata
 annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
@@
 next two steps of compilation are more interesting.
-We state them based on different \systemname operations.
+We state them based on different \lang operations.
 
@@
   *(mode(p) $\wedge$ m,p);
-  verify(q, not_null && q_lo <= 0 && 0 <= q_hi);
+  verify(q, not_null(m, q_lo, q_hi) 
+             && q_lo <= 0 && 0 <= q_hi);
   *(mode(q) $\wedge$ m,q)=1;
@@
 to track the lower and upper bounds potentially changed for pointer arithmetic and NT-array bound widening.
-We also associated with every \cmode-mode pointer variable according to its type;
+We also associated with every $\cmode$-mode pointer variable according to its type;
 we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
-In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9), 
+In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
 there is an inserted verification step similar to \Cref{fig:const-type},
-which checks a pointer is well defined in the heap for every element in its defined range.
+which checks a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
+Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
 In implementation, we actually optimize the verification away and substitute it
@@
 %
-In the implementation,
+In the \systemname implementation,
 $\euncheckedtext$ and $\echeckedtext$ blocks 
@@
 but we insert checks to check if the pointer address is not null in 
-the function store and if the type is correctly represented, 
+the function store instead of heap, and whether or not the type is correctly represented, 
 for both $\cmode$ and $\tmode$ mode pointers 
 \footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
-Fpr example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
-we place a check \code{verify_fun(cmp, not_null && type_match)},
+For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
+we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
 and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
@@
 For compiling dependent function calls,
-\Cref{fig:compilationexample1} also shows an example.
+\Cref{fig:compilationexample1} provides a hint.
 Notice that the bounds for the array pointer \code{p} are not passed as
@@
 
-The current version of the Redex model has been tested against $20000$
-expressions with depth less than $10$. Each expression can
+The current version of the Redex model has been tested against $23000$
+expressions with depth less than $11$. Each expression can
 reduce multiple steps, and we test simulation between every two

### 2022-08-02 sandp23/introduction.tex
finished introduction.

@@
 \section{Introduction}
-Vulnerabilities due to memory corruption, especially spatial memory corruptions, are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe} despite a large body of work that tries
-to prevent them~\cite{song2019sanitizing}.
+
+Vulnerabilities due to memory corruption, especially spatial memory corruptions, 
+are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798} 
+despite a large body of work that tries to prevent them~\cite{song2019sanitizing}.
+Several industrial and research efforts---including CCured~\cite{Necula2005},
+Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
+explored means to compile C programs 
+to automatically enforce spatial safety. These
+approaches all impose performance overheads deemed too high for
+deployment use. 
+Recently, \citet{Elliott2018,li22checkedc} introduced and formalized \checkedc, an
+open-source extension to C with new types and
+annotations whose use can ensure a program’s spatial safety.
+Importantly, \checkedc supports development that is 
+incremental and compositional. Code regions (e.g.,
+functions or whole files) designated as \emph{checked} enforce
+spatial safety in a manner preserved by composition with
+other checked regions. But not all regions must be checked: Checked
+C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
+may coexist in the same code, which permits a deliberate (and
+semi-automated) refactoring process. Parts of the FreeBSD kernel have
+been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
+overhead seems low enough for practical deployment.
+
+\ignore{
 Converting these existing programs into safe languages, such as Java, requires considerable effort and is not (yet) feasible for performance-critical environments such as Embedded Operating Systems.
 Although high-performant safe system languages such as Rust~\cite{rustlang} and Go exist, these a too different to constitute a practical target~\cite{zeng2019identifying}, and rewriting in these languages requires considerable effort.
+
 Existing automated conversion tools~\cite{c2rust,c2rusttalk, emre2021c2rust} are able to take only small steps (i.e., to unsafe, non-idiomatic Rust) and still~\emph{require considerable work from developers}.
@@
 for deployment (between 60\%-200\%)}, especially in low-powered IoT devices where the usage of C-based system software (e.g., FreeRTOS) is still rampant.
+
+\input{figures/existingmem}
 
@@
 The backward compatibility also enables incremental or partial conversion where Checked annotations can be provided to a few pointers.
-Such partially annotated programs still enjoy the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
+Such partially annotated programs still enjoy the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe core regions).
 ~\fig{fig:memprotections} shows the summary of these existing approaches along with Checked C.
 We argue that~\emph{converting to Checked C provides a reasonable and practical approach to adding safety to existing system software written in C}.
+}
 
-\input{figures/existingmem}
+% One possible solution is Checked C
+% \begin{itemize}
+% \item Think of it as migratory/gradual typing, but for C.
+% \item Distinct pointer types, but which are backward binary- and source-compatible.
+% \item Checked regions, containing only checked pointers and restricted
+%   idioms, aim to ensure spatial safety
+% \item Implemented as an extension to Clang/LLVM. Good performance
+%   (compared to ASAN etc.)
+% \end{itemize}
 
+The guarantee \checkedc provides focuses on checked core regions,
+where any stuck (i.e., ill-defined)
+state reached by a well-typed program amounts to a spatial safety
+violation; such a state can always be attributed to, i.e., 
+\emph{blamed on}, the execution of code that is not in a checked core
+region. But what are the guarantees of unchecked (unsafe) core regions?
+Programmers might not expect to convert every code to \checkedc,
+but leave some functions in unsafe regions,
+in which case the unsafe region executions should not affect other parts.
+In this paper, we cover this gap by introducing \systemname,
+combining \checkedc with program partitioning mechanism, 
+making three main contributions.
+
+\myparagraph{Executions Cannot Crash}
+Our first contribution is a clear program partition between checked and unchecked core regions,
+and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
+We restrict the checked and unchecked pointers usage only in checked and unchecked core regions, respectively;
+and develop tainted pointers for the communication between the two regions.
+Heap is partitioned into two parts: a \textit{checked memory region} containing all checked pointers
+ and an \textit{unchecked memory region} that is sandboxed.
+Tainted pointers live in the unchecked memory region and is verified every time when using.
+
+\myparagraph{Formaling the Type System, Semantics and Compiler}
+%
+Our second contribution is a core formalism named \lang, which extends
+\citet{li22checkedc} with the non-crashing guarantee and other new features below.
+We prove formally the \textit{non-crashing theorem}, i.e., 
+a well-typed \lang program can never crash due to spatial safety violations.
+We maintain the compiler formalism with the new features extended in \systemname.
+Especially, we maintain the bound-check insertions for array accesses 
+without ``fat'' pointer compilation scheme, while keeping the non-crashing guarantee.
+To certify the simulation relation between the \lang semantics and the compiler,
+we utilize the model-based randomized testing.
+This is done by a conversion tool that converts expressions from \lang into
+actual \checkedc code that can be compiled by the Clang \checkedc
+compiler. We build a random generator of programs largely based on the
+typing rules of \lang and make sure that, both statically and
+dynamically, \lang and Clang Checked C are consistent after
+conversion.  
+To the best of our knowledge, 
+\lang is the first language and compiler formalism with the program partitioning mechanism.
+
+\myparagraph{Supporting Checked Pointer Callbacks and No Checked Pointer Exposure}
+Our third contribution is an added up feature
+to support checked (function) pointer callbacks in unchecked core regions.
+In implementing a multi-threaded system, users might want 
+to provide third party interface permitting third parties 
+to write implement some new program features, while keeping those programs in unchecked core regions.
+Moreover, they do want to provide them a (function) pointer pointing to checked data fields.
+
+However, accessing a checked pointer in an unchecked region violates the program partition principle of \systemname.
+To overcome the discord, we develop two mechanisms in \systemname 
+and maintain a stronger \textit{non-exposure} guarantee on of the non-crashing guarantee; 
+i.e., no checked pointer addresses can be observed in an unchecked region.
+The first mechanism is nested checked and unchecked code regions.
+Users can context switch between checked and unchecked code regions 
+by nested using the keywords $\echeckedtext$ $\euncheckedtext$.
+The type system ensures that no checked pointers can be accessed across the context switching.
+The second one is that a call to a checked pointer in unchecked code regions 
+must be surrounded by a \textit{tainted shell}; 
+i.e., a tainted function pointer that points to a checked regions possibly holding checked pointers.
+In this case, no checked pointer address can be observed in the unchecked code regions.
+In \systemname, for every checked function, 
+we automatically compiler a tainted version by surrounding the function without a tainted shell.
+
+We begin with a review of \checkedc (Section~\ref{sec:overview}) and introduce new features in \systemname,
+present our main contributions
+(Sections~\ref{sec:formal}--\ref{sec:evaluation}), and conclude with a
+discussion of 
+related and future work (Sections~\ref{sec:related},
+\ref{sec:conclude}). All code and proof artifacts (both for Coq and
+Redex) can be found at \url{https://github.com/plum-umd/checkedc}.
+
+\ignore{
 \noindent
@@
 }
+}
 

### 2022-08-03 sandp23/background.tex

@@
 
-This section discusses the key merge features between Checked-C and RLBox and the problems we are trying to solve.
+This section describes \checkedc and new features \systemname provides.
+
+\subsection{\checkedc Overview}
+
+Development of Checked C was initiated by Microsoft Research in 2015 but starting in
+late 2021 was forked and is now actively managed by the Secure Software
+Development Project (SSDP). Details can be found in a prior
+overview~\cite{Elliott2018} or the formalism~\cite{li22checkedc}.
+
+\myparagraph{Checked Pointer Types}
+\checkedc introduces three varieties of \emph{checked pointer}:
+\begin{itemize}
+\item \code{ptr<}$T$\code{>} types a pointer that is either null or
+  points to a single object of type $T$.
+\item \code{array_ptr<}$T$\code{>} types a pointer that is either null
+  or points to an array of $T$ objects. The array width is defined
+  by a \emph{bounds} expression, discussed below.
+\item \code{nt_array_ptr<}$T$\code{>} is like
+  \code{array_ptr<}$T$\code{>} except that the bounds expression
+  defines the \emph{minimum} array width---additional objects may
+  be available past the upper bound, up to a null terminator.
+\end{itemize}
+A bounds expression used with the latter two pointer types has three
+forms:
+\begin{itemize}
+\item \code{count(}$e$\code{)} where $e$ defines the array's
+  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
+  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
+  expression $e$ must be side-effect free and may only refer to
+  variables whose addresses are not taken, or adjacent \code{struct}
+  fields.
+\item \code{byte_count(}$e$\code{)} is like \code{count}, but
+  expresses arithmetic using bytes, no objects; i.e.,
+  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
+  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
+\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
+  pointers that bound the accessible region $[e_l,e_h)$ (the
+  expressions are similarly restricted). Bounds
+  \code{count(}$e$\code{)} is shorthand for
+  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
+  expression is useful for supporting pointer arithmetic.
+\end{itemize}
+  Dropping the bounds expression on an \code{nt_array_ptr} is equivalent
+  to the bounds being \code{count(0)}.
+
+The \checkedc compiler will instrument loads and stores of checked
+pointers to confirm the pointer is non-null, and the access is within
+the specified bounds. For pointers $p$ of type
+\code{nt_array_ptr<}$T$\code{>}, such a check could spuriously fail if
+the index is past $p$'s specified upper bound, but before the null
+terminator. To address this problem, Checked C supports \emph{bounds
+  widening}.
+If $p$'s bounds expression is \code{bounds}$(e_l$,$e_h)$ a program may read from (but not
+write to) $e_h$; when the compiler notices that a non-null character
+is read at the upper bound, it will extend that bound to $e_h+1$.
+% \liyi{solved: II.B.: LaTeX issue, your inline monospaced code snippets generate a space on
+%   the following line (i.e. "a program may read from (but not write to)" is not
+%   aligned with the left boundary of the paragraph).}
+
+\myparagraph{Spatial Safety and Backward Compatibility}
+%
+\checkedc is backward compatible with legacy C in the sense that all
+legacy code will type-check and compile. However, only code that
+appears in \emph{checked code regions}, which we call \emph{checked code},
+is spatially safe. Checked regions can be designated at the level of
+files, functions, or individual code blocks, the first with a
+\code{#pragma} and the latter two using the \code{checked}
+keyword.\footnote{You can also designate \emph{unchecked} regions
+  within checked ones.}  Within checked regions, both
+legacy pointers and certain unsafe idioms (e.g., \emph{variadic} function
+calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
+satisfies these conditions, and will type-check in a checked region.
+
+How should we think about code that contains both checked and legacy
+components? \citet{li22checkedc} proved, for a simple
+formalization of \checkedc, that \emph{checked code cannot be blamed}:
+Any spatial safety violation is caused by the execution of unchecked
+code.
+
+\myparagraph{Converting C to \checkedc}
+%
+The safety guarantees of Checked C come with certain restrictions.
+For instance, as shown below, Checked C programs cannot use address-taken variables in a bounds expression as the bounds relations may not hold because of possible modifications through pointers.
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+...
+array_ptr<int> p : count (n) = NULL;
+|\textcolor{red}{\faTimes}|..,&n,.
+\end{minted}
+Consequently, converting existing C programs to Checked C might require refactoring, e.g., modifying the above program to not use~\inlinecode{&n} expression, which might require considerable effort~\cite{duanrefactoring} depending on the program's complexity.
+Recently, Machiry et al. developed~\threec~\cite{machiry2022c} that tries to automatically convert a program to Checked C by adding appropriate pointer annotations.
+However, as described in \threec, complete automated conversion is infeasible and requires the developer to convert some code regions manually.
+Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
 
@@
 
+\subsection{\systemname Transcendence}
+
+Here, we discuss three key new features in \systemname with examples.
+
 \myparagraph{Maintaining Non-crashing}
-Previously, the main guarantee of \checkedc \cite{li22checkedc} was the blame theorem,
-i.e., if there is a crash, the source is from some unchecked regions.
+Previously, the main guarantee of \checkedc \cite{li22checkedc} was the blame theorem.
+The sources of crashing in \checkedc are (1) unchecked regions crash themselves;
+  and (2) the misuse of checked pointers in unchecked regions.
 For example, at \Cref{fig:checkedc-example-3} line 31,
@@
 
-The philosophy is that unchecked regions are written in C and unstable,
-and programmers will eventually remove them by converting them to \checkedc code.
-The reality is that users might not want to convert everything.
-For some insignificant or handy history C code, such as library functions,
-they might want to keep it as long as no crashing.
-
-The sources of crashing in \checkedc are (1) unchecked regions crash themselves;
-  and (2) the misuse of checked pointers in unchecked regions.
-Enlightened by RLBox \cite{rlbox-paper},
-we sandbox the unchecked regions and then we utilize the \checkedc type system
-to disallow checked pointers to be used in an unchecked region. 
-To achieve the communication between checked and unchecked regions,
-we create tainted pointers that can be shared by different regions,
-whose data are stored in the sandboxed heap.
-Users are required to copy data to tainted pointers that are shared in unchecked regions.
+Enlightened by program partitioning mechanism,
+we sandbox the unchecked code regions and utilize the \checkedc type system
+to disallow checked pointers to be used in an unchecked code region. 
+To achieve the communication between checked and unchecked code regions,
+Tainted pointers can be shared by different regions,
+whose data are stored in the sandboxed unchecked heap.
+Users are required to copy checked data to tainted pointers before they are shared in unchecked regions.
 For example,  we copy the checked pointer data to the tainted pointer \code{tp}
@@
 
-In short, banning checked pointer uses in unchecked regions and providing tainted shared pointers
-are the solution for crashing. 
-In \systemname, we proved the non-crashing theorem based on the \systemname type system,
-.i.e, any program execution is stopped in a predictable manner
-where either the execution terminates or is stopped at a place where an exception handler exists.
-In fact, our work is \textit{the first formalism of RLBox},
-and formally shows the security guarantee that RLBox provides 
-is actually the non-crashing theorem.
+In \systemname, we prove the non-crashing theorem;
+.i.e, any well-typed \systemname program program can never crash due to spatial safety violations.
 
 \myparagraph{Formalism Function Pointers}
-In C, manipulating function pointers is a major way of implementing high order functions.
-Previously, we assumed all function calls are through calling by name to a global map.
+In C, manipulating function pointers is a major way of 
+implementing high order functions as well as accessing data stored in classes.
+Previously, \checkedc assumed all function calls are called by name to a global map.
 In \systemname, we formalize function pointers and maintain the \systemname type soundness.
-To the best of our knowledge,
-this is the first work of formalizing C function pointers with security guarantee.
 
 \Cref{fig:checkedc-example-1} defines a string sorting algorithm
-depending on the input function pointer \code{cmp} defining the generic order for strings,
+depending on the input function pointer \code{cmp} that provides the generic order for strings,
 and \code{compare_1} is an example \code{cmp} function
@@
 \myparagraph{Not Exposing Checked Pointer Addresses}
-The design for the non-crashing property bans the checked pointer uses in unchecked regions.
+The non-crashing guarantee in \systemname bans the checked pointer manipulation in unchecked code regions.
 Thus, there is no reason to permit checked pointer variable assignments in unchecked regions;
@@
 For example, the call to function $g$ at \Cref{fig:checkedc-example-3} line 41 lives in an unchecked region,
-and $g$ might use some mechanism, such as derandomizing ASLR \cite{shacham-aslr},
+and $g$ might use some mechanisms, such as derandomizing ASLR \cite{shacham-aslr},
 to achieve the checked pointer address.
@@
 For example, \code{tainted_compare_1} and \code{tainted_stringsort} in \Cref{fig:checkedc-example-2} 
-are the tainted shells of the checked functions \code{compare_1} and \code{stringsort}.
+are the tainted shells of the checked functions \code{compare_1} and \code{stringsort}, respectively.
 In the tainted shells, the arguments are tainted versions of the corresponding arguments in the checked functions.
-Inside the shell body, create checked pointer copies of the tainted arguments, and call the checked functions.
-In addition, once a checked function returns,
-if the output is a checked pointer, we also its the data to a new tainted pointer and exit the shell.
+Inside the shell body, we create checked pointer copies of the tainted arguments, and call the checked functions.
+In addition, once the checked function returns,
+if the output is a checked pointer, we copy its the data to a new tainted pointer and return it.
 \Cref{fig:checkedc-example-3} line 42 is an example of serving the function call living in an unchecked region
@@
 Conceptually, the shell is run in a checked region.
-One can imagine that a tainted shell is a safe closure that contains a checked block.
-Once the closure is called, the system is turned to a checked region,
-so that the checked function call at \Cref{fig:checkedc-example-2} line 14 is safe, 
+Essentially, a tainted shell is a safe closure that contains a checked block.
+Once the closure is called, the system is turned to a checked region.
+For example, \Cref{fig:checkedc-example-2} line 25 surrounds the tainted shell body with a $\echeckedtext$ block indicating
+that we context-switch from the unchecked to checked code region;
+thus, the checked function call at \Cref{fig:checkedc-example-2} line 14 is safe, 
 even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.

### 2022-08-03 sandp23/introduction.tex

@@
 The backward compatibility also enables incremental or partial conversion where Checked annotations can be provided to a few pointers.
-Such partially annotated programs still enjoy the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe core regions).
+Such partially annotated programs still enjoy the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe code regions).
 ~\fig{fig:memprotections} shows the summary of these existing approaches along with Checked C.
@@
 
-The guarantee \checkedc provides focuses on checked core regions,
+The guarantee \checkedc provides focuses on checked code regions,
 where any stuck (i.e., ill-defined)
@@
 \emph{blamed on}, the execution of code that is not in a checked core
-region. But what are the guarantees of unchecked (unsafe) core regions?
+region. But what are the guarantees of unchecked (unsafe) code regions?
 Programmers might not expect to convert every code to \checkedc,
@@
 \myparagraph{Executions Cannot Crash}
-Our first contribution is a clear program partition between checked and unchecked core regions,
+Our first contribution is a clear program partition between checked and unchecked code regions,
 and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
-We restrict the checked and unchecked pointers usage only in checked and unchecked core regions, respectively;
+We restrict the checked and unchecked pointers usage only in checked and unchecked code regions, respectively;
 and develop tainted pointers for the communication between the two regions.
@@
 Our third contribution is an added up feature
-to support checked (function) pointer callbacks in unchecked core regions.
+to support checked (function) pointer callbacks in unchecked code regions.
 In implementing a multi-threaded system, users might want 
 to provide third party interface permitting third parties 
-to write implement some new program features, while keeping those programs in unchecked core regions.
+to write implement some new program features, while keeping those programs in unchecked code regions.
 Moreover, they do want to provide them a (function) pointer pointing to checked data fields.
+To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
 

### 2022-08-07 sandp23/compilation.tex
update.

@@
 In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
-if the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
+If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
 
 This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
-with no loss of expressiveness;
+with no loss of expressiveness
 as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
@@
   light on how compilation can be implemented in the real Checked C
-  compiler, while eschewing many important details (\elang has many 
+  compiler, while eschewing many vital details (\elang has many 
   differences with LLVM IR).
@@
 Compilation is defined by extending \lang's
-typing judgment thusly:
+typing judgment as follows:
 \[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
@@
 each variable \code{p} to a pair of \emph{shadow
-  variables} that keep \code{p}'s up-to-date upper and lower bounds; 
-these may differ from the bounds in \code{p}'s type due to bounds
+  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
+These may differ from the bounds in \code{p}'s type due to bounds
 widening.\footnote{Since lower bounds are never widened, the
@@
 %
-A-normal form conversion is straightforward: compound expressions are
-handled by storing results of subexpressions into temporary variables,
+A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
 as in the following example.
@@
 to track the lower and upper bounds potentially changed for pointer arithmetic and NT-array bound widening.
-We also associated with every $\cmode$-mode pointer variable according to its type;
-we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
+Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
+Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
 In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
 there is an inserted verification step similar to \Cref{fig:const-type},
-which checks a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
+which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
 Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
-In implementation, we actually optimize the verification away and substitute it
-with the bounds and null-pointer checks.
-Because tainted pointer is checked every time before it is used,
-so that we only need to check the top pointer is 
-well defined without recursively looking at the sub-terms in a nested pointer case.
 The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
@@
 a \texttt{callback} to a checked block code $e$ inside a sandbox.
-In \systemname, We adopt an aggressive execution scheme that
+In \systemname, we adopt an aggressive execution scheme that
 directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
@@
 %
-Function pointers are dealt with similarly to normal pointers,
+Function pointers are managed similarly to normal pointers,
 but we insert checks to check if the pointer address is not null in 
@@
 testing. Redex allows us
-  to specify compilation as logical rules (essentially, an extension
+  to specify compilation as logical rules (an extension
   of typing), but then execute it algorithmically to
@@
 
-Our random generator (discussed in the next section) never produces
-$\euncheckedtext$ expressions (whose behavior could be undefined), so we
-can only test a the simulation theorem 
-as it applies to checked code. This limitation makes it
+As our random generator never generates
+$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
+as it relates to checked code. This limitation makes it
 unnecessary to state the other direction of the simulation theorem

### 2022-08-07 sandp23/formal.tex
update.

@@
 variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
-be updated; we can model mutable stack variables as pointers into the
+be updated. We can model mutable stack variables as pointers into the
 mutable heap.
 As mentioned, value $\evalue{0}{\tau}$
-represents a $\enull$ pointer when $\tau$ is a pointer type;
-correspondingly, $\heap(m,0)$ should always be undefined.
+represents a $\enull$ pointer when $\tau$ is a pointer type.
+Correspondingly, $\heap(m,0)$ should always be undefined.
 The relation steps to a \emph{result} $r$,
 which is either an expression or a $\enull$ or $\ebounds$ failure,
-representing a null-pointer dereference or out-of-bounds access,
+represent a null-pointer dereference or out-of-bounds access,
 respectively. Such failures are a \emph{good} outcome; stuck states
 (non-value expressions that cannot transition to a result $r$)
-characterize undefined behavior.
+characterizing undefined behavior.
 %
@@
 \textsc{S-DefNull}, and \textsc{S-Cast}.
-The first five define deference and assignment operations---illustrate how the semantics checks bounds.
+The first five define deference and assignment operations and illustrate how the semantics checks bounds.
 Rule \textsc{S-DefNull} transitions attempted null-pointer
@@
 %
-The semantics handles variable scopes using the special $\erettext$
+The semantics manages variable scopes using the special $\erettext$
 form. \textsc{S-Let} evaluates to a configuration whose expression is
 $\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
-and remember $x$ and its the new value $\evalue{n}{\tau}$
+and remember $x$ and its new value $\evalue{n}{\tau}$
 in $e$'s scope that is defined by the $\erettext$ operation.
 Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
-we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope;
-after one-step evaluation is completed, 
+we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope.
+After one-step evaluation is completed, 
 we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
@@
 %
-Rules \textsc{T-Def} and \textsc{T-AssignArr} check array
-dereference and assignment operations resp., returning the type of
-pointed-to objects; rules for pointers for other object types are
+\textsc{T-Def} and \textsc{T-AssignArr} rules examine array
+dereference and assignment operations respectively, returning the type of
+pointed-to objects. Rules for pointers for other object types are
 similar.
 The condition $m\le m'$ ensures that checked and unchecked pointers 
-can only be deferenced in checked and unchecked regions, resp.;
+can only be dereferenced in checked and unchecked regions, respectively;
 The type rules do not attempt to reason whether the access is in bounds;
@@
 $\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and it is built on the type equality
-($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$); the many
-rules ensure the relation is transitive. Most of the rules handle
+$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and is built on the type equality
+($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
+rules ensure the relation is transitive. Most of the rules manage
 casting between array pointer types. The second rule 
@@
 if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
-There is another casting rule in \Cref{app:main} states that
+There is another casting rule in \Cref{app:main} stating that
  users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
@@
 contain variables, determining assumptions like $b_l \leq b_l'$
-requires reasoning about those variables' possible values. The type
+requires reasoning about the probable values of these variables'. The type
 system uses $\Theta$ to make such reasoning more precise.
@@
 may have a type that refers to one of its bound variables (e.g., if
-the result is a checked pointer with a variable-defined bound); if so,
-we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
+the result is a checked pointer with a variable-defined bound). If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
 Note that we restrict the expression $e_1$ to syntactically match the
@@
 evaluating a let binding (rule \textsc{S-Let} in
-Fig.~\ref{fig:semantics}); this rule is needed for the preservation
+Fig.~\ref{fig:semantics}). This rule is needed for the preservation
 proof.
@@
 non-exposure, and non-crashing.
-These proofs have been carried out in our Coq model.
+These proofs have been conducted in our Coq model.
 
@@
 \begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
-A type environment $\Gamma$ is well-formed iff every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
+A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
 \end{defi}
@@
 \begin{defi}[Heap Well-formedness]
-For every $m$, A heap $\heap$ is well-formed iff (i) $\heap(m,0)$ is undefined, and
+For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
 (ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
@@
 \begin{defi}[Stack Well-formedness]
-A stack snapshot $\varphi$ is well-formed iff
+A stack snapshot $\varphi$ is well-formed if
 for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
@@
 stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
-\varphi$---iff for every variable $x$, $\Theta(x)$ is defined implies
+\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
 $\Gamma(x) = \tau$ for some $\tau$ and 
@@
 \begin{defi}[Checked Stack-Heap Consistency]
-A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---iff
+A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
 for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
@@
 \begin{defi}[Checked Heap-Heap Consistency]
-A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---iff
+A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
 for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.

### 2022-08-07 sandp23/background.tex
update.

@@
 
-Development of Checked C was initiated by Microsoft Research in 2015 but starting in
-late 2021 was forked and is now actively managed by the Secure Software
+\checkedc Development began in 2015 by Microsoft Research, but it was forked
+in late 2021 and is now actively managed by the Secure Software
 Development Project (SSDP). Details can be found in a prior
@@
 calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
-satisfies these conditions, and will type-check in a checked region.
+satisfies these conditions and will type-check in a checked region.
 
-How should we think about code that contains both checked and legacy
+How should we approach code that has both checked and legacy
 components? \citet{li22checkedc} proved, for a simple
@@
 However, as described in \threec, complete automated conversion is infeasible and requires the developer to convert some code regions manually.
-Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
+Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked, or safe regions).
 
@@
 For example, at \Cref{fig:checkedc-example-3} line 31,
-we call a unchecked function \code{f} with a checked null-terminated array (NT-array) pointer argument.
+we called an unchecked function \code{f} with a checked null-terminated array (NT-array) pointer argument.
 At line 8, depending on the NT-array size, \code{free(s[10])} might crash.
@@
 requires a verification on it.
-This is handled by inserting additional checks
+This is managed by inserting additional checks
 and creating exception handling before the use by the \systemname compiler.
@@
 and allows the client to access some server resources by calling back the pointer.
-This is a common usage between a web-browser and untrusted third party libraries.
+This is a common usage between a web-browser and untrusted third-party libraries.
 The function call in \Cref{fig:checkedc-example-3} line 34 is one such usage.
@@
 
-To prevent the checked pointer address leak,
-we prevent any unchecked regions acknowledge checked pointer variables.
+we prevent any unchecked regions from acknowledging checked pointer variables to avoid the checked pointer address leak.
 In addition, to facilitate checked function callbacks, 
@@
 In addition, once the checked function returns,
-if the output is a checked pointer, we copy its the data to a new tainted pointer and return it.
+if the output is a checked pointer, we copy its data to a new tainted pointer and return it.
 \Cref{fig:checkedc-example-3} line 42 is an example of serving the function call living in an unchecked region
 with a tainted shell pointer argument \code{tainted_compare_1}.
-Even if $g$ derndomizes its address (line 14), the shell address is in the sandbox and has no harm,
+Even if $g$ derandomizes its address (line 14), the shell address is in the sandbox and has no harm,
 and calling the shell never exposes any checked pointer information outside of the shell.

### 2022-08-07 sandp23/introduction.tex
update DynCast.
update.

@@
-\section{Introduction}
+\section{Introduction}\label{sec:intros}
 
@@
 despite a large body of work that tries to prevent them~\cite{song2019sanitizing}.
-Several industrial and research efforts---including CCured~\cite{Necula2005},
-Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
-explored means to compile C programs 
-to automatically enforce spatial safety. These
-approaches all impose performance overheads deemed too high for
-deployment use. 
-Recently, \citet{Elliott2018,li22checkedc} introduced and formalized \checkedc, an
+Several industrial and research efforts, including CCured~\cite{Necula2005},
+Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
+have investigated ways to compile C programs 
+to automatic spatial safety enforcement.
+These approaches all impose performance overheads deemed too high for deployment use. 
+Recently, \citet{Elliott2018} and \citet{li22checkedc} introduced and formalized \checkedc, an
 open-source extension to C with new types and
@@
 may coexist in the same code, which permits a deliberate (and
-semi-automated) refactoring process. Parts of the FreeBSD kernel have
-been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
-overhead seems low enough for practical deployment.
+semi-automated) refactoring process. 
+Parts of the FreeBSD kernel have
+been successfully migrated to \checkedc~\cite{duanrefactoring}, and the overall performance
+overhead is minimal enough for practical deployment.
 
@@
 state reached by a well-typed program amounts to a spatial safety
-violation; such a state can always be attributed to, i.e., 
-\emph{blamed on}, the execution of code that is not in a checked core
-region. But what are the guarantees of unchecked (unsafe) code regions?
+violation.
+Such a state can always be attributed to, i.e., 
+\emph{blamed on}, the execution of code that is not in a checked core region. 
+However, what are the guarantees that unchecked (unsafe) code regions can provide?
 Programmers might not expect to convert every code to \checkedc,
@@
 \myparagraph{Executions Cannot Crash}
-Our first contribution is a clear program partition between checked and unchecked code regions,
+Our first contribution is a clear program partition between checked and unchecked code regions
 and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
@@
 and develop tainted pointers for the communication between the two regions.
-Heap is partitioned into two parts: a \textit{checked memory region} containing all checked pointers
- and an \textit{unchecked memory region} that is sandboxed.
-Tainted pointers live in the unchecked memory region and is verified every time when using.
+Heap is partitioned into two parts: a \textit{checked heap} containing all checked pointers
+ and an \textit{unchecked heap} that is sandboxed.
+Tainted pointers exist in the unchecked heap region and must be verified each time being used.
 
-\myparagraph{Formaling the Type System, Semantics and Compiler}
+\myparagraph{Formalizing the Type System, Semantics and Compiler}
 %
@@
 actual \checkedc code that can be compiled by the Clang \checkedc
-compiler. We build a random generator of programs largely based on the
-typing rules of \lang and make sure that, both statically and
-dynamically, \lang and Clang Checked C are consistent after
-conversion.  
+compiler. We create a program random generator 
+based on the
+typing rules of \lang and ensure that \lang and Clang Checked C are consistent after conversion, both statically and
+dynamically.  
 To the best of our knowledge, 
-\lang is the first language and compiler formalism with the program partitioning mechanism.
+\lang is the first C(-like) language and compiler formalism with the program partitioning mechanism.
 
 \myparagraph{Supporting Checked Pointer Callbacks and No Checked Pointer Exposure}
-Our third contribution is an added up feature
+Our third contribution is an added-up feature
 to support checked (function) pointer callbacks in unchecked code regions.
-In implementing a multi-threaded system, users might want 
-to provide third party interface permitting third parties 
-to write implement some new program features, while keeping those programs in unchecked code regions.
+When designing a multi-threaded system, users might want 
+to provide a third party interface that allows third party developers to create new program features, while keeping these programs in unchecked code regions.
 Moreover, they do want to provide them a (function) pointer pointing to checked data fields.
-To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
 
 However, accessing a checked pointer in an unchecked region violates the program partition principle of \systemname.
-To overcome the discord, we develop two mechanisms in \systemname 
+To resolve the conflict, we develop two mechanisms in \systemname 
 and maintain a stronger \textit{non-exposure} guarantee on of the non-crashing guarantee; 
-i.e., no checked pointer addresses can be observed in an unchecked region.
-The first mechanism is nested checked and unchecked code regions.
+that is, no checked pointer addresses can be observed in an unchecked code region.
+The first mechanism allows nested checked and unchecked code regions.
 Users can context switch between checked and unchecked code regions 
@@
 must be surrounded by a \textit{tainted shell}; 
-i.e., a tainted function pointer that points to a checked regions possibly holding checked pointers.
-In this case, no checked pointer address can be observed in the unchecked code regions.
+i.e., a tainted function pointer that points to a checked region possibly holding checked pointers.
+In this case, no checked pointer address will be observed in the unchecked code regions.
 In \systemname, for every checked function, 
-we automatically compiler a tainted version by surrounding the function without a tainted shell.
+we automatically compile a tainted version by surrounding the function without a tainted shell.
+To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
 

### 2022-08-08 sandp23/background.tex
Adding mods to intro

--- 2022-07-17 sandp23/background.tex
@@
-\section{Overview and Transcendence}\label{sec:overview}
-
-This section discusses the key merge features between Checked-C and RLBox and the problems we are trying to solve.
-
-\begin{figure}[t]
-{\small
-  \begin{lstlisting}[xleftmargin=4 mm]
-//in checked region
-
-int compare_1(nt_array_ptr<char> x: count (0),
-  nt_array_ptr<char> y : count (0)) { 
-  int len_x = strlen(x);
-  int len_y = strlen(y);
-  return sum(x,len_x) < sum(y,len_y);
-  }
-...
-
-int stringsort(
- nt_array_ptr<nt_array_ptr<char>> s : count (n),
- ptr<(int)(nt_array_ptr<char>,
-  nt_array_ptr<char>)> cmp, int n) {
-  int i, j, gap;
-  int didswap;
-
-  for(gap = n / 2; gap > 0; gap /= 2) {
-    {
-    do  {
-        didswap = 0;
-        for(i = 0; i < n - gap; i++)
-           {
-             j = i + gap;
-             if((*cmp)(s[i], s[j]) > 0)
-             {
-                int len = strlen(s[i]);
-                nt_array_ptr<char> 
-                  tmp : count (len) = s[i];
-                s[i] = s[j];
-                s[j] = tmp;
-                didswap = 1;
-             }
-           }
-        } while(didswap);
-    }
-  }
-
-  return 0;
-}
-  \end{lstlisting}
-}
-\caption{Checked stringsort Code}
-\label{fig:checkedc-example-1}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{lstlisting}[xleftmargin=4 mm]
-//in checked region, tainted version
-int tainted_compare_1(
-  nt_array_tptr<char> x : count (0),
-  nt_array_tptr<char> y : count (0)) {
- checked (x,y) {
-  int len_x = strlen(x);
-  int len_y = strlen(y);
-  nt_array_ptr<char> tx : count (len_x)
-   = malloc(nt_array<char>, len_x);
-  nt_array_ptr<char> ty : count (len_y)
-   = malloc(nt_array<char>, len_y);
-  safe_memcpy(tx,x,len_x);
-  safe_memcpy(ty,y,len_y);
-  return compare_1(tx,ty);
- }
-}
-...
-
-//calling the function turns
-//an unchecked region to a checked region.
-int tainted_stringsort(nt_array_tptr
-     <nt_array_tptr<char>> s : count (n),
- tptr<(int)(nt_array_tptr<char>,
-  nt_array_tptr<char>)> cmp, int n) {
- checked (s,cmp,n) {
-  int i;
-  nt_array_ptr<nt_array_ptr<char>> p : count (n)
-    = malloc(nt_array<nt_array_ptr<char>>, n);
-  for(i = 0; i < n; i++) {
-    int len = strlen s[i];
-    nt_array_ptr<char> tmp : count (len)
-      = new malloc(nt_array<char>, len);
-    safe_memcpy(tmp,s[i],len);
-    p[i] = tmp;
-  }
-  ptr<(int)(nt_array_ptr<char> : count (0),
-  nt_array_ptr<char> : count (0))>
-    cfun = find_check(cmp);
-
-  return stringsort(p,cfun);
- }
-}
-  \end{lstlisting}
-}
-\caption{Tainted stringsort Code}
-\label{fig:checkedc-example-2}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{lstlisting}[xleftmargin=4 mm]
-//in unchecked region
-int f(char ** s, int (*cmp)(char *,char *),
-  int (*sort)(char **, int (*)(char *,char *),
-               int), int n) {
-...
-
-  int i = sort(s,cmp,n);
-  free(s[10]);
-...
-}
-
-int g(int (*cmp)(char *,char *)) {
-...
-  int real_addr = derandomize(cmp);
-...
-}
-
-int main(int n) {
-  nt_array_ptr<nt_array_ptr<char>> p : count(n)
-    = malloc(nt_array<nt_array_ptr<char>>, n);
-
-  nt_array_tptr<nt_array_tptr<char>>
-     tp : count(n) =
-       tmalloc(nt_array<nt_array_tptr<char>>, n);
-...
-  safe_memcpy(tp, p, n);
-
-  unchecked {
-    if (BAD) // a flag to call different funs.
-      //input checked pointers
-      f(p, compare_1, stringsort); 
-    else 
-      //input tainted pointers
-      f(tp, tainted_compare_1, tainted_stringsort);
-  }
-
-  if (!BAD) safe_memcpy(p, tp, n);
-  p[10] = "crash?";
-
-  unchecked {
-    if (BAD) g(compare_1)
-      else g(tainted_compare_1);
-  }
-
-  return 0;
-}
-  \end{lstlisting}
-}
-\caption{Tainted Pointer Usage in Calling Unchecked Fun}
-\label{fig:checkedc-example-3}
-\end{figure}
-
-\myparagraph{Maintaining Non-crashing}
-Previously, the main guarantee of \checkedc \cite{li22checkedc} was the blame theorem,
-i.e., if there is a crash, the source is from some unchecked regions.
-For example, at \Cref{fig:checkedc-example-3} line 31,
-we call a unchecked function \code{f} with a checked null-terminated array (NT-array) pointer argument.
-At line 8, depending on the NT-array size, \code{free(s[10])} might crash.
-Even if it does not crash, line 38 is doomed because of the \code{free} call.
-
-The philosophy is that unchecked regions are written in C and unstable,
-and programmers will eventually remove them by converting them to \checkedc code.
-The reality is that users might not want to convert everything.
-For some insignificant or handy history C code, such as library functions,
-they might want to keep it as long as no crashing.
-
-The sources of crashing in \checkedc are (1) unchecked regions crash themselves;
-  and (2) the misuse of checked pointers in unchecked regions.
-Enlightened by RLBox \cite{rlbox-paper},
-we sandbox the unchecked regions and then we utilize the \checkedc type system
-to disallow checked pointers to be used in an unchecked region. 
-To achieve the communication between checked and unchecked regions,
-we create tainted pointers that can be shared by different regions,
-whose data are stored in the sandboxed heap.
-Users are required to copy data to tainted pointers that are shared in unchecked regions.
-For example,  we copy the checked pointer data to the tainted pointer \code{tp}
-at \Cref{fig:checkedc-example-3} line 26,
-and input the tainted pointer to the unchecked function at line 34.
-At line 8, even if statement might crash, since tainted pointers are stored
-in the sandboxed heap, it can be recovered.
-At line 37, the use of a tainted pointer in a checked region
-requires a verification on it.
-This is handled by inserting additional checks
-and creating exception handling before the use by the \systemname compiler.
-Thus, the checked pointer \code{p} is safely used at line 38.
-
-In short, banning checked pointer uses in unchecked regions and providing tainted shared pointers
-are the solution for crashing. 
-In \systemname, we proved the non-crashing theorem based on the \systemname type system,
-.i.e, any program execution is stopped in a predictable manner
-where either the execution terminates or is stopped at a place where an exception handler exists.
-In fact, our work is \textit{the first formalism of RLBox},
-and formally shows the security guarantee that RLBox provides 
-is actually the non-crashing theorem.
-
-\myparagraph{Formalism Function Pointers}
-In C, manipulating function pointers is a major way of implementing high order functions.
-Previously, we assumed all function calls are through calling by name to a global map.
-In \systemname, we formalize function pointers and maintain the \systemname type soundness.
-To the best of our knowledge,
-this is the first work of formalizing C function pointers with security guarantee.
-
-\Cref{fig:checkedc-example-1} defines a string sorting algorithm
-depending on the input function pointer \code{cmp} defining the generic order for strings,
-and \code{compare_1} is an example \code{cmp} function
-that adds the ASCII numbers of characters in the two strings and compare the results.
-In addition, function pointers enable the callback mechanism,
-i.e., a server sends a function pointer to a client in an unchecked region,
-and allows the client to access some server resources by calling back the pointer.
-This is a common usage between a web-browser and untrusted third party libraries.
-The function call in \Cref{fig:checkedc-example-3} line 34 is one such usage.
-
-We also utilize \systemname subtyping relation to permit function pointer static auto-casting.
-Function pointer type information might contain array pointer bound information,
-for which it is inconvenient to coincide the defined types for a function implementation and the function pointer type. 
-For example, the \code{cmp} argument in \code{stringsort} (\Cref{fig:checkedc-example-1})
-has type \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>},
-meaning that the function takes two NT-array pointers with arbitrary size and outputs an integer.
-The function \code{compare_1}'s pointer has type \code{ptr<(int)(nt_array_ptr<char> : count (0), nt_array_ptr<char> : count (0) )>}
-To use \code{compare_1} in \code{stringsort}, the type is auto-cast to the \code{cmp}'s type.
-In general, if function pointer $x$ has type $* (tl \to t)$, and  $y$ has $* (tl' \to t')$,
-in order to use $x$ as $y$, $tl'$ should be a subtype of $tl$ and $t$ subtypes to $t'$.
-
-\myparagraph{Not Exposing Checked Pointer Addresses}
-The design for the non-crashing property bans the checked pointer uses in unchecked regions.
-Thus, there is no reason to permit checked pointer variable assignments in unchecked regions;
-especially, this might expose a checked pointer address to untrusted parties.
-For example, the call to function $g$ at \Cref{fig:checkedc-example-3} line 41 lives in an unchecked region,
-and $g$ might use some mechanism, such as derandomizing ASLR \cite{shacham-aslr},
-to achieve the checked pointer address.
-Thus, it enables a third party to access any checked heap and function data by simple pointer arithmetic.
-
-To prevent the checked pointer address leak,
-we prevent any unchecked regions acknowledge checked pointer variables.
-In addition, to facilitate checked function callbacks, 
-the \systemname compiler compiles every checked function
-with an additional tainted shell function.
-Users are required to serve unchecked regions with the tainted shell pointer instead of the original checked function pointer.
-For example, \code{tainted_compare_1} and \code{tainted_stringsort} in \Cref{fig:checkedc-example-2} 
-are the tainted shells of the checked functions \code{compare_1} and \code{stringsort}.
-In the tainted shells, the arguments are tainted versions of the corresponding arguments in the checked functions.
-Inside the shell body, create checked pointer copies of the tainted arguments, and call the checked functions.
-In addition, once a checked function returns,
-if the output is a checked pointer, we also its the data to a new tainted pointer and exit the shell.
-\Cref{fig:checkedc-example-3} line 42 is an example of serving the function call living in an unchecked region
-with a tainted shell pointer argument \code{tainted_compare_1}.
-Even if $g$ derndomizes its address (line 14), the shell address is in the sandbox and has no harm,
-and calling the shell never exposes any checked pointer information outside of the shell.
-
-Conceptually, the shell is run in a checked region.
-One can imagine that a tainted shell is a safe closure that contains a checked block.
-Once the closure is called, the system is turned to a checked region,
-so that the checked function call at \Cref{fig:checkedc-example-2} line 14 is safe, 
-even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.
-In the \systemname formalism, we formalize a \textit{checked} block on top of the existing unchecked regions,
-and the transition of a tainted shell call creates a checked block containing the shell body.
-We also make sure that no arguments in these tainted shell contains any checked pointers,
-as well as no output is of a checked type.
-
-% \paragraph*{\textbf{NT-array Pointer Formalization}} NT-array pointers point to NT-arrays whose endings are determined by a null (\code{'\0'}).
-The length of a NT-array is usually not fixed.
-The two uses of NT-array pointers in \checkedc are branching and \code{strlen} operations.
-\code{if (*x) }$e_1$\code{ else }$e_2$ branches to $e_1$/$e_2$ depending one the \code{x}'s data, while \code{strlen(x)} computes the length of the NT-array pointed to by \code{x}.
-% \liyi{move to formal and later section.}
-
-% \paragraph*{\textbf{Dependent Functions}} \checkedc allows users to declare dependent functions.
-Fig.~\ref{fig:checkedc-example} shows an example usage of dependent functions.
-The integer argument \code{x} is used to specify the bounds of the two NT-array pointers.
-It means that when we call the \code{memcpy} function, the upper-bounds of the two NT-array pointers must be no less than \code{x}.
-If the programmer makes the mistake of writing \lstinline{i <= x} in the for loop, the runtime checks will ensure that the indexing of\lstinline|x| in either \code{a} or \code{b} results in an error.
-% Because of the dependent function features, we are able to utilize the fact that the two remaining length of the NT-array pointers are greater than \code{x}, so that the \code{for} is guaranteed to execute properly.
-
-% \ignore{
-% \liyi{Good information about pointer types here:
-% https://github.com/microsoft/checkedc/wiki/New-pointer-and-array-types}
-
-% This code creates an array pointer of length $10$, and reads the pointer $(x+y)$. 
-% If the integer value $y$ is greater than $10$, executing the program in C results in undefined behavior. 
-% In \checkedc, however, we links every pointer in a program with their bound information statically, and inserts checks to dynamically verify if the usage of the pointer violates the spatial safety.
-For example, In compiling the above fragment, we inserts the bound's check for $x$ in the read operation $(x+y)$, and if $y$ results in a value $11$, the execution of the read operation $(x+y)$ results in an dynamic error.
-
-% There are many existing language development trying to guarantee the C spatial safety property through different approaches.
-For example, Cyclone \cite{Jim2002} and Cerberus \cite{cerberus} are the fat pointer approaches for guaranteeing the spatial safety and eliminating undefined behaviors in C.
-Rust \cite{Rust2016}, on the other hand, tries to restrict the possible way of writing programs to guarantee such property statically.
-% The problem of these two approaches are effectiveness of executing programs and user practicality of writing programs.
-The fat-pointer approach completely relies on dynamic checks to guarantee the safety properties but it might pay a huge price on execution efficiency.
-In our experiment (Sec.~\ref{sec:evaluation}), we found that the execution of fat pointer approaches cost the speed a 50\% overhead.
-% On the other hand, Rust uses a static approach by restricting the usage of pointers to guarantee the properties, which results in a user UN-friendly system.  
-% The development of \checkedc is an investigation on keeping the balance between execution efficiency and user-friendly systems.
-We develop a system to utilize static type systems to allow the \checkedc compiler to insert dynamic checks to guarantee the safety property.
-The efficiency of \checkedc is in the same level of Rust while we maintain a user-friendly system and inherited most of the operation features from C.
-
-% Another feature of \checkedc is the division of the checked and unchecked blocks.
-When users try to rewrite their C programs into \checkedc programs, they can develop the code incrementally by rewriting sub-parts of their programs to \checkedc code fragments and placing them as the checked blocks, while keeping the rest C programs as unchecked C programs.
-\checkedc guarantees that if there is any safety violation, this comes from the unchecked blocks (the blame theory).
-
-% Here is the contributions of the paper.
-First, we develop the code formalization of \checkedc and show that the type system in \checkedc is type-sound and satisfies the blame theory with respect to the \checkedc semantics.
-Specifically, we formalize null-terminated array pointers.
-To our best knowledge, this is the first work defining such feature in C-related languages.
-% Second, in formalizing the \checkedc language, we investigate the balance between program execution efficiency and user practicality.
-We re-investigate the \checkedc type system with subtyping relations to make it more efficient and user-friendly.
-% Second, we also develop random testing tools in Redex \cite{pltredex} to test the \checkedc compiler and make sure it is properly developed.
-We are able to find and reproduce many bugs/faults in the \checkedc compiler \liyi{numbers?}.
-The random testing tool is able to generate tens of thousands of programs to properly validate the \checkedc compiler.
-Third, we also spent a great amount of work in developing experiments to compare \checkedc and other existing C-like languages that guarantee the spatial safety and eliminating undefined behaviors.
-We found that \checkedc keeps the best balance between execution efficiency and user practicality.
-% }
+\section{Background}
+\subsection{Program Partitioning}
+\label{subsec:background:programpart}

### 2022-08-08 sandp23/introduction.tex
Adding mods to intro

@@
 % \end{itemize}
+\noindent
+\textbf{No safety against unchecked code.} The spatial safety guarantee~\checkedc provides applies to completely converted programs, i.e., programs that uses only checked types.
+Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
+But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+Although the blameless proof exists~\cite{ruef2019achieving} for~\checkedc, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
+Consider the following example:
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+// Checked code
+int func(array_ptr<char> p : count(5)) {
+|\textcolor{red}{\faChainBroken}|..p[4]..
+}
+// unchecked code
+...
+str = "he";
+...
+|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5); 
+...
+char ptr[16];
+...
+len <- derived from user input
+...
+|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
+\end{minted}
+Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function with a buffer of 2 elements.
+This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
+Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
+We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+This mechanism is called program partitioning~\cite{rul2009towards}.
+Existing mechanisms are not suitable as they are based on process isolation and have high overhead, and are data-centric~\emph{data-centric} (\sect{subsec:background:programpart}).
+Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
+But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
+We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
+
+ and there has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
+Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
+E.g., Given sensitive data in a program, the goal is to partition functions into two parts or partitions based on whether a function can access the sensitive data.
+The performance overhead of these approaches is dominated by marshaling costs and depends on the usage of sensitive data.
+The overhead of state-of-the-art approaches~\cite{lind2017glamdring, liu2017ptrsplit} is prohibitive and varies from 37\%-163\%.
+But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
+We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
 

### 2022-08-09 sandp23/introduction.tex
New intro

@@
 
- and there has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
-Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
-E.g., Given sensitive data in a program, the goal is to partition functions into two parts or partitions based on whether a function can access the sensitive data.
-The performance overhead of these approaches is dominated by marshaling costs and depends on the usage of sensitive data.
-The overhead of state-of-the-art approaches~\cite{lind2017glamdring, liu2017ptrsplit} is prohibitive and varies from 37\%-163\%.
-But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
-We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
+In this work, we propose a type-directed code-centric program partitioning approach.
+Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
+The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.
 
-The guarantee \checkedc provides focuses on checked code regions,
-where any stuck (i.e., ill-defined)
-state reached by a well-typed program amounts to a spatial safety
-violation.
-Such a state can always be attributed to, i.e., 
-\emph{blamed on}, the execution of code that is not in a checked core region. 
-However, what are the guarantees that unchecked (unsafe) code regions can provide?
-Programmers might not expect to convert every code to \checkedc,
-but leave some functions in unsafe regions,
-in which case the unsafe region executions should not affect other parts.
-In this paper, we cover this gap by introducing \systemname,
-combining \checkedc with program partitioning mechanism, 
-making three main contributions.
+The developer starts by marking desired (i.e., unchecked) functions and pointers used in these functions as tainted.
+Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
+\begin{itemize}
+\item~\ucregion or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (i.e., functions and pointers) and can only access tainted pointers.
+\item~\cregion or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
+The functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
+\end{itemize}
+Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (e.g., WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
 
-\myparagraph{Executions Cannot Crash}
+The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
+As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
+We avoid marshaling by allocating all tainted pointers (i.e., tainted buffers) in~\ucregion and thus can be accessed in both partitions.
+Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
+Our compiler avoids these attacks by ensuring, either statically or through dynamic checks that tainted pointers can only point to~\ucregion address space.
+
+In summary, we make the following three main contributions.
+
+\myparagraph{Type-Directed Program Partitioning}
 Our first contribution is a clear program partition between checked and unchecked code regions

### 2022-08-10 sandp23/introduction.tex
Adding intro

@@
 \myparagraph{Type-Directed Program Partitioning}
-Our first contribution is a clear program partition between checked and unchecked code regions
-and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
-We restrict the checked and unchecked pointers usage only in checked and unchecked code regions, respectively;
-and develop tainted pointers for the communication between the two regions.
-Heap is partitioned into two parts: a \textit{checked heap} containing all checked pointers
- and an \textit{unchecked heap} that is sandboxed.
-Tainted pointers exist in the unchecked heap region and must be verified each time being used.
+We present a type-directed program partition technique to separate checked and unchecked code regions and ensure that the code in checked region~\textit{cannot crash} due to spatial safety violations.
+Our type system restricts checked and tainted (or unchecked) pointers usage only in~\cregion and~\ucregion, respectively.
+Our modular design enables us to use existing sandboxed techniques to enforce memory and execution isolation.
 
@@
 %
-Our second contribution is a core formalism named \lang, which extends
-\citet{li22checkedc} with the non-crashing guarantee and other new features below.
-We prove formally the \textit{non-crashing theorem}, i.e., 
-a well-typed \lang program can never crash due to spatial safety violations.
-We maintain the compiler formalism with the new features extended in \systemname.
-Especially, we maintain the bound-check insertions for array accesses 
-without ``fat'' pointer compilation scheme, while keeping the non-crashing guarantee.
-To certify the simulation relation between the \lang semantics and the compiler,
-we utilize the model-based randomized testing.
-This is done by a conversion tool that converts expressions from \lang into
-actual \checkedc code that can be compiled by the Clang \checkedc
-compiler. We create a program random generator 
-based on the
-typing rules of \lang and ensure that \lang and Clang Checked C are consistent after conversion, both statically and
-dynamically.  
-To the best of our knowledge, 
-\lang is the first C(-like) language and compiler formalism with the program partitioning mechanism.
+We developed a core formalism named~\lang, which extends
+\citet{li22checkedc}~\checkedc formalism with the non-crashing guarantee and other new features below.
+We prove formally the~\emph{non-crashing theorem}, i.e., 
+a well-typed~\lang program can never crash due to spatial safety violations.
+We utilize the model-based randomized testing (CITE) to certify the simulation relation between the~\lang semantics and the compiler.
+Specifically, we use a conversion tool that converts expressions from~\lang into actual~\checkedc code that can be compiled by the \checkedc compiler. We create a random program generator 
+based on the typing rules of~\lang and ensure that~\lang and~\checkedc compiler are consistent after conversion, both statically and dynamically.  
+To the best of our knowledge,~\lang is the first C(-like) language and compiler formalism with the program partitioning mechanism.
 

### 2022-08-10 sandp23/introduction.tex
Adding intro

@@
 
-\myparagraph{Supporting Checked Pointer Callbacks and No Checked Pointer Exposure}
+\myparagraph{Supporting callbacks to~\cregion with no Checked Pointer Exposure}
+Although we disallow access to~\cregion from~\ucregion, there can be cases where such access is needed. Specifically, when~\cregion wants to provide access to certain shared checked data to~\ucregion.
+To enable this, we support callback functions in~\cregion that can be invoked from~\ucregion through function pointers.
+However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of \systemname.
+
+-- NEED TO FINISH THIS--
 Our third contribution is an added-up feature

### 2022-08-16 sandp23/background.tex
cleanup, review background

--- 2022-08-07 sandp23/background.tex
@@
 \section{Overview and Transcendence}\label{sec:overview}
-
-This section describes \checkedc and new features \systemname provides.
+This section describes \checkedc{} and new features \systemname{} provides.
 
 \subsection{\checkedc Overview}
-
-\checkedc Development began in 2015 by Microsoft Research, but it was forked
+% 
+\checkedc{} development began in 2015 by Microsoft Research, but it was forked
 in late 2021 and is now actively managed by the Secure Software
@@
 overview~\cite{Elliott2018} or the formalism~\cite{li22checkedc}.
+%
 
 \myparagraph{Checked Pointer Types}
-\checkedc introduces three varieties of \emph{checked pointer}:
+\checkedc{} introduces three varieties of \emph{checked pointer}:
 \begin{itemize}
@@
 is read at the upper bound, it will extend that bound to $e_h+1$.
-% \liyi{solved: II.B.: LaTeX issue, your inline monospaced code snippets generate a space on
+% pp\liyi{solved: II.B.: LaTeX issue, your inline monospaced code snippets generate a space on
 %   the following line (i.e. "a program may read from (but not write to)" is not
@@
 %
-\checkedc is backward compatible with legacy C in the sense that all
-legacy code will type-check and compile. However, only code that
-appears in \emph{checked code regions}, which we call \emph{checked code},
-is spatially safe. Checked regions can be designated at the level of
-files, functions, or individual code blocks, the first with a
-\code{#pragma} and the latter two using the \code{checked}
-keyword.\footnote{You can also designate \emph{unchecked} regions
-  within checked ones.}  Within checked regions, both
-legacy pointers and certain unsafe idioms (e.g., \emph{variadic} function
-calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
-satisfies these conditions and will type-check in a checked region.
+\checkedc is backward compatible with legacy C in the sense that all legacy code
+will type-check and compile. However, only code that appears in \emph{checked
+  code regions}, which we call \emph{checked code}, is spatially safe. Checked
+regions can be designated at the level of files, functions, or individual code
+blocks, the first with a \code{#pragma} and the latter two using the
+\code{checked} keyword.\footnote{You can also designate \emph{unchecked} regions
+  within checked ones.}  Within checked regions, both legacy pointers and
+certain unsafe idioms (e.g., \emph{variadic} function calls) are disallowed. The
+code in Fig.~\ref{fig:checkedc-example} satisfies these conditions and will
+type-check in a checked region.
 
-How should we approach code that has both checked and legacy
-components? \citet{li22checkedc} proved, for a simple
-formalization of \checkedc, that \emph{checked code cannot be blamed}:
-Any spatial safety violation is caused by the execution of unchecked
-code.
+How should we approach code that has both checked and legacy components?
+\citet{li22checkedc} proved, for a simple formalization of \checkedc, that
+\emph{checked code cannot be blamed}: Any spatial safety violation is caused by
+the execution of unchecked code.
+%
+\mz{You might want to add that spatial safety violation is still there with the
+  amalgamation of both checked and unchecked code which gave rise to the
+  motivation of \systemname.
+}
+% 
+%
 
@@
 %
-The safety guarantees of Checked C come with certain restrictions.
-For instance, as shown below, Checked C programs cannot use address-taken variables in a bounds expression as the bounds relations may not hold because of possible modifications through pointers.
+The safety guarantees of Checked C come with certain restrictions. For instance,
+as shown below, Checked C programs cannot use address-taken variables in a
+bounds expression as the bounds relations may not hold because of possible
+modifications through pointers.
+% 
 \begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 \end{minted}
-Consequently, converting existing C programs to Checked C might require refactoring, e.g., modifying the above program to not use~\inlinecode{&n} expression, which might require considerable effort~\cite{duanrefactoring} depending on the program's complexity.
-Recently, Machiry et al. developed~\threec~\cite{machiry2022c} that tries to automatically convert a program to Checked C by adding appropriate pointer annotations.
-However, as described in \threec, complete automated conversion is infeasible and requires the developer to convert some code regions manually.
-Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked, or safe regions).
+% 
+Consequently, converting existing C programs to Checked C might require
+refactoring, e.g., eliminate~\inlinecode{&n} from the program above without
+changing its functionality.
+% 
+This might require considerable effort~\cite{duanrefactoring} depending on the
+program's complexity.
+% 
+Recently, Machiry et al. developed~\threec~\cite{machiry2022c} that tries to
+automatically convert a program to Checked C by adding appropriate pointer
+annotations.  However, as described in \threec, completely automated conversion
+is \emph{infeasible}, and it requires the developer to convert some code regions
+manually.  
+% 
+Although the backward compatibility of Checked C helps a partially annotated
+program to enjoy spatial memory safety on those regions using only Checked
+pointers (i.e., checked, or safe regions), \mz{then what?}.
 
@@
 Previously, the main guarantee of \checkedc \cite{li22checkedc} was the blame theorem.
-The sources of crashing in \checkedc are (1) unchecked regions crash themselves;
+The sources of crashing in \checkedc{} are (1) \mzr{crashes in unchecked regions}
 and (2) the misuse of checked pointers in unchecked regions.
-For example, at \Cref{fig:checkedc-example-3} line 31,
-we called an unchecked function \code{f} with a checked null-terminated array (NT-array) pointer argument.
+% 
+For example, at \Cref{fig:checkedc-example-3} line 31, we passed a checked
+null-terminated array (NT-array) pointer into an unchecked function \code{f}.
+% 
 At line 8, depending on the NT-array size, \code{free(s[10])} might crash.
-Even if it does not crash, line 38 is doomed because of the \code{free} call.
-
-Enlightened by program partitioning mechanism,
-we sandbox the unchecked code regions and utilize the \checkedc type system
-to disallow checked pointers to be used in an unchecked code region. 
-To achieve the communication between checked and unchecked code regions,
-Tainted pointers can be shared by different regions,
-whose data are stored in the sandboxed unchecked heap.
-Users are required to copy checked data to tainted pointers before they are shared in unchecked regions.
+% 
+\mzu{Even if it does not crash, line 38 is doomed because of the \code{free} call.}
+% 
+\mz{                            %
+  % 
+  That's not true:
+  %
+  Line 38 assigns the address of a constant char array to \code{p[10]} which
+  could be valid since \code{p} is not yet freed.
+  % 
+}
+% 
+Enlightened by program partitioning mechanism, \mzr{we address the crashes
+  caused by spatial safety violation by sandboxing} the unchecked code regions
+and utilize the \checkedc{} type system to rule out checked pointers from any
+unchecked regions.
+%
+\mzr{ % 
+  % 
+  Besides, we incorporate a new form of pointer, \emph{tainted pointer}, into
+  our existing \checkedc{} system to allow the interaction between checked and
+  unchecked code regions.
+  %
+  A tainted pointer can be used in both checked and unchecked region; but it can
+  only point to \emph{unchecked} heap memory so that the integrity of checked
+  region would not be compromised by writes issued by unchecked code.
+  %
+  Hence to pass a piece of data from a checked region to an unchecked one using
+  a tainted pointer, instead of exposing the address directly, the user would
+  need to copy it from checked region to the unchecked heap and create a tainted
+  pointer to refer to the copy.
+% 
+}
+% 
 For example, we copy the checked pointer data to the tainted pointer \code{tp}
@@
 in the sandboxed heap, it can be recovered.
+\mzu{
+  % 
   At line 37, the use of a tainted pointer in a checked region
   requires a verification on it.
+  % 
   This is managed by inserting additional checks
   and creating exception handling before the use by the \systemname compiler.
+  % 
+}
+\mz{
+  Might as well state separately that dynamic checkes are instructmented upon
+  dereferncing a tainted pointer in checked region.  
+}
+% 
 Thus, the checked pointer \code{p} is safely used at line 38.
+%
 
-In \systemname, we prove the non-crashing theorem;
-.i.e, any well-typed \systemname program program can never crash due to spatial safety violations.
+% 
+\mz{The example itself is a bit confusing.
+  %
+  There are a few simplifications can be made here:
+  % 
+  (1) introduce a macro: \code{#define nt_ptr nt_array_ptr} to make the example
+  less verbose;
+  % 
+  (2) split untangle BAD and GOOD code into 2 fragments, listed side-by-side.
+  %
+  (3) use more descriptive function name instead of \code{f}, \code{g}
+}
+%
 
-\myparagraph{Formalism Function Pointers}
-In C, manipulating function pointers is a major way of 
-implementing high order functions as well as accessing data stored in classes.
-Previously, \checkedc assumed all function calls are called by name to a global map.
-In \systemname, we formalize function pointers and maintain the \systemname type soundness.
+% 
+In \systemname{}, we prove the non-crashing theorem: any well-typed
+\systemname{} program program can \emph{never crash} due to spatial safety
+violations, \mzr{i.e., spatial safety violation can be caught at runtime.}
+%
 
-\Cref{fig:checkedc-example-1} defines a string sorting algorithm
-depending on the input function pointer \code{cmp} that provides the generic order for strings,
-and \code{compare_1} is an example \code{cmp} function
-that adds the ASCII numbers of characters in the two strings and compare the results.
-In addition, function pointers enable the callback mechanism,
-i.e., a server sends a function pointer to a client in an unchecked region,
-and allows the client to access some server resources by calling back the pointer.
-This is a common usage between a web-browser and untrusted third-party libraries.
-The function call in \Cref{fig:checkedc-example-3} line 34 is one such usage.
+%
+\mz{``catching a SSV at runtime'' may sound more appealing and more of
+  an application?
+  %
+  Assume audience here are still general; they could prefer more application
+  over high-level ideas?
+}
+%
 
-We also utilize \systemname subtyping relation to permit function pointer static auto-casting.
-Function pointer type information might contain array pointer bound information,
-for which it is inconvenient to coincide the defined types for a function implementation and the function pointer type. 
-For example, the \code{cmp} argument in \code{stringsort} (\Cref{fig:checkedc-example-1})
-has type \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>},
-meaning that the function takes two NT-array pointers with arbitrary size and outputs an integer.
-The function \code{compare_1}'s pointer has type \code{ptr<(int)(nt_array_ptr<char> : count (0), nt_array_ptr<char> : count (0) )>}
-To use \code{compare_1} in \code{stringsort}, the type is auto-cast to the \code{cmp}'s type.
-In general, if function pointer $x$ has type $* (tl \to t)$, and  $y$ has $* (tl' \to t')$,
-in order to use $x$ as $y$, $tl'$ should be a subtype of $tl$ and $t$ subtypes to $t'$.
+% 
+\myparagraph{\mzs{Formalism}\mzr{Formalizing} Function Pointers}
+% 
+In C, manipulating function pointers is a major way of implementing high order
+functions as well as accessing data stored in structs.  In previous works,
+\checkedc{} assumed all function calls are called by name to a global map.
+\mz{Global map can be confusing.}
+% 
+In \systemname{}, we \mzr{extend the existing formalism with} \mzs{formalize}
+function pointers and prove that the type soundness \systemname{} is preserved.
+%
 
+%
+\mz{I'm trying to make it simpler. Let's rename \code{compare_1} to
+  \code{cmp_sum}. I assume the minumun knowledge of C for readers so that they
+  know sum of string is by the ASCII encoding. }
+% 
+\Cref{fig:checkedc-example-1} defines a string sorting algorithm depending on
+the input function pointer \code{cmp} \mzr{as a string comparator}
+\mzs{that provides the generic order for strings}; and,
+\mzr{                           %
+  %
+  \code{cmp_sum} is an example of \code{cmp} function that compares strings by
+  the sum of all characters in each of them.
+  % 
+} %
+\mzs{                           %
+  % 
+  \code{compare_1} is an example \code{cmp} function that adds the
+  ASCII numbers of characters in the two strings and compare the results.
+  % 
+} %
+%
+
+% 
+In addition, function pointers enable the callback mechanism, i.e., a server
+sends a function pointer to a client in an unchecked region, and allows the
+client to access some server resources by calling back the pointer \mzr{on
+  demand}.  This is a common usage between a web-browser and untrusted
+third-party libraries.  The function call in \Cref{fig:checkedc-example-3} line
+34 is one such usage.
+% 
+\mz{It sounds good.
+  % 
+  But, we might as well avoid using server/client here.
+  % 
+  They could potentially divert our motivation from memory-oriented to a network
+  related one or even sync-vs-async discussion.
+  %
+}
+%
+
+%
+\mzr{                           %
+  % 
+  Simply adding function pointer type into the existing is not enough.
+  % 
+  Continuing with the example in \Cref{fig:checkedc-example-1}, we can find out
+  that \code{cmp} argument in \code{stringsort} is not exactly the same with the
+  signature of \code{cmp_sum}: \code{cmp} has type
+  \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>} which takes stwo
+  NT-array pointers with arbitrary size and outputs an integer, but
+  \code{cmp_sum} \uline{ (Wait, what's the exact difference?) We may need to
+    make it clear here}. 
+  %
+  To make function pointers usable, we also defines a subtyping relation in
+  \systemname{} that permits function pointer subsumption.
+  %
+  In this example, the of type \code{cmp_sum} is automatically casted to the
+  type \code{cmp}. 
+  % 
+  In general, if function pointer $x$
+  has type $* (tl \to t)$, and $y$ has $* (tl' \to t')$, in order to use $x$ as
+  $y$, $tl'$ should be a subtype of $tl$ and $t$ subtypes to $t'$.
+}
+%
+
+% 
+\mzs{
+  % 
+  We also utilize \systemname{} subtyping relation to permit function pointer
+  static auto-casting.  Function pointer type information might contain array
+  pointer bound information, for which it is inconvenient to coincide the
+  defined types for a function implementation and the function pointer type.
+  For example, the \code{cmp} argument in \code{stringsort}}
+% 
+(\Cref{fig:checkedc-example-1})
+% 
+\mzs{
+  % 
+  has type
+  \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>}, meaning that the
+  function takes two NT-array pointers with arbitrary size and outputs an
+  integer.  The function pointer to {\code{compare_1}} has type
+  \code{ptr<(int)(nt_array_ptr<char> : count (0), nt_array_ptr<char> : count (0)
+    )>.}
+  % 
+}
+% 
+\mzs{                           %
+  To use {\code{compare_1}} in \code{stringsort}, the
+  type is auto-cast to the \code{cmp}'s type.  In general, if function pointer $x$
+  has type $* (tl \to t)$, and $y$ has $* (tl' \to t')$, in order to use $x$ as
+  $y$, $tl'$ should be a subtype of $tl$ and $t$ subtypes to $t'$.
+  % 
+}
+  % 
+% 
+\mz{This could make the subtyping sound more natural.}
+%
+
+% 
 \myparagraph{Not Exposing Checked Pointer Addresses}
-The non-crashing guarantee in \systemname bans the checked pointer manipulation in unchecked code regions.
-Thus, there is no reason to permit checked pointer variable assignments in unchecked regions;
-especially, this might expose a checked pointer address to untrusted parties.
-For example, the call to function $g$ at \Cref{fig:checkedc-example-3} line 41 lives in an unchecked region,
-and $g$ might use some mechanisms, such as derandomizing ASLR \cite{shacham-aslr},
-to achieve the checked pointer address.
-Thus, it enables a third party to access any checked heap and function data by simple pointer arithmetic.
+% 
+The non-crashing guarantee in \systemname{} bans the checked pointer
+manipulation in unchecked code regions. \mzu{Thus, there is no reason to permit
+checked pointer variable assignments in unchecked regions; especially, this
+might expose a checked pointer address to untrusted parties.}  For example, the
+call to function $g$ at \Cref{fig:checkedc-example-3} line 41 lives in an
+unchecked region, and $g$ might use some mechanisms, such as derandomizing ASLR
+\cite{shacham-aslr}, to achieve the checked pointer address.  Thus, it enables a
+third party to access any checked heap and function data by simple pointer
+arithmetic.
+% 
+\mz{
+  Is it rather a consequence of forbidding checked pointer manipulation?
+}
+%
 
-we prevent any unchecked regions from acknowledging checked pointer variables to avoid the checked pointer address leak.
-In addition, to facilitate checked function callbacks, 
-the \systemname compiler compiles every checked function
-with an additional tainted shell function.
-Users are required to serve unchecked regions with the tainted shell pointer instead of the original checked function pointer.
-For example, \code{tainted_compare_1} and \code{tainted_stringsort} in \Cref{fig:checkedc-example-2} 
-are the tainted shells of the checked functions \code{compare_1} and \code{stringsort}, respectively.
-In the tainted shells, the arguments are tainted versions of the corresponding arguments in the checked functions.
-Inside the shell body, we create checked pointer copies of the tainted arguments, and call the checked functions.
-In addition, once the checked function returns,
-if the output is a checked pointer, we copy its data to a new tainted pointer and return it.
-\Cref{fig:checkedc-example-3} line 42 is an example of serving the function call living in an unchecked region
-with a tainted shell pointer argument \code{tainted_compare_1}.
-Even if $g$ derandomizes its address (line 14), the shell address is in the sandbox and has no harm,
-and calling the shell never exposes any checked pointer information outside of the shell.
+% 
+We prevent any unchecked regions from acknowledging checked pointer variables to
+avoid the checked pointer address leak.  In addition, to facilitate checked
+function callbacks, the \systemname{} compiler compiles every checked function
+with an additional tainted shell function.  Users are required to serve
+unchecked regions with the tainted shell pointer instead of the original checked
+function pointer.  For example, \code{tainted_compare_1} and
+\code{tainted_stringsort} in \Cref{fig:checkedc-example-2} are the tainted
+shells of the checked functions \code{compare_1} and \code{stringsort},
+respectively.  In the tainted shells, \mzr{the types of} \mzs{arguments}
+\mzr{parameters} are the tainted \mzr{counterparts} \mzs{\versions of the
+  corresponding arguments} in the \mzr{original} checked \mzs{functions}
+\mzr{ones}.  
+% 
+Inside the shell body, \mzr{the compiler} \mzs{we} \mzu{creates checked pointer
+  copies of the tainted arguments, and call the checked functions}.  \mz{This
+  part is unclear. What do you call copy? Copying pointers doesn't make sense:
+  pointers are just memory address. It's copying data that really helps. } 
+% 
+In addition, once the checked function returns, if the output is a checked
+pointer, we copy its data to a new tainted pointer and return it.
+\Cref{fig:checkedc-example-3} line 42 is an example of serving the function call
+living in an unchecked region with a tainted shell pointer argument
+\code{tainted_compare_1}.  Even if $g$ derandomizes its address (line 14), the
+shell address is in the sandbox and has no harm, and calling the shell never
+exposes any checked pointer information outside of the shell.
+%
 
-Conceptually, the shell is run in a checked region.
-Essentially, a tainted shell is a safe closure that contains a checked block.
-Once the closure is called, the system is turned to a checked region.
-For example, \Cref{fig:checkedc-example-2} line 25 surrounds the tainted shell body with a $\echeckedtext$ block indicating
-that we context-switch from the unchecked to checked code region;
-thus, the checked function call at \Cref{fig:checkedc-example-2} line 14 is safe, 
-even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.
-In the \systemname formalism, we formalize a \textit{checked} block on top of the existing unchecked regions,
-and the transition of a tainted shell call creates a checked block containing the shell body.
-We also make sure that no arguments in these tainted shell contains any checked pointers,
-as well as no output is of a checked type.
+% 
+Conceptually, the shell is run in a checked region.  Essentially, a tainted
+shell is a safe closure that contains a checked block.  Once the closure is
+called, the system is turned to a checked region.  For example,
+\Cref{fig:checkedc-example-2} line 25 surrounds the tainted shell body with a
+$\echeckedtext$ block indicating that we context-switch from the unchecked to
+checked code region; thus, the checked function call at
+\Cref{fig:checkedc-example-2} line 14 is safe, even if it is called by $g$ in
+\Cref{fig:checkedc-example-3} , because it lives in the checked region.  In the
+\systemname formalism, we formalize a \textit{checked} block on top of the
+existing unchecked regions, and the transition of a tainted shell call creates a
+checked block containing the shell body.  We also make sure that no arguments in
+these tainted shell contains any checked pointers, as well as no output is of a
+checked type.
+%
+
+\mz{Shell vs. Wrapper? Shell reminds me of bash. }
 

### 2022-08-16 sandp23/formal.tex

@@
 \section{Formalization}\label{sec:formal}
+\liyi{This section might say too much, we might cut more here, and put more items in the compilation and implementation. }
 
@@
 language presented in Fig.~\ref{fig:checkc-syn}.
+\liyi{why can we just say in.. }
 
@@
 $\omega$ classify multi-word values such as arrays, null-terminated
-arrays, functions, and single-word-size values.
+arrays, functions, and single-word-size values. \liyi{Is this a good way? the differences are mainly for the result types vs type elements. }
 %
@@
 and a type ($\omega$) denoting valid values that can be pointed to.
+\liyi{a little too much? Should we assume people know what pointer type is?}
+\liyi{never mention $m$. }
 Array types include both the type of
@@
 otherwise (we elide $\cdot$ when writing types).
+\liyi{should have a better way. }
 \lang function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
@@
 that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
-We have a well-formed requirement for a function type,
-such that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+\liyi{need to fix, do not understand. }
+We have a well-formed requirement for a function type;
+that is, all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
 Here is the
@@
 As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
-$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
+$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the firs two examples above could
 be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
 $\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
+\liyi{we might just cut nt-array ptrs. it seems to me that nt-array ptrs are not important. }
 
@@
 
+\liyi{should we bring out the difference? since readers might not even know the previous paper.}
 Compared to the former \checkedc model \cite{li22checkedc},
@@
 for maintaining non-exposure.
-Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi'} ...)}{\xi}$, $\xi'\le \xi$.
+Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
 It is worth noting that pointer modes are a three point partial order ($\le$),
@@
 such that $\tmode$ is projected as $\umode$.
+\liyi{not clear. }
 Second, $\emalloc{\xi}{\omega}$ includes a mode flag
-$\xi$ for allocating different pointers in different heap mode regions.
+$\xi$ for allocating different pointers in different heaps.
 We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
 Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
+\liyi{might need to bring out what is the difference of previous ones.}
 Fourth, $\echeckedtext$ blocks are added to the system, 
-which permits the nested switching between $\echeckedtext$ (represented by context mode $\cmode$)
+which permits the nested context-switching between $\echeckedtext$ (represented by context mode $\cmode$)
 and $\euncheckedtext$ (represented by context mode $\umode$) code regions.
-One example usage of the nested switching is the checked function callbacks inside 
+One example usage of the nested context-switching is the checked function callbacks inside 
 an unsafe region in \Cref{fig:checkedc-example-2} and \ref{fig:checkedc-example-3}.
@@
 in \checkedc, so we omit them.
+\liyi{last sentence can be cut. }
 
@@
 transition relation with the judgment $ (\varphi,\heap,e)
-\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
+\longrightarrow_m (\varphi',\heap',r)$.
+\liyi{do we need to say judgment here? }
+ Here, $\varphi$ is a
 \emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
-$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ regions), each of which
+$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ heaps), each of which
 maps addresses (integer literals) to values $\evalue{n}{\tau}$.
-A $\cmode$pointer is mapped to a heap location in the $\cmode$ region, 
-while a $\tmode$ and $\umode$ pointer represents a $\umode$ region location.
-We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ region,
+A $\cmode$ pointer is mapped to a heap location in the $\cmode$ heap, 
+while a $\tmode$ and $\umode$ pointer represents a $\umode$ heap location.
+We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ heap,
 and $\heapup{m}{n}{\evalue{n'}{\tau}}$ 
-to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ region.
+to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ heap.
 It is worth noting that \systemname is not a fat-pointer system;
@@
 variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
-be updated. We can model mutable stack variables as pointers into the
+be updated. 
+We can model mutable stack variables as pointers into the
 mutable heap.
+\liyi{why we need this sentence? }
 As mentioned, value $\evalue{0}{\tau}$
@@
 The second rule describes the exception handling 
-for possible crashing behaviors in unchecked region.
+for possible crashing behaviors in unchecked regions.
 A $\umode$ mode operation can non-deterministically crash
@@
 and continues with the existing program state.
+\liyi{is non-deterministically crashed or non-determinstically transitioned to some place. }
 Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
@@
 There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
+\liyi{might not need this. }
 
@@
 describes the context mode determination in each evaluation step based on the context $E$.
+\liyi{understood? }
 For any program execution, the function starts the mode computation with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
@@
 the function recursively traverses the sub-context to find the context mode.
+\liyi{may be too low-level detailed. describe it in a high level. }
 
@@
 \textsc{S-DefNull}, and \textsc{S-Cast}.
-The first five define deference and assignment operations and illustrate how the semantics checks bounds.
+The first five define the semantics of deference and assignment operations.
 Rule \textsc{S-DefNull} transitions attempted null-pointer
@@
 \textsc{S-DefT} and \textsc{S-AssignArrT} are similar rules to \textsc{S-DefC} and \textsc{S-AssignArrC} for tainted pointers.
-Any dynamic heap use of a tainted pointer requires a verification.
-Performing such a verification equates to performing a type check for a pointer constant in \Cref{fig:const-type}.
+Any dynamic heap access of a tainted pointer requires a \textit{verification}.
+Performing such a verification equates to performing a literal type check for a pointer constant in \Cref{fig:const-type}.
 We explain this shortly in \Cref{sec:type-system}.
 For now, the verification step, e.g. $\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$ in \textsc{S-DefC},
-means we verify that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.
+refers to that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.
 
@@
 
+\liyi{Yeah, the block semantics is easy, but is there anything to say? I mean this is the essentially item.}
+
 \myparagraph{Binding and Function Calls}
@@
 in $e$'s scope that is defined by the $\erettext$ operation.
+\liyi{understood? }
 Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
@@
 Similar to \heap, the global function store $\Xi$ is also partitioned into
-two parts ($\cmode$ and $\umode$ regions), each of which
+two parts ($\cmode$ and $\umode$ stores), each of which
 maps addresses (integer literals) to the function data described above.
@@
 $\overline{n}$, but annotated with the parameter types
-$\overline{\tau}$ (this will be safe for type-correct programs). The
-function body $e$ is wrapped in a static cast
+$\overline{\tau}$ (this will be safe for type-correct programs). 
+The function body $e$ is wrapped in a static cast
 $(\tau[\overline{n} / \overline{x}])$ which is the function's return
@@
 argument list.
+\liyi{This has appeared in the previous paper. A better description or more items? }
 Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
@@
 Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
+\liyi{might be cut because no more description on this. }
 Rule \textsc{S-FunT} defines the tainted version of function call semantics.
@@
 $\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
-are equivalent, if we can find a same length, as $\overline{x}$ and $\overline{y}$, variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
+are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
 and the substitution results are equal.
@@
  users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
+\liyi{why? }
 
@@
 system uses $\Theta$ to make such reasoning more precise.
+\liyi{why?}
 $\Theta$ is a map from variables $x$ to
@@
 describes type assumptions for constants appearing in a program.
-$\cmode(\tau)$ judges a constant pointer 
+$\cmode(\tau)$ judges that a constant pointer 
 in an unchecked region cannot be of a checked type,
@@
 which is similar to the dynamic verification process in \Cref{sec:semantics}. 
-
-Given a constant $\evalue{n}{\tau}$, 
-the verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
-if the constant is valid, 
+\liyi{what dynamic verification process? }
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
+validate the constant $\evalue{n}{\tau}$, 
 where $\heap(m)$ is the initial heap that the constant resides on and
@@
 A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
+\liyi{Do we discuss this $\Xi(m)$ before? }
 A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
 and the address stores a function with the right type.
+\liyi{how? }
 The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
 where every element in the pointer range ($[0,\size(\omega))$) should be well typed.
+\liyi{only well typed? Can we have a better way to say this? }
 
@@
 Tainted pointers are verified through the validity check in dynamic execution as we mentioned above.
+\liyi{This is saying that the validity checks and verification are the same, but the word ''verification'' could be misleading, so maybe change it? }
+
+\liyi{I think unchecked/checked blocks and function calls below can be merged, 
+   because this section is small and might not describe the needed items. }
 
@@
 the types for $\overline{x}$ and the final return type $\tau$ have no checked pointers.
+\liyi{why no? This should be emphasized more. }
 A $\echeckedtext$ or $\euncheckedtext$ block represents 
@@
 We need to make sure no checked pointers are information exposed to unsafe code regions.
+\liyi{example? }
 
@@
 Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
-type dependency. In particular, the result of evaluating a $\elettext$
+type dependency. 
+\liyi{The type dependency means the scope of type variables. Might not need, because we have said this in previous paper. }
+In particular, the result of evaluating a $\elettext$
 may have a type that refers to one of its bound variables (e.g., if
@@
 proof.
+\liyi{why? }
 % After the evaluation of a let binding a variable $x$ concludes,
@@
 also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
+\liyi{should keep it uniformly with respect to the one above. 
+    It is also not so clear and too much wording here. }
 Consider the \code{safe_strcat} function in
@@
 
+\liyi{Can we cut some?}
 Type soundness relies on several \emph{well-formedness}:
@@
 snapshots together.
+\liyi{Tell people why only checked heap/store matter? }
 
@@
 
+\liyi{maybe some intuition why this is important. }
 Using the above theorem, we first show the non-exposure theorem,

### 2022-08-16 sandp23/background.tex

@@
 %
+\mz{I changed a lot in this paragraph. Please compare with the previous commit.}
+% 
 Enlightened by program partitioning mechanism, \mzr{we address the crashes
@@
 respectively.  In the tainted shells, \mzr{the types of} \mzs{arguments}
-\mzr{parameters} are the tainted \mzr{counterparts} \mzs{\versions of the
+\mzr{parameters} are the tainted \mzr{counterparts} \mzs{versions of the
   corresponding arguments} in the \mzr{original} checked \mzs{functions}
@@
 In addition, once the checked function returns, if the output is a checked
-pointer, we copy its data to a new tainted pointer and return it.
-\Cref{fig:checkedc-example-3} line 42 is an example of serving the function call
-living in an unchecked region with a tainted shell pointer argument
-\code{tainted_compare_1}.  Even if $g$ derandomizes its address (line 14), the
-shell address is in the sandbox and has no harm, and calling the shell never
-exposes any checked pointer information outside of the shell.
+pointer, we copy its data to a new tainted pointer and return it \mzr{so that
+  the tainted caller only observes tainted pointer that points to data in
+  unchecked heap}.  \Cref{fig:checkedc-example-3} line 42 is an example of
+\mzr{passing a tainted shell pointer as an argument to the function issued from
+  an unchecked region} \mzs{serving the function call living in an unchecked
+  region with a tainted shell pointer argument \code{tainted_compare_1}}.  Even
+if $g$ derandomizes its address (line 14), the shell address \mzs{is} in the
+sandbox \mzs{and} has no harm \mzs{, and} \mzr{because} calling the shell never
+\mzr{leaks} \mzs{exposes} any checked pointer information outside of the shell.
 %
@@
 shell is a safe closure that contains a checked block.  Once the closure is
-called, the system is turned to a checked region.  For example,
-\Cref{fig:checkedc-example-2} line 25 surrounds the tainted shell body with a
-$\echeckedtext$ block indicating that we context-switch from the unchecked to
-checked code region; thus, the checked function call at
+called, the system \mzr{switches into the checked mode} \mzs{is turned to a
+  checked region}.  For example, \Cref{fig:checkedc-example-2} line 25 surrounds
+the tainted shell body with a $\echeckedtext$ block indicating that we
+\mzu{context-switch} \mz{potential term abuse; ``switch'' would suffice} from
+the unchecked to checked code region; thus, the checked function call at
 \Cref{fig:checkedc-example-2} line 14 is safe, even if it is called by $g$ in
-\Cref{fig:checkedc-example-3} , because it lives in the checked region.  In the
-\systemname formalism, we formalize a \textit{checked} block on top of the
+\Cref{fig:checkedc-example-3}, because it lives in the checked region.  In the
+\systemname{} formalism, we formalize a \textit{checked} block on top of the
 existing unchecked regions, and the transition of a tainted shell call creates a
-checked block containing the shell body.  We also make sure that no arguments in
-these tainted shell contains any checked pointers, as well as no output is of a
-checked type.
+checked block containing the shell body.  We also make sure that no
+arguments\mzr{, as well as the output, }in \mzr{any} \mzs{these} tainted shell
+contains any checked pointers\mzs{, as well as no output is of a checked type}.
 %

### 2022-08-16 sandp23/introduction.tex

--- 2022-08-09 sandp23/introduction.tex
@@
 % \end{itemize}
-\noindent
-\textbf{No safety against unchecked code.} The spatial safety guarantee~\checkedc provides applies to completely converted programs, i.e., programs that uses only checked types.
-Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
-But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
-Although the blameless proof exists~\cite{ruef2019achieving} for~\checkedc, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
-Consider the following example:
-\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
-// Checked code
-int func(array_ptr<char> p : count(5)) {
-|\textcolor{red}{\faChainBroken}|..p[4]..
+
+The guarantee \checkedc provides focuses on checked code regions,
+where any stuck (i.e., ill-defined)
+state reached by a well-typed program amounts to a spatial safety
+violation.
+Such a state can always be attributed to, i.e., 
+\emph{blamed on}, the execution of code that is not in a checked core region. 
+However, what are the guarantees that unchecked (unsafe) code regions can provide?
+Programmers might not expect to convert every code to \checkedc,
+but leave some functions in unsafe regions,
+in which case the unsafe region executions should not affect other parts.
+In this paper, we cover this gap by introducing \systemname,
+combining \checkedc with program partitioning mechanism, 
+making three main contributions.
+%
+
+\myparagraph{Executions Cannot Crash}
+\mz{This paragraph sounds incoherent to me.}
+% 
+Our first contribution is a clear program partition between checked and
+unchecked code regions \mzs{and ensures} \mzr{so} that a \systemname program execution
+\textit{cannot crash} due to spatial safety violations.
+\mzr{
+  % 
+  In \systemname, heap is partitioned into two parts: a \textit{checked heap}
+  containing all checked pointers and an \textit{unchecked heap} that is
+  sandboxed.
+  % 
+  Then, we restricted pointer usages so that checked pointers can only refer to
+  checked heap region and unchecked ones can only be used in unchecked heap
+  region.
+  %
+  To communicate safely between those two isolated regions, we developed tainted
+  pointers that is always verified upon accessing data and program in unchecked
+  region.
+  % 
 }
-// unchecked code
-...
-str = "he";
-...
-|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5); 
-...
-char ptr[16];
-...
-len <- derived from user input
-...
-|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
-\end{minted}
-Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function with a buffer of 2 elements.
-This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
-We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
-This mechanism is called program partitioning~\cite{rul2009towards}.
-Existing mechanisms are not suitable as they are based on process isolation and have high overhead, and are data-centric~\emph{data-centric} (\sect{subsec:background:programpart}).
-Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
-But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
-We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
-
-In this work, we propose a type-directed code-centric program partitioning approach.
-Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
-The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.
-
-The developer starts by marking desired (i.e., unchecked) functions and pointers used in these functions as tainted.
-Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
-\begin{itemize}
-\item~\ucregion or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (i.e., functions and pointers) and can only access tainted pointers.
-\item~\cregion or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
-The functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
-\end{itemize}
-Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (e.g., WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
-
-The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
-As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
-We avoid marshaling by allocating all tainted pointers (i.e., tainted buffers) in~\ucregion and thus can be accessed in both partitions.
-Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
-Our compiler avoids these attacks by ensuring, either statically or through dynamic checks that tainted pointers can only point to~\ucregion address space.
-
-In summary, we make the following three main contributions.
-
-\myparagraph{Type-Directed Program Partitioning}
-Our first contribution is a clear program partition between checked and unchecked code regions
-and ensures that a \systemname program execution \textit{cannot crash} due to spatial safety violations.
-We restrict the checked and unchecked pointers usage only in checked and unchecked code regions, respectively;
-and develop tainted pointers for the communication between the two regions.
-Heap is partitioned into two parts: a \textit{checked heap} containing all checked pointers
- and an \textit{unchecked heap} that is sandboxed.
-Tainted pointers exist in the unchecked heap region and must be verified each time being used.
+\mzs{ 
+  % 
+  We restrict the checked and unchecked pointers usage only in checked and
+  unchecked code regions, respectively; and develop tainted pointers for the
+  communication between the two regions.  Heap is partitioned into two parts: a
+  \textit{checked heap} containing all checked pointers and an \textit{unchecked
+    heap} that is sandboxed.  Tainted pointers exist in the unchecked heap
+  region and must be verified each time being used.
+  % 
+}
 
@@
 In this case, no checked pointer address will be observed in the unchecked code regions.
-In \systemname, for every checked function, 
-we automatically compile a tainted version by surrounding the function without a tainted shell.
+% 
+\mzu{In \systemname, for every checked function, 
+we automatically compile a tainted version by surrounding the function without a
+tainted shell.}
+\mz{What does this mean?}
+% 
 To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
 
-We begin with a review of \checkedc (Section~\ref{sec:overview}) and introduce new features in \systemname,
-present our main contributions
+We begin with a review of \checkedc{} \mzr{and introduction of} new features in
+\systemname (Section~\ref{sec:overview}), present our main contributions
 (Sections~\ref{sec:formal}--\ref{sec:evaluation}), and conclude with a
-discussion of 
-related and future work (Sections~\ref{sec:related},
-\ref{sec:conclude}). All code and proof artifacts (both for Coq and
-Redex) can be found at \url{https://github.com/plum-umd/checkedc}.
+discussion of related and future work (Sections~\ref{sec:related},
+\ref{sec:conclude}). All code and proof artifacts (both for Coq and Redex) can
+be found at \url{https://github.com/plum-umd/checkedc}.
 

### 2022-08-16 sandp23/background.tex
fix typo

@@
 \mz{
-  Might as well state separately that dynamic checkes are instructmented upon
+  Might as well state separately that dynamic checks are instructmented upon
   dereferncing a tainted pointer in checked region.  
@@
   % 
-  Continuing with the example in \Cref{fig:checkedc-example-1}, we can find out
-  that \code{cmp} argument in \code{stringsort} is not exactly the same with the
-  signature of \code{cmp_sum}: \code{cmp} has type
-  \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>} which takes stwo
+  Continuing with the example in \Cref{fig:checkedc-example-1}, we could notice
+  that the type of the argument \code{cmp} in \code{stringsort} is not exactly
+  the same with the signature of \code{cmp_sum}: \code{cmp} has type
+  \code{ptr<(int)(nt_array_ptr<char>, nt_array_ptr<char>)>} which takes two
   NT-array pointers with arbitrary size and outputs an integer, but
-  \code{cmp_sum} \uline{ (Wait, what's the exact difference?) We may need to
+  \code{cmp_sum} \ldots \uline{ (Wait, what's the exact difference?) We may need to
     make it clear here}.
   %
-  To make function pointers usable, we also defines a subtyping relation in
+  To make function pointers usable, we also define a subtyping relation in
   \systemname{} that permits function pointer subsumption.
   %
-  In this example, the of type \code{cmp_sum} is automatically casted to the
-  type \code{cmp}. 
+  In this example, the type of \code{cmp_sum} is automatically casted to the
+  type of \code{cmp}. 
   % 
@@
 shells of the checked functions \code{compare_1} and \code{stringsort},
-respectively.  In the tainted shells, \mzr{the types of} \mzs{arguments}
-\mzr{parameters} are the tainted \mzr{counterparts} \mzs{versions of the
-  corresponding arguments} in the \mzr{original} checked \mzs{functions}
-\mzr{ones}.  
+respectively.  In the tainted shells, \mzr{the parameter types} \mzs{arguments}
+are the tainted \mzr{counterparts of} \mzs{versions of the
+  corresponding arguments in} the \mzr{original} checked functions.
 % 
@@
   copies of the tainted arguments, and call the checked functions}.  \mz{This
-  part is unclear. What do you call copy? Copying pointers doesn't make sense:
-  pointers are just memory address. It's copying data that really helps. } 
+  part is unclear. What are you trying to copy? Copying pointers doesn't make
+  differences: pointers are just memory address. It's copying data that really
+  helps. }
 % 
@@
 pointer, we copy its data to a new tainted pointer and return it \mzr{so that
-  the tainted caller only observes tainted pointer that points to data in
-  unchecked heap}.  \Cref{fig:checkedc-example-3} line 42 is an example of
-\mzr{passing a tainted shell pointer as an argument to the function issued from
-  an unchecked region} \mzs{serving the function call living in an unchecked
-  region with a tainted shell pointer argument \code{tainted_compare_1}}.  Even
-if $g$ derandomizes its address (line 14), the shell address \mzs{is} in the
-sandbox \mzs{and} has no harm \mzs{, and} \mzr{because} calling the shell never
-\mzr{leaks} \mzs{exposes} any checked pointer information outside of the shell.
+  the tainted caller only observes a tainted pointer that points to data in
+  unchecked heap. This prevents checked pointer address exposure.}
+  \Cref{fig:checkedc-example-3} line 42 is an example of \mzr{passing a tainted
+    shell pointer as an argument to the function issued from an unchecked
+    region} \mzs{serving the function call living in an unchecked region with a
+    tainted shell pointer argument \code{tainted_compare_1}}.  Even if $g$
+  derandomizes its address (line 14), the shell address \mzs{is} in the sandbox
+  \mzs{and} has no harm \mzs{, and} \mzr{because} calling the shell never
+  \mzr{leaks} \mzs{exposes} any checked pointer information outside of the
+  shell.
 %

### 2022-08-16 sandp23/compilation.tex
review compilation

@@
 The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
-for verifying pointers being not null and the bounds are within their limits.
+for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
 The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.
@@
 i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
+\lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
 Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
-In the compiled LLVM code, pointer access operations have different syntax when the modes are different.
-\Cref{fig:flagtable} lists the different compiled syntax for a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
+In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different.
+\lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
+\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
 The columns represent different pointer modes and the rows represent context modes.
@@
 In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
-$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$.
+$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
 The algorithm sheds
@@
   lower-bound shadow variable is unnecessary; we include it for uniformity.} 
+\lc{These two sentences are confusing. }
+
 % When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
@@
 Due to space constraints, we explain the rules for compilation by
-example, using a C-like syntax; the complete rules are given in
+examples, using a C-like syntax; the complete rules are given in
 \iftr
@@
 %
-In every declaration (or the beginning of a function body) of a pointer,
-if the poniter is an (NT-)array one,
-we first allocates two \emph{shadow variables}
-to track the lower and upper bounds potentially changed for pointer arithmetic and NT-array bound widening.
+In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
+if the poniter is an (NT-)array,
+we first allocate two \emph{shadow variables}
+to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
 Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.

### 2022-08-22 sandp23/formal.tex
remark on formalism

@@
 \section{Formalization}\label{sec:formal}
-\liyi{This section might say too much, we might cut more here, and put more items in the compilation and implementation. }
+\liyi{This section might say too much, we might cut more here, and put more
+  items in the compilation and implementation. }
+
+\mz{                            %
+  I am wondering if it's still a good idea to talk about syntax, semantics
+  and types separately?
+  %
+  The key of the work is about tainted pointer, functions and blocks instead of
+  the overview of the entire system.
+  %
+  Tearing those topics apart into 3 aspects can lead to incoherence.
+  %
+}
 
@@
 \liyi{should have a better way. }
-\lang function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
+%
+
+\mz{Let's zero in on the function pointer and the mode context formalism.
+We can say something like ``bounds and array ptrs have been studied in [...]''
+already.
+}
+
+% 
+\lang{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
 reflect its dependent function declarations,
 where $\overline{x}$ represents
-a list of \tint type variables in a dependent function header
+a list of \tint{} type variables in a dependent function header
 that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
@@
 
+\mz{
+  I would say that let's just focus on uncheckd and checked.
+  %
+  We can make an simple example that contains a function pointer with dependent
+  parameters; then, explain in great detail what that means.
+  % 
+  Instead of rephrasing the meaning of symbols, we might as well direct the reader
+  to the appendix/previous work for tech details on the last paper. 
+}
+
 \liyi{should we bring out the difference? since readers might not even know the previous paper.}
@@
 
+\mz{
+  I don't see the thesis of these first a few paragraphs. 
+  %
+  And, what's the motivation of mode? How on earth this setup solves the
+  problem?
+  %
+  If we're explaining reduction in context, shall we delay the introduction of
+  this context mode now?
+}
+
 % For a term $e$, $C$ is a context if any
@@
 \liyi{Yeah, the block semantics is easy, but is there anything to say? I mean this is the essentially item.}
+\mz{
+  Explain how context mode works here. 
+}
 
@@
  the function definition in $n$'s location in the global function store $\Xi$,
+ %
+
+ \mz{
+   The ``global function store'' is what's missing in the background section.  
+ }
+
+% 
 % LEO: This is the first occurence of \Xi. It is explained,
@@
 function body. 
+\mz{
+  ``Function data'' is ``function definition'' right?
+  %
+  This sentence is verbose and contains repetitous words? 
+}
+
 Similar to \heap, the global function store $\Xi$ is also partitioned into
@@
 
-The \systemname functions are dependent functions.
+The \systemname{} functions are dependent functions.
 Recall that array
@@
 \liyi{This has appeared in the previous paper. A better description or more items? }
+
 Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
@@
 such check is deferred to the semantics.
-
+\mz{
+  What about tainted pointers?
+}
 % Subtyping and casting operations are briefly introduced in
@@
 and the alpha equivalence of two function types in \Cref{fig:checkc-subtype}.
+\mz{
+  What's ``type construct equivalent relation''?
+}
 Two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
@@
 ($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
-rules ensure the relation is transitive. Most of the rules manage
+rules ensure the relation is transitive. \mz{?} Most of the rules manage
 casting between array pointer types. The second rule 
@@
 \liyi{why? }
+
+\mz{
+  What changes are maded to bound equality compared to CheckedC?
+}
 
@@
 cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+\mz{Or simply a checked pointer is not permitted in an unchecked region? }
 In rule \textsc{T-ConstC}, we requires a static 
@@
 The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
-where every element in the pointer range ($[0,\size(\omega))$) should be well typed.
+where every element in the pointer range ($[0,\size(\omega))$) should be well
+typed.
+\mz{
+  Name those judgements, instead of using ``last rule''.
+  If \(\mathcal{H}(m, n + i)\) refers to the location on \(m\) mode heap,
+  I would prefer \(\mathcal{H}_m\).
+}
 \liyi{only well typed? Can we have a better way to say this? }
+\mz{Is there a typo in the type rule? Is that supposed to be \(n : \tau\) rather
+than \(n : \tau : \tau\)?}
 
@@
 \liyi{This is saying that the validity checks and verification are the same, but the word ''verification'' could be misleading, so maybe change it? }
+\mz{Validation instead? 
+  %
+  If for this entire section you want to emphasize what constant pointers are
+  allowed in some specific regions, validation might be much appropriate.
+  % 
+  It might also be worth stated at the beginning that you have two kinds of
+  validation procedures: static one for checked, and instrumented dynamic checks
+  for unchecked region.  }
 
@@
 \liyi{example? }
+\mz{
+  Giving a counterexample is sometimes more convincing.
+}
 
@@
 
+\mz{
+  Audience might be interested in how the save/load operation in tainted
+  pointers faciliates the interaction between checked and unchecked pointer.
+  %
+  However, I'm wondering if there's any point in the proof have something to do
+  with this idea.
+  %
+  If there is, it's worth talking about than listing well-formedness and
+  consistency.
+  %
+  Also, state the exception mechanism we are using in representing 
+  results in reduction. 
+}
 % Before we present our main theorems, we need to first
@@
 \end{thm}
+\mz{
+  It's in the footnote. But, we should explain the definition of ``stuck'' in the
+  text. It's not crystal clear.
+}
+
 %{\em Proof:} By induction on the number of steps of the \checkedc

### 2022-11-22 sandp23/formal.tex

@@
 cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
-\mz{Or simply a checked pointer is not permitted in an unchecked region? }
+\mz{Or, simply, a checked pointer is not permitted in an unchecked region? }
 In rule \textsc{T-ConstC}, we requires a static 
@@
   %
-  If for this entire section you want to emphasize what constant pointers are
-  allowed in some specific regions, validation might be much appropriate.
-  % 
-  It might also be worth stated at the beginning that you have two kinds of
+  If what you want to emphasize is what kind of constant pointers are allowed in
+  some specific regions, validation might be more appropriate in these paragraphs.
+  % 
+  It might also be worth stateing at the beginning that you have two kinds of
   validation procedures: static one for checked, and instrumented dynamic checks

### 2022-11-22 sandp23/background.tex
Make file modifications

@@
+\section{Background}
+\subsection{Program Partitioning}
+\label{subsec:background:programpart}
+
 \section{Overview and Transcendence}\label{sec:overview}
@@
 % 
-\mz{The example itself is a bit confusing.
+%\mz{The example itself is a bit confusing.
   %
-  There are a few simplifications can be made here:
+%  There are a few simplifications can be made here:
   % 
-  (1) introduce a macro: \code{#define nt_ptr nt_array_ptr} to make the example
-  less verbose;
+%  (1) introduce a macro: \code{#define nt_ptr nt_array_ptr} to make the example
+%  less verbose;
   % 
-  (2) split untangle BAD and GOOD code into 2 fragments, listed side-by-side.
+%  (2) split untangle BAD and GOOD code into 2 fragments, listed side-by-side.
   %
-  (3) use more descriptive function name instead of \code{f}, \code{g}
-}
+%  (3) use more descriptive function name instead of \code{f}, \code{g}
+%}
 %

### 2022-11-22 sandp23/overview.tex
Modifying

\section{Overview}
\label{sec:overview}
We need to explain the overall flow of our system here.

1) Start with a simple example and say what steps developer follows to convert the code to RLBoxed Checked C

2) Say what our system does with the converted code

### 2022-11-28 sandp23/overview.tex
Adding overview

@@
 \label{sec:overview}
-We need to explain the overall flow of our system here.
 
-1) Start with a simple example and say what steps developer follows to convert the code to RLBoxed Checked C
+\begin{listing}[t!]
+  \begin{tabular}{c c}
+    \begin{minipage}[b]{.22\textwidth}
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/orig1.c}
+    \end{minipage} &
+    \begin{minipage}[b]{.22\textwidth}
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/orig2.c}
+    \end{minipage} %\\
+   %(a) Original C code & (b) After initial conversion.\\
+  \end{tabular}
+\caption{C program snippet of a simple network server with an arbitrary memory write vulnerability indicated by~\realbug.}
+\label{lst:comb}
+\end{listing}
 
-2) Say what our system does with the converted code
+\begin{listing}[t!]
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/firstrun.c}
+\caption{Pointers in~\lst{lst:comb} annotated (manually or through automated tools like~\threec) with Checked Types.}
+\label{lst:initialconv}
+\end{listing}
+
+\begin{listing}[t!]
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/humanannotations.c}
+\caption{Initial annotations of tainted functions in~\lst{lst:comb}.}
+\label{lst:humantaint}
+\end{listing}
+
+\begin{listing}[t!]
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/humanadjustments.c}
+\caption{Additional tainted pointers annotations to~\lst{lst:comb} according to typing rules.}
+\label{lst:humanadjust}
+\end{listing}
+
+\begin{listing}[t!]
+  \begin{tabular}{c c}
+    \begin{minipage}[b]{.22\textwidth}
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/conv1.c}
+    \end{minipage} &
+    \begin{minipage}[b]{.22\textwidth}
+\inputminted[mathescape, highlightlines={1-19}, highlightcolor=taintcolor, escapeinside=||, fontsize=\tiny{}]{c}{examples/conv2.c}
+    \end{minipage} %\\
+   %(a) Original C code & (b) After initial conversion.\\
+  \end{tabular}
+\caption{Final annotated program of~\lst{lst:comb} with~\textcolor{taintcolor}{tainted} and~\textcolor{checkcolor}{checked} types.
+The~\colorbox{taintcolor}{highlighted} functions will be executed in a sandbox.}
+\label{lst:final}
+\end{listing}
+
+\input{figures/overview}
+
+We use the example of a simple network server in~\lst{lst:comb} to explain the overview of developers using~\systemname{}.
+The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
+The server runs in a loop and calls~\code{handle_request}, which handles each network request.
+The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} will be called to handle the request.
+As we can see, there is an arbitrary memory write vulnerability (indicated by~\realbug) in~\code{process_req1} because of using~\code{i} as an index into the array~\code{msg} without any sanity check.
+The variable~\code{i} can take any integer value as it is parsed from~\code{msg}, whose contents are read from socket in~\code{read_msg} indicated by~\rootcause. 
+The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the overall program.
+Ideally, the developer must convert the entire code to Checked C so that the entire program achieves spatial memory safety.
+However, Checked C has restrictions that can result in a considerable conversion effort.
+For instance,~\code{void*} pointers are not directly supported by Checked C.
+Consequently, the developer needs to convert functions using~\code{void*} pointers into generic versions -- this could be tedious.
+To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
+However, as explained in~\sect{}, the unchecked code can also affect the safety of checked pointers.
+To handle this, the developer marks risky functions with unchecked pointers~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded.
+The developer starts using~\oursystem with the resulting partially annotated program with checked and tainted types~\ie~\lst{lst:initialconv} and~\ref{lst:humantaint} atop of~\lst{lst:comb}.
+The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (CITE).
+The developer uses our type checker to identify the additional annotations required to have a well-typed program and adds them as shown in~\lst{}.
+The resulting well-typed program as shown in~\lst{} will be passed to our source level program partitioner along with certain configuration parameters of the target sandbox.
+Our partitioner splits the provided program into two sets of source files:
+with the necessary source changes required to communiate with sandboxed code.
+These sets of source files will be compiled with the corresponding compilers to get the corresponding object files.
+The c-region object file will have the necessary runtime checks enforcing~\systemname guarantees.
+The uc-region object file will be produced according to the corresponding sandbox mechanism.
+Finally, these two object files will be linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions will be executed in a sandbox (uc-region) and the rest of the functions as regular code (c-region).

### 2022-11-28 sandp23/examples/conv1.c

@@ -1,0 +1,35 @@
+|\entrypoint| // Entry point
+void server_loop(int sock_fd) {
+ unsigned b_z;
+ |\textcolor{checkcolor}{\_Array\_ptr}|<struct queue> p : 
+              count(b_z) = NULL;
+ ...
+ while(1) {
+  ...
+  p = malloc(sizeof(struct queue) * b_z);
+  ...
+  if (handle_request(sock_fd)) {
+    ...  
+  }
+ }
+}
+
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_\_tainted}|;
+ int rc = -1;
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, 
+                  MAX_MSG_SIZE);
+ if (r_len > 0) {
+  switch(buff[0]) {
+   case REQ1: 
+    rc = process_req1(buff); 
+    break;
+   case REQ2: 
+    rc = process_req2(buff); 
+    break;
+   ...
+  }
+ }
+ return rc;
+}

### 2022-11-28 sandp23/examples/conv2.c

@@ -1,0 +1,37 @@
+size_t read_msg(int sock_fd, char *msg, 
+                size_t sz) __tainted {
+ size_t nr;
+ nr = read(sock_fd, (void*)msg, sz);
+ ...
+ |\complexcode| // complex code
+ return nr;
+}
+
+int process_req1(char *msg, size_t m_l) 
+                 __tainted {
+ int rc = -1, i;
+ if (m_l > MIN_SIZE) {
+   sscanf(msg, "%d", &i);
+   msg[i] = ...
+ }
+ // complex code
+ return rc;
+}
+
+int process_req2(|\textcolor{taintcolor}{\_t\_Array\_ptr}|<char> msg : 
+                 count(m_l), 
+                 size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum |\textcolor{checkcolor}{\_Checked}|[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}

### 2022-11-28 sandp23/examples/firstrun.c

@@ -1,0 +1,13 @@
+|\entrypoint| // Entry point
+void server_loop(int sock_fd) {
+ ...
+ |\textcolor{checkcolor}{\_Array\_ptr}|<struct queue> p : count(b_z) = NULL;
+ ...
+}
+
+int process_req2(_Array_ptr<char> msg: count(m_l), 
+                 size_t m_l) {
+ ...
+ char anum |\textcolor{checkcolor}{\_Checked}|[MAX_MSG_SIZE] = {0};
+ ...
+}

### 2022-11-28 sandp23/examples/humanannotations.c

@@ -1,0 +1,8 @@
+size_t read_msg(int sock_fd, char *msg, 
+                size_t sz) |\textcolor{taintcolor}{\_\_tainted}| |\useradded| {
+ ...
+}
+
+int process_req1(char *msg, size_t m_l) |\textcolor{taintcolor}{\_\_tainted}| |\useradded| {
+ ...
+}

### 2022-11-28 sandp23/examples/orig2.c

@@ -1,0 +1,34 @@
+size_t read_msg(int sock_fd, char *msg, 
+                size_t sz) {
+ size_t nr;
+ nr = read(sock_fd, |\rootcause|(void*)msg, sz);
+ ...
+ |\complexcode| // complex code
+ return nr;
+}
+
+int process_req1(char *msg, size_t m_l) {
+ int rc = -1, i;
+ if (m_l > MIN_SIZE) {
+   sscanf(msg, "%d", &i);
+   |\realbug| msg[i] = ...
+ }
+ |\complexcode| // complex code
+ return rc;
+}
+
+int process_req2(char *msg, size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}

### 2022-11-28 sandp23/examples/orig3.c

@@ -1,0 +1,15 @@
+int process_req2(char *msg, size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}

### 2022-11-28 sandp23/examples/originalprogram.c

@@ -1,0 +1,68 @@
+size_t read_msg(int sock_fd, char *msg, size_t sz) {
+ size_t nr;
+ nr = read(sock_fd, (void*)msg, sz);
+ ...
+ // complex code
+ return nr;
+}
+
+int process_req1(char *msg, size_t m_l) {
+ int rc = -1;
+ //complex code
+ if (m_l > MIN_SIZE) {
+   //complex code
+   sscanf(msg, "%d", &i);
+   msg[i] = ...
+ }
+ //complex code
+ return rc;
+}
+
+int process_req2(char *msg, size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}
+
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE];
+ int rc = -1;
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, MAX_MSG_SIZE);
+ if (r_len > 0) {
+  switch(buff[0]) {
+   case REQ1: 
+    rc = process_req1(buff); 
+    break;
+   case REQ2: 
+    rc = process_req2(buff); 
+    break;
+   ...
+  }
+ }
+ return rc;
+}
+
+void server_loop(int sock_fd) {
+ unsigned b_z;
+ struct queue *q;
+ ...
+ while(1) {
+  ...
+  p = malloc(sizeof(struct queue) * b_z);
+  ...
+  if (handle_request(sock_fd)) {
+    ...  
+  }
+ }
+}

### 2022-11-28 sandp23/introduction.tex
Adding overview
More refactoring on intro and overview

@@
 \begin{itemize}
-\item~\ucregion or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (i.e., functions and pointers) and can only access tainted pointers.
+\item~\ucregion or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (\ie functions and pointers) and can only access tainted pointers.
 \item~\cregion or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
-The functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
+The functions in~\cregion can invoke any function in~\ucregion and access all its data but not the other way around, except for call-back functions, which we will discuss later.
 \end{itemize}
-Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (e.g., WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
+Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
 
@@
 As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
-We avoid marshaling by allocating all tainted pointers (i.e., tainted buffers) in~\ucregion and thus can be accessed in both partitions.
+We avoid marshaling by allocating all tainted pointers (\ie tainted buffers) in~\ucregion and thus can be accessed in both partitions.
 Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
@@
 We present a type-directed program partition technique to separate checked and unchecked code regions and ensure that the code in checked region~\textit{cannot crash} due to spatial safety violations.
-Our type system restricts checked and tainted (or unchecked) pointers usage only in~\cregion and~\ucregion, respectively.
+Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted (or unchecked) pointers.
 Our modular design enables us to use existing sandboxed techniques to enforce memory and execution isolation.
-
-\myparagraph{Formalizing the Type System, Semantics and Compiler}
-%
-We developed a core formalism named~\lang, which extends
-\citet{li22checkedc}~\checkedc formalism with the non-crashing guarantee and other new features below.
-We prove formally the~\emph{non-crashing theorem}, i.e., 
-a well-typed~\lang program can never crash due to spatial safety violations.
-We utilize the model-based randomized testing (CITE) to certify the simulation relation between the~\lang semantics and the compiler.
-Specifically, we use a conversion tool that converts expressions from~\lang into actual~\checkedc code that can be compiled by the \checkedc compiler. We create a random program generator 
-based on the typing rules of~\lang and ensure that~\lang and~\checkedc compiler are consistent after conversion, both statically and dynamically.  
-To the best of our knowledge,~\lang is the first C(-like) language and compiler formalism with the program partitioning mechanism.
 
@@
 %
-To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
 
+\myparagraph{Formalizing the Type System, Semantics and Compiler}
+%
+We developed a core formalism named~\lang, which extends
+\citet{li22checkedc}~\checkedc formalism with the non-crashing guarantee and other new features below.
+We formally prove the~\emph{non-crashing theorem},~\ie 
+a well-typed~\lang program can never crash due to spatial safety violations.
+We utilize the model-based randomized testing (CITE) to certify the simulation relation between the~\lang semantics and the compiler.
+Specifically, we use a conversion tool that converts expressions from~\lang into actual~\checkedc code that can be compiled by the~\checkedc compiler. We create a random program generator 
+based on the typing rules of~\lang and ensure that~\lang and~\checkedc compiler are consistent after conversion, both statically and dynamically.  
+To the best of our knowledge,~\lang is the first C(-like) language and compiler formalism with the program partitioning mechanism.
+%To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
+
+We evaluated~\systemname by partioning~\numprog{} large real-world programs to demonstrate its effectiveness.
+Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.
+
+\aravind{Fix the following text after all the sections of the papers are finalized.}
 We begin with a review of \checkedc{} \mzr{and introduction of} new features in
@@
 % highlightlines={1-3}, highlightcolor=red
-\ignore{
+%\ignore{
 \begin{listing}[t!]
@@
 \end{listing}
-}
+%}
 }

### 2022-11-28 sandp23/overview.tex
More refactoring on intro and overview

@@
 To handle this, the developer marks risky functions with unchecked pointers~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded.
-The developer starts using~\oursystem with the resulting partially annotated program with checked and tainted types~\ie~\lst{lst:initialconv} and~\ref{lst:humantaint} atop of~\lst{lst:comb}.
+The developer starts using~\systemname{} with the resulting partially annotated program with checked and tainted types~\ie~\lst{lst:initialconv} and~\ref{lst:humantaint} atop of~\lst{lst:comb}.
 The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (CITE).
@@
 These sets of source files will be compiled with the corresponding compilers to get the corresponding object files.
-The c-region object file will have the necessary runtime checks enforcing~\systemname guarantees.
+The c-region object file will have the necessary runtime checks enforcing~\systemname{} guarantees.
 The uc-region object file will be produced according to the corresponding sandbox mechanism.

### 2022-11-28 sandp23/background.tex
Minor

@@
 
+\noindent
 \myparagraph{Checked Pointer Types}
@@
 \begin{itemize}
-\item \code{ptr<}$T$\code{>} types a pointer that is either null or
+\item \code{ptr<}$T$\code{>} ($\ptr$) types a pointer that is either null or
   points to a single object of type $T$.
-\item \code{array_ptr<}$T$\code{>} types a pointer that is either null
+\item \code{array_ptr<}$T$\code{>} ($\arr$) types a pointer that is either null
   or points to an array of $T$ objects. The array width is defined
   by a \emph{bounds} expression, discussed below.
-\item \code{nt_array_ptr<}$T$\code{>} is like
+\item \code{nt_array_ptr<}$T$\code{>} ($\ntarr$) is like
   \code{array_ptr<}$T$\code{>} except that the bounds expression
@@
 \end{itemize}
-A bounds expression used with the latter two pointer types has three
-forms:
-\begin{itemize}
-\item \code{count(}$e$\code{)} where $e$ defines the array's
-  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
-  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
-  expression $e$ must be side-effect free and may only refer to
-  variables whose addresses are not taken, or adjacent \code{struct}
-  fields.
-\item \code{byte_count(}$e$\code{)} is like \code{count}, but
-  expresses arithmetic using bytes, no objects; i.e.,
-  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
-  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
-\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
-  pointers that bound the accessible region $[e_l,e_h)$ (the
-  expressions are similarly restricted). Bounds
-  \code{count(}$e$\code{)} is shorthand for
-  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
-  expression is useful for supporting pointer arithmetic.
-\end{itemize}
-  Dropping the bounds expression on an \code{nt_array_ptr} is equivalent
-  to the bounds being \code{count(0)}.
+Both $\arr$ and $\ntarr$ pointers have an associated bounds which defines the
+range of memory referenced by the pointer.
+The three different ways to declare bounds and the corresponding memory range is:
+\begin{footnotesize}
+\begin{tabular}{ll}
+\arrT{|$T$|} \inlinecode{p: count(|$n$|)}
+  &
+$[\inlinecode{p}, \inlinecode{p}+\inlinecode{sizeof}(T) \times n) $ \\
+%$[\inlinecode{p}, )$ \\
+\arrT{|$T$|} \inlinecode{p: byte_count(|$b$|)}
+
+  &
+    $[\code{p}, \code{p}+b)$ \\
+
+\arrT{|$T$|} \inlinecode{p: bounds(|$x, y$|)}
+
+  &
+    $[x, y)    $\\  
+\end{tabular}
+\end{footnotesize}
+The bounds can be declared for $\ntarr$ as well, but the memory range can extend further to the right,
+until a~\code{NULL} terminator is reached (\ie the~\code{NULL} is not within the bounds).
+
+\subsection{Ensuring Memory Safety}
+
+%A bounds expression used with the latter two pointer types has three
+%forms:
+%\begin{itemize}
+%\item \code{count(}$e$\code{)} where $e$ defines the array's
+%  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
+%  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
+%  expression $e$ must be side-effect free and may only refer to
+%  variables whose addresses are not taken, or adjacent \code{struct}
+%  fields.
+%\item \code{byte_count(}$e$\code{)} is like \code{count}, but
+%  expresses arithmetic using bytes, no objects; i.e.,
+%  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
+%  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
+%\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
+%  pointers that bound the accessible region $[e_l,e_h)$ (the
+%  expressions are similarly restricted). Bounds
+%  \code{count(}$e$\code{)} is shorthand for
+%  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
+%  expression is useful for supporting pointer arithmetic.
+%\end{itemize}
+%
+%\begin{footnotesize}
+%\begin{tabular}{ll}
+%\arrT{|$T$|} \inlinecode{p: count(|$n$|)}
+%  &
+%$[\inlinecode{p}, \inlinecode{p}+\inlinecode{sizeof}(T) \times n) $ \\
+%%$[\inlinecode{p}, )$ \\
+%\arrT{|$T$|} \inlinecode{p: byte_count(|$b$|)}
+%
+%  &
+%    $[\code{p}, \code{p}+b)$ \\
+%
+%
+%\arrT{|$T$|} \inlinecode{p: bounds(|$x, y$|)}
+%
+%  &
+%    $[x, y)    $\\  
+%\end{tabular}
+%\end{footnotesize}
+%
+%  Dropping the bounds expression on an \code{nt_array_ptr} is equivalent
+%  to the bounds being \code{count(0)}.
 

### 2022-12-13 sandp23/background.tex
Adding back

@@
 until a~\code{NULL} terminator is reached (\ie the~\code{NULL} is not within the bounds).
+
+\noindent
+\myparagraph{Ensuring Spatial Memory Safety}
+The \checkedc compiler will instrument loads and stores of checked
+pointers to confirm the pointer is non-null, and additionally for $\arr$ and $\ntarr$ pointers the access is within
+the specified bounds.
+For instance, in the code \inlinecode{if (n>0) a[n-1] =} $...$ the write
+is via address $\alpha = \inlinecode{a + sizeof(int)}\!\times\!\code{(n-1)}$. If the
+bounds of \code{a} are \code{count(u)}, the
+inserted check will confirm that prior to dereference
+$\inlinecode{a} \leq \alpha < \inlinecode{a + sizeof(int)} \!\times\!
+\inlinecode{u}$.
+Failed checks throw an exception.
+
+Oftentimes, inserted checks can be optimized away by LLVM.
+Consider the above code to be enclosed in another condition, such as,~\inlinecode{if (n<u) if (n>0) a[n-1] =}.
+In such cases, the inserted check can be
+removed as the outer condition $\inlinecode{n<u}$ already ensures that the
+dereference is within bounds.
+Dual~\etal~\cite{duanrefactoring} found essentially no
+overhead when running Checked C-converted portions of the FreeBSD kernel.
+
+\noindent
+\myparagraph{Backward Compatibility}
 

### 2022-12-13 sandp23/background.tex
Adding

@@
 
-\subsection{\checkedc Overview}
+\subsection{\checkedc}
 % 
@@
 \myparagraph{Backward Compatibility}
+\checkedc is backward compatible with legacy C in the sense that all legacy code will type-check and compile.
+However, the compiler adds the aforementioned spatial safety checks to only checked pointers.
+The spatial safety guarantee is partial when the code is not fully ported.
+Specifically, only code that appears in \emph{checked
+  code regions}, which we call \emph{checked code}, is guaranteed to be spatially safe.
+Checked regions can be designated at the level of files, functions, or individual code blocks using the
+\code{checked} keyword.\footnote{You can also designate~\emph{unchecked} regions  within checked ones.}
+ Within checked regions, both legacy pointers and
+certain unsafe idioms (\eg \emph{variadic} function calls) are disallowed.
+
+\myparagraph{Converting C to \checkedc}
+%
+The safety guarantees of Checked C come with certain restrictions. For instance,
+as shown below, Checked C programs cannot use address-taken variables in a
+bounds expression as the bounds relations may not hold because of possible
+modifications through pointers.
+% 
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+...
+array_ptr<int> p : count (n) = NULL;
+|\textcolor{red}{\faTimes}|..,&n,.
+\end{minted}
+% 
+Consequently, converting existing C programs to Checked C might require
+refactoring,~\eg eliminate~\inlinecode{&n} from the program above without
+changing its functionality.
+% 
+This might require considerable effort~\cite{duanrefactoring} depending on the
+program's complexity.
+% 
+Recently, Machiry~\etal developed~\threec~\cite{machiry2022c} that tries to
+automatically convert a program to Checked C by adding appropriate pointer
+annotations.
+However, as described in \threec, completely automated conversion
+is \emph{infeasible}, and it requires the developer to convert some code regions
+manually.
 

### 2022-12-14 sandp23/background.tex
Adding CCS template

@@
-\section{Background}
-\subsection{Program Partitioning}
-\label{subsec:background:programpart}
-
-\section{Overview and Transcendence}\label{sec:overview}
-This section describes \checkedc{} and new features \systemname{} provides.
+\section{Background and Motivation}
+%\section{Overview and Transcendence}\label{sec:overview}
+%This section describes \checkedc{} and new features \systemname{} provides.
+This section presents a brief background of~\checkedc and motivation for~\systemname{}.
 
@@
 manually.
+
+\subsection{No Safety Against Unchecked Code}
+\label{subsec:nosafetyagsintuncheckedcode}
+The spatial safety guarantee~\checkedc provides applies to completely converted programs,~\ie programs that uses only checked types.
+Although, the backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
+The unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
+Consider the following example:
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+// Checked code
+int func(array_ptr<char> p : count(5)) {
+|\textcolor{red}{\faChainBroken}|..p[4]..
+}
+// unchecked code
+...
+str = "he";
+...
+func(|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5)); 
+...
+char ptr[16];
+...
+len <- derived from user input
+...
+|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
+\end{minted}
+Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function (Line 9) with a buffer of 2 elements.
+This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
+Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
+We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+
+\subsection{Program Partitioning}
+\label{subsec:background:programpart}
+Program partitioning~\cite{rul2009towards} is a well-known technique to divide a program into multiple isolated parts or partitions.
+There has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
+Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
+Specifically, given sensitive data in a program, the goal is to partition functions into two parts or partitions based on whether a function can access the sensitive data.
+The performance overhead of these approaches is dominated by marshaling costs and depends on the usage of sensitive data.
+The overhead of state-of-the-art approaches, such as Glamdring~\cite{lind2017glamdring} and  PtrSplit~
+\cite{liu2017ptrsplit}, is prohibitive and varies from 37\%-163\%.
+
+\iffalse
 
@@
 \mz{Shell vs. Wrapper? Shell reminds me of bash. }
+\fi
 

### 2022-12-14 sandp23/introduction.tex
Modifying overview

@@
 
+\myparagraph{\systemname Type System, Formalism and Compiler}
+We present a type system that integrates tainted types with~\checkedc and ensures the~\emph{non-crashing property},~\ie 
+a well-typed~\systemname program can never crash due to spatial safety violations of tainted pointers.
+We extend~\checkedc compiler to support our type system.
+We formalize our system by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing guarantee and other new features below.
+We formally prove the~\emph{non-crashing theorem} and use model-based randomized testing (CITE) to certify the simulation relation between the~\systemname semantics and the compiler.
+To the best of our knowledge,~\systemname is the first C(-like) language and compiler formalism with the program partitioning mechanism.
+
 \myparagraph{Type-Directed Program Partitioning}
 We present a type-directed program partition technique to separate checked and unchecked code regions and ensure that the code in checked region~\textit{cannot crash} due to spatial safety violations.
-Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted (or unchecked) pointers.
+Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted pointers.
 Our modular design enables us to use existing sandboxed techniques to enforce memory and execution isolation.
@@
 
+\iffalse
 \myparagraph{Formalizing the Type System, Semantics and Compiler}
@@
 %To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
+\fi
 

### 2022-12-14 sandp23/overview.tex
Modifying overview

@@
 The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} will be called to handle the request.
+
+\myparagraph{The Vulnerability}
 As we can see, there is an arbitrary memory write vulnerability (indicated by~\realbug) in~\code{process_req1} because of using~\code{i} as an index into the array~\code{msg} without any sanity check.
 The variable~\code{i} can take any integer value as it is parsed from~\code{msg}, whose contents are read from socket in~\code{read_msg} indicated by~\rootcause.
+
+\myparagraph{Goal} 
 The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the overall program.
@@
 To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
-However, as explained in~\sect{}, the unchecked code can also affect the safety of checked pointers.
+However, as explained in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
+
+\myparagraph{Using tainted types}
 To handle this, the developer marks risky functions with unchecked pointers~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded.
@@
 The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (CITE).
-The developer uses our type checker to identify the additional annotations required to have a well-typed program and adds them as shown in~\lst{}.
-The resulting well-typed program as shown in~\lst{} will be passed to our source level program partitioner along with certain configuration parameters of the target sandbox.
-Our partitioner splits the provided program into two sets of source files:
-with the necessary source changes required to communiate with sandboxed code.
+The developer uses our type checker to identify the additional annotations (\usermods) required to have a well-typed program and adds them as shown in~\lst{lst:humanadjust}.
+The resulting well-typed program as shown in~\lst{lst:final} will be passed to our source level program partitioner along with certain configuration parameters of the target sandbox.
+
+\myparagraph{Partitioning}
+Our partitioner splits the provided program into two sets of source files with the necessary source changes required to communicate with sandboxed code.
 These sets of source files will be compiled with the corresponding compilers to get the corresponding object files.
 The c-region object file will have the necessary runtime checks enforcing~\systemname{} guarantees.
-The uc-region object file will be produced according to the corresponding sandbox mechanism.
-Finally, these two object files will be linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions will be executed in a sandbox (uc-region) and the rest of the functions as regular code (c-region).
+The~\ucregion object file will be produced according to the corresponding sandbox mechanism.
+Finally, these two object files will be linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions will be executed in a sandbox (\ucregion) and the rest of the functions as regular code (\cregion).

### 2022-12-14 sandp23/introduction.tex
Adding

@@
 
-Vulnerabilities due to memory corruption, especially spatial memory corruptions, 
+Vulnerabilities due to memory corruption, especially spatial memory corruption, 
 are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798} 
@@
 These approaches all impose performance overheads deemed too high for deployment use. 
-Recently, \citet{Elliott2018} and \citet{li22checkedc} introduced and formalized \checkedc, an
+Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
 open-source extension to C with new types and
 annotations whose use can ensure a program’s spatial safety.
-Importantly, \checkedc supports development that is 
-incremental and compositional. Code regions (e.g.,
-functions or whole files) designated as \emph{checked} enforce
-spatial safety in a manner preserved by composition with
-other checked regions. But not all regions must be checked: Checked
-C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
-may coexist in the same code, which permits a deliberate (and
-semi-automated) refactoring process. 
-Parts of the FreeBSD kernel have
-been successfully migrated to \checkedc~\cite{duanrefactoring}, and the overall performance
-overhead is minimal enough for practical deployment.
+Specifically,~\checkedc extends C with checked pointer types, which are restricted by the compiler to spatially safe uses.
+The checked pointers are represented as system-level memory words,~\ie without ``fattening'' metadata, ensuring backward compatibility.
+This enables incremental conversion, where regions of code can be converted to~\checkedc and ensure spatial safety within those regions.
+However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the unconverted or unchecked code can violate guarantees provided in Checked regions.
+%
+We need to ensure that~\emph{code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
 
@@
 % \end{itemize}
+\iffalse
 \noindent
@@
 We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+\fi
 This mechanism is called program partitioning~\cite{rul2009towards}.

### 2022-12-22 sandp23/evaluation.tex
Adding evaluation section

@@
 % \liyi{Deena? Leo? }
+Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program-benchmarks.
+MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions and sandboxed execution of functions.
+We further go on to evaluate \systemname on six real-world programs ranging across diversified domains to evaluate real-world performance when using checkcbox comprehensively or conservatively.
+Our evaluation highlights \systemname's capabilities in required conversion efforts, run-time and memory overhead.
+All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test was sampled as the mean of ten consecutive iterations.
 
-\begin{itemize}
+\subsection{Micro-Benchmarks}
+\subsubsection{Memory Access in SBX}
+In order to evaluate memory access performance between the checked and the WASM region, we craft a test case which involves a simple pointer arithmetic operation enclosed in a loop 100k iterations long.
+Considering the mean of ten executions of this test, we found sandboxed test case to execute 156.6\% slower than when the same test case is executed in the checked region.
+This can directly be attributed to the fact that WASM compiler toolchains do not support optimization similar in effect to LLVM/GCC.
 
-\item provide evidence that the \systemname compiler is efficient. Compare the compiler with respect to other work, like RLBox, also the previous \checkedc compiler.
+\subsubsection{Indirect-Call}
+This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call into the SBX) , initiated in Checked Region, to return (to the Checked Region).
+This metric is evaluated against the time taken for a checked Callee (local call) to return in the Checked Region.
+Observed overhead is once again a consequence of unoptimized WASM Sandbox and stack of indirect calls.
+\subsubsection{Pointer Access}
+Finally, we evaluate run-time overhead between the checked region pointers (Generic-C Pointers) and tainted pointers.
+This test involves 100k read/write/arithmetic operations on the pointer.
+We have found the overhead to be 34\%, which is the result of Offset to pointer converters and Taintedness sanity checks inserted by \systemname at every access to the tainted pointer.
+\subsubsection{Verdict on Micro-Benchmarks}
+The observed performance deficit is the execution of sandboxed code is directly tied to the limitation of the Web-Assembly binary and the instrumentation inserted by \systemname.
+\systemname is aimed at best use in encapsulating selective unsafe/unchecked regions rather that significant or massive sections of performance intensive code.
+When used judiciously alongside checked-c, overhead offered by  \systemname is masked to be insignificant whilst offering complete spatial safety.
+\begin{center}
+\begin{tabular}{||c c||} 
+ \hline
+ Test-Name & \systemname overhead \\ [0.5ex] 
+ \hline\hline
+ Memory Access in SBX & +156\% \\
+ Indirect-Call & +398\% \\ 
+ Pointer Access & +34.16\% \\ [1ex]
+ \hline 
+\end{tabular}
+\end{center}
 
-\item provide user experience of \systemname.
-We restrict the use of checked pointers compared to previous checked-c compiler.
-Is the restriction arrangable.
-We can say that the tainted shells are auto-matically generated, so we have a mechanism for auto-generating tainted pointers if necessary.
+\subsection{Program Run-time Benchmarks}
+We have manually converted six programs from checked-c/Generic-C to \systemname and evaluate the performance metrics by making use of a comprehensive pre-included test-suite for each program.
+Our conversion is aimed at encapsulating the root causes for CVE's by making the least possible \systemname annotations to fix the bug.
+Our intuition behind this goal is to show that \systemname is capable of enforcing spatial safety in all of the scenarios leading to modern day vulnerabilities.
+However, this intuition would also imply that the bugs are predicted before they are discovered, which is a chronological impossible.
+Hence in search of uniformity, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on bug locality.
+Since most of the bugs pertaining to these CVE's are fixed, we first reproduce the issue by undoing the fix and then attempt to reproduce the same by retrofitting the vulnerable code with \systemname annotations.
+Once we annnotate a generous region of the program, We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
+Our timing scope excludes marshalling activities that may occur within a test-case environment within an intuition that the user would never need to perform marshalling if the user propagates the tainted\-ness from the call-site up until the point in the program where these arguments are declared.
+However, we do account for marshalling that might happen within the source code.
+We use Valgrind's "massif" memory profiler benchmark the peak memory usage of the Heap.
+Unlike the Runtime performance which is recorded in percentage, we record Peak Memory usage as a relative offset, with the reason being many program executables are extremely small in comparison to the constant overhead from the sandbox.
+Unlike Runtime overhead which is directly proportional to the extent to which pointers are marked as tainted, peak heap-memory usage of \systemname is linearly related to the unconverted version by a fixed constant (81 KB approx).
+This constant comprises of a 10KB (approx) sandbox, and a 70 KB (approx) custom Tainted wrapper functions for stdlib.
+This can further be optimized away by only compiling custom Tainted wrappers for those stdlib functions that are use by the tainted pointers in the program.
+However, since Tainted pointers are allocated in WASM Sandbox's shadow memory region, valgrind's memory figures do not account for these allocations.
 
-\item if we have space, we can re-introduce random testing a little, saying that how it helps us to develop the compiler.
+\subsubsection{Parsons}
+We have converted parsons to \systemname in two different variants.
+The first variant, parsons\_wasm involves moving most of the parsing functions to the sandbox, whilst marking all pointers as tainted.
+Sandboxed functions executing in the sandbox interact with the checked region by making indirect calls through RLBOX's callback mechanism.
+With the second variant, parsons\_tainted, we do not move any of the functions to the sandbox, however, we still mark all pointers as tainted and propagate the taintedness to every single use of these pointers.
+The test-suite itself consists of 328 tests comprehensively testing the JSON parser's functionality.
+The benchmarking for both of these forks is recorded using the mean difference between the \systemname and generic-C variants when executing 10 consecutive iterations of the test-suite.
+Since parsons also has a checked-c version, we have evaluated checkcbox-parsons against both the checkedc and generic-c unconverted versions.
+parsons\_wasm expectedly shows 200-266\% runtime overhead when evaluated against checked-c and generic-c respectively.
+However, evaluating parsons\_tainted against checked-c shows \systemname to be faster.
+This is because \systemname by itself performs lighter run-time instrumentation on tainted pointers compared to the run-time bounds checking performed on checked pointers by checked-c.
+Furthermore, we only see an average 9.5 KiB memory as compared to the anticipated 82 KiB overhead as valgrind does not consider the WASM Shadow memory allocated to the tainted pointers.
 
-\item we can then talk about the possible bugs we find in the \checkedc compiler for function pointer or the RLBox bugs.
+\subsubsection{Libpng}
+\systemname changes for Libpng involves encapsulating CVE-2018-144550 and a buffer overflow not registered with the NVD (buffer overflow in compare\_read()).
+CVE-2018-144550 is recorded as a stack-based buffer overflow caused by an Out-of-bounds write in the function get\_token() in pnm2png.
+Consequently we marked the input argument "token\_buf" as a tainted pointer and propagated its tainted-ness throughout the program.
+However, in interests of making changes comprehensive, our changes for libpng involve reading, writing and operating (interlace, intrapixel, etc) on image data that is directly stored in the Tainted region.
+That is, rows of image bytes are read into tainted pointers and the taintedness for the row\_bytes is propagated throughout the program.
+This way, manually written tainted version (t\_png\_do\_read\_intrapixel()) for each of the pointer arithmetic heavy functions (png\_do\_read\_intrapixel())are spatially memory safe.
+All our changes extend to png2pnm and pnm2png.
+To evaluate png2pnm, we take the mean of 10 iterations of a test script that runs png2pnm on 52 png files located within the libpng's pngsuite.
+To test pnm2png, we take the mean of 10 iterations of pnm2png in converting a 52MB 5184x3456 pixels large pnm image file to png.
+We observe that valgrind's massif memory profiler reports lower Heap space consumption for \systemname converted code as compared to the the unconverted code.
+As mentioned above, the memory profiler's outcome does not account for the space consumed on the heap by the Sandbox's shadow memory.
+Consequenlty, when evaluating pnm2png, \systemname's heap consumption was 52 MB lower as the entire image was loaded onto the shadow memory.
 
-\end{itemize}
+\subsubsection{Proftpd}
+Checkcbox changes for ProFTPD were aimed at encapsulating CVE-2010-4221.
+CVE-2010-4221 is recorded as a vulnerability within "pr\_netio\_telnet\_gets()" function, that is triggered when processing user input containing the Telnet IAC (Interpret As Command) escape sequence.
+This vulnerabilty can be exploited to cause a stack-based buffer overflow by sending specially crafted input to the FTP or FTPS service.
+We mark the user input to the unsafe function (char*) as tainted (\_TPtr<char>), and propagate the tainted-ness forward through its return value and as arguments of its callers and callees enclosed within its definition scope.
+ProFTPD is pre-included with a test-suite that tests each of APIs comprehensively.
+Hence, our changes for the above function were policed by 24 unique tests, which we use to benchmark the performance.
+For each test, our benchmark samples the delta between call and return time of pr\_netio\_telnet\_gets().
+This sampling is repeated 10 times and its mean is reflected in the tabulation.
+Although, \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where the major metric to performance is network bandwidth.
 
+\subsubsection{MicroHTTPD}
+MicroHTTPD is a good demonstration of the practical difficulties in converting a program to \systemname.
+Our changes were attempted at sandboxing memory vulnerabilities CVE-2021-3466 and CVE-2013-7039.
+CVE-2021-3466 is described as a vulnerability from a buffer overflow that occurs in an unguarded memcpy which copies data into a structure pointer (struct MHD\_PostProcessor pp) which is type-casted to a char buffer (char *kbuf = (char *) \&pp[1]).
+Our changes would require making the memcpy safe by marking this pointer as tainted.
+However, this would either require marshalling the structure pointer (from checked to tainted memory) or would require marking every reference to this structure pointer as tainted, which in turn implies, marking every pointer member of this structure as tainted.
+Marshalling data between structure pointers is not easy due to the non-linearity of its members unlike a char* thereby requiring substantial marshalling glue logic that did not align with our conversion goals.
+Consequently, the above CVE stands un-handled by \systemname.
+CVE-2013-7039 is described as a stack buffer overflow in the MHD\_digest\_auth\_check() that could be used as an exploit.
+Our changes involved marking the user input data arguments of this function as tainted pointers.
+However, in interests to keep changes simple, we do not propagate the tainted-ness of these function arguments to all its callees and just perform marshalling at the call-site.
+We went a little further to future-proof and moved many of the internal functions (like MHD\_str\_pct\_decode\_strict\_() and MHD\_http\_unescape()), that perform a lot of pointer arithmetic based on user-input, into execute in the sandbox.
+The 172\% rise in RunTime can be directly tied to the limitations of WebAssembly, as most of the string processing functions are executing within the WASM sandbox.
+The Average memory overhead is less than the expected as two of the ten test cases do not make use of the Tainted stdlib wrappers, consequently, the peak memory usage for these test cases is almost the same.
+
+\subsubsection{UFTPD}
+Chekcbox changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
+CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which lead to a Denial Of Service in versions before 2.12.
+Consequently, we sandboxed this function.
+CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which required us to sandbox this function as well.
+For evaluation, we manually write a script for 3 Tests that triggers "quote CWD", "quote PORT", and ftp "get file" requests 10 times in a loop.
+We thereby record an entry for each of these tests as a mean of recorded timestamps for 10 executions for checkcbox and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the below table.
+
+\subsubsection{Tiny-bignum}
+\systemname changes for Tiny-bignum was comprehensive and involved marking all the pointers as tainted.
+Furthermore, in likes of an unsafe use of sprintf() that could lead to a buffer overflow within the bignum\_to\_string(), we have moved this function into the WASM sandbox.
+Ideally, a major portion of conversion efforts to \systemname is attributed to understanding the codebase and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling.
+However tiny-bignum was completely ported to \systemname under 4 hours as the codebase was quick to understand and with the eliminated problem of tainted-propagation as a result of marking all pointers as tainted.
+Tiny-bignum's test suite consists of 4 Test cases that test functionality to scale on big numbers subject to all of the supported unary and binary operations.
+
+\begin{center}
+\begin{tabular}{||c | p{1.3cm} | p{1.8cm}||} 
+ \hline
+ Program & Run-Time & Avg Peak Memory \\ [0.5ex] 
+ \hline\hline
+ ProFTPD & +47.8\% & +82.1 KiB  \\
+ MicroHTTPD & +171.7\% & +24.7 KiB\\ 
+ UFTPD & +2.7\% & +86.3 KiB \\ 
+ Tiny-Bignum (vs checked-c) & -23.4\% &  +81.1 KiB \\
+ Tiny-Bignum &  +121.7\% &  +81.1 KiB \\
+ parsons\_wasm (vs checked-c) & +200.2\% & +9.8 KiB\\ 
+ parsons\_tainted (vs checked-c) & -8.7\% & +9.2 KiB\\
+ parsons\_wasm & +266.5\% & +9.8 KiB\\ 
+ parsons\_tainted & +10.45\% & +9.2 KiB\\ 
+ LibPNG (png2pnm) & +11.41\% & -102.5 KiB\\ 
+ LibPNG (pnm2png) & +46.5\% & -51.6 MiB\\ [1ex]
+ \hline 
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{||p{2 cm} | p{1.2cm} | p{1.2cm} | p{1.0cm} | p{1.0cm}||} 
+ \hline
+ Program & Pointers Annotated  & Lines Sandboxed & CVEs fixed & Time To Port \\ [0.5ex] 
+ \hline\hline
+ ProFTPD & 6 & 0 & - & 1 Day  \\
+ \hline
+ MicroHTTPD & 139 & 650 & CVE-2013-7039 & 4 Days \\
+ \hline
+ UFTPD & 146 & 90 & CVE-2020-14149, CVE-2020-5204 & 3 Days \\ 
+ \hline
+ LibPNG & 248 &  0 & CVE-2018-144550 & 8 Days \\
+ \hline
+ Tiny-Bignum &  69 &  30 & - & 4 Hours \\
+ \hline
+ parsons\_wasm & 364 & 800 &  - & 3 Days \\ 
+ \hline
+ parsons\_tainted & 378 & 0 & - & 8 Days \\ [1ex]
+ \hline 
+\end{tabular}
+\end{center}
 % 

### 2022-12-26 sandp23/formal.tex
Adding CCS template
update.

--- 2022-08-16 sandp23/formal.tex
@@
-\section{Formalization}\label{sec:formal}
-\liyi{This section might say too much, we might cut more here, and put more items in the compilation and implementation. }
+\section{\systemname Type System and Formalism}\label{sec:formal}
 
@@
 \end{array}
-\\
 \]
@@
 language presented in Fig.~\ref{fig:checkc-syn}.
-\liyi{why can we just say in.. }
 
@@
 $\omega$ classify multi-word values such as arrays, null-terminated
-arrays, functions, and single-word-size values. \liyi{Is this a good way? the differences are mainly for the result types vs type elements. }
-%
+arrays, functions, and single-word-size values. 
 Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
@@
 and a type ($\omega$) denoting valid values that can be pointed to.
-\liyi{a little too much? Should we assume people know what pointer type is?}
-\liyi{never mention $m$. }
 Array types include both the type of
@@
 otherwise (we elide $\cdot$ when writing types).
-\liyi{should have a better way. }
-\lang function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
+
+\lang{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
 reflect its dependent function declarations,
 where $\overline{x}$ represents
-a list of \tint type variables in a dependent function header
+a list of \tint{} type variables in a dependent function header
 that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
-\liyi{need to fix, do not understand. }
 We have a well-formed requirement for a function type;
@@
 $\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-\liyi{we might just cut nt-array ptrs. it seems to me that nt-array ptrs are not important. }
 
@@
 
-\liyi{should we bring out the difference? since readers might not even know the previous paper.}
 Compared to the former \checkedc model \cite{li22checkedc},
@@
 such that $\tmode$ is projected as $\umode$.
-\liyi{not clear. }
 Second, $\emalloc{\xi}{\omega}$ includes a mode flag
@@
 Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
-\liyi{might need to bring out what is the difference of previous ones.}
 Fourth, $\echeckedtext$ blocks are added to the system, 
@@
 in \checkedc, so we omit them.
-\liyi{last sentence can be cut. }
 
@@
 
-\subsection{Semantics}\label{sec:semantics}
+\subsection{Semantics and Typing}\label{sec:semantics}
 
@@
 
-The operational semantics for \lang is defined as a small-step
+Here, we discuss the \lang type system and operational semantics; 
+mainly focusing on the new changes on top of \checkedc in \cite{li22checkedc} with function pointers, modes, and function calls.
+The other type and semantic rules about (NT)-arrays are given in \cite{li22checkedc} and \Cref{sec:literal-pointer-typing}.
+
+The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+which states that in a type environment $\Gamma$ (mapping variables to
+their types) and a predicate environment $\Theta$ (mapping integer-typed
+variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
+in context mode $m$. Key rules for this judgment are given in
+Fig.~\ref{fig:type-system-1},
+while the operational semantics for \lang is defined as a small-step
 transition relation with the judgment $ (\varphi,\heap,e)
 \longrightarrow_m (\varphi',\heap',r)$.
-\liyi{do we need to say judgment here? }
  Here, $\varphi$ is a
@@
 maps addresses (integer literals) to values $\evalue{n}{\tau}$.
+
+\myparagraph{Pointers, Contexts, and Modes}
 A $\cmode$ pointer is mapped to a heap location in the $\cmode$ heap, 
@@
 mutable heap.
-\liyi{why we need this sentence? }
 As mentioned, value $\evalue{0}{\tau}$
@@
 and continues with the existing program state.
-\liyi{is non-deterministically crashed or non-determinstically transitioned to some place. }
 Evaluation contexts $E$ define a standard left-to-right evaluation order. (We explain the
 $\ret{x}{\mu}{e}$ syntax shortly.)
-There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
-\liyi{might not need this. }
-
-%\footnote{This approach is that of the PLT Redex model of \lang; the Coq
-%development uses a slightly simpler syntax to achieve the same
-%effect.}
-% \review{the special case raises questions, e.g. why is this syntax-driven and
-%   not type-driven? }
-% \liyi{This describes the semantic transition rules.
-We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3.
-like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.
-}
-% \mwh{Don't follow the above. I don't see this ``context transition
-%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
-% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
-The $\mode$ function  at the bottom of Fig.~\ref{fig:c-context}
-describes the context mode determination in each evaluation step based on the context $E$.
-\liyi{understood? }
-For any program execution, the function starts the mode computation with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
-The result context mode depends on where $\Box$ locates.
-If it occurs within $E$ in $(\eunchecked{\overline{x}}{E})$
-that has no surrounding $\echeckedtext$ block,
-the mode is $\umode$; otherwise, the mode is $\cmode$.
-$\mode'(\alpha({E}),m)=\mode'(E,m)$ represent other construct cases 
-that are not $\echeckedtext$ and $\euncheckedtext$; in such case, 
-the function recursively traverses the sub-context to find the context mode.
-\liyi{may be too low-level detailed. describe it in a high level. }
-
-% For a term $e$, $C$ is a context if any
-% only if it contains a $\Box$ hole term in one of the subterm position
-% in $e$. $\Box$ indicates the break point where we split $e$ into $C$
-% and a subterm $e'$ such that $C[e'] = e$.  For a function call
-% $\ecall{f}{\overline{e}}$, a valid context ($\overline{C}$) is defined
-% as any one of the elements in the list $\overline{e}$ being a
-% context. We show a rule of the upper level \checkedc semantics. The
-% rule splits a term $e$ into a context $C$ and redex $e'$.  If subterm
-% $e'$ is transitioned to $e_a$, $C[e']$ is transitioned to
-% $C[e_a]$.
+%There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
 
@@
     {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-AssignArrC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\cmode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}
-
+\quad
         \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
@@
     {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-AssignArrT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h)
-               \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_1}{\tau} }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\umode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}
-
+\quad
     \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-
+\quad
     \inferrule[S-Cast]
@@
 
+    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
     \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}
 
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
 
@@
    (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-
-    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
+\quad
     \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
@@
 \end{DIFnomarkup}
+
+The $\mode$ function  at the bottom of Fig.~\ref{fig:c-context}
+describes the context mode determination in each evaluation step based on the context $E$.
+For any program execution, the function starts the mode computation with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
+The result context mode depends on where $\Box$ locates.
+If it occurs within $E$ in $(\eunchecked{\overline{x}}{E})$
+that has no surrounding $\echeckedtext$ block,
+the mode is $\umode$; otherwise, the mode is $\cmode$.
+$\mode'(\alpha({E}),m)=\mode'(E,m)$ represent other construct cases 
+that are not $\echeckedtext$ and $\euncheckedtext$; in such case, 
+the function recursively traverses the sub-context to find the context mode.
+
+Fig.~\ref{fig:semantics} shows selected rules for the computation relation.
+The rules for pointer related operations---\textsc{S-DefC},
+\textsc{S-DefT}, \textsc{S-DefNull}, and \textsc{S-Cast}.
+The type rule for deference operations is given as rule \rulelab{T-Def} in \Cref{fig:type-system-1}.
+The first three define the semantics of deference and assignment operations.
+Rule \textsc{S-DefNull} transitions attempted null-pointer
+dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a $\cmode$-mode
+non-null (single) pointer.
+When $\enull$ is returned by the
+computation relation, the evaluation relation halts the entire
+evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
+does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
+%\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
+%dereference) is within the bounds designated by the pointer's annotation
+%and strictly less than the upper bound. 
+\textsc{S-DefT} is similar to \textsc{S-DefC} for tainted pointers.
+Any dynamic heap access of a tainted pointer requires a \textit{verification}.
+Performing such a verification equates to performing a literal type check for a pointer constant in \Cref{fig:const-type}.
+We explain this shortly below for \emph{constant validity checks}.
+For now, the verification step, e.g. $\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$ in \textsc{S-DefC},
+refers to that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.
+
+Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
+by \textsc{S-Cast}. In a type-correct program, such casts are
+confirmed safe by the type system no matter
+if the target is a $\tmode$ or $\cmode$ pointer. To evaluate a cast, the rule
+updates the type annotation on $n$. Before doing so, it must
+``evaluate'' any variables that occur in $\tau$ according to their
+bindings in $\varphi$. For example, if $\tau$ was
+$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
+produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
+%The full formalism, including \kw{struct}
+%and null-terminated bound widening pointer operations, is given in \Cref{app:main}.
+
+%\footnote{This approach is that of the PLT Redex model of \lang; the Coq
+%development uses a slightly simpler syntax to achieve the same
+%effect.}
+% \review{the special case raises questions, e.g. why is this syntax-driven and
+%   not type-driven? }
+% \liyi{This describes the semantic transition rules.
+We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3.
+like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.
+}
+% \mwh{Don't follow the above. I don't see this ``context transition
+%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
+% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
+
+\begin{DIFnomarkup}
+\begin{figure*}[t]
+{\small
+  \begin{mathpar}
+   \inferrule[T-ConstU]
+       { \neg \cmode(\tau)}
+       {\Gamma;\Theta\vdash_u \evalue{n}{\tau} : \tau}
+
+   \inferrule[T-ConstC]
+       {\Theta;\heap;\emptyset \vdash_c n : \tau}
+       {\Gamma;\Theta\vdash_c \evalue{n}{\tau} : \tau}
+
+    \inferrule[T-Def]
+              {\xi \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
+              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
+
+     \inferrule[T-CastPtr]
+               {\Gamma;\Theta \vdash_m e : \tau' \\
+                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
+               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
+
+   \inferrule[T-Let]
+    { x\not\in \fv(\tau') \\
+        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
+          \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'}
+
+   \inferrule[T-LetInt]
+    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\
+        \Gamma;\Theta \vdash_m e_1 : \tint \\\\
+           \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
+\qquad
+   \inferrule[T-RetInt]
+    { \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
+    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tint}}{e} : \tau}
+
+    \inferrule[T-Checked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
+              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
+
+    \inferrule[T-Unchecked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
+              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
+
+\inferrule[T-Fun]
+    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
+        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\\\
+         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
+         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
+             \overline{\tau'} \sqsubseteq_{\Theta}
+               \overline{\tau}[\overline{e'} / \overline{x}]}
+    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
+
+  \end{mathpar}
+}
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
+% \\[0.2em]
+% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
+% \end{array}
+% $
+% \end{center}
+% }
+{\footnotesize
+$
+\cmode(\tint)=\texttt{false}
+\qquad
+\cmode(\tptr{\omega}{\cmode})=\texttt{true}
+\qquad
+\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
+$
+}
+\caption{Selected type rules}
+\label{fig:type-system-1}
+\end{figure*}
+\end{DIFnomarkup}
+
+\begin{DIFnomarkup}
+\begin{figure}
+{\small
+\[\hspace*{-1.2em}
+\begin{array}{l}
+\textcolor{blue}{\text{Bound Inequality and Equality:}}\\[0.3em]
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
+     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
+     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
+     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
+     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
+     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
+    \end{array}
+  \\[1.5em]
+\textcolor{blue}{\text{Type Equility:}}\\
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     && \tint & =_{\Theta} & \tint\\
+     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
+     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
+             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
+
+    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
+
+ &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
+                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
+    \end{array}
+  \\[0.7em]
+\textcolor{blue}{\text{Subtype:}}\\[0.3em]
+
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
+
+    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
+\\[0.2em]
+\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
+
+    \end{array}
+\end{array}
+  \]
+}
+{\footnotesize
+\[
+\begin{array}{l}
+n'+n = add(n',n)
+\qquad
+(x+n')+n = x+add(n',n)\\
+\textit{cond}(\overline{x},\tau,\overline{y},\tau')
+=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
+  \wedge \overline{y}\cupdot\overline{z}
+  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
+\\\qquad\qquad\qquad\qquad\qquad
+  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
+\end{array}
+\]
+}
+  \caption{Type Equality and Subtyping}
+  \label{fig:checkc-subtype}
+\end{figure}
+\end{DIFnomarkup}
+
+\myparagraph{Type Equality and Subtyping and Casting}
+%
+In \lang, type equality $\tau=_{\Theta}\tau'$
+is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
+and the alpha equivalence of two function types in \Cref{fig:checkc-subtype};
+i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
+$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
+are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
+and the substitution results are equal.
+
+ The \textsc{T-CastPtr} rule
+permits casting from an expression of type $\tau'$ to a checked pointer when
+$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
+$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and is built on the type equality
+($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
+rules ensure the relation is transitive. Most of the rules manage
+casting between array pointer types. The rule 
+($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
+\tarrayptr{b_l}{b_h}{\tau}{m}$) permits treating a singleton
+pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
+Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
+%There is another casting rule in \Cref{app:main} stating that
+% users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
+
+Since bounds expressions may
+contain variables, determining assumptions like $b_l \leq b_l'$
+requires reasoning about the probable values of these variables'. The type
+system uses $\Theta$ to make such reasoning more precise.
+$\Theta$ is a map from variables $x$ to
+equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
+It maps variables to equations that are recorded along the type checking procedure.
+If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
+$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
+such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
+\iftr
+  Appendix~\ref{app:le}.
+\else
+  the supplemental report~\cite{checkedc-tech-report}.
+\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
+
+\begin{DIFnomarkup}
+ \begin{figure}[t]
+ {\small
+
+ \begin{mathpar}
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m n : \tint}
+
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
+       {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
+
+   \inferrule
+       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
+            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       { \xi \le m 
+     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
+       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
+
+   \inferrule
+       {\neg\funptr(\omega)\\ \xi \le m\\
+        \forall i \in [0,\size(\omega)) \;.\;
+            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+ \end{mathpar}
+ }
+{\footnotesize
+\[
+\begin{array}{l} 
+\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
+\qquad
+\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
+\end{array}
+\]
+}
+ \caption{Verification/Type Rules for Constants}
+ \label{fig:const-type}
+ \end{figure}
+\end{DIFnomarkup}
+
+\myparagraph{Constant Validity}
+Rules \textsc{T-ConstU} and \textsc{T-ConstC}
+describes type assumptions for constants appearing in a program.
+$\cmode(\tau)$ judges that a constant pointer 
+in an unchecked region cannot be of a checked type,
+which represents the assumption that programmers 
+cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+In rule \textsc{T-ConstC}, we requires a static 
+verification procedure for validating a constant pointer, 
+which is similar to the dynamic verification process in \Cref{sec:semantics}.
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
+validate the constant $\evalue{n}{\tau}$, 
+where $\heap(m)$ is the initial heap that the constant resides on and
+$\sigma$ is a set of constant assumed to be checked.
+A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
+A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
+and the address stores a function with the right type.
+The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
+where every element in the pointer range ($[0,\size(\omega))$) should be well
+typed.
+
+A checked pointer checks validity in type step as rule \textsc{T-ConstC},
+while a tainted/unchecked pointer does not check for such during the type checking.
+Tainted pointers are validated through the validity check in dynamic execution as we mentioned above.
 
@@
 
-Fig.~\ref{fig:semantics} shows selected rules for the computation relation.
-
-\ignore{
- in
-Fig.~\ref{fig:strcat-ex},
-  which defines a 
-  safe version of \code{strcat} (using actual Checked C syntax).  The
-  function takes a target 
-  pointer \code{dst} of capacity \code{n}, where the first null
-  character (determined by \code{strlen}) is at index \code{x} where
-  $0 \leq $\code{x}$ \leq n$. It concatenates the \code{src} buffer to
-  the end of \code{dst} as long as \code{dst} has sufficient space.}
-
 % Below, we introduce low-level transition semantics for some case operations.
 The design of the low-level individual operation semantics is carefully engineered to perform match our compiler's behavior, such as correctly characterizing the bound widening behaviors for NT-array pointers, even though it is written in terms of fat-pointer formalization.
-
-\myparagraph{Checked and Tainted Pointer Operations}
-%
-The rules for pointer related operations---\textsc{S-DefC},
-\textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
-\textsc{S-DefNull}, and \textsc{S-Cast}.
-The first five define the semantics of deference and assignment operations.
-Rule \textsc{S-DefNull} transitions attempted null-pointer
-dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a $\cmode$-mode
-non-null (single) pointer.
-When $\enull$ is returned by the
-computation relation, the evaluation relation halts the entire
-evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
-does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
-\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
-dereference) is within the bounds designated by the pointer's annotation
-and strictly less than the upper bound.
-
-\textsc{S-DefT} and \textsc{S-AssignArrT} are similar rules to \textsc{S-DefC} and \textsc{S-AssignArrC} for tainted pointers.
-Any dynamic heap access of a tainted pointer requires a \textit{verification}.
-Performing such a verification equates to performing a literal type check for a pointer constant in \Cref{fig:const-type}.
-We explain this shortly in \Cref{sec:type-system}.
-For now, the verification step, e.g. $\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$ in \textsc{S-DefC},
-refers to that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.
-
-Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
-by \textsc{S-Cast}. In a type-correct program, such casts are
-confirmed safe by the type system no matter
-if the target is a $\tmode$ or $\cmode$ pointer. To evaluate a cast, the rule
-updates the type annotation on $n$. Before doing so, it must
-``evaluate'' any variables that occur in $\tau$ according to their
-bindings in $\varphi$. For example, if $\tau$ was
-$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
-produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
-The full formalism, including \kw{struct}
-and null-terminated bound widening pointer operations, is given in \Cref{app:main}.
-
 % \review{- on page 6, section IIIC, paragraph "Pointer Access" mentions that checked pointers cannot be dereferenced in unchecked blocks - this looks funny, shouldn't it be the other way around?
@@
 %
-Semantically, $\euncheckedtext$ and $\echeckedtext$ blocks act as classical C blocks
+During the type checking,
+Both $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$
+check all free variables in $e$ are within $\overline{x}$;
+the types for $\overline{x}$ and the final return type $\tau$ have no checked pointers.
+A $\echeckedtext$ or $\euncheckedtext$ block represents 
+the context switching from a checked to an checked region, or vice versa.
+We need to make sure no checked pointers are information exposed to unsafe code regions.
 as rules \textsc{S-Unchecked} and \textsc{S-Checked} in \Cref{fig:semantics}.
 
-\liyi{Yeah, the block semantics is easy, but is there anything to say? I mean this is the essentially item.}
-
-\myparagraph{Binding and Function Calls}
+\myparagraph{Let Bindings and Dependent Function Pointers}
 %
+Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
+type dependency. 
+In particular, the result of evaluating a $\elettext$
+may have a type that refers to one of its bound variables (e.g., if
+the result is a checked pointer with a variable-defined bound). 
+If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
+Note that we restrict the expression $e_1$ to syntactically match the
+structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
+
+Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
+$\erettext$ does not appear in source programs but is introduced by the semantics when
+evaluating a let binding (rule \textsc{S-Let} in
+Fig.~\ref{fig:semantics}). 
+%\liyi{why? }
+% After the evaluation of a let binding a variable $x$ concludes,
+%we need to restore any prior binding of $x$, which is either
+%$\bot$ (meaning that there is no $x$ originally) or some value
+%$\evalue{n}{\tau}$.
+
+Rule \textsc{T-Fun} is the dependent function call rule. 
+Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
+from a type-check for $e$ and the types $\overline{\tau'}$ from the argument type checks for $\overline{e}$,
+we confirm that each of $\overline{\tau'}$ is
+a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$,
+which replaces possible integer bound variables $\overline{x}$ with bound expressions $\overline{e'}$.
+The final result type is the defined target type $\tau$ appearing in the function pointer type
+also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
+Consider the \code{safe_strcat} function in
+Fig.~\ref{fig:strcat-ex}; its parameter type for \code{dst} 
+depends on \code{n}. The \textsc{T-Fun} rule will substitute 
+\code{n} with the argument at a call-site.
+
 The semantics manages variable scopes using the special $\erettext$
@@
 in $e$'s scope that is defined by the $\erettext$ operation.
-\liyi{understood? }
 Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
@@
  the function definition in $n$'s location in the global function store $\Xi$,
+% 
 % LEO: This is the first occurence of \Xi. It is explained,
@@
 function body.
+
 Similar to \heap, the global function store $\Xi$ is also partitioned into
@@
 argument list.
-\liyi{This has appeared in the previous paper. A better description or more items? }
+
 Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
@@
 Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
-\liyi{might be cut because no more description on this. }
 Rule \textsc{S-FunT} defines the tainted version of function call semantics.
@@
 
-\subsection{Typing}\label{sec:type-system}
-
-%% DVH: covered above.
-%% Any \checkedc expression is a word type object ($\tau$ in Fig.~\ref{fig:checkc-syn}), which is either an \tint{} or a pointer.
-A pointer can be either a word type $\tau$, an array $\tarray{b}{b}{\tau}$, or an NT-array $\tntarray{b}{b}{\tau}$.
-Each pointer type is associated with a mode $m$ indicating whether the pointer is checked (\cmode) or unchecked ($\umode$).
-
-We now turn to the \lang type system.
-% The design of the type
-% system has been carefully constructed to ensure the expected
-% properties of progress and preservation, but also so that the
-% type-based compilation strategy detailed later is correct and
-% economical in its representation of pointers.
-%
-The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
-which states that in a type environment $\Gamma$ (mapping variables to
-their types) and a predicate environment $\Theta$ (mapping integer-typed
-variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-in context mode $m$. Key rules for this judgment are given in
-Fig.~\ref{fig:type-system-1}.
-% \review{text says $u < c$ but this seems to contradict the paragraph right
-%   after, because if $m = u$ then $u <= c$ and it seems like T-DefArr *does*
-%   allow me to dereference a checked pointer in unchecked mode...? Or am I
-%   missing something?}
-% \liyi{ Why it will be a problem in dereferencing a checked pointer in unchecked mode?
-%    unchecked regions contain history C code that is not translated to Checked-C syntax, but it could also mean that the translation is under construction, so that we are free to put checked-C code there, and we gaurantee that the checked-pointers in unchecked-mode are also spatially safe.
-The opposite is not true, where you cannot use (dereference/assign) a unchecked-pointer in a checked mode.
-}
-All remaining rules
-are given in
-\iftr
-Appendix~\ref{sec:literal-pointer-typing}~and~\ref{rem-type}.
-\else
-the supplemental report~\cite{checkedc-tech-report}.
-\fi
-
-\begin{DIFnomarkup}
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-   \inferrule[T-ConstU]
-       { \neg \cmode(\tau)}
-       {\Gamma;\Theta\vdash_u \evalue{n}{\tau} : \tau}
-
-   \inferrule[T-ConstC]
-       {\Theta;\heap;\emptyset \vdash_c n : \tau}
-       {\Gamma;\Theta\vdash_c \evalue{n}{\tau} : \tau}
-
-    \inferrule[T-Def]
-              {\xi \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-
-    \inferrule[T-AssignArr]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{\xi}\\\\
-                \Gamma; \Theta \vdash_m e_2 : \tau' \\\\
-                \tau'\sqsubseteq_{\Theta} \tau\\
-                \xi \leq m}
-              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
-
-     \inferrule[T-CastPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
-
-   \inferrule[T-Let]
-    { x\not\in \fv(\tau') \\
-        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
-          \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
-             }
-    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'}
-
-   \inferrule[T-LetInt]
-    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\\\
-        \Gamma;\Theta \vdash_m e_1 : \tint \\\\
-           \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
-             }
-    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
-
-   \inferrule[T-RetInt]
-    { \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
-    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tint}}{e} : \tau}
-
-    \inferrule[T-Checked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
-              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
-
-    \inferrule[T-Unchecked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
-
-\inferrule[T-Fun]
-    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\\\
-         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
-         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
-             \overline{\tau'} \sqsubseteq_{\Theta}
-               \overline{\tau}[\overline{e'} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
-
-  \end{mathpar}
-}
-% {\footnotesize
-% \begin{center}
-% $
-% \begin{array}{l}
-% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
-% \\[0.2em]
-% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
-% \end{array}
-% $
-% \end{center}
-% }
-{\footnotesize
-\[
-\begin{array}{l} 
-\cmode(\tint)=\texttt{false}
-\qquad
-\cmode(\tptr{\omega}{\cmode})=\texttt{true}
-\qquad
-\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
-\end{array}
-\\
-\]
-}
-\caption{Selected type rules}
-\label{fig:type-system-1}
-\end{figure*}
-\end{DIFnomarkup}
-
-\myparagraph{Pointer Access}
-%
-\textsc{T-Def} and \textsc{T-AssignArr} rules examine array
-dereference and assignment operations respectively, returning the type of
-pointed-to objects. Rules for pointers for other object types are
-similar.
-The condition $m\le m'$ ensures that checked and unchecked pointers 
-can only be dereferenced in checked and unchecked regions, respectively;
-The type rules do not attempt to reason whether the access is in bounds;
-such check is deferred to the semantics.
-
-% Subtyping and casting operations are briefly introduced in
-% Sec.~\ref{sec:intros}~and~\ref{sec:overview}.  Subtyping is useful in
-% static casting operations that allow users to view a pointer in one
-% type as another, such as casting an NT-array pointer to an array
-% one. \checkedc provides a set of safe static casting operations that
-% have no cost in execution.  Moreover, subtyping acts as oracles for
-% bound widening and dynamic casting operations; thus, \checkedc is
-% different from a complete static array pointer bound system.  For
-% example, if $e$ has type $\tau'$ and $\varphi$ is the current stack
-% snapshot, the semantics of $\edyncast{\tau}{e}$ does not transition to
-% an error state when $\varphi(\tau')\sqsubseteq\varphi(\tau)$.  In a
-% function call, for every argument, \lang permits users to input a
-% subtype entity and we prove that this does not affect the correctness
-% of the program.
-
-\myparagraph{Type Equality and Subtyping and Casting}
-%
-In \lang, type equality $\tau=_{\Theta}\tau'$
-is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
-and the alpha equivalence of two function types in \Cref{fig:checkc-subtype}.
-Two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
-$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
-$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
-are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
-and the substitution results are equal.
-
- The \textsc{T-CastPtr} rule
-permits casting from an expression of type $\tau'$ to a checked pointer when
-$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and is built on the type equality
-($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
-rules ensure the relation is transitive. Most of the rules manage
-casting between array pointer types. The second rule 
-$0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
-\tarrayptr{b_l}{b_h}{\tau}{m}$ permits treating a singleton
-pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
-Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
-if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
-There is another casting rule in \Cref{app:main} stating that
- users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
-\liyi{why? }
-
-Since bounds expressions may
-contain variables, determining assumptions like $b_l \leq b_l'$
-requires reasoning about the probable values of these variables'. The type
-system uses $\Theta$ to make such reasoning more precise.
-\liyi{why?}
-$\Theta$ is a map from variables $x$ to
-equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
-It maps variables to equations that are recorded along the type checking procedure.
-If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
-$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
-such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
-\iftr
-  Appendix~\ref{app:le}.
-\else
-  the supplemental report~\cite{checkedc-tech-report}.
-\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
-
-\begin{DIFnomarkup}
-\begin{figure}
-{\small
-\[\hspace*{-1.2em}
-\begin{array}{l}
-\text{Bound Inequality and Equality:}\\[0.3em]
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
-     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
-     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
-     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
-     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
-     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
-    \end{array}
-  \\[0.7em]
-\text{Type Equility:}\\[0.3em]
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-     && \tint & =_{\Theta} & \tint\\
-     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
-     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
-             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
-
-    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
-
- &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
-                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
-    \end{array}
-  \\[0.7em]
-\text{Subtype:}\\[0.3em]
-
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
-
-    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
-    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
-    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
-    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
-\\[0.2em]
-\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
-
-    \end{array}
-\end{array}
-  \]
-}
-{\footnotesize
-\[
-\begin{array}{l}
-n'+n = add(n',n)
-\qquad
-(x+n')+n = x+add(n',n)\\
-\textit{cond}(\overline{x},\tau,\overline{y},\tau')
-=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
-  \wedge \overline{y}\cupdot\overline{z}
-  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
-\\\qquad\qquad\qquad\qquad\qquad
-  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
-\end{array}
-\]
-}
-  \caption{Type Equality and Subtyping}
-  \label{fig:checkc-subtype}
-\end{figure}
-\end{DIFnomarkup}
-
-\begin{DIFnomarkup}
- \begin{figure}[t]
- {\small
-
- \begin{mathpar}
-   \inferrule
-       {}
-       {\Theta;\heap;\sigma \vdash_m n : \tint}
-
-   \inferrule
-       {}
-       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
-
-   \inferrule
-       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
-       {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
-
-   \inferrule
-       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
-
-   \inferrule
-       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
-            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
-
-   \inferrule
-       { \xi \le m 
-     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
-       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
-
-   \inferrule
-       {\neg\funptr(\omega)\\ \xi \le m\\
-        \forall i \in [0,\size(\omega)) \;.\;
-            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
- \end{mathpar}
- }
-{\footnotesize
-\[
-\begin{array}{l} 
-\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
-\qquad
-\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
-\end{array}
-\\
-\]
-}
- \caption{Verification/Type Rules for Constants}
- \label{fig:const-type}
- \end{figure}
-\end{DIFnomarkup}
-
-\myparagraph{Constant Validity}
-Rules \textsc{T-ConstU} and \textsc{T-ConstC}
-describes type assumptions for constants appearing in a program.
-$\cmode(\tau)$ judges that a constant pointer 
-in an unchecked region cannot be of a checked type,
-which represents an assumption that programmers 
-cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
-In rule \textsc{T-ConstC}, we requires a static 
-verification procedure for validating a constant pointer, 
-which is similar to the dynamic verification process in \Cref{sec:semantics}. 
-\liyi{what dynamic verification process? }
-
-The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
-validate the constant $\evalue{n}{\tau}$, 
-where $\heap(m)$ is the initial heap that the constant resides on and
-$\sigma$ is a set of constant assumed to be checked.
-A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
-\liyi{Do we discuss this $\Xi(m)$ before? }
-A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
-and the address stores a function with the right type.
-\liyi{how? }
-The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
-where every element in the pointer range ($[0,\size(\omega))$) should be well typed.
-\liyi{only well typed? Can we have a better way to say this? }
-
-A checked pointer checks validity in type step as rule \textsc{T-ConstC},
-while a tainted/unchecked pointer does not check for such during the type checking.
-Tainted pointers are verified through the validity check in dynamic execution as we mentioned above.
-\liyi{This is saying that the validity checks and verification are the same, but the word ''verification'' could be misleading, so maybe change it? }
-
-\liyi{I think unchecked/checked blocks and function calls below can be merged, 
-   because this section is small and might not describe the needed items. }
-
-\myparagraph{Unchecked and Checked Blocks}
-%
-During the type checking,
-Both $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$
-check all free variables in $e$ are within $\overline{x}$;
-the types for $\overline{x}$ and the final return type $\tau$ have no checked pointers.
-\liyi{why no? This should be emphasized more. }
-A $\echeckedtext$ or $\euncheckedtext$ block represents 
-the context switching from a checked to an checked region, or vice versa.
-We need to make sure no checked pointers are information exposed to unsafe code regions.
-\liyi{example? }
-
-\myparagraph{Let Bindings and Dependent Function Pointers}
-%
-Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
-type dependency. 
-\liyi{The type dependency means the scope of type variables. Might not need, because we have said this in previous paper. }
-In particular, the result of evaluating a $\elettext$
-may have a type that refers to one of its bound variables (e.g., if
-the result is a checked pointer with a variable-defined bound). If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
-Note that we restrict the expression $e_1$ to syntactically match the
-structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
-
-Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
-$\erettext$ does not appear in source programs but is introduced by the semantics when
-evaluating a let binding (rule \textsc{S-Let} in
-Fig.~\ref{fig:semantics}). This rule is needed for the preservation
-proof.
-\liyi{why? }
-% After the evaluation of a let binding a variable $x$ concludes,
-%we need to restore any prior binding of $x$, which is either
-%$\bot$ (meaning that there is no $x$ originally) or some value
-%$\evalue{n}{\tau}$.
-
-Rule \textsc{T-Fun} is the dependent function call rule. 
-Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
-from a type-check for $e$ and the types $\overline{\tau'}$ from the argument type checks for $\overline{e}$,
-we confirm that each of $\overline{\tau'}$ is
-a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$,
-which replaces possible integer bound variables $\overline{x}$ with bound expressions $\overline{e'}$.
-The final result type is the defined target type $\tau$ appearing in the function pointer type
-also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
-\liyi{should keep it uniformly with respect to the one above. 
-    It is also not so clear and too much wording here. }
-Consider the \code{safe_strcat} function in
-Fig.~\ref{fig:strcat-ex}; its parameter type for \code{dst} 
-depends on \code{n}. The \textsc{T-Fun} rule will substitute 
-\code{n} with the argument at a call-site.
-
-\subsection{Type Soundness, Non-exposure, Non-crashing}\label{sec:theorem}
+\subsection{Meta Theories}\label{sec:theorem}
 
@@
 These proofs have been conducted in our Coq model.
-
-\liyi{Can we cut some?}
 Type soundness relies on several \emph{well-formedness}:
@@
 snapshots together.
-\liyi{Tell people why only checked heap/store matter? }
 
@@
 
-\liyi{maybe some intuition why this is important. }
 Using the above theorem, we first show the non-exposure theorem,
@@
 \end{thm}
+
 %{\em Proof:} By induction on the number of steps of the \checkedc

### 2022-12-30 sandp23/evaluation.tex

--- 2022-07-16 sandp23/evaluation.tex
@@
 % \liyi{Deena? Leo? }
+Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program-benchmarks.
+MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
+We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.
+We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
+Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted\-ness of a pointer is propagated from the call site up until its declaration.
+We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
+Unlike the Runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
+This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLib functions that are in use by the tainted pointers in the program.
+Valgrind's memory figures do not account for Sandboxed allocations which happen in the shadow memory.
+All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.
 
-\begin{itemize}
+\subsection{Micro-Benchmarks}
+\subsubsection{Memory Access in SBX}
+Memory access performance between the checked and the WASM region involves crafting a test case that involves a simple pointer arithmetic operation enclosed in a loop of 100k iterations.
+156.6\% overhead with \systemname is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.
 
-\item provide evidence that the \systemname compiler is efficient. Compare the compiler with respect to other work, like RLBox, also the previous \checkedc compiler.
+\subsubsection{Indirect-Call}
+This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call from the checked region into the SBX) to return.
+This metric is evaluated against the time taken for a checked Callee (call within local scope) to return.
+Observed overhead is once again a consequence of code executing in the WASM Sandbox and call indirection.
 
-\item provide user experience of \systemname.
-We restrict the use of checked pointers compared to previous checked-c compiler.
-Is the restriction arrangable.
-We can say that the tainted shells are auto-matically generated, so we have a mechanism for auto-generating tainted pointers if necessary.
+\subsubsection{Pointer Access}
+Pointer overhead during run-time between the pointers in the checked and Sandboxed regions is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
+34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.
 
-\item if we have space, we can re-introduce random testing a little, saying that how it helps us to develop the compiler.
+\begin{center}
+\begin{tabular}{||c c||} 
+ \hline
+ Test-Name & \systemname overhead \\ [0.5ex] 
+ \hline\hline
+ Memory Access in SBX & +156\% \\
+ Indirect-Call & +398\% \\ 
+ Pointer Access & +34.16\% \\ [1ex]
+ \hline 
+\end{tabular}
+\end{center}
 
-\item we can then talk about the possible bugs we find in the \checkedc compiler for function pointer or the RLBox bugs.
+\subsection{Program Run-time Benchmarks}
+We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program.
+In our conversion efforts, we intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern-day memory vulnerabilities by re-introducing fixed memory bugs and then annotating the relevant buggy code with tainted pointers.
+However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible.
+Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
+The extent of conversion qualitatively dictates the converted program's conversion efforts required and the run-time performance.
+Consequently, each of the conversions for the six programs follows a varied approach.
+Parsons is annotated comprehensively in two variants parsons\_wasm and parsons\_tainted.
+parsons\_wasm has most of its input parsing functions moved into the sandbox, whilst having all its pointers marked as tainted.
+These sandboxed functions interact with the checked region by making indirect calls through RLBOX's callback mechanism.
+However, with parsons\_tainted, we do not move any of the functions to the sandbox but still mark all the pointers as tainted.
+The test suite itself consists of 328 tests comprehensively testing the JSON parser's functionality.
+Benchmarks for both of these forks are recorded using the mean difference between the \systemname and generic-C/checked-C variants when executing 10 consecutive iterations of the test suite.
+parsons\_wasm expectedly shows 200/266\% runtime overhead when evaluated against checked-c and generic-c respectively due to the performance limitation of WebAssembly.
+However, evaluating parsons\_tainted against checked-c shows \systemname to be faster because \systemname by itself performs lighter run-time-instrumentation on tainted pointers as compared to the run-time bounds checking performed on checked pointers by checked-c.
+Furthermore, we only see an average peak memory of 9.5 KiB as compared to the anticipated 82 KiB overhead as Valgrind does not consider the WASM Shadow memory allocated to the tainted pointers.
 
-\end{itemize}
+We intended to keep \systemname changes for Lib-png comparatively narrow in scope by beginning with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read().
+However, we also annotate sections of Lib-png that involve reading, writing, and image processing (interlace, intrapixel, etc) on user-input image data as tainted.
+That is, rows of image bytes are read into tainted pointers and the taintedness for the row\_bytes is propagated throughout the program.
+All our changes extend to the png2pnm and pnm2png executables.
+To evaluate png2pnm, we take the mean of 10 iterations of a test script that runs png2pnm on 52 png files located within the libpng's pngsuite.
+To test pnm2png, we take the mean of 10 iterations of pnm2png in converting a 52MB 5184x3456 pixels large pnm image file to png.
+Valgrind's reported lower Heap space consumption for \systemname converted code is due to the discounted space consumed on the heap by the Sandbox's shadow memory.
+Consequently, when evaluating pnm2png, \systemname's heap consumption was 52 MB lower as the entire image was loaded onto the shadow memory.
 
+\systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
+We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr\<char\>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
+Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance.
+For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets().
+This sampling is repeated 10 times and its mean value is reflected in the below table.
+Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
+
+MicroHTTPD is a good demonstration of the practical difficulties in converting a program to \systemname.
+Our conversion for this program was aimed at sandboxing memory vulnerabilities CVE-2021-3466 and CVE-2013-7039.
+CVE-2021-3466 is described as a vulnerability from a buffer overflow that occurs in an unguarded "memcpy" which copies data into a structure pointer (struct MHD\_PostProcessor pp) which is type-casted to a char buffer (char *kbuf = (char *) \&pp[1]).
+Our changes would require making the "memcpy" safe by marking this pointer as tainted.
+However, this would either require marshaling the data pointed by this structure (and its sub-structure pointer members) pointer or would require marking every reference to this structure pointer as tainted, which in turn requires every pointer member of this structure to be tainted.
+Marshalling data between structure pointers is not easy and demands substantial marshaling code due to the spatial non-linearity of its pointer members unlike a char*.
+This did not align with our conversion goals which were aimed at making minimal changes.
+Consequently, the above CVE stands un-handled by \systemname.
+Our changes for CVE-2013-7039 involve marking the user input data arguments of this function as tainted pointers and in the interests of seeking minimal conversion changes, we do not propagate the tainted-ness on these functions.
+Following up on the chronological impossibility of sandboxing bugs before they are discovered and the general programmer intuiting, we moved many of the core internal functions (like MHD\_str\_pct\_decode\_strict\_() and MHD\_http\_unescape()) into the sandbox.
+
+\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
+CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function.
+CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function.
+Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.
+For evaluation, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
+Following this we record an entry for each of these tests as a mean of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the below table.
+
+Motivated by its size and simplicity, we intended to make comprehensive \systemname changes for Tiny-bignum by marking all the pointers as tainted.
+Furthermore, in likes of an unsafe use of sprintf() that could possibly lead to a buffer overflow within the bignum\_to\_string(), we move this function into the WASM sandbox.
+Ideally, a major portion of conversion efforts to \systemname is attributed to understanding the codebase and finding the precise extent to which we choose to propagate the taintedness or to stop and give up and marshall the data, thereby, requiring only 4 hours to convert tiny-bignum to \systemname.
+The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.
+
+\begin{center}
+\begin{tabular}{||c | p{1.3cm} | p{1.8cm}||} 
+ \hline
+ Program & Run-Time & Avg Peak Memory \\ [0.5ex] 
+ \hline\hline
+ ProFTPD & +47.8\% & +82.1 KiB  \\
+ MicroHTTPD & +171.7\% & +24.7 KiB\\ 
+ UFTPD & +2.7\% & +86.3 KiB \\ 
+ Tiny-Bignum (vs checked-c) & -23.4\% &  +81.1 KiB \\
+ Tiny-Bignum &  +121.7\% &  +81.1 KiB \\
+ parsons\_wasm (vs checked-c) & +200.2\% & +9.8 KiB\\ 
+ parsons\_tainted (vs checked-c) & -8.7\% & +9.2 KiB\\
+ parsons\_wasm & +266.5\% & +9.8 KiB\\ 
+ parsons\_tainted & +10.45\% & +9.2 KiB\\ 
+ LibPNG (png2pnm) & +11.41\% & -102.5 KiB\\ 
+ LibPNG (pnm2png) & +46.5\% & -51.6 MiB\\ [1ex]
+ \hline 
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{||p{2 cm} | p{1.2cm} | p{1.2cm} | p{1.0cm} | p{1.0cm}||} 
+ \hline
+ Program & Pointers Annotated  & Lines Sandboxed & CVEs fixed & Time To Port \\ [0.5ex] 
+ \hline\hline
+ ProFTPD & 6 & 0 & - & 1 Day  \\
+ \hline
+ MicroHTTPD & 139 & 650 & 1 & 4 Days \\
+ \hline
+ UFTPD & 146 & 90 & 2 & 3 Days \\ 
+ \hline
+ LibPNG & 248 &  0 & 1 & 8 Days \\
+ \hline
+ Tiny-Bignum &  69 &  30 & - & 4 Hours \\
+ \hline
+ parsons\_wasm & 364 & 800 &  - & 3 Days \\ 
+ \hline
+ parsons\_tainted & 378 & 0 & - & 8 Days \\ [1ex]
+ \hline 
+\end{tabular}
+\end{center}
 % 

### 2022-12-30 sandp23/overview.tex
add comments
update.

@@
-\section{Overview}
+\section{\systemname Compiler Overview}
 \label{sec:overview}
@@
 
-We use the example of a simple network server in~\lst{lst:comb} to explain the overview of developers using~\systemname{}.
+\Cref{fig:overview} provides the compiler structure of \systemname.
+The \systemname type checker compiles a \systemname program by partitioning the different program regions, i.e., $c$ and $uc$ regions, as well as different pointers, i.e., checked, unchecked and tainted pointers.
+Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler.
+
+\Cref{sec:formal} provides the formalism of the different compiler components.
+Here,
+we use the example of a simple network server in~\lst{lst:comb} to explain the \systemname compiler.
 The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
@@
 The variable~\code{i} can take any integer value as it is parsed from~\code{msg}, whose contents are read from socket in~\code{read_msg} indicated by~\rootcause.
+
+\liyi{The example looks good. But we are focusing on converting C to \checkedc here? }
 
@@
 However, as explained in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
+
+\liyi{Can we just assume that programmers know \checkedc and will use \checkedc, and then we added tainted type? }
 

### 2022-12-30 sandp23/formal.tex
update.

@@
-\section{\systemname Type System and Formalism}\label{sec:formal}
+\section{\systemname Formalism}\label{sec:formal}
 
@@
 This section describes the formal model of \systemname, named
-\lang, making precise its syntax, semantics, and type system. It also
-develops \lang's meta-theory, including the type soundness, non-exposure, and non-crashing
-theorems.
+\lang, making precise its syntax, semantics, type system, and compilation. It also
+develops \lang's meta-theory, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
 
@@
 In this case, we also do not need to increase the upper-bound of the type of $x$.}
+
+\subsection{Compilation}\label{sec:compilation}
+
+As we have shown in \Cref{fig:overview}, the \systemname compiler utilizes the sandbox mechanism \cite{rul2009towards} and the \checkedc compiler \cite{li22checkedc} to compile programs.
+Here, we introduce how \systemname compiles a program into these two components.
+
+\begin{figure}[t!]
+{\small
+\hspace*{-0.5em}
+\begin{tabular}{|c|c|c|c|}
+\hline
+& \cmode & \tmode & \umode \\
+\hline
+& \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} \\
+\hline
+\cmode & $\estar{x}$ / $\getstar{\cmode}{x}$ 
+ & $\texttt{sand\_get}(x)$ / $\getstar{\umode}{x}$ &  $\times$ \\
+\hline
+\umode & $\times$
+ & $\estar{x}$ / $\getstar{\umode}{x}$ &  $\estar{x}$ / $\getstar{\umode}{x}$ \\
+\hline
+\end{tabular}
+
+}
+\caption{Compiled Targets for Dereference}
+\label{fig:flagtable}
+\end{figure}
+
+In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
+i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
+Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
+In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different. 
+\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
+The columns represent different pointer modes and the rows represent context modes.
+For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation.
+In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
+If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
+
+This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
+with no loss of expressiveness
+as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
+For the compiler formalism, 
+we present a compilation algorithm that converts from
+\lang to \elang, an untyped language without metadata
+annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
+In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
+$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$.
+The algorithm sheds
+  light on how compilation can be implemented in the real Checked C
+  compiler, while eschewing many vital details (\elang has many 
+  differences with LLVM IR).
+
+%This section shows how \systemname deals with 
+%annotations can be safely erased: using static information a compiler
+%can insert code to manage and check bounds metadata, with no loss of
+%expressiveness. We present a compilation algorithm that converts from
+%\lang to \elang, an untyped language without metadata
+%annotations. The syntax and semantics \elang
+  %closely mirrors that of \lang; it differs only in that literals lack
+  %type annotations and its operational rules perform no
+  %bounds and null checks, which are instead inserted during
+  %compilation. Our compilation algorithm is evidence that \lang's
+  %semantics, despite its apparent use of fat pointers, faithfully
+  %represents Checked C's intended behavior. The algorithm also sheds
+  %light on how compilation can be implemented in the real Checked C
+  %compiler, while eschewing many important details (\elang has many 
+  %differences with LLVM IR).
+
+Compilation is defined by extending \lang's
+typing judgment as follows:
+\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
+There is now a \elang output $\dot e$ and an input $\rho$, which maps
+each (NT-)array pointer variable to its mode and
+each variable \code{p} to a pair of \emph{shadow
+  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
+These may differ from the bounds in \code{p}'s type due to bounds
+widening.\footnote{Since lower bounds are never widened, the
+  lower-bound shadow variable is unnecessary; we include it for uniformity.}
+
+% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
+% complete judgment, implicitly assuming that $e$ is a well-typed and closed
+% term.
+
+We formalize rules for this judgment in PLT Redex~\cite{pltredex},
+following and extending our Coq development for \lang. To give
+confidence that compilation is correct, we use Redex's property-based
+random testing support to show that compiled-to $\dot e $ simulates
+$e$, for all $e$.
+
+\myparagraph{Checked and Unchecked Blocks}
+%
+In the \systemname implementation,
+$\euncheckedtext$ and $\echeckedtext$ blocks 
+are compiled as context switching functions provided by the sandbox mechanism.
+We compile $\eunchecked{\overline{x}}{e}$ to 
+$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
+to execute expression $e$ with the arguments $\overline{x}$.
+$\echecked{\overline{x}}{e}$ is compiled to 
+$\texttt{callback}(\overline{x},e)$, where we perform 
+a \texttt{callback} to a checked block code $e$ inside a sandbox.
+In \systemname, we adopt an aggressive execution scheme that
+directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
+In the formalism, we rely on the type system to 
+guarantee the context switching without creating the extra function calls for simplicity.
+
+%Fig.~\ref{fig:compilationexample} shows how an invocation of
+%\code{strlen} on a null-terminated string is compiled into C
+%code. Each dereference of a checked pointer requires a null check
+%(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
+%compiler makes explicit: Line~$3$ of the generated code has the null
+%check on pointer \code{p} due to the \code{strlen},
+%  and a similar check happens
+%  at line~$8$ due to the pointer arithmetic on \code{p}.
+%Dereferences also require bounds checks: line~$2$ checks \code{p} is
+%in bounds before computing \code{strlen(p)}, while line~$10$ does
+%likewise before computing \code{*(p+1)}.
+
+\myparagraph{Function Pointers and Calls}
+%
+Function pointers are managed similarly to normal pointers,
+but we insert checks to check if the pointer address is not null in 
+the function store instead of heap, and whether or not the type is correctly represented, 
+for both $\cmode$ and $\tmode$ mode pointers 
+\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
+For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
+we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
+and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
+The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
+is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
+
+\begin{figure}[t!]
+  \begin{small}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+int deref_array(n : int,
+     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
+     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
+  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
+  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
+    * p;
+    * q = 1;
+}
+...
+/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
+/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\begin{frame}
+
+\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
+
+\end{frame}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+deref_array(int n, int* p, int * q) {
+  //m is the current context mode
+  let p_lo = 0; let p_hi = n; 
+  let q_lo = 0; let q_hi = n; 
+  /* runtime checks */
+  assert(p_lo <= 0 && 0 <= p_hi);
+  assert(p != 0);
+  *(mode(p) $\wedge$ m,p);
+  verify(q, not_null(m, q_lo, q_hi) 
+             && q_lo <= 0 && 0 <= q_hi);
+  *(mode(q) $\wedge$ m,q)=1;
+}
+...
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\end{small}
+    \caption{Compilation Example for Dependent Functions}
+\label{fig:compilationexample1}
+\end{figure}
+
+For compiling dependent function calls,
+\Cref{fig:compilationexample1} provides a hint.
+Notice that the bounds for the array pointer \code{p} are not passed as
+arguments. Instead, they are initialized according to \code{p}'s
+type---see line~4 of the original \lang program at the top of the figure.
+Line~$3$ of the generated code
+sets the lower bound  to \code{0} and the
+upper bound to \code{n}.
 
@@
 
+% \review{ 
+% My interpretation of section IV.C is that the authors have a pen-and-paper
+%   proof of Theorem 4, and that random testing using the Redex models has been
+%   used to gain confidence in such a proof. Is this correct?}
+% \liyi{Yes, this is correct.}
+% \mwh{No, not correct: We have no pen-and-paper proof, just the
+%   model. Clarify here}
+% showing that
+% extensive random testing fails to falsify the bisimilarity
+% property. \mwh{Better way to state the previous?}
+
+In the \systemname compiler, we also restore the compiler simulation theorem, such that executing a \systemname program is simulated by executing the compiled LLVM program, which is introduced in the \checkedc formalism \cite{li22checkedc} and \Cref{appx:comp1}.
+
+We formalize both the compilation procedure and the simulation
+theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
+and then attempt to falsify it via Redex's support for random
+testing. Redex allows us
+  to specify compilation as logical rules (an extension
+  of typing), but then execute it algorithmically to
+  automatically test whether simulation holds. This process revealed
+  several bugs in compilation and the theorem statement.
+%
+  % us gain confidence that our original pen and paper proof of
+  % simulation remains true with the addition of variable bounds. }
+We ultimately plan to prove simulation in the Coq model.
+
+%Turning to the simulation theorem: We first introduce notation
+%used to specify the theorem.
+We use the notation $\gg$ to
+indicate the \emph{erasure} of stack and heap---the rhs is the same as
+the lhs but with type annotations removed:
+\begin{equation*}
+  \begin{split}
+    \heap  \gg & \dot \heap \\
+    \varphi \gg & \dot \varphi
+  \end{split}
+\end{equation*}
+In addition, when $\Gamma;\emptyset\vdash
+\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
+\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
+and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively.
+$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
+Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
+reduction relation of $\elang$. Unlike the $\lang$, the semantics of
+$\elang$ does not distinguish checked and unchecked regions.
+
+Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
+the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
+that is similar to the one by~\citet{merigoux2021catala}.
+
+An ordinary simulation property would
+replace the middle and bottom parts of the figure with the
+following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
+  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
+Instead, we relate two erased configurations using the relation $\sim$,
+which only requires that the two configurations will eventually reduce
+to the same state.
+
+\ignore{
+ We formulate our simulation theorem differently
+because the standard simulation theorem imposes a very strong
+syntactic restriction to the compilation strategy. Very often, $(\dot
+\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
+semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
+but we are unable to syntactically equate the two configurations due
+to the extra binders generated for dynamic checks and ANF
+transformation. In earlier versions of the Redex model, we attempted
+to change the compilation rules so the configurations could match
+syntactically. However, the approach scaled poorly as we added
+additional rules. 
+This slight relaxation on the equivalence relation
+between target configurations allows us to specify compilation more
+naturally without having to worry about syntactic constraints.
+}
+
+% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
+\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
+and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
+
+\begin{itemize}
+
+\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\varphi,\dot \heap,\dot e)$ and $(\dot
+\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+\dot \heap,\dot e)$.
+
+\item if $r_1 = \ebounds$ or $\enull$, then we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+\varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+\end{itemize}
+\end{thm}
+
+% when $r_1 = e_1$ for
+% some $e_1$ and
+% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
+% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
+% some $e_1$ and
+% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
+% , then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+% \end{thm}
+
+As our random generator never generates
+$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
+as it relates to checked code. This limitation makes it
+unnecessary to state the other direction of the simulation theorem
+where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
+that $e_0$ will never enter a stuck state if it is well-typed in
+checked mode.
+
+The current version of the Redex model has been tested against $23000$
+expressions with depth less than $11$. Each expression can
+reduce multiple steps, and we test simulation between every two
+adjacent steps to cover a wider range of programs, particularly the
+ones that have a non-empty heap.
+
+\begin{figure}[t]
+{\small
+\[
+\begin{array}{c}
+\begin{tikzpicture}[
+            > = stealth, % arrow head style
+            shorten > = 1pt, % don't touch arrow head to node
+            auto,
+            node distance = 3cm
+        ]
+
+\begin{scope}[every node/.style={draw}]
+    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
+    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
+    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
+    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
+    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
+\end{scope}
+\begin{scope}[every edge/.style={draw=black}]
+
+    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
+    \path [<->] (A) edge node {$\gg$} (C);
+    \path [<->] (B) edge node {$\gg$} (D);
+    \path [dashed,<->] (C) edge node {$\sim$} (D);
+    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
+    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
+\end{scope}
+
+\end{tikzpicture}
+\end{array}
+\]
+}
+\caption{Simulation between \lang and \elang }
+\label{fig:checkedc-simulation-ref}
+\end{figure}
+
 %{\em Proof:} By induction on the number of steps of the \checkedc

### 2023-01-03 sandp23/appendix.tex

@@
-\section{Appendix}
+\section{Appendix}\label{app:main}
 
@@
 \subsection{The Compilation Rules}\label{appx:comp1}
-Fig.~\ref{fig:syntaxerased} and Fig.~\ref{fig:semanticserased} shows
-the syntax for $\elang$, the target language for compilation. We syntactically
-restrict the expressions to be in A-normal form to simplify the
-presentation of the compilation rules. In the Redex model, we
-occasionally break this constraint to speed up the performance of
-random testing by removing unnecessary let bindings.
-To allow explicit runtime
-checks, we include $\ebounds$ and $\enull$ as part of $\elang$
-expressions which, once evaluated, result in an corresponding error state. $\eassignstack{x}{\dot a}$ is a new syntactic form that
-modifies the stack variable $x$ with the result of $\dot a$. It is
-essential for bounds widening. $\leq$ and $-$ are introduced
-to operate on bounds and decide whether we need to halt with a bounds error
-or widen a null-terminated string.
-
-\begin{figure}[h]
+
+% \review{
+% The first reviewer said:
+% \begin{itemize}
+% \item I had a really hard time understanding precisely the "formalized" compilation
+%   from CoreChkC to CoreC. Specifically: is CoreC intended to model LLVM-IR, or
+%   is it a subset of C? Is CheckedC compiling to C or is it a new frontend like
+%   `clang`? See remark about undefined behavior (UB) below which got me even more
+%   confused. \mwh{CoreC is meant to model CoreChkC but with
+%   annotations, and checks removed; it is not meant to model C or LLVM-IR.}
+% \item Is the compilation scheme from CoreChkC to CoreC is faithful to the actual
+%   compilation scheme of the checkedc-clang compiler to ...? I feel like the
+%   paper is missing an actual description of what happens in the compiler to
+%   allow us to connect the dots and understand how the formalization illuminates
+%   the implementation. \mwh{We did not model compilation on the real
+%   implementation; our purpose was to show that annotations in CoreChkC
+%   do not necessitate fat pointers in an implementation; that said,
+%   our formalization does show how a real implementation can be
+%   carried out}
+% \item The paper doesn't seem upfront about what is *shown* (theorem in Coq) and what
+%   is *tested* (via PLT-Redex), and thus remains a
+%   conjecture. \mwh{Updated the intro and the individual sections}
+% \item  Missing discussion of why Coq vs. PLT-Redex, effort involved, any plans to
+%   formally prove compilation from CoreChkC to CoreC, any hopes of integrating
+%   that in the official implementation, etc. \mwh{Added note at the
+%   start of section 3.}
+% \end{itemize} }
+
+% \review{IV: ghost variables in other contexts (e.g. Why3, Dafny) are used for things
+%   that do not exist at run-time, but this doesn't seem to be the case here.}
+% \yiyun{Agreed: We changed the name to ``shadow variables'' to avoid confusion}
+% \review{Do your bug report and github links break anonymity?}
+
+% \review{From reviewer C:
+% \begin{itemize}
+% \item I'd like to have seen a bit more motivation for using PLT redex:
+%   what aspects made the use of this tool preferable to formulating the
+%   compilation in Coq and using Quickchick to do random testing of the
+%   simulation result.
+%   \mwh{Added some text to the end of III.A and IV.C}
+% \yiyun{Some reasons I can think of:\begin{itemize}
+%     \item Redex is highly optimized for specifying judgments that are algorithmic. By writing down a typing relation, we can immediately obtain a typechecker
+%     that is executable.
+Same applies for the small-step evaluation relation.
+Translating the relations into functions in Coq is definitely doable but time-consuming, especially since compilation is embedded as part of the typing rules.
+It is also hard to see whether the function we define really corresponds to the relation unless we formally prove it.
+This issue is particularly relevant at the early stage of the development when the compilation rules were buggy and the simulation property was violated often as we added new generator cases.
+%     In Redex, we don't have any formal guarantee either, but at least we can more easily see the correspondence because Redex is able to convert the relation into an executable version so we can specify the relation literally.
+This feature of Redex helped us speed up our development significantly when our compilation rules were constantly changing.
+% \end{itemize} }
+% \item there's a funny change of line spacing in column 2 of page 6, about 2/3 down.
+% \end{itemize}
+% }
+
+The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
+that track array bound information, which is used in premises
+of rules like \textsc{S-DefArray} and
+  \textsc{S-AssignArr} to prevent spatial safety violations.
+The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
+for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
+The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.
+
+\begin{figure}[t!]
 {\small
-  \[\begin{array}{llcl}
-      \text{Atoms}       & \dot a & ::= & n \mid x \\
-
-      \text{C-Expressions} & \dot c & ::= & \dot a \mid \estrlen{\dot a}  \mid \emalloc {\dot  a}  \mid \ecall{f} {\overline{\dot a}}  \\
-                         & & \mid &  \mid  \dot a \circ \dot a \mid \estar{\dot a} \\
-                         & & \mid & \eassign{\dot a}{\dot a} \mid \eassignstack{x}{\dot a}  \mid \eif{\dot a}{\dot e}{\dot e} \\
-                         & & \mid & \ebounds \mid \enull \\
-      \text{Expressions}  & \dot e & ::= & \dot c \mid \elet{x}{\dot c}{\dot e}\\
-      \text{Binops} & \circ & ::= & + \mid - \mid \leq  \\
-      \text{Closure} & \dot C & ::= & \hole \mid \elet{x}{\dot a}{\dot C} \\
-       & & \mid & \eif{\dot a}{\dot e}{\dot C} \mid \eif{\dot a}{\dot    C}{\dot e} \\
-      \text{Bounds Map} & \rho & \in & \texttt{Var} \rightharpoonup \texttt{Var} \times \texttt{Var}
+\hspace*{-0.5em}
+\begin{tabular}{|c|c|c|c|}
+\hline
+& \cmode & \tmode & \umode \\
+\hline
+& \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} \\
+\hline
+\cmode & $\estar{x}$ / $\getstar{\cmode}{x}$ 
+ & $\texttt{sand\_get}(x)$ / $\getstar{\umode}{x}$ &  $\times$ \\
+\hline
+\umode & $\times$
+ & $\estar{x}$ / $\getstar{\umode}{x}$ &  $\estar{x}$ / $\getstar{\umode}{x}$ \\
+\hline
+\end{tabular}
+
+}
+\caption{Compiled Targets for Dereference}
+\label{fig:flagtable}
+\end{figure}
+
+In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
+i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
+\lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
+Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
+In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different.
+\lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
+\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
+The columns represent different pointer modes and the rows represent context modes.
+For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation.
+In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
+If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
+
+This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
+with no loss of expressiveness
+as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
+For the compiler formalism, 
+we present a compilation algorithm that converts from
+\lang to \elang, an untyped language without metadata
+annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
+In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
+$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
+The algorithm sheds
+  light on how compilation can be implemented in the real Checked C
+  compiler, while eschewing many vital details (\elang has many 
+  differences with LLVM IR).
+
+%This section shows how \systemname deals with 
+%annotations can be safely erased: using static information a compiler
+%can insert code to manage and check bounds metadata, with no loss of
+%expressiveness. We present a compilation algorithm that converts from
+%\lang to \elang, an untyped language without metadata
+%annotations. The syntax and semantics \elang
+  %closely mirrors that of \lang; it differs only in that literals lack
+  %type annotations and its operational rules perform no
+  %bounds and null checks, which are instead inserted during
+  %compilation. Our compilation algorithm is evidence that \lang's
+  %semantics, despite its apparent use of fat pointers, faithfully
+  %represents Checked C's intended behavior. The algorithm also sheds
+  %light on how compilation can be implemented in the real Checked C
+  %compiler, while eschewing many important details (\elang has many 
+  %differences with LLVM IR).
+
+Compilation is defined by extending \lang's
+typing judgment as follows:
+\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
+There is now a \elang output $\dot e$ and an input $\rho$, which maps
+each (NT-)array pointer variable to its mode and
+each variable \code{p} to a pair of \emph{shadow
+  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
+These may differ from the bounds in \code{p}'s type due to bounds
+widening.\footnote{Since lower bounds are never widened, the
+  lower-bound shadow variable is unnecessary; we include it for uniformity.} 
+\lc{These two sentences are confusing. }
+
+% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
+% complete judgment, implicitly assuming that $e$ is a well-typed and closed
+% term.
+
+We formalize rules for this judgment in PLT Redex~\cite{pltredex},
+following and extending our Coq development for \lang. To give
+confidence that compilation is correct, we use Redex's property-based
+random testing support to show that compiled-to $\dot e $ simulates
+$e$, for all $e$.
+
+% We developed a \checkedc compiler to compile a \checkedc program to a C program.
+% \mwh{We formalized compilation from CoreChkC to a version of CoreChkC
+%   but with the metadata removed, right? This is not a Checked C
+%   compiler. You go on to see stuff about CompCert, CLight,
+%   etc. This is confusing. We should be talking about how this relates
+%   to what was just presented. Pick definitive names for things. }
+% Given a \checkedc program $e$, we build a compilation process ($\gg$), such that $e \gg \dot e$, where $\dot e$ is the corresponding C program for $e$ in A-normal form (ANF). 
+% The compilation process ($\gg$) relies on the type checking step $\Gamma;\Theta\vdash_m e:\tau$. 
+% Especially, it relies on $\Gamma$ to provide the type information for variables in $e$. 
+% We utilize CompCert/CLight syntax and semantics \cite{Leroy:2009:FVC:1666192.1666216,Blazy2009} as our translation target language.
+%  Besides, we defined two data structures in the CompCert format for representing $\enull$ and $\ebounds$ states.
+% We write $\xrightarrow{c}$ for the semantics of CLight. \mwh{Of our
+%   target language, which I presume does not have all the features that
+%   CLight has. Should mention up front that we did all of this in PLT Redex.}
+
+Due to space constraints, we explain the rules for compilation by
+examples, using a C-like syntax; the complete rules are given in
+\iftr
+Appendix~\ref{appx:comp1}.
+\else
+the supplemental report~\cite{checkedc-tech-report}.
+\fi
+Each rule performs up to three tasks: (a) conversion of $e$ to
+A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
+and (c) generate right pointer accessing expressions based on modes.
+%
+A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
+as in the following example.
+
+{\vspace*{-0.5em}
+{\small
+\begin{center}
+$
+\begin{array}{l}
+$\code{let y=(x+1)+(6+1)}$
+\;
+\begin{frame}
+
+\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
+
+\end{frame}
+\;
+\begin{array}{l}
+$\code{let a=x+1;}$\\
+$\code{let b=6+1;}$\\
+$\code{let y=a+b}$\\
+\end{array}
+\end{array}
+$
+\end{center}
+}
+}
+
+This simplifies the management of effects from subexpressions. The
+next two steps of compilation are more interesting.
+We state them based on different \lang operations.
+
+% \review{Fig 9: if this is actual C code, then your null-check at line 6 will be
+%   eliminated by the compiler. At line 3, you performed a pointer addition, which
+%   is only defined when `p` is non-null. So, either `p` is non-null, and the
+%   NULL-check can be eliminated; or, `p` is NULL, but line 3 was undefined
+%   behavior, meaning the compiler is allowed to do anything, notably eliminate
+%   the NULL-check. This is where I am super confused, and either:
+%   - CoreC is not really the C language, and has different semantics...? but is
+%     this well-defined in the context of LLVM?
+%   - there is a problem that was not caught by the PLT-Redex-based testing.
+% \yiyun{We have clarified at the start of IV that CoreC is an untyped
+%   variant of CoreChkC, and does not aim to represent C per se, or LLVM
+%   IR. We aimed to avoid confusion by rewriting the examples in a way that is more closely
+%     related to the syntax presented in Fig 3. Pointer arithmetic between 0 and a non-zero
+%     index is always valid because CoreC there is technically only
+%     integer arithmetic.}}
+
+\myparagraph{Pointer Accesses and Modes}
+%
+In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
+if the poniter is an (NT-)array,
+we first allocate two \emph{shadow variables}
+to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
+Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
+Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
+In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
+there is an inserted verification step similar to \Cref{fig:const-type},
+which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
+Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
+The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
+ and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
+are computed based on the meet 
+operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).
+
+\myparagraph{Checked and Unchecked Blocks}
+%
+In the \systemname implementation,
+$\euncheckedtext$ and $\echeckedtext$ blocks 
+are compiled as context switching functions provided by the sandbox mechanism.
+$\eunchecked{\overline{x}}{e}$ is compiled to 
+$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
+to execute expression $e$ with the arguments $\overline{x}$.
+$\echecked{\overline{x}}{e}$ is compiled to 
+$\texttt{callback}(\overline{x},e)$, where we perform 
+a \texttt{callback} to a checked block code $e$ inside a sandbox.
+In \systemname, we adopt an aggressive execution scheme that
+directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
+In the formalism, we rely on the type system to 
+guarantee the context switching without creating the extra function calls for simplicity.
+
+%Fig.~\ref{fig:compilationexample} shows how an invocation of
+%\code{strlen} on a null-terminated string is compiled into C
+%code. Each dereference of a checked pointer requires a null check
+%(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
+%compiler makes explicit: Line~$3$ of the generated code has the null
+%check on pointer \code{p} due to the \code{strlen},
+%  and a similar check happens
+%  at line~$8$ due to the pointer arithmetic on \code{p}.
+%Dereferences also require bounds checks: line~$2$ checks \code{p} is
+%in bounds before computing \code{strlen(p)}, while line~$10$ does
+%likewise before computing \code{*(p+1)}.
+
+\myparagraph{Function Pointers and Calls}
+%
+Function pointers are managed similarly to normal pointers,
+but we insert checks to check if the pointer address is not null in 
+the function store instead of heap, and whether or not the type is correctly represented, 
+for both $\cmode$ and $\tmode$ mode pointers 
+\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
+For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
+we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
+and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
+The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
+is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
+
+For compiling dependent function calls,
+\Cref{fig:compilationexample1} provides a hint.
+Notice that the bounds for the array pointer \code{p} are not passed as
+arguments. Instead, they are initialized according to \code{p}'s
+type---see line~4 of the original \lang program at the top of the figure.
+Line~$3$ of the generated code
+sets the lower bound  to \code{0} and the
+upper bound to \code{n}.
+
+\subsection{Metatheory}
+\label{sec:meta}
+
+% \review{ 
+% My interpretation of section IV.C is that the authors have a pen-and-paper
+%   proof of Theorem 4, and that random testing using the Redex models has been
+%   used to gain confidence in such a proof. Is this correct?}
+% \liyi{Yes, this is correct.}
+% \mwh{No, not correct: We have no pen-and-paper proof, just the
+%   model. Clarify here}
+% showing that
+% extensive random testing fails to falsify the bisimilarity
+% property. \mwh{Better way to state the previous?}
+
+We formalize both the compilation procedure and the simulation
+theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
+and then attempt to falsify it via Redex's support for random
+testing. Redex allows us
+  to specify compilation as logical rules (an extension
+  of typing), but then execute it algorithmically to
+  automatically test whether simulation holds. This process revealed
+  several bugs in compilation and the theorem statement.
+%
+  % us gain confidence that our original pen and paper proof of
+  % simulation remains true with the addition of variable bounds. }
+We ultimately plan to prove simulation in the Coq model.
+
+%Turning to the simulation theorem: We first introduce notation
+%used to specify the theorem.
+We use the notation $\gg$ to
+indicate the \emph{erasure} of stack and heap---the rhs is the same as
+the lhs but with type annotations removed:
+\begin{equation*}
+  \begin{split}
+    \heap  \gg & \dot \heap \\
+    \varphi \gg & \dot \varphi
+  \end{split}
+\end{equation*}
+In addition, when $\Gamma;\emptyset\vdash
+\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
+\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
+and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively.
+$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
+Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
+reduction relation of $\elang$. Unlike the $\lang$, the semantics of
+$\elang$ does not distinguish checked and unchecked regions.
+
+Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
+the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
+that is similar to the one by~\citet{merigoux2021catala}.
+
+An ordinary simulation property would
+replace the middle and bottom parts of the figure with the
+following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
+  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
+Instead, we relate two erased configurations using the relation $\sim$,
+which only requires that the two configurations will eventually reduce
+to the same state.
+
+\ignore{
+ We formulate our simulation theorem differently
+because the standard simulation theorem imposes a very strong
+syntactic restriction to the compilation strategy. Very often, $(\dot
+\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
+semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
+but we are unable to syntactically equate the two configurations due
+to the extra binders generated for dynamic checks and ANF
+transformation. In earlier versions of the Redex model, we attempted
+to change the compilation rules so the configurations could match
+syntactically. However, the approach scaled poorly as we added
+additional rules. 
+This slight relaxation on the equivalence relation
+between target configurations allows us to specify compilation more
+naturally without having to worry about syntactic constraints.
+}
+
+% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
+\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
+and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
+
+\begin{itemize}
+
+\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\varphi,\dot \heap,\dot e)$ and $(\dot
+\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+\dot \heap,\dot e)$.
+
+\item if $r_1 = \ebounds$ or $\enull$, then we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+\varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+\end{itemize}
+\end{thm}
+
+% when $r_1 = e_1$ for
+% some $e_1$ and
+% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
+% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
+% some $e_1$ and
+% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
+% , then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+% \end{thm}
+
+As our random generator never generates
+$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
+as it relates to checked code. This limitation makes it
+unnecessary to state the other direction of the simulation theorem
+where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
+that $e_0$ will never enter a stuck state if it is well-typed in
+checked mode.
+
+The current version of the Redex model has been tested against $23000$
+expressions with depth less than $11$. Each expression can
+reduce multiple steps, and we test simulation between every two
+adjacent steps to cover a wider range of programs, particularly the
+ones that have a non-empty heap.
+
+\begin{figure}[t]
+{\small
+\[
+\begin{array}{c}
+\begin{tikzpicture}[
+            > = stealth, % arrow head style
+            shorten > = 1pt, % don't touch arrow head to node
+            auto,
+            node distance = 3cm
+        ]
+
+\begin{scope}[every node/.style={draw}]
+    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
+    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
+    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
+    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
+    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
+\end{scope}
+\begin{scope}[every edge/.style={draw=black}]
+
+    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
+    \path [<->] (A) edge node {$\gg$} (C);
+    \path [<->] (B) edge node {$\gg$} (D);
+    \path [dashed,<->] (C) edge node {$\sim$} (D);
+    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
+    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
+\end{scope}
+
+\end{tikzpicture}
 \end{array}
@@
 }
-  \caption{\elang{} Syntax}
-  \label{fig:syntaxerased}
-{\small
-  \[\begin{array}{lll}
-    \dot \mu & ::= & n \mid \bot \\
-    \dot c & ::= & \ldots  \mid \eret{x}{\dot \mu}{\dot e} \\
-    \dot H & \in & \mathbb{Z} \rightharpoonup \mathbb{Z} \\
-    \dot r & ::= & \dot e \mid \enull \mid \ebounds \\
-    \dot E & ::= & \hole \mid  \elet{x}{\dot E}{\dot e}  \mid \eret{x}{i}{\dot E} \\
-             & \mid & \eif{\dot E}{\dot e}{\dot e} \mid \estrlen{\dot E} \\
-             & \mid & \emalloc{\dot E} \mid \ecall{f}{\overline{\dot
-                      E}} \mid \dot E \circ \dot a \mid n \circ \dot E \\
-             & \mid & \estar{\dot E} \mid \eassign{\dot E}{\dot a}
-                      \mid \eassign{n}{\dot E} \mid
-                      \eassignstack{x}{\dot E} \\
-      \overline{\dot E } & ::= & \dot E \mid n,\overline{\dot E} \mid
-                                 \overline{\dot E}, \dot a
-
-  \end{array} \]
-}
-  \caption{\elang Semantic Defs}
-  \label{fig:semanticserased}
+\caption{Simulation between \lang and \elang }
+\label{fig:checkedc-simulation-ref}
 \end{figure}
 
-$\elang$ does not include any annotations. We remove structs from
-$\elang$ because we can always statically convert expressions of the form
-$\eamper{n:\tau}{f}$ into $n + n_f$, where $n_f$ is the statically
-determined offset of $f$ within the struct. We ellide the semantics of
-$\elang$ because it is self-evident and mirrors the semantics
-$\lang$. The difference is that in $\elang$, only $\ebounds$ and
-$\enull$ can step into an error state. All failed dereferences and
-assignments would result in a stuck state and therefore we rely on the
-compiler to explicitly insert checks for checked pointers.
-
-Fig.~\ref{fig:compilation} and Fig.~\ref{fig:compilation2}
- shows the rules for the compilation judgment for expressions,
-\[\Gamma;\rho \vdash e \gg \dot C, \dot a\]
-The judgment is presented differently from the one in
-Sec.~\ref{sec:compilation}, which was simplified for presentation purposes. First,
-we remove $\Theta$ and $m$ because these
-parameters are only used for checking and have no
-impact on compilation. Second, the judgment includes two
-outputs, a closure $\dot C$ and an atom expression $\dot a$, instead of
-a single $\elang$ expression $\dot e$. $\dot C$ can be intuitively understood as a
-partially constructed program or context. Whereas $\dot E$ is used for
-evaluation, $\dot C$
-is used purely as a device for compilation. As an example,
-when compiling $(1 : \tint) \plus (2 : \tint)$, 
-we would first create a fresh variable $x$, and then produce two outputs:
-\[    \dot C = \elet{x}{1 \plus 2}{\hole}\]
-\[    \dot a = x\]
-To obtain the compiled expression $\dot e$, we plug $\dot a$ into
-$\dot C$ using the usual notation $\dot C[\dot a]$. We can also use $\dot C$ to represent runtime checks, which
-usually take the form $\elet{x}{\dot c}{\hole}$, where $\dot c$
-contains the check whose evaluation must not trigger $\ebounds$ or $\enull$ for the program to
-continue (see Fig.~\ref{fig:compilationhelpers} for the metafunctions that create
-those checks).
-
-This unconventional output format enables us to separate the
-evaluation of the term and the computation that relies on the term's
-evaluated result. Since effects and reduction (except for variables) happen only
-within closures, we can precisely control the order in which effects
-and evaluation happen by composing the contexts in a specific order.
-Given two closures $\dot C_1$
-and $\dot C_2$, we write $\dot C_1[\dot C_2]$ to denote the meta
-operation of plugging $\dot C_2$ into $\dot C_1$. We also use
-$\dot C_{a;b;c}$ as a shorthand for $\dot C_a[\dot C_b[\dot C_c]]$. In
-the \textsc{C-Ind} rule,
-we first evaluate the expressions that correspond to $e_1$ and $e_2$
-through $\dot C_1$ and $\dot C_2$, and then perform a null check and
-an addition through $\dot C_n$ and $\dot C_3$. Finally, we dereference
-the result through $\dot C_4$ before returning the pair $\dot C_4,\dot
-x_4$, propagating the flexibility to the compilation rule that recursively
-calls \textsc{C-Ind}.
-
-Fig.~\ref{fig:compilationhelpers} shows the metafunctions that create
-closures representing dynamic checks. These functions first examine
-whether the pointer is a checked. If the pointer is unchecked, an empty closure $\hole$
-will be returned, because there is no need to perform a check. For bounds
-checking, there is a special case for NT-array pointers, where the
-bounds are retrived from the \textcolor{purple}{shadow} variables (found by
-looking up $\rho$) on the stack  rather than
-using the bounds specified in the type annotation. This is how we
-achieve the same precise runtime behavior as $\lang$ in our compiled expressions.
-
-Fig.~\ref{fig:compilationhelpers2} shows the metafunctions related to
-bounds widening. $\vdash_{extend}$ takes $\rho$,
-a checked NT-array pointer variable $x$, and its bounds $(b_l,b_h)$ as inputs, and returns an
-extended $\rho'$ that maps $x$ to two fresh variables $x_l$, $x_h$,
-together with a closure $\dot C$ that initializes $x_l$ and $x_h$ to
-$b_l$ and $b_h$ respectively. This function is used in the
-\textsc{C-Let} rule to extend $\rho$ before compiling the body of the
-$\elettext$ binding. The updated $\rho'$ can be used for generating
-precise bounds checks, and for inserting expressions that can potentially
-widen the upper bounds, as seen in the $\vdash_{widenstr}$
-metafunction used in the \textsc{C-Str} compilation rule.
-
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    { x_l, x_h = \fresh \\
-      \rho' = \rho[x \mapsto (x_l, x_h)]\\
-    \dot C = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}}
-  {\cextend{\rho}{x}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot C, \rho'}\\}
-
-  \inferrule[]
-  {x_l,x_h = \rho(x) \\
-    x_w = \fresh \\
-  \dot C = \elet{x_w}{\eif{x_h}{0}{\eassignstack{x_h}{1}}}{\hole}}
-{\ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C}}
-
-\inferrule[]
-{e \notin dom(\rho)}
-{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{m}}{\hole}}
-
-\inferrule[]
-{x_l,x_h = \rho(e) \\
-  x_a = \fresh \\
-  \dot C = \elet{x_a}{\eif{\dot a \leq x_h}{0}{\eassignstack{x_h}{\dot a}}}{\hole}}
-{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{\cmode}}{\dot C}}
-
-  \end{mathpar}
-
-}
- \caption{Metafunctions for Widening}
-\label{fig:compilationhelpers2}
-\end{figure*}
-
-\begin{figure*}[h]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    {x = \fresh \\
-      \dot C = \elet{x}{\eif{\dot a}{0}{\enull}}{\hole}}
-    {\echecknull{\dot a}{c}{\dot C}}
-
-    \inferrule[]
-    {}
-    {\echecknull{\dot a}{u}{\hole}} \\
-
-    \inferrule[]
-    {}
-    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}
-
-    \inferrule[]
-    { x_l,x_h = \rho(e) \\
-      x_{cl},x_{ch} = \fresh\\
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
-    }
-    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}
-
-    \inferrule[]
-    { e \notin dom(\rho) \\
-      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
-      \dot C_l = \elet{x_l}{b_l}{\hole}\\
-      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
-    }
-    {\echeckbounds{\rho}{e}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}
-
-    \inferrule[]
-    { e \notin dom(\rho) \\
-      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
-      \dot C_l = \elet{x_l}{b_l}{\hole}\\
-      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
-    }
-    {\echeckbounds{\rho}{e}{\tarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}
-
-    \inferrule[]
-    {}
-    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}
-
-    \inferrule[]
-    { x_l,x_h = \rho(e) \\
-      x_{cl},x_{ch} = \fresh\\
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
-    }
-    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}
-
-    \inferrule[]
-    { e \notin dom(\rho) \\
-      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
-      \dot C_l = \elet{x_l}{b_l}{\hole}\\
-      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
-      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
-      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
-    }
-    {\echeckboundsw{\rho}{e}{\tallarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}
-
-\inferrule[]
-{e \notin dom(\rho) \\
-  x_l,x_l',x_h,x_h' = \fresh \\
-  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
-  \dot C_2 = \elet{x_l'}{b_l'}{\elet{x_h'}{b_h'}{\hole}}\\
-  \dot C_3 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
-  \dot C_4 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
-}
-{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3;4}}}
-
-\inferrule[]
-{x_l',x_h' =\rho(e) \\
-  x_l,x_h = \fresh \\
-  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
-  \dot C_2 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
-  \dot C_3 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
-}
-{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3}}}
-  \end{mathpar}
-}
-\caption{Metafunctions for Dynamic Checks}
-\label{fig:compilationhelpers}
-\end{figure*}
-
-% the closure $\dot
-% C_{1;2;n;3;b;4}$ encodes the idea that $e_1$ and $e_2$ are evaluated to values
-% before a null check (included in $\dot C_n$) is performed on
-% $\dot a_1$ , the result of evaluating $e_1$. $\dot C_3$ adds $\dot
-% a_1$ and $\dot a_2$ before a bounds check (included in $\dot C_b$) is performed on $\dot a_1$
-% with $\dot a_2$ as the index. Finally, we use $\dot C_4$ 
-% to store the result of the dereference in $x_4$ and return the pair
-% $\dot C_4,x_4$ as the result of the compilation. 
-% Further checks and
-% widening can be performed using $x_4$ and we don't have to worry about
-% the effects being triggered multiple times because the evaluation only
-% happens once in $\dot C_4$.
-
-% insert runtime checks and bounds widening
-% code in a straightforward way. Take the \textbf{C-Deref} rule as an example,
-% if we were to output $\dot e$ directly, we would have to rebind the
-% expression to a variable before we can perform the null check and
-% bounds check since $\dot e$ is not always a pure expression. Adding
-% $\elettext$ bindings in such an ad-hoc fashion for avoiding side
-% effects can clutter the compilation rules. Our approach of
-% outputting a context and an atom threads the ANF transformation into
-% each individual rule. Fig.~\ref{fig:compilationhelpers} shows helper
-% functions that take atom expressions and return a context that
-% contains the checks. The advantage of using explicit contexts becomes
-% evident when we start to compose them. For the simulation property to
-% hold, we want to ensure that the effects happen in the exact same
-% order.  In
-% the C-Deref rule, we first evaluate the expression that corresponds to
-% $e$ through $\dot C_1$, and then perform a bounds check followed by a
-% null check through $\dot C_b$ and $\dot C_n$. Finally, we dereference
-% the expression through $\dot C_2$, potentially widen the upper bound
-% by checking the dereferenced
-% result.
-
-% One big difference is that
-% our compilation strategy is not parameterized by the heap. Given a
-% literal annotated with a type, the compilation judgment simply
-% discards the type without validating whether the literal's type is
-% valid. This kind of relaxation not only makes the compilation rules
-% simpler (we no longer need to keep track of $\defscope$ and the mode),
-% but also allows us to formulate a stronger simulation theorem that
-% applies even after the program state is corrupted. More concretely,
-% the compilation judgment satisfies preservation regardless of whether
-% an unchecked step is being taken. On the other hand, if we were to
-% embed the compilation into the full-fledged typechecking rules, we
-% would not always be able to compile the expression after a step has
-% been taken.
-
-% When our source
-% program is stuck because of the execution of unchecked code, we can
-% conclude that the compiled program will also get stuck with the
-% same erased heap state. It is possible to embed our compilation
-% strategy directly into the typing judgment, but our simulation would
-% be limited to states that are not yet corrupted to be untypable.
-
-% Since the erased language does not have type annotations, we need
-% to explicitly represent bounds of null-terminated arrays as stack
-% variables to support bounds widening at runtime. Every time a checked
-% null-terminated array pointer is bound to a variable, we generate two
-% fresh new variables that are assigned the lower and upper bounds of
-% the array. The upper bound can be mutated into
-% a larger bound during execution. We use $\rho$ to keep track of the
-% bounds variables that belong to null-terminated array variables.
-
-% The compilation judgment does not produce expressions
-% directly. Instead, it outputs a tuple of a context $\dot C$ and an atom
-% expression $\dot a$ such that $\dot C[\dot a]$ corresponds to the
-% compiled target expression. In the case where the atom is a variable, $\dot
-% C$ will contain a let binding that defines the variable. This unconventional
-% output format enables us to insert runtime checks and bounds widening
-% code in a straightforward way. Take the C-Deref rule as an example,
-% if we were to output $\dot e$ directly, we would have to bind the
-% expression to a variable before we can perform the null check and
-% bounds check since $\dot e$ is not always a pure expression. Adding
-% $\kw{let}$ bindings in such an ad-hoc fashion for avoiding side
-% effects can clutter the compilation rules. Our approach of
-% outputting a context and an atom threads the ANF transformation into
-% each individual rule. Fig.~\ref{fig:compilationhelpers} shows helper
-% functions that take atom expressions and return a context that
-% contains the checks. The advantage of using explicit contexts becomes
-% evident when we start to compose them. For the simulation property to
-% hold, we want to ensure that the effects happen in the exact same
-% order. Since effects and reduction (except for variables?) happen only
-% within contexts, we can precisely control the order in which effects
-% and evaluation happen by composing the contexts in a specific order. In
-% the C-Deref rule, we first evaluate the expression that corresponds to
-% $e$ through $\dot C_1$, and then perform a bounds check followed by a
-% null check through $\dot C_b$ and $\dot C_n$. Finally, we dereference
-% the expression through $\dot C_2$, potentially widen the upper bound
-% by checking the dereferenced
-% result.
-
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-
-    \inferrule[C-Const]
-      {}
-      {\Gamma;\rho\vdash \evalue{n}{\tau} \gg \hole, n : \tau}
-
-    \inferrule[C-Var]
-      {x : \tau \in \Gamma}
-      {\Gamma;\rho \vdash x \gg \hole, x : \tau}
-
-    \inferrule[C-Cast]
-              {\Gamma;\rho \vdash e \gg  \dot C , \dot a : \tau'}
-              {\Gamma;\rho \vdash \ecast{\tau}{e} \gg \dot C, \dot a  : \tau}
-
-      \inferrule[C-DynCast]
-      {\Gamma;\rho \vdash e \gg \dot C_1, \dot a : \tptr{\tallarrayb{\bvar'}{\tau}}{m} \\
-      \echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\tptr{\tallarrayb{\bvar'}{\tau}}{m}}{\dot C_{b}}}
-      {\Gamma;\rho \vdash \edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{e}
-            \gg \dot C_{1;b}, \dot a : \tptr{\tallarrayb{\bvar}{\tau}}{m}}
-
-      \inferrule[C-Str]
-      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 :  \tptr{\tntarrayb{\bvar}{\tau_a}}{m} \\
-        \echecknull{\dot a_1}{m}{\dot C_{n}} \\
-        \echeckbounds{\rho}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{0}{\dot C_{b}} \\
-        x_2 = \fresh \\
-          \dot C_2 = \elet{x_2}{\estrlen{\dot a_1}}{\hole} \\
-      \ewidenstrlen{\rho}{e}{\dot a_1}{\tptr{\tntarrayb{\bvar}{\tau_a}}{m}}{\dot C_{w}} } 
-                {\Gamma;\rho \vdash \estrlen{e} \gg \dot C_{1;n;b;2;w}, x_2 : \tint}
-
-      \inferrule[C-LetStr]
-      {\Gamma(y) = \tntarrayptr{b_l}{b_h}{\tau_a}{\cmode} \\ x \not\in \fv(\tau) \\
-        \Gamma; \rho \vdash \estrlen{y} \gg \dot C_1, \dot a_1 : \tint\\
-        \dot C_2 = \elet{x}{\dot a_1}{\hole} \\
-        \Gamma[x\mapsto \tint,y\mapsto [\tntarrayptr{b_l}{x}{\tau_a}{\cmode}]];\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau}
-      %   x_3 = \fresh \\
-      %   \dot C_3 = \elet{x_3}{\dot a_2}{\hole}}
-                {\Gamma;\rho \vdash \elet{x}{\estrlen{y}}{e} \gg \dot C_{1;2;3},\dot a_3 : \tau}
-
-      \inferrule[C-If ]
-      {\Gamma; \rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
-        \Gamma; \rho \vdash e_1 \gg \dot C_2, \dot a_2 : \tau_2 \\
-        \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
-        x_4 = \fresh \\
-        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_2[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}
-      }
-        {\Gamma; \rho \vdash \eif{e_1}{e_2}{e_3} \gg \dot C_{1;4}, x_4 : \tau_2 \sqcup \tau_3}
-
-        \inferrule[C-IfNT]
-        {\Gamma; \rho \vdash x : \tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c} \\
-          b_h = 0 \Rightarrow \Gamma' = \Gamma[x \mapsto \tptr{\tntarrayb{(b_l,1)}{\tau}}{c}] \\
-          b_h \neq 0 \Rightarrow \Gamma' = \Gamma \\
-          \Gamma; \rho \vdash \estar{x} \gg \dot C_1, \dot a_1 : \tau_1\\
-          \Gamma'; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau_2\\
-          \Gamma;\rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau_3 \\
-        \ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C_{w}} \\
-        x_4 = \fresh \\
-        \dot C_4 = \elet{x_4}{\eif{\dot a_1}{\dot C_{2;w}[\dot a_2]}{\dot C_3[\dot a_3]}}{\hole}}
-        {\Gamma;\rho \vdash \eif{\estar{x}}{e_1}{e_2} \gg \dot C_{1;4}, x_4 : \tau_1 \sqcup \tau_2}
-
-   \inferrule[C-Let]
-    { (x\in \fv(\tau') \Rightarrow e_1 \in Bound) \\
-        \Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tau_1 \\ 
-        \cextend{\rho}{x}{\tau_1}{\dot C_2,\rho'}\\
-        \dot C_3 = \elet{x}{\dot a_1}{\hole}\\
-        \Gamma[x\mapsto \tau];\rho' \vdash e_4 \gg \dot C_4, \dot a_4  : \tau_4}
-    {\Gamma;\rho' \vdash \elet{x}{e_1}{e_4} \gg \dot C_{1;2;3;4},\dot a_4 : \tau_4[\tau_1 = \tint \Rightarrow x \mapsto e_1]}
-
-  \inferrule[C-Ret]
-    {\Gamma(x)\neq \bot \\
-      \Gamma;\rho \vdash e \gg \dot C_1, \dot a_1 : \tau \\
-      x_2 = \fresh \\
-      \mu \gg \dot \mu \\
-    \dot C_2 = \elet{x_2}{\eret{x}{\dot \mu}{\dot C_1[\dot a_1]}}{\hole}}
-  {\Gamma;\rho \vdash \eret{x}{\mu}{e} \gg \dot C_2, x_2 : \tau}
-
-    \inferrule[C-Fun]
-    {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\
-      % \dot \Xi(f) = (\overline{x})\; \dot C_f; \dot a_f \\
-      (\forall e_i\in\overline{e}\;\;\tau_i\in\overline{\tau}\;.\; \Gamma; \rho \vdash {e_i} \gg \dot C_i, \dot a_i  : \\ \tau_i' \wedge
-      \tau_i' \sqsubseteq 
-      \tau_i[\overline{e} / \overline{x}]%  \\
-      ) \\
-      x_f = \fresh \\
-      \dot C_f = \elet{x_f}{f(\overline{a})}{\hole}}
-    {\Gamma; \rho \vdash f(\overline{e}) \gg \overline{\dot C}[\dot C_f] , x_f  : \tau[\overline{e} / \overline{x}]}
-
-  \inferrule[C-Def]
-  {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m} \\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    x_2 = \fresh \\
-    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}
-   }
-   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;2}, x_2 : \tau}
-
-   \inferrule[C-DefArr]
-   {\Gamma;\rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tallarrayptr{b_l}{b_h}{\tau}{m} \\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    \echeckbounds{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
-    x_2 = \fresh \\
-    \dot C_2 = \elet{x_2}{\estar{\dot a_1}}{\hole}}
-   {\Gamma;\rho \vdash \estar{e_1} \gg \dot C_{1;n;b;2},x_2 : \tau}
-
-    \inferrule[C-Mac]
-    {\dot C_{1}, \dot a_1 = \esizeof{\omega} \\
-      x_2 = \fresh \\
-    \dot C_2 = \elet{x_2}{\emalloc{\dot a_1}}{\hole}}
-     {\Gamma; \rho \vdash \emalloc{\omega} \gg \dot C_{1;2}, x_2 : \tptr{\omega}{\cmode}}
-
-  \end{mathpar}
-}
-\caption{Compilation}
-\label{fig:compilation}
-\end{figure*}
-
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[C-Add]
-              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tint \\
-                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint\\
-                x_3 = \fresh \\
-              \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}}
-              {\Gamma; \rho \vdash \dot C_3, x_3 : \tint }
-
-   \inferrule[C-Ind] 
-              {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
-                \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
-                \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-                \echeckbounds{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
-                x_3, x_4 = \fresh\\
-                \dot C_3 = \elet{x_3}{\dot a_1 \plus \dot a_2}{\hole}\\
-                \dot C_4 = \elet{x_4}{\estar{x_3}}{\hole}
-                }
-                {\Gamma; \rho \vdash \estar{(\ebinop{e_1}{e_2})} \gg \dot C_{1;2;n;3;b;4} , x_4: \tau}
-
-  \inferrule[C-Assign]
-  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tau}{m'} \\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
-    \tau'\sqsubseteq \tau \\
-    x_3 = \fresh \\
-    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
-  }
-  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;3}, x_3 : \tau}
-
-  \inferrule[C-AssignArr]
-  {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
-    \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-    \echeckboundsw{\rho}{e_1}{\tallarrayptr{b_l}{b_h}{\tau}{m}}{0}{\dot C_{b}}\\
-    \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tau' \\
-    x_3 = \fresh \\
-    \dot C_3 = \elet{x_3}{\eassign{\dot a_1}{\dot a_2}}{\hole} \\
-    \tau'\sqsubseteq \tau}
-  {\Gamma; \rho \vdash \eassign{e_1}{e_2} \gg \dot C_{1;2;n;b;3},x_3 : \tau}
-
-    \inferrule[C-IndAssign]
-    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m} \\
-      \Gamma; \rho \vdash e_2 \gg \dot C_2, \dot a_2 : \tint \\
-      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-      \echeckboundsw{\rho}{e_1}{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\dot a_2}{\dot C_{b}} \\
-      \Gamma; \rho \vdash e_3 \gg \dot C_3, \dot a_3 : \tau' \\
-      x_4, x_5 = \fresh \\
-      \dot C_4 = \elet{x_4}{\dot a_1 \plus \dot a_2}{\hole}\\
-      \dot C_5 = \elet{x_5}{\eassign{x_4}{x_3}{\hole}}
-      \tau'\sqsubseteq \tau}
-    {\Gamma; \rho \vdash \eassign{(e_1 \plus e_2)}{e_3} \gg \dot C_{1;2;n;3;4;b;5} : \tau}
-
-    \inferrule [C-Struct]
-    {\Gamma; \rho \vdash e_1 \gg \dot C_1, \dot a_1  :  {\tptr{\tstruct{T}}{m}} \\
-      D(T) = \tau_0~f_0\ldots; \tau_j~f; ...  \\
-      \echecknull{\dot a_1}{m}{\dot C_{n}}\\
-      x_2 = \fresh\\
-    \dot C_2 = \elet{x_2}{\dot a_1 \plus \dot j}{\hole}}
-    {\Gamma; \rho \vdash \eamper{e_1}{f} \gg \dot C_2, x_2  : \tptr{\tau_f}{m}}
-
-    \inferrule[C-Unchecked]
-              {\Gamma;\rho \vdash e \gg \dot C, \dot a : \tau}
-              {\Gamma;\rho \vdash \eunchecked{e} \gg \dot C, \dot a : \tau}
-
-  \end{mathpar}
-    % \begin{prooftree}
-    %   \hypo{
-    %     \Gamma; \rho \vdash_u e \gg \dot C, a : \tau
+% \ignore
+% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
+% % The diagram also indicates the simulation relation ($\sim$) among our formalization of the \checkedc Type System ($\vdash$), Semantics ($\longrightarrow$), Compilation ($\gg$), as well as the compiled C semantics ($\xrightarrow{c}$).
+% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
+% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist.
+The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.
+
+%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C.
+The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
+% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).
+
+% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
+% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
+% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation.
+We generate tens of thousands cases and cover all possible corner cases of the compilation.
 % }
-    %   \infer1[C-Unchecked]{\Gamma; \rho \vdash \eunchecked{e} \gg \dot C, a : \tau}
-    % \end{prooftree}\\\\
-
-    % \begin{prooftree}
-    %   \hypo{
-    %     \begin{matrix}
-    %         \Gamma; \rho \vdash e \gg \dot C_1, a_1 :  {\tptr[m]{\tstruct{T}}} \\
-    %         % \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
-    %         D(T) = \tau_0\ldots; \tau_j~f; ...\\
-    %         x_2 = \fresh \\
-    %         \dot C_2 = \elet{x_2}{a_1 + j}{\hole} \\
-    %         \echecknull{a_1}{m}{\dot C_{null}}
-    %     \end{matrix}
+% \ignore{
+% \begin{figure*}[t]
+%   \begin{subfigure}[b]{1\textwidth}
+%     \begin{lstlisting}
+% int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
+%   return *p;
 % }
-    %   \infer1[C-Amper]{\Gamma; \rho \vdash \eamper{e}{f} \gg \dot C_{1;null;2}, x_2  : \tptr[m]{\tau_f}}
-    % \end{prooftree} \\ \\
-
-    % \begin{prooftree}
-    %   \hypo{
-    %     \begin{matrix}
-    %       \Gamma; \rho \vdash e_1 \gg \dot C_1, a_1 : \tptr[m']{\omega} \\
-    %       \Gamma; \rho \vdash e_2 \gg \dot C_2, a_2 : \tau \\
-    %       m' = u \Rightarrow m = u \\
-    %       \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}\\
-    %       \echecknull{a_1}{m}{\dot C_{n}} \\
-    %       % \echeckbounds{\rho}{a_1}{\tptr[m]{(\tntarray{le}{he}{\tau})}}{\dot C_{b}} \\
-    %       x_3 = \fresh\\
-    %       \dot C_3 = \elet{x_3}{\eassign{a_1}{a_2}}{\hole} \\
-    %     \end{matrix}
+
+% ...
+% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
+% deref_array(5, p0);
+%     \end{lstlisting}
+%     \label{fig:chkcexamplederef}
+%     \caption{\texttt{deref\_array} in \checkedc}
+%   \end{subfigure}
+
+%   \begin{subfigure}[b]{1\textwidth}
+% \begin{lstlisting}
+% int deref_array(size_t n, int *p) {
+%   /* statically determine the definitions p_lo and p_hi */
+%   int *p_lo = p;
+%   int *p_hi = p + n;
+%   /* runtime checks */
+%   assert(p_lo <= p && p <= p_hi);
+%   assert(p != NULL);
+%   /* widening */
+%   int p_derefed = *p;
+%   if (p_derefed != '\0') {
+%     if (p_hi == p) {
+%       ++p_hi;
 %     }
-    %   \infer1[C-Assign]{\Gamma; \rho \vdash \dot C_{1;b;n;2;3}, x_3 : \tau}
-    % \end{prooftree}\\\\
-
-    % \begin{prooftree}
-    %   \hypo{
-    %     \begin{matrix}
-    %       \Gamma; \rho \vdash e_1 : \tptr[m']{\omega} \\
-    %       \Gamma; \rho \vdash e_2 : \tint \\
-    %       \Gamma; \rho \vdash e_3 : \tau \\
-    %       m' = u \Rightarrow m = u \\
-    %       \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-    %     \end{matrix}
 %   }
-    %   \infer1[C-IndAssign]{\Gamma; \rho \vdash \eassign{(e_1 + e_2)}{e_3} : \tau}
-    % \end{prooftree}
-}
-\caption{Compilation (Continued)}
-\label{fig:compilation2}
-\end{figure*}
+%   return p_derefed;
+% }
+% ...
+% /* int *p0 */
+% /* rho(p0) = p_lo, p_hi */
+% deref_array(5, p0);
+% \end{lstlisting}
+% \caption{\texttt{deref\_array} in C}
+%     \label{fig:cexamplederef}
+%   \end{subfigure}
+
+%   \begin{subfigure}[b]{\textwidth}
+%     \begin{lstlisting}
+% /* nt_array_ptr<int> p : bounds(e0, e1) */
+% /* rho(p) = p_lo, p_hi */
+% size_t x = strlen(p);
+% if(x >= 1) {
+%   *(p+1);
+% }
+%     \end{lstlisting}
+%     \caption{\texttt{strlen} in \checkedc}
+%     \label{fig:chkcexamplestrlen}
+%   \end{subfigure}
+
+%   \begin{subfigure}[b]{\textwidth}
+%     \begin{lstlisting}
+% /* nt_array_ptr<int> p : bounds(e0, e1) */
+% /* rho(p) = p_lo, p_hi */
+% /* runtime checks omitted */
+% ...
+% size_t x = strlen(p);
+% int *p_hi_new = p + x;
+% p_hi = max(p_hi, p_hi_new);
+% if (x >= 1) {
+%   /* null check for pointer arithmetic */
+%   assert(p != NULL);
+%   int *p_1 = p + 1;
+%   /* null check for dereference */
+%   assert(p_1 != NULL);
+%   /* bounds check for dereference */
+%   /* note how we are using p_hi rather than p + x */
+%   assert(p_lo <= p_1 && p_1 <= p_hi);
+%   ...
+% }
+%     \end{lstlisting}
+%     \caption{\texttt{strlen} in C}
+%     \label{fig:cexamplestrlen}
+%   \end{subfigure}
+% \caption{Compilation of null-terminated string dereference}
+% \label{fig:compilationexample}
+% \end{figure*}
+% }
+
+% \ignore{
+% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
+% dereference of null-terminated strings by inserting explicit checks and
+% bounds widening code. For readability, we present the example in real
+% \checkedc and C syntax. At
+% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
+% bounds are defined in terms of the arguments \code{p} and \code{n}
+% according to the bounds annotations obtained during typechecking. This
+% avoids the need to pass in bounds as extra arguments, thus maintaining
+% compatibility with C code. When we call \code{deref_array} on
+% \code{p0}, a string with size 5 (excluding the null character), there
+% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
+% the stack. It is possible that \code{p_hi} has been widened
+% before we reach line 21. The programmer will have to perform a
+% \code{dynamic_bounds_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
+% attempts to widen the bound when the dereferenced result is not
+% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
+% \code{deref_array}, in contrast to the T-IfNT rule, which only
+% remembers the widened bounds within the scope of the then branch of
+% the if statement.
+
+% \todo[inline]{Widening can happen at every
+%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
+% shows how an invocation of \code{strlen} on null-termianted strings
+% is compiled into C code. We perform the same runtime checks that
+% happen during dereferencing. The widening code at line
+% 8 updates \code{p}'s upper bound only if the result of
+% \code{strlen} is larger than the value of the upper bound stored on
+% the stack. This is another scenario where the runtime can be more precise
+% than the statically determined bounds information.
+
+% }

### 2023-01-03 sandp23/background.tex

@@
 
+\liyi{can we reuse the old text for the spatial safety? }
+
 \noindent
@@
 
+\liyi{If we are not going to talk about converting C to \checkedc, can we just move this to future work? }
+
 \myparagraph{Converting C to \checkedc}
@@
 
+\liyi{Does the program too small? Will reviewer complain about the program size? }
 \subsection{No Safety Against Unchecked Code}
@@
 
+\liyi{I do not think you need three sub-sections for these topics. They can be in the same place. }
 \subsection{Program Partitioning}

### 2023-01-03 sandp23/introduction.tex

@@
 Specifically,~\checkedc extends C with checked pointer types, which are restricted by the compiler to spatially safe uses.
+\liyi{why we need to mention C to \checkedc conversion?
+Is this paper about introducing CheckedCBox or about the conversion from C to CheckedCBox.
+I am actually against putting CheckedCBox conversion stuff in the paper, because it is kind of confusion.
+We have the compilation from CheckedCBox to LLVM and have the conversion from C to CheckedCBox.
+}
 The checked pointers are represented as system-level memory words,~\ie without ``fattening'' metadata, ensuring backward compatibility.
@@
 We need to ensure that~\emph{code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+\liyi{Can we strengthen the sentence a little by saying something fancy? }
 
@@
 Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
+\liyi{why? Can we not say this? It seems that the whole reason people should care our work is we did it in \checkedc, but \checkedc itself is the selling product here.... }
 But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
@@
 The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.
+\liyi{really? I think they can exchange data in the unchecked region, and then the sandbox guarantee the correctness? }
 
@@
 Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
+\liyi{can we just name them unchecked and checked regions, and then referring to them as \ucregion and~\cregion? We need to make it clear that we are re-define c and uc regions. }
 \begin{itemize}
@@
 \aravind{Fix the following text after all the sections of the papers are finalized.}
+\liyi{the following might be removed if space is needed, since it is usually not quite useful.}
 We begin with a review of \checkedc{} \mzr{and introduction of} new features in

### 2023-01-03 sandp23/evaluation.tex
Minor changes

@@
 % \liyi{Deena? Leo? }
-Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program-benchmarks.
+Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks.
 MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
@@
 \begin{center}
+\label{fig:micrbenchmarks}
 \begin{tabular}{||c c||} 
@@
 \subsection{Program Run-time Benchmarks}
+\myparagraph{Overview}
 We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program.
-In our conversion efforts, we intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern-day memory vulnerabilities by re-introducing fixed memory bugs and then annotating the relevant buggy code with tainted pointers.
+We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then annotating the relevant buggy code with tainted pointers.
 However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible.
@@
 Consequently, each of the conversions for the six programs follows a varied approach.
+
+\myparagraph{Parsons}
 Parsons is annotated comprehensively in two variants parsons\_wasm and parsons\_tainted.
@@
 
-We intended to keep \systemname changes for Lib-png comparatively narrow in scope by beginning with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read().
+\myparagraph{LibPNG}
+\systemname changes for libPNG is narrow in scope and begins with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read().
 However, we also annotate sections of Lib-png that involve reading, writing, and image processing (interlace, intrapixel, etc) on user-input image data as tainted.
@@
 
+\myparagraph{ProFTPD}
 \systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
-We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr\<char\>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
+We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr$l$char$g$) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
 Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance.
@@
 
-MicroHTTPD is a good demonstration of the practical difficulties in converting a program to \systemname.
+\myparagraph{MicroHTTPD}
+MicroHTTPD demonstrates the practical difficulties in converting a program to \systemname.
 Our conversion for this program was aimed at sandboxing memory vulnerabilities CVE-2021-3466 and CVE-2013-7039.
@@
 
+\myparagraph{UFTPD}
 \systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
@@
 
+\myparagraph{Tiny-bignum}
 Motivated by its size and simplicity, we intended to make comprehensive \systemname changes for Tiny-bignum by marking all the pointers as tainted.
@@
 \begin{center}
+\label{fig:prgrbenchmarks}
 \begin{tabular}{||c | p{1.3cm} | p{1.8cm}||} 
@@
  \hline
+
 \end{tabular}

### 2023-01-04 sandp23/evaluation.tex

@@
 % \liyi{Deena? Leo? }
-Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks.
+Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks.
 MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
 We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.
+
+\myparagraph{Benchmarking Methodology}
 We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
@@
 \subsubsection{Pointer Access}
-Pointer overhead during run-time between the pointers in the checked and Sandboxed regions is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
+Run-time Pointer overhead is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
 34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.
@@
 We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program.
-We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then annotating the relevant buggy code with tainted pointers.
+We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code .
 However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible.
 Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
-The extent of conversion qualitatively dictates the converted program's conversion efforts required and the run-time performance.
+Conversion style and extent qualitatively dictate the converted program's conversion efforts required and the run-time performance.
 Consequently, each of the conversions for the six programs follows a varied approach.
@@
 \systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
-We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr$l$char$g$) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
+We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
 Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance.
@@
 \myparagraph{Tiny-bignum}
-Motivated by its size and simplicity, we intended to make comprehensive \systemname changes for Tiny-bignum by marking all the pointers as tainted.
-Furthermore, in likes of an unsafe use of sprintf() that could possibly lead to a buffer overflow within the bignum\_to\_string(), we move this function into the WASM sandbox.
-Ideally, a major portion of conversion efforts to \systemname is attributed to understanding the codebase and finding the precise extent to which we choose to propagate the taintedness or to stop and give up and marshall the data, thereby, requiring only 4 hours to convert tiny-bignum to \systemname.
+Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive.
+Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf().
+Given that significant \systemname conversion efforts is attributed to understanding the source-code and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling the data between regions, Tiny-bignum only required 4 hours.
 The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

### 2023-01-04 sandp23/overview.tex

@@
 The \systemname type checker compiles a \systemname program by partitioning the different program regions, i.e., $c$ and $uc$ regions, as well as different pointers, i.e., checked, unchecked and tainted pointers.
-Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler.
+Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler within the checked region.
 
 \Cref{sec:formal} provides the formalism of the different compiler components.
+Converting a C/checked-c program to \systemname is a two-step process that involves identifying unsafe/unchecked regions described in \Cref{subsec:identifyregionstosbx} followed by making the \systemname changes described in \Cref{subsec:moveregionstosbx}.
+\subsection{Identify Unchecked Regions}
+\label{subsec:identifyregionstosbx}
 Here,
-we use the example of a simple network server in~\lst{lst:comb} to explain the \systemname compiler.
+we use the example of a simple network server in~\lst{lst:comb} to explain changes required by the \systemname compiler.
 The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
@@
 \liyi{The example looks good. But we are focusing on converting C to \checkedc here? }
-
+\subsection{\systemname Annotations}
+\label{subsec:moveregionstosbx}
 \myparagraph{Goal} 

### 2023-01-05 sandp23/overview.tex
Minor edits

--- 2022-12-14 sandp23/overview.tex
@@
 
-We use the example of a simple network server in~\lst{lst:comb} to explain the overview of developers using~\systemname{}.
+The~\Cref{fig:overview} shows the interaction between various components of~\systemname{}.
+Given an appropriately annotated source program and a sandbox configuration, our system creates an executable such that all tainted functions and data will reside in a sandbox and the~\cregion (non-sandboxed) region is not affected by the code in the sandbox.
+
+We will present an overview of the~\systemname{} from a developer perspective and the details of individual components in the later sections.
+
+%\Cref{fig:overview} provides the compiler structure of \systemname.
+The \systemname type checker compiles a \systemname program by partitioning the different program regions, i.e., $c$ and $uc$ regions, as well as different pointers, i.e., checked, unchecked and tainted pointers.
+Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler within the checked region.
+
+%\Cref{sec:formal} provides the formalism of the different compiler components.
+%Converting a C/checked-c program to \systemname is a two-step process that involves identifying unsafe/unchecked regions described in \Cref{subsec:identifyregionstosbx} followed by making the \systemname changes described in \Cref{subsec:moveregionstosbx}.
+\subsection{Running Example}
+\label{subsec:identifyregionstosbx}
+Here,
+we use the example of a simple network server in~\lst{lst:comb} to explain changes required by the \systemname compiler.
 The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
@@
 
-\myparagraph{Using tainted types}
-To handle this, the developer marks risky functions with unchecked pointers~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded.
+\liyi{Can we just assume that programmers know \checkedc and will use \checkedc, and then we added tainted type? }
+
+\subsection{\systemname Annotations}
+\label{subsec:moveregionstosbx}
+%\myparagraph{Using tainted types}
+To handle this, the developer marks risky functions with unchecked pointers,~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded.
 The developer starts using~\systemname{} with the resulting partially annotated program with checked and tainted types~\ie~\lst{lst:initialconv} and~\ref{lst:humantaint} atop of~\lst{lst:comb}.
@@
 
-\myparagraph{Partitioning}
+\subsection{Partitioning}
 Our partitioner splits the provided program into two sets of source files with the necessary source changes required to communicate with sandboxed code.
@@
 Finally, these two object files will be linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions will be executed in a sandbox (\ucregion) and the rest of the functions as regular code (\cregion).
+
+\subsection{Alternative Use Cases}
+\input{figures/newusecase}
+
+Tainting all functions that are not checked may lead to a large number of transitions to the sandbox, which could add unnecessary overhead.
+In our example, in~\lst{lst:final}, marking~\code{read_msg} function as tainted would cause the system to transition to and from the sandbox every time a message is read, which could be inefficient.
+
+However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain checked pointers as well.
+%
+This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them.
+In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
+
+For our example in~\lst{lst:final}, we can make only the function~\code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
+
+The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
+It is best to annotate all functions with checked types for security purposes.
+However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted.
+This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
+
+%newusecase.tex

### 2023-01-08 sandp23/appendix.tex
update.

@@
 % }
+
+\subsection{Additional Type and Semantic Rules}\label{appx:add-type-sem}
+
+\begin{DIFnomarkup}
+\begin{figure}
+{\small
+\[\hspace*{-1.2em}
+\begin{array}{l}
+\textcolor{blue}{\text{Bound Inequality and Equality:}}\\[0.3em]
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
+     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
+     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
+     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
+     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
+     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
+    \end{array}
+  \\[1.5em]
+\textcolor{blue}{\text{Type Equility:}}\\
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     && \tint & =_{\Theta} & \tint\\
+     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
+     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
+             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
+
+    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
+
+ &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
+                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
+    \end{array}
+  \\[0.7em]
+\textcolor{blue}{\text{Subtype:}}\\[0.3em]
+
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
+
+    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
+\\[0.2em]
+\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
+
+    \end{array}
+\end{array}
+  \]
+}
+{\footnotesize
+\[
+\begin{array}{l}
+n'+n = add(n',n)
+\qquad
+(x+n')+n = x+add(n',n)\\
+\textit{cond}(\overline{x},\tau,\overline{y},\tau')
+=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
+  \wedge \overline{y}\cupdot\overline{z}
+  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
+\\\qquad\qquad\qquad\qquad\qquad
+  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
+\end{array}
+\]
+}
+  \caption{Type Equality and Subtyping}
+  \label{fig:checkc-subtype}
+\end{figure}
+\end{DIFnomarkup}
+
+Since bounds expressions may
+contain variables, determining assumptions like $b_l \leq b_l'$
+requires reasoning about the probable values of these variables'. The type
+system uses $\Theta$ to make such reasoning more precise.
+$\Theta$ is a map from variables $x$ to
+equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
+It maps variables to equations that are recorded along the type checking procedure.
+If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
+$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
+such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
+\iftr
+  Appendix~\ref{app:le}.
+\else
+  the supplemental report~\cite{checkedc-tech-report}.
+\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
+
+\begin{figure}[t]
+{\small
+{\captionsetup[lstlisting]{margin = 8 mm}
+  \begin{lstlisting}[xleftmargin=8 mm]
+nt_array_ptr<char> safe_strcat
+   (nt_array_ptr<char> dst : count(n),
+    nt_array_ptr<char> src : count(0), int n) {
+  int x = strlen(dst);
+  int y = strlen(src);
+  nt_array_ptr<char> c : count(n) =
+    dyn_bounds_cast
+           <nt_array_ptr<char>>(dst,count(n));
+    // sets c == dst with bound n (not x)
+  if (x+y < n) {
+    for (int i = 0; i < y; ++i)
+      *(c+x+i) = *(src+i);
+    *(c+x+y) = '\0';
+    return dst;
+  }
+  return null;
+}
+  \end{lstlisting}
+}
+}
+\caption{Implementation of safe \code{strcat}}
+\label{fig:strcat-ex}
+\end{figure}
+
+Recall that array
+bounds in types may refer to in-scope variables; e.g., parameter
+\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
+2-3 in \Cref{fig:strcat-ex}. 
+All dependent bound variables appearing in function headers 
+are bounded by a variable list in the function pointer type, which is checked in the type rule.
+Semantically,
+the call is expanded into a \texttt{let} which binds
+parameter variables $\overline{x}$ to the actual arguments
+$\overline{n}$, but annotated with the parameter types
+$\overline{\tau}$ (this will be safe for type-correct programs). 
+The function body $e$ is wrapped in a static cast
+$(\tau[\overline{n} / \overline{x}])$ which is the function's return
+type but with any integer parameter variables $\overline{x}$ appearing in that
+type, as type bound variables,
+substituted with the call's actual arguments $\overline{n}$.
+To see why this is needed, suppose that \code{safe_strcat} in
+Fig.~\ref{fig:strcat-ex} is defined to return a
+\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
+perform a \code{safe_strcat} function call as
+\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
+function returns a value with type \code{nt_array_ptr<int>:count(10)}
+because we substitute bound variable \code{n} in the 
+defined return type with \code{10} from the function call's
+argument list.
+
+\begin{figure}[t!]
+  \begin{small}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+int deref_array(n : int,
+     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
+     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
+  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
+  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
+    * p;
+    * q = 1;
+}
+...
+/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
+/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\begin{frame}
+
+\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
+
+\end{frame}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+deref_array(int n, int* p, int * q) {
+  //m is the current context mode
+  let p_lo = 0; let p_hi = n; 
+  let q_lo = 0; let q_hi = n; 
+  /* runtime checks */
+  assert(p_lo <= 0 && 0 <= p_hi);
+  assert(p != 0);
+  *(mode(p) $\wedge$ m,p);
+  verify(q, not_null(m, q_lo, q_hi) 
+             && q_lo <= 0 && 0 <= q_hi);
+  *(mode(q) $\wedge$ m,q)=1;
+}
+...
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\end{small}
+    \caption{Compilation Example for Dependent Functions}
+\label{fig:compilationexample1}
+\end{figure}
+
+For compiling dependent function calls,
+\Cref{fig:compilationexample1} provides a hint.
+Notice that the bounds for the array pointer \code{p} are not passed as
+arguments. Instead, they are initialized according to \code{p}'s
+type---see line~4 of the original \lang program at the top of the figure.
+Line~$3$ of the generated code
+sets the lower bound  to \code{0} and the
+upper bound to \code{n}.
+
+\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
+A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
+\end{defi}
+
+\begin{defi}[Heap Well-formedness]
+For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
+(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
+contains no free variables. 
+\end{defi}
+
+\begin{defi}[Stack Well-formedness]
+A stack snapshot $\varphi$ is well-formed if
+for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
+contains no free variables. 
+\end{defi}
+
+We also need to introduce a notion of
+\emph{consistency}, relating heap environments before and after a
+reduction step, and type environments, predicate sets, and stack
+snapshots together.
+
+\begin{defi}[Stack Consistency]
+A type environment $\Gamma$, variable predicate set $\Theta$, and
+stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
+\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
+$\Gamma(x) = \tau$ for some $\tau$ and 
+$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
+\end{defi}
+
+\begin{defi}[Checked Stack-Heap Consistency]
+A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
+for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
+\end{defi}
+
+\begin{defi}[Checked Heap-Heap Consistency]
+A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
+for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
+\end{defi}

### 2023-01-08 sandp23/formal.tex
update.

@@
 \\[0.5em]
-
 \begin{array}{llcllcl}
-
-\text{Context Mode:} & m & ::= & \cmode \mid \umode \\[0.5em]
-
-\text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\[0.5em]
-
+\text{Context Mode:} & m & ::= & \cmode \mid \umode \\
+\text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\
 \text{Bound:} & b & ::= & n \mid x \plus n \\
-              & \bvar & ::= & (b,b) \\[0.5em]
+              & \bvar & ::= & (b,b) \\
 
      \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{ \xi}
-\\[0.5em]
-
+\\
 \text{Type Flag:}&\kappa &::=& nt \mid \cdot
-\\[0.5em]
-
+\\
 \text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau} \mid \tfun{\overline{x}}{\overline{\tau}}{\tau}
-\\[0.5em]
-
+\\
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}  \\[0.2em]
-&&\mid& \estrlen{x} \mid\estar{e}\mid\eassign{e}{e}  \\[0.2em]
-&&\mid& \elet{x}{e}{e} \mid \eif{e}{e}{e}
-\\[0.2em]
-&&\mid&  \emalloc{\xi}{\omega} \mid \ecall{e}{\overline{e}}
-\\[0.2em]
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}  \\
+&&\mid& \estrlen{x} \mid \emalloc{\xi}{\omega} \mid\estar{e}\mid\eassign{e}{e}  \\
+&&\mid& \elet{x}{e}{e} \mid \eif{e}{e}{e} \mid \ecall{e}{\overline{e}}
+\\
 &&\mid&\eunchecked{\overline{x}}{e}
 \mid \echecked{\overline{x}}{e}
-
 \end{array}
@@
 
+\ignore{
 \begin{figure}[t]
@@
 \end{figure}
+}
 
@@
 
-This section describes the formal model of \systemname, named
+This section describes the formal core model of \systemname, named
 \lang, making precise its syntax, semantics, type system, and compilation. It also
-develops \lang's meta-theory, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
+develops \lang's meta-theories, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
 
@@
 
-The syntax of \lang is given by the expression-based
-language presented in Fig.~\ref{fig:checkc-syn}.
-
+The \lang syntax of is given in Fig.~\ref{fig:checkc-syn}.
 There are two type notions in \lang.  Types $\tau$ classify
@@
 \begin{array}{rcl}
-$\code{array_tptr<}$\tau$\code{> : count(}$n$\code{)}$
+$\code{_t_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
 &\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
 \\[0.2em]
-$\code{nt_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
+$\code{_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
 &\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
@@
 \\[0.2em]
-$\code{tptr<(int)(nt_array_tptr<}$\tau$\code{> : count(}$n$\code{),}$\\
-\qquad\qquad$\code{nt_array_tptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
+$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{),}$\\
+\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
 \\[0.2em]
@@
 % \mwh{Added reference back to Section 2-B}
-
 Integer literals $n$ are annotated with a type $\tau$ which can be either
@@
 
-\code{dyn_bounds_cast<array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}
-
 \noindent
-becomes $\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$.
+{\footnotesize
+\code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}
+}
+becomes 
+{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}
+.
 
@@
 there are four differences.
-First, the \systemname type annotations have well-formed restrictions in \Cref{fig:wftypes}, 
-for maintaining non-exposure.
+First, the \systemname type annotations have well-formed restrictions for maintaining non-exposure.
 Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
 It is worth noting that pointer modes are a three point partial order ($\le$),
-where $\tmode$ is the infimum, and
-$\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
+where $\tmode$ is the infimum, and $\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
 such that $\tmode$ is projected as $\umode$.
@@
 One example usage of the nested context-switching is the checked function callbacks inside 
-an unsafe region in \Cref{fig:checkedc-example-2} and \ref{fig:checkedc-example-3}.
+an unsafe region in \Cref{lst:humanadjust} and \Cref{lst:humantaint}.
 To guarantee the non-exposure safety,
@@
 }
-{\footnotesize
-\[
-\begin{array}{l} 
-\mode(E)=\mode'(E,\cmode)
-\\[0.2em]
-\mode'(\Box,m)=m
-\\[0.2em]
-\mode'(\eunchecked{\overline{x}}{E},m)=\mode'(E,\umode)
-\\[0.2em]
-\mode'(\echecked{\overline{x}}{E},m)=\mode'(E,\cmode)
-\\[0.2em]
-\mode'(\alpha({E}),m)=\mode'(E,m)\;\;{[\emph{owise}]}
-\end{array}
-\]
-}
   \caption{\lang Semantics: Evaluation}
@@
 
-Here, we discuss the \lang type system and operational semantics; 
+Here, we discuss the \lang type system (\Cref{fig:type-system-1}) and operational semantics (\Cref{fig:semantics}); 
 mainly focusing on the new changes on top of \checkedc in \cite{li22checkedc} with function pointers, modes, and function calls.
@@
 The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the
-stepped redex within $e$ was in a checked ($\cmode$) or
-unchecked ($\umode$) region.
+stepped redex within $e$ was in a $\cmode$ or $\umode$ region.
 
 The rules for the main operational semantics
-judgment---\emph{evaluation}---are given at the middle of
+judgment \emph{evaluation} are given at the bottom of
 Fig.~\ref{fig:c-context}.
@@
 shortly.
+The $\mode$ function  at the bottom of Fig.~\ref{fig:c-context}
+determines the context mode that the expression $e'$ locates based on the context $E$.
+In \Cref{lst:humantaint}, the function call expression \code{read_msg} has $\umode$ mode since it is inside a tainted function.
 The second rule describes the exception handling 
@@
 \end{DIFnomarkup}
-
-The $\mode$ function  at the bottom of Fig.~\ref{fig:c-context}
-describes the context mode determination in each evaluation step based on the context $E$.
-For any program execution, the function starts the mode computation with $\cmode$ ($\mode(E)=\mode'(E,\cmode)$).
-The result context mode depends on where $\Box$ locates.
-If it occurs within $E$ in $(\eunchecked{\overline{x}}{E})$
-that has no surrounding $\echeckedtext$ block,
-the mode is $\umode$; otherwise, the mode is $\cmode$.
-$\mode'(\alpha({E}),m)=\mode'(E,m)$ represent other construct cases 
-that are not $\echeckedtext$ and $\euncheckedtext$; in such case, 
-the function recursively traverses the sub-context to find the context mode.
 
@@
 refers to that the value $n_a$ is well-defined in $\heap(m,n_a)$ and has type $\tau$, if $\tau$ is a pointer.
-
 Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
@@
     {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\\\
+        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
          \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
@@
 
-\begin{DIFnomarkup}
-\begin{figure}
-{\small
-\[\hspace*{-1.2em}
-\begin{array}{l}
-\textcolor{blue}{\text{Bound Inequality and Equality:}}\\[0.3em]
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
-     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
-     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
-     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
-     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
-     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
-    \end{array}
-  \\[1.5em]
-\textcolor{blue}{\text{Type Equility:}}\\
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-     && \tint & =_{\Theta} & \tint\\
-     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
-     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
-             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
-
-    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
-
- &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
-                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
-    \end{array}
-  \\[0.7em]
-\textcolor{blue}{\text{Subtype:}}\\[0.3em]
-
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
-
-    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
-    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
-    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
-    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
-\\[0.2em]
-\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
-
-    \end{array}
-\end{array}
-  \]
-}
-{\footnotesize
-\[
-\begin{array}{l}
-n'+n = add(n',n)
-\qquad
-(x+n')+n = x+add(n',n)\\
-\textit{cond}(\overline{x},\tau,\overline{y},\tau')
-=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
-  \wedge \overline{y}\cupdot\overline{z}
-  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
-\\\qquad\qquad\qquad\qquad\qquad
-  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
-\end{array}
-\]
-}
-  \caption{Type Equality and Subtyping}
-  \label{fig:checkc-subtype}
-\end{figure}
-\end{DIFnomarkup}
-
 \myparagraph{Type Equality and Subtyping and Casting}
 %
-In \lang, type equality $\tau=_{\Theta}\tau'$
+In \lang, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
+We provide some example descriptions here.
+Type equality $\tau=_{\Theta}\tau'$
 is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
-and the alpha equivalence of two function types in \Cref{fig:checkc-subtype};
+and the alpha equivalence of two function types;
 i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
@@
 
- The \textsc{T-CastPtr} rule
+The \textsc{T-CastPtr} rule in \Cref{fig:type-system-1}
 permits casting from an expression of type $\tau'$ to a checked pointer when
 $\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and is built on the type equality
-($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
-rules ensure the relation is transitive. Most of the rules manage
-casting between array pointer types. The rule 
-($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
+$\sqsubseteq$ is built on the type equality ($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). 
+The rule  ($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
 \tarrayptr{b_l}{b_h}{\tau}{m}$) permits treating a singleton
@@
 % users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
-
-Since bounds expressions may
-contain variables, determining assumptions like $b_l \leq b_l'$
-requires reasoning about the probable values of these variables'. The type
-system uses $\Theta$ to make such reasoning more precise.
-$\Theta$ is a map from variables $x$ to
-equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
-It maps variables to equations that are recorded along the type checking procedure.
-If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
-$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
-such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
-\iftr
-  Appendix~\ref{app:le}.
-\else
-  the supplemental report~\cite{checkedc-tech-report}.
-\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
 
@@
 \myparagraph{Constant Validity}
-Rules \textsc{T-ConstU} and \textsc{T-ConstC}
-describes type assumptions for constants appearing in a program.
-$\cmode(\tau)$ judges that a constant pointer 
-in an unchecked region cannot be of a checked type,
-which represents the assumption that programmers 
+Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
+describe type assumptions for constants appearing in a program.
+$\neg \cmode(\tau)$ judges that a constant pointer 
+in an unchecked region cannot be of a checked type.
+The restriction ensures that programmers 
 cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
 In rule \textsc{T-ConstC}, we requires a static 
-verification procedure for validating a constant pointer, 
-which is similar to the dynamic verification process in \Cref{sec:semantics}.
-
-The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
-validate the constant $\evalue{n}{\tau}$, 
+verification procedure for validating a constant pointer in \Cref{fig:const-type}.
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
+validates the constant $\evalue{n}{\tau}$, 
 where $\heap(m)$ is the initial heap that the constant resides on and
@@
 typed.
-
 A checked pointer checks validity in type step as rule \textsc{T-ConstC},
 while a tainted/unchecked pointer does not check for such during the type checking.
-Tainted pointers are validated through the validity check in dynamic execution as we mentioned above.
-
-\begin{figure}[t]
-{\small
-{\captionsetup[lstlisting]{margin = 8 mm}
-  \begin{lstlisting}[xleftmargin=8 mm]
-nt_array_ptr<char> safe_strcat
-   (nt_array_ptr<char> dst : count(n),
-    nt_array_ptr<char> src : count(0), int n) {
-  int x = strlen(dst);
-  int y = strlen(src);
-  nt_array_ptr<char> c : count(n) =
-    dyn_bounds_cast
-           <nt_array_ptr<char>>(dst,count(n));
-    // sets c == dst with bound n (not x)
-  if (x+y < n) {
-    for (int i = 0; i < y; ++i)
-      *(c+x+i) = *(src+i);
-    *(c+x+y) = '\0';
-    return dst;
-  }
-  return null;
-}
-  \end{lstlisting}
-}
-}
-\caption{Implementation of safe \code{strcat}}
-\label{fig:strcat-ex}
-\end{figure}
+Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.
 
@@
 check all free variables in $e$ are within $\overline{x}$;
-the types for $\overline{x}$ and the final return type $\tau$ have no checked pointers.
+the types for $\overline{x}$ and the final return type $\tau$ of $e$ have no checked pointers.
+Otherwise, it violates the non-exposure safety.
+For example, \code{read_msg} in the \code{handle_request} function is tainted in \Cref{lst:humantaint},
+if any argument for \code{handle_request} is a checked pointer, 
+it means that we are exposing a checked pointer address to unsafe regions.
 A $\echeckedtext$ or $\euncheckedtext$ block represents 
@@
 %
-Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
+Rules \textsc{T-Let} and \textsc{T-LetInt} in in \Cref{fig:type-system-1} type a $\elettext$ expression, which also admits
 type dependency. 
-In particular, the result of evaluating a $\elettext$
+In particular, the result of evaluating a $\elettext$ expression
 may have a type that refers to one of its bound variables (e.g., if
@@
 structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
-
 Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
@@
 
-Rule \textsc{T-Fun} is the dependent function call rule. 
+Rule \textsc{T-Fun} in \Cref{fig:type-system-1} is the dependent function call rule. 
 Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
@@
 also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
-Consider the \code{safe_strcat} function in
-Fig.~\ref{fig:strcat-ex}; its parameter type for \code{dst} 
-depends on \code{n}. The \textsc{T-Fun} rule will substitute 
+Consider the \code{process_req2} function in
+Fig.~\ref{lst:final}; its parameter type for \code{msg} 
+depends on \code{m_1}. The \textsc{T-Fun} rule will substitute 
 \code{n} with the argument at a call-site.
-
 The semantics manages variable scopes using the special $\erettext$
@@
 
-Function calls are handled by \textsc{S-FunC} and \textsc{S-FunT},
+\textsc{S-FunC} and \textsc{S-FunT} are
 for $\cmode$ and $\tmode$ mode function pointers, respectively. 
@@
  the function definition in $n$'s location in the global function store $\Xi$,
-% 
-% LEO: This is the first occurence of \Xi. It is explained,
-% but reads a bit awkward as it's unclear this is its "introduction"
 which maps function pointers to
@@
 function body. 
-
 Similar to \heap, the global function store $\Xi$ is also partitioned into
@@
 
-The \systemname{} functions are dependent functions.
-Recall that array
-bounds in types may refer to in-scope variables; e.g., parameter
-\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
-2-3 in \Cref{fig:strcat-ex}. 
-%All dependent bound variables appearing in function headers 
-%are bounded by a variable list in the function pointer type, which is checked in the type rule.
-Semantically,
-the call is expanded into a \texttt{let} which binds
-parameter variables $\overline{x}$ to the actual arguments
-$\overline{n}$, but annotated with the parameter types
-$\overline{\tau}$ (this will be safe for type-correct programs). 
-The function body $e$ is wrapped in a static cast
-$(\tau[\overline{n} / \overline{x}])$ which is the function's return
-type but with any integer parameter variables $\overline{x}$ appearing in that
-type, as type bound variables,
-substituted with the call's actual arguments $\overline{n}$.
-To see why this is needed, suppose that \code{safe_strcat} in
-Fig.~\ref{fig:strcat-ex} is defined to return a
-\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
-perform a \code{safe_strcat} function call as
-\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
-function returns a value with type \code{nt_array_ptr<int>:count(10)}
-because we substitute bound variable \code{n} in the 
-defined return type with \code{10} from the function call's
-argument list.
-
+\systemname{} has dependent functions, whose semantic explaination is given in \Cref{appx:add-type-sem}.
 Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
@@
 
-This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
+We now show how \lang deals with pointer modes, mode switching and function pointer compilations, 
 with no loss of expressiveness
@@
 \footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
-For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
-we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
-and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
+For example, in compiling the \code{read_msg} function in \Cref{lst:humantaint},
+we place a check \code{verify_fun(read_msg, not_null(c, p_lo, p_hi) && type_match)},
 The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
 is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
-
-\begin{figure}[t!]
-  \begin{small}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-int deref_array(n : int,
-     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
-     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
-  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
-  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
-    * p;
-    * q = 1;
-}
-...
-/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
-/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
-deref_array(5, p0, q0);
-    \end{lstlisting}
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-deref_array(int n, int* p, int * q) {
-  //m is the current context mode
-  let p_lo = 0; let p_hi = n; 
-  let q_lo = 0; let q_hi = n; 
-  /* runtime checks */
-  assert(p_lo <= 0 && 0 <= p_hi);
-  assert(p != 0);
-  *(mode(p) $\wedge$ m,p);
-  verify(q, not_null(m, q_lo, q_hi) 
-             && q_lo <= 0 && 0 <= q_hi);
-  *(mode(q) $\wedge$ m,q)=1;
-}
-...
-deref_array(5, p0, q0);
-    \end{lstlisting}
-\end{small}
-    \caption{Compilation Example for Dependent Functions}
-\label{fig:compilationexample1}
-\end{figure}
-
-For compiling dependent function calls,
-\Cref{fig:compilationexample1} provides a hint.
-Notice that the bounds for the array pointer \code{p} are not passed as
-arguments. Instead, they are initialized according to \code{p}'s
-type---see line~4 of the original \lang program at the top of the figure.
-Line~$3$ of the generated code
-sets the lower bound  to \code{0} and the
-upper bound to \code{n}.
+The other compilation rules are given in \Cref{appx:add-type-sem}.
 
@@
 These proofs have been conducted in our Coq model.
-Type soundness relies on several \emph{well-formedness}:
-
-\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
-A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
-\end{defi}
-
-\begin{defi}[Heap Well-formedness]
-For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
-(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
-contains no free variables. 
-\end{defi}
-
-\begin{defi}[Stack Well-formedness]
-A stack snapshot $\varphi$ is well-formed if
-for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
-contains no free variables. 
-\end{defi}
-
-We also need to introduce a notion of
-\emph{consistency}, relating heap environments before and after a
-reduction step, and type environments, predicate sets, and stack
-snapshots together.
-
-\begin{defi}[Stack Consistency]
-A type environment $\Gamma$, variable predicate set $\Theta$, and
-stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
-\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
-$\Gamma(x) = \tau$ for some $\tau$ and 
-$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
-\end{defi}
-
-\begin{defi}[Checked Stack-Heap Consistency]
-A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
-for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
-\end{defi}
-
-\begin{defi}[Checked Heap-Heap Consistency]
-A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
-for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
-\end{defi}
-
-%Moreover, as a program evaluates, its expression may contain literals
-%$\evalue{n}{\tau}$ where $\tau$ is a pointer type, i.e., $n$ is an
-%index in $\heap$ (perhaps because $n$ was chosen by
-%$\mathtt{malloc}$). The normal type-checking judgment for $e$ is
-%implicitly parameterized by $\heap$, and the rules for type-checking
-%literals confirm that pointed-to heap cells are compatible with
-%(subtypes of) the pointer's type annotation; in turn this check may
-%precipitate checking the type consistency of the heap itself. We
-%follow the same approach as \citet{ruef18checkedc-incr}, and show the
-%rules in
-%  Fig.~\ref{fig:const-type}; the judgment
-%  $\Theta;\heap;\sigma \vdash_m n : \tau$ is used to confirm literal
-%  well-typing, where $\sigma$ is a set of pointer literals already
-%  checked in $\heap$ (to allow pointer cycles). See
-%  \iftr
-%  Appendix~\ref{sec:literal-pointer-typing}
-%  \else
-%  the supplemental report~\cite{checkedc-tech-report}
-%  \fi
-%   for further discussion.
-
+Type soundness relies on several \emph{well-formedness} given in \ref{li22checkedc} and \Cref{appx:add-type-sem}.
 Progress states that a \lang program can always make a move:
@@
 
-In the \systemname compiler, we also restore the compiler simulation theorem, such that executing a \systemname program is simulated by executing the compiled LLVM program, which is introduced in the \checkedc formalism \cite{li22checkedc} and \Cref{appx:comp1}.
+In the \systemname compiler, we also restore the compiler simulation theorem, such that executing a \systemname program is simulated by executing the compiled LLVM program.
+The details are given in the \checkedc formalism \cite{li22checkedc} and \Cref{appx:comp1}.
 
 We formalize both the compilation procedure and the simulation
-theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
+theorem in the semantic model we developed in Redex (same as our Coq model),
 and then attempt to falsify it via Redex's support for random
@@
 
-\ignore{
- We formulate our simulation theorem differently
-because the standard simulation theorem imposes a very strong
-syntactic restriction to the compilation strategy. Very often, $(\dot
-\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
-semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
-but we are unable to syntactically equate the two configurations due
-to the extra binders generated for dynamic checks and ANF
-transformation. In earlier versions of the Redex model, we attempted
-to change the compilation rules so the configurations could match
-syntactically. However, the approach scaled poorly as we added
-additional rules. 
-This slight relaxation on the equivalence relation
-between target configurations allows us to specify compilation more
-naturally without having to worry about syntactic constraints.
-}
-
 % The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
@@
 
-\item if $r_1 = \ebounds$ or $\enull$, then we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\item if $r_1 = \ebounds$ or $\enull$, then $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
 \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot

### 2023-01-08 sandp23/formal.tex
update.

@@
 In addition, when $\Gamma;\emptyset\vdash
-\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
+\varphi$ and $\varphi$ is well-typed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
 \dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
 and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively. 
-$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
-Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
-reduction relation of $\elang$. Unlike the $\lang$, the semantics of
-$\elang$ does not distinguish checked and unchecked regions.
-
-Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
-the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
-that is similar to the one by~\citet{merigoux2021catala}.
-
-An ordinary simulation property would
+%$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
+%Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
+%reduction relation of $\elang$. Unlike the $\lang$, the semantics of
+%$\elang$ does not distinguish checked and unchecked regions.
+We show an overview of the simulation theorem in \Cref{fig:checkedc-simulation-ref}.
+The simulation theorem is specified in a way that is similar to the one by~\citet{merigoux2021catala}.
+
+Previous ordinary simulation properties would
 replace the middle and bottom parts of the figure with the
@@
 
-As our random generator never generates
-$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
-as it relates to checked code. This limitation makes it
-unnecessary to state the other direction of the simulation theorem
-where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
-that $e_0$ will never enter a stuck state if it is well-typed in
-checked mode.
-
-The current version of the Redex model has been tested against $23000$
-expressions with depth less than $11$. Each expression can
+The current version of the Redex model has been tested against $27500$
+test cases with depth less than $15$. Each expression can
 reduce multiple steps, and we test simulation between every two

### 2023-01-08 sandp23/formal.tex
Modifying first section of formalism

@@
 
-This section describes the formal core model of \systemname, named
-\lang, making precise its syntax, semantics, type system, and compilation. It also
-develops \lang's meta-theories, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
+This section describes the formal core model of~\systemname,
+named~\lang.
+We precisely present its syntax, semantics, type
+system, and compilation.
+The formal model also enables us to develop~\lang’s meta-
+theories, including the type soundness, non-exposure, non-crashing,
+and the compiler simulation theorems, which we present in~\apdx{sec:theorem}.
+
+%This section describes the formal core model of \systemname, named
+%\lang, making precise its syntax, semantics, type system, and compilation. It also
+%develops \lang's meta-theories, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
 
 \subsection{Syntax}\label{sec:syntax}
-
-The \lang syntax of is given in Fig.~\ref{fig:checkc-syn}.
-There are two type notions in \lang.  Types $\tau$ classify
-word-sized values including integers and pointers, while types
-$\omega$ classify multi-word values such as arrays, null-terminated
-arrays, functions, and single-word-size values. 
+The~\fig{fig:checkc-syn} shows a simplified syntax of~\lang.
+\subsubsection{Types ($\omega$)}
+At a high level, we classify types as word-size values, multi-word values, or functions.
+The word-size values can be either integers or pointers.
 Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
-($\xi$, the difference between context and pointer modes is introduced shortly below)
-that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
+($\xi$, the difference between context and pointer modes is introduced shortly below) that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
 and a type ($\omega$) denoting valid values that can be pointed to.
-Array types include both the type of
-elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
-lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
-limited to integer literals $n$ and expressions $x + n$.
-Whether an array pointer is null terminated or not is determined by annotation
-$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide $\cdot$ when writing types).
-
-\lang{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
-reflect its dependent function declarations,
-where $\overline{x}$ represents
-a list of \tint{} type variables in a dependent function header
-that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
-We have a well-formed requirement for a function type;
-that is, all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
-Here is the
-corresponding \systemname syntax for these types:
+
+Multi-word values represent arrays and null-terminated arrays, which include both the type of elements ($\tau$) and a bound ($\bvar$).
+The bounds can be either integer literals $n$, expressions $x + n$, or an explicit specification of lower and upper bound,~\ie $(b_l,b_h)$.
+We denote null-terminated arrays using $\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$ otherwise (we elide $\cdot$ when writing types).
+The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
+An example representation of an array and null-terminated array in~\lang syntax is shown below:
 \[\hspace*{-0.5em}
@@
 \end{array}
-\\[0.2em]
+\end{array}
+\]
+To simplify the representation, we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
+$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
+be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
+$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
+
+Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
+where $\overline{x}$ represents a list of \tint{} type variables that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
+An example of a function pointer type is shown below:
+\[\hspace*{-0.5em}
+\begin{array}{l}
 $\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{),}$\\
-\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
+\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>: count(}$n$\code{))>}$
 \\[0.2em]
@@
 \]
-As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
-$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the firs two examples above could
-be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
-$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-
-\lang expressions include literals ($n\!:\!\tau$), variables ($x$),
- addition ($\ebinop{e_1}{e_2}$), static casts ($\ecast{\tau}{e}$), 
-dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}},
-the \texttt{strlen} operation ($\estrlen{x}$),
-pointer dereference and assignment ($\estar{e}$)
-and ($\eassign{e_1}{e_2}$), resp.),
-let binding ($\elet{x}{e_1}{e_2}$),
-conditionals ($\eif{e}{e_1}{e_2}$),
+Note that the well-formed requirement (\fig{fig:wffunctions} in Appendix) of a function type ensures that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+\aravind{The syntax of well-formed requirements for functions is incorrect. It looks like the forall qualifier is missing.}
+
+%The \lang syntax of is given in Fig.~\ref{fig:checkc-syn}.
+%There are two type notions in \lang.  Types $\tau$ classify
+%word-sized values including integers and pointers, while types
+%$\omega$ classify multi-word values such as arrays, null-terminated
+%arrays, functions, and single-word-size values. 
+%Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
+%($\xi$, the difference between context and pointer modes is introduced shortly below)
+%that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
+%and a type ($\omega$) denoting valid values that can be pointed to.
+%Array types include both the type of
+%elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
+%lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
+%limited to integer literals $n$ and expressions $x + n$.
+%Whether an array pointer is null terminated or not is determined by annotation
+%$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
+%otherwise (we elide $\cdot$ when writing types).
+%
+%\lang{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
+%reflect its dependent function declarations,
+%where $\overline{x}$ represents
+%a list of \tint{} type variables in a dependent function header
+%that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
+%We have a well-formed requirement for a function type;
+%that is, all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+%Here is the
+%corresponding \systemname syntax for these types:
+%\[\hspace*{-0.5em}
+%\begin{array}{l}
+%\begin{array}{rcl}
+%$\code{_t_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
+%&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
+%\\[0.2em]
+%$\code{_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
+%&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
+%\end{array}
+%\\[0.2em]
+%$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{),}$\\
+%\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
+%\\[0.2em]
+%\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
+%\end{array}
+%\]
+%As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
+%$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the firs two examples above could
+%be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
+%$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
+\subsubsection{Expressions ($e$)}
+\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$)($n\!:\!\tau$), pointer dereference and assignment ($\estar{e}$) and ($\eassign{e_1}{e_2}$), resp.)
+along with expressions that require special handling, such as,
+ static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}}, the \texttt{strlen} operation ($\estrlen{x}$),
 memory allocation ($\emalloc{\xi}{\omega}$), 
@@
 unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
-% \review{III.A.: the "dynamic cast" terminology may be briefly confused for C++'s
-%   RTTI-based dynamic cast feature}
-% \mwh{Added reference back to Section 2-B}
-Integer literals $n$ are annotated with a type $\tau$ which can be either
-$\tint$, or $\tptr{\omega}{\xi}$ in the case $n$ is being used as 
-a heap address (this is useful for the semantics);
-$\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer, as usual. 
-The $\texttt{strlen}$ expression operates on variables $x$
-rather than arbitrary expressions to simplify managing
-bounds information in the type system; the more general case can be
-encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
-casts; e.g., the following %
+For instance, a dynamic bounds cast expression
 
@@
 
-Compared to the former \checkedc model \cite{li22checkedc},
-there are four differences.
-First, the \systemname type annotations have well-formed restrictions for maintaining non-exposure.
-Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
-It is worth noting that pointer modes are a three point partial order ($\le$),
-where $\tmode$ is the infimum, and $\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
-such that $\tmode$ is projected as $\umode$.
-Second, $\emalloc{\xi}{\omega}$ includes a mode flag
-$\xi$ for allocating different pointers in different heaps.
+We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
+For instance, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer.
+The~\code{strlen} expression plays an important role in the type-checking of null-terminated arrays.
+For simplifying type checking, we assume~\code{strlen} to operate on variables $x$ rather than arbitrary expressions.
+The heap allocation expression $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions (\cregion or \ucregion).
 We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
-Fourth, $\echeckedtext$ blocks are added to the system, 
-which permits the nested context-switching between $\echeckedtext$ (represented by context mode $\cmode$)
-and $\euncheckedtext$ (represented by context mode $\umode$) code regions.
-One example usage of the nested context-switching is the checked function callbacks inside 
-an unsafe region in \Cref{lst:humanadjust} and \Cref{lst:humantaint}.
-To guarantee the non-exposure safety,
- we extend the $\echeckedtext$ and $\euncheckedtext$ block syntax to be 
-$\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$:
-$\overline{x}$ restricts all free variables appearing in $e$, and they cannot be checked pointers.
-
-\lang aims to be simple enough to work with, but powerful enough to
+The $\echeckedtext$ and $\euncheckedtext$ expressions are used to represent the corresponding code regions.
+To guarantee the non-exposure property, we extend the base syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$.
+Wherein $\overline{x}$ restricts all free variables appearing in $e$ to be checked or unchecked pointers according to the corresponding region.
+
+\lang aims to be simple enough to work with but powerful enough to
 encode realistic \systemname idioms. For example, mutable local
@@
 and loops can be encoded as recursive function calls. \code{struct}s are
-not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
-actually in our model, and developed in
-\iftr
-Appendix~\ref{appx:struct}.
-\else
-the supplemental report~\cite{checkedc-tech-report}.
-\fi
-C-style \code{union}s have no safe typing
-in \checkedc, so we omit them.
+not included in~\fig{fig:checkc-syn} for space reasons, but they are
+supported as shown in~\apdx{appx:struct}.
+C-style \code{union}s have no safe typing in \checkedc, so we omit them.
+
+Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
+
+%\lang expressions include literals ($n\!:\!\tau$), variables ($x$),
+% addition ($\ebinop{e_1}{e_2}$), static casts ($\ecast{\tau}{e}$), 
+%dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}},
+%the \texttt{strlen} operation ($\estrlen{x}$),
+%pointer dereference and assignment ($\estar{e}$)
+%and ($\eassign{e_1}{e_2}$), resp.),
+%let binding ($\elet{x}{e_1}{e_2}$),
+%conditionals ($\eif{e}{e_1}{e_2}$),
+%memory allocation ($\emalloc{\xi}{\omega}$), 
+%function calls ($\ecall{e}{\overline{e}}$),
+%unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
+%% \review{III.A.: the "dynamic cast" terminology may be briefly confused for C++'s
+%%   RTTI-based dynamic cast feature}
+%% \mwh{Added reference back to Section 2-B}
+%Integer literals $n$ are annotated with a type $\tau$ which can be either
+%$\tint$, or $\tptr{\omega}{\xi}$ in the case $n$ is being used as 
+%a heap address (this is useful for the semantics);
+%$\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer, as usual. 
+%The $\texttt{strlen}$ expression operates on variables $x$
+%rather than arbitrary expressions to simplify managing
+%bounds information in the type system; the more general case can be
+%encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
+%casts; e.g., the following %
+%
+%\noindent
+%{\footnotesize
+%\code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}
+%}
+%becomes 
+%{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}
+%. 
+%
+%Compared to the former \checkedc model \cite{li22checkedc},
+%there are four differences.
+%First, the \systemname type annotations have well-formed restrictions for maintaining non-exposure.
+%Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
+%It is worth noting that pointer modes are a three point partial order ($\le$),
+%where $\tmode$ is the infimum, and $\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
+%such that $\tmode$ is projected as $\umode$.
+%Second, $\emalloc{\xi}{\omega}$ includes a mode flag
+%$\xi$ for allocating different pointers in different heaps.
+%We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
+%Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
+%Fourth, $\echeckedtext$ blocks are added to the system, 
+%which permits the nested context-switching between $\echeckedtext$ (represented by context mode $\cmode$)
+%and $\euncheckedtext$ (represented by context mode $\umode$) code regions.
+%One example usage of the nested context-switching is the checked function callbacks inside 
+%an unsafe region in \Cref{lst:humanadjust} and \Cref{lst:humantaint}.
+%To guarantee the non-exposure safety,
+% we extend the $\echeckedtext$ and $\euncheckedtext$ block syntax to be 
+%$\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$:
+%$\overline{x}$ restricts all free variables appearing in $e$, and they cannot be checked pointers.
+%
+%\lang aims to be simple enough to work with, but powerful enough to
+%encode realistic \systemname idioms. For example, mutable local
+%variables can be encoded as immutable locals that point to the heap;
+%the use of \code{&} can be simulated with \code{malloc};
+%and loops can be encoded as recursive function calls. \code{struct}s are
+%not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
+%actually in our model, and developed in
+%\iftr
+%Appendix~\ref{appx:struct}.
+%\else
+%the supplemental report~\cite{checkedc-tech-report}.
+%\fi
+%C-style \code{union}s have no safe typing
+%in \checkedc, so we omit them.
+
+\aravind{Why we have only function pointers in function call?}
 
@@
 The other compilation rules are given in \Cref{appx:add-type-sem}.
-
-\subsection{Meta Theories}\label{sec:theorem}
-
-% Before we present our main theorems, we need to first
-% discuss the meaning what a pointer being well-typed in a given heap
-% snapshot $\heap$ means, which is captured by rules in
-% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
-% simply checks if a given variable has the defined type in $\Gamma$;
-% the constant rule ($\textsc{T-Const}$) is slightly more involved.
-% First, it ensures that the type annotation $\tau$ does not contain any
-% free variables. More importantly, it ensures that the pointer points
-% to a location that makes sense in a given heap.
-%  
-%  
-%  The $\size$ function in Fig.~\ref{fig:const-type}
-% refers to the \code{sizeof} function in C computing the number of
-% bytes for a type.
-%  
-%  
-%  Second, we
-% require that any constant ($\evalue{n}{\tau}$) should make sense in
-% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
-% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
-% constant set containing the constants that have been verified by the
-% relation. For every constant $\evalue{n}{\tau}$, it is either an
-% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
-% zero-valued number ($n=0$), checked in $\sigma$
-% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
-% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
-% location $n+i$ in the range of the pointer (if $\omega$ is a word
-% type, range is $[0,1)$; if $\omega$ is an array type
-%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
-%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
-%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
-%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
-%       \vdash n_a : \tau_a$.
-%  
-%  
-% \begin{figure}[t]
-% {\small
-% \text{Type Rules for Constants and Variables:}
-% \begin{mathpar}
-%   \inferrule[T-Var]
-%       {x : \tau \in \Gamma}
-%       {\Gamma;\Theta \vdash_m x : \tau}
-%  
-%   \inferrule[T-Const]
-%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
-%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
-% \end{mathpar}
-%     
-% \text{Rules for Checking Constant Pointers In Heap:}
-% \begin{mathpar}
-%   \inferrule
-%       {}
-%       {\sigma \vdash n : \tint}
-%  
-%   \inferrule
-%       {}
-%       {\sigma \vdash n : \tptr{\omega}{\umode}}
-%  
-%   \inferrule
-%       {}
-%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
-%  
-%   \inferrule
-%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
-%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-%  
-%   \inferrule
-%       {\forall i \in [0,\size(\omega)) .
-%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
-%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-% \end{mathpar}
-% }
-% \caption{Type Rules for Checking Constants/Variables}
-% \label{fig:const-type}
-% \end{figure}
-
-% \review{
-%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
-%   something, I didn't see such a definition in the paper.}
-% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
-%   assumption of well typing; we have added more details about that.}
-
-In this subsection, we focus on our main meta-theoretic results about
-\lang: type soundness (progress and preservation),
-non-exposure, and non-crashing.
-These proofs have been conducted in our Coq model.
-Type soundness relies on several \emph{well-formedness} given in \ref{li22checkedc} and \Cref{appx:add-type-sem}.
-Progress states that a \lang program can always make a move:
-
-\begin{thm}[Progress]\label{thm:progress}
-
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
-one of the following holds:
-
-\begin{itemize}
-
-\item $e$ is a value ($\evalue{n}{\tau}$).
-
-\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.
-
-\end{itemize}
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-
-\noindent
-There are two forms of preservation regarding the checked and unchecked regions.
-Checked Preservation states that a reduction step preserves both the
-type and consistency of the program being reduced.
-Unchecked Preservation states that any evaluation happens at unchecked region does not affect the checked heap.
-
-\begin{thm}[Checked Preservation]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
-checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$ and $\tau'$ that are well formed, checked region consistent
-($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
-well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
-$\tau'\sqsubseteq_{\Theta} \tau$.
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-%\smallskip
-\begin{thm}[Unchecked Preservation]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
-\end{thm}
-
-Using the above theorem, we first show the non-exposure theorem,
-where code in unchecked region cannot observe a valid checked pointer address.
-
-\begin{thm}[Non-Exposure]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
-where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
-thus, it is not a checked pointer.
-\end{thm}
-
-We now state our main result, {\em non-crashing},
-which suggests that a well-typed program can never be \emph{stuck} (expression
-$e$ is a non-value that cannot take a step\footnote{Note that
-  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
-  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
-  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
-  safety.}).
-
-% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r.
-Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
-% \liyi{It is a typo.
-We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region.
-This is a bit due to the space limitation.
-The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement.
-}
-
-\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
-  program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are well-formed and consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
-if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
-\tau$ for some $\tau$) and there exists
-$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
-$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
-\end{thm}
-
-% \review{ 
-% My interpretation of section IV.C is that the authors have a pen-and-paper
-%   proof of Theorem 4, and that random testing using the Redex models has been
-%   used to gain confidence in such a proof. Is this correct?}
-% \liyi{Yes, this is correct.}
-% \mwh{No, not correct: We have no pen-and-paper proof, just the
-%   model. Clarify here}
-% showing that
-% extensive random testing fails to falsify the bisimilarity
-% property. \mwh{Better way to state the previous?}
-
-In the \systemname compiler, we also restore the compiler simulation theorem, such that executing a \systemname program is simulated by executing the compiled LLVM program.
-The details are given in the \checkedc formalism \cite{li22checkedc} and \Cref{appx:comp1}.
-
-We formalize both the compilation procedure and the simulation
-theorem in the semantic model we developed in Redex (same as our Coq model),
-and then attempt to falsify it via Redex's support for random
-testing. Redex allows us
-  to specify compilation as logical rules (an extension
-  of typing), but then execute it algorithmically to
-  automatically test whether simulation holds. This process revealed
-  several bugs in compilation and the theorem statement.
-%
-  % us gain confidence that our original pen and paper proof of
-  % simulation remains true with the addition of variable bounds. }
-We ultimately plan to prove simulation in the Coq model.
-
-%Turning to the simulation theorem: We first introduce notation
-%used to specify the theorem.
-We use the notation $\gg$ to
-indicate the \emph{erasure} of stack and heap---the rhs is the same as
-the lhs but with type annotations removed:
-\begin{equation*}
-  \begin{split}
-    \heap  \gg & \dot \heap \\
-    \varphi \gg & \dot \varphi
-  \end{split}
-\end{equation*}
-In addition, when $\Gamma;\emptyset\vdash
-\varphi$ and $\varphi$ is well-typed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
-\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
-and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively. 
-%$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
-%Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
-%reduction relation of $\elang$. Unlike the $\lang$, the semantics of
-%$\elang$ does not distinguish checked and unchecked regions.
-We show an overview of the simulation theorem in \Cref{fig:checkedc-simulation-ref}.
-The simulation theorem is specified in a way that is similar to the one by~\citet{merigoux2021catala}.
-
-Previous ordinary simulation properties would
-replace the middle and bottom parts of the figure with the
-following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
-  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
-Instead, we relate two erased configurations using the relation $\sim$,
-which only requires that the two configurations will eventually reduce
-to the same state.
-
-% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
-\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
-and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
-
-\begin{itemize}
-
-\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\varphi,\dot \heap,\dot e)$ and $(\dot
-\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-\dot \heap,\dot e)$.
-
-\item if $r_1 = \ebounds$ or $\enull$, then $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-\varphi_1$, $\heap_1 \gg \dot \heap_1$.
-
-\end{itemize}
-\end{thm}
-
-% when $r_1 = e_1$ for
-% some $e_1$ and
-% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
-% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \varphi,\dot \heap,\dot e)$ and $(\dot
-% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
-
-% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
-% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
-% some $e_1$ and
-% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
-% , then
-% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \varphi,\dot \heap,\dot e)$ and $(\dot
-% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
-% \end{thm}
-
-The current version of the Redex model has been tested against $27500$
-test cases with depth less than $15$. Each expression can
-reduce multiple steps, and we test simulation between every two
-adjacent steps to cover a wider range of programs, particularly the
-ones that have a non-empty heap.
-
-\begin{figure}[t]
-{\small
-\[
-\begin{array}{c}
-\begin{tikzpicture}[
-            > = stealth, % arrow head style
-            shorten > = 1pt, % don't touch arrow head to node
-            auto,
-            node distance = 3cm
-        ]
-
-\begin{scope}[every node/.style={draw}]
-    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
-    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
-    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
-    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
-    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
-\end{scope}
-\begin{scope}[every edge/.style={draw=black}]
-
-    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
-    \path [<->] (A) edge node {$\gg$} (C);
-    \path [<->] (B) edge node {$\gg$} (D);
-    \path [dashed,<->] (C) edge node {$\sim$} (D);
-    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
-    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
-\end{scope}
-
-\end{tikzpicture}
-\end{array}
-\]
-}
-\caption{Simulation between \lang and \elang }
-\label{fig:checkedc-simulation-ref}
-\end{figure}
-
-%{\em Proof:} By induction on the number of steps of the \checkedc
-%evaluation ($\longrightarrow_m^*$), using progress and preservation to
-%maintain the invariance of the assumptions.
-
-% \review{
-%   add a paragraph that discusses what are the main changes from [21] in terms
-%   of the technical development (if there are any), e.g., are there any new
-%   challenges that needed to be solved while proving the blame theorem for this
-%   paper's semantics?
-% }
-%  Compared to \citet{ruef18checkedc-incr}, proofs for
-%  \lang were made challenging by the addition of dependently typed
-%  functions and dynamic arrays, and the need to handle bounds widening for NT
-%  array pointers. These features required changes in the runtime
-%%  semantics (adding a stack, and dynamically changing bounds) and in
- % compile-time knowledge of them (to soundly typing widened bounds).
-
-% \ignore{
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \hypo{\evalue{n}{\tau} \in \defscope}
-%     \infer1[T-VConst]{\Gamma;\defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%   \end{prooftree}
-%   \qquad
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma;\defscope \vdash_m e : \tau \\
-%         \Gamma;\defscope \vdash_m e_1 : \tau_1 
-%       \end{matrix}
-%     }
-%     \hypo{
-%       \begin{matrix}
-% \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2 \\
-%             \Gamma;\defscope \vdash_m e_2 : \tau_2
-%           \end{matrix}
-%         }
-%     \infer2[T-If]{\Gamma;\defscope \vdash_m \eif{e}{e_1}{e_2} : \tau_3}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma;\defscope \vdash_m x : \tptr[c]{(\tntarray{l}{0}{\tau})}, l \leq 0 \\
-%         \Gamma, x : \tptr[c]{(\tntarray{l}{1}{\tau})};\defscope \vdash_m e_1 : \tau_1 \\
-%         \Gamma;\defscope \vdash_m e_2 : \tau_2 \\
-%         \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2
-%       \end{matrix}
-%     }
-%     \infer1[T-IfNT]{\Gamma;\defscope \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_3}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         F(f) = \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e  \\
-%         \Gamma; \defscope \vdash_m \tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}] ~~~ 0 \leq i \leq j\\
-%         \Gamma; \defscope \vdash_m e_i : \tau_i' ~~~ 0 \leq i \leq j\\
-%         \Gamma; \defscope \vdash_m \subtype{\tau_i'}{\tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}]}  ~~~ 0 \leq i < j
-%       \end{matrix}
-%     }
-%     \infer1[T-VCall]{\Gamma; \defscope \vdash_m f(\overline{e}) : \tau_j[x_0,\ldots,x_{j-1} \mapsto e_0,\ldots,e_{j-1}]}
-%   \end{prooftree} \\ \\
-% % \inferrule*[lab=T-PtrC]
-% % {
-% %   \tau = \tptr[c]{\omega} \\
-% %   \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\\\
-% %   \wt[\Gamma][\defscope, n^\tau]{H(n+k)}{\tau_k} ~~~ 0 \leq k < j
-% % %  \Gamma,n^\tau \proves H(n+k) : \tau_k~~~0 \leq k < j
-% % }
-
-%   \begin{prooftree}
-%     \hypo{
-%           \Gamma; \defscope \vdash_m e_1 : \tau_1
-%       }
-%       \hypo{        \Gamma, x = e_1 : \tau_1; \defscope \vdash_m e_2 : \tau_2
-% }
-%     \infer2[T-Let]{\Gamma; \defscope \vdash_m \elet{x}{e_1}{e_2} : \tau[x \mapsto e_1]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m \estrlen{y} : \tau_1\\
-%         \Gamma; \defscope \vdash_m y : \tptr[c]{(\tntarray{le}{\_}{\tau_3})}\\
-%         \Gamma, x = \estrlen{y} : \tau_1, \\ y : \tptr[c]{(\tntarray{le}{x}{\tau_3})} ;  \defscope  \vdash_m e_2 : \tau_2
-%         \end{matrix}
-%         }
-%         \infer1[T-LetStr]{\Gamma; \defscope \vdash_m \elet{x}{\estrlen{y}}{e_2} : \tau[x \mapsto e_1]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%         \Gamma; \defscope \vdash_m e : \tptr[c]{(\tntarray{le}{he}{\tau})}
-%       }
-%         \infer1[T-Str]{\Gamma; \defscope \vdash_m \estrlen{y} : \tint}
-%       \end{prooftree} \\ \\
-
-%       \begin{prooftree}
-%         \hypo{
-%           \begin{matrix}
-%             \tau = \tint \vee \tau = \tptr[u]{\omega}~\vee  n=0~ \vee \\ \tau = \tptr[c]{(\tarray{0}{0}{\tau'})} \vee \\
-%             \tau = \tptr[c]{(\tntarray{0}{0}{\tau'})}
-%         \end{matrix}
-%       }
-%       \infer1[T-Base]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%     \end{prooftree} \\ \\
-
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%         \tau = \tptr[c]{\omega} \\
-%         \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\
-%         \Gamma;\defscope, \evalue{n}{\tau}  \vdash_m {H(n+k)} : {\tau_k} ~~~ 0 \leq k < j
-%       \end{matrix}
-%       }
-%       \infer1[T-PtrC]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%     \end{prooftree} \\ \\
-
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%             \Gamma; \defscope \vdash_m e :  {\tptr[m]{\tstruct{T}}} \\
-%             % \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
-%             D(T) = ...; \tau_f~f; ...
-%         \end{matrix}
-%       }
-%       \infer1[T-Amper]{\Gamma; \defscope \vdash_m \eamper{e}{f} : \tptr[m]{\tau_f}}
-%     \end{prooftree} \\ \\
-
-%     \todo[inline]{YL: how to express le - n as a metafunction?}
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}\\
-%         \Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tint \\
-%         le' = le - n, he' = he - n
-%       \end{matrix}
-%     }
-%       \infer1[T-BinopInd]{\Gamma; \defscope \vdash_m (e_1 \plus \evalue{n}{\tau} ) : \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})}}
-%     \end{prooftree}
-% \caption{Typing}
-% \label{fig:typing}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \todo[inline]{no sizeof check because we don't know statically whether the allocation would be null}
-%   \begin{prooftree}
-%     \hypo{
-%       \omega = \tgarray{\alpha}{le}{he}{\tau} \Rightarrow le = 0
-%     }
-%     \infer1[T-Malloc]{\Gamma; \defscope \vdash_m \emalloc{\omega} : \tptr[c]{\omega}}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \Gamma; \defscope \vdash_u e : \tau
-%     }
-%     \infer1[T-Unchecked]{\Gamma; \defscope \vdash_m \eunchecked{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \todo[inline]{any constraints on m'?}
-%   \begin{prooftree}
-%     \hypo{\Gamma; \defscope \vdash_m e :  \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})} }
-%     \infer1[T-DynCast]{\Gamma; \defscope \vdash_m \edyncast{\tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}}{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{\Gamma; \defscope \vdash_m e : \tau'}
-%     \infer1[T-Cast]{\Gamma; \defscope \vdash_m \ecast{\tau}{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\tgarray{\alpha}{le}{he}{\tau}} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tint \\
-%         m' = u \Rightarrow m = u \\
-%       \end{matrix}
-%     }
-%     \infer1[T-Index]{\Gamma; \defscope \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tau \\
-%         m' = u \Rightarrow m = u \\
-%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-%       \end{matrix}
-%     }
-%     \infer1[T-Assign]{\Gamma; \defscope \vdash_m \eassign{e_1}{e_2} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tint \\
-%         \Gamma; \defscope \vdash_m e_3 : \tau \\
-%         m' = u \Rightarrow m = u \\
-%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-%       \end{matrix}
-%     }
-%     \infer1[T-IndAssign]{\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
-%   \end{prooftree}
-
-% \caption{Typing Cont.}
-% \label{fig:typing2}
-% \end{figure}
-% \ignore{
-
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j} \\
-%         \Gamma, x_0 = \kw{none} : \tau_0, \ldots, x_{j-1} = \kw{none} : \tau_{j-1}; \defscope \vdash e : \tau_{r}
-%       \end{matrix}
-%     }
-%     \infer1[WF-Fun]{\Gamma;\defscope \vdash \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \infer0[WF-Nil]{\Gamma \vdash \circ}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma \vdash \tau_0 \\
-%         \Gamma, x_0 = \kw{none} :\tau_0 \vdash x_1:\tau_{1}, \ldots, x_{j}:\tau_{j}
-%       \end{matrix}
-%     }
-%     \infer1[WF-Cons]{\Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \infer0[WF-Int]{\Gamma \vdash \tint}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{   \Gamma \vdash le }
-%      \hypo{   \Gamma \vdash he }
-%       \hypo{  \Gamma \vdash \tau}
-
-%     \infer3[WF-Array]{\Gamma \vdash \tptr[m]{(\tgarray{\alpha}{le}{he}{\tau})}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{T \in dom(D)}
-%     \infer1[WF-Struct]{\Gamma \vdash \tptr[m]{\tstruct{T}}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{\Gamma \vdash \tau}
-%     \infer1[WF-Ptr]{\Gamma \vdash \tptr[m]{\tau}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \infer0[WFB-Int]{\Gamma \vdash i}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{x = e? : \tint \in \Gamma}
-%     \infer1[WFB-Var]{\Gamma \vdash x \plus i}
-%   \end{prooftree} \\\\
-
-%   \caption{Well-Formedness}
-%   \label{fig:wf}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \infer0[Sub-Nt]{\subtype{\tptr[c]{(\tntarray{le}{he}{\tau})}}{\tptr[c]{(\tarray{le}{he}{\tau})}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \infer0[Sub-Refl]{\subtype{\tau}{\tau}}
-%   \end{prooftree} \\ \\
-
-%   \todo[inline]{what about ntarrays?}
-%   \begin{prooftree}
-%     \infer0[Sub-Ptr]{\subtype{\tptr[c]{\tau}}{\tptr{\tarray{0}{1}{\tau}}}}
-%   \end{prooftree} \\ \\
-
-%   \todo[inline]{the subtyping relation is not anti-symmetric..}
-%   \begin{prooftree}
-%     \infer0[Sub-Arr]{\subtype{\tptr[c]{\tarray{0}{1}{\tau}}}{\tptr{\tau}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{\rboundle{le_0}{le_1}}
-%     \hypo{\rboundle{he_1}{he_0}}
-%     \infer2[Sub-Subsume]{\subtype{\tptr{(\tgarray{\alpha}{le}{he}{\tau})}{\cmode}}{ \tptr{(\tgarray{\alpha}{ le_1} {he_1} {\tau})}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     % check t-amper below
-%     \hypo{D(T) = \tau_f~f; ...}
-%     \hypo{\subtype{\tau_f}{\tau}}
-%     \infer2{\subtype{\tptr[c]{\tstruct{T}}}{\tau}}
-%   \end{prooftree}
-%   \caption{Subtyping}
-%   \label{fig:sub}
-% \end{figure}
-% }
-% }
-
-% \ignore{
-
-% \begin{figure*}[t]
-%   \begin{lstlisting}
-% int foo(nt_array_ptr<char> p : count(0)) {
-%   if (* p) {
-%     dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
-%   }
-%   dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
-%   return 0;
-% }
-%   \end{lstlisting}
-% \caption{The example where clang Checked C fails at run-time}
-% \label{fig:clangbad1}
-% \end{figure*}
-
-% \begin{figure*}[t]
-%   \begin{lstlisting}
-% /* nt_array_ptr<char> p : bounds(p,p) */
-% size_t cnt = 0;
-
-% while(*(p+cnt)) {
-%   ++cnt;
-% }
-% dyn_bounds_cast<nt_array_ptr<char>>(p, count(cnt));
-%   \end{lstlisting}
-% \caption{A useful program that the Checked C spec doesn't allow at run-time}
-% \label{fig:clangbad2}
-% \end{figure*}
-% }
-
-% \ignore{
-% The Clang CheckedC implementation uses statically determined
-% bounds to insert run-time checks. In Fig.~\ref{fig:clangbad1}, the
-% \code{dyn_bounds_cast} at line 3 will always succeed, because the
-% compiler knows that within the scope of then branch, the pointer
-% \code{p} must have at least one element. The same cast at line 5,
-% however, will always fail, since there is no way to tell statically
-% whether the program has entered the then branch before. The compiler
-% will check whether the \code{count(1)} bounds specification is
-% contained within the earlier \code{count(0)} specification, resulting
-% in a run-time failure even when we pass in a non-empty string.
-
-% Our formalization diverges from this run-time behavior and instead keeps
-% track of the bounds on the stack. After entering the then branch, we
-% increment the upper bound for \code{p}, effectively making the
-% updated bounds information available even after we exit the if
-% statement. The cast at line 5 will be checking the new bounds against
-% the incremented bounds for non-empty strings.
-
-% Fig.~\ref{fig:clangbad2} gives a more practical example of why keeping
-% track of the bounds on the stack is useful. The program snippet
-% implements the functionality of the \code{strlen} function using a
-% while loop and a \code{cnt} variable. Even though the type system is
-% unable to reason about the while loop, as long as the runtime system
-% updates the bounds in-place, the user can apply a
-% \code{dyn_bounds_cast} to soundly recover the more precise bounds
-% information.
-% }
-
-% \inferrule*[lab=T-Amper]
-% {
-%   \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
-%   D(T) = ...; \tau_f~f; ...
-% }
-% {
-%   \Gamma \proves \eamper{e}{f} : \tptr[m]{\tau_f}
-% }

### 2023-01-08 sandp23/appendix.tex

@@
 \section{Appendix}\label{app:main}
+
+\subsection{Meta Theories}\label{sec:theorem}
+
+% Before we present our main theorems, we need to first
+% discuss the meaning what a pointer being well-typed in a given heap
+% snapshot $\heap$ means, which is captured by rules in
+% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
+% simply checks if a given variable has the defined type in $\Gamma$;
+% the constant rule ($\textsc{T-Const}$) is slightly more involved.
+% First, it ensures that the type annotation $\tau$ does not contain any
+% free variables. More importantly, it ensures that the pointer points
+% to a location that makes sense in a given heap.
+%  
+%  
+%  The $\size$ function in Fig.~\ref{fig:const-type}
+% refers to the \code{sizeof} function in C computing the number of
+% bytes for a type.
+%  
+%  
+%  Second, we
+% require that any constant ($\evalue{n}{\tau}$) should make sense in
+% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
+% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
+% constant set containing the constants that have been verified by the
+% relation. For every constant $\evalue{n}{\tau}$, it is either an
+% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
+% zero-valued number ($n=0$), checked in $\sigma$
+% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
+% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
+% location $n+i$ in the range of the pointer (if $\omega$ is a word
+% type, range is $[0,1)$; if $\omega$ is an array type
+%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
+%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
+%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
+%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
+%       \vdash n_a : \tau_a$.
+%  
+%  
+% \begin{figure}[t]
+% {\small
+% \text{Type Rules for Constants and Variables:}
+% \begin{mathpar}
+%   \inferrule[T-Var]
+%       {x : \tau \in \Gamma}
+%       {\Gamma;\Theta \vdash_m x : \tau}
+%  
+%   \inferrule[T-Const]
+%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
+%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
+% \end{mathpar}
+%     
+% \text{Rules for Checking Constant Pointers In Heap:}
+% \begin{mathpar}
+%   \inferrule
+%       {}
+%       {\sigma \vdash n : \tint}
+%  
+%   \inferrule
+%       {}
+%       {\sigma \vdash n : \tptr{\omega}{\umode}}
+%  
+%   \inferrule
+%       {}
+%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
+%  
+%   \inferrule
+%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
+%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
+%  
+%   \inferrule
+%       {\forall i \in [0,\size(\omega)) .
+%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
+%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
+% \end{mathpar}
+% }
+% \caption{Type Rules for Checking Constants/Variables}
+% \label{fig:const-type}
+% \end{figure}
+
+% \review{
+%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
+%   something, I didn't see such a definition in the paper.}
+% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
+%   assumption of well typing; we have added more details about that.}
+
+In this subsection, we focus on our main meta-theoretic results about
+\lang: type soundness (progress and preservation),
+non-exposure, and non-crashing.
+These proofs have been conducted in our Coq model.
+Type soundness relies on several \emph{well-formedness} given in \ref{li22checkedc} and \Cref{appx:add-type-sem}.
+Progress states that a \lang program can always make a move:
+
+\begin{thm}[Progress]\label{thm:progress}
+
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed, consistent
+($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
+one of the following holds:
+
+\begin{itemize}
+
+\item $e$ is a value ($\evalue{n}{\tau}$).
+
+\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.
+
+\end{itemize}
+\end{thm}
+%{\em Proof:} By induction on the typing derivation.
+
+\noindent
+There are two forms of preservation regarding the checked and unchecked regions.
+Checked Preservation states that a reduction step preserves both the
+type and consistency of the program being reduced.
+Unchecked Preservation states that any evaluation happens at unchecked region does not affect the checked heap.
+
+\begin{thm}[Checked Preservation]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed, consistent
+($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
+checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
+$\Gamma'$ and $\tau'$ that are well formed, checked region consistent
+($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
+well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
+$\tau'\sqsubseteq_{\Theta} \tau$.
+\end{thm}
+%{\em Proof:} By induction on the typing derivation.
+%\smallskip
+\begin{thm}[Unchecked Preservation]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
+\end{thm}
+
+Using the above theorem, we first show the non-exposure theorem,
+where code in unchecked region cannot observe a valid checked pointer address.
+
+\begin{thm}[Non-Exposure]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
+where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
+thus, it is not a checked pointer.
+\end{thm}
+
+We now state our main result, {\em non-crashing},
+which suggests that a well-typed program can never be \emph{stuck} (expression
+$e$ is a non-value that cannot take a step\footnote{Note that
+  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
+  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
+  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
+  safety.}).
+
+% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r.
+Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
+% \liyi{It is a typo.
+We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region.
+This is a bit due to the space limitation.
+The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement.
+}
+
+\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
+  program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are well-formed and consistent
+($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
+if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
+\tau$ for some $\tau$) and there exists
+$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
+$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
+\end{thm}
+
+% \review{ 
+% My interpretation of section IV.C is that the authors have a pen-and-paper
+%   proof of Theorem 4, and that random testing using the Redex models has been
+%   used to gain confidence in such a proof. Is this correct?}
+% \liyi{Yes, this is correct.}
+% \mwh{No, not correct: We have no pen-and-paper proof, just the
+%   model. Clarify here}
+% showing that
+% extensive random testing fails to falsify the bisimilarity
+% property. \mwh{Better way to state the previous?}
+
+In the \systemname compiler, we also restore the compiler simulation theorem, such that executing a \systemname program is simulated by executing the compiled LLVM program.
+The details are given in the \checkedc formalism \cite{li22checkedc} and \Cref{appx:comp1}.
+
+We formalize both the compilation procedure and the simulation
+theorem in the semantic model we developed in Redex (same as our Coq model),
+and then attempt to falsify it via Redex's support for random
+testing. Redex allows us
+  to specify compilation as logical rules (an extension
+  of typing), but then execute it algorithmically to
+  automatically test whether simulation holds. This process revealed
+  several bugs in compilation and the theorem statement.
+%
+  % us gain confidence that our original pen and paper proof of
+  % simulation remains true with the addition of variable bounds. }
+We ultimately plan to prove simulation in the Coq model.
+
+%Turning to the simulation theorem: We first introduce notation
+%used to specify the theorem.
+We use the notation $\gg$ to
+indicate the \emph{erasure} of stack and heap---the rhs is the same as
+the lhs but with type annotations removed:
+\begin{equation*}
+  \begin{split}
+    \heap  \gg & \dot \heap \\
+    \varphi \gg & \dot \varphi
+  \end{split}
+\end{equation*}
+In addition, when $\Gamma;\emptyset\vdash
+\varphi$ and $\varphi$ is well-typed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
+\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
+and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively. 
+%$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
+%Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
+%reduction relation of $\elang$. Unlike the $\lang$, the semantics of
+%$\elang$ does not distinguish checked and unchecked regions.
+We show an overview of the simulation theorem in \Cref{fig:checkedc-simulation-ref}.
+The simulation theorem is specified in a way that is similar to the one by~\citet{merigoux2021catala}.
+
+Previous ordinary simulation properties would
+replace the middle and bottom parts of the figure with the
+following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
+  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
+Instead, we relate two erased configurations using the relation $\sim$,
+which only requires that the two configurations will eventually reduce
+to the same state.
+
+% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
+\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
+and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
+
+\begin{itemize}
+
+\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\varphi,\dot \heap,\dot e)$ and $(\dot
+\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+\dot \heap,\dot e)$.
+
+\item if $r_1 = \ebounds$ or $\enull$, then $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+\varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+\end{itemize}
+\end{thm}
+
+% when $r_1 = e_1$ for
+% some $e_1$ and
+% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
+% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
+% some $e_1$ and
+% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
+% , then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+% \end{thm}
+
+The current version of the Redex model has been tested against $27500$
+test cases with depth less than $15$. Each expression can
+reduce multiple steps, and we test simulation between every two
+adjacent steps to cover a wider range of programs, particularly the
+ones that have a non-empty heap.
+
+\begin{figure}[t]
+{\small
+\[
+\begin{array}{c}
+\begin{tikzpicture}[
+            > = stealth, % arrow head style
+            shorten > = 1pt, % don't touch arrow head to node
+            auto,
+            node distance = 3cm
+        ]
+
+\begin{scope}[every node/.style={draw}]
+    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
+    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
+    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
+    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
+    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
+\end{scope}
+\begin{scope}[every edge/.style={draw=black}]
+
+    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
+    \path [<->] (A) edge node {$\gg$} (C);
+    \path [<->] (B) edge node {$\gg$} (D);
+    \path [dashed,<->] (C) edge node {$\sim$} (D);
+    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
+    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
+\end{scope}
+
+\end{tikzpicture}
+\end{array}
+\]
+}
+\caption{Simulation between \lang and \elang }
+\label{fig:checkedc-simulation-ref}
+\end{figure}
+
+%{\em Proof:} By induction on the number of steps of the \checkedc
+%evaluation ($\longrightarrow_m^*$), using progress and preservation to
+%maintain the invariance of the assumptions.
+
+% \review{
+%   add a paragraph that discusses what are the main changes from [21] in terms
+%   of the technical development (if there are any), e.g., are there any new
+%   challenges that needed to be solved while proving the blame theorem for this
+%   paper's semantics?
+% }
+%  Compared to \citet{ruef18checkedc-incr}, proofs for
+%  \lang were made challenging by the addition of dependently typed
+%  functions and dynamic arrays, and the need to handle bounds widening for NT
+%  array pointers. These features required changes in the runtime
+%%  semantics (adding a stack, and dynamically changing bounds) and in
+ % compile-time knowledge of them (to soundly typing widened bounds).
+
+% \ignore{
+% \begin{figure}[t!]
+%   \begin{prooftree}
+%     \hypo{\evalue{n}{\tau} \in \defscope}
+%     \infer1[T-VConst]{\Gamma;\defscope \vdash_m \evalue{n}{\tau}  : \tau}
+%   \end{prooftree}
+%   \qquad
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma;\defscope \vdash_m e : \tau \\
+%         \Gamma;\defscope \vdash_m e_1 : \tau_1 
+%       \end{matrix}
+%     }
+%     \hypo{
+%       \begin{matrix}
+% \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2 \\
+%             \Gamma;\defscope \vdash_m e_2 : \tau_2
+%           \end{matrix}
+%         }
+%     \infer2[T-If]{\Gamma;\defscope \vdash_m \eif{e}{e_1}{e_2} : \tau_3}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma;\defscope \vdash_m x : \tptr[c]{(\tntarray{l}{0}{\tau})}, l \leq 0 \\
+%         \Gamma, x : \tptr[c]{(\tntarray{l}{1}{\tau})};\defscope \vdash_m e_1 : \tau_1 \\
+%         \Gamma;\defscope \vdash_m e_2 : \tau_2 \\
+%         \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2
+%       \end{matrix}
+%     }
+%     \infer1[T-IfNT]{\Gamma;\defscope \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_3}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         F(f) = \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e  \\
+%         \Gamma; \defscope \vdash_m \tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}] ~~~ 0 \leq i \leq j\\
+%         \Gamma; \defscope \vdash_m e_i : \tau_i' ~~~ 0 \leq i \leq j\\
+%         \Gamma; \defscope \vdash_m \subtype{\tau_i'}{\tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}]}  ~~~ 0 \leq i < j
+%       \end{matrix}
+%     }
+%     \infer1[T-VCall]{\Gamma; \defscope \vdash_m f(\overline{e}) : \tau_j[x_0,\ldots,x_{j-1} \mapsto e_0,\ldots,e_{j-1}]}
+%   \end{prooftree} \\ \\
+% % \inferrule*[lab=T-PtrC]
+% % {
+% %   \tau = \tptr[c]{\omega} \\
+% %   \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\\\
+% %   \wt[\Gamma][\defscope, n^\tau]{H(n+k)}{\tau_k} ~~~ 0 \leq k < j
+% % %  \Gamma,n^\tau \proves H(n+k) : \tau_k~~~0 \leq k < j
+% % }
+
+%   \begin{prooftree}
+%     \hypo{
+%           \Gamma; \defscope \vdash_m e_1 : \tau_1
+%       }
+%       \hypo{        \Gamma, x = e_1 : \tau_1; \defscope \vdash_m e_2 : \tau_2
+% }
+%     \infer2[T-Let]{\Gamma; \defscope \vdash_m \elet{x}{e_1}{e_2} : \tau[x \mapsto e_1]}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma; \defscope \vdash_m \estrlen{y} : \tau_1\\
+%         \Gamma; \defscope \vdash_m y : \tptr[c]{(\tntarray{le}{\_}{\tau_3})}\\
+%         \Gamma, x = \estrlen{y} : \tau_1, \\ y : \tptr[c]{(\tntarray{le}{x}{\tau_3})} ;  \defscope  \vdash_m e_2 : \tau_2
+%         \end{matrix}
+%         }
+%         \infer1[T-LetStr]{\Gamma; \defscope \vdash_m \elet{x}{\estrlen{y}}{e_2} : \tau[x \mapsto e_1]}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \hypo{
+%         \Gamma; \defscope \vdash_m e : \tptr[c]{(\tntarray{le}{he}{\tau})}
+%       }
+%         \infer1[T-Str]{\Gamma; \defscope \vdash_m \estrlen{y} : \tint}
+%       \end{prooftree} \\ \\
+
+%       \begin{prooftree}
+%         \hypo{
+%           \begin{matrix}
+%             \tau = \tint \vee \tau = \tptr[u]{\omega}~\vee  n=0~ \vee \\ \tau = \tptr[c]{(\tarray{0}{0}{\tau'})} \vee \\
+%             \tau = \tptr[c]{(\tntarray{0}{0}{\tau'})}
+%         \end{matrix}
+%       }
+%       \infer1[T-Base]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
+%     \end{prooftree} \\ \\
+
+%     \begin{prooftree}
+%       \hypo{
+%         \begin{matrix}
+%         \tau = \tptr[c]{\omega} \\
+%         \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\
+%         \Gamma;\defscope, \evalue{n}{\tau}  \vdash_m {H(n+k)} : {\tau_k} ~~~ 0 \leq k < j
+%       \end{matrix}
+%       }
+%       \infer1[T-PtrC]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
+%     \end{prooftree} \\ \\
+
+%     \begin{prooftree}
+%       \hypo{
+%         \begin{matrix}
+%             \Gamma; \defscope \vdash_m e :  {\tptr[m]{\tstruct{T}}} \\
+%             % \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
+%             D(T) = ...; \tau_f~f; ...
+%         \end{matrix}
+%       }
+%       \infer1[T-Amper]{\Gamma; \defscope \vdash_m \eamper{e}{f} : \tptr[m]{\tau_f}}
+%     \end{prooftree} \\ \\
+
+%     \todo[inline]{YL: how to express le - n as a metafunction?}
+%     \begin{prooftree}
+%       \hypo{
+%         \begin{matrix}
+%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}\\
+%         \Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tint \\
+%         le' = le - n, he' = he - n
+%       \end{matrix}
+%     }
+%       \infer1[T-BinopInd]{\Gamma; \defscope \vdash_m (e_1 \plus \evalue{n}{\tau} ) : \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})}}
+%     \end{prooftree}
+% \caption{Typing}
+% \label{fig:typing}
+% \end{figure}
+
+% \begin{figure}[t!]
+%   \todo[inline]{no sizeof check because we don't know statically whether the allocation would be null}
+%   \begin{prooftree}
+%     \hypo{
+%       \omega = \tgarray{\alpha}{le}{he}{\tau} \Rightarrow le = 0
+%     }
+%     \infer1[T-Malloc]{\Gamma; \defscope \vdash_m \emalloc{\omega} : \tptr[c]{\omega}}
+%   \end{prooftree}\\\\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \Gamma; \defscope \vdash_u e : \tau
+%     }
+%     \infer1[T-Unchecked]{\Gamma; \defscope \vdash_m \eunchecked{e} : \tau}
+%   \end{prooftree}\\\\
+
+%   \todo[inline]{any constraints on m'?}
+%   \begin{prooftree}
+%     \hypo{\Gamma; \defscope \vdash_m e :  \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})} }
+%     \infer1[T-DynCast]{\Gamma; \defscope \vdash_m \edyncast{\tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}}{e} : \tau}
+%   \end{prooftree}\\\\
+
+%   \begin{prooftree}
+%     \hypo{\Gamma; \defscope \vdash_m e : \tau'}
+%     \infer1[T-Cast]{\Gamma; \defscope \vdash_m \ecast{\tau}{e} : \tau}
+%   \end{prooftree}\\\\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\tgarray{\alpha}{le}{he}{\tau}} \\
+%         \Gamma; \defscope \vdash_m e_2 : \tint \\
+%         m' = u \Rightarrow m = u \\
+%       \end{matrix}
+%     }
+%     \infer1[T-Index]{\Gamma; \defscope \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
+%         \Gamma; \defscope \vdash_m e_2 : \tau \\
+%         m' = u \Rightarrow m = u \\
+%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
+%       \end{matrix}
+%     }
+%     \infer1[T-Assign]{\Gamma; \defscope \vdash_m \eassign{e_1}{e_2} : \tau}
+%   \end{prooftree}\\\\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
+%         \Gamma; \defscope \vdash_m e_2 : \tint \\
+%         \Gamma; \defscope \vdash_m e_3 : \tau \\
+%         m' = u \Rightarrow m = u \\
+%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
+%       \end{matrix}
+%     }
+%     \infer1[T-IndAssign]{\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
+%   \end{prooftree}
+
+% \caption{Typing Cont.}
+% \label{fig:typing2}
+% \end{figure}
+% \ignore{
+
+% \begin{figure}[t!]
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j} \\
+%         \Gamma, x_0 = \kw{none} : \tau_0, \ldots, x_{j-1} = \kw{none} : \tau_{j-1}; \defscope \vdash e : \tau_{r}
+%       \end{matrix}
+%     }
+%     \infer1[WF-Fun]{\Gamma;\defscope \vdash \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \infer0[WF-Nil]{\Gamma \vdash \circ}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \hypo{
+%       \begin{matrix}
+%         \Gamma \vdash \tau_0 \\
+%         \Gamma, x_0 = \kw{none} :\tau_0 \vdash x_1:\tau_{1}, \ldots, x_{j}:\tau_{j}
+%       \end{matrix}
+%     }
+%     \infer1[WF-Cons]{\Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j}}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \infer0[WF-Int]{\Gamma \vdash \tint}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \hypo{   \Gamma \vdash le }
+%      \hypo{   \Gamma \vdash he }
+%       \hypo{  \Gamma \vdash \tau}
+
+%     \infer3[WF-Array]{\Gamma \vdash \tptr[m]{(\tgarray{\alpha}{le}{he}{\tau})}}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \hypo{T \in dom(D)}
+%     \infer1[WF-Struct]{\Gamma \vdash \tptr[m]{\tstruct{T}}}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \hypo{\Gamma \vdash \tau}
+%     \infer1[WF-Ptr]{\Gamma \vdash \tptr[m]{\tau}}
+%   \end{prooftree} \\\\
+
+%   \begin{prooftree}
+%     \infer0[WFB-Int]{\Gamma \vdash i}
+%   \end{prooftree}\\\\
+
+%   \begin{prooftree}
+%     \hypo{x = e? : \tint \in \Gamma}
+%     \infer1[WFB-Var]{\Gamma \vdash x \plus i}
+%   \end{prooftree} \\\\
+
+%   \caption{Well-Formedness}
+%   \label{fig:wf}
+% \end{figure}
+
+% \begin{figure}[t!]
+%   \begin{prooftree}
+%     \infer0[Sub-Nt]{\subtype{\tptr[c]{(\tntarray{le}{he}{\tau})}}{\tptr[c]{(\tarray{le}{he}{\tau})}}}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \infer0[Sub-Refl]{\subtype{\tau}{\tau}}
+%   \end{prooftree} \\ \\
+
+%   \todo[inline]{what about ntarrays?}
+%   \begin{prooftree}
+%     \infer0[Sub-Ptr]{\subtype{\tptr[c]{\tau}}{\tptr{\tarray{0}{1}{\tau}}}}
+%   \end{prooftree} \\ \\
+
+%   \todo[inline]{the subtyping relation is not anti-symmetric..}
+%   \begin{prooftree}
+%     \infer0[Sub-Arr]{\subtype{\tptr[c]{\tarray{0}{1}{\tau}}}{\tptr{\tau}}}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     \hypo{\rboundle{le_0}{le_1}}
+%     \hypo{\rboundle{he_1}{he_0}}
+%     \infer2[Sub-Subsume]{\subtype{\tptr{(\tgarray{\alpha}{le}{he}{\tau})}{\cmode}}{ \tptr{(\tgarray{\alpha}{ le_1} {he_1} {\tau})}}}
+%   \end{prooftree} \\ \\
+
+%   \begin{prooftree}
+%     % check t-amper below
+%     \hypo{D(T) = \tau_f~f; ...}
+%     \hypo{\subtype{\tau_f}{\tau}}
+%     \infer2{\subtype{\tptr[c]{\tstruct{T}}}{\tau}}
+%   \end{prooftree}
+%   \caption{Subtyping}
+%   \label{fig:sub}
+% \end{figure}
+% }
+% }
+
+% \ignore{
+
+% \begin{figure*}[t]
+%   \begin{lstlisting}
+% int foo(nt_array_ptr<char> p : count(0)) {
+%   if (* p) {
+%     dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
+%   }
+%   dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
+%   return 0;
+% }
+%   \end{lstlisting}
+% \caption{The example where clang Checked C fails at run-time}
+% \label{fig:clangbad1}
+% \end{figure*}
+
+% \begin{figure*}[t]
+%   \begin{lstlisting}
+% /* nt_array_ptr<char> p : bounds(p,p) */
+% size_t cnt = 0;
+
+% while(*(p+cnt)) {
+%   ++cnt;
+% }
+% dyn_bounds_cast<nt_array_ptr<char>>(p, count(cnt));
+%   \end{lstlisting}
+% \caption{A useful program that the Checked C spec doesn't allow at run-time}
+% \label{fig:clangbad2}
+% \end{figure*}
+% }
+
+% \ignore{
+% The Clang CheckedC implementation uses statically determined
+% bounds to insert run-time checks. In Fig.~\ref{fig:clangbad1}, the
+% \code{dyn_bounds_cast} at line 3 will always succeed, because the
+% compiler knows that within the scope of then branch, the pointer
+% \code{p} must have at least one element. The same cast at line 5,
+% however, will always fail, since there is no way to tell statically
+% whether the program has entered the then branch before. The compiler
+% will check whether the \code{count(1)} bounds specification is
+% contained within the earlier \code{count(0)} specification, resulting
+% in a run-time failure even when we pass in a non-empty string.
+
+% Our formalization diverges from this run-time behavior and instead keeps
+% track of the bounds on the stack. After entering the then branch, we
+% increment the upper bound for \code{p}, effectively making the
+% updated bounds information available even after we exit the if
+% statement. The cast at line 5 will be checking the new bounds against
+% the incremented bounds for non-empty strings.
+
+% Fig.~\ref{fig:clangbad2} gives a more practical example of why keeping
+% track of the bounds on the stack is useful. The program snippet
+% implements the functionality of the \code{strlen} function using a
+% while loop and a \code{cnt} variable. Even though the type system is
+% unable to reason about the while loop, as long as the runtime system
+% updates the bounds in-place, the user can apply a
+% \code{dyn_bounds_cast} to soundly recover the more precise bounds
+% information.
+% }
+
+% \inferrule*[lab=T-Amper]
+% {
+%   \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
+%   D(T) = ...; \tau_f~f; ...
+% }
+% {
+%   \Gamma \proves \eamper{e}{f} : \tptr[m]{\tau_f}
+% }
 

### 2023-01-08 sandp23/evaluation.tex

@@
 \myparagraph{Benchmarking Methodology}
-We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
-Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted\-ness of a pointer is propagated from the call site up until its declaration.
+We use C's~\code{<time.h>} library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
+Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
 We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.

### 2023-01-08 sandp23/introduction.tex

@@
 Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
-The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.
-\liyi{really? I think they can exchange data in the unchecked region, and then the sandbox guarantee the correctness? }
+The checked types follow the standard Checked C typing rules.
+However, our type-system disallows any unsafe interactions between taint and untainted types.
 
-The developer starts by marking desired (i.e., unchecked) functions and pointers used in these functions as tainted.
+The developer starts by marking desired (\ie unchecked) functions and pointers used in these functions as tainted.
 Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
@@
 \myparagraph{Supporting callbacks to~\cregion with no Checked Pointer Exposure}
-Although we disallow access to~\cregion from~\ucregion, there can be cases where such access is needed. Specifically, when~\cregion wants to provide access to certain shared checked data to~\ucregion.
+Although we disallow access to~\cregion from~\ucregion, there can be cases where such access is needed.
+Specifically, when~\cregion wants to provide access to certain shared checked data to~\ucregion.
 To enable this, we support callback functions in~\cregion that can be invoked from~\ucregion through function pointers.
-However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of \systemname.
+%However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of \systemname.
+However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of~\systemname and leads to other attacks~\cite{hauser2019sleak}.
+
+We handle this by using indirection.
+Specifically, instead of directly accessing the~\cregion callbacks, the ~\ucregion accesses them using a protected trampoline function, which will direct the execution to the appropriate callback function.
+In addition, the trampoline function itself will be referenced using an opaque index rather than its virtual address, and we will utilize existing sandboxing techniques to implement this.
+
+\ignore{
 
@@
 \mz{What does this mean?}
+}
 %

### 2023-01-08 sandp23/overview.tex

@@
 
-For our example in~\lst{lst:final}, we can make only the function~\code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
+For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
 

### 2023-01-08 sandp23/evaluation.tex
update.

@@
 Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks.
-MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
-We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.
+Micro-benchmarking involves evaluating performance on operations containing tainted pointers, context switching between checked and sandboxed regions, and sandboxed function execution.
+We evaluate \systemname using six real-world programs 
+pertaining to diversified domains for evaluating real-world run-time and memory performance.
 
 \myparagraph{Benchmarking Methodology}
-We use C's~\code{<time.h>} library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
+We use C's~\code{<time.h>} library to evaluate the runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
 Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
 We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
-Unlike the Runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
+Unlike the runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
 This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLib functions that are in use by the tainted pointers in the program.

### 2023-01-08 sandp23/introduction.tex
update.

@@
 These approaches all impose performance overheads deemed too high for deployment use. 
+\liyi{Should we say this? Is the paper about overhead reducing? }
 Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
 open-source extension to C with new types and
-annotations whose use can ensure a program’s spatial safety.
-Specifically,~\checkedc extends C with checked pointer types, which are restricted by the compiler to spatially safe uses.
-\liyi{why we need to mention C to \checkedc conversion?
-Is this paper about introducing CheckedCBox or about the conversion from C to CheckedCBox.
-I am actually against putting CheckedCBox conversion stuff in the paper, because it is kind of confusion.
-We have the compilation from CheckedCBox to LLVM and have the conversion from C to CheckedCBox.
-}
+annotations to ensure a program’s spatial safety.
+Specifically, \checkedc extends C with checked pointer types, which are restricted by the compiler to spatially safe uses.
 The checked pointers are represented as system-level memory words,~\ie without ``fattening'' metadata, ensuring backward compatibility.
@@
 We need to ensure that~\emph{code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
-\liyi{Can we strengthen the sentence a little by saying something fancy? }
-
-\ignore{
-Converting these existing programs into safe languages, such as Java, requires considerable effort and is not (yet) feasible for performance-critical environments such as Embedded Operating Systems.
-Although high-performant safe system languages such as Rust~\cite{rustlang} and Go exist, these a too different to constitute a practical target~\cite{zeng2019identifying}, and rewriting in these languages requires considerable effort.
-
-Existing automated conversion tools~\cite{c2rust,c2rusttalk, emre2021c2rust} are able to take only small steps (i.e., to unsafe, non-idiomatic Rust) and still~\emph{require considerable work from developers}.
-There are other retrofitting techniques such as CCured~\cite{necula2005ccured},
-Softbound~\cite{nagarakatte2009softbound}, Low Fat
-pointers~\cite{duck2016heap}, and Address Sanitizer
-(ASAN)~\cite{serebryany2012addresssanitizer} aim to enforce spatial safety
-automatically~\emph{without any developer effort}, by analyzing a C program and compiling it to include
-run-time safety checks. Unfortunately, the resulting~\emph{run-time overhead is too high
-for deployment (between 60\%-200\%)}, especially in low-powered IoT devices where the usage of C-based system software (e.g., FreeRTOS) is still rampant.
-
-\input{figures/existingmem}
-
-\noindent
-\textbf{Checked C.} Tarditi et al. developed~\emph{Checked C}, a safe dialect of C that prevents spatial memory issues (temporal safety is underway~\cite{checkedc:temporal}) with no overhead.
-Checked C extends C with \emph{checked pointer
-  types} which are restricted by the compiler to spatially safe uses (temporal safety is underway~\cite{checkedc:temporal}).
-Such pointers have one of three possible types, \ptrT{T}, \arrT{T},
-or~\ntarrT{T} (\ptr, \arr, and \ntarr{} for short), representing a
-pointer to a single element, array of elements, or null-terminated
-array of elements of type $T$, respectively. The latter two have an
-associated \emph{bounds annotation}; e.g., a declaration \arrT{int}
-\code{p} : \code{count(n)} says that \code{p} is a pointer to an \code{int} array whose size is \code{n}.
-The compiler uses these bounds annotations to add dynamic checks prior to checked pointer accesses, to prevent spatial safety violations.
-These run-time checks can often be proved redundant and removed by LLVM, yielding good performance.
-Specifically, Tarditi et al.~\cite{tarditi2018checked} reported average
-run-time overheads of 8.6\% on a small benchmark suite, and
-Duan et al.~\cite{duanrefactoring} found essentially~\emph{no overhead} when running
-Checked C-converted portions of the FreeBSD kernel.
-Furthermore, Checked C is backward compatible as the compiler  represents checked pointers as system-level memory words, i.e., without
-``fattening'' metadata, ensuring backward compatibility. 
-The backward compatibility also enables incremental or partial conversion where Checked annotations can be provided to a few pointers.
-Such partially annotated programs still enjoy the benefits of spatial memory safety on those regions using only Checked pointers (i.e., checked or safe code regions).
-~\fig{fig:memprotections} shows the summary of these existing approaches along with Checked C.
-We argue that~\emph{converting to Checked C provides a reasonable and practical approach to adding safety to existing system software written in C}.
-}
 
@@
 \noindent
-\textbf{No safety against unchecked code.} The spatial safety guarantee~\checkedc provides applies to completely converted programs, i.e., programs that uses only checked types.
+\textbf{No safety against unchecked code.} The \checkedc spatial safety guarantee applies to completely converted programs, i.e., programs that uses only checked types.
 Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
@@
 The developer starts by marking desired (\ie unchecked) functions and pointers used in these functions as tainted.
-Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
-\liyi{can we just name them unchecked and checked regions, and then referring to them as \ucregion and~\cregion? We need to make it clear that we are re-define c and uc regions. }
+Second,~\systemname partitions the given program into two partitions (\umode and~\cmode regions)  of different privileges:
+
 \begin{itemize}
-\item~\ucregion or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (\ie functions and pointers) and can only access tainted pointers.
-\item~\cregion or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
+\item~\umode \emph{region} or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (\ie functions and pointers) and can only access tainted pointers.
+\item~\cmode \emph{region} or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cmode region.
 The functions in~\cregion can invoke any function in~\ucregion and access all its data but not the other way around, except for call-back functions, which we will discuss later.
@@
 The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
-As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
+As functions in the~\ucregion can only access tainted types (or unchecked types),~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
 We avoid marshaling by allocating all tainted pointers (\ie tainted buffers) in~\ucregion and thus can be accessed in both partitions.
 Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
-Our compiler avoids these attacks by ensuring, either statically or through dynamic checks that tainted pointers can only point to~\ucregion address space.
+Our compiler avoids these attacks by ensuring, through dynamic checks that tainted pointers validly point to \ucregion address space.
+Such checks are statically generated by our compiler.
 
@@
 In addition, the trampoline function itself will be referenced using an opaque index rather than its virtual address, and we will utilize existing sandboxing techniques to implement this.
+In our \systemname formalism, we formally verified that our formalism is \emph{non-exposure}, \ie, no \cregion pointer addresses will be leaked in \ucregion code.
 

### 2023-01-08 sandp23/background.tex
update.

@@
 \subsection{\checkedc}
+\label{subsec:checkedc}
 % 
@@
 
+\liyi{Do we need to spend more than half page introducing checked-c? OR just introduce it briefly like what we had in the formalism paper? }
+
 \noindent
@@
 \begin{itemize}
-\item \code{ptr<}$T$\code{>} ($\ptr$) types a pointer that is either null or
+\item \code{_Ptr<}$T$\code{>} ($\ptr$) types a pointer that is either null or
   points to a single object of type $T$.
-\item \code{array_ptr<}$T$\code{>} ($\arr$) types a pointer that is either null
+\item \code{_Array_ptr<}$T$\code{>} ($\arr$) types a pointer that is either null
   or points to an array of $T$ objects. The array width is defined
   by a \emph{bounds} expression, discussed below.
-\item \code{nt_array_ptr<}$T$\code{>} ($\ntarr$) is like
-  \code{array_ptr<}$T$\code{>} except that the bounds expression
+\item \code{_NT_Array_ptr<}$T$\code{>} ($\ntarr$) is like
+  \code{_Array_ptr<}$T$\code{>} except that the bounds expression
   defines the \emph{minimum} array width---additional objects may
@@
 For instance, in the code \inlinecode{if (n>0) a[n-1] =} $...$ the write
-is via address $\alpha = \inlinecode{a + sizeof(int)}\!\times\!\code{(n-1)}$. If the
-bounds of \code{a} are \code{count(u)}, the
+is via address $\alpha = \inlinecode{a + sizeof(int)}\!\times\!\code{(n-1)}$. 
+If the bounds of \code{a} are \code{count(u)}, the
 inserted check will confirm that prior to dereference
@@
 
+\liyi{why we need the following? I suggest we can cut. }
 Oftentimes, inserted checks can be optimized away by LLVM.
@@
 ...
-array_ptr<int> p : count (n) = NULL;
+_Array_ptr<int> p : count (n) = NULL;
 |\textcolor{red}{\faTimes}|..,&n,.
@@
 
-\liyi{I do not think you need three sub-sections for these topics. They can be in the same place. }
 \subsection{Program Partitioning}

### 2023-01-08 sandp23/overview.tex
update.

@@
 
-The~\Cref{fig:overview} shows the interaction between various components of~\systemname{}.
-Given an appropriately annotated source program and a sandbox configuration, our system creates an executable such that all tainted functions and data will reside in a sandbox and the~\cregion (non-sandboxed) region is not affected by the code in the sandbox.
-
+\Cref{fig:overview} shows the interaction between various components of~\systemname{}.
+Given an appropriately annotated source program and a sandbox configuration, our system creates an executable such that all tainted functions and data will reside in a sandbox and the~\cregion (non-sandboxed) is not affected by the code in the sandbox.
 We will present an overview of the~\systemname{} from a developer perspective and the details of individual components in the later sections.
@@
 
-\liyi{The example looks good. But we are focusing on converting C to \checkedc here? }
-
 \myparagraph{Goal} 
@@
 To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
-However, as explained in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
-
-\liyi{Can we just assume that programmers know \checkedc and will use \checkedc, and then we added tainted type? }
+However, as shown in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
 
@@
 These sets of source files will be compiled with the corresponding compilers to get the corresponding object files.
-The c-region object file will have the necessary runtime checks enforcing~\systemname{} guarantees.
+The \cregion object file will have the necessary runtime checks enforcing~\systemname{} guarantees.
 The~\ucregion object file will be produced according to the corresponding sandbox mechanism.
@@
 
+\liyi{This paper should just introduce the language and the compiler, right?
+I do not think the following is necessary.
+Instead, we can have an overview section introduce the compiler with examples; especially highlight the callback.
+}
 \subsection{Alternative Use Cases}

### 2023-01-08 sandp23/introduction.tex
[review] comments/revision on the first paragraph

@@
 Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
-have investigated ways to compile C programs 
-to automatic spatial safety enforcement.
+have investigated ways to
+\mzu{compile C programs 
+  to automatic spatial safety enforcement.}
+\mz{You cannot compile a program to enforcement.}
+% 
 These approaches all impose performance overheads deemed too high for deployment use. 
+% 
 \liyi{Should we say this? Is the paper about overhead reducing? }
 Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
-open-source extension to C with new types and
-annotations to ensure a program’s spatial safety.
-Specifically, \checkedc extends C with checked pointer types, which are restricted by the compiler to spatially safe uses.
-The checked pointers are represented as system-level memory words,~\ie without ``fattening'' metadata, ensuring backward compatibility.
-This enables incremental conversion, where regions of code can be converted to~\checkedc and ensure spatial safety within those regions.
-However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the unconverted or unchecked code can violate guarantees provided in Checked regions.
+open-source extension to C
+% 
+\mzs{ with new types and annotations to ensure a program’s spatial safety.
+  Specifically, \checkedc extends C with checked pointer types, which are
+  restricted by the compiler to spatially safe uses.}
+% 
+\mzr{ with additional annotations and types, checked pointer types in
+  particular, to ensure a program’s spatial safety.}
+% 
+\mz{Redundancy?}
+% 
+The checked pointers are represented as system-level memory words\mzs{,~\ie} without ``fattening'' metadata, ensuring backward compatibility.
+\mz{\ie{} sounds indirect and redundant to me.}
+% 
+\mzu{This enables incremental conversion, where regions of code can be converted to~\checkedc and ensure spatial safety within those regions.}
+\mz{This sentence is incoherent. I don't understand the connection between
+  incremental conversion and regions of code. 
+  % 
+  I guess your point is that since you don't introduce extra metadata, 
+  checked and unchecked region can coexist and interact seamlessly.
+  % 
+}
+% 
+However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
+unconverted or unchecked code can violate guarantees provided in \mzu{Checked}
+\mz{either tt it or use lowercase?}
+regions.
 %

### 2023-01-09 sandp23/introduction.tex
[review] quick review on intro

@@
 % 
-\mzr{ with additional annotations and types, checked pointer types in
-  particular, to ensure a program’s spatial safety.}
+\mzr{to ensure a program’s spatial safety by introducing additional annotations and
+  enforcing new types, checked pointer types in particular.}
 % 
@@
 \fi
+% 
 This mechanism is called program partitioning~\cite{rul2009towards}.
@@
 However, our type-system disallows any unsafe interactions between taint and untainted types.
+% 
+\mz{It sounds hard to follow. The subjects in these sentences are  ``our
+  system''. ``checked types'' and ``our type-system''.
+  %
+  For readers, the connection among checked, tainted and unchecked seems weak.
+  % 
+} 
+% 
+% Specifically, in our system,~\systemname, extending CheckedC,
+% functions and pointers that need to be isolated from the original program can be
+% marked with ~\textbf{tainted} (\taintt) types.
+%
 
+% 
 The developer starts by marking desired (\ie unchecked) functions and pointers used in these functions as tainted.
-Second,~\systemname partitions the given program into two partitions (\umode and~\cmode regions)  of different privileges:
+Second,~\systemname partitions the given program into two partitions (\umode and~\cmode regions)  of different \mzu{privileges}:
+% 
+\mz{need a bit more description (4-5 words) before introducing the word
+  ``privilege''}
+%
 
@@
 \end{itemize}
+% 
+\mz{                            %
+  This paragraph is unclear. 
+  % 
+  How does regions look like? Can one region nested in another?
+  %
+  If not, ``This partition contains the remaining (untainted) code and data and
+  has complete access to~\cmode region.'' creates a sense of recursive
+  definition of \cregion.
+}
+%
+
+% 
 Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
@@
 The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
-As functions in the~\ucregion can only access tainted types (or unchecked types),~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
-We avoid marshaling by allocating all tainted pointers (\ie tainted buffers) in~\ucregion and thus can be accessed in both partitions.
+% 
+\mzu{As functions in the~\ucregion can only access tainted types (or unchecked
+types),~\cregion functions should use tainted types to pass pointer arguments
+to~\ucregion functions.
+% 
+We avoid marshaling by allocating all tainted pointers (\ie tainted buffers)
+in~\ucregion and thus can be accessed in both partitions.}
+%
+\mz{I don't quite understand the logical consequence here. Why does allocating
+  tainted points avoids marchaling. 
+  % 
+  I see that ``allocating all tainted pointers'' makes ptrs accessable in both
+  regions, but avoiding marchaling doesn't seem to cause that.
+  %
+  The question I should probably ask is what marshaling means? 
+  % 
+  Does the procedure that copies a checked buffer to a tainted one counts as
+  marshaling?
+  % 
+}
+% 
 Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
@@
 To the best of our knowledge,~\systemname is the first C(-like) language and compiler formalism with the program partitioning mechanism.
+%
+\mz{uh}
 
@@
 Specifically, instead of directly accessing the~\cregion callbacks, the ~\ucregion accesses them using a protected trampoline function, which will direct the execution to the appropriate callback function.
-In addition, the trampoline function itself will be referenced using an opaque index rather than its virtual address, and we will utilize existing sandboxing techniques to implement this.
-In our \systemname formalism, we formally verified that our formalism is \emph{non-exposure}, \ie, no \cregion pointer addresses will be leaked in \ucregion code.
+In addition, the trampoline function itself will be referenced using an opaque index rather than its virtual address, and \mzu{we will utilize existing sandboxing techniques to implement this}.\mz{tense?}
+% 
+In our \systemname formalism, we formally verified that our formalism is
+\emph{non-exposure}, \ie, no \cregion pointer addresses will be leaked in
+\ucregion code.
 

### 2023-01-09 sandp23/formal.tex
Adding

@@
 We denote null-terminated arrays using $\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$ otherwise (we elide $\cdot$ when writing types).
-The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
 An example representation of an array and null-terminated array in~\lang syntax is shown below:
@@
 
+\emph{Disallowing unsafe types.}
+The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
+These requirements add additional restrictions for type creation and restrict the creation of unsafe types.
+For instance, we cannot create a tainted array whose elements are checked pointers,~\eg Although syntactically valid, the type~\code{_t_Array_ptr<_Ptr<int>>} is~\emph{not} well-formed and is not allowed in~\lang.
+However, we can have a checked array whose elements are tainted pointers,~\eg~\code{_t_Array_ptr<_t_Ptr<int>>} is a valid type.
+
 Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
@@
 \]
-Note that the well-formed requirement (\fig{fig:wffunctions} in Appendix) of a function type ensures that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+The function type also has well-formed requirements (\fig{fig:wffunctions} in Appendix), which disallows nesting checked or unchecked pointers inside tainted pointers.
+Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
 \aravind{The syntax of well-formed requirements for functions is incorrect. It looks like the forall qualifier is missing.}
@@
 \aravind{Why we have only function pointers in function call?}
+\aravind{Talk about the ret expression after adding it in the syntax.}
 

### 2023-01-09 sandp23/background.tex
update

@@
 The spatial safety guarantee~\checkedc provides applies to completely converted programs,~\ie programs that uses only checked types.
-Although, the backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
+Although,\lc{why use "Although" here?
+seems like this is not an opposite/new idea} the backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
 The unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+\lc{I'm confused, does the partially checked program "enjoy spatial memory safety" or not?}
 Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.

### 2023-01-09 sandp23/introduction.tex
update

@@
 This mechanism is called program partitioning~\cite{rul2009towards}.
-Existing mechanisms are not suitable as they are based on process isolation and have high overhead, and are data-centric~\emph{data-centric} (\sect{subsec:background:programpart}).
+Existing mechanisms are not suitable as they are based on process isolation and have high overhead, and are data-centric~\emph{data-centric} \lc{duplicate?}(\sect{subsec:background:programpart}).
 Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
@@
 } 
+\lc{it's  a little bit hard to follow. it talks about tainted -> checked -> unsafe interactions but doesn't explain how these are related}
 % 
@@
 The developer starts by marking desired (\ie unchecked) functions and pointers used in these functions as tainted.
+\lc{what's "these functions" in "used in these functions as tainted"?
+(and "Second" seems wired here)}
 Second,~\systemname partitions the given program into two partitions (\umode and~\cmode regions)  of different \mzu{privileges}:
@@
   definition of \cregion.
-}
+}\lc{<- same question, maybe giving an abstract example?}
 %

### 2023-01-09 sandp23/formal.tex
Adding type checking

@@
 
+\subsection{Type Checking}
+\label{sec:typechecking}
+Our type checker restricts the usage of tainted and checked pointer types to ensure that tainted pointers do not affect checked types, along with enforcing~\checkedc typing rules~\cite{li22checkedc}.
+A few selected type rules are presented in~\fig{fig:type-system-1}.
+The complete set of typing rules and special handling of (NT)-arrays are provided in~\apdx{sec:literal-pointer-typing}.
+
+Each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+which states that in a type environment $\Gamma$ (mapping variables to
+their types) and a predicate environment $\Theta$ (mapping integer-typed
+variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
+in context mode $m$.
+
 \subsection{Semantics and Typing}\label{sec:semantics}

### 2023-01-09 sandp23/formal.tex
[review] till function types

@@
 system, and compilation.
-The formal model also enables us to develop~\lang’s meta-
+% 
+\mzs{The formal model also enables us to}
+% 
+\mzr{From those, we} develop~\lang’s meta-
 theories, including the type soundness, non-exposure, non-crashing,
@@
 \subsubsection{Types ($\omega$)}
-At a high level, we classify types as word-size values, multi-word values, or functions.
-The word-size values can be either integers or pointers.
-Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
-($\xi$, the difference between context and pointer modes is introduced shortly below) that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
-and a type ($\omega$) denoting valid values that can be pointed to.
-
-Multi-word values represent arrays and null-terminated arrays, which include both the type of elements ($\tau$) and a bound ($\bvar$).
-The bounds can be either integer literals $n$, expressions $x + n$, or an explicit specification of lower and upper bound,~\ie $(b_l,b_h)$.
-We denote null-terminated arrays using $\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$ otherwise (we elide $\cdot$ when writing types).
+At a high level, we classify types as word-size value \mzr{type}s, multi-word
+value \mzr{type}s, or function \mzr{type}s.
+%
+\mz{
+  % 
+  You cannot classify a type as values. You can only classify types into other 3
+  categories, namely, word, multi-word and function types.
+  % 
+}
+% 
+\mzs{The}\mzr{A} word-size value\mzs{s} can be either \mzr{of type integer} or
+\mzr{of type pointer}.
+% 
+\mzr{Each} pointer type\mzs{s} ($\tptr{\omega}{\xi}$) \mzr{is represented by} a
+pointer mode annotation ($\xi$, the difference between context and pointer modes
+is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
+or unchecked (\umode), and a type ($\omega$) denoting \mzr{the valid type of
+  values it can refer to}.
+%
+
+% 
+\mzr{The multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
+arrays and null-terminated arrays is constructed by the type of elements in the
+array ($\tau$), an array bound ($\bvar$) and an array flag ($\kappa$).}
+% 
+\mzr{A} bound can be either \mzr{an} integer literal $n$, \mzr{an} expressions
+$x + n$, or an explicit specification of lower and upper bound\mzr{s},~\ie
+$(b_l,b_h)$.
+%
+\mzr{ %
+  Additionally, to distinguish null-terminated arrays from other arrays, we set
+  \(\kappa\) to be \textit{nt} for nt-arrays and \(\cdot\) for other arrays (for
+  convenience, we will elide \(\cdot\) in the text).
+  % 
+}
+%
+
+% 
 An example representation of an array and null-terminated array in~\lang syntax is shown below:
+% 
 \[\hspace*{-0.5em}
@@
 \]
-To simplify the representation, we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
-$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the above examples could
-be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
+\mzr{For simplicity}\mzs{To simplify the representation}, we write
+$\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean $\tptr{\tarray{0}{b}{\tau}}{\cmode}$,
+so the above examples could be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
 $\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
 
-\emph{Disallowing unsafe types.}
+\emph{\mzr{Rejecting}\mzs{Disallowing} unsafe types.}
 The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
-These requirements add additional restrictions for type creation and restrict the creation of unsafe types.
-For instance, we cannot create a tainted array whose elements are checked pointers,~\eg Although syntactically valid, the type~\code{_t_Array_ptr<_Ptr<int>>} is~\emph{not} well-formed and is not allowed in~\lang.
-However, we can have a checked array whose elements are tainted pointers,~\eg~\code{_t_Array_ptr<_t_Ptr<int>>} is a valid type.
-
+% 
+These requirements \mzs{add additional restrictions for type creation and restrict the creation of unsafe types.}
+\mzr{prevents unsafe types from being constructed}.
+% 
+\mzr{
+  % 
+  Consider the type ~\code{_t_Array_ptr<_Ptr<int>>} which describes a tainted
+  array of checked pointers.
+  % 
+  Syntactically valid, this type is however \emph{illegal} in~\lang{} and
+  rejected by the well-formedness requirement. 
+}
+% 
+\mz{Made some changes to shift the emphasis in hope of improving clearness.}
+% 
+However, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_t_Array_ptr<_t_Ptr<int>>} is a valid type.
+%
+
+% 
 Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,

### 2023-01-09 sandp23/formal.tex
Adding all

--- 2023-01-09 sandp23/formal.tex
@@
 
+In \lang, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
+We provide some example descriptions here.
+Type equality $\tau=_{\Theta}\tau'$
+is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
+and the alpha equivalence of two function types;
+i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
+$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
+are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
+and the substitution results are equal.
+\aravind{For the above paragraph, give concrete example, For example: array is a subtype of ptr, etc}
+
+\aravind{We should explain how our type checker restricts unsafe interactions between checked and tainted pointer types.}
+
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+a -> checked pointer
+b -> Tainted pointer
+c -> Unchecked pointer.
+
+_Ptr<int> a;
+_t_Ptr<int> b;
+int *c;
+
+a = b; // Not allowed.
+b = a; // Not allowed.
+
+a = (_Ptr<int>)c; // This is okay.
+a = (_Ptr<int>)b; // Not allowed.
+\end{minted}
+
 \subsection{Semantics and Typing}\label{sec:semantics}

### 2023-01-10 sandp23/formal.tex
Adding all
[review] function pointer

--- 2023-01-09 sandp23/formal.tex
@@
   Syntactically valid, this type is however \emph{illegal} in~\lang{} and
-  rejected by the well-formedness requirement. 
+  is rejected by the well-formedness requirement. 
 }
@@
 Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
-where $\overline{x}$ represents a list of \tint{} type variables that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
+where $\overline{x}$ represents a list of \tint{} type variables that bind \mzu{bound}\mz{free?} variables appearing in $\overline{\tau}$ and $\tau$.
 An example of a function pointer type is shown below:

### 2023-01-10 sandp23/formal.tex

@@
 
+In \lang, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
+We provide some example descriptions here.
+Type equality $\tau=_{\Theta}\tau'$
+is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
+and the alpha equivalence of two function types;
+i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
+$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
+are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
+and the substitution results are equal.
+\aravind{For the above paragraph, give concrete example, For example: array is a subtype of ptr, etc}
+
+The type checker rules are designed such that any well-typed (\ie follows all the rules) is isolated by sandboxed code. We present the corresponding formal proof in~\apdx{}.
+
+One of the important type cheker rules is disallowing unsafe interactions between tainted and untainted (c/u) pointers.
+\aravind{Explain about the LET rule}
+\aravind{We should explain how our type checker restricts unsafe interactions between checked and tainted pointer types.}
+
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+a -> checked pointer
+b -> Tainted pointer
+c -> Unchecked pointer.
+
+_Ptr<int> a;
+_t_Ptr<int> b;
+int *c;
+
+a = b; // Not allowed.
+b = a; // Not allowed.
+
+a = (_Ptr<int>)c; // This is okay.
+a = (_Ptr<int>)b; // Not allowed.
+\end{minted}
+
 \subsection{Semantics and Typing}\label{sec:semantics}

### 2023-01-10 sandp23/introduction.tex
[review] fix typo

@@
   % 
-  How does regions look like? Can one region nested in another?
+  What do regions look like? Can one region be nested in another?
   %
-  If not, ``This partition contains the remaining (untainted) code and data and
+  ``This partition contains the remaining (untainted) code and data and
   has complete access to~\cmode region.'' creates a sense of recursive

### 2023-01-12 sandp23/appendix.tex
update.
Adding Implementation section and modifying evaluation section

@@
 \end{defi}
+
+\subsection{Additional Program evaluations}\label{appx:add-prog-eval}
+
+\subsubsection{parsons}
+Parsons is annotated comprehensively in two variants parsons\_wasm and parsons\_tainted.
+parsons\_wasm has most of its input parsing functions moved into the sandbox, whilst having all its pointers marked as tainted.
+These sandboxed functions interact with the checked region by making indirect calls through RLBOX's callback mechanism.
+However, with parsons\_tainted, we do not move any of the functions to the sandbox but still mark all the pointers as tainted.
+The test suite itself consists of 328 tests comprehensively testing the JSON parser's functionality.
+Benchmarks for both of these forks are recorded using the mean difference between the \systemname and generic-C/checked-C variants when executing 10 consecutive iterations of the test suite.
+parsons\_wasm expectedly shows 200/266\% runtime overhead when evaluated against checked-c and generic-c respectively due to the performance limitation of WebAssembly.
+However, evaluating parsons\_tainted against checked-c shows \systemname to be faster because \systemname by itself performs lighter run-time-instrumentation on tainted pointers as compared to the run-time bounds checking performed on checked pointers by checked-c.
+Furthermore, we only see an average peak memory of 9.5 KiB as compared to the anticipated 82 KiB overhead as Valgrind does not consider the WASM Shadow memory allocated to the tainted pointers.
+
+\subsubsection{LibPNG}
+\systemname changes for libPNG is narrow in scope and begins with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read().
+However, we also annotate sections of Lib-png that involve reading, writing, and image processing (interlace, intrapixel, etc) on user-input image data as tainted.
+That is, rows of image bytes are read into tainted pointers and the taintedness for the row\_bytes is propagated throughout the program.
+All our changes extend to the png2pnm and pnm2png executables.
+To evaluate png2pnm, we take the mean of 10 iterations of a test script that runs png2pnm on 52 png files located within the libpng's pngsuite.
+To test pnm2png, we take the mean of 10 iterations of pnm2png in converting a 52MB 5184x3456 pixels large pnm image file to png.
+Valgrind's reported lower Heap space consumption for \systemname converted code is due to the discounted space consumed on the heap by the Sandbox's shadow memory.
+Consequently, when evaluating pnm2png, \systemname's heap consumption was 52 MB lower as the entire image was loaded onto the shadow memory.
+
+\subsubsection{MicroHTTPD}
+MicroHTTPD demonstrates the practical difficulties in converting a program to \systemname.
+Our conversion for this program was aimed at sandboxing memory vulnerabilities CVE-2021-3466 and CVE-2013-7039.
+CVE-2021-3466 is described as a vulnerability from a buffer overflow that occurs in an unguarded "memcpy" which copies data into a structure pointer (struct MHD\_PostProcessor pp) which is type-casted to a char buffer (char *kbuf = (char *) \&pp[1]).
+Our changes would require making the "memcpy" safe by marking this pointer as tainted.
+However, this would either require marshaling the data pointed by this structure (and its sub-structure pointer members) pointer or would require marking every reference to this structure pointer as tainted, which in turn requires every pointer member of this structure to be tainted.
+Marshalling data between structure pointers is not easy and demands substantial marshaling code due to the spatial non-linearity of its pointer members unlike a char*.
+This did not align with our conversion goals which were aimed at making minimal changes.
+Consequently, the above CVE stands un-handled by \systemname.
+Our changes for CVE-2013-7039 involve marking the user input data arguments of this function as tainted pointers and in the interests of seeking minimal conversion changes, we do not propagate the tainted-ness on these functions.
+Following up on the chronological impossibility of sandboxing bugs before they are discovered and the general programmer intuiting, we moved many of the core internal functions (like MHD\_str\_pct\_decode\_strict\_() and MHD\_http\_unescape()) into the sandbox.
+
+\subsubsection{Tiny-bignum}
+Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive.
+Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf().
+Given that significant \systemname conversion efforts is attributed to understanding the source-code and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling the data between regions, Tiny-bignum only required 4 hours.
+The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

### 2023-01-12 sandp23/evaluation.tex
Adding Implementation section and modifying evaluation section

--- 2023-01-04 sandp23/evaluation.tex
@@
+\section{\systemname Implementation} \label{sec:implementation}
+\subsection{Overview}
+\systemname is implemented as a \systemname compiler and a \systemname source-level partitioning assistant (CheckMate).
+Given a well-typed \systemname program as an input, CheckMate performs the automation of moving the \systemname qualified function definitions to new "unchecked" source files, whilst inserting STD library headers needed for the successful compilation of these "unchecked" sources to an object file with a Sandbox-specific compiler.
+Since \systemname is irrelevant to the Sandbox-specific compiler, CheckMate internally runs an OCAML translator to cleanse away the \systemname annotations by converting these tainted definitions to generic-C.
+Now with the \systemname qualified functions and variables re-defined under the "unchecked" library object, the \systemname qualified functions definitions in the "checked" region are modified by the CheckMate accordingly to allow communication between the two regions.
+\subsection{\systemname Function Qualifiers}
+\systemname function qualifiers are introduced with a purpose of semantic policing to allow the developer to perform dependency-resolved program partitioning into checked and unchecked regions from within the checked sources.
+Each of the newly introduced qualifiers "\_Tainted", "\_Mirror", "\_TLIB" and "\_Callback" enforce different semantic-ruling at compile time to allow "unchecked" and "checked" functions to communicate and access global data in a region-exclusive manner.
+Although checked functions are allowed to call and access "unchecked" data within the checked region, "unchecked" region can only access checked data through a specialized callback mechanism.
+Each of these function qualifiers differently control casting, caller, callee and data-access rules within its qualifier-scope.
+For example, \_TLIB qualifier allows tainted pointer parameters to be cast to non-tainted pointer arguments at the call-site of a qualified function with developer's discretion that the qualified function does not leak tainted addresses.
+This type-checker relaxation is granted to enable developer to avoid marshalling costs when passing tainted pointers to external library functions.
+
+\subsection{CheckMate S2S transformation}
+\subsubsection{\textbf{Need for Callback}}
+Here, we use the example of a simple \systemname annotated string authentication program snippet in ~\lst{lst:callbackBefore} with~\code{StringProc} as its entry point (\entrypoint).
+~\code{StringProc} was found to internally perform a lot of unsafe string parsing and manipulation on the user input and thereby has been marked \_Tainted.
+However, ~\code{StringProc} internally calls ~\code{StringAuth} that performs final authentication on user input by referencing classified global checked data.
+In interests of maintaining confidentiality of global checked data from the unchecked region whilst satisfying this control and data dependency, ~\code{StringAuth} has been qualified as a \_Callback, thereby bestowing it with WASM Sandbox's addition sanity checks at its call-sites within the unchecked region.
+
+\subsubsection{\textbf{CheckMate S2S transformation on Checked Sources}}
+\myparagraph{WebAssembly Sandbox:}
+Unchecked region implemented as a WebAssembly sandbox requires registering ~\code{StringAuth} with its expected WebAssembly callee signature as a requirement to safeguard the control-flow between the two regions.
+Within the WebAssembly sandbox, all functions are defined with an extra sandbox pointer argument and all pointer members are unsigned integers.
+
+\myparagraph{CheckMate Instrumentation: }
+Consequently, ~\code{StringAuth}'s corresponding call-site signature in the tainted region would be 
+\newline
+~\code{_T_StringAuth}, which is an indirected trampoline call to the actual callback function.
+CheckMate identifies ~\code{StringAuth} to be a \_Callback qualified callee and correspondingly generates callback trampoline ~\code{StringAuth} and callback registration code
+\newline
+~\code{registerCallback_StringAuth} within the checked region as show in ~\lst{lst:callbackAfter}. 
+~\code{registerCallback_StringAuth} registers the WebAssembly-specific callback function ~\code{_T_StringAuth} by calling ~\code{\_SBXREG\_} with callback-specific argument count and argument signature.
+CheckMate replaces \_Tainted qualified ~\code{StringProc}'s definition with an indirect call to sandbox specific implementation ~\code{w2c_StringProc}.
+The \_Mirror qualifier is used mark a function in checked region as a local to both the checked and the unchecked regions.
+Consequently, CheckMate does not modify \_Mirror qualified ~\code{isStringMatching} and simply duplicates this function in the unchecked region.
+
+\subsubsection{\textbf{CheckMate S2S generation of Tainted Sources}}
+Implementing \systemname's unchecked region as a WebAssembly sandbox requires compiling unchecked functions and unchecked variables checked-region-exclusively with a sandbox-specific compiler (WebAssembly).
+Consequently, CheckMate copies each of the \_Mirror and \_Tainted qualified function definitions to a new source file and runs an OCAML parser to convert \systemname code to generic-C as shown in ~\lst{lst:callbackAfterTainted} which shows the CheckMate generated code for ~\code{StringProc}.
+After running CheckMate, each of newly generated tainted sources can be compiled to an "unchecked" static library using a sandbox-specific compiler.
+CheckMate's OCAML parser also modifies each of the tainted STDLIB functions from ~\code{t_strncpy} under ~\lst{lst:callbackBefore} to ~\code{strncpy} under ~\lst{lst:callbackAfterTainted}
+
+\begin{listing}
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_before_checkmate.c}
+\caption{Callback Annotations before running CheckMate}
+\label{lst:callbackBefore}
+\end{listing}
+\begin{listing}[t!]
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate.c}
+\caption{CheckMate generated Callback Annotations in Checked Sources }
+\label{lst:callbackAfter}
+\end{listing}
+\begin{listing}[t!]
+\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate_tainted.c}
+\caption{CheckMate generated Unchecked Sources }
+\label{lst:callbackAfterTainted}
+\end{listing}
+
+\subsection{\systemname Dynamic Instrumentation}
+\systemname compiler is implemented to be sandbox-agnostic and differs in inserted instrumentation based on the sandbox type specified by a compiler flag.
+However, the scope of this paper restricts discussion to the 32-Bit WebAssembly sandbox wherein which pointers are represented as 4-Byte offsets from a base sandbox heap.
+Similar analogy has been applied in \systemname for WebAssembly wherein all tainted pointers in the checked region store zero-extended offsets, that upon access, are transformed to full addresses by the function \ref{O2P}.
+\subsubsection{\textbf{Overview}}
+\systemname compiler inserts run-time instrumentation on every access to a tainted pointer to assert pointer taintedness and to generate dereferencible address from the pointer offset.
+Pointer taintedness is an invariant that prevents checked memory addresses from leaking into the tainted region by establishing the tainted pointer to only point into the Sandbox memory.
+Since WebAssembly sandbox is implemented in 32-Bit, specialized struct type is created and instrumentation is inserted by the \systemname compiler at every pointer access to account for pointer-swizzling across different pointer representations.
+
+\subsubsection{\textbf{Pointer-swizzling}}
+WebAssmebly implementation of \systemname motivates Tainted pointers in checked region to implemented as a zero-extended 32-bit offset into the Sandbox.
+Every access to the tainted pointer calls for a pointer generation mechanism that generates a fully de-referencible pointer address as
+shown by the equation\ref{O2P}.
+
+Since \systemname does not require qualified functions to mandatorily execute in the unchecked region, tainted pointers may never have to cross into the sandbox.
+In such cases, we need not swizzle and the tainted pointers are implemented as 64-Bit pointers that contain zero extended 4-Byte pointer offsets into the Sandbox memory.
+However, Need for swizzling emerges when indexing into multi-level tainted pointers or tainted pointers to structures that are passed into the sandbox region.
+“\_Decoy” is a pointer variable qualifier that informs the compiler to enforce 64-Bit to 32-bit pointer swizzling throughout the qualified variable's lifetime in the checked region.
+Pointer-Swizzling is implemented in Clang-CodeGen as a modified GEP-instruction that considers \_Decoy when generating swizzled addresses causing \_Decoy qualified variables to be accessed as 32-bit pointers in the checked region.
+\begin{equation}
+  \_T\_ptr\_ADDR = \_T\_ptr\_OFFSET + SBX\_HEAP\_BASE
+\label{O2P}
+\end{equation}
+
+\subsubsection{\textbf{Tainted Structures}}
+Tainted Pointers to structures crossing into the sandbox are expected to only allow encapsulation of pointer members that are of the tainted type.
+Consequently, Tstructs have been introduced as a specialized struct type that enforces compile-time semantics to allow prevent its pointer members from being of non-tainted type.
+\systemname also restricts tainted pointers to only point to structures of Tstruct type.
+These semantic restrictions of Tstructs are put in place to prevent checked pointer from leaking into the unchecked regions as members of pointer structures.
+Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout.
+
 \section{Evaluation}\label{sec:evaluation}
@@
 % \liyi{Deena? Leo? }
-Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks.
+Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks with the use of WebAssembly Sandbox.
+Consequently, all further references to sandbox refer to unchecked code in the WebAssembly Sandbox.
 MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
@@
 
+\subsection{Program Run-time Benchmarks}
+\begin{figure*}[t]
+\includegraphics[width=1.0\linewidth]{images/program_runtime_benchmark.png}
+\caption{Run-time benchmarking of ~\systemname.}
+\label{fig:runtime}
+\end{figure*}
+
 \myparagraph{Benchmarking Methodology}
-We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs).
-Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted\-ness of a pointer is propagated from the call site up until its declaration.
-We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
-Unlike the Runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
-This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLib functions that are in use by the tainted pointers in the program.
-Valgrind's memory figures do not account for Sandboxed allocations which happen in the shadow memory.
+\systemname is evaluated in terms of performance and conversion efforts similar to that of RLBOX \cite{rlbox-paper} by recording run-time and memory overhead incurred by a \systemname in comparison to an unconverted program in generic-C/checked-C.
+However, the scope of \systemname changes made on each of the programs is varied and the converted programs themselves are relatively tiny.
+Since each of the program is evaluated on a pre-included test-suite that runs on a single instance of the program, sandbox creation cost is a one-time constant due to \systemname implementing RLBOX's \cite{rlbox-paper} WebAssembly sandbox creation API.
+Run-time performance is evaluated with C's~\code{<time.h>} library by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs) and measuring latency delta as a percentage.
+Timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
+Memory overhead is measured with Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
+Unlike the Runtime performance, Peak Memory is recorded as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
+This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLIB functions that are in use by the tainted pointers in the program.
+Valgrind's memory figures do not account for Sandbox's Heap allocations which are linearly proportional to sandboxed code and tainted pointers.
 All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.
@@
 \subsection{Micro-Benchmarks}
-\subsubsection{Memory Access in SBX}
+
+\begin{figure}[h!]
+\includegraphics[width=1.0\linewidth]{images/microbenchmark.png}
+\caption{\systemname Micro-Benchmarks}
+\label{fig:microbenchmarks}
+\end{figure}
+
+Figure \ref{fig:microbenchmarks} shows calls to the sandbox and sandboxed code execution to have significantly higher overhead in comparison to that of Tainted pointers.
+This observation suggests for sandboxing less performance intensive code and reducing the indirect calls between the two regions.
+Consequently, unsafe regions of code can be annotated cost-effectively by choosing to annotate the unsafe pointer references within the demarked unsafe function as tainted at the cost of losing out on the Sandbox's safety features.
+
+\subsubsection{\textbf{Memory Access in SBX}}
 Memory access performance between the checked and the WASM region involves crafting a test case that involves a simple pointer arithmetic operation enclosed in a loop of 100k iterations.
-156.6\% overhead with \systemname is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.
 
-\subsubsection{Indirect-Call}
+\myparagraph{Results}
+156.6\% overhead as shown in Figure \ref{fig:microbenchmarks} is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.
+
+\subsubsection{\textbf{Indirect-Calls}}
 This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call from the checked region into the SBX) to return.
 This metric is evaluated against the time taken for a checked Callee (call within local scope) to return.
-Observed overhead is once again a consequence of code executing in the WASM Sandbox and call indirection.
 
-\subsubsection{Pointer Access}
+\myparagraph{Result}
+Observed overhead is a consequence of control transfer overhead between the checked and unchecked region as mentioned in \cite{rlbox-paper}
+
+\subsubsection{\textbf{Pointer Accesses}}
 Run-time Pointer overhead is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
+
+\myparagraph{Result}
 34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.   
+\myparagraph{Result:}
+Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
+\begin{figure*}[t]
+\includegraphics[width=1.0\linewidth]{images/program_memory_benchmark.png}
+\caption{Memory benchmarking of ~\systemname.}
+\label{fig:memory}
+\end{figure*}
 
-\begin{center}
-\label{fig:micrbenchmarks}
-\begin{tabular}{||c c||} 
+\begin{table*}[htb]
+  \centering
+  \begin{tabular}{|p{2.6 cm }|p{1.6 cm}|p{1.9cm}|p{1.7cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
     \hline
- Test-Name & \systemname overhead \\ [0.5ex] 
- \hline\hline
- Memory Access in SBX & +156\% \\
- Indirect-Call & +398\% \\ 
- Pointer Access & +34.16\% \\ [1ex]
+    \textbf{Task} & \textbf{ProFTPD} & \textbf{MicroHTTPD} & \textbf{UFTPD} & \textbf{LibPNG} & \textbf{Tiny Bignum} & \textbf{parsons (wasm)} & \textbf{parsons (tainted) }\\
+    \hline
+     \textbf{Pointers Annotated} & 6 & 139 & 146 & 248 & 69 & 364 & 378 \\
+     \textbf{Lines Sandboxed} & 0 & 650 & 90 & 0 & 30 & 800 & 0  \\
+     \textbf{CVEs fixed} & CVE-2010-4221 & CVE-2013-7039 & CVE-2020-14149, CVE-2020-5204 & CVE-2018-144550 & - & - & -  \\
+     \textbf{Time To Port} & 1 Day & 2 Days & 2 Days & 3 Days & 4 Hours & 1 Day & 2 Days \\
     \hline
   \end{tabular}
-\end{center}
+  \caption{Conversion Efforts}
+\end{table*}
 
-\subsection{Program Run-time Benchmarks}
+\subsection{Program Benchmarks}
 \myparagraph{Overview}
@@
 Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
-Conversion style and extent qualitatively dictate the converted program's conversion efforts required and the run-time performance.
 Consequently, each of the conversions for the six programs follows a varied approach.
 
-\myparagraph{Parsons}
-Parsons is annotated comprehensively in two variants parsons\_wasm and parsons\_tainted.
-parsons\_wasm has most of its input parsing functions moved into the sandbox, whilst having all its pointers marked as tainted.
-These sandboxed functions interact with the checked region by making indirect calls through RLBOX's callback mechanism.
-However, with parsons\_tainted, we do not move any of the functions to the sandbox but still mark all the pointers as tainted.
-The test suite itself consists of 328 tests comprehensively testing the JSON parser's functionality.
-Benchmarks for both of these forks are recorded using the mean difference between the \systemname and generic-C/checked-C variants when executing 10 consecutive iterations of the test suite.
-parsons\_wasm expectedly shows 200/266\% runtime overhead when evaluated against checked-c and generic-c respectively due to the performance limitation of WebAssembly.
-However, evaluating parsons\_tainted against checked-c shows \systemname to be faster because \systemname by itself performs lighter run-time-instrumentation on tainted pointers as compared to the run-time bounds checking performed on checked pointers by checked-c.
-Furthermore, we only see an average peak memory of 9.5 KiB as compared to the anticipated 82 KiB overhead as Valgrind does not consider the WASM Shadow memory allocated to the tainted pointers.
+\myparagraph{Observations}
+Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
+\begin{itemize}
+  \item Run-time Overhead is proportional to the extent of 
+annotated pointers and sandboxed code.
+  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
+  \item Marshalling data might sometimes be cost-effective instead of propagating the taintedness of a pointer
+  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
+  \item Conversion style and extent qualitatively dictate the converted program's conversion efforts and is modest at a couple of days. 
+\end{itemize}
 
-\myparagraph{LibPNG}
-\systemname changes for libPNG is narrow in scope and begins with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read().
-However, we also annotate sections of Lib-png that involve reading, writing, and image processing (interlace, intrapixel, etc) on user-input image data as tainted.
-That is, rows of image bytes are read into tainted pointers and the taintedness for the row\_bytes is propagated throughout the program.
-All our changes extend to the png2pnm and pnm2png executables.
-To evaluate png2pnm, we take the mean of 10 iterations of a test script that runs png2pnm on 52 png files located within the libpng's pngsuite.
-To test pnm2png, we take the mean of 10 iterations of pnm2png in converting a 52MB 5184x3456 pixels large pnm image file to png.
-Valgrind's reported lower Heap space consumption for \systemname converted code is due to the discounted space consumed on the heap by the Sandbox's shadow memory.
-Consequently, when evaluating pnm2png, \systemname's heap consumption was 52 MB lower as the entire image was loaded onto the shadow memory.
+\subsubsection{\textbf{ProFTPD}}
+\systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
+We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
 
-\myparagraph{ProFTPD}
-\systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
-We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
+\myparagraph{Testbench:}
 Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance.
@@
 This sampling is repeated 10 times and its mean value is reflected in the below table.
-Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
 
-\myparagraph{MicroHTTPD}
-MicroHTTPD demonstrates the practical difficulties in converting a program to \systemname.
-Our conversion for this program was aimed at sandboxing memory vulnerabilities CVE-2021-3466 and CVE-2013-7039.
-CVE-2021-3466 is described as a vulnerability from a buffer overflow that occurs in an unguarded "memcpy" which copies data into a structure pointer (struct MHD\_PostProcessor pp) which is type-casted to a char buffer (char *kbuf = (char *) \&pp[1]).
-Our changes would require making the "memcpy" safe by marking this pointer as tainted.
-However, this would either require marshaling the data pointed by this structure (and its sub-structure pointer members) pointer or would require marking every reference to this structure pointer as tainted, which in turn requires every pointer member of this structure to be tainted.
-Marshalling data between structure pointers is not easy and demands substantial marshaling code due to the spatial non-linearity of its pointer members unlike a char*.
-This did not align with our conversion goals which were aimed at making minimal changes.
-Consequently, the above CVE stands un-handled by \systemname.
-Our changes for CVE-2013-7039 involve marking the user input data arguments of this function as tainted pointers and in the interests of seeking minimal conversion changes, we do not propagate the tainted-ness on these functions.
-Following up on the chronological impossibility of sandboxing bugs before they are discovered and the general programmer intuiting, we moved many of the core internal functions (like MHD\_str\_pct\_decode\_strict\_() and MHD\_http\_unescape()) into the sandbox.
-
-\myparagraph{UFTPD}
+\subsubsection{\textbf{UFTPD}}
 \systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
@@
 Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.
+
+\myparagraph{Testbench:}
 For evaluation, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
@@
 
-\myparagraph{Tiny-bignum}
-Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive.
-Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf().
-Given that significant \systemname conversion efforts is attributed to understanding the source-code and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling the data between regions, Tiny-bignum only required 4 hours.
-The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.
+\myparagraph{Result:}
+Although, \systemname-UFTPD consists of modest changes in annotations and sandboxed code, observed overhead is significantly less at 10\% due to the less performance intensive sandboxed code and FTP-protocol overshadowed overhead from \systemname instrumentation and sandbox.
 
-\begin{center}
-\label{fig:prgrbenchmarks}
-\begin{tabular}{||c | p{1.3cm} | p{1.8cm}||} 
- \hline
- Program & Run-Time & Avg Peak Memory \\ [0.5ex] 
- \hline\hline
- ProFTPD & +47.8\% & +82.1 KiB  \\
- MicroHTTPD & +171.7\% & +24.7 KiB\\ 
- UFTPD & +2.7\% & +86.3 KiB \\ 
- Tiny-Bignum (vs checked-c) & -23.4\% &  +81.1 KiB \\
- Tiny-Bignum &  +121.7\% &  +81.1 KiB \\
- parsons\_wasm (vs checked-c) & +200.2\% & +9.8 KiB\\ 
- parsons\_tainted (vs checked-c) & -8.7\% & +9.2 KiB\\
- parsons\_wasm & +266.5\% & +9.8 KiB\\ 
- parsons\_tainted & +10.45\% & +9.2 KiB\\ 
- LibPNG (png2pnm) & +11.41\% & -102.5 KiB\\ 
- LibPNG (pnm2png) & +46.5\% & -51.6 MiB\\ [1ex]
- \hline
-
-\end{tabular}
-\end{center}
-
-\begin{center}
-\begin{tabular}{||p{2 cm} | p{1.2cm} | p{1.2cm} | p{1.0cm} | p{1.0cm}||} 
- \hline
- Program & Pointers Annotated  & Lines Sandboxed & CVEs fixed & Time To Port \\ [0.5ex] 
- \hline\hline
- ProFTPD & 6 & 0 & - & 1 Day  \\
- \hline
- MicroHTTPD & 139 & 650 & 1 & 4 Days \\
- \hline
- UFTPD & 146 & 90 & 2 & 3 Days \\ 
- \hline
- LibPNG & 248 &  0 & 1 & 8 Days \\
- \hline
- Tiny-Bignum &  69 &  30 & - & 4 Hours \\
- \hline
- parsons\_wasm & 364 & 800 &  - & 3 Days \\ 
- \hline
- parsons\_tainted & 378 & 0 & - & 8 Days \\ [1ex]
- \hline 
-\end{tabular}
-\end{center}
 % 

### 2023-01-12 sandp23/overview.tex
Adding all
Adding Implementation section and modifying evaluation section

--- 2023-01-08 sandp23/overview.tex
@@
 \label{subsec:identifyregionstosbx}
-Here,
-we use the example of a simple network server in~\lst{lst:comb} to explain changes required by the \systemname compiler.
-The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
+to explain changes required by the \systemname compiler.The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
 The server runs in a loop and calls~\code{handle_request}, which handles each network request.

### 2023-01-12 sandp23/examples/callback_mechanism_sample_before_checkmate.c

@@ -1,0 +1,52 @@
+#include <string_tainted.h> //tainted wrapper of STDLIB string.h
+...
+
+_Callback _TPtr<char> StringAuth(_TPtr<const char> input , size_t len) {
+  ...
+  //complex Function Body
+  /*
+  "_Callback" allows access to global checked data
+  
+  "_Callback" restricts function signature to
+  only contain pointers of Tainted type
+  Since Callbacks execute in checked region, they have no 
+  restrictions on callees.
+  */
+  ...
+}
+
+_Tainted _TPtr<char> StringProc(_TPtr<_TPtr<const char>> user_input,
+_TPtr<_TPtr<char>(_TPtr<const char> input, size_t len)>StringAuth) {
+  ...
+  //complex Function Body	
+  
+   /*
+   "_Tainted" restricts access to global data 
+   
+   "_Tainted" restricts function signature to 
+   only contain pointers of Tainted type 
+  
+   Since "_Tainted" functions are destined to execute in unchecked
+   region, callees of this function must be "_Mirror", "_Callback" or 
+   "_Tainted" qualified. 
+  */
+  ...
+  t_strncpy(local_str, user_input[i], len); // tainted wrapper for strncpy
+  ...
+  return StringAuth(one_past_start, string_len);
+}
+
+_Mirror int isStringMatching(char* str1 : itype (_Ptr<char>), 
+		char* str2 : itype(_Ptr<char>))
+{
+  ...
+  //complex Function body
+  /*
+   * "_Mirror" restricts access to global data
+   * 
+   * Since "_Mirror" functions are locally replicated in both 
+   * checked and unchecked regions, they have no restriction on 
+   * type signature and callees.
+   */
+  ...
+}

### 2023-01-12 sandp23/formal.tex

@@
 \subsubsection{Expressions ($e$)}
-\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$)($n\!:\!\tau$), pointer dereference and assignment ($\estar{e}$) and ($\eassign{e_1}{e_2}$), resp.)
+\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$)
+\mzu{($n\!:\!\tau$)}\mz{what's this?}, pointer dereference and assignment ($\estar{e}$) and ($\eassign{e_1}{e_2}$), resp.
 along with expressions that require special handling, such as,
@@
 For instance, a dynamic bounds cast expression
-
-\noindent
+\begin{center}
   {\footnotesize
-\code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}
-}
+    \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,
+      count(}$n$\code{))} }
+\end{center}
 becomes 
-{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}
-.
-
-We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
+{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}. 
+%
+
+% 
+\mzu{We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.}
+\mz{This sentence is unclear. I suppose that you mean an integer literal
+  can be both of interger type and pointer type depending on the context.  }
 For instance, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer.
 The~\code{strlen} expression plays an important role in the type-checking of null-terminated arrays.
-For simplifying type checking, we assume~\code{strlen} to operate on variables $x$ rather than arbitrary expressions.
+For simplifying type checking, we assume~\code{strlen} to operate on variables \mzs{$x$} rather than arbitrary expressions.
+% 
 The heap allocation expression $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions (\cregion or \ucregion).
 We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-The $\echeckedtext$ and $\euncheckedtext$ expressions are used to represent the corresponding code regions.
+The $\echeckedtext$ and $\euncheckedtext$ expressions are used to
+\mzs{represent the corresponding} \mzr{delimit} code regions.
 To guarantee the non-exposure property, we extend the base syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$.
-Wherein $\overline{x}$ restricts all free variables appearing in $e$ to be checked or unchecked pointers according to the corresponding region.
-
-\lang aims to be simple enough to work with but powerful enough to
+\mzu{Wherein $\overline{x}$ restricts all free variables appearing in $e$ to be checked or unchecked pointers according to the corresponding region.}
+\mz{Is this sentence grammatical?}
+%
+
+% 
+\lang{} aims to be simple enough to work with but powerful enough to
 encode realistic \systemname idioms. For example, mutable local
   variables can be encoded as immutable locals that point to the heap;
-the use of \code{&} can be simulated with \code{malloc};
+\mzu{the use of \code{&} can be simulated with \code{malloc};}
+\mz{This doesn't make sense if you use \code{&} to create alias. You don't want
+  to allocate memory.}
+% 
 and loops can be encoded as recursive function calls. \code{struct}s are
@@
 A few selected type rules are presented in~\fig{fig:type-system-1}.
+% 
+\mz{Move the figure here.}
+% 
 The complete set of typing rules and special handling of (NT)-arrays are provided in~\apdx{sec:literal-pointer-typing}.
@@
 \aravind{Explain about the LET rule}
+
 \aravind{We should explain how our type checker restricts unsafe interactions between checked and tainted pointer types.}
+
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+a -> checked pointer
+b -> Tainted pointer
+c -> Unchecked pointer.
+
+_Ptr<int> a;
+_t_Ptr<int> b;
+int *c;
+
+a = b; // Not allowed.
+b = a; // Not allowed.
+\end{minted}
+
+The \textsc{T-CastPtr} rule in \Cref{fig:type-system-1}
+permits casting from an expression of type $\tau'$ to a checked pointer when
+$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
+$\sqsubseteq$ is built on the type equality ($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). 
+The rule  ($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
+\tarrayptr{b_l}{b_h}{\tau}{m}$) permits treating a singleton
+pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
+Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
+
+\aravind{Explain this with .}
 
@@
 thus, in every heap update, the value type annotation remains the same through program executions.
+% 
+\mz{Does a ``non-fat pointer'' system cause the type preservation of heap
+  values?
+  % 
+  The story seems to be that you design the system in a certain way s.t. heap
+  value types are unchanged, which allows you to erase the type.
+  % 
+  And, from the type erasure property, we know that we don't need fat pointers.
+}
+% 
 Additionally, for both stack and heap, 
@@
 Correspondingly, $\heap(m,0)$ should always be undefined.
-The relation steps to a \emph{result} $r$,
-which is either an expression or a $\enull$ or $\ebounds$ failure,
-represent a null-pointer dereference or out-of-bounds access,
-respectively. Such failures are a \emph{good} outcome; stuck states
+% 
+The relation steps to a \emph{result} $r$, which is \mzs{either} \mzr{one of} an
+expression, a $\enull$ or $\ebounds$ failure, represent \mzr{an expression right
+  after the reduction}, a null-pointer dereference or out-of-bounds access,
+respectively.
+% 
+Such failures are a \emph{good} outcome; stuck states
 (non-value expressions that cannot transition to a result $r$)
 characterizing undefined behavior.
+%
 % 
@@
 it means that we are exposing a checked pointer address to unsafe regions.
-A $\echeckedtext$ or $\euncheckedtext$ block represents 
+% 
+\mzu{A $\echeckedtext$ or $\euncheckedtext$ block represents 
   the context switching from a checked to an checked region, or vice versa.
-We need to make sure no checked pointers are information exposed to unsafe code regions.
+}
+% 
+\mz{An action + checked/unchecked block represents the switching.}
+% 
+We need to make sure no checked pointers are \mzs{information} exposed to unsafe code regions.
 as rules \textsc{S-Unchecked} and \textsc{S-Checked} in \Cref{fig:semantics}.
+% 
+\mz{This paragraph is not that cohesive.}
+%
 

### 2023-01-12 sandp23/overview.tex

--- 2023-01-08 sandp23/overview.tex
@@
 This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
+\aravind{Change title of this section and say that this mode has no guarentees and move this before evaluation.}
 

### 2023-01-12 sandp23/evaluation.tex

@@
 % \liyi{Deena? Leo? }
-Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks with the use of WebAssembly Sandbox.
+Our evaluation of \systemname consists of a set of tests that can be classified into Micro-benchmarks and Program Benchmarks that evaluate \systemname on WebAssembly sandbox.
 Consequently, all further references to sandbox refer to unchecked code in the WebAssembly Sandbox.

### 2023-01-12 sandp23/formal.tex
Adding

@@
 \aravind{Explain about the LET rule}
+%
+Rules \textsc{T-Let} and \textsc{T-LetInt} in in \Cref{fig:type-system-1} type a $\elettext$ expression, which also admits
+type dependency. 
+In particular, the result of evaluating a $\elettext$ expression
+may have a type that refers to one of its bound variables (e.g., if
+the result is a checked pointer with a variable-defined bound). 
+If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
+Note that we restrict the expression $e_1$ to syntactically match the
+structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
+Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
+$\erettext$ does not appear in source programs but is introduced by the semantics when
+evaluating a let binding (rule \textsc{S-Let} in
+Fig.~\ref{fig:semantics}). 
+%\liyi{why? }
+% After the evaluation of a let binding a variable $x$ concludes,
+%we need to restore any prior binding of $x$, which is either
+%$\bot$ (meaning that there is no $x$ originally) or some value
+%$\evalue{n}{\tau}$.
 
@@
 
-\aravind{Explain this with .}
+\aravind{Explain how the cast rule allows/prevents the casts.}
 
@@
 
-\subsection{Semantics and Typing}\label{sec:semantics}
+\myparagraph{Constant Validity}
+Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
+describe type assumptions for constants appearing in a program.
+$\neg \cmode(\tau)$ judges that a constant pointer 
+in an unchecked region cannot be of a checked type.
+The restriction ensures that programmers 
+cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+In rule \textsc{T-ConstC}, we requires a static 
+verification procedure for validating a constant pointer in \Cref{fig:const-type}.
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
+validates the constant $\evalue{n}{\tau}$, 
+where $\heap(m)$ is the initial heap that the constant resides on and
+$\sigma$ is a set of constant assumed to be checked.
+A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
+A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
+and the address stores a function with the right type.
+The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
+where every element in the pointer range ($[0,\size(\omega))$) should be well
+typed.
+A checked pointer checks validity in type step as rule \textsc{T-ConstC},
+while a tainted/unchecked pointer does not check for such during the type checking.
+Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.
+
+\myparagraph{Unchecked and Checked Blocks}
+%
+During the type checking,
+Both $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$
+check all free variables in $e$ are within $\overline{x}$;
+the types for $\overline{x}$ and the final return type $\tau$ of $e$ have no checked pointers.
+Otherwise, it violates the non-exposure safety.
+For example, \code{read_msg} in the \code{handle_request} function is tainted in \Cref{lst:humantaint},
+if any argument for \code{handle_request} is a checked pointer, 
+it means that we are exposing a checked pointer address to unsafe regions.
+% 
+\mzu{A $\echeckedtext$ or $\euncheckedtext$ block represents 
+  the context switching from a checked to an checked region, or vice versa.
+}
+% 
+\mz{An action + checked/unchecked block represents the switching.}
+% 
+We need to make sure no checked pointers are \mzs{information} exposed to unsafe code regions.
+as rules \textsc{S-Unchecked} and \textsc{S-Checked} in \Cref{fig:semantics}.
+% 
+\mz{This paragraph is not that cohesive.}
+%
+
+\myparagraph{Dependent Function Pointers}
+Rule \textsc{T-Fun} in \Cref{fig:type-system-1} is the dependent function call rule. 
+Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
+from a type-check for $e$ and the types $\overline{\tau'}$ from the argument type checks for $\overline{e}$,
+we confirm that each of $\overline{\tau'}$ is
+a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$,
+which replaces possible integer bound variables $\overline{x}$ with bound expressions $\overline{e'}$.
+The final result type is the defined target type $\tau$ appearing in the function pointer type
+also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
+Consider the \code{process_req2} function in
+Fig.~\ref{lst:final}; its parameter type for \code{msg} 
+depends on \code{m_1}. The \textsc{T-Fun} rule will substitute 
+\code{n} with the argument at a call-site.
+The semantics manages variable scopes using the special $\erettext$
+form. \textsc{S-Let} evaluates to a configuration whose expression is
+$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
+and remember $x$ and its new value $\evalue{n}{\tau}$
+in $e$'s scope that is defined by the $\erettext$ operation.
+Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
+we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope.
+After one-step evaluation is completed, 
+we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
+and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
+This procedure continues until $e'$ becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
+the literal.
+
+\textsc{S-FunC} and \textsc{S-FunT} are
+for $\cmode$ and $\tmode$ mode function pointers, respectively. 
+A call to a function pointer $n$ retrieves
+ the function definition in $n$'s location in the global function store $\Xi$,
+which maps function pointers to
+function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
+$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
+is the parameter list of variables and their types, 
+$\xi$ determines the mode of the function, and $e$ is the
+function body. 
+Similar to \heap, the global function store $\Xi$ is also partitioned into
+two parts ($\cmode$ and $\umode$ stores), each of which
+maps addresses (integer literals) to the function data described above.
+
+\systemname{} has dependent functions, whose semantic explaination is given in \Cref{appx:add-type-sem}.
+Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
+  annotations $\overline{\tau_a}$ with
+  $\overline{\tau}$ (after instantiation) from the function's
+  signature. Using $\overline{\tau_a}$ when executing the body of
+the function has no impact on the soundness of \lang, but will violate
+Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
+Rule \textsc{S-FunT} defines the tainted version of function call semantics.
+In such case, the verification process 
+$\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}$
+makes sure that the function in the global store is well-defined and has the right type.
+
+\subsection{Semantics}\label{sec:semantics}
 
@@
 
-Here, we discuss the \lang type system (\Cref{fig:type-system-1}) and operational semantics (\Cref{fig:semantics}); 
+Here, we discuss the \lang operational semantics (\Cref{fig:semantics}); 
 mainly focusing on the new changes on top of \checkedc in \cite{li22checkedc} with function pointers, modes, and function calls.
@@
 
-The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
-which states that in a type environment $\Gamma$ (mapping variables to
-their types) and a predicate environment $\Theta$ (mapping integer-typed
-variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-in context mode $m$. Key rules for this judgment are given in
-Fig.~\ref{fig:type-system-1},
-while the operational semantics for \lang is defined as a small-step
+%The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+%which states that in a type environment $\Gamma$ (mapping variables to
+%their types) and a predicate environment $\Theta$ (mapping integer-typed
+%variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
+%in context mode $m$. Key rules for this judgment are given in
+%Fig.~\ref{fig:type-system-1},
+
+The operational semantics for \lang is defined as a small-step
 transition relation with the judgment $ (\varphi,\heap,e)
@@
 
+\begin{DIFnomarkup}
+ \begin{figure}[t]
+ {\small
+
+ \begin{mathpar}
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m n : \tint}
+
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
+       {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
+
+   \inferrule
+       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
+            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       { \xi \le m 
+     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
+       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
+
+   \inferrule
+       {\neg\funptr(\omega)\\ \xi \le m\\
+        \forall i \in [0,\size(\omega)) \;.\;
+            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+ \end{mathpar}
+ }
+{\footnotesize
+\[
+\begin{array}{l} 
+\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
+\qquad
+\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
+\end{array}
+\]
+}
+ \caption{Verification/Type Rules for Constants}
+ \label{fig:const-type}
+ \end{figure}
+\end{DIFnomarkup}
+
+% SKIPPING THIS
+\iffalse
+
 \myparagraph{Type Equality and Subtyping and Casting}
@@
 
+\fi
 \subsection{Compilation}\label{sec:compilation}

### 2023-01-12 sandp23/implementation.tex
Splitting Implementation

--- 2023-01-12 sandp23/evaluation.tex
@@
 Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout.
-
-\section{Evaluation}\label{sec:evaluation}
-
-% \review{While I found the idea behind section V very interesting, the current version
-%   of this section lacks some details that would help in better understanding (1) 
-%   how the approach works, and (2) the overall scope of the approach. 
-%   $\\$
-%   For instance, the authors state that, following [19], they try to "exercise
-%   interesting patterns" by adding "admissible but redundant typing rules" like
-%   G-ASTR. There are a few points that are unclear here: (1) are these rules
-%   discovered manually or automatically (starting from the Redex semantics)?, (2)
-%   are there any guiding principles for coming up with rules that lead to
-%   interesting cases?
-%   $\\$
-%   Later, the authors refer to "generation rules modified to be slightly more
-%   permissive" to generate "a little" ill-typed terms. Again, are these rules
-%   obtained automatically or defined manually? If the latter, did you follow any
-%   methodology to derive such rules? Are these rules the same as the "admissible
-%   but redundant typing rules" from above?}
-% \liyi{Deena? Leo? }
-Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks with the use of WebAssembly Sandbox.
-Consequently, all further references to sandbox refer to unchecked code in the WebAssembly Sandbox.
-MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
-We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.
-
-\subsection{Program Run-time Benchmarks}
-\begin{figure*}[t]
-\includegraphics[width=1.0\linewidth]{images/program_runtime_benchmark.png}
-\caption{Run-time benchmarking of ~\systemname.}
-\label{fig:runtime}
-\end{figure*}
-
-\myparagraph{Benchmarking Methodology}
-\systemname is evaluated in terms of performance and conversion efforts similar to that of RLBOX \cite{rlbox-paper} by recording run-time and memory overhead incurred by a \systemname in comparison to an unconverted program in generic-C/checked-C.
-However, the scope of \systemname changes made on each of the programs is varied and the converted programs themselves are relatively tiny.
-Since each of the program is evaluated on a pre-included test-suite that runs on a single instance of the program, sandbox creation cost is a one-time constant due to \systemname implementing RLBOX's \cite{rlbox-paper} WebAssembly sandbox creation API.
-Run-time performance is evaluated with C's~\code{<time.h>} library by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs) and measuring latency delta as a percentage.
-Timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
-Memory overhead is measured with Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
-Unlike the Runtime performance, Peak Memory is recorded as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
-This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLIB functions that are in use by the tainted pointers in the program.
-Valgrind's memory figures do not account for Sandbox's Heap allocations which are linearly proportional to sandboxed code and tainted pointers.
-All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.
-
-\subsection{Micro-Benchmarks}
-
-\begin{figure}[h!]
-\includegraphics[width=1.0\linewidth]{images/microbenchmark.png}
-\caption{\systemname Micro-Benchmarks}
-\label{fig:microbenchmarks}
-\end{figure}
-
-Figure \ref{fig:microbenchmarks} shows calls to the sandbox and sandboxed code execution to have significantly higher overhead in comparison to that of Tainted pointers.
-This observation suggests for sandboxing less performance intensive code and reducing the indirect calls between the two regions.
-Consequently, unsafe regions of code can be annotated cost-effectively by choosing to annotate the unsafe pointer references within the demarked unsafe function as tainted at the cost of losing out on the Sandbox's safety features.
-
-\subsubsection{\textbf{Memory Access in SBX}}
-Memory access performance between the checked and the WASM region involves crafting a test case that involves a simple pointer arithmetic operation enclosed in a loop of 100k iterations.
-
-\myparagraph{Results}
-156.6\% overhead as shown in Figure \ref{fig:microbenchmarks} is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.
-
-\subsubsection{\textbf{Indirect-Calls}}
-This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call from the checked region into the SBX) to return.
-This metric is evaluated against the time taken for a checked Callee (call within local scope) to return.
-
-\myparagraph{Result}
-Observed overhead is a consequence of control transfer overhead between the checked and unchecked region as mentioned in \cite{rlbox-paper}
-
-\subsubsection{\textbf{Pointer Accesses}}
-Run-time Pointer overhead is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
-
-\myparagraph{Result}
-34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.   
-\myparagraph{Result:}
-Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
-\begin{figure*}[t]
-\includegraphics[width=1.0\linewidth]{images/program_memory_benchmark.png}
-\caption{Memory benchmarking of ~\systemname.}
-\label{fig:memory}
-\end{figure*}
-
-\begin{table*}[htb]
-  \centering
-  \begin{tabular}{|p{2.6 cm }|p{1.6 cm}|p{1.9cm}|p{1.7cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-    \hline
-    \textbf{Task} & \textbf{ProFTPD} & \textbf{MicroHTTPD} & \textbf{UFTPD} & \textbf{LibPNG} & \textbf{Tiny Bignum} & \textbf{parsons (wasm)} & \textbf{parsons (tainted) }\\
-    \hline
-     \textbf{Pointers Annotated} & 6 & 139 & 146 & 248 & 69 & 364 & 378 \\
-     \textbf{Lines Sandboxed} & 0 & 650 & 90 & 0 & 30 & 800 & 0  \\
-     \textbf{CVEs fixed} & CVE-2010-4221 & CVE-2013-7039 & CVE-2020-14149, CVE-2020-5204 & CVE-2018-144550 & - & - & -  \\
-     \textbf{Time To Port} & 1 Day & 2 Days & 2 Days & 3 Days & 4 Hours & 1 Day & 2 Days \\
-    \hline
-  \end{tabular}
-  \caption{Conversion Efforts}
-\end{table*}
-
-\subsection{Program Benchmarks}
-\myparagraph{Overview}
-We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program.
-We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code .
-However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible.
-Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
-Consequently, each of the conversions for the six programs follows a varied approach.
-
-\myparagraph{Observations}
-Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
-\begin{itemize}
-  \item Run-time Overhead is proportional to the extent of 
-annotated pointers and sandboxed code.
-  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
-  \item Marshalling data might sometimes be cost-effective instead of propagating the taintedness of a pointer
-  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
-  \item Conversion style and extent qualitatively dictate the converted program's conversion efforts and is modest at a couple of days. 
-\end{itemize}
-
-\subsubsection{\textbf{ProFTPD}}
-\systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
-We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
-
-\myparagraph{Testbench:}
-Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance.
-For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets().
-This sampling is repeated 10 times and its mean value is reflected in the below table.
-
-\subsubsection{\textbf{UFTPD}}
-\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
-CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function.
-CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function.
-Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.
-
-\myparagraph{Testbench:}
-For evaluation, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
-Following this we record an entry for each of these tests as a mean of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the below table.
-
-\myparagraph{Result:}
-Although, \systemname-UFTPD consists of modest changes in annotations and sandboxed code, observed overhead is significantly less at 10\% due to the less performance intensive sandboxed code and FTP-protocol overshadowed overhead from \systemname instrumentation and sandbox.
-
-% 
-%\leo{The following is extremely
-%  weak. ``Most of them'', were there any that weren't? Which ones? For
-%  the ones that were, mention github issues.}  The random generator,
-%equipped with the conversion tool, successfully found a few minor
-%errors in the clang compiler, most of them were already issues in the
-%git bug reports. For example, we discovered that while the ternary
-%operator is implemented in the compiler it cannot handle complex
-%bounds types in the branches. The static analysis is not sophisticated
-%enough to properly detect that both branches have the same type. While
-%not precisely a bug, the clang compiler does not permit memory for
-%null terminated arrays to be allocated with calloc. Although calloc
-%fills all spaces in memory with null, the compiler does not recognize
-%this and claims that it is an unsafe cast.
-
-% Recall that our
-% formal model makes liberal use of bounds annotations in literals and
-% the heap.
-
-% In order to get a better understanding of the formalism we wrote it in
-% redex. This allowed us to make sure that expressions were well-typed
-% and evaluated to what we expected. It also was helpful for use in
-% prototyping; new features could first be added to the redex model to
-% see how they interacted with the existing language. This model was
-% slightly larger than the Coq model and there are some differences in
-% the type systems. We included top level functions and conditional
-% expressions. All of these extra expressions are still expressible in
-% the coq model, for example functions can be represented as nested let
-% expressions. In the Coq model variables are stored on a stack while in
-% the Redex model the variables are simply looked up in the context. In
-% general the Redex model is easier to modify and slightly closer to the
-% actual Checked-C specifications. Instead of using the model for a
-% static proof, we used it to increase our certainty of the accuracy of
-% the model.
-
-% \item Describe the random testing generator setup and the properties
-%   to test.
-% \yiyun{Deena's description of the implementation details. I tried
-%   integrating the ones that I find relevant/interesting to the text above. Maybe we can
-%   add more if we have some space to fill in.}
-% In order for our guarantee of safety to hold, we need to know that our
-% model acurately reflects the CheckedC clang compiler. Safety is proved
-% for the Coq model, but it is significantly smaller than the actual
-% language. The Redex model is a combination of both. It is written in
-% the same style as the formalism but has slightly more of Checked-C's
-% extra features. If expressions from the Redex model display the same
-% behavior as equivalent programs in Checked-C then we have greater
-% certainty that our model is useful. We built a random testing
-% generator to increase this certainty.
-
-  % \item Describe the bug findings from the random testing against the Checked-C compiler.
-%   \leo{This is now integrated above}
-% The generator was helpful in finding bugs in the redex model.
-Several things failed to typecheck that should have been well typed, and the generator was able to catch them.
-The generated code also found a few minor errors in the clang compiler, most of them were already issues in the git bug reports.
-For example we discovered that while the ternary operator is implemented in the compiler it cannot handle complex bounds types in the branches.
-The static analysis is not sophisticated enough to properly detect that both branches have the same type.
-While not precisely a bug, the clang compiler does not permit memory for null terminated arrays to be allocated with calloc.
-Although calloc fills all spaces in memory with null, the compiler does not recognize this and claims that it is an unsafe cast.
-In the Redex model there is no issue with this.
-A few other minor things were brought to light in the implementation of the generator.
-The main use was to increase certainty that the behavior in the formal model accurately matched the clang compiler.
-%  
-% % \end{itemize}
-
-% \begin{itemize}
-%  
-% \item Show that why the formal semantics/type-system defined for Checked-C is useful. 
-% Since we have certainty that our model reflects the clang compiler the model is very useful.
-Proofs are easier on the smaller model, so we can show  that certain things are true for it.
-Since the Redex model is between the formalism and the clang version we can have certainty that properties we expect are actually true for the clang version.
-%  
-% \begin{itemize}
-% \item Show some bug findings. 
-% \item Show the properties that we can guarantee for Checked-C based on the type-system and blame theorem.
-% \item Maybe other useful tools that can be extracted from the Redex model.
-%  
-% \end{itemize}
-%  
-% \end{itemize}

### 2023-01-13 sandp23/formal.tex
update.

@@
   \small \centering
-  \[ \hspace*{-1.2em}
+  \[ \hspace*{-1.5em}
+\setlength{\arraycolsep}{4pt}
 \begin{array}{l}
@@
        \text{Variables:}~ x
-& \text{Integers:}~n::=\mathbb{Z} 
+& \quad\text{Integers:}~n::=\mathbb{Z} 
 \end{array}
@@
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}  \\
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}
+\mid \ret{x}{\evalue{n}{\tau}}{e}
+  \\
 &&\mid& \estrlen{x} \mid \emalloc{\xi}{\omega} \mid\estar{e}\mid\eassign{e}{e}  \\
@@
 % 
-\mzs{The formal model also enables us to}
-% 
-\mzr{From those, we} develop~\lang’s meta-
+From those, we develop~\lang’s meta-
 theories, including the type soundness, non-exposure, non-crashing,
-and the compiler simulation theorems, which we present in~\apdx{sec:theorem}.
+and the compiler simulation theorems.
 
@@
 \subsection{Syntax}\label{sec:syntax}
-The~\fig{fig:checkc-syn} shows a simplified syntax of~\lang.
-\subsubsection{Types ($\omega$)}
-At a high level, we classify types as word-size value \mzr{type}s, multi-word
-value \mzr{type}s, or function \mzr{type}s.
-%
-\mz{
-  % 
-  You cannot classify a type as values. You can only classify types into other 3
-  categories, namely, word, multi-word and function types.
-  % 
-}
-% 
-\mzs{The}\mzr{A} word-size value\mzs{s} can be either \mzr{of type integer} or
-\mzr{of type pointer}.
-% 
-\mzr{Each} pointer type\mzs{s} ($\tptr{\omega}{\xi}$) \mzr{is represented by} a
+The~\fig{fig:checkc-syn} shows the syntax of~\lang.
+
+\myparagraph{Types}
+At a high level, we classify types as word-size value, multi-word
+value, or function types. A word-size value can be either integer or pointer.
+Every pointer type ($\tptr{\omega}{\xi}$) includes a
 pointer mode annotation ($\xi$, the difference between context and pointer modes
 is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
-or unchecked (\umode), and a type ($\omega$) denoting \mzr{the valid type of
-  values it can refer to}.
-%
-
-% 
-\mzr{The multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
+or unchecked (\umode), and a type ($\omega$) denoting the valid value type it is pointed to.
+A multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
 arrays and null-terminated arrays is constructed by the type of elements in the
-array ($\tau$), an array bound ($\bvar$) and an array flag ($\kappa$).}
-% 
-\mzr{A} bound can be either \mzr{an} integer literal $n$, \mzr{an} expressions
-$x + n$, or an explicit specification of lower and upper bound\mzr{s},~\ie
-$(b_l,b_h)$.
-%
-\mzr{ %
-  Additionally, to distinguish null-terminated arrays from other arrays, we set
-  \(\kappa\) to be \textit{nt} for nt-arrays and \(\cdot\) for other arrays (for
-  convenience, we will elide \(\cdot\) in the text).
-  % 
-}
-%
-
-% 
-An example representation of an array and null-terminated array in~\lang syntax is shown below:
+array ($\tau$), an array bound ($\bvar$) comprised of an upper and
+lower bound on the size of the array ($(b_l,b_h)$), and an array flag ($\kappa$).
+Bounds $b$ are
+limited to integer literals $n$ and expressions $x + n$.
+Whether an array pointer is null terminated or not is determined by annotation
+$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
+otherwise (we elide $\cdot$ when writing types).
+An example representation of an array and null-terminated array in~\lang is shown below:
 % 
@@
 \]
-\mzr{For simplicity}\mzs{To simplify the representation}, we write
+
+For simplicity, we write
 $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean $\tptr{\tarray{0}{b}{\tau}}{\cmode}$,
@@
 
-\emph{\mzr{Rejecting}\mzs{Disallowing} unsafe types.}
+\myparagraph{Disallowing Unsafe Types}
 The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
 % 
-These requirements \mzs{add additional restrictions for type creation and restrict the creation of unsafe types.}
-\mzr{prevents unsafe types from being constructed}.
-% 
-\mzr{
-  % 
+These requirements prevents unsafe types from being constructed.
   Consider the type ~\code{_t_Array_ptr<_Ptr<int>>} which describes a tainted
   array of checked pointers.
-  % 
   Syntactically valid, this type is however \emph{illegal} in~\lang{} and
   is rejected by the well-formedness requirement. 
-}
-% 
-\mz{Made some changes to shift the emphasis in hope of improving clearness.}
-% 
-However, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_t_Array_ptr<_t_Ptr<int>>} is a valid type.
+However, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
 %
@@
 Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
-where $\overline{x}$ represents a list of \tint{} type variables that bind \mzu{bound}\mz{free?} variables appearing in $\overline{\tau}$ and $\tau$.
+where $\overline{x}$ represents a list of \tint{} type variables that bind type variables appearing in $\overline{\tau}$ and $\tau$.
 An example of a function pointer type is shown below:
@@
 \begin{array}{l}
-$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{),}$\\
-\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>: count(}$n$\code{))>}$
+$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau_1$\code{> : count(}$n$\code{),}$\\
+\qquad\qquad$\code{_t_NT_Array_ptr<}$\tau_2$\code{>: count(}$n$\code{), int}$\;n$\code{)>}$
 \\[0.2em]
-\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
+\Leftrightarrow\; $\tptr{(\tfun{n}{\tint \times \tntarrayptr{0}{n}{\tau_2}{\tmode} \times \tntarrayptr{0}{n}{\tau_1}{\tmode}}{\tint})}{\tmode}$
 \end{array}
 \]
-The function type also has well-formed requirements (\fig{fig:wffunctions} in Appendix), which disallows nesting checked or unchecked pointers inside tainted pointers.
+The function type also has well-formed requirements (\fig{fig:wffunctions} in Appendix), which disallows nesting checked pointers inside tainted pointers.
 Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
-\aravind{The syntax of well-formed requirements for functions is incorrect. It looks like the forall qualifier is missing.}
 
@@
 %$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-\subsubsection{Expressions ($e$)}
+\myparagraph{Expressions}
 \lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$)

### 2023-01-13 sandp23/implementation.tex
Implementation overhaul

@@
-\section{\systemname Implementation} \label{sec:implementation}
-\subsection{Overview}
-\systemname is implemented as a \systemname compiler and a \systemname source-level partitioning assistant (CheckMate).
+\section{Implementation} \label{sec:implementation}
+As mentioned in~\sect{sec:overview}, our implementation of~\systemname{} has two main components,~\ie source level program partitioner (\sourcerewriter) and compiler.
+Given a well-typed~\systemname{} program and a sandbox configuration,~\sourcerewriter performs source-to-source transformation and splits the program into two parts (\ie two sets of source files) -- checked and tainted source files corresponding to~\cregion and~\ucregion respectively.
+Our compiler will compile the checked source files and adds the necessary dynamic instruction.
+The tainted source files need to be compiled with the target sandbox compiler.
+Finally, these two sets of object files will be linked along with any sandbox library to produce the final executable that enforces our security guarantees.
+
+\subsection{\sourcerewriter} \label{subsec:checkmate}
+The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{}).
+However, we use a small OCAML program (\numocamllines{}) module to remove annotations and make the code compilable with the sandbox compiler (CITE CORRESPONDING IMPL SECTION).
+
+\subsubsection{Additional Function Qualifiers}
+In addition to the~\code{_Tainted} qualifier that marks functions to be in~\ucregion, we provide a few other qualifiers that enable developers to provide additional information and ease the partitioning process.
+Specifically, we provide three additional qualifiers:~\code{_Callback},~\code{_Mirror}, and~\code{_TLIB}.
+
+\noindent\emph{\_Callback}:
+Developers should use this qualifier to mark callback functions,~\ie functions in~\cregion, that can be called from the tainted region. 
+The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\sect{}).
+
+\noindent\emph{\_Mirror}: 
+This qualifier copies the corresponding function both into~\cregion and~\ucregion code. Such that callers from each region will invoke the corresponding copy.
+This helps handle certain simple utility functions that might be called from both~\cregion and~\ucregion code.
+\aravind{Give an example from our evaluation program. (E.g., function foo in program1)}.
+
+Placing such common functions in either of the regions restricts their usage from the other region.
+For instance, if we place~\code{foo} in~\ucregion, then all parameters of it should be tainted.
+Consequently, callers from~\cregion should use tainted pointers to invoke~\code{foo}, making the arguments and other dependent pointers tainted, leading to over-tainting.
+Similarly, placing~\code{foo} in~\cregion will make it hard for the callers in~\ucregion as they should now use the callback mechanism to invoke~\code{foo}.
+
+Developers can use the~\code{_Mirror} qualifier, which copies the function into both the region so that callers in each region will invoke the corresponding copy avoiding unnecessary tainting.
+
+\noindent\emph{\_TLIB}:
+This qualifier relaxes type-checking rules on library functions avoiding unnecessary copying of tainted data.
+Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.
+For instance, consider that in~\cregion, we want to string concat~\code{n} characters from tainted pointer~\code{t_str} to~\code{dst} (a checked pointer).
+We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
+We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then call~\code{strncpy} as~\code{strncat(dst, tmp, n);}.
+Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
+If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use~\code{strncat} with tainted source pointer in~\cregion.
+The~\code{_TLIB} qualifier enables this as shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+// Put your example annotation here.
+strncpy(dst, t_str, n);
+\end{minted}
+\aravind{Fix the above code}
+Once annotated, our type-checker allows tainted arguments to~\code{strncat} functions.
+
+%\subsection{Overview}
+%\systemname is implemented as a \systemname compiler and a \systemname source-level partitioning assistant (CheckMate).
 Given a well-typed \systemname program as an input, CheckMate performs the automation of moving the \systemname qualified function definitions to new "unchecked" source files, whilst inserting STD library headers needed for the successful compilation of these "unchecked" sources to an object file with a Sandbox-specific compiler.
@@
 Now with the \systemname qualified functions and variables re-defined under the "unchecked" library object, the \systemname qualified functions definitions in the "checked" region are modified by the CheckMate accordingly to allow communication between the two regions.
-\subsection{\systemname Function Qualifiers}
-\systemname function qualifiers are introduced with a purpose of semantic policing to allow the developer to perform dependency-resolved program partitioning into checked and unchecked regions from within the checked sources.
+%\subsection{\systemname Function Qualifiers}
+%\systemname function qualifiers are introduced with a purpose of semantic policing to allow the developer to perform dependency-resolved program partitioning into checked and unchecked regions from within the checked sources.
 Each of the newly introduced qualifiers "\_Tainted", "\_Mirror", "\_TLIB" and "\_Callback" enforce different semantic-ruling at compile time to allow "unchecked" and "checked" functions to communicate and access global data in a region-exclusive manner.

### 2023-01-13 sandp23/implementation.tex
Adding

@@
 Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.
+
 For instance, consider that in~\cregion, we want to string concat~\code{n} characters from tainted pointer~\code{t_str} to~\code{dst} (a checked pointer).
@@
 \aravind{Fix the above code}
-Once annotated, our type-checker allows tainted arguments to~\code{strncat} functions.
+The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.
 

### 2023-01-13 sandp23/formal.tex
update.

@@
 
-\myparagraph{Types}
+\myparagraph{Type Syntax}
 At a high level, we classify types as word-size value, multi-word
@@
 \myparagraph{Expressions}
-\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$)
-\mzu{($n\!:\!\tau$)}\mz{what's this?}, pointer dereference and assignment ($\estar{e}$) and ($\eassign{e_1}{e_2}$), resp.
+\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
+pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
 along with expressions that require special handling, such as,
@@
 unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
-For instance, a dynamic bounds cast expression
-\begin{center}
+For example, a dynamic bounds cast
   {\footnotesize
-    \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,
-      count(}$n$\code{))} }
-\end{center}
+    \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
 becomes 
@@
 % 
-\mzu{We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.}
-\mz{This sentence is unclear. I suppose that you mean an integer literal
-  can be both of interger type and pointer type depending on the context.  }
+We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
 For instance, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer.
-The~\code{strlen} expression plays an important role in the type-checking of null-terminated arrays.
-For simplifying type checking, we assume~\code{strlen} to operate on variables \mzs{$x$} rather than arbitrary expressions.
-% 
 The heap allocation expression $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions (\cregion or \ucregion).
 We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-The $\echeckedtext$ and $\euncheckedtext$ expressions are used to
-\mzs{represent the corresponding} \mzr{delimit} code regions.
-To guarantee the non-exposure property, we extend the base syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$.
-\mzu{Wherein $\overline{x}$ restricts all free variables appearing in $e$ to be checked or unchecked pointers according to the corresponding region.}
-\mz{Is this sentence grammatical?}
-%
+The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit code regions.
+To guarantee the non-exposure property, we extend the \checkedc syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$,
+where $\overline{x}$ represents all variables that are allowed to communicate between instructions outside and inside of the block $e$, and cannot contain checked pointers.
+$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
 
@@
 \lang{} aims to be simple enough to work with but powerful enough to
-encode realistic \systemname idioms. For example, mutable local
-  variables can be encoded as immutable locals that point to the heap;
-\mzu{the use of \code{&} can be simulated with \code{malloc};}
-\mz{This doesn't make sense if you use \code{&} to create alias. You don't want
-  to allocate memory.}
-% 
-and loops can be encoded as recursive function calls. \code{struct}s are
+encode realistic \systemname idioms. For example,
+loops can be encoded as recursive function calls. \code{struct}s are
 not included in~\fig{fig:checkc-syn} for space reasons, but they are
-supported as shown in~\apdx{appx:struct}.
+supported as shown in~\cite{li22checkedc}.
 C-style \code{union}s have no safe typing in \checkedc, so we omit them.
-
 Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
@@
 
-\aravind{Why we have only function pointers in function call?}
-\aravind{Talk about the ret expression after adding it in the syntax.}
-
 % \mwh{NOTE: COULD WORK THE FOLLOWING INTO THE ABOVE DESCRIPTION OF
@@
     \begin{array}{lll}
-e & ::= & \ldots \mid \ret{x}{\evalue{n}{\tau}}{e}\\
 r & ::= & e \mid \enull \mid \ebounds\\

### 2023-01-13 sandp23/implementation.tex
Adding implememtation

@@
 
+We created the sandbox library once and for all for each sandbox.
+This library abstracts the sandbox-specific details and exposes a uniform interface (header file) to be used in~\cregion.
+For instance,~\code{_SBX_()} gets an opaque pointer to the target sandbox.
+We will present our implementation using WASM~\cite{bosamiya2020webassembly} as our target sandbox.
+However, as we discuss in~\sect{}, our implementation can be easily extended to other sandbox mechanisms.
+
 \subsection{\sourcerewriter} \label{subsec:checkmate}
 The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{}).
-However, we use a small OCAML program (\numocamllines{}) module to remove annotations and make the code compilable with the sandbox compiler (CITE CORRESPONDING IMPL SECTION).
+However, we use a small OCAML program (\numocamllines{}) to remove annotations and make the code compilable with the sandbox compiler (CITE CORRESPONDING IMPL SECTION).
 
@@
 
+\subsubsection{Generating~\cregion Source Partition}
+We copy all non-tainted functions into~\cregion source files and make the following modifications to enable interaction with~\ucregion (\ie sandboxed code).
+
+We created a library once and for all for each sandbox.
+This library abstracts the sandbox-specific details and exposes a uniform interface (header file) to be used in~\cregion.
+For instance,~\code{_SBX_()} gets an opaque pointer to the target sandbox.
+
+\noindent\emph{Handling Tainted Buffers in~\cregion}:
+All buffers that are marked as tainted in~\cregion should be allocated in the sandbox. For instance,~\code{buff} in function~\code{handle_request} of~\lst{lst:final}.
+We perform source rewriting and make the buffer into a pointer variable and allocate the corresponding number of bytes using the sandbox-specific allocator.
+For instance, we perform the following rewriting for~\code{buff} in~\lst{lst:final} with WASM sandbox.
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+// char buff[MAX_MSG_SIZE] __Tainted;
+char *buff = ....
+\end{minted}
+\aravind{Arun, please fix the above example on how exactly we replace it with WASM allocator.}
+Note that we also add the necessary deallocator calls at function exit points.
+
+\noindent\emph{Handling Calls to Tainted Functions}:
+In~\cregion, we also need to modify calls to tainted functions as they execute inside the sandbox (separate address space) and thus cannot be invoked as regular functions.
 %\subsection{Overview}

### 2023-01-13 sandp23/limitations.tex
Adding implememtation

\section{Limitations}
\label{sec:limitations}
We cannot handling function pointer calls to tainted function.

### 2023-01-14 sandp23/implementation.tex
Minor

@@
 For instance, we perform the following rewriting for~\code{buff} in~\lst{lst:final} with WASM sandbox.
-\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-// char buff[MAX_MSG_SIZE] __Tainted;
-char *buff = ....
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
+- char buff[MAX_MSG_SIZE] __Tainted;
++ char *buff = ....
 \end{minted}
@@
 In~\cregion, we also need to modify calls to tainted functions as they execute inside the sandbox (separate address space) and thus cannot be invoked as regular functions.
+We rewrite all calls to tainted functions with corresponding sandbox invocations.
+For instance, we modify the call to~\code{process_req1} in~\lst{lst:final} as below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
+case REQ1: 
+- rc = process_req1(buff); 
++ rc = .....
+\end{minted}
+\aravind{Arun, please fix the above example with the exact details.}
 %\subsection{Overview}

### 2023-01-14 sandp23/implementation.tex
Minor

@@
 \subsubsection{Additional Function Qualifiers}
+\label{subsubsec:addfuncqual}
 In addition to the~\code{_Tainted} qualifier that marks functions to be in~\ucregion, we provide a few other qualifiers that enable developers to provide additional information and ease the partitioning process.
@@
 \aravind{Arun, please fix the above example with the exact details.}
+
+\noindent\emph{Handling~\texttt{} Qualifiers}:
+As mentioned in~\sect{subsubsec:addfuncqual}, functions with these qualifiers can be called from~\ucregion.
+Consider the following~\code{StringAuth} function that checks whether the provided user is authenticated and needs to be invoked from~\ucregion and hence a callback.
+
+Note that the function accepts a tainted array pointer as its parameter, enabling~\ucregion (tainted) region to invoke it.
+
+For each such function, we create a corresponding trampoline function that serves as the entry point for the callback function, as shown below:
+
+The trampoline function performs the necessary conversion of pointer arguments and eventually invokes the callback.
+
+We also add the code to register this trampoline function with the sandbox. The registration function for WASM sandbox is as shown below:
+
+This registration function creates an opaque handle for the trampoline function and enables~\ucregion to call the trampoline using the corresponding handle.
+
+Lastly, we change all instances where the callback function pointer is given to tainted functions to use the handle for the corresponding trampoline a shown below:
+
 %\subsection{Overview}

### 2023-01-14 sandp23/implementation.tex
Minor

@@
 
-\noindent\emph{Handling~\texttt{} Qualifiers}:
+\noindent\emph{Handling~\texttt{\_Callback} Qualifiers}:
 As mentioned in~\sect{subsubsec:addfuncqual}, functions with these qualifiers can be called from~\ucregion.
 Consider the following~\code{StringAuth} function that checks whether the provided user is authenticated and needs to be invoked from~\ucregion and hence a callback.
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+_Callback _TPtr<char> StringAuth(_T_Array_Ptr<const char> usertoken : count(len),
+                                 size_t len) {
+...
+// Checks whether usertoken is authenticated
+/*
+ These functions will be restricted to only accept
+ tainted parameters.
+*/
+...
+}
+\end{minted}
 

### 2023-01-14 sandp23/implementation.tex
Minor

@@
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-_Callback _TPtr<char> StringAuth(_T_Array_Ptr<const char> usertoken : count(len),
+_Callback _TPtr<char> StringAuth(
+               _T_Array_Ptr<const char> usertoken : count(len),
                size_t len) {
@@
 \end{minted}
-
-Note that the function accepts a tainted array pointer as its parameter, enabling~\ucregion (tainted) region to invoke it.
+These callback functions are only allowed to use tainted parameters as they will be called from a tainted region.
 
 For each such function, we create a corresponding trampoline function that serves as the entry point for the callback function, as shown below:
-
-The trampoline function performs the necessary conversion of pointer arguments and eventually invokes the callback.
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
++ unsigned int _T_StringAuth(unsigned sandbox,
++               unsigned int arg_1,
++               unsigned long int arg_2) {
++   // Perform necessary conversion of arguments.
++   // uname <- conver arg_1
++   // len <- arg_2
++   ret = StringAuth(uname, len);
++   // ret_val <- ret
++   return ret_val;
++ }
+\end{minted}
+The trampoline function handles the invocations from sandbox (and hence the extra parameter~\code{sandbox}), performs the necessary conversion of pointer arguments, and eventually invokes the callback.
 
 We also add the code to register this trampoline function with the sandbox. The registration function for WASM sandbox is as shown below:
-
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
++ void registerCallback_StringAuth(void){
++ //callback function signature {ret <- int, arg_1 <- int, arg_2 <- long}
++ int ret_param_types[] = {0, 0, 1};
++ // 2 <- arg count, 1 <- ret count
++ __StringAuth__C_ = _SBXREG_((void*)_T_StringAuth,2,1, ret_param_types);
++ }
+\end{minted}
 This registration function creates an opaque handle for the trampoline function and enables~\ucregion to call the trampoline using the corresponding handle.
@@
 Lastly, we change all instances where the callback function pointer is given to tainted functions to use the handle for the corresponding trampoline a shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
+// How do we change the call-sites where the function pointer is passed
+// as the call-back.
+\end{minted}
+\aravind{Arun, please check above code example to see if they make sense.}
+
+\subsubsection{Generating~\ucregion Source Partition}
 

### 2023-01-14 sandp23/implementation.tex
Adding implementation

@@
 For instance,~\code{_SBX_()} gets an opaque pointer to the target sandbox.
-We will present our implementation using WASM~\cite{bosamiya2020webassembly} as our target sandbox.
+We will present our implementation using WebAssembly (WASM~\cite{bosamiya2020webassembly}) as our target sandbox.
 However, as we discuss in~\sect{}, our implementation can be easily extended to other sandbox mechanisms.
@@
 The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{}).
-However, we use a small OCAML program (\numocamllines{}) to remove annotations and make the code compilable with the sandbox compiler (CITE CORRESPONDING IMPL SECTION).
+However, we use a small OCAML program (\numocamllines{}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
 
@@
 \subsubsection{Generating~\cregion Source Partition}
+\label{subsub:gencregion}
 We copy all non-tainted functions into~\cregion source files and make the following modifications to enable interaction with~\ucregion (\ie sandboxed code).
@@
 \subsubsection{Generating~\ucregion Source Partition}
+\label{subsubsec:genucregion}
+We create~\ucregion partition by first copying all functions marked as~\code{_Tained} and~\code{_Mirror}.
+Next, we convert the partition into regular C by replacing all~\systemname{} tainted types with the corresponding C types.
+For instance:~\code{_T_Array_ptr<char>} will be replaced with~\code{char *}.
+The~\code{read_msg} function signature in~\lst{} will be modified as shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+size_t read_msg(int sock_fd, char *msg,
+-                             size_t sz) __Tainted {
++                             size_t sz) {
+\end{minted}
+
+Finally, we convert all calls to~\cregion callbacks with sandbox-specific invocations.
+For instance, for the~\code{StringAuth} call back (\sect{subsub:gencregion}), we modify~\code{...UNCHECKED FUNC..} as shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
+SAY HOW WE MODIFY UNCHECKED CODE
+TO INVOKE CALLBACK FUNCTIONS.
+\end{minted}
+\aravind{Arun, please fix the above example}
 
@@
 
-\subsection{CheckMate S2S transformation}
-\subsubsection{\textbf{Need for Callback}}
-Here, we use the example of a simple \systemname annotated string authentication program snippet in ~\lst{lst:callbackBefore} with~\code{StringProc} as its entry point (\entrypoint).
+%\subsection{CheckMate S2S transformation}
+%\subsubsection{\textbf{Need for Callback}}
+%Here, we use the example of a simple \systemname annotated string authentication program snippet in ~\lst{lst:callbackBefore} with~\code{StringProc} as its entry point (\entrypoint).
 ~\code{StringProc} was found to internally perform a lot of unsafe string parsing and manipulation on the user input and thereby has been marked \_Tainted.
@@
 In interests of maintaining confidentiality of global checked data from the unchecked region whilst satisfying this control and data dependency, ~\code{StringAuth} has been qualified as a \_Callback, thereby bestowing it with WASM Sandbox's addition sanity checks at its call-sites within the unchecked region.
-
-\subsubsection{\textbf{CheckMate S2S transformation on Checked Sources}}
-\myparagraph{WebAssembly Sandbox:}
-Unchecked region implemented as a WebAssembly sandbox requires registering ~\code{StringAuth} with its expected WebAssembly callee signature as a requirement to safeguard the control-flow between the two regions.
-Within the WebAssembly sandbox, all functions are defined with an extra sandbox pointer argument and all pointer members are unsigned integers.
-
-\myparagraph{CheckMate Instrumentation: }
-Consequently, ~\code{StringAuth}'s corresponding call-site signature in the tainted region would be 
-\newline
-~\code{_T_StringAuth}, which is an indirected trampoline call to the actual callback function.
+%
+%\subsubsection{\textbf{CheckMate S2S transformation on Checked Sources}}
+%\myparagraph{WebAssembly Sandbox:}
+%Unchecked region implemented as a WebAssembly sandbox requires registering ~\code{StringAuth} with its expected WebAssembly callee signature as a requirement to safeguard the control-flow between the two regions.
+%Within the WebAssembly sandbox, all functions are defined with an extra sandbox pointer argument and all pointer members are unsigned integers. 
+%
+%\myparagraph{CheckMate Instrumentation: }
+%Consequently, ~\code{StringAuth}'s corresponding call-site signature in the tainted region would be 
+%\newline
+%~\code{_T_StringAuth}, which is an indirected trampoline call to the actual callback function.
 CheckMate identifies ~\code{StringAuth} to be a \_Callback qualified callee and correspondingly generates callback trampoline ~\code{StringAuth} and callback registration code
-\newline
-~\code{registerCallback_StringAuth} within the checked region as show in ~\lst{lst:callbackAfter}. 
-~\code{registerCallback_StringAuth} registers the WebAssembly-specific callback function ~\code{_T_StringAuth} by calling ~\code{\_SBXREG\_} with callback-specific argument count and argument signature.
-CheckMate replaces \_Tainted qualified ~\code{StringProc}'s definition with an indirect call to sandbox specific implementation ~\code{w2c_StringProc}.
+%\newline
+%~\code{registerCallback_StringAuth} within the checked region as show in ~\lst{lst:callbackAfter}. 
+%~\code{registerCallback_StringAuth} registers the WebAssembly-specific callback function ~\code{_T_StringAuth} by calling ~\code{\_SBXREG\_} with callback-specific argument count and argument signature.
+%CheckMate replaces \_Tainted qualified ~\code{StringProc}'s definition with an indirect call to sandbox specific implementation ~\code{w2c_StringProc}.
 The \_Mirror qualifier is used mark a function in checked region as a local to both the checked and the unchecked regions.
 Consequently, CheckMate does not modify \_Mirror qualified ~\code{isStringMatching} and simply duplicates this function in the unchecked region.
-
-\subsubsection{\textbf{CheckMate S2S generation of Tainted Sources}}
-Implementing \systemname's unchecked region as a WebAssembly sandbox requires compiling unchecked functions and unchecked variables checked-region-exclusively with a sandbox-specific compiler (WebAssembly).
+%
+%\subsubsection{\textbf{CheckMate S2S generation of Tainted Sources}}
+%Implementing \systemname's unchecked region as a WebAssembly sandbox requires compiling unchecked functions and unchecked variables checked-region-exclusively with a sandbox-specific compiler (WebAssembly).
 Consequently, CheckMate copies each of the \_Mirror and \_Tainted qualified function definitions to a new source file and runs an OCAML parser to convert \systemname code to generic-C as shown in ~\lst{lst:callbackAfterTainted} which shows the CheckMate generated code for ~\code{StringProc}.
-After running CheckMate, each of newly generated tainted sources can be compiled to an "unchecked" static library using a sandbox-specific compiler.
-CheckMate's OCAML parser also modifies each of the tainted STDLIB functions from ~\code{t_strncpy} under ~\lst{lst:callbackBefore} to ~\code{strncpy} under ~\lst{lst:callbackAfterTainted}
-
-\begin{listing}
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_before_checkmate.c}
-\caption{Callback Annotations before running CheckMate}
-\label{lst:callbackBefore}
-\end{listing}
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate.c}
-\caption{CheckMate generated Callback Annotations in Checked Sources }
-\label{lst:callbackAfter}
-\end{listing}
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate_tainted.c}
-\caption{CheckMate generated Unchecked Sources }
-\label{lst:callbackAfterTainted}
-\end{listing}
+%After running CheckMate, each of newly generated tainted sources can be compiled to an "unchecked" static library using a sandbox-specific compiler.
+%CheckMate's OCAML parser also modifies each of the tainted STDLIB functions from ~\code{t_strncpy} under ~\lst{lst:callbackBefore} to ~\code{strncpy} under ~\lst{lst:callbackAfterTainted}
+%
+%\begin{listing}
+%\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_before_checkmate.c}
+%\caption{Callback Annotations before running CheckMate}
+%\label{lst:callbackBefore}
+%\end{listing}
+%\begin{listing}[t!]
+%\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate.c}
+%\caption{CheckMate generated Callback Annotations in Checked Sources }
+%\label{lst:callbackAfter}
+%\end{listing}
+%\begin{listing}[t!]
+%\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate_tainted.c}
+%\caption{CheckMate generated Unchecked Sources }
+%\label{lst:callbackAfterTainted}
+%\end{listing}
 

### 2023-01-14 sandp23/evaluation.tex

@@
-\section{\systemname Implementation} \label{sec:implementation}
-\subsection{Overview}
-\systemname is implemented as a \systemname compiler and a \systemname source-level partitioning assistant (CheckMate).
-Given a well-typed \systemname program as an input, CheckMate performs the automation of moving the \systemname qualified function definitions to new "unchecked" source files, whilst inserting STD library headers needed for the successful compilation of these "unchecked" sources to an object file with a Sandbox-specific compiler.
-Since \systemname is irrelevant to the Sandbox-specific compiler, CheckMate internally runs an OCAML translator to cleanse away the \systemname annotations by converting these tainted definitions to generic-C.
-Now with the \systemname qualified functions and variables re-defined under the "unchecked" library object, the \systemname qualified functions definitions in the "checked" region are modified by the CheckMate accordingly to allow communication between the two regions.
-\subsection{\systemname Function Qualifiers}
-\systemname function qualifiers are introduced with a purpose of semantic policing to allow the developer to perform dependency-resolved program partitioning into checked and unchecked regions from within the checked sources.
-Each of the newly introduced qualifiers "\_Tainted", "\_Mirror", "\_TLIB" and "\_Callback" enforce different semantic-ruling at compile time to allow "unchecked" and "checked" functions to communicate and access global data in a region-exclusive manner.
-Although checked functions are allowed to call and access "unchecked" data within the checked region, "unchecked" region can only access checked data through a specialized callback mechanism.
-Each of these function qualifiers differently control casting, caller, callee and data-access rules within its qualifier-scope.
-For example, \_TLIB qualifier allows tainted pointer parameters to be cast to non-tainted pointer arguments at the call-site of a qualified function with developer's discretion that the qualified function does not leak tainted addresses.
-This type-checker relaxation is granted to enable developer to avoid marshalling costs when passing tainted pointers to external library functions.
-
-\subsection{CheckMate S2S transformation}
-\subsubsection{\textbf{Need for Callback}}
-Here, we use the example of a simple \systemname annotated string authentication program snippet in ~\lst{lst:callbackBefore} with~\code{StringProc} as its entry point (\entrypoint).
-~\code{StringProc} was found to internally perform a lot of unsafe string parsing and manipulation on the user input and thereby has been marked \_Tainted.
-However, ~\code{StringProc} internally calls ~\code{StringAuth} that performs final authentication on user input by referencing classified global checked data.
-In interests of maintaining confidentiality of global checked data from the unchecked region whilst satisfying this control and data dependency, ~\code{StringAuth} has been qualified as a \_Callback, thereby bestowing it with WASM Sandbox's addition sanity checks at its call-sites within the unchecked region.
-
-\subsubsection{\textbf{CheckMate S2S transformation on Checked Sources}}
-\myparagraph{WebAssembly Sandbox:}
-Unchecked region implemented as a WebAssembly sandbox requires registering ~\code{StringAuth} with its expected WebAssembly callee signature as a requirement to safeguard the control-flow between the two regions.
-Within the WebAssembly sandbox, all functions are defined with an extra sandbox pointer argument and all pointer members are unsigned integers.
-
-\myparagraph{CheckMate Instrumentation: }
-Consequently, ~\code{StringAuth}'s corresponding call-site signature in the tainted region would be 
-\newline
-~\code{_T_StringAuth}, which is an indirected trampoline call to the actual callback function.
-CheckMate identifies ~\code{StringAuth} to be a \_Callback qualified callee and correspondingly generates callback trampoline ~\code{StringAuth} and callback registration code
-\newline
-~\code{registerCallback_StringAuth} within the checked region as show in ~\lst{lst:callbackAfter}. 
-~\code{registerCallback_StringAuth} registers the WebAssembly-specific callback function ~\code{_T_StringAuth} by calling ~\code{\_SBXREG\_} with callback-specific argument count and argument signature.
-CheckMate replaces \_Tainted qualified ~\code{StringProc}'s definition with an indirect call to sandbox specific implementation ~\code{w2c_StringProc}.
-The \_Mirror qualifier is used mark a function in checked region as a local to both the checked and the unchecked regions.
-Consequently, CheckMate does not modify \_Mirror qualified ~\code{isStringMatching} and simply duplicates this function in the unchecked region.
-
-\subsubsection{\textbf{CheckMate S2S generation of Tainted Sources}}
-Implementing \systemname's unchecked region as a WebAssembly sandbox requires compiling unchecked functions and unchecked variables checked-region-exclusively with a sandbox-specific compiler (WebAssembly).
-Consequently, CheckMate copies each of the \_Mirror and \_Tainted qualified function definitions to a new source file and runs an OCAML parser to convert \systemname code to generic-C as shown in ~\lst{lst:callbackAfterTainted} which shows the CheckMate generated code for ~\code{StringProc}.
-After running CheckMate, each of newly generated tainted sources can be compiled to an "unchecked" static library using a sandbox-specific compiler.
-CheckMate's OCAML parser also modifies each of the tainted STDLIB functions from ~\code{t_strncpy} under ~\lst{lst:callbackBefore} to ~\code{strncpy} under ~\lst{lst:callbackAfterTainted}
-
-\begin{listing}
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_before_checkmate.c}
-\caption{Callback Annotations before running CheckMate}
-\label{lst:callbackBefore}
-\end{listing}
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate.c}
-\caption{CheckMate generated Callback Annotations in Checked Sources }
-\label{lst:callbackAfter}
-\end{listing}
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/callback_mechanism_sample_after_checkmate_tainted.c}
-\caption{CheckMate generated Unchecked Sources }
-\label{lst:callbackAfterTainted}
-\end{listing}
-
-\subsection{\systemname Dynamic Instrumentation}
-\systemname compiler is implemented to be sandbox-agnostic and differs in inserted instrumentation based on the sandbox type specified by a compiler flag.
-However, the scope of this paper restricts discussion to the 32-Bit WebAssembly sandbox wherein which pointers are represented as 4-Byte offsets from a base sandbox heap.
-Similar analogy has been applied in \systemname for WebAssembly wherein all tainted pointers in the checked region store zero-extended offsets, that upon access, are transformed to full addresses by the function \ref{O2P}.
-\subsubsection{\textbf{Overview}}
-\systemname compiler inserts run-time instrumentation on every access to a tainted pointer to assert pointer taintedness and to generate dereferencible address from the pointer offset.
-Pointer taintedness is an invariant that prevents checked memory addresses from leaking into the tainted region by establishing the tainted pointer to only point into the Sandbox memory.
-Since WebAssembly sandbox is implemented in 32-Bit, specialized struct type is created and instrumentation is inserted by the \systemname compiler at every pointer access to account for pointer-swizzling across different pointer representations.
-
-\subsubsection{\textbf{Pointer-swizzling}}
-WebAssmebly implementation of \systemname motivates Tainted pointers in checked region to implemented as a zero-extended 32-bit offset into the Sandbox.
-Every access to the tainted pointer calls for a pointer generation mechanism that generates a fully de-referencible pointer address as
-shown by the equation\ref{O2P}.
-
-Since \systemname does not require qualified functions to mandatorily execute in the unchecked region, tainted pointers may never have to cross into the sandbox.
-In such cases, we need not swizzle and the tainted pointers are implemented as 64-Bit pointers that contain zero extended 4-Byte pointer offsets into the Sandbox memory.
-However, Need for swizzling emerges when indexing into multi-level tainted pointers or tainted pointers to structures that are passed into the sandbox region.
-“\_Decoy” is a pointer variable qualifier that informs the compiler to enforce 64-Bit to 32-bit pointer swizzling throughout the qualified variable's lifetime in the checked region.
-Pointer-Swizzling is implemented in Clang-CodeGen as a modified GEP-instruction that considers \_Decoy when generating swizzled addresses causing \_Decoy qualified variables to be accessed as 32-bit pointers in the checked region.
-\begin{equation}
-  \_T\_ptr\_ADDR = \_T\_ptr\_OFFSET + SBX\_HEAP\_BASE
-\label{O2P}
-\end{equation}
-
-\subsubsection{\textbf{Tainted Structures}}
-Tainted Pointers to structures crossing into the sandbox are expected to only allow encapsulation of pointer members that are of the tainted type.
-Consequently, Tstructs have been introduced as a specialized struct type that enforces compile-time semantics to allow prevent its pointer members from being of non-tainted type.
-\systemname also restricts tainted pointers to only point to structures of Tstruct type.
-These semantic restrictions of Tstructs are put in place to prevent checked pointer from leaking into the unchecked regions as members of pointer structures.
-Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout.
-
 \section{Evaluation}\label{sec:evaluation}

### 2023-01-14 sandp23/formal.tex

@@
 
-\subsection{Type Checking}
+\subsection{Typing and Semantics}
 \label{sec:typechecking}
+
+The \lang system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
 Our type checker restricts the usage of tainted and checked pointer types to ensure that tainted pointers do not affect checked types, along with enforcing~\checkedc typing rules~\cite{li22checkedc}.
-A few selected type rules are presented in~\fig{fig:type-system-1}.
-% 
-\mz{Move the figure here.}
-% 
-The complete set of typing rules and special handling of (NT)-arrays are provided in~\apdx{sec:literal-pointer-typing}.
-
+
+As partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{T-}$X$),
 Each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
@@
 in context mode $m$.
-
-In \lang, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
-We provide some example descriptions here.
-Type equality $\tau=_{\Theta}\tau'$
-is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
-and the alpha equivalence of two function types;
-i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
-$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
-$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
-are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
-and the substitution results are equal.
-\aravind{For the above paragraph, give concrete example, For example: array is a subtype of ptr, etc}
-
-The type checker rules are designed such that any well-typed (\ie follows all the rules) is isolated by sandboxed code. We present the corresponding formal proof in~\apdx{}.
-
-One of the important type cheker rules is disallowing unsafe interactions between tainted and untainted (c/u) pointers.
-\aravind{Explain about the LET rule}
-%
-Rules \textsc{T-Let} and \textsc{T-LetInt} in in \Cref{fig:type-system-1} type a $\elettext$ expression, which also admits
-type dependency. 
-In particular, the result of evaluating a $\elettext$ expression
-may have a type that refers to one of its bound variables (e.g., if
-the result is a checked pointer with a variable-defined bound). 
-If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
-Note that we restrict the expression $e_1$ to syntactically match the
-structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
-Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
-$\erettext$ does not appear in source programs but is introduced by the semantics when
-evaluating a let binding (rule \textsc{S-Let} in
-Fig.~\ref{fig:semantics}). 
-%\liyi{why? }
-% After the evaluation of a let binding a variable $x$ concludes,
-%we need to restore any prior binding of $x$, which is either
-%$\bot$ (meaning that there is no $x$ originally) or some value
-%$\evalue{n}{\tau}$.
-
-\aravind{We should explain how our type checker restricts unsafe interactions between checked and tainted pointer types.}
+The operational semantics for \lang is defined as a small-step
+transition relation with the judgment $ (\varphi,\heap,e)
+\longrightarrow_m (\varphi',\heap',r)$, as partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{S-}$X$).
+ Here, $\varphi$ is a
+\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
+$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ heaps), each of which
+maps addresses (integer literals) to values $\evalue{n}{\tau}$.
+The complete set of typing rules and special handling of (NT)-arrays are provided in~\apdx{sec:literal-pointer-typing}.
+
+\myparagraph{Modes, Static Casting, and Subtyping}
+In \lang, Context modes $m$ appearing in a type rule determine the code region 
+that permits pointer dereferences and value-assignments, depending on the pointer modes.
+We define a three point lattice $\xi_1 \le \xi_2$ to describe such permission, where $\tmode \le \xi$ and $m \le m$.
+This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can perform such operations in $\cmode$ and $\umode$ regions, respectively.
+Pointer modes are also useful in determining if a nested pointer has a valid type.
+For example, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, we require $\xi_2\le \xi_1$ to maintain non-exposure.
+
+\lang also has static casting operations. As described in rule \textsc{T-CastPtr},
+an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
+if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2} $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
+In \lang, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
 
 \begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-a -> checked pointer
-b -> Tainted pointer
-c -> Unchecked pointer.
-
-_Ptr<int> a;
-_t_Ptr<int> b;
-int *c;
-
-a = b; // Not allowed.
-b = a; // Not allowed.
+//_Ptr<int> a; _t_Ptr<int> b; int *c;
+c = (int *)b; // This is okay.
+a = (_Ptr<int>)b; // Not allowed.
 \end{minted}
 
-The \textsc{T-CastPtr} rule in \Cref{fig:type-system-1}
-permits casting from an expression of type $\tau'$ to a checked pointer when
-$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
-$\sqsubseteq$ is built on the type equality ($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). 
-The rule  ($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
-\tarrayptr{b_l}{b_h}{\tau}{m}$) permits treating a singleton
+In the above example, a $\tmode$ mode pointer can be cast to $\umode$ mode but casting $\tmode$ mode to $\cmode$ mode is disallowed.
+The complete subtyping relation was described in \Cref{fig:checkc-subtype}.
+\ignore{
+For example, the rule ($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{\xi}\sqsubseteq
+\tarrayptr{b_l}{b_h}{\tau}{\xi}$) permits treating a singleton
 pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
 Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
-if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
-
-\aravind{Explain how the cast rule allows/prevents the casts.}
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).}
+Notice that \texttt{let} statements are immutable in \lang, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \lang.
 
 \begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-a -> checked pointer
-b -> Tainted pointer
-c -> Unchecked pointer.
-
-_Ptr<int> a;
-_t_Ptr<int> b;
-int *c;
-
-a = b; // Not allowed.
-b = a; // Not allowed.
-
-a = (_Ptr<int>)c; // This is okay.
-a = (_Ptr<int>)b; // Not allowed.
+//_Ptr<int> x;  _t_Ptr<int> y;
+x = y; // Not allowed.
 \end{minted}

### 2023-01-14 sandp23/appendix.tex
update.

@@
 
-    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
+    \path [->] (A) edge node {$\longrightarrow_{\cmode}$} (B);
     \path [<->] (A) edge node {$\gg$} (C);
@@
 The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.
+
+\myparagraph{Constant Validity}
+Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
+describe type assumptions for constants appearing in a program.
+$\neg \cmode(\tau)$ judges that a constant pointer 
+in an unchecked region cannot be of a checked type.
+The restriction ensures that programmers 
+cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+In rule \textsc{T-ConstC}, we requires a static 
+verification procedure for validating a constant pointer in \Cref{fig:const-type}.
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
+validates the constant $\evalue{n}{\tau}$, 
+where $\heap(m)$ is the initial heap that the constant resides on and
+$\sigma$ is a set of constant assumed to be checked.
+A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
+A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
+and the address stores a function with the right type.
+The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
+where every element in the pointer range ($[0,\size(\omega))$) should be well
+typed.
+A checked pointer checks validity in type step as rule \textsc{T-ConstC},
+while a tainted/unchecked pointer does not check for such during the type checking.
+Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.

### 2023-01-14 sandp23/formal.tex
update.

@@
 
+\begin{DIFnomarkup}
+\begin{figure*}[t]
+{\small
+  \begin{mathpar}
+     \inferrule[T-CastPtr]
+               {\Gamma;\Theta \vdash_m e : \tau' \\
+                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
+               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
+
+    \inferrule[S-Cast]
+              {}
+              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
+
+\\\\
+    \inferrule[T-Def]
+              {\xi \leq m \\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
+              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
+\quad
+        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+\quad
+        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
+         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+\quad
+    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
+
+    \inferrule[T-Checked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
+              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
+
+    \inferrule[T-Unchecked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
+              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
+
+    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
+\inferrule[T-Fun]
+    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
+        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
+         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
+         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
+             \overline{\tau'} \sqsubseteq_{\Theta}
+               \overline{\tau}[\overline{e'} / \overline{x}]}
+    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
+
+    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
+        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+\quad
+    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
+                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
+        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+
+  \end{mathpar}
+}
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
+% \\[0.2em]
+% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
+% \end{array}
+% $
+% \end{center}
+% }
+{\footnotesize
+$
+\cmode(\tint)=\texttt{false}
+\qquad
+\cmode(\tptr{\omega}{\cmode})=\texttt{true}
+\qquad
+\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
+$
+}
+\caption{Selected type rules}
+\label{fig:type-system-1}
+\end{figure*}
+\end{DIFnomarkup}
+
 The \lang system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
@@
 
+We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ heap,
+and $\heapup{m}{n}{\evalue{n'}{\tau}}$ 
+to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ heap.
+While heap bindings can change, stack bindings are immutable---once
+variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
+be updated. 
+%We can model mutable stack variables as pointers into the mutable heap.
+As mentioned, value $\evalue{0}{\tau}$
+represents a $\enull$ pointer when $\tau$ is a pointer type.
+Correspondingly, $\heap(m,0)$ should always be undefined.
+% 
+The relation steps to a \emph{result} $r$, which is either an
+expression, a $\enull$ or $\ebounds$ failure, represent an expression right
+  after the reduction, a null-pointer dereference or out-of-bounds access,
+respectively.
+% 
+Such failures are a \emph{good} outcome; stuck states
+(non-value expressions that cannot transition to a result $r$)
+characterizing undefined behavior.
+%The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the stepped redex within $e$ was in a $\cmode$ or $\umode$ region.
+
+The rules for the main operational semantics
+judgment \emph{evaluation} are given at the bottom of
+Fig.~\ref{fig:c-context}.
+The first rule takes an expression $e$, decomposes
+it into an \emph{evaluation context} $E$ and a sub-expression $e'$
+(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
+$e$), and then evaluates $e'$ according to the \emph{computation}
+  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
+whose rules are given along with type rules in Fig.~\ref{fig:type-system-1}, discussed
+shortly.
+The $\mode$ function in Fig.~\ref{fig:c-context}
+determines the context mode that the expression $e'$ locates based on the context $E$.
+In \Cref{lst:humantaint}, the function call expression \code{read_msg} has $\umode$ mode since it is inside a tainted function.
+The second rule describes the exception handling 
+for possible crashing behaviors in unchecked regions.
+A $\umode$ mode operation can non-deterministically crash
+and the \systemname sandbox mechanism recovers
+the program to a safe point ($\evalue{0}{\tau}$)
+and continues with the existing program state.
+Evaluation contexts $E$ define a standard left-to-right evaluation order. 
+%(We explain the $\ret{x}{\mu}{e}$ syntax shortly.)
+%There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
+
 \myparagraph{Modes, Static Casting, and Subtyping}
@@
 that permits pointer dereferences and value-assignments, depending on the pointer modes.
-We define a three point lattice $\xi_1 \le \xi_2$ to describe such permission, where $\tmode \le \xi$ and $m \le m$.
+We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ since $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
 This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can perform such operations in $\cmode$ and $\umode$ regions, respectively.
-Pointer modes are also useful in determining if a nested pointer has a valid type.
-For example, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, we require $\xi_2\le \xi_1$ to maintain non-exposure.
-
-\lang also has static casting operations. As described in rule \textsc{T-CastPtr},
+%Pointer modes are also useful in determining if a nested pointer has a valid type.
+For example, in a nested pointer $\tptr{(...
+\tptr{\tau}{\xi_2} ...)}{\xi_1}$, we require $\xi_2\le \xi_1$ to maintain non-exposure.
+
+\lang also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
 an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
@@
 \begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-//_Ptr<int> a; _t_Ptr<int> b; int *c;
-c = (int *)b; // This is okay.
-a = (_Ptr<int>)b; // Not allowed.
+//_Ptr<int> x; _t_Ptr<int> y; int *z;
+z = (int *)y; // This is okay.
+x = (_Ptr<int>)y; // Not allowed.
 \end{minted}
@@
 The complete subtyping relation was described in \Cref{fig:checkc-subtype}.
-\ignore{
-For example, the rule ($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{\xi}\sqsubseteq
-\tarrayptr{b_l}{b_h}{\tau}{\xi}$) permits treating a singleton
-pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
-Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
-if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).}
 Notice that \texttt{let} statements are immutable in \lang, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \lang.
@@
 
-\myparagraph{Constant Validity}
-Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
-describe type assumptions for constants appearing in a program.
-$\neg \cmode(\tau)$ judges that a constant pointer 
-in an unchecked region cannot be of a checked type.
-The restriction ensures that programmers 
-cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
-In rule \textsc{T-ConstC}, we requires a static 
-verification procedure for validating a constant pointer in \Cref{fig:const-type}.
-
-The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
-validates the constant $\evalue{n}{\tau}$, 
-where $\heap(m)$ is the initial heap that the constant resides on and
-$\sigma$ is a set of constant assumed to be checked.
-A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
-A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
-and the address stores a function with the right type.
-The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
-where every element in the pointer range ($[0,\size(\omega))$) should be well
-typed.
-A checked pointer checks validity in type step as rule \textsc{T-ConstC},
-while a tainted/unchecked pointer does not check for such during the type checking.
-Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.
+\myparagraph{Pointer Dereference}
+The type and semantic rules for pointer dereference (\textsc{T-Def}, \textsc{S-DefC},
+\textsc{S-DefT}, \textsc{S-DefNull} in \Cref{fig:type-system-1})
+reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in evaluation stage.
+The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
+With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a \enull pointer is used,
+\lang captures the runtime error.
+Type and semantic rules for array types and pointer assignments are given in \Cref{appx:add-type-sem}.
+
+Rules \textsc{S-DefC} and \textsc{S-DefT} are for $\cmode$ and $\tmode$ mode pointer dereference, respectively.
+In addition to the no \enull check in $\cmode$ mode pointer dereference,
+any dynamic heap access of a tainted pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
 
@@
 
-\systemname{} has dependent functions, whose semantic explaination is given in \Cref{appx:add-type-sem}.
+\systemname{} has dependent functions, whose semantic explanation is given in \Cref{appx:add-type-sem}.
 Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
@@
 %There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
-
-\begin{DIFnomarkup}
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
-         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-\quad
-    \inferrule[S-Cast]
-              {}
-              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
-
-    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{\evalue{n'}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n'}{\tau'})}
-
-        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e})}
-
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
-    \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}
-
-    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
-    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-\quad
-    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
-                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-
- % \inferrule[S-IfNTNot]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}} \\ \heap(n)\neq 0\\ 0 < n_h}
- %            {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
-
-\end{mathpar}
-}
-% {\footnotesize
-% \begin{center}
-% $
-% \begin{array}{l}
-% \tau[\overline{n} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall n_i\in\overline{n}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \Rightarrow \tau[n_i / x_i]\\[0.2em]
-% \mathtt{let}\;\overline{x}=\overline{e}\;\mathtt{in}...\triangleq \mathtt{let}\;x_0=e_0\;\mathtt{in}\;\mathtt{let}\;x_1=e_1\;\mathtt{in}...
-% \end{array}
-% $
-% \end{center}
-% }
-\caption{\lang Semantics: Computation (Selected Rules)}
-\label{fig:semantics}
-\end{figure*}
-\end{DIFnomarkup}
 
@@
 \begin{DIFnomarkup}
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-   \inferrule[T-ConstU]
-       { \neg \cmode(\tau)}
-       {\Gamma;\Theta\vdash_u \evalue{n}{\tau} : \tau}
-
-   \inferrule[T-ConstC]
-       {\Theta;\heap;\emptyset \vdash_c n : \tau}
-       {\Gamma;\Theta\vdash_c \evalue{n}{\tau} : \tau}
-
-    \inferrule[T-Def]
-              {\xi \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-
-     \inferrule[T-CastPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
-
-   \inferrule[T-Let]
-    { x\not\in \fv(\tau') \\
-        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
-          \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
-             }
-    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'}
-
-   \inferrule[T-LetInt]
-    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\
-        \Gamma;\Theta \vdash_m e_1 : \tint \\\\
-           \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
-             }
-    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
-\qquad
-   \inferrule[T-RetInt]
-    { \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
-    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tint}}{e} : \tau}
-
-    \inferrule[T-Checked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
-              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
-
-    \inferrule[T-Unchecked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
-
-\inferrule[T-Fun]
-    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
-         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
-         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
-             \overline{\tau'} \sqsubseteq_{\Theta}
-               \overline{\tau}[\overline{e'} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
-
-  \end{mathpar}
-}
-% {\footnotesize
-% \begin{center}
-% $
-% \begin{array}{l}
-% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
-% \\[0.2em]
-% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
-% \end{array}
-% $
-% \end{center}
-% }
-{\footnotesize
-$
-\cmode(\tint)=\texttt{false}
-\qquad
-\cmode(\tptr{\omega}{\cmode})=\texttt{true}
-\qquad
-\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
-$
-}
-\caption{Selected type rules}
-\label{fig:type-system-1}
-\end{figure*}
-\end{DIFnomarkup}
-
-\begin{DIFnomarkup}
  \begin{figure}[t]

### 2023-01-14 sandp23/formal.tex
update.

@@
 \myparagraph{Unchecked and Checked Blocks}
-%
-During the type checking,
-Both $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$
-check all free variables in $e$ are within $\overline{x}$;
-the types for $\overline{x}$ and the final return type $\tau$ of $e$ have no checked pointers.
-Otherwise, it violates the non-exposure safety.
+The execution of a $\echeckedtext$ or $\euncheckedtext$ block represents 
+the context switching from a checked to an checked region, or vice versa,
+with its type and semantic rules given in \Cref{fig:type-system-1}.
+In this context switching, to guarantee the checked pointer non-exposure property, 
+checked pointers are not allowed to pass through, which is guaranteed by the predicates 
+$\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$,
+as well as the check that all free variables in the block content $e$ are in $\overline{x}$.
 For example, \code{read_msg} in the \code{handle_request} function is tainted in \Cref{lst:humantaint},
 if any argument for \code{handle_request} is a checked pointer, 
-it means that we are exposing a checked pointer address to unsafe regions.
-% 
-\mzu{A $\echeckedtext$ or $\euncheckedtext$ block represents 
-  the context switching from a checked to an checked region, or vice versa.
-}
-% 
-\mz{An action + checked/unchecked block represents the switching.}
-% 
-We need to make sure no checked pointers are \mzs{information} exposed to unsafe code regions.
-as rules \textsc{S-Unchecked} and \textsc{S-Checked} in \Cref{fig:semantics}.
-% 
-\mz{This paragraph is not that cohesive.}
-%
+it means that we are exposing a checked pointer address to unsafe regions, 
+which means that users might randomly manipulate the data region pointed to by the checked pointer
+and \lang forbids such behavior.
 
 \myparagraph{Dependent Function Pointers}
-Rule \textsc{T-Fun} in \Cref{fig:type-system-1} is the dependent function call rule. 
 Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
@@
 we confirm that each of $\overline{\tau'}$ is
-a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$,
+a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$ 
+in rule \textsc{T-Fun} (\Cref{fig:type-system-1}),
 which replaces possible integer bound variables $\overline{x}$ with bound expressions $\overline{e'}$.
@@
 \code{n} with the argument at a call-site.
+\ignore{
 The semantics manages variable scopes using the special $\erettext$
@@
 the literal. 
+}
 
@@
 maps addresses (integer literals) to the function data described above.
-
-\systemname{} has dependent functions, whose semantic explanation is given in \Cref{appx:add-type-sem}.
-Note that the \textsc{S-FunC} and \textsc{S-FunT} rules replace the
-  annotations $\overline{\tau_a}$ with
-  $\overline{\tau}$ (after instantiation) from the function's
-  signature. Using $\overline{\tau_a}$ when executing the body of
-the function has no impact on the soundness of \lang, but will violate
-Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
 Rule \textsc{S-FunT} defines the tainted version of function call semantics.
@@
 
+\ignore{
 \subsection{Semantics}\label{sec:semantics}
@@
 The other compilation rules are given in \Cref{appx:add-type-sem}.
+}

### 2023-01-14 sandp23/formal.tex
update section 4.

@@
 
+\subsection{Meta Theories}\label{sec:theorem}
+
+% Before we present our main theorems, we need to first
+% discuss the meaning what a pointer being well-typed in a given heap
+% snapshot $\heap$ means, which is captured by rules in
+% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
+% simply checks if a given variable has the defined type in $\Gamma$;
+% the constant rule ($\textsc{T-Const}$) is slightly more involved.
+% First, it ensures that the type annotation $\tau$ does not contain any
+% free variables. More importantly, it ensures that the pointer points
+% to a location that makes sense in a given heap.
+%  
+%  
+%  The $\size$ function in Fig.~\ref{fig:const-type}
+% refers to the \code{sizeof} function in C computing the number of
+% bytes for a type.
+%  
+%  
+%  Second, we
+% require that any constant ($\evalue{n}{\tau}$) should make sense in
+% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
+% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
+% constant set containing the constants that have been verified by the
+% relation. For every constant $\evalue{n}{\tau}$, it is either an
+% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
+% zero-valued number ($n=0$), checked in $\sigma$
+% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
+% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
+% location $n+i$ in the range of the pointer (if $\omega$ is a word
+% type, range is $[0,1)$; if $\omega$ is an array type
+%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
+%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
+%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
+%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
+%       \vdash n_a : \tau_a$.
+%  
+%  
+% \begin{figure}[t]
+% {\small
+% \text{Type Rules for Constants and Variables:}
+% \begin{mathpar}
+%   \inferrule[T-Var]
+%       {x : \tau \in \Gamma}
+%       {\Gamma;\Theta \vdash_m x : \tau}
+%  
+%   \inferrule[T-Const]
+%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
+%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
+% \end{mathpar}
+%     
+% \text{Rules for Checking Constant Pointers In Heap:}
+% \begin{mathpar}
+%   \inferrule
+%       {}
+%       {\sigma \vdash n : \tint}
+%  
+%   \inferrule
+%       {}
+%       {\sigma \vdash n : \tptr{\omega}{\umode}}
+%  
+%   \inferrule
+%       {}
+%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
+%  
+%   \inferrule
+%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
+%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
+%  
+%   \inferrule
+%       {\forall i \in [0,\size(\omega)) .
+%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
+%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
+% \end{mathpar}
+% }
+% \caption{Type Rules for Checking Constants/Variables}
+% \label{fig:const-type}
+% \end{figure}
+
+% \review{
+%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
+%   something, I didn't see such a definition in the paper.}
+% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
+%   assumption of well typing; we have added more details about that.}
+
+In this subsection, we focus on our main meta-theoretic results about
+\lang: type soundness (progress and preservation),
+non-exposure, and non-crashing.
+These proofs have been conducted in our Coq model.
+Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{appx:add-type-sem}.
+Progress states that a \lang program can always make a move:
+
+\begin{thm}[Progress]\label{thm:progress}
+
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed, consistent
+($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
+one of the following holds:
+
+\begin{itemize}
+
+\item $e$ is a value ($\evalue{n}{\tau}$).
+
+\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.
+
+\end{itemize}
+\end{thm}
+%{\em Proof:} By induction on the typing derivation.
+
+\noindent
+There are two forms of preservation regarding the checked and unchecked regions.
+Checked Preservation states that a reduction step preserves both the
+type and consistency of the program being reduced.
+Unchecked Preservation states that any evaluation happens at unchecked region does not affect the checked heap.
+
+\begin{thm}[Checked Preservation]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed, consistent
+($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
+checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
+$\Gamma'$ and $\tau'$ that are well formed, checked region consistent
+($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
+well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
+$\tau'\sqsubseteq_{\Theta} \tau$.
+\end{thm}
+%{\em Proof:} By induction on the typing derivation.
+%\smallskip
+\begin{thm}[Unchecked Preservation]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
+\end{thm}
+
+$\heap(\cmode)$ refers to all $\cmode$ mode pointers in heap $\heap$.
+Using the above theorem, we first show the non-exposure theorem,
+where code in unchecked region cannot observe a valid checked pointer address.
+
+\begin{thm}[Non-Exposure]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are all are well-formed and well
+typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
+$\heap'$ and $e'$, such that $(\varphi,\heap,e)
+\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
+where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
+thus, it is not a checked pointer.
+\end{thm}
+
+We now state our main result, {\em non-crashing},
+which suggests that a well-typed program can never be \emph{stuck} (expression
+$e$ is a non-value that cannot take a step\footnote{Note that
+  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
+  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
+  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
+  safety.}).
+
+% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r.
+Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
+% \liyi{It is a typo.
+We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region.
+This is a bit due to the space limitation.
+The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement.
+}
+
+\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
+  program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+that are well-formed and consistent
+($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
+if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
+\tau$ for some $\tau$) and there exists
+$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
+$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
+\end{thm}
+
+%{\em Proof:} By induction on the number of steps of the \checkedc
+%evaluation ($\longrightarrow_m^*$), using progress and preservation to
+%maintain the invariance of the assumptions.
+
 \ignore{

### 2023-01-14 sandp23/formal.tex
[review] rewrite dependnt FP

@@
 
-\myparagraph{Dependent Function Pointers}
-Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
-from a type-check for $e$ and the types $\overline{\tau'}$ from the argument type checks for $\overline{e}$,
-we confirm that each of $\overline{\tau'}$ is
-a subtype of the corresponding one in $\overline{\tau}[\overline{e'} / \overline{x}]$ 
-in rule \textsc{T-Fun} (\Cref{fig:type-system-1}),
-which replaces possible integer bound variables $\overline{x}$ with bound expressions $\overline{e'}$.
-The final result type is the defined target type $\tau$ appearing in the function pointer type
-also with such replacement, written as $\tau[\overline{e'} / \overline{x}]$.
+\myparagraph{Dependent Function Pointers} %
+The rule \textsc{T-Fun} (\Cref{fig:type-system-1}) states the type judgement for
+dependent function pointer application where we represent the result of
+replacing all integer bound variables $\overline{x}$ in the type \(\tau\) with
+with bound expressions $\overline{e'}$ by $\tau[\overline{e'} / \overline{x}]$
+and write $\overline{\tau}[\overline{e'} / \overline{x}]$ to lift the
+substitution to every type \(\overline{\tau}\).
+% 
+Given an expression $e$ of function pointer type
+($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$) and arguments
+$\overline{e}$ of types $\overline{\tau'}$,
+% 
+the result of the application will
+be of type $\tau[\overline{e'} / \overline{x}]$ 
+% 
+if for each pair of \(\tau'\) and \(\tau''\) in \(\overline{\tau'}\) and
+$\overline{\tau}[\overline{e'} / \overline{x}]$, \(\tau'\) is a subtype of
+\(\tau''\).
+% 
+\mz{I rewrote the sentences above.}
+% 
 Consider the \code{process_req2} function in
 Fig.~\ref{lst:final}; its parameter type for \code{msg} 
-depends on \code{m_1}. The \textsc{T-Fun} rule will substitute 
-\code{n} with the argument at a call-site.
+depends on \code{m_1}. The \textsc{T-Fun} rule will
+\mzu{substitute \code{n} with the argument}
+\mz{\code{n} seems unbound here.}
+at a call-site.
 \ignore{

### 2023-01-14 sandp23/implementation.tex
Fixing implementation

@@
 
+\subsection{Compiler}
+The~\systemname compiler includes the type checker (\sect{}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
+We implemented the compiler on top of the existing~\checkedc compiler.
+
+For all tainted pointer access, our instrumentation adds a check (executed at runtime) to see we check if the address is within the valid range for the sandbox. An example of the check is shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+_TPtr<int> t = ...
++ if (!is_in_range(t, SBX_LOW(), SBX_HIGH()))
++  handle_violation();
+*t = 1;
+\end{minted}
+We show the added check in C for clarify. However, our instrumentation works on LLVM IR.
+
+Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
+\cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers.
+
+\iffalse
+
 %\subsection{Overview}
@@
 Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout.
+\fi

### 2023-01-14 sandp23/implementation.tex
Minor

@@
 
+\subsection{Supporting General Partioning Use Case}
+\input{figures/newusecase}
+
+Tainting all functions that are not checked may lead to a large number of transitions to the sandbox, which could add unnecessary overhead.
+In our example, in~\lst{lst:final}, marking~\code{read_msg} function as tainted would cause the system to transition to and from the sandbox every time a message is read, which could be inefficient.
+
+However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain checked pointers as well.
+%
+This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them.
+In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
+
+For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
+
+The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
+It is best to annotate all functions with checked types for security purposes.
+However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted.
+This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
+\aravind{Change title of this section and say that this mode has no guarentees and move this before evaluation.}
+
 \iffalse

### 2023-01-14 sandp23/overview.tex
Minor

@@
 However, as shown in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
-
+\aravind{Fix the running example code with the annotations according to the implementation.}
 \subsection{\systemname Annotations}
@@
 }
-\subsection{Alternative Use Cases}
-\input{figures/newusecase}
-
-Tainting all functions that are not checked may lead to a large number of transitions to the sandbox, which could add unnecessary overhead.
-In our example, in~\lst{lst:final}, marking~\code{read_msg} function as tainted would cause the system to transition to and from the sandbox every time a message is read, which could be inefficient.
-
-However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain checked pointers as well.
-%
-This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them.
-In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
-
-For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
-
-The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
-It is best to annotate all functions with checked types for security purposes.
-However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted.
-This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
-\aravind{Change title of this section and say that this mode has no guarentees and move this before evaluation.}
 

### 2023-01-14 sandp23/implementation.tex
Minor

@@
 
-However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain checked pointers as well.
+However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain regular (unchecked) pointers as well.
 %

### 2023-01-15 sandp23/implementation.tex
Minor

@@
 We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
-We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then call~\code{strncpy} as~\code{strncat(dst, tmp, n);}.
+We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then invoke~\code{strncat(dst, tmp, n);}.
 Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
@@
 // Put your example annotation here.
-strncpy(dst, t_str, n);
+// First, manually check the memory is in tainted region.
+// if yes, then call strncpy.
+if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
+  handle_violation();
+// our type checker ignores this because 
+// the above _TLIB annotation.
+strncat(dst, t_str, n);
 \end{minted}
-\aravind{Fix the above code}
+\aravind{Arun, Fix the above code. i.e., Add \_TLIB annotation to strncpy }
 The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.

### 2023-01-15 sandp23/implementation.tex
Minor
Adding Implementation

@@
 The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.
+Note that developer needs to manually add the check before calling~\code{strcat}.
 
@@
 
+\iffalse
 \noindent\emph{Handling Tainted Buffers in~\cregion}:
@@
 Note that we also add the necessary deallocator calls at function exit points.
+\fi
 
@@
 
+Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
+For instance, for the code in~\lst{lst:final}, our type checker emits an error saying that ``~\emph{buffer~\code{buff} is tainted, but it is allocated in~\cregion}''.
+Consequently, the developer has to make the following changes to allocate~\code{buff} in the sandbox:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+- char buff[MAX_MSG_SIZE] _Tainted;
++ _t_Array_ptr<char> buff: count(MAX_MSG_SIZE)
++    = sbx_allocate(MAX_MSG_SIZE);
+...
++ sbx_free(buff)
+return rc;
+\end{minted}
+
 For all tainted pointer access, our instrumentation adds a check (executed at runtime) to see we check if the address is within the valid range for the sandbox. An example of the check is shown below:
@@
 This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
-\aravind{Change title of this section and say that this mode has no guarentees and move this before evaluation.}
-
+We add the additional constraint of not allowing casting from taint types to generic types to enable this generic use case without violating the safety guarantees of~\systemname.
+Specifically, we disallow the following casting in~\cregion:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+//_t_Ptr<int> y; int *z;
+|\faRemove| z = (int *)y; // Not allowed.
+\end{minted}
 \iffalse

### 2023-01-15 sandp23/implementation.tex
Minor additions to implementation

@@
 In~\cregion, we also need to modify calls to tainted functions as they execute inside the sandbox (separate address space) and thus cannot be invoked as regular functions.
-We rewrite all calls to tainted functions with corresponding sandbox invocations.
-For instance, we modify the call to~\code{process_req1} in~\lst{lst:final} as below:
+However, modifying every call site of tainted functions is tedious and also requires precise pointer analysis~\cite{milanova2002precise} to handle indirect calls through function pointers.
+
+We handle this by~\emph{indirection}: Instead of modifying the call sites, we modify the body of tainted functions to invoke the corresponding function in the sandbox.
+For instance, we modify the body of tainted function~\code{process_req1} (\lst{lst:final}) in~\cregion as below:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
-case REQ1: 
-- rc = process_req1(buff); 
-+ rc = .....
+int process_req1(char *msg, size_t m_l) _Tainted {
+- int rc = -1, i;
+- if (m_l > MIN_SIZE) {
+- ...
++ HOW DO WE HANDLE SANDBOX CALL.
+}
 \end{minted}
-\aravind{Arun, please fix the above example with the exact details.}
+This ensures that all calls (even through function pointers) to the tainted function~\code{process_req1} are redirected to the sandbox.
+\aravind{Arun, add the call on how to call sandbox function.}
 
@@
 For instance:~\code{_T_Array_ptr<char>} will be replaced with~\code{char *}.
-The~\code{read_msg} function signature in~\lst{} will be modified as shown below:
+The~\code{read_msg} function signature in~\lst{lst:final} will be modified as shown below:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
 size_t read_msg(int sock_fd, char *msg,
--                             size_t sz) __Tainted {
+-                             size_t sz) _Tainted {
 +                             size_t sz) {

### 2023-01-15 sandp23/implementation.tex
Adding

@@
 We will present our implementation using WebAssembly (WASM~\cite{bosamiya2020webassembly}) as our target sandbox.
-However, as we discuss in~\sect{}, our implementation can be easily extended to other sandbox mechanisms.
+%However, as we discuss in~\sect{}, our implementation can be easily extended to other sandbox mechanisms.
 
@@
 Developers should use this qualifier to mark callback functions,~\ie functions in~\cregion, that can be called from the tainted region. 
-The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\sect{}).
+The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\sect{subsub:gencregion}).
 

### 2023-01-15 sandp23/appendix.tex
update appx.

@@
 \section{Appendix}\label{app:main}
+
+\subsection{Well-formedness and Subtype}
+\label{app:le}
+
+\begin{figure}[h]
+{\small
+  \begin{mathpar}
+
+  \inferrule[]
+  {}
+  {m \vdash \tint}
+
+  \inferrule[]
+  {\xi \wedge m\vdash \tau \\ \xi \le m}
+  {m \vdash \tptr{\tallarrayb{\bvar}{\tau}}{\xi}}
+
+  \inferrule[]
+  {\xi \wedge m \vdash \tau\\ \xi \le m}
+  {m \vdash \tptr{\tau}{\xi}}
+
+  \inferrule[]
+  {\xi \wedge m \vdash \tau\\ \xi \le m \\\\ \fv(\overline{\tau})\cup\fv(\tau)\subseteq \overline{x}}
+  {m \vdash \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi})}
+  \end{mathpar}
+}
+{\footnotesize
+\[
+\begin{array}{l} 
+\tmode \wedge \cmode = \umode \qquad \xi \wedge \umode = \umode
+\qquad \cmode \wedge m = m 
+\qquad  m_1 \wedge m_2 = m_2 \wedge m_1
+\end{array}
+\]
+}
+ \caption{Well-formedness for Nested Pointers}
+\label{fig:wftypes}
+\end{figure}
+
+\begin{figure}[h]
+{\small
+  \begin{mathpar}
+
+    \inferrule[]
+    {}
+    {\Gamma \vdash n}
+
+    \inferrule[]
+    {x:\tint \in \Gamma}
+    {\Gamma \vdash x + n}
+
+    \inferrule[]
+    {\Gamma \vdash b_l\\
+    \Gamma \vdash b_h}
+  {\Gamma \vdash (b_l,b_h)}
+
+  \inferrule[]
+  {}
+  {\Gamma \vdash \tint}
+
+  \inferrule[]
+  {\Gamma \vdash \bvar \\
+  \Gamma \vdash \tau}
+  {\Gamma \vdash \tptr{\tallarrayb{\bvar}{\tau}}{m}}
+
+  \inferrule[]
+  {\Gamma \vdash \tau}
+  {\Gamma \vdash \tptr{\tau}{m}}
+
+  \inferrule[]
+  {T \in D}
+  {\Gamma \vdash \tptr{\tstruct{T}}{m}}
+
+  \inferrule[]
+  {\Gamma \vdash \tau}
+  {\Gamma \vdash \tptr{\tau}{m}}
+
+  \inferrule[]
+  {\forall \tau_i\in \overline{\tau}\,.\,\Gamma[\forall x\in\overline{x}\,.\, x\mapsto \tint] \vdash \tau_i\\\\
+   \Gamma[\forall x\in\overline{x}\,.\, x\mapsto \tint] \vdash \tau }
+  {\Gamma \vdash \tfun{\overline{x}}{\overline{\tau}}{\tau}}
+  \end{mathpar}
+}
+ \caption{Well-formedness for Types and Bounds}
+\label{fig:wftypesandbounds}
+\end{figure}
+
+\Cref{fig:wftypes} defines the well-formedness for nested pointers, guaranteeing that no tainted ($\tmode$) pointer has a checked ($\cmode$) element field.
+Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
+\Cref{fig:wftypesandbounds} defines the well-formedness for bounds appearing in a type.
+All bound variables appearing in a type must be either an instance of the type environment $\Gamma$,
+or a bound variable appearing in a function pointer.
+
+\begin{DIFnomarkup}
+\begin{figure}
+{\small
+\[\hspace*{-1.2em}
+\begin{array}{l}
+\textcolor{blue}{\text{Bound Inequality and Equality:}}\\[0.3em]
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
+     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
+     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
+     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
+     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
+     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
+    \end{array}
+  \\[1.5em]
+\textcolor{blue}{\text{Type Equility:}}\\
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+     && \tint & =_{\Theta} & \tint\\
+     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
+     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
+             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
+
+    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
+
+ &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
+                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
+    \end{array}
+  \\[0.7em]
+\textcolor{blue}{\text{Subtype:}}\\[0.3em]
+
+  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
+    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
+
+    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
+    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
+    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
+\\[0.2em]
+\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
+
+    \end{array}
+\end{array}
+  \]
+}
+{\footnotesize
+\[
+\begin{array}{l}
+n'+n = add(n',n)
+\qquad
+(x+n')+n = x+add(n',n)\\
+\textit{cond}(\overline{x},\tau,\overline{y},\tau')
+=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
+  \wedge \overline{y}\cupdot\overline{z}
+  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
+\\\qquad\qquad\qquad\qquad\qquad
+  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
+\end{array}
+\]
+}
+  \caption{Type Equality and Subtyping}
+  \label{fig:checkc-subtype}
+\end{figure}
+\end{DIFnomarkup}
+
+In \lang, type equality $\tau=_{\Theta}\tau'$
+is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
+and the alpha equivalence of two function types in \Cref{fig:checkc-subtype};
+i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
+$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
+are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
+and the substitution results are equal.
+
+ The \textsc{T-CastPtr} rule
+permits casting from an expression of type $\tau'$ to a checked pointer when
+$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
+$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and is built on the type equality
+($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
+rules ensure the relation is transitive. Most of the rules manage
+casting between array pointer types. The rule 
+($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
+\tarrayptr{b_l}{b_h}{\tau}{m}$) permits treating a singleton
+pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
+Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
+%There is another casting rule in \Cref{app:main} stating that
+% users are free to cast types in unchecked code regions, since unchecked regions can contain C code.
+
+The subtyping relation given in Fig.~\ref{fig:checkc-subtype} involves
+dependent bounds, i.e., bounds that may refer to variables. To decide
+premises $b \leq_{\Theta} b'$ in \Cref{fig:checkc-subtype}, we need a decision procedure that accounts for
+the possible values of these variables. This process considers
+$\Theta$, tracked by the typing judgment, and $\varphi$, the current
+stack snapshot (when performing subtyping as part of the type
+preservation proof).
+
+Since bounds expressions may
+contain variables, determining assumptions like $b_l \leq_{\Theta} b_l'$
+requires reasoning about the probable values of these variables'. The type
+system uses $\Theta$ to make such reasoning more precise.
+$\Theta$ is a map from variables $x$ to
+equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
+It maps variables to equations that are recorded along the type checking procedure.
+If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
+$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
+such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
+
+$\sqsubseteq$ is parameterized by
+$\Theta$, which provides the range of allowed values for a bound
+variable; thus, more $\sqsubseteq$ relation is provable. For example,
+rule \rulelab{T-LetInt} inserts a predicate $\teq{e_1}$ for variable $x$.
+Assume that $e_1$ is equal to $0$,
+when $x$ is used as a type variable in type $\tntarrayptr{0}{x}{\tint}{\cmode}$ in $e_2$,
+the subtyping relation $\tntarrayptr{0}{x}{\tint}{\cmode} \sqsubseteq
+\tntarrayptr{0}{0}{\tint}{\cmode}$ is provable when we know
+\code{x}$\teq{0}$.
+
+To capture bound variables in dependent types, the \checkedc subtyping
+relation ($\sqsubseteq$) is parameterized by a restricted stack
+snapshot $\varphi|_{\rho}$ and the predicate map $\Theta$, where
+$\varphi$ is a stack and $\rho$ is a set of
+variables. $\varphi|_{\rho}$ means to restrict the domain of $\varphi$
+to the variable set $\rho$. Clearly, we have the relation:
+$\varphi|_{\rho} \subseteq \varphi$. $\sqsubseteq$
+being parameterized by $\varphi|_{\rho}$ refers to that when we
+compare two bounds $b \le_{\Theta} b'$, we actually do
+$\varphi|_{\rho}(b) \le_{\Theta} \varphi|_{\rho}(b')$ by interpreting the
+variables in $b$ and $b'$ with possible values in $\varphi|_{\rho}$.
+Let's define a subset relation $\preceq$ for two restricted stack
+snapshot $\varphi|_{\rho}$ and $\varphi'|_{\rho}$:
+
+\begin{DIFnomarkup}
+\begin{figure*}[t]
+{\small
+  \begin{mathpar}
+   \inferrule[T-ConstU]
+       { \neg \cmode(\tau)}
+       {\Gamma;\Theta\vdash_u \evalue{n}{\tau} : \tau}
+\quad
+   \inferrule[T-ConstC]
+       {\Theta;\heap;\emptyset \vdash_c n : \tau}
+       {\Gamma;\Theta\vdash_c \evalue{n}{\tau} : \tau}
+\quad      
+   \inferrule[T-Let]
+    { x\not\in \fv(\tau') \\
+        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
+          \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'}
+\quad
+   \inferrule[T-LetInt]
+    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\
+        \Gamma;\Theta \vdash_m e_1 : \tint \\\\
+           \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
+\quad
+   \inferrule[T-RetInt]
+    { \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
+    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tint}}{e} : \tau}
+
+    \inferrule[T-Mac]
+              {\xi\le m}
+              {\Gamma; \Theta \vdash_m \emalloc{\xi}{\omega} : \tptr{\omega}{\xi}}
+
+    \inferrule[T-Add]
+              {\Gamma; \Theta \vdash_m e_1 : \tint \\
+                \Gamma; \Theta \vdash_m e_2 : \tint}
+              {\Gamma; \Theta \vdash_m (e_1 \plus e_2) : \tint }
+
+    \inferrule[T-Ind] 
+              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{\xi} \\
+                \Gamma; \Theta \vdash_m e_2 : \tint \\
+                \xi \leq m}              
+              {\Gamma; \Theta \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
+
+    \inferrule[T-Assign]
+              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tau}{\xi} \\
+                \Gamma; \Theta \vdash_m e_2 : \tau' \\\\
+                \tau'\sqsubseteq_{\Theta} \tau \\
+                \xi \leq m}
+              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
+
+    \inferrule[T-AssignArr]
+              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{\xi}\\\\
+                \Gamma; \Theta \vdash_m e_2 : \tau' \\
+                \tau'\sqsubseteq_{\Theta} \tau\\
+                \xi \leq m}
+              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
+
+   \inferrule[T-IndAssign]
+              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{\xi}\\
+                \Gamma; \Theta \vdash_m e_2 : \tint \\\\
+                \Gamma; \Theta \vdash_m e_3 : \tau' \\
+                \tau'\sqsubseteq_{\Theta} \tau \\
+                \xi \leq m}
+              {\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
+
+  \end{mathpar}
+}
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
+% \\[0.2em]
+% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
+% \end{array}
+% $
+% \end{center}
+% }
+\caption{Remaining \lang Type Rules (extends Fig.~\ref{fig:type-system-1})}
+\label{fig:rem-type-system}
+\end{figure*}
+\end{DIFnomarkup}
+
+\begin{defi}[Subset of Stack Snapshots]
+  Given two $\varphi|_{\rho}$ and $\varphi'|_{\rho}$,
+  $\varphi|_{\rho} \preceq \varphi'|_{\rho}$, iff for $x\in\rho$ and
+  $y$,
+  $(x,y) \in \varphi|_{\rho} \Rightarrow (x,y) \in \varphi'|_{\rho}$.
+\end{defi}
+
+For every two restricted stack snapshots $\varphi|_{\rho}$ and
+$\varphi'|_{\rho}$, such that
+$\varphi|_{\rho} \preceq \varphi'|_{\rho}$, we have the following
+theorem in \checkedc (proved in Coq):
+
+\begin{thm}[Stack Snapshot Theorem]
+  Given two types $\tau$ and $\tau'$, two restricted stack snapshots
+  $\varphi|_{\rho}$ and $\varphi'|_{\rho}$, if
+  $\varphi|_{\rho}\preceq \varphi'|_{\rho}$, and
+  $\tau \sqsubseteq \tau'$ under the parameterization of
+  $\varphi|_{\rho}$, then $\tau \sqsubseteq \tau'$ under the
+  parameterization of $\varphi'|_{\rho}$.
+\end{thm}
+
+Clearly, for every $\varphi|_{\rho}$, we have
+$\emptyset \preceq \varphi|_{\rho}$. The type checking stage is a
+compile-time process, so $\varphi|_{\rho}$
+is $\emptyset$ at the type checking stage. Stack snapshots are needed
+for proving type preserving, as variables in bounds expressions are
+evaluated away.
+
+\subsection{Other Type Rules and Literal Validity Checks}\label{rem-type}
+
+Here we show the type rules for other \lang operations in Fig.~\ref{fig:rem-type-system}.
+Rule \textsc{T-Mac} deals with
+$\emalloctext$ operations. There is a well-formedness check to require
+that the possible bound variables in $\omega$ must be in the domain of
+$\Gamma$ (see Fig.~\ref{fig:wftypesandbounds}).
+This is similar to the well-formedness assumption of the type environment (Definition~\ref{type-wellformed}).
+Rule \textsc{T-Add} deals with binary operations whose sub-terms are integer expressions.
+The other rules are explained as follows.
+
+\myparagraph{Pointer Access}
+%
+The \textsc{T-AssignArr} rule examines array assignment operations, returning the type of
+pointed-to objects. Rules for pointers for other object types are
+similar, such that Rule \textsc{T-Assign} assigns a value to a non-array pointer location.
+The condition $\xi\le m$ ensures that checked and unchecked pointers 
+can only be dereferenced in checked and unchecked regions, respectively;
+The type rules do not attempt to reason whether the access is in bounds;
+such check is deferred to the semantics.
+Rule \textsc{T-Ind} serves the case for pointer arithmetic.
+For simplicity, in the \checkedc formalization, we do not allow arbitrary pointer arithmetic.
+The only pointer arithmetic operations allowed are the forms shown in rules \textsc{T-Ind} and \textsc{T-IndAssign} in Fig.~\ref{fig:rem-type-system}.
+The predicate $\tau'\sqsubseteq_{\Theta} \tau$ requires that the value being assigned is a subtype of the pointer type.
+The \textsc{T-IndAssign} rule is an extended assignment operation for handling assignments for array/NT-array pointers with pointer arithmetic.
+
+% Subtyping and casting operations are briefly introduced in
+% Sec.~\ref{sec:intros}~and~\ref{sec:overview}.  Subtyping is useful in
+% static casting operations that allow users to view a pointer in one
+% type as another, such as casting an NT-array pointer to an array
+% one. \checkedc provides a set of safe static casting operations that
+% have no cost in execution.  Moreover, subtyping acts as oracles for
+% bound widening and dynamic casting operations; thus, \checkedc is
+% different from a complete static array pointer bound system.  For
+% example, if $e$ has type $\tau'$ and $\varphi$ is the current stack
+% snapshot, the semantics of $\edyncast{\tau}{e}$ does not transition to
+% an error state when $\varphi(\tau')\sqsubseteq\varphi(\tau)$.  In a
+% function call, for every argument, \lang permits users to input a
+% subtype entity and we prove that this does not affect the correctness
+% of the program.
+
+\begin{DIFnomarkup}
+ \begin{figure}[t]
+ {\small
+
+ \begin{mathpar}
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m n : \tint}
+
+   \inferrule
+       {}
+       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
+       {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
+
+   \inferrule
+       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
+            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+
+   \inferrule
+       { \xi \le m 
+     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
+       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
+
+   \inferrule
+       {\neg\funptr(\omega)\\ \xi \le m\\
+        \forall i \in [0,\size(\omega)) \;.\;
+            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+ \end{mathpar}
+ }
+{\footnotesize
+\[
+\begin{array}{l} 
+\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
+\qquad
+\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
+\end{array}
+\]
+}
+ \caption{Verification/Type Rules for Constants}
+ \label{fig:const-type}
+ \end{figure}
+\end{DIFnomarkup}
+
+\myparagraph{Literal Constant Validity}
+Rules \textsc{T-ConstU} and \textsc{T-ConstC}
+describes type assumptions for literals appearing in a program.
+$\cmode(\tau)$ judges that a literal pointer 
+in an unchecked region cannot be of a checked type,
+which represents an assumption that programmers 
+cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+In rule \textsc{T-ConstC}, we requires a static 
+verification procedure for validating a literal pointer, 
+which is similar to the dynamic verification process in \Cref{sec:typechecking}.
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$ checks (\Cref{fig:const-type})
+validate the literal $\evalue{n}{\tau}$, 
+where $\heap(m)$ is the initial heap that the literal resides on and
+$\sigma$ is a set of literals assumed to be checked.
+A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
+A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
+and the address stores a function with the right type.
+The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
+where every element in the pointer range ($[0,\size(\omega))$) should be well
+typed.
+
+\begin{figure*}[t]
+{\small
+\begin{mathpar}
+
+\inferrule [S-Var]{} {(\varphi,\heap,x)\longrightarrow (\varphi,\heap,\varphi(x))}
+
+    \inferrule[S-DefArrayC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
+    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefArrayT]{\heap(\umode,n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h) 
+               \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}}
+    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefArrayBound]{0 \not\in [n_l,n_h)}
+     { (\varphi,\heap,\estar{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\xi}}}) \longrightarrow (\varphi,\heap,\ebounds)}
+\quad
+    \inferrule[S-DefNTArrayBound]{0 \notin [n_l,n_h]}
+    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\xi}}}) \longrightarrow (\varphi,\heap,\ebounds)}
+\quad
+    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{\evalue{n'}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n'}{\tau'})}
+
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e})}
+
+    \inferrule[S-AssignNull]{}
+      {(\varphi,\heap,\eassign{\evalue{0}{\tptr{\omega}{\xi}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\enull)}
+
+    \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}
+
+    \inferrule[S-AssignArrC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\cmode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}
+
+    \inferrule[S-AssignArrT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) 
+               \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}}
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\tmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heapup{\umode}{n}{\evalue{n_1}{\tau_a}},\evalue{n_1}{\tau})}
+
+    \inferrule[S-AssignArrBound]{0 \not\in [n_l,n_h) }
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\ebounds)}
+
+    \inferrule[S-AssignC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
+      {(\varphi,\heap,\eassign{\evalue{n}{\tptr{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}
+
+  \inferrule[S-Malloc]{\varphi(\omega)=\omega_a \\ \mathtt{alloc}(\heap,\xi,\omega_a)=(n,\heap')}
+   { (\varphi,\heap,\emalloc{\xi}{\omega}) \longrightarrow (\varphi,\heap',\evalue{n}{\tptr{\omega_a}{\xi}})}
+
+  \inferrule[S-MallocBound]{\varphi(\omega)=\tallarray{n_l}{n_h}{\tau}\\ (n_l \neq 0 \vee n_h \le 0)}
+    { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\ebounds)}
+
+  \inferrule[S-IfNTNotC]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}} \\ \heap(\cmode,n)\neq 0\\ 0 < n_h}
+             {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
+
+    \inferrule[S-IfT]{n \neq 0 }
+    {(\varphi,\heap,\eif{\evalue{n}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
+
+    \inferrule[S-IfF]{}
+    {(\varphi,\heap,\eif{\evalue{0}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}
+
+    \inferrule[S-Add]{n = n_1 + n_2}
+    {(\varphi,\heap,\evalue{n_1}{\tint} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, n)}
+
+    \inferrule[S-AddArr]{n = n_1 + n_2\\ n_l' = n_l - n_2 \\ n_h' = n_h - n_2}
+    {(\varphi,\heap,\evalue{n_1}{\tallarrayptr{n_l}{n_h}{\tau}{\xi}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau}{\xi}})}
+
+n    \inferrule[S-AddArrNull]{}
+    {(\varphi,\heap,\evalue{0}{\tallarrayptr{n_l}{n_h}{\tau}{\xi}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \enull)}
+
+\end{mathpar}
+
+}
+\caption{Remaining \lang Semantics Rules (extends Fig.~\ref{fig:type-system-1})}
+\label{fig:rem-semantics}
+\end{figure*}
+
+A checked pointer checks validity in type step as rule \textsc{T-ConstC},
+while a tainted/unchecked pointer does not check for such during the type checking.
+Tainted pointers are verified through the validity check in dynamic execution as we mentioned above.
+
+If the literal's type is an integer, an unchecked pointer, or a null
+pointer, it is well typed, as shown by the top three rules in
+Fig.~\ref{fig:const-type}. However, if it is a checked pointer
+$\tptr{\omega}{\cmode}$, we need to ensure that what it points to in
+the heap is of the appropriate pointed-to type ($\omega$), and also
+recursively ensure that any literal pointers reachable this way are
+also well-typed. This is captured by the bottom rule in the figure,
+which states that for every location $n+i$ in the pointers' range
+%
+$[n, n+\size(\omega))$, where $\size$ yields the size of its argument,
+  then the value at the location $\heap(n+i)$ is also well-typed.
+  However, as heap snapshots can contain cyclic structures (which
+  would lead to infinite typing deriviations), we use a scope $\sigma$
+  to assume that the original pointer is well-typed when checking the
+  types of what it points to. The middle rule then accesses the scope
+  to tie the knot and keep the derivation finite, just like in
+  \citet{ruef18checkedc-incr}.
+
+\myparagraph{Let Bindings}
+%
+Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
+type dependency. 
+In particular, the result of evaluating a $\elettext$
+may have a type that refers to one of its bound variables (e.g., if
+the result is a checked pointer with a variable-defined bound). If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
+Note that we restrict the expression $e_1$ to syntactically match the
+structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).
+Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
+$\erettext$ does not appear in source programs but is introduced by the semantics when
+evaluating a let binding (rule \textsc{S-Let} in
+Fig.~\ref{fig:rem-semantics}).
+
+\subsection{Other Semantic Rules}\label{sec:rem-semantics}
+
+Fig.~\ref{fig:rem-semantics} shows the remaining semantic rules for
+$\lang$. We explain a selected few rules in this subsection.
+% other few low-level semantic rules for variable and dereference and $\emalloctext$ operations in \checkedc. Other operations are defined in the same manner.
+
+Rule \textsc{S-Var} loads the value for $x$ in stack $\varphi$.
+Rule \textsc{S-DefArray} dereferences an array pointer, which is similar to the Rule \textsc{S-DefNTArray} in Fig.~\ref{fig:semantics} (dealing with null-terminated array pointers).
+The only difference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
+Rules \textsc{DefArrayBound} and \textsc{DefNTArrayBound} describe an error case for a dereference operation.
+If we are dereferencing an array/NT-array pointer and the mode is $\cmode$, $0$ must be in the range from $n_l$ to $n_h$ (meaning that the dereference is in-bound); if not, the system results in a $\ebounds$ error.
+Obviously, the dereference of an array/NT-array pointer also experiences a $\enull$ state transition if $n\le 0$.
+
+Rules \textsc{S-Malloc} and \textsc{S-MallocBound} describe the $\emalloctext$ semantics.
+Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$, and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation.
+The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\cmode}}$ and new updated heap.
+It is possible for $\emalloctext$ to transition to a $\ebounds$ error if the $\omega_a$ is an array/NT-array type $\tallarray{n_l}{n_h}{\tau}$, and either $n_l \neq 0$ or $n_h \le 0$.
+This can happen when the bound variable is evaluated to a bound constant that is not desired.
+
+\subsection{Additional Type and Semantic Rules}\label{appx:add-type-sem}
+
+Since bounds expressions may
+contain variables, determining assumptions like $b_l \leq b_l'$
+requires reasoning about the probable values of these variables'. The type
+system uses $\Theta$ to make such reasoning more precise.
+$\Theta$ is a map from variables $x$ to
+equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
+It maps variables to equations that are recorded along the type checking procedure.
+If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
+$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
+such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
+\iftr
+  Appendix~\ref{app:le}.
+\else
+  the supplemental report~\cite{checkedc-tech-report}.
+\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
+
+\begin{figure}[t]
+{\small
+{\captionsetup[lstlisting]{margin = 8 mm}
+  \begin{lstlisting}[xleftmargin=8 mm]
+nt_array_ptr<char> safe_strcat
+   (nt_array_ptr<char> dst : count(n),
+    nt_array_ptr<char> src : count(0), int n) {
+  int x = strlen(dst);
+  int y = strlen(src);
+  nt_array_ptr<char> c : count(n) =
+    dyn_bounds_cast
+           <nt_array_ptr<char>>(dst,count(n));
+    // sets c == dst with bound n (not x)
+  if (x+y < n) {
+    for (int i = 0; i < y; ++i)
+      *(c+x+i) = *(src+i);
+    *(c+x+y) = '\0';
+    return dst;
+  }
+  return null;
+}
+  \end{lstlisting}
+}
+}
+\caption{Implementation of safe \code{strcat}}
+\label{fig:strcat-ex}
+\end{figure}
+
+Recall that array
+bounds in types may refer to in-scope variables; e.g., parameter
+\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
+2-3 in \Cref{fig:strcat-ex}. 
+All dependent bound variables appearing in function headers 
+are bounded by a variable list in the function pointer type, which is checked in the type rule.
+Semantically,
+the call is expanded into a \texttt{let} which binds
+parameter variables $\overline{x}$ to the actual arguments
+$\overline{n}$, but annotated with the parameter types
+$\overline{\tau}$ (this will be safe for type-correct programs). 
+The function body $e$ is wrapped in a static cast
+$(\tau[\overline{n} / \overline{x}])$ which is the function's return
+type but with any integer parameter variables $\overline{x}$ appearing in that
+type, as type bound variables,
+substituted with the call's actual arguments $\overline{n}$.
+To see why this is needed, suppose that \code{safe_strcat} in
+Fig.~\ref{fig:strcat-ex} is defined to return a
+\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
+perform a \code{safe_strcat} function call as
+\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
+function returns a value with type \code{nt_array_ptr<int>:count(10)}
+because we substitute bound variable \code{n} in the 
+defined return type with \code{10} from the function call's
+argument list.
+
+\begin{figure}[t!]
+  \begin{small}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+int deref_array(n : int,
+     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
+     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
+  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
+  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
+    * p;
+    * q = 1;
+}
+...
+/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
+/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\begin{frame}
+
+\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
+
+\end{frame}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+deref_array(int n, int* p, int * q) {
+  //m is the current context mode
+  let p_lo = 0; let p_hi = n; 
+  let q_lo = 0; let q_hi = n; 
+  /* runtime checks */
+  assert(p_lo <= 0 && 0 <= p_hi);
+  assert(p != 0);
+  *(mode(p) $\wedge$ m,p);
+  verify(q, not_null(m, q_lo, q_hi) 
+             && q_lo <= 0 && 0 <= q_hi);
+  *(mode(q) $\wedge$ m,q)=1;
+}
+...
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\end{small}
+    \caption{Compilation Example for Dependent Functions}
+\label{fig:compilationexample1}
+\end{figure}
+
+For compiling dependent function calls,
+\Cref{fig:compilationexample1} provides a hint.
+Notice that the bounds for the array pointer \code{p} are not passed as
+arguments. Instead, they are initialized according to \code{p}'s
+type---see line~4 of the original \lang program at the top of the figure.
+Line~$3$ of the generated code
+sets the lower bound  to \code{0} and the
+upper bound to \code{n}.
+
+\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
+A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
+\end{defi}
+
+\begin{defi}[Heap Well-formedness]
+For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
+(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
+contains no free variables. 
+\end{defi}
+
+\begin{defi}[Stack Well-formedness]
+A stack snapshot $\varphi$ is well-formed if
+for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
+contains no free variables. 
+\end{defi}
+
+We also need to introduce a notion of
+\emph{consistency}, relating heap environments before and after a
+reduction step, and type environments, predicate sets, and stack
+snapshots together.
+
+\begin{defi}[Stack Consistency]
+A type environment $\Gamma$, variable predicate set $\Theta$, and
+stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
+\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
+$\Gamma(x) = \tau$ for some $\tau$ and 
+$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
+\end{defi}
+
+\begin{defi}[Checked Stack-Heap Consistency]
+A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
+for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
+\end{defi}
+
+\begin{defi}[Checked Heap-Heap Consistency]
+A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
+for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
+\end{defi}
+
+\myparagraph{Constant Validity}
+Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
+describe type assumptions for constants appearing in a program.
+$\neg \cmode(\tau)$ judges that a constant pointer 
+in an unchecked region cannot be of a checked type.
+The restriction ensures that programmers 
+cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
+In rule \textsc{T-ConstC}, we requires a static 
+verification procedure for validating a constant pointer in \Cref{fig:const-type}.
+
+The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
+validates the constant $\evalue{n}{\tau}$, 
+where $\heap(m)$ is the initial heap that the constant resides on and
+$\sigma$ is a set of constant assumed to be checked.
+A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
+A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
+and the address stores a function with the right type.
+The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
+where every element in the pointer range ($[0,\size(\omega))$) should be well
+typed.
+A checked pointer checks validity in type step as rule \textsc{T-ConstC},
+while a tainted/unchecked pointer does not check for such during the type checking.
+Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.
 
@@
 
-\subsection{Typing Rules for Literal Pointers}\label{sec:literal-pointer-typing}
-
-The typing of integer literals, which can also be pointers to the
-heap, was presented in Sec.~\ref{sec:theorem} in Fig.~\ref{fig:const-type}. Here
-we describe these rules further.
-
-The variable type rule
-($\textsc{T-Var}$) simply checks if a given variable has the defined
-type in $\Gamma$; the constant rule ($\textsc{T-Const}$) is slightly
-more involved.  First, it ensures that the type annotation $\tau$ does
-not contain any free variables. More importantly, it ensures that the
-literal itself is well typed using an auxilliary typing relation
-$\heap;\sigma \vdash n : \tau$.
-
-If the literal's type is an integer, an unchecked pointer, or a null
-pointer, it is well typed, as shown by the top three rules in
-Fig.~\ref{fig:const-type}. However, if it is a checked pointer
-$\tptr{\omega}{\cmode}$, we need to ensure that what it points to in
-the heap is of the appropriate pointed-to type ($\omega$), and also
-recursively ensure that any literal pointers reachable this way are
-also well-typed. This is captured by the bottom rule in the figure,
-which states that for every location $n+i$ in the pointers' range
-%
-$[n, n+\size(\omega))$, where $\size$ yields the size of its argument,
-  then the value at the location $\heap(n+i)$ is also well-typed.
-  However, as heap snapshots can contain cyclic structures (which
-  would lead to infinite typing deriviations), we use a scope $\sigma$
-  to assume that the original pointer is well-typed when checking the
-  types of what it points to. The middle rule then accesses the scope
-  to tie the knot and keep the derivation finite, just like in
-  \citet{ruef18checkedc-incr}.
-
-\subsection{Other Semantic Rules}\label{sec:rem-semantics}
-
-Fig.~\ref{fig:rem-semantics} shows the remaining semantic rules for
-$\lang$. We explain a selected few rules in this subsection.
-% other few low-level semantic rules for variable and dereference and $\emalloctext$ operations in \checkedc. Other operations are defined in the same manner.
-
-\begin{figure*}[t]
-{\small
-\begin{mathpar}
-
-\inferrule [S-Var]{} {(\varphi,\heap,x)\longrightarrow (\varphi,\heap,\varphi(x))}
-
-    \inferrule[S-DefArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-DefArrayBound]{0 \not\in [n_l,n_h)}
-     { (\varphi,\heap,\estar{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-    \inferrule[S-DefNTArrayBound]{0 \notin [n_l,n_h]}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-    \inferrule[S-Assign]{\heap(n)=\evalue{n_a}{\tau_a} }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tptr{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}
-
-    \inferrule[S-AssignNull]{}
-      {(\varphi,\heap,\eassign{\evalue{0}{\tptr{\omega}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\enull)}
-
-    \inferrule[S-AssignArrBound]{0 \not\in [n_l,n_h) }
-      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-  \inferrule[S-Malloc]{\varphi(\omega)=\omega_a \\ \mathtt{alloc}(\heap,\omega_a)=(n,\heap')}
-   { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\evalue{n}{\tptr{\omega_a}{\cmode}})}
-
-  \inferrule[S-MallocBound]{\varphi(\omega)=\tallarray{n_l}{n_h}{\tau}\\ (n_l \neq 0 \vee n_h \le 0)}
-    { (\varphi,\heap,\emalloc{\omega}) \longrightarrow (\varphi,\heap',\ebounds)}
-
-    \inferrule[S-IfT]{n \neq 0 }
-    {(\varphi,\heap,\eif{\evalue{n}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
-
-    \inferrule[S-IfF]{}
-    {(\varphi,\heap,\eif{\evalue{0}{\tau}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}
-
-    \inferrule[S-Unchecked]{}
-    {(\varphi,\heap,\eunchecked{\evalue{n}{\tau}} \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
-    \inferrule[S-Str]{
-        0 \in [n_l,n_h]
-\\ n_a \le n_h
-\\ \heap(n+n_a) = 0 
-\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
-    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tint})}
-
-    \inferrule[S-StrBounds]{
-        0 \notin [n_l,n_h]
-}
-    {(\varphi,\heap,\estrlen{\evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\ebounds)}
-
-    \inferrule[S-StrNull]{}
-    {(\varphi,\heap,\estrlen{\evalue{0}{\tarrayptr{n_l}{n_h}{\tau}{c}}}) \longrightarrow (\varphi,\heap,\enull)}
-
-    \inferrule[S-Add]{n = n_1 + n_2}
-    {(\varphi,\heap,\evalue{n_1}{\tint} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, n)}
-
-    \inferrule[S-AddArr]{n = n_1 + n_2\\ n_l' = n_l - n_2 \\ n_h' = n_h - n_2}
-    {(\varphi,\heap,\evalue{n_1}{\tallarrayptr{n_l}{n_h}{\tau}{m}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau}{m}})}
-
-n    \inferrule[S-AddArrNull]{}
-    {(\varphi,\heap,\evalue{0}{\tallarrayptr{n_l}{n_h}{\tau}{c}} \plus \evalue{n_2}{\tint}) \longrightarrow (\varphi,\heap, \enull)}
-
-\end{mathpar}
-
-}
-\caption{Remaining \lang Semantics Rules (extends Fig.~\ref{fig:semantics})}
-\label{fig:rem-semantics}
-\end{figure*}
-
-Rule \textsc{S-Var} loads the value for $x$ in stack $\varphi$.
-Rule \textsc{S-DefArray} dereferences an array pointer, which is similar to the Rule \textsc{S-DefNTArray} in Fig.~\ref{fig:semantics} (dealing with null-terminated array pointers).
-The only difference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
-Rules \textsc{DefArrayBound} and \textsc{DefNTArrayBound} describe an error case for a dereference operation.
-If we are dereferencing an array/NT-array pointer and the mode is $\cmode$, $0$ must be in the range from $n_l$ to $n_h$ (meaning that the dereference is in-bound); if not, the system results in a $\ebounds$ error.
-Obviously, the dereference of an array/NT-array pointer also experiences a $\enull$ state transition if $n\le 0$.
-
-Rules \textsc{S-Malloc} and \textsc{S-MallocBound} describe the $\emalloctext$ semantics.
-Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$, and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation.
-The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\cmode}}$ and new updated heap.
-It is possible for $\emalloctext$ to transition to a $\ebounds$ error if the $\omega_a$ is an array/NT-array type $\tallarray{n_l}{n_h}{\tau}$, and either $n_l \neq 0$ or $n_h \le 0$.
-This can happen when the bound variable is evaluated to a bound constant that is not desired.
-
-\subsection{Subtyping for dependent types}
-\label{app:le}
-
-The subtyping relation given in Fig.~\ref{fig:checkc-subtype} involves
-dependent bounds, i.e., bounds that may refer to variables. To decide
-premises $b \leq b'$, we need a decision procedure that accounts for
-the possible values of these variables. This process considers
-$\Theta$, tracked by the typing judgment, and $\varphi$, the current
-stack snapshot (when performing subtyping as part of the type
-preservation proof).
-
-\begin{defi}[Inequality]
-
-\begin{itemize}
-
-\item $n \le m$ if $n$ is less than or equal to $m$.
-\item $x+n \le x + m$ if $n$ is less than or equal to $m$.
-\item All other cases result in $\efalse$.
-
-\end{itemize}
-\end{defi}
-
-To capture bound variables in dependent types, the \checkedc subtyping
-relation ($\sqsubseteq$) is parameterized by a restricted stack
-snapshot $\varphi|_{\rho}$ and the predicate map $\Theta$, where
-$\varphi$ is a stack and $\rho$ is a set of
-variables. $\varphi|_{\rho}$ means to restrict the domain of $\varphi$
-to the variable set $\rho$. Clearly, we have the relation:
-$\varphi|_{\rho} \subseteq \varphi$. $\sqsubseteq$
-being parameterized by $\varphi|_{\rho}$ refers to that when we
-compare two bounds $b \le b'$, we actually do
-$\varphi|_{\rho}(b) \le \varphi|_{\rho}(b')$ by interpreting the
-variables in $b$ and $b'$ with possible values in $\varphi|_{\rho}$.
-Let's define a subset relation $\preceq$ for two restricted stack
-snapshot $\varphi|_{\rho}$ and $\varphi'|_{\rho}$:
-
-\begin{defi}[Subset of Stack Snapshots]
-  Given two $\varphi|_{\rho}$ and $\varphi'|_{\rho}$,
-  $\varphi|_{\rho} \preceq \varphi'|_{\rho}$, iff for $x\in\rho$ and
-  $y$,
-  $(x,y) \in \varphi|_{\rho} \Rightarrow (x,y) \in \varphi'|_{\rho}$.
-\end{defi}
-
-For every two restricted stack snapshots $\varphi|_{\rho}$ and
-$\varphi'|_{\rho}$, such that
-$\varphi|_{\rho} \preceq \varphi'|_{\rho}$, we have the following
-theorem in \checkedc (proved in Coq):
-
-\begin{thm}[Stack Snapshot Theorem]
-  Given two types $\tau$ and $\tau'$, two restricted stack snapshots
-  $\varphi|_{\rho}$ and $\varphi'|_{\rho}$, if
-  $\varphi|_{\rho}\preceq \varphi'|_{\rho}$, and
-  $\tau \sqsubseteq \tau'$ under the parameterization of
-  $\varphi|_{\rho}$, then $\tau \sqsubseteq \tau'$ under the
-  parameterization of $\varphi'|_{\rho}$.
-\end{thm}
-
-Clearly, for every $\varphi|_{\rho}$, we have
-$\emptyset \preceq \varphi|_{\rho}$. The type checking stage is a
-compile-time process, so $\varphi|_{\rho}$
-is $\emptyset$ at the type checking stage. Stack snapshots are needed
-for proving type preserving, as variables in bounds expressions are
-evaluated away.
-
-\begin{figure}[h!]
-{\small
-  \begin{mathpar}
-    \inferrule[T-Def]
-              {\Gamma;\Theta \vdash_m e : \tptr{\tau}{m'} \\
-                m \leq m'}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-
-    \inferrule[T-Mac]
-              {}
-              {\Gamma; \Theta \vdash_m \emalloc{\omega} : \tptr{\omega}{\cmode}}
-
-    \inferrule[T-Add]
-              {\Gamma; \Theta \vdash_m e_1 : \tint \\
-                \Gamma; \Theta \vdash_m e_2 : \tint}
-              {\Gamma; \Theta \vdash_m (e_1 \plus e_2) : \tint }
-
-    \inferrule[T-Ind] 
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'} \\
-                \Gamma; \Theta \vdash_m e_2 : \tint \\
-                m \leq m'}              
-              {\Gamma; \Theta \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
-
-    \inferrule[T-Assign]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tau}{m'} \\
-                \Gamma; \Theta \vdash_m e_2 : \tau' \\
-                \tau'\sqsubseteq \tau \\
-                m \leq m'}
-              {\Gamma; \Theta \vdash_m \eassign{e_1}{e_2} : \tau}
-
-   \inferrule[T-IndAssign]
-              {\Gamma; \Theta \vdash_m e_1 : \tptr{\tallarrayb{\bvar}{\tau}}{m'}\\
-                \Gamma; \Theta \vdash_m e_2 : \tint \\
-                \Gamma; \Theta \vdash_m e_3 : \tau' \\
-                \tau'\sqsubseteq \tau \\
-                m \leq m'}
-              {\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
-
-  \end{mathpar}
-}
-\caption{Remaining \lang Type Rules (extends Fig.~\ref{fig:type-system-1})}
-\label{fig:rem-type-system}
-\end{figure}
-
-As mentioned in the main text, $\sqsubseteq$ is also parameterized by
-$\Theta$, which provides the range of allowed values for a bound
-variable; thus, more $\sqsubseteq$ relation is provable. For example,
-in Fig.~\ref{fig:strcat-ex}, the \code{strlen} operation in line 4
-turns the type of \code{dst} to be $\tntarrayptr{0}{x}{\tint}{\cmode}$
-and extends the upper bound to \code{x}. In the \code{strlen} type
-rule, it also inserts a predicate \code{x}$\ge 0$ in $\Theta$; thus,
-the cast operation in line 16 is valid because
-$\tntarrayptr{0}{x}{\tint}{\cmode} \sqsubseteq
-\tntarrayptr{0}{0}{\tint}{\cmode}$ is provable when we know
-\code{x}$\ge 0$.
-
-Note that if $\varphi$ and $\Theta$ are $\emptyset$, we do only the
-syntactic $\le$ comparison; otherwise, we apply $\varphi$ to both
-sides of $\sqsubseteq$, and then determine the $\le$ comparasion based
-on a Boolean predicate decision procedure on top of $\Theta$. This
-process allows us to type check both an input expression and the
-intermediate expression after evaluating an expression.
-
-\subsection{Other Type Rules}\label{rem-type}
-
-Here we show the type rules for other \checkedc operations in Fig.~\ref{fig:rem-type-system}.
-Rule \textsc{T-Def} is for dereferencing a non-array pointer. 
-The statement $m \leq m'$ ensures that no unchecked pointers are used in checked regions.
-Rule \textsc{T-Mac} deals with
-$\emalloctext$ operations. There is a well-formedness check to require
-that the possible bound variables in $\omega$ must be in the domain of
-$\Gamma$ (see Fig.~\ref{fig:wftypesandbounds}).
-This is similar to the well-formedness assumption of the type environment (Definition~\ref{type-wellformed}) Rule \textsc{T-Add} deals with binary operations whose sub-terms are integer expressions, while rule \textsc{T-Ind} serves the case for pointer arithmetic.
-For simplicity, in the \checkedc formalization, we do not allow arbitrary pointer arithmetic.
-The only pointer arithmetic operations allowed are the forms shown in rules \textsc{T-Ind} and \textsc{T-IndAssign} in Fig.~\ref{fig:rem-type-system}.
-Rule \textsc{T-Assign} assigns a value to a non-array pointer location.
-The predicate $\tau'\sqsubseteq \tau$ requires that the value being assigned is a subtype of the pointer type.
-The \textsc{T-IndAssign} rule is an extended assignment operation for handling assignments for array/NT-array pointers with pointer arithmetic.
-Rule \textsc{T-Unchecked} type checks \code{unchecked} blocks.
-
 \subsection{Struct Pointers}\label{appx:struct}
@@
   \label{fig:checkc-struct}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-
-    \inferrule[]
-    {}
-    {\Gamma \vdash n}
-
-    \inferrule[]
-    {x:\tint \in \Gamma}
-    {\Gamma \vdash x + n}
-
-    \inferrule[]
-    {\Gamma \vdash b_l\\
-    \Gamma \vdash b_h}
-  {\Gamma \vdash (b_l,b_h)}
-
-  \inferrule[]
-  {}
-  {\Gamma \vdash \tint}
-
-  \inferrule[]
-  {\Gamma \vdash \bvar \\
-  \Gamma \vdash \tau}
-  {\Gamma \vdash \tptr{\tallarrayb{\bvar}{\tau}}{m}}
-
-  \inferrule[]
-  {\Gamma \vdash \tau}
-  {\Gamma \vdash \tptr{\tau}{m}}
-
-  \inferrule[]
-  {T \in D}
-  {\Gamma \vdash \tptr{\tstruct{T}}{m}}
-  \end{mathpar}
-}
- \caption{Well-formedness for Types and Bounds}
-\label{fig:wftypesandbounds}
-\end{figure}
-
-% \inferrule[T-Fun]
-%     {\Xi(f) = \tau\;(\overline{x}:\overline{\tau})\;e \\\\
-%         \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
-%              \overline{\tau'} \sqsubseteq 
-%                \overline{\tau}[\overline{e} / \overline{x}]}
-%     {\Gamma; \Theta \vdash_m f(\overline{e}) : \tau[\overline{e} / \overline{x}]}
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    {\Gamma \vdash \overline{x}:\overline{\tau} \\
-      \Gamma[\overline{x} \mapsto \overline{\tau}] \vdash \tau \\
-    \Gamma[\overline{x} \mapsto \overline{\tau}]; \Theta  \vdash_{\cmode} e : \tau}
-    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e}
-
-    \inferrule[]
-    {}
-    {\Gamma \vdash \cdot}
-
-    \inferrule[]
-    {\Gamma \vdash \tau \\
-    \Gamma[x \mapsto \tau] \vdash \overline{x}:\overline{\tau}}
-    {\Gamma \vdash x:\tau, \overline{x}:\overline{\tau}}
-  \end{mathpar}
-}
- \caption{Well-formedness for functions}
-\label{fig:wffunctions}
 \end{figure}
@@
 
-\subsection{Additional Type and Semantic Rules}\label{appx:add-type-sem}
-
-\begin{DIFnomarkup}
-\begin{figure}
-{\small
-\[\hspace*{-1.2em}
-\begin{array}{l}
-\textcolor{blue}{\text{Bound Inequality and Equality:}}\\[0.3em]
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-     n \le n' &\Rightarrow& n &\le_{\Theta} & n'\\
-     n \le n' &\Rightarrow& x+n &\le_{\Theta} & x+n'\\
-     n \le n' \wedge \Theta(x)=\tgez &\Rightarrow& n &\le_{\Theta} & x+n'\\
-     \Theta(x)=\teq{b} \wedge b+n\le_{\Theta}b'  &\Rightarrow& x+n & \le_{\Theta} & b'\\
-     \Theta(x)=\teq{b}\wedge b'\le_{\Theta}b+n  &\Rightarrow& b' & \le_{\Theta} & x+n\\
-     b \le_{\Theta} b' \wedge b' \le_{\Theta} b  &\Rightarrow& b & =_{\Theta} & b'
-    \end{array}
-  \\[1.5em]
-\textcolor{blue}{\text{Type Equility:}}\\
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-     && \tint & =_{\Theta} & \tint\\
-     \omega =_{\Theta} \omega' &\Rightarrow& \tptr{\omega}{\xi} & =_{\Theta} & \tptr{\omega'}{\xi}\\
-     \bvar =_{\Theta} \bvar' \wedge  \tau =_{\Theta} \tau'
-             &\Rightarrow& \tallarrayb{\bvar}{\tau} & =_{\Theta} & \tallarrayb{\bvar'}{\tau'}\\
-
-    \textit{cond}(\overline{x},\overline{\tau}\to\tau,\overline{y},\overline{\tau'}\to\tau')
-
- &\Rightarrow& \tfun{\overline{x}}{\overline{\tau}}{\tau} & 
-                         =_{\Theta} & \tfun{\overline{y}}{\overline{\tau'}}{\tau'}\\
-    \end{array}
-  \\[0.7em]
-\textcolor{blue}{\text{Subtype:}}\\[0.3em]
-
-  \begin{array}{r@{~}c@{~}l@{~}c@{~}l}
-    \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
-
-    0\le_{\Theta} b_l \wedge b_h \le_{\Theta} 1 &\Rightarrow& \tptr{\tau}{m}&\sqsubseteq_{\Theta}& \tarrayptr{b_l}{b_h}{\tau}{m}\\[0.2em]
-    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
-    b_l \le_{\Theta} 0 \wedge 1 \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tptr{\tau}{m}\\[0.2em]
-    %% b_l \le b_l' \wedge b_h' \le b_h &\Rightarrow&  \tarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq&  \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tntarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tarrayptr{b_l'}{b_h'}{\tau}{m}\\[0.6em]
-    b_l \le_{\Theta} b_l' \wedge b_h' \le_{\Theta} b_h &\Rightarrow& \tallarrayptr{b_l}{b_h}{\tau}{m} &\sqsubseteq_{\Theta}& \tallarrayptr{b_l'}{b_h'}{\tau}{m}
-\\[0.2em]
-\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}\wedge \tau\sqsubseteq_{\Theta}\tau' &\Rightarrow& \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} &\sqsubseteq_{\Theta}& \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}
-
-    \end{array}
-\end{array}
-  \]
-}
-{\footnotesize
-\[
-\begin{array}{l}
-n'+n = add(n',n)
-\qquad
-(x+n')+n = x+add(n',n)\\
-\textit{cond}(\overline{x},\tau,\overline{y},\tau')
-=\exists\overline{z}\;.\;\overline{x}\cupdot\overline{z}
-  \wedge \overline{y}\cupdot\overline{z}
-  \wedge \size(\overline{x})=\size(\overline{y})=\size(\overline{z})
-\\\qquad\qquad\qquad\qquad\qquad
-  \wedge \tau[\overline{z}/\overline{x}]= \tau'[\overline{z}/\overline{x}]
-\end{array}
-\]
-}
-  \caption{Type Equality and Subtyping}
-  \label{fig:checkc-subtype}
-\end{figure}
-\end{DIFnomarkup}
-
 Since bounds expressions may
@@
 The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.
-
-\myparagraph{Constant Validity}
-Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
-describe type assumptions for constants appearing in a program.
-$\neg \cmode(\tau)$ judges that a constant pointer 
-in an unchecked region cannot be of a checked type.
-The restriction ensures that programmers 
-cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
-In rule \textsc{T-ConstC}, we requires a static 
-verification procedure for validating a constant pointer in \Cref{fig:const-type}.
-
-The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
-validates the constant $\evalue{n}{\tau}$, 
-where $\heap(m)$ is the initial heap that the constant resides on and
-$\sigma$ is a set of constant assumed to be checked.
-A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
-A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
-and the address stores a function with the right type.
-The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
-where every element in the pointer range ($[0,\size(\omega))$) should be well
-typed.
-A checked pointer checks validity in type step as rule \textsc{T-ConstC},
-while a tainted/unchecked pointer does not check for such during the type checking.
-Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.

### 2023-01-15 sandp23/appendix.tex
update appx semantics and compilation.

@@
 
+\myparagraph{Checked and Tainted Pointer Operations}
 Rule \textsc{S-Var} loads the value for $x$ in stack $\varphi$.
-Rule \textsc{S-DefArray} dereferences an array pointer, which is similar to the Rule \textsc{S-DefNTArray} in Fig.~\ref{fig:semantics} (dealing with null-terminated array pointers).
-The only difference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
+Rules \textsc{S-DefArrayC} and \textsc{S-DefArrayT} dereference an $\cmode$ and $\tmode$ array pointer, respectively.
+In \lang, the difference between array and NT-array dereference is that the range of $0$ is at $[n_l,n_h)$ not $[n_l,n_h]$, meaning that one cannot dereference the upper-bound position in an array.
 Rules \textsc{DefArrayBound} and \textsc{DefNTArrayBound} describe an error case for a dereference operation.
-If we are dereferencing an array/NT-array pointer and the mode is $\cmode$, $0$ must be in the range from $n_l$ to $n_h$ (meaning that the dereference is in-bound); if not, the system results in a $\ebounds$ error.
+If we are dereferencing an array/NT-array pointer and the mode is $\cmode$ (or $\tmode$), $0$ must be in the range from $n_l$ to $n_h$ (meaning that the dereference is in-bound); if not, the system results in a $\ebounds$ error.
 Obviously, the dereference of an array/NT-array pointer also experiences a $\enull$ state transition if $n\le 0$.
@@
 Rules \textsc{S-Malloc} and \textsc{S-MallocBound} describe the $\emalloctext$ semantics.
-Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$, and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation.
-The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\cmode}}$ and new updated heap.
+Given a valid type $\omega_a$ that contains no free variables, $\mathtt{alloc}$ function returns an address pointing at the first position of an allocated space whose size is equal to the size of $\omega_a$ for a specific mode $\xi$,
+and a new heap snapshot $\heap'$ that marks the allocated space for the new allocation. 
+If $\xi$ is $\cmode$, the new $\heap$ allocation is in the $\cmode$ region, while $\tmode$ and $\umode$ mode allocation creates new pieces in $\umode$ region.
+The $\emalloctext$ is transitioned to the address $n$ with the type ${\tptr{\omega_a}{\xi}}$ and new updated heap.
 It is possible for $\emalloctext$ to transition to a $\ebounds$ error if the $\omega_a$ is an array/NT-array type $\tallarray{n_l}{n_h}{\tau}$, and either $n_l \neq 0$ or $n_h \le 0$.
 This can happen when the bound variable is evaluated to a bound constant that is not desired.
+
+Rules \textsc{S-AssignArrC} and \textsc{S-AssignArrT} are for assignment operations.
+\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
+dereference) is within the bounds designated by the pointer's annotation
+and strictly less than the upper bound. 
+Rule \textsc{S-AssignArrT} is similar to \textsc{S-AssignArrC} for tainted pointers.
+Any dynamic heap access of a tainted pointer requires a \textit{verification}.
+Performing such a verification equates to performing a literal type check for a pointer constant in \Cref{fig:const-type}.
+
+\myparagraph{Let Bindings}
+%
+The semantics manages variable scopes using the special $\erettext$
+form. \textsc{S-Let} evaluates to a configuration whose expression is
+$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
+and remember $x$ and its new value $\evalue{n}{\tau}$
+in $e$'s scope that is defined by the $\erettext$ operation.
+Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
+we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope.
+After one-step evaluation is completed, 
+we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
+and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
+This procedure continues until $e'$ becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
+the literal.
+
+\subsection{Compilation Formalism}\label{sec:compilation}
+
+% \review{
+% The first reviewer said:
+% \begin{itemize}
+% \item I had a really hard time understanding precisely the "formalized" compilation
+%   from CoreChkC to CoreC. Specifically: is CoreC intended to model LLVM-IR, or
+%   is it a subset of C? Is CheckedC compiling to C or is it a new frontend like
+%   `clang`? See remark about undefined behavior (UB) below which got me even more
+%   confused. \mwh{CoreC is meant to model CoreChkC but with
+%   annotations, and checks removed; it is not meant to model C or LLVM-IR.}
+% \item Is the compilation scheme from CoreChkC to CoreC is faithful to the actual
+%   compilation scheme of the checkedc-clang compiler to ...? I feel like the
+%   paper is missing an actual description of what happens in the compiler to
+%   allow us to connect the dots and understand how the formalization illuminates
+%   the implementation. \mwh{We did not model compilation on the real
+%   implementation; our purpose was to show that annotations in CoreChkC
+%   do not necessitate fat pointers in an implementation; that said,
+%   our formalization does show how a real implementation can be
+%   carried out}
+% \item The paper doesn't seem upfront about what is *shown* (theorem in Coq) and what
+%   is *tested* (via PLT-Redex), and thus remains a
+%   conjecture. \mwh{Updated the intro and the individual sections}
+% \item  Missing discussion of why Coq vs. PLT-Redex, effort involved, any plans to
+%   formally prove compilation from CoreChkC to CoreC, any hopes of integrating
+%   that in the official implementation, etc. \mwh{Added note at the
+%   start of section 3.}
+% \end{itemize} }
+
+% \review{IV: ghost variables in other contexts (e.g. Why3, Dafny) are used for things
+%   that do not exist at run-time, but this doesn't seem to be the case here.}
+% \yiyun{Agreed: We changed the name to ``shadow variables'' to avoid confusion}
+% \review{Do your bug report and github links break anonymity?}
+
+% \review{From reviewer C:
+% \begin{itemize}
+% \item I'd like to have seen a bit more motivation for using PLT redex:
+%   what aspects made the use of this tool preferable to formulating the
+%   compilation in Coq and using Quickchick to do random testing of the
+%   simulation result.
+%   \mwh{Added some text to the end of III.A and IV.C}
+% \yiyun{Some reasons I can think of:\begin{itemize}
+%     \item Redex is highly optimized for specifying judgments that are algorithmic. By writing down a typing relation, we can immediately obtain a typechecker
+%     that is executable.
+Same applies for the small-step evaluation relation.
+Translating the relations into functions in Coq is definitely doable but time-consuming, especially since compilation is embedded as part of the typing rules.
+It is also hard to see whether the function we define really corresponds to the relation unless we formally prove it.
+This issue is particularly relevant at the early stage of the development when the compilation rules were buggy and the simulation property was violated often as we added new generator cases.
+%     In Redex, we don't have any formal guarantee either, but at least we can more easily see the correspondence because Redex is able to convert the relation into an executable version so we can specify the relation literally.
+This feature of Redex helped us speed up our development significantly when our compilation rules were constantly changing.
+% \end{itemize} }
+% \item there's a funny change of line spacing in column 2 of page 6, about 2/3 down.
+% \end{itemize}
+% }
+
+The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
+that track array bound information, which is used in premises
+of rules like \textsc{S-DefArray} and
+  \textsc{S-AssignArr} to prevent spatial safety violations.
+The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
+for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
+The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.
+
+\begin{figure}[t!]
+{\small
+\hspace*{-0.5em}
+\begin{tabular}{|c|c|c|c|}
+\hline
+& \cmode & \tmode & \umode \\
+\hline
+& \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} \\
+\hline
+\cmode & $\estar{x}$ / $\getstar{\cmode}{x}$ 
+ & $\texttt{sand\_get}(x)$ / $\getstar{\umode}{x}$ &  $\times$ \\
+\hline
+\umode & $\times$
+ & $\estar{x}$ / $\getstar{\umode}{x}$ &  $\estar{x}$ / $\getstar{\umode}{x}$ \\
+\hline
+\end{tabular}
+
+}
+\caption{Compiled Targets for Dereference}
+\label{fig:flagtable}
+\end{figure}
+
+In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
+i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
+\lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
+Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
+In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different.
+\lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
+\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
+The columns represent different pointer modes and the rows represent context modes.
+For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation.
+In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
+If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
+
+This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
+with no loss of expressiveness
+as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
+For the compiler formalism, 
+we present a compilation algorithm that converts from
+\lang to \elang, an untyped language without metadata
+annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
+In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
+$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
+The algorithm sheds
+  light on how compilation can be implemented in the real Checked C
+  compiler, while eschewing many vital details (\elang has many 
+  differences with LLVM IR).
+
+%This section shows how \systemname deals with 
+%annotations can be safely erased: using static information a compiler
+%can insert code to manage and check bounds metadata, with no loss of
+%expressiveness. We present a compilation algorithm that converts from
+%\lang to \elang, an untyped language without metadata
+%annotations. The syntax and semantics \elang
+  %closely mirrors that of \lang; it differs only in that literals lack
+  %type annotations and its operational rules perform no
+  %bounds and null checks, which are instead inserted during
+  %compilation. Our compilation algorithm is evidence that \lang's
+  %semantics, despite its apparent use of fat pointers, faithfully
+  %represents Checked C's intended behavior. The algorithm also sheds
+  %light on how compilation can be implemented in the real Checked C
+  %compiler, while eschewing many important details (\elang has many 
+  %differences with LLVM IR).
+
+Compilation is defined by extending \lang's
+typing judgment as follows:
+\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
+There is now a \elang output $\dot e$ and an input $\rho$, which maps
+each (NT-)array pointer variable to its mode and
+each variable \code{p} to a pair of \emph{shadow
+  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
+These may differ from the bounds in \code{p}'s type due to bounds
+widening.\footnote{Since lower bounds are never widened, the
+  lower-bound shadow variable is unnecessary; we include it for uniformity.} 
+\lc{These two sentences are confusing. }
+
+% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
+% complete judgment, implicitly assuming that $e$ is a well-typed and closed
+% term.
+
+We formalize rules for this judgment in PLT Redex~\cite{pltredex},
+following and extending our Coq development for \lang. To give
+confidence that compilation is correct, we use Redex's property-based
+random testing support to show that compiled-to $\dot e $ simulates
+$e$, for all $e$.
+
+% We developed a \checkedc compiler to compile a \checkedc program to a C program.
+% \mwh{We formalized compilation from CoreChkC to a version of CoreChkC
+%   but with the metadata removed, right? This is not a Checked C
+%   compiler. You go on to see stuff about CompCert, CLight,
+%   etc. This is confusing. We should be talking about how this relates
+%   to what was just presented. Pick definitive names for things. }
+% Given a \checkedc program $e$, we build a compilation process ($\gg$), such that $e \gg \dot e$, where $\dot e$ is the corresponding C program for $e$ in A-normal form (ANF). 
+% The compilation process ($\gg$) relies on the type checking step $\Gamma;\Theta\vdash_m e:\tau$. 
+% Especially, it relies on $\Gamma$ to provide the type information for variables in $e$. 
+% We utilize CompCert/CLight syntax and semantics \cite{Leroy:2009:FVC:1666192.1666216,Blazy2009} as our translation target language.
+%  Besides, we defined two data structures in the CompCert format for representing $\enull$ and $\ebounds$ states.
+% We write $\xrightarrow{c}$ for the semantics of CLight. \mwh{Of our
+%   target language, which I presume does not have all the features that
+%   CLight has. Should mention up front that we did all of this in PLT Redex.}
+
+\subsection{Approach}
+
+Due to space constraints, we explain the rules for compilation by
+examples, using a C-like syntax; the complete rules are given in
+\iftr
+Appendix~\ref{appx:comp1}.
+\else
+the supplemental report~\cite{checkedc-tech-report}.
+\fi
+Each rule performs up to three tasks: (a) conversion of $e$ to
+A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
+and (c) generate right pointer accessing expressions based on modes.
+%
+A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
+as in the following example.
+
+{\vspace*{-0.5em}
+{\small
+\begin{center}
+$
+\begin{array}{l}
+$\code{let y=(x+1)+(6+1)}$
+\;
+\begin{frame}
+
+\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
+
+\end{frame}
+\;
+\begin{array}{l}
+$\code{let a=x+1;}$\\
+$\code{let b=6+1;}$\\
+$\code{let y=a+b}$\\
+\end{array}
+\end{array}
+$
+\end{center}
+}
+}
+
+This simplifies the management of effects from subexpressions. The
+next two steps of compilation are more interesting.
+We state them based on different \lang operations.
+
+\begin{figure}[t!]
+  \begin{small}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+int deref_array(n : int,
+     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
+     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
+  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
+  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
+    * p;
+    * q = 1;
+}
+...
+/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
+/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\begin{frame}
+
+\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
+
+\end{frame}
+\begin{lstlisting}[mathescape,xleftmargin=4 mm]
+deref_array(int n, int* p, int * q) {
+  //m is the current context mode
+  let p_lo = 0; let p_hi = n; 
+  let q_lo = 0; let q_hi = n; 
+  /* runtime checks */
+  assert(p_lo <= 0 && 0 <= p_hi);
+  assert(p != 0);
+  *(mode(p) $\wedge$ m,p);
+  verify(q, not_null(m, q_lo, q_hi) 
+             && q_lo <= 0 && 0 <= q_hi);
+  *(mode(q) $\wedge$ m,q)=1;
+}
+...
+deref_array(5, p0, q0);
+    \end{lstlisting}
+\end{small}
+    \caption{Compilation Example for Dependent Functions}
+\label{fig:compilationexample1}
+\end{figure}
+
+% \review{Fig 9: if this is actual C code, then your null-check at line 6 will be
+%   eliminated by the compiler. At line 3, you performed a pointer addition, which
+%   is only defined when `p` is non-null. So, either `p` is non-null, and the
+%   NULL-check can be eliminated; or, `p` is NULL, but line 3 was undefined
+%   behavior, meaning the compiler is allowed to do anything, notably eliminate
+%   the NULL-check. This is where I am super confused, and either:
+%   - CoreC is not really the C language, and has different semantics...? but is
+%     this well-defined in the context of LLVM?
+%   - there is a problem that was not caught by the PLT-Redex-based testing.
+% \yiyun{We have clarified at the start of IV that CoreC is an untyped
+%   variant of CoreChkC, and does not aim to represent C per se, or LLVM
+%   IR. We aimed to avoid confusion by rewriting the examples in a way that is more closely
+%     related to the syntax presented in Fig 3. Pointer arithmetic between 0 and a non-zero
+%     index is always valid because CoreC there is technically only
+%     integer arithmetic.}}
+
+\myparagraph{Pointer Accesses and Modes}
+%
+In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
+if the poniter is an (NT-)array,
+we first allocate two \emph{shadow variables}
+to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
+Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
+Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
+In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
+there is an inserted verification step similar to \Cref{fig:const-type},
+which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
+Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
+The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
+ and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
+are computed based on the meet 
+operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).
+
+\myparagraph{Checked and Unchecked Blocks}
+%
+In the \systemname implementation,
+$\euncheckedtext$ and $\echeckedtext$ blocks 
+are compiled as context switching functions provided by the sandbox mechanism.
+$\eunchecked{\overline{x}}{e}$ is compiled to 
+$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
+to execute expression $e$ with the arguments $\overline{x}$.
+$\echecked{\overline{x}}{e}$ is compiled to 
+$\texttt{callback}(\overline{x},e)$, where we perform 
+a \texttt{callback} to a checked block code $e$ inside a sandbox.
+In \systemname, we adopt an aggressive execution scheme that
+directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
+In the formalism, we rely on the type system to 
+guarantee the context switching without creating the extra function calls for simplicity.
+
+%Fig.~\ref{fig:compilationexample} shows how an invocation of
+%\code{strlen} on a null-terminated string is compiled into C
+%code. Each dereference of a checked pointer requires a null check
+%(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
+%compiler makes explicit: Line~$3$ of the generated code has the null
+%check on pointer \code{p} due to the \code{strlen},
+%  and a similar check happens
+%  at line~$8$ due to the pointer arithmetic on \code{p}.
+%Dereferences also require bounds checks: line~$2$ checks \code{p} is
+%in bounds before computing \code{strlen(p)}, while line~$10$ does
+%likewise before computing \code{*(p+1)}.
+
+\myparagraph{Function Pointers and Calls}
+%
+Function pointers are managed similarly to normal pointers,
+but we insert checks to check if the pointer address is not null in 
+the function store instead of heap, and whether or not the type is correctly represented, 
+for both $\cmode$ and $\tmode$ mode pointers 
+\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
+For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
+we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
+and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
+The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
+is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
+
+For compiling dependent function calls,
+\Cref{fig:compilationexample1} provides a hint.
+Notice that the bounds for the array pointer \code{p} are not passed as
+arguments. Instead, they are initialized according to \code{p}'s
+type---see line~4 of the original \lang program at the top of the figure.
+Line~$3$ of the generated code
+sets the lower bound  to \code{0} and the
+upper bound to \code{n}.
+
+\subsection{Metatheory}
+\label{sec:meta}
+
+% \review{ 
+% My interpretation of section IV.C is that the authors have a pen-and-paper
+%   proof of Theorem 4, and that random testing using the Redex models has been
+%   used to gain confidence in such a proof. Is this correct?}
+% \liyi{Yes, this is correct.}
+% \mwh{No, not correct: We have no pen-and-paper proof, just the
+%   model. Clarify here}
+% showing that
+% extensive random testing fails to falsify the bisimilarity
+% property. \mwh{Better way to state the previous?}
+
+We formalize both the compilation procedure and the simulation
+theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
+and then attempt to falsify it via Redex's support for random
+testing. Redex allows us
+  to specify compilation as logical rules (an extension
+  of typing), but then execute it algorithmically to
+  automatically test whether simulation holds. This process revealed
+  several bugs in compilation and the theorem statement.
+%
+  % us gain confidence that our original pen and paper proof of
+  % simulation remains true with the addition of variable bounds. }
+We ultimately plan to prove simulation in the Coq model.
+
+%Turning to the simulation theorem: We first introduce notation
+%used to specify the theorem.
+We use the notation $\gg$ to
+indicate the \emph{erasure} of stack and heap---the rhs is the same as
+the lhs but with type annotations removed:
+\begin{equation*}
+  \begin{split}
+    \heap  \gg & \dot \heap \\
+    \varphi \gg & \dot \varphi
+  \end{split}
+\end{equation*}
+In addition, when $\Gamma;\emptyset\vdash
+\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
+\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
+and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively.
+$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
+Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
+reduction relation of $\elang$. Unlike the $\lang$, the semantics of
+$\elang$ does not distinguish checked and unchecked regions.
+
+Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
+the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
+that is similar to the one by~\citet{merigoux2021catala}.
+
+An ordinary simulation property would
+replace the middle and bottom parts of the figure with the
+following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
+  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
+Instead, we relate two erased configurations using the relation $\sim$,
+which only requires that the two configurations will eventually reduce
+to the same state.
+
+\ignore{
+ We formulate our simulation theorem differently
+because the standard simulation theorem imposes a very strong
+syntactic restriction to the compilation strategy. Very often, $(\dot
+\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
+semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
+but we are unable to syntactically equate the two configurations due
+to the extra binders generated for dynamic checks and ANF
+transformation. In earlier versions of the Redex model, we attempted
+to change the compilation rules so the configurations could match
+syntactically. However, the approach scaled poorly as we added
+additional rules. 
+This slight relaxation on the equivalence relation
+between target configurations allows us to specify compilation more
+naturally without having to worry about syntactic constraints.
+}
+
+% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
+\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
+and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
+
+\begin{itemize}
+
+\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\varphi,\dot \heap,\dot e)$ and $(\dot
+\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+\dot \heap,\dot e)$.
+
+\item if $r_1 = \ebounds$ or $\enull$, then we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+\varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+\end{itemize}
+\end{thm}
+
+% when $r_1 = e_1$ for
+% some $e_1$ and
+% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+
+% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
+% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
+% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
+% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
+% some $e_1$ and
+% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
+% , then
+% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
+% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \varphi,\dot \heap,\dot e)$ and $(\dot
+% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
+% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
+% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
+% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
+% \end{thm}
+
+As our random generator never generates
+$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
+as it relates to checked code. This limitation makes it
+unnecessary to state the other direction of the simulation theorem
+where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
+that $e_0$ will never enter a stuck state if it is well-typed in
+checked mode.
+
+The current version of the Redex model has been tested against $23000$
+expressions with depth less than $11$. Each expression can
+reduce multiple steps, and we test simulation between every two
+adjacent steps to cover a wider range of programs, particularly the
+ones that have a non-empty heap.
+
+\begin{figure}[t]
+{\small
+\[
+\begin{array}{c}
+\begin{tikzpicture}[
+            > = stealth, % arrow head style
+            shorten > = 1pt, % don't touch arrow head to node
+            auto,
+            node distance = 3cm
+        ]
+
+\begin{scope}[every node/.style={draw}]
+    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
+    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
+    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
+    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
+    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
+\end{scope}
+\begin{scope}[every edge/.style={draw=black}]
+
+    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
+    \path [<->] (A) edge node {$\gg$} (C);
+    \path [<->] (B) edge node {$\gg$} (D);
+    \path [dashed,<->] (C) edge node {$\sim$} (D);
+    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
+    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
+\end{scope}
+
+\end{tikzpicture}
+\end{array}
+\]
+}
+\caption{Simulation between \lang and \elang }
+\label{fig:checkedc-simulation-ref}
+\end{figure}
+
+% \ignore
+% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
+% % The diagram also indicates the simulation relation ($\sim$) among our formalization of the \checkedc Type System ($\vdash$), Semantics ($\longrightarrow$), Compilation ($\gg$), as well as the compiled C semantics ($\xrightarrow{c}$).
+% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
+% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist.
+The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.
+
+%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C.
+The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
+% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).
+
+% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
+% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
+% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation.
+We generate tens of thousands cases and cover all possible corner cases of the compilation.
+% }
+% \ignore{
+% \begin{figure*}[t]
+%   \begin{subfigure}[b]{1\textwidth}
+%     \begin{lstlisting}
+% int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
+%   return *p;
+% }
+
+% ...
+% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
+% deref_array(5, p0);
+%     \end{lstlisting}
+%     \label{fig:chkcexamplederef}
+%     \caption{\texttt{deref\_array} in \checkedc}
+%   \end{subfigure}
+
+%   \begin{subfigure}[b]{1\textwidth}
+% \begin{lstlisting}
+% int deref_array(size_t n, int *p) {
+%   /* statically determine the definitions p_lo and p_hi */
+%   int *p_lo = p;
+%   int *p_hi = p + n;
+%   /* runtime checks */
+%   assert(p_lo <= p && p <= p_hi);
+%   assert(p != NULL);
+%   /* widening */
+%   int p_derefed = *p;
+%   if (p_derefed != '\0') {
+%     if (p_hi == p) {
+%       ++p_hi;
+%     }
+%   }
+%   return p_derefed;
+% }
+% ...
+% /* int *p0 */
+% /* rho(p0) = p_lo, p_hi */
+% deref_array(5, p0);
+% \end{lstlisting}
+% \caption{\texttt{deref\_array} in C}
+%     \label{fig:cexamplederef}
+%   \end{subfigure}
+
+%   \begin{subfigure}[b]{\textwidth}
+%     \begin{lstlisting}
+% /* nt_array_ptr<int> p : bounds(e0, e1) */
+% /* rho(p) = p_lo, p_hi */
+% size_t x = strlen(p);
+% if(x >= 1) {
+%   *(p+1);
+% }
+%     \end{lstlisting}
+%     \caption{\texttt{strlen} in \checkedc}
+%     \label{fig:chkcexamplestrlen}
+%   \end{subfigure}
+
+%   \begin{subfigure}[b]{\textwidth}
+%     \begin{lstlisting}
+% /* nt_array_ptr<int> p : bounds(e0, e1) */
+% /* rho(p) = p_lo, p_hi */
+% /* runtime checks omitted */
+% ...
+% size_t x = strlen(p);
+% int *p_hi_new = p + x;
+% p_hi = max(p_hi, p_hi_new);
+% if (x >= 1) {
+%   /* null check for pointer arithmetic */
+%   assert(p != NULL);
+%   int *p_1 = p + 1;
+%   /* null check for dereference */
+%   assert(p_1 != NULL);
+%   /* bounds check for dereference */
+%   /* note how we are using p_hi rather than p + x */
+%   assert(p_lo <= p_1 && p_1 <= p_hi);
+%   ...
+% }
+%     \end{lstlisting}
+%     \caption{\texttt{strlen} in C}
+%     \label{fig:cexamplestrlen}
+%   \end{subfigure}
+% \caption{Compilation of null-terminated string dereference}
+% \label{fig:compilationexample}
+% \end{figure*}
+% }
+
+% \ignore{
+% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
+% dereference of null-terminated strings by inserting explicit checks and
+% bounds widening code. For readability, we present the example in real
+% \checkedc and C syntax. At
+% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
+% bounds are defined in terms of the arguments \code{p} and \code{n}
+% according to the bounds annotations obtained during typechecking. This
+% avoids the need to pass in bounds as extra arguments, thus maintaining
+% compatibility with C code. When we call \code{deref_array} on
+% \code{p0}, a string with size 5 (excluding the null character), there
+% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
+% the stack. It is possible that \code{p_hi} has been widened
+% before we reach line 21. The programmer will have to perform a
+% \code{dynamic_bounds_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
+% attempts to widen the bound when the dereferenced result is not
+% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
+% \code{deref_array}, in contrast to the T-IfNT rule, which only
+% remembers the widened bounds within the scope of the then branch of
+% the if statement.
+
+% \todo[inline]{Widening can happen at every
+%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
+% shows how an invocation of \code{strlen} on null-termianted strings
+% is compiled into C code. We perform the same runtime checks that
+% happen during dereferencing. The widening code at line
+% 8 updates \code{p}'s upper bound only if the result of
+% \code{strlen} is larger than the value of the upper bound stored on
+% the stack. This is another scenario where the runtime can be more precise
+% than the statically determined bounds information.
+
+% }
 
@@
 
-\subsection{Struct Pointers}\label{appx:struct}
-
-\checkedc has \kw{struct} types and \kw{struct} pointers. Fig.~\ref{fig:checkc-struct} contains the syntax of \kw{struct} types as well as new subtyping relations built on the \kw{struct} values.
-For a \kw{struct} typed value, \checkedc has a special operation for it, which is $\eamper{e}{f}$.
-This operation indexes the $f$-th position \kw{struct}~$T$ item, if the expression $e$ is evaluated to a \kw{struct} pointer $\tptr{\tstruct{T}}{m}$.
-Rule \textsc{T-Struct} in Fig.~\ref{fig:checkc-struct} describes its typing behavior.
-Rules \textsc{S-StructChecked} and \textsc{S-StructUnChecked} describe the semantic behaviors of $\eamper{e}{f}$ on a given \kw{struct} \code{checked}/\code{unchecked} pointers, while rule \textsc{S-StructNull} describes a \code{checked} \kw{struct} null-pointer case.
-In our Coq/Redex formalization, we include the \kw{struct} values and the operation $\eamper{e}{f}$. We omit it in the main text due to the paper length limitation.
-
-\begin{figure}
-{\small
-$\begin{array}{l}
-\text{  Struct Syntax: }\\[0.5em]
-  \begin{array}{ll}
- \mathtt{Type} & \tstruct{T}
-\\[0.2em]
-     \text{Structdefs} & D \; \in \; T \rightharpoonup fs \\[0.2em]
-      \text{Fields} & fs \; ::= \; \tau~\mathtt{f} \mid \tau~\mathtt{f}; fs 
-    \end{array}\\[2em]
-\text{  Struct Subtype: }\\[0.5em]
-\begin{array}{l}
-    D(T) = fs \wedge fs(0) = \tnat \Rightarrow  \tptr{\tstruct{T}}{m} \sqsubseteq \tptr{\tnat}{m}\\[0.5em]
-    D(T) = fs \wedge fs(0) = \tnat \wedge 0 \le b_l \wedge b_h \le 1 \\[0.2em]
- \qquad\qquad \Rightarrow 
-       \tptr{\tstruct{T}}{m} \sqsubseteq \tarrayptr{b_l}{b_h}{\tnat}{m}
-    \end{array}
-\\[3em]
-\text{  Struct Type Rule: }\\[0.5em]
-    \end{array}
-$
-
-{
-\begin{mathpar}
-  \inferrule [T-Struct]
-  {\Gamma; \Theta \vdash_m e :  {\tptr{\tstruct{T}}{m}} \\
-    D(T) = fs \\ fs(f)=\tau_f}
-  {\Gamma; \Theta \vdash_m \eamper{e}{f} : \tptr{\tau_f}{m}}
-
-\end{mathpar}
-}
-{
-$\begin{array}{l}
-\text{Struct Semantics: }
-\end{array}
-$
-}
-{
-\begin{mathpar}
-  \inferrule [S-StructChecked]
-  {n > 0 \\ D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
-  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\cmode}})}
-
-  \inferrule [S-StructNull]
-  {n = 0}
-  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\cmode}}}{f}) \longrightarrow (\varphi,\heap,\enull)}
-
-  \inferrule [S-StructUnChecked]
-  {D(T) = fs \\ fs(f)=\tau_a \\ n_a=\mathtt{index}(fs,f)}
-  {(\varphi,\heap,\eamper{\evalue{n}{\tptr{\tstruct{T}}{\umode}}}{f}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tptr{\tau_a}{\umode}})}
-
-\end{mathpar}
-}
-}
-  \caption{\lang Struct Definitions}
-  \label{fig:checkc-struct}
-\end{figure}
-
-\begin{figure}[t]
-{\small
-  \begin{mathpar}
-    \inferrule[]
-    {\Gamma \vdash \tau}
-    {\Gamma \vdash \tau~\mathtt{f} }
-
-    \inferrule[]
-    {\Gamma \vdash \tau\\
-     \Gamma \vdash fs}
-    {\Gamma \vdash \tau~\mathtt{f}; fs }
-
-  \end{mathpar}
-}
- \caption{Well-formedness for structs}
-\label{fig:wfstructs}
-
-{\small
-  \begin{mathpar}
-
-    \inferrule[]
-    {\Gamma[\overline{x} \mapsto \overline{\tau}]; \emptyset  \vdash
-      e \gg \dot e : \tau}
-    {\Gamma \vdash \tau\;(\overline{x}:\overline{\tau})\;e \gg
-      (\overline{x})\;\dot e}
-  \end{mathpar}
-}
- \caption{Compilation Rules for Functions}
-\label{fig:compilefunctions}
-\end{figure}
-
 \subsection{The Compilation Rules}\label{appx:comp1}
@@
 
-% \review{Fig 9: if this is actual C code, then your null-check at line 6 will be
-%   eliminated by the compiler. At line 3, you performed a pointer addition, which
-%   is only defined when `p` is non-null. So, either `p` is non-null, and the
-%   NULL-check can be eliminated; or, `p` is NULL, but line 3 was undefined
-%   behavior, meaning the compiler is allowed to do anything, notably eliminate
-%   the NULL-check. This is where I am super confused, and either:
-%   - CoreC is not really the C language, and has different semantics...? but is
-%     this well-defined in the context of LLVM?
-%   - there is a problem that was not caught by the PLT-Redex-based testing.
-% \yiyun{We have clarified at the start of IV that CoreC is an untyped
-%   variant of CoreChkC, and does not aim to represent C per se, or LLVM
-%   IR. We aimed to avoid confusion by rewriting the examples in a way that is more closely
-%     related to the syntax presented in Fig 3. Pointer arithmetic between 0 and a non-zero
-%     index is always valid because CoreC there is technically only
-%     integer arithmetic.}}
-
-\myparagraph{Pointer Accesses and Modes}
-%
-In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
-if the poniter is an (NT-)array,
-we first allocate two \emph{shadow variables}
-to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
-Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
-Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
-In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
-there is an inserted verification step similar to \Cref{fig:const-type},
-which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
-Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
-The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
- and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
-are computed based on the meet 
-operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).
-
-\myparagraph{Checked and Unchecked Blocks}
-%
-In the \systemname implementation,
-$\euncheckedtext$ and $\echeckedtext$ blocks 
-are compiled as context switching functions provided by the sandbox mechanism.
-$\eunchecked{\overline{x}}{e}$ is compiled to 
-$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
-to execute expression $e$ with the arguments $\overline{x}$.
-$\echecked{\overline{x}}{e}$ is compiled to 
-$\texttt{callback}(\overline{x},e)$, where we perform 
-a \texttt{callback} to a checked block code $e$ inside a sandbox.
-In \systemname, we adopt an aggressive execution scheme that
-directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
-In the formalism, we rely on the type system to 
-guarantee the context switching without creating the extra function calls for simplicity.
-
-%Fig.~\ref{fig:compilationexample} shows how an invocation of
-%\code{strlen} on a null-terminated string is compiled into C
-%code. Each dereference of a checked pointer requires a null check
-%(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
-%compiler makes explicit: Line~$3$ of the generated code has the null
-%check on pointer \code{p} due to the \code{strlen},
-%  and a similar check happens
-%  at line~$8$ due to the pointer arithmetic on \code{p}.
-%Dereferences also require bounds checks: line~$2$ checks \code{p} is
-%in bounds before computing \code{strlen(p)}, while line~$10$ does
-%likewise before computing \code{*(p+1)}.
-
-\myparagraph{Function Pointers and Calls}
-%
-Function pointers are managed similarly to normal pointers,
-but we insert checks to check if the pointer address is not null in 
-the function store instead of heap, and whether or not the type is correctly represented, 
-for both $\cmode$ and $\tmode$ mode pointers 
-\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
-For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
-we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
-and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
-The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
-is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
-
-For compiling dependent function calls,
-\Cref{fig:compilationexample1} provides a hint.
-Notice that the bounds for the array pointer \code{p} are not passed as
-arguments. Instead, they are initialized according to \code{p}'s
-type---see line~4 of the original \lang program at the top of the figure.
-Line~$3$ of the generated code
-sets the lower bound  to \code{0} and the
-upper bound to \code{n}.
-
 \subsection{Metatheory}

### 2023-01-15 sandp23/appendix.tex
finished appx.

@@
 
-\subsection{Compilation Formalism}\label{sec:compilation}
+\subsection{Compilation Formalism}\label{appx:comp1}
 
@@
 that track array bound information, which is used in premises
-of rules like \textsc{S-DefArray} and
-  \textsc{S-AssignArr} to prevent spatial safety violations.
+of rules like \textsc{S-DefArrayC} and
+  \textsc{S-AssignArrC} to prevent spatial safety violations.
 The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
-for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
+for verifying pointers are not null and the bounds are within their limits.
 The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.
@@
 i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions. 
-\lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
 Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
 In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different.
-\lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
 \Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
@@
 with no loss of expressiveness
-as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
+as the \checkedc contains the erase of annotations in \cite{li22checkedc}.
 For the compiler formalism, 
@@
 In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
-$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
-The algorithm sheds
-  light on how compilation can be implemented in the real Checked C
-  compiler, while eschewing many vital details (\elang has many 
-  differences with LLVM IR).
-
-%This section shows how \systemname deals with 
-%annotations can be safely erased: using static information a compiler
-%can insert code to manage and check bounds metadata, with no loss of
-%expressiveness. We present a compilation algorithm that converts from
-%\lang to \elang, an untyped language without metadata
-%annotations. The syntax and semantics \elang
-  %closely mirrors that of \lang; it differs only in that literals lack
-  %type annotations and its operational rules perform no
-  %bounds and null checks, which are instead inserted during
-  %compilation. Our compilation algorithm is evidence that \lang's
-  %semantics, despite its apparent use of fat pointers, faithfully
-  %represents Checked C's intended behavior. The algorithm also sheds
-  %light on how compilation can be implemented in the real Checked C
-  %compiler, while eschewing many important details (\elang has many 
-  %differences with LLVM IR).
+$\emalloc{m}{\omega}$, and $\ecall{e}{\overline{e}}$.
+The algorithm sheds light on how compilation can be implemented in the real \systemname
+  compiler, while eschewing many vital details (\elang has many differences with LLVM IR).
 
@@
   lower-bound shadow variable is unnecessary; we include it for uniformity.}
-\lc{These two sentences are confusing. }
 
@@
 
-\subsection{Approach}
-
-Due to space constraints, we explain the rules for compilation by
+\myparagraph{Approach}
+%
+Here, we explain the rules for compilation by
 examples, using a C-like syntax; the complete rules are given in
-\iftr
-Appendix~\ref{appx:comp1}.
-\else
-the supplemental report~\cite{checkedc-tech-report}.
-\fi
+\cite{checkedc-tech-report}.
 Each rule performs up to three tasks: (a) conversion of $e$ to
@@
 %
-In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
+In every declaration of a pointer,
 if the poniter is an (NT-)array,
@@
 $\euncheckedtext$ and $\echeckedtext$ blocks 
-are compiled as context switching functions provided by the sandbox mechanism.
+are compiled as context switching functions provided by sandbox.
 $\eunchecked{\overline{x}}{e}$ is compiled to 
@@
 In \systemname, we adopt an aggressive execution scheme that
-directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
+directly learns pointer addresses from compiled assembly to make the $\code{_Callback}$ happen.
 In the formalism, we rely on the type system to 
@@
 \footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
-For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
-we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
-and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
 The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
 is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
-
 For compiling dependent function calls,
@@
 
-\subsection{Metatheory}
+\subsection{Constraints and Metatheory}
 \label{sec:meta}
 
-% \review{ 
-% My interpretation of section IV.C is that the authors have a pen-and-paper
-%   proof of Theorem 4, and that random testing using the Redex models has been
-%   used to gain confidence in such a proof. Is this correct?}
-% \liyi{Yes, this is correct.}
-% \mwh{No, not correct: We have no pen-and-paper proof, just the
-%   model. Clarify here}
-% showing that
-% extensive random testing fails to falsify the bisimilarity
-% property. \mwh{Better way to state the previous?}
+Here, we first show some Well-formedness and consistency definitions that are required in \Cref{sec:theorem}, and then show the simulation theorem for the \lang compiler.
+Type soundness relies on several \emph{well-formedness}:
+
+\begin{definition}[Type Environment Well-formedness]\label{type-wellformed}
+A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
+\end{definition}
+
+\begin{definition}[Heap Well-formedness]
+For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
+(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
+contains no free variables. 
+\end{definition}
+
+\begin{definition}[Stack Well-formedness]
+A stack snapshot $\varphi$ is well-formed if
+for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
+contains no free variables. 
+\end{definition}
+
+We also need to introduce a notion of
+\emph{consistency}, relating heap environments before and after a
+reduction step, and type environments, predicate sets, and stack
+snapshots together.
+
+\begin{definition}[Stack Consistency]
+A type environment $\Gamma$, variable predicate set $\Theta$, and
+stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
+\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
+$\Gamma(x) = \tau$ for some $\tau$ and 
+$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
+\end{definition}
+
+\begin{definition}[Checked Stack-Heap Consistency]
+A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
+for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
+\end{definition}
+
+\begin{definition}[Checked Heap-Heap Consistency]
+A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
+for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
+\end{definition}
 
@@
 
+\begin{figure}[t]
+{\small
+\[
+\begin{array}{c}
+\begin{tikzpicture}[
+            > = stealth, % arrow head style
+            shorten > = 1pt, % don't touch arrow head to node
+            auto,
+            node distance = 3cm
+        ]
+
+\begin{scope}[every node/.style={draw}]
+    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
+    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
+    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
+    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
+    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
+\end{scope}
+\begin{scope}[every edge/.style={draw=black}]
+
+    \path [->] (A) edge node {$\longrightarrow_{\cmode}$} (B);
+    \path [<->] (A) edge node {$\gg$} (C);
+    \path [<->] (B) edge node {$\gg$} (D);
+    \path [dashed,<->] (C) edge node {$\sim$} (D);
+    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
+    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
+\end{scope}
+
+\end{tikzpicture}
+\end{array}
+\]
+}
+\caption{Simulation between \lang and \elang }
+\label{fig:checkedc-simulation-ref}
+\end{figure}
+
 Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
@@
 to the same state.
-
-\ignore{
- We formulate our simulation theorem differently
-because the standard simulation theorem imposes a very strong
-syntactic restriction to the compilation strategy. Very often, $(\dot
-\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
-semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
-but we are unable to syntactically equate the two configurations due
-to the extra binders generated for dynamic checks and ANF
-transformation. In earlier versions of the Redex model, we attempted
-to change the compilation rules so the configurations could match
-syntactically. However, the approach scaled poorly as we added
-additional rules. 
-This slight relaxation on the equivalence relation
-between target configurations allows us to specify compilation more
-naturally without having to worry about syntactic constraints.
-}
 
@@
 
-The current version of the Redex model has been tested against $23000$
-expressions with depth less than $11$. Each expression can
+The current version of the Redex model has been tested against $21500$
+expressions with depth less than $12$. Each expression can
 reduce multiple steps, and we test simulation between every two
@@
 ones that have a non-empty heap.
-
-\begin{figure}[t]
-{\small
-\[
-\begin{array}{c}
-\begin{tikzpicture}[
-            > = stealth, % arrow head style
-            shorten > = 1pt, % don't touch arrow head to node
-            auto,
-            node distance = 3cm
-        ]
-
-\begin{scope}[every node/.style={draw}]
-    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
-    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
-    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
-    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
-    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
-\end{scope}
-\begin{scope}[every edge/.style={draw=black}]
-
-    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
-    \path [<->] (A) edge node {$\gg$} (C);
-    \path [<->] (B) edge node {$\gg$} (D);
-    \path [dashed,<->] (C) edge node {$\sim$} (D);
-    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
-    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
-\end{scope}
-
-\end{tikzpicture}
-\end{array}
-\]
-}
-\caption{Simulation between \lang and \elang }
-\label{fig:checkedc-simulation-ref}
-\end{figure}
-
-% \ignore
-% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
-% % The diagram also indicates the simulation relation ($\sim$) among our formalization of the \checkedc Type System ($\vdash$), Semantics ($\longrightarrow$), Compilation ($\gg$), as well as the compiled C semantics ($\xrightarrow{c}$).
-% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
-% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist.
-The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.
-
-%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C.
-The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
-% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).
-
-% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
-% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
-% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation.
-We generate tens of thousands cases and cover all possible corner cases of the compilation.
-% }
-% \ignore{
-% \begin{figure*}[t]
-%   \begin{subfigure}[b]{1\textwidth}
-%     \begin{lstlisting}
-% int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
-%   return *p;
-% }
-
-% ...
-% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
-% deref_array(5, p0);
-%     \end{lstlisting}
-%     \label{fig:chkcexamplederef}
-%     \caption{\texttt{deref\_array} in \checkedc}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{1\textwidth}
-% \begin{lstlisting}
-% int deref_array(size_t n, int *p) {
-%   /* statically determine the definitions p_lo and p_hi */
-%   int *p_lo = p;
-%   int *p_hi = p + n;
-%   /* runtime checks */
-%   assert(p_lo <= p && p <= p_hi);
-%   assert(p != NULL);
-%   /* widening */
-%   int p_derefed = *p;
-%   if (p_derefed != '\0') {
-%     if (p_hi == p) {
-%       ++p_hi;
-%     }
-%   }
-%   return p_derefed;
-% }
-% ...
-% /* int *p0 */
-% /* rho(p0) = p_lo, p_hi */
-% deref_array(5, p0);
-% \end{lstlisting}
-% \caption{\texttt{deref\_array} in C}
-%     \label{fig:cexamplederef}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{\textwidth}
-%     \begin{lstlisting}
-% /* nt_array_ptr<int> p : bounds(e0, e1) */
-% /* rho(p) = p_lo, p_hi */
-% size_t x = strlen(p);
-% if(x >= 1) {
-%   *(p+1);
-% }
-%     \end{lstlisting}
-%     \caption{\texttt{strlen} in \checkedc}
-%     \label{fig:chkcexamplestrlen}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{\textwidth}
-%     \begin{lstlisting}
-% /* nt_array_ptr<int> p : bounds(e0, e1) */
-% /* rho(p) = p_lo, p_hi */
-% /* runtime checks omitted */
-% ...
-% size_t x = strlen(p);
-% int *p_hi_new = p + x;
-% p_hi = max(p_hi, p_hi_new);
-% if (x >= 1) {
-%   /* null check for pointer arithmetic */
-%   assert(p != NULL);
-%   int *p_1 = p + 1;
-%   /* null check for dereference */
-%   assert(p_1 != NULL);
-%   /* bounds check for dereference */
-%   /* note how we are using p_hi rather than p + x */
-%   assert(p_lo <= p_1 && p_1 <= p_hi);
-%   ...
-% }
-%     \end{lstlisting}
-%     \caption{\texttt{strlen} in C}
-%     \label{fig:cexamplestrlen}
-%   \end{subfigure}
-% \caption{Compilation of null-terminated string dereference}
-% \label{fig:compilationexample}
-% \end{figure*}
-% }
-
-% \ignore{
-% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
-% dereference of null-terminated strings by inserting explicit checks and
-% bounds widening code. For readability, we present the example in real
-% \checkedc and C syntax. At
-% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
-% bounds are defined in terms of the arguments \code{p} and \code{n}
-% according to the bounds annotations obtained during typechecking. This
-% avoids the need to pass in bounds as extra arguments, thus maintaining
-% compatibility with C code. When we call \code{deref_array} on
-% \code{p0}, a string with size 5 (excluding the null character), there
-% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
-% the stack. It is possible that \code{p_hi} has been widened
-% before we reach line 21. The programmer will have to perform a
-% \code{dynamic_bounds_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
-% attempts to widen the bound when the dereferenced result is not
-% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
-% \code{deref_array}, in contrast to the T-IfNT rule, which only
-% remembers the widened bounds within the scope of the then branch of
-% the if statement.
-
-% \todo[inline]{Widening can happen at every
-%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
-% shows how an invocation of \code{strlen} on null-termianted strings
-% is compiled into C code. We perform the same runtime checks that
-% happen during dereferencing. The widening code at line
-% 8 updates \code{p}'s upper bound only if the result of
-% \code{strlen} is larger than the value of the upper bound stored on
-% the stack. This is another scenario where the runtime can be more precise
-% than the statically determined bounds information.
-
-% }
-
-\subsection{Additional Type and Semantic Rules}\label{appx:add-type-sem}
-
-Since bounds expressions may
-contain variables, determining assumptions like $b_l \leq b_l'$
-requires reasoning about the probable values of these variables'. The type
-system uses $\Theta$ to make such reasoning more precise.
-$\Theta$ is a map from variables $x$ to
-equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
-It maps variables to equations that are recorded along the type checking procedure.
-If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
-$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
-such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
-\iftr
-  Appendix~\ref{app:le}.
-\else
-  the supplemental report~\cite{checkedc-tech-report}.
-\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
-
-\begin{figure}[t]
-{\small
-{\captionsetup[lstlisting]{margin = 8 mm}
-  \begin{lstlisting}[xleftmargin=8 mm]
-nt_array_ptr<char> safe_strcat
-   (nt_array_ptr<char> dst : count(n),
-    nt_array_ptr<char> src : count(0), int n) {
-  int x = strlen(dst);
-  int y = strlen(src);
-  nt_array_ptr<char> c : count(n) =
-    dyn_bounds_cast
-           <nt_array_ptr<char>>(dst,count(n));
-    // sets c == dst with bound n (not x)
-  if (x+y < n) {
-    for (int i = 0; i < y; ++i)
-      *(c+x+i) = *(src+i);
-    *(c+x+y) = '\0';
-    return dst;
-  }
-  return null;
-}
-  \end{lstlisting}
-}
-}
-\caption{Implementation of safe \code{strcat}}
-\label{fig:strcat-ex}
-\end{figure}
-
-Recall that array
-bounds in types may refer to in-scope variables; e.g., parameter
-\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
-2-3 in \Cref{fig:strcat-ex}. 
-All dependent bound variables appearing in function headers 
-are bounded by a variable list in the function pointer type, which is checked in the type rule.
-Semantically,
-the call is expanded into a \texttt{let} which binds
-parameter variables $\overline{x}$ to the actual arguments
-$\overline{n}$, but annotated with the parameter types
-$\overline{\tau}$ (this will be safe for type-correct programs). 
-The function body $e$ is wrapped in a static cast
-$(\tau[\overline{n} / \overline{x}])$ which is the function's return
-type but with any integer parameter variables $\overline{x}$ appearing in that
-type, as type bound variables,
-substituted with the call's actual arguments $\overline{n}$.
-To see why this is needed, suppose that \code{safe_strcat} in
-Fig.~\ref{fig:strcat-ex} is defined to return a
-\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
-perform a \code{safe_strcat} function call as
-\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
-function returns a value with type \code{nt_array_ptr<int>:count(10)}
-because we substitute bound variable \code{n} in the 
-defined return type with \code{10} from the function call's
-argument list.
-
-\begin{figure}[t!]
-  \begin{small}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-int deref_array(n : int,
-     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
-     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
-  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
-  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
-    * p;
-    * q = 1;
-}
-...
-/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
-/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
-deref_array(5, p0, q0);
-    \end{lstlisting}
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-deref_array(int n, int* p, int * q) {
-  //m is the current context mode
-  let p_lo = 0; let p_hi = n; 
-  let q_lo = 0; let q_hi = n; 
-  /* runtime checks */
-  assert(p_lo <= 0 && 0 <= p_hi);
-  assert(p != 0);
-  *(mode(p) $\wedge$ m,p);
-  verify(q, not_null(m, q_lo, q_hi) 
-             && q_lo <= 0 && 0 <= q_hi);
-  *(mode(q) $\wedge$ m,q)=1;
-}
-...
-deref_array(5, p0, q0);
-    \end{lstlisting}
-\end{small}
-    \caption{Compilation Example for Dependent Functions}
-\label{fig:compilationexample1}
-\end{figure}
-
-For compiling dependent function calls,
-\Cref{fig:compilationexample1} provides a hint.
-Notice that the bounds for the array pointer \code{p} are not passed as
-arguments. Instead, they are initialized according to \code{p}'s
-type---see line~4 of the original \lang program at the top of the figure.
-Line~$3$ of the generated code
-sets the lower bound  to \code{0} and the
-upper bound to \code{n}.
-
-\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
-A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
-\end{defi}
-
-\begin{defi}[Heap Well-formedness]
-For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
-(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
-contains no free variables. 
-\end{defi}
-
-\begin{defi}[Stack Well-formedness]
-A stack snapshot $\varphi$ is well-formed if
-for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
-contains no free variables. 
-\end{defi}
-
-We also need to introduce a notion of
-\emph{consistency}, relating heap environments before and after a
-reduction step, and type environments, predicate sets, and stack
-snapshots together.
-
-\begin{defi}[Stack Consistency]
-A type environment $\Gamma$, variable predicate set $\Theta$, and
-stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
-\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
-$\Gamma(x) = \tau$ for some $\tau$ and 
-$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
-\end{defi}
-
-\begin{defi}[Checked Stack-Heap Consistency]
-A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
-for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
-\end{defi}
-
-\begin{defi}[Checked Heap-Heap Consistency]
-A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
-for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
-\end{defi}
-
-\myparagraph{Constant Validity}
-Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
-describe type assumptions for constants appearing in a program.
-$\neg \cmode(\tau)$ judges that a constant pointer 
-in an unchecked region cannot be of a checked type.
-The restriction ensures that programmers 
-cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
-In rule \textsc{T-ConstC}, we requires a static 
-verification procedure for validating a constant pointer in \Cref{fig:const-type}.
-
-The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
-validates the constant $\evalue{n}{\tau}$, 
-where $\heap(m)$ is the initial heap that the constant resides on and
-$\sigma$ is a set of constant assumed to be checked.
-A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
-A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
-and the address stores a function with the right type.
-The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
-where every element in the pointer range ($[0,\size(\omega))$) should be well
-typed.
-A checked pointer checks validity in type step as rule \textsc{T-ConstC},
-while a tainted/unchecked pointer does not check for such during the type checking.
-Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.
-
-\subsection{Meta Theories}\label{sec:theorem}
-
-% Before we present our main theorems, we need to first
-% discuss the meaning what a pointer being well-typed in a given heap
-% snapshot $\heap$ means, which is captured by rules in
-% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
-% simply checks if a given variable has the defined type in $\Gamma$;
-% the constant rule ($\textsc{T-Const}$) is slightly more involved.
-% First, it ensures that the type annotation $\tau$ does not contain any
-% free variables. More importantly, it ensures that the pointer points
-% to a location that makes sense in a given heap.
-%  
-%  
-%  The $\size$ function in Fig.~\ref{fig:const-type}
-% refers to the \code{sizeof} function in C computing the number of
-% bytes for a type.
-%  
-%  
-%  Second, we
-% require that any constant ($\evalue{n}{\tau}$) should make sense in
-% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
-% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
-% constant set containing the constants that have been verified by the
-% relation. For every constant $\evalue{n}{\tau}$, it is either an
-% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
-% zero-valued number ($n=0$), checked in $\sigma$
-% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
-% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
-% location $n+i$ in the range of the pointer (if $\omega$ is a word
-% type, range is $[0,1)$; if $\omega$ is an array type
-%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
-%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
-%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
-%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
-%       \vdash n_a : \tau_a$.
-%  
-%  
-% \begin{figure}[t]
-% {\small
-% \text{Type Rules for Constants and Variables:}
-% \begin{mathpar}
-%   \inferrule[T-Var]
-%       {x : \tau \in \Gamma}
-%       {\Gamma;\Theta \vdash_m x : \tau}
-%  
-%   \inferrule[T-Const]
-%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
-%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
-% \end{mathpar}
-%     
-% \text{Rules for Checking Constant Pointers In Heap:}
-% \begin{mathpar}
-%   \inferrule
-%       {}
-%       {\sigma \vdash n : \tint}
-%  
-%   \inferrule
-%       {}
-%       {\sigma \vdash n : \tptr{\omega}{\umode}}
-%  
-%   \inferrule
-%       {}
-%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
-%  
-%   \inferrule
-%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
-%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-%  
-%   \inferrule
-%       {\forall i \in [0,\size(\omega)) .
-%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
-%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-% \end{mathpar}
-% }
-% \caption{Type Rules for Checking Constants/Variables}
-% \label{fig:const-type}
-% \end{figure}
-
-% \review{
-%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
-%   something, I didn't see such a definition in the paper.}
-% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
-%   assumption of well typing; we have added more details about that.}
-
-In this subsection, we focus on our main meta-theoretic results about
-\lang: type soundness (progress and preservation),
-non-exposure, and non-crashing.
-These proofs have been conducted in our Coq model.
-Type soundness relies on several \emph{well-formedness} given in \ref{li22checkedc} and \Cref{appx:add-type-sem}.
-Progress states that a \lang program can always make a move:
-
-\begin{thm}[Progress]\label{thm:progress}
-
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
-one of the following holds:
-
-\begin{itemize}
-
-\item $e$ is a value ($\evalue{n}{\tau}$).
-
-\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.
-
-\end{itemize}
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-
-\noindent
-There are two forms of preservation regarding the checked and unchecked regions.
-Checked Preservation states that a reduction step preserves both the
-type and consistency of the program being reduced.
-Unchecked Preservation states that any evaluation happens at unchecked region does not affect the checked heap.
-
-\begin{thm}[Checked Preservation]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
-checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$ and $\tau'$ that are well formed, checked region consistent
-($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
-well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
-$\tau'\sqsubseteq_{\Theta} \tau$.
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-%\smallskip
-\begin{thm}[Unchecked Preservation]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
-\end{thm}
-
-Using the above theorem, we first show the non-exposure theorem,
-where code in unchecked region cannot observe a valid checked pointer address.
-
-\begin{thm}[Non-Exposure]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
-where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
-thus, it is not a checked pointer.
-\end{thm}
-
-We now state our main result, {\em non-crashing},
-which suggests that a well-typed program can never be \emph{stuck} (expression
-$e$ is a non-value that cannot take a step\footnote{Note that
-  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
-  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
-  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
-  safety.}).
-
-% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r.
-Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
-% \liyi{It is a typo.
-We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region.
-This is a bit due to the space limitation.
-The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement.
-}
-
-\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
-  program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are well-formed and consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
-if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
-\tau$ for some $\tau$) and there exists
-$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
-$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
-\end{thm}
-
-% \review{ 
-% My interpretation of section IV.C is that the authors have a pen-and-paper
-%   proof of Theorem 4, and that random testing using the Redex models has been
-%   used to gain confidence in such a proof. Is this correct?}
-% \liyi{Yes, this is correct.}
-% \mwh{No, not correct: We have no pen-and-paper proof, just the
-%   model. Clarify here}
-% showing that
-% extensive random testing fails to falsify the bisimilarity
-% property. \mwh{Better way to state the previous?}
-
-In the \systemname compiler, we also restore the compiler simulation theorem, such that executing a \systemname program is simulated by executing the compiled LLVM program.
-The details are given in the \checkedc formalism \cite{li22checkedc} and \Cref{appx:comp1}.
-
-We formalize both the compilation procedure and the simulation
-theorem in the semantic model we developed in Redex (same as our Coq model),
-and then attempt to falsify it via Redex's support for random
-testing. Redex allows us
-  to specify compilation as logical rules (an extension
-  of typing), but then execute it algorithmically to
-  automatically test whether simulation holds. This process revealed
-  several bugs in compilation and the theorem statement.
-%
-  % us gain confidence that our original pen and paper proof of
-  % simulation remains true with the addition of variable bounds. }
-We ultimately plan to prove simulation in the Coq model.
-
-%Turning to the simulation theorem: We first introduce notation
-%used to specify the theorem.
-We use the notation $\gg$ to
-indicate the \emph{erasure} of stack and heap---the rhs is the same as
-the lhs but with type annotations removed:
-\begin{equation*}
-  \begin{split}
-    \heap  \gg & \dot \heap \\
-    \varphi \gg & \dot \varphi
-  \end{split}
-\end{equation*}
-In addition, when $\Gamma;\emptyset\vdash
-\varphi$ and $\varphi$ is well-typed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
-\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
-and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively. 
-%$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
-%Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
-%reduction relation of $\elang$. Unlike the $\lang$, the semantics of
-%$\elang$ does not distinguish checked and unchecked regions.
-We show an overview of the simulation theorem in \Cref{fig:checkedc-simulation-ref}.
-The simulation theorem is specified in a way that is similar to the one by~\citet{merigoux2021catala}.
-
-Previous ordinary simulation properties would
-replace the middle and bottom parts of the figure with the
-following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
-  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
-Instead, we relate two erased configurations using the relation $\sim$,
-which only requires that the two configurations will eventually reduce
-to the same state.
-
-% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
-\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
-and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
-
-\begin{itemize}
-
-\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\varphi,\dot \heap,\dot e)$ and $(\dot
-\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-\dot \heap,\dot e)$.
-
-\item if $r_1 = \ebounds$ or $\enull$, then $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-\varphi_1$, $\heap_1 \gg \dot \heap_1$.
-
-\end{itemize}
-\end{thm}
-
-% when $r_1 = e_1$ for
-% some $e_1$ and
-% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
-% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \varphi,\dot \heap,\dot e)$ and $(\dot
-% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
-
-% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
-% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
-% some $e_1$ and
-% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
-% , then
-% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \varphi,\dot \heap,\dot e)$ and $(\dot
-% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
-% \end{thm}
-
-The current version of the Redex model has been tested against $27500$
-test cases with depth less than $15$. Each expression can
-reduce multiple steps, and we test simulation between every two
-adjacent steps to cover a wider range of programs, particularly the
-ones that have a non-empty heap.
-
-\begin{figure}[t]
-{\small
-\[
-\begin{array}{c}
-\begin{tikzpicture}[
-            > = stealth, % arrow head style
-            shorten > = 1pt, % don't touch arrow head to node
-            auto,
-            node distance = 3cm
-        ]
-
-\begin{scope}[every node/.style={draw}]
-    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
-    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
-    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
-    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
-    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
-\end{scope}
-\begin{scope}[every edge/.style={draw=black}]
-
-    \path [->] (A) edge node {$\longrightarrow_{\cmode}$} (B);
-    \path [<->] (A) edge node {$\gg$} (C);
-    \path [<->] (B) edge node {$\gg$} (D);
-    \path [dashed,<->] (C) edge node {$\sim$} (D);
-    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
-    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
-\end{scope}
-
-\end{tikzpicture}
-\end{array}
-\]
-}
-\caption{Simulation between \lang and \elang }
-\label{fig:checkedc-simulation-ref}
-\end{figure}
-
-%{\em Proof:} By induction on the number of steps of the \checkedc
-%evaluation ($\longrightarrow_m^*$), using progress and preservation to
-%maintain the invariance of the assumptions.
-
-% \review{
-%   add a paragraph that discusses what are the main changes from [21] in terms
-%   of the technical development (if there are any), e.g., are there any new
-%   challenges that needed to be solved while proving the blame theorem for this
-%   paper's semantics?
-% }
-%  Compared to \citet{ruef18checkedc-incr}, proofs for
-%  \lang were made challenging by the addition of dependently typed
-%  functions and dynamic arrays, and the need to handle bounds widening for NT
-%  array pointers. These features required changes in the runtime
-%%  semantics (adding a stack, and dynamically changing bounds) and in
- % compile-time knowledge of them (to soundly typing widened bounds).
-
-% \ignore{
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \hypo{\evalue{n}{\tau} \in \defscope}
-%     \infer1[T-VConst]{\Gamma;\defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%   \end{prooftree}
-%   \qquad
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma;\defscope \vdash_m e : \tau \\
-%         \Gamma;\defscope \vdash_m e_1 : \tau_1 
-%       \end{matrix}
-%     }
-%     \hypo{
-%       \begin{matrix}
-% \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2 \\
-%             \Gamma;\defscope \vdash_m e_2 : \tau_2
-%           \end{matrix}
-%         }
-%     \infer2[T-If]{\Gamma;\defscope \vdash_m \eif{e}{e_1}{e_2} : \tau_3}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma;\defscope \vdash_m x : \tptr[c]{(\tntarray{l}{0}{\tau})}, l \leq 0 \\
-%         \Gamma, x : \tptr[c]{(\tntarray{l}{1}{\tau})};\defscope \vdash_m e_1 : \tau_1 \\
-%         \Gamma;\defscope \vdash_m e_2 : \tau_2 \\
-%         \Gamma \vdash_m \tau_3 = \tau_1 \vee \tau_2
-%       \end{matrix}
-%     }
-%     \infer1[T-IfNT]{\Gamma;\defscope \vdash_m \eif{\estar{x}}{e_1}{e_2} : \tau_3}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         F(f) = \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e  \\
-%         \Gamma; \defscope \vdash_m \tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}] ~~~ 0 \leq i \leq j\\
-%         \Gamma; \defscope \vdash_m e_i : \tau_i' ~~~ 0 \leq i \leq j\\
-%         \Gamma; \defscope \vdash_m \subtype{\tau_i'}{\tau_i[x_0,\ldots,x_{i-1} \mapsto e_0,\ldots,e_{i-1}]}  ~~~ 0 \leq i < j
-%       \end{matrix}
-%     }
-%     \infer1[T-VCall]{\Gamma; \defscope \vdash_m f(\overline{e}) : \tau_j[x_0,\ldots,x_{j-1} \mapsto e_0,\ldots,e_{j-1}]}
-%   \end{prooftree} \\ \\
-% % \inferrule*[lab=T-PtrC]
-% % {
-% %   \tau = \tptr[c]{\omega} \\
-% %   \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\\\
-% %   \wt[\Gamma][\defscope, n^\tau]{H(n+k)}{\tau_k} ~~~ 0 \leq k < j
-% % %  \Gamma,n^\tau \proves H(n+k) : \tau_k~~~0 \leq k < j
-% % }
-
-%   \begin{prooftree}
-%     \hypo{
-%           \Gamma; \defscope \vdash_m e_1 : \tau_1
-%       }
-%       \hypo{        \Gamma, x = e_1 : \tau_1; \defscope \vdash_m e_2 : \tau_2
-% }
-%     \infer2[T-Let]{\Gamma; \defscope \vdash_m \elet{x}{e_1}{e_2} : \tau[x \mapsto e_1]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m \estrlen{y} : \tau_1\\
-%         \Gamma; \defscope \vdash_m y : \tptr[c]{(\tntarray{le}{\_}{\tau_3})}\\
-%         \Gamma, x = \estrlen{y} : \tau_1, \\ y : \tptr[c]{(\tntarray{le}{x}{\tau_3})} ;  \defscope  \vdash_m e_2 : \tau_2
-%         \end{matrix}
-%         }
-%         \infer1[T-LetStr]{\Gamma; \defscope \vdash_m \elet{x}{\estrlen{y}}{e_2} : \tau[x \mapsto e_1]}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{
-%         \Gamma; \defscope \vdash_m e : \tptr[c]{(\tntarray{le}{he}{\tau})}
-%       }
-%         \infer1[T-Str]{\Gamma; \defscope \vdash_m \estrlen{y} : \tint}
-%       \end{prooftree} \\ \\
-
-%       \begin{prooftree}
-%         \hypo{
-%           \begin{matrix}
-%             \tau = \tint \vee \tau = \tptr[u]{\omega}~\vee  n=0~ \vee \\ \tau = \tptr[c]{(\tarray{0}{0}{\tau'})} \vee \\
-%             \tau = \tptr[c]{(\tntarray{0}{0}{\tau'})}
-%         \end{matrix}
-%       }
-%       \infer1[T-Base]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%     \end{prooftree} \\ \\
-
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%         \tau = \tptr[c]{\omega} \\
-%         \tau_0, ..., \tau_{j-1} = \mathrm{types}(D,\omega)\\
-%         \Gamma;\defscope, \evalue{n}{\tau}  \vdash_m {H(n+k)} : {\tau_k} ~~~ 0 \leq k < j
-%       \end{matrix}
-%       }
-%       \infer1[T-PtrC]{\Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tau}
-%     \end{prooftree} \\ \\
-
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%             \Gamma; \defscope \vdash_m e :  {\tptr[m]{\tstruct{T}}} \\
-%             % \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
-%             D(T) = ...; \tau_f~f; ...
-%         \end{matrix}
-%       }
-%       \infer1[T-Amper]{\Gamma; \defscope \vdash_m \eamper{e}{f} : \tptr[m]{\tau_f}}
-%     \end{prooftree} \\ \\
-
-%     \todo[inline]{YL: how to express le - n as a metafunction?}
-%     \begin{prooftree}
-%       \hypo{
-%         \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}\\
-%         \Gamma; \defscope \vdash_m \evalue{n}{\tau}  : \tint \\
-%         le' = le - n, he' = he - n
-%       \end{matrix}
-%     }
-%       \infer1[T-BinopInd]{\Gamma; \defscope \vdash_m (e_1 \plus \evalue{n}{\tau} ) : \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})}}
-%     \end{prooftree}
-% \caption{Typing}
-% \label{fig:typing}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \todo[inline]{no sizeof check because we don't know statically whether the allocation would be null}
-%   \begin{prooftree}
-%     \hypo{
-%       \omega = \tgarray{\alpha}{le}{he}{\tau} \Rightarrow le = 0
-%     }
-%     \infer1[T-Malloc]{\Gamma; \defscope \vdash_m \emalloc{\omega} : \tptr[c]{\omega}}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \Gamma; \defscope \vdash_u e : \tau
-%     }
-%     \infer1[T-Unchecked]{\Gamma; \defscope \vdash_m \eunchecked{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \todo[inline]{any constraints on m'?}
-%   \begin{prooftree}
-%     \hypo{\Gamma; \defscope \vdash_m e :  \tptr[m']{(\tgarray{\alpha}{le'}{he'}{\tau})} }
-%     \infer1[T-DynCast]{\Gamma; \defscope \vdash_m \edyncast{\tptr[m']{(\tgarray{\alpha}{le}{he}{\tau})}}{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{\Gamma; \defscope \vdash_m e : \tau'}
-%     \infer1[T-Cast]{\Gamma; \defscope \vdash_m \ecast{\tau}{e} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\tgarray{\alpha}{le}{he}{\tau}} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tint \\
-%         m' = u \Rightarrow m = u \\
-%       \end{matrix}
-%     }
-%     \infer1[T-Index]{\Gamma; \defscope \vdash_m \estar{(\ebinop{e_1}{e_2})} : \tau}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tau \\
-%         m' = u \Rightarrow m = u \\
-%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-%       \end{matrix}
-%     }
-%     \infer1[T-Assign]{\Gamma; \defscope \vdash_m \eassign{e_1}{e_2} : \tau}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma; \defscope \vdash_m e_1 : \tptr[m']{\omega} \\
-%         \Gamma; \defscope \vdash_m e_2 : \tint \\
-%         \Gamma; \defscope \vdash_m e_3 : \tau \\
-%         m' = u \Rightarrow m = u \\
-%         \omega = \tau \vee \omega = \tgarray{\alpha}{le}{he}{\tau}
-%       \end{matrix}
-%     }
-%     \infer1[T-IndAssign]{\Gamma; \defscope \vdash_m \eassign{(e_1 \plus e_2)}{e_3} : \tau}
-%   \end{prooftree}
-
-% \caption{Typing Cont.}
-% \label{fig:typing2}
-% \end{figure}
-% \ignore{
-
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j} \\
-%         \Gamma, x_0 = \kw{none} : \tau_0, \ldots, x_{j-1} = \kw{none} : \tau_{j-1}; \defscope \vdash e : \tau_{r}
-%       \end{matrix}
-%     }
-%     \infer1[WF-Fun]{\Gamma;\defscope \vdash \tau_{j}\;(x_0:\tau_0, \ldots, x_{j-1}:\tau_{j-1})\;e}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \infer0[WF-Nil]{\Gamma \vdash \circ}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{
-%       \begin{matrix}
-%         \Gamma \vdash \tau_0 \\
-%         \Gamma, x_0 = \kw{none} :\tau_0 \vdash x_1:\tau_{1}, \ldots, x_{j}:\tau_{j}
-%       \end{matrix}
-%     }
-%     \infer1[WF-Cons]{\Gamma \vdash x_0:\tau_0, \ldots, x_{j}:\tau_{j}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \infer0[WF-Int]{\Gamma \vdash \tint}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{   \Gamma \vdash le }
-%      \hypo{   \Gamma \vdash he }
-%       \hypo{  \Gamma \vdash \tau}
-
-%     \infer3[WF-Array]{\Gamma \vdash \tptr[m]{(\tgarray{\alpha}{le}{he}{\tau})}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{T \in dom(D)}
-%     \infer1[WF-Struct]{\Gamma \vdash \tptr[m]{\tstruct{T}}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \hypo{\Gamma \vdash \tau}
-%     \infer1[WF-Ptr]{\Gamma \vdash \tptr[m]{\tau}}
-%   \end{prooftree} \\\\
-
-%   \begin{prooftree}
-%     \infer0[WFB-Int]{\Gamma \vdash i}
-%   \end{prooftree}\\\\
-
-%   \begin{prooftree}
-%     \hypo{x = e? : \tint \in \Gamma}
-%     \infer1[WFB-Var]{\Gamma \vdash x \plus i}
-%   \end{prooftree} \\\\
-
-%   \caption{Well-Formedness}
-%   \label{fig:wf}
-% \end{figure}
-
-% \begin{figure}[t!]
-%   \begin{prooftree}
-%     \infer0[Sub-Nt]{\subtype{\tptr[c]{(\tntarray{le}{he}{\tau})}}{\tptr[c]{(\tarray{le}{he}{\tau})}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \infer0[Sub-Refl]{\subtype{\tau}{\tau}}
-%   \end{prooftree} \\ \\
-
-%   \todo[inline]{what about ntarrays?}
-%   \begin{prooftree}
-%     \infer0[Sub-Ptr]{\subtype{\tptr[c]{\tau}}{\tptr{\tarray{0}{1}{\tau}}}}
-%   \end{prooftree} \\ \\
-
-%   \todo[inline]{the subtyping relation is not anti-symmetric..}
-%   \begin{prooftree}
-%     \infer0[Sub-Arr]{\subtype{\tptr[c]{\tarray{0}{1}{\tau}}}{\tptr{\tau}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     \hypo{\rboundle{le_0}{le_1}}
-%     \hypo{\rboundle{he_1}{he_0}}
-%     \infer2[Sub-Subsume]{\subtype{\tptr{(\tgarray{\alpha}{le}{he}{\tau})}{\cmode}}{ \tptr{(\tgarray{\alpha}{ le_1} {he_1} {\tau})}}}
-%   \end{prooftree} \\ \\
-
-%   \begin{prooftree}
-%     % check t-amper below
-%     \hypo{D(T) = \tau_f~f; ...}
-%     \hypo{\subtype{\tau_f}{\tau}}
-%     \infer2{\subtype{\tptr[c]{\tstruct{T}}}{\tau}}
-%   \end{prooftree}
-%   \caption{Subtyping}
-%   \label{fig:sub}
-% \end{figure}
-% }
-% }
-
-% \ignore{
-
-% \begin{figure*}[t]
-%   \begin{lstlisting}
-% int foo(nt_array_ptr<char> p : count(0)) {
-%   if (* p) {
-%     dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
-%   }
-%   dyn_bounds_cast<nt_array_ptr<char>>(p, count(1));
-%   return 0;
-% }
-%   \end{lstlisting}
-% \caption{The example where clang Checked C fails at run-time}
-% \label{fig:clangbad1}
-% \end{figure*}
-
-% \begin{figure*}[t]
-%   \begin{lstlisting}
-% /* nt_array_ptr<char> p : bounds(p,p) */
-% size_t cnt = 0;
-
-% while(*(p+cnt)) {
-%   ++cnt;
-% }
-% dyn_bounds_cast<nt_array_ptr<char>>(p, count(cnt));
-%   \end{lstlisting}
-% \caption{A useful program that the Checked C spec doesn't allow at run-time}
-% \label{fig:clangbad2}
-% \end{figure*}
-% }
-
-% \ignore{
-% The Clang CheckedC implementation uses statically determined
-% bounds to insert run-time checks. In Fig.~\ref{fig:clangbad1}, the
-% \code{dyn_bounds_cast} at line 3 will always succeed, because the
-% compiler knows that within the scope of then branch, the pointer
-% \code{p} must have at least one element. The same cast at line 5,
-% however, will always fail, since there is no way to tell statically
-% whether the program has entered the then branch before. The compiler
-% will check whether the \code{count(1)} bounds specification is
-% contained within the earlier \code{count(0)} specification, resulting
-% in a run-time failure even when we pass in a non-empty string.
-
-% Our formalization diverges from this run-time behavior and instead keeps
-% track of the bounds on the stack. After entering the then branch, we
-% increment the upper bound for \code{p}, effectively making the
-% updated bounds information available even after we exit the if
-% statement. The cast at line 5 will be checking the new bounds against
-% the incremented bounds for non-empty strings.
-
-% Fig.~\ref{fig:clangbad2} gives a more practical example of why keeping
-% track of the bounds on the stack is useful. The program snippet
-% implements the functionality of the \code{strlen} function using a
-% while loop and a \code{cnt} variable. Even though the type system is
-% unable to reason about the while loop, as long as the runtime system
-% updates the bounds in-place, the user can apply a
-% \code{dyn_bounds_cast} to soundly recover the more precise bounds
-% information.
-% }
-
-% \inferrule*[lab=T-Amper]
-% {
-%   \Gamma \proves e : \tptr[m]{\tstruct{T}} \\\\
-%   D(T) = ...; \tau_f~f; ...
-% }
-% {
-%   \Gamma \proves \eamper{e}{f} : \tptr[m]{\tau_f}
-% }
-
-\subsection{Differences with the Coq and Redex Models}\label{app:model-diffs}
-
-% \begin{figure}[h]
-% {\small
-% {\captionsetup[lstlisting]{margin = 8 mm}
-%   \begin{lstlisting}[xleftmargin=8 mm]
-% Inductive expression : Type :=
-%   | ELit : Z -> type -> expression
-%   | EVar : var -> expression
-%   | EStrlen : var -> expression
-%   | ECall : funid -> list expression -> expression
-%   | ERet : var -> Z* type -> expression -> expression
-%   | EDynCast : type -> expression -> expression
-%   | ELet : var -> expression -> expression -> expression
-%   | EMalloc : type -> expression
-%   | ECast : type -> expression -> expression
-%   | EPlus : expression -> expression -> expression
-%   | EFieldAddr : expression -> field -> expression
-%   | EDeref : expression -> expression (*  * e *)
-%   | EAssign : expression -> expression -> expression (* *e = e *)
-%   | EIfDef : var -> expression -> expression -> expression (* if *x then e1 else e2. *)
-%   | EIf : expression -> expression -> expression -> expression (* if e1 then e2 else e3. *)
-%   | EUnchecked : expression -> expression.
-%   \end{lstlisting}
-% }
-% }
-% \caption{Expression Syntax in Coq}
-% \label{fig:coq-syntax}
-% \end{figure}
-
-The Coq and Redex models of \lang may be found at \url{https://github.com/plum-umd/checkedc}.
-The Coq model's syntax is slightly different from that in
-Fig.~\ref{fig:checkc-syn}. In particular, the arguments in a function call
-are restricted to variables and constants, according to
-a separate well-formedness condition. A function call \code{f(e)} can
-always be written in \code{let x = e in f(x)} to cope.
-% In the Coq model, we do not restrict the function argument syntax. Instead, we achieve the restriction
-% by restricting the type rule for each function argument element to only deal with constants and variables. 
-% To avoid recursive inductive relations and simplify proofs, we define different inductive relations for typing function argument lists and elements.
-In the inductive relation of function argument elements, we only define rules for typing constants and variables, and the rules are copied from the main inductive type rules for constants and variables.
-%MWH: The above is nonsensical. The answer is: We have an expresion
-%well-formedness condition that restricts this syntax, among other things.
-In addition,
-conditionals have two syntactic forms: \code{EIf} is a normal
-conditional, and \code{EIfDef} is one whose boolean guard is of the
-form \estar{x}. By syntactically distinguishing these two
-cases, the Coq model does not need the \emph{[prefer]} rule for
-\code{if (*x)}$...$ forms as in
-Fig.~\ref{fig:c-context}. The Redex model \emph{does} prioritize such
-forms but not the same way as in the figure. It uses a variation of
-the \textsc{S-Var} rule: The 
-modified rule is equipped with a precondition that is false whenever
-\textsc{S-IfNTT} is applicable.
-% \mwh{Do we need
-%   to talk about other semantic judgments that are not in the paper,
-%   e.g., well-formedness? Maybe not, but should make sure the code is
-%   commented to clarify that these things are not strictly needed --
-%   well-formedness is subsumed by well typing (if that's true)?} \liyi{I will add comments to make sure it is clear in Coq.}
-% \yiyun{No, well-formedness is not subsumed by the typing relation in
-%   the Coq model. I
-%   checked the latest coq code; they must be explicitly stated in the theorems.}
-% \mwh{I mean: We might have it in the coq model, but I suspect it is
-%   not necessary. If it is, we should be saying something in the main
-%   paper.
-
-The Coq model uses a runtime stack $\varphi$ as described at the start
-of Sec.~\ref{sec:semantics}.
-The Redex model introduces let bindings during evaluation to
-simulate a runtime stack. For example, consider the expression 
-$e \equiv \elet{x}{(\evalue{5}{\tint})}{\ebinop{x}{x}}$. Expression $e$ first steps to
-$\elet{x}{(\evalue{5}{\tint})}{\ebinop{(\evalue{5}{\tint})}{x}}$, which in
-turns steps to
-$\elet{x}{(\evalue{5}{\tint})}{\ebinop{(\evalue{5}{\tint})}{(\evalue{5}{\tint})}}$. Since
-the rhs of $x$ is a value, the let binding in $e$ effectively
-functions as a stack that maps from $x$ to $\evalue{5}{\tint}$. The
-let form remains in the expression and lazily replaces the variables
-in its body. The let form can be removed from the expression only if its
-body is evaluated to a value, e.g., 
-$\elet{x}{(\evalue{5}{\tint})}{(\evalue{10}{\tint})}$
-steps to 
-$\evalue{10}{\tint}$. The rule for popping let bindings in this manner
-corresponds to the \textsc{S-Ret} rule in Fig.~\ref{fig:semantics}.
-Leveraging let bindings adds complexity to the semantics but
-simplifies typing/consistency and term generation during randomized testing.
-%  \mwh{I don't understand the last sentence. Nowhere in the
-%   paper do we talk about ANF. Also: Isn't it the case that there are
-%   rules for let bindings do not use substitution, but rather ``look
-%   up'' the binding on the definition chain? E.g.,
-%   \code{let x=5 in x+x} does not step to \code{5+5} but instead steps
-%   to \code{let x=5 in 5+5} and then to \code{let x=5 in 10} ? }
-% \liyi{Yiyun: please see here. If you do not know, please explain to me and then we can make a correct statement. }
-% \yiyun{It is supposed to refer to the ANF transformation of the erased language, not the source
-%   language. I moved it to the discussion of ANF form below in \ref{appx:comp1}. }
-
-\subsection{The Compilation Rules}\label{appx:comp1}
-
-% \review{
-% The first reviewer said:
-% \begin{itemize}
-% \item I had a really hard time understanding precisely the "formalized" compilation
-%   from CoreChkC to CoreC. Specifically: is CoreC intended to model LLVM-IR, or
-%   is it a subset of C? Is CheckedC compiling to C or is it a new frontend like
-%   `clang`? See remark about undefined behavior (UB) below which got me even more
-%   confused. \mwh{CoreC is meant to model CoreChkC but with
-%   annotations, and checks removed; it is not meant to model C or LLVM-IR.}
-% \item Is the compilation scheme from CoreChkC to CoreC is faithful to the actual
-%   compilation scheme of the checkedc-clang compiler to ...? I feel like the
-%   paper is missing an actual description of what happens in the compiler to
-%   allow us to connect the dots and understand how the formalization illuminates
-%   the implementation. \mwh{We did not model compilation on the real
-%   implementation; our purpose was to show that annotations in CoreChkC
-%   do not necessitate fat pointers in an implementation; that said,
-%   our formalization does show how a real implementation can be
-%   carried out}
-% \item The paper doesn't seem upfront about what is *shown* (theorem in Coq) and what
-%   is *tested* (via PLT-Redex), and thus remains a
-%   conjecture. \mwh{Updated the intro and the individual sections}
-% \item  Missing discussion of why Coq vs. PLT-Redex, effort involved, any plans to
-%   formally prove compilation from CoreChkC to CoreC, any hopes of integrating
-%   that in the official implementation, etc. \mwh{Added note at the
-%   start of section 3.}
-% \end{itemize} }
-
-% \review{IV: ghost variables in other contexts (e.g. Why3, Dafny) are used for things
-%   that do not exist at run-time, but this doesn't seem to be the case here.}
-% \yiyun{Agreed: We changed the name to ``shadow variables'' to avoid confusion}
-% \review{Do your bug report and github links break anonymity?}
-
-% \review{From reviewer C:
-% \begin{itemize}
-% \item I'd like to have seen a bit more motivation for using PLT redex:
-%   what aspects made the use of this tool preferable to formulating the
-%   compilation in Coq and using Quickchick to do random testing of the
-%   simulation result.
-%   \mwh{Added some text to the end of III.A and IV.C}
-% \yiyun{Some reasons I can think of:\begin{itemize}
-%     \item Redex is highly optimized for specifying judgments that are algorithmic. By writing down a typing relation, we can immediately obtain a typechecker
-%     that is executable.
-Same applies for the small-step evaluation relation.
-Translating the relations into functions in Coq is definitely doable but time-consuming, especially since compilation is embedded as part of the typing rules.
-It is also hard to see whether the function we define really corresponds to the relation unless we formally prove it.
-This issue is particularly relevant at the early stage of the development when the compilation rules were buggy and the simulation property was violated often as we added new generator cases.
-%     In Redex, we don't have any formal guarantee either, but at least we can more easily see the correspondence because Redex is able to convert the relation into an executable version so we can specify the relation literally.
-This feature of Redex helped us speed up our development significantly when our compilation rules were constantly changing.
-% \end{itemize} }
-% \item there's a funny change of line spacing in column 2 of page 6, about 2/3 down.
-% \end{itemize}
-% }
-
-The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
-that track array bound information, which is used in premises
-of rules like \textsc{S-DefArray} and
-  \textsc{S-AssignArr} to prevent spatial safety violations.
-The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
-for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
-The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.
-
-\begin{figure}[t!]
-{\small
-\hspace*{-0.5em}
-\begin{tabular}{|c|c|c|c|}
-\hline
-& \cmode & \tmode & \umode \\
-\hline
-& \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} \\
-\hline
-\cmode & $\estar{x}$ / $\getstar{\cmode}{x}$ 
- & $\texttt{sand\_get}(x)$ / $\getstar{\umode}{x}$ &  $\times$ \\
-\hline
-\umode & $\times$
- & $\estar{x}$ / $\getstar{\umode}{x}$ &  $\estar{x}$ / $\getstar{\umode}{x}$ \\
-\hline
-\end{tabular}
-
-}
-\caption{Compiled Targets for Dereference}
-\label{fig:flagtable}
-\end{figure}
-
-In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
-i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
-\lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
-Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
-In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different.
-\lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
-\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
-The columns represent different pointer modes and the rows represent context modes.
-For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation.
-In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
-If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
-
-This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
-with no loss of expressiveness
-as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
-For the compiler formalism, 
-we present a compilation algorithm that converts from
-\lang to \elang, an untyped language without metadata
-annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
-In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
-$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
-The algorithm sheds
-  light on how compilation can be implemented in the real Checked C
-  compiler, while eschewing many vital details (\elang has many 
-  differences with LLVM IR).
-
-%This section shows how \systemname deals with 
-%annotations can be safely erased: using static information a compiler
-%can insert code to manage and check bounds metadata, with no loss of
-%expressiveness. We present a compilation algorithm that converts from
-%\lang to \elang, an untyped language without metadata
-%annotations. The syntax and semantics \elang
-  %closely mirrors that of \lang; it differs only in that literals lack
-  %type annotations and its operational rules perform no
-  %bounds and null checks, which are instead inserted during
-  %compilation. Our compilation algorithm is evidence that \lang's
-  %semantics, despite its apparent use of fat pointers, faithfully
-  %represents Checked C's intended behavior. The algorithm also sheds
-  %light on how compilation can be implemented in the real Checked C
-  %compiler, while eschewing many important details (\elang has many 
-  %differences with LLVM IR).
-
-Compilation is defined by extending \lang's
-typing judgment as follows:
-\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
-There is now a \elang output $\dot e$ and an input $\rho$, which maps
-each (NT-)array pointer variable to its mode and
-each variable \code{p} to a pair of \emph{shadow
-  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
-These may differ from the bounds in \code{p}'s type due to bounds
-widening.\footnote{Since lower bounds are never widened, the
-  lower-bound shadow variable is unnecessary; we include it for uniformity.} 
-\lc{These two sentences are confusing. }
-
-% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
-% complete judgment, implicitly assuming that $e$ is a well-typed and closed
-% term.
-
-We formalize rules for this judgment in PLT Redex~\cite{pltredex},
-following and extending our Coq development for \lang. To give
-confidence that compilation is correct, we use Redex's property-based
-random testing support to show that compiled-to $\dot e $ simulates
-$e$, for all $e$.
-
-% We developed a \checkedc compiler to compile a \checkedc program to a C program.
-% \mwh{We formalized compilation from CoreChkC to a version of CoreChkC
-%   but with the metadata removed, right? This is not a Checked C
-%   compiler. You go on to see stuff about CompCert, CLight,
-%   etc. This is confusing. We should be talking about how this relates
-%   to what was just presented. Pick definitive names for things. }
-% Given a \checkedc program $e$, we build a compilation process ($\gg$), such that $e \gg \dot e$, where $\dot e$ is the corresponding C program for $e$ in A-normal form (ANF). 
-% The compilation process ($\gg$) relies on the type checking step $\Gamma;\Theta\vdash_m e:\tau$. 
-% Especially, it relies on $\Gamma$ to provide the type information for variables in $e$. 
-% We utilize CompCert/CLight syntax and semantics \cite{Leroy:2009:FVC:1666192.1666216,Blazy2009} as our translation target language.
-%  Besides, we defined two data structures in the CompCert format for representing $\enull$ and $\ebounds$ states.
-% We write $\xrightarrow{c}$ for the semantics of CLight. \mwh{Of our
-%   target language, which I presume does not have all the features that
-%   CLight has. Should mention up front that we did all of this in PLT Redex.}
-
-Due to space constraints, we explain the rules for compilation by
-examples, using a C-like syntax; the complete rules are given in
-\iftr
-Appendix~\ref{appx:comp1}.
-\else
-the supplemental report~\cite{checkedc-tech-report}.
-\fi
-Each rule performs up to three tasks: (a) conversion of $e$ to
-A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
-and (c) generate right pointer accessing expressions based on modes.
-%
-A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
-as in the following example.
-
-{\vspace*{-0.5em}
-{\small
-\begin{center}
-$
-\begin{array}{l}
-$\code{let y=(x+1)+(6+1)}$
-\;
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\;
-\begin{array}{l}
-$\code{let a=x+1;}$\\
-$\code{let b=6+1;}$\\
-$\code{let y=a+b}$\\
-\end{array}
-\end{array}
-$
-\end{center}
-}
-}
-
-This simplifies the management of effects from subexpressions. The
-next two steps of compilation are more interesting.
-We state them based on different \lang operations.
-
-\subsection{Metatheory}
-\label{sec:meta}
-
-% \review{ 
-% My interpretation of section IV.C is that the authors have a pen-and-paper
-%   proof of Theorem 4, and that random testing using the Redex models has been
-%   used to gain confidence in such a proof. Is this correct?}
-% \liyi{Yes, this is correct.}
-% \mwh{No, not correct: We have no pen-and-paper proof, just the
-%   model. Clarify here}
-% showing that
-% extensive random testing fails to falsify the bisimilarity
-% property. \mwh{Better way to state the previous?}
-
-We formalize both the compilation procedure and the simulation
-theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
-and then attempt to falsify it via Redex's support for random
-testing. Redex allows us
-  to specify compilation as logical rules (an extension
-  of typing), but then execute it algorithmically to
-  automatically test whether simulation holds. This process revealed
-  several bugs in compilation and the theorem statement.
-%
-  % us gain confidence that our original pen and paper proof of
-  % simulation remains true with the addition of variable bounds. }
-We ultimately plan to prove simulation in the Coq model.
-
-%Turning to the simulation theorem: We first introduce notation
-%used to specify the theorem.
-We use the notation $\gg$ to
-indicate the \emph{erasure} of stack and heap---the rhs is the same as
-the lhs but with type annotations removed:
-\begin{equation*}
-  \begin{split}
-    \heap  \gg & \dot \heap \\
-    \varphi \gg & \dot \varphi
-  \end{split}
-\end{equation*}
-In addition, when $\Gamma;\emptyset\vdash
-\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
-\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
-and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively.
-$\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
-Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
-reduction relation of $\elang$. Unlike the $\lang$, the semantics of
-$\elang$ does not distinguish checked and unchecked regions.
-
-Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
-the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
-that is similar to the one by~\citet{merigoux2021catala}.
-
-An ordinary simulation property would
-replace the middle and bottom parts of the figure with the
-following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
-  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
-Instead, we relate two erased configurations using the relation $\sim$,
-which only requires that the two configurations will eventually reduce
-to the same state.
-
-\ignore{
- We formulate our simulation theorem differently
-because the standard simulation theorem imposes a very strong
-syntactic restriction to the compilation strategy. Very often, $(\dot
-\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
-semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
-but we are unable to syntactically equate the two configurations due
-to the extra binders generated for dynamic checks and ANF
-transformation. In earlier versions of the Redex model, we attempted
-to change the compilation rules so the configurations could match
-syntactically. However, the approach scaled poorly as we added
-additional rules. 
-This slight relaxation on the equivalence relation
-between target configurations allows us to specify compilation more
-naturally without having to worry about syntactic constraints.
-}
-
-% The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
-\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-if $\heap_0 \vdash \varphi_0$, $(\varphi_0,\heap_0,e_0)\gg_c (\dot \varphi_0,\dot \heap_0, \dot e_0)$,
-and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-\rightarrow_c (\varphi_1, \heap_1, r_1)$, then the following facts hold:
-
-\begin{itemize}
-
-\item if there exists $e_1$ such that $r=e_1$ and $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-$(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\varphi,\dot \heap,\dot e)$ and $(\dot
-\varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-\dot \heap,\dot e)$.
-
-\item if $r_1 = \ebounds$ or $\enull$, then we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-\dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-\varphi_1$, $\heap_1 \gg \dot \heap_1$.
-
-\end{itemize}
-\end{thm}
-
-% when $r_1 = e_1$ for
-% some $e_1$ and
-% $(\varphi_1, \heap_1, e_1) \gg (\dot \varphi_1, \dot \heap_1, \dot e_1)$, then
-% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \varphi,\dot \heap,\dot e)$ and $(\dot
-% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
-
-% \begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
-% For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 
-% if $\emptyset;\emptyset;\emptyset \vdash_\cmode e_0 \gg \dot e_0 :\tau_0$,
-% and if there exists some $r_1$ such that $(\varphi_0, \heap_0, e_0)
-% \rightarrow_\cmode (\varphi_1, \heap_1, r_1)$, when $r_1 = e_1$ for
-% some $e_1$ and
-% $\emptyset;\emptyset;\emptyset \vdash_\cmode e_1 \gg \dot e_1 :\tau_1$ where $\tau_1 \sqsubseteq \tau_0$
-% , then
-% there exists some $\dot \varphi$,$\dot \heap$, $\dot e$, such that
-% $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \varphi,\dot \heap,\dot e)$ and $(\dot
-% \varphi_1,\dot \heap_1,\dot e_1) \xrightarrow{\cdot}^* (\dot \varphi,
-% \dot \heap,\dot e)$. When $r_1 = \ebounds$ or $\enull$, we have $(\dot \varphi_0, \dot \heap_0,\dot e_0) \xrightarrow{\cdot}^* (\dot
-% \dot \varphi_1,\dot \heap_1, r_1)$ where $\varphi_1 \gg \dot
-% \varphi_1$, $\heap_1 \gg \dot \heap_1$.
-% \end{thm}
-
-As our random generator never generates
-$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
-as it relates to checked code. This limitation makes it
-unnecessary to state the other direction of the simulation theorem
-where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
-that $e_0$ will never enter a stuck state if it is well-typed in
-checked mode.
-
-The current version of the Redex model has been tested against $23000$
-expressions with depth less than $11$. Each expression can
-reduce multiple steps, and we test simulation between every two
-adjacent steps to cover a wider range of programs, particularly the
-ones that have a non-empty heap.
-
-\begin{figure}[t]
-{\small
-\[
-\begin{array}{c}
-\begin{tikzpicture}[
-            > = stealth, % arrow head style
-            shorten > = 1pt, % don't touch arrow head to node
-            auto,
-            node distance = 3cm
-        ]
-
-\begin{scope}[every node/.style={draw}]
-    \node (A) at (0,1.5) {$\varphi_0,\heap_0, e_0$};
-    \node (B) at (4,1.5) {$\varphi_1, \heap_1 ,e_1$};
-    \node (C) at (0,0) {$\dot \varphi_0, \dot \heap_0 ,\dot e_0$};
-    \node (D) at (4,0) {$\dot \varphi_1, \dot \heap_1, \dot e_1$};
-    \node (E) at (2,-1.5) {$\dot \varphi,\dot \heap ,\dot e$};
-\end{scope}
-\begin{scope}[every edge/.style={draw=black}]
-
-    \path [->] (A) edge node {$\longrightarrow_{\cmode}$} (B);
-    \path [<->] (A) edge node {$\gg$} (C);
-    \path [<->] (B) edge node {$\gg$} (D);
-    \path [dashed,<->] (C) edge node {$\sim$} (D);
-    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
-    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
-\end{scope}
-
-\end{tikzpicture}
-\end{array}
-\]
-}
-\caption{Simulation between \lang and \elang }
-\label{fig:checkedc-simulation-ref}
-\end{figure}
-
-% \ignore
-% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
-% % The diagram also indicates the simulation relation ($\sim$) among our formalization of the \checkedc Type System ($\vdash$), Semantics ($\longrightarrow$), Compilation ($\gg$), as well as the compiled C semantics ($\xrightarrow{c}$).
-% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
-% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist.
-The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.
-
-%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C.
-The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
-% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).
-
-% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
-% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
-% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation.
-We generate tens of thousands cases and cover all possible corner cases of the compilation.
-% }
-% \ignore{
-% \begin{figure*}[t]
-%   \begin{subfigure}[b]{1\textwidth}
-%     \begin{lstlisting}
-% int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
-%   return *p;
-% }
-
-% ...
-% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
-% deref_array(5, p0);
-%     \end{lstlisting}
-%     \label{fig:chkcexamplederef}
-%     \caption{\texttt{deref\_array} in \checkedc}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{1\textwidth}
-% \begin{lstlisting}
-% int deref_array(size_t n, int *p) {
-%   /* statically determine the definitions p_lo and p_hi */
-%   int *p_lo = p;
-%   int *p_hi = p + n;
-%   /* runtime checks */
-%   assert(p_lo <= p && p <= p_hi);
-%   assert(p != NULL);
-%   /* widening */
-%   int p_derefed = *p;
-%   if (p_derefed != '\0') {
-%     if (p_hi == p) {
-%       ++p_hi;
-%     }
-%   }
-%   return p_derefed;
-% }
-% ...
-% /* int *p0 */
-% /* rho(p0) = p_lo, p_hi */
-% deref_array(5, p0);
-% \end{lstlisting}
-% \caption{\texttt{deref\_array} in C}
-%     \label{fig:cexamplederef}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{\textwidth}
-%     \begin{lstlisting}
-% /* nt_array_ptr<int> p : bounds(e0, e1) */
-% /* rho(p) = p_lo, p_hi */
-% size_t x = strlen(p);
-% if(x >= 1) {
-%   *(p+1);
-% }
-%     \end{lstlisting}
-%     \caption{\texttt{strlen} in \checkedc}
-%     \label{fig:chkcexamplestrlen}
-%   \end{subfigure}
-
-%   \begin{subfigure}[b]{\textwidth}
-%     \begin{lstlisting}
-% /* nt_array_ptr<int> p : bounds(e0, e1) */
-% /* rho(p) = p_lo, p_hi */
-% /* runtime checks omitted */
-% ...
-% size_t x = strlen(p);
-% int *p_hi_new = p + x;
-% p_hi = max(p_hi, p_hi_new);
-% if (x >= 1) {
-%   /* null check for pointer arithmetic */
-%   assert(p != NULL);
-%   int *p_1 = p + 1;
-%   /* null check for dereference */
-%   assert(p_1 != NULL);
-%   /* bounds check for dereference */
-%   /* note how we are using p_hi rather than p + x */
-%   assert(p_lo <= p_1 && p_1 <= p_hi);
-%   ...
-% }
-%     \end{lstlisting}
-%     \caption{\texttt{strlen} in C}
-%     \label{fig:cexamplestrlen}
-%   \end{subfigure}
-% \caption{Compilation of null-terminated string dereference}
-% \label{fig:compilationexample}
-% \end{figure*}
-% }
-
-% \ignore{
-% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
-% dereference of null-terminated strings by inserting explicit checks and
-% bounds widening code. For readability, we present the example in real
-% \checkedc and C syntax. At
-% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
-% bounds are defined in terms of the arguments \code{p} and \code{n}
-% according to the bounds annotations obtained during typechecking. This
-% avoids the need to pass in bounds as extra arguments, thus maintaining
-% compatibility with C code. When we call \code{deref_array} on
-% \code{p0}, a string with size 5 (excluding the null character), there
-% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
-% the stack. It is possible that \code{p_hi} has been widened
-% before we reach line 21. The programmer will have to perform a
-% \code{dynamic_bounds_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
-% attempts to widen the bound when the dereferenced result is not
-% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
-% \code{deref_array}, in contrast to the T-IfNT rule, which only
-% remembers the widened bounds within the scope of the then branch of
-% the if statement.
-
-% \todo[inline]{Widening can happen at every
-%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
-% shows how an invocation of \code{strlen} on null-termianted strings
-% is compiled into C code. We perform the same runtime checks that
-% happen during dereferencing. The widening code at line
-% 8 updates \code{p}'s upper bound only if the result of
-% \code{strlen} is larger than the value of the upper bound stored on
-% the stack. This is another scenario where the runtime can be more precise
-% than the statically determined bounds information.
-
-% }
-
-Since bounds expressions may
-contain variables, determining assumptions like $b_l \leq b_l'$
-requires reasoning about the probable values of these variables'. The type
-system uses $\Theta$ to make such reasoning more precise.
-$\Theta$ is a map from variables $x$ to
-equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
-It maps variables to equations that are recorded along the type checking procedure.
-If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
-$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
-such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
-\iftr
-  Appendix~\ref{app:le}.
-\else
-  the supplemental report~\cite{checkedc-tech-report}.
-\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.
-
-\begin{figure}[t]
-{\small
-{\captionsetup[lstlisting]{margin = 8 mm}
-  \begin{lstlisting}[xleftmargin=8 mm]
-nt_array_ptr<char> safe_strcat
-   (nt_array_ptr<char> dst : count(n),
-    nt_array_ptr<char> src : count(0), int n) {
-  int x = strlen(dst);
-  int y = strlen(src);
-  nt_array_ptr<char> c : count(n) =
-    dyn_bounds_cast
-           <nt_array_ptr<char>>(dst,count(n));
-    // sets c == dst with bound n (not x)
-  if (x+y < n) {
-    for (int i = 0; i < y; ++i)
-      *(c+x+i) = *(src+i);
-    *(c+x+y) = '\0';
-    return dst;
-  }
-  return null;
-}
-  \end{lstlisting}
-}
-}
-\caption{Implementation of safe \code{strcat}}
-\label{fig:strcat-ex}
-\end{figure}
-
-Recall that array
-bounds in types may refer to in-scope variables; e.g., parameter
-\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
-2-3 in \Cref{fig:strcat-ex}. 
-All dependent bound variables appearing in function headers 
-are bounded by a variable list in the function pointer type, which is checked in the type rule.
-Semantically,
-the call is expanded into a \texttt{let} which binds
-parameter variables $\overline{x}$ to the actual arguments
-$\overline{n}$, but annotated with the parameter types
-$\overline{\tau}$ (this will be safe for type-correct programs). 
-The function body $e$ is wrapped in a static cast
-$(\tau[\overline{n} / \overline{x}])$ which is the function's return
-type but with any integer parameter variables $\overline{x}$ appearing in that
-type, as type bound variables,
-substituted with the call's actual arguments $\overline{n}$.
-To see why this is needed, suppose that \code{safe_strcat} in
-Fig.~\ref{fig:strcat-ex} is defined to return a
-\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
-perform a \code{safe_strcat} function call as
-\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
-function returns a value with type \code{nt_array_ptr<int>:count(10)}
-because we substitute bound variable \code{n} in the 
-defined return type with \code{10} from the function call's
-argument list.
-
-\begin{figure}[t!]
-  \begin{small}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-int deref_array(n : int,
-     p :  $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\cmode}$,
-     q : $\color{green!40!black}\tntarrayptr{0}{n}{\tint}{\tmode}$) {
-  /* $\color{purple!40!black}\rho$(p) = p_lo,p_hi,p_m */
-  /* $\color{purple!40!black}\rho$(q) = q_lo,q_hi,q_m */
-    * p;
-    * q = 1;
-}
-...
-/* p0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\cmode}$ */
-/* q0 : $\color{purple!40!black}\tntarrayptr{0}{5}{\tint}{\tmode}$ */
-deref_array(5, p0, q0);
-    \end{lstlisting}
-\begin{frame}
-
-\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);
-
-\end{frame}
-\begin{lstlisting}[mathescape,xleftmargin=4 mm]
-deref_array(int n, int* p, int * q) {
-  //m is the current context mode
-  let p_lo = 0; let p_hi = n; 
-  let q_lo = 0; let q_hi = n; 
-  /* runtime checks */
-  assert(p_lo <= 0 && 0 <= p_hi);
-  assert(p != 0);
-  *(mode(p) $\wedge$ m,p);
-  verify(q, not_null(m, q_lo, q_hi) 
-             && q_lo <= 0 && 0 <= q_hi);
-  *(mode(q) $\wedge$ m,q)=1;
-}
-...
-deref_array(5, p0, q0);
-    \end{lstlisting}
-\end{small}
-    \caption{Compilation Example for Dependent Functions}
-\label{fig:compilationexample1}
-\end{figure}
-
-For compiling dependent function calls,
-\Cref{fig:compilationexample1} provides a hint.
-Notice that the bounds for the array pointer \code{p} are not passed as
-arguments. Instead, they are initialized according to \code{p}'s
-type---see line~4 of the original \lang program at the top of the figure.
-Line~$3$ of the generated code
-sets the lower bound  to \code{0} and the
-upper bound to \code{n}.
-
-\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
-A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
-\end{defi}
-
-\begin{defi}[Heap Well-formedness]
-For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
-(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
-contains no free variables. 
-\end{defi}
-
-\begin{defi}[Stack Well-formedness]
-A stack snapshot $\varphi$ is well-formed if
-for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
-contains no free variables. 
-\end{defi}
-
-We also need to introduce a notion of
-\emph{consistency}, relating heap environments before and after a
-reduction step, and type environments, predicate sets, and stack
-snapshots together.
-
-\begin{defi}[Stack Consistency]
-A type environment $\Gamma$, variable predicate set $\Theta$, and
-stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
-\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
-$\Gamma(x) = \tau$ for some $\tau$ and 
-$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
-\end{defi}
-
-\begin{defi}[Checked Stack-Heap Consistency]
-A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
-for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
-\end{defi}
-
-\begin{defi}[Checked Heap-Heap Consistency]
-A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
-for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
-\end{defi}
 

### 2023-01-15 sandp23/appendix.tex
finished appx.

@@
 
-\subsubsection{parsons}
+Here, we provide the description of additional program evaluations.
+
+\myparagraph{parsons}
 Parsons is annotated comprehensively in two variants parsons\_wasm and parsons\_tainted.
@@
 
-\subsubsection{LibPNG}
+\myparagraph{LibPNG}
 \systemname changes for libPNG is narrow in scope and begins with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read().
@@
 
-\subsubsection{MicroHTTPD}
+\myparagraph{MicroHTTPD}
 MicroHTTPD demonstrates the practical difficulties in converting a program to \systemname.
@@
 
-\subsubsection{Tiny-bignum}
+\myparagraph{Tiny-bignum}
 Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive.

### 2023-01-15 sandp23/formal.tex
update.

@@
 From those, we develop~\lang’s meta-
-theories, including the type soundness, non-exposure, non-crashing,
-and the compiler simulation theorems.
+theories, including the type soundness, non-exposure, and non-crashing theorems.
 
@@
 \subsection{Syntax}\label{sec:syntax}
-The~\fig{fig:checkc-syn} shows the syntax of~\lang.
+\Cref{fig:checkc-syn} shows the syntax of~\lang.
 
@@
 \myparagraph{Disallowing Unsafe Types}
-The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
-% 
-These requirements prevents unsafe types from being constructed.
+The well-formedness of these types are presented in~\Cref{app:le}.
+It prevents unsafe types from being constructed.
   Consider the type ~\code{_t_Array_ptr<_Ptr<int>>} which describes a tainted
   array of checked pointers.
-  Syntactically valid, this type is however \emph{illegal} in~\lang{} and
-  is rejected by the well-formedness requirement. 
+  Syntactically, this type is however \emph{illegal} in~\lang{} and
+  is rejected by the well-formedness. 
 However, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
@@
 \]
-The function type also has well-formed requirements (\fig{fig:wffunctions} in Appendix), which disallows nesting checked pointers inside tainted pointers.
+The function type also has well-formed requirements (\Cref{app:le}), which disallows nesting checked pointers inside tainted pointers.
 Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
@@
 along with expressions that require special handling, such as,
- static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}}, the \texttt{strlen} operation ($\estrlen{x}$),
+ static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \cite{li22checkedc}}, the \texttt{strlen} operation ($\estrlen{x}$),
 memory allocation ($\emalloc{\xi}{\omega}$), 
@@
 }
-\caption{Selected type rules}
+\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.
+First line is for cast operations, second line is for pointer dereferences, third line is for checked/unchecked blocks, and the rest is for function calls.}
 \label{fig:type-system-1}
@@
 maps addresses (integer literals) to values $\evalue{n}{\tau}$.
-The complete set of typing rules and special handling of (NT)-arrays are provided in~\apdx{sec:literal-pointer-typing}.
+The complete set of typing rules and special handling of (NT)-arrays are provided in~\Cref{rem-type,sec:rem-semantics}.
 
@@
 determines the context mode that the expression $e'$ locates based on the context $E$.
-In \Cref{lst:humantaint}, the function call expression \code{read_msg} has $\umode$ mode since it is inside a tainted function.
+In \Cref{lst:humantaint}, the function call \code{handle_request} has $\umode$ context mode (is in $\umode$ region) since it is inside an unchecked function \code{server_loop}.
 The second rule describes the exception handling 
 for possible crashing behaviors in unchecked regions.
-A $\umode$ mode operation can non-deterministically crash
+Operations in $\umode$ region can non-deterministically crash
 and the \systemname sandbox mechanism recovers
@@
 In the above example, a $\tmode$ mode pointer can be cast to $\umode$ mode but casting $\tmode$ mode to $\cmode$ mode is disallowed.
-The complete subtyping relation was described in \Cref{fig:checkc-subtype}.
+The complete subtyping relation was described in \Cref{app:le}.
 Notice that \texttt{let} statements are immutable in \lang, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \lang.
@@
 \lang captures the runtime error.
-Type and semantic rules for array types and pointer assignments are given in \Cref{appx:add-type-sem}.
+Type and semantic rules for array types and pointer assignments are given in \Cref{rem-type,sec:rem-semantics}.
 
@@
 \myparagraph{Dependent Function Pointers} %
-The rule \textsc{T-Fun} (\Cref{fig:type-system-1}) states the type judgement for
-dependent function pointer application where we represent the result of
+Rule \textsc{T-Fun} (\Cref{fig:type-system-1}) states the type judgment for
+dependent function pointer application, where we represent the result of
 replacing all integer bound variables $\overline{x}$ in the type \(\tau\) with
@@
 \(\tau''\).
-% 
-\mz{I rewrote the sentences above.}
-% 
 Consider the \code{process_req2} function in
 Fig.~\ref{lst:final}; its parameter type for \code{msg} 
-depends on \code{m_1}. The \textsc{T-Fun} rule will
-\mzu{substitute \code{n} with the argument}
-\mz{\code{n} seems unbound here.}
-at a call-site.
-\ignore{
-The semantics manages variable scopes using the special $\erettext$
-form. \textsc{S-Let} evaluates to a configuration whose expression is
-$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
-and remember $x$ and its new value $\evalue{n}{\tau}$
-in $e$'s scope that is defined by the $\erettext$ operation.
-Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
-we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope.
-After one-step evaluation is completed, 
-we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
-and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
-This procedure continues until $e'$ becomes a literal
-$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
-the literal. 
-}
+depends on \code{m_1}.
+Its function pointer type is 
+$\tptr{\tfun{\code{m_1}}{\tint,\tntarray{0}{\code{m_1}}{\texttt{char}}}{\tint}}{\tmode}$.
 
@@
 These proofs have been conducted in our Coq model.
-Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{appx:add-type-sem}.
+Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{sec:meta}.
 Progress states that a \lang program can always make a move:

### 2023-01-15 sandp23/background.tex
Adding all

@@
 \myparagraph{Converting C to \checkedc}
-%
-The safety guarantees of Checked C come with certain restrictions. For instance,
-as shown below, Checked C programs cannot use address-taken variables in a
-bounds expression as the bounds relations may not hold because of possible
-modifications through pointers.
-% 
-\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-...
-_Array_ptr<int> p : count (n) = NULL;
-|\textcolor{red}{\faTimes}|..,&n,.
-\end{minted}
-% 
-Consequently, converting existing C programs to Checked C might require
-refactoring,~\eg eliminate~\inlinecode{&n} from the program above without
-changing its functionality.
-% 
-This might require considerable effort~\cite{duanrefactoring} depending on the
-program's complexity.
-% 
-Recently, Machiry~\etal developed~\threec~\cite{machiry2022c} that tries to
-automatically convert a program to Checked C by adding appropriate pointer
-annotations.
-However, as described in \threec, completely automated conversion
-is \emph{infeasible}, and it requires the developer to convert some code regions
-manually.
+It is not possible to fully automate the conversion of C code to~\checkedc due to the requirement for semantic reasoning and other modifications such as refactoring.
+We provide more details on this in~\apdx{app:convertctocc}.
+
+%%
+%The safety guarantees of Checked C come with certain restrictions. For instance,
+%as shown below, Checked C programs cannot use address-taken variables in a
+%bounds expression as the bounds relations may not hold because of possible
+%modifications through pointers.
+%% 
+%\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+%...
+%_Array_ptr<int> p : count (n) = NULL;
+%|\textcolor{red}{\faTimes}|..,&n,.
+%\end{minted}
+%% 
+%Consequently, converting existing C programs to Checked C might require
+%refactoring,~\eg eliminate~\inlinecode{&n} from the program above without
+%changing its functionality.
+%% 
+%This might require considerable effort~\cite{duanrefactoring} depending on the
+%program's complexity.
+%% 
+%Recently, Machiry~\etal developed~\threec~\cite{machiry2022c} that tries to
+%automatically convert a program to Checked C by adding appropriate pointer
+%annotations.
+%However, as described in \threec, completely automated conversion
+%is \emph{infeasible}, and it requires the developer to convert some code regions
+%manually.
 
@@
 The spatial safety guarantee~\checkedc provides applies to completely converted programs,~\ie programs that uses only checked types.
-Although,\lc{why use "Although" here?
-seems like this is not an opposite/new idea} the backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
-The unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+The backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
+But, these unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
 \lc{I'm confused, does the partially checked program "enjoy spatial memory safety" or not?}

### 2023-01-15 sandp23/evaluation.tex
Adding all

@@
 % \liyi{Deena? Leo? }
-Our evaluation of \systemname consists of a set of tests that can be classified into Micro-benchmarks and Program Benchmarks that evaluate \systemname on WebAssembly sandbox.
+We evaluate~\systemname in terms of the following three aspects:
+\begin{itemize}
+\item\textbf{Performance Overhead:} What is the performance overhead (both runtime and memory) in using~\systemname?
+\item\textbf{Conversion Effort:} How much developer effort is needed to annotate parts of programs to run in~\systemname?
+\item\textbf{Security Impact:} How effective is the isolation provided by~\systemname{} in preventing security vulnerabilities?
+\end{itemize}
+
+%Our evaluation of \systemname consists of a set of tests that can be classified into Micro-benchmarks and Program Benchmarks that evaluate \systemname on WebAssembly sandbox.
 Consequently, all further references to sandbox refer to unchecked code in the WebAssembly Sandbox.
-MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
-We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.
+%MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
+%We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.
+
+\subsection{Dataset}
+\input{tables/datasettable}
+Network-facing programs such as servers directly interact with external input, often process complex data types, and are more susceptible to security issues.
+We primarily focus on network servers as they can benefit most from our partitioning approach.
+We use WebAssembly (WASM) as our target sandbox. Consequently, we also want any of the selected programs to be compilable with the WASM sandbox.
+We selected network servers that we could (with minimal effort) compile with the WASM compiler.
+We also selected a few standalone programs suggested by the ~\checkedc team, which are good candidates to evaluate modifications to~\checkedc.
+The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
+\aravind{Arun, please fix the dataset table with details of all programs.}
+
+\subsection{Experimental Setup}
 

### 2023-01-15 sandp23/implementation.tex
Adding all

@@
 However, we use a small OCAML program (\numocamllines{}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
+\aravind{Arun, please fix the number of lines you changed for clang frontend and ocaml program}
 
@@
 \subsection{Compiler}
-The~\systemname compiler includes the type checker (\sect{}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
-We implemented the compiler on top of the existing~\checkedc compiler.
+The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
+We implemented the compiler (\numcompilerlines{}) on top of the existing~\checkedc compiler.
+\aravind{Arun, please fix the number of changes you made for the compiler}
 
@@
 
+The~\ucregion source files will be compiled with sandbox compiler,~\ie NAMEOFSANDBOXCOMPILER (CITATION).
 Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
 \cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers. 
+\aravind{Arun, please fill up the name of the sandbox compiler and add citation.}
 

### 2023-01-15 sandp23/overview.tex
update.
Adding all

@@
 The developer starts using~\systemname{} with the resulting partially annotated program with checked and tainted types~\ie~\lst{lst:initialconv} and~\ref{lst:humantaint} atop of~\lst{lst:comb}.
-The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (CITE).
+The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (\sect{sec:typechecking}).
 The developer uses our type checker to identify the additional annotations (\usermods) required to have a well-typed program and adds them as shown in~\lst{lst:humanadjust}.
@@
 
-\liyi{This paper should just introduce the language and the compiler, right?
-I do not think the following is necessary.
-Instead, we can have an overview section introduce the compiler with examples; especially highlight the callback.
-}
-
 %newusecase.tex

### 2023-01-15 sandp23/tables/datasettable.tex
Adding all

{
\footnotesize
\begin{table}[]
\begin{tabular}{c|c|c|r}
\toprule
\textbf{ID} & \textbf{Program} & \textbf{Description}          & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Size\\ (SLoc)\end{tabular}}} \\ 
\midrule
\rowcolor{black!15} 1           & ProFTPD          & High performance FTP Server   & XXX K                                                                               \\ 
2           & MicroHTTPD       & Simple HTTPD Server           & YY K                                                                                \\ 
\rowcolor{black!15} 3           & UFTPD            & UDP based FTP Server          & ZZZ K                                                                               \\ 
4           & png2pnm          & Program to convert png to pnm & UUU K                                                                               \\ 
\bottomrule
\end{tabular}
\caption{Evaluation Dataset.}
\label{table:dataset}
\end{table}
}

### 2023-01-15 sandp23/examples/conv1.c

@@ -15,18 +15,19 @@ void server_loop(int sock_fd) {
 }
 
 int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_\_tainted}|;
+ char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_Tainted}|;
  int rc = -1;
  size_t r_len;
  r_len = read_msg(sock_fd, buff, 
                   MAX_MSG_SIZE);
- if (r_len > 0) {
+ if (r_len > 0 && 
+     r_len < MAX_MSG_SIZE) {
   switch(buff[0]) {
    case REQ1: 
-    rc = process_req1(buff); 
+    rc = process_req1(buff, r_len); 
     break;
    case REQ2: 
-    rc = process_req2(buff); 
+    rc = process_req2(buff, r_len); 
     break;
    ...
   }

### 2023-01-15 sandp23/examples/conv2.c

@@ -1,5 +1,5 @@
 size_t read_msg(int sock_fd, char *msg, 
-                size_t sz) __tainted {
+                size_t sz) _Tainted {
  size_t nr;
  nr = read(sock_fd, (void*)msg, sz);
  ...
@@ -8,7 +8,7 @@ size_t read_msg(int sock_fd, char *msg,
 }
 
 int process_req1(char *msg, size_t m_l) 
-                 __tainted {
+                 _Tainted {
  int rc = -1, i;
  if (m_l > MIN_SIZE) {
    sscanf(msg, "%d", &i);

### 2023-01-15 sandp23/examples/humanannotations.c

@@ -1,8 +1,8 @@
 size_t read_msg(int sock_fd, char *msg, 
-                size_t sz) |\textcolor{taintcolor}{\_\_tainted}| |\useradded| {
+                size_t sz) |\textcolor{taintcolor}{\_Tainted}| |\useradded| {
  ...
 }
 
-int process_req1(char *msg, size_t m_l) |\textcolor{taintcolor}{\_\_tainted}| |\useradded| {
+int process_req1(char *msg, size_t m_l) |\textcolor{taintcolor}{\_Tainted}| |\useradded| {
  ...
 }

### 2023-01-15 sandp23/examples/orig1.c

@@ -19,13 +19,14 @@ int handle_request(int sock_fd) {
  size_t r_len;
  r_len = read_msg(sock_fd, buff, 
                   MAX_MSG_SIZE);
- if (r_len > 0) {
+ if (r_len > 0 && 
+     r_len < MAX_MSG_SIZE) {
   switch(buff[0]) {
    case REQ1: 
-    rc = process_req1(buff); 
+    rc = process_req1(buff, r_len); 
     break;
    case REQ2: 
-    rc = process_req2(buff); 
+    rc = process_req2(buff, r_len); 
     break;
    ...
   }

### 2023-01-15 sandp23/evaluation.tex
update.

@@
 However, the scope of \systemname changes made on each of the programs is varied and the converted programs themselves are relatively tiny.
-Since each of the program is evaluated on a pre-included test-suite that runs on a single instance of the program, sandbox creation cost is a one-time constant due to \systemname implementing RLBOX's \cite{rlbox-paper} WebAssembly sandbox creation API.
+Since each of the program is evaluated on a pre-defined test-suite that runs on a single instance of the program, sandbox creation cost is a one-time constant due to \systemname implementing RLBOX's \cite{rlbox-paper} WebAssembly sandbox API.
+\liyi{I do not understand the following sentence, what is test-bench? }
 Run-time performance is evaluated with C's~\code{<time.h>} library by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs) and measuring latency delta as a percentage.
+\liyi{need to rewrite}
 Timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
+\liyi{What is Valgrind? }
 Memory overhead is measured with Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
@@
 \end{figure}
+
+\liyi{What is SBX, and how it is related to indirect-calls, pointer accesses and what is micro-benchimarks doing? Should we elaborate a little? }
 
@@
 
+\liyi{results?}
 \subsubsection{\textbf{ProFTPD}}

### 2023-01-15 sandp23/formal.tex
update.

@@
 named~\lang.
-We precisely present its syntax, semantics, type
-system, and compilation.
-% 
-From those, we develop~\lang’s meta-
-theories, including the type soundness, non-exposure, and non-crashing theorems.
+We present its syntax, semantics, type system,
+as well as \lang’s meta-theories, including the type soundness, non-exposure, and non-crashing theorems.
 
@@
 At a high level, we classify types as word-size value, multi-word
-value, or function types. A word-size value can be either integer or pointer.
+value, or function types. A word-size value can be either an integer or pointer.
 Every pointer type ($\tptr{\omega}{\xi}$) includes a
@@
   array of checked pointers.
-  Syntactically, this type is however \emph{illegal} in~\lang{} and
-  is rejected by the well-formedness. 
-However, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
-%
-
+  This is \emph{illegal} in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region;
+  thus, it is rejected by the well-formedness. 
+  Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
 %
@@
 For example, a dynamic bounds cast
-  {\footnotesize
+  { \footnotesize
     \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
-becomes 
-{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}. 
+is formalized as 
+{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \lang. 
 %
@@
 
-The \lang system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
+The \lang type system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
 Our type checker restricts the usage of tainted and checked pointer types to ensure that tainted pointers do not affect checked types, along with enforcing~\checkedc typing rules~\cite{li22checkedc}.
@@
 As partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{T-}$X$),
-Each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
 which states that in a type environment $\Gamma$ (mapping variables to
@@
 variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-in context mode $m$.
+in context mode $m$, indicating that the code is in $m$ region.
 The operational semantics for \lang is defined as a small-step
@@
 The $\mode$ function in Fig.~\ref{fig:c-context}
-determines the context mode that the expression $e'$ locates based on the context $E$.
-In \Cref{lst:humantaint}, the function call \code{handle_request} has $\umode$ context mode (is in $\umode$ region) since it is inside an unchecked function \code{server_loop}.
+determines the context mode, i.e., region, that the expression $e'$ locates based on the context $E$.
+In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.
 The second rule describes the exception handling 
-for possible crashing behaviors in unchecked regions.
+for possible crashing behaviors in $\umode$ regions.
 Operations in $\umode$ region can non-deterministically crash
@@
 In \lang, Context modes $m$ appearing in a type rule determine the code region 
-that permits pointer dereferences and value-assignments, depending on the pointer modes.
-We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ since $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
+that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
+We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
 This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can perform such operations in $\cmode$ and $\umode$ regions, respectively.
@@
 an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
-if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2} $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
+if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2}, i.e., $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
 In \lang, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
@@
 \textsc{S-DefT}, \textsc{S-DefNull} in \Cref{fig:type-system-1})
-reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in evaluation stage.
+reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
 The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
-With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a \enull pointer is used,
+With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a $\enull$ pointer is used,
 \lang captures the runtime error.
@@
 
-Rules \textsc{S-DefC} and \textsc{S-DefT} are for $\cmode$ and $\tmode$ mode pointer dereference, respectively.
-In addition to the no \enull check in $\cmode$ mode pointer dereference,
-any dynamic heap access of a tainted pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
+Rules \textsc{S-DefC} and \textsc{S-DefT} are for $\cmode$ and $\tmode$ mode pointer dereferences, respectively.
+In addition to the no $\enull$ check in $\cmode$ mode pointer dereference,
+any dynamic heap access of a tainted pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to that the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
 
@@
 The execution of a $\echeckedtext$ or $\euncheckedtext$ block represents 
-the context switching from a checked to an checked region, or vice versa,
+the context switching from a $\cmode$ to an $\umode$ region, or vice versa,
 with its type and semantic rules given in \Cref{fig:type-system-1}.
@@
 as well as the check that all free variables in the block content $e$ are in $\overline{x}$.
-For example, \code{read_msg} in the \code{handle_request} function is tainted in \Cref{lst:humantaint},
-if any argument for \code{handle_request} is a checked pointer, 
-it means that we are exposing a checked pointer address to unsafe regions, 
-which means that users might randomly manipulate the data region pointed to by the checked pointer
-and \lang forbids such behavior.
+For example, \code{StringAuth} in \Cref{subsub:gencregion} is a trampoline function that disallows
+checked pointers as arguments and return values.
+The use of the function in the following \code{_T_StringAuth}, which is in $\umode$ region,
+cannot legally acknowledges any checked pointers; otherwise, we might expose a checked pointer address to unsafe code regions.
+In \systemname, we actually permits the accesses of checked pointers inside \code{StringAuth},
+since the function body of a trampoline function is in $\cmode$ region.
+More information is given in \Cref{subsub:gencregion}.
 
@@
 Consider the \code{process_req2} function in
-Fig.~\ref{lst:final}; its parameter type for \code{msg} 
+Fig.~\ref{lst:final}, whose parameter type for \code{msg} 
 depends on \code{m_1}.
@@
 $\tptr{\tfun{\code{m_1}}{\tint,\tntarray{0}{\code{m_1}}{\texttt{char}}}{\tint}}{\tmode}$.
-
-\textsc{S-FunC} and \textsc{S-FunT} are
+In \code{handle_request}, the call \code{process_req2(buff, r_len)} binds variable \code{m_1} to \code{r_len}.
+After the call returns, \code{m_1}'s scope is ended, so we need to substitute it with \code{r_len} in the final return type
+because it might contain \code{m_1}.
+
+\textsc{S-FunC} and \textsc{S-FunT} define the semantics
 for $\cmode$ and $\tmode$ mode function pointers, respectively. 
@@
 maps addresses (integer literals) to the function data described above.
-Rule \textsc{S-FunT} defines the tainted version of function call semantics.
-In such case, the verification process 
+Rule \textsc{S-FunT} defines the tainted version of function call
+with the verification process 
 $\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}$
@@
 
-In this subsection, we focus on our main meta-theoretic results about
+Here, we discuss our main meta-theoretic results for
 \lang: type soundness (progress and preservation),
@@
 Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{sec:meta}.
-Progress states that a \lang program can always make a move:
+The progress theorem below states that a \lang program can always make a move.
 
@@
 \noindent
-There are two forms of preservation regarding the checked and unchecked regions.
+There are two forms of preservation regarding the $\cmode$ and $\umode$ regions.
 Checked Preservation states that a reduction step preserves both the
-type and consistency of the program being reduced.
-Unchecked Preservation states that any evaluation happens at unchecked region does not affect the checked heap.
+type and consistency of the program being reduced, while
+unchecked Preservation states that any evaluation happens at $\umode$ region does not affect the $\cmode$ mode heap.
 
@@
 \longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
-checked region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$ and $\tau'$ that are well formed, checked region consistent
+$\cmode$ region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
+$\Gamma'$ and $\tau'$ that are well formed, $\cmode$ region consistent
 ($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
@@
 
-$\heap(\cmode)$ refers to all $\cmode$ mode pointers in heap $\heap$.
 Using the above theorem, we first show the non-exposure theorem,
-where code in unchecked region cannot observe a valid checked pointer address.
+where code in $\umode$ region cannot observe a valid checked pointer address.
 

### 2023-01-16 sandp23/evaluation.tex
Arun- Implementation Section

--- 2023-01-14 sandp23/evaluation.tex
@@
   \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
-  \item Marshalling data might sometimes be cost-effective instead of propagating the taintedness of a pointer
+  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
   \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
-  \item Conversion style and extent qualitatively dictate the converted program's conversion efforts and is modest at a couple of days. 
+  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
+  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
 \end{itemize}

### 2023-01-16 sandp23/implementation.tex
Arun- Implementation Section

--- 2023-01-15 sandp23/implementation.tex
@@
 Developers should use this qualifier to mark callback functions,~\ie functions in~\cregion, that can be called from the tainted region. 
-The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\sect{}).
+The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\ref{subsub:gencregion}).
 
 \noindent\emph{\_Mirror}: 
-This qualifier copies the corresponding function both into~\cregion and~\ucregion code. Such that callers from each region will invoke the corresponding copy.
-This helps handle certain simple utility functions that might be called from both~\cregion and~\ucregion code.
-\aravind{Give an example from our evaluation program. (E.g., function foo in program1)}.
+This qualifier makes ~\sourcerewriter copy the corresponding function into both ~\cregion and~\ucregion code such that callers from each region can invoke the corresponding copy.
+This helps handle certain simple utility functions that might be called from both ~\cregion and~\ucregion code.
+For example, ~\code{json_value_get_type} in our evaluation of parsons\_wasm has callers from both the ~\cregion and ~\ucregion.
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+/* JSON Value API */
+_Mirror JSON_Value_Type json_value_get_type  
+      (_TPtr<const TJSON_Value> value) {
+    return value ? value->type : JSONError;
+}
+
+_Mirror _TPtr<TJSON_Object> json_value_get_object
+(_TPtr<const TJSON_Value> value) {
+    return json_value_get_type(value) == JSONObject ?
+    value->value.object : NULL;
+}
+
+\end{minted}
 
@@
 
-Developers can use the~\code{_Mirror} qualifier, which copies the function into both the region so that callers in each region will invoke the corresponding copy avoiding unnecessary tainting.
+Developers can use the ~\code{_Mirror} qualifier, which makes ~\sourcerewriter copy the function into both regions so that callers in each region will invoke the corresponding copy, thereby, avoiding unnecessary tainting.
+
+However, ~\code{_Mirror} qualifier semantically asserts the function to only have callees that are either qualified with ~\code{_Tainted} or ~\code {_Mirror}.
+For example, ~\code{json_value_get_object}'s only callee
+\newline
+~\code{json_value_get_type} is ~\code{\_Mirror} qualified.
+Similarly, \systemname enforces ~\code{_Mirror} qualified functions to only access global checked data that are marked constant.
+These control-flow and data-flow compile-time semantics are enforced by \systemname to assert region locality and exclusivity amongst qualified function scopes in both the regions.
 
 \noindent\emph{\_TLIB}:
-This qualifier relaxes type-checking rules on library functions avoiding unnecessary copying of tainted data.
+This qualifier relaxes type-checking rules on library functions, thereby, avoiding unnecessary copying of tainted data.
 Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.
@@
 We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
-We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then invoke~\code{strncat(dst, tmp, n);}.
+We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then call~\code{strncpy} as~\code{strncat(dst, tmp, n);}.
 Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
 If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use~\code{strncat} with tainted source pointer in~\cregion.
-The~\code{_TLIB} qualifier enables this as shown below:
+Consequently, ~\code{\_TLIB} qualified functions originally implemented in ~\cregion are expected to be non memory-modifying and thereby, its safe use is left at developer's discretion as \systemname does not enforce any semantics for this.
+For example, malloc cannot be qualified with ~\code{_TLIB} and thereby is implemented as an indirect call to sandbox-specific allocator.
+
+The~\code{_TLIB} qualifier enables the qualified function to be used as shown below:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-// Put your example annotation here.
-// First, manually check the memory is in tainted region.
-// if yes, then call strncpy.
-if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
-  handle_violation();
-// our type checker ignores this because 
-// the above _TLIB annotation.
+/* Append SRC onto DEST.  */
+_TLIB extern char *strcat (char *__restrict __dest,
+            const char *__restrict __src)
+...        
 strncat(dst, t_str, n);
 \end{minted}
-\aravind{Arun, Fix the above code. i.e., Add \_TLIB annotation to strncpy }
 The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.
-Note that developer needs to manually add the check before calling~\code{strcat}.
 
@@
 
-\iffalse
 \noindent\emph{Handling Tainted Buffers in~\cregion}:
-All buffers that are marked as tainted in~\cregion should be allocated in the sandbox. For instance,~\code{buff} in function~\code{handle_request} of~\lst{lst:final}.
+All buffers that are marked as tainted in~\cregion should be manually allocated in the sandbox. For instance,~\code{buff} in function~\code{handle_request} of~\lst{lst:final}.
 We perform source rewriting and make the buffer into a pointer variable and allocate the corresponding number of bytes using the sandbox-specific allocator.
@@
 - char buff[MAX_MSG_SIZE] __Tainted;
-+ char *buff = ....
++ _TPtr<char> buff = (_TPtr<char>)t_malloc(MAX_MSG_SIZE);  
 \end{minted}
-\aravind{Arun, please fix the above example on how exactly we replace it with WASM allocator.}
 Note that we also add the necessary deallocator calls at function exit points.
-\fi
 
-\noindent\emph{Handling Calls to Tainted Functions}:
-In~\cregion, we also need to modify calls to tainted functions as they execute inside the sandbox (separate address space) and thus cannot be invoked as regular functions.
-However, modifying every call site of tainted functions is tedious and also requires precise pointer analysis~\cite{milanova2002precise} to handle indirect calls through function pointers.
+\noindent\emph{Handling Tainted Functions}:
+In~\cregion, we also need to modify tainted functions as they need to execute inside the sandbox (separate address space).
+Consequently, We rewrite the tainted function's body with its corresponding sandbox invocation, thereby, routing all ~\cregion calls to tainted functions to their respective sandboxed implementations in ~\ucregion.
+For instance, we modify the function body of ~\code{process_req1} in~\lst{lst:final} as below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
+int process_req1(_T_Array_ptr<char> msg :
+count(m_l), size_t m_l)
+__tainted {
+- int rc = -1, i;
+- // complex code
+- return rc;
++ return w2c_process_req1(msg, m_l);
+}
 
-We handle this by~\emph{indirection}: Instead of modifying the call sites, we modify the body of tainted functions to invoke the corresponding function in the sandbox.
-For instance, we modify the body of tainted function~\code{process_req1} (\lst{lst:final}) in~\cregion as below:
-\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
-int process_req1(char *msg, size_t m_l) _Tainted {
-- int rc = -1, i;
-- if (m_l > MIN_SIZE) {
-- ...
-+ HOW DO WE HANDLE SANDBOX CALL.
-}
 \end{minted}
-This ensures that all calls (even through function pointers) to the tainted function~\code{process_req1} are redirected to the sandbox.
-\aravind{Arun, add the call on how to call sandbox function.}
 
@@
 As mentioned in~\sect{subsubsec:addfuncqual}, functions with these qualifiers can be called from~\ucregion.
-Consider the following~\code{StringAuth} function that checks whether the provided user is authenticated and needs to be invoked from~\ucregion and hence a callback.
+Consider the following~\code{StringAuth} function that checks whether the provided user input ~\code{usertoken} is authenticated by accessing checked data.
+Since this needs to be invoked from ~\ucregion it is annotated as a ~\code{\_Callback}.
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
-+ unsigned int _T_StringAuth(unsigned sandbox,
++ unsigned int _T_StringAuth(void* sandbox,
 +               unsigned int arg_1,
 +               unsigned long int arg_2) {
-+   // Perform necessary conversion of arguments.
++   // Perform necessary Type-conversion of arguments.
 +   // uname <- conver arg_1
@@
 \end{minted}
+
 The trampoline function handles the invocations from sandbox (and hence the extra parameter~\code{sandbox}), performs the necessary conversion of pointer arguments, and eventually invokes the callback.
@@
 
-Lastly, we change all instances where the callback function pointer is given to tainted functions to use the handle for the corresponding trampoline a shown below:
-\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
-// How do we change the call-sites where the function pointer is passed
-// as the call-back.
+Lastly, we change the tainted function's body to include an indirect call to the sandbox's implementation of the tainted function.
+However, instead of passing the callback function pointer directly from the argument list, we pass the generated trampoline handle ~\code{__StringAuth__C_} as shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+_Tainted _TPtr<char> StringProc(_TPtr<_TPtr<const char>> user_input,
+_TPtr<_TPtr<char>(_TPtr<const char> input, size_t len)>StringAuth) {
+-  ...
+-  //complex Function Body       
+- return StringAuth(one_past_start, string_len);
++ return w2c_StringProc(_SBX_(), (unsigned int)string, __StringAuth__C_);
+
+}
 \end{minted}
-\aravind{Arun, please check above code example to see if they make sense.}
 
@@
 \label{subsubsec:genucregion}
-We create~\ucregion partition by first copying all functions marked as~\code{_Tained} and~\code{_Mirror}.
+We create~\ucregion partition by first copying all functions and variables marked as~\code{_Tained} and~\code{_Mirror}.
 Next, we convert the partition into regular C by replacing all~\systemname{} tainted types with the corresponding C types.
+\newline
 For instance:~\code{_T_Array_ptr<char>} will be replaced with~\code{char *}.
-The~\code{read_msg} function signature in~\lst{lst:final} will be modified as shown below:
+The~\code{read_msg} function signature in \lst{lst:final} will be modified as shown below in the ~\ucregion:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
-size_t read_msg(int sock_fd, char *msg,
--                             size_t sz) _Tainted {
+- size_t read_msg(int sock_fd, _T_Array_ptr<char> msg,
+-                             size_t sz) __Tainted {
++ size_t read_msg(int sock_fd, char* msg,
 +                             size_t sz) {
@@
 
-Finally, we convert all calls to~\cregion callbacks with sandbox-specific invocations.
+% ** We Do NOT modify callback function in checked sources **
+% Finally, we convert all calls to~\cregion callbacks with sandbox-specific invocations.
 For instance, for the~\code{StringAuth} call back (\sect{subsub:gencregion}), we modify~\code{...UNCHECKED FUNC..} as shown below:
-\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
-SAY HOW WE MODIFY UNCHECKED CODE
-TO INVOKE CALLBACK FUNCTIONS.
-\end{minted}
-\aravind{Arun, please fix the above example}
-
-\subsection{Compiler}
-The~\systemname compiler includes the type checker (\sect{}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
-We implemented the compiler on top of the existing~\checkedc compiler.
-
-Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
-For instance, for the code in~\lst{lst:final}, our type checker emits an error saying that ``~\emph{buffer~\code{buff} is tainted, but it is allocated in~\cregion}''.
-Consequently, the developer has to make the following changes to allocate~\code{buff} in the sandbox:
-\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
-- char buff[MAX_MSG_SIZE] _Tainted;
-+ _t_Array_ptr<char> buff: count(MAX_MSG_SIZE)
-+    = sbx_allocate(MAX_MSG_SIZE);
-...
-+ sbx_free(buff)
-return rc;
-\end{minted}
-
-For all tainted pointer access, our instrumentation adds a check (executed at runtime) to see we check if the address is within the valid range for the sandbox. An example of the check is shown below:
-\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
-_TPtr<int> t = ...
-+ if (!is_in_range(t, SBX_LOW(), SBX_HIGH()))
-+  handle_violation();
-*t = 1;
-\end{minted}
-We show the added check in C for clarify. However, our instrumentation works on LLVM IR.
-
-Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
-\cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers.
-
-\subsection{Supporting General Partioning Use Case}
-\input{figures/newusecase}
-
-Tainting all functions that are not checked may lead to a large number of transitions to the sandbox, which could add unnecessary overhead.
-In our example, in~\lst{lst:final}, marking~\code{read_msg} function as tainted would cause the system to transition to and from the sandbox every time a message is read, which could be inefficient.
-
-However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain regular (unchecked) pointers as well.
-%
-This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them.
-In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
-
-For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
-
-The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
-It is best to annotate all functions with checked types for security purposes.
-However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted.
-This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
-We add the additional constraint of not allowing casting from taint types to generic types to enable this generic use case without violating the safety guarantees of~\systemname.
-Specifically, we disallow the following casting in~\cregion:
-\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
-//_t_Ptr<int> y; int *z;
-|\faRemove| z = (int *)y; // Not allowed.
-\end{minted}
-\iffalse
+% \begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
+% SAY HOW WE MODIFY UNCHECKED CODE
+% TO INVOKE CALLBACK FUNCTIONS.
+% \end{minted}
+% \aravind{Arun, please fix the above example}
 
@@
 Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout.
-\fi

### 2023-01-16 sandp23/examples/MirrorExample.c

@@ -1,0 +1,5 @@
+/* JSON Value API */
+_Mirror JSON_Value_Type json_value_get_type  
+      (_TPtr<const TJSON_Value> value) {
+    return value ? value->type : JSONError;
+}

### 2023-01-16 sandp23/examples/conv1.c

@@ -23,10 +23,10 @@ int handle_request(int sock_fd) {
  if (r_len > 0) {
   switch(buff[0]) {
    case REQ1: 
-    rc = process_req1(buff); 
+    rc = process_req1(buff, r_len); 
     break;
    case REQ2: 
-    rc = process_req2(buff); 
+    rc = process_req2(buff, r_len); 
     break;
    ...
   }

### 2023-01-16 sandp23/examples/conv2.c

@@ -7,7 +7,8 @@ size_t read_msg(int sock_fd, char *msg,
  return nr;
 }
 
-int process_req1(char *msg, size_t m_l) 
+int process_req1(_T_Array_ptr<char> msg :
+                 count(m_l), size_t m_l) 
                  __tainted {
  int rc = -1, i;
  if (m_l > MIN_SIZE) {
@@ -18,7 +19,7 @@ int process_req1(char *msg, size_t m_l)
  return rc;
 }
 
-int process_req2(|\textcolor{taintcolor}{\_t\_Array\_ptr}|<char> msg : 
+int process_req2(|\textcolor{taintcolor}{\_T\_Array\_ptr}|<char> msg : 
                  count(m_l), 
                  size_t m_l) {
  size_t i = 0, j = 0;

### 2023-01-16 sandp23/examples/orig1.c

@@ -22,10 +22,10 @@ int handle_request(int sock_fd) {
  if (r_len > 0) {
   switch(buff[0]) {
    case REQ1: 
-    rc = process_req1(buff); 
+    rc = process_req1(buff, r_len); 
     break;
    case REQ2: 
-    rc = process_req2(buff); 
+    rc = process_req2(buff, r_len); 
     break;
    ...
   }

### 2023-01-16 sandp23/appendix.tex
finished appx.

@@
 \section{Appendix}\label{app:main}
+
+\subsection{Converting C to \checkedc}
+\label{app:convertctocc}
+%
+The safety guarantees of Checked C come with certain restrictions. For instance,
+as shown below, Checked C programs cannot use address-taken variables in a
+bounds expression as the bounds relations may not hold because of possible
+modifications through pointers.
+% 
+\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+...
+_Array_ptr<int> p : count (n) = NULL;
+|\textcolor{red}{\faTimes}|..,&n,.
+\end{minted}
+% 
+Consequently, converting existing C programs to Checked C might require
+refactoring,~\eg eliminate~\inlinecode{&n} from the program above without
+changing its functionality.
+% 
+This might require considerable effort~\cite{duanrefactoring} depending on the
+program's complexity.
+% 
+Recently, Machiry~\etal developed~\threec~\cite{machiry2022c} that tries to
+automatically convert a program to Checked C by adding appropriate pointer
+annotations.
+However, as described in \threec, completely automated conversion
+is \emph{infeasible}, and it requires the developer to convert some code regions
+manually.
 
@@
 Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
-\Cref{fig:wftypesandbounds} defines the well-formedness for bounds appearing in a type.
+\Cref{fig:wftypesandbounds} defines the well-formedness for bounds appearing in a type, which requires that any bound must be $\tint$ type and has instance in $\Gamma$.
 All bound variables appearing in a type must be either an instance of the type environment $\Gamma$,
@@
     \tau =_{\Theta} \tau'&\Rightarrow&\tau &\sqsubseteq_{\Theta}& \tau'\\[0.2em]
+
+    && \tptr{\tau}{\tmode}&\sqsubseteq_{\Theta}& \tptr{\tau}{\umode}\\[0.2em]
 

### 2023-01-16 sandp23/formal.tex

@@
 and write $\overline{\tau}[\overline{e'} / \overline{x}]$ to lift the
-substitution to every type \(\overline{\tau}\).
+substitution to every type in \(\overline{\tau}\).
 % 

### 2023-01-16 sandp23/implementation.tex

@@
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-/* Append SRC onto DEST.  */
+
 _TLIB extern char *strcat (char *__restrict __dest,
-            const char *__restrict __src)
+            const char *__restrict __src); // In the header file
 ...
+// First, manually check the memory is in tainted region.
+// if yes, then call strncpy.
+if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
+  handle_violation();
+// our type checker ignores this because 
+// the above _TLIB annotation.
 strncat(dst, t_str, n);
@@
 Note that we also add the necessary deallocator calls at function exit points.
+\fi
 
-\noindent\emph{Handling Tainted Functions}:
-In~\cregion, we also need to modify tainted functions as they need to execute inside the sandbox (separate address space).
-Consequently, We rewrite the tainted function's body with its corresponding sandbox invocation, thereby, routing all ~\cregion calls to tainted functions to their respective sandboxed implementations in ~\ucregion.
-For instance, we modify the function body of ~\code{process_req1} in~\lst{lst:final} as below:
+\noindent\emph{Handling Calls to Tainted Functions}:
+In~\cregion, we also need to modify calls to tainted functions as they execute inside the sandbox (separate address space) and thus cannot be invoked as regular functions.
+However, modifying every call site of tainted functions is tedious and also requires precise pointer analysis~\cite{milanova2002precise} to handle indirect calls through function pointers.
+
+We handle this by~\emph{indirection}: Instead of modifying the call sites, we modify the body of tainted functions to invoke the corresponding function in the sandbox.
+For instance, we modify the body of tainted function~\code{process_req1} (\lst{lst:final}) in~\cregion as below:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
-int process_req1(_T_Array_ptr<char> msg :
-count(m_l), size_t m_l)
-__tainted {
+int process_req1(char *msg, size_t m_l) _Tainted {
 - int rc = -1, i;
-- // complex code
-- return rc;
+- if (m_l > MIN_SIZE) {
+- ...
 + return w2c_process_req1(msg, m_l);
 }
-
 \end{minted}
+This ensures that all calls (even through function pointers) to the tainted function~\code{process_req1} are redirected to the sandbox.
 
@@
 For instance:~\code{_T_Array_ptr<char>} will be replaced with~\code{char *}.
-The~\code{read_msg} function signature in \lst{lst:final} will be modified as shown below in the ~\ucregion:
+
+The~\code{read_msg} function signature in \lst{lst:final} will be modified in the ~\ucregion as shown below:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
@@
 % \aravind{Arun, please fix the above example}
+
+\subsection{Compiler}
+The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
+We implemented the compiler (\numcompilerlines{}) on top of the existing~\checkedc compiler.
+\aravind{Arun, please fix the number of changes you made for the compiler}
+
+Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
+For instance, for the code in~\lst{lst:final}, our type checker emits an error saying that ``~\emph{buffer~\code{buff} is tainted, but it is allocated in~\cregion}''.
+Consequently, the developer has to make the following changes to allocate~\code{buff} in the sandbox:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+- char buff[MAX_MSG_SIZE] _Tainted;
++ _t_Array_ptr<char> buff: count(MAX_MSG_SIZE)
++    = sbx_allocate(MAX_MSG_SIZE);
+...
++ sbx_free(buff)
+return rc;
+\end{minted}
+
+For all tainted pointer access, our instrumentation adds a check (executed at runtime) to see we check if the address is within the valid range for the sandbox. An example of the check is shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+_TPtr<int> t = ...
++ if (!is_in_range(t, SBX_LOW(), SBX_HIGH()))
++  handle_violation();
+*t = 1;
+\end{minted}
+We show the added check in C for clarify. However, our instrumentation works on LLVM IR.
+
+The~\ucregion source files will be compiled with sandbox compiler,~\ie NAMEOFSANDBOXCOMPILER (CITATION).
+Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
+\cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers. 
+\aravind{Arun, please fill up the name of the sandbox compiler and add citation.}
+
+\subsection{Supporting General Partioning Use Case}
+\input{figures/newusecase}
+
+Tainting all functions that are not checked may lead to a large number of transitions to the sandbox, which could add unnecessary overhead.
+In our example, in~\lst{lst:final}, marking~\code{read_msg} function as tainted would cause the system to transition to and from the sandbox every time a message is read, which could be inefficient.
+
+However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain regular (unchecked) pointers as well.
+%
+This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them.
+In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
+
+For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
+
+The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
+It is best to annotate all functions with checked types for security purposes.
+However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted.
+This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
+We add the additional constraint of not allowing casting from taint types to generic types to enable this generic use case without violating the safety guarantees of~\systemname.
+Specifically, we disallow the following casting in~\cregion:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
+//_t_Ptr<int> y; int *z;
+|\faRemove| z = (int *)y; // Not allowed.
+\end{minted}
+\iffalse
 

### 2023-01-16 sandp23/evaluation.tex
Fixing eval

--- 2023-01-15 sandp23/evaluation.tex
@@
 We selected network servers that we could (with minimal effort) compile with the WASM compiler.
-We also selected a few standalone programs suggested by the ~\checkedc team, which are good candidates to evaluate modifications to~\checkedc.
+We also selected a few standalone programs suggested by the ~\checkedc team~\cite{benchmarkcc}, which are good candidates to evaluate modifications to~\checkedc.
 The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
@@
 \subsection{Experimental Setup}
+All our experiments are performed on a 6-Core Intel i7-10700H machine with 40 GB of RAM, running Ubuntu 20.04.3 LTS.
+We use WASM as our target sandbox and use a similar configuration as that of the recent work~\cite{rlbox-paper}.
+We use Valgrind's "massif" memory profiler~\cite{seward2008valgrind} to measure the memory usage and consider the peak heap usage of an application as its memory consumption.
+We measure runtime using the difference in elapsed clock cycles using~\code{clock()} API from POSIX's~\code{<time.h>}.
+We perform every measurement ten times and use the average as the final result.
 
-\subsection{Program Run-time Benchmarks}
+%\subsection{Program Run-time Benchmarks}
 \begin{figure*}[t]
@@
 
-\myparagraph{Benchmarking Methodology}
-\systemname is evaluated in terms of performance and conversion efforts similar to that of RLBOX \cite{rlbox-paper} by recording run-time and memory overhead incurred by a \systemname in comparison to an unconverted program in generic-C/checked-C.
-However, the scope of \systemname changes made on each of the programs is varied and the converted programs themselves are relatively tiny.
-Since each of the program is evaluated on a pre-defined test-suite that runs on a single instance of the program, sandbox creation cost is a one-time constant due to \systemname implementing RLBOX's \cite{rlbox-paper} WebAssembly sandbox API.
-\liyi{I do not understand the following sentence, what is test-bench? }
-Run-time performance is evaluated with C's~\code{<time.h>} library by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs) and measuring latency delta as a percentage.
-\liyi{need to rewrite}
-Timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
-\liyi{What is Valgrind? }
-Memory overhead is measured with Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
+%\myparagraph{Benchmarking Methodology}
+%\systemname is evaluated in terms of performance and conversion efforts similar to that of RLBOX \cite{rlbox-paper} by recording run-time and memory overhead incurred by a \systemname in comparison to an unconverted program in generic-C/checked-C.
+%However, the scope of \systemname changes made on each of the programs is varied and the converted programs themselves are relatively tiny.
+%Since each of the program is evaluated on a pre-defined test-suite that runs on a single instance of the program, sandbox creation cost is a one-time constant due to \systemname implementing RLBOX's \cite{rlbox-paper} WebAssembly sandbox API.
+%\liyi{I do not understand the following sentence, what is test-bench? }
+%Run-time performance is evaluated with C's~\code{<time.h>} library by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs) and measuring latency delta as a percentage.
+%\liyi{need to rewrite}
+%Timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted-ness of a pointer is propagated from the call site up until its declaration.
+%\liyi{What is Valgrind? }
+%Memory overhead is measured with Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap.
 Unlike the Runtime performance, Peak Memory is recorded as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx).
@@
 Valgrind's memory figures do not account for Sandbox's Heap allocations which are linearly proportional to sandboxed code and tainted pointers.
-All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.
+%All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.
+\subsection{Performance Overhead}
+Recent work~\cite{jangda2019not} shows that code executed as part of WASM sandbox incurs significant runtime overhead,~\ie$\sim$200\%.
+To better understand our runtime overhead, we first perform micro benchmarking of additional sandbox-related operations involved in~\systemname{}.
 
-\subsection{Micro-Benchmarks}
+\subsubsection{Micro-Benchmarking}
+\aravind{Arun, Draw Figure 7 using Latex and use the symbols below.}
+The~\fig{FIXTHIS} shows the results of our micro-benchmarking.
+We measure the following operations as part of this:
+
+\noindent\emph{Memory access in WASM Sandbox ($SBX_{m}$):} 
+All memory access in a sandbox goes through additional verification by the sandbox runtime resulting in runtime overhead.
+We perform 100K memory access (read and write) in a loop, measure the time inside the sandbox, and compare it with the time executed as a regular program.
+
+Our results (\fig{FIXTHIS}) show that we incur 156.6\% overhead for memory access in the WASM sandbox compared to that in a normal program.
+This is inline with the observations of the recent work~\cite{jangda2019not}.
+
+\noindent\emph{Sandbox Roundtrip ($SBX_{RT}$):}
+Here, we measure the time it takes to make a round trip between~\cregion and sandbox (\ucregion) compared to a regular function call and return.
+We create a no-op function as shown below:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+void noop() { return; }
+\end{minted}
+We place this~\code{noop} function in the sandbox and measure the time it takes to call and return from it:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+s = clock();
+sandbox_noop();
+e = clock();
+\end{minted}
+We compare the time with the regular call,~\ie when we place~\code{noop} in~\cregion.
+
+As shown in~\fig{FIXTHIS}, we incur an overhead of $\sim 400\%$. This is also in-line with the performance numbers reported by prior works~\cite{jangda2019not, rlbox-paper}.
+This is expected because transitioning to/from sandbox requires a context switch which is relatively more expensive than a regular function call (\ie\code{call} and~\code{ret} instruction).
+
+\noindent\emph{Tainted Pointer Access in~\cregion ($TP_{c}$):}
+As explained in~\sect{subsec:compilerimple}, we need to perform pointer swizzling to convert the sandbox-specific representation of tainted pointers to raw addresses.
+In addition, our instrumentation also checks at runtime that all tainted pointers are within the sandbox address range before accessing them.
+We measure this additional overhead by comparing tainted pointer access with regular pointer access.
+
+As shown in~\fig{FIXTHIS}, we incur 34\% overhead in accessing tainted pointers in~\cregion.
+Most of this is because of the validation check, which requires executing two additional compare instructions for every memory access.
 
@@
 
-\liyi{What is SBX, and how it is related to indirect-calls, pointer accesses and what is micro-benchimarks doing? Should we elaborate a little? }
-
-Figure \ref{fig:microbenchmarks} shows calls to the sandbox and sandboxed code execution to have significantly higher overhead in comparison to that of Tainted pointers.
+%\subsection{Micro-Benchmarks}
+%
+%
+%
+%\liyi{What is SBX, and how it is related to indirect-calls, pointer accesses and what is micro-benchimarks doing? Should we elaborate a little? }
+%
+%Figure \ref{fig:microbenchmarks} shows calls to the sandbox and sandboxed code execution to have significantly higher overhead in comparison to that of Tainted pointers.
 This observation suggests for sandboxing less performance intensive code and reducing the indirect calls between the two regions.
 Consequently, unsafe regions of code can be annotated cost-effectively by choosing to annotate the unsafe pointer references within the demarked unsafe function as tainted at the cost of losing out on the Sandbox's safety features.
-
-\subsubsection{\textbf{Memory Access in SBX}}
-Memory access performance between the checked and the WASM region involves crafting a test case that involves a simple pointer arithmetic operation enclosed in a loop of 100k iterations.
-
-\myparagraph{Results}
-156.6\% overhead as shown in Figure \ref{fig:microbenchmarks} is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.
-
-\subsubsection{\textbf{Indirect-Calls}}
-This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call from the checked region into the SBX) to return.
+%
+%\subsubsection{\textbf{Memory Access in SBX}}
+%Memory access performance between the checked and the WASM region involves crafting a test case that involves a simple pointer arithmetic operation enclosed in a loop of 100k iterations. 
+%
+%\myparagraph{Results}
+%156.6\% overhead as shown in Figure \ref{fig:microbenchmarks} is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.
+%
+%
+%\subsubsection{\textbf{Indirect-Calls}}
+%This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call from the checked region into the SBX) to return.
 This metric is evaluated against the time taken for a checked Callee (call within local scope) to return.
-
-\myparagraph{Result}
-Observed overhead is a consequence of control transfer overhead between the checked and unchecked region as mentioned in \cite{rlbox-paper}
-
-\subsubsection{\textbf{Pointer Accesses}}
-Run-time Pointer overhead is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
-
-\myparagraph{Result}
-34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.   
-\myparagraph{Result:}
-Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
+%
+%\myparagraph{Result}
+%Observed overhead is a consequence of control transfer overhead between the checked and unchecked region as mentioned in \cite{rlbox-paper}
+%
+%\subsubsection{\textbf{Pointer Accesses}}
+%Run-time Pointer overhead is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer.
+%
+%\myparagraph{Result}
+%34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.   
+%\myparagraph{Result:}
+%Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
 \begin{figure*}[t]

### 2023-01-16 sandp23/implementation.tex
Fixing eval

--- 2023-01-15 sandp23/implementation.tex
@@
 \subsection{Compiler}
+\label{subsec:compilerimple}
 The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.

### 2023-01-16 sandp23/implementation.tex
update.

@@
 \section{Implementation} \label{sec:implementation}
-As mentioned in~\sect{sec:overview}, our implementation of~\systemname{} has two main components,~\ie source level program partitioner (\sourcerewriter) and compiler.
-Given a well-typed~\systemname{} program and a sandbox configuration,~\sourcerewriter performs source-to-source transformation and splits the program into two parts (\ie two sets of source files) -- checked and tainted source files corresponding to~\cregion and~\ucregion respectively.
-Our compiler will compile the checked source files and adds the necessary dynamic instruction.
+As mentioned in~\Cref{sec:overview}, our implementation of~\systemname{} has two main components,~\ie source level program partitioner (\sourcerewriter) and compiler.
+Given a well-typed~\systemname{} program and a sandbox configuration,~\sourcerewriter performs source-to-source transformation and splits the program into two parts (\ie two sets of source files) -- checked and tainted source files corresponding to~$\cmode$ and~\ucregion respectively.
+Our compiler translates the checked source files and adds the necessary dynamic instruction.
 The tainted source files need to be compiled with the target sandbox compiler.
@@
 We will present our implementation using WebAssembly (WASM~\cite{bosamiya2020webassembly}) as our target sandbox.
-%However, as we discuss in~\sect{}, our implementation can be easily extended to other sandbox mechanisms.
+We also formalize the implementation and show a simulation theorem in \Cref{appx:comp1}.
 
@@
 Developers should use this qualifier to mark callback functions,~\ie functions in~\cregion, that can be called from the tainted region. 
-The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\ref{subsub:gencregion}).
+The~\sourcerewriter inserts appropriate sandbox dependent mechanisms to enable this (\ref{subsub:gencregion}).
 

### 2023-01-16 sandp23/implementation.tex
Minor

--- 2023-01-16 sandp23/implementation.tex
@@
 \begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
-- size_t read_msg(int sock_fd, _T_Array_ptr<char> msg,
--                             size_t sz) __Tainted {
-+ size_t read_msg(int sock_fd, char* msg,
+size_t read_msg(int sock_fd, char* msg,
+-                             size_t sz) _Tainted {
 +                             size_t sz) {

### 2023-01-16 sandp23/implementation.tex
add ACM
update.

@@
 We copy all non-tainted functions into~\cregion source files and make the following modifications to enable interaction with~\ucregion (\ie sandboxed code).
-
 We created a library once and for all for each sandbox.
@@
 
-\subsection{Supporting General Partioning Use Case}
+\subsection{Having Non-sandboxed Pointers}
 \input{figures/newusecase}
@@
 However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain regular (unchecked) pointers as well.
-%
 This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them. 
-In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
-
 For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.
 
+In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.
 The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
 It is best to annotate all functions with checked types for security purposes.
-However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted.
-This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
+However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted,
+which maintains the assurance and overhead balance.
 We add the additional constraint of not allowing casting from taint types to generic types to enable this generic use case without violating the safety guarantees of~\systemname.

### 2023-01-16 sandp23/limitations.tex
Updating conclusion

@@
-\section{Limitations}
+\section{Limitations and Future Work}
 \label{sec:limitations}
-We cannot handling function pointer calls to tainted function.
+Despite the effectiveness of~\systemname{}, it has the following limitations:
+
+\noindent\textbf{Sandbox Dependency:} \systemname{} assumes the availability of a sandbox and consequently inherits all the limitations of the corresponding sandbox.~\eg Programs should be compilable with the sandbox compiler.
+Also, as shown in~\sect{subsec:programoverhead}, the performance of the partitioned applications mainly depends on the sandbox.
+However, our implementation is not dependent on one specific sandbox and can be easily extended to other sandboxes. As part of our future work, we will extend our implementation to other sandboxes.
+
+\noindent\textbf{Annotation Effort:} Currently, all~\code{taint} annotations have to be done manually -- such that these annotations satisfy our type checker rules (\sect{sec:typechecking}).
+This could be tedious based on the complexity of the sandboxed function, its parameter complexity, and its dependency on other functions.
+We plan to develop an automated annotation tool such that, given the initial annotations (\lst{lst:humantaint}), our tool will automatically add all the required annotations (\lst{lst:humanadjust}) according to our type rules.

### 2023-01-16 sandp23/evaluation.tex
Adding limitations and other eval
Fixing eval

@@
 \begin{itemize}
+\item\textbf{Conversion Effort:} How much developer effort is needed to annotate parts of programs to run in~\systemname?
 \item\textbf{Performance Overhead:} What is the performance overhead (both runtime and memory) in using~\systemname?
-\item\textbf{Conversion Effort:} How much developer effort is needed to annotate parts of programs to run in~\systemname?
 \item\textbf{Security Impact:} How effective is the isolation provided by~\systemname{} in preventing security vulnerabilities?
@@
 %All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.
+
+\subsection{Conversion Effort}
+The flexibility of~\systemname{} (\sect{subsec:otherusecases}) enables an application to be partitioned in several ways -- with varying levels of overhead and assurance.
+We explore the following three ways:
+
+\noindent\emph{Checked C and Tainted Partitions ($CTP$):} This is the most comprehensive use case, where the~\cregion partition contains completely annotated Checked C code, and~\ucregion contains one or more tainted functions.
+This provides the complete spatial safety of the code in~\cregion{} including isolation from~\ucregion{}.
+
+\noindent\emph{Only Tainted Partition ($TP$):}
+This is the general partitioning (\sect{subsec:otherusecases}) use case without Checked C.
+This is similar to $CTP$, but~\cregion code does not use any Checked C annotations.
+This provides only the isolation guarantee without spatial safety.
+
+\noindent\emph{Only Tainted Pointers ($T_{Pr}$):}
+In this use case, we only use tainted pointers, and all code lies in~\cregion.
+This is needed for data isolation-only cases, where the developer might want to just isolate certain data (\eg user input) from the rest of the program.
+As explained in~\sect{subsec:compilerimple}, all access through tainted pointers will be dynamically checked to ensure that every access is within sandbox.
+This provides partial spatial safety by ensuring that any spatial violation involving tainted pointer cannot affect~\cregion.
+
+\input{tables/conversioneffort}
+\aravind{Arun, Please fix Table 2 with all details also remove Table 3} 
+\subsubsection{Conversion Methodology}
+We partitioned each program in our dataset using one of the above three methods.
+
+Our goal is to isolate risky input processing routines from the application code.
+We manually analyzed the source code of each application to identify which functions handle the input data and how.
+We also looked into previously reported vulnerabilities to identify risky input processing routined.
+We picked one of the above methods to partition based on how the input data is processed.
+\tbl{} shows the summary of our partitioning of the dataset.
+For~\texttt{ProFTPD}, we use $T_{Pr}$ method and mark the input data as trained, and consequently, five other pointers need to be marked as tainted according to the type rules.
+This resulted in a total of 6 pointer annotations.
+There is no code in~\ucregion as used $T_{Pr}$ method and used only tainted pointers.
+We followed the same approach for~\texttt{LibPNG} (\texttt{png2pnm} and~\texttt{pnm2png}).
+However, in this case, we had to annotate a lot more pointers (248) because of the complex nature of~\texttt{libPNG}'s internal structures.
+
+For~\texttt{MicroHTTPD} and~\texttt{UFTPD}, we used $TP$ method and marked all of their direct input handled methods as tainted and consequently moved to the sandbox.
+We also had to annotate several intermediate pointers as tainted.
+For~\texttt{TinyBignum} and~\texttt{Parsons}, we followed $CTP$ and marked all input processing routines as tainted and placed them in the sandbox.
+The rest of the non-sandboxed code is annotated completely using~\checkedc types and placed in~\cregion.
+
+\subsubsection{Conversion Effort}
+\araivnd{Arun, fix the average hours}
+The last column shows the number of hours required to partition each application.
+On average, it takes $\sim$ XX hours for each application. However, the exact time depends on the complexity of the code and pointers used in an application.
+Although the absolute time is high, partitioning is just a one-time effort for each application.
+We start by annotating functions and then iteratively fixing type-checker errors.
+Most of the time (80\%) in all applications is spent just running the type-checker.
+The type-checker stops at the first error without giving information about all the other variable that needs to be fixed. 
+For instance, in the following code:
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
+_t_Ptr<int> y = ...; 
+int *z;
+int *x;
+|\textcolor{red}{\faRemove}| z = y;
+x = z;
+\end{minted}
+The type-checker displays an error only at the first assignment.
+However, to correctly fix it, we need to annotate both~\code{x} and~\code{y}.
+If $N$ pointers need to be annotated, then in the worst case, we might have to run the type-checker $N$ times, annotating one additional pointer in every run.
+We plan to fix this in our future work and make the type-checker display all the affected pointer variables instead of just the type errors.
+
 \subsection{Performance Overhead}
@@
 Most of this is because of the validation check, which requires executing two additional compare instructions for every memory access.
+
+\subsubsection{Overhead on Dataset}
+\label{subsec:programoverhead}
 

### 2023-01-16 sandp23/implementation.tex
edit 5.3
Fixing eval
Shortening Implenentation section

@@
 \subsection{\sourcerewriter} \label{subsec:checkmate}
-The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{}).
-However, we use a small OCAML program (\numocamllines{}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
+The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{3028}).
+However, we use a small OCAML program (\numocamllines{680}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
 \aravind{Arun, please fix the number of lines you changed for clang frontend and ocaml program}
@@
 This qualifier makes ~\sourcerewriter copy the corresponding function into both ~\cregion and~\ucregion code such that callers from each region can invoke the corresponding copy.
-This helps handle certain simple utility functions that might be called from both ~\cregion and~\ucregion code.
-For example, ~\code{json_value_get_type} in our evaluation of parsons\_wasm has callers from both the ~\cregion and ~\ucregion.
+This helps handle certain simple utility functions that might be called from both regions. For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions. 
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-/* JSON Value API */
-_Mirror JSON_Value_Type json_value_get_type  
-      (_TPtr<const TJSON_Value> value) {
-    return value ? value->type : JSONError;
-}
-
-_Mirror _TPtr<TJSON_Object> json_value_get_object
-(_TPtr<const TJSON_Value> value) {
-    return json_value_get_type(value) == JSONObject ?
-    value->value.object : NULL;
+_Mirror int append_string(_TPtr<char> buf,
+const char* appendStr : itype(_Nt_array_ptr<const char>),
+_TPtr<char> buf_start, size_t buf_len) {
+/* Qualifier Rules:
+1.) No access to global data NOT marked "const"
+2.) Callees must be _Tainted or _Mirror
+*/
+...
 }
@@
 \end{minted}
-
-Placing such common functions in either of the regions restricts their usage from the other region.
-For instance, if we place~\code{foo} in~\ucregion, then all parameters of it should be tainted.
-Consequently, callers from~\cregion should use tainted pointers to invoke~\code{foo}, making the arguments and other dependent pointers tainted, leading to over-tainting.
-Similarly, placing~\code{foo} in~\cregion will make it hard for the callers in~\ucregion as they should now use the callback mechanism to invoke~\code{foo}.
-
-Developers can use the ~\code{_Mirror} qualifier, which makes ~\sourcerewriter copy the function into both regions so that callers in each region will invoke the corresponding copy, thereby, avoiding unnecessary tainting.
-
-However, ~\code{_Mirror} qualifier semantically asserts the function to only have callees that are either qualified with ~\code{_Tainted} or ~\code {_Mirror}.
-For example, ~\code{json_value_get_object}'s only callee
-\newline
-~\code{json_value_get_type} is ~\code{\_Mirror} qualified.
-Similarly, \systemname enforces ~\code{_Mirror} qualified functions to only access global checked data that are marked constant.
-These control-flow and data-flow compile-time semantics are enforced by \systemname to assert region locality and exclusivity amongst qualified function scopes in both the regions.
+Placing ~\code{append_string} in ~\ucregion would require marking ~\code{appendStr} as tainted, thereby, requiring callers from ~\cregion to pass tainted parameter to ~\code{appendStr} at call-site.
+This would require either costly marshalling or making the arguments and other dependent pointers tainted, leading to over-tainting.
+Similarly, placing ~\code{append_string} only in ~\cregion will make it hard for the callers in ~\ucregion as they should now use the callback mechanism to invoke~\code{append_string}.
+Qualifying ~\code{append_string} with~\code{_Mirror} makes ~\sourcerewriter duplicate the function to ~\ucregion so that callers in each region will invoke their corresponding local copy, thereby, avoiding unnecessary tainting.
+"\_Mirror" enforces control-flow and data-flow compile-time semantic rules as shown above to assert region locality and exclusivity amongst between both the regions.
 
 \noindent\emph{\_TLIB}:
-This qualifier relaxes type-checking rules on library functions, thereby, avoiding unnecessary copying of tainted data.
-Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.
-
-For instance, consider that in~\cregion, we want to string concat~\code{n} characters from tainted pointer~\code{t_str} to~\code{dst} (a checked pointer).
-We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
-We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then invoke~\code{strncat(dst, tmp, n);}.
-Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
-If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use~\code{strncat} with tainted source pointer in~\cregion.
-Consequently, ~\code{\_TLIB} qualified functions originally implemented in ~\cregion are expected to be non memory-modifying and thereby, its safe use is left at developer's discretion as \systemname does not enforce any semantics for this.
-For example, malloc cannot be qualified with ~\code{_TLIB} and thereby is implemented as an indirect call to sandbox-specific allocator.
-
-The~\code{_TLIB} qualifier enables the qualified function to be used as shown below:
+This qualifier relaxes type-checking rules on non-memory modifying library functions allowing developers to use the function freely in ~\cregion.
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-
-_TLIB extern char *strcat (char *__restrict __dest,
-            const char *__restrict __src); // In the header file
-...
 // First, manually check the memory is in tainted region.
@@
 \end{minted}
-The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.
-Note that developer needs to manually add the check before calling~\code{strcat}.
+
+In the above code, passing tainted pointer ~\code{t_str} to unqualified ~\code{strncat} is disallowed by our type-checker as it expects both the pointers to be non-tainted as we are in ~\cregion.
+If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use ~\code{strncat} with tainted source pointer in ~\cregion, instead of implementing the logic for string concatenation in ~\ucregion or marshalling data between the two regions.
+\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
+- extern char *strncat (char *__restrict __dest,
++ _TLIB extern char *strncat (char *__restrict __dest,
+const char *__restrict __src, size_t __n); // In the header file
+}
+\end{minted}
+
+Consequently, annotating ~\code{strncat} as shown above relaxes type-checking for all the arguments to its calls.
+However, \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying, thereby, requiring developers to be careful to while using ~\code{\_TLIB} to prevent memory address leaks.
 
@@
 
-\subsection{Having Non-sandboxed Pointers}
+\subsection{Balance Assurance and Overhead}
 \input{figures/newusecase}

### 2023-01-16 sandp23/implementation.tex
update.

@@
 \noindent\emph{\_Mirror}: 
-This qualifier makes ~\sourcerewriter copy the corresponding function into both ~\cregion and~\ucregion code such that callers from each region can invoke the corresponding copy.
-This helps handle certain simple utility functions that might be called from both regions. For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions. 
+This qualifier permits copying the corresponding function into both ~\cregion and~\ucregion, which permits the handling of certain simple utility functions that might be called from both regions.
+For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions.
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 \end{minted}
-Placing ~\code{append_string} in ~\ucregion would require marking ~\code{appendStr} as tainted, thereby, requiring callers from ~\cregion to pass tainted parameter to ~\code{appendStr} at call-site.
-This would require either costly marshalling or making the arguments and other dependent pointers tainted, leading to over-tainting.
-Similarly, placing ~\code{append_string} only in ~\cregion will make it hard for the callers in ~\ucregion as they should now use the callback mechanism to invoke~\code{append_string}.
-Qualifying ~\code{append_string} with~\code{_Mirror} makes ~\sourcerewriter duplicate the function to ~\ucregion so that callers in each region will invoke their corresponding local copy, thereby, avoiding unnecessary tainting.
-"\_Mirror" enforces control-flow and data-flow compile-time semantic rules as shown above to assert region locality and exclusivity amongst between both the regions.
+Qualifying ~\code{append_string} with~\code{_Mirror} duplicates the function in both regions,
+so that calling \code{append_string} permits the avoidance of over-tainting \code{appendStr},
+as \code{appendStr} can be unchecked and checked pointers in $\umode$ and $\cmode$ regions, respectively.
+"\_Mirror" enforces control-flow and data-flow compile-time semantic rules as shown above to assert region locality and exclusivity amongst between both the regions. \liyi{fix the sentence.}
 
 \noindent\emph{\_TLIB}:
-This qualifier relaxes type-checking rules on non-memory modifying library functions allowing developers to use the function freely in ~\cregion.
+This qualifier relaxes type-checking rules on library functions without memory modification allowing developers to use the function freely in ~\cregion.
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 
-In the above code, passing tainted pointer ~\code{t_str} to unqualified ~\code{strncat} is disallowed by our type-checker as it expects both the pointers to be non-tainted as we are in ~\cregion.
+Passing tainted pointer ~\code{t_str} to unqualified ~\code{strncat} above is disallowed by our type-checker as it expects both the pointers to be non-tainted since we are in ~\cregion.
 If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use ~\code{strncat} with tainted source pointer in ~\cregion, instead of implementing the logic for string concatenation in ~\ucregion or marshalling data between the two regions.
@@
 \subsection{Balance Assurance and Overhead}
+\label{subsec:otherusecases}
 \input{figures/newusecase}

### 2023-01-16 sandp23/evaluation.tex
Adding charts

@@
 \subsubsection{Conversion Effort}
-\araivnd{Arun, fix the average hours}
+\aravind{Arun, fix the average hours}
 The last column shows the number of hours required to partition each application.
@@
 
+\aravind{Arun, Fix Figure 8 and Remove Figure 6.}
+
+\aravind{Arun, Fix Figure 9 and Remove Figure 10.}
+
 \begin{figure}[h!]
@@
 \end{figure}
+
+\input{figures/runtimechart}
+
+\input{figures/memoryoverhead}
 

### 2023-01-17 sandp23/background.tex

@@
 Oftentimes, inserted checks can be optimized away by LLVM.
-Consider the above code to be enclosed in another condition, such as,~\inlinecode{if (n<u) if (n>0) a[n-1] =}.
-In such cases, the inserted check can be
-removed as the outer condition $\inlinecode{n<u}$ already ensures that the
-dereference is within bounds.
-Dual~\etal~\cite{duanrefactoring} found essentially no
+For instance, Dual~\etal~\cite{duanrefactoring} found essentially no
 overhead when running Checked C-converted portions of the FreeBSD kernel.
+%Consider the above code to be enclosed in another condition, such as,~\inlinecode{if (n<u) if (n>0) a[n-1] =}.
+%In such cases, the inserted check can be
+%removed as the outer condition $\inlinecode{n<u}$ already ensures that the
+%dereference is within bounds.
+%Dual~\etal~\cite{duanrefactoring} found essentially no
+%overhead when running Checked C-converted portions of the FreeBSD kernel.
 

### 2023-01-17 sandp23/figures/memoryoverhead.tex

\begin{figure}
\pgfplotsset{every x tick label/.append style={font=\tiny, rotate=30}}
\pgfplotsset{every y tick label/.append style={font=\tiny}}
\begin{tikzpicture}

\begin{axis}  
[  
    ybar,  
    enlargelimits=0.04,  
    ylabel={Memory Overhead (KiB)}, % the ylabel must precede a # symbol.  
    %xlabel={Programs},  
    symbolic x coords={ProFTPD, TinyBigNum, UFTPD, MicroHTTPD, png2pnm, pnm2png, Parsons}, % these are the specification of coordinates on the x-axis.  
    xtick=data,  
    bar width=10pt,
    width=7cm,
    x tick label/.append style={font=\tiny, rotate=30},
    y tick label/.append style={font=\tiny},
    axis background/.style={fill=gray!10},
    x label style={at={(axis description cs:0.5,-0.1)},anchor=north, font=\footnotesize},
    xlabel={Programs},
    ylabel style={font=\footnotesize},
     nodes near coords, % this command is used to mention the y-axis points on the top of the particular bar.  
    nodes near coords align={vertical},  
    ]  
\addplot coordinates {(ProFTPD,50) (TinyBigNum,90) (UFTPD, 120) (MicroHTTPD,80) (png2pnm, 60) (pnm2png, 150) (Parsons, 280)};

\end{axis}  
\end{tikzpicture} 
\caption{Memory Overhead of Partioned Programs.}
\label{fig:memoryoverhead}
\end{figure}

### 2023-01-17 sandp23/implementation.tex

@@
 // our type checker ignores this because 
-// the above _TLIB annotation.
+// the _TLIB annotation below.
 strncat(dst, t_str, n);
@@
 
-Passing tainted pointer ~\code{t_str} to unqualified ~\code{strncat} above is disallowed by our type-checker as it expects both the pointers to be non-tainted since we are in ~\cregion.
-If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use ~\code{strncat} with tainted source pointer in ~\cregion, instead of implementing the logic for string concatenation in ~\ucregion or marshalling data between the two regions.
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
@@
 
-Consequently, annotating ~\code{strncat} as shown above relaxes type-checking for all the arguments to its calls.
-However, \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying, thereby, requiring developers to be careful to while using ~\code{\_TLIB} to prevent memory address leaks.
+Passing tainted pointer \code{t_str} to unqualified ~\code{strncat} above is disallowed without having additional sandboxing for \code{t_str}.
+If a user ascertains that \code{t_str} has the right buffer size for \code{strncat}, she might label \code{strncat} with \code{\_TLIB}, so that \code{t_str} is treated as an unchecked pointer; such annotation relaxes type-checking for all the arguments to its calls.
+It is worth noting that \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying; therefore, using \code{\_TLIB} requires users' awareness of memory address leaks.
 

### 2023-01-17 sandp23/evaluation.tex
Adding al

@@
 The rest of the non-sandboxed code is annotated completely using~\checkedc types and placed in~\cregion.
+
+We ensured that the partitioned programs retained their expected functionality by verifying using corresponding test suites.
 
@@
 
+\subsection{Security Impact}
+\label{subsec:securityimpact}
+The last column of~\tbl{TODO} shows the list of all spatial safety vulnerabilities in the functions that have tainted types or are isolated in the~\ucregion{}.
+As expected,~\emph{all vulnerabilities} are prevented by~\systemname{}.
+The symbols~\vulprevented{} and~\vulisolated{} indicate whether the vulnerability was detected by our dynamic instrumentation or isolated in the sandbox, respectively.
+This shows that~\systemname{} provides an effective mechanism to prevent spatial safety vulnerabilities.
+
+\aravind{Fix the conversion effort table by the correct symbols}
+
 \subsection{Program Benchmarks}

### 2023-01-17 sandp23/evaluation.tex
[review] inline, in-line -> in line

--- 2023-01-16 sandp23/evaluation.tex
@@
 Our results (\fig{FIXTHIS}) show that we incur 156.6\% overhead for memory access in the WASM sandbox compared to that in a normal program.
-This is inline with the observations of the recent work~\cite{jangda2019not}.
+This is in line with the observations of the recent work~\cite{jangda2019not}.
 
@@
 
-As shown in~\fig{FIXTHIS}, we incur an overhead of $\sim 400\%$. This is also in-line with the performance numbers reported by prior works~\cite{jangda2019not, rlbox-paper}.
+As shown in~\fig{FIXTHIS}, we incur an overhead of $\sim 400\%$. This is also in line with the performance numbers reported by prior works~\cite{jangda2019not, rlbox-paper}.
 This is expected because transitioning to/from sandbox requires a context switch which is relatively more expensive than a regular function call (\ie\code{call} and~\code{ret} instruction).

### 2023-01-17 sandp23/evaluation.tex
Fixing eval section

--- 2023-01-17 sandp23/evaluation.tex
@@
 The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
-\aravind{Arun, please fix the dataset table with details of all programs.}
 
@@
 We use Valgrind's "massif" memory profiler~\cite{seward2008valgrind} to measure the memory usage and consider the peak heap usage of an application as its memory consumption.
-We measure runtime using the difference in elapsed clock cycles using~\code{clock()} API from POSIX's~\code{<time.h>}.
+We measure runtime using the difference in elapsed clock cycles using~\code{clock()} API from POSIX's~\code{<time.h>} and linux's ~\code{time} command.
 We perform every measurement ten times and use the average as the final result.
-
-%\subsection{Program Run-time Benchmarks}
-\begin{figure*}[t]
-\includegraphics[width=1.0\linewidth]{images/program_runtime_benchmark.png}
-\caption{Run-time benchmarking of ~\systemname.}
-\label{fig:runtime}
-\end{figure*}
 
@@
 \input{tables/conversioneffort} 
-\aravind{Arun, Please fix Table 2 with all details also remove Table 3} 
 \subsubsection{Conversion Methodology}
@@
 We picked one of the above methods to partition based on how the input data is processed.
-\tbl{} shows the summary of our partitioning of the dataset.
+Table ~\ref{table:conversioneffort} shows the summary of our partitioning of the dataset.
 For~\texttt{ProFTPD}, we use $T_{Pr}$ method and mark the input data as trained, and consequently, five other pointers need to be marked as tainted according to the type rules.
@@
 \subsubsection{Conversion Effort}
-\aravind{Arun, fix the average hours}
 The last column shows the number of hours required to partition each application.
@@
 \subsubsection{Micro-Benchmarking}
-\aravind{Arun, Draw Figure 7 using Latex and use the symbols below.}
-The~\fig{FIXTHIS} shows the results of our micro-benchmarking.
+The figure ~\ref{fig:microbenchmarks} shows the results of our micro-benchmarking.
 We measure the following operations as part of this:
@@
 
-Our results (\fig{FIXTHIS}) show that we incur 156.6\% overhead for memory access in the WASM sandbox compared to that in a normal program.
-This is in line with the observations of the recent work~\cite{jangda2019not}.
+Our results (~\fig{fig:microbenchmarks}) show that we incur 156.6\% overhead for memory access in the WASM sandbox compared to that in a normal program.
+This is inline with the observations of the recent work~\cite{jangda2019not}.
 
@@
 
-As shown in~\fig{FIXTHIS}, we incur an overhead of $\sim 400\%$. This is also in line with the performance numbers reported by prior works~\cite{jangda2019not, rlbox-paper}.
+As shown in ~\fig{fig:microbenchmarks}, we incur an overhead of $\sim 400\%$. This is also in-line with the performance numbers reported by prior works~\cite{jangda2019not, rlbox-paper}.
 This is expected because transitioning to/from sandbox requires a context switch which is relatively more expensive than a regular function call (\ie\code{call} and~\code{ret} instruction).
@@
 
-As shown in~\fig{FIXTHIS}, we incur 34\% overhead in accessing tainted pointers in~\cregion.
+As shown in~\fig{fig:microbenchmarks}, we incur 34\% overhead in accessing tainted pointers in~\cregion.
 Most of this is because of the validation check, which requires executing two additional compare instructions for every memory access.
@@
 \label{subsec:programoverhead}
-
-\aravind{Arun, Fix Figure 8 and Remove Figure 6.}
-
-\aravind{Arun, Fix Figure 9 and Remove Figure 10.}
-
-\begin{figure}[h!]
-\includegraphics[width=1.0\linewidth]{images/microbenchmark.png}
-\caption{\systemname Micro-Benchmarks}
-\label{fig:microbenchmarks}
-\end{figure}
-
 \input{figures/runtimechart}
@@
 %Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
-\begin{figure*}[t]
-\includegraphics[width=1.0\linewidth]{images/program_memory_benchmark.png}
-\caption{Memory benchmarking of ~\systemname.}
-\label{fig:memory}
-\end{figure*}
 
-\begin{table*}[htb]
-  \centering
-  \begin{tabular}{|p{2.6 cm }|p{1.6 cm}|p{1.9cm}|p{1.7cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-    \hline
-    \textbf{Task} & \textbf{ProFTPD} & \textbf{MicroHTTPD} & \textbf{UFTPD} & \textbf{LibPNG} & \textbf{Tiny Bignum} & \textbf{parsons (wasm)} & \textbf{parsons (tainted) }\\
-    \hline
-     \textbf{Pointers Annotated} & 6 & 139 & 146 & 248 & 69 & 364 & 378 \\
-     \textbf{Lines Sandboxed} & 0 & 650 & 90 & 0 & 30 & 800 & 0  \\
-     \textbf{CVEs fixed} & CVE-2010-4221 & CVE-2013-7039 & CVE-2020-14149, CVE-2020-5204 & CVE-2018-144550 & - & - & -  \\
-     \textbf{Time To Port} & 1 Day & 2 Days & 2 Days & 3 Days & 4 Hours & 1 Day & 2 Days \\
-    \hline
-  \end{tabular}
-  \caption{Conversion Efforts}
-\end{table*}
-
-\subsection{Program Benchmarks}
-\myparagraph{Overview}
-We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program.
+\systemname is evaluated on six programs with the use of a pre-included test suite for each program.
 We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code .
@@
 
-\myparagraph{Observations}
-Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
-\begin{itemize}
-  \item Run-time Overhead is proportional to the extent of 
-annotated pointers and sandboxed code.
-  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
-  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
-  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
-  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
-  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
-\end{itemize}
-
-\liyi{results?}
 \subsubsection{\textbf{ProFTPD}}
@@
 
+\subsubsection{\textbf{Results}}
+Our evaluation of \systemname on six programs yields the following results:
+\begin{itemize}
+  \item Run-time Overhead is proportional to the extent of 
+annotated pointers and sandboxed code.
+  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
+  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
+  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
+  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
+  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
+\end{itemize}
 % 

### 2023-01-17 sandp23/figures/runtimechart.tex
Fixing eval section

--- 2023-01-17 sandp23/figures/memoryoverhead.tex
@@
     enlargelimits=0.04,  
-    ylabel={Memory Overhead (KiB)}, % the ylabel must precede a # symbol.  
+    ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  
+    %xlabel={Micro-Benchmark Tests},  
+    symbolic x coords={Memory Access in SBX, Indirect Calls, Pointer Access}, % these are the specification of coordinates on the x-axis.  
+    xtick=data,  
+    bar width=10pt,
+    width=7cm,
+    x tick label/.append style={font=\tiny, rotate=30},
+    y tick label/.append style={font=\tiny},
+    axis background/.style={fill=gray!10},
+    x label style={at={(axis description cs:0.5,-0.1)},anchor=north, font=\footnotesize},
+    xlabel={tests},
+    ylabel style={font=\footnotesize},
+     nodes near coords, % this command is used to mention the y-axis points on the top of the particular bar.  
+    nodes near coords align={vertical},  
+    ]  
+\addplot coordinates {(Memory Access in SBX,156) (Indirect Calls, 398) (Pointer Access, 34.16)};
+
+\end{axis}  
+\end{tikzpicture} 
+\caption{\systemname Micro-Benchmarks.}
+\label{fig:microbenchmarks}
+\end{figure}
+
+\begin{figure}
+\pgfplotsset{every x tick label/.append style={font=\tiny, rotate=30}}
+\pgfplotsset{every y tick label/.append style={font=\tiny}}
+\begin{tikzpicture}
+
+\begin{axis}  
+[  
+    ybar,  
+    enlargelimits=0.04,  
+    ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  
     %xlabel={Programs},  
@@
     ]  
-\addplot coordinates {(ProFTPD,50) (TinyBigNum,90) (UFTPD, 120) (MicroHTTPD,80) (png2pnm, 60) (pnm2png, 150) (Parsons, 280)};
+\addplot coordinates {(ProFTPD,4.2) (TinyBigNum,122) (UFTPD, 120) (MicroHTTPD,3.8) (png2pnm, 11) (pnm2png, 46.5) (Parsons, 267)};
 
@@
 \end{tikzpicture} 
-\caption{Memory Overhead of Partioned Programs.}
-\label{fig:memoryoverhead}
+\caption{Runtime Overhead of Partioned Programs.}
+\label{fig:runtimeoverhead}
 \end{figure}

### 2023-01-17 sandp23/tables/datasettable.tex
Fixing eval section

@@
 \midrule
-\rowcolor{black!15} 1           & ProFTPD          & High performance FTP Server   & XXX K                                                                               \\ 
-2           & MicroHTTPD       & Simple HTTPD Server           & YY K                                                                                \\ 
-\rowcolor{black!15} 3           & UFTPD            & UDP based FTP Server          & ZZZ K                                                                               \\ 
-4           & png2pnm          & Program to convert png to pnm & UUU K                                                                               \\ 
+\rowcolor{black!15} 1           & ProFTPD          & High performance FTP Server   & 556 K                                                                               \\ 
+2           & MicroHTTPD       & Simple HTTPD Server           & 122 K                                                                                \\ 
+\rowcolor{black!15} 3           & UFTPD            & UDP based FTP Server          & 3 K                                                                               \\ 
+4           & LibPNG          & Program to convert between png and pnm & 76 K                                                                               \\ 
+\rowcolor{black!15}5           & TinyBigNum          & Multiple-precision integer implementation & 1.6 K                                                                               \\ 
+6           & Parsons          & JSON parsing library & 3.1 K                                                                               \\
+
 \bottomrule

### 2023-01-17 sandp23/evaluation.tex

@@
 
+We ensured that the partitioned programs retained their expected functionality by verifying using corresponding test suites.
+
 \subsubsection{Conversion Effort}
@@
 %Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.
+
+\begin{figure*}[t]
+\includegraphics[width=1.0\linewidth]{images/program_memory_benchmark.png}
+\caption{Memory benchmarking of ~\systemname.}
+\label{fig:memory}
+\end{figure*}
+
+\begin{table*}[htb]
+  \centering
+  \begin{tabular}{|p{2.6 cm }|p{1.6 cm}|p{1.9cm}|p{1.7cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+    \hline
+    \textbf{Task} & \textbf{ProFTPD} & \textbf{MicroHTTPD} & \textbf{UFTPD} & \textbf{LibPNG} & \textbf{Tiny Bignum} & \textbf{parsons (wasm)} & \textbf{parsons (tainted) }\\
+    \hline
+     \textbf{Pointers Annotated} & 6 & 139 & 146 & 248 & 69 & 364 & 378 \\
+     \textbf{Lines Sandboxed} & 0 & 650 & 90 & 0 & 30 & 800 & 0  \\
+     \textbf{CVEs fixed} & CVE-2010-4221 & CVE-2013-7039 & CVE-2020-14149, CVE-2020-5204 & CVE-2018-144550 & - & - & -  \\
+     \textbf{Time To Port} & 1 Day & 2 Days & 2 Days & 3 Days & 4 Hours & 1 Day & 2 Days \\
+    \hline
+  \end{tabular}
+  \caption{Conversion Efforts}
+\end{table*}
+
+\subsection{Security Impact}
+\label{subsec:securityimpact}
+The last column of~\tbl{TODO} shows the list of all spatial safety vulnerabilities in the functions that have tainted types or are isolated in the~\ucregion{}.
+As expected,~\emph{all vulnerabilities} are prevented by~\systemname{}.
+The symbols~\vulprevented{} and~\vulisolated{} indicate whether the vulnerability was detected by our dynamic instrumentation or isolated in the sandbox, respectively.
+This shows that~\systemname{} provides an effective mechanism to prevent spatial safety vulnerabilities.
+
+\aravind{Fix the conversion effort table by the correct symbols}
+
+\subsection{Program Benchmarks}
+\myparagraph{Overview}
+We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program.
+We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code .
+However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible.
+Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
+Consequently, each of the conversions for the six programs follows a varied approach.
+
+\myparagraph{Observations}
+Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
+\begin{itemize}
+  \item Run-time Overhead is proportional to the extent of 
+annotated pointers and sandboxed code.
+  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
+  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
+  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
+  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
+  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
+\end{itemize}
 

### 2023-01-17 sandp23/implementation.tex

@@
 \subsection{\sourcerewriter} \label{subsec:checkmate}
-The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{3028}).
-However, we use a small OCAML program (\numocamllines{680}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
-\aravind{Arun, please fix the number of lines you changed for clang frontend and ocaml program}
+The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (3028 SLoc).
+However, we use a small OCAML program (680 SLoc) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
 
@@
 \noindent\emph{\_Mirror}: 
-This qualifier permits copying the corresponding function into both ~\cregion and~\ucregion, which permits the handling of certain simple utility functions that might be called from both regions.
+This qualifier permits copying the corresponding function into both ~\cregion and~\ucregion, which permits the handling of certain simple utility functions that are called from both regions.
 For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions.
@@
 \end{minted}
-Qualifying ~\code{append_string} with~\code{_Mirror} duplicates the function in both regions,
-so that calling \code{append_string} permits the avoidance of over-tainting \code{appendStr},
-as \code{appendStr} can be unchecked and checked pointers in $\umode$ and $\cmode$ regions, respectively.
-"\_Mirror" enforces control-flow and data-flow compile-time semantic rules as shown above to assert region locality and exclusivity amongst between both the regions. \liyi{fix the sentence.}
+Qualifying ~\code{append_string} with~\code{_Mirror} duplicates the function in both regions, allowing calls to \code{append_string} with parameter to \code{appendStr} as an unchecked or checked pointer within $\umode$ and $\cmode$ regions, respectively.
+Consequently complexity from over-tainting is avoided as ~\code {appendStr} need not be tainted in ~\cregion and neither are callbacks required to access ~\code{append_string} from ~\ucregion.
+"\_Mirror" enforces control-flow and data-flow compile-time semantic rules to ensure all variable and function call dependencies of mirrored functions required for ~\ucregion's compilation are resolved.
 
 \noindent\emph{\_TLIB}:
-This qualifier relaxes type-checking rules on library functions without memory modification allowing developers to use the function freely in ~\cregion.
+This qualifier relaxes type-checking rules on library functions, allowing developers to use the function freely in ~\cregion.
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 
-Passing tainted pointer \code{t_str} to unqualified ~\code{strncat} above is disallowed without having additional sandboxing for \code{t_str}.
-If a user ascertains that \code{t_str} has the right buffer size for \code{strncat}, she might label \code{strncat} with \code{\_TLIB}, so that \code{t_str} is treated as an unchecked pointer; such annotation relaxes type-checking for all the arguments to its calls.
+Passing tainted pointer \code{t_str} to unqualified ~\code{strncat} above is disallowed without having additional ~\ucregion implementation for \code{strncat}.
+If a user ascertains that \code{t_str} has the right buffer size for \code{strncat}, she might label \code{strncat} with \code{\_TLIB}, so that \code{t_str} can be treated as an checked pointer parameter; such annotation relaxes type-checking for all the arguments to its calls.
 It is worth noting that \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying; therefore, using \code{\_TLIB} requires users' awareness of memory address leaks.
@@
 The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
-We implemented the compiler (\numcompilerlines{}) on top of the existing~\checkedc compiler.
-\aravind{Arun, please fix the number of changes you made for the compiler}
-
+We implemented the compiler as a clang library (1812K SLoc) on top of the existing~\checkedc compiler.
 Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
@@
 - char buff[MAX_MSG_SIZE] _Tainted;
-+ _t_Array_ptr<char> buff: count(MAX_MSG_SIZE)
++ _T_Array_ptr<char> buff: count(MAX_MSG_SIZE)
 +    = sbx_allocate(MAX_MSG_SIZE);
@@
 
-The~\ucregion source files will be compiled with sandbox compiler,~\ie NAMEOFSANDBOXCOMPILER (CITATION).
+The~\ucregion source files will be compiled with sandbox compiler,~\ie WASI-SDK (~\cite{wasi-clang}).
 Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
 \cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers.
-\aravind{Arun, please fill up the name of the sandbox compiler and add citation.}
 

### 2023-01-17 sandp23/appendix.tex

@@
 Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf().
-Given that significant \systemname conversion efforts is attributed to understanding the source-code and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling the data between regions, Tiny-bignum only required 4 hours.
-The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

### 2023-01-17 sandp23/evaluation.tex

--- 2023-01-17 sandp23/evaluation.tex
@@
 
+We ensured that the partitioned programs retained their expected functionality by verifying using corresponding test suites.
+
 \subsubsection{Conversion Effort}
 The last column shows the number of hours required to partition each application.
-On average, it takes $\sim$ XX hours for each application. However, the exact time depends on the complexity of the code and pointers used in an application.
+On average, it takes $\sim$ 3.5 hours for each application. However, the exact time depends on the complexity of the code and pointers used in an application.
 Although the absolute time is high, partitioning is just a one-time effort for each application.
@@
 
-\subsubsection{Overhead on Dataset}
-\label{subsec:programoverhead}
-\input{figures/runtimechart}
-
-\input{figures/memoryoverhead}
-
 %\subsection{Micro-Benchmarks}
@@
 
-\systemname is evaluated on six programs with the use of a pre-included test suite for each program.
-We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code .
-However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible.
+\subsubsection{Overhead on Dataset}
+\label{subsec:programoverhead}
+\input{figures/runtimechart}
+
+\input{figures/memoryoverhead}
+\systemname is evaluated on six programs on the basis of conversion efforts and performance with the use of a pre-included test suite for each program.
+We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code.
+However, this would require the bugs to be predicted before they are discovered, which is chronologically impossible.
 Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
-Consequently, each of the conversions for the six programs follows a varied approach.
 
-\subsubsection{\textbf{ProFTPD}}
+\myparagraph{\textbf{ProFTPD}}
 \systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
@@
 \myparagraph{Testbench:}
-Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance.
-For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets().
-This sampling is repeated 10 times and its mean value is reflected in the below table.
+ProFTPD was evaluated by sampling the time taken to execute its comprehensive test-suite.
+The delta in run-time between \systemname ProFTPD and its corresponding unconverted version's sample was calculated as a percentage.
+This percentage was recorded over ten iterations and its median value has been reflected in the table \ref{table:dataset}.
 
-\subsubsection{\textbf{UFTPD}}
+\myparagraph{\textbf{UFTPD}}
 \systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
@@
 \myparagraph{Testbench:}
-For evaluation, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
-Following this we record an entry for each of these tests as a mean of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the below table.
+Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
+Following this we record an entry for each of these tests as a median of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the table \ref{table:dataset}.
 
@@
 
-\subsubsection{\textbf{Results}}
+\subsubsection{\textbf{\systemname Dataset Results}}
 Our evaluation of \systemname on six programs yields the following results:
@@
   \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days.   
-  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
 \end{itemize}
+
+\subsection{Security Impact}
+\label{subsec:securityimpact}
+The last column of the Table ~\ref{table:conversioneffort} shows the list of all spatial safety vulnerabilities in the functions that have tainted types or are isolated in the~\ucregion{}.
+As expected,~\emph{all vulnerabilities} are prevented by~\systemname{}.
+
+The symbols~\vulprevented{} and~\vulisolated{} indicate whether the vulnerability was detected by our dynamic instrumentation or isolated in the sandbox, respectively.
+This shows that~\systemname{} provides an effective mechanism to prevent spatial safety vulnerabilities.
+
+\aravind{Fix the conversion effort table by the correct symbols}
+
 % 

### 2023-01-17 sandp23/evaluation.tex
fixing wording in section 6.

@@
 We also selected a few standalone programs suggested by the ~\checkedc team~\cite{benchmarkcc}, which are good candidates to evaluate modifications to~\checkedc.
-The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
+\Cref{table:dataset} shows the program selected as part of our evaluation dataset.
 
 \subsection{Experimental Setup}
-All our experiments are performed on a 6-Core Intel i7-10700H machine with 40 GB of RAM, running Ubuntu 20.04.3 LTS.
-We use WASM as our target sandbox and use a similar configuration as that of the recent work~\cite{rlbox-paper}.
-We use Valgrind's "massif" memory profiler~\cite{seward2008valgrind} to measure the memory usage and consider the peak heap usage of an application as its memory consumption.
-We measure runtime using the difference in elapsed clock cycles using~\code{clock()} API from POSIX's~\code{<time.h>} and linux's ~\code{time} command.
-We perform every measurement ten times and use the average as the final result.
+All experiments are performed on a 6-Core Intel i7-10700H machine with 40 GB of RAM, running Ubuntu 20.04.3 LTS.
+We use WASM as our target sandbox and use a similar configuration as that of the recent work~\cite{rlbox-paper};
+and Valgrind's "massif" memory profiler~\cite{seward2008valgrind} to measure the memory usage and consider the peak heap usage of an application as its memory consumption.
+We measure runtime using the difference in elapsed clock cycles using~\code{clock()} API from POSIX's~\code{<time.h>} and linux's ~\code{time} command, and perform every measurement ten times and use the average as the final result.
 
@@
 The delta in run-time between \systemname ProFTPD and its corresponding unconverted version's sample was calculated as a percentage.
-This percentage was recorded over ten iterations and its median value has been reflected in the table \ref{table:dataset}.
+This percentage was recorded over ten iterations and its median value has been reflected in \Cref{table:dataset}.
 

### 2023-01-17 sandp23/evaluation.tex
update.

@@
 Our goal is to isolate risky input processing routines from the application code.
-We manually analyzed the source code of each application to identify which functions handle the input data and how.
-We also looked into previously reported vulnerabilities to identify risky input processing routined.
-We picked one of the above methods to partition based on how the input data is processed.
-Table ~\ref{table:conversioneffort} shows the summary of our partitioning of the dataset.
-For~\texttt{ProFTPD}, we use $T_{Pr}$ method and mark the input data as trained, and consequently, five other pointers need to be marked as tainted according to the type rules.
-This resulted in a total of 6 pointer annotations.
-There is no code in~\ucregion as used $T_{Pr}$ method and used only tainted pointers.
-We followed the same approach for~\texttt{LibPNG} (\texttt{png2pnm} and~\texttt{pnm2png}).
-However, in this case, we had to annotate a lot more pointers (248) because of the complex nature of~\texttt{libPNG}'s internal structures.
+We manually analyze each application's source code to identify which functions handle the input data and how.
+We also look into previously reported vulnerabilities to identify risky input processing routines.
+We picke one of the above methods to partition based on how the input data is processed.
+Table ~\ref{table:conversioneffort} is a summary of our dataset partitioning.
+For~\texttt{ProFTPD}, $T_{Pr}$ method is used and the input data is marked as tainted.
+Consequently, five other pointers need to be marked as tainted according to the type rules.
+This results in a total of 6 pointer annotations.
+There is no code in~\ucregion as used $T_{Pr}$ method with only tainted pointers being used.
+We follow the same approach for~\texttt{LibPNG} (\texttt{png2pnm} and~\texttt{pnm2png}).
+However, in this case, we have to annotate much more pointers (248) due to the complicated \texttt{libPNG}'s internal structures.
 
-For~\texttt{MicroHTTPD} and~\texttt{UFTPD}, we used $TP$ method and marked all of their direct input handled methods as tainted and consequently moved to the sandbox.
-We also had to annotate several intermediate pointers as tainted.
-For~\texttt{TinyBignum} and~\texttt{Parsons}, we followed $CTP$ and marked all input processing routines as tainted and placed them in the sandbox.
+For~\texttt{MicroHTTPD} and~\texttt{UFTPD}, $TP$ method is used and we mark all of the direct input handled methods as tainted, which are consequently moved to the sandbox, with annotating several intermediate pointers as tainted.
+For~\texttt{TinyBignum} and~\texttt{Parsons},
+we follow $CTP$ and mark all input processing routines as tainted and place them in the sandbox.
 The rest of the non-sandboxed code is annotated completely using~\checkedc types and placed in~\cregion.
@@
 \subsubsection{Conversion Effort}
-The last column shows the number of hours required to partition each application.
-On average, it takes $\sim$ 3.5 hours for each application. However, the exact time depends on the complexity of the code and pointers used in an application.
-Although the absolute time is high, partitioning is just a one-time effort for each application.
+The second last column shows the hour numbers for partitioning applications.
+On average, it takes $\sim$ 3.5 hours for each partitioning. However, the exact time depends on the complexity of the application and the pointer usage.
+Although the absolute time is high, partitioning is a one-time effort for each application.
 We start by annotating functions and then iteratively fixing type-checker errors.
-Most of the time (80\%) in all applications is spent just running the type-checker.
-The type-checker stops at the first error without giving information about all the other variable that needs to be fixed. 
-For instance, in the following code:
+Most of the time (80\%) is spent on running the type-checker.
+The type-checker stops at the first error without giving information about other variables that needs to be fixed. 
+For an instance, in the following code:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 \end{minted}
-The type-checker displays an error only at the first assignment.
+The type-checker displays an error only for the first assignment.
 However, to correctly fix it, we need to annotate both~\code{x} and~\code{y}.
 If $N$ pointers need to be annotated, then in the worst case, we might have to run the type-checker $N$ times, annotating one additional pointer in every run.
-We plan to fix this in our future work and make the type-checker display all the affected pointer variables instead of just the type errors.
+We plan to fix this in our future work by making the conversion procedure automatic.
+%make the type-checker display all the affected pointer variables instead of just the type errors.
 
@@
 Recent work~\cite{jangda2019not} shows that code executed as part of WASM sandbox incurs significant runtime overhead,~\ie$\sim$200\%.
-To better understand our runtime overhead, we first perform micro benchmarking of additional sandbox-related operations involved in~\systemname{}.
+To better understand our runtime overhead, we first perform micro benchmarking of additional sandbox-related operations in~\systemname{}.
 
 \subsubsection{Micro-Benchmarking}
-The figure ~\ref{fig:microbenchmarks} shows the results of our micro-benchmarking.
+\Cref{fig:microbenchmarks} shows our micro-benchmarking result.
 We measure the following operations as part of this:
@@
 \noindent\emph{Memory access in WASM Sandbox ($SBX_{m}$):} 
-All memory access in a sandbox goes through additional verification by the sandbox runtime resulting in runtime overhead.
-We perform 100K memory access (read and write) in a loop, measure the time inside the sandbox, and compare it with the time executed as a regular program.
-
-Our results (~\fig{fig:microbenchmarks}) show that we incur 156.6\% overhead for memory access in the WASM sandbox compared to that in a normal program.
+All memory accesses in a sandbox need additional verification by the sandbox runtime, which results in runtime overhead.
+We perform 100K memory accesses (read and write) in a loop, measure the time inside the sandbox, and compare it with the time executed as a regular program.
+%
+The results (~\fig{fig:microbenchmarks}) show that we incur 156.6\% overhead for memory accesses in the WASM sandbox compared to that in a normal program.
 This is inline with the observations of the recent work~\cite{jangda2019not}.
@@
 \noindent\emph{Sandbox Roundtrip ($SBX_{RT}$):}
-Here, we measure the time it takes to make a round trip between~\cregion and sandbox (\ucregion) compared to a regular function call and return.
-We create a no-op function as shown below:
+We measure the time to make a round trip between~\cregion and sandbox (\ucregion) compared to a regular function call and return.
+We create a no-op function below:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 \end{minted}
-We place this~\code{noop} function in the sandbox and measure the time it takes to call and return from it:
+We place this~\code{noop} function in the sandbox and measure the time to call and return from it:
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
@@
 \end{minted}
-We compare the time with the regular call,~\ie when we place~\code{noop} in~\cregion.
+We compare the time with a regular call when \code{noop} is in~\cregion.
 
-As shown in ~\fig{fig:microbenchmarks}, we incur an overhead of $\sim 400\%$. This is also in-line with the performance numbers reported by prior works~\cite{jangda2019not, rlbox-paper}.
-This is expected because transitioning to/from sandbox requires a context switch which is relatively more expensive than a regular function call (\ie\code{call} and~\code{ret} instruction).
+As shown in ~\fig{fig:microbenchmarks}, we incur an overhead of $\sim 400\%$. This is also inline with the performance reported by prior works~\cite{jangda2019not, rlbox-paper}.
+This is expected because transitions to/from sandbox require context switches which are more expensive than regular function calls (\ie\code{call} and~\code{ret} instructions).
 
@@
 In addition, our instrumentation also checks at runtime that all tainted pointers are within the sandbox address range before accessing them.
-We measure this additional overhead by comparing tainted pointer access with regular pointer access.
+We measure this additional overhead by comparing tainted pointer accesses with regular pointer accesses.
 
-As shown in~\fig{fig:microbenchmarks}, we incur 34\% overhead in accessing tainted pointers in~\cregion.
-Most of this is because of the validation check, which requires executing two additional compare instructions for every memory access.
+As in~\fig{fig:microbenchmarks}, we incur 34\% overhead in accessing tainted pointers in~\cregion,
+ due to additional validation checks, which require executing two compare instructions for every memory access.
 

### 2023-01-17 sandp23/figures/runtimechart.tex
update.

@@
     %xlabel={Micro-Benchmark Tests},  
-    symbolic x coords={Memory Access in SBX, Indirect Calls, Pointer Access}, % these are the specification of coordinates on the x-axis.  
+    symbolic x coords={Memory Access in SBX, SBX Roundtrip, Tainted-ptr Access}, % these are the specification of coordinates on the x-axis.  
     xtick=data,  
@@
     ]  
-\addplot coordinates {(Memory Access in SBX,156) (Indirect Calls, 398) (Pointer Access, 34.16)};
+\addplot coordinates {(Memory Access in SBX,156) (SBX Roundtrip, 398) (Tainted-ptr Access, 34.16)};
 
@@
     ]  
-\addplot coordinates {(ProFTPD,4.2) (TinyBigNum,122) (UFTPD, 120) (MicroHTTPD,3.8) (png2pnm, 11) (pnm2png, 46.5) (Parsons, 267)};
+\addplot coordinates {(ProFTPD,4.2) (TinyBigNum,121.7) (UFTPD, 2.7) (MicroHTTPD,3.8) (png2pnm, 11.47) (pnm2png, 46.5) (Parsons, 267)};
 

### 2023-01-18 sandp23/evaluation.tex
Fixing Eval section and Appendix

--- 2023-01-17 sandp23/evaluation.tex
@@
 The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
+In our conversion, we re-introduce corrected memory bugs and then perform \systemname annotations on the relevant buggy code.
+However, this by itself is based on the assumption that the bugs are predicted before they are discovered, which is chronologically impossible.
+Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without input on the bug locality.
 
 \subsection{Experimental Setup}
+\label{experimentalsetup}
 All our experiments are performed on a 6-Core Intel i7-10700H machine with 40 GB of RAM, running Ubuntu 20.04.3 LTS.
@@
 \begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-_t_Ptr<int> y = ...; 
+_TPtr<int> y = ...; 
 int *z;
@@
 \input{figures/memoryoverhead}
-\systemname is evaluated on six programs on the basis of conversion efforts and performance with the use of a pre-included test suite for each program.
-We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code.
-However, this would require the bugs to be predicted before they are discovered, which is chronologically impossible.
-Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
 
-\myparagraph{\textbf{ProFTPD}}
-\systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
-We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
+ProFTPD's run-time overhead is expected as our conversion was small and strictly encapsulates CVE-2010-4221.
+Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" on a 4.0K file, following which, we record overhead as described in ~\ref{experimentalsetup}.
+Run-time overhead for UFTPD is expected as the sandboxed code was less performance intensive and FTP-protocol by itself overshadows \systemname's overhead.
 
-\myparagraph{Testbench:}
-ProFTPD was evaluated by sampling the time taken to execute its comprehensive test-suite.
-The delta in run-time between \systemname ProFTPD and its corresponding unconverted version's sample was calculated as a percentage.
-This percentage was recorded over ten iterations and its median value has been reflected in the table \ref{table:dataset}.
-
-\myparagraph{\textbf{UFTPD}}
-\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
-CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function.
-CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function.
-Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.
-
-\myparagraph{Testbench:}
-Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
-Following this we record an entry for each of these tests as a median of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the table \ref{table:dataset}.
-
-\myparagraph{Result:}
-Although, \systemname-UFTPD consists of modest changes in annotations and sandboxed code, observed overhead is significantly less at 10\% due to the less performance intensive sandboxed code and FTP-protocol overshadowed overhead from \systemname instrumentation and sandbox.
-
-\subsubsection{\textbf{\systemname Dataset Results}}
-Our evaluation of \systemname on six programs yields the following results:
+Our evaluation of \systemname's performance results on the dataset yields the observations:
 \begin{itemize}
@@
 annotated pointers and sandboxed code.
-  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
-  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
-  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
-  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days.   
+  \item Marshalling can be avoided if Taintedness of a pointer is propagated across its data-flow.
+  \item Marshalling might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout its data-flow. 
 \end{itemize}
@@
 This shows that~\systemname{} provides an effective mechanism to prevent spatial safety vulnerabilities.
-
-\aravind{Fix the conversion effort table by the correct symbols}
 

### 2023-01-18 sandp23/evaluation.tex

--- 2023-01-17 sandp23/evaluation.tex
@@
 We also selected a few standalone programs suggested by the ~\checkedc team~\cite{benchmarkcc}, which are good candidates to evaluate modifications to~\checkedc.
-\Cref{table:dataset} shows the program selected as part of our evaluation dataset.
+The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset. 
+In our conversion, we re-introduce corrected memory bugs and then perform \systemname annotations on the relevant buggy code.
+However, this is based on the assumption that the bugs are predicted before they are discovered, which is chronologically impossible.
+Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations 
+without input on the bug locality.
 
 \subsection{Experimental Setup}
+\label{experimentalsetup}
 All experiments are performed on a 6-Core Intel i7-10700H machine with 40 GB of RAM, running Ubuntu 20.04.3 LTS.
@@
 \input{figures/memoryoverhead}
-\systemname is evaluated on six programs on the basis of conversion efforts and performance with the use of a pre-included test suite for each program.
-We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code.
-However, this would require the bugs to be predicted before they are discovered, which is chronologically impossible.
-Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality.
 
-\myparagraph{\textbf{ProFTPD}}
-\systemname changes for ProFTPD were limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221.
-We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope.
+ProFTPD's run-time overhead is expected as our conversion was small and strictly encapsulates CVE-2010-4221.
+Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" on a 4.0K file, following which, we record overhead as described in ~\ref{experimentalsetup}.
+Run-time overhead for UFTPD is expected as the sandboxed code was less performance intensive and FTP-protocol by itself overshadows \systemname's overhead.
 
-\myparagraph{Testbench:}
-ProFTPD was evaluated by sampling the time taken to execute its comprehensive test-suite.
-The delta in run-time between \systemname ProFTPD and its corresponding unconverted version's sample was calculated as a percentage.
-This percentage was recorded over ten iterations and its median value has been reflected in \Cref{table:dataset}.
-
-\myparagraph{\textbf{UFTPD}}
-\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
-CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function.
-CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function.
-Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.
-
-\myparagraph{Testbench:}
-Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop.
-Following this we record an entry for each of these tests as a median of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the table \ref{table:dataset}.
-
-\myparagraph{Result:}
-Although, \systemname-UFTPD consists of modest changes in annotations and sandboxed code, observed overhead is significantly less at 10\% due to the less performance intensive sandboxed code and FTP-protocol overshadowed overhead from \systemname instrumentation and sandbox.
-
-\subsubsection{\textbf{\systemname Dataset Results}}
-Our evaluation of \systemname on six programs yields the following results:
+Our evaluation of \systemname's performance results on the dataset yields the observations:
 \begin{itemize}

### 2023-01-18 sandp23/appendix.tex

@@
 Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf().
+
+\myparagraph{\textbf{ProFTPD}:}
+\systemname changes for ProFTPD were lim-
+ited was small and exactly encapsulates CVE-2010-4221 by marking the user input to the unsafe function
+"pr\_netio\_telnet\_gets()" as tainted.
+Although we propagate the taintedness of the above function's argument, run-time overhead measured by following the methodology in ~\ref{experimentalsetup} was minimal as the data-flow graph for the pointer was small and thereby, required less pointers to be annotated.
+
+\myparagraph{\textbf{UFTPD}}
+\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204.
+CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function.
+CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function.
+Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic.

### 2023-01-18 sandp23/introduction.tex
update introduction.

--- 2023-01-09 sandp23/introduction.tex
@@
 are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798} 
-despite a large body of work that tries to prevent them~\cite{song2019sanitizing}.
+despite many efforts that tried to prevent them~\cite{song2019sanitizing}.
 Several industrial and research efforts, including CCured~\cite{Necula2005},
 Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
-have investigated ways to
-\mzu{compile C programs 
-  to automatic spatial safety enforcement.}
-\mz{You cannot compile a program to enforcement.}
-% 
+have investigated ways to better compile C programs with automatic spatial safety enforcement.
 These approaches all impose performance overheads deemed too high for deployment use. 
 % 
-\liyi{Should we say this? Is the paper about overhead reducing? }
+%\liyi{Should we say this? Is the paper about overhead reducing? }
 Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
-open-source extension to C
+open-source extension to C,
+to ensure a program’s spatial safety by introducing additional annotations and
+enforcing new types, checked pointer types in particular.
 %
-\mzs{ with new types and annotations to ensure a program’s spatial safety.
-  Specifically, \checkedc extends C with checked pointer types, which are
-  restricted by the compiler to spatially safe uses.}
-% 
-\mzr{to ensure a program’s spatial safety by introducing additional annotations and
-  enforcing new types, checked pointer types in particular.}
-% 
-\mz{Redundancy?}
-% 
-The checked pointers are represented as system-level memory words\mzs{,~\ie} without ``fattening'' metadata, ensuring backward compatibility.
-\mz{\ie{} sounds indirect and redundant to me.}
-% 
-\mzu{This enables incremental conversion, where regions of code can be converted to~\checkedc and ensure spatial safety within those regions.}
-\mz{This sentence is incoherent. I don't understand the connection between
-  incremental conversion and regions of code. 
-  % 
-  I guess your point is that since you don't introduce extra metadata, 
-  checked and unchecked region can coexist and interact seamlessly.
-  % 
-}
-% 
+The checked pointers are represented as system-level memory words without ``fattening'' metadata, 
+and ensuring backward compatibility, i.e.,
+users are able to split code into checked and unchecked regions and incrementally convert
+C code in the unchecked region to Checked C code in the checked one.
 However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
-unconverted or unchecked code can violate guarantees provided in \mzu{Checked}
-\mz{either tt it or use lowercase?}
+unconverted or unchecked code can violate guarantees provided in checked
 regions.
 %
-We need to ensure that~\emph{code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+We need to ensure that~\emph{code executed as part of 
+unchecked regions does not violate the safety guarantees in checked regions  with the use of program partitioning mechanism \cite{rul2009towards}}.
 
@@
 %
-This mechanism is called program partitioning~\cite{rul2009towards}.
-Existing mechanisms are not suitable as they are based on process isolation and have high overhead, and are data-centric~\emph{data-centric} (\sect{subsec:background:programpart}).
-Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
-\liyi{why? Can we not say this? It seems that the whole reason people should care our work is we did it in \checkedc, but \checkedc itself is the selling product here.... }
+
+Existing such mechanisms are not suitable as they are based on process isolation and have high overhead, and are \emph{data-centric} (\sect{subsec:background:programpart}).
+%Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
+%\liyi{why? Can we not say this? It seems that the whole reason people should care our work is we did it in \checkedc, but \checkedc itself is the selling product here.... }
 But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
@@
 
-In this work, we propose a type-directed code-centric program partitioning approach.
-Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
-The checked types follow the standard Checked C typing rules.
-However, our type-system disallows any unsafe interactions between taint and untainted types.
-% 
-\mz{It sounds hard to follow. The subjects in these sentences are  ``our
-  system''. ``checked types'' and ``our type-system''.
-  %
-  For readers, the connection among checked, tainted and unchecked seems weak.
-  % 
-} 
+Here, we propose a type-directed code-centric program partitioning approach.
+Specifically, our system,~\systemname, extends \checkedc's checked and unchecked pointer types---representing safe and unsafe program pieces---with \textbf{tainted} (\taintt) types running on an isolated sandbox mechanism,
+forbids the communication between checked and unchecked type entities, and enforces the communication between checked and unchecked types
+through the uses of tainted types with additional validity checks. 
+%which can be used to mark functions and pointers that need to be isolated from the original program, while checked types follow the standard \checkedc typing rules.
+%However, our type-system disallows any unsafe interactions between taint and untainted types.
 % 
@@
 
-% 
-The developer starts by marking desired (\ie unchecked) functions and pointers used in these functions as tainted.
-Second,~\systemname partitions the given program into two partitions (\umode and~\cmode regions)  of different \mzu{privileges}:
-% 
-\mz{need a bit more description (4-5 words) before introducing the word
-  ``privilege''}
-%
+The developer starts by marking desired (\ie unchecked) functions and pointers used in functions as tainted. 
+Then, \systemname partitions the given program into two partitions (\umode and~\cmode regions)  of different privileges:
 
@@
 \end{itemize}
-% 
-\mz{                            %
-  This paragraph is unclear. 
-  % 
-  How does regions look like? Can one region nested in another?
-  %
-  If not, ``This partition contains the remaining (untainted) code and data and
-  has complete access to~\cmode region.'' creates a sense of recursive
-  definition of \cregion.
-}
-%
 
-% 
-Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
+The $\cmode$ region code is executed by compiling the code to a normal \checkedc compiler, while the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler adds necessary instrumentations to facilitate the communication between code in~$\cmode$ and~$\umode$ regions.
 
-The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
-% 
-\mzu{As functions in the~\ucregion can only access tainted types (or unchecked
-types),~\cregion functions should use tainted types to pass pointer arguments
-to~\ucregion functions.
-% 
-We avoid marshaling by allocating all tainted pointers (\ie tainted buffers)
-in~\ucregion and thus can be accessed in both partitions.}
-%
-\mz{I don't quite understand the logical consequence here. Why does allocating
-  tainted points avoids marchaling. 
-  % 
-  I see that ``allocating all tainted pointers'' makes ptrs accessable in both
-  regions, but avoiding marchaling doesn't seem to cause that.
-  %
-  The question I should probably ask is what marshaling means? 
-  % 
-  Does the procedure that copies a checked buffer to a tainted one counts as
-  marshaling?
-  % 
-}
+The combination of tainted types and privileged partitions 
+enables us to enforce isolation and provide memory safety without transforming all unchecked C code to \checkedc code,
+because unchecked types can stay in \ucregion, and \cregion code can access tainted type entities that are allocated 
+in \ucregion.
 % 
@@
 \myparagraph{\systemname Type System, Formalism and Compiler}
-We present a type system that integrates tainted types with~\checkedc and ensures the~\emph{non-crashing property},~\ie 
-a well-typed~\systemname program can never crash due to spatial safety violations of tainted pointers.
+We present a type system that integrates tainted types with~\checkedc and ensures the~\emph{non-crashing} and \emph{non-exposure} guarantees,~\ie a well-typed~\systemname program can never crash due to spatial safety violations,
+as well as \ucregion code cannot directly observe a checked pointer address.
+%no \cregion pointer addresses will be leaked in
+%\ucregion code
 We extend~\checkedc compiler to support our type system.
-We formalize our system by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing guarantee and other new features below.
-We formally prove the~\emph{non-crashing theorem} and use model-based randomized testing (CITE) to certify the simulation relation between the~\systemname semantics and the compiler.
+We formalize our system by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing and non-exposure guarantees.
+We formally prove theorems related to the two guarantees and use model-based randomized testing (CITE) to certify the simulation relation between the~\systemname semantics and its compiler formalism.
 To the best of our knowledge,~\systemname is the first C(-like) language and compiler formalism with the program partitioning mechanism.
-%
-\mz{uh}
 
 \myparagraph{Type-Directed Program Partitioning}
-We present a type-directed program partition technique to separate checked and unchecked code regions and ensure that the code in checked region~\textit{cannot crash} due to spatial safety violations.
-Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted pointers.
-Our modular design enables us to use existing sandboxed techniques to enforce memory and execution isolation.
+We present a type-directed program partition technique to separate $\cmode$ and $\umode$ code regions and ensure the above guarantees.
+%Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted pointers.
+Our modular design enables us to use existing sandboxed techniques to enforce memory and execution isolation, with the implementation of tainted pointers in the \systemname compiler.
 
@@
 %However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of \systemname.
-However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of~\systemname and leads to other attacks~\cite{hauser2019sleak}.
+However, knowing the address of~\cregion functions in~\ucregion violates the non-exposure guarantee and leads to other attacks~\cite{hauser2019sleak}.
 
 We handle this by using indirection.
-Specifically, instead of directly accessing the~\cregion callbacks, the ~\ucregion accesses them using a protected trampoline function, which will direct the execution to the appropriate callback function.
-In addition, the trampoline function itself will be referenced using an opaque index rather than its virtual address, and \mzu{we will utilize existing sandboxing techniques to implement this}.\mz{tense?}
+Specifically, instead of directly accessing the~\cregion callbacks, the ~\ucregion accesses them using a tainted-typed protected trampoline function, which directs the execution to the appropriate callback function.
+In addition, the trampoline function itself is referenced using an opaque index rather than its virtual address, implemented through
+existing sandboxing techniques.
 % 
-In our \systemname formalism, we formally verified that our formalism is
-\emph{non-exposure}, \ie, no \cregion pointer addresses will be leaked in
-\ucregion code.
+%In our \systemname formalism, we formally verified that our formalism is
+%\emph{non-exposure}, \ie, no \cregion pointer addresses will be leaked in
+%\ucregion code.
 
@@
 
-We evaluated~\systemname by partioning~\numprog{} large real-world programs to demonstrate its effectiveness.
+We evaluated~\systemname by partioning six large real-world programs to demonstrate its effectiveness.
 Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.
 
+\ignore{
 \aravind{Fix the following text after all the sections of the papers are finalized.}
@@
 be found at \url{https://github.com/plum-umd/checkedc}.
+}
 

### 2023-01-18 sandp23/evaluation.tex
Fixing eval

@@
 The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
-In our conversion, we re-introduce corrected memory bugs and then perform \systemname annotations on the relevant buggy code.
-However, this is based on the assumption that the bugs are predicted before they are discovered, which is chronologically impossible.
-Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations 
-without input on the bug locality.
 
@@
 
-As in~\fig{fig:microbenchmarks}, we incur 34\% overhead in accessing tainted pointers in~\cregion,
+As shown in~\fig{fig:microbenchmarks}, we incur 34\% overhead in accessing tainted pointers in~\cregion,
  due to additional validation checks, which require executing two compare instructions for every memory access.
@@
 \input{figures/memoryoverhead}
+The~\fig{fig:runtimeoverhead} shows the overhead of the partitioned programs.
+It is interesting to see a lot of variance in the runtime overhead.
+Our analysis shows that the runtime overhead is proportional to the execution time spent in the sandbox and the number of transitions between~\cregion and the sandbox.
+This is expected because of the overhead in sandbox execution, as shown by our micro-benchmarks experiments (\fig{fig:microbenchmarks}).
+The~\tbl{table:conversioneffort} shows the number of lines in the sandbox for each of the partitioned programs.
+For~\texttt{ProFTPD}, we used only tainted pointers with no code in the sandbox, and there were also no transitions to/from the sandbox.
+Consequently, the overhead is very less 4.2\%.
+For~\texttt{UFTPD} and~\texttt{MicroHTTPD}, we partitioned only request handlers (that are relatively small) into the sandbox.
+These handlers parse the message and return the actions that need to be performed.
+The server only invokes these handles on particular requests resulting in less number of transitions into the sandbox.
+As expected, the overhead is also very less in these cases.
+For~\texttt{TinyBigNum} and~\texttt{Parsons}, the overhead is very high because of the relatively large amount of code in sandbox region.
+In both these applications, we place the frequently used parsing functions in the sandbox resulting in a lot of sandbox transitions with most of them in a loop.
+As expected, the overhead is very high.
+The case is slightly different in~\texttt{pnm2png} and~\texttt{png2pnm} where we made the entire~\code{png} structure tainted, which resulted in dynamic checks every time any part of the~\code{png} struct is accessed, resulting in slightly higher overhead.
+In summary:
+\begin{itemize}
+  \item The runtime overhead of the partioned program directly depends on that of the target sandbox.
+  \item Run-time Overhead is proportional to the extent of annotated pointers, sandboxed code, and transitions between~\cregion and sandbox.
+  \item Tainting heavily used pointers leads to high overhead.
+\end{itemize}
 
-ProFTPD's run-time overhead is expected as our conversion was small and strictly encapsulates CVE-2010-4221.
-Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" on a 4.0K file, following which, we record overhead as described in ~\ref{experimentalsetup}.
-Run-time overhead for UFTPD is expected as the sandboxed code was less performance intensive and FTP-protocol by itself overshadows \systemname's overhead.
+%ProFTPD's run-time overhead is expected as our conversion was small and strictly encapsulates CVE-2010-4221.
+%Since UFTPD's test-suite lacks coverage for some of the \systemname changes, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" on a 4.0K file, following which, we record overhead as described in ~\ref{experimentalsetup}.
+%Run-time overhead for UFTPD is expected as the sandboxed code was less performance intensive and FTP-protocol by itself overshadows \systemname's overhead.
 
-Our evaluation of \systemname's performance results on the dataset yields the observations:
-\begin{itemize}
-  \item Run-time Overhead is proportional to the extent of 
-annotated pointers and sandboxed code.
-  \item Marshalling can be avoided if Taintedness of a pointer is propagated across its data-flow.
-  \item Marshalling might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout its data-flow. 
-\end{itemize}
+%Our evaluation of \systemname's performance results on the dataset yields the observations:
+%\begin{itemize}
+%  \item Run-time Overhead is proportional to the extent of 
+%annotated pointers and sandboxed code.
+%  \item Marshalling can be avoided if Taintedness of a pointer is propagated across its data-flow.
+%  \item Marshalling might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout its data-flow. 
+%\end{itemize}
 
@@
 The last column of the Table ~\ref{table:conversioneffort} shows the list of all spatial safety vulnerabilities in the functions that have tainted types or are isolated in the~\ucregion{}.
+We re-introduced these bugs in the annotated program and checked whether these bugs could be triggered by the corresponding crashing input or exploit (if available).
+We also manually verified whether the bug can be triggered or prevented by~\systemname{}.
 As expected,~\emph{all vulnerabilities} are prevented by~\systemname{}.

### 2023-01-18 sandp23/tables/datasettable.tex
Fixing eval

@@
 {
-\footnotesize
+\scriptsize
 \begin{table}[]
@@
 \rowcolor{black!15} 3           & UFTPD            & UDP based FTP Server          & 3 K                                                                               \\ 
-4           & LibPNG          & Program to convert between png and pnm & 76 K                                                                               \\ 
+4           & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}LibPNG\\ (png2pnm and pnm2png)\end{tabular}}           & Program to convert between png and pnm & 76 K                                                                               \\
 \rowcolor{black!15}5           & TinyBigNum          & Multiple-precision integer implementation & 1.6 K                                                                               \\ 

### 2023-01-18 sandp23/evaluation.tex
Adding all

@@
 
-\input{figures/memoryoverhead}
+%\input{figures/memoryoverhead}
 The~\fig{fig:runtimeoverhead} shows the overhead of the partitioned programs.
@@
 \end{itemize}
+
+\subsubsection{Memory Overhead}
+All programs have a constant memory overhead ($\sim$81 KB) mainly for sandbox and a few variables related to creating sandbox and other helper functions.
+However, similar to the original~\checkedc,~\systemname{} itself doesn't add any memory overhead.
+This is expected because, similar to~\checkedc types, ~\code{taint}ed pointers also do not have metadata and have the same runtime representation as regular pointers.
 

### 2023-01-18 sandp23/figures/runtimechart.tex
fix figure 7,8

@@
     ybar,  
-    enlargelimits=0.04,  
+    enlargelimits=0.12,  
     ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  
@@
     xtick=data,  
-    bar width=10pt,
-    width=7cm,
+    bar width=8pt,
+    width=6cm,
     x tick label/.append style={font=\tiny, rotate=30},
@@
     x label style={at={(axis description cs:0.5,-0.1)},anchor=north, font=\footnotesize},
-    xlabel={tests},
+    xlabel={},
     ylabel style={font=\footnotesize},
@@
     x label style={at={(axis description cs:0.5,-0.1)},anchor=north, font=\footnotesize},
-    xlabel={Programs},
+    xlabel={},
     ylabel style={font=\footnotesize},

### 2023-01-18 sandp23/figures/runtimechart.tex
fix figure 7,8

@@
     ybar,  
-    enlargelimits=0.12,  
+    enlargelimits=0.03,
+    enlarge x limits=0.12,
     ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  
@@
     ybar,  
-    enlargelimits=0.12,  
+    enlargelimits=0.01,  
+    enlarge x limits=0.12,
     ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  

### 2023-01-18 sandp23/tables/datasettable.tex
move tables around.

@@
-{
-\scriptsize
 \begin{table}[]
+{\scriptsize
 \begin{tabular}{c|c|c|r}
@@
 \end{tabular}
+}
 \caption{Evaluation Dataset.}
 \label{table:dataset}
+
+{\tiny
+\begin{tabular}{c|c|c|c|r|c}
+\toprule
+\textbf{Program} & \textbf{\begin{tabular}[c]{@{}c@{}}Partition\\ Methodology\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Pointers\\ Annotated\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Lines in\\ \ucregion \\ (Sandbox)\end{tabular}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Time\\ Taken\\ (hours)\end{tabular}}} & \textbf{\begin{tabular}[c]{@{}c@{}}CVEs\\ Prevented\end{tabular}} \\
+\midrule
+\rowcolor{black!15} ProFTPD          &     $TP$                                                                     &       6                                                                &                                                                     N/A &                                                                                      1        & CVE-2010-4221 (\vulprevented{})                                                           \\ %\hline
+png2pnm          & \multirow{2}{*}{$TP$}                                                        & \multirow{2}{*}{248}                                                     & \multirow{2}{*}{N/A}                                                    & \multirow{2}{*}{}                                                                            & \multirow{2}{*}{CVE-2018-144550 (\vulprevented{})}                                             \\ %\cline{1-1}
+pnm2png          &                                                                          &                                                                       &                                                                      &                                                     8                                         &                                                                   \\ %\hline
+\rowcolor{black!15} MicroHTTPD       &     $TP$                                                                     &          139                                                             &                                                                     450 &                                                                                       3       &                    N/A                                               \\ %\hline
+UFTPD             &        $T_{Pr}$                                                                  &  146                                                                     &                                                                     90 &                                                           3                                   & \begin{tabular}[c]{@{}c@{}}CVE-2020-14149 (\vulprevented{}) \\ CVE-2020-5204 (\vulisolated{})\end{tabular}             \\ %\hline
+\rowcolor{black!15} TinyBigNum       &      $CTP$                                                                    &   69                                                                    &                                                                     30 &                                                                                        2      &                             N/A                                      \\ %\hline
+Parsons          &     $CTP$                                                                     &      364                                                                 &                                                                     800 &                                                      5                                        &                                                  N/A                 \\
+\bottomrule
+\end{tabular}
+}
+\caption{Summary of Conversion Efforts \& Security Impact.}
+\label{table:conversioneffort}
 \end{table}
-}

### 2023-01-18 sandp23/evaluation.tex

@@
 \subsection{Dataset}
-\input{tables/datasettable}
 Network-facing programs such as servers directly interact with external input, often process complex data types, and are more susceptible to security issues.
@@
 
-\input{tables/conversioneffort} 
+\input{tables/datasettable}
+%\input{tables/conversioneffort} 
 \subsubsection{Conversion Methodology}

### 2023-01-18 sandp23/implementation.tex

@@
 \subsection{\sourcerewriter} \label{subsec:checkmate}
-The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (3028 SLoc).
+The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (3K SLoc).
 However, we use a small OCAML program (680 SLoc) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
@@
 The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
-We implemented the compiler as a clang library (1812K SLoc) on top of the existing~\checkedc compiler.
+We implemented the compiler as a clang library (1.8K SLoc) on top of the existing~\checkedc compiler.
 Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.

### 2023-01-18 sandp23/evaluation.tex
update.

@@
 %\input{figures/memoryoverhead}
-The~\fig{fig:runtimeoverhead} shows the overhead of the partitioned programs.
-It is interesting to see a lot of variance in the runtime overhead.
-Our analysis shows that the runtime overhead is proportional to the execution time spent in the sandbox and the number of transitions between~\cregion and the sandbox.
-This is expected because of the overhead in sandbox execution, as shown by our micro-benchmarks experiments (\fig{fig:microbenchmarks}).
-The~\tbl{table:conversioneffort} shows the number of lines in the sandbox for each of the partitioned programs.
-For~\texttt{ProFTPD}, we used only tainted pointers with no code in the sandbox, and there were also no transitions to/from the sandbox.
-Consequently, the overhead is very less 4.2\%.
-For~\texttt{UFTPD} and~\texttt{MicroHTTPD}, we partitioned only request handlers (that are relatively small) into the sandbox.
-These handlers parse the message and return the actions that need to be performed.
-The server only invokes these handles on particular requests resulting in less number of transitions into the sandbox.
-As expected, the overhead is also very less in these cases.
-For~\texttt{TinyBigNum} and~\texttt{Parsons}, the overhead is very high because of the relatively large amount of code in sandbox region.
-In both these applications, we place the frequently used parsing functions in the sandbox resulting in a lot of sandbox transitions with most of them in a loop.
-As expected, the overhead is very high.
-The case is slightly different in~\texttt{pnm2png} and~\texttt{png2pnm} where we made the entire~\code{png} structure tainted, which resulted in dynamic checks every time any part of the~\code{png} struct is accessed, resulting in slightly higher overhead.
+\Cref{fig:runtimeoverhead} shows the overhead of the partitioned programs.
+%It is interesting to see a lot of variance in the runtime overhead.
+The runtime overhead is proportional to the execution time 
+spent in the sandbox and the number of transitions between~\cregion and the sandbox,
+which coincides with the sandbox execution overhead observation in the micro-benchmarks experiments (\Cref{fig:microbenchmarks}).
+\Cref{table:conversioneffort} shows the numbers of lines in the sandbox for our partitioned programs.
+For~\texttt{ProFTPD}, we used only tainted pointers without code in the sandbox and transitions to/from the sandbox.
+Consequently, the overhead is less than 4.2\%.
+For~\texttt{UFTPD} and~\texttt{MicroHTTPD}, we sandboxed only request handlers (relatively small)
+that parse messages and return the actions that need to be performed.
+The server only invokes these handlers on particular requests resulting in less transitions interacting with the sandbox.
+As expected, the overhead is also little in these cases.
+For~\texttt{TinyBigNum} and~\texttt{Parsons}, the overhead is high because of the relatively large amount of code in sandbox region.
+In both applications, we place the frequently used parsing functions in the sandbox 
+resulting in a lot of sandbox transitions with most of them in a loop.
+The case is slightly different in~\texttt{pnm2png} and~\texttt{png2pnm},
+where we made the entire~\code{png} structure tainted, 
+which resulted in dynamic checks every time when the~\code{png} struct is accessed.
+The result indicates a higher overhead than the non-sandboxing cases and less overhead than the heavy sandboxing cases.
 In summary:
 \begin{itemize}
-  \item The runtime overhead of the partioned program directly depends on that of the target sandbox.
-  \item Run-time Overhead is proportional to the extent of annotated pointers, sandboxed code, and transitions between~\cregion and sandbox.
-  \item Tainting heavily used pointers leads to high overhead.
+  \item Runtime overhead is proportional to the extent of annotated pointers, sandboxed code, and transitions between~\cregion and sandbox.
+  \item Using tainted pointers with compiler generated dynamic checks is preferred than placing code in sandbox.
+  \item Different C to \systemname conversion methods affect the overhead performance, which is our major future work.
 \end{itemize}
@@
 All programs have a constant memory overhead ($\sim$81 KB) mainly for sandbox and a few variables related to creating sandbox and other helper functions.
-However, similar to the original~\checkedc,~\systemname{} itself doesn't add any memory overhead.
-This is expected because, similar to~\checkedc types, ~\code{taint}ed pointers also do not have metadata and have the same runtime representation as regular pointers.
+However, similar to the original~\checkedc,~\systemname{} itself does not add any memory overhead,
+because the compilation of \code{tainted} pointers do not come with any metadata.
+%This is expected because, similar to~\checkedc types, ~\code{taint}ed pointers also do not have metadata and have the same runtime representation as regular pointers.
 

### 2023-01-18 sandp23/background.tex
Fixing intro

@@
 Failed checks throw an exception.
+Oftentimes, inserted checks can be optimized away by LLVM resulting in almost no runtime overhead~\cite{duanrefactoring}.
 
-\liyi{why we need the following? I suggest we can cut. }
-Oftentimes, inserted checks can be optimized away by LLVM.
-For instance, Dual~\etal~\cite{duanrefactoring} found essentially no
-overhead when running Checked C-converted portions of the FreeBSD kernel.
+%\liyi{why we need the following? I suggest we can cut. }
+%Oftentimes, inserted checks can be optimized away by LLVM~\cite{duanrefactoring}.
+%For instance, Dual~\etal~\cite{duanrefactoring} found essentially no
+%overhead when running Checked C-converted portions of the FreeBSD kernel.
 %Consider the above code to be enclosed in another condition, such as,~\inlinecode{if (n<u) if (n>0) a[n-1] =}.
@@
 \label{subsec:nosafetyagsintuncheckedcode}
-The spatial safety guarantee~\checkedc provides applies to completely converted programs,~\ie programs that uses only checked types.
-The backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
-But, these unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+The spatial safety guarantee~\checkedc provides applies to completely converted programs,~\ie programs that uses~\emph{only} checked types and no regular pointer types.
+A partially annotated program can still enjoy spatial safety but only on those code regions using only Checked pointers (\ie checked or safe regions).
+For instance, in the below example, there can be no spatial safety violations in the function~\code{func} as it uses only checked pointers.
+However, the other unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+
+%The backward compatibility of~\checkedc helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (\ie checked or safe regions).
+%But, these unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
 \lc{I'm confused, does the partially checked program "enjoy spatial memory safety" or not?}
@@
 func(|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5)); 
-...
-char ptr[16];
-...
-len <- derived from user input
-...
-|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
 \end{minted}
@@
 This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
+Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions (\eg buffer overflow) in unchecked regions can take down the complete program despite having checked regions.
 We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.

### 2023-01-18 sandp23/introduction.tex
Adding all
Fixing intro

@@
 open-source extension to C,
-to ensure a program’s spatial safety by introducing additional annotations and
-enforcing new types, checked pointer types in particular.
+to ensure a program’s spatial safety by introducing new pointer types,~\ie checked pointer types.
 %
-The checked pointers are represented as system-level memory words without ``fattening'' metadata, 
-and ensuring backward compatibility, i.e.,
-users are able to split code into checked and unchecked regions and incrementally convert
-C code in the unchecked region to Checked C code in the checked one.
+The checked pointers are represented as system-level memory words without ``fattening'' metadata~\cite{duck2016heap}, 
+and ensuring backward compatibility,~\ie developers can use checked and regular (unchecked) pointers with the same program.
+%users are able to split code into checked and unchecked regions and incrementally convert
+%C code in the unchecked region to Checked C code in the checked one.
 However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
@@
 
-The $\cmode$ region code is executed by compiling the code to a normal \checkedc compiler, while the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler adds necessary instrumentations to facilitate the communication between code in~$\cmode$ and~$\umode$ regions.
+The $\cmode$ region code is executed as a regular program, while the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler adds necessary instrumentations to facilitate the communication between code in~$\cmode$ and~$\umode$ regions.
 
@@
 % 
-Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
-Our compiler avoids these attacks by ensuring, through dynamic checks that tainted pointers validly point to \ucregion address space.
+Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks~\cite{rajani2016access, machiry2017boomerang},~\eg by using a valid~\cregion address in a tainted pointer.
+Our compiler avoids these attacks by ensuring using dynamic checks that tainted pointers validly point to~\ucregion address space.
 Such checks are statically generated by our compiler.
@@
 \myparagraph{\systemname Type System, Formalism and Compiler}
-We present a type system that integrates tainted types with~\checkedc and ensures the~\emph{non-crashing} and \emph{non-exposure} guarantees,~\ie a well-typed~\systemname program can never crash due to spatial safety violations,
+We present a type system that integrates tainted types with~\checkedc and provides additional guarentees,~\ie the~\emph{non-crashing} and \emph{non-exposure} guarantees,~\ie a well-typed~\systemname program can never crash due to spatial safety violations,
 as well as \ucregion code cannot directly observe a checked pointer address.
@@
 
-We evaluated~\systemname by partioning six large real-world programs to demonstrate its effectiveness.
+We evaluated~\systemname by partioning seven large real-world programs to demonstrate its effectiveness.
 Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.

### 2023-01-18 sandp23/evaluation.tex
update.
Adding no-op sandbox experiment

@@
 that parse messages and return the actions that need to be performed.
-The server only invokes these handlers on particular requests resulting in less transitions interacting with the sandbox.
-As expected, the overhead is also little in these cases.
+The server only invokes these handlers on particular requests resulting in less transitions with the sandbox.
+As expected, the overhead is also less in these cases.
 For~\texttt{TinyBigNum} and~\texttt{Parsons}, the overhead is high because of the relatively large amount of code in sandbox region.
@@
 which resulted in dynamic checks every time when the~\code{png} struct is accessed.
-The result indicates a higher overhead than the non-sandboxing cases and less overhead than the heavy sandboxing cases.
-In summary:
-\begin{itemize}
-  \item Runtime overhead is proportional to the extent of annotated pointers, sandboxed code, and transitions between~\cregion and sandbox.
-  \item Using tainted pointers with compiler generated dynamic checks is preferred than placing code in sandbox.
-  \item Different C to \systemname conversion methods affect the overhead performance, which is our major future work.
-\end{itemize}
+In summary, our results indicate that the runtime overhead largely depends on the sandbox.
+
+\noindent\textbf{Overhead of only~\systemname{}.}
+To verify the impact of the sandbox on the overall program runtime, we perform a NO-OP sandbox experiment.
+Our goal is to measure the runtime overhead introduced ONLY by~\systemname{}.
+We perform this experiment by skipping sandboxing.
+Specifically, we run the completely annotated program as a regular application without any partitioning.
+However, we compile the annotated program with~\systemname compiler, which will add the relevant runtime checks (\sect{subsec:compilerimple}).
+We modify the instrumentation on~\code{taint}ed pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
+
+The~\fig{XXX} shows the runtime overhead. The overhead is negligible in all cases.
+This shows that~\systemname{} fundamentally has the least overhead.
+\aravind{Arun, we should do this experiment. This should be fairly easy.}
+
+%The result indicates a higher overhead than the non-sandboxing cases but much less overhead than the heavy sandboxing cases.
+%In summary:
+%\begin{itemize}
+%  \item Runtime overhead is proportional to the extent of annotated pointers, sandboxed code, and transitions between~\cregion and sandbox.
+%  \item Using tainted pointers with compiler generated dynamic checks is preferred than placing code in sandbox.
+%  \item Different C to \systemname conversion methods affect the overhead performance, which is our major future work.
+%\end{itemize}
 

### 2023-01-18 sandp23/background.tex
update section 1 and 2.

@@
 %This section describes \checkedc{} and new features \systemname{} provides.
-This section presents a brief background of~\checkedc and motivation for~\systemname{}.
+Here, we brief \checkedc and the motivation for~\systemname{}.
 
@@
 Development Project (SSDP). Details can be found in a prior
-overview~\cite{Elliott2018} or the formalism~\cite{li22checkedc}.
-%
-
-\liyi{Do we need to spend more than half page introducing checked-c? OR just introduce it briefly like what we had in the formalism paper? }
+overview~\cite{Elliott2018} and the formalism~\cite{li22checkedc}.
 
@@
 The bounds can be declared for $\ntarr$ as well, but the memory range can extend further to the right,
-until a~\code{NULL} terminator is reached (\ie the~\code{NULL} is not within the bounds).
-
-\liyi{can we reuse the old text for the spatial safety? }
+until a~\code{NULL} terminator is reached (\ie \code{NULL} is not within the bounds).
 
@@
 \myparagraph{Ensuring Spatial Memory Safety}
-The \checkedc compiler will instrument loads and stores of checked
-pointers to confirm the pointer is non-null, and additionally for $\arr$ and $\ntarr$ pointers the access is within
-the specified bounds.
-For instance, in the code \inlinecode{if (n>0) a[n-1] =} $...$ the write
+The \checkedc compiler instruments loads and stores of checked
+pointers to confirm the pointer is non-null, and additionally the access to $\arr$ and $\ntarr$ pointers is within
+their specified bounds.
+For example, in the code \inlinecode{if (n>0) a[n-1] =} $...$ the write
 is via address $\alpha = \inlinecode{a + sizeof(int)}\!\times\!\code{(n-1)}$. 
 If the bounds of \code{a} are \code{count(u)}, the
-inserted check will confirm that prior to dereference
+inserted check confirms
 $\inlinecode{a} \leq \alpha < \inlinecode{a + sizeof(int)} \!\times\!
-\inlinecode{u}$.
+\inlinecode{u}$ prior to dereference.
 Failed checks throw an exception.
@@
 \myparagraph{Backward Compatibility}
-\checkedc is backward compatible with legacy C in the sense that all legacy code will type-check and compile.
+\checkedc is backward compatible with legacy C as all legacy code will type-check and compile.
 However, the compiler adds the aforementioned spatial safety checks to only checked pointers.
@@
 Specifically, only code that appears in \emph{checked
-  code regions}, which we call \emph{checked code}, is guaranteed to be spatially safe.
-Checked regions can be designated at the level of files, functions, or individual code blocks using the
-\code{checked} keyword.\footnote{You can also designate~\emph{unchecked} regions  within checked ones.}
- Within checked regions, both legacy pointers and
+  code regions} (\cregion), is guaranteed to be spatially safe.
+$\cmode$ regions can be designated at the level of files, functions, or individual code blocks using the
+\code{checked} keyword.\footnote{You can also designate~\emph{unchecked} regions (\ucregion) within checked ones.}
+ Within $\cmode$ regions, both legacy pointers and
 certain unsafe idioms (\eg \emph{variadic} function calls) are disallowed.
-
-\liyi{If we are not going to talk about converting C to \checkedc, can we just move this to future work? }
 
@@
 
-\liyi{Does the program too small? Will reviewer complain about the program size? }
-\subsection{No Safety Against Unchecked Code}
+\subsection{No Safety Against $\umode$ Regions}
 \label{subsec:nosafetyagsintuncheckedcode}
-The spatial safety guarantee~\checkedc provides applies to completely converted programs,~\ie programs that uses~\emph{only} checked types and no regular pointer types.
-A partially annotated program can still enjoy spatial safety but only on those code regions using only Checked pointers (\ie checked or safe regions).
-For instance, in the below example, there can be no spatial safety violations in the function~\code{func} as it uses only checked pointers.
+\checkedc provides spatial safety guarantees for completely converted programs,~\ie programs that uses~\emph{only} checked types and no regular pointer types.
+A partially annotated program can still enjoy spatial safety only if checked pointers do not communicate with any unchecked ones.
+For instance, in the example below, there are no spatial safety violations in the function~\code{func} as it uses only checked pointers.
 However, the other unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
@@
 \lc{I'm confused, does the partially checked program "enjoy spatial memory safety" or not?}
-Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
+Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in $\cmode$ regions but rather states that $\cmode$ regions~\emph{cannot be blamed for any spatial safety violations}.
 Consider the following example:
 \begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
-// Checked code
+// c region code
 int func(array_ptr<char> p : count(5)) {
@@
 }
-// unchecked code
+// u region code
 ...
@@
 \end{minted}
-Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function (Line 9) with a buffer of 2 elements.
-This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions (\eg buffer overflow) in unchecked regions can take down the complete program despite having checked regions.
-We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
+Here, the $\cmode$ region function~\inlinecode{func} expects a pointer to a buffer of five elements, but the $\umode$ region code
+invokes the function (Line 9) with a buffer of 2 elements.
+This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the $\cmode$ region, but of course, the blame or the root cause is in the $\umode$ region (\textcolor{red}{\faBug}).
+Furthermore, since $\cmode$ and $\umode$ regions execute in the same address space, spatial memory corruptions (\eg buffer overflow) in $\umode$ regions can take down the complete program despite having $\cmode$ regions.
+We need~\emph{an isolation mechanism to ensure that code executed as part of $\umode$  regions does not violate the safety guarantees in $\cmode$ regions}.
 

### 2023-01-19 sandp23/figures/runtimechart.tex
update figures.

--- 2023-01-18 sandp23/figures/runtimechart.tex
@@
     ybar,  
-    enlargelimits=0.12,  
+    %ymode=log,
+    enlargelimits=0.03,
+    enlarge x limits=0.12,
     ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  
+     axis x line*=bottom,
+     axis y line*=left,
     %xlabel={Micro-Benchmark Tests},  
@@
 \end{tikzpicture} 
+\vspace*{-1.5em}
 \caption{\systemname Micro-Benchmarks.}
@@
 [  
-    ybar,  
-    enlargelimits=0.12,  
+    ybar = 0.8,  
+    enlargelimits=0.01,  
+    enlarge x limits=0.12,
     ylabel={Runtime Overhead (\%)}, % the ylabel must precede a # symbol.  
+     axis x line*=bottom,
+     axis y line*=left,
     %xlabel={Programs},  
@@
     bar width=8pt,
-    width=6cm,
+    %legend style={font=\tiny},
+    legend style={font=\tiny, at={(0.7,1.0)}},
+    width=7cm,
+    height=5cm,
     x tick label/.append style={font=\tiny, rotate=30},
@@
     axis background/.style={fill=gray!10},
-    x label style={at={(axis description cs:0.5,-0.1)},anchor=north, font=\footnotesize},
+    x label style={at={(axis description cs:0.5,-0.1)},anchor=north, font=\tiny},
     xlabel={},
     ylabel style={font=\footnotesize},
+    every node near coord/.append style={font=\tiny},
      nodes near coords, % this command is used to mention the y-axis points on the top of the particular bar.  
@@
     ]  
-\addplot coordinates {(ProFTPD,4.2) (TinyBigNum,121.7) (UFTPD, 2.7) (MicroHTTPD,3.8) (png2pnm, 11.47) (pnm2png, 46.5) (Parsons, 267)};
+\addplot coordinates {(ProFTPD,4.2) (TinyBigNum,121.7) (UFTPD, 2.7) (MicroHTTPD,3.8) (png2pnm, 11.4) (pnm2png, 46.5) (Parsons, 267)};
+
+\addplot coordinates {(ProFTPD,4) (TinyBigNum,11) (UFTPD, 2.7) (MicroHTTPD,3.8) (png2pnm, 11.4) (pnm2png, 10) (Parsons, 2)};
+
+\legend{CheckedCBox, CheckedCBox (w NO-OP Sandbox)}
 
@@
 \end{tikzpicture} 
+\vspace*{-1.5em}
 \caption{Runtime Overhead of Partioned Programs.}

### 2023-01-19 sandp23/formal.tex
update section 3.

@@
 $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean $\tptr{\tarray{0}{b}{\tau}}{\cmode}$,
-so the above examples could be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
+so the above examples could be rewritten as $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
 $\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
@@
 It prevents unsafe types from being constructed.
-  Consider the type ~\code{_t_Array_ptr<_Ptr<int>>} which describes a tainted
-  array of checked pointers.
-  This is \emph{illegal} in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region;
-  thus, it is rejected by the well-formedness. 
+  Consider the type ~\code{_t_Array_ptr<_Ptr<int>>}, which describes a tainted array of checked pointers.
+  This is not well-formed in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the tainted ($\tmode$) array is used. 
   Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
@@
  static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \cite{li22checkedc}}, the \texttt{strlen} operation ($\estrlen{x}$),
-memory allocation ($\emalloc{\xi}{\omega}$), 
+memory allocations ($\emalloc{\xi}{\omega}$), 
 function calls ($\ecall{e}{\overline{e}}$),
@@
 For example, a dynamic bounds cast
-  { \footnotesize
-    \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
+  { \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
 is formalized as 
-{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \lang. 
+{$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \lang. 
 %
@@
 We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
-For instance, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer.
-The heap allocation expression $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions (\cregion or \ucregion).
+For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
+The heap allocation $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions, $\cmode$ mode pointer in \cregion or $\umode$ and $\tmode$ mode pointers in \ucregion.
 We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
 The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit code regions.
-To guarantee the non-exposure property, we extend the \checkedc syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$,
-where $\overline{x}$ represents all variables that are allowed to communicate between instructions outside and inside of the block $e$, and cannot contain checked pointers.
+To guarantee the non-exposure property, we extend the \checkedc syntax to include $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$ blocks,
+where $\overline{x}$ represents all variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain checked pointers.
 $\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
@@
               {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
-
-\\\\
+  \end{mathpar}
+
+  \begin{mathpar}
     \inferrule[T-Def]
@@
     \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-
+  \end{mathpar}
+
+  \begin{mathpar}
     \inferrule[T-Checked]
@@
     \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
+  \end{mathpar}
+
+  \begin{mathpar}
 \inferrule[T-Fun]
@@
     {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
-
+  \end{mathpar}
+  \begin{mathpar}
     \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
@@
 transition relation with the judgment $ (\varphi,\heap,e)
-\longrightarrow_m (\varphi',\heap',r)$, as partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{S-}$X$).
+\longrightarrow_m (\varphi',\heap',r)$, as shown in \Cref{fig:c-context}.
  Here, $\varphi$ is a
 \emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
-$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ heaps), each of which
+$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ heap regions), each of which
 maps addresses (integer literals) to values $\evalue{n}{\tau}$.
@@
   relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
-whose rules are given along with type rules in Fig.~\ref{fig:type-system-1}, discussed
+whose rules are given along with type rules in Fig.~\ref{fig:type-system-1} (labeled as \rulelab{S-}$X$), discussed
 shortly.
@@
 We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
-This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can perform such operations in $\cmode$ and $\umode$ regions, respectively.
+This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can only perform such operations in $\cmode$ and $\umode$ regions, respectively.
 %Pointer modes are also useful in determining if a nested pointer has a valid type.
@@
 In addition to the no $\enull$ check in $\cmode$ mode pointer dereference,
-any dynamic heap access of a tainted pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to that the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
+any dynamic heap access of a tainted ($\tmode$) pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to that the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
 
@@
 with its type and semantic rules given in \Cref{fig:type-system-1}.
-In this context switching, to guarantee the checked pointer non-exposure property, 
-checked pointers are not allowed to pass through, which is guaranteed by the predicates 
+In this context switching, to guarantee the checked ($\cmode$) pointer non-exposure property, 
+checked pointers are not allowed to go cross different regions, which is guaranteed by the predicates 
 $\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$,
@@
 the result of the application will
-be of type $\tau[\overline{e'} / \overline{x}]$ 
-% 
+be of type $\tau[\overline{e'} / \overline{x}]$;
 if for each pair of \(\tau'\) and \(\tau''\) in \(\overline{\tau'}\) and
@@
 Similar to \heap, the global function store $\Xi$ is also partitioned into
-two parts ($\cmode$ and $\umode$ stores), each of which
+two parts ($\cmode$ and $\umode$ store regions), each of which
 maps addresses (integer literals) to the function data described above.
@@
 
-Using the above theorem, we first show the non-exposure theorem,
-where code in $\umode$ region cannot observe a valid checked pointer address.
+Using the above theorems, we first show the non-exposure theorem,
+where code in $\umode$ region cannot observe a valid checked ($\cmode$) pointer address.
 

### 2023-01-19 sandp23/introduction.tex

@@
 open-source extension to C,
-to ensure a program’s spatial safety by introducing new pointer types,~\ie checked pointer types.
+to ensure a program’s spatial safety by introducing new pointer types,~\ie checked ($\cmode$) pointer types.
 %
 The checked pointers are represented as system-level memory words without ``fattening'' metadata~\cite{duck2016heap}, 
-and ensuring backward compatibility,~\ie developers can use checked and regular (unchecked) pointers with the same program.
+and ensuring backward compatibility,~\ie developers can use checked and regular (unchecked $\umode$) pointers within the same program.
 %users are able to split code into checked and unchecked regions and incrementally convert
@@
 However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
-unconverted or unchecked code can violate guarantees provided in checked
+unconverted or unchecked ($\umode$) code can violate guarantees provided in $\cmode$
 regions.
@@
 We need to ensure that~\emph{code executed as part of 
-unchecked regions does not violate the safety guarantees in checked regions  with the use of program partitioning mechanism \cite{rul2009towards}}.
+unchecked ($\umode$) regions does not lead to the safety violations in checked ($\cmode$) regions  with the use of program partitioning mechanism \cite{rul2009towards}}.
 
@@
 %\liyi{why? Can we not say this? It seems that the whole reason people should care our work is we did it in \checkedc, but \checkedc itself is the selling product here.... }
-But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
-We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
+But in our case, we want a low-overhead code-centric partitioning, where the $\umode$ region code (or functions) should be isolated (or partitioned) from $\cmode$ one.
+We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing $\cmode$ region code can maintain spatial safety.
 
@@
 
-The developer starts by marking desired (\ie unchecked) functions and pointers used in functions as tainted. 
+The developer starts by marking desired (\ie unchecked, $\umode$) functions and pointers used in functions as tainted. 
 Then, \systemname partitions the given program into two partitions (\umode and~\cmode regions) of different privileges:
@@
 \begin{itemize}
-\item~\umode \emph{region} or~\emph{tainted region} (low privilege tainted region): This partition contains tainted types (\ie functions and pointers) and can only access tainted pointers.
+\item~\umode \emph{region} (low privilege tainted region, extended from the unchecked region in \checkedc): this partition contains tainted types (\ie functions and pointers) and can only access tainted and unchecked pointers.
 \item~\cmode \emph{region} or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cmode region.
@@
 
-The $\cmode$ region code is executed as a regular program, while the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), and our compiler adds necessary instrumentations to facilitate the communication between code in~$\cmode$ and~$\umode$ regions.
+The $\cmode$ region code is executed as a regular program, while the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), with additional instrumentations to facilitate the communication between code in~$\cmode$ and~$\umode$ regions.
 
 The combination of tainted types and privileged partitions
-enables us to enforce isolation and provide memory safety without transforming all unchecked C code to \checkedc code,
+enforces isolation and provides memory safety without transforming all unchecked C code to \checkedc code,
 because unchecked types can stay in \ucregion, and \cregion code can access tainted type entities that are allocated 
@@
 \myparagraph{\systemname Type System, Formalism and Compiler}
-We present a type system that integrates tainted types with~\checkedc and provides additional guarentees,~\ie the~\emph{non-crashing} and \emph{non-exposure} guarantees,~\ie a well-typed~\systemname program can never crash due to spatial safety violations,
+We present a type system that integrates tainted types with~\checkedc and provides additional guarentees---the~\emph{non-crashing} and \emph{non-exposure} guarantees,~\ie a well-typed~\systemname program can never crash due to spatial safety violations,
 as well as \ucregion code cannot directly observe a checked pointer address.
@@
 %\ucregion code
-We extend~\checkedc compiler to support our type system.
-We formalize our system by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing and non-exposure guarantees.
+We extend the \checkedc compiler to support the type system and
+formalize it by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing and non-exposure guarantees.
 We formally prove theorems related to the two guarantees and use model-based randomized testing (CITE) to certify the simulation relation between the~\systemname semantics and its compiler formalism.
@@
 %Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted pointers.
-Our modular design enables us to use existing sandboxed techniques to enforce memory and execution isolation, with the implementation of tainted pointers in the \systemname compiler.
+Our modular design enables us to use existing sandbox techniques to enforce memory and execution isolation, with the implementation of tainted pointers in the \systemname compiler.
 
 \myparagraph{Supporting callbacks to~\cregion with no Checked Pointer Exposure}
-Although we disallow access to~\cregion from~\ucregion, there can be cases where such access is needed.
+Although we disallow access to~\cregion from~\ucregion directly, there can be cases where such access is needed.
 Specifically, when~\cregion wants to provide access to certain shared checked data to~\ucregion.

### 2023-01-20 sandp23/evaluation.tex
update.

@@
 We also selected a few standalone programs suggested by the ~\checkedc team~\cite{benchmarkcc}, which are good candidates to evaluate modifications to~\checkedc.
-The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
+\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
 
@@
 %\input{figures/memoryoverhead}
-\Cref{fig:runtimeoverhead} shows the overhead of the partitioned programs.
+The first set of bars in~\Cref{fig:runtimeoverhead} shows the overhead of the partitioned programs.
 %It is interesting to see a lot of variance in the runtime overhead.
@@
 
-The~\fig{XXX} shows the runtime overhead. The overhead is negligible in all cases.
+The second set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
 This shows that~\systemname{} fundamentally has the least overhead.
-\aravind{Arun, we should do this experiment. This should be fairly easy.}
+%\aravind{Arun, we should do this experiment. This should be fairly easy.}
 

### 2023-01-20 sandp23/evaluation.tex
update.

@@
 However, we compile the annotated program with~\systemname compiler, which will add the relevant runtime checks (\sect{subsec:compilerimple}).
-We modify the instrumentation on~\code{taint}ed pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
+We modify the instrumentation on~\code{tainted} pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
 
-The second set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
+The marked \textcolor{red}{red} set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
 This shows that~\systemname{} fundamentally has the least overhead.

### 2023-01-20 sandp23/evaluation.tex
Adding experiemnt results

--- 2023-01-20 sandp23/evaluation.tex
@@
 We also selected a few standalone programs suggested by the ~\checkedc team~\cite{benchmarkcc}, which are good candidates to evaluate modifications to~\checkedc.
-\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
+The~\tbl{table:dataset} shows the program selected as part of our evaluation dataset.
 
@@
 We modify the instrumentation on~\code{taint}ed pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
-
+~\paragraph{Results:}
+Our evaluation of ~\systemname{} with NO-OP Sandbox was performed on two programs with the largest overheads in our dataset.
+With parsons, we see an average run-time overhead of 1.94\% as is expected as parsons (($T_{Pr}$)) gave us an overhead of only 9\%.
+With TinyBigNum, we see an overhead of 54.7\%, and is expected referring to the figure ~\ref{fig:microbenchmarks}, as TinyBigNum is pointer intensive with pointer operations inside nested loops.
 The second set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
 This shows that~\systemname{} fundamentally has the least overhead.
-%\aravind{Arun, we should do this experiment. This should be fairly easy.}
 

### 2023-01-20 sandp23/evaluation.tex

@@
 
+We modify the instrumentation on~\code{tainted} pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
+
+The marked \textcolor{red}{red} set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
+This shows that~\systemname{} fundamentally has the least overhead.
+
 %The result indicates a higher overhead than the non-sandboxing cases but much less overhead than the heavy sandboxing cases.

### 2023-01-20 sandp23/evaluation.tex
rearrange my prev commit
update.

--- 2023-01-20 sandp23/evaluation.tex
@@
 We modify the instrumentation on~\code{taint}ed pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
-~\paragraph{Results:}
+
 Our evaluation of ~\systemname{} with NO-OP Sandbox was performed on two programs with the largest overheads in our dataset.
@@
 With TinyBigNum, we see an overhead of 54.7\%, and is expected referring to the figure ~\ref{fig:microbenchmarks}, as TinyBigNum is pointer intensive with pointer operations inside nested loops.
-The second set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
 This shows that~\systemname{} fundamentally has the least overhead.
+%The second set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
+%This shows that~\systemname{} fundamentally has the least overhead.
+%We modify the instrumentation on~\code{tainted} pointers to check for a valid pointer (instead of within sandbox bounds) -- this will add the same amount of checks as in sandboxing case, but the comparison values will be different.
+%The marked \textcolor{red}{red} set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.
+%This shows that~\systemname{} fundamentally has the least overhead.
 

### 2023-01-20 sandp23/introduction.tex
update.

@@
 formalize it by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing and non-exposure guarantees.
-We formally prove theorems related to the two guarantees and use model-based randomized testing (CITE) to certify the simulation relation between the~\systemname semantics and its compiler formalism.
+We formally prove theorems related to the two guarantees and use model-based randomized testing \citet{Pierce:SF4} to certify the simulation relation between the~\systemname semantics and its compiler formalism.
 To the best of our knowledge,~\systemname is the first C(-like) language and compiler formalism with the program partitioning mechanism.

### 2023-01-20 sandp23/evaluation.tex
Fixing eval and intro

@@
 
-Our evaluation of ~\systemname{} with NO-OP Sandbox was performed on two programs with the largest overheads in our dataset.
+We performed our experiment on~\texttt{Persons} and~\texttt{TinyBigNum}, the two programs with the largest overheads in our dataset.
+As expected, the runtime overhead of~\texttt{Persons} is 1.94\% -- which shows that the sandbox is the main contributor to the overhead.
+For~\texttt{TinyBigNum}, the overhead is a bit higher at 54.7\%, our analysis shows that the overhead is because we tainted the main input buffer, which is processed in loops.
+This leads to additional checking for every loop iteration resulting in higher overhead.
+Another reason is that in the current implementation, our instrumentation is performed at the end after all optimization passes; thus, none of the instrumentation is optimized.
+We plan to move our instrumentation before the optimization passes and exploit them to optimize further and decrease the runtime overhead.
+
+%Our evaluation of~\systemname{} with NO-OP Sandbox was performed on two programs with the largest overheads in our dataset.
 With parsons, we see an average run-time overhead of 1.94\% as is expected as parsons (($T_{Pr}$)) gave us an overhead of only 9\%.
 With TinyBigNum, we see an overhead of 54.7\%, and is expected referring to the figure ~\ref{fig:microbenchmarks}, as TinyBigNum is pointer intensive with pointer operations inside nested loops.
-This shows that~\systemname{} fundamentally has the least overhead.
+%This shows that~\systemname{} fundamentally has the least overhead.
 %The second set of bars in~\fig{fig:runtimeoverhead} shows the runtime overhead. The overhead is negligible in all cases.

### 2023-01-20 sandp23/introduction.tex
Fixing eval and intro

--- 2023-01-19 sandp23/introduction.tex
@@
 We extend the \checkedc compiler to support the type system and
-formalize it by extending~\checkedc formalism~\citet{li22checkedc} with the non-crashing and non-exposure guarantees.
-We formally prove theorems related to the two guarantees and use model-based randomized testing (CITE) to certify the simulation relation between the~\systemname semantics and its compiler formalism.
+formalize it by extending~\checkedc formalism~\cite{li22checkedc} with the non-crashing and non-exposure guarantees.
+We formally prove theorems related to the two guarantees and use model-based randomized testing \cite{Pierce:SF4} to certify the simulation relation between the~\systemname semantics and its compiler formalism.
 To the best of our knowledge,~\systemname is the first C(-like) language and compiler formalism with the program partitioning mechanism.

### 2023-03-06 csf2024/appendix.tex
update.

--- 2023-01-18 sandp23/appendix.tex
@@
 
-    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{\evalue{n'}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n'}{\tau'})}
-
-        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e})}
+    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}
+
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
 
@@
       {(\varphi,\heap,\eassign{\evalue{0}{\tptr{\omega}{\xi}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap,\enull)}
-
-    \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}
 
@@
 form. \textsc{S-Let} evaluates to a configuration whose expression is
-$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
-and remember $x$ and its new value $\evalue{n}{\tau}$
-in $e$'s scope that is defined by the $\erettext$ operation.
-Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
-we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope.
-After one-step evaluation is completed, 
-we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
-and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
-This procedure continues until $e'$ becomes a literal
-$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
-the literal.
+$\ret{x}{\varphi(x)}{e})$.
+It keeps the old value of $x$ in the stack of $\texttt{ret}$ 
+and updates $x$'s value in $\varphi$ with the new value of the let binding.
+After $e$ is evaluated to a value $\evalue{n}{\tau}$,
+we restore $\varphi$'s value back to the one stored in the stack of $\texttt{ret}$ as $\mu$,
+which is explain in rule \rulelab{S-Ret}.
 

### 2023-03-06 csf2024/compilation.tex
update.

--- 2022-08-16 sandp23/compilation.tex
@@
 % The two theorems are translation preservation and simulation. We donate $\xrightarrow{c}$ as the transition semantics of CLight.
-\begin{thm}[Simulation ($\sim$)]\label{simulation-thm}
+\begin{thm}\label{simulation-thm}[Simulation ($\sim$)]
 For \lang expressions $e_0$, stacks $\varphi_0$, $\varphi_1$, and heap snapshots $\heap_0$, $\heap_1$, 

### 2023-03-06 csf2024/formalaa.tex
update.

--- 2023-01-15 sandp23/formal.tex
@@
   \small \centering
-  \[ \hspace*{-1.5em}
-\setlength{\arraycolsep}{4pt}
+  \[ \hspace*{-2.2em}
+\setlength{\arraycolsep}{3pt}
 \begin{array}{l}
@@
        \text{Variables:}~ x
-& \quad\text{Integers:}~n::=\mathbb{Z} 
+&\quad \text{Integers:}~n::=\mathbb{Z} 
 \end{array}
 \\[0.5em]
+
 \begin{array}{llcllcl}
@@
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}
-\mid \ret{x}{\evalue{n}{\tau}}{e}
-  \\
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}  \\
 &&\mid& \estrlen{x} \mid \emalloc{\xi}{\omega} \mid\estar{e}\mid\eassign{e}{e}  \\
@@
 \mid \echecked{\overline{x}}{e}
+\mid \ret{x}{\evalue{n}{\tau}}{e}
 \end{array}
@@
 named~\lang.
-We present its syntax, semantics, type system,
-as well as \lang’s meta-theories, including the type soundness, non-exposure, and non-crashing theorems.
+We precisely present its syntax, semantics, type
+system, and compilation.
+% 
+\mzs{The formal model also enables us to}
+% 
+\mzr{From those, we} develop~\lang’s meta-
+theories, including the type soundness, non-exposure, non-crashing,
+and the compiler simulation theorems, which we present in~\apdx{sec:theorem}.
 
@@
 \subsection{Syntax}\label{sec:syntax}
-\Cref{fig:checkc-syn} shows the syntax of~\lang.
-
-\myparagraph{Type Syntax}
-At a high level, we classify types as word-size value, multi-word
-value, or function types. A word-size value can be either an integer or pointer.
-Every pointer type ($\tptr{\omega}{\xi}$) includes a
+The~\fig{fig:checkc-syn} shows a simplified syntax of~\lang.
+\subsubsection{Types ($\omega$)}
+At a high level, we classify types as word-size value \mzr{type}s, multi-word
+value \mzr{type}s, or function \mzr{type}s.
+%
+\mz{
+  % 
+  You cannot classify a type as values. You can only classify types into other 3
+  categories, namely, word, multi-word and function types.
+  % 
+}
+% 
+\mzs{The}\mzr{A} word-size value\mzs{s} can be either \mzr{of type integer} or
+\mzr{of type pointer}.
+% 
+\mzr{Each} pointer type\mzs{s} ($\tptr{\omega}{\xi}$) \mzr{is represented by} a
 pointer mode annotation ($\xi$, the difference between context and pointer modes
 is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
-or unchecked (\umode), and a type ($\omega$) denoting the valid value type it is pointed to.
-A multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
+or unchecked (\umode), and a type ($\omega$) denoting \mzr{the valid type of
+  values it can refer to}.
+%
+
+% 
+\mzr{The multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
 arrays and null-terminated arrays is constructed by the type of elements in the
-array ($\tau$), an array bound ($\bvar$) comprised of an upper and
-lower bound on the size of the array ($(b_l,b_h)$), and an array flag ($\kappa$).
-Bounds $b$ are
-limited to integer literals $n$ and expressions $x + n$.
-Whether an array pointer is null terminated or not is determined by annotation
-$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide $\cdot$ when writing types).
-An example representation of an array and null-terminated array in~\lang is shown below:
+array ($\tau$), an array bound ($\bvar$) and an array flag ($\kappa$).}
+% 
+\mzr{A} bound can be either \mzr{an} integer literal $n$, \mzr{an} expressions
+$x + n$, or an explicit specification of lower and upper bound\mzr{s},~\ie
+$(b_l,b_h)$.
+%
+\mzr{ %
+  Additionally, to distinguish null-terminated arrays from other arrays, we set
+  \(\kappa\) to be \textit{nt} for nt-arrays and \(\cdot\) for other arrays (for
+  convenience, we will elide \(\cdot\) in the text).
+  % 
+}
+%
+
+% 
+An example representation of an array and null-terminated array in~\lang syntax is shown below:
 % 
@@
 \]
-
-For simplicity, we write
+\mzr{For simplicity}\mzs{To simplify the representation}, we write
 $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean $\tptr{\tarray{0}{b}{\tau}}{\cmode}$,
@@
 
-\myparagraph{Disallowing Unsafe Types}
-The well-formedness of these types are presented in~\Cref{app:le}.
-It prevents unsafe types from being constructed.
+\emph{\mzr{Rejecting}\mzs{Disallowing} unsafe types.}
+The well-formed requirements of these types are presented in~\fig{fig:wftypesandbounds} (Appendix).
+% 
+These requirements \mzs{add additional restrictions for type creation and restrict the creation of unsafe types.}
+\mzr{prevents unsafe types from being constructed}.
+% 
+\mzr{
+  % 
   Consider the type ~\code{_t_Array_ptr<_Ptr<int>>} which describes a tainted
   array of checked pointers.
-  This is \emph{illegal} in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region;
-  thus, it is rejected by the well-formedness. 
-  Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
+  % 
+  Syntactically valid, this type is however \emph{illegal} in~\lang{} and
+  is rejected by the well-formedness requirement. 
+}
+% 
+\mz{Made some changes to shift the emphasis in hope of improving clearness.}
+% 
+However, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_t_Array_ptr<_t_Ptr<int>>} is a valid type.
+%
+
 % 
 Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
-where $\overline{x}$ represents a list of \tint{} type variables that bind type variables appearing in $\overline{\tau}$ and $\tau$.
+where $\overline{x}$ represents a list of \tint{} type variables that bind \mzu{bound}\mz{free?} variables appearing in $\overline{\tau}$ and $\tau$.
 An example of a function pointer type is shown below:
@@
 \begin{array}{l}
-$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau_1$\code{> : count(}$n$\code{),}$\\
-\qquad\qquad$\code{_t_NT_Array_ptr<}$\tau_2$\code{>: count(}$n$\code{), int}$\;n$\code{)>}$
+$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{),}$\\
+\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>: count(}$n$\code{))>}$
 \\[0.2em]
-\Leftrightarrow\; $\tptr{(\tfun{n}{\tint \times \tntarrayptr{0}{n}{\tau_2}{\tmode} \times \tntarrayptr{0}{n}{\tau_1}{\tmode}}{\tint})}{\tmode}$
+\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
 \end{array}
 \]
-The function type also has well-formed requirements (\Cref{app:le}), which disallows nesting checked pointers inside tainted pointers.
+The function type also has well-formed requirements (\fig{fig:wffunctions} in Appendix), which disallows nesting checked or unchecked pointers inside tainted pointers.
 Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+\aravind{The syntax of well-formed requirements for functions is incorrect. It looks like the forall qualifier is missing.}
 
@@
 %$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-\myparagraph{Expressions}
-\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
-pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
+\subsubsection{Expressions ($e$)}
+\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$)
+\mzu{($n\!:\!\tau$)}\mz{what's this?}, pointer dereference and assignment ($\estar{e}$) and ($\eassign{e_1}{e_2}$), resp.
 along with expressions that require special handling, such as,
- static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \cite{li22checkedc}}, the \texttt{strlen} operation ($\estrlen{x}$),
+ static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}}, the \texttt{strlen} operation ($\estrlen{x}$),
 memory allocation ($\emalloc{\xi}{\omega}$), 
@@
 unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
-For example, a dynamic bounds cast
-  { \footnotesize
-    \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
-is formalized as 
-{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \lang. 
-%
-
-% 
-We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
+For instance, a dynamic bounds cast expression
+\begin{center}
+  {\footnotesize
+    \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,
+      count(}$n$\code{))} }
+\end{center}
+becomes 
+{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}. 
+%
+
+% 
+\mzu{We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.}
+\mz{This sentence is unclear. I suppose that you mean an integer literal
+  can be both of interger type and pointer type depending on the context.  }
 For instance, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer.
+The~\code{strlen} expression plays an important role in the type-checking of null-terminated arrays.
+For simplifying type checking, we assume~\code{strlen} to operate on variables \mzs{$x$} rather than arbitrary expressions.
+% 
 The heap allocation expression $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions (\cregion or \ucregion).
 We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit code regions.
-To guarantee the non-exposure property, we extend the \checkedc syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$,
-where $\overline{x}$ represents all variables that are allowed to communicate between instructions outside and inside of the block $e$, and cannot contain checked pointers.
-$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
+The $\echeckedtext$ and $\euncheckedtext$ expressions are used to
+\mzs{represent the corresponding} \mzr{delimit} code regions.
+To guarantee the non-exposure property, we extend the base syntax to be $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$.
+\mzu{Wherein $\overline{x}$ restricts all free variables appearing in $e$ to be checked or unchecked pointers according to the corresponding region.}
+\mz{Is this sentence grammatical?}
+%
 
@@
 \lang{} aims to be simple enough to work with but powerful enough to
-encode realistic \systemname idioms. For example,
-loops can be encoded as recursive function calls. \code{struct}s are
+encode realistic \systemname idioms. For example, mutable local
+  variables can be encoded as immutable locals that point to the heap;
+\mzu{the use of \code{&} can be simulated with \code{malloc};}
+\mz{This doesn't make sense if you use \code{&} to create alias. You don't want
+  to allocate memory.}
+% 
+and loops can be encoded as recursive function calls. \code{struct}s are
 not included in~\fig{fig:checkc-syn} for space reasons, but they are
-supported as shown in~\cite{li22checkedc}.
+supported as shown in~\apdx{appx:struct}.
 C-style \code{union}s have no safe typing in \checkedc, so we omit them.
+
 Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
@@
 
+\aravind{Why we have only function pointers in function call?}
+\aravind{Talk about the ret expression after adding it in the syntax.}
+
 % \mwh{NOTE: COULD WORK THE FOLLOWING INTO THE ABOVE DESCRIPTION OF
@@
 
-\subsection{Typing and Semantics}
+\subsection{Type Checking}
 \label{sec:typechecking}
-
-\begin{DIFnomarkup}
-\begin{figure*}[t]
-{\small
-  \begin{mathpar}
-     \inferrule[T-CastPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
-
-    \inferrule[S-Cast]
-              {}
-              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
-
-\\\\
-    \inferrule[T-Def]
-              {\xi \leq m \\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-\quad
-        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
-         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-
-    \inferrule[T-Checked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
-              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
-
-    \inferrule[T-Unchecked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
-
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-
-\inferrule[T-Fun]
-    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
-         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
-         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
-             \overline{\tau'} \sqsubseteq_{\Theta}
-               \overline{\tau}[\overline{e'} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
-
-    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-\quad
-    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
-                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-
-  \end{mathpar}
-}
-% {\footnotesize
-% \begin{center}
-% $
-% \begin{array}{l}
-% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
-% \\[0.2em]
-% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
-% \end{array}
-% $
-% \end{center}
-% }
-{\footnotesize
-$
-\cmode(\tint)=\texttt{false}
-\qquad
-\cmode(\tptr{\omega}{\cmode})=\texttt{true}
-\qquad
-\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
-$
-}
-\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.
-First line is for cast operations, second line is for pointer dereferences, third line is for checked/unchecked blocks, and the rest is for function calls.}
-\label{fig:type-system-1}
-\end{figure*}
-\end{DIFnomarkup}
-
-The \lang type system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
 Our type checker restricts the usage of tainted and checked pointer types to ensure that tainted pointers do not affect checked types, along with enforcing~\checkedc typing rules~\cite{li22checkedc}.
-
-As partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{T-}$X$),
-each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+A few selected type rules are presented in~\fig{fig:type-system-1}.
+% 
+\mz{Move the figure here.}
+% 
+The complete set of typing rules and special handling of (NT)-arrays are provided in~\apdx{sec:literal-pointer-typing}.
+
+Each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
 which states that in a type environment $\Gamma$ (mapping variables to
@@
 variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-in context mode $m$, indicating that the code is in $m$ region.
-The operational semantics for \lang is defined as a small-step
-transition relation with the judgment $ (\varphi,\heap,e)
-\longrightarrow_m (\varphi',\heap',r)$, as partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{S-}$X$).
- Here, $\varphi$ is a
-\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
-$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ heaps), each of which
-maps addresses (integer literals) to values $\evalue{n}{\tau}$.
-The complete set of typing rules and special handling of (NT)-arrays are provided in~\Cref{rem-type,sec:rem-semantics}.
-
-We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ heap,
-and $\heapup{m}{n}{\evalue{n'}{\tau}}$ 
-to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ heap.
-While heap bindings can change, stack bindings are immutable---once
-variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
-be updated. 
-%We can model mutable stack variables as pointers into the mutable heap.
-As mentioned, value $\evalue{0}{\tau}$
-represents a $\enull$ pointer when $\tau$ is a pointer type.
-Correspondingly, $\heap(m,0)$ should always be undefined.
-% 
-The relation steps to a \emph{result} $r$, which is either an
-expression, a $\enull$ or $\ebounds$ failure, represent an expression right
-  after the reduction, a null-pointer dereference or out-of-bounds access,
-respectively.
-% 
-Such failures are a \emph{good} outcome; stuck states
-(non-value expressions that cannot transition to a result $r$)
-characterizing undefined behavior.
-%The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the stepped redex within $e$ was in a $\cmode$ or $\umode$ region.
-
-The rules for the main operational semantics
-judgment \emph{evaluation} are given at the bottom of
-Fig.~\ref{fig:c-context}.
-The first rule takes an expression $e$, decomposes
-it into an \emph{evaluation context} $E$ and a sub-expression $e'$
-(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
-$e$), and then evaluates $e'$ according to the \emph{computation}
-  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
-whose rules are given along with type rules in Fig.~\ref{fig:type-system-1}, discussed
-shortly.
-The $\mode$ function in Fig.~\ref{fig:c-context}
-determines the context mode, i.e., region, that the expression $e'$ locates based on the context $E$.
-In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.
-The second rule describes the exception handling 
-for possible crashing behaviors in $\umode$ regions.
-Operations in $\umode$ region can non-deterministically crash
-and the \systemname sandbox mechanism recovers
-the program to a safe point ($\evalue{0}{\tau}$)
-and continues with the existing program state.
-Evaluation contexts $E$ define a standard left-to-right evaluation order. 
-%(We explain the $\ret{x}{\mu}{e}$ syntax shortly.)
-%There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
-
-\myparagraph{Modes, Static Casting, and Subtyping}
-In \lang, Context modes $m$ appearing in a type rule determine the code region 
-that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
-We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
-This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can perform such operations in $\cmode$ and $\umode$ regions, respectively.
-%Pointer modes are also useful in determining if a nested pointer has a valid type.
-For example, in a nested pointer $\tptr{(...
-\tptr{\tau}{\xi_2} ...)}{\xi_1}$, we require $\xi_2\le \xi_1$ to maintain non-exposure.
-
-\lang also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
-an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
-if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2}, i.e., $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
-In \lang, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
+in context mode $m$.
+
+In \lang, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
+We provide some example descriptions here.
+Type equality $\tau=_{\Theta}\tau'$
+is a type construct equivalent relation defined by the bound equality ($=_{\Theta}$) in (NT-)array pointer types
+and the alpha equivalence of two function types;
+i.e., two (NT-)array pointer types $\tallarrayb{\bvar}{\tau} $ and $ \tallarrayb{\bvar'}{\tau'}$ are equivalent, if 
+$\bvar =_{\Theta} \bvar'$ and $\tau=_{\Theta}\tau'$; two function types 
+$\tfun{\overline{x}}{\overline{\tau}}{\tau} $ and $ \tfun{\overline{y}}{\overline{\tau'}}{\tau'}$
+are equivalent, if we can find a same length (as $\overline{x}$ and $\overline{y}$) variable list $\overline{z}$ that is substituted for $\overline{x}$ and $\overline{y}$ in $\overline{\tau} \to {\tau}$ and $\overline{\tau'} \to {\tau'}$, resp.,
+and the substitution results are equal.
+\aravind{For the above paragraph, give concrete example, For example: array is a subtype of ptr, etc}
+
+The type checker rules are designed such that any well-typed (\ie follows all the rules) is isolated by sandboxed code. We present the corresponding formal proof in~\apdx{}.
+
+One of the important type cheker rules is disallowing unsafe interactions between tainted and untainted (c/u) pointers.
+\aravind{Explain about the LET rule}
+
+\aravind{We should explain how our type checker restricts unsafe interactions between checked and tainted pointer types.}
 
 \begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-//_Ptr<int> x; _t_Ptr<int> y; int *z;
-z = (int *)y; // This is okay.
-x = (_Ptr<int>)y; // Not allowed.
+a -> checked pointer
+b -> Tainted pointer
+c -> Unchecked pointer.
+
+_Ptr<int> a;
+_t_Ptr<int> b;
+int *c;
+
+a = b; // Not allowed.
+b = a; // Not allowed.
 \end{minted}
 
-In the above example, a $\tmode$ mode pointer can be cast to $\umode$ mode but casting $\tmode$ mode to $\cmode$ mode is disallowed.
-The complete subtyping relation was described in \Cref{app:le}.
-Notice that \texttt{let} statements are immutable in \lang, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \lang.
+The \textsc{T-CastPtr} rule in \Cref{fig:type-system-1}
+permits casting from an expression of type $\tau'$ to a checked pointer when
+$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
+$\sqsubseteq$ is built on the type equality ($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). 
+The rule  ($0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
+\tarrayptr{b_l}{b_h}{\tau}{m}$) permits treating a singleton
+pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
+Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
+
+\aravind{Explain this with .}
 
 \begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-//_Ptr<int> x;  _t_Ptr<int> y;
-x = y; // Not allowed.
+a -> checked pointer
+b -> Tainted pointer
+c -> Unchecked pointer.
+
+_Ptr<int> a;
+_t_Ptr<int> b;
+int *c;
+
+a = b; // Not allowed.
+b = a; // Not allowed.
+
+a = (_Ptr<int>)c; // This is okay.
+a = (_Ptr<int>)b; // Not allowed.
 \end{minted}
 
-\myparagraph{Pointer Dereference}
-The type and semantic rules for pointer dereference (\textsc{T-Def}, \textsc{S-DefC},
-\textsc{S-DefT}, \textsc{S-DefNull} in \Cref{fig:type-system-1})
-reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
-The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
-With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a $\enull$ pointer is used,
-\lang captures the runtime error.
-Type and semantic rules for array types and pointer assignments are given in \Cref{rem-type,sec:rem-semantics}.
-
-Rules \textsc{S-DefC} and \textsc{S-DefT} are for $\cmode$ and $\tmode$ mode pointer dereferences, respectively.
-In addition to the no $\enull$ check in $\cmode$ mode pointer dereference,
-any dynamic heap access of a tainted pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to that the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
-
-\myparagraph{Unchecked and Checked Blocks}
-The execution of a $\echeckedtext$ or $\euncheckedtext$ block represents 
-the context switching from a $\cmode$ to an $\umode$ region, or vice versa,
-with its type and semantic rules given in \Cref{fig:type-system-1}.
-In this context switching, to guarantee the checked pointer non-exposure property, 
-checked pointers are not allowed to pass through, which is guaranteed by the predicates 
-$\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$,
-as well as the check that all free variables in the block content $e$ are in $\overline{x}$.
-For example, \code{StringAuth} in \Cref{subsub:gencregion} is a trampoline function that disallows
-checked pointers as arguments and return values.
-The use of the function in the following \code{_T_StringAuth}, which is in $\umode$ region,
-cannot legally acknowledges any checked pointers; otherwise, we might expose a checked pointer address to unsafe code regions.
-In \systemname, we actually permits the accesses of checked pointers inside \code{StringAuth},
-since the function body of a trampoline function is in $\cmode$ region.
-More information is given in \Cref{subsub:gencregion}.
-
-\myparagraph{Dependent Function Pointers} %
-Rule \textsc{T-Fun} (\Cref{fig:type-system-1}) states the type judgment for
-dependent function pointer application, where we represent the result of
-replacing all integer bound variables $\overline{x}$ in the type \(\tau\) with
-with bound expressions $\overline{e'}$ by $\tau[\overline{e'} / \overline{x}]$
-and write $\overline{\tau}[\overline{e'} / \overline{x}]$ to lift the
-substitution to every type \(\overline{\tau}\).
-% 
-Given an expression $e$ of function pointer type
-($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$) and arguments
-$\overline{e}$ of types $\overline{\tau'}$,
-% 
-the result of the application will
-be of type $\tau[\overline{e'} / \overline{x}]$ 
-% 
-if for each pair of \(\tau'\) and \(\tau''\) in \(\overline{\tau'}\) and
-$\overline{\tau}[\overline{e'} / \overline{x}]$, \(\tau'\) is a subtype of
-\(\tau''\).
-Consider the \code{process_req2} function in
-Fig.~\ref{lst:final}, whose parameter type for \code{msg} 
-depends on \code{m_1}.
-Its function pointer type is 
-$\tptr{\tfun{\code{m_1}}{\tint,\tntarray{0}{\code{m_1}}{\texttt{char}}}{\tint}}{\tmode}$.
-In \code{handle_request}, the call \code{process_req2(buff, r_len)} binds variable \code{m_1} to \code{r_len}.
-After the call returns, \code{m_1}'s scope is ended, so we need to substitute it with \code{r_len} in the final return type
-because it might contain \code{m_1}.
-
-\textsc{S-FunC} and \textsc{S-FunT} define the semantics
-for $\cmode$ and $\tmode$ mode function pointers, respectively. 
-A call to a function pointer $n$ retrieves
- the function definition in $n$'s location in the global function store $\Xi$,
-which maps function pointers to
-function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
-$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
-is the parameter list of variables and their types, 
-$\xi$ determines the mode of the function, and $e$ is the
-function body. 
-Similar to \heap, the global function store $\Xi$ is also partitioned into
-two parts ($\cmode$ and $\umode$ stores), each of which
-maps addresses (integer literals) to the function data described above.
-Rule \textsc{S-FunT} defines the tainted version of function call
-with the verification process 
-$\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}$
-makes sure that the function in the global store is well-defined and has the right type.
-
-\subsection{Meta Theories}\label{sec:theorem}
-
-% Before we present our main theorems, we need to first
-% discuss the meaning what a pointer being well-typed in a given heap
-% snapshot $\heap$ means, which is captured by rules in
-% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
-% simply checks if a given variable has the defined type in $\Gamma$;
-% the constant rule ($\textsc{T-Const}$) is slightly more involved.
-% First, it ensures that the type annotation $\tau$ does not contain any
-% free variables. More importantly, it ensures that the pointer points
-% to a location that makes sense in a given heap.
-%  
-%  
-%  The $\size$ function in Fig.~\ref{fig:const-type}
-% refers to the \code{sizeof} function in C computing the number of
-% bytes for a type.
-%  
-%  
-%  Second, we
-% require that any constant ($\evalue{n}{\tau}$) should make sense in
-% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
-% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
-% constant set containing the constants that have been verified by the
-% relation. For every constant $\evalue{n}{\tau}$, it is either an
-% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
-% zero-valued number ($n=0$), checked in $\sigma$
-% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
-% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
-% location $n+i$ in the range of the pointer (if $\omega$ is a word
-% type, range is $[0,1)$; if $\omega$ is an array type
-%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
-%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
-%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
-%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
-%       \vdash n_a : \tau_a$.
-%  
-%  
-% \begin{figure}[t]
-% {\small
-% \text{Type Rules for Constants and Variables:}
-% \begin{mathpar}
-%   \inferrule[T-Var]
-%       {x : \tau \in \Gamma}
-%       {\Gamma;\Theta \vdash_m x : \tau}
-%  
-%   \inferrule[T-Const]
-%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
-%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
-% \end{mathpar}
-%     
-% \text{Rules for Checking Constant Pointers In Heap:}
-% \begin{mathpar}
-%   \inferrule
-%       {}
-%       {\sigma \vdash n : \tint}
-%  
-%   \inferrule
-%       {}
-%       {\sigma \vdash n : \tptr{\omega}{\umode}}
-%  
-%   \inferrule
-%       {}
-%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
-%  
-%   \inferrule
-%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
-%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-%  
-%   \inferrule
-%       {\forall i \in [0,\size(\omega)) .
-%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
-%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
-% \end{mathpar}
-% }
-% \caption{Type Rules for Checking Constants/Variables}
-% \label{fig:const-type}
-% \end{figure}
-
-% \review{
-%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
-%   something, I didn't see such a definition in the paper.}
-% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
-%   assumption of well typing; we have added more details about that.}
-
-Here, we discuss our main meta-theoretic results for
-\lang: type soundness (progress and preservation),
-non-exposure, and non-crashing.
-These proofs have been conducted in our Coq model.
-Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{sec:meta}.
-The progress theorem below states that a \lang program can always make a move.
-
-\begin{thm}[Progress]\label{thm:progress}
-
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
-one of the following holds:
-
-\begin{itemize}
-
-\item $e$ is a value ($\evalue{n}{\tau}$).
-
-\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.
-
-\end{itemize}
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-
-\noindent
-There are two forms of preservation regarding the $\cmode$ and $\umode$ regions.
-Checked Preservation states that a reduction step preserves both the
-type and consistency of the program being reduced, while
-unchecked Preservation states that any evaluation happens at $\umode$ region does not affect the $\cmode$ mode heap.
-
-\begin{thm}[Checked Preservation]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
-$\cmode$ region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$ and $\tau'$ that are well formed, $\cmode$ region consistent
-($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
-well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
-$\tau'\sqsubseteq_{\Theta} \tau$.
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-%\smallskip
-\begin{thm}[Unchecked Preservation]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
-\end{thm}
-
-Using the above theorem, we first show the non-exposure theorem,
-where code in $\umode$ region cannot observe a valid checked pointer address.
-
-\begin{thm}[Non-Exposure]
-For any \lang program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
-where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
-thus, it is not a checked pointer.
-\end{thm}
-
-We now state our main result, {\em non-crashing},
-which suggests that a well-typed program can never be \emph{stuck} (expression
-$e$ is a non-value that cannot take a step\footnote{Note that
-  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
-  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
-  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
-  safety.}).
-
-% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r.
-Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
-% \liyi{It is a typo.
-We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region.
-This is a bit due to the space limitation.
-The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement.
-}
-
-\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
-  program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are well-formed and consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
-if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
-\tau$ for some $\tau$) and there exists
-$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
-$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
-\end{thm}
-
-%{\em Proof:} By induction on the number of steps of the \checkedc
-%evaluation ($\longrightarrow_m^*$), using progress and preservation to
-%maintain the invariance of the assumptions.
-
-\ignore{
-\subsection{Semantics}\label{sec:semantics}
+\subsection{Semantics and Typing}\label{sec:semantics}
 
@@
 
-Here, we discuss the \lang operational semantics (\Cref{fig:semantics}); 
+Here, we discuss the \lang type system (\Cref{fig:type-system-1}) and operational semantics (\Cref{fig:semantics}); 
 mainly focusing on the new changes on top of \checkedc in \cite{li22checkedc} with function pointers, modes, and function calls.
@@
 
-%The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
-%which states that in a type environment $\Gamma$ (mapping variables to
-%their types) and a predicate environment $\Theta$ (mapping integer-typed
-%variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
-%in context mode $m$. Key rules for this judgment are given in
-%Fig.~\ref{fig:type-system-1},
-
-The operational semantics for \lang is defined as a small-step
+The typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+which states that in a type environment $\Gamma$ (mapping variables to
+their types) and a predicate environment $\Theta$ (mapping integer-typed
+variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
+in context mode $m$. Key rules for this judgment are given in
+Fig.~\ref{fig:type-system-1},
+while the operational semantics for \lang is defined as a small-step
 transition relation with the judgment $ (\varphi,\heap,e)
@@
 %There are other rules for describing the halts of evaluation to $\enull$ and $\ebounds$ states in \Cref{app:main}.
+
+\begin{DIFnomarkup}
+\begin{figure*}[t]
+{\small
+  \begin{mathpar}
+        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+\quad
+        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
+         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+\quad
+    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
+\quad
+    \inferrule[S-Cast]
+              {}
+              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
+
+    \inferrule[S-RetEnd]{}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{\evalue{n'}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n'}{\tau'})}
+
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e})}
+
+    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-RetCon]{ (\varphi[x\mapsto \evalue{n}{\tau}],\heap,e) \longrightarrow (\varphi',\heap',e')}{(\varphi,\heap,\ret{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi'[x\mapsto \varphi(x)],\heap',\ret{x}{\varphi'(x)}{e'})}
+
+    \inferrule[S-Checked]{}{(\varphi,\heap,\echecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
+        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+\quad
+    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
+                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
+        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+
+ % \inferrule[S-IfNTNot]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}} \\ \heap(n)\neq 0\\ 0 < n_h}
+ %            {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
+
+\end{mathpar}
+}
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \tau[\overline{n} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall n_i\in\overline{n}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \Rightarrow \tau[n_i / x_i]\\[0.2em]
+% \mathtt{let}\;\overline{x}=\overline{e}\;\mathtt{in}...\triangleq \mathtt{let}\;x_0=e_0\;\mathtt{in}\;\mathtt{let}\;x_1=e_1\;\mathtt{in}...
+% \end{array}
+% $
+% \end{center}
+% }
+\caption{\lang Semantics: Computation (Selected Rules)}
+\label{fig:semantics}
+\end{figure*}
+\end{DIFnomarkup}
 
@@
 \begin{DIFnomarkup}
- \begin{figure}[t]
+\begin{figure*}[t]
 {\small
-
   \begin{mathpar}
-   \inferrule
-       {}
-       {\Theta;\heap;\sigma \vdash_m n : \tint}
-
-   \inferrule
-       {}
-       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
-
-   \inferrule
-       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
-       {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
-
-   \inferrule
-       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
-
-   \inferrule
-       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
-            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
-
-   \inferrule
-       { \xi \le m 
-     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
-       \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
-
-   \inferrule
-       {\neg\funptr(\omega)\\ \xi \le m\\
-        \forall i \in [0,\size(\omega)) \;.\;
-            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+   \inferrule[T-ConstU]
+       { \neg \cmode(\tau)}
+       {\Gamma;\Theta\vdash_u \evalue{n}{\tau} : \tau}
+
+   \inferrule[T-ConstC]
+       {\Theta;\heap;\emptyset \vdash_c n : \tau}
+       {\Gamma;\Theta\vdash_c \evalue{n}{\tau} : \tau}
+
+    \inferrule[T-Def]
+              {\xi \leq m \\\\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
+              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
+
+     \inferrule[T-CastPtr]
+               {\Gamma;\Theta \vdash_m e : \tau' \\
+                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
+               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
+
+   \inferrule[T-Let]
+    { x\not\in \fv(\tau') \\
+        \Gamma;\Theta \vdash_m e_1 : \tau \\\\
+          \Gamma[x\mapsto \tau];\Theta \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'}
+
+   \inferrule[T-LetInt]
+    { x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} \\
+        \Gamma;\Theta \vdash_m e_1 : \tint \\\\
+           \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{e_1}] \vdash_m e_2 : \tau'
+             }
+    {\Gamma;\Theta \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
+\qquad
+   \inferrule[T-RetInt]
+    { \Gamma[x\mapsto \tint];\Theta[x\mapsto \teq{n}] \vdash_m e : \tau}
+    {\Gamma;\Theta \vdash_m \eret{x}{\evalue{n}{\tint}}{e} : \tau}
+
+    \inferrule[T-Checked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
+              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
+
+    \inferrule[T-Unchecked]
+              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
+              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
+
+\inferrule[T-Fun]
+    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
+        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
+         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
+         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
+             \overline{\tau'} \sqsubseteq_{\Theta}
+               \overline{\tau}[\overline{e'} / \overline{x}]}
+    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
+
   \end{mathpar}
 }
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
+% \\[0.2em]
+% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
+% \end{array}
+% $
+% \end{center}
+% }
 {\footnotesize
-\[
-\begin{array}{l} 
-\funptr(\tfun{\overline{x}}{\overline{\tau}}{\tau}) = \texttt{true}
+$
+\cmode(\tint)=\texttt{false}
 \qquad
-\funptr(\omega) = \texttt{false}\;\;{[\emph{owise}]}
-\end{array}
-\]
-}
- \caption{Verification/Type Rules for Constants}
- \label{fig:const-type}
- \end{figure}
+\cmode(\tptr{\omega}{\cmode})=\texttt{true}
+\qquad
+\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
+$
+}
+\caption{Selected type rules}
+\label{fig:type-system-1}
+\end{figure*}
 \end{DIFnomarkup}
-
-% SKIPPING THIS
-\iffalse
 
@@
 
-\fi
 \subsection{Compilation}\label{sec:compilation}
@@
 The other compilation rules are given in \Cref{appx:add-type-sem}.
-}

### 2023-03-06 csf2024/limitations.tex
update.

--- 2023-01-16 sandp23/limitations.tex
@@
 \label{sec:limitations}
-Despite the effectiveness of~\systemname{}, it has the following limitations:
+Despite the effectiveness of~\systemname{}, it has limitations:
 
@@
 Also, as shown in~\sect{subsec:programoverhead}, the performance of the partitioned applications mainly depends on the sandbox.
-However, our implementation is not dependent on one specific sandbox and can be easily extended to other sandboxes. As part of our future work, we will extend our implementation to other sandboxes.
+However, our implementation is not dependent on one specific sandbox and can be easily extended to other sandboxes. As a future work, we will extend our implementation to other sandboxes.
 

### 2023-03-06 csf2024/examples/conv1.c

@@ -1,0 +1,36 @@
+|\entrypoint| // Entry point
+void server_loop(int sock_fd) {
+ unsigned b_z;
+ |\textcolor{checkcolor}{\_Array\_ptr}|<struct queue> p : 
+              count(b_z) = NULL;
+ ...
+ while(1) {
+  ...
+  p = malloc(sizeof(struct queue) * b_z);
+  ...
+  if (handle_request(sock_fd)) {
+    ...  
+  }
+ }
+}
+
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_Tainted}|;
+ int rc = -1;
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, 
+                  MAX_MSG_SIZE);
+ if (r_len > 0 && 
+     r_len < MAX_MSG_SIZE) {
+  switch(buff[0]) {
+   case REQ1: 
+    rc = process_req1(buff, r_len); 
+    break;
+   case REQ2: 
+    rc = process_req2(buff, r_len); 
+    break;
+   ...
+  }
+ }
+ return rc;
+}

### 2023-03-06 csf2024/examples/conv2.c

@@ -1,0 +1,39 @@
+size_t read_msg(int sock_fd, char *msg, 
+                size_t sz) _Tainted {
+ size_t nr;
+ nr = read(sock_fd, (void*)msg, sz);
+ ...
+ |\complexcode| // complex code
+ return nr;
+}
+
+int process_req1(_T_Array_ptr<char> msg :
+                 count(m_l), size_t m_l) 
+                 _Tainted {
+
+ int rc = -1, i;
+ if (m_l > MIN_SIZE) {
+   sscanf(msg, "%d", &i);
+   msg[i] = ...
+ }
+ // complex code
+ return rc;
+}
+
+int process_req2(|\textcolor{taintcolor}{\_T\_Array\_ptr}|<char> msg : 
+                 count(m_l), 
+                 size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum |\textcolor{checkcolor}{\_Checked}|[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}

### 2023-03-06 csf2024/examples/humanadjustments.c

@@ -1,0 +1,9 @@
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_Tainted} \usermods|;
+ ...
+}
+
+int process_req2(|\usermods| |\textcolor{taintcolor}{\_T\_Array\_ptr}|<char> msg: count(m_l), 
+                 size_t m_l) {
+ ...
+}

### 2023-03-06 csf2024/examples/humanannotations.c

@@ -1,0 +1,8 @@
+size_t read_msg(int sock_fd, char *msg, 
+                size_t sz) |\textcolor{taintcolor}{\_Tainted}| |\useradded| {
+ ...
+}
+
+int process_req1(char *msg, size_t m_l) |\textcolor{taintcolor}{\_Tainted}| |\useradded| {
+ ...
+}

### 2023-03-06 csf2024/examples/orig1.c

@@ -1,0 +1,35 @@
+|\entrypoint| // Entry point
+void server_loop(int sock_fd) {
+ unsigned b_z;
+ struct queue *p;
+ ...
+ while(1) {
+  ...
+  p = malloc(sizeof(struct queue) * b_z);
+  ...
+  if (handle_request(sock_fd)) {
+    ...  
+  }
+ }
+}
+
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE];
+ int rc = -1;
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, 
+                  MAX_MSG_SIZE);
+ if (r_len > 0 && 
+     r_len < MAX_MSG_SIZE) {
+  switch(buff[0]) {
+   case REQ1: 
+    rc = process_req1(buff, r_len); 
+    break;
+   case REQ2: 
+    rc = process_req2(buff, r_len); 
+    break;
+   ...
+  }
+ }
+ return rc;
+}

### 2023-03-08 csf2024/formal.tex
update.
update syntax.

--- 2023-01-19 sandp23/formal.tex
@@
 \begin{array}{llcllcl}
-\text{Context Mode:} & m & ::= & \cmode \mid \umode \\
-\text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\
-\text{Bound:} & b & ::= & n \mid x \plus n \\
-              & \bvar & ::= & (b,b) \\
-
-     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{ \xi}
-\\
-\text{Type Flag:}&\kappa &::=& nt \mid \cdot
-\\
-\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau} \mid \tfun{\overline{x}}{\overline{\tau}}{\tau}
-\\
+\text{Context Mode:} & m & ::= & \cmode \mid \umode \\[0.2em]
+\text{Memory Mode:} & \xi & ::= & m \mid \tmodered \\[0.2em]
+\text{Bound:} & b & ::= & n \mid x \plus n \\[0.2em]
+              & \bvar & ::= & (b,b) \\[0.2em]
+
+     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{\xi}
+\\[0.2em]
+\text{Type:}&\omega &::=& \tau \mid \tarrayb{\bvar}{\tau} \mid \textcolor{red}{\tfun{\overline{x}}{\overline{\tau}}{\tau}}
+\\[0.2em]
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}
-\mid \ret{x}{\evalue{n}{\tau}}{e}
-  \\
-&&\mid& \estrlen{x} \mid \emalloc{\xi}{\omega} \mid\estar{e}\mid\eassign{e}{e}  \\
-&&\mid& \elet{x}{e}{e} \mid \eif{e}{e}{e} \mid \ecall{e}{\overline{e}}
-\\
-&&\mid&\eunchecked{\overline{x}}{e}
-\mid \echecked{\overline{x}}{e}
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e}
+  \\[0.2em]
+&&\mid& \estar{e}\mid\eassign{e}{e} \mid \textcolor{red}{\emalloc{\xi}{\omega}} \\[0.2em]
+&&\mid&\eif{e}{e}{e} \mid  \elet{x}{e}{e} \mid \textcolor{red}{\ecall{e}{\overline{e}}}
 \end{array}
@@
 
-\ignore{
+\begin{figure}
+{
+\begin{center}
+\includegraphics[width=0.35\textwidth]{memory.png}
+\end{center}
+}
+\caption{
+    \systemname separate heap layouts
+}
+  \label{fig:checkc-memory}
+\end{figure}
+
+\begin{figure}
+{
+  \small \centering
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\diagbox{\text{Context}}{\text{Memory}}
+& $\texttt{malloc}$ & $\estar{e}$ & $\eassign{e}{e}$ & $\ecall{e}{\overline{e}}$\\[0.1em]
+\hline
+\cmode&$\cmodenor$,$\tmodenor$ & $\cmodenor$,$\tmodenor$ & $\cmodenor$,$\tmodenor$ & $\cmodefun$,$\tmodefun$\\[0.1em]
+\hline
+\umode&$\umode$ & $\tmodenor$,$\umode$ & $\tmodenor$,$\umode$ & $\tmodefun$,$\umode$\\[0.1em]
+\hline
+    \end{tabular}
+}
+  \caption{Permitted Memory Modes in Context Regions}
+  \label{fig:checkc-mode}
+\end{figure}
+
 \begin{figure}[t]
@@
   {m \vdash \tint}
-
+\qquad
   \inferrule[]
   {\xi \wedge m\vdash \tau \\ \xi \le m}
-  {m \vdash \tptr{\tallarrayb{\bvar}{\tau}}{\xi}}
-
+  {m \vdash \tptr{\tarrayb{\bvar}{\tau}}{\xi}}
+\qquad
   \inferrule[]
@@
   \inferrule[]
-  {\xi \wedge m \vdash \tau\\ \xi \le m \\\\ \fv(\overline{\tau})\cup\fv(\tau)\subseteq \overline{x}}
+  {\forall \tau'\in \overline{\tau}\cup\{\tau\}\,.\,\xi \wedge m \vdash \tau'\\ \xi \le m \\ \fv(\overline{\tau})\cup\fv(\tau)\subseteq \overline{x}}
   {m \vdash \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi})}
@@
 \[
-\begin{array}{l} 
+\begin{array}{c} 
 \tmode \wedge \cmode = \umode \qquad \xi \wedge \umode = \umode
@@
 \end{figure}
-}
 
@@
 
-This section describes the formal core model of~\systemname,
-named~\lang.
+This section describes \systemname's formal core model, \lang.
 We present its syntax, semantics, type system,
-as well as \lang’s meta-theories, including the type soundness, non-exposure, and non-crashing theorems.
+as well as \lang’s meta-theories, including the non-exposure, type soundness, and non-crashing theorems.
 
@@
 \subsection{Syntax}\label{sec:syntax}
-\Cref{fig:checkc-syn} shows the syntax of~\lang.
-
-\myparagraph{Type Syntax}
-At a high level, we classify types as word-size value, multi-word
-value, or function types. A word-size value can be either an integer or pointer.
+\Cref{fig:checkc-syn} shows the syntax of~\lang, with marked red items being new added features different from the \checkedc model \cite{li22checkedc}.
+
+\myparagraph{Types}
+%
+\systemname and \checkedc are similar in terms of type syntax, except for function types.
+We classify types as word-size value, multi-word
+value, or newly introduced function types (explained shortly below). A word-size value can be either an integer or pointer.
 Every pointer type ($\tptr{\omega}{\xi}$) includes a
-pointer mode annotation ($\xi$, the difference between context and pointer modes
-is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
-or unchecked (\umode), and a type ($\omega$) denoting the valid value type it is pointed to.
+memory mode (explained shortly below), and a type ($\omega$) denoting the valid value type it is pointed to.
 A multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
-arrays and null-terminated arrays is constructed by the type of elements in the
-array ($\tau$), an array bound ($\bvar$) comprised of an upper and
-lower bound on the size of the array ($(b_l,b_h)$), and an array flag ($\kappa$).
+arrays is constructed by the type of elements in the
+array ($\tau$), a bound ($\bvar$) comprised of an upper and
+lower bound on the size of the array $(b_l,b_h)$, and a flag ($\kappa$).
 Bounds $b$ are
 limited to integer literals $n$ and expressions $x + n$.
-Whether an array pointer is null terminated or not is determined by annotation
-$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide $\cdot$ when writing types).
-An example representation of an array and null-terminated array in~\lang is shown below:
+An example representation of an array in~\lang is shown below:
 % 
@@
 &\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
-\\[0.2em]
-$\code{_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
 \end{array}
@@
 $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean $\tptr{\tarray{0}{b}{\tau}}{\cmode}$,
-so the above examples could be rewritten as $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
-$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-
-\myparagraph{Disallowing Unsafe Types}
-The well-formedness of these types are presented in~\Cref{app:le}.
-It prevents unsafe types from being constructed.
-  Consider the type ~\code{_t_Array_ptr<_Ptr<int>>}, which describes a tainted array of checked pointers.
-  This is not well-formed in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the tainted ($\tmode$) array is used. 
+so the above examples could be rewritten as $\tptr{\tarrayb{n}{\tau}}{\tmode}$.
+There are other syntactic categories for modeling null-terminated array and \code{struct} types in \Cref{app:main}.
+
+\myparagraph{Modes and Well-formed Types}
+%
+In \systemname, we divide \checkedc modes to context and memory modes.
+In terms of program contexts, modes refer to the partition of a program into checked (\cmode) and unchecked (\umode) regions,
+as the context modes in \Cref{fig:checkc-syn}.
+In terms of pointers, memory modes classify the partitioned memory ranges that particular pointers point to, either checked (\cmode), tainted (\tmode), or unchecked (\umode) memory ranges.
+More specifically, we partition each of the above ranges to a sub-range (subscripted with $\texttt{f}$) storing static function code data, and a sub-range (subscripted with $\texttt{n}$) storing normal heap data,
+so function pointers and heap pointers point to different sub-ranges conceptually.
+Different operations in different context regions are allowed to access different pointer modes, shown in \Cref{fig:checkc-mode}.
+A \texttt{malloc} operation in \cmode region produces a pointer accessing $\cmodenor$ or $\tmodenor$ memory range, while a \texttt{malloc} in \umode region produces a $\umode$ range pointer.
+Function calls ($\ecall{e}{\overline{e}}$) in \cmode region are allowed to access $\cmodefun$ or $\tmodefun$ range pointers,
+while calls in \umode region are permitted to access $\umode$ range pointers.
+
+\systemname nested pointers also has well-formedness type restrictions in~\Cref{fig:wftypes},
+to prevent unsafe pointers, which might leak \cmode mode pointer address to \umode regions, from being constructed.
+Consider the pointer type ~\code{_t_Array_ptr<_Ptr<int>>}, which describes a tainted array of checked pointers.
+This is not well-formed in~\lang because it potentially exposes the checked pointer addresses in a $\umode$ region (a violation of the non-exposure property in \Cref{sec:theorem}), when the tainted ($\tmode$) array is used.
   Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
-%
+
 Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
@@
 An example of a function pointer type is shown below:
+{\small
 \[\hspace*{-0.5em}
@@
 \]
-The function type also has well-formed requirements (\Cref{app:le}), which disallows nesting checked pointers inside tainted pointers.
+}
+The function type also has well-formed requirements (\Cref{fig:wftypes}), which disallows nesting checked pointers inside tainted pointers.
 Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
 
-%The \lang syntax of is given in Fig.~\ref{fig:checkc-syn}.
-%There are two type notions in \lang.  Types $\tau$ classify
-%word-sized values including integers and pointers, while types
-%$\omega$ classify multi-word values such as arrays, null-terminated
-%arrays, functions, and single-word-size values. 
-%Pointer types ($\tptr{\omega}{\xi}$) include a pointer mode annotation 
-%($\xi$, the difference between context and pointer modes is introduced shortly below)
-%that is either checked (\cmode), tainted (\tmode), or unchecked (\umode),
-%and a type ($\omega$) denoting valid values that can be pointed to.
-%Array types include both the type of
-%elements ($\tau$) and a bound ($\bvar$) comprised of an upper and
-%lower bound on the size of the array ($(b_l,b_h)$). Bounds $b$ are
-%limited to integer literals $n$ and expressions $x + n$.
-%Whether an array pointer is null terminated or not is determined by annotation
-%$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-%otherwise (we elide $\cdot$ when writing types).
-%
-%\lang{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
-%reflect its dependent function declarations,
-%where $\overline{x}$ represents
-%a list of \tint{} type variables in a dependent function header
-%that bind bound variables appearing in $\overline{\tau}$ and $\tau$.
-%We have a well-formed requirement for a function type;
-%that is, all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
-%Here is the
-%corresponding \systemname syntax for these types:
-%\[\hspace*{-0.5em}
-%\begin{array}{l}
-%\begin{array}{rcl}
-%$\code{_t_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-%&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
-%\\[0.2em]
-%$\code{_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-%&\Leftrightarrow& \tntarrayptr{0}{n}{\tau}{\cmode}
-%\end{array}
-%\\[0.2em]
-%$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau$\code{> : count(}$n$\code{),}$\\
-%\qquad\qquad$\code{_NT_Array_ptr<}$\tau$\code{>)>: count(}$n$\code{))>}$
-%\\[0.2em]
-%\Leftrightarrow\;\; $\tptr{(\tfun{n}{ \tntarrayptr{0}{n}{\tau}{\tmode} \times \tntarrayptr{0}{n}{\tau}{\tmode}}{\tint})}{\tmode}$
-%\end{array}
-%\]
-%As a convention we write $\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean
-%$\tptr{\tarray{0}{b}{\tau}}{\cmode}$, so the firs two examples above could
-%be rewritten $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
-%$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
 \myparagraph{Expressions}
@@
 along with expressions that require special handling, such as,
- static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \cite{li22checkedc}}, the \texttt{strlen} operation ($\estrlen{x}$),
+ static casts ($\ecast{\tau}{e}$),
 memory allocations ($\emalloc{\xi}{\omega}$), 
-function calls ($\ecall{e}{\overline{e}}$),
-unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
-For example, a dynamic bounds cast
-  { \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
-is formalized as 
-{$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \lang. 
-%
-
-% 
+function calls ($\ecall{e}{\overline{e}}$).
 We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
 For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
-The heap allocation $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions, $\cmode$ mode pointer in \cregion or $\umode$ and $\tmode$ mode pointers in \ucregion.
-We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit code regions.
-To guarantee the non-exposure property, we extend the \checkedc syntax to include $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$ blocks,
-where $\overline{x}$ represents all variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain checked pointers.
-$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
-
-% 
+
+Compared with \checkedc, \lang's \texttt{malloc} operation includes a mode $\xi$ to model the allocations of pointers with different modes; $e$ in a function call $\ecall{e}{\overline{e}}$ represents a function pointer whose pointer mode indicates the memory range the function text locates, as well as the context-switch of context regions, explained shortly in \Cref{sec:typechecking}.
+
 \lang{} aims to be simple enough to work with but powerful enough to
 encode realistic \systemname idioms. For example,
-loops can be encoded as recursive function calls. \code{struct}s are
-not included in~\fig{fig:checkc-syn} for space reasons, but they are
-supported as shown in~\cite{li22checkedc}.
+loops can be encoded as recursive function calls. 
+More operations, such as \code{struct}s are given in \Cref{app:main}.
+\ignore{
 C-style \code{union}s have no safe typing in \checkedc, so we omit them.
 Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
+}
 
@@
 E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
-&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid \eif{E}{e}{e}\\[0.2em]
+&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\mid \eif{E}{e}{e}\\[0.2em]
 &&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 

### 2023-03-11 csf2024/formal.tex
update.

@@
 
-Compared with \checkedc, \lang's \texttt{malloc} operation includes a mode $\xi$ to model the allocations of pointers with different modes; $e$ in a function call $\ecall{e}{\overline{e}}$ represents a function pointer whose pointer mode indicates the memory range the function text locates, as well as the context-switch of context regions, explained shortly in \Cref{sec:typechecking}.
+Compared with \checkedc, \lang's \texttt{malloc} operation includes a mode $\xi$ to model the allocations of pointers with different modes; $e$ in a function call $\ecall{e}{\overline{e}}$ represents a function pointer whose pointer mode indicates the memory range the function text locates, as well as the context-switch of context regions, explained shortly in \Cref{sec:semantics}.
 
@@
 More operations, such as \code{struct}s are given in \Cref{app:main}.
-\ignore{
-C-style \code{union}s have no safe typing in \checkedc, so we omit them.
-Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
-}
-
-%\lang expressions include literals ($n\!:\!\tau$), variables ($x$),
-% addition ($\ebinop{e_1}{e_2}$), static casts ($\ecast{\tau}{e}$), 
-%dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \Cref{app:main}},
-%the \texttt{strlen} operation ($\estrlen{x}$),
-%pointer dereference and assignment ($\estar{e}$)
-%and ($\eassign{e_1}{e_2}$), resp.),
-%let binding ($\elet{x}{e_1}{e_2}$),
-%conditionals ($\eif{e}{e_1}{e_2}$),
-%memory allocation ($\emalloc{\xi}{\omega}$), 
-%function calls ($\ecall{e}{\overline{e}}$),
-%unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
-%% \review{III.A.: the "dynamic cast" terminology may be briefly confused for C++'s
-%%   RTTI-based dynamic cast feature}
-%% \mwh{Added reference back to Section 2-B}
-%Integer literals $n$ are annotated with a type $\tau$ which can be either
-%$\tint$, or $\tptr{\omega}{\xi}$ in the case $n$ is being used as 
-%a heap address (this is useful for the semantics);
-%$\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents the $\enull$ pointer, as usual. 
-%The $\texttt{strlen}$ expression operates on variables $x$
-%rather than arbitrary expressions to simplify managing
-%bounds information in the type system; the more general case can be
-%encoded with a \code{let}. We use a less verbose syntax for dynamic bounds
-%casts; e.g., the following %
-%
-%\noindent
-%{\footnotesize
-%\code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{, count(}$n$\code{))}
-%}
-%becomes 
-%{\footnotesize$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$}
-%. 
-%
-%Compared to the former \checkedc model \cite{li22checkedc},
-%there are four differences.
-%First, the \systemname type annotations have well-formed restrictions for maintaining non-exposure.
-%Mainly, in a nested pointer $\tptr{(... \tptr{\tau}{\xi_2} ...)}{\xi_1}$, $\xi_2\le \xi_1$.
-%It is worth noting that pointer modes are a three point partial order ($\le$),
-%where $\tmode$ is the infimum, and $\xi\wedge m$ is a special meet operation that projects pointer modes onto context modes,
-%such that $\tmode$ is projected as $\umode$.
-%Second, $\emalloc{\xi}{\omega}$ includes a mode flag
-%$\xi$ for allocating different pointers in different heaps.
-%We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-%Third, the first expression $e$ in a function call ($\ecall{e}{\overline{e}}$) represents a function pointer.
-%Fourth, $\echeckedtext$ blocks are added to the system, 
-%which permits the nested context-switching between $\echeckedtext$ (represented by context mode $\cmode$)
-%and $\euncheckedtext$ (represented by context mode $\umode$) code regions.
-%One example usage of the nested context-switching is the checked function callbacks inside 
-%an unsafe region in \Cref{lst:humanadjust} and \Cref{lst:humantaint}.
-%To guarantee the non-exposure safety,
-% we extend the $\echeckedtext$ and $\euncheckedtext$ block syntax to be 
-%$\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$:
-%$\overline{x}$ restricts all free variables appearing in $e$, and they cannot be checked pointers.
-%
-%\lang aims to be simple enough to work with, but powerful enough to
-%encode realistic \systemname idioms. For example, mutable local
-%variables can be encoded as immutable locals that point to the heap;
-%the use of \code{&} can be simulated with \code{malloc};
-%and loops can be encoded as recursive function calls. \code{struct}s are
-%not in Fig.~\ref{fig:checkc-syn} for space reasons, but they are
-%actually in our model, and developed in
-%\iftr
-%Appendix~\ref{appx:struct}.
-%\else
-%the supplemental report~\cite{checkedc-tech-report}.
-%\fi
-%C-style \code{union}s have no safe typing
-%in \checkedc, so we omit them.
-
-% \mwh{NOTE: COULD WORK THE FOLLOWING INTO THE ABOVE DESCRIPTION OF
-%   POINTERS: Array and NT-array types have two relative bounds, whose structures
-% can be either an integer or a variable plus an integer. For example,
-% if we have the expression
-% $x\texttt{=}\emalloc{\tarrayb{(0,10)}{\tint}}$, $x$ then has the type
-% $\tptr{\tarrayb{10}{\tint}}{\cmode}$, which represents an array
-% pointer of size $10$. The bounds ($(b_l,b_h)$ in
-% $\evalue{p}{\tptr{\tarrayb{(b_l,b_h)}{\tau}}{m}}$) indicate relative
-% offsets from this pointer of the accessible memory, i.e., $p+b_l$ and
-% $p+b_h$. As such, a pointer can only be directly dereferenced if 0 is
-% included within the annotated range.  If following the $\emalloctext$
-% operation, we execute $*(x-1)$ and $*(x\plus 10)$. The two expressions
-% are not valid, because the type of the two expressions $(x-1)$ and
-% $(x\plus 10)$ are $\tarrayptr{1}{11}{\tint}{\cmode}$ and
-% $\tarrayptr{-10}{0}{\tint}{\cmode}$ and $0$ in these two cases are not
-% in the ranges: $[1,11)$ and $[-10,0)$. Thus, in order for a $\cmode$
-% pointer with type $\tarrayptr{b_l}{b_h}{\tint}{\cmode}$ to be
-% accessible in \checkedc, $0$ must be in the range of $[b_l,b_h)$.}
+
+\subsection{Semantics}\label{sec:semantics}
 
@@
     \begin{array}{lll}
-r & ::= & e \mid \enull \mid \ebounds\\
-E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
-&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\mid \eif{E}{e}{e}\\[0.2em]
-&&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 
-\eunchecked{\overline{x}}{E}
-\mid \echecked{\overline{x}}{E}
-
+\mu & ::= & \evalue{n}{\tau} \mid \bot\\[0.1em]
+e & ::= & \ldots \mid \ret{x}{\mu}{e} \mid \textcolor{red}{\toa{m}{e}}\\[0.1em]
+r & ::= & e \mid \enull \mid \ebounds\\[0.1em]
+E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}
+\mid\eassign{\evalue{n}{\tau}}{E}\\[0.1em]
+&\mid& \eif{E}{e}{e}\mid \elet{x}{E}{e}\mid\ret{x}{\mu}{E}\\[0.1em]
+&\mid& \textcolor{red}{\ecall{E}{\overline{e}}} \mid \textcolor{red}{\ecall{\evalue{n}{\tau}}{\overline{E}}}
+\mid \textcolor{red}{\toa{m}{E}}\\[0.1em]
+\overline{E} &::=& E \mid  \evalue{n}{\tau},\overline{E} \mid \overline{E}, e
 \end{array}
@@
   \begin{mathpar}
-    \inferrule{ m=\mode(E) \\
+    \inferrule[S-Frame]{ m=\mode(E) \\
       e=E[e'] \\
@@
 
-    \inferrule{ \umode=\mode(E) \\
-      e=E[e'] \\
-      \tau=\type(e')}
-    {(\varphi,\heap,e)\longrightarrow_{\umode} (\varphi,\heap,E[\evalue{0}{\tau}])}
+\textcolor{red}{
+    \inferrule[S-Random]{ \umode=\mode(E) \\ \tau=\type(e)}
+    {(\varphi,\heap,E[e])\longrightarrow_{\umode} (\varphi,\heap,E[\evalue{0}{\tau}])}
+}
 
@@
 \end{figure}
+
+The operational semantics for \lang is defined as a small-step
+transition relation with the judgment $ (\varphi,\heap,e)_m
+\longrightarrow (\varphi',\heap',r)_{m'}$. Here, $\varphi$ is a
+\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
+$\heap$ is a \emph{heap} mapping addresses (integer literals) to
+values $\evalue{n}{\tau}$; for both we ensure
+$\fv(\tau)=\emptyset$.
+While heap bindings can change, stack bindings are immutable---once
+variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
+be updated; we can model mutable stack variables as pointers into the
+mutable heap.
+As mentioned, value $\evalue{0}{\tau}$
+represents a $\enull$ pointer when $\tau$ is a pointer type;
+correspondingly, $\heap(0)$ should always be undefined.
+The relation steps to a \emph{result} $r$,
+which is either an expression or a $\enull$ or $\ebounds$ failure,
+representing a null-pointer dereference or out-of-bounds access,
+respectively. Such failures are a \emph{good} outcome; stuck states
+(non-value expressions that cannot transition to a result $r$)
+characterize undefined behavior.
+%
+Context modes $m$ and $m'$ indicate the current and next context regions that $e$ resides.
+
+The rules for the main operational semantics
+judgment---\emph{evaluation}---are given as the two rules in Fig.~\ref{fig:c-context}.
+The first rule takes an expression $e$, decomposes
+it into an \emph{evaluation context} $E$ and a sub-expression $e'$
+(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
+$e$), and then evaluates $e'$ according to the \emph{computation}
+  relation $(\varphi,\heap,e') \longrightarrow_m (\varphi,\heap,e'')$,
+whose rules are given in Fig.~\ref{fig:semantics}, discussed
+shortly.
+
+in contrast to \checkedc formalism \cite{li22checkedc},
+every transition \lang step requires the current context mode $m$ as an input, and the calculation of the next context mode $m'$ based  by the $\nexta$ function in Fig.~\ref{fig:c-context}, which takes $m$ and the context $E$ as inputs.
+The context form $\ecall{(\evalue{n}{\tptr{\tau}{\xi}})}{\overline{e},\Box})$ indicates that the next
+
+In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.
+The second rule describes the exception handling 
+for possible crashing behaviors in $\umode$ regions.
+Operations in $\umode$ region can non-deterministically crash
+and the \systemname sandbox mechanism recovers
+the program to a safe point ($\evalue{0}{\tau}$)
+and continues with the existing program state.
+Evaluation contexts $E$ define a standard left-to-right evaluation order.
+
+% \review{the special case raises questions, e.g. why is this syntax-driven and
+%   not type-driven? }
+% \liyi{This describes the semantic transition rules.
+We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3.
+like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.
+}
+% \mwh{Don't follow the above. I don't see this ``context transition
+%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
+% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
+The $\mode$ function
+determines the mode when evaluating $e'$ based on the context $E$:
+if the $\Box$ occurs within $(\eunchecked{E'})$ inside $E$, then the mode is
+$\umode$; otherwise, it is $\cmode$. Evaluation contexts $E$ define a
+standard left-to-right evaluation order. (We explain the
+$\ret{x}{\mu}{e}$ syntax shortly.)
+% For a term $e$, $C$ is a context if any
+% only if it contains a $\Box$ hole term in one of the subterm position
+% in $e$. $\Box$ indicates the break point where we split $e$ into $C$
+% and a subterm $e'$ such that $C[e'] = e$.  For a function call
+% $\ecall{f}{\overline{e}}$, a valid context ($\overline{C}$) is defined
+% as any one of the elements in the list $\overline{e}$ being a
+% context. We show a rule of the upper level \checkedc semantics. The
+% rule splits a term $e$ into a context $C$ and redex $e'$.  If subterm
+% $e'$ is transitioned to $e_a$, $C[e']$ is transitioned to
+% $C[e_a]$.
+
+\begin{DIFnomarkup}
+\begin{figure*}[t]
+{\small
+  \begin{mathpar}
+        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
+
+        \inferrule[S-DefNTArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h]}
+    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-AssignArr]{\heap(n)=\evalue{n_a}{\tau_a}\\ 0 \in [n_l,n_h) }
+      {(\varphi,\heap,\eassign{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\cmode}}}{\evalue{n_1}{\tau_1}}) \longrightarrow (\varphi,\heap[n \mapsto \evalue{n_1}{\tau}],\evalue{n_1}{\tau})}
+
+    \inferrule[S-Cast]
+              {}
+              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
+
+  \inferrule[S-DynCast]{
+\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{m})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{m} \\ n_l' \le n_l\ \\ n_h \le n_h'}
+    { (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{m}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{m}}}) \longrightarrow (\varphi',\heap',\evalue{n}{\tallarrayptr{n_l'}{n_h'}{\tau_b}{m}})}
+
+ \inferrule[S-DynCastBound]{\varphi(\tptr{\tallarrayb{\bvar}{\tau}}{\cmode})=\tallarrayptr{n_l'}{n_h'}{\tau_b}{\cmode} \\ n_l' > n_l \vee n_h > n_h'}{ (\varphi,\heap,\edyncast{\tptr{\tallarrayb{\bvar}{\tau}}{\cmode}}{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau_a}{\cmode}}})
+ \longrightarrow (\varphi',\heap',\ebounds)}
+
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
+
+    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-Fun]{ \Xi(f) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
+        {(\varphi,\heap,\ecall{f}{{\evalue{\overline{n}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+
+     % \inferrule[S-IfNTF]{\varphi(x)=\evalue{n}{\tau} \\ \heap(n)=0}
+     %      {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_2)}
+      \inferrule[S-StrWiden]{
+         \varphi(x) = \evalue{n}{\tarrayptr{n_l}{n_h}{\tau}{\cmode}}\\ 0 \in [n_l,n_h] \\ n_a > n_h
+\\ \heap(n+n_a) = 0 
+\\\\ (\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i) = \evalue{n_i}{\tau_i} \wedge n_i \neq 0))}
+    {(\varphi,\heap,\estrlen{x}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tarrayptr{n_l}{n_a}{\tau}{\cmode}}],\heap,\evalue{n_a}{\tint})}
+
+     \inferrule[S-IfNTT]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{0}{\tau}{\cmode}} \\ \heap(n)\neq 0}
+    {(\varphi,\heap,\eif{*x}{e_1}{e_2}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tntarrayptr{n_l}{1}{\tau}{\cmode}}],\heap,e_1)}
+
+ % \inferrule[S-IfNTNot]{\varphi(x)=\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}} \\ \heap(n)\neq 0\\ 0 < n_h}
+ %            {(\varphi,\heap,\eif{\estar{x}}{e_1}{e_2}) \longrightarrow (\varphi,\heap,e_1)}
+
+\end{mathpar}
+}
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \tau[\overline{n} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall n_i\in\overline{n}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \Rightarrow \tau[n_i / x_i]\\[0.2em]
+% \mathtt{let}\;\overline{x}=\overline{e}\;\mathtt{in}...\triangleq \mathtt{let}\;x_0=e_0\;\mathtt{in}\;\mathtt{let}\;x_1=e_1\;\mathtt{in}...
+% \end{array}
+% $
+% \end{center}
+% }
+\caption{\lang Semantics: Computation (Selected Rules)}
+\label{fig:semantics}
+\end{figure*}
+\end{DIFnomarkup}
+
+% \review{Fig4: it was hard to tell which cases were stuck states, or could reduce
+%   owing to a rule that was not shown}
+% \mwh{Stuck states are those where the expression is a non-value and
+%   not $\enull$ or $\ebounds$. Updated III-B and III-D.}
+% \review{Fig4: can the rules be presented in the same order they were introduced in
+%   the paper?}
+% \liyi{ Reordered }
+
+% \review{Fig4: S-FUN: $\vec\tau_a$ seems unused; why?}
+% \liyi{the list of $\tau_a$ is a list of input argument types. These
+%   are used during type checking, but not during evaluation (as is
+%   typical).}
+
+% LEO: This has an overfull line...?
+\begin{figure}[t]
+{\small
+{\captionsetup[lstlisting]{margin = 8 mm}
+  \begin{lstlisting}[xleftmargin=8 mm]
+nt_array_ptr<char> safe_strcat
+   (nt_array_ptr<char> dst : count(n),
+    nt_array_ptr<char> src : count(0), int n) {
+  int x = strlen(dst);
+  int y = strlen(src);
+  nt_array_ptr<char> c : count(n) =
+    dyn_bounds_cast
+           <nt_array_ptr<char>>(dst,count(n));
+    // sets c == dst with bound n (not x)
+  if (x+y < n) {
+    for (int i = 0; i < y; ++i)
+      *(c+x+i) = *(src+i);
+    *(c+x+y) = '\0';
+    return dst;
+  }
+  return null;
+}
+  \end{lstlisting}
+}
+}
+\caption{Implementation of safe \code{strcat}}
+\label{fig:strcat-ex}
+\end{figure}
+
+Fig.~\ref{fig:semantics} shows selected rules for the computation
+relation; we explain them with the help of the example in
+Fig.~\ref{fig:strcat-ex},
+  which defines a 
+  safe version of \code{strcat} (using actual Checked C syntax).  The
+  function takes a target 
+  pointer \code{dst} of capacity \code{n}, where the first null
+  character (determined by \code{strlen}) is at index \code{x} where
+  $0 \leq $\code{x}$ \leq n$. It concatenates the \code{src} buffer to
+  the end of \code{dst} as long as \code{dst} has sufficient space.
+
+% Below, we introduce low-level transition semantics for some case operations.
+The design of the low-level individual operation semantics is carefully engineered to perform match our compiler's behavior, such as correctly characterizing the bound widening behaviors for NT-array pointers, even though it is written in terms of fat-pointer formalization.
+
+\myparagraph{Pointer accesses}
+%
+The rules for dereference and assignment operations---\textsc{S-Def},
+\textsc{S-DefNull}, \textsc{S-DefNTArray}, and
+\textsc{S-AssignArr}---illustrate how the semantics checks bounds.
+Rule \textsc{S-DefNull} transitions attempted null-pointer
+dereferences to $\enull$, whereas \textsc{S-Def} dereferences a
+non-null (single) pointer.
+  When $\enull$ is returned by the
+computation relation, the evaluation relation halts the entire
+evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
+does likewise when $\ebounds$ is returned (see below).
+
+% \review{- on page 6, section IIIC, paragraph "Pointer Access" mentions that checked pointers cannot be dereferenced in unchecked blocks - this looks funny, shouldn't it be the other way around?
+The Coq code contains the hypothesis m'=Unchecked -> m Unchecked in various rules of definition well-typed (BoundCheckedC, line 669; rule TyDeref in the code seems closest to the figure's T-DefArr and T-Def in the appendix, although it's a bit concerning that there's no 1-to-1 correspondence of the rules in the code and the paper).
+% }
+% \liyi{Yeah. It is another typo. It should be the other way around.  }
+
+\textsc{S-AssignArr} assigns to an array as long as 0 (the point of
+dereference) is within the bounds designated by the pointer's annotation
+and strictly less than the upper bound. For the assignment rule,
+arrays are treated uniformly whether they are null-terminated or not
+($\kappa$ can be $\cdot$ or $nt$)---the semantics does not search past
+the current position for a null terminator. The program
+can widen the bounds as needed, if they currently precede the null
+terminator: \textsc{S-DefNTArray}, which dereferences an NT array
+pointer, allows an upper bound of $0$, since the program may read, but
+not write, the null terminator. A separate rule (not shown) handles
+normal arrays.
+
+\myparagraph{Casts}
+%
+Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
+by \textsc{S-Cast}. In a type-correct program, such casts are
+confirmed safe by the type system. To evaluate a cast, the rule
+updates the type annotation on $n$. Before doing so, it must
+``evaluate'' any variables that occur in $\tau$ according to their
+bindings in $\varphi$. For example, if $\tau$ was
+$\tarrayptr{0}{x+3}{\tint}{\cmode}$, then $\varphi(\tau)$ would
+produce $\tarrayptr{0}{5}{\tint}{\cmode}$ if $\varphi(x) = 2$.
+
+Dynamic casts are accounted for by \textsc{S-DynCast} and
+\textsc{S-DynCastBound}. In a type-correct program, such casts are
+assumed correct by the type system, and later confirmed by the
+semantics. As such, a dynamic cast will cause a $\ebounds$ failure if
+the cast-to type is incompatible with the type of the target pointer,
+as per the $n_l' > n_l \vee n_h > n_h'$ condition in
+\textsc{S-DynCastBound}. An example use of dynamic casts is given on
+line~7 in Fig.~\ref{fig:strcat-ex}. The values of \code{x} and
+\code{n} might not be known statically, so the type system cannot
+confirm that \code{x <= n}; the dynamic cast assumes this inequality
+holds, but then checks it at run-time.
+
+\myparagraph{Binding and Function Calls}
+%
+The semantics handles variable scopes using the special $\erettext$
+form. \textsc{S-Let} evaluates to a configuration whose stack
+is $\varphi$ extended with a binding for $x$, and whose expression is
+$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
+$\varphi(x)$; if it had no previous binding, $\varphi(x) =
+\bot$. Evaluation proceeds on $e$ until it becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
+binding (or $\bot$) in the new stack, and evaluates to
+$n\!:\!\tau$.
+
+Function calls are handled by \textsc{S-Fun}. Recall that array
+bounds in types may refer to in-scope variables; e.g., parameter
+\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
+2-3 in Fig.~\ref{fig:strcat-ex}. A call to function $f$ causes $f$'s
+definition to be retrieved from $\Xi$,
+% LEO: This is the first occurence of \Xi. It is explained,
+% but reads a bit awkward as it's unclear this is its "introduction"
+which maps function names to
+forms $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e$, where
+$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
+is the parameter list of variables and their types, and $e$ is the
+function body. The call is expanded into a \texttt{let} which binds
+parameter variables $\overline{x}$ to the actual arguments
+$\overline{n}$, but annotated with the parameter types
+$\overline{\tau}$ (this will be safe for type-correct programs). The
+function body $e$ is wrapped in a static cast
+$(\tau[\overline{n} / \overline{x}])$ which is the function's return
+type but with any parameter variables $\overline{x}$ appearing in that
+type substituted with the call's actual arguments $\overline{n}$. To
+see why this is needed, suppose that \code{safe_strcat} in
+Fig.~\ref{fig:strcat-ex} is defined to return a
+\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
+perform a \code{safe_strcat} function call as
+\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
+function returns a value with type \code{nt_array_ptr<int>:count(10)}
+because we substitute bound variable \code{n} in the 
+defined return type with \code{10} from the function call's
+argument list.
+Note that the \textsc{S-Fun} rule replaces the
+  annotations $\overline{\tau_a}$ with
+  $\overline{\tau}$ (after instantiation) from the function's
+  signature. Using $\overline{\tau_a}$ when executing the body of
+the function has no impact on the soundness of \lang, but will violate
+Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
+
+\myparagraph{Bounds Widening}
+%
+Bounds widening occurs when branching on a dereference of a NT array
+pointer, or when performing $\estrlentext$. The latter is most useful
+when assigned to a local variable so that subsequent code can use the
+result, e.g., $e$ in $\elet{x}{\estrlen{y}}{e}$.  Lines 4 and 5 in
+Fig.~\ref{fig:strcat-ex} are examples. 
+The widened upper bound
+precipitated by $\estrlen{y}$ is extended beyond the lifetime of $x$, as long as $y$ is live.
+% Even if $x$ returns, if $y$'s upper bound keeps the widened bound that is caused by $\estrlen{y}$.
+For example, \code{x}'s scope in line 4 at runtime is
+the whole function body in \code{safe_strcat} because the lifetime of the pointer \code{dst} is in the function body.
+% Allowing widening to persist
+% outside the scope is sound, % but would precipitate the need for
+% % fat pointers during compilation, which we want to avoid.
+% and we take this approach to give more precise run-time behavior. We
+% show in Sec.~\ref{sec:compilation}
+This is different from the \checkedc specification, 
+which only allows bound widening to happen within the scope of \code{x}, and restoring old bound values once \code{x} dies.
+We allow widening to persist outside the scope at run-time as long as
+we are within the stack frame, and we show
+this does not necessarily require the use of fat pointers in
+Sec.~\ref{sec:compilation}.
+
+Rule \textsc{S-StrWiden} implements $\estrlentext$ widening. The
+predicate
+$\forall i. n \le i < n+n_a \Rightarrow (\exists n_i\;t_i. \heap(n+i)
+= \evalue{n_i}{\tau_i} \wedge n_i \neq 0))$ aims to find a position
+$n+n_a$ in the NT array that stores a null character, where no
+character as indexes between $n$ and $n+n_a$ contains one. (This rule
+handles the case when $n_a > n_h$, the $n_a \le n_h$ case is handled
+by a normal $\estrlentext$ rule; see 
+\iftr
+Appendix~\ref{sec:rem-semantics}.)
+\else
+the supplemental report~\cite{checkedc-tech-report}.)
+\fi
+
+Rule \textsc{S-IfNTT} performs bounds widening on $x$ when the
+dereference $\estar{x}$ is not at the null terminator, but the
+pointer's upper bound is 0 (i.e., it's at the end of its known
+range). $x$'s upper bound is incremented to 1, and this count persists
+as long as $x$ is live.  For example, \code{s}'s increment (lines
+5--8) is live until the return of the function in
+Fig.~\ref{fig:checkedc-example}; thus, line 11 is valid because
+\code{s}'s upper bound is properly extended. %  For this rule and
 

### 2023-03-27 csf2024/formal.tex
update

@@
 &\mid& \eif{E}{e}{e}\mid \elet{x}{E}{e}\mid\ret{x}{\mu}{E}\\[0.1em]
-&\mid& \textcolor{red}{\ecall{E}{\overline{e}}} \mid \textcolor{red}{\ecall{\evalue{n}{\tau}}{\overline{E}}}
+&\mid& \textcolor{red}{\ecall{E}{\overline{e}}} \mid \textcolor{red}{\ecall{(\evalue{n}{\tau})}{\overline{E}}}
 \mid \textcolor{red}{\toa{m}{E}}\\[0.1em]
@@
 The operational semantics for \lang is defined as a small-step
-transition relation with the judgment $ (\varphi,\heap,e)_m
-\longrightarrow (\varphi',\heap',r)_{m'}$. Here, $\varphi$ is a
+transition relation with the judgment $ (\varphi,\heap,e)
+\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
 \emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
@@
 %
-Context modes $m$ and $m'$ indicate the current and next context regions that $e$ resides.
+The context mode $m$ indicates whether the
+stepped redex within $e$ was in a checked ($\cmode$) or
+unchecked ($\umode$) region.
 
 The rules for the main operational semantics
-judgment---\emph{evaluation}---are given as the two rules in Fig.~\ref{fig:c-context}.
+judgment---\emph{evaluation}---are given at the bottom of
+Fig.~\ref{fig:c-context}.
 The first rule takes an expression $e$, decomposes
@@
 $e$), and then evaluates $e'$ according to the \emph{computation}
-  relation $(\varphi,\heap,e') \longrightarrow_m (\varphi,\heap,e'')$,
+  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
 whose rules are given in Fig.~\ref{fig:semantics}, discussed
-shortly.
-
-in contrast to \checkedc formalism \cite{li22checkedc},
-every transition \lang step requires the current context mode $m$ as an input, and the calculation of the next context mode $m'$ based  by the $\nexta$ function in Fig.~\ref{fig:c-context}, which takes $m$ and the context $E$ as inputs.
-The context form $\ecall{(\evalue{n}{\tptr{\tau}{\xi}})}{\overline{e},\Box})$ indicates that the next
-
-In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.
-The second rule describes the exception handling 
-for possible crashing behaviors in $\umode$ regions.
-Operations in $\umode$ region can non-deterministically crash
-and the \systemname sandbox mechanism recovers
-the program to a safe point ($\evalue{0}{\tau}$)
-and continues with the existing program state.
-Evaluation contexts $E$ define a standard left-to-right evaluation order.
-
-% \review{the special case raises questions, e.g. why is this syntax-driven and
-%   not type-driven? }
-% \liyi{This describes the semantic transition rules.
-We are using context evaluation framework to define the transition rules as the $E$ definition in Fig.3.
-like $\frac{x \Rightarrow y}{x+z \Rightarrow y + z}$, I don't know how type-driven can help us define translation rules.
-}
-% \mwh{Don't follow the above. I don't see this ``context transition
-%   rule'' anywhere, and I'm not sure how it would fire, if we had it.}
-% \liyi{The comment seems to confuse the meaning of the text about the if-then-else rules. Making the rule specific will help. }
+shortly. Compared with \checkedc, we add three marked red contexts in \Cref{fig:c-context}.
+$\ecall{E}{\overline{e}}$ and $\ecall{(\evalue{n}{\tau})}{\overline{E}}$ are the contexts for function calls,
+since every call requires a function pointer having the address $n$ here,
+while the \checkedc formalism's function call takes a global function variable name \cite{li22checkedc}.
+$\toa{m}{E}$ is the context form of mode switching expression $\toa{m}{e}$.
+
 The $\mode$ function
 determines the mode when evaluating $e'$ based on the context $E$:
-if the $\Box$ occurs within $(\eunchecked{E'})$ inside $E$, then the mode is
+if the $\Box$ occurs within $\toa{m}{E'}$ inside $E$ and $E'$ does not contain another $\totext$ expression, then the mode is
 $\umode$; otherwise, it is $\cmode$. Evaluation contexts $E$ define a
-standard left-to-right evaluation order. (We explain the
-$\ret{x}{\mu}{e}$ syntax shortly.)
-% For a term $e$, $C$ is a context if any
-% only if it contains a $\Box$ hole term in one of the subterm position
-% in $e$. $\Box$ indicates the break point where we split $e$ into $C$
-% and a subterm $e'$ such that $C[e'] = e$.  For a function call
-% $\ecall{f}{\overline{e}}$, a valid context ($\overline{C}$) is defined
-% as any one of the elements in the list $\overline{e}$ being a
-% context. We show a rule of the upper level \checkedc semantics. The
-% rule splits a term $e$ into a context $C$ and redex $e'$.  If subterm
-% $e'$ is transitioned to $e_a$, $C[e']$ is transitioned to
-% $C[e_a]$.
+standard left-to-right evaluation order.
+Two expressions $\ret{x}{\mu}{e}$ and $\toa{m}{e}$, which does not appear in source program syntax in \Cref{fig:checkc-syn}, are introduced in evaluating programs.
+The former is used for remembers $x$ was previously bound to
+$\varphi(x)$ in a $\elettext$ binding, which is described in \citet{li22checkedc} and \Cref{app:main}.
+The latter ($\toa{m}{e}$) is a new $\lang$ syntax for switching the context mode in $e$ as $m$, which is generated when a function is called; explained shortly below.
+
+The second rule in \Cref{fig:c-context} is newly introduced in \lang, to help the proof of the non-crashing theorem.
+In any given time, executing a unchecked region might non-deterministically cause the program to crash.
+We assume that there is a recovery mechanism for recover every such crashing, only in unchecked regions, to an initial state $\evalue{0}{\tau}$;
+thus, non-crashing shows that every unchecked region crashing does not affect the program evaluation in the checked regions.
 

### 2023-03-27 sandp23/formal.tex
update

@@
 %   \end{center}
-%   \caption{\lang: Syntax}
+%   \caption{\systemname: Syntax}
 % \end{figure}
@@
 %   \end{center}
-%   \caption{\lang: Typing}
+%   \caption{\systemname: Typing}
 % \end{figure}
@@
               & \bvar & ::= & (b,b) \\
-
      \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{ \xi}
 \\
-\text{Type Flag:}&\kappa &::=& nt \mid \cdot
-\\
-\text{Type:}&\omega &::=& \tau \mid \tallarrayb{\bvar}{\tau} \mid \tfun{\overline{x}}{\overline{\tau}}{\tau}
+\text{Type:}&\omega &::=& \tau \mid \tarrayb{\bvar}{\tau}
 \\
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \edyncast{\tau}{e}
-\mid \ret{x}{\evalue{n}{\tau}}{e}
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid\estar{e}\mid\eassign{e}{e}
   \\
-&&\mid& \estrlen{x} \mid \emalloc{\xi}{\omega} \mid\estar{e}\mid\eassign{e}{e}  \\
-&&\mid& \elet{x}{e}{e} \mid \eif{e}{e}{e} \mid \ecall{e}{\overline{e}}
-\\
-&&\mid&\eunchecked{\overline{x}}{e}
-\mid \echecked{\overline{x}}{e}
+&&\mid& \emalloc{\xi}{\omega} \mid \efree{e} \mid \elet{x}{e}{e}   \\
+&&\mid& \ret{x}{\evalue{n}{\tau}}{e} \mid \eif{e}{e}{e} \\
+&&\mid&\eunchecked{\overline{x}}{e} \mid \echecked{\overline{x}}{e}
 \end{array}
@@
   \]
-  \caption{\lang Syntax}
+  \caption{\systemname Syntax}
   \label{fig:checkc-syn}
@@
 
-This section describes the formal core model of~\systemname,
-named~\lang.
+This section describes the formal core model of~\systemname.
 We present its syntax, semantics, type system,
-as well as \lang’s meta-theories, including the type soundness, non-exposure, and non-crashing theorems.
+as well as \systemname’s meta-theories, including the type soundness and non-exposure theorems.
 
 %This section describes the formal core model of \systemname, named
-%\lang, making precise its syntax, semantics, type system, and compilation. It also
-%develops \lang's meta-theories, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
+%\systemname, making precise its syntax, semantics, type system, and compilation. It also
+%develops \systemname's meta-theories, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
 
 \subsection{Syntax}\label{sec:syntax}
-\Cref{fig:checkc-syn} shows the syntax of~\lang.
+\Cref{fig:checkc-syn} shows the syntax of~\systemname.
 
@@
 otherwise (we elide $\cdot$ when writing types).
-An example representation of an array and null-terminated array in~\lang is shown below:
+An example representation of an array and null-terminated array in~\systemname is shown below:
 % 
@@
   Consider the type ~\code{_t_Array_ptr<_Ptr<int>>}, which describes a tainted array of checked pointers.
-  This is not well-formed in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the tainted ($\tmode$) array is used. 
+  This is not well-formed in~\systemname{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the tainted ($\tmode$) array is used. 
   Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
@@
 
-%The \lang syntax of is given in Fig.~\ref{fig:checkc-syn}.
-%There are two type notions in \lang.  Types $\tau$ classify
+%The \systemname syntax of is given in Fig.~\ref{fig:checkc-syn}.
+%There are two type notions in \systemname.  Types $\tau$ classify
 %word-sized values including integers and pointers, while types
@@
 %
-%\lang{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
+%\systemname{} function types ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$)
 %reflect its dependent function declarations,
@@
 \myparagraph{Expressions}
-\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
+\systemname expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
 pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
@@
 is formalized as 
-{$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \lang. 
+{$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \systemname. 
 %
@@
 % 
-\lang{} aims to be simple enough to work with but powerful enough to
+\systemname{} aims to be simple enough to work with but powerful enough to
 encode realistic \systemname idioms. For example,
@@
 C-style \code{union}s have no safe typing in \checkedc, so we omit them.
-Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
-
-%\lang expressions include literals ($n\!:\!\tau$), variables ($x$),
+Although the base syntax of~\systemname is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
+
+%\systemname expressions include literals ($n\!:\!\tau$), variables ($x$),
 % addition ($\ebinop{e_1}{e_2}$), static casts ($\ecast{\tau}{e}$), 
@@
 %
-%\lang aims to be simple enough to work with, but powerful enough to
+%\systemname aims to be simple enough to work with, but powerful enough to
 %encode realistic \systemname idioms. For example, mutable local
@@
 r & ::= & e \mid \enull \mid \ebounds\\
-E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
-&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid \eif{E}{e}{e}\\[0.2em]
-&&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 
-\eunchecked{\overline{x}}{E}
-\mid \echecked{\overline{x}}{E}
+E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}
+    \mid \eassign{\evalue{n}{\tau}}{E}\\[0.2em]
+&&\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\mid \eif{E}{e}{e}\\[0.2em]
+&& \mid \efree{E}\mid \eunchecked{\overline{x}}{E} \mid \echecked{\overline{x}}{E}
 
@@
 }
-  \caption{\lang Semantics: Evaluation}
+  \caption{\systemname Semantics: Evaluation}
   \label{fig:c-context}
@@
 
-The \lang type system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
+The \systemname type system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
 Our type checker restricts the usage of tainted and checked pointer types to ensure that tainted pointers do not affect checked types, along with enforcing~\checkedc typing rules~\cite{li22checkedc}.
@@
 in context mode $m$, indicating that the code is in $m$ region.
-The operational semantics for \lang is defined as a small-step
+The operational semantics for \systemname is defined as a small-step
 transition relation with the judgment $ (\varphi,\heap,e)
@@
 \myparagraph{Modes, Static Casting, and Subtyping}
-In \lang, Context modes $m$ appearing in a type rule determine the code region 
+In \systemname, Context modes $m$ appearing in a type rule determine the code region 
 that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
@@
 
-\lang also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
+\systemname also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
 an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
 if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2}, i.e., $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
-In \lang, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
+In \systemname, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
 
@@
 The complete subtyping relation was described in \Cref{app:le}.
-Notice that \texttt{let} statements are immutable in \lang, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \lang.
+Notice that \texttt{let} statements are immutable in \systemname, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \systemname.
 
@@
 \textsc{S-DefT}, \textsc{S-DefNull} in \Cref{fig:type-system-1})
-reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
+reflect the key \systemname feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
 The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
 With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a $\enull$ pointer is used,
-\lang captures the runtime error.
+\systemname captures the runtime error.
 Type and semantic rules for array types and pointer assignments are given in \Cref{rem-type,sec:rem-semantics}.
@@
 Here, we discuss our main meta-theoretic results for
-\lang: type soundness (progress and preservation),
+\systemname: type soundness (progress and preservation),
 non-exposure, and non-crashing.
@@
 Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{sec:meta}.
-The progress theorem below states that a \lang program can always make a move.
+The progress theorem below states that a \systemname program can always make a move.
 
@@
 
-For any \lang program $e$, heap $\heap$, stack
+For any \systemname program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 \begin{thm}[Checked Preservation]
-For any \lang program $e$, heap $\heap$, stack
+For any \systemname program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 \begin{thm}[Unchecked Preservation]
-For any \lang program $e$, heap $\heap$, stack
+For any \systemname program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 \begin{thm}[Non-Exposure]
-For any \lang program $e$, heap $\heap$, stack
+For any \systemname program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
@@
 
-\begin{thm}[Non-Crashing]\label{thm:blame} For any \lang
+\begin{thm}[Non-Crashing]\label{thm:blame} For any \systemname
   program $e$, heap $\heap$, stack
@@
 % The semantics
-% gives an independent account of spatial safety in \lang by
+% gives an independent account of spatial safety in \systemname by
 % checking pointer bounds based on the annotations carried on types at
@@
 
-Here, we discuss the \lang operational semantics (\Cref{fig:semantics}); 
+Here, we discuss the \systemname operational semantics (\Cref{fig:semantics}); 
 mainly focusing on the new changes on top of \checkedc in \cite{li22checkedc} with function pointers, modes, and function calls.
@@
 
-The operational semantics for \lang is defined as a small-step
+The operational semantics for \systemname is defined as a small-step
 transition relation with the judgment $ (\varphi,\heap,e)
@@
 
-%\footnote{This approach is that of the PLT Redex model of \lang; the Coq
+%\footnote{This approach is that of the PLT Redex model of \systemname; the Coq
 %development uses a slightly simpler syntax to achieve the same
@@
 %
-In \lang, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
+In \systemname, the type equability ($=_{\Theta}$) and subtype ($\sqsubseteq$) relations are given in \Cref{fig:checkc-subtype}.
 We provide some example descriptions here.
@@
   signature. Using $\overline{\tau_a}$ when executing the body of
-the function has no impact on the soundness of \lang, but will violate
+the function has no impact on the soundness of \systemname, but will violate
 Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
@@
 In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different. 
-\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang).
+\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \systemname).
 The columns represent different pointer modes and the rows represent context modes.
 For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation.
-In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
+In \systemname, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$.
 If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.
 
-We now show how \lang deals with pointer modes, mode switching and function pointer compilations, 
+We now show how \systemname deals with pointer modes, mode switching and function pointer compilations, 
 with no loss of expressiveness
@@
 we present a compilation algorithm that converts from
-\lang to \elang, an untyped language without metadata
+\systemname to \elang, an untyped language without metadata
 annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
@@
 %expressiveness. We present a compilation algorithm that converts from
-%\lang to \elang, an untyped language without metadata
+%\systemname to \elang, an untyped language without metadata
 %annotations. The syntax and semantics \elang
-  %closely mirrors that of \lang; it differs only in that literals lack
+  %closely mirrors that of \systemname; it differs only in that literals lack
   %type annotations and its operational rules perform no
   %bounds and null checks, which are instead inserted during
-  %compilation. Our compilation algorithm is evidence that \lang's
+  %compilation. Our compilation algorithm is evidence that \systemname's
   %semantics, despite its apparent use of fat pointers, faithfully
@@
 
-Compilation is defined by extending \lang's
+Compilation is defined by extending \systemname's
 typing judgment as follows:
@@
 We formalize rules for this judgment in PLT Redex~\cite{pltredex},
-following and extending our Coq development for \lang. To give
+following and extending our Coq development for \systemname. To give
 confidence that compilation is correct, we use Redex's property-based

### 2023-03-29 sandp23/formal.tex
update.

@@
-\section{\systemname Formalism}\label{sec:formal}
+\section{Formalism}\label{sec:formal}
 
@@
 \text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\
-\text{Bound:} & b & ::= & n \mid x \plus n \\
-              & \bvar & ::= & (b,b) \\
-     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{ \xi}
-\\
-\text{Type:}&\omega &::=& \tau \mid \tarrayb{\bvar}{\tau}
+     \text{Type:}& \tau &::=& \tint\mid \tptr{\tau}{ \xi}
 \\
@@
 &&\mid& \ret{x}{\evalue{n}{\tau}}{e} \mid \eif{e}{e}{e} \\
-&&\mid&\eunchecked{\overline{x}}{e} \mid \echecked{\overline{x}}{e}
+&&\mid&\eunchecked{\overline{x}}{e}\mid \echecked{\overline{x}}{e}
 \end{array}
@@
   \]
-  \caption{\systemname Syntax}
+  \caption{\lang Syntax}
   \label{fig:checkc-syn}
@@
 
-This section describes the formal core model of~\systemname.
-We present its syntax, semantics, type system,
-as well as \systemname’s meta-theories, including the type soundness and non-exposure theorems.
+This section describes our formal model of \systemname, called
+\lang, making precise its syntax, semantics, and type system,
+as well as \systemname’s meta-theories, including the type preservation, non-exposure, and clean separation theorems.
+One thing worth noting is that \lang and its meta-theories are built and verified, despite the underling sandbox mechanisms,
+i.e., both implementing both separate heap and WASM sandboxes result in the same properties.
 
@@
 \subsection{Syntax}\label{sec:syntax}
-\Cref{fig:checkc-syn} shows the syntax of~\systemname.
-
-\myparagraph{Type Syntax}
-At a high level, we classify types as word-size value, multi-word
-value, or function types. A word-size value can be either an integer or pointer.
-Every pointer type ($\tptr{\omega}{\xi}$) includes a
+\Cref{fig:checkc-syn} shows the syntax of~\lang, 
+which is based on a subset of the Compcert formalism (C formalism) \cite{Blazy2009}, with some features coming from the \checkedc formalism \cite{li22checkedc}.
+We classify types as integers or pointers.
+Every pointer type ($\tptr{\tau}{\xi}$) includes a
 pointer mode annotation ($\xi$, the difference between context and pointer modes
 is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
-or unchecked (\umode), and a type ($\omega$) denoting the valid value type it is pointed to.
-A multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
-arrays and null-terminated arrays is constructed by the type of elements in the
-array ($\tau$), an array bound ($\bvar$) comprised of an upper and
-lower bound on the size of the array ($(b_l,b_h)$), and an array flag ($\kappa$).
-Bounds $b$ are
-limited to integer literals $n$ and expressions $x + n$.
-Whether an array pointer is null terminated or not is determined by annotation
-$\kappa$, which is $nt$ for null-terminated arrays, and $\cdot$
-otherwise (we elide $\cdot$ when writing types).
-An example representation of an array and null-terminated array in~\systemname is shown below:
-% 
-\[\hspace*{-0.5em}
-\begin{array}{l}
-\begin{array}{rcl}
-$\code{_t_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
-&\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
-\end{array}
-\end{array}
-\]
-
-For simplicity, we write
-$\tptr{\tarrayb{b}{\tau}}{\cmode}$ to mean $\tptr{\tarray{0}{b}{\tau}}{\cmode}$,
-so the above examples could be rewritten as $\tptr{\tarrayb{n}{\tau}}{\cmode}$ and
-$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
+or unchecked (\umode), and a content type ($\tau$) denoting the valid type it is pointed to.
+
+\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
+pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
+ static casts ($\ecast{\tau}{e}$),
+memory allocations ($\emalloc{\xi}{\omega}$) and free ($\efree{e}$), 
+along with expressions that require special handling, such as, tainted blocks ($\eunchecked{\overline{x}}{e}$),
+and checked blocks ($\echecked{\overline{x}}{e}$).
+
+We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
+For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
+The heap allocation $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in $\xi$ memory heap regions, i.e., heaps are separated into three different regions, classified by $\xi$.
+The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit \textit{code} regions,
+where the former represents the $\cmode$ code region in a program that are trusted by users, and the latter represents non-trust $\umode$ code region.
+To guarantee the clean separation property, we extend the C syntax to include $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$ blocks,
+where $\overline{x}$ represents all free variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain checked pointers.
+$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
+
+\ignore{
+\lang{} aims to be simple enough to work with but powerful enough to
+encode realistic \lang idioms. For example,
+loops can be encoded as recursive function calls. \code{struct}s are
+not included in~\fig{fig:checkc-syn} for space reasons, but they are
+supported as shown in~\cite{li22checkedc}.
+C-style \code{union}s have no safe typing in \checkedc, so we omit them.
+Although the base syntax of~\lang is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.}
 
@@
 It prevents unsafe types from being constructed.
-  Consider the type ~\code{_t_Array_ptr<_Ptr<int>>}, which describes a tainted array of checked pointers.
-  This is not well-formed in~\systemname{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the tainted ($\tmode$) array is used. 
-  Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
-%
-Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
-where $\overline{x}$ represents a list of \tint{} type variables that bind type variables appearing in $\overline{\tau}$ and $\tau$.
-An example of a function pointer type is shown below:
-\[\hspace*{-0.5em}
-\begin{array}{l}
-$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau_1$\code{> : count(}$n$\code{),}$\\
-\qquad\qquad$\code{_t_NT_Array_ptr<}$\tau_2$\code{>: count(}$n$\code{), int}$\;n$\code{)>}$
-\\[0.2em]
-\Leftrightarrow\; $\tptr{(\tfun{n}{\tint \times \tntarrayptr{0}{n}{\tau_2}{\tmode} \times \tntarrayptr{0}{n}{\tau_1}{\tmode}}{\tint})}{\tmode}$
-\end{array}
-\]
-The function type also has well-formed requirements (\Cref{app:le}), which disallows nesting checked pointers inside tainted pointers.
-Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+  Consider the type ~\code{_t_ptr<_Ptr<int>>}, which describes a tainted pointer of a checked pointer.
+  This is not well-formed in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the unchecked ($\umode$) pointer is used. 
+  Nevertheless, we can have a checked pointer whose element is a tainted pointers\mzs{,}\mzr{:}~\eg~\code{_ptr<_t_Ptr<int>>} is a valid type.
 
@@
 %$\tptr{\tntarrayb{n}{\tau}}{\cmode}$, respectively.
-\myparagraph{Expressions}
-\systemname expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
-pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
-along with expressions that require special handling, such as,
- static casts ($\ecast{\tau}{e}$), dynamic casts ($\edyncast{\tau}{e}$) \footnote{assumed at compile-time and verified at run-time, see \cite{li22checkedc}}, the \texttt{strlen} operation ($\estrlen{x}$),
-memory allocations ($\emalloc{\xi}{\omega}$), 
-function calls ($\ecall{e}{\overline{e}}$),
-unchecked blocks ($\eunchecked{\overline{x}}{e}$), and checked blocks ($\echecked{\overline{x}}{e}$).
-For example, a dynamic bounds cast
-  { \code{dyn_bounds_cast<_Array_ptr<}$\tau$\code{>>(}$e$\code{,count(}$n$\code{))} }
-is formalized as 
-{$\edyncast{\tptr{\tarrayb{n}{\tau}}{\cmode}}{e}$} in \systemname. 
-%
-
-% 
-We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
-For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
-The heap allocation $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in different regions, $\cmode$ mode pointer in \cregion or $\umode$ and $\tmode$ mode pointers in \ucregion.
-We disallow $\omega$ to be a function type ($\tfun{\overline{x}}{\overline{\tau}}{\tau}$).
-The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit code regions.
-To guarantee the non-exposure property, we extend the \checkedc syntax to include $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$ blocks,
-where $\overline{x}$ represents all variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain checked pointers.
-$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
-
-% 
-\systemname{} aims to be simple enough to work with but powerful enough to
-encode realistic \systemname idioms. For example,
-loops can be encoded as recursive function calls. \code{struct}s are
-not included in~\fig{fig:checkc-syn} for space reasons, but they are
-supported as shown in~\cite{li22checkedc}.
-C-style \code{union}s have no safe typing in \checkedc, so we omit them.
-Although the base syntax of~\systemname is similar to that of~\checkedc model~\cite{li22checkedc}, there are considerable enhancements to support tainted types ($t$ in $\xi$), special heap handling (\ie $\emalloc{\xi}{\omega}$), and explicit specification of pointers in the $\echeckedtext$ and $\euncheckedtext$ regions.
 
@@
 {\small
-$\hspace*{-1.2em}
+$\hspace*{-1.5em}
     \begin{array}{l}
@@
     \mid \eassign{\evalue{n}{\tau}}{E}\\[0.2em]
-&&\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\mid \eif{E}{e}{e}\\[0.2em]
-&& \mid \efree{E}\mid \eunchecked{\overline{x}}{E} \mid \echecked{\overline{x}}{E}
+&&\mid \efree{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\\[0.2em]
+&& \mid \eif{E}{e}{e} \mid \eunchecked{\overline{x}}{E} \mid \echecked{\overline{x}}{E}
 
@@
 }
-  \caption{\systemname Semantics: Evaluation}
+  \caption{\lang Semantics: Evaluation}
   \label{fig:c-context}
@@
 
-\subsection{Typing and Semantics}
+\subsection{Semantics}
 \label{sec:typechecking}
@@
 {\small
-  \begin{mathpar}
-     \inferrule[T-CastPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
-
-    \inferrule[S-Cast]
-              {}
-              {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
-  \end{mathpar}
-
+$\textcolor{blue}{\text{Dereference Operations:}}\\$
   \begin{mathpar}
     \inferrule[T-Def]
-              {\xi \leq m \\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-\quad
-        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
-         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
+              {\xi \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\xi}}
+              {\Gamma \vdash_m \estar{e} : \tau}
+
+        \inferrule[S-DefT]{\heap(\tmode,n)=\evalue{n_a}{\tau_a} \\ n \not\in \Theta \\ n < |\heap(\tmode)|}
+    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
+
+        \inferrule[S-Def]{\xi \neq \tmode \\ \heap(\xi,n)=\evalue{n_a}{\tau_a} }
+    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefNull]{}{(\varphi,\Theta,\heap,\estar{\evalue{0}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\enull)}
+
+    \inferrule[S-DefBound]{n \in \Theta \vee n \ge |\heap(\tmode)|}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
   \end{mathpar}
 
+$\textcolor{blue}{\text{Malloc and Free:}}\\$
+  \begin{mathpar}
+    \inferrule[T-Malloc]
+              {\xi \leq m}
+              {\Gamma \vdash_m \emalloc{\xi}{\tau} : \tptr{\tau}{\xi}}
+
+    \inferrule[T-Free]
+              {\xi \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\xi}}
+              {\Gamma \vdash_m \efree{e} : \tint}
+
+  \inferrule[S-Malloc]{\mathtt{alloc}(\heap,\Theta,\xi,\tau)=(n,\heap')}
+   { (\varphi,\Theta,\heap,\emalloc{\xi}{\tau}) \longrightarrow (\varphi,\Theta',\heap',\evalue{n}{\tptr{\tau}{\xi}})}
+
+  \inferrule[S-Free]{\xi \neq \tmode}
+   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{0}{\tint})}
+
+  \inferrule[S-FreeT]{}
+   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\{n\}\cup\Theta,\heap,\evalue{0}{\tint})}
+
+  \end{mathpar}
+
+$\textcolor{blue}{\text{Checked and Unchecked Context Switches:}}\\$
   \begin{mathpar}
@@
               {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
+                     \\\\\fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
               {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
@@
               {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
+                \\\\ \fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
               {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
@@
     \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-  \end{mathpar}
-
-  \begin{mathpar}
-\inferrule[T-Fun]
-    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
-         \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
-         \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
-             \overline{\tau'} \sqsubseteq_{\Theta}
-               \overline{\tau}[\overline{e'} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
-  \end{mathpar}
-  \begin{mathpar}
-    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-\quad
-    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
-                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-
   \end{mathpar}
@@
 {\footnotesize
+\begin{center}
 $
+\begin{array}{l}
+\tmode \le \xi
+\qquad
+m \le m
+\\[0.2em]
 \cmode(\tint)=\texttt{false}
@@
 \cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
+\end{array}
 $
-}
-\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.
-First line is for cast operations, second line is for pointer dereferences, third line is for checked/unchecked blocks, and the rest is for function calls.}
+\end{center}
+}
+\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.}
 \label{fig:type-system-1}
@@
 
-The \systemname type system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
-Our type checker restricts the usage of tainted and checked pointer types to ensure that tainted pointers do not affect checked types, along with enforcing~\checkedc typing rules~\cite{li22checkedc}.
+The \lang type system is a flow-sensitive, gradual type one that generates additional dynamic checks that are inserted in the typing checking stage and executed in the semantic evaluation stage.
+Our type checker restricts the usage of tainted pointer types to ensure that tainted pointer uses in the $\umode$ regions do not affect the $\cmode$ region execution.
 
 As partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{T-}$X$),
-each typing judgment has the form $\Gamma;\Theta\vdash_m e : \tau$,
+each typing judgment has the form $\Gamma\vdash_m e : \tau$,
 which states that in a type environment $\Gamma$ (mapping variables to
-their types) and a predicate environment $\Theta$ (mapping integer-typed
-variables to Boolean predicates), expression $e$ will have type $\tau$ if evaluated
+their types), expression $e$ will have type $\tau$ if evaluated
 in context mode $m$, indicating that the code is in $m$ region.
-The operational semantics for \systemname is defined as a small-step
-transition relation with the judgment $ (\varphi,\heap,e)
-\longrightarrow_m (\varphi',\heap',r)$, as shown in \Cref{fig:c-context}.
+The operational semantics for \lang is defined as a small-step
+transition relation with the judgment $ (\varphi,\Theta,\heap,e)
+\longrightarrow_m (\varphi',\Theta',\heap',r)$, as shown in \Cref{fig:c-context}.
  Here, $\varphi$ is a
-\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
-$\heap$ is a \emph{heap} that is partitioned into two parts ($\cmode$ and $\umode$ heap regions), each of which
+\emph{stack} mapping from variables to values $\evalue{n}{\tau}$,
+$\Theta$ is a set of $\tmode$ mode memory addresses that are freed at the point of executing $e$,
+and $\heap$ is a \emph{heap} that is partitioned into three parts ($\cmode$, $\tmode$, and $\umode$ heap regions), each of which
 maps addresses (integer literals) to values $\evalue{n}{\tau}$.
-The complete set of typing rules and special handling of (NT)-arrays are provided in~\Cref{rem-type,sec:rem-semantics}.
-
-We wrote $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ heap,
-and $\heapup{m}{n}{\evalue{n'}{\tau}}$ 
-to update location $n$ with the value $\evalue{n'}{\tau}$ in the $m$ heap.
+We wrote $\heap(m)$ to retrieve the $m$ heap and $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ heap.
+The complete set of typing rules are provided in~\Cref{rem-type,sec:rem-semantics}.
+
 While heap bindings can change, stack bindings are immutable---once
@@
 expression, a $\enull$ or $\ebounds$ failure, represent an expression right
-  after the reduction, a null-pointer dereference or out-of-bounds access,
+  after the reduction, a null-pointer dereference or heap-location-not-existence access,
 respectively.
@@
 determines the context mode, i.e., region, that the expression $e'$ locates based on the context $E$.
+If the context hole locates in an $\euncheckedtext$, i.e., $E[\eunchecked{\Box}]$, the context region mode is $\umode$,
+while if the hole locates in a $\echeckedtext$, i.e., $E[\echecked{\Box}]$, its context mode is $\cmode$.
 In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.
+
 The second rule describes the exception handling 
@@
 Operations in $\umode$ region can non-deterministically crash
-and the \systemname sandbox mechanism recovers
+and the \lang sandbox mechanism recovers
 the program to a safe point ($\evalue{0}{\tau}$)
@@
 
-\myparagraph{Modes, Static Casting, and Subtyping}
-In \systemname, Context modes $m$ appearing in a type rule determine the code region 
+In \lang, Context modes $m$ appearing in a type rule determine the code region 
 that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
@@
 This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can only perform such operations in $\cmode$ and $\umode$ regions, respectively.
-%Pointer modes are also useful in determining if a nested pointer has a valid type.
-For example, in a nested pointer $\tptr{(...
-\tptr{\tau}{\xi_2} ...)}{\xi_1}$, we require $\xi_2\le \xi_1$ to maintain non-exposure.
-
-\systemname also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
+\ignore{
+In \lang, we also provide a static casting operation in \Cref{rem-type},
+
+\lang also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
 an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
-if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2}, i.e., $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
-In \systemname, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
-
-\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-//_Ptr<int> x; _t_Ptr<int> y; int *z;
-z = (int *)y; // This is okay.
-x = (_Ptr<int>)y; // Not allowed.
-\end{minted}
-
-In the above example, a $\tmode$ mode pointer can be cast to $\umode$ mode but casting $\tmode$ mode to $\cmode$ mode is disallowed.
-The complete subtyping relation was described in \Cref{app:le}.
-Notice that \texttt{let} statements are immutable in \systemname, so the following code is not possible, because variables \code{x} and \code{y} must have the same type in \systemname.
-
-\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-//_Ptr<int> x;  _t_Ptr<int> y;
-x = y; // Not allowed.
-\end{minted}
+if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2}, i.e., $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.}
 
 \myparagraph{Pointer Dereference}
-The type and semantic rules for pointer dereference (\textsc{T-Def}, \textsc{S-DefC},
-\textsc{S-DefT}, \textsc{S-DefNull} in \Cref{fig:type-system-1})
-reflect the key \systemname feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
+The type and semantic rules for pointer dereference (\rulelab{T-Def}, \rulelab{S-DefT},
+\rulelab{S-Def}, \rulelab{S-DefBound}, \rulelab{S-DefNull} in \Cref{fig:type-system-1})
+reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
 The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
-With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a $\enull$ pointer is used,
-\systemname captures the runtime error.
-Type and semantic rules for array types and pointer assignments are given in \Cref{rem-type,sec:rem-semantics}.
-
-Rules \textsc{S-DefC} and \textsc{S-DefT} are for $\cmode$ and $\tmode$ mode pointer dereferences, respectively.
-In addition to the no $\enull$ check in $\cmode$ mode pointer dereference,
+
+Rules \rulelab{S-DefT} and \rulelab{S-Def} are for $\tmode$ mode and other mode pointer dereferences, respectively.
+Operation $\heap(\xi,n)$ not only accesses the $n$-th address location,
+but also performs a $\enull$ check because $0$ is undefined in all heaps.
+In addition to the no $\enull$ check in rule \rulelab{S-Def},
+we perform checks in \rulelab{S-DefT} to ensure the existence of $\tmode$ mode pointers.
 any dynamic heap access of a tainted ($\tmode$) pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to that the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
+Predicate $n \not\in \Theta$ means that location $n$ is not freed in the $\tmode$ heap region,
+while $n < |\heap(\tmode)|$ makes sure that $n$ does not go out of the high bound of the $\tmode$ heap region,
+where $|\heap(\tmode)|$ refers to the cardinality of the $\tmode$ heap region.
+If dynamic checks are violated, rules \rulelab{S-DefNull} and \rulelab{S-DefBound} are performed, meaning that the evaluation is transitioned into a runtime error that is correctly captured by the \systemname compiler.
+
+\myparagraph{Malloc and Free}
+The second block (\rulelab{T-Malloc}, \rulelab{T-Free},
+\rulelab{S-Malloc}, \rulelab{S-Free}, \rulelab{S-FreeT}) in \Cref{fig:type-system-1} defines
+the type and semantic rules for $\texttt{malloc}$ and $\texttt{free}$ operations.
+Rules $\rulelab{T-Malloc}$ and $\rulelab{T-Free}$ make sure that $\cmode$ and $\umode$ mode pointers are allocated and freed only in the $\cmode$ and $\umode$ regions, respectively, while $\tmode$ mode pointers are allowed to allocate and free in both regions.
+The $\mathtt{alloc}(\heap,\Theta,\xi,\tau)$ function in rule \rulelab{S-Malloc} allocates a new pointer in the $\xi$ heap region.
+If $\xi$ is $\tmode$ and $\Theta$ is not empty, the function picks a memory space in $\Theta$ as the new pointer;
+otherwise, the function picks the $|\heap(\xi)|$ as the new pointer and increments $\heap(\xi)$'s cardinality.
+Rules \rulelab{S-FreeT} and \rulelab{S-Free} define the \texttt{free} operation for $\tmode$ mode and other mode pointers, respectively.
+Rule \rulelab{S-Free} "forgets" about the freed pointer, while \rulelab{S-FreeT} puts the freed $\tmode$ mode pointer address in the $\Theta$ set.
 
@@
 with its type and semantic rules given in \Cref{fig:type-system-1}.
-In this context switching, to guarantee the checked ($\cmode$) pointer non-exposure property, 
+In this context switching, to guarantee the checked ($\cmode$) pointer non-exposure property, i.e.,
 checked pointers are not allowed to go cross different regions, which is guaranteed by the predicates 
-$\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$,
-as well as the check that all free variables in the block content $e$ are in $\overline{x}$.
+$\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$ \footnote{$\cmode(-)$ checks the mode of a pointer.},
+as well as the check that all free variables in the block content $e$ are in $\overline{x}$ ($\fv(e)\subseteq\overline{x}$).
 For example, \code{StringAuth} in \Cref{subsub:gencregion} is a trampoline function that disallows
@@
 cannot legally acknowledges any checked pointers; otherwise, we might expose a checked pointer address to unsafe code regions.
-In \systemname, we actually permits the accesses of checked pointers inside \code{StringAuth},
+In \lang, we actually permits the accesses of checked pointers inside \code{StringAuth},
 since the function body of a trampoline function is in $\cmode$ region.
 More information is given in \Cref{subsub:gencregion}.
-
-\myparagraph{Dependent Function Pointers} %
-Rule \textsc{T-Fun} (\Cref{fig:type-system-1}) states the type judgment for
-dependent function pointer application, where we represent the result of
-replacing all integer bound variables $\overline{x}$ in the type \(\tau\) with
-with bound expressions $\overline{e'}$ by $\tau[\overline{e'} / \overline{x}]$
-and write $\overline{\tau}[\overline{e'} / \overline{x}]$ to lift the
-substitution to every type in \(\overline{\tau}\).
-% 
-Given an expression $e$ of function pointer type
-($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$) and arguments
-$\overline{e}$ of types $\overline{\tau'}$,
-% 
-the result of the application will
-be of type $\tau[\overline{e'} / \overline{x}]$;
-if for each pair of \(\tau'\) and \(\tau''\) in \(\overline{\tau'}\) and
-$\overline{\tau}[\overline{e'} / \overline{x}]$, \(\tau'\) is a subtype of
-\(\tau''\).
-Consider the \code{process_req2} function in
-Fig.~\ref{lst:final}, whose parameter type for \code{msg} 
-depends on \code{m_1}.
-Its function pointer type is 
-$\tptr{\tfun{\code{m_1}}{\tint,\tntarray{0}{\code{m_1}}{\texttt{char}}}{\tint}}{\tmode}$.
-In \code{handle_request}, the call \code{process_req2(buff, r_len)} binds variable \code{m_1} to \code{r_len}.
-After the call returns, \code{m_1}'s scope is ended, so we need to substitute it with \code{r_len} in the final return type
-because it might contain \code{m_1}.
-
-\textsc{S-FunC} and \textsc{S-FunT} define the semantics
-for $\cmode$ and $\tmode$ mode function pointers, respectively. 
-A call to a function pointer $n$ retrieves
- the function definition in $n$'s location in the global function store $\Xi$,
-which maps function pointers to
-function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
-$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
-is the parameter list of variables and their types, 
-$\xi$ determines the mode of the function, and $e$ is the
-function body. 
-Similar to \heap, the global function store $\Xi$ is also partitioned into
-two parts ($\cmode$ and $\umode$ store regions), each of which
-maps addresses (integer literals) to the function data described above.
-Rule \textsc{S-FunT} defines the tainted version of function call
-with the verification process 
-$\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}$
-makes sure that the function in the global store is well-defined and has the right type.
 
@@
 Here, we discuss our main meta-theoretic results for
-\systemname: type soundness (progress and preservation),
-non-exposure, and non-crashing.
+\lang: type preservation,
+non-exposure, and clean separation.
 These proofs have been conducted in our Coq model.
-Type soundness relies on several \emph{well-formedness} given in \cite{li22checkedc} and \Cref{sec:meta}.
-The progress theorem below states that a \systemname program can always make a move.
-
-\begin{thm}[Progress]\label{thm:progress}
-
-For any \systemname program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$),
-one of the following holds:
-
-\begin{itemize}
-
-\item $e$ is a value ($\evalue{n}{\tau}$).
-
-\item there exists $\varphi'$ $\heap'$ $r$, such that $(\varphi,\heap,e) \longrightarrow_m (\varphi',\heap',r)$.
-
-\end{itemize}
-\end{thm}
-%{\em Proof:} By induction on the typing derivation.
-
-\noindent
-There are two forms of preservation regarding the $\cmode$ and $\umode$ regions.
-Checked Preservation states that a reduction step preserves both the
-type and consistency of the program being reduced, while
-unchecked Preservation states that any evaluation happens at $\umode$ region does not affect the $\cmode$ mode heap.
-
-\begin{thm}[Checked Preservation]
-For any \systemname program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed, consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$) and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\cmode} (\varphi',\heap',e')$, then $\heap'$ is
-$\cmode$ region consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
-$\Gamma'$ and $\tau'$ that are well formed, $\cmode$ region consistent
-($\Gamma';\Theta\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
-well typed ($\Gamma';\Theta \vdash_{\cmode} e: \tau'$), where
+Type preservation relies on several \emph{well-formedness} given in \Cref{sec:meta}.
+The \lang does not have type progress, referring to that a well-typed program can always make a move, 
+because C does not have type progress.
+However, \lang has type preservation, in the sense that a reduction step preserves both the
+type and consistency of the program being reduced.
+
+\begin{thm}[Type Preservation]
+For any \lang program $e$, heap $\heap$, stack
+$\varphi$, type environment $\Gamma$ being well-formed and consistent
+($\Gamma\vdash \varphi$ and $\heap \vdash \varphi$) and well
+typed ($\Gamma\vdash_{m} e : \tau$ for some $\tau$), if there exists $\varphi'$, $\Theta'$
+$\heap'$ and $e'$, such that $(\varphi,\Theta,\heap,e)
+\longrightarrow_{m} (\varphi',\Theta',\heap',e')$, then $\heap'$ is
+consistent with $\heap$ ($\heap \triangleright \heap'$) and there exists
+$\Gamma'$ and $\tau'$ that are well formed, consistent
+($\Gamma'\vdash \varphi'$ and $\heap' \vdash \varphi'$) and
+well typed ($\Gamma' \vdash_{m} e: \tau'$), where
 $\tau'\sqsubseteq_{\Theta} \tau$.
@@
 %\smallskip
-\begin{thm}[Unchecked Preservation]
-For any \systemname program $e$, heap $\heap$, stack
+
+We first show the non-exposure theorem,
+where code in $\umode$ region cannot observe a valid checked ($\cmode$) pointer address.
+
+\begin{thm}[Non-Exposure]
+For any \lang program $e$, heap $\heap$, stack
 $\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
 that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$, then $\heap'(\cmode)=\heap(\cmode)$.
-\end{thm}
-
-Using the above theorems, we first show the non-exposure theorem,
-where code in $\umode$ region cannot observe a valid checked ($\cmode$) pointer address.
-
-\begin{thm}[Non-Exposure]
-For any \systemname program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
-that are all are well-formed and well
-typed ($\Gamma;\Theta\vdash_{\cmode} e : \tau$ for some $\tau$), if there exists $\varphi'$,
-$\heap'$ and $e'$, such that $(\varphi,\heap,e)
-\longrightarrow_{\umode} (\varphi',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
+typed ($\Gamma\vdash_{m} e : \tau$ for some $\tau$), if there exists $\varphi'$, $\Theta'$,
+$\heap'$ and $e'$, such that $(\varphi,\Theta,\heap,e)
+\longrightarrow_{\umode} (\varphi',\Theta',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
 where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
@@
 
-We now state our main result, {\em non-crashing},
-which suggests that a well-typed program can never be \emph{stuck} (expression
-$e$ is a non-value that cannot take a step\footnote{Note that
-  $\ebounds$ and $\enull$ are \emph{not} stuck expressions---they represent a
-  program terminated by a failed run-time check. A program that tries to access $\heap{n}$
-  but $\heap$ is undefined at $n$ will be stuck, and violates spatial
-  safety.}).
-
-% \review{- There appears to be a slight discrepancy between the blame theorem in Coq and the one in the paper: the paper mentions some e', which I believe should be r.
-Also, the Coq code has a further disjunct m=Unchecked in the conclusion.}
-% \liyi{It is a typo.
-We will add the thing back that we show that either user uses a unchecked mode to evaluate $e$ or $e$ lives in a context that is an unchecked region.
-This is a bit due to the space limitation.
-The semantic rules allow users to input the mode $m$ of evaluating an expression, I just forgot to include the $m$ in the result of the proof statement.
-}
-
-\begin{thm}[Non-Crashing]\label{thm:blame} For any \systemname
+The essence of non-exposure represents how trampoline functions are organized to support the callback mechanism in \systemname.
+
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+// u region code
+int func_t(_t_Ptr<char> p : count(5)) {
+  checked (p) { //c region code
+   char * q;
+   * q = * p;
+   func(q)
+   ... 
+  }
+}
+\end{minted}
+
+In the code above, we want to call a $\cmode$ function~\inlinecode{func} in line 7, but the calling happens in a $\umode$ region, which does not allow to touch any $\cmode$ region code.
+We provide a trampoline function that provides a tainted shell \inlinecode{func_t} for \inlinecode{func}.
+In \systemname, we generate these shells for every $\cmode$ region functions, and perform a dynamic check to disallow any function calls to $\cmode$ region functions in $\umode$ region.
+
+We define a state to be \emph{stuck} and \emph{critically stuck} below, and then show our main result, {\em clean separation},
+which suggests that a well-typed program can never be critically stuck.
+
+\begin{defi}[Critically Stuck]\label{type-stuck}
+For a program $e$ and environment tuple $(\varphi,\Theta,\heap,e)$ and a context mode $m$, we define it to be stuck as that there is no transition tuple $(\varphi',\Theta',\heap',r)$, such that $(\varphi,\Theta,\heap,e) \longrightarrow_m (\varphi',\Theta',\heap',r)$;
+it is critically stuck, when $(\varphi,\Theta,\heap,e)$ is stuck, and $e$ is of the three situations:
+\begin{itemize}
+
+\item $e=\estar{\evalue{n}{\tptr{\tau}{\tmode}}}$.
+
+\item $e=\eassign{\evalue{n}{\tptr{\tau}{\tmode}}}{\evalue{n'}{\tptr{\tau}{m'}}}$,
+and $\estar{\evalue{n'}{\tptr{\tau}{m'}}}$ can make a move.
+
+\item $e=\eassign{\evalue{n}{\tptr{\tau}{m}}}{\evalue{n'}{\tptr{\tau}{\tmode}}}$,
+and $\estar{\evalue{n}{\tptr{\tau}{m}}}$ can make a move.
+
+\end{itemize}
+\end{defi}
+
+\begin{thm}[Clean Separation]\label{thm:blame} For any \lang
   program $e$, heap $\heap$, stack
-$\varphi$, type environment $\Gamma$, and variable predicate set $\Theta$
+$\varphi$, type environment $\Gamma$, and set $\Theta$
 that are well-formed and consistent
-($\Gamma;\Theta\vdash \varphi$ and $\heap \vdash \varphi$),
-if $e$ is well-typed ($\varphi;\Theta\vdash_{\cmode} e :
+($\Gamma\vdash \varphi$ and $\heap \vdash \varphi$),
+if $e$ is type-preserved ($\varphi\vdash_{m} e :
 \tau$ for some $\tau$) and there exists
-$\varphi_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
-$(\varphi,\heap,e) \longrightarrow_{m_1} (\varphi_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\heap_k,r)$, then $r$ can never be \emph{stuck}.
+$\varphi_i$, $\Theta_i$, $\heap_i$, $e_i$, and $m_i$ for $i\in [1,k]$, such that
+$(\varphi,\Theta,\heap,e) \longrightarrow_{m_1} (\varphi_1,\Theta_1,\heap_1,e_1)\longrightarrow_{m_2} ...\longrightarrow_{m_k} (\varphi_k,\Theta_k,\heap_k,r)$, then $r$ can never be \emph{critically stuck}.
 \end{thm}
 
-%{\em Proof:} By induction on the number of steps of the \checkedc
-%evaluation ($\longrightarrow_m^*$), using progress and preservation to
-%maintain the invariance of the assumptions.
+Clean separation suggests that $\cmode$ and $\umode$ code regions are completely separated, because the tainted pointers, the only types of pointers that communicate the $\cmode$ and $\umode$ code regions, do not cause any problem in $\cmode$ code region.
+Any bugs in the region are caused by $\cmode$ mode pointers that are only allowed in the $\cmode$ code region.
+If we allow \checkedc as the $\cmode$ code region, which admits the blame theorem, i.e., any bug happening in a $\cmode$ code region is because of a $\umode$ code region bug, the system then admits the non-crashing property, i.e., $\cmode$ code regions never crash.
 

### 2023-03-30 csf2024/appendix.tex
update sandp.

@@
 
-    \inferrule[S-DefArrayC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
+        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} }
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
+
+        \inferrule[S-DefNTArray]{\heap(n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h]}
     {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\cmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-
-    \inferrule[S-DefArrayT]{\heap(\umode,n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h) 
-               \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}}
-    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
 
@@
 
-\begin{definition}[Type Environment Well-formedness]\label{type-wellformed}
+\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
 A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
-\end{definition}
-
-\begin{definition}[Heap Well-formedness]
+\end{defi}
+
+\begin{defi}[Heap Well-formedness]
 For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
@@
 contains no free variables. 
-\end{definition}
-
-\begin{definition}[Stack Well-formedness]
+\end{defi}
+
+\begin{defi}[Stack Well-formedness]
 A stack snapshot $\varphi$ is well-formed if
@@
 contains no free variables. 
-\end{definition}
+\end{defi}
 
@@
 
-\begin{definition}[Stack Consistency]
+\begin{defi}[Stack Consistency]
 A type environment $\Gamma$, variable predicate set $\Theta$, and
@@
 $\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
-\end{definition}
-
-\begin{definition}[Checked Stack-Heap Consistency]
+\end{defi}
+
+\begin{defi}[Checked Stack-Heap Consistency]
 A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
 for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
-\end{definition}
-
-\begin{definition}[Checked Heap-Heap Consistency]
+\end{defi}
+
+\begin{defi}[Checked Heap-Heap Consistency]
 A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
 for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
-\end{definition}
+\end{defi}
 

### 2023-04-01 csf2024/formal.tex
update sandp.
update semantics.

--- 2023-03-08 csf2024/formal.tex
@@
-\section{\systemname Formalism}\label{sec:formal}
-
-% \begin{itemize}
-% \item Describe the types for checked-C as a graph. \dvh{I don't know what this means.}
-
-% \item Describe the subtyping for checked-C types.
-% \begin{itemize}
-% \item State that the subtypes in Checked-C are transitive.
-% \end{itemize}
-
-% \item Describe the syntax of Checked-C
-
-% \item Describe the semantics of Checked-C
-
-% \item Describe the type system of Checked-C
-
-% \item Describe the progress and preservation theorems, and outline the proofs.
-
-% \item Describe the blame theorem and proof.
-
-% \end{itemize}
-
-% \ignore{
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[height=6in]{syntax.pdf}
-%   \end{center}
-%   \caption{\lang: Syntax}
-% \end{figure}
-
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[height=6in]{types.pdf}
-%   \end{center}
-%   \caption{\lang: Typing}
-% \end{figure}
-% }
-% \liyi{main text begins here. }
-
-% \review{
-% While reading the semantics, I found the fact that S-Def and S-DefNull are
-%   applicable non-deterministically if n is 0 a bit confusing. Only when
-%   reading the meta-theory section I realized that this is not a concrete issue
-%   because well-formed heaps are such that $\mathcal{H}(0)$ is never defined. It
-%   might be worth pointing this out early on. }
-% \mwh{Done.}
+\section{Formalism}\label{sec:formal}
 
@@
 \text{Context Mode:} & m & ::= & \cmode \mid \umode \\[0.2em]
-\text{Memory Mode:} & \xi & ::= & m \mid \tmodered \\[0.2em]
+\text{Memory Mode:} & \kappa & ::= & m \mid \tmodered \\[0.2em]
 \text{Bound:} & b & ::= & n \mid x \plus n \\[0.2em]
@@
 
-     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{\xi}
+     \text{Word Type:}& \tau &::=& \tint\mid \tptr{\omega}{\kappa}
 \\[0.2em]
@@
 \text{Expression:}& e & ::= & 
-\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e}
+\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid \estar{e}\mid\eassign{e}{e} 
   \\[0.2em]
-&&\mid& \estar{e}\mid\eassign{e}{e} \mid \textcolor{red}{\emalloc{\xi}{\omega}} \\[0.2em]
+&&\mid& \textcolor{red}{\emalloc{\kappa}{\omega}} \mid \textcolor{red}{\toa{m}{\overline{x}}{e}} \\[0.2em]
 &&\mid&\eif{e}{e}{e} \mid  \elet{x}{e}{e} \mid \textcolor{red}{\ecall{e}{\overline{e}}}
@@
 \hline
-\umode&$\umode$ & $\tmodenor$,$\umode$ & $\tmodenor$,$\umode$ & $\tmodefun$,$\umode$\\[0.1em]
+\umode&$\tmodenor$,$\umode$ & $\tmodenor$,$\umode$ & $\tmodenor$,$\umode$ & $\tmodefun$,$\umode$\\[0.1em]
 \hline
@@
   \inferrule[]
-  {\xi \wedge m\vdash \tau \\ \xi \le m}
-  {m \vdash \tptr{\tarrayb{\bvar}{\tau}}{\xi}}
+  {\kappa \wedge m\vdash \tau \\ \kappa \le m}
+  {m \vdash \tptr{\tarrayb{\bvar}{\tau}}{\kappa}}
 \qquad
   \inferrule[]
-  {\xi \wedge m \vdash \tau\\ \xi \le m}
-  {m \vdash \tptr{\tau}{\xi}}
+  {\kappa \wedge m \vdash \tau\\ \kappa \le m}
+  {m \vdash \tptr{\tau}{\kappa}}
 
   \inferrule[]
-  {\forall \tau'\in \overline{\tau}\cup\{\tau\}\,.\,\xi \wedge m \vdash \tau'\\ \xi \le m \\ \fv(\overline{\tau})\cup\fv(\tau)\subseteq \overline{x}}
-  {m \vdash \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi})}
+  {\forall \tau'\in \overline{\tau}\cup\{\tau\}\,.\,\kappa \wedge m \vdash \tau'\\ \kappa \le m \\ \fv(\overline{\tau})\cup\fv(\tau)\subseteq \overline{x}}
+  {m \vdash \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\kappa})}
   \end{mathpar}
@@
 \[
-\begin{array}{c} 
-\tmode \wedge \cmode = \umode \qquad \xi \wedge \umode = \umode
+\begin{array}{l} 
+m \le m \qquad \tmode \le \kappa
+\\[0.2em]
+\tmode \wedge \cmode = \umode \qquad \kappa \wedge \umode = \umode
 \qquad \cmode \wedge m = m 
 \qquad  m_1 \wedge m_2 = m_2 \wedge m_1
-\\[0.2em]
-\xi \le \xi \qquad \tmode \le \xi
+
 \end{array}
@@
 
-%% \dvh{I don't understand the variable grammar.  What is $T$?  What is $\eta$?  I think $\cmode$ and $\umode$ should be in tt font.}
-%% \liyi{T and $\eta$ can be moved to the appendix, they are useful only for struct types.}
-
-% \review{
-% - Furthermore, inspecting the code also suggests that the expression type (line 155) does not contain constructors for function calls (and I don't see a way to define functions either), conditionals, or strlen, and doesn't distinguish between the two forms of casting.
-All this contradicts figure 2, and should be clarified
-% }
-% \liyi{It is in the CheckedC.v file, 392. }
-% \mwh{How is this answer helping the reviewer since you've added
-%   nothing to the text? Maybe we should add something to an appendix
-%   that matches the formalism shown in the paper to definitions in the
-%   Coq file?}
-% \liyi{The detailed explanation is in the appendix.}
-
-This section describes \systemname's formal core model, \lang.
+This section describes \lang, the formal core model for both \systemname and \systemnamea.
 We present its syntax, semantics, type system,
-as well as \lang’s meta-theories, including the non-exposure, type soundness, and non-crashing theorems.
-
-%This section describes the formal core model of \systemname, named
-%\lang, making precise its syntax, semantics, type system, and compilation. It also
-%develops \lang's meta-theories, including the type soundness, non-exposure, non-crashing, and the compiler simulation theorems.
+as well as \lang’s meta-theories, including the non-exposure, type soundness, and non-crashing theorems,
+which are verified without the assumptions of the underlining sandbox structure.
+\lang utilizes dynamic checks that are inserted in the compilation time to ensure memory safety, i.e.,
+the \lang semantics (\Cref{sec:semantics}) describes the dynamic checks to perform in transitions,
+while its type system (\Cref{sec:typechecking}) indicates the places needing for inserting such checks.
 
@@
 value, or newly introduced function types (explained shortly below). A word-size value can be either an integer or pointer.
-Every pointer type ($\tptr{\omega}{\xi}$) includes a
+Every pointer type ($\tptr{\omega}{\kappa}$) includes a
 memory mode (explained shortly below), and a type ($\omega$) denoting the valid value type it is pointed to.
-A multi-word value type ($ \tallarrayb{\bvar}{\tau}$) that ranges over 
+A multi-word value type ($ \tarrayb{\bvar}{\tau}$) that ranges over 
 arrays is constructed by the type of elements in the
-array ($\tau$), a bound ($\bvar$) comprised of an upper and
-lower bound on the size of the array $(b_l,b_h)$, and a flag ($\kappa$).
+array ($\tau$), and a bound ($\bvar$) comprised of an upper and
+lower bound on the size of the array $(b_l,b_h)$.
 Bounds $b$ are
@@
 \begin{array}{rcl}
-$\code{_t_Array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
+$\code{t_array_ptr<}$\tau$\code{> : count(}$n$\code{)}$
 &\Leftrightarrow& \tarrayptr{0}{n}{\tau}{\tmode}
@@
 
-\myparagraph{Modes and Well-formed Types}
-%
-In \systemname, we divide \checkedc modes to context and memory modes.
-In terms of program contexts, modes refer to the partition of a program into checked (\cmode) and unchecked (\umode) regions,
-as the context modes in \Cref{fig:checkc-syn}.
-In terms of pointers, memory modes classify the partitioned memory ranges that particular pointers point to, either checked (\cmode), tainted (\tmode), or unchecked (\umode) memory ranges.
-More specifically, we partition each of the above ranges to a sub-range (subscripted with $\texttt{f}$) storing static function code data, and a sub-range (subscripted with $\texttt{n}$) storing normal heap data,
-so function pointers and heap pointers point to different sub-ranges conceptually.
-Different operations in different context regions are allowed to access different pointer modes, shown in \Cref{fig:checkc-mode}.
-A \texttt{malloc} operation in \cmode region produces a pointer accessing $\cmodenor$ or $\tmodenor$ memory range, while a \texttt{malloc} in \umode region produces a $\umode$ range pointer.
-Function calls ($\ecall{e}{\overline{e}}$) in \cmode region are allowed to access $\cmodefun$ or $\tmodefun$ range pointers,
-while calls in \umode region are permitted to access $\umode$ range pointers.
+\myparagraph{Modes and Type Well-formedness}
+%
+In \checkedc, modes can act as a context division to partition code into checked (\cmode) and unchecked (\umode) regions, and a pointer separation to divide pointers into checked (\cmode) and unchecked (\umode) pointers.
+In \systemname, we keep the context division, while upgrade mode's pointer separation aspect to a memory partition mechanism that separates heap into three disjoint areas: checked (\cmode), tainted (\tmode), or unchecked (\umode) heap regions.
+
+More specifically, we partition each of the above heap regions to a subrange (subscripted by $\texttt{f}$) storing static function code data, and a subrange (subscripted by $\texttt{n}$) storing normal heap data,
+so function pointers and heap pointers point to different subranges conceptually.
+Different operations in different context regions are allowed to access pointers with different modes, shown in \Cref{fig:checkc-mode}.
+For example, \texttt{malloc} operations in \cmode region produce pointers accessing $\cmodenor$ or $\tmodenor$ memory subranges, while a \texttt{malloc} in $\umode$ region produces a pointer accessing $\tmodenor$ and $\umode$ subranges.
+Function calls ($\ecall{e}{\overline{e}}$) in \cmode region are allowed to access pointers in $\cmodefun$ or $\tmodefun$ subranges,
+while calls in \umode region are permitted to access pointers in the $\umode$ heap region.
 
@@
 to prevent unsafe pointers, which might leak \cmode mode pointer address to \umode regions, from being constructed.
-Consider the pointer type ~\code{_t_Array_ptr<_Ptr<int>>}, which describes a tainted array of checked pointers.
+Consider the pointer type ~\code{t_array_ptr<ptr<int>>}, which describes a tainted array of checked pointers.
 This is not well-formed in~\lang because it potentially exposes the checked pointer addresses in a $\umode$ region (a violation of the non-exposure property in \Cref{sec:theorem}), when the tainted ($\tmode$) array is used.
-  Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{_Array_ptr<_t_Ptr<int>>} is a valid type.
-
-Function types are represented using a dependent function declarations,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
+Nevertheless, we can have a checked array whose elements are tainted pointers\mzs{,}\mzr{:}~\eg~\code{array_ptr<t_ptr<int>>} is a valid type.
+
+Function types represent dependent functions,~\ie $\tfun{\overline{x}}{\overline{\tau}}{\tau}$,
 where $\overline{x}$ represents a list of \tint{} type variables that bind type variables appearing in $\overline{\tau}$ and $\tau$.
@@
 \begin{array}{l}
-$\code{_t_ptr<(int)(_t_NT_Array_ptr<}$\tau_1$\code{> : count(}$n$\code{),}$\\
-\qquad\qquad$\code{_t_NT_Array_ptr<}$\tau_2$\code{>: count(}$n$\code{), int}$\;n$\code{)>}$
+$\code{t_ptr<(int)(t_array_ptr<}$\tau_1$\code{> : count(}$n$\code{),}$\\
+\qquad\qquad$\code{t_array_ptr<}$\tau_2$\code{>: count(}$n$\code{), int}$\;n$\code{)>}$
 \\[0.2em]
-\Leftrightarrow\; $\tptr{(\tfun{n}{\tint \times \tntarrayptr{0}{n}{\tau_2}{\tmode} \times \tntarrayptr{0}{n}{\tau_1}{\tmode}}{\tint})}{\tmode}$
+\Leftrightarrow\; $\tptr{(\tfun{n}{\tint \times \tarrayptr{0}{n}{\tau_2}{\tmode} \times \tarrayptr{0}{n}{\tau_1}{\tmode}}{\tint})}{\tmode}$
 \end{array}
@@
 }
-The function type also has well-formed requirements (\Cref{fig:wftypes}), which disallows nesting checked pointers inside tainted pointers.
-Furthermore, these requirements also ensure that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
+The function type also has well-formedness definition (\Cref{fig:wftypes}), which disallows nesting checked pointers inside tainted pointers.
+Furthermore, the well-formedness also ensures that all variables in $\overline{\tau}$ and $\tau$ are bounded by $\overline{x}$.
 
@@
  static casts ($\ecast{\tau}{e}$),
-memory allocations ($\emalloc{\xi}{\omega}$), 
+memory allocations ($\emalloc{\kappa}{\omega}$), 
 function calls ($\ecall{e}{\overline{e}}$).
-We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
-For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
-
-Compared with \checkedc, \lang's \texttt{malloc} operation includes a mode $\xi$ to model the allocations of pointers with different modes; $e$ in a function call $\ecall{e}{\overline{e}}$ represents a function pointer whose pointer mode indicates the memory range the function text locates, as well as the context-switch of context regions, explained shortly in \Cref{sec:typechecking}.
+We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\kappa}$), enabling the use of fixed addresses as pointers.
+For example, $\evalue{0}{\tptr{\omega}{\kappa}}$ (for any $\kappa$ and $\omega$) represents a $\enull$ pointer.
+
+Compared with \checkedc, \lang's \texttt{malloc} operation includes a mode $\kappa$ to model the allocations of pointers with different modes; $e$ in a function call $\ecall{e}{\overline{e}}$ represents a function pointer whose pointer mode indicates the memory range the function text locates, as well as the context-switch of context regions, explained shortly in \Cref{sec:semantics}.
 
@@
     \begin{array}{lll}
-r & ::= & e \mid \enull \mid \ebounds\\
-E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid \edyncast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}\\[0.2em]
-&&\mid\eassign{\evalue{n}{\tau}}{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\mid \eif{E}{e}{e}\\[0.2em]
-&&\mid \ecall{E}{\overline{e}} \mid \ecall{\evalue{n}{\tau}}{\overline{E}} \mid 
-\eunchecked{\overline{x}}{E}
-\mid \echecked{\overline{x}}{E}
-
+\mu & ::= & \evalue{n}{\tau} \mid \bot\\[0.1em]
+e & ::= & \ldots \mid \ret{x}{\mu}{e}\\[0.1em]
+r & ::= & e \mid \enull \mid \ebounds\\[0.1em]
+E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}
+\mid\eassign{\evalue{n}{\tau}}{E}\\[0.1em]
+&\mid& \eif{E}{e}{e}\mid \elet{x}{E}{e}\mid\ret{x}{\mu}{E}\\[0.1em]
+&\mid& \textcolor{red}{\ecall{E}{\overline{e}}} \mid \textcolor{red}{\ecall{(\evalue{n}{\tau})}{\overline{E}}}
+\mid \textcolor{red}{\toa{m}{\overline{x}}{e}}\\[0.1em]
+\overline{E} &::=& E \mid  \evalue{n}{\tau},\overline{E} \mid \overline{E}, e
 \end{array}
@@
 
-    \inferrule{ \umode=\mode(E) \\
-      e=E[e'] \\
-      \tau=\type(e')}
-    {(\varphi,\heap,e)\longrightarrow_{\umode} (\varphi,\heap,E[\evalue{0}{\tau}])}
+\textcolor{red}{
+    \inferrule[S-Recov]{ \umode=\mode(E) \\ \tau=\type(e)}
+    {(\varphi,\heap,E[e])\longrightarrow_{\umode} (\varphi,\heap,E[\evalue{0}{\tau}])}
+}
 
@@
 
-\subsection{Typing and Semantics}
-\label{sec:typechecking}
+The operational semantics for \lang is defined as a small-step
+transition relation with the judgment $ (\varphi,\heap,e)
+\longrightarrow_m (\varphi',\heap',r)$. Here, $\varphi$ is a
+\emph{stack} mapping from variables to values $\evalue{n}{\tau}$ and
+$\heap$ is a \emph{heap} that is partitioned into three parts ($\cmode$, $\tmode$, and $\umode$ heap regions), each of which
+maps addresses (integer literals) to values $\evalue{n}{\tau}$; for all we ensure
+$\fv(\tau)=\emptyset$.
+
+While heap bindings can change, stack bindings are immutable---once
+variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
+be updated. 
+We wrote $\heap(\kappa)$ to access the $\kappa$ heap region in $\heap$ and
+$\heap(\kappa,n)$ to retrieve the $n$-location heap value in the $\kappa$ heap.
+As mentioned, value $\evalue{0}{\tau}$
+represents a $\enull$ pointer when $\tau$ is a pointer type.
+Correspondingly, $\heap(m,0)$ should always be undefined.
+For $\kappa=\cmode \vee \kappa = \tmode$, there is a bound value $\eta_{\kappa}$,
+such that range $(0,\eta_{\kappa})$ in $\heap(\kappa)$ is reserved for function code storage, 
+where we define function $\Xi(\kappa,n)$ to access the function code appearing in the $n$-th location of $\kappa$ heap region.
+
+The relation steps to a \emph{result} $r$,
+which is either an expression or a $\enull$ or $\ebounds$ failure,
+representing a null-pointer dereference or out-of-bounds access,
+respectively. Such failures are a \emph{good} outcome; stuck states
+(non-value expressions that cannot transition to a result $r$)
+characterize undefined behavior.
+%
+The context mode $m$ indicates whether the
+stepped redex within $e$ was in a $\cmode$ or $\umode$ code region.
+
+The rules for the main operational semantics
+judgment---\emph{evaluation}---are given at the bottom of
+Fig.~\ref{fig:c-context}.
+The first rule takes an expression $e$, decomposes
+it into an \emph{evaluation context} $E$ and a sub-expression $e'$
+(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
+$e$), and then evaluates $e'$ according to the \emph{computation}
+  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
+whose rules are given in Fig.~\ref{fig:semantics}, discussed
+shortly. 
+Evaluation contexts $E$ define a
+standard left-to-right evaluation order.
+For example, context $\ecall{(\evalue{n}{\tau})}{\overline{E}}$ means that after the function pointer expression $e$ in $\ecall{e}{\overline{e'}}$ is evaluated to a function pointer $\evalue{n}{\tau}$, we now evaluate each argument expression, appearing in $\overline{e'}$, in sequence, such redex is marked as $\overline{E}$.
+
+Compared with the \checkedc formalism, other than the additional contexts in \Cref{fig:c-context},
+we add a new marked red rule \rulelab{S-Recov},
+referring to that every $\umode$ region transition step might crash, because it represents the unsafe code,
+and we have a mechanism to recover the crash as $\evalue{0}{\tau}$ \footnote{Function $\type$ represents the call of the type judgment in \Cref{sec:typechecking} to provide the type of the expression.}.
+The $\mode$ function
+determines the mode when evaluating $e'$ based on the context $E$:
+if the $\Box$ occurs within $\toa{m}{\overline{x}}{E'}$ inside $E$ and $E'$ does not contain another $\totext$ expression, 
+then the mode is $m$.
+Expression $\ret{x}{\mu}{e}$ does not appear in source program syntax in \Cref{fig:checkc-syn}, are introduced in evaluating programs,
+which is used for remembering $x$ was previously bound to
+$\varphi(x)$ in a $\elettext$ binding, explained shortly below.
+
+%The second rule in \Cref{fig:c-context} is newly introduced in \lang, to help the proof of the non-crashing theorem.
+%In any given time, executing a unchecked region might non-deterministically cause the program to crash.
+We assume that there is a recovery mechanism for recover every such crashing, only in unchecked regions, to an initial state $\evalue{0}{\tau}$;
+%thus, non-crashing shows that every unchecked region crashing does not affect the program evaluation in the checked regions.
 
@@
   \begin{mathpar}
-     \inferrule[T-CastPtr]
-               {\Gamma;\Theta \vdash_m e : \tau' \\
-                 \tau' \sqsubseteq_{\Theta} \tptr{\tau}{\xi}}
-               {\Gamma;\Theta \vdash_m \ecast{\tptr{\tau}{\xi}}{e} : \tptr{\tau}{\xi}}
+    \inferrule[S-DefArr]{\kappa \neq \tmode\\
+      n\in[\eta_{\kappa},|\heap(\kappa)|)\\\heap(\kappa,n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)}
+    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\kappa}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+        \inferrule[S-Def]{\heap(n)=\evalue{n_a}{\tau_a} \\ n\in[\eta_{\kappa},|\heap(\kappa)|)}
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\kappa}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefArrT]{n\in[\eta_{\kappa},|\heap(\kappa)|)\\\heap(\kappa,n)=\evalue{n_a}{\tau_a} \\ 0 \in [n_l,n_h)\\\heap;\emptyset \vdash \evalue{n_a}{\tau}}
+    {(\varphi,\heap,\estar{\evalue{n}{\tntarrayptr{n_l}{n_h}{\tau}{\kappa}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+        \inferrule[S-DefT]{\heap(n)=\evalue{n_a}{\tau_a} \\ n\in[\eta_{\kappa},|\heap(\kappa)|)\\\heap;\emptyset \vdash \evalue{n_a}{\tau}}
+    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\kappa}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
+
+    \inferrule[S-DefArrBound]{0 \not\in [n_l,n_h)}
+     { (\varphi,\heap,\estar{\evalue{n}{\tallarrayptr{n_l}{n_h}{\tau}{\kappa}}}) \longrightarrow (\varphi,\heap,\ebounds)}
+
+    \inferrule[S-DefNull]{n\not\in[\eta_{\kappa},|\heap(\kappa)|)}{(\varphi,\heap,\estar{\evalue{n}{\tptr{\omega}{\kappa}}}) \longrightarrow (\varphi,\heap,\enull)}
 
@@
               {(\varphi,\heap,\ecast{\tau}{\evalue{n}{\tau'}}) \longrightarrow (\varphi,\heap,\evalue{n}{\varphi(\tau)})}
+
+  \inferrule[S-Malloc]{\varphi(\omega)=\omega_a \\ alloc(\heap,\kappa,\omega_a)=(n,\heap')}
+   { (\varphi,\heap,\emalloc{\kappa}{\omega}) \longrightarrow (\varphi,\heap',\evalue{n}{\tptr{\omega_a}{\kappa}})}
+
+        \inferrule[S-Let]{}{(\varphi,\heap,\elet{x}{\evalue{n}{\tau}}{e}) \longrightarrow (\varphi[x\mapsto \evalue{n}{\tau}],\heap,\ret{x}{\varphi(x)}{e})}
+
+    \inferrule[S-Fun]{ \kappa\neq \umode\\n\in(0,\eta_{\kappa})\\\Xi(\kappa,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
+        {(\varphi,\heap,\ecall{(\evalue{n}{\tptr{\tau}{\kappa})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+
+    \inferrule[S-Ret]{}{(\varphi,\heap,\ret{x}{\mu}{\evalue{n}{\tau}}) \longrightarrow (\varphi[x\mapsto \mu],\heap,\evalue{n}{\tau})}
+
+    \inferrule[S-FunU]{ n\in(0,|\heap(\umode)|)\\ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e}
+        {(\varphi,\heap,\ecall{(\evalue{n}{\tptr{\tau}{\umode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
+   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
+
+    \inferrule[S-Switch]
+              {}
+              {(\varphi,\heap,\toa{m}{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
 \end{mathpar}
-
+}
+% {\footnotesize
+% \begin{center}
+% $
+% \begin{array}{l}
+% \tau[\overline{n} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall n_i\in\overline{n}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \Rightarrow \tau[n_i / x_i]\\[0.2em]
+% \mathtt{let}\;\overline{x}=\overline{e}\;\mathtt{in}...\triangleq \mathtt{let}\;x_0=e_0\;\mathtt{in}\;\mathtt{let}\;x_1=e_1\;\mathtt{in}...
+% \end{array}
+% $
+% \end{center}
+% }
+\caption{\lang Semantics: Computation (Selected Rules)}
+\label{fig:semantics}
+\end{figure*}
+\end{DIFnomarkup}
+
+\ignore{
+\begin{figure}[t]
+{\small
+{\captionsetup[lstlisting]{margin = 8 mm}
+  \begin{lstlisting}[xleftmargin=8 mm]
+nt_array_ptr<char> safe_strcat
+   (nt_array_ptr<char> dst : count(n),
+    nt_array_ptr<char> src : count(0), int n) {
+  int x = strlen(dst);
+  int y = strlen(src);
+  nt_array_ptr<char> c : count(n) =
+    dyn_bounds_cast
+           <nt_array_ptr<char>>(dst,count(n));
+    // sets c == dst with bound n (not x)
+  if (x+y < n) {
+    for (int i = 0; i < y; ++i)
+      *(c+x+i) = *(src+i);
+    *(c+x+y) = '\0';
+    return dst;
+  }
+  return null;
+}
+  \end{lstlisting}
+}
+}
+\caption{Implementation of safe \code{strcat}}
+\label{fig:strcat-ex}
+\end{figure}
+}
+
+\Cref{fig:semantics} shows selected rules for showing the key guarantees that \systemname and \systemnamea provide
+; we explain them with the help of the example in
+\Cref{sec:overview}.
+\ignore{
+  which defines a 
+  safe version of \code{strcat} (using actual Checked C syntax).  The
+  function takes a target 
+  pointer \code{dst} of capacity \code{n}, where the first null
+  character (determined by \code{strlen}) is at index \code{x} where
+  $0 \leq $\code{x}$ \leq n$. It concatenates the \code{src} buffer to
+  the end of \code{dst} as long as \code{dst} has sufficient space.
+}
+% Below, we introduce low-level transition semantics for some case operations.
+The design of the low-level individual operation semantics is carefully engineered to perform match our compiler's behavior, such as correctly characterizing the bound widening behaviors for NT-array pointers, even though it is written in terms of fat-pointer formalization.
+
+\myparagraph{Pointer accesses, Casts, and Mallocs.}
+%
+We describe rules (\rulelab{S-DefArr}, \rulelab{S-Def}, \rulelab{S-DefArrBound}, \rulelab{S-DefNull}, and \rulelab{S-Cast}) for dereference and cast operations to illustrate how the dynamic checks are performed in \lang.
+For every heap region $\heap(\kappa)$, every dereference and assignment operation (e.g., rule \rulelab{S-Def}) ensures that the pointer address is within the low and high bound range ($[\eta_{\kappa},|\heap(\kappa)|)$) of the heap region,
+where the low bound refers to the bound $\eta_{\kappa}$ splitting the function code and heap data subranges
+and the high bound is the cardinality of the heap ($|\heap(\kappa)|$).
+If the check is not satisfied, the semantics produces a $\enull$ state in \rulelab{S-DefNull}.
+When $\enull$ is returned by the
+computation relation, the evaluation relation halts the entire
+evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
+does likewise when $\ebounds$ is returned (see below).
+
+\rulelab{S-DefArr} dereferences an element in an array as long as $0$ (the point of
+dereference) is within the bounds designated by the pointer's annotation
+and strictly less than the upper bound ($0 \in [n_l,n_h)$),
+while the location address $n$ should be within the bounds of the specific heap region ($n\in[\eta_{\kappa},|\heap(\kappa)|)$).
+The former check is a $\ebounds$ check that verifies the element dereference is within the array bound,
+while the latter is a $\enull$ check that ensures that the pointer exists in the correct heap region.
+The dereference semantics and type rules are general enough to be used in both $\cmode$ and $\umode$ code regions.
+As we mentioned in \Cref{sec:background}, $\umode$ code regions contain mostly C code that does not have array annotations.
+In this case, a $\umode$ mode array pointer is cast and viewed as a normal C pointer, where we apply rule \rulelab{S-Def} to deal with its dereference.
+
+Rules \rulelab{S-DefArrT} and \rulelab{S-DefT} define the dereference operations for $\tmode$ mode pointers.
+Here, we perform an additional type verification $\heap;\emptyset \vdash \evalue{n_a}{\tau}$ making sure that the memory location context $n_a$ has the declared type $\tau$ appearing in the operation $\estar{\evalue{n}{\tptr{\tau}{\kappa}}}$.
+Such check not only verifies if $n_a$ has type $\tau$, but also if $n_a$ is a nested pointer,
+the check further verifies if all addresses within $n_a$'s legal scopes are well-typed.
+In \Cref{lst:final}, function \code{read_msg} casts a $\umode$ mode pointer and assigns it to a tainted array pointer and returns to the checked function \code{process_req2}; such behavior is legal in a $\umode$ code region.
+\code{process_req2} catches the bug in line 10 through the type verification above, because pointer $x$ cannot have type $\tptr{\tarrayptr{0}{5}{\tint}{\tmode}}{\tmode}$ since it is illegal to cast type $\tptr{\tptr{\umode}{\tint}}{\tmode}$ to it.
+
+Static casts of a literal $n\!:\!\tau'$ to a type $\tau$ are handled
+by \rulelab{S-Cast}. In a type-correct program, such casts are
+confirmed safe by the type system. To evaluate a cast, the rule
+updates the type annotation on $n$. Before doing so, it must
+``evaluate'' any variables that occur in $\tau$ according to their
+bindings in $\varphi$. As an instance, if $\tau$ was
+$\tarrayptr{0}{x+2}{\tint}{\cmode}$, then $\varphi(\tau)$ would
+produce $\tarrayptr{0}{4}{\tint}{\cmode}$ if $\varphi(x) = 2$.
+Rule \rulelab{S-Malloc} creates a new memory pointer $\tptr{\omega_a}{\kappa}$
+in heap region $\heap(\kappa)$ through function $alloc$, which returns a neq heap $\heap'$ that creates a new memory space in $\heap(\kappa)$ and increments $\heap(\kappa)$'s cardinality, as $|\heap'(\kappa)|=|\heap(\kappa)|+$\code{sizeof}$(\omega_a)$;
+$\varphi(\omega)$ evaluates potential bound variables in $\omega$ and returns a new type $\omega_a$ with no variables.
+
+\myparagraph{Binding, Function Calls, and Context Switching}
+We handle variable scopes using the special $\erettext$
+operation. \textsc{S-Let} evaluates to a configuration whose stack
+is $\varphi$ extended with a binding for $x$, and whose expression is
+$\ret{x}{\varphi(x)}{e})$ which remembers $x$ was previously bound to
+$\varphi(x)$; if it had no previous binding, $\varphi(x) =
+\bot$. Evaluation proceeds on $e$ until it becomes a literal
+$n\!:\!\tau$, in which case \textsc{S-Ret} restores the saved
+binding (or $\bot$) in the new stack, and evaluates to
+$n\!:\!\tau$.
+
+Function calls are handled by rules \rulelab{S-FunU} and \rulelab{S-Fun},
+for $\umode$ and non-$\umode$ mode functions \footnote{A $\kappa$ mode function means the function pointer is labeled with mode $\kappa$ and its code is stored in $\kappa$ heap region.}, respectively.
+When calling a function pointer $\evalue{n}{\tptr{\tau}{\kappa}}$,
+we extra the $n$-th location code data from $\Xi(\kappa)$, referring to the function code fields in the $\kappa$ heap region and makes a memory address to $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;e$, where
+$\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
+is the parameter list of variables and their types, and $e$ is the
+function body.
+We place a bound check in rule \rulelab{S-Fun} to 
+ensure that $\cmode$ and $\tmode$ mode function calls only access addresses in the range $(0,\eta_{\kappa})$,
+the code data field ranges for $\cmode$ and $\tmode$, respectively;
+while $\umode$ function calls (\rulelab{S-FunU}) are allowed to access all $\umode$ heap region.
+Similar to \checkedc, \lang function calls are dependent, i.e.,
+array bounds in types may refer to in-scope variables; e.g., 
+\code{process_req1}'s bound \code{count(m_1)} refers to parameter \code{m_1} in \Cref{lst:final}
+The call is expanded into a \texttt{let} which binds
+parameter variables $\overline{x}$ to the actual arguments
+$\overline{n}$, but annotated with the parameter types
+$\overline{\tau}$ (this will be safe for type-correct programs). The
+function body $e$ is wrapped in a static cast
+$(\tau[\overline{n} / \overline{x}])$ which is the function's return
+type but with any parameter variables $\overline{x}$ appearing in that
+type substituted with the call's actual arguments $\overline{n}$. 
+\ignore{
+To
+see why this is needed, suppose that \code{safe_strcat} in
+Fig.~\ref{fig:strcat-ex} is defined to return a
+\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
+perform a \code{safe_strcat} function call as
+\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
+function returns a value with type \code{nt_array_ptr<int>:count(10)}
+because we substitute bound variable \code{n} in the 
+defined return type with \code{10} from the function call's
+argument list.
+}
+Note that the \textsc{S-Fun} rule replaces the
+  annotations $\overline{\tau_a}$ with
+  $\overline{\tau}$ (after instantiation) from the function's
+  signature. Using $\overline{\tau_a}$ when executing the body of
+the function has no impact on the soundness of \lang, but will violate
+Theorem~\ref{simulation-thm}, which we introduce in Sec.~\ref{sec:compilation}.
+Rule \rulelab{S-Swtich} defines the context mode switch block's semantics, which gets rid of the $\texttt{to}$ block once its context is a value.
+
+\subsection{Typing}
+\label{sec:typechecking}
+
+\begin{DIFnomarkup}
+\begin{figure*}[t]
+{\small
   \begin{mathpar}
+    \inferrule[T-DefArr]
+              {\kappa\le m \\ \Gamma;\Theta \vdash_{m} e : \tptr{\tarrayb{\bvar}{\tau}}{\kappa}}
+              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
+
     \inferrule[T-Def]
-              {\xi \leq m \\\Gamma;\Theta \vdash_m e : \tptr{\tau}{\xi}}
-              {\Gamma;\Theta \vdash_m \estar{e} : \tau}
-\quad
-        \inferrule[S-DefC]{\heap(\cmode,n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{c}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-        \inferrule[S-DefT]{\heap(\umode,n)=\evalue{n_a}{\tau_a}
-         \\  \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau} }
-    {(\varphi,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\heap,\evalue{n_a}{\tau})}
-\quad
-    \inferrule[S-DefNull]{}{(\varphi,\heap,\estar{\evalue{0}{\tptr{\omega}{\cmode}}}) \longrightarrow (\varphi,\heap,\enull)}
-  \end{mathpar}
-
-  \begin{mathpar}
-    \inferrule[T-Checked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                     \\\\\fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
-              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
-
-    \inferrule[T-Unchecked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                \\\\ \fv(e)\in\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
-
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
-  \end{mathpar}
-
-  \begin{mathpar}
+              {\kappa \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\kappa}}
+              {\Gamma \vdash_m \estar{e} : \tau}
+
+     \inferrule[T-Cast]
+               {\Gamma \vdash_{\cmode} e : \tau' \\
+                 \tau' \sqsubseteq \tptr{\tau}{\kappa}}
+               {\Gamma \vdash_{\cmode} \ecast{\tptr{\tau}{\kappa}}{e} : \tptr{\tau}{\kappa}}
+
+    \inferrule[T-CastU]
+              {\kappa \le \umode \\
+                \Gamma;\Theta \vdash_{\umode} e : \tau'}
+              {\Gamma;\Theta \vdash_{\umode} \ecast{\tptr{\tau}{\kappa}}{e} : \tptr{\tau}{\kappa}}
+
+    \inferrule[T-Malloc]
+              {\kappa\le m}
+              {\Gamma \vdash_m \emalloc{\kappa}{\omega} : \tptr{\omega}{\kappa}}
+
+   \inferrule[T-Let]
+    { x\not\in \fv(\tau') \\
+        \Gamma \vdash_m e_1 : \tau \\\\
+          \Gamma[x\mapsto \tau] \vdash_m e_2 : \tau'
+             }
+    {\Gamma \vdash_m \elet{x}{e_1}{e_2} : \tau'}
+
+   \inferrule[T-LetInt]
+    {\Gamma \vdash_m e_1 : \tint \\\\
+           \Gamma[x\mapsto \tint] \vdash_m e_2 : \tau'\\\\
+        x\in \fv(\tau') \Rightarrow e_1 \in \text{Bound} 
+             }
+    {\Gamma \vdash_m \elet{x}{e_1}{e_2} : \tau'[e_1 / x]}
+
+\inferrule[T-Ret]
+    {\Gamma(x)\neq \bot \\
+          \Gamma;\Theta \vdash_m e : \tau}
+    {\Gamma;\Theta \vdash_m \eret{x}{\mu}{e} : \tau}
+
+    \inferrule[T-Switch]
+              {\forall x\in\{\overline{x}\}.\,\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
+                     \\\\\fv(e)\subseteq\overline{x}\\\Gamma \vdash_{m'} e : \tau}
+              {\Gamma \vdash_m \toa{m'}{\overline{x}}{e} : \tau}
+
 \inferrule[T-Fun]
-    {\Gamma;\Theta \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \\
-        \Gamma; \Theta \vdash_m \overline{e} : \overline{\tau'} \\
+    {\Gamma \vdash_m e : \tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\kappa} \\
+        \Gamma \vdash_m \overline{e} : \overline{\tau'} \\
          \overline{e'}=\{e'|(e',\tint)\in (\overline{e} : \overline{\tau'})\}\\\\
          \forall e'\;.\;e' \in \overline{e'} \Rightarrow e'\in \text{Bound}\\
-             \overline{\tau'} \sqsubseteq_{\Theta}
+             \overline{\tau'} \sqsubseteq
                \overline{\tau}[\overline{e'} / \overline{x}]}
-    {\Gamma; \Theta \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
-  \end{mathpar}
-  \begin{mathpar}
-    \inferrule[S-FunC]{ \Xi(\cmode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\cmode,e)}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\cmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-\quad
-    \inferrule[S-FunT]{ \Xi(\umode,n) = \tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\tmode,e)
-                  \\ \emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n}{\tptr{\tau}{\tmode}}}
-        {(\varphi,\heap,\ecall{\evalue{n}{(\tptr{\tau}{\tmode})}}{{\evalue{\overline{n_a}}{\overline{\tau_a}}}}) \longrightarrow
-   (\varphi,\heap, \mathtt{let}\;\overline{x}={\evalue{\overline{n}}{(\overline{\tau}[\overline{n} / \overline{x}])}}\;\mathtt{in}\;\ecast{\tau[\overline{n} / \overline{x}]}{e})}
-
+    {\Gamma \vdash_m e(\overline{e}) : \tau[\overline{e'} / \overline{x}]}
   \end{mathpar}
@@
 {\footnotesize
+\begin{center}
 $
@@
 \qquad
-\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
+\cmode(\tptr{\omega}{\kappa})=\texttt{false}\;\;{[\emph{owise}]}
 $
-}
-\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.
-First line is for cast operations, second line is for pointer dereferences, third line is for checked/unchecked blocks, and the rest is for function calls.}
+\end{center}
+}
+\caption{Selected type rules}
 \label{fig:type-system-1}
@@
 that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
-We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
+We define a three point lattice $\kappa_1 \le \kappa_2$ \footnote{In typing rule, the lattice is usually used as $\kappa \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \kappa$ and $m \le m$.
 This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can only perform such operations in $\cmode$ and $\umode$ regions, respectively.
@@
 For example, in a nested pointer $\tptr{(...
-\tptr{\tau}{\xi_2} ...)}{\xi_1}$, we require $\xi_2\le \xi_1$ to maintain non-exposure.
+\tptr{\tau}{\kappa_2} ...)}{\kappa_1}$, we require $\kappa_2\le \kappa_1$ to maintain non-exposure.
 
 \lang also provides static casting operations. As described in rule \textsc{T-CastPtr} in \Cref{fig:type-system-1},
-an pointer typed expression of type $\tptr{\tau_1}{\xi_1}$ can be casted to another pointer type \tptr{\tau_2}{\xi_2},
-if \tptr{\tau_1}{\xi_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\xi_2}, i.e., $\tptr{\tau_1}{\xi_1} \sqsubseteq \tptr{\tau_2}{\xi_2}$.
-In \lang, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\xi_1$ and $\xi_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
+an pointer typed expression of type $\tptr{\tau_1}{\kappa_1}$ can be casted to another pointer type \tptr{\tau_2}{\kappa_2},
+if \tptr{\tau_1}{\kappa_1} subtypes ($\sqsubseteq_{\Theta}$) to \tptr{\tau_2}{\kappa_2}, i.e., $\tptr{\tau_1}{\kappa_1} \sqsubseteq \tptr{\tau_2}{\kappa_2}$.
+In \lang, except that we can cast a $\tmode$ mode pointer to a $\umode$ mode one, all subtyping relations are between two types with the same mode, meaning that $\kappa_1$ and $\kappa_2$ above are mostly the same and the above mode lattice ($\le$) has no business with subtyping.
 
@@
 reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
-The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
+The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\kappa \le m$).
 With the dynamic checks inserted by the compiler, rule \textsc{S-DefNull} ensure that if a $\enull$ pointer is used,
@@
 Given an expression $e$ of function pointer type
-($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$) and arguments
+($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\kappa}$) and arguments
 $\overline{e}$ of types $\overline{\tau'}$,
@@
 which maps function pointers to
-function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\xi,e)$, where
+function data $\tau\;(\evalue{\overline{x}}{\overline{\tau}})\;(\kappa,e)$, where
 $\tau$ is the return type, $(\evalue{\overline{x}}{\overline{\tau}})$
 is the parameter list of variables and their types, 
-$\xi$ determines the mode of the function, and $e$ is the
+$\kappa$ determines the mode of the function, and $e$ is the
 function body. 
@@
        {}
-       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\xi}}
+       {\Theta;\heap;\sigma \vdash_m 0 : \tptr{\omega}{\kappa}}
 
    \inferrule
-       {(m = \cmode \Rightarrow \xi \neq \cmode) \\\\ (m=\umode \Rightarrow \xi = \umode)}
+       {(m = \cmode \Rightarrow \kappa \neq \cmode) \\\\ (m=\umode \Rightarrow \kappa = \umode)}
        {\Theta;\heap;\sigma \vdash_{\cmode} n : \tptr{\omega}{\tmode}}
@@
    \inferrule
-       {(\evalue{n}{\tptr{\omega}{\xi}})\in \sigma}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+       {(\evalue{n}{\tptr{\omega}{\kappa}})\in \sigma}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\kappa}}
 
    \inferrule
-       {\tptr{\omega'}{\xi'} \sqsubseteq_{\Theta} \tptr{\omega}{\xi} 
-            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\xi'}}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+       {\tptr{\omega'}{\kappa'} \sqsubseteq_{\Theta} \tptr{\omega}{\kappa} 
+            \\ \Theta;\heap;\sigma \vdash_m n : \tptr{\omega'}{\kappa'}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\kappa}}
 
    \inferrule
-       { \xi \le m 
-     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\xi,e)
+       { \kappa \le m 
+     \\\Xi(m,n)=\tau\;(\evalue{\overline{x'}}{\overline{\tau}})\;(\kappa,e)
        \\  \overline{x} = \{x|(x:\tint) \in (\overline{x'}:\overline{\tau}) \}}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\xi}}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{(\tfun{\overline{x}}{\overline{\tau}}{\tau})}{\kappa}}
 
    \inferrule
-       {\neg\funptr(\omega)\\ \xi \le m\\
+       {\neg\funptr(\omega)\\ \kappa \le m\\
         \forall i \in [0,\size(\omega)) \;.\;
-            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\xi})) \}\vdash_m \heap(m,n+i)}
-       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\xi}}
+            \Theta;\heap;(\sigma \cup \{(n:\tptr{\omega}{\kappa})) \}\vdash_m \heap(m,n+i)}
+       {\Theta;\heap;\sigma \vdash_m n : \tptr{\omega}{\kappa}}
  \end{mathpar}
@@
 pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
-Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
+Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\kappa} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\kappa}$), 
 if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
@@
 Rule \textsc{T-Fun} in \Cref{fig:type-system-1} is the dependent function call rule. 
-Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi}$)
+Given a function pointer type ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\kappa}$)
 from a type-check for $e$ and the types $\overline{\tau'}$ from the argument type checks for $\overline{e}$,

### 2023-04-01 csf2024/overview.tex
update semantics

\section{Overview}
\label{sec:overview}

\begin{figure*}[t]
  \begin{tabular}{c c}
    \begin{minipage}[b]{.48\textwidth}
\inputminted[mathescape, escapeinside=||, fontsize=\small{}]{c}{examples/conv1.c}
    \end{minipage} &
    \begin{minipage}[b]{.48\textwidth}
\inputminted[mathescape, highlightlines={1-19}, highlightcolor=taintcolor, escapeinside=||, fontsize=\small{}]{c}{examples/conv2.c}
    \end{minipage} %\\
   %(a) Original C code & (b) After initial conversion.\\
  \end{tabular}
\caption{Final annotated program of~\lst{lst:comb} with~\textcolor{taintcolor}{tainted} and~\textcolor{checkcolor}{checked} types. The~\colorbox{taintcolor}{highlighted} functions will be executed in a sandbox.}
\label{lst:final}
\end{figure*}

%\input{figures/overview}

\Cref{fig:overview} shows the interaction between various components of~\systemname{}.
Given an appropriately annotated source program and a sandbox configuration, \systemname{} creates an executable such that all tainted functions and data reside in a sandbox and the~\cregion (non-sandboxed) is not affected by the code in the sandbox.
Here, we brief \systemname{} from a developer perspective and the details of individual components in the later sections.

%\Cref{fig:overview} provides the compiler structure of \systemname. The \systemname type checker compiles a \systemname program by partitioning the different program regions, i.e., $c$ and $uc$ regions, as well as different pointers, i.e., checked, unchecked and tainted pointers. Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler within the checked region.

%\Cref{sec:formal} provides the formalism of the different compiler components.
%Converting a C/checked-c program to \systemname is a two-step process that involves identifying unsafe/unchecked regions described in \Cref{subsec:identifyregionstosbx} followed by making the \systemname changes described in \Cref{subsec:moveregionstosbx}.  
\subsection{Running Example}
\label{subsec:identifyregionstosbx}

To explain changes required by the \systemname compiler, \Cref{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
The server runs in a loop and calls~\code{handle_request}, which handles a network request.
The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} is called to handle the request.

\myparagraph{The Vulnerability}
There is an arbitrary memory write vulnerability (indicated by~\realbug) in~\code{process_req1} because of using~\code{i} as an index into the array~\code{msg} without any sanity check.
The variable~\code{i} can take any integer value as it is parsed from~\code{msg}, whose content is read from socket in~\code{read_msg} indicated by~\rootcause.

\myparagraph{Goal} 
The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the other part of the program.
Ideally, the developer can convert the entire code to \checkedc so that we achieve full spatial memory safety.
However, there might be considerable conversion efforts in \checkedc.
For instance,~\code{void*} pointers are not directly supported by \checkedc.
Consequently, the developer needs to convert functions using~\code{void*} pointers into generic versions -- this could be tedious.
To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
However, as shown in~\sect{subsec:nosafetyagsintuncheckedcode}, $\umode$ region code can also affect the safety of checked pointers.
%\aravind{Fix the running example code with the annotations according to the implementation.}

\subsection{\systemname Annotations}
\label{subsec:moveregionstosbx}
%\myparagraph{Using tainted types}
To overcome the above difficulty, The developer marks risky functions with unchecked pointers,~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded by using with \systemname{},
 and results in the partially annotated program with checked and tainted types~\ie~\Cref{lst:initialconv,lst:humantaint} atop of~\lst{lst:comb}.
The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (\sect{sec:typechecking}).
The developer uses our type checker to identify the additional required annotations (\usermods) and adds them as shown in~\lst{lst:humanadjust}.
The resulting well-typed program as shown in~\lst{lst:final} is passed to our source level program partitioner along with certain configuration parameters of the target sandbox.

\subsection{Partitioning}
Our partitioner splits the provided program into two sets of source files with the necessary source changes required to communicate with sandboxed code.
These sets of source files are compiled with the corresponding compilers to get the corresponding object files.
The \cregion object file has the necessary runtime checks enforcing~\systemname{} guarantees.
The~\ucregion object file is produced according to the corresponding sandbox mechanism.
Finally, these two object files are linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions are executed in a sandbox (\ucregion) and the rest of the functions as regular code (\cregion).

%newusecase.tex

### 2023-04-01 sandp23/formal.tex
update sandp.
update semantics

@@
 \begin{figure*}[t]
-{\small
+{\footnotesize
 $\textcolor{blue}{\text{Dereference Operations:}}\\$
@@
 
-        \inferrule[S-DefT]{\heap(\tmode,n)=\evalue{n_a}{\tau_a} \\ n \not\in \Theta \\ n < |\heap(\tmode)|}
+        \inferrule[S-Def]{\xi \neq \tmode \\ \heap(n)=\evalue{n_a}{\tptr{\tau_a}{\xi}} }
+    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
+
+        \inferrule[S-DefT]{\heap(n)=\evalue{n_a}{\tptr{\tau_a}{\tmode}} \\ n \not\in \Theta}
     {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
 
-        \inferrule[S-Def]{\xi \neq \tmode \\ \heap(\xi,n)=\evalue{n_a}{\tau_a} }
-    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
-
     \inferrule[S-DefNull]{}{(\varphi,\Theta,\heap,\estar{\evalue{0}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\enull)}
-
-    \inferrule[S-DefBound]{n \in \Theta \vee n \ge |\heap(\tmode)|}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
+\qquad
+    \inferrule[S-DefBoundA]{n \in \Theta}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
+\qquad
+    \inferrule[S-DefBoundB]{\heap(n)=\evalue{n_a}{\tptr{\tau_a}{\xi}}\\\xi\neq \xi'}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi'}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
   \end{mathpar}
@@
 \emph{stack} mapping from variables to values $\evalue{n}{\tau}$,
-$\Theta$ is a set of $\tmode$ mode memory addresses that are freed at the point of executing $e$,
-and $\heap$ is a \emph{heap} that is partitioned into three parts ($\cmode$, $\tmode$, and $\umode$ heap regions), each of which
-maps addresses (integer literals) to values $\evalue{n}{\tau}$.
-We wrote $\heap(m)$ to retrieve the $m$ heap and $\heap(m,n)$ to retrieve the $n$-location heap value in the $m$ heap.
+$\Theta$ is a set of $\tmode$ mode memory addresses that are freed at the point of executing $e$
+and $\heap$ is a \emph{heap} mapping addresses (integer literals) to
+values $\evalue{n}{\ptr{\tau}{\xi}}$; for both we ensure
+$\fv(\tau)=\emptyset$.
+In a heap value $\evalue{n}{\ptr{\tau}{\xi}}$,
+$\tau$ represents the value's content type, while $\xi$ indicates the mode of the heap location;
+we classify a heap location into $\cmode$, $\tmode$, and $\umode$ modes, corresponding to the three kinds of pointer modes.
 The complete set of typing rules are provided in~\Cref{rem-type,sec:rem-semantics}.
@@
 %We can model mutable stack variables as pointers into the mutable heap.
-As mentioned, value $\evalue{0}{\tau}$
+As mentioned, value $\evalue{0}{\tptr{\tau}{\xi}}$
 represents a $\enull$ pointer when $\tau$ is a pointer type.
-Correspondingly, $\heap(m,0)$ should always be undefined.
+Correspondingly, $\heap(0)$ should always be undefined.
 % 
@@
 \myparagraph{Pointer Dereference}
-The type and semantic rules for pointer dereference (\rulelab{T-Def}, \rulelab{S-DefT},
-\rulelab{S-Def}, \rulelab{S-DefBound}, \rulelab{S-DefNull} in \Cref{fig:type-system-1})
+The type and semantic rules for pointer dereference (the first block in \Cref{fig:type-system-1})
 reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
@@
 Rules \rulelab{S-DefT} and \rulelab{S-Def} are for $\tmode$ mode and other mode pointer dereferences, respectively.
-Operation $\heap(\xi,n)$ not only accesses the $n$-th address location,
+Operation $\heap(n)$ not only accesses the $n$-th address location,
 but also performs a $\enull$ check because $0$ is undefined in all heaps.
-In addition to the no $\enull$ check in rule \rulelab{S-Def},
-we perform checks in \rulelab{S-DefT} to ensure the existence of $\tmode$ mode pointers.
-any dynamic heap access of a tainted ($\tmode$) pointer requires a \textit{verification} ($\emptyset;\heap ; \emptyset \vdash_{\umode}\evalue{n_a}{\tau}$), which refers to that the pointer value $n_a$ is well-defined in $\heap(m,n_a)$ and has right type $\tau$.
-Predicate $n \not\in \Theta$ means that location $n$ is not freed in the $\tmode$ heap region,
-while $n < |\heap(\tmode)|$ makes sure that $n$ does not go out of the high bound of the $\tmode$ heap region,
-where $|\heap(\tmode)|$ refers to the cardinality of the $\tmode$ heap region.
-If dynamic checks are violated, rules \rulelab{S-DefNull} and \rulelab{S-DefBound} are performed, meaning that the evaluation is transitioned into a runtime error that is correctly captured by the \systemname compiler.
+The pointer type ($\evalue{n}{\tptr{\tau}{\tmode}}$) in a dereference operation must have the same mode
+as the one labeled in the type in the heap value $(\evalue{n_a}{\tptr{\tau_a}{\xi}}$).
+The heap value type $\tptr{\tau_a}{\xi}$ is not an additional structure existed in the heap structure,
+but an annotation lives in the compiler and is compiled as a dynamic check to ensure that pointers only access heap locations with right modes. We explain the dynamic check generation shortly below.
+In addition to the above checks in rule \rulelab{S-Def},
+we perform checks in \rulelab{S-DefT} $n \not\in \Theta$ to ensure that location $n$ is not freed in the $\tmode$ heap region.
+If the above dynamic checks are violated, rules \rulelab{S-DefNull}, \rulelab{S-DefBoundA}, and \rulelab{S-DefBoundB} are performed, meaning that the evaluation is transitioned into a runtime error that is correctly captured by the \systemname compiler.
 
@@
 otherwise, the function picks the $|\heap(\xi)|$ as the new pointer and increments $\heap(\xi)$'s cardinality.
+The new location is labeled with a type annotation $\tptr{\tau}{\xi}$, indicating that the location is $\xi$ mode.
+Every time when a deference and assignment is applied on a $\tmode$ pointer, the \systemname compiler applies a dynamic check to verify if the location is labeled with $\tmode$.
+We can think of the dynamic check as $n \in \heap(\tmode)$, checking if a location $n$ is in a set $\heap(\tmode)$ containing all the $\tmode$ addresses.
+When a $\texttt{malloc}$ function is called to create a new $\tmode$ pointer, we expand set $\heap(\tmode)$ to include the new address.
+See \Cref{sec:compilation}.
 Rules \rulelab{S-FreeT} and \rulelab{S-Free} define the \texttt{free} operation for $\tmode$ mode and other mode pointers, respectively.
@@
 
-\item $e=\eassign{\evalue{n}{\tptr{\tau}{\tmode}}}{\evalue{n'}{\tptr{\tau}{m'}}}$,
-and $\estar{\evalue{n'}{\tptr{\tau}{m'}}}$ can make a move.
-
-\item $e=\eassign{\evalue{n}{\tptr{\tau}{m}}}{\evalue{n'}{\tptr{\tau}{\tmode}}}$,
-and $\estar{\evalue{n}{\tptr{\tau}{m}}}$ can make a move.
+\item $e=\eassign{\evalue{n}{\tptr{\tau}{\tmode}}}{\evalue{n'}{\tau'}}$.
 

### 2023-04-02 csf2024/introduction.tex
update intro.

--- 2023-01-20 sandp23/introduction.tex
@@
 \section{Introduction}\label{sec:intros}
+
+\begin{itemize}
+
+\item Vulnerabilities due to memory corruption are still issues, with examples.
+
+\item We want program partitioning with privilege separation.
+With a picture, program is split into trusted part and unsafe or high-risk code is separated into an untrusted and least privileged partition.
+
+\item most program partitioning is "data-centric".
+Nowadays, there is a new group doing "code-centric" program partitioning, which we called it as type-directed program partitioning, such as RLBox and others.
+It usually contains a type of variables needed to communicate between safe and unsafe regions.
+
+\item however, type-directed program partitioning has no formalism.
+We do not know what they can actually ensure for memory safety.
+1) they obviously cannot ensure spatial safety.
+2) clean separation might be too strong because many these implementations have ways to allow users not to follow validity checks for communication data.
+3) trusted part might contain bugs creates another problem.
+
+\item We present \systemname, as a formalism for type-directed program partitioning to answer these concerns.
+The system is based on \checkedc because of the third concern above.
+In privilege separation, it is obvious that trusted code is viewed as the part that does not have problem, while programs are in untrusted code, which is similar to the concept of checked and unchecked code regions in \checkedc and make it a perfect model to formalize type-directed program partitioning.
+
+\item introduce \checkedc and two code regions.
+
+\item state the design principle of \systemname by adding tainted type to \checkedc and clearly separate the two parts, so that \checkedc's different code regions are turned from a waiting for conversion separation to a type-direction program partitioning.
+
+\item state three contributions of the paper: 1) the first formalism for type-directed program partitioning with the theorems, clean separation and non-crashing, stating clearly what these mechanism can actually provide.
+2) Such formalism is independent with respect to the underlining sandbox mechanism, we find out the independent property and implement a separate-heap based sandbox mechanism to achieve a great overhead compared to the original wasm sandbox based implementation [cite] that faithfully represents the RLBox implementation.
+3) we formalize the compiler and show that the simulation relation between the compiler and our formalism.
+
+\item Summary Visualization.
+
+\end{itemize}
 
@@
 despite many efforts that tried to prevent them~\cite{song2019sanitizing}.
+\liyi{add some example sentences here. like }
+Software vulnerabilities remain a critical issue of computer security,
+despite decades of research on identifying, preventing, and fixing
+them. Although vulnerabilities often exist in non-privileged code,
+exploiting such vulnerabilities can allow adversaries to gain access
+to privileged code that co-exists with non-privileged code in the
+same program. For example, such vulnerabilities were discovered
+recently in both Linux and Windows applications .
+
+\ignore{
 Several industrial and research efforts, including CCured~\cite{Necula2005},
@@
 regions.
+}
 %
@@
 
-We evaluated~\systemname by partioning seven large real-world programs to demonstrate its effectiveness.
+We evaluated~\systemname~\footnote{Our implementation is available open source at~\oururl{}.} by partioning seven large real-world programs to demonstrate its effectiveness.
 Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.

### 2023-04-05 sandp23/introduction.tex
Adding
Addiong all

@@
 Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
-have investigated ways to better compile C programs with automatic spatial safety enforcement.
-These approaches all impose performance overheads deemed too high for deployment use. 
+have investigated ways to better compile C programs with automatic safety enforcement.
+These approaches impose performance overheads deemed too high for deployment use, especially for resource-constrained embedded systems.
+%
+Previous works~\cite{meng2021bran, du2019leopard} show that vulnerabilities are not uniformly distributed in real programs.
+Some components/functions are likelier to have a certain class of vulnerabilities.
+\eg{} spatial safety issues in string processing functions.
+have.
+
+Program partitioning is a well-known technique that enables separating potentially vulnerable functions (untrusted region) from the rest of the program (trusted region).
+%
+Several automated program partitioning techniques~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} exist, but they have a considerable performance overhead.
+
+One of the main reasons for this is combining partitioning with~\acf{SFI}.
+Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region.\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
+However, this extreme partitioning is not always necessary and feasible in certain resource constrained environments.
+For instance, we want to isolate a string processing function only from spatial violations,~\ie any spatial violations (\eg buffer overread) in that function should not affect the program.
+
+But existing partitioning techniques are specialized for a specific~\ac{SFI}~\cite{wahbe1993efficient} technique.
+For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling.
+Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
+In summary, existing techniques combine policy (\ie{}~\emph{what} to partition) with mechanism (\ie{}~\emph{how} to partition).
+We argue that existing program partitioning mechanisms are restrictive and inflexible for different use cases (\eg isolation from spatial safety vulnerabilities).
+We need a technique that enables developers to flexibility partition programs and enforce isolation -- balancing safety guarantees and performance overhead.
+
 % 

### 2023-04-06 sandp23/introduction.tex
Adding all

@@
 One of the main reasons for this is combining partitioning with~\acf{SFI}.
-Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region.\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
+Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region,~\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
 However, this extreme partitioning is not always necessary and feasible in certain resource constrained environments.
@@
 We need a technique that enables developers to flexibility partition programs and enforce isolation -- balancing safety guarantees and performance overhead.
+
+In this paper, we present~\systemname{}, a Type Directed Flexible Program Partitioning system.
+The main design principle of the system is to separate partitioning (\ie policy) from enforcing the separation between partitions (\ie mechanism).
+Once a program is partitioned, users can choose enforcement mechanisms based on the desired guarantees and tolerable performance overhead.
+For instance, for the same program, the user may use complete~\ac{SFI} for desktop systems v/s isolation against only spatial safety issues on embedded systems.
+
+We propose a type directed partitioning through untrusted types.
+Specifically, we introduce~\textbf{tainted} (\taintt) qualifiers that can be used to annotate pointer types or functions. 
+The tainted types form the untrusted partition or~\umode~\emph{region} (can only access tainted types), and the non-tainted types will be the trusted partition or~\cmode~\emph{region} (has complete access).
+Through static type checking, our type system ensures that tainted types cannot directly affect untainted types.
+After successful type checking,~\systemname{} creates two sets of source files~\ucregion{} and~\cregion{} containing tainted and untainted entities, respectively.
+
+These sets of source files will be compiled using~\systemname{} compiler, which enforces the desired isolation through instrumentation.
+Developers can choose a specific isolation mechanism.
+We explore four different isolation mechanisms with varying guarantees and performance overhead (CITE SECTION).
+For instance,~\isoheap{} provides safety against spatial and temporal violations of tainted pointers with almost no performance overhead.
+Whereas~\sandbox{} isolation provides complete~\ac{SFI} with~\avgsandboxoverhead{} runtime overhead.
+
+%Our dynamic instrumentation ensures that tainted pointer accesses are valid according to the desired isolation mechanism.
+\noindent\textbf{Formal Guarantee.} We formally prove the~\emph{clean separation theorem} that guarantees that tainted types are isolated from the rest of the programs in a well-typed program.
+The specific isolation guarantee depends on the target~\ac{SFI} mechanism.
+
+In addition to the flexibility, the type based partitioning has the following advantages:
+
+\noindent\textbf{Interactive Partitioning.} Our type system enables interactive partitioning, wherein developers start by annotating a risky pointer (\ie{} user data), and our type system will identify all the other additional pointers (depending on the tainted pointer) that need tainted or sanitized (\eg by explicitly copying to untainted buffer).
+
+\noindent\textbf{Integrating with Other Mechanisms.} Using types enables us to integrate easily with existing type based safety mechanisms.
+We demonstrate this by combining~\systemname{} with~\checkedc, which we call~\ccflex.
+\checkedc~\cite{Elliott2018} provides backward-compatible spatial safety by using checked pointer types. 
+But, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked (or wild) pointers in a~\checkedc program are prone to cross-language attacks~\cite{mergendahl2022cross} and can violate the whole program's spatial safety.
+However, using~\systemname{}, one can mark all unchecked (or wild) pointers as tainted, thus partitioning away unchecked types.
+These unchecked components can be executed in an isolated sandbox, thereby preventing cross-language attacks and upholding~\checkedc guarantees.
+
+Our evaluation shows that ---complete this. \aravind{Complete this.}
+
+In summary following are our contributions:
 

### 2023-04-06 sandp23/introduction.tex
Intro first draft ready

@@
 Through static type checking, our type system ensures that tainted types cannot directly affect untainted types.
-After successful type checking,~\systemname{} creates two sets of source files~\ucregion{} and~\cregion{} containing tainted and untainted entities, respectively.
+After successful type checking,~\systemname{} creates two sets of source files:~\ucregion{} and~\cregion{}, containing tainted and untainted entities, respectively.
 
@@
 We explore four different isolation mechanisms with varying guarantees and performance overhead (CITE SECTION).
-For instance,~\isoheap{} provides safety against spatial and temporal violations of tainted pointers with almost no performance overhead.
+For instance,~\isoheap{} provides safety against spatial and temporal violations of tainted pointers with almost~\emph{no} performance overhead.
 Whereas~\sandbox{} isolation provides complete~\ac{SFI} with~\avgsandboxoverhead{} runtime overhead.
@@
 In summary following are our contributions:
+
+\myparagraph{\systemname Type System and Formalism}
+We designed and implemented a flexible program partition technique using tainted types.
+Our type system guarantees that the tainted pointers cannot affect untainted pointers and partition the given code into~\ucregion (containing only tainted entities) and~\cregion (CITE SECTION).
+We provide formal proof of the~\emph{clean separation theorem} that guarantees that code in ~\cregion will be separated from~\ucregion according to the target isolation mechanism (CITE SECTION).
+We also integerated~\systemname{} with~\checkedc that provides additional~\emph{non-crashing} and \emph{non-exposure} guarantees (CITE SECTION).
+
+\myparagraph{Flexible Isolation Mechanisms and Implementation}
+We designed four isolation mechanisms with varying guarantees and performance overhead, providing flexibility in choosing a desired isolation mechanism (CITE SECTION).
+We implemented our type system (CITE SECTION) and flexible partitioning (CITE SECTION) by modifying~\clang{} compiler.
+
+\myparagraph{Evaluation}
+Our evaluation on ...
+
+We made our implementation and datasets open-source to foster future research.
+
+\iffalse
 
@@
 
+\fi
+
 %\input{figures/taintedsplit}

### 2023-04-07 sandp23/background.tex
Adding

@@
 \section{Background and Motivation}
+\aravind{I am working on this.}
 %\section{Overview and Transcendence}\label{sec:overview}
 %This section describes \checkedc{} and new features \systemname{} provides.
+
+\subsection{Memory Safety}
+This property ensures that all memory accesses a program makes to valid areas containing the program's data~\cite{payer2019software}.
+Languages that guarantee (or do not) this property are called memory-safe (or memory-unsafe) languages.~\eg Java is a memory-safe language, whereas C/C++ is not.
+There are two main classes of memory safety violations: spatial and temporal.
+Spatial violations occur when a program dereferences a pointer that points outside of the associated object.~\eg buffer overread, underread, array out-of-bounds access, null-ptr dereference, etc.
+Temporal violations occur when a program dereferences a pointer that points to an invalid object which was previously object.~\eg use-after-free, double-free, etc.
+
+\subsection{Program Partioning}
+\label{subsec:background:programpart}
+Program partitioning~\cite{rul2009towards} is a well-known technique for dividing a program into multiple isolated parts or partitions.
+There has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
+Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
+Specifically, given sensitive data in a program, all functions accessing sensitive data will be considered part of the trusted partition, and all other functions will be in the untrusted partition.
+For instance, in the following code~\code{key} is marked as sensitive and functions that access it,~\ie{}~\code{initkey} and~\code{encrypt} are considered to be part of trusted parition and the other function,~\ie{}~\code{main} will be in the untrusted partition.
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+char __attribute__((annotate("sensitive"))) *key;
+
+void initkey(int sz) {
+ ...
+ for (i=0; i<sz; i++) key[i]= ...;
+}
+
+void encrypt(char *ptxt, int sz) {
+ ..
+ for (i=0; i<sz; i++)
+  ctxt[i]=ptxt[i] ^ key[i];
+ ..
+}
+
+int main() {
+ ..
+ initkey(strlen(txt));
+ encrypt(txt, strlen(txt));
+ ... 
+}
+\end{minted}
+
+As we can see, partitioning is achieved easily in the existing techniques by identifying which functions access sensitive data.
+These partitions can be separated in various ways,~\eg{} as different processes communicating through~\ac{RPC}~\cite{nelson1981remote}, parent and child processes communicating through shared memory, etc.
+
+The main novelty of existing techniques is using various program analysis and dynamic instrumentation methods to facilitate transparent communication between functions in trusted and untrusted partitions.
+Because of their focus on inter-partition interactions, the existing techniques are specialized for the target isolation mechanism.
+For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling,~\eg{} sending data pointed by~\code{ptxt} and~\code{sz} during calls to~\code{encrypt} (\ie line 18 in the above example) and synchronizing the data back to~\code{ctxt} on return.
+Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
+Similarly,~\glamdring~\cite{lind2017glamdring} is specialized for partitioning using SGX-enclaves.
+
+However, these isolation mechanisms may not be available on all systems.
+Furthermore, the existing partioning techniques focus on sensitive data use-case -- where the main goal is to protect the integrity and confidentiality of the sensitive data.
+Such isolation mechanisms might be unnecessary or too expensive for security use cases.
+For instance, consider the case of isolating functions (\eg input parsing) that process tainted data. 
+Here, we do not care about the integrity or confidentiality of the tainted data but rather want to restrict the effect of these functions on the overall program.
+
+Consequently, developers might want to use lightweight isolation mechanisms with low-performance overhead and limited guarantees (\eg isolation against spatial safety issues).
+The existing techniques do not support such flexible use cases because of their tight coupling with the isolation mechanism.
+
+% SFI: mechanisms and its use cases.
+
+% Checked C
+
+% Motivation: vulnerabilities.
 Here, we brief \checkedc and the motivation for~\systemname{}.

### 2023-04-07 sandp23/introduction.tex
Adding

@@
 
-One of the main reasons for this is combining partitioning with~\acf{SFI}.
+One of the main reasons for this is combining partitioning with~\acf{SFI}~\cite{wahbe1993efficient}.
 Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region,~\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
@@
 For instance, we want to isolate a string processing function only from spatial violations,~\ie any spatial violations (\eg buffer overread) in that function should not affect the program.
+%
+But existing partitioning techniques are specialized for a specific~\ac{SFI} technique (\sect{subsec:background:programpart}).
+%For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling.
+%Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
 
-But existing partitioning techniques are specialized for a specific~\ac{SFI}~\cite{wahbe1993efficient} technique.
-For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling.
-Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
-In summary, existing techniques combine policy (\ie{}~\emph{what} to partition) with mechanism (\ie{}~\emph{how} to partition).
+In other words, existing techniques combine policy (\ie{}~\emph{what} to partition) with mechanism (\ie{}~\emph{how} to partition).
 We argue that existing program partitioning mechanisms are restrictive and inflexible for different use cases (\eg isolation from spatial safety vulnerabilities).

### 2023-04-07 sandp23/formal.tex
update formalism

@@
 \begin{array}{llcllcl}
-\text{Context Mode:} & m & ::= & \cmode \mid \umode \\
-\text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\
-     \text{Type:}& \tau &::=& \tint\mid \tptr{\tau}{ \xi}
+\text{Mode:} & m & ::= & \cmode \mid \umode \\
+     \text{Type:}& \tau &::=& \tint\mid \tptr{\tau}{ m}
 \\
@@
   \\
-&&\mid& \emalloc{\xi}{\omega} \mid \efree{e} \mid \elet{x}{e}{e}   \\
+&&\mid& \emalloc{m}{\omega} \mid \efree{e} \mid \elet{x}{e}{e}   \\
 &&\mid& \ret{x}{\evalue{n}{\tau}}{e} \mid \eif{e}{e}{e} \\
-&&\mid&\eunchecked{\overline{x}}{e}\mid \echecked{\overline{x}}{e}
+&&\mid&\toa{m}{\overline{x}}{e}
 \end{array}
@@
 We classify types as integers or pointers.
-Every pointer type ($\tptr{\tau}{\xi}$) includes a
-pointer mode annotation ($\xi$, the difference between context and pointer modes
-is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
-or unchecked (\umode), and a content type ($\tau$) denoting the valid type it is pointed to.
+Every pointer type ($\tptr{\tau}{m}$) includes a
+ mode annotation that is either tainted (\umode) and non-tainted (\cmode),
+and a content type ($\tau$) denoting the valid type it is pointed to.
 
@@
  static casts ($\ecast{\tau}{e}$),
-memory allocations ($\emalloc{\xi}{\omega}$) and free ($\efree{e}$), 
-along with expressions that require special handling, such as, tainted blocks ($\eunchecked{\overline{x}}{e}$),
-and checked blocks ($\echecked{\overline{x}}{e}$).
-
-We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
-For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
-The heap allocation $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in $\xi$ memory heap regions, i.e., heaps are separated into three different regions, classified by $\xi$.
-The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit \textit{code} regions,
-where the former represents the $\cmode$ code region in a program that are trusted by users, and the latter represents non-trust $\umode$ code region.
-To guarantee the clean separation property, we extend the C syntax to include $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$ blocks,
-where $\overline{x}$ represents all free variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain checked pointers.
+memory allocations ($\emalloc{m}{\tau}$) and free ($\efree{e}$), 
+along with expressions that require special handling, such as, context switching blocks ($\toa{m}{\overline{x}}{e}$).
+
+We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\tau}{m}$), enabling the use of fixed addresses as pointers.
+For example, $\evalue{0}{\tptr{\tau}{m}}$ (for any $m$ and $\tau$) represents a $\enull$ pointer.
+The heap allocation $\emalloc{m}{\tau}$ includes a mode flag $m$ for allocating memory in $m$ memory heap regions, i.e., heaps are separated into two different regions, classified by $m$.
+Expression $\toa{m}{\overline{x}}{e}$ is used to delimit \textit{code} regions, where $m$ indicates that $e$ locates in the $m$ code region.
+To guarantee the clean separation property, we extend the C syntax to include $\toa{m}{\overline{x}}{e}$ blocks,
+which represents the flag assigning to a function to indicate if the function is tainted or not.
+$\overline{x}$ represents all free variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain $\cmode$ mode pointers.
 $\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.
@@
 It prevents unsafe types from being constructed.
-  Consider the type ~\code{_t_ptr<_Ptr<int>>}, which describes a tainted pointer of a checked pointer.
-  This is not well-formed in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the unchecked ($\umode$) pointer is used. 
-  Nevertheless, we can have a checked pointer whose element is a tainted pointers\mzs{,}\mzr{:}~\eg~\code{_ptr<_t_Ptr<int>>} is a valid type.
+  Consider the type ~\code{_t_ptr<_Ptr<int>>}, which describes a tainted pointer of a $\cmode$ mode pointer.
+  This is not well-formed in~\lang{} because it potentially exposes the $\cmode$ mode pointer addresses in a $\umode$ region when the tainted ($\umode$) pointer is used. 
+  Nevertheless, we can have a $\cmode$ mode pointer whose element is a tainted pointers\mzs{,}\mzr{:}~\eg~\code{_ptr<_t_Ptr<int>>} is a valid type.
 
@@
 &&\mid \efree{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\\[0.2em]
-&& \mid \eif{E}{e}{e} \mid \eunchecked{\overline{x}}{E} \mid \echecked{\overline{x}}{E}
+&& \mid \eif{E}{e}{e} \mid \toa{m}{\overline{x}}{E}
 
@@
     \inferrule[T-Def]
-              {\xi \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\xi}}
+              {m' \leq m \\\Gamma \vdash_m e : \tptr{\tau}{m'}}
               {\Gamma \vdash_m \estar{e} : \tau}
 
-        \inferrule[S-Def]{\xi \neq \tmode \\ \heap(n)=\evalue{n_a}{\tptr{\tau_a}{\xi}} }
-    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
+        \inferrule[S-Def]{ \heap(n)=\evalue{n_a}{\tptr{\tau_a}{\umode}} }
+    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\umode}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}
 
@@
 
-    \inferrule[S-DefNull]{}{(\varphi,\Theta,\heap,\estar{\evalue{0}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\enull)}
+    \inferrule[S-DefNull]{}{(\varphi,\Theta,\heap,\estar{\evalue{0}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\Theta,\heap,\enull)}
 \qquad
-    \inferrule[S-DefBoundA]{n \in \Theta}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
+    \inferrule[S-DefBoundA]{n \in \Theta}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\umode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
 \qquad
-    \inferrule[S-DefBoundB]{\heap(n)=\evalue{n_a}{\tptr{\tau_a}{\xi}}\\\xi\neq \xi'}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi'}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
+    \inferrule[S-DefBoundB]{\heap(n)=\evalue{n_a}{\tptr{\tau_a}{m}}\\ m\neq m'}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{m'}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
   \end{mathpar}
@@
     \inferrule[T-Malloc]
-              {\xi \leq m}
-              {\Gamma \vdash_m \emalloc{\xi}{\tau} : \tptr{\tau}{\xi}}
+              {m' \leq m}
+              {\Gamma \vdash_m \emalloc{m'}{\tau} : \tptr{\tau}{m'}}
 
     \inferrule[T-Free]
-              {\xi \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\xi}}
+              {m' \leq m \\\Gamma \vdash_m e : \tptr{\tau}{m'}}
               {\Gamma \vdash_m \efree{e} : \tint}
 
-  \inferrule[S-Malloc]{\mathtt{alloc}(\heap,\Theta,\xi,\tau)=(n,\heap')}
-   { (\varphi,\Theta,\heap,\emalloc{\xi}{\tau}) \longrightarrow (\varphi,\Theta',\heap',\evalue{n}{\tptr{\tau}{\xi}})}
-
-  \inferrule[S-Free]{\xi \neq \tmode}
-   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{0}{\tint})}
+  \inferrule[S-Malloc]{\mathtt{alloc}(\heap,\Theta,m,\tau)=(n,\heap')}
+   { (\varphi,\Theta,\heap,\emalloc{m}{\tau}) \longrightarrow (\varphi,\Theta',\heap',\evalue{n}{\tptr{\tau}{m}})}
+
+  \inferrule[S-Free]{}
+   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\cmode}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{0}{\tint})}
 
   \inferrule[S-FreeT]{}
-   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\{n\}\cup\Theta,\heap,\evalue{0}{\tint})}
+   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\umode}}}) \longrightarrow (\varphi,\{n\}\cup\Theta,\heap,\evalue{0}{\tint})}
 
@@
 
-$\textcolor{blue}{\text{Checked and Unchecked Context Switches:}}\\$
+$\textcolor{blue}{\text{Context Switches:}}\\$
   \begin{mathpar}
-    \inferrule[T-Checked]
+    \inferrule[T-Region]
               {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                     \\\\\fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
-              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}
-
-    \inferrule[T-Unchecked]
-              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
-                \\\\ \fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
-              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}
-
-    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
+                     \\\fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_{m'} e : \tau}
+              {\Gamma;\Theta \vdash_m \toa{m'}{\overline{x}}{e} : \tau}
+
+    \inferrule[S-Region]{}{(\varphi,\heap,\toa{m}{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
   \end{mathpar}
@@
 \begin{array}{l}
-\tmode \le \xi
+\umode \le \cmode
 \qquad
@@
 \emph{stack} mapping from variables to values $\evalue{n}{\tau}$,
-$\Theta$ is a set of $\tmode$ mode memory addresses that are freed at the point of executing $e$
+$\Theta$ is a set of tainted ($\umode$ mode) memory addresses that are freed at the point of executing $e$
 and $\heap$ is a \emph{heap} mapping addresses (integer literals) to
-values $\evalue{n}{\ptr{\tau}{\xi}}$; for both we ensure
+values $\evalue{n}{\ptr{\tau}{m}}$; for both we ensure
 $\fv(\tau)=\emptyset$.
-In a heap value $\evalue{n}{\ptr{\tau}{\xi}}$,
-$\tau$ represents the value's content type, while $\xi$ indicates the mode of the heap location;
-we classify a heap location into $\cmode$, $\tmode$, and $\umode$ modes, corresponding to the three kinds of pointer modes.
+In a heap value $\evalue{n}{\ptr{\tau}{m}}$,
+$\tau$ represents the value's content type, while $m$ indicates the mode of the heap location;
+we classify a heap location into $\cmode$ and $\umode$ modes, corresponding to the two kinds of pointer modes.
 The complete set of typing rules are provided in~\Cref{rem-type,sec:rem-semantics}.
@@
 %We can model mutable stack variables as pointers into the mutable heap.
-As mentioned, value $\evalue{0}{\tptr{\tau}{\xi}}$
+As mentioned, value $\evalue{0}{\tptr{\tau}{m}}$
 represents a $\enull$ pointer when $\tau$ is a pointer type.
@@
 determines the context mode, i.e., region, that the expression $e'$ locates based on the context $E$.
-If the context hole locates in an $\euncheckedtext$, i.e., $E[\eunchecked{\Box}]$, the context region mode is $\umode$,
-while if the hole locates in a $\echeckedtext$, i.e., $E[\echecked{\Box}]$, its context mode is $\cmode$.
-In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.
+If the context hole locates in an $\totext(m)$, i.e., $E[\toa{m}{\overline{x}}{\Box}]$, the context region mode is $m$.
+In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an tainted function \code{server_loop}.
 
@@
 
-In \lang, Context modes $m$ appearing in a type rule determine the code region 
+In \lang, modes $m$ appearing in a type rule determine the code region 
 that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
-We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
-This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can only perform such operations in $\cmode$ and $\umode$ regions, respectively.
+We define a two point lattice $m_1 \le m_2$, with $\umode \le \cmode$, to describe such permission.
+This means that a $\umode$ mode pointer can be dereferenced and value-assigned in any region, while $\cmode$ mode pointers can only perform such operations in $\cmode$ region.
 \ignore{
@@
 reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
-The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).
-
-Rules \rulelab{S-DefT} and \rulelab{S-Def} are for $\tmode$ mode and other mode pointer dereferences, respectively.
+The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($m' \le m$).
+
+Rules \rulelab{S-DefT} and \rulelab{S-Def} are for $\umode$ mode and other mode pointer dereferences, respectively.
 Operation $\heap(n)$ not only accesses the $n$-th address location,
 but also performs a $\enull$ check because $0$ is undefined in all heaps.
-The pointer type ($\evalue{n}{\tptr{\tau}{\tmode}}$) in a dereference operation must have the same mode
-as the one labeled in the type in the heap value $(\evalue{n_a}{\tptr{\tau_a}{\xi}}$).
-The heap value type $\tptr{\tau_a}{\xi}$ is not an additional structure existed in the heap structure,
+The pointer type ($\evalue{n}{\tptr{\tau}{\umode}}$) in a dereference operation must have the same mode
+as the one labeled in the type in the heap value $(\evalue{n_a}{\tptr{\tau_a}{m}}$).
+The heap value type $\tptr{\tau_a}{m}$ is not an additional structure existed in the heap structure,
 but an annotation lives in the compiler and is compiled as a dynamic check to ensure that pointers only access heap locations with right modes. We explain the dynamic check generation shortly below.
 In addition to the above checks in rule \rulelab{S-Def},
-we perform checks in \rulelab{S-DefT} $n \not\in \Theta$ to ensure that location $n$ is not freed in the $\tmode$ heap region.
+we perform checks in \rulelab{S-DefT} $n \not\in \Theta$ to ensure that location $n$ is not freed in the $\umode$ heap region.
 If the above dynamic checks are violated, rules \rulelab{S-DefNull}, \rulelab{S-DefBoundA}, and \rulelab{S-DefBoundB} are performed, meaning that the evaluation is transitioned into a runtime error that is correctly captured by the \systemname compiler.
@@
 the type and semantic rules for $\texttt{malloc}$ and $\texttt{free}$ operations.
-Rules $\rulelab{T-Malloc}$ and $\rulelab{T-Free}$ make sure that $\cmode$ and $\umode$ mode pointers are allocated and freed only in the $\cmode$ and $\umode$ regions, respectively, while $\tmode$ mode pointers are allowed to allocate and free in both regions.
-The $\mathtt{alloc}(\heap,\Theta,\xi,\tau)$ function in rule \rulelab{S-Malloc} allocates a new pointer in the $\xi$ heap region.
-If $\xi$ is $\tmode$ and $\Theta$ is not empty, the function picks a memory space in $\Theta$ as the new pointer;
-otherwise, the function picks the $|\heap(\xi)|$ as the new pointer and increments $\heap(\xi)$'s cardinality.
-The new location is labeled with a type annotation $\tptr{\tau}{\xi}$, indicating that the location is $\xi$ mode.
-Every time when a deference and assignment is applied on a $\tmode$ pointer, the \systemname compiler applies a dynamic check to verify if the location is labeled with $\tmode$.
-We can think of the dynamic check as $n \in \heap(\tmode)$, checking if a location $n$ is in a set $\heap(\tmode)$ containing all the $\tmode$ addresses.
-When a $\texttt{malloc}$ function is called to create a new $\tmode$ pointer, we expand set $\heap(\tmode)$ to include the new address.
+Rules $\rulelab{T-Malloc}$ and $\rulelab{T-Free}$ make sure that $\cmode$ mode pointers are allocated and freed only in the $\cmode$ region, while $\umode$ mode pointers are allowed to allocate and free in both regions.
+The $\mathtt{alloc}(\heap,\Theta,m,\tau)$ function in rule \rulelab{S-Malloc} allocates a new pointer in the $m$ heap region.
+If $m$ is $\umode$ and $\Theta$ is not empty, the function picks a memory space in $\Theta$ as the new pointer;
+otherwise, the function picks the $|\heap|$ as the new pointer and increments $\heap$'s cardinality.
+The new location is labeled with a type annotation $\tptr{\tau}{m}$, indicating that the location is $m$ mode.
+Every time when a deference and assignment is applied on a $\umode$ pointer, the \systemname compiler applies a dynamic check to verify if the location is labeled with $\umode$.
+We can think of the dynamic check as $n \in \heap(\umode)$, checking if a location $n$ is in a set $\heap(\umode)$ containing all the $\umode$ addresses.
+When a $\texttt{malloc}$ function is called to create a new $\umode$ pointer, we expand set $\heap(\umode)$ to include the new address.
 See \Cref{sec:compilation}.
-Rules \rulelab{S-FreeT} and \rulelab{S-Free} define the \texttt{free} operation for $\tmode$ mode and other mode pointers, respectively.
-Rule \rulelab{S-Free} "forgets" about the freed pointer, while \rulelab{S-FreeT} puts the freed $\tmode$ mode pointer address in the $\Theta$ set.
-
-\myparagraph{Unchecked and Checked Blocks}
-The execution of a $\echeckedtext$ or $\euncheckedtext$ block represents 
+Rules \rulelab{S-FreeT} and \rulelab{S-Free} define the \texttt{free} operation for $\umode$ mode and other mode pointers, respectively.
+Rule \rulelab{S-Free} "forgets" about the freed pointer, while \rulelab{S-FreeT} puts the freed $\umode$ mode pointer address in the $\Theta$ set.
+
+\myparagraph{Context Switching Blocks}
+The execution of a $\totext$ block represents 
 the context switching from a $\cmode$ to an $\umode$ region, or vice versa,
 with its type and semantic rules given in \Cref{fig:type-system-1}.
-In this context switching, to guarantee the checked ($\cmode$) pointer non-exposure property, i.e.,
-checked pointers are not allowed to go cross different regions, which is guaranteed by the predicates 
+In this context switching, to guarantee the $\cmode$ mode pointer non-exposure property, i.e.,
+$\cmode$ pointers are not allowed to go cross different regions, which is guaranteed by the predicates 
 $\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$ \footnote{$\cmode(-)$ checks the mode of a pointer.},
@@
 For example, \code{StringAuth} in \Cref{subsub:gencregion} is a trampoline function that disallows
-checked pointers as arguments and return values.
+$\cmode$ pointers as arguments and return values.
 The use of the function in the following \code{_T_StringAuth}, which is in $\umode$ region,
-cannot legally acknowledges any checked pointers; otherwise, we might expose a checked pointer address to unsafe code regions.
-In \lang, we actually permits the accesses of checked pointers inside \code{StringAuth},
+cannot legally acknowledges any $\cmode$ pointers; otherwise, we might expose a $\cmode$ pointer address to unsafe code regions.
+In \lang, we actually permits the accesses of $\cmode$ pointers inside \code{StringAuth},
 since the function body of a trampoline function is in $\cmode$ region.
@@
 We first show the non-exposure theorem,
-where code in $\umode$ region cannot observe a valid checked ($\cmode$) pointer address.
+where code in $\umode$ region cannot observe a valid $\cmode$ pointer address.
 
@@
 \longrightarrow_{\umode} (\varphi',\Theta',\heap',e')$ and $e=E[\alpha(x)]$ and $\mode(E)=\umode$,
-where $\alpha(x)$ is some expression (not $\echeckedtext$ nor $\euncheckedtext$) containing variable $x$; 
-thus, it is not a checked pointer.
+where $\alpha(x)$ is some expression (not $\totext$) containing variable $x$; 
+thus, it is not a $\cmode$ pointer.
 \end{thm}
@@
 int func_t(_t_Ptr<char> p : count(5)) {
-  checked (p) { //c region code
+  to(c) (p) { //c region code
    char * q;
@@
 
-\item $e=\estar{\evalue{n}{\tptr{\tau}{\tmode}}}$.
-
-\item $e=\eassign{\evalue{n}{\tptr{\tau}{\tmode}}}{\evalue{n'}{\tau'}}$.
+\item $e=\estar{\evalue{n}{\tptr{\tau}{\umode}}}$.
+
+\item $e=\eassign{\evalue{n}{\tptr{\tau}{\umode}}}{\evalue{n'}{\tau'}}$.
 

### 2023-04-07 sandp23/background.tex
Adding all

@@
 \section{Background and Motivation}
+This section presents the necessary background and motivation for a system like~\systemname{}.
 \aravind{I am working on this.}
@@
 There are two main classes of memory safety violations: spatial and temporal.
-Spatial violations occur when a program dereferences a pointer that points outside of the associated object.~\eg buffer overread, underread, array out-of-bounds access, null-ptr dereference, etc.
-Temporal violations occur when a program dereferences a pointer that points to an invalid object which was previously object.~\eg use-after-free, double-free, etc.
+\emph{Spatial violations} occur when a program dereferences a pointer that points outside of the associated object.~\eg buffer overread, underread, array out-of-bounds access, null-ptr dereference, etc.
+\emph{Temporal violations} occur when a program dereferences a pointer that points to an invalid object which was previously object.~\eg use-after-free, double-free, etc.
 
@@
 Furthermore, the existing partioning techniques focus on sensitive data use-case -- where the main goal is to protect the integrity and confidentiality of the sensitive data.
-Such isolation mechanisms might be unnecessary or too expensive for security use cases.
+This extreme isolation also incurs significant performance overhead (37\% - 163\%)~\cite{lind2017glamdring, liu2017ptrsplit}.
+But such isolation mechanisms might be unnecessary or too expensive for security use cases.
 For instance, consider the case of isolating functions (\eg input parsing) that process tainted data. 

### 2023-04-08 sandp23/background.tex
finishing up background

@@
 \section{Background and Motivation}
-This section presents the necessary background and motivation for a system like~\systemname{}.
+This section presents the necessary background and motivation for our work.
 \aravind{I am working on this.}
@@
 Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
-Similarly,~\glamdring~\cite{lind2017glamdring} is specialized for partitioning using SGX-enclaves.
+Similarly,~\glamdring~\cite{lind2017glamdring} is specialized for partitioning using SGX-enclaves~\cite{costan2016intel}.
 
-However, these isolation mechanisms may not be available on all systems.
-Furthermore, the existing partioning techniques focus on sensitive data use-case -- where the main goal is to protect the integrity and confidentiality of the sensitive data.
-This extreme isolation also incurs significant performance overhead (37\% - 163\%)~\cite{lind2017glamdring, liu2017ptrsplit}.
-But such isolation mechanisms might be unnecessary or too expensive for security use cases.
-For instance, consider the case of isolating functions (\eg input parsing) that process tainted data. 
-Here, we do not care about the integrity or confidentiality of the tainted data but rather want to restrict the effect of these functions on the overall program.
+\subsection{Need for Selective Fault Isolation}
+The existing partitioning techniques focus on sensitive data use-case -- where the main goal is to protect the integrity and confidentiality of the sensitive data.
+Consequently, they try to achieve complete~\acf{SFI},~\ie all executions, data, or faults within a partition do not affect other partitions.
+This extreme isolation incurs significant performance overhead (37\% - 163\%)~\cite{lind2017glamdring, liu2017ptrsplit}.
+Some isolation mechanisms (\eg Intel SGX) may not be available on all systems.
+Such extreme isolation may not be needed or unnecessary for practical security use cases.
 
-Consequently, developers might want to use lightweight isolation mechanisms with low-performance overhead and limited guarantees (\eg isolation against spatial safety issues).
-The existing techniques do not support such flexible use cases because of their tight coupling with the isolation mechanism.
+As shown by the recent study from Microsoft, 70\% of the vulnerabilities still occur due to spatial safety violations.
+Several works show that certain functions are more prone to have a certain class of vulnerabilities than others.
+Moreover, certain vulnerabilities may be impossible to occur in a given function.
+For instance, a function that cannot invoke (directly or transitively) any heap operations (\eg{} malloc/free) cannot contain temporal violations,~\eg use-after-free.
+Instead of complete isolation, we need~\emph{\acf{SeFI}} from a program partition.
+Formally, we say that a partition is $\phi$-\ac{SeFI} if and only if any violation of the property $\phi$ is contained with the partition and does not affect other partition(s).
+For instance,~\emph{spatial safety-\ac{SeFI}} from a partition provides isolation against spatial safety violations in the partition.
+
+\subsection{\checkedc}
+\label{subsec:checkedc}
+\label{subsec:nosafetyagsintuncheckedcode}
+Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
+open-source extension to C,
+to ensure a program’s spatial safety by introducing new pointer types,~\ie checked pointer types.
+Which are represented as system-level memory words without ``fattening'' metadata~\cite{duck2016heap}, 
+and ensuring backward compatibility,~\ie developers can use checked and
+regular (unchecked or wild) pointers within the same program.
+
+\noindent
+\myparagraph{Checked Pointer Types}
+\checkedc{} introduces three varieties of \emph{checked pointer}:
+\begin{itemize}
+\item \code{_Ptr<}$T$\code{>} ($\ptr$) types a pointer that is either null or
+  points to a single object of type $T$.
+\item \code{_Array_ptr<}$T$\code{>} ($\arr$) types a pointer that is either null
+  or points to an array of $T$ objects. The array width is defined
+  by a \emph{bounds} expression, discussed below.
+\item \code{_NT_Array_ptr<}$T$\code{>} ($\ntarr$) is like
+  \code{_Array_ptr<}$T$\code{>} except that the bounds expression
+  defines the \emph{minimum} array width---additional objects may
+  be available past the upper bound, up to a null terminator.
+\end{itemize}
+Both $\arr$ and $\ntarr$ pointers have an associated bounds which defines the
+range of memory referenced by the pointer.
+The three different ways to declare bounds and the corresponding memory range is:
+\begin{footnotesize}
+\begin{tabular}{ll}
+\arrT{|$T$|} \inlinecode{p: count(|$n$|)}
+  &
+$[\inlinecode{p}, \inlinecode{p}+\inlinecode{sizeof}(T) \times n) $ \\
+%$[\inlinecode{p}, )$ \\
+\arrT{|$T$|} \inlinecode{p: byte_count(|$b$|)}
+
+  &
+    $[\code{p}, \code{p}+b)$ \\
+
+\arrT{|$T$|} \inlinecode{p: bounds(|$x, y$|)}
+
+  &
+    $[x, y)    $\\  
+\end{tabular}
+\end{footnotesize}
+The bounds can be declared for $\ntarr$ as well, but the memory range can extend further to the right,
+until a~\code{NULL} terminator is reached (\ie \code{NULL} is not within the bounds).
+
+\noindent
+\myparagraph{Ensuring Spatial Memory Safety}
+The \checkedc compiler instruments loads and stores of checked
+pointers to confirm the pointer is non-null, and additionally the access to $\arr$ and $\ntarr$ pointers is within
+their specified bounds.
+For example, in the code \inlinecode{if (n>0) a[n-1] =} $...$ the write
+is via address $\alpha = \inlinecode{a + sizeof(int)}\!\times\!\code{(n-1)}$. 
+If the bounds of \code{a} are \code{count(u)}, the
+inserted check confirms
+$\inlinecode{a} \leq \alpha < \inlinecode{a + sizeof(int)} \!\times\!
+\inlinecode{u}$ prior to dereference.
+Failed checks throw an exception.
+Oftentimes, inserted checks can be optimized away by LLVM resulting in almost no runtime overhead~\cite{duanrefactoring}.
+
+\myparagraph{Backward Compatibility}
+\checkedc is backward compatible with legacy C as all legacy code will type-check and compile.
+However, the compiler adds the aforementioned spatial safety checks to only checked pointers.
+The spatial safety guarantee is partial when the code is not fully ported.
+%Specifically, only code that appears in~\emph{checked code regions}, is guaranteed to be spatially safe.
+
+\myparagraph{No Safety Against Unchecked Pointers}
+A partially annotated program can still enjoy spatial safety only if checked pointers do not communicate with any unchecked ones.
+For instance, in the example below, there are no spatial safety violations in the function~\code{func} as it uses only checked pointers.
+However, the other unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
+
+Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in checked regions but rather states that checked regions~\emph{cannot be blamed for any spatial safety violations}.
+Consider the following example:
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+// c region code
+int func(array_ptr<char> p : count(5)) {
+|\textcolor{red}{\faChainBroken}|..p[4]..
+}
+// u region code
+...
+str = "he";
+...
+func(|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5)); 
+\end{minted}
+Here, the checked function~\inlinecode{func} expects a pointer to a buffer of five elements, but the unchecked region code invokes the function (Line 9) with a buffer of 2 elements.
+This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the checked function, but of course, the blame or the root cause is in the unchecked function (\textcolor{red}{\faBug}).
+Furthermore, since checked and unchecked functions execute in the same address space, spatial memory corruptions (\eg buffer overflow) in unchecked functions (or regions) can take down the complete program despite
+having checked regions.
+We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked functions does not violate the safety guarantees in checked functions}.
 
@@
 % Motivation: vulnerabilities.
-Here, we brief \checkedc and the motivation for~\systemname{}.
 
+\iffalse
 \subsection{\checkedc}
@@
 \cite{liu2017ptrsplit}, is prohibitive and varies from 37\%-163\%.
-
+\fi
 \iffalse

### 2023-04-08 sandp23/background.tex
Finishing up background

@@
 This section presents the necessary background and motivation for our work.
-\aravind{I am working on this.}
 %\section{Overview and Transcendence}\label{sec:overview}
@@
 For instance, in the following code~\code{key} is marked as sensitive and functions that access it,~\ie{}~\code{initkey} and~\code{encrypt} are considered to be part of trusted parition and the other function,~\ie{}~\code{main} will be in the untrusted partition.
-\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\scriptsize{}]{c}
 char __attribute__((annotate("sensitive"))) *key;
@@
 
-As shown by the recent study from Microsoft, 70\% of the vulnerabilities still occur due to spatial safety violations.
-Several works show that certain functions are more prone to have a certain class of vulnerabilities than others.
+As shown by the recent study from Microsoft~\cite{microsoftmemsafe}, 70\% of the vulnerabilities still occur due to spatial safety violations.
+Several works~\cite{meng2021bran, du2019leopard, ding2022velvet, situ2021vulnerable} show that certain functions are more prone to have a certain class of vulnerabilities than others.
 Moreover, certain vulnerabilities may be impossible to occur in a given function.
@@
 Consider the following example:
-\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
-// c region code
+\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\scriptsize{}]{c}
+// code with checked c annotations
 int func(array_ptr<char> p : count(5)) {
@@
 }
-// u region code
+// unchecked (wild) code
 ...

### 2023-04-08 sandp23/examples/backup/callback_mechanism_sample_after_checkmate.c
Finishing up overview

@@ -1,0 +1,52 @@
+#include <string_tainted.h>
+
+_Callback _TPtr<char> StringAuth(_TPtr<const char> input , size_t len) {
+   ...
+   //complex Function Body
+   ...
+}
+
+/*
+ * CheckMate Inserted Callback Registration
+ */
+
+void registerCallback_StringAuth(void){
+    //callback function signature {ret <- int, arg_1 <- int, arg_2 <- long}
+    int ret_param_types[] = {0, 0, 1}; 
+    // 2 <- arg count, 1 <- ret count 
+    __StringAuth__C_ = _SBXREG_((void*)_T_StringAuth,2,1, ret_param_types);
+}
+
+/*
+ * CheckMate Inserted Trampoline for Directing Callbacks from WASM sandbox
+ */
+
+unsigned int _T_StringAuth(unsigned sandbox,
+                                 unsigned int arg_1,
+                                 unsigned long int arg_2) {
+    return StringAuth(arg_1, (size_t)arg_2);
+}
+
+_Tainted _TPtr<char> StringProc(_TPtr<_TPtr<const char>> string,
+_TPtr<_TPtr<char>(_TPtr<const char> input, size_t len)>StringAuth) {
+	
+   // StringAuth as a function ptr arg is only used as a 
+   //_Callback signature to CheckMate and for type-checker
+   return w2c_StringProc(_SBX_(), (unsigned int)string, __StringAuth__C_);
+}
+
+_Mirror int isStringMatching(char* str1 : itype (_Ptr<char>),
+                char* str2 : itype(_Ptr<char>))
+{
+  ...
+  //complex Function body
+  /*
+   * "_Mirror" restricts access to global data
+   *
+   * Since "_Mirror" functions are locally replicated in both
+   * checked and unchecked regions, they have no restriction on
+   * type signature and callees.
+   */
+  ...
+}
+

### 2023-04-08 sandp23/examples/backup/conv1.c

@@ -1,29 +1,34 @@
 |\entrypoint| // Entry point
 void server_loop(int sock_fd) {
  unsigned b_z;
- struct queue *p;
+ |\textcolor{checkcolor}{\_Array\_ptr}|<struct queue> p : 
+              count(b_z) = NULL;
  ...
  while(1) {
   ...
   p = malloc(sizeof(struct queue) * b_z);
   ...
-  if (handle_request(sock_fd)) {...}
+  if (handle_request(sock_fd)) {
+    ...  
+  }
  }
 }
 
 int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE] |\textcolor{blue}{\encircle{\textbf{3}}} \faGears\usermods\textcolor{taintcolor}{\_Tainted}|;
+ char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_Tainted}|;
  int rc = -1;
- ssize_t r_len;
- r_len = read(sock_fd, buff, 
-              MAX_MSG_SIZE);
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, 
+                  MAX_MSG_SIZE);
  if (r_len > 0 && 
      r_len < MAX_MSG_SIZE) {
   switch(buff[0]) {
    case REQ1: 
     rc = process_req1(buff, r_len); 
     break;
-   case REQ2:
+   case REQ2: 
+    rc = process_req2(buff, r_len); 
+    break;
    ...
   }
  }

### 2023-04-08 sandp23/examples/backup/orig1.c

@@ -7,23 +7,27 @@ void server_loop(int sock_fd) {
   ...
   p = malloc(sizeof(struct queue) * b_z);
   ...
-  if (handle_request(sock_fd)) {...}
+  if (handle_request(sock_fd)) {
+    ...  
+  }
  }
 }
 
 int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE] |\textcolor{blue}{\encircle{\textbf{3}}} \faGears\usermods\textcolor{taintcolor}{\_Tainted}|;
+ char buff[MAX_MSG_SIZE];
  int rc = -1;
- ssize_t r_len;
- r_len = read(sock_fd, buff, 
-              MAX_MSG_SIZE);
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, 
+                  MAX_MSG_SIZE);
  if (r_len > 0 && 
      r_len < MAX_MSG_SIZE) {
   switch(buff[0]) {
    case REQ1: 
     rc = process_req1(buff, r_len); 
     break;
-   case REQ2:
+   case REQ2: 
+    rc = process_req2(buff, r_len); 
+    break;
    ...
   }
  }

### 2023-04-08 sandp23/examples/backup/originalprogram.c

@@ -1,0 +1,68 @@
+size_t read_msg(int sock_fd, char *msg, size_t sz) {
+ size_t nr;
+ nr = read(sock_fd, (void*)msg, sz);
+ ...
+ // complex code
+ return nr;
+}
+
+int process_req1(char *msg, size_t m_l) {
+ int rc = -1;
+ //complex code
+ if (m_l > MIN_SIZE) {
+   //complex code
+   sscanf(msg, "%d", &i);
+   msg[i] = ...
+ }
+ //complex code
+ return rc;
+}
+
+int process_req2(char *msg, size_t m_l) {
+ size_t i = 0, j = 0;
+ int rc = -1;
+ char anum[MAX_MSG_SIZE] = {0};
+ if (msg) {
+  while(i < m_l && j < MAX_MSG_SIZE-1) {
+   ...
+   if (isalnum(msg[i]))
+    anum[j++] = msg[i]; 
+   i++;
+  }
+   rc = process_data(anum);
+ }
+ return rc;
+}
+
+int handle_request(int sock_fd) {
+ char buff[MAX_MSG_SIZE];
+ int rc = -1;
+ size_t r_len;
+ r_len = read_msg(sock_fd, buff, MAX_MSG_SIZE);
+ if (r_len > 0) {
+  switch(buff[0]) {
+   case REQ1: 
+    rc = process_req1(buff, r_len); 
+    break;
+   case REQ2: 
+    rc = process_req2(buff, r_len); 
+    break;
+   ...
+  }
+ }
+ return rc;
+}
+
+void server_loop(int sock_fd) {
+ unsigned b_z;
+ struct queue *q;
+ ...
+ while(1) {
+  ...
+  p = malloc(sizeof(struct queue) * b_z);
+  ...
+  if (handle_request(sock_fd)) {
+    ...  
+  }
+ }
+}

### 2023-04-08 sandp23/examples/orig2.c

@@ -1,34 +1,32 @@
-size_t read_msg(int sock_fd, char *msg, 
-                size_t sz) {
- size_t nr;
- nr = read(sock_fd, |\rootcause|(void*)msg, sz);
- ...
- |\complexcode| // complex code
- return nr;
-}
-
-int process_req1(char *msg, size_t m_l) {
+int process_req1(char *msg |\textcolor{blue}{\encircle{\textbf{2}}} \faGears\usermods\textcolor{taintcolor}{\_Tainted}| , 
+                 size_t m_l) 
+                 |\textcolor{blue}{\encircle{\textbf{1}}} \usermods\textcolor{taintcolor}{\_Tainted}| {
  int rc = -1, i;
  if (m_l > MIN_SIZE) {
-   sscanf(msg, "%d", &i);
-   |\realbug| msg[i] = ...
- }
- |\complexcode| // complex code
- return rc;
-}
-
-int process_req2(char *msg, size_t m_l) {
- size_t i = 0, j = 0;
- int rc = -1;
- char anum[MAX_MSG_SIZE] = {0};
- if (msg) {
-  while(i < m_l && j < MAX_MSG_SIZE-1) {
+  msg += sscanf(msg, "%d", &i);
+  if (i > 0) {
+    char *cp1 |\textcolor{blue}{\encircle{\textbf{2}}} \faGears\usermods\textcolor{taintcolor}{\_Tainted}|;
+    char *cp2 |\textcolor{blue}{\encircle{\textbf{2}}} \faGears\usermods\textcolor{taintcolor}{\_Tainted}|;
+    char tmp[CMD_S] |\textcolor{blue}{\encircle{\textbf{2}}} \faGears\usermods\textcolor{taintcolor}{\_Tainted}|;
+    for ( cp1 = msg, cp2 = tmp;
+        *cp1 != '\0' && 
+        cp2 - tmp < CMD_S - 1;
+        ++cp1, ++cp2 )
+    {
+     switch ( *cp1 )
+     {
+       case '<':
+        // We write 3 charecters into
+        // cp2 but we ony checked for 1
+        // charecter.
+        *cp2++ = '&';
+        |\realbug| *cp2++ = 'l';
+        |\realbug| *cp2++ = 't';
+        |\realbug| *cp2 = ';';
+      break;
+       ...
+     } }
    ...
-   if (isalnum(msg[i]))
-    anum[j++] = msg[i]; 
-   i++;
-  }
-   rc = process_data(anum);
- }
+  } }
  return rc;
 }

### 2023-04-16 csf2024/introduction.tex
update introduction.

--- 2023-01-20 sandp23/introduction.tex
@@
 
-Vulnerabilities due to memory corruption, especially spatial memory corruption, 
-are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798} 
-despite many efforts that tried to prevent them~\cite{song2019sanitizing}.
-Several industrial and research efforts, including CCured~\cite{Necula2005},
-Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012},
-have investigated ways to better compile C programs with automatic spatial safety enforcement.
-These approaches all impose performance overheads deemed too high for deployment use. 
-% 
-%\liyi{Should we say this? Is the paper about overhead reducing? }
-Recently,~\citet{Elliott2018} and~\citet{li22checkedc} introduced and formalized \checkedc, an
-open-source extension to C,
-to ensure a program’s spatial safety by introducing new pointer types,~\ie checked ($\cmode$) pointer types.
-%
-The checked pointers are represented as system-level memory words without ``fattening'' metadata~\cite{duck2016heap}, 
-and ensuring backward compatibility,~\ie developers can use checked and regular (unchecked $\umode$) pointers within the same program.
-%users are able to split code into checked and unchecked regions and incrementally convert
-%C code in the unchecked region to Checked C code in the checked one.
-However, as we explain in~\sect{subsec:nosafetyagsintuncheckedcode}, the
-unconverted or unchecked ($\umode$) code can violate guarantees provided in $\cmode$
-regions.
-%
-We need to ensure that~\emph{code executed as part of 
-unchecked ($\umode$) regions does not lead to the safety violations in checked ($\cmode$) regions  with the use of program partitioning mechanism \cite{rul2009towards}}.
+Vulnerabilities due to memory corruption are still a major issue for C programs~\cite{cvetrend, microsoftmemsafe, Zeng:2013:SRF:2534766.2534798}  despite many efforts that tried to prevent them~\cite{song2019sanitizing}.
+These vulnerabilities are not uniformly distributed in real programs~\cite{meng2021bran, du2019leopard},
+as some components are likelier to have a certain class of vulnerabilities.
+\eg{} spatial safety issues in string processing functions.
 
-% One possible solution is Checked C
-% \begin{itemize}
-% \item Think of it as migratory/gradual typing, but for C.
-% \item Distinct pointer types, but which are backward binary- and source-compatible.
-% \item Checked regions, containing only checked pointers and restricted
-%   idioms, aim to ensure spatial safety
-% \item Implemented as an extension to Clang/LLVM. Good performance
-%   (compared to ASAN etc.)
-% \end{itemize}
-\iffalse
-\noindent
-\textbf{No safety against unchecked code.} The \checkedc spatial safety guarantee applies to completely converted programs, i.e., programs that uses only checked types.
-Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
-But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
-Although the blameless proof exists~\cite{ruef2019achieving} for~\checkedc, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
-Consider the following example:
-\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
-// Checked code
-int func(array_ptr<char> p : count(5)) {
-|\textcolor{red}{\faChainBroken}|..p[4]..
-}
-// unchecked code
-...
-str = "he";
-...
-|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5); 
-...
-char ptr[16];
-...
-len <- derived from user input
-...
-|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
-\end{minted}
-Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function with a buffer of 2 elements.
-This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
-We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
-\fi
-%
+Program partitioning enables separating potentially vulnerable functions (untrusted region) from the rest of the program (trusted region).
+Several program partitioning mechanisms~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit,rlbox-paper,10.1145/3492321.3519582,privtrans,cqual-kernel-ptr,10.1145/3371100} exist.
+These techniques usually combine program partitioning with SFI to guarantee robust safety \cite{10.1145/3371100}, such that code in the untrusted region cannot affect code in the trusted region,~\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a trusted execution environment~\cite{lind2017glamdring}.
 
-Existing such mechanisms are not suitable as they are based on process isolation and have high overhead, and are \emph{data-centric} (\sect{subsec:background:programpart}).
-%Furthermore, these techniques are hard to engineer to co-exist with~\checkedc.
-%\liyi{why? Can we not say this? It seems that the whole reason people should care our work is we did it in \checkedc, but \checkedc itself is the selling product here.... }
-But in our case, we want a low-overhead code-centric partitioning, where the $\umode$ region code (or functions) should be isolated (or partitioned) from $\cmode$ one.
-We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing $\cmode$ region code can maintain spatial safety.
+Unfortunately, the robust safety statement is too vague in expressing the exact safety it provides, and it causes the systems based on it --- such as the ones above --- provide stronger guarantees than they should be and cause considerable performance overhead, but also provide ways where users can reduce overhead by essentially violating the robust safety.
+For example, RLBox \cite{rlbox-paper} and PKRU-Safe \cite{10.1145/3492321.3519582} have a $100\%$ - $200\%$ overhead, and RLbox permits users to bypass its verification procedures that guarantee robust safety.
+The ambiguity and its associated performance drawback decelerates the mechanism's development \cite{sandboxfu}.
 
-Here, we propose a type-directed code-centric program partitioning approach.
-Specifically, our system,~\systemname, extends \checkedc's checked and unchecked pointer types---representing safe and unsafe program pieces---with \textbf{tainted} (\taintt) types running on an isolated sandbox mechanism,
-forbids the communication between checked and unchecked type entities, and enforces the communication between checked and unchecked types
-through the uses of tainted types with additional validity checks. 
-%which can be used to mark functions and pointers that need to be isolated from the original program, while checked types follow the standard \checkedc typing rules.
-%However, our type-system disallows any unsafe interactions between taint and untainted types.
-% 
-% Specifically, in our system,~\systemname, extending CheckedC,
-% functions and pointers that need to be isolated from the original program can be
-% marked with ~\textbf{tainted} (\taintt) types.
-%
+In all these program partitioning mechanisms, the program partitioning parts are clear and similar.
+They provided a type system or user defined annotations to permit the identification of untrusted data, perform some memory separation techniques (separate heaps or sandbox mechanism) to separate the allocations of trusted and untrusted data, and insert dynamic checks to restrict the communications of these two data.
+The differences in these mechanisms appeared in the SFI part for separating completely the trusted and untrusted regions that respectively contain trusted and untrusted data, as well as the user defined verification procedures for the permission of some untrusted data usage in trusted regions, which are also the performance bottlenecks.
+To balance safety guarantees and performance overhead, we first need to investigate the guarantees that program partitioning can provide, without the involvement of SFI.
 
-The developer starts by marking desired (\ie unchecked, $\umode$) functions and pointers used in functions as tainted. 
-Then, \systemname partitions the given program into two partitions (\umode and~\cmode regions) of different privileges:
+In a program partitioning mechanism, a program is split into two regions: a \umode~\emph{region} representing the untrusted partition, and the \cmode~\emph{region} representing the trusted partition.
+Typically, there is a special types of pointers, which we named \textbf{tainted} (\tmode) mode pointers, being introduced to be used in communicating the \umode and \cmode regions, i.e.,
+untainted pointers cannot be used in \umode region, while there are checks requiring for \tmode mode pointers to be used in \cmode region.
+Users annotate pointer types and there is a static procedure, such as type checking, then creates two sets of source files:~\ucregion{} and~\cregion{}, containing tainted and untainted entities, respectively.
+In addition, untainted pointers that are used in \cmode region are typically viewed as "good" pointers, named \emph{checked} (\cmode mode) pointers.
+Users classify \cmode mode pointers and \cmode region in a program because they consider these entities are safe to use, which indicates that the guarantee assumptions in \cmode and \umode regions are different and the safety property should reflect such differences.
 
-\begin{itemize}
-\item~\umode \emph{region} (low privilege tainted region, extended from the unchecked region in \checkedc): this partition contains tainted types (\ie functions and pointers) and can only access tainted and unchecked pointers.
-\item~\cmode \emph{region} or~\emph{safe region} (high privilege untainted or checked region): This partition contains the remaining (untainted) code and data and has complete access to~\cmode region.
-The functions in~\cregion can invoke any function in~\ucregion and access all its data but not the other way around, except for call-back functions, which we will discuss later.
-\end{itemize}
+We present \systemname, as a formalism for the above type-directed program partitioning mechanism based on \checkedc,
+and show that the memory safety the mechanism provides is the \emph{non-crashing} property, i.e.,
+\cmode region never experiences crashing due to memory safety issues regardless the execution of \umode region.
+We also show that the system provides \emph{clean separation} property as a corollary of the non-crashing property,
+i.e., any memory violation in \umode region does not affect the execution of \cmode region.
 
-The $\cmode$ region code is executed as a regular program, while the~\ucregion partition will be executed in an existing sandboxed environment (\eg WASM sandbox), with additional instrumentations to facilitate the communication between code in~$\cmode$ and~$\umode$ regions.
+The system is based on \checkedc because of the third concern above.
+In privilege separation, it is obvious that trusted code is viewed as the part that does not have problem, while programs are in untrusted code, which is similar to the concept of checked and unchecked code regions in \checkedc and make it a perfect model to formalize type-directed program partitioning.
+We make precise about the guarantees that a program partition mechanism can provide and have the following contributions.
 
-The combination of tainted types and privileged partitions
-enforces isolation and provides memory safety without transforming all unchecked C code to \checkedc code,
-because unchecked types can stay in \ucregion, and \cregion code can access tainted type entities that are allocated 
-in \ucregion.
-% 
-Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks~\cite{rajani2016access, machiry2017boomerang},~\eg by using a valid~\cregion address in a tainted pointer.
-Our compiler avoids these attacks by ensuring using dynamic checks that tainted pointers validly point to~\ucregion address space.
-Such checks are statically generated by our compiler.
-
-In summary, we make the following three main contributions.
-
-\myparagraph{\systemname Type System, Formalism and Compiler}
+\myparagraph{\systemname Type System and Formalism}
 We present a type system that integrates tainted types with~\checkedc and provides additional guarentees---the~\emph{non-crashing} and \emph{non-exposure} guarantees,~\ie a well-typed~\systemname program can never crash due to spatial safety violations,
@@
 
-\myparagraph{Type-Directed Program Partitioning}
-We present a type-directed program partition technique to separate $\cmode$ and $\umode$ code regions and ensure the above guarantees.
-%Our type system restricts checked pointers usage only in~\cregion and ensures that~\ucregion can only access tainted pointers.
-Our modular design enables us to use existing sandbox techniques to enforce memory and execution isolation, with the implementation of tainted pointers in the \systemname compiler.
+We designed and implemented a flexible program partition technique using tainted types.
+Our type system guarantees that the tainted pointers cannot affect untainted pointers and partition the given code into~\ucregion (containing only tainted entities) and~\cregion (CITE SECTION).
+We provide formal proof of the~\emph{clean separation theorem} that guarantees that code in ~\cregion will be separated from~\ucregion according to the target isolation mechanism (CITE SECTION).
+We also integerated~\systemname{} with~\checkedc that provides additional~\emph{non-crashing} and \emph{non-exposure} guarantees (CITE SECTION).
 
-\myparagraph{Supporting callbacks to~\cregion with no Checked Pointer Exposure}
-Although we disallow access to~\cregion from~\ucregion directly, there can be cases where such access is needed.
-Specifically, when~\cregion wants to provide access to certain shared checked data to~\ucregion.
-To enable this, we support callback functions in~\cregion that can be invoked from~\ucregion through function pointers.
-%However, knowing the address of~\cregion functions in~\ucregion violates the program partition principle of \systemname.
-However, knowing the address of~\cregion functions in~\ucregion violates the non-exposure guarantee and leads to other attacks~\cite{hauser2019sleak}.
+\myparagraph{Sound and Better Compilation} Our second
+contribution is a formalization of bounds-check insertion for array
+accesses (Section~\ref{sec:compilation}). Our operational semantics
+annotates each pointer with metadata that describes its bounds, and
+the assignment and dereference rules have premises to confirm the
+access is in bounds. An obvious compilation scheme (taken by
+Cyclone~\cite{Jim2002,GrossmanMJHWC02}, CCured~\cite{Necula2005}, and
+earlier works) would be to translate annotated pointers to multi-word
+objects: one word for the pointer, and 1-2 words to describe its lower
+and upper bounds. Inserted checks reference these bounds. While
+convenient, such ``fat'' pointers are expensive, and break backward
+binary compatibility with legacy pointers.
 
-We handle this by using indirection.
-Specifically, instead of directly accessing the~\cregion callbacks, the ~\ucregion accesses them using a tainted-typed protected trampoline function, which directs the execution to the appropriate callback function.
-In addition, the trampoline function itself is referenced using an opaque index rather than its virtual address, implemented through
-existing sandboxing techniques.
-% 
-%In our \systemname formalism, we formally verified that our formalism is
-%\emph{non-exposure}, \ie, no \cregion pointer addresses will be leaked in
-%\ucregion code.
+  To show that pointer annotations can be safely
+  erased, and thus fat pointers are not needed, we formalize a
+  translation of \lang to \elang, which is an 
+  untyped version of \lang that drops metadata annotations, and
+  lacks bounds/null checks in the semantics rules. Instead,
+  the compilation process inserts null/bounds checks explicitly, leveraging
+  compile-time type information. While we do not definitively prove
+  it, we provide strong evidence that compilation is correct. We use PLT
+  Redex~\cite{pltredex} to mechanize (a generalization of) \lang, 
+  \elang, and compilation between the two, and we use randomized testing 
+to validate that the compiled program \emph{simulates} the
+original. In addition to demonstrating the technical point that metadata
+annotations in the \lang formalism do not necessitate fat pointers,
+compilation also sheds light on the actual Checked C compilation
+process. 
+% \review{ Beginning of page 2: "we show", then "give confidence". I am left confused as
+%   to whether the CoreChkC --> CoreC compilation is shown correct (theorem with
+%   proof), or carefully debugged and tested in plt-redex (which is fine too, I
+%   just find the choice of terminology confusing)}
+% \liyi{typo, should be "validate through randomized testing"}
 
-\ignore{
+As far as we are aware, \lang is the first formalism to cleanly
+separate bounds-checking compilation from the core semantics; prior
+work~\cite{Feng2006,Condit2007} merged the two, conflating
+\emph{meaning} with \emph{mechanism}. In carrying out the
+formalization, we discovered that our compilation approach for
+null-terminated array pointers is more expressive than that proposed
+in the Checked C specification~\cite{checkedc}
+(Section~\ref{sec:disc}); we would not have discovered this
+improvement had we not separated checks from semantics.
 
--- NEED TO FINISH THIS--
-Our third contribution is an added-up feature
-to support checked (function) pointer callbacks in unchecked code regions.
-When designing a multi-threaded system, users might want 
-to provide a third party interface that allows third party developers to create new program features, while keeping these programs in unchecked code regions.
-Moreover, they do want to provide them a (function) pointer pointing to checked data fields.
+With the understanding of the true guarantees from program partitioning, we find a better compilation strategy.
+Instead of sanitizing the entire \umode region, as many previous mechanisms did \cite{rlbox-paper},
+we can only sanitize the \tmode pointers by providing extra dynamic checks in the compiler.
+Compared to a typical $100\%$ - $200\%$ overhead in these works \cite{rlbox-paper} as our faithful RLBox representation \systemnamea,
+we show that the new compilation implementation, named \systemnameh, only having an average $5\%$ overhead compared to the original \checkedc compiler.
 
-However, accessing a checked pointer in an unchecked region violates the program partition principle of \systemname.
-To resolve the conflict, we develop two mechanisms in \systemname 
-and maintain a stronger \textit{non-exposure} guarantee on of the non-crashing guarantee; 
-that is, no checked pointer addresses can be observed in an unchecked code region.
-The first mechanism allows nested checked and unchecked code regions.
-Users can context switch between checked and unchecked code regions 
-by nested using the keywords $\echeckedtext$ $\euncheckedtext$.
-The type system ensures that no checked pointers can be accessed across the context switching.
-The second one is that a call to a checked pointer in unchecked code regions 
-must be surrounded by a \textit{tainted shell}; 
-i.e., a tainted function pointer that points to a checked region possibly holding checked pointers.
-In this case, no checked pointer address will be observed in the unchecked code regions.
-% 
-\mzu{In \systemname, for every checked function, 
-we automatically compile a tainted version by surrounding the function without a
-tainted shell.}
-\mz{What does this mean?}
+\myparagraph{Extending \checkedc}
+We formalize \systemname in \checkedc \cite{li22checkedc} because it has a formalism and compilation.
+More importantly, \checkedc has checked and unchecked regions coinciding with the \cmode and \umode regions in program partitioning,
+and it provides spatial safety for \cmode region and guarantees that any memory violation is from \umode region,
+which is similar to the concept that \cmode region contains only "good" pointers.
+Based on the \checkedc formalism, we provide extra \tmode mode pointers to communicate the \cmode and \umode regions in \checkedc, while keeping the original \cmode and \umode mode pointers to be only used in \cmode and \umode regions, respectively.
+By upgrading the \checkedc formalism to \systemname, we not only build the foundation for program partitioning, but also extends \checkedc
+to support safeties for third party libraries.
+Previously, users are required to rewrite C code to \checkedc completely to guarantee memory safeties.
+With \systemname, users can view C library functions as \umode region, and utilize \systemname to validate the \tmode mode pointers returning from these \umode region, without painfully rewriting the C library functions to \checkedc.
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{relationship.pdf}
+\caption{
+    \lang models' relationship to \checkedc
 }
-%
+  \label{fig:model-relation}
+\end{figure}
 
-\iffalse
-\myparagraph{Formalizing the Type System, Semantics and Compiler}
-%
-We developed a core formalism named~\lang, which extends
-\citet{li22checkedc}~\checkedc formalism with the non-crashing guarantee and other new features below.
-We formally prove the~\emph{non-crashing theorem},~\ie 
-a well-typed~\lang program can never crash due to spatial safety violations.
-We utilize the model-based randomized testing (CITE) to certify the simulation relation between the~\lang semantics and the compiler.
-Specifically, we use a conversion tool that converts expressions from~\lang into actual~\checkedc code that can be compiled by the~\checkedc compiler. We create a random program generator 
-based on the typing rules of~\lang and ensure that~\lang and~\checkedc compiler are consistent after conversion, both statically and dynamically.  
-To the best of our knowledge,~\lang is the first C(-like) language and compiler formalism with the program partitioning mechanism.
-%To the best of our knowledge, \systemname is the first work of formalizing C function pointers with security guarantee.
-\fi
+\myparagraph{Summary Visualization} The
+  relationship among our contributions is visualized in
+  Fig.~\ref{fig:model-relation}. With the Coq model of \lang we prove
+  soundness (and with it, \emph{blame}) of the \checkedc type system
+  and semantics. With the Redex model, we use randomized testing to
+  validate both type soundness and compilation correctness, where the
+  latter shows how compilation need not output fat pointers despite
+  the use of pointer annotations in the \lang model. The Redex \lang
+  model is also the basis of randomized testing of the correctness of
+  the \checkedc compiler implementation, both its type checker and the
+  semantics of its emitted code, at least for the subset of the
+  language in the Redex model. The Redex model's syntax
+  is slightly richer than the Coq version: conditional guards and
+  function arguments may be arbitrary expressions, where the Coq
+  version limits them to constants and variables, making handling of
+  dependent types a bit simpler. We find a useful synergy between the
+  Coq and Redex models for carrying out a language development. The
+  richer, executable Redex model is useful for quickly modeling and
+  testing new features, both formally and against a real
+  implementation. Once solidified, new features can be added to the
+  Coq model (perhaps somewhat simplified) for final proofs of
+  correctness.
 
-We evaluated~\systemname by partioning seven large real-world programs to demonstrate its effectiveness.
-Our evaluation shows that~\systemname provides a flexible, low-overhead program partitioning mechanism and guarantees spatial memory safety.
-
-\ignore{
-\aravind{Fix the following text after all the sections of the papers are finalized.}
-\liyi{the following might be removed if space is needed, since it is usually not quite useful.}
-We begin with a review of \checkedc{} \mzr{and introduction of} new features in
-\systemname (Section~\ref{sec:overview}), present our main contributions
+We begin with a review of \checkedc (Section~\ref{sec:overview}),
+present our main contributions
 (Sections~\ref{sec:formal}--\ref{sec:evaluation}), and conclude with a
-discussion of related and future work (Sections~\ref{sec:related},
-\ref{sec:conclude}). All code and proof artifacts (both for Coq and Redex) can
-be found at \url{https://github.com/plum-umd/checkedc}.
-}
-
-\ignore{
-\noindent
-\textbf{Converting C to Checked C.} The safety guarantees of Checked C come with certain restrictions.
-For instance, as shown below, Checked C programs cannot use address-taken variables in a bounds expression as the bounds relations may not hold because of possible modifications through pointers.
-\begin{minted}[xleftmargin=30pt, mathescape, escapeinside=||, fontsize=\footnotesize]{c}
-...
-array_ptr<int> p : count (n) = NULL;
-|\textcolor{red}{\faTimes}|..,&n,.
-\end{minted}
-Consequently, converting existing C programs to Checked C might require refactoring, e.g., modifying the above program to not use~\inlinecode{&n} expression, which might require considerable effort~\cite{duanrefactoring} depending on the program's complexity.
-Recently, Machiry et al. developed~\threec~\cite{machiry2022c} that tries to automatically convert a program to Checked C by adding appropriate pointer annotations.
-However, as described in \threec, complete automated conversion is infeasible and requires the developer to convert some code regions manually.
-Although, the backward compatibility of Checked C helps a partially annotated program to enjoy spatial memory safety on those regions using only Checked pointers (i.e., checked or safe regions).
-
-\noindent
-\textbf{No safety against unchecked code.} But, the unconverted code regions (or unsafe regions) can affect pointers in safe regions and violate certain assumptions leading to vulnerabilities, as demonstrated by cross-language attacks~\cite{mergendahlcross}.
-Although the blameless proof exists~\cite{ruef2019achieving}, it does not state that spatial safety violations cannot happen in Checked regions but rather states that Checked regions~\emph{cannot be blamed for any spatial safety violations}.
-Consider the following example:
-\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
-// Checked code
-int func(array_ptr<char> p : count(5)) {
-|\textcolor{red}{\faChainBroken}|..p[4]..
-}
-// unchecked code
-...
-str = "he";
-...
-|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5); 
-...
-char ptr[16];
-...
-len <- derived from user input
-...
-|\textcolor{red}{\faBug}| memcpy(ptr, buff, len); // buffer overflow
-\end{minted}
-Here, the checked function~\inlinecode{func} expected a pointer to a buffer of five elements, but unchecked code violated it and invoked the function with a buffer of 2 elements.
-This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the Checked region, but of course, the blame or the root cause is in the unchecked region (\textcolor{red}{\faBug}).
-Furthermore, since checked and unchecked regions execute in the same address space, spatial memory corruptions in unchecked regions (Line 15) can take down the complete program despite having checked regions.
-We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked regions does not violate the safety guarantees in checked regions}.
-
-This mechanism is called program partitioning~\cite{rul2009towards}, and there has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
-Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
-E.g., Given sensitive data in a program, the goal is to partition functions into two parts or partitions based on whether a function can access the sensitive data.
-The performance overhead of these approaches is dominated by marshaling costs and depends on the usage of sensitive data.
-The overhead of state-of-the-art approaches~\cite{lind2017glamdring, liu2017ptrsplit} is prohibitive and varies from 37\%-163\%.
-But in our case, we want a low-overhead code-centric partitioning, where the unchecked code (or functions) should be isolated (or partitioned) from checked code.
-We also want the technique to co-exist and be compatible with Checked C guarantees such that the partition containing checked code should still enjoy its spatial safety.
-
-In this work, we propose a type-directed code-centric program partitioning approach.
-Specifically, our system,~\systemname, extends Checked c using~\textbf{tainted} (\taintt) types, which can be used to mark functions and pointers that need to be isolated from the original program.
-The tainted types long with Checked C allow annotating pointer along two dimensions, i.e., (i) taintedness: a pointer can be either tainted or not (untainted), and (ii) checkedness: it can be either checked or not.
-The checked types follow the standard Checked C typing rules. However, our type-system disallows all interactions between taint and untainted types.
-The developer starts by marking desired (unchecked or risky) functions and pointers used in these functions as tainted.
-Second,~\systemname partitions the given program into two partitions (\ucregion and~\cregion)  of different privileges:
-\begin{itemize}
-\item~\ucregion (low privilege tainted region): This partition contains tainted types (i.e., functions and pointers) and can only access tainted pointers.
-\item~\cregion (high privilege untainted region): This partition contains the remaining (untainted) code and data and has complete access to~\cregion{}.
-The functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
-\end{itemize}
-Finally, during program execution, the~\ucregion partition will be executed in an existing sandboxed environment (e.g., WASM sandbox), and our compiler will add the necessary instrumentation to facilitate the communication between code in~\cregion{} and~\ucregion{}.
-
-The combination of tainted types and privileged partitions enables us to enforce isolation and provide memory safety without marshaling costs.
-As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
-We avoid marshaling by allocating all tainted pointers (i.e., tainted buffers) in~\ucregion and thus can be accessed in both partitions.
-Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
-Our compiler avoids these attacks by ensuring, either statically or through dynamic checks, that tainted pointers can only point to~\ucregion address space.
-
-%In this work, we propose~\systemname, an isolation mechanism that enables unchecked regions or, in general, any arbitrary set of functions to be executed in an isolated address space (\ucregion).
-%So that checked regions can still enjoy true spatial memory safety even in the presence of unchecked or unsafe regions.
-%\systemname extends Checked c using~\textbf{tainted} (\taintt) types, representing the code and data that belong to the~\ucregion.
-%
-%\noindent
-%\textbf{Automated Partitioning.} A pointer (checked or unchecked) can be marked as tainted; similarly, a function can be marked as tainted.
-%Along with enforcing spatial safety constraints of checked tainted pointers, the~\systemname compiler will bundle all the tainted types (functions and pointers) so that they remain in an isolated sandbox (\ucregion), and non-tainted types will remain in the process address space (\cregion).
-%
-%\noindent
-%\textbf{Memory Access Restrictions.} The functions in the~\ucregion can only access tainted pointers, i.e., memory within the corresponding sandbox region.
-%However, the functions in~\cregion have no-such restrictions and can access both c- and uc-region's memory.
-%Similarly, the functions in~\cregion can invoke any function in~\ucregion but not the other way around, except for call-back functions, which we will discuss later.
-%
-%\noindent
-%\textbf{Marshaling free communication.} As functions in the~\ucregion can only access tainted types,~\cregion functions should use tainted types to pass pointer arguments to~\ucregion functions. 
-%We avoid marshaling as all tainted buffers (i.e., regions pointed by tainted pointers) are allocated in~\ucregion and thus can be accessed in both c- and uc-region.
-%
-%\noindent
-%\textbf{Additional checks.} Although memory isolation prevents direct violations,~\ucregion code can still affect~\cregion through tainted pointers by confused deputy attacks.
-%However, our compiler avoids these attacks by ensuring, either statically or through dynamic checks, that tainted pointers in~\cregion can only point to~\ucregion address space, and our type system enforces that tainted pointers cannot be assigned to non-tainted pointers.
-%
-%%\iffalse
-%Recently, Narayan et al. proposed RLBox, which enables sandboxing libraries by making data- and control-flow at the sandbox interface explicit, using types.
-%Specifically, RLBox defines a tainted type (\code{tainted<...>}), which is a C++ template class and can be used to have any type tainted (e.g.,~\code{tainted<int>},~\code{tainted<float*>}, etc).
-%The tainted types represent values or pointers to buffers that can be accessed from a sandbox.
-%The program or unsandboxed code can use only tainted type variables to communicate (i.e., parameters and return values) with sandboxed functions.
-%Although RLBox provides a practical technique for sandboxing libraries, it has the following drawbacks:
-%\begin{itemize}
-%\item\textbf{Marshalling Cost.} Any non-scalar value (e.g., a pointer to a struct) that need to be passed to a sandboxed function should be first copied into a tainted buffer (\emph{marshalling}).
-Then the corresponding tainted value (i.e., tainted pointer) should be passed to the sandboxed function.
-%This can lead to considerable overhead for large objects.
-%\item\textbf{Developer Written Sanitization.} The unsandboxed code can access tainted type variables only through certain special (i.e., sanitization) routines.
-%This forces developers to sanitize every value that comes from sandboxed library functions.
-%However, the correctness of the sanitization is entrusted to the developer.
-%Consequently, any mistake in sanitization methods can still lead to vulnerabilities.
-%\item\textbf{Library Level Sandboxing.} The unit of sandboxing in RLBox is the whole library. However, we want to sandbox a set of functions in a program.
-%\end{itemize}
-%\fi
-
-% How to highlighgt code in minted
-% highlightlines={1-3}, highlightcolor=red
-%\ignore{
-\begin{listing}[t!]
-  \begin{tabular}{c c}
-    \begin{minipage}[b]{.22\textwidth}
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/orig1.c}
-    \end{minipage} &
-    \begin{minipage}[b]{.22\textwidth}
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/orig2.c}
-    \end{minipage} \\
-   (a) Original C code & (b) After initial conversion.\\
-  \end{tabular}
-\caption{(Contrived) Example demonstrating various phases of.}
-\label{lst:comb}
-\end{listing}
-
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/firstrun.c}
-\caption{(Contrived) Example demonstrating various phases of.}
-\label{lst:initialconv}
-\end{listing}
-
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/humanannotations.c}
-\caption{(Contrived) Example demonstrating various phases of.}
-\label{lst:humantaint}
-\end{listing}
-
-\begin{listing}[t!]
-\inputminted[mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/humanadjustments.c}
-\caption{(Contrived) Example demonstrating various phases of.}
-\label{lst:humanadjust}
-\end{listing}
-%}
-}
-
-%\input{figures/taintedsplit}
-
-%\begin{listing*}[t!]
-%\inputminted[linenos, mathescape, escapeinside=||, fontsize=\tiny{}]{c}{examples/originalprogram.c}
-%\caption{Simple server example.}
-%\label{lst:comb}
-%\end{listing*}
+discussion of 
+related and future work (Sections~\ref{sec:related},
+\ref{sec:conclude}). All code and proof artifacts (both for Coq and
+Redex) can be found at \url{https://github.com/plum-umd/checkedc}.

### 2023-04-22 sandp23/formal.tex
update.

@@
 \lang, making precise its syntax, semantics, and type system,
-as well as \systemname’s meta-theories, including the type preservation, non-exposure, and clean separation theorems.
+with the \systemname’s meta-theories shown in \Cref{}, including the type preservation, non-exposure, and clean separation theorems.
 One thing worth noting is that \lang and its meta-theories are built and verified, despite the underling sandbox mechanisms,

### 2023-04-22 csf2024/examples/MirrorExample.c

@@ -1,5 +1,0 @@
-/* JSON Value API */
-_Mirror JSON_Value_Type json_value_get_type  
-      (_TPtr<const TJSON_Value> value) {
-    return value ? value->type : JSONError;
-}

### 2023-04-22 csf2024/examples/callback_mechanism_sample_after_checkmate.c

@@ -1,52 +1,0 @@
-#include <string_tainted.h>
-
-_Callback _TPtr<char> StringAuth(_TPtr<const char> input , size_t len) {
-   ...
-   //complex Function Body
-   ...
-}
-
-/*
- * CheckMate Inserted Callback Registration
- */
-
-void registerCallback_StringAuth(void){
-    //callback function signature {ret <- int, arg_1 <- int, arg_2 <- long}
-    int ret_param_types[] = {0, 0, 1}; 
-    // 2 <- arg count, 1 <- ret count 
-    __StringAuth__C_ = _SBXREG_((void*)_T_StringAuth,2,1, ret_param_types);
-}
-
-/*
- * CheckMate Inserted Trampoline for Directing Callbacks from WASM sandbox
- */
-
-unsigned int _T_StringAuth(unsigned sandbox,
-                                 unsigned int arg_1,
-                                 unsigned long int arg_2) {
-    return StringAuth(arg_1, (size_t)arg_2);
-}
-
-_Tainted _TPtr<char> StringProc(_TPtr<_TPtr<const char>> string,
-_TPtr<_TPtr<char>(_TPtr<const char> input, size_t len)>StringAuth) {
-	
-   // StringAuth as a function ptr arg is only used as a 
-   //_Callback signature to CheckMate and for type-checker
-   return w2c_StringProc(_SBX_(), (unsigned int)string, __StringAuth__C_);
-}
-
-_Mirror int isStringMatching(char* str1 : itype (_Ptr<char>),
-                char* str2 : itype(_Ptr<char>))
-{
-  ...
-  //complex Function body
-  /*
-   * "_Mirror" restricts access to global data
-   *
-   * Since "_Mirror" functions are locally replicated in both
-   * checked and unchecked regions, they have no restriction on
-   * type signature and callees.
-   */
-  ...
-}
-

### 2023-04-22 csf2024/examples/callback_mechanism_sample_after_checkmate_tainted.c

@@ -1,20 +1,0 @@
-// CheckMate inserts headers
-
-char * StringProc(const char * string,
-char * (*StringAuth)(const char * input,size_t len)) {
-  ...
-  //Complex code definition converted to generic-C
-  ...
-  strncpy(local_str, user_input[i], len); //CheckMate modified 
-  ...
-  return StringAuth(one_past_start, string_len);
-}
-
-int isStringMatching(char* str1 ,
-                char* str2)
-{
-  ...
-  //complex Function body
-  ...
-}
-

### 2023-04-22 csf2024/examples/callback_mechanism_sample_before_checkmate.c

@@ -1,52 +1,0 @@
-#include <string_tainted.h> //tainted wrapper of STDLIB string.h
-...
-
-_Callback _TPtr<char> StringAuth(_TPtr<const char> input , size_t len) {
-  ...
-  //complex Function Body
-  /*
-  "_Callback" allows access to global checked data
-  
-  "_Callback" restricts function signature to
-  only contain pointers of Tainted type
-  Since Callbacks execute in checked region, they have no 
-  restrictions on callees.
-  */
-  ...
-}
-
-_Tainted _TPtr<char> StringProc(_TPtr<_TPtr<const char>> user_input,
-_TPtr<_TPtr<char>(_TPtr<const char> input, size_t len)>StringAuth) {
-  ...
-  //complex Function Body	
-  
-   /*
-   "_Tainted" restricts access to global data 
-   
-   "_Tainted" restricts function signature to 
-   only contain pointers of Tainted type 
-  
-   Since "_Tainted" functions are destined to execute in unchecked
-   region, callees of this function must be "_Mirror", "_Callback" or 
-   "_Tainted" qualified. 
-  */
-  ...
-  t_strncpy(local_str, user_input[i], len); // tainted wrapper for strncpy
-  ...
-  return StringAuth(one_past_start, string_len);
-}
-
-_Mirror int isStringMatching(char* str1 : itype (_Ptr<char>), 
-		char* str2 : itype(_Ptr<char>))
-{
-  ...
-  //complex Function body
-  /*
-   * "_Mirror" restricts access to global data
-   * 
-   * Since "_Mirror" functions are locally replicated in both 
-   * checked and unchecked regions, they have no restriction on 
-   * type signature and callees.
-   */
-  ...
-}

### 2023-04-22 csf2024/examples/conv1.c

@@ -1,36 +1,0 @@
-|\entrypoint| // Entry point
-void server_loop(int sock_fd) {
- unsigned b_z;
- |\textcolor{checkcolor}{\_Array\_ptr}|<struct queue> p : 
-              count(b_z) = NULL;
- ...
- while(1) {
-  ...
-  p = malloc(sizeof(struct queue) * b_z);
-  ...
-  if (handle_request(sock_fd)) {
-    ...  
-  }
- }
-}
-
-int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_Tainted}|;
- int rc = -1;
- size_t r_len;
- r_len = read_msg(sock_fd, buff, 
-                  MAX_MSG_SIZE);
- if (r_len > 0 && 
-     r_len < MAX_MSG_SIZE) {
-  switch(buff[0]) {
-   case REQ1: 
-    rc = process_req1(buff, r_len); 
-    break;
-   case REQ2: 
-    rc = process_req2(buff, r_len); 
-    break;
-   ...
-  }
- }
- return rc;
-}

### 2023-04-22 csf2024/examples/conv2.c

@@ -1,39 +1,0 @@
-size_t read_msg(int sock_fd, char *msg, 
-                size_t sz) _Tainted {
- size_t nr;
- nr = read(sock_fd, (void*)msg, sz);
- ...
- |\complexcode| // complex code
- return nr;
-}
-
-int process_req1(_T_Array_ptr<char> msg :
-                 count(m_l), size_t m_l) 
-                 _Tainted {
-
- int rc = -1, i;
- if (m_l > MIN_SIZE) {
-   sscanf(msg, "%d", &i);
-   msg[i] = ...
- }
- // complex code
- return rc;
-}
-
-int process_req2(|\textcolor{taintcolor}{\_T\_Array\_ptr}|<char> msg : 
-                 count(m_l), 
-                 size_t m_l) {
- size_t i = 0, j = 0;
- int rc = -1;
- char anum |\textcolor{checkcolor}{\_Checked}|[MAX_MSG_SIZE] = {0};
- if (msg) {
-  while(i < m_l && j < MAX_MSG_SIZE-1) {
-   ...
-   if (isalnum(msg[i]))
-    anum[j++] = msg[i]; 
-   i++;
-  }
-   rc = process_data(anum);
- }
- return rc;
-}

### 2023-04-22 csf2024/examples/firstrun.c

@@ -1,13 +1,0 @@
-|\entrypoint| // Entry point
-void server_loop(int sock_fd) {
- ...
- |\textcolor{checkcolor}{\_Array\_ptr}|<struct queue> p : count(b_z) = NULL;
- ...
-}
-
-int process_req2(_Array_ptr<char> msg: count(m_l), 
-                 size_t m_l) {
- ...
- char anum |\textcolor{checkcolor}{\_Checked}|[MAX_MSG_SIZE] = {0};
- ...
-}

### 2023-04-22 csf2024/examples/humanadjustments.c

@@ -1,9 +1,0 @@
-int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE] |\textcolor{taintcolor}{\_Tainted} \usermods|;
- ...
-}
-
-int process_req2(|\usermods| |\textcolor{taintcolor}{\_T\_Array\_ptr}|<char> msg: count(m_l), 
-                 size_t m_l) {
- ...
-}

### 2023-04-22 csf2024/examples/humanannotations.c

@@ -1,8 +1,0 @@
-size_t read_msg(int sock_fd, char *msg, 
-                size_t sz) |\textcolor{taintcolor}{\_Tainted}| |\useradded| {
- ...
-}
-
-int process_req1(char *msg, size_t m_l) |\textcolor{taintcolor}{\_Tainted}| |\useradded| {
- ...
-}

### 2023-04-22 csf2024/examples/orig1.c

@@ -1,35 +1,0 @@
-|\entrypoint| // Entry point
-void server_loop(int sock_fd) {
- unsigned b_z;
- struct queue *p;
- ...
- while(1) {
-  ...
-  p = malloc(sizeof(struct queue) * b_z);
-  ...
-  if (handle_request(sock_fd)) {
-    ...  
-  }
- }
-}
-
-int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE];
- int rc = -1;
- size_t r_len;
- r_len = read_msg(sock_fd, buff, 
-                  MAX_MSG_SIZE);
- if (r_len > 0 && 
-     r_len < MAX_MSG_SIZE) {
-  switch(buff[0]) {
-   case REQ1: 
-    rc = process_req1(buff, r_len); 
-    break;
-   case REQ2: 
-    rc = process_req2(buff, r_len); 
-    break;
-   ...
-  }
- }
- return rc;
-}

### 2023-04-22 csf2024/examples/orig2.c

@@ -1,34 +1,0 @@
-size_t read_msg(int sock_fd, char *msg, 
-                size_t sz) {
- size_t nr;
- nr = read(sock_fd, |\rootcause|(void*)msg, sz);
- ...
- |\complexcode| // complex code
- return nr;
-}
-
-int process_req1(char *msg, size_t m_l) {
- int rc = -1, i;
- if (m_l > MIN_SIZE) {
-   sscanf(msg, "%d", &i);
-   |\realbug| msg[i] = ...
- }
- |\complexcode| // complex code
- return rc;
-}
-
-int process_req2(char *msg, size_t m_l) {
- size_t i = 0, j = 0;
- int rc = -1;
- char anum[MAX_MSG_SIZE] = {0};
- if (msg) {
-  while(i < m_l && j < MAX_MSG_SIZE-1) {
-   ...
-   if (isalnum(msg[i]))
-    anum[j++] = msg[i]; 
-   i++;
-  }
-   rc = process_data(anum);
- }
- return rc;
-}

### 2023-04-22 csf2024/examples/orig3.c

@@ -1,15 +1,0 @@
-int process_req2(char *msg, size_t m_l) {
- size_t i = 0, j = 0;
- int rc = -1;
- char anum[MAX_MSG_SIZE] = {0};
- if (msg) {
-  while(i < m_l && j < MAX_MSG_SIZE-1) {
-   ...
-   if (isalnum(msg[i]))
-    anum[j++] = msg[i]; 
-   i++;
-  }
-   rc = process_data(anum);
- }
- return rc;
-}

### 2023-04-22 csf2024/examples/originalprogram.c

@@ -1,68 +1,0 @@
-size_t read_msg(int sock_fd, char *msg, size_t sz) {
- size_t nr;
- nr = read(sock_fd, (void*)msg, sz);
- ...
- // complex code
- return nr;
-}
-
-int process_req1(char *msg, size_t m_l) {
- int rc = -1;
- //complex code
- if (m_l > MIN_SIZE) {
-   //complex code
-   sscanf(msg, "%d", &i);
-   msg[i] = ...
- }
- //complex code
- return rc;
-}
-
-int process_req2(char *msg, size_t m_l) {
- size_t i = 0, j = 0;
- int rc = -1;
- char anum[MAX_MSG_SIZE] = {0};
- if (msg) {
-  while(i < m_l && j < MAX_MSG_SIZE-1) {
-   ...
-   if (isalnum(msg[i]))
-    anum[j++] = msg[i]; 
-   i++;
-  }
-   rc = process_data(anum);
- }
- return rc;
-}
-
-int handle_request(int sock_fd) {
- char buff[MAX_MSG_SIZE];
- int rc = -1;
- size_t r_len;
- r_len = read_msg(sock_fd, buff, MAX_MSG_SIZE);
- if (r_len > 0) {
-  switch(buff[0]) {
-   case REQ1: 
-    rc = process_req1(buff, r_len); 
-    break;
-   case REQ2: 
-    rc = process_req2(buff, r_len); 
-    break;
-   ...
-  }
- }
- return rc;
-}
-
-void server_loop(int sock_fd) {
- unsigned b_z;
- struct queue *q;
- ...
- while(1) {
-  ...
-  p = malloc(sizeof(struct queue) * b_z);
-  ...
-  if (handle_request(sock_fd)) {
-    ...  
-  }
- }
-}

### 2023-12-01 sandp23/review-answer.md
update.

# Overview

We have uploaded the full version of the paper, which includes an
appendix. The final version of the paper will not have the appendix
but instead will cite a supplemental report.

For the reviewers' convenience we have included a diff between the
original and updated paper version at the end of the PDF. We hope this
will make it easier to see the changes made to the main body of the
paper.

# Major questions

## Formalization status of Theorem 4

Theorem 4, the simulation property, is formalized in PLT Redex, but not
proved; rather, we use randomized testing to give confidence that it
is correct. We don't have a pen and paper proof for it (and did
not claim that we did, in the text).

## Intended guarantee of Theorem 4

Simulation is intended to show that the use of fat pointers in
CoreChkC is not more expressive than a language that does not rely on
fat pointers, as is the expectation for the Checked C language.
The compilation procedure sheds light on how Checked C can be
compiled so as to not need fat pointers. The spatial safety property
is guaranteed by Theorem 3, the blame theorem, and is independent of
what simulation is trying to establish.

## Undefinedness in code examples in Section IV

The target language of compilation, CoreC, is an untyped version of
CoreChkC, i.e., without metadata annotations. We did not intend to
design CoreC as a model of C or an intermediate language such as
LLVM IR. Rather, CoreC can be seen as a device for establishing the fact
that the more expressive semantics of CoreChkC (w.r.t the CheckedC
specification) does not require the use of fat pointers.  The
`deref_array` and `strlen` examples in Section IV (Figs 10-11) are
presented in C syntax to make them easier to understand. As such,
performing pointer arithmetic between a address-0 (null) pointer and a
non-zero integer is defined in CoreC even though it would not be in
actual C, as Reviewer C pointed out. To avoid confusion, we have
dropped type annotations in the examples and made them closer to
actual CoreC, including the use of relative bounds rather than
absolute ones.

## Dereferencing of (un)checked pointers in (un)checked regions

The ordering `u<c` is not a typo, it was the text that was wrong. In
particular, the condition in the rules enforces that unchecked
pointers cannot be used in checked regions; the reverse is
acceptable.

Note that the blame theorem states that whenever the program gets
stuck, we can blame unchecked code. This does not necessarily mean we
always step into a stuck state from the unchecked region. Allowing
access to checked pointers in unchecked regions can allow unchecked
code to corrupt the addresses that checked pointers are assigned
to. The program can get stuck when we later access a corrupted checked pointer
in the checked region, but blame still holds, because the state would
not have been corrupted in the first place if there was no unchecked 
code to corrupt checked pointers. The theorem statements have been
tightened to make this apparent (and make them match the Coq model).

## Discrepancies between the paper's calculus and the Coq formalization

We have added to the paper's supplemental report a dedicated section
that explains the differences between the paper's calculus and the Coq
formalization. For example, Coq syntactically distinguishes the `if`
form for potential ntarray widening and the plain `if` form for
everything else. We have also added a fair bit of extra detail to the
metatheory in Section III.

## Details on model-driven random testing

Our model driven testing approach follows the process recommended by
Pałka et al. [19]. We have added a couple of paragraphs to add more
details, and to clarify that this process is manual, not automatic.

# Other changes

We strived to take care of all of the requests made by the reviewers,
which can be see in the diff at the end of the PDF. Some notes about
specified requests below.

## Reviewer A

> clarify relationship between formalization and implementation

We have added a figure to section I that visualizes the relationship
between our contributions. We also added text to the end of 3.1 to
introduce the Coq and Redex models, and to explain some of their
differences and motivation. We have updated the abstract, conclusions,
and elsewhere to be clearer about what has been proved and what has
been tested.

> correct and address mistakes, typos, confusing terminology and code snippets

We have updated all of the examples. The `strcat` example has been
split into two figures, one in section 3 and one in seciton 4, and
renamed `safe_strcat` to make its intended semantics more clear
(thanks for the suggestion). The two examples intend to implement the
same functionaity, but differ according the semantics of bounds
widening; see the discussion at the end of Section IV-B.

> address pressing questions in the paper.

We have addressed the pressing questions as described in the main
response above, We have gone over all of the reviewer's comments and
attempted to address them; see the diff.

## Reviewer B

> Add details in section 5 (see above)

Done (see above).

> Clarify the status of Theorem 4

Done (see above). Per the reviewer's request, we have also added some
discussion about challenges in the proof, compared to Ruef et al. And
we have clarified the handling of null pointers in the formalism, and
some confusing details in the metatheory.

## Reviewer C

> confirmation that blame theorem issue is essentially just a typo.

Yes this was a typo. We fleshed out the metatheory section be more
precise.

> clarify the policy rules wrt accessing (un)checked pointers in
> (un)checked regions

See above.

> remove, or at least point out and justify the discrepancies between the Coq code and the paper, eg concerning Fig2

We added some discussion to the end of section 3.1 about both the Coq
and Redex models, and added a section to the Appendix about this.

### 2023-12-01 sandp23/checkedc-formal/background.tex
update.

--- 2022-06-25 sandp23/old.tex
@@
-\documentclass[conference]{IEEEtran}
-
-\makeatletter
-\def\endthebibliography{\def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}\endlist
-}
-\makeatother
-
-\IEEEpeerreviewmaketitle
-
-\usepackage{filecontents}
-\usepackage[noadjust]{cite}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{ebproof}
-\usepackage{todonotes}
-\usepackage{amsfonts}
-\usepackage{algorithmic}
-\usepackage{textcomp}
-\usepackage{mathpartir}
-\usepackage{color}
-\usepackage{xcolor}
-\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
-
-\usepackage{subcaption} \usepackage{epsfig}
-\usepackage{listings}
-\usepackage[override]{cmtt}
-\usepackage{graphicx}
-\usepackage{balance}
-\usepackage{xspace}
-\usepackage{booktabs}
-\usepackage{url}
-\usepackage{xparse}
-\usepackage{enumitem}
-\usepackage{tikz}
-\usetikzlibrary{chains,fit,shapes,positioning,calc,arrows.meta,shapes.arrows}
-
-\usepackage[numbers]{natbib}
-\bibliographystyle{plainnat}
-
-\lstdefinestyle{customc}{
-  belowcaptionskip=1\baselineskip,
-  breaklines=true,
-  numbers=left,
-  xleftmargin=\parindent,
-  language=C,
-  columns=flexible,      
-  showstringspaces=false,
-  basicstyle=\small\sffamily,
-  otherkeywords={uint},
-  literate={{<-}{{$\leftarrow\,$}}2
-            {->}{{$\rightarrow\,$}}2
-            {<=}{{$\leq\,$}}2},
-  numberstyle=\tiny\rmfamily,
-  keywordstyle=\bfseries\color{green!40!black},
-  commentstyle=\itshape\color{purple!40!black},
-  identifierstyle=\color{blue!80!black},
-}
-\lstset{language=C,style=customc}
-
-\newenvironment{DIFnomarkup}{}{}
-\newcommand{\myparagraph}[1]{\textbf{#1}.\xspace}
-
-\newcommand{\code}[1]{\lstinline|#1|}
-%\newcommand{\ignore}[1]{{}}
-\newcommand{\kwchecked}{\lstinline|\_Checked|}
-\newcommand{\kwunchecked}{\lstinline|\_Unchecked|}
-\newcommand{\kwdynamiccheck}{\lstinline|\_Dynamic\_check|}
-
-\newcommand{\var}[1]{#1}
-\newcommand{\Arrayptr}[1]{\lstinline|_Array_ptr<|{#1}\lstinline|>|}
-\newcommand{\ArrayptrT}{\Arrayptr{$\tau$}}
-\newcommand{\Ptr}[1]{\lstinline|_Ptr<|{#1}\lstinline|>|}
-\newcommand{\PtrName}{\lstinline|_Ptr|}
-\newcommand{\PtrT}{\Ptr{$\tau$}}
-\newcommand{\Ntarrayptr}[1]{\lstinline|_Nt_array_ptr<|{#1}\lstinline|>|}
-\newcommand{\NtarrayptrT}{\Ntarrayptr{$\tau$}}
-\newcommand{\Checkedarr}[2]{{#1}~\lstinline|_Checked[|{#2}\lstinline|]|}
-\newcommand{\CheckedarrT}[1]{\Checkedarr{$\tau$}{#1}}
-
-\newif\ifsubmit\submitfalse
-
-\ifsubmit
-\newcommand{\mwh}[1]{}
-\newcommand{\ashe}[1]{}
-\newcommand{\dtarditi}[1]{}
-\newcommand{\dvh}[1]{}
-\newcommand{\leo}[1]{}
-\newcommand{\liyi}[1]{}
-\newcommand{\yiyun}[1]{}
-\else
-\newcommand{\mwh}[1]{\textcolor{red}{Mike: #1}}
-\newcommand{\dtarditi}[1]{\textcolor{purple}{David: #1}}
-\newcommand{\leo}[1]{\textcolor{green!80!blue}{Leo: #1}}
-\newcommand{\liyi}[1]{\textbf{\textcolor{orange}{[ Liyi: #1 ]}}}
-\newcommand{\yiyun}[1]{\textcolor{cyan}{Yiyun: #1}}
-\newcommand{\dvh}[1]{\textcolor{magenta}{DVH: #1}}
-\fi
-
-\newcommand{\lang}{\textsc{CoreChkC}\xspace}
-\newcommand{\elang}{\textsc{CoreC}\xspace}
-
-\newtheorem{defi}{Definition}
-\newtheorem{thm}{Theorem}
-\newtheorem{prop}{Proposition}
-
-\newcommand{\CoreChkC}{\textsc{CoreChkC}\xspace}
-\newcommand{\checkedc}{\text{Checked C}\xspace}
-
- \newcommand{\rulelab}[1]{{\small \textsc{#1}}}
-\newcommand{\steps}{\ensuremath{\longrightarrow}}
-\newcommand{\tbool}{\texttt{bool}}
-\newcommand{\tsizeof}{\texttt{sizeof}}
-\newcommand{\vbool}{\texttt{Bool}}
-\newcommand{\verror}{\texttt{Error}}
-\newcommand{\vval}{\mathpzc{val}}
-\newcommand{\tfixed}{\texttt{fixedp}}
-\newcommand{\tnat}{\texttt{nat}}
-\newcommand{\tarr}[2]{\texttt{array}~{#1}~{#2}}
-\newcommand{\econst}[2]{({#1}){#2}}
-\newcommand{\eindex}[2]{{#1}\texttt{[}{#2}\texttt{]}}
-\newcommand{\sassign}[4]{{#1} \leftarrow {#3}~{#2}~{#4}}
-\newcommand{\ssassign}[3]{{#1} \xleftarrow{#2} {#3}}
-\newcommand{\sif}[3]{\texttt{if}~{#1}~{#2}~{#3}}
-\newcommand{\sfor}[3]{\texttt{for}~{#1}~{#2}~{#3}}
-\newcommand{\scall}[3]{{#1}\leftarrow {#2}~{#3}}
-\newcommand{\sseq}[2]{{#1}\,\texttt{;}\,{#2}}
-\newcommand{\sinv}[1]{\texttt{inv}~{#1}}
-\newcommand{\inst}[3][ ]{\texttt{#2}^{#1}~{#3}}
-\newcommand{\insttwo}[4][ ]{\texttt{#2}^{#1}~{#3}~{#4}}
-\newcommand{\instthree}[5][ ]{\texttt{#2}^{#1}~{#3}~{#4}~{#5}}
-\newcommand{\iskip}[1]{\inst{ID}{#1}}
-\newcommand{\inot}[1]{\inst{X}{#1}}
-\newcommand{\ictrl}[2]{\insttwo{CU}{#1}{#2}}
-\newcommand{\irz}[3][ ]{\insttwo[#1]{RZ}{#2}{#3}}
-\newcommand{\isr}[3][ ]{\insttwo[#1]{SR}{#2}{#3}}
-\newcommand{\icnot}[2]{\insttwo{CNOT}{#1}{#2}}
-\newcommand{\ilshift}[1]{\inst{Lshift}{#1}}
-\newcommand{\irshift}[1]{\inst{Rshift}{#1}}
-\newcommand{\irev}[1]{\inst{Rev}{#1}}
-\newcommand{\iqft}[2][ ]{\inst[#1]{QFT}{#2}}
-\newcommand{\ihad}[1]{\inst{H}{#1}}
-
-\newcommand{\iseq}[2]{{#1}\,\texttt{;}\,{#2}}
-\newcommand{\inval}[2]{\insttwo{Nval}{#1}{#2}}
-\newcommand{\ihval}[3]{\instthree{Hval}{#1}{#2}{#3}}
-\newcommand{\iqval}[2]{\insttwo{Qval}{#1}{#2}}
-
-\newcommand{\instr}{\iota}
-
-\newcommand{\kw}[1]{\ensuremath{\mathtt{#1}}}
-\newcommand{\subtype}[2]{\ensuremath{\vdash{#1}\prec{#2}}}
-\newcommand{\estrlen}[1]{\ensuremath{\kw{strlen}({#1})}}
-\newcommand{\estrlentext}{\ensuremath{\kw{strlen}}}
-
-\newcommand{\tarray}[3]{\tarrayb{({#1},{#2})}{#3}}
-\newcommand{\tarrayb}[2]{\ensuremath{[{#1}~{#2}]}}
-\newcommand{\tntarray}[3]{\tntarrayb{({#1},{#2})}{#3}}
-\newcommand{\tntarrayb}[2]{\tarrayb{#1}{#2}_{nt}}
-\newcommand{\tallarrayb}[2]{\ensuremath{[{#1}~{#2}]_{\kappa}}}
-\newcommand{\tallarraybc}[2]{\ensuremath{[{#1}~{#2}]_{\textcolor{cyan}{\kappa}}}}
-
-\newcommand{\tallarray}[3]{\tallarrayb{({#1},{#2})}{#3}}
-
-\newcommand{\tptr}[2]{\ensuremath{\mathtt{ptr}^{#2}~{#1}}}
-\newcommand{\tntptr}[4]{\ensuremath{[({#1},{#2})~{#3}]^{#4}_{nt}*}}
-
-\newcommand{\tarrayptr}[4]{{\tptr{\tarray{#1}{#2}{#3}}{#4}}}
-\newcommand{\tntarrayptr}[4]{{\tptr{\tntarray{#1}{#2}{#3}}{#4}}}
-\newcommand{\tallarrayptr}[4]{{\tptr{\tallarray{#1}{#2}{#3}}{#4}}}
-
-\newcommand{\tgarray}[4]{\ensuremath{\mathtt{#1}~{#2}~{#3}~{#4}}}
-\newcommand{\tstruct}[1]{\ensuremath{\kw{struct}~{#1}}}
-
-\newcommand{\evalue}[2]{\ensuremath{{#1}\!:\!{#2}}}
-
-\newcommand{\emalloc}[1]{\ensuremath{\kw{malloc}({#1})}}
-\newcommand{\emalloctext}{\ensuremath{\kw{malloc}}}
-\newcommand{\ecall}[2]{\ensuremath{{#1}({#2})}}
-\newcommand{\ret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
-\newcommand{\rettext}{\ensuremath{\kw{ret}}}
-\newcommand{\ecast}[2]{\ensuremath{\kw{(}{#1}\kw{)}{#2}}}
-\newcommand{\edyncast}[2]{\ensuremath{\langle{#1}\rangle{#2}}}
-\newcommand{\edcast}[2]{\ensuremath{\kw{(D},{#1}\kw{)}{#2}}}
-\newcommand{\elet}[3]{\ensuremath{\kw{let}~#1\, \texttt{=}\, #2~\kw{in}\;{#3}}}
-\newcommand{\elettext}{\ensuremath{\kw{let}}}
-\newcommand{\ebinop}[2]{\ensuremath{#1 \plus #2}}
-\newcommand{\eassign}[2]{\ensuremath{\texttt{*}{#1}\,\texttt{=}\, {#2}}}
-\newcommand{\eassignstack}[2]{\ensuremath{{#1}\,\texttt{=}\, {#2}}}
-\newcommand{\efield}[2]{\ensuremath{{#1}\kw{\rightarrow}{#2}}}
-\newcommand{\estar}[1]{\ensuremath{\texttt{*}{#1}}}
-\newcommand{\eamper}[2]{\ensuremath{\kw{\&}{#1}\kw{\rightarrow}{#2}}}
-\newcommand{\eunchecked}[1]{\ensuremath{\kw{unchecked}\;{#1}}}
-\newcommand{\euncheckedtext}{\ensuremath{\kw{unchecked}}}
-\newcommand{\eif}[3]{\ensuremath{\kw{if\;}(#1)\;{#2}\;\kw{else}\;{#3}}}
-\newcommand{\eifa}[3]{\ensuremath{\kw{IF\;}(#1)\;{#2}\;\kw{ELSE}\;{#3}}}
-\newcommand{\eifatext}{\ensuremath{\kw{IF}}}
-
-\newcommand{\eiftext}{\ensuremath{\kw{if}}}
-\newcommand{\ethentext}{\ensuremath{\kw{then}}}
-\newcommand{\eelse}{\ensuremath{\kw{else}}}
-\newcommand{\ebreak}{\ensuremath{\kw{break}}}
-\newcommand{\ebounds}{\ensuremath{\kw{bounds}}}
-\newcommand{\enull}{\ensuremath{\kw{null}}}
-\newcommand{\hole}{\ensuremath{\Box}}
-\newcommand{\defscope}{\sigma}
-\newcommand{\wf}[1]{wf \, #1}
-\newcommand{\tint}{\ensuremath{\mathtt{int}}}
-\newcommand{\efor}{\ensuremath{\mathtt{for}}}
-\newcommand{\ewhile}{\ensuremath{\mathtt{while}}}
-\newcommand{\etrue}{\ensuremath{\mathtt{true}}}
-\newcommand{\efalse}{\ensuremath{\mathtt{false}}}
-\newcommand{\edplus}{\ensuremath{\texttt{++}}}
-\newcommand{\edminus}{\ensuremath{\texttt{--}}}
-
-\newcommand{\ememcpy}{\ensuremath{\mathtt{memcpy}}}
-\newcommand{\ereturn}{\ensuremath{\mathtt{return}}}
-\newcommand{\emain}{\ensuremath{\mathtt{main}}}
-\newcommand{\estrcat}{\ensuremath{\mathtt{strcat}}}
-\newcommand{\estrcatbad}{\ensuremath{\mathtt{strcat\_b}}}
-\newcommand{\heap}{\ensuremath{\mathpzc{H}}}
-\newcommand{\eret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
-\newcommand{\erettext}{\ensuremath{\kw{ret}}}
-
-\newcommand{\plus}{\mathbin{\texttt{+}}}
-
-\newcommand{\fv}{\mathit{FV}}
-\newcommand{\dom}{\mathit{dom}}
-
-\newcommand{\fm}{\mathit{FM}}
-\newcommand{\size}{\mathit{size}}
-
-\newcommand{\cmode}{\texttt{c}}
-\newcommand{\umode}{\texttt{u}}
-\newcommand{\bvar}{\ensuremath{\beta}}
-\newcommand{\mode}{\textit{mode}}
-
-\newcommand{\cextend}[4]{\ensuremath{#4 = ~ \vdash_{extend} #1, #2, #3}}
-\newcommand{\fresh}{\ensuremath{\kw{fresh}}}
-\newcommand{\echecknull}[3]{\ensuremath{#3 = ~\vdash_{null}#1, #2}}
-\newcommand{\echeckboundsdyn}[5]{\ensuremath{#5 = ~ \vdash_{boundsD}#1, #2, #3, #4}}
-\newcommand{\echeckbounds}[5]{\ensuremath{#5 = ~ \vdash_{boundsR}#1, #2, #3, #4}}
-\newcommand{\echeckboundsw}[5]{\ensuremath{#5 = ~ \vdash_{boundsW}#1, #2, #3, #4}}
-\newcommand{\esizeof}[1]{\ensuremath{\kw{sizeof}(#1)}}
-\newcommand{\ewidenstrlen}[5]{\ensuremath{#5 = ~ \vdash_{widenstr}#1, #2, #3, #4}}
-\newcommand{\ewidenderef}[4]{\ensuremath{#4 = ~ \vdash_{widenderef}#1, #2, #3}}
-\newcommand{\eleq}[2]{\ensuremath{#1 \leq #2}}
-
-\newcommand{\rboundle}[2]{\ensuremath{\vdash_{bounds}#1 \leq #2}}
-
-\def\codesize{\normalsize}
-\definecolor{programs}{gray}{0.1}
-
-\lstset{
-  basicstyle=\ttfamily,  
-}
-
-\makeatletter
-\lst@CCPutMacro
-    \lst@ProcessOther {"22}{\lst@ifupquote \textquotedbl
-                                     \else \char34\relax \fi}
-    \@empty\z@\@empty
-\makeatother
-
-\lstloadlanguages{C}
-
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
-
-\title{A Formal Model of Checked C}
-
-% \ignore{
-% \author{\IEEEauthorblockN{1\textsuperscript{st} Leonidas Lampropoulos}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{University of Maryland}\\
-% llampro@cs.umd.edu}
-% \and
-% \IEEEauthorblockN{1\textsuperscript{st} Michael Hicks}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{University of Maryland}\\
-% llampro@cs.umd.edu}
-% }
-% \author{Leonidas Lampropoulos \and
-% Michael Hicks}
-% }
-\maketitle
-
-\begin{abstract}
-
-  We present a formal model of Checked C, a dialect of C
-  that aims to enforce spatial memory safety. Our model pays particular
-  attention to the semantics of dynamically sized, potentially null-terminated
-  arrays.
-We formalize this model in Coq, and prove that any spatial memory
-  safety errors can be \emph{blamed} on portions of the program
-  labeled \emph{unchecked}; this is a Checked C feature that supports
-  incremental porting and backward compatibility.
-Our model develops an operational semantics that uses fat pointers to
-  guarantee spatial safety. However, we formalize a compilation scheme
-  that can yield thin pointers, with bounds information managed using
-  inserted code. We show that the generated code faithfully simulates
-  the original.
-Finally, we build an executable version of our model in PLT Redex, 
-  and use a custom random generator
-  for well-typed and almost-well-typed terms to find
-  inconsistencies between our model and the Clang Checked C
-  implementation. We find these steps to be a useful way to co-develop a language
-  (Checked C is still in development) and a core model of it.
-\end{abstract}
-
-\section{Introduction}\label{sec:intros}
-
-The C programming language remains extremely popular despite the
-emergence of new, modern languages. Unfortunately, C programs lack
-spatial memory safety, which makes them susceptible to a host
-of devastating vulnerabilities, including buffer overflows and
-out-of-bounds reads/writes. Despite their long history, buffer
-overflows and other spatial safety violations are among the most
-prevalent and dangerous vulnerabilities on the Internet today \cite{Zeng:2013:SRF:2534766.2534798}.
-
-Several industrial and research efforts---including CCured~\cite{Necula2005},
-Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
-explored means to compile C programs 
-to automatically enforce spatial safety. These
-approaches all impose performance overheads that are deemed too high for
-use in deployment. Recently, Microsoft introduced \checkedc, an
-open-source extension to C with new types and
-annotations whose use can ensure a program’s spatial
-safety~\cite{Elliott2018}. Importantly, \checkedc supports development that is
-incremental and compositional. Code regions (e.g.,
-functions or whole files) designated as \emph{checked} are sure to enforce
-spatial safety, a property which is preserved via composition with
-other checked regions. But not all regions must be checked: Checked
-C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
-may coexist in the same code, which permits a deliberate (and
-semi-automated) refactoring process. Parts of the FreeBSD kernel have
-been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
-overhead seems low enough for practical deployment.
-
-While \checkedc promises to enforce spatial safety, we might wonder
-whether its design and implementation deliver on this promise, or even
-what ``spatial safety'' means when a program contains both checked and
-unchecked code. In prior work, \citet{ruef18checkedc-incr} developed a core
-formalization of \checkedc and with it proved that \emph{checked code cannot 
-  be blamed}: any spatial safety violation can only be attributed to
-code that is not in a checked region. While their work is a good start,
-it fails to model important aspects of \checkedc's functionality,
-particularly those involving pointers to arrays. In this paper, we
-cover this gap, making three main contributions.
-
-\myparagraph{Dynamically bounded and null-terminated arrays}
-Our first contribution is a core formalism called \lang, which extends
-\citet{ruef18checkedc-incr} with several new features, most notably 
-\emph{dynamically bounded arrays} (Section~\ref{sec:formal}). 
-Dynamically bounded arrays are those whose size is
-known only at run time, as designated by in-scope variables using
-dependent types. A pointer's accessible memory is bounded both above
-and below, to admit arbitrary pointer arithmetic.
-
-\lang also models
-\emph{null-terminated} arrays, a kind of dynamically bounded array whose upper bound 
-defines the array's \emph{minimum} length---additional space is available 
-up to a null terminator. For example, the \checkedc type
-\lstinline{nt_array_ptr<char> p:count(n)} says that
-\lstinline{p} has length \emph{at least} \lstinline{n} (excluding the
-null terminator), but further capacity is present if \lstinline|p[n]| is not null. \checkedc (and \lang) supports flow-sensitive
-\emph{bounds widening}: statements of the form \lstinline|if (*p) | $s$,
-where \lstinline{p}'s type is \lstinline{nt_array_ptr<T> count(0)},
-typecheck statement $s$ under the assumption that \lstinline{p} has
-type \lstinline{nt_array_ptr<T> count(1)}, 
- i.e., one more than it was,
-since the character at the current-known length is
-non-null. Similarly, the call \lstinline|n = strlen(p)| will widen
-\lstinline|p|'s bounds to \lstinline|n|. Subtyping permits treating
-null-terminated arrays as normal arrays of the same size (which
-does not include, and thereby protects, the null terminator).
-
-We prove, in Coq, a blame theorem for \lang.  As far as we are aware,
-ours is the first formalized type system and proof of soundness for
-pointers to null-terminated arrays with expandable bounds.
-
-\myparagraph{Sound compilation of checked pointers} Our second
-contribution is a formalization of bounds-check insertion for array
-accesses (Section~\ref{sec:compilation}). Our operational semantics
-annotates each pointer with metadata that describes its bounds, and
-the assignment and dereference rules have premises to confirm the
-access is in bounds. An obvious compilation scheme (taken by
-Cyclone~\cite{Jim2002,GrossmanMJHWC02}, CCured~\cite{Necula2005}, and
-earlier works) would be to translate annotated pointers to multi-word
-objects: one word for the pointer, and 1-2 words to describe its lower
-and upper bounds. Inserted checks reference these bounds. While
-convenient, such ``fat'' pointers are expensive, and break backward
-binary compatibility with legacy pointers. We formalize \checkedc's
-compilation approach, which uses a single machine word for the
-pointer, and adds checks involving the declared bounds (e.g., in a
-dependent type) or additional stack-allocated \emph{ghost variables}
-to accommodate bounds widening. We show that the compiled program
-\emph{simulates} the original by mechanizing \lang and the compilation
-judgment in PLT Redex~\cite{pltredex}, and use its random testing
-feature to give confidence that simulation holds.
-
-As far as we are aware, \lang is the first formalism to cleanly
-separate bounds-checking compilation from the core semantics; prior
-work merged the two, conflating \emph{meaning} with
-\emph{mechanism}~\cite{Feng2006,Condit2007}. In carrying out the 
-formalization, we discovered that our compilation approach is 
-more expressive than that proposed in the Checked C
-specification~\cite{checkedc} (Section~\ref{sec:disc}); we 
-would not have discovered this improvement had we not separated checks from 
-semantics.
-
-\myparagraph{Model-based randomized testing} Our third
-and final contribution is a strategy and implementation of model-based
-randomized testing (Section~\ref{sec:evaluation}).  To check the
-correctness of our formal model, we compare the behavior between the
-existing Clang \checkedc implementation and our own model. This is
-done by a conversion tool that converts expressions from \lang into
-actual \checkedc code that can be compiled by the Clang \checkedc
-compiler. We build a random generator of programs largely based on the
-typing rules of \lang and make sure that, both statically and
-dynamically, \lang and Clang Checked C are consistent after
-conversion.  This helped rapidly prototype the model and uncovered
-several issues in the Checked C compiler.
-
-We begin with a review of \checkedc (Section~\ref{sec:overview}),
-present our main contributions (Sections~\ref{sec:formal}--\ref{sec:evaluation}),
-and conclude with a discussion of
-related and future work (Sections~\ref{sec:related}, \ref{sec:conclude}).
-
 \section{\checkedc Overview}\label{sec:overview}
@@
 This section describes Checked C, which extends C with new pointer
-types and annotations that ensure spatial safety. More
-details can be found in a prior overview~\cite{Elliott2018} or the
-full specification~\cite{checkedc}.
-Checked C is implemented as a fork of Clang/LLVM and is freely
-available.\footnote{\url{https://github.com/Microsoft/checkedc-clang}}
+types and annotations that ensure spatial safety. Development of
+Checked C was initiated by Microsoft Research in 2015 but starting in
+late 2021 was forked and is now actively managed by the Secure Software
+Development Project (SSDP). Details can be found in a prior
+overview~\cite{Elliott2018} or the full specification~\cite{checkedc}.
+Checked C is implemented as an extension of Clang/LLVM and the SSDP
+fork is freely available at \url{https://github.com/secure-sw-dev}.
 
@@
 
+% {\captionsetup[lstlisting]{margin = 8 mm}
+%   \begin{lstlisting}[xleftmargin=8 mm]
+% array_ptr<int> memcpy(int x,
+%      array_ptr<int> a : count(x),
+%      array_ptr<int> b : count(x)) : count(x) {
+%  for(int i = 0; i < x; ++i) {
+%     (*a = *b); a++; b++;
+%  }
+%  return a;
 }
 
-\caption{Parsing a string of UTF16 hex characters in \checkedc}
+% static int idx = 0;
+% int buf (int * reg,
+%           nt_array_ptr<int> x,
+%           nt_array_ptr<int> y) {
+%   if (idx < n)
+%     unchecked { * x = &(reg + idx); idx++; }
+%   else bug();
+%   memcpy(n,x,y);
+%   return 0;
+% }
+% }
+\caption{Parsing a String of UTF16 Hex Characters in \checkedc}
 \label{fig:checkedc-example}
@@
 
+% Fig.~\ref{fig:checkedc-example}, all dereferences of pointer \code{a}
+% are within the range $[$\code{0}$,$\code{x}$)$ and therefore
+% safe. Moreover, if we were to remove the condition \code{i<x} from the
+% \code{for} loop, attempting to dereference an out-of-bounds memory
+% location, the inserted checks guarantee that the system will raise an
+% out-of-bound exception instead of crashing with a segmentation fault
+% or, worse, allowing the access to
+% happen. %if a out-of-bound error happens, the system causes an out-of-bound exception.
+
+% To insert such checks, both array and NT-array pointers are associated
+% with two values, their \emph{bounds}, defining the range of memory
+% referenced by the pointers. An array pointer $p$ with
+% bounds $b_l$ and $b_h$ allows access to the memory region
+% $[p+b_l,p+b_h)$.
+
 \subsection{Checked Pointer Types}
+%
 \checkedc introduces three varieties of \emph{checked pointer}:
@@
 \end{itemize}
+  Dropping the bounds expression on an \code{nt_array_ptr} is equivalent
+  to the bounds being \code{count(0)}.
+
 The \checkedc compiler will instrument loads and stores of checked
@@
 terminator. To address this problem, Checked C supports \emph{bounds
-  widening}. If $p$ has bounds expression
-\code{bounds(}$e_l$,$e_h$\code{)} a program may read from (but not
+  widening}.
+If $p$'s bounds expression is \code{bounds}$(e_l$,$e_h)$ a program may read from (but not
 write to) $e_h$; when the compiler notices that a non-null character
 is read at the upper bound, it will extend that bound to $e_h+1$.
+% \liyi{solved: II.B.: LaTeX issue, your inline monospaced code snippets generate a space on
+%   the following line (i.e. "a program may read from (but not write to)" is not
+%   aligned with the left boundary of the paragraph).}
 
 \subsection{Example}
+\label{sec:example}
 
@@
 parser, \url{https://github.com/kgabis/parson}} 
-The function \code{parse_utf16_hex} on lines 1-17 takes as its argument
-null-terminated pointer \code{s} from which it attempts to read four
+The function \code{parse_utf16_hex} on lines 1-15 takes a 
+null-terminated pointer \code{s} as its argument, from which it attempts to read four
 characters. These are interpreted as hex digits and converted to an
 \code{uint} returned via parameter \code{result}. At first,
-\code{s} has no specific bounds annotation, which we can interpret as
+\code{s}~has no specific bounds annotation, which we can interpret as
 \code{count(0)}; this means that \code{s[0]} may be read on line
 5. The true branch of the conditional (which extends all the way to
-the brace on line 15) is thus typechecked with \code{s} given a
+the brace on line 13) is thus type-checked with \code{s} given a
 \emph{widened} bound of \code{count(1)}. Likewise, the conditionals on
@@
 
-The \code{parse} function on lines 18-26 repeatedly invokes
+The \code{parse} function on lines 16-26 repeatedly invokes
 \code{parse_utf16_hex} with its parameter \code{s}, and fills out
@@
 arithmetic. We specify its bounds as
-\code{bounds(p,p+n)} to support this: even as \code{q} changes, 
-variables \code{p} and \code{n} do not. Converting from an
+\code{bounds(p,p+n)} to support this: even as \code{q} changes, variables \code{p} and \code{n} (and therefore also \code{q}'s bounds) do not. Converting from an
 \code{array_ptr<uint>} to a \code{ptr<uint>}, done for the call on
-line 25, requires proving the array has size at least 1. While this is true
+line 23, requires proving the array has size at least 1. While this is true
 because of the loop condition \code{q < p+n}, which is \code{q}'s
 upper bound,  the compiler is not smart enough to figure this
-out. To convince it, we manually insert a bounds check using
-\code{dyn_bounds_cast}.
+out. To convince it, we manually insert a \emph{dynamic cast} via
+\code{dyn_bounds_cast}, which is trusted at compile-time but confirmed
+with a dynamic check at run-time.
 
@@
 
+% For soundness, variables used in a bound expression may
+% neither be modified nor have their address taken until the pointer
+% containing the bound expression is out of the scope. As such, some
+% legacy idioms may be unsupported. \yiyun{Mention how this restriction further motivates our dynamic bounds tracking?} In \checkedc, variables with
+% checked pointer types or containing checked pointers must be
+% initialized when they are declared.
+
+% \leo{The following is no longer needed, I think. It doesn't add anything new now}
+% Fig.~\ref{fig:checkedc-example} \textbf{(a)} shows a usage of a \code{checked} NT-array pointers with dependent functions.
+We provide a special \code{memcpy} function to copy the data from NT-array pointer \code{b} to \code{a}.
+The execution of the function is guaranteed to the spatially safe because the argument types for \code{a} and \code{a} are set to \code{nt_array_ptr<int> : count(x)} (meaning that the address range is $[$\code{0}$,$\code{x}$)$), so it guarantees that the two input argument NT-arrays to have at least length \code{x}; thus, the \code{for} execution is valid all the time.
+%  Sec.~\ref{sec:formal} discusses how we formally define these pointer bounds with slightly different syntax.
 \subsection{Other features}
@@
 \subsection{Spatial Safety and Backward Compatibility}
+%
 \checkedc is backward compatible with legacy C in the sense that all
-legacy code will typecheck and compile. However, only code that
+legacy code will type-check and compile. However, only code that
 appears in \emph{checked regions}, which we call \emph{checked code},
@@
   within checked ones.}  Within checked regions, both
-legacy pointers and certain unsafe idioms (e.g., variadic function
+legacy pointers and certain unsafe idioms (e.g., \emph{variadic} function
 calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
-satisfies these conditions, and will typecheck in a checked region.
+satisfies these conditions, and will type-check in a checked region.
 
@@
 formalization of \checkedc, that \emph{checked code cannot be blamed}:
-Any spatial safety violation owes to the execution of unchecked
+Any spatial safety violation is caused by the execution of unchecked
 code. In this paper we extend that result to a richer formalization of
 \checkedc.
+
+% \paragraph*{\textbf{NT-array Pointer Formalization}} NT-array pointers point to NT-arrays whose endings are determined by a null (\code{'\0'}).
+The length of a NT-array is usually not fixed.
+The two uses of NT-array pointers in \checkedc are branching and \code{strlen} operations.
+\code{if (*x) }$e_1$\code{ else }$e_2$ branches to $e_1$/$e_2$ depending one the \code{x}'s data, while \code{strlen(x)} computes the length of the NT-array pointed to by \code{x}.
+% \liyi{move to formal and later section.}
+
+% \paragraph*{\textbf{Dependent Functions}} \checkedc allows users to declare dependent functions.
+Fig.~\ref{fig:checkedc-example} shows an example usage of dependent functions.
+The integer argument \code{x} is used to specify the bounds of the two NT-array pointers.
+It means that when we call the \code{memcpy} function, the upper-bounds of the two NT-array pointers must be no less than \code{x}.
+If the programmer makes the mistake of writing \lstinline{i <= x} in the for loop, the runtime checks will ensure that the indexing of\lstinline|x| in either \code{a} or \code{b} results in an error.
+% Because of the dependent function features, we are able to utilize the fact that the two remaining length of the NT-array pointers are greater than \code{x}, so that the \code{for} is guaranteed to execute properly.
 

### 2023-12-01 sandp23/checkedc-formal/conclusion.tex
update.

\section{Conclusion and Future Work}
\label{sec:conclude}

This paper presented \CoreChkC, a formalization of an extended core of
the Checked C language which aims to provide spatial memory
safety. \CoreChkC models dynamically sized and
null-terminated arrays with dependently typed bounds that can
additionally be widened at runtime. We prove, in Coq, the key safety
property of Checked C for our formalization, {\em blame}: if a mix of
checked and unchecked code gives rise to a spatial memory safety
violation, then this violation originated in an unchecked part of the
code. We also show how programs written in \CoreChkC (whose
semantics leverage fat pointers) can be compiled to \elang (which does
not) while preserving their behavior. We developed a version of \lang
written in PLT Redex, and used a custom term generator in conjunction with Redex's
randomized testing framework to give confidence that compilation is
correct. We also used this framework to cross-check \lang
against the \checkedc compiler, finding multiple inconsistencies
in the process.

As future work, we wish to extend \CoreChkC to model more of Checked
C, with our Redex-based testing framework guiding the process. The
most interesting Checked C feature not yet modeled is \emph{interop
  types} (itypes), which are used to simplify interactions with
unchecked code via function calls. A function whose
parameters are itypes can be passed checked or unchecked pointers
depending on whether the caller is in a checked region. This feature
allows for a more modular C-to-Checked C porting process, but complicates reasoning
about blame. A more ambitious next step would be to extend an existing
formally verified framework for C, such as CompCert~\cite{compcert} or VeLLVM
\cite{Zhao:2012:FLI:2103621.2103709}, with Checked C features, towards
producing a verified-correct Checked C compiler. We believe that
\lang's Coq and Redex models lay the foundation for such a step, but
substantial engineering work remains.

### 2023-12-01 sandp23/checkedc-formal/introduction.tex
update.

% \review{Request from Reviewer A:
% \begin{itemize}
% \item clarify relationship between formalization and implementation
% \item correct and address mistakes, typos, confusing terminology and code snippets
% \item address pressing questions in the paper.
% \end{itemize}}

% \review{Request from Reviewer B:
% \begin{itemize}
% \item Add details in section 5 (see above)
% \item Clarify the status of Theorem 4
% \end{itemize}}

% \review{Request from Reviewer C:
% \begin{itemize}
% \item confirmation that blame theorem issue is essentially just a typo.
% \item  clarify the policy rules wrt accessing (un)checked pointers in (un)checked regions
% \item  remove, or at least point out and justify the discrepancies between the Coq code and the paper, eg concerning Fig2
% \end{itemize}}

% \review{Request from the conditions:
% Minimal requirements for acceptance ("shepherding conditions"):
% \begin{itemize}
% \item please clarify the formalization status of Theorem 4 and the intended guarantee w.r.t. NULL 
%   dereferencing, clarify whether the code examples are intended to be actual C -- cf the discussion
%   of  pointer arithmetic and undefined behavior in review 1 -- and tweak them if necessary) 
% \item please clarify the policy of dereferencing of (un)checked pointers in (un)checked regions as pointed out by two reviewers, and address the discrepancies between the paper's calculus and the Coq formalization
% \item please include more details on the proposed model-driven random testing approach from section 5 (cf review B)
% \end{itemize}
% }

\section{Introduction}\label{sec:intros}

% \review{End of I.: a diagram would be helpful understanding how III, IV and V relate
%   to each other and to clang-checkedc}

% \review{The paper builds on the operational semantics from [21]. Even though the
%   authors explained at high-level the main extensions w.r.t. [21], it'd be great
%   to add a paragraph that discusses what are the main changes from [21] in terms
%   of the technical development (if there are any), e.g., are there any new
%   challenges that needed to be solved while proving the blame theorem for this
%   paper's semantics?}

The C programming language remains extremely popular despite the
emergence of new, modern languages. Unfortunately, C programs lack
spatial memory safety, which makes them susceptible to a host
of devastating vulnerabilities, including buffer overflows and
out-of-bounds reads/writes. Despite their long history, buffer
overflows and other spatial safety violations are among the most
prevalent and dangerous vulnerabilities on the Internet today \cite{Zeng:2013:SRF:2534766.2534798}.

% One possible solution is Checked C
% \begin{itemize}
% \item Think of it as migratory/gradual typing, but for C.
% \item Distinct pointer types, but which are backward binary- and source-compatible.
% \item Checked regions, containing only checked pointers and restricted
%   idioms, aim to ensure spatial safety
% \item Implemented as an extension to Clang/LLVM. Good performance
%   (compared to ASAN etc.)
% \end{itemize}

Several industrial and research efforts---including CCured~\cite{Necula2005},
Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
explored means to compile C programs 
to automatically enforce spatial safety. These
approaches all impose performance overheads deemed too high for
deployment use. Recently, \citet{Elliott2018} introduced \checkedc, an
open-source extension to C with new types and
annotations whose use can ensure a program’s spatial safety.
Importantly, \checkedc supports development that is 
incremental and compositional. Code regions (e.g.,
functions or whole files) designated as \emph{checked} enforce
spatial safety in a manner preserved by composition with
other checked regions. But not all regions must be checked: Checked
C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
may coexist in the same code, which permits a deliberate (and
semi-automated) refactoring process. Parts of the FreeBSD kernel have
been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
overhead seems low enough for practical deployment.

While \checkedc promises to enforce spatial safety, we might wonder
whether its design and implementation deliver on this promise, or even
what ``spatial safety'' means when a program contains both checked and
unchecked code. In prior work, \citet{ruef18checkedc-incr} developed a core
formalization of \checkedc and with it proved a \emph{soundness} theorem for
checked code: any stuck (i.e., ill-defined)
state reached by a well-typed program amounts to a spatial safety
violation; such a state can always be attributed to, i.e., 
\emph{blamed on}, the execution of code that is not in a checked
region. While their work is a good start, 
it fails to model important aspects of \checkedc's functionality,
particularly those involving pointers to arrays. In this paper, we
cover this gap, making three main contributions.

\myparagraph{Dynamically bounded and null-terminated arrays}
%
Our first contribution is a core formalism called \lang, which extends
\citet{ruef18checkedc-incr} with several new features, most notably 
\emph{dynamically bounded arrays} (Section~\ref{sec:formal}). 
% First, we add important language features previously not modeled and
Dynamically bounded arrays are those whose size is
known only at run time, as designated by in-scope variables using
dependent types. A pointer's accessible memory is bounded both above
and below, to admit arbitrary pointer arithmetic.

\lang also models
\emph{null-terminated} arrays, a kind of dynamically bounded array
whose upper bound
defines the array's \emph{minimum} length---additional space is available 
up to a null terminator. For example, the \checkedc type
\lstinline{nt_array_ptr<char> p:count(n)} says that
\lstinline{p} has length \emph{at least} \lstinline{n} (excluding the
null terminator), but further capacity is present if %
\lstinline|p[n]| is not null. \checkedc (and \lang) supports flow-sensitive
\emph{bounds widening}: statements of the form \lstinline|if (*p) | $s$,
where \lstinline{p}'s type is \lstinline{nt_array_ptr<T> count(0)},
typecheck statement $s$ under the assumption that \lstinline{p} has
type \lstinline{nt_array_ptr<T> count(1)}, 
 i.e., one more than it was,
since the character at the current-known length is
non-null. Similarly, the call \lstinline|n = strlen(p)| will widen
\lstinline|p|'s bounds to \lstinline|n|. Subtyping permits treating
null-terminated arrays as normal arrays of the same size (which
does not include, and thereby protects, the null
terminator).\footnote{
    See Sec.~\ref{sec:related} for a careful comparison of
    \citet{ruef18checkedc-incr} and \lang.
}

We prove, in Coq, a blame theorem for \lang.  As far as we are aware,
ours is the first formalized type system and proof of soundness for
pointers to null-terminated arrays with expandable bounds.

% \yiyun{What about strlen? Clang doesn't do null check/bounds check on
%   strlen. We found that issue while developing our model rather than
%   through random testing. If we don't have enough bugs
%   revealed from random testing, maybe we should instead focus on the
%   under-specified or unclear behaviors from the checkedc spec that we
%   found in the process of designing the typing/semantic rules}
% \mwh{Mentioned strlen above. How should we change what's written?}

\myparagraph{Sound compilation of checked pointers} Our second
contribution is a formalization of bounds-check insertion for array
accesses (Section~\ref{sec:compilation}). Our operational semantics
annotates each pointer with metadata that describes its bounds, and
the assignment and dereference rules have premises to confirm the
access is in bounds. An obvious compilation scheme (taken by
Cyclone~\cite{Jim2002,GrossmanMJHWC02}, CCured~\cite{Necula2005}, and
earlier works) would be to translate annotated pointers to multi-word
objects: one word for the pointer, and 1-2 words to describe its lower
and upper bounds. Inserted checks reference these bounds. While
convenient, such ``fat'' pointers are expensive, and break backward
binary compatibility with legacy pointers.

  To show that pointer annotations can be safely
  erased, and thus fat pointers are not needed, we formalize a
  translation of \lang to \elang, which is an 
  untyped version of \lang that drops metadata annotations, and
  lacks bounds/null checks in the semantics rules. Instead,
  the compilation process inserts null/bounds checks explicitly, leveraging
  compile-time type information. While we do not definitively prove
  it, we provide strong evidence that compilation is correct. We use PLT
  Redex~\cite{pltredex} to mechanize (a generalization of) \lang, 
  \elang, and compilation between the two, and we use randomized testing 
to validate that the compiled program \emph{simulates} the
original. In addition to demonstrating the technical point that metadata
annotations in the \lang formalism do not necessitate fat pointers,
compilation also sheds light on the actual Checked C compilation
process. 
% \review{ Beginning of page 2: "we show", then "give confidence". I am left confused as
%   to whether the CoreChkC --> CoreC compilation is shown correct (theorem with
%   proof), or carefully debugged and tested in plt-redex (which is fine too, I
%   just find the choice of terminology confusing)}
% \liyi{typo, should be "validate through randomized testing"}

As far as we are aware, \lang is the first formalism to cleanly
separate bounds-checking compilation from the core semantics; prior
work~\cite{Feng2006,Condit2007} merged the two, conflating
\emph{meaning} with \emph{mechanism}. In carrying out the
formalization, we discovered that our compilation approach for
null-terminated array pointers is more expressive than that proposed
in the Checked C specification~\cite{checkedc}
(Section~\ref{sec:disc}); we would not have discovered this
improvement had we not separated checks from semantics.
%\leo{Either checks from semantics or the checks from the semantics.
%  Not sure what I prefer phrasing wise...}

% The second contribution is a presentation of the semantics of
% dynamically bounded arrays that simplifies formal reasoning. 
% \mwh{I don't follow what's here. Want to change this to be about the
%   clean separation of semantics metadata from obligating the compiler
%   to include it as fat pointers.}
% We design our type system to closely model the behavior of the
% \checkedc specification. However, we choose to diverge slightly from the
% \checkedc semantics by utilizing stack variables to keep additional
% bounds information that cannot be expressed with a simple type
% system that does not support theorem proving. Due to
% the limitation that \checkedc functions must be binary compatible with
% C, we cannot explicitly pass bounds to functions. We make a
% comprimise and restrict the more accurate bounds information to be
% accessible only within the stack frame where the checked array pointer
% is defined. As we will see, despite this restriction, the user can
% write significantly more expressive code that would not be
% possible with the existing \checkedc specification.
% As far as we know, prior systems \yiyun{what
%   systems? the current checkedc spec is clearly one of them. What
%   about Deputy?} 
% \yiyun{widening of null-terminated array pointers at the current
%   stacck frame}
% take a more conservative approach and insert checks completely based
% on bounds determined at compile time. 
%  
% \mwh{merge the following and the above into one para/contribution}
% By introducing this new mechanism of bounds tracking through stack
% variables, we run the risk of accidentally using ``fat pointers'' in
% our formalism, where every the heap and literals are annotated with
% types. To ensure our system does not access bounds information that is
% not supposed to be accessible at runtime, we compile from \lang into
% \elang, where the expressions are free of bounds annotations and the
% dynamic checks and widening expressions are made explicit. We prove
% the blame theory using the annotated language \lang, and then prove
% that \lang are \elang are bisimilar. \yiyun{Maybe discuss its
%   impact on proof engineering? We don't have the full proofs yet so not
%   sure how to comment on that.}

% define the semantics of the source language via compilation to a
% target where the checks are made manifest. We'd prefer the semantics
% are made directly manifest as operational rules, and we followed this
% approach, taken by the earlier formal work. However, we found that the
% addition of dynamic sizes complicates this approach---seems to require
% ``fat pointers.'' So: We developed a compiler to a system without
% bounds annotations and proved it bisimilar to the one with
% annotations.

\myparagraph{Model-based randomized testing} Our third and final
contribution is a strategy and implementation of model-based
randomized testing (Section~\ref{sec:evaluation}).  To check the
correctness of our formal model, we compare the behavior between the
existing Clang \checkedc implementation and our own model. This is
done by a conversion tool that converts expressions from \lang into
actual \checkedc code that can be compiled by the Clang \checkedc
compiler. We build a random generator of programs largely based on the
typing rules of \lang and make sure that, both statically and
dynamically, \lang and Clang Checked C are consistent after
conversion.  This helped rapidly prototype the model and uncovered
several issues in the Checked C compiler.

\begin{figure}
\includegraphics[width=0.5\textwidth]{relationship.pdf}
\caption{
    \lang models' relationship to \checkedc
}
  \label{fig:model-relation}
\end{figure}

  \myparagraph{Summary Visualization} The
  relationship among our contributions is visualized in
  Fig.~\ref{fig:model-relation}. With the Coq model of \lang we prove
  soundness (and with it, \emph{blame}) of the \checkedc type system
  and semantics. With the Redex model, we use randomized testing to
  validate both type soundness and compilation correctness, where the
  latter shows how compilation need not output fat pointers despite
  the use of pointer annotations in the \lang model. The Redex \lang
  model is also the basis of randomized testing of the correctness of
  the \checkedc compiler implementation, both its type checker and the
  semantics of its emitted code, at least for the subset of the
  language in the Redex model. The Redex model's syntax
  is slightly richer than the Coq version: conditional guards and
  function arguments may be arbitrary expressions, where the Coq
  version limits them to constants and variables, making handling of
  dependent types a bit simpler. We find a useful synergy between the
  Coq and Redex models for carrying out a language development. The
  richer, executable Redex model is useful for quickly modeling and
  testing new features, both formally and against a real
  implementation. Once solidified, new features can be added to the
  Coq model (perhaps somewhat simplified) for final proofs of
  correctness.

We begin with a review of \checkedc (Section~\ref{sec:overview}),
present our main contributions
(Sections~\ref{sec:formal}--\ref{sec:evaluation}), and conclude with a
discussion of 
related and future work (Sections~\ref{sec:related},
\ref{sec:conclude}). All code and proof artifacts (both for Coq and
Redex) can be found at \url{https://github.com/plum-umd/checkedc}.

% In this paper, we develop a NULL-terminated (NT) checked pointer type
% in Checked C, whose syntax is \lstinline{nt_array_ptr<T>} (for any type
% \lstinline{T}) \yiyun{we didn't model anything besides a single
%   numeric type but what's the meaning of an ntarray when T /=
%   char? If it was int, then it implicitly rules out the possibility of
% storing 0}. Programmers familiar with C's standard string library will
% be quite familiar with NT pointers. But spatially safe NT pointers are
% not present in modern type-safe languages (such as Java, Scala,
% Haskell, Rust, etc.), which typically pair dynamically-sized buffers
% with metadata to indicate the buffer's length and capacity. While some
% prior safe-C implementations (such as CCured, Cyclone, and Deputy)
% have supported NT pointers, no prior work of which we are aware has
% formalized and proved them safe, nor has it assessed \yiyun{did we
%   assess anything?} their performance
% against non-NT alternatives. We fill these gaps in this work.

% In this paper we formalize Checked C with checked NT pointers and
% re-prove our compositional proof of blame (done in the Coq proof
% assistant); our formalism also includes lightweight dependent and
% function types, and other extensions. We have also implemented checked
% NT pointers in the Checked C compiler, and evaluated their
% effectiveness on a series of benchmarks; we use these benchmarks to
% show the advantages of our support compared to refactoring the code to
% use normal arrays. \yiyun{I don't think we ever evaluated nt-array
%   pointer against ordinary array pointers. It's more of a legacy code
%   issue since C uses nt-array in its stdlib} Our implementation, code, and benchmarks are (or
% will be) available in the Checked C Github repository.

% Intro:
% \begin{itemize}
% \item NULL-terminated arrays are a feature long-present in
%   C. Low-overhead way of implementing expandable buffers.
% \item But: Spatially-safe languages have typically eschewed such
%   arrays in favor of traditional, bounds-checked arrays. For example,
%   Java Strings are not zero-terminated (I think)
% \item Some considerations in past languages, like Cyclone, CCured, and
%   Deputy, but never formalized and proved correct.
% \item Checked C is backward compatible with C, so we need to
%   incorporate NT arrays.
% \end{itemize}

### 2023-12-01 sandp23/discussion.tex
update.

\section{Difference in the \checkedc Formalization}\label{sec:discussion}

\begin{figure}[t]
{\small
{\captionsetup[lstlisting]{margin = 8 mm}

 \begin{lstlisting}[xleftmargin=8 mm]
void memcpy (nt_array_ptr<int> a : count(0),
         nt_array_ptr<int> b : count(0)) {
   while (* a) {
     if (* b)
        * a = * b;
     else break;
     a++; b++;
   }
}
  \end{lstlisting}

}
}
\caption{Bound Widening Examples}
\label{fig:bound-widening}
\end{figure}

% In the left of program \textbf{(c)} (Fig.~\ref{fig:checkedc-example}), we 
% assume that the lower/upper bounds of \code{x} are given as \code{0} in the beginning of the program. 
% In a fat-pointer framework, this program might not reach failure because the \code{x} type is adjusted dynamically and the upper bound keeps growing.
% In a classical language that guarantees spatial safety statically, however, this program results in dynamic bound check errors, because once we access \code{x} and increase the pointer address by $1$, the pointer reaches the upper bound given by the type.

There are some differences between our \checkedc formalization and the \checkedc specification/implementation. One of the examples is the use of $\Theta$ in the type system. The difference has been described in Sec.~\ref{subtype-type}.
The other difference is the bound widening scope of the branching operations.
Bound widening refers to that when we are accessing the final element in an NT-array indicated by the upper bound of the NT-array, we find that the element is not null; clearly, there is more elements after the element, so the upper bound can extended dynamically and we are able to access the elements after the final one.
In a fat-pointer framework, this is easy to implement. In a static bound system, however, there is impossible to implement the bound widening behaviors in any context because all bound information is statically guided. 
The \checkedc specification allows bound widening behaviors in a limited format. 
\code{strncat} in Fig.~\ref{fig:strncat-ex} is an example bound widening in \checkedc. Without the \texttt{strlen} usage in line 5, \code{b}'s upper bound is not extended to \code{x}. In that case, line 13 results in a $\ebounds$ error.

Our formalization is different from the \checkedc specification and implementation.
One cool thing that our formalization can do is a second version of the \code{memcpy} implementation in Fig.~\ref{fig:bound-widening}. 
This function takes two NT-array pointers, and copies the data in NT-array \code{b} to \code{\a} as mush as we can until we hit a null-character in either \code{a} or \code{b}.
The current \checkedc specification and compiler do not support this program, but they have difference. 
The \checkedc specification allows bound widening in the scope of the \code{true} branch, meaning that the \code{b}'s upper bound is $1$ in line 5 (in the \code{true} branch) in the first loop iteration; so does the pointer \code{a} in the while loop. In each iteration, \code{a}'s upper bound is extended by $1$. The problem is that the scope of the bound widening in the specification is lost after jumping out of the branching scope. After line 5, \code{b}'s upper bound is back to $0$. The \code{b++} in line 7 shifts \code{b}'s pointer address; thus, in the second iteration, the line 5 code results in a $\ebounds$ error because \code{b}'s dereference address is bigger than the upper bound. 
On the other hand, the \checkedc compiler results in a $\ebounds$ error for \code{a}'s pointer in line 3 in the second iteration, because the compiler's bound widening behavior in a \code{while} loop is only restricted to a local loop body, and does not extended to further loop iteration. 
Our formalization extends the bound widening behavior after the \code{true} branch scope; thus, executing the \code{memcpy} in Fig.~\ref{fig:bound-widening} does not reach an runtime error, and NT-array \code{a} copies \code{b}'s data as long as the length of \code{a} is greater than \code{b}.

We implement this behavior different from both the \checkedc specification and compiler because of two reasons. First, the \checkedc specification and compiler disagree here so there is no formally "good" behavior. Second, we believe that this formalization permits more interesting programs, and sometimes, more efficient programs.

In our compiler implementation, we utilize ghost variables, which is introduced in Sec.~\ref{sec:compilation}, to equalize the behavior in our formalization and implementation. Thus, the behavior of our formalization and implementation matches tightly (proved by Theorem~|ref{simulation-thm}).
The lack of shadow variables become more of an issue when we extend the
\checkedc specification with other bounds widening primitives such as \code{strlen}. 
For example, if users are not careful and move the \code{strlen} expression for pointer \code{y} in line 5 to later  are not able to Fig.~\ref{fig:strncat-ex}, it might causes the compiler ill-recognize the scope of the \code{strlen} bound widening effect; thus, the execution in line 13 fails.
With the shadow variables, it certainly allows us to mutate \code{y}'s bound widening scope and designs more effective compilation strategy to permit more intereting programs, like Fig.~\ref{fig:bound-widening}.
The use of shadow variables also allow an easy design of the compiler. 
complex static reasoning can be
difficult to implement. At the time of writing, the Clang \checkedc
implementation does not yet support the type of static reasoning to
allow functions like \code{memcpy} in Fig.~\ref{fig:bound-widening} to work properly.
Additionally, forcing the
user to declare a bound variable explicitly and guide the typechecker
to prove that it's a valid bound can result in extremely verbose
code.
Finally, shadow variables resides in the stack so they do not create any significant overhead for the execution of a program.

\liyi{below are more wording from Yiyun that I think we can drop. Below means to the end of the section. }

The
 \code{strcat\_checked} function from Fig.~\ref{fig:bound-widening}
 takes an extra argument \code{n} that represents the maximum capacity
 of the NT-array \code{s0}. We first use the \code{strlen} function to
 locate the end of \code{s0}, and then start copying from \code{s1} to
 \code{s0} through a while loop. With the \code{strlen} widening
 extension introduced by the Clang \checkedc compiler, the bounds of
 \code{s0} will be casted into \code{count(i)} after
 line 14. This well-intended casting behavior can end up shrinking the
bounds of \code{s0} when \code{strlen(s0)} is smaller than \code{n},
which is almost always the case when the \code{strcat\_checked}
function is called. This will force the programmer to expand the
call to \code{strlen} into an explicit \code{while} loop to prevent
the \checkedc compiler from casting the bounds, adding unnecessary
complexity. On the other hand, with shadow variables, we ensure at
runtime the ghost variable representing the upper bound can only get
larger. In our formal semantics, the
ghost variable will be updated only if the return value of \code{strlen} is
greater than the variable's current value. The subsequent indexing
into the array will thus be checked against the upper bound that is equal
to \code{max(n,i)}, avoiding the runtime error.
Sadly, while our formal model is better-behaved at runtime, we do not
leverage the existence of shadow variables in our type
system. This is because the original \checkedc specification is
not designed around the idea of shadow variables. If we were to extend the
specification with shadow variables, we could expose special syntax for
programmers to read those variables so they don't have to rely on
runtime casts to indirectly retrieve that information. In the type
system, we can use the ghost variable to represent the largest known
bounds of an array and extend the subtyping rules to allow shrinking
to smaller bounds that can be inferred statically.

The other significant difference is the use of shadow variables in our implementation of the \checkedc compilation.
The usefulness of shadow variables is not immediately obvious. In
this section, we provide examples that demonstrate the limitation of the
current \checkedc specification due to the lack of shadow variables.

In Fig.~\ref{fig:bound-widening}, we define the function \code{foo}
which takes an NT-array pointer of size $0$ as its argument. The
function initializes the variable \code{cnt} to \code{0} and
increments \code{cnt} only if the value pointed by \code{p} is not a
null character. Finally, after exiting the if statement, we attempt to
dynamically cast bounds of \code{p} to be \code{count(cnt)} before
returning.

It is also worth mentioning that the bookkeeping of shadow variables do
introduce hidden overhead, compared to an alternative approach that
relies on static proofs. In the \checkedc specification, it is
possible to take advantage of its data-flow analysis and assign more
precise bounds without relying on shadow variables. The function
\code{bar} from Figure.~\ref{fig:bound-widening} implements the same
functionality as \code{foo} but includes additional static bounds casting in the
\code{if} statement. The type system can prove that
\code{count(cnt)} is a valid bounds annotation after the if statement,
avoiding the overhead of an additional \code{dynamic_cast} as in \code{foo}.

However, we argue that there are strong reasons that justify
why we should include shadow variables. First,  Second,  \yiyun{I'm not familiar with the conversion tool. is this accurate?} This also
defeats the purpose of using \checkedc for gradually migrating legacy C code,
which, in practice, is more likely done by automatic
conversion tools that are unable to perform any sophisticated
theorem proving. Finally, the introduction of shadow variables can
always be made optional or simply used in conjunction with a static reasoning
system. \yiyun{I think I phrased this rather poorly. Would be helpful
  if someone could help me rephrase this a little} We expect to introduce
shadow variables to the \checkedc specification but defer some of the
more subtle design decisions that might have an impact on performance
until we receive more benchmark results from existing C to \checkedc
conversion tools \yiyun{should I cite the 3C paper here?}.
\ignore{
As we discuss in Section~\ref{sec:overview}, the \checkedc compiler
is able to statically determine that the upper bound of \code{p} must
be at least \code{1} in the then branch of the if statement. However,
once the program exits the if statement, the upper bound of \code{p}
is conservatively reverted back to \code{0}. Without using ghost
variables to keep track of the updated bounds, the \checkedc compiler
would attempt to check whether \code{count(cnt)} is included within
the statically determined imprecise bounds \code{count(0)} at line 6,
causing a runtime error when we pass in a non-empty string.
}

%\code{strcat_checked} function

% This limitation can also affect programs that are more practical. For
% example, at Line 16, we compute \code{strlen(p)} manually through a
% \code{while} loop. The \code{dynamic_bounds_cast} fail for the reason
% as the one we see in \code{foo}.

% The Clang \checkedc implementation uses statically determined
% bounds to insert runtime checks. In the left of Fig.~\ref{fig:checkedc-example} \textbf{(c)}, the
% \code{dynamic_bounds_cast} at line 3 will always succeed, because the
% compiler knows that within the scope of then branch, the pointer
% \code{p} must have at least one element. The same cast at line 6,
% however, will always fail, since there is no way to tell statically
% whether the program has entered the then branch before. The compiler
% will check whether the \code{count(1)} bounds specification is
% contained within the earlier \code{count(0)} specification, resulting
% in a runtime failure even when we pass in a non-empty string.

% Our formalization diverges from this runtime behavior and instead keeps
% track of the bounds on the stack. After entering the then branch, we
% increment the upper bound for \code{p}, effectively making the
% updated bounds information available even after we exit the if
% statement. The cast at line 6 will be checking the new bounds against
% the incremented bounds for non-empty strings.

% In our formalization, every pointer literal is annotated with bounds
% information. During compilation, the bounds annotations are converted
% into shadow variables that are updated during widening and accessed
% for runtime checks. This allows our formal model to successfully
% run functions defined like \code{foo}. However, the \checkedc
% specification is designed based on the absence of ghost
% variables and therefore does not provide primitive operations that
% allow programmers to interact with them.

% Fig.~\ref{fig:checkedc-example} \textbf{(e)} gives a more practical example of why keeping
% track of the bounds on the stack is useful. The program snippet
% implements the functionality of the \code{strlen} function using a
% while loop and a \code{cnt} variable. Even though the type system is
% unable to reason about the while loop statically (which would require
% inductive reasoning), as long as the runtime system updates the bounds
% in-place, the user can apply a
% \code{dynamic_bounds_cast} to soundly recover the more precise bounds
% information.

% \yiyun{That's actually not right. The fresh bounds variables are generated during the creation of x, before strlen is called. I'll rewrite this later.}
% The \code{strlen} formalization is a special case of the branching operations with an NT-array pointer Boolean guard.
% In this case, \checkedc generates a new fresh local variable in the local stack in its type rule, which records the length from the current position of a pointer (\code{x}) to the first null-character. We do not know the value of \code{x} in the type checking stage; instead, in the compiler, \checkedc insert dynamic checks of \code{x} into necessary places that use the upper bound information of \code{x}.
% Fig.~\ref{fig:checkedc-example} \textbf{(b)} calls \code{strlen} to get the length of the NT-array pointer \code{x}. Since the input lower/upper bounds for \code{x} are \code{0}, the result of executing \code{strlen} is always the length between the current position of \code{x} and the first null character in the list because the length is always greater than \code{0}. In general, calling \code{strlen} on an NT-array pointer might not return the above size. If the the length between the current position of \code{x} and the first null character is less then the high-bound size of the pointer type, then the result is the greater of the two.

% \yiyun{strcmp would be a useful example here? It would be assigned a signature \lstinline{nt_array_ptr<char> p0:bounds(p0,p0)}, \lstinline{nt_array_ptr<char> p1:bounds(p1,p1)} }
% \paragraph{\textbf{Subtyping and Casting Operations}} 
% It is a common usage in a program to use a large size array as a smaller one, i.e., using a \code{array_ptr<int> : count(5)} typed array as a \code{array_ptr<int> : count(2)} typed one. \checkedc provides a static casting operation $\ecast{\tau}{x}$ to cast a subtype into its supertype. Every type checked casting operation is always executed successfully without causing any error states and with a "zero" cost. One example use of static casting is given in \code{strncat} in Fig.~\ref{fig:strncat-ex}.
% The \code{strlen(a)} in line 4 updates the upper-bound of \code{a} to be \code{x}. Then, the statement \code{(nt_array_ptr<int> : count(0)) a} casts the upper-bound of pointer \code{a} from \code{x} to \code{0}.

% \yiyun{TODO: Rewrite using the static bound cast syntax from page 39 of the checekdc spec. Also, this type of static reasoning that the n >= 0 is easy to do, but I don't think it's implemented in Redex. Maybe we should at least mention this strlen extension is not part of the checkedc spec?}
% \liyi{described in formal typing system.}

\mwh{HERE IS OLD TEXT FROM BACKGROUND. MIGHT GO HERE}

\yiyun{I rewrote this paragraph to express more explicitly why the static bounds are not precise.}
The bounds inferred at compile time are not always precise for null-terminated arrays. \yiyun{Should we rephrase this or remove the same example from the introduction?} Consider the statement \lstinline|if (*p) {..._0} {..._1} ..._2|
where \lstinline{p}'s type is \lstinline{nt_array_ptr<T> count(0)}, the type checker can infer that \code{p} has the bounds annotation \code{count(1)} in \lstinline{..._1} because the \lstinline{*p} cannot possibly be the null terminator in the then branch. However, what type should we assign to \lstinline{p} at \lstinline{..._2}? With the limited expressiveness of the \checkedc type system, it can only soundly infer the more conservative bounds annotation \lstinline{count(0)}.
We will later discuss how we keep track of the updated bounds information at runtime and also allow programmers to reflect the updated bounds to the type system via dynamic bounds casts.

\yiyun{move the following text somewhere else}
It is still possible to keep track of the updated bounds information at runtime.
To do so, we store the upper bound of the pointer \lstinline{p} as a stack variable. Inside the then branch, the stack variable will be incremented, effectively tracking the information even after the program exits the if statement. Later, we will see an practical example that makes use of this extra bit of information.

\yiyun{I want to justify our decision of using stack variables by pointing out the lack of clarity of the checkedc spec on the runtime behavior. Should I include this discussion in a later section?}
It is worth noting that the original \checkedc specification is very unclear ...
% The above pointer bound description discusses the case when we know exactly the lower/upper bounds of a pointer $p$. The pointer bound information of \checkedc is detected through the \checkedc type system and statically inserted bound checks. It is unlikely we can always know the exactly bounds for a pointer $p$. 
% For example,
% a given array/NT-array pointer \code{a} is declared as \code{array_ptr<int> a : count(x)}/\code{nt_array_ptr<int> a : count(x)}. The \code{count(x)} here means that the lower-bound of the pointer is \code{0}, and the upper-bound is \code{x}.
% For an array pointer, even though the bound values of the lower/upper bounds (\code{0} and \code{x}) can change during evaluation, the array length (\code{x-0}) is fixed.

% For a NT-array pointer, the situation is different. 
% The interpretation of the bounds of an NT-array pointer is similar to an array pointer, but the
% range can extend further to the right, until a null terminator
% is reached (i.e., the null is not within the bounds).
% During this process, it is possible that the upper bound \code{x} is extended and the NT-array length is "extended" due to the discovery of the upper-bound array position is not a null character.

### 2023-12-01 sandp23/figures/memoryoverhead.tex
update.

@@
     ybar,  
+    ymin=-200, ymax=+200,
     enlargelimits=0.04,  
@@
     ]  
-\addplot coordinates {(ProFTPD,50) (TinyBigNum,90) (UFTPD, 120) (MicroHTTPD,80) (png2pnm, 60) (pnm2png, 150) (Parsons, 280)};
+\addplot coordinates {(ProFTPD,82.1) (TinyBigNum,81.1) (UFTPD, 86.3) (MicroHTTPD,82.3) (png2pnm, -102.5) (pnm2png, -51000) (Parsons, 9.8)};
 

### 2023-12-01 sandp23/figures/newusecase.tex
update.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=0.75]
%\node[text width=3cm] at (1.5,-0.5) 
%    {some text spanning three};
%\draw[step=5cm,gray,very thin] (0,0) grid (5,5);
\fill[taintcolor] (2.5,2.5) rectangle (5,5);
\fill[green!10!white] (0,2.5) rectangle (2.5,5);
\fill[green!40!white] (2.5,0) rectangle (5,2.5);
\fill[green!40!white] (0,0) rectangle (2.5,2.5);

\Text[x=3.8cm,y=4.5cm,fontsize=\scriptsize]{\bf High Risk and}
\Text[x=3.8cm,y=4.2cm,fontsize=\scriptsize]{\bf Hard to Convert}
\Text[x=3.8cm,y=3.9cm,fontsize=\scriptsize]{\bf Functions.}
\Text[x=3.8cm,y=3.6cm,fontsize=\tiny]{(Isolated in}
\Text[x=3.8cm,y=3.3cm,fontsize=\tiny]{the Sandbox)}

\Text[x=1.2cm,y=4.5cm,fontsize=\scriptsize]{\bf Low Risk and}
\Text[x=1.2cm,y=4.2cm,fontsize=\scriptsize]{\bf Hard to Convert.}
\Text[x=1.2cm,y=3.9cm,fontsize=\scriptsize]{\bf Functions.}
\Text[x=1.2cm,y=3.6cm,fontsize=\tiny]{(Partially annotated}
\Text[x=1.3cm,y=3.3cm,fontsize=\tiny]{and execute in c-region)}
\Text[x=1.2cm,y=2.9cm,fontsize=\tiny]{}
%\Text[x=1.2cm,y=4.3cm,fontsize=\tiny]{(Will be partially converted)}

\Text[x=2.3cm,y=1.8cm,fontsize=\scriptsize]{\bf Easy to Convert Functions.}
\Text[x=2.3cm,y=1.5cm,fontsize=\tiny]{(Will be fully converted to Checked C}
\Text[x=2.3cm,y=1.2cm,fontsize=\tiny]{with all necessary annotations}
\Text[x=2.3cm,y=0.9cm,fontsize=\tiny]{ and execute in c-region)}

%\Text[x=3.8cm,y=2.1cm,fontsize=\tiny]{Retrofitting}
%\Text[x=3.8cm,y=1.8cm,fontsize=\tiny]{Techniques.}
%\Text[x=3.8cm,y=1.5cm,fontsize=\tiny]{(ASAN, CCured,}
%\Text[x=3.8cm,y=1.3cm,fontsize=\tiny]{etc)}

%\Text[x=0.9cm,y=0.6cm,fontsize=\tiny]{\bf \faFlagCheckered}

%\Text[x=2.0cm,y=1.4cm,fontsize=\tiny]{\bf \footnotesize iC3C}

%\draw [{Implies}-,double,line width=1pt] (1.0,0.8) -- (1.9,2.1);

\draw[-Stealth,thick] (0,0) -- (5.1,0) node[anchor=north west] {\bf \tiny high};
\Text[x=0cm,y=-0.2cm,fontsize=\tiny]{\bf least}
\Text[x=2.6cm,y=-0.2cm,fontsize=\footnotesize]{Functions' Risk Level}
\Text[x=-0.2,y=2.6,rotation=90,fontsize=\footnotesize]{Conversion effort}
\draw[-Stealth,thick] (0,0) -- (0,5.1) node[anchor=south east] {\bf \tiny high};
\end{tikzpicture}
\caption{Flexibile partioning using~\systemname{}.}
\label{fig:mixedmodesupport}
\end{center}
\end{figure}

### 2023-12-01 sandp23/implementation.tex
update.

@@
 
-The trampoline function handles the invocations from sandbox (and hence the extra parameter~\code{sandbox}), performs the necessary conversion of pointer arguments, and eventually invokes the callback.
+The trampoline function handles the invocations from sandbox (and hence the extra parameter~\code{sandbox}), performs necessary pointer argument conversion, and eventually invokes the callback.
 
@@
 We create~\ucregion partition by first copying all functions and variables marked as~\code{_Tained} and~\code{_Mirror}.
-Next, we convert the partition into regular C by replacing all~\systemname{} tainted types with the corresponding C types.
+Next, we convert the partition into regular C by replacing all tainted types with the corresponding C types.
 \newline
@@
 The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
-We implemented the compiler as a clang library (1.8K SLoc) on top of the existing~\checkedc compiler.
+We implemented the compiler as a library (1.8K SLoc) on top of the existing~\checkedc compiler.
 Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.

### 2023-12-01 sandp23/introduction.tex
update.

@@
 
-These sets of source files will be compiled using~\systemname{} compiler, which enforces the desired isolation through instrumentation.
-Developers can choose a specific isolation mechanism.
-We explore four different isolation mechanisms with varying guarantees and performance overhead (CITE SECTION).
+%These sets of source files will be compiled using~\systemname{} compiler, which enforces the desired isolation through instrumentation.
+%Developers can choose a specific isolation mechanism.
+These sets of source files will be compiled using~\systemname{} compiler, which enforces a selected isolation mechanism.
+We introduce the concept of~\emph{\acf{SeFI}} and explore four different isolation mechanisms with varying guarantees and performance overhead (CITE SECTION).
 For instance,~\isoheap{} provides safety against spatial and temporal violations of tainted pointers with almost~\emph{no} performance overhead.
@@
 \myparagraph{Flexible Isolation Mechanisms and Implementation}
-We designed four isolation mechanisms with varying guarantees and performance overhead, providing flexibility in choosing a desired isolation mechanism (CITE SECTION).
+We introduce the concept of~\emph{\acf{SeFI}} and designed four isolation mechanisms with varying guarantees and performance overhead, providing flexibility in choosing a desired isolation mechanism (CITE SECTION).
 We implemented our type system (CITE SECTION) and flexible partitioning (CITE SECTION) by modifying~\clang{} compiler.

### 2023-12-01 sandp23/overview.tex
Fixing intro
update.

@@
 \Cref{fig:overview} shows the interaction between various components of~\systemname{}.
-Given an appropriately annotated source program and a sandbox configuration, our system creates an executable such that all tainted functions and data will reside in a sandbox and the~\cregion (non-sandboxed) is not affected by the code in the sandbox.
-We will present an overview of the~\systemname{} from a developer perspective and the details of individual components in the later sections.
+Given an appropriately annotated source program and a sandbox configuration, \systemname{} creates an executable such that all tainted functions and data reside in a sandbox and the~\cregion (non-sandboxed) is not affected by the code in the sandbox.
+Here, we brief \systemname{} from a developer perspective and the details of individual components in the later sections.
 
@@
 \label{subsec:identifyregionstosbx}
-to explain changes required by the \systemname compiler.The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
-The server runs in a loop and calls~\code{handle_request}, which handles each network request.
-The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} will be called to handle the request.
+
+To explain changes required by the \systemname compiler, \Cref{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
+The server runs in a loop and calls~\code{handle_request}, which handles a network request.
+The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} is called to handle the request.
 
 \myparagraph{The Vulnerability}
-As we can see, there is an arbitrary memory write vulnerability (indicated by~\realbug) in~\code{process_req1} because of using~\code{i} as an index into the array~\code{msg} without any sanity check.
-The variable~\code{i} can take any integer value as it is parsed from~\code{msg}, whose contents are read from socket in~\code{read_msg} indicated by~\rootcause.
+There is an arbitrary memory write vulnerability (indicated by~\realbug) in~\code{process_req1} because of using~\code{i} as an index into the array~\code{msg} without any sanity check.
+The variable~\code{i} can take any integer value as it is parsed from~\code{msg}, whose content is read from socket in~\code{read_msg} indicated by~\rootcause.
 
 \myparagraph{Goal} 
-The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the overall program.
-Ideally, the developer must convert the entire code to Checked C so that the entire program achieves spatial memory safety.
-However, Checked C has restrictions that can result in a considerable conversion effort.
-For instance,~\code{void*} pointers are not directly supported by Checked C.
+The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the other part of the program.
+Ideally, the developer can convert the entire code to \checkedc so that we achieve full spatial memory safety.
+However, there might be considerable conversion efforts in \checkedc.
+For instance,~\code{void*} pointers are not directly supported by \checkedc.
 Consequently, the developer needs to convert functions using~\code{void*} pointers into generic versions -- this could be tedious.
 To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
-However, as shown in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
-\aravind{Fix the running example code with the annotations according to the implementation.}
+However, as shown in~\sect{subsec:nosafetyagsintuncheckedcode}, $\umode$ region code can also affect the safety of checked pointers.
+%\aravind{Fix the running example code with the annotations according to the implementation.}
+
 \subsection{\systemname Annotations}
@@
 %\myparagraph{Using tainted types}
-To handle this, the developer marks risky functions with unchecked pointers,~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded.
-The developer starts using~\systemname{} with the resulting partially annotated program with checked and tainted types~\ie~\lst{lst:initialconv} and~\ref{lst:humantaint} atop of~\lst{lst:comb}.
+To overcome the above difficulty, The developer marks risky functions with unchecked pointers,~\ie~\code{read_msg} and~\code{process_req1} as~\code{tainted} as shown in~\lst{lst:humantaint} indicated by~\useradded by using with \systemname{},
+ and results in the partially annotated program with checked and tainted types~\ie~\Cref{lst:initialconv,lst:humantaint} atop of~\lst{lst:comb}.
 The initial tainted functions might require other pointers to be marked as tainted according to our typing rules (\sect{sec:typechecking}).
-The developer uses our type checker to identify the additional annotations (\usermods) required to have a well-typed program and adds them as shown in~\lst{lst:humanadjust}.
-The resulting well-typed program as shown in~\lst{lst:final} will be passed to our source level program partitioner along with certain configuration parameters of the target sandbox.
+The developer uses our type checker to identify the additional required annotations (\usermods) and adds them as shown in~\lst{lst:humanadjust}.
+The resulting well-typed program as shown in~\lst{lst:final} is passed to our source level program partitioner along with certain configuration parameters of the target sandbox.
 
@@
 Our partitioner splits the provided program into two sets of source files with the necessary source changes required to communicate with sandboxed code.
-These sets of source files will be compiled with the corresponding compilers to get the corresponding object files.
-The \cregion object file will have the necessary runtime checks enforcing~\systemname{} guarantees.
-The~\ucregion object file will be produced according to the corresponding sandbox mechanism.
-Finally, these two object files will be linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions will be executed in a sandbox (\ucregion) and the rest of the functions as regular code (\cregion).
+These sets of source files are compiled with the corresponding compilers to get the corresponding object files.
+The \cregion object file has the necessary runtime checks enforcing~\systemname{} guarantees.
+The~\ucregion object file is produced according to the corresponding sandbox mechanism.
+Finally, these two object files are linked along with the necessary sandbox libraries to produce the final executable such that all the tainted functions are executed in a sandbox (\ucregion) and the rest of the functions as regular code (\cregion).
 
